### 2021-09-03 src/data.prep.Rmd
initial commit

---
title: "data.prep"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(tidyverse)
library(rprojroot)

# This is the project path
path <- find_rstudio_root_file()
```

The internet archive data is saved on /hdd

```{r}
data.folder <- "/hdd/internet_archive/archive/data/"
data.path <- paste0(data.folder, "all.csv")

all <- fread(data.path)
#glimpse(all)
unique(all$year)

all$id <- seq.int(nrow(all))

# for test at the end
test.all <- dim(all[1])

# data.path_ <- paste0(data.folder, "all_sample.csv")
# all_ <-fread(data.path_)
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- fread(path.lookup)
#glimpse(lookup)
```
Merge

```{r}
setkey(lookup, pcds)
setkey(all, pc)
all.msoa <- lookup[all]

sapply(all.msoa, function(x) sum(is.na(x)))
```

The next snippet drops URLS from the same host referring to same postcode, which
were observed more than once per year. 
In other words, it drops URLs, which were archived multiple times per year.

```{r}
all.msoa <- unique(all.msoa, by = c("host", "year", "pcds"))
```

Aggregate

```{r}
msoa <- all.msoa[, .(count.url=.N), by = .(year, msoa11cd)]
```

Test to see if i have missed any of the original urls

```{r}
test.msoa <- sum(msoa$count.url)
ifelse(test.all==test.msoa, yes = "OK", no = "problem")

missing <- msoa[is.na(msoa$msoa11cd)]
print(sum(missing$count.url)) #
```

Drop the observations without MSOA code

```{r}
msoa <- msoa[!is.na(msoa11cd)]
```

Export

```{r}
path.out <- paste0(path, "/data/temp/msoa.csv")
fwrite(msoa, path.out) # for all obs
```

### 2022-04-13 src/data.prep.2011.2012.Rmd
update

---
title: "data.prep.2011-2012"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(urltools)

# This is the project path
path <- find_rstudio_root_file()
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  select(pcds, msoa11cd) %>% 
  dplyr::rename(pc = pcds)
  # the below is for a 'fuzzy' pc match for missing pc
  #mutate(pc_1 = str_sub(pc, start = 1, end = nchar(pc)-1))

#glimpse(lookup)
#problems(lookup)
#lookup[400296:400300,]
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd

# old pc file to fill a few hundred NA
old.pc.path <- "/hdd/internet_archive/archive/ONSPD_FEB_2017_UK.csv"
old.pc <- read_csv(old.pc.path) %>% 
  dplyr::select(pcds, msoa11) %>% 
  rename(msoa11.help = msoa11)

```

Download the 2011-2012 data.

**TODO** move to second drive.

```{r eval=F}
pre <- "https://www.webarchive.org.uk/datasets/ukwa.ds.2/geo/2011-201304/arcs/part-"
post <- ".bz2"  
numbers <- 0:49 
numbers <- ifelse(numbers < 10, paste0("0000", numbers), paste0("000", numbers))
urls <- lapply(numbers, function(x) paste0(pre, numbers, post))
urls <- urls[1]

for (url in urls) {
    #download.file(url, destfile = basename(url), method="curl", extra="-k")
    download.file(url, destfile = paste0(path,"/data/temp/", basename(url)))
}
```

Loop over the 50 .bz2 files of raw data for 2011-2012 and create a tibble.

```{r}

#bz2.path <- paste0(path,"/data/temp/part-00000.bz2")

all2011_12 <- tibble()

files <- list.files(path=paste0(path,"/data/temp"), pattern="*.bz2", full.names=TRUE, recursive=FALSE)

# files <-c("/home/nw19521/projects/web.diffusion/data/temp/part-00000.bz2",
#            "/home/nw19521/projects/web.diffusion/data/temp/part-00001.bz2")
# 
# x <- files[2]

#lapply(files, function(x) {

for (x in files){
  df <- read_delim(x, delim = "\t", col_names = F) %>% 
    rename(url = X1,
         pc = X2) %>%
    mutate(year = str_sub(url, start = 1, end = 4),
           newurl = substring(url,16),
           domain.ext = domain(newurl)) %>% 
    #suffix = suffix_extract(domain.ext)[,4])
    left_join(lookup, by = "pc", suffix =c("","")) %>% 
    left_join(old.pc, by=c("pc" = "pcds")) %>% 
    mutate(msoa11cd = ifelse(is.na(msoa11cd), msoa11.help,
                             msoa11cd)) %>% 
    select(-msoa11.help) %>% 
    filter(!is.na(msoa11cd))

  help <- suffix_extract(df$domain.ext) %>% 
    select(host, domain, suffix)

  #file.name = paste0(path,"/data/temp/data20112012_", substr(x[1], 56, 57), ".csv")

  df <- df %>% 
    bind_cols(help) %>% 
    select(-newurl, -domain.ext) %>% 
    filter(suffix == "co.uk")  # keep the .co.uk
    #write_csv(file = file.name)

  all2011_12 <- all2011_12 %>% bind_rows(df)

  rm(help)
 }

path.all2011_12 <- paste0(path, "/data/temp/all2011_2012.csv")
write_csv(all2011_12, path.all2011_12)
```
Create PC frequencies and bring them in all2011_2021.

As per the `loop_read_bb_new.R`:

`f <- a[, uniqueN(pc), by = .(host,year)][order(-V1)]`
`f.old <- a[, .N, by = .(host,year)][order(-N)] # old f`

I need to replicate V1 as this is what I have been using in other papers too.

```{r}
# # help code
# f <- test %>% 
#   group_by(domain.ext, year) %>%
#   summarise(V1 = n_distinct(pc))
# 
# f.old <- test %>% 
#   group_by(domain.ext, year) %>%
#   summarise(N = n())
# # equivalent to:
# test %>% count(domain.ext, year)
# 
# left_join(test, f, by = c("domain.ext" = "domain.ext",
#                           "year" = "year"))

f <- all2011_12 %>% 
  group_by(host, year) %>%
  summarise(V1 = n_distinct(pc))

all2011_12 <- all2011_12 %>% 
  left_join(f, by = c("host" = "host",
                      "year" = "year"))
#sapply(all2011_12, function(x) sum(is.na(x)))
```
Aggregate

```{r}
all2011_12 <- all2011_12 %>%

  # strips out the timestamp from the url
  mutate(url.no.date = substring(url, 16)) %>%

  # OLD 
  # Drops URLS from the same host referring to same postcode, which
  # are observed more than once per year. 
  # In other words, it drops URLs, which were archived multiple times per year.
  #distinct(host, pc,  .keep_all = TRUE) %>%

  # Drops URLs (no timestamp) referring to same postcode, which
  # are observed more than once per year. 
  # In other words, it drops pairs of URLs and postcodes, which were archived multiple times per year.
  distinct(url.no.date, pc, year, .keep_all = TRUE) %>% 
  # from 162,716,272 to 106,755,879 rows

  # Aggregate:
  # Counts how many unique pairs of url.no.date and pc exist in every MSOA per year.
  # In other words, it provides the number of georeferenced archived webpages per MSOA and year
  count(msoa11cd, year) %>%

  # Drops 2013 as we only have data for the 1st quarter
  filter(year!=2013)

# exports one file per year
all2011_12 %>% 
  group_by(year) %>%
  group_map(~ write_csv(.x, paste0(path,"/data/temp/msoa", .y$year, ".csv")))
```

Test with other PC files from https://geoportal.statistics.gov.uk/.
None of the below decreases NA substantially. The old file adds 
helps a bit, so I use this above.

```{r eval=F}
# I can only find a few more hunderd PC from the 46k missing.
old.pc.path <- "/hdd/internet_archive/archive/ONSPD_FEB_2017_UK.csv"
old.pc <- read_csv(old.pc.path)
missing_ <- left_join(missing, old.pc, by=c("pc" = "pcds"))
sapply(missing_, function(x) sum(is.na(x)))

# New pc file
#https://geoportal.statistics.gov.uk/datasets/national-statistics-postcode-lookup-august-2011/about
new.pc.file <- "/home/nw19521/projects/web.diffusion/data/raw/NSPL_AUG_2011_UK/Data/NSPL_AUG_2011.csv"
new.pc <- read_csv(new.pc.file, col_names = F)
missing_ <- left_join(missing, new.pc, by=c("pc" = "X1"))
sapply(missing_, function(x) sum(is.na(x)))

# NHS pc file
nhs.pc.file <- "/home/nw19521/projects/web.diffusion/data/raw/NHSPD_UK_FULL/Data/nhg21aug.csv"
nhs.pc <- read_csv(nhs.pc.file, col_names = F)
missing_ <- left_join(missing, nhs.pc, by=c("pc" = "X2"))
sapply(missing_, function(x) sum(is.na(x)))
```

### 2022-04-13 src/data.prep.Rmd
update

@@
 
-library(data.table)
 library(tidyverse)
@@
 
-The internet archive data is saved on /hdd
+The internet archive data is saved on /hdd.
+Split to yearly files.
 
@@
 
-all <- fread(data.path)
-#glimpse(all)
-unique(all$year)
-
-all$id <- seq.int(nrow(all))
-
-# for test at the end
-test.all <- dim(all[1])
-
-# data.path_ <- paste0(data.folder, "all_sample.csv")
-# all_ <-fread(data.path_)
+read_csv(data.path) %>% 
+  #distinct(host, year)
+  group_by(year) %>%
+  #group_walk(~ write_csv(.x, paste0(.y$year, ".csv"))) #
+  group_map(~ write_csv(.x, paste0(path,"/data/temp/msoa", .y$year, ".csv")))
 ```
-
 Postcode lookup.
@@
 path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
-lookup <- fread(path.lookup)
+lookup <- read_csv(path.lookup) %>% 
+  select(pcds, msoa11cd) %>% 
+  dplyr::rename(pc = pcds)
 #glimpse(lookup)
+# The problems refer to Welsh LAD names. Not a problem for the analysis.
+#sapply(lookup, function(x) sum(is.na(x)))
+# 10332 missing msoa11cd
 ```
-Merge
+
+Bring MSOA code and then aggregate per MSOA.
 
 ```{r}
-setkey(lookup, pcds)
-setkey(all, pc)
-all.msoa <- lookup[all]
 
-sapply(all.msoa, function(x) sum(is.na(x)))
-```
+# all <- read_csv(data.path)
+# all %>% 
+#   #mutate(id = 1:n()) %>% #seq.int(nrow())) %>% 
+#   #mutate(id = id(.)) %>% 
+#   left_join(lookup, by = "pc", suffix =c("","")) %>% 
+#   distinct(host, year, pc, .keep_all = TRUE) %>% 
+#   group_by(msoa11cd) %>% 
+#   summarise(count.url=sum(N))
+#   write_csv(paste0(path,"/data/temp/data", "X", ".csv"))
 
-The next snippet drops URLS from the same host referring to same postcode, which
-were observed more than once per year. 
-In other words, it drops URLs, which were archived multiple times per year.
+files <- list.files(path=paste0(path,"/data/temp"), pattern="*.csv", full.names=TRUE, recursive=FALSE)
+lapply(files, function(x) {
+  read_csv(x, col_names = T) %>% 
+    #mutate(id = id(.)) %>%  
+    left_join(lookup, by = "pc", suffix =c("","")) %>%
 
-```{r}
-all.msoa <- unique(all.msoa, by = c("host", "year", "pcds"))
-```
+    # strips out the timestamp from the url
+    mutate(url.no.date = substring(url, 16)) %>%
 
-Aggregate
+    # OLD
+    # Drops URLS from the same host referring to same postcode, which
+    # are observed more than once per year. 
+    # In other words, it drops URLs, which were archived multiple times per year.
+    #distinct(host, pc,  .keep_all = TRUE) %>%
 
-```{r}
-msoa <- all.msoa[, .(count.url=.N), by = .(year, msoa11cd)]
+    # Drops URLs (no timestamp) referring to same postcode, which
+    # are observed more than once per year. 
+    # In other words, it drops pairs of URLs and postcodes, which were archived multiple times per year.
+    distinct(url.no.date, pc, .keep_all = TRUE) %>%
+
+    # OLD Aggregate
+    #group_by(msoa11cd) %>% 
+    #summarise(count.url=sum(N)) %>%
+
+    # Aggregate:
+    # Counts how many unique pairs of url.no.date and pc exist in every MSOA per year.
+    # In other words, it provides the number of georeferenced archived webpages per MSOA and year
+    count(msoa11cd) %>%
+
+    write_csv(x)
+})
+
+# for test at the end
+#test.all <- dim(all[1])
+
+# data.path_ <- paste0(data.folder, "all_sample.csv")
+# all_ <-fread(data.path_)
 ```
@@
 ```
-
-Drop the observations without MSOA code
-
-```{r}
-msoa <- msoa[!is.na(msoa11cd)]
-```
-
-Export
-
-```{r}
-path.out <- paste0(path, "/data/temp/msoa.csv")
-fwrite(msoa, path.out) # for all obs
-```

### 2022-04-13 src/diffusion.Rmd
update

---
title: "diffusion"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(urltools)
library(ggrepel)
library(tsbox)
library(dtwclust)
library(imputeTS)
library(rgdal)
library(broom)
library(cowplot)

#lisa
library(sp)
library(sf)
library(fastmap)
library(skimr)
library(tmap)
library(tmaptools)
library(spdep)
library(rgeos)
library(rgeoda)

options(scipen=999)

# This is the project path
path <- find_rstudio_root_file()
```

Loading yearly files

```{r}
years <- 1996:2012
path.files <- paste0(path, "/data/temp/msoa")
file.names <- lapply(years, function(x) paste0(path.files, years, ".csv"))
file.names <- unlist(file.names[1])

df <- data.frame()
year <- 1995
for (i in file.names){
  x <- read_csv(i)
  year <- year + 1
  x$year <- year
  df <- rbind(a,x)
}

df <- df %>% 
  arrange(msoa11cd, year) %>% 
  drop_na() %>% 
  glimpse()

sapply(df, function(x) sum(is.na(x)))

# **NEDD TO SEE WHY I HAVE DUPLICATES**
df <- df %>% distinct(msoa11cd, year, .keep_all = TRUE)

total <- df %>% 
  group_by(year) %>% 
  summarise(n = sum(n)) %>% 
  glimpse()
```
The MSOA plot takes ages to be drawn!!!

```{r}
for.plot <- all[all$YEAR>1999 & all$YEAR<2011,]
for.plot[, outlier := ifelse(for.plot$counturlcouk>150000 & for.plot$YEAR==2008, for.plot$regionme, "")] #this line is from the previous version
#for.plot <- for.plot[for.plot$outlier=="OK",]

tspag = 
  ggplot(df, aes(x=year, y=n, group = msoa11cd, colour = factor(msoa11cd))) + 
  geom_line() + guides(colour=FALSE) + xlab("Year") +
  ylab("N. of .co.uk webpages, NUTS3") +
  geom_text_repel(aes(label=msoa11cd), cex = 4) + #this line is from the previous version
  #xlim(2000, 2010) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(labels = scales::number_format(accuracy = 1))
  #theme(plot.margin = unit(c(1,3,1,1), "lines")) +
  #geom_dl(aes(label = outlier), method = list(dl.combine("first.points", "last.points"), cex = 0.8)) + 
  #scale_x_continuous(breaks=c(1996:2010 )) 
  #scale_color_discrete(labels = outlier) 
  #scale_fill_discrete(name="test")
plot(tspag)

ggplot(data=df, aes(x=year,y=n,group=msoa11cd, colour = msoa11cd)) + geom_line(alpha=I(1/7),size=1)

outliers <-df %>% filter(n>900000) %>% 
  select(msoa11cd) %>% 
  pull()

df %>% 
  #rowwise() %>% 
  filter(n<500000) %>% 
  #filter(! msoa11cd %in% outliers)
  ggplot(aes(x=year,y=n,group=msoa11cd)) + geom_line(alpha=I(1/7),size=.5) +
  scale_y_continuous(labels = scales::comma)

total %>% mutate(n.max = n/max(n)) %>% 
ggplot(aes(x=year, y=n.max)) + 
  geom_line() + guides(colour=FALSE) + xlab("Year") +
  ylab("N. of .co.uk webpages, NUTS3") +
  #geom_text_repel(aes(label=msoa11cd), cex = 4) + #this line is from the previous version
  #xlim(2000, 2010) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(labels = scales::number_format(accuracy = 1))

```

time series clusters

```{r}
# turn to a wide format
ts.wide <- df %>% 
  rename(time = year) %>%
  select(time, msoa11cd, n) %>% 
  #mutate(year = )
  ts_wide()

# impute with means of each TS
# TODO check sensitivity
ts.wide <- na_mean(ts.wide)
#sapply(ts.wide, function(x) sum(is.na(x)))

# extract names
ts.wide.names <- names(ts.wide[,-1])
ts.wide <- ts.wide[,-1]

# standardise
ts.wide <- zscore(ts.wide)

# convert to list for tsclust
ts.wide <- split(ts.wide, rep(1:ncol(ts.wide), each = nrow(ts.wide)))

# 5 - 20 cluster sollutions
pc_k <- tsclust(ts.wide, k = c(5:15), seed = 94L,
                type="partitional",
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))

# decide k, see RJ-2019-023 for max/min
names(pc_k) <- paste0("k_", c("5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15"))
cvis <- sapply(pc_k, cvi, type = "internal")
cvis
# max Sil
# max SF
# max CH
# min DB
# min DBstar
# max D
# min COP

# loop for k decision
pref.cvi <- NULL

max.cvis <- c(1:3,6)
for (i in max.cvis){
  a <- which.max(cvis[i,])
  pref.cvi <- c(pref.cvi, a)
}

min.cvis <- c(4:5,7)
for (i in min.cvis){
  a <- which.min(cvis[i,])
  pref.cvi <- c(pref.cvi, a)
}

print(pref.cvi) # cvis[2,] is always 0

pc_k <- tsclust(ts.wide, k = 14, seed = 94L, 
                type="partitional", 
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))
plot(pc_k)

clusters <- data.frame(LAD=ts.wide.names,
                           cluster=pc_k@cluster)

```

## Mapping the clusters

```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# get MSOA 
# source: https://geoportal.statistics.gov.uk/datasets/middle-layer-super-output-areas-december-2011-boundaries-generalised-clipped-bgc-ew-v3/explore?location=52.849644%2C-2.489527%2C7.08

path.geo <- paste0(path, "/data/raw/Middle_Layer_Super_Output_Areas_(December_2011)_Boundaries_Generalised_Clipped_(BGC)_EW_V3.geojson")
msoa <- readOGR(path.geo)
# UK BFC: la <- readOGR("https://opendata.arcgis.com/datasets/1d78d47c87df4212b79fe2323aae8e08_0.geojson")

# source: https://data.gov.uk/dataset/7c387c64-d25f-474a-b07e-b933578caea2/local-authority-districts-april-2019-boundaries-uk-bfe

# spatial transformations
msoa <- spTransform(la, CRS("+init=epsg:4326"))

# ggplot2 maps
msoa.f <- fortify(msoa, region = "MSOA11CD")
msoa.f <- merge(msoa.f, clusters, by.x = "id", by.y = "LAD")
msoa.f <- msoa.f[order(msoa.f$order),] # if i don't order merge.nuts.f loses order and the map has gaps

ggplot(msoa.f, aes(x = long, y = lat)) +
  geom_polygon(aes( group = group, fill = as.factor(cluster))) +
  #theme_nothing(legend = TRUE) +
  labs(title = "Clusters") +
  scale_fill_viridis_d()+
  theme_void()

ggplot(la.f, aes(x = long, y = lat)) +
  geom_polygon(aes( group = group, fill = as.factor(cluster.down))) +
  theme_nothing(legend = TRUE) +
  labs(title = "Download clusters") +
  scale_fill_viridis_d()+
  theme_void()

cluster.upL <- c(6, 9, 11, 12, 13)
cluster.upS <- c(1, 2, 3, 7, 8, 10)

la.f_ <- la.f #plot the background

cluster.upL.plot <- la.f %>% 
  filter(cluster.up %in% cluster.upL) %>% 
  ggplot(aes(x = long, y = lat)) +
  geom_polygon(data = la.f_, color='grey', fill='white', aes(x=long, y=lat, group=group, colour = 'grey30')) +
  geom_polygon(aes( group = group, fill = as.factor(cluster.up))) +
  labs(title = "Larger clusters, upload speeds") + 
  scale_fill_viridis_d() +
  guides(fill=guide_legend(title="Clusters")) +
  theme_void()

cluster.upS.plot <- la.f %>% 
  filter(cluster.up %in% cluster.upS) %>% 
  ggplot(aes(x = long, y = lat)) +
  geom_polygon(data = la.f_, color='grey', fill='white', aes(x=long, y=lat, group=group, colour = 'grey30')) +
  geom_polygon(aes( group = group, fill = as.factor(cluster.up))) +
  labs(title = "Smaller clusters, upload speeds") + 
  scale_fill_viridis_d() +
  guides(fill=guide_legend(title="Clusters")) +
  theme_void()

plot_grid(cluster.upL.plot, cluster.upS.plot, nrow = 1,
          align = "h", axis = "l", 
          #labels = c(".co.uk / pop., 2000", ".co.uk / pop., 2010", "non .co.uk / pop., 2000", "non .co.uk / pop., 2010", ""), 
          #label_size = 12,
          greedy = T)
          #rel_heights = c(1, 1, .3)) # this is the relative position of the 3 lines

#ggsave("all_maps2.png") # this is the colour map for the preprint
out.path.map <- paste0(path, "/paper/v1/figures/map.up.clusters.png")
ggsave(out.path.map)#,
       #width = 210, #A4
       #height = 297, #A4
       #units = "mm")
```

## maps

```{r}

t = 1996

maps <- msoa
maps.df <- fortify(maps, region = "MSOA11CD")

maps.df <- maps.df %>% left_join(df[df$year==t,], by = c("id"="msoa11cd")) %>% 
  mutate(n = ifelse(is.na(n), 0, n),
         year = ifelse(is.na(year), 1996, year))

#plot
options(scipen=10000)

base <- ggplot(data = maps.df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed() + 
  geom_polygon(color = NA, fill = "white")

ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
)

#N.web
base + 
  geom_polygon(aes(fill = n), color = NA) +
  theme_bw() +
  ditch_the_axes + ggtitle(t) +
  theme(legend.title=element_text(size=5), legend.text=element_text(size=4)) +
  guides(fill = guide_colorbar(barwidth = 0.5, barheight = 2.5)) +
  scale_fill_gradientn(colours=c('light gray','dark blue'),
                     name='N. of web pages',na.value=NA)#, trans = "log10") #  breaks = c(100, 1000, 10000)
  #coord_map(xlim = c(200000, 300000),ylim = c(200000, 300000)) # zoom in
ggsave("C:/net27/COSMIC/archive/gis/msoa_Nweb2001.png", dpi = 300, width=4, height=4)

```

## lisa

```{r}

# 0 for NAs and the same for year
test <- msoa
test@data <- test@data %>% left_join(df[df$year==1996,], by = c("MSOA11CD"="msoa11cd")) %>% 
  mutate(n = ifelse(is.na(n), 0, n),
         year = ifelse(is.na(year), 1996, year))

#spatial weight matrix
nb<- poly2nb(test)
listw<- nb2listw(nb, style = "W", zero.policy = T)

#Moran's I
globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
globalMoran

#LISA
lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
#lmoran

# centers the local Moran's around the mean
lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)

# centers the variable of interest around its mean

lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)

signif <- 0.05
#lmoran

lmoran <- lmoran%>% 
  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%

#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))

mun_merge_new<- merge(test, lmoran, by="MSOA11CD")

# R p value map
breaks = c(0, 1, 2, 3, 4, 5)

tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, palette=  c("white","red","blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha=0.4)), labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), title="") +
  tm_legend(text.size = 1) +
 # tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
  # tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
    tm_borders(alpha=.5) +
   tm_layout( frame = FALSE,  title = "LISA with the R p-values ")

```

### 2022-04-13 src/domain1112.Rmd
update

---
title: "domain1112"
output: html_document
---

The following creates a file with a unique pc per domain per year. 
The `all.csv` was based on all pc per host per year.
Here, I am using the domain, which in essense is the 3rd level DN.

It is based on the `/archive/data/domain.r` but for the last 2011-2012 data.

```{r setup, include=FALSE}
library(data.table)
library(kableExtra)
library(rprojroot)

path <- find_rstudio_root_file()
```

```{r}

data.path2011_2012 <- paste0(path, "/data/temp/all2011_2012.csv")
#add comments
all1112 <- read_csv(data.path2011_2012)
all1112 <- as.data.table(all1112)

f <- all1112[, uniqueN(pc), by = .(domain,year)][order(-V1)]

setkey(all1112, domain, year) # table, column
setkey(f, domain, year) #

# perform the join, eliminating not matched rows from Right
dim(all1112)
dim(f)
all1112<-all1112[f] #, nomatch=0

dim(all1112)
setnames(all1112, old=("V1"), new=("V1.domain"))

# unique rows: 1 row per domain, postcode and year. No URLs
all1112.unique <- unique(all1112, by = c("domain", "pc", "year"))
dim(all1112.unique)
all1112.unique$url <- NULL
all1112.unique$host <-NULL
all1112.unique$V1 <-NULL
all1112.unique$N <- NULL
all1112.unique$msoa11cd <- NULL
all1112.unique$suffix <- NULL

# export data
path.out <- paste0(path, "/data/temp/domain_pc_year1112.csv")
fwrite(all1112.unique, path.out)
```

### 2022-04-13 src/growth_lisa.Rmd
update

---
title: "diffusion, websites with 1 pc"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(rgdal)
library(spdep)
library(tmap)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  select(pcds, msoa11cd) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive data is saved on /hdd.

```{r}
n = 1 #number of unique postcodes

data.folder <- "/hdd/internet_archive/archive/data/"
data.path <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df <- read_csv(data.path) %>% 
  filter(V1.domain == n,
         year > 1995) %>% 
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, msoa11cd) %>%
  summarise(n = sum(V1.domain)) %>% 
  group_by(msoa11cd) %>%                   # turn on for growth
  mutate(growth = (n - lag(n))/lag(n)) %>% # turn on for growth 
  filter(year != 1996 )
```

## maps

```{r}
# get MSOA 
# source: https://geoportal.statistics.gov.uk/datasets/middle-layer-super-output-areas-december-2011-boundaries-generalised-clipped-bgc-ew-v3/explore?location=52.849644%2C-2.489527%2C7.08

path.geo <- paste0(path, "/data/raw/Middle_Layer_Super_Output_Areas_(December_2011)_Boundaries_Generalised_Clipped_(BGC)_EW_V3.geojson")
msoa <- readOGR(path.geo)
# UK BFC: la <- readOGR("https://opendata.arcgis.com/datasets/1d78d47c87df4212b79fe2323aae8e08_0.geojson") 
# source: https://data.gov.uk/dataset/7c387c64-d25f-474a-b07e-b933578caea2/local-authority-districts-april-2019-boundaries-uk-bfe

# spatial transformations
msoa <- spTransform(msoa, CRS("+init=epsg:4326"))
```

```{r, eval=F}
maps <- msoa
maps.df <- fortify(maps, region = "MSOA11CD")

base <- maps.df %>% left_join(df[df$year==t,], by = c("id"="msoa11cd")) %>% 
  # mutate(n = ifelse(is.na(n), 0, n),
  #        year = ifelse(is.na(year), 1996, year))
  ggplot(mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed() + 
  geom_polygon(color = NA, fill = "white")

ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
)

#N.web
base + 
  geom_polygon(aes(fill = growth), color = NA) +
  theme_bw() +
  ditch_the_axes + ggtitle(t) +
  theme(legend.title=element_text(size=5), legend.text=element_text(size=4)) +
  guides(fill = guide_colorbar(barwidth = 0.5, barheight = 2.5)) +
  scale_fill_gradient2(midpoint = 0, mid="#eee8d5", high="#dc322f", low="#268bd2")

  scale_fill_gradientn(colours=c('light gray','dark blue'),
                     name='N. of web pages',na.value=NA)#, trans = "log10") #  breaks = c(100, 1000, 10000)
  #coord_map(xlim = c(200000, 300000),ylim = c(200000, 300000)) # zoom in

#ggsave("C:/net27/COSMIC/archive/gis/msoa_Nweb2001.png", dpi = 300, width=4, height=4)

```

## lisa

```{r}

# 0 for NAs and the same for year

path.out <- paste0(path, "/outputs/lisa_growth_msoa")

for (t in 1997:2010){
test <- msoa
test@data <- test@data %>% left_join(df[df$year==t,], by = c("MSOA11CD"="msoa11cd")) #%>% 
  # mutate(n = ifelse(is.na(n), 0, n),
  #        year = ifelse(is.na(year), 1996, year))

#spatial weight matrix
nb<- poly2nb(test)
listw<- nb2listw(nb, style = "W", zero.policy = T)

#Moran's I
globalMoran <- moran.test(test$growth, listw, zero.policy = T, na.action = na.exclude)
globalMoran

#LISA
lmoran<- cbind(test@data, localmoran(test$growth, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
#lmoran

# centers the local Moran's around the mean
lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
lmoran$lag.growth <- lag.listw(listw,lmoran$growth, NAOK = TRUE)

# centers the variable of interest around its mean

lmoran$ns <- lmoran$growth - mean(lmoran$growth, na.rm = TRUE) 
lmoran$lag.growth <- lmoran$lag.growth - mean(lmoran$lag.growth, na.rm = TRUE)

signif <- 0.05
#lmoran

lmoran <- lmoran%>% 
  mutate(quadrant= ifelse(ns>0 & lag.growth > 0, 1, 0)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.growth < 0, 2, quadrant)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.growth > 0, 3, quadrant)) %>% 
  mutate(quadrant= ifelse(ns>0 & lag.growth < 0, 4, quadrant)) %>%   
  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%

#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))

mun_merge_new<- merge(test, lmoran, by="MSOA11CD")

# R p value map
breaks = c(0, 1, 2, 3, 4, 5)

map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, palette=  c("white","red","blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha=0.4)), labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), title="LISA for growth, MSOA") +
  tm_legend(text.size = 1) +
 # tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
  # tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
    tm_borders(alpha=.5) +
   tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))

}
```

Spatial regression

```{r}
library(splm)
m <- spml(log(n)~lag(log(n)), data = df, listw = listw, lag = TRUE, spatial.error = "b", 
          model = "within", effect = "individual", method = "eigen", na.action = na.fail)

```

### 2022-04-13 src/growth_lisa_oa.Rmd
update

---
title: "diffusion, websites with 1 pc"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(rgdal)
library(spdep)
library(tmap)
library(maptools)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  select(pcds, oa11cd) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

```{r}
n = 1 #number of unique postcodes

data.folder <- "/hdd/internet_archive/archive/data/"
data.path <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path)

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012)

df <- bind_rows(df9610, df1112) %>% 
  #arrange(year, domain) %>% 
  filter(V1.domain == n,
         year > 1995) %>% 
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, oa11cd) %>%
  summarise(n = sum(V1.domain)) %>% 
  group_by(oa11cd) %>%                   # turn on for growth
  mutate(growth = (n - lag(n))/lag(n)) %>% # turn on for growth 
  filter(year != 1996)

rm(df9610, df1112)

# tests for Scotland
# df %>% filter(substr(oa11cd, 1,1) =="S",
#               year==2010) %>%
#   summarise(mean(n, na.rm = T))
# 
# df %>% filter(oa11cd=="E00000001",
#               year==2010)
# 
# test@data %>% filter(substr(id, 1,1) =="S",
#                      n!=0)
# 
# df %>% filter(substr(oa11cd, 1,1) =="W",
#               year==2003)
# 
# test@data %>% filter(id=="S00088961",
#                      year==2010)
```

## maps

```{r}
# get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))

# keep in the data slot only the ONS Output Area id, renaming it as 'id'
oa.ew <- oa.ew[, 'OA11CD']
colnames(oa.ew@data) <- c('id')

# reassign the polygon IDs
oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.ew)

# get OA for Scotland
path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries

# spatial transformations
oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))

# Scotland (follows same steps as EW, see notes above)
oa.sc <- oa.sc[, 'code']
colnames(oa.sc@data) <- c('id')

# reassign the polygon IDs
oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.sc)

# build oa for GB
oa.gb <- spRbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)

# get GB
path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
gb <- readOGR(path.gb)
# spatial transformations
gb <- spTransform(gb, CRS("+init=epsg:4326"))

```

## lisa

```{r}

path.out <- paste0(path, "/outputs/lisa_growth_pc", n, "_oa")
morani <- data.frame()

#spatial weight matrix
sf_use_s2(FALSE)
nb <- poly2nb(oa.gb)
listw<- nb2listw(nb, style = "W", zero.policy = T)

for (t in 1997:2012){
test <- oa.gb
test@data <- test@data %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) #%>% 
  # mutate(n = ifelse(is.na(n), 0, n),
  #        year = ifelse(is.na(year), 1996, year))

#spatial weight matrix
#moved out of the loop

#Moran's I
globalMoran <- moran.test(test$growth, listw, zero.policy = T, na.action = na.exclude)
a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
morani <- rbind(a, morani)

#LISA
lmoran<- cbind(test@data, localmoran(test$growth, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
#lmoran

# centers the local Moran's around the mean
lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
lmoran$lag.growth <- lag.listw(listw,lmoran$growth, NAOK = TRUE)

# centers the variable of interest around its mean

lmoran$ns <- lmoran$growth - mean(lmoran$growth, na.rm = TRUE) 
lmoran$lag.growth <- lmoran$lag.growth - mean(lmoran$lag.growth, na.rm = TRUE)

signif <- 0.05
#lmoran

lmoran <- lmoran%>% 
  mutate(quadrant= ifelse(ns>0 & lag.growth > 0, 1, 0)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.growth < 0, 2, quadrant)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.growth > 0, 3, quadrant)) %>% 
  mutate(quadrant= ifelse(ns>0 & lag.growth < 0, 4, quadrant)) %>%   
  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%

#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))

mun_merge_new<- merge(test, lmoran, by="id")

# R p value map
breaks = c(0, 1, 2, 3, 4, 5)

map <- tm_shape(mun_merge_new) + 
  tm_fill(col = "quadrant", breaks = breaks, palette = c("white","red","blue",
                                                         rgb(0,0,1,alpha=0.4),
                                                         rgb(1,0,0,alpha=0.4)), 
          labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
          title="LISA for growth, OA") +
  tm_legend(text.size = 1) +
  # tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
  # tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
  # tm_borders(alpha=.5) +
  tm_borders(lwd = 0) +
  tm_shape(gb, simplify = .2) + tm_borders(alpha=1, lwd = .5) +
  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values")
 tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
}
```

Moran I

```{r}
morani %>% as_tibble() %>% 
  rename('morani'=V1,
                   p = V2,
                   year = V3) %>% 
  kable()
```

Spatial regression

```{r}
library(splm)
m <- spml(log(n)~lag(log(n)), data = df, listw = listw, lag = TRUE, spatial.error = "b", 
          model = "within", effect = "individual", method = "eigen", na.action = na.fail)

```

### 2022-04-13 src/hdbscan.Rmd
update

---
title: "hdbscan"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(rgdal)
library(spdep)
library(dbscan)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  select(pcds, msoa11cd, dointr, doterm) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd

path.lookup <- "/hdd/internet_archive/archive/ONSPD_FEB_2017_UK.csv"
lookup <- read_csv(path.lookup) %>% 
  select(pcds, oseast1m, osnrth1m) %>% 
  dplyr::rename(pc = pcds)

```

The internet archive data is saved on /hdd.

```{r}
n = 1 #number of unique postcodes

data.folder9610 <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder9610, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

data.path1112 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610) 
df1112 <- read_csv(data.path1112)

df <- bind_rows(df9610, df1112) %>% 
  filter(V1.domain == n,
         year > 1995) %>% 
  left_join(lookup, by = "pc", suffix =c("",""))

rm(df9610, df1112)
```

## Boundary data

```{r}
# get MSOA 
# source: https://geoportal.statistics.gov.uk/datasets/middle-layer-super-output-areas-december-2011-boundaries-generalised-clipped-bgc-ew-v3/explore?location=52.849644%2C-2.489527%2C7.08

# path.geo <- paste0(path, "/data/raw/Middle_Layer_Super_Output_Areas_(December_2011)_Boundaries_Generalised_Clipped_(BGC)_EW_V3.geojson")
# msoa <- readOGR(path.geo)
# # UK BFC: la <- readOGR("https://opendata.arcgis.com/datasets/1d78d47c87df4212b79fe2323aae8e08_0.geojson") 
# # source: https://data.gov.uk/dataset/7c387c64-d25f-474a-b07e-b933578caea2/local-authority-districts-april-2019-boundaries-uk-bfe
# 
# # spatial transformations
# msoa <- spTransform(msoa, CRS("+init=epsg:4326"))

# get GB
path.gb <- paste0(path, "/data/raw/Countries_(December_2020)_UK_BUC.geojson")
gb <- readOGR(path.gb) 
# spatial transformations
gb <- spTransform(gb, CRS("+init=epsg:4326"))
gb <- subset(gb, CTRY20NM != "Northern Ireland")
gb.df <- fortify(gb, region = "CTRY20NM")

# UK boundaries
# path.uk <- paste0(path, "/data/raw/Countries_(December_2020)_UK_BUC.geojson")
# uk <- readOGR(path.uk)
# uk <- spTransform(uk, CRS("+init=epsg:4326"))
# uk.df <- fortify(uk, region = "CTRY20CD")
# 
# # find points not in GB
# help2 <- df %>% 
#   #slice_sample(n = 10000) %>% 
#   filter(!is.na(oseast1m), #6k rows with missing east/north for all years
#               year == t) %>%
#   st_as_sf(coords = c("oseast1m", "osnrth1m"), crs = 27700) %>%
#   st_transform(4326) %>%
#   st_coordinates() %>%
#   as_tibble() 
# 
# help2 <- df %>% 
#   filter(!is.na(oseast1m), #6k rows with missing east/north for all years
#               year == t) %>%
#   bind_cols(help2)
# 
# 
# points <- st_as_sf(help2, coords = c("X", "Y"), crs = 27700)
# gb_ <- st_as_sf(gb) #, crs = 27700)
# gb_ <- st_set_crs(gb_, 27700)
# 
# plot(gb_, col = "red")
# plot(points, pch = 19, cex = 1, add = T)
# 
# pip <- points %>%
#   st_join(gb_)
# 
# unique(pip$ctry14nm)
# missing <- pip %>% filter(is.na(ctry14nm)) %>% 
#   mutate(pc2dig = substr(pc, 1,2)) 
# unique(missing$pc2dig)
# 
# missing <- pip %>% filter(is.na(ctry14nm)) %>% 
#   filter(substr(pc, 1,2) !="BT")
# in 2010 58 points exl. NI have a postcode but not XY

# cities
cities <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  arrange(pop) %>% tail(10) 
tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)

```

## hdbscan

```{r}
t = 1998

x <- df %>%
    #slice_sample(n = 10000) %>% 
    filter(!is.na(oseast1m), #6k rows with missing east/north for all years
           year == t,
           substr(pc, 1,2) !="BT") %>% #drop Northern Ireland, for 2010 c. 4k obs.
    st_as_sf(coords = c("oseast1m", "osnrth1m"), crs = 27700) %>%
    st_transform(4326) %>%
    st_coordinates() %>%
    as_tibble()

kNNdistplot(x, k = 2)               # it runs on the console
abline(h=.3, col = "red", lty=2)

res <- dbscan(x, eps = 0.1, minPts = 3)

plot(x, col = res$cluster + 1L, pch = res$cluster + 1L)
hullplot(x, res)

res <- optics(x, eps = 10, minPts = 10)
```

## hex heatmaps

```{r}
# old code: counts per hexagon based on xy
#xy point

path.out <- paste0(path, "/outputs/hex", n)

for (t in 1996:2012){
  p <- df %>%
    #slice_sample(n = 10000) %>% 
    filter(!is.na(oseast1m), #6k rows with missing east/north for all years
           year == t,
           substr(pc, 1,2) !="BT") %>% #drop Northern Ireland, for 2010 c. 4k obs.
    st_as_sf(coords = c("oseast1m", "osnrth1m"), crs = 27700) %>%
    st_transform(4326) %>%
    st_coordinates() %>%
    as_tibble() %>% 
    ggplot(aes(x=X,y=Y)) +

    stat_binhex(bins=150) + 
    geom_polygon(data = gb.df, aes(x=long, y = lat, group = group), #gb.df
                 fill=NA, color = "black", size = 0.1, alpha = 0.8)+
    scale_fill_gradientn(colours=c('light gray','blue'),
                         name='Frequency',na.value=NA) + #, trans = "log10") +
    coord_equal() +
    labs(x=NULL, y=NULL) +
    geom_point(data = cities, aes(x=long, y=lat), colour = "black", size = .1) + 
    geom_text(size = 2, check_overlap = F, nudge_y = .2, data = cities, aes(x=long, y=lat, label = name)) +
    #theme(legend.key.size = unit(0.05, "cm")) + # legend.key.width=unit(1, "cm"), 
    ggmap::theme_nothing(legend = T) +
    theme(legend.title=element_text(size=5.5), legend.text=element_text(size=4)) +
    #theme(legend.position=c(0.075,0.28))
    guides(fill = guide_colorbar(barwidth = 0.5, barheight = 2.5)) +
    ggtitle(t)
  ggsave(p, filename = paste(path.out, t, ".png", sep = ""))
}
```

### 2022-04-13 src/hex_heatmaps.Rmd
update

--- 2022-04-13 src/hdbscan.Rmd
@@
 ---
-title: "hdbscan"
+title: "diffusion, websites with 1 pc"
 date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 library(spdep)
-library(dbscan)
+library(tmap)
 
@@
 
-## hdbscan
-
-```{r}
-t = 1998
-
-x <- df %>%
-    #slice_sample(n = 10000) %>% 
-    filter(!is.na(oseast1m), #6k rows with missing east/north for all years
-           year == t,
-           substr(pc, 1,2) !="BT") %>% #drop Northern Ireland, for 2010 c. 4k obs.
-    st_as_sf(coords = c("oseast1m", "osnrth1m"), crs = 27700) %>%
-    st_transform(4326) %>%
-    st_coordinates() %>%
-    as_tibble()
-
-kNNdistplot(x, k = 2)               # it runs on the console
-abline(h=.3, col = "red", lty=2)
-
-res <- dbscan(x, eps = 0.1, minPts = 3)
-
-plot(x, col = res$cluster + 1L, pch = res$cluster + 1L)
-hullplot(x, res)
-
-res <- optics(x, eps = 10, minPts = 10)
-```
-
 ## hex heatmaps

### 2022-04-13 src/level_lisa.Rmd
update

--- 2022-04-13 src/growth_lisa.Rmd
@@
 
+Replace `msoa11cd1` with `oa11cd` for OA.
+
 ```{r}
@@
   group_by(year, msoa11cd) %>%
-  summarise(n = sum(V1.domain)) %>% 
-  group_by(msoa11cd) %>%                   # turn on for growth
-  mutate(growth = (n - lag(n))/lag(n)) %>% # turn on for growth 
-  filter(year != 1996 )
+  summarise(n = sum(V1.domain)) 
 ```
@@
 
-```{r, eval=F}
-maps <- msoa
-maps.df <- fortify(maps, region = "MSOA11CD")
-
-base <- maps.df %>% left_join(df[df$year==t,], by = c("id"="msoa11cd")) %>% 
-  # mutate(n = ifelse(is.na(n), 0, n),
-  #        year = ifelse(is.na(year), 1996, year))
-  ggplot(mapping = aes(x = long, y = lat, group = group)) + 
-  coord_fixed() + 
-  geom_polygon(color = NA, fill = "white")
-
-ditch_the_axes <- theme(
-  axis.text = element_blank(),
-  axis.line = element_blank(),
-  axis.ticks = element_blank(),
-  panel.border = element_blank(),
-  panel.grid = element_blank(),
-  axis.title = element_blank()
-)
-
-#N.web
-base + 
-  geom_polygon(aes(fill = growth), color = NA) +
-  theme_bw() +
-  ditch_the_axes + ggtitle(t) +
-  theme(legend.title=element_text(size=5), legend.text=element_text(size=4)) +
-  guides(fill = guide_colorbar(barwidth = 0.5, barheight = 2.5)) +
-  scale_fill_gradient2(midpoint = 0, mid="#eee8d5", high="#dc322f", low="#268bd2")
-
-  scale_fill_gradientn(colours=c('light gray','dark blue'),
-                     name='N. of web pages',na.value=NA)#, trans = "log10") #  breaks = c(100, 1000, 10000)
-  #coord_map(xlim = c(200000, 300000),ylim = c(200000, 300000)) # zoom in
-
-#ggsave("C:/net27/COSMIC/archive/gis/msoa_Nweb2001.png", dpi = 300, width=4, height=4)
-
-```
-
 ## lisa
@@
 
-path.out <- paste0(path, "/outputs/lisa_growth_msoa")
+path.out <- paste0(path, "/outputs/lisa_n_msoa")
 
-for (t in 1997:2010){
+for (t in 1996:2010){
 test <- msoa
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("MSOA11CD"="msoa11cd")) #%>% 
-  # mutate(n = ifelse(is.na(n), 0, n),
-  #        year = ifelse(is.na(year), 1996, year))
+test@data <- test@data %>% left_join(df[df$year==t,], by = c("MSOA11CD"="msoa11cd")) %>% 
+  mutate(n = ifelse(is.na(n), 0, n),
+         year = ifelse(is.na(year), t, year))
 
@@
 #Moran's I
-globalMoran <- moran.test(test$growth, listw, zero.policy = T, na.action = na.exclude)
+globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
 globalMoran
@@
 #LISA
-lmoran<- cbind(test@data, localmoran(test$growth, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
+lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
 #lmoran
@@
 lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
-lmoran$lag.growth <- lag.listw(listw,lmoran$growth, NAOK = TRUE)
+lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
 
@@
 
-lmoran$ns <- lmoran$growth - mean(lmoran$growth, na.rm = TRUE) 
-lmoran$lag.growth <- lmoran$lag.growth - mean(lmoran$lag.growth, na.rm = TRUE)
+lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
+lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
 
@@
 lmoran <- lmoran%>% 
-  mutate(quadrant= ifelse(ns>0 & lag.growth > 0, 1, 0)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.growth < 0, 2, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.growth > 0, 3, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns>0 & lag.growth < 0, 4, quadrant)) %>%   
+  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
+  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
+  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
+  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
   #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
@@
 
-map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, palette=  c("white","red","blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha=0.4)), labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), title="LISA for growth, MSOA") +
+map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, palette=  c("white","red","blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha=0.4)), labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), title="LISA for level, MSOA") +
   tm_legend(text.size = 1) +

### 2022-04-13 src/level_lisa_oa.Rmd
update

---
title: "diffusion, websites with 1 pc"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(rgdal)
library(rgeos)
library(spdep)
library(tmap)
library(maptools)
library(knitr)
library(REAT)
library(tidygeocoder)
library(geosphere)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  select(pcds, oa11cd) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

```{r}
n = 1 #number of unique postcodes. 
m = 12

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610)

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012)

df <- bind_rows(df9610, df1112) %>% 
  filter(
    V1.domain < m,             # for multiple pc
    #V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, oa11cd) %>%
  summarise(n = n()) %>%            # for multiple pc  
  #summarise(n = sum(V1.domain)) %>% # for n == 1 
  ungroup()

# tests for Scotland
# df %>% filter(substr(oa11cd, 1,1) =="S",
#               year==2010) %>% summarise(mean(n, na.rm = T))
# 
# df %>% filter(oa11cd=="S00090182")
# 
# test@data %>% filter(substr(id, 1,1) =="S",
#                      n!=0)
# 
# df %>% filter(substr(oa11cd, 1,1) =="W",
#               year==2003)
```

## spatial data

```{r}
# get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))

# keep in the data slot only the ONS Output Area id, renaming it as 'id'
oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
colnames(oa.ew@data) <- c('id', 'area')

# reassign the polygon IDs
oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.ew)

# # get OA for Scotland
# path.geo.sc <- paste0(path, "/data/raw/SG_DataZoneBdry_2011")
# oa.sc <- readOGR(dsn=path.geo.sc, layer = "SG_DataZone_Bdry_2011")
# # source: https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011
# 
# # spatial transformations
# oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
# 
# # Scotland (follows same steps as EW, see notes above)
# oa.sc <- oa.sc[, 'DataZone']
# colnames(oa.sc@data) <- c('id')
# 
# # reassign the polygon IDs
# oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
# 
# # check the CRS has changed correctly, and the data slot has shrink to only the ID
# summary(oa.sc)

path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries

# spatial transformations
oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))

# Scotland (follows same steps as EW, see notes above)
oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
colnames(oa.sc@data) <- c('id', 'area')

# reassign the polygon IDs
oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.sc)

# build oa for GB
oa.gb <- spRbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)

# oa.gb$geometry <- oa.gb$geometry %>%
#   s2::s2_rebuild() %>%
#   sf::st_as_sfc()

# get GB
path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
gb <- readOGR(path.gb)
# spatial transformations
gb <- spTransform(gb, CRS("+init=epsg:4326"))

```

## lisa maps

```{r}

#path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_oa") # for unique PC
path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_oa")  # for multiple PC

morani <- data.frame()

#spatial weight matrix
sf_use_s2(FALSE)
nb <- poly2nb(oa.gb)
listw <- nb2listw(nb, style = "W", zero.policy = T)

for (t in 1996:2012){
test <- oa.gb
test@data <- test@data %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
         year = ifelse(is.na(year), t, year))

#spatial weight matrix
#moved out of the loop

#Moran's I
globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
morani <- rbind(a, morani)

#LISA
lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
#lmoran

# centers the local Moran's around the mean
lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)

# centers the variable of interest around its mean

lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)

signif <- 0.05
#lmoran

lmoran <- lmoran%>% 
  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%

#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))

mun_merge_new<- merge(test, lmoran, by="id")

# R p value map
breaks = c(0, 1, 2, 3, 4, 5)

cities <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  arrange(pop) %>% tail(10) 
tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)

map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, 
                                         palette = c("white","red","blue",
                                                     rgb(0,0,1,alpha=0.4),
                                                     rgb(1,0,0,alpha=0.4)), 
                                         labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
                                         title="LISA for level, OA") +
  tm_legend(text.size = 1) +
  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
  #tm_borders(alpha=.1) +
  tm_borders(lwd = 0) +
  tm_shape(gb, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
}
```

Moran I

```{r}
morani %>% as_tibble() %>% 
  rename('morani'=V1,
                   p = V2,
                   year = V3) %>% 
  kable()

```

## Gini

```{r}

gini.df <- data.frame()

for (t in 1996:2012){
    a <- df %>% filter(year==t)
  b <- as.data.frame(gini(a$n))
  b$year = t
  gini.df <- rbind(gini.df, b)
}
```

## Getis-Ord

```{r}

t= 1998
local_g <- localG(OA.Census$Qualification, nb_lw)
local_g <- localG(df[df$year==t,]$n, listw)

local_g <- cbind(OA.Census, as.matrix(local_g))
names(local_g)[6] <- "gstat"

```

## Distances to cities and retail centres

First set up the destinations. Origins are all the OA.

Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)

```{r}
# oa centroids
# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
#   mutate(id = row.names())
oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
  as.data.frame() %>% 
  add_rownames(var = "id")

# cities
city.names <- c("London, UK", 
"Birmingham, UK",
"Glasgow, UK",
"Liverpool, UK",
"Bristol, UK",
"Manchester, UK",
"Sheffield, UK",
"Leeds, UK",
"Edinburgh, UK",
"Leicester, UK")

cities <- geo(city.names, no_query = F, method = "arcgis")

# retail centres
geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
retail <-readOGR(geo.path) 
#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
  as.data.frame() %>% 
  add_rownames(var = "id") #%>%

retail.major.cetres <- retail.major.cetres.help %>% 
  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
  st_transform(4326) %>%
  st_coordinates() %>%
  as_tibble() %>% 
  bind_cols(retail.major.cetres.help$id) %>% 
  rename(id = '...3') %>% 
  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
```

Distance to cities

```{r}
dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist <- round((dist/1000),2) %>% 
  as_tibble()

city.names <- city.names %>% stringr::str_remove(", UK")
names(dist) <- city.names

dist <- dist %>% bind_cols(oa.gb.c$id) %>% 
  rename(oa11cd = last_col()) %>% 
  relocate(oa11cd)

# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
# dist$distMet <- apply(dist[,2:11], 1, min)

dist <- transform(dist, dist = do.call(pmin, dist[-1]))

# NOT USED AS IT SCREWS UP THE REGRESSIONS
# Complete panel, N = 3871903, 17 year * 227759 OA
# df <- df %>% bind_rows(dist %>% select(oa11cd)) %>% 
#   mutate(year = ifelse(is.na(year),1000, year),
#          country = substring(oa11cd, 1, 1)) %>% 
#   filter(country != "L" & country != "M" & country != "N") %>% 
#   filter(!is.na(country)) %>% 
#   complete(year, oa11cd, fill = list(n = 0)) %>% 
#   filter(year!=1000,
#          !is.na(oa11cd)) %>% 
#   select(-country)

# tests
# test$country <- substring(test$oa11cd, 1, 1)
# unique(test$country)
# length(unique(test$oa11cd))

# Join with distance and area
df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,
         density = n / area) %>% 
  relocate(density, .after = area)
sapply(df, function(x) sum(is.na(x)))
```

Distance to retail centres

```{r}
dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail <- round((dist.retail/1000),2) %>% 
  as_tibble()

# retail.names <- retail.major.cetres$RC_Name
# names(dist.retail) <- retail.names

dist.retail <- dist.retail %>% bind_cols(oa.gb.c$id) %>% 
  rename(oa11cd = last_col()) %>% 
  relocate(oa11cd)

# Minimum distance 
dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
dist.retail <- dist.retail %>% select(oa11cd, dist.retail)

# Join with complete panel
df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd"))

sapply(df, function(x) sum(is.na(x)))
```

Website density regressions to London

```{r}
dens.London <- data.frame()
out.model <- data.frame()

for (t in 1996:2012){
  model <- lm(log(density+0.001) ~ log(London), data = df[df$year==t,])
  r2 <- as.data.frame(summary(model)$r.squared)
  r2$year = t
  dens.London <- rbind(r2, dens.London)

  out.model_ <- tidy(model, conf.int = TRUE) %>% 
    slice(2) %>% 
    mutate_if(is.numeric, round, 2)
  out.model <- rbind(out.model_, out.model)
}

names(dens.London)[1] <- "rsquared"
ggplot(dens.London) +
  geom_line(aes(x=year, y = rsquared))

library(broom)
out.model <- out.model %>% 
  mutate(year = 1996:2012)

ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
#ggplot(out.model, aes(x=year, y = estimate)) +  
           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
                       width = 0.2,size  = 1,
                       position = "dodge", color="turquoise4") +
  #geom_hline(yintercept = 0, color = "red", size = 1) +
  geom_point() + coord_flip() 
```

Website density regressions to the nearest city

```{r}
dens.distMet <- data.frame()
out.modelMet <- data.frame()

for (t in 1996:2012){
  model <- lm(log(density+0.001) ~ log(dist), data = df[df$year==t,])
  r2 <- as.data.frame(summary(model)$r.squared)
  r2$year = t
  dens.distMet <- rbind(r2, dens.distMet)

  out.model_ <- tidy(model, conf.int = TRUE) %>% 
    slice(2) %>% 
    mutate_if(is.numeric, round, 2)
  out.modelMet <- rbind(out.model_, out.modelMet)
}

names(dens.distMet)[1] <- "rsquared"
ggplot(dens.distMet) +
  geom_line(aes(x=year, y = rsquared))

out.modelMet <- out.modelMet %>% 
  mutate(year = 1996:2012)

ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
                       width = 0.2,size  = 1,
                       position = "dodge", color="turquoise4") +
  #geom_hline(yintercept = 0, color = "red", size = 1) +
  geom_point() + coord_flip() 
```

Website density regressions to the nearest retail centre

```{r}
dens.dist.retail <- data.frame()
out.model.retail <- data.frame()

for (t in 1996:2012){
  model <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = df[df$year==t,])
  r2 <- as.data.frame(summary(model)$r.squared)
  r2$year = t
  dens.dist.retail <- rbind(r2, dens.dist.retail)

  out.model_ <- tidy(model, conf.int = TRUE) %>% 
    slice(2) %>% 
    mutate_if(is.numeric, round, 2)
  out.model.retail <- rbind(out.model_, out.model.retail)
}

names(dens.dist.retail)[1] <- "rsquared"
ggplot(dens.dist.retail) +
  geom_line(aes(x=year, y = rsquared))

out.model.retail <- out.model.retail %>% 
  mutate(year = 1996:2012)

ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
                       width = 0.2,size  = 1,
                       position = "dodge", color="turquoise4") +
  #geom_hline(yintercept = 0, color = "red", size = 1) +
  geom_point() + coord_flip()

# library(pscl)
# m1 <- zeroinfl((n) ~ log(dist.retail+0.001),
#   data = df[df$year==t,], dist = "negbin", EM = TRUE)
# summary(m1)
# 
# df %>% filter(year == 1996) %>% 
#   group_by(group = cut(n, breaks = c(0, 1, 2, 10, 100, 1000, 10000), include.lowest = T, right = F)) %>%  #seq(0, max(n), 100))) %>%
#   summarise(n = sum(n))
```

Website density regressions to the nearest retail centre with spatial lag

```{r}
library(spatialreg)

model <- lagsarlm(log(density+0.1) ~ log(dist.retail+0.1), data = df[df$year==t,], 
                  listw, tol.solve = 1e-30, zero.policy = T)

dens.dist.retail <- data.frame()
for (t in 1996:2012){
  model <- lm(log(density) ~ log(dist.retail+0.1), data = df[df$year==t,])
  r2 <- as.data.frame(summary(model)$r.squared)
  r2$year = t
  dens.dist.retail <- rbind(r2, dens.dist.retail)
}
names(dens.dist.retail)[1] <- "rsquared"
ggplot(dens.dist.retail) +
  geom_line(aes(x=year, y = rsquared))

```

## HH clusters

```{r}
hh <- mun_merge_new@data %>% filter(quadrant==1) %>% 
  select(id, quadrant) %>% 
  left_join(df[df$year==2000,], by =c("id" = "oa11cd")) %>% 
  select(id, year, n, area, density, London, dist, dist.retail)

hh.London <- lm(log(density+0.001) ~ log(London), data = hh)
summary(hh.London)
hh.dist <- lm(log(density+0.001) ~ log(dist), data = hh)
summary(hh.dist)
hh.retail <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = hh)
summary(hh.retail)

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610) %>% 
  filter(V1.domain == n,
         year == 1999) %>% 
  left_join(lookup, by = "pc", suffix =c("",""))

hh %>% left_join(df9610, by =c("id"="oa11cd"))

```

Spatial regression

```{r}
library(splm)
m <- spml(log(n)~lag(log(n)), data = df, listw = listw, lag = TRUE, spatial.error = "b", 
          model = "within", effect = "individual", method = "eigen", na.action = na.fail)

```

### 2022-04-13 src/point_distance.Rmd
update

--- 2022-04-13 src/level_lisa_oa.Rmd
@@
 ---
-title: "diffusion, websites with 1 pc"
+title: "point.distance"
 date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 library(rgdal)
-library(rgeos)
 library(spdep)
-library(tmap)
-library(maptools)
-library(knitr)
-library(REAT)
-library(tidygeocoder)
-library(geosphere)
+library(dbscan)
 
@@
 # 10332 missing msoa11cd
+
+# path.lookup <- "/hdd/internet_archive/archive/ONSPD_FEB_2017_UK.csv"
+# lookup <- read_csv(path.lookup) %>% 
+#   select(pcds, oseast1m, osnrth1m) %>% 
+#   dplyr::rename(pc = pcds)
+
 ```
 
-The internet archive $1996$-$2010$ data is saved on /hdd.
-The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.
+The internet archive data is saved on /hdd.
 
 ```{r}
-n = 1 #number of unique postcodes. 
-m = 12
+n = 1 #number of unique postcodes
 
-data.folder <- "/hdd/internet_archive/archive/data/"
-data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
+data.folder9610 <- "/hdd/internet_archive/archive/data/"
+data.path9610 <- paste0(data.folder9610, "domain_pc_year.csv")
 #Created by domain.R, which uses domain instead of host.
@@
 
-df9610 <- read_csv(data.path9610)
-
-data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
+data.path1112 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
 #Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
@@
 
-df1112 <- read_csv(data.path2011_2012)
+df9610 <- read_csv(data.path9610) 
+df1112 <- read_csv(data.path1112)
 
 df <- bind_rows(df9610, df1112) %>% 
-  filter(
-    V1.domain < m,             # for multiple pc
-    #V1.domain == n,             # for n == 1
-    year > 1995 & year <2013) %>%   
-  left_join(lookup, by = "pc", suffix =c("","")) %>% 
-  group_by(year, oa11cd) %>%
-  summarise(n = n()) %>%            # for multiple pc  
-  #summarise(n = sum(V1.domain)) %>% # for n == 1 
-  ungroup()
+  filter(V1.domain == n,
+         year > 1995) %>% 
+  left_join(lookup, by = "pc", suffix =c("",""))
 
-# tests for Scotland
-# df %>% filter(substr(oa11cd, 1,1) =="S",
-#               year==2010) %>% summarise(mean(n, na.rm = T))
-# 
-# df %>% filter(oa11cd=="S00090182")
-# 
-# test@data %>% filter(substr(id, 1,1) =="S",
-#                      n!=0)
-# 
-# df %>% filter(substr(oa11cd, 1,1) =="W",
-#               year==2003)
+rm(df9610, df1112)
 ```
 
-## spatial data
+## Boundary data
 
 ```{r}
+# get GB
+path.gb <- paste0(path, "/data/raw/Countries_(December_2020)_UK_BUC.geojson")
+gb <- readOGR(path.gb) 
+# spatial transformations
+gb <- spTransform(gb, CRS("+init=epsg:4326"))
+gb <- subset(gb, CTRY20NM != "Northern Ireland")
+gb.df <- fortify(gb, region = "CTRY20NM")
+
+# cities
+cities <- maps::world.cities %>% 
+  filter(country.etc=="UK") %>% 
+  arrange(pop) %>% tail(10) 
+tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
+
 # get OA for England and Wales
@@
 summary(oa.ew)
-
-# # get OA for Scotland
-# path.geo.sc <- paste0(path, "/data/raw/SG_DataZoneBdry_2011")
-# oa.sc <- readOGR(dsn=path.geo.sc, layer = "SG_DataZone_Bdry_2011")
-# # source: https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011
-# 
-# # spatial transformations
-# oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-# 
-# # Scotland (follows same steps as EW, see notes above)
-# oa.sc <- oa.sc[, 'DataZone']
-# colnames(oa.sc@data) <- c('id')
-# 
-# # reassign the polygon IDs
-# oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-# 
-# # check the CRS has changed correctly, and the data slot has shrink to only the ID
-# summary(oa.sc)
 
@@
 rm(oa.ew, oa.sc)
-
-# oa.gb$geometry <- oa.gb$geometry %>%
-#   s2::s2_rebuild() %>%
-#   sf::st_as_sfc()
-
-# get GB
-path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
-gb <- readOGR(path.gb)
-# spatial transformations
-gb <- spTransform(gb, CRS("+init=epsg:4326"))
-
-```
-
-## lisa maps
-
-```{r}
-
-#path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_oa") # for unique PC
-path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_oa")  # for multiple PC
-
-morani <- data.frame()
-
-#spatial weight matrix
-sf_use_s2(FALSE)
-nb <- poly2nb(oa.gb)
-listw <- nb2listw(nb, style = "W", zero.policy = T)
-
-for (t in 1996:2012){
-test <- oa.gb
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
-  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
-         year = ifelse(is.na(year), t, year))
-
-#spatial weight matrix
-#moved out of the loop
-
-#Moran's I
-globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
-a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
-morani <- rbind(a, morani)
-
-#LISA
-lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
-#lmoran
-
-# centers the local Moran's around the mean
-lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
-lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
-
-# centers the variable of interest around its mean
-
-lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
-lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
-
-signif <- 0.05
-#lmoran
-
-lmoran <- lmoran%>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
-  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
-  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%
-
-#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
-# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
-#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))
-
-mun_merge_new<- merge(test, lmoran, by="id")
-
-# R p value map
-breaks = c(0, 1, 2, 3, 4, 5)
-
-cities <- maps::world.cities %>% 
-  filter(country.etc=="UK") %>% 
-  arrange(pop) %>% tail(10) 
-tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
-
-map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, 
-                                         palette = c("white","red","blue",
-                                                     rgb(0,0,1,alpha=0.4),
-                                                     rgb(1,0,0,alpha=0.4)), 
-                                         labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
-                                         title="LISA for level, OA") +
-  tm_legend(text.size = 1) +
-  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
-  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
-  #tm_borders(alpha=.1) +
-  tm_borders(lwd = 0) +
-  tm_shape(gb, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
-  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
-  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
-tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
-}
-```
-
-Moran I
-
-```{r}
-morani %>% as_tibble() %>% 
-  rename('morani'=V1,
-                   p = V2,
-                   year = V3) %>% 
-  kable()
-
-```
-
-## Gini
-
-```{r}
-
-gini.df <- data.frame()
-
-for (t in 1996:2012){
-    a <- df %>% filter(year==t)
-  b <- as.data.frame(gini(a$n))
-  b$year = t
-  gini.df <- rbind(gini.df, b)
-}
-```
-
-## Getis-Ord
-
-```{r}
-
-t= 1998
-local_g <- localG(OA.Census$Qualification, nb_lw)
-local_g <- localG(df[df$year==t,]$n, listw)
-
-local_g <- cbind(OA.Census, as.matrix(local_g))
-names(local_g)[6] <- "gstat"
-
 ```
@@
 
-# NOT USED AS IT SCREWS UP THE REGRESSIONS
-# Complete panel, N = 3871903, 17 year * 227759 OA
-# df <- df %>% bind_rows(dist %>% select(oa11cd)) %>% 
-#   mutate(year = ifelse(is.na(year),1000, year),
-#          country = substring(oa11cd, 1, 1)) %>% 
-#   filter(country != "L" & country != "M" & country != "N") %>% 
-#   filter(!is.na(country)) %>% 
-#   complete(year, oa11cd, fill = list(n = 0)) %>% 
-#   filter(year!=1000,
-#          !is.na(oa11cd)) %>% 
-#   select(-country)
-
 # tests
@@
 # Join with distance and area
-df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
-  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
-  relocate(area, .after = n) %>% 
-  mutate(area = area / 1000000,
-         density = n / area) %>% 
-  relocate(density, .after = area)
-sapply(df, function(x) sum(is.na(x)))
 ```
@@
 dist.retail <- dist.retail %>% select(oa11cd, dist.retail)
+```
+
+Joins
+
+**IMPORTANT** these are distants not to points but to the centroids of the OA they fall within.
+
+```{r}
+
+df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
+  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
+  relocate(area, .after = n) %>% 
+  mutate(area = area / 1000000,
+         density = n / area) %>% 
+  relocate(density, .after = area)
+sapply(df, function(x) sum(is.na(x)))
 
@@
 sapply(df, function(x) sum(is.na(x)))
-```
-
-Website density regressions to London
-
-```{r}
-dens.London <- data.frame()
-out.model <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(London), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.London <- rbind(r2, dens.London)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model <- rbind(out.model_, out.model)
-}
-
-names(dens.London)[1] <- "rsquared"
-ggplot(dens.London) +
-  geom_line(aes(x=year, y = rsquared))
-
-library(broom)
-out.model <- out.model %>% 
-  mutate(year = 1996:2012)
-
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-#ggplot(out.model, aes(x=year, y = estimate)) +  
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() 
-```
-
-Website density regressions to the nearest city
-
-```{r}
-dens.distMet <- data.frame()
-out.modelMet <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.distMet <- rbind(r2, dens.distMet)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.modelMet <- rbind(out.model_, out.modelMet)
-}
-
-names(dens.distMet)[1] <- "rsquared"
-ggplot(dens.distMet) +
-  geom_line(aes(x=year, y = rsquared))
-
-out.modelMet <- out.modelMet %>% 
-  mutate(year = 1996:2012)
-
-ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() 
-```
-
-Website density regressions to the nearest retail centre
-
-```{r}
-dens.dist.retail <- data.frame()
-out.model.retail <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model.retail <- rbind(out.model_, out.model.retail)
-}
-
-names(dens.dist.retail)[1] <- "rsquared"
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared))
-
-out.model.retail <- out.model.retail %>% 
-  mutate(year = 1996:2012)
-
-ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip()
-
-# library(pscl)
-# m1 <- zeroinfl((n) ~ log(dist.retail+0.001),
-#   data = df[df$year==t,], dist = "negbin", EM = TRUE)
-# summary(m1)
-# 
-# df %>% filter(year == 1996) %>% 
-#   group_by(group = cut(n, breaks = c(0, 1, 2, 10, 100, 1000, 10000), include.lowest = T, right = F)) %>%  #seq(0, max(n), 100))) %>%
-#   summarise(n = sum(n))
-```
-
-Website density regressions to the nearest retail centre with spatial lag
-
-```{r}
-library(spatialreg)
-
-model <- lagsarlm(log(density+0.1) ~ log(dist.retail+0.1), data = df[df$year==t,], 
-                  listw, tol.solve = 1e-30, zero.policy = T)
-
-dens.dist.retail <- data.frame()
-for (t in 1996:2012){
-  model <- lm(log(density) ~ log(dist.retail+0.1), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-}
-names(dens.dist.retail)[1] <- "rsquared"
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared))
 
@@
 
-## HH clusters
+## distance to London plot
 
 ```{r}
-hh <- mun_merge_new@data %>% filter(quadrant==1) %>% 
-  select(id, quadrant) %>% 
-  left_join(df[df$year==2000,], by =c("id" = "oa11cd")) %>% 
-  select(id, year, n, area, density, London, dist, dist.retail)
+df %>%
+  #mutate(text = fct_reorder(text, value)) %>%
+  ggplot( aes(x=London, color=year, fill=year)) +
+    geom_histogram(alpha=0.6, binwidth = 5) +
+    #scale_fill_viridis(discrete=F) +
+    #scale_color_viridis(discrete=F) +
+    #theme_ipsum() +
 
-hh.London <- lm(log(density+0.001) ~ log(London), data = hh)
-summary(hh.London)
-hh.dist <- lm(log(density+0.001) ~ log(dist), data = hh)
-summary(hh.dist)
-hh.retail <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = hh)
-summary(hh.retail)
-
-data.folder <- "/hdd/internet_archive/archive/data/"
-data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
-#Created by domain.R, which uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df9610 <- read_csv(data.path9610) %>% 
-  filter(V1.domain == n,
-         year == 1999) %>% 
-  left_join(lookup, by = "pc", suffix =c("",""))
-
-hh %>% left_join(df9610, by =c("id"="oa11cd"))
-
+  theme(
+      legend.position="none",
+      panel.spacing = unit(0.1, "lines"),
+      strip.text.x = element_text(size = 8)
+    ) +
+    xlab("") +
+    ylab("domains") +
+    facet_wrap(~year, scales = "free_y")
 ```
 
-Spatial regression
+## distance to cities plot
 
 ```{r}
-library(splm)
-m <- spml(log(n)~lag(log(n)), data = df, listw = listw, lag = TRUE, spatial.error = "b", 
-          model = "within", effect = "individual", method = "eigen", na.action = na.fail)
+df %>%
+  #mutate(text = fct_reorder(text, value)) %>%
+  ggplot( aes(x=dist, color=year, fill=year)) +
+  geom_histogram(alpha=0.6, binwidth = 5) +
+    #scale_fill_viridis(discrete=F) +
+    #scale_color_viridis(discrete=F) +
+    #theme_ipsum() +
 
+  # geom_histogram(aes(y = stat(count) / sum(count)), alpha=0.6, binwidth = 5) +
+  # scale_y_continuous(labels = scales::percent)+
+
+  theme(
+      legend.position="none",
+      panel.spacing = unit(0.1, "lines"),
+      strip.text.x = element_text(size = 8)
+    ) +
+    xlab("") +
+    ylab("domains") +
+    facet_wrap(~year, scales = "free_y")
 ```
+
+## distance to retail centres
+
+```{r}
+df %>%
+  #mutate(text = fct_reorder(text, value)) %>%
+  ggplot( aes(x=dist.retail, color=year, fill=year)) +
+    geom_histogram(alpha=0.6, binwidth = 5) +
+    #scale_fill_viridis(discrete=F) +
+    #scale_color_viridis(discrete=F) +
+    #theme_ipsum() +
+
+  theme(
+      legend.position="none",
+      panel.spacing = unit(0.1, "lines"),
+      strip.text.x = element_text(size = 8)
+    ) +
+    xlab("") +
+    ylab("domains") +
+    facet_wrap(~year, scales = "free_y")
+```

### 2022-04-13 src/s_la.Rmd
update

---
title: "s.function"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
library(rprojroot)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

## Load data

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

```{r}
n = 1 #number of unique postcodes. 
m = 12

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610)

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012)

df <- bind_rows(df9610, df1112) %>% 
  filter(
    V1.domain < m,             # for multiple pc
    #V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, ladcd) %>%
  summarise(n = n()) %>%            # for multiple pc  
  #summarise(n = sum(V1.domain)) %>% # for n == 1 
  ungroup()

# tests for Scotland
# df %>% filter(substr(oa11cd, 1,1) =="S",
#               year==2010) %>% summarise(mean(n, na.rm = T))
# 
# df %>% filter(oa11cd=="S00090182")
# 
# test@data %>% filter(substr(id, 1,1) =="S",
#                      n!=0)
# 
# df %>% filter(substr(oa11cd, 1,1) =="W",
#               year==2003)
```

## Help, not really used

```{r, eval=FALSE}

df.test <- df %>% filter(oa11cd == "E00000007" |
                         oa11cd == "E00000021" |
                         oa11cd == "E00000025") %>% 
  mutate(rownum = row_number(),
         value2 = n/max(n))

# function to select unique groups
sample_n_groups = function(grouped_df, size, replace = FALSE, weight=NULL) {
  grp_var <- grouped_df %>% 
    groups %>%
    unlist %>% 
    as.character
  random_grp <- grouped_df %>% 
    summarise() %>% 
    sample_n(size, replace, weight) %>% 
    mutate(unique_id = 1:NROW(.))
  grouped_df %>% 
    right_join(random_grp, by=grp_var) %>% 
    group_by_(grp_var) 
}

df.test <- df %>% group_by(oa11cd) %>% 
  sample_n_groups(10000) %>% 
  arrange(oa11cd, year) %>% 
  select(-unique_id)

fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.test)

fd.nls <- df.test %>% 
  group_by(oa11cd) %>%
  select(year, oa11cd, n) %>% 
  do(fitloop = tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .))) %>% 
  unnest(fitloop)

fd.nls %>% filter(term=="xmid") %>% 
  summarise(rng = range(estimate))
```

## Test for S function estimation

To interpret the coefficients, check [SSlogis](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/SSlogis).
This is a self-starting model, which estimates the starting values.

```{r}
df.test <- df %>% filter(ladcd == "S12000033")

fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.test)
summary(fit)

df.test$prednls = predict(fit)

ggplot(df.test, aes(x = year, y = n) ) +
  geom_point() +
  geom_line(aes(y = prednls), size = 1) +
  scale_x_continuous("year", labels = as.character(df.test$year), breaks = df.test$year) +
  ylab("N. of websites")
  #scale_x_date(breaks = df.test$year) 
```

## S function estimation for all the country

```{r}
df.country <- df %>% group_by(year) %>% 
  summarise(n = sum(n))

fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.country)
summary(fit)

df.country$prednls = predict(fit)

ggplot(df.country, aes(x = year, y = n) ) +
  geom_point() +
  geom_line(aes(y = prednls), size = 1) +
  scale_x_continuous("year", labels = as.character(df.country$year), breaks = df.country$year) +
  ylab("N. of websites") +
  ggtitle("Country")
  #scale_x_date(breaks = df.test$year) 
```

## Loop for S function estimation for Local Authorities

```{r}
a.la <- as_tibble()
all.la <- df %>% filter(ladcd != "NA") %>% distinct(ladcd) %>% simplify2array()
simplify2array(all.la)

for (i in all.la){
  data <- df %>% filter(ladcd==i) %>%
    select(year, ladcd, n) 
    #lm(n~year, data = .)
    #tryCatch(tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .)), , error=function(e) NULL)
  tryCatch(b.la <- tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = data)), error=function(e) NULL)
  tryCatch(a.la <-rbind(a.la,b.la), error=function(e) NULL)
}

a.la %>% filter(term=="xmid") %>% 
  mutate(estimate = round(estimate, 0)) %>% 
  select(estimate) %>% 
  summarise(r = range(estimate),
            median = median(estimate))

a.la %>% filter(term=="xmid") %>% 
  mutate(estimate = round(estimate, 0)) %>% 
  filter(p.value < 0.01) %>%    # (227759 - 212782)/227759  # 6% not sig 
  select(estimate) %>% 
  ggplot(aes(x=estimate)) + 
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = 2003, colour = "red")
  scale_x_continuous("year", labels = round(a$estimate, 0), breaks = round(a$estimate, 0)) 
  #scale_x_discrete(labels = as.factor(a$estimate))

a.la <- a.la %>% filter(term=="xmid") %>% 
  bind_cols(as.tibble(all.la)) %>% 
  mutate(estimate = round(estimate, 0)) %>% 
  filter(p.value < 0.01) %>%    # 376 / 376, all sig
  select(ladcd, estimate)
  #filter(estimate>2011)
```

write the object with all the models

```{r eval=F}
path.out <- paste0(path, "/data/temp/s_la.csv")
write_csv(a.la, path.out)
```

## Map the xmid

Spatial data

```{r}
# get LA for the UK
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
la <- spTransform(la, CRS("+init=epsg:4326"))

la.f <- fortify(la, region = "LAD21CD")

# cities
cities <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  arrange(pop) %>% tail(10) 
tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
```

```{r}

la.f <- la.f %>% left_join(a.la, by = c("id" = "ladcd")) %>% 
  mutate(fast = ifelse(estimate < 2003, "fast", "slow")) %>% 
  arrange(order)  # if i don't order merge.nuts.f loses order and the map has gaps

# oa.gb.f %>% mutate(fast = ifelse(estimate < 2003, 1, 0)) %>% 
#   arrange(order) %>% 
ggplot(data = la.f, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = estimate)) + #fill = fast, 
  #theme_nothing(legend = TRUE) +
  labs(title = "Clusters") +
  scale_fill_continuous(type = "viridis") +
  theme_void()

ggplot(data = la.f, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = fast)) + #fill = fast, 
  #theme_nothing(legend = TRUE) +
  labs(title = "Clusters") +
  geom_point(data = cities, aes(x=long, y=lat), colour = "black", size = .1) + 
    geom_text(size = 2, check_overlap = F, nudge_y = .2, data = cities, aes(x=long, y=lat, label = name)) +
  #scale_fill_continuous(type = "viridis") +
  theme_void()

```

```{r}
la.f %>% left_join(df %>% filter(year==2012), by = c("id" = "ladcd")) %>% 
  arrange(order) %>% # if i don't order merge.nuts.f loses order and the map has gaps
  ggplot(aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = n)) + #fill = fast, 
  #theme_nothing(legend = TRUE) +
  labs(title = "Clusters") +
  scale_fill_continuous(type = "viridis") +
  theme_void()

```

### 2022-04-13 src/s_oa.Rmd
update

--- 2022-04-13 src/level_lisa_oa.Rmd
@@
 ---
-title: "diffusion, websites with 1 pc"
+title: "s.function"
 date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 library(tidyverse)
+library(broom)
 library(rprojroot)
-library(rgdal)
-library(rgeos)
-library(spdep)
-library(tmap)
-library(maptools)
-library(knitr)
-library(REAT)
-library(tidygeocoder)
-library(geosphere)
 
@@
 ```
+
+## Load data
 
@@
   filter(
-    V1.domain < m,             # for multiple pc
-    #V1.domain == n,             # for n == 1
+    #V1.domain < m,             # for multiple pc
+    V1.domain == n,             # for n == 1
     year > 1995 & year <2013) %>%   
@@
   group_by(year, oa11cd) %>%
-  summarise(n = n()) %>%            # for multiple pc  
-  #summarise(n = sum(V1.domain)) %>% # for n == 1 
+  #summarise(n = n()) %>%            # for multiple pc  
+  summarise(n = sum(V1.domain)) %>% # for n == 1 
   ungroup()
@@
 
-## spatial data
+## Help, not really used
+
+```{r, eval=FALSE}
+
+df.test <- df %>% filter(oa11cd == "E00000007" |
+                         oa11cd == "E00000021" |
+                         oa11cd == "E00000025") %>% 
+  mutate(rownum = row_number(),
+         value2 = n/max(n))
+
+# function to select unique groups
+sample_n_groups = function(grouped_df, size, replace = FALSE, weight=NULL) {
+  grp_var <- grouped_df %>% 
+    groups %>%
+    unlist %>% 
+    as.character
+  random_grp <- grouped_df %>% 
+    summarise() %>% 
+    sample_n(size, replace, weight) %>% 
+    mutate(unique_id = 1:NROW(.))
+  grouped_df %>% 
+    right_join(random_grp, by=grp_var) %>% 
+    group_by_(grp_var) 
+}
+
+df.test <- df %>% group_by(oa11cd) %>% 
+  sample_n_groups(10000) %>% 
+  arrange(oa11cd, year) %>% 
+  select(-unique_id)
+
+fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.test)
+
+fd.nls <- df.test %>% 
+  group_by(oa11cd) %>%
+  select(year, oa11cd, n) %>% 
+  do(fitloop = tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .))) %>% 
+  unnest(fitloop)
+
+fd.nls %>% filter(term=="xmid") %>% 
+  summarise(rng = range(estimate))
+```
+
+## Test for S function estimation
+
+To interpret the coefficients, check [SSlogis](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/SSlogis).
+This is a self-starting model, which estimates the starting values.
+
+```{r}
+df.test <- df %>% filter(oa11cd == "S00103828")
+
+fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.test)
+summary(fit)
+
+df.test$prednls = predict(fit)
+
+ggplot(df.test, aes(x = year, y = n) ) +
+  geom_point() +
+  geom_line(aes(y = prednls), size = 1) +
+  scale_x_continuous("year", labels = as.character(df.test$year), breaks = df.test$year) +
+  ylab("N. of websites")
+  #scale_x_date(breaks = df.test$year) 
+```
+
+## S function estimation for all the country
+
+```{r}
+df.country <- df %>% group_by(year) %>% 
+  summarise(n = sum(n))
+
+fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.country)
+summary(fit)
+
+df.country$prednls = predict(fit)
+
+ggplot(df.country, aes(x = year, y = n) ) +
+  geom_point() +
+  geom_line(aes(y = prednls), size = 1) +
+  scale_x_continuous("year", labels = as.character(df.country$year), breaks = df.country$year) +
+  ylab("N. of websites") +
+  ggtitle("Country")
+  #scale_x_date(breaks = df.test$year) 
+```
+
+## Loop for S function estimation for OA
+
+```{r}
+a <- as_tibble()
+all.oa <- unique(df$oa11cd)
+for (oa in all.oa){
+  data <- df %>% filter(oa11cd==oa) %>%
+    select(year, oa11cd, n) 
+    #lm(n~year, data = .)
+    #tryCatch(tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .)), , error=function(e) NULL)
+  tryCatch(b <- tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = data)), error=function(e) NULL)
+  tryCatch(a <-rbind(a,b), error=function(e) NULL)
+}
+
+a %>% filter(term=="xmid") %>% 
+  mutate(estimate = round(estimate, 0)) %>% 
+  select(estimate) %>% 
+  summarise(r = range(estimate))
+
+a %>% filter(term=="xmid") %>% 
+  mutate(estimate = round(estimate, 0)) %>% 
+  filter(p.value < 0.01) %>%    # (227759 - 212782)/227759  # 6% not sig 
+  select(estimate) %>% 
+  ggplot(aes(x=estimate)) + 
+  geom_histogram(binwidth = 1) +
+  geom_vline(xintercept = 2003, colour = "red")
+  scale_x_continuous("year", labels = round(a$estimate, 0), breaks = round(a$estimate, 0)) 
+  #scale_x_discrete(labels = as.factor(a$estimate))
+
+a <- a %>% filter(term=="xmid") %>% 
+    bind_cols(all.oa) %>% 
+    mutate(estimate = round(estimate, 0)) %>% 
+    filter(p.value < 0.01) %>%    # (227759 - 212782)/227759  # 6% not sig 
+    rename(id = ...6) %>% 
+    select(id, estimate)
+```
+
+write the object with all the models
+
+```{r}
+path.out <- paste0(path, "/data/temp/s.csv")
+write_csv(a, path.out)
+```
+
+## Map the xmid
+
+Spatial data
 
@@
 
+gb.df <- fortify(gb)
 ```
-
-## lisa maps
 
@@
 
-#path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_oa") # for unique PC
-path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_oa")  # for multiple PC
+# spatial transformations
+oa.gb <- spTransform(oa.gb, CRS("+init=epsg:4326"))
 
-morani <- data.frame()
+# ggplot2 maps
+oa.gb.f <- fortify(oa.gb, region = "id")
 
-#spatial weight matrix
-sf_use_s2(FALSE)
-nb <- poly2nb(oa.gb)
-listw <- nb2listw(nb, style = "W", zero.policy = T)
+oa.gb.f <- oa.gb.f %>% left_join(a, by = "id") %>% 
+  mutate(fast = ifelse(estimate < 2003, "fast", "slow")) %>% 
+  arrange(order)  # if i don't order merge.nuts.f loses order and the map has gaps
 
-for (t in 1996:2012){
-test <- oa.gb
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
-  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
-         year = ifelse(is.na(year), t, year))
-
-#spatial weight matrix
-#moved out of the loop
-
-#Moran's I
-globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
-a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
-morani <- rbind(a, morani)
-
-#LISA
-lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
-#lmoran
-
-# centers the local Moran's around the mean
-lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
-lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
-
-# centers the variable of interest around its mean
-
-lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
-lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
-
-signif <- 0.05
-#lmoran
-
-lmoran <- lmoran%>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
-  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
-  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%
-
-#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
-# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
-#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))
-
-mun_merge_new<- merge(test, lmoran, by="id")
-
-# R p value map
-breaks = c(0, 1, 2, 3, 4, 5)
-
-cities <- maps::world.cities %>% 
-  filter(country.etc=="UK") %>% 
-  arrange(pop) %>% tail(10) 
-tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
-
-map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, 
-                                         palette = c("white","red","blue",
-                                                     rgb(0,0,1,alpha=0.4),
-                                                     rgb(1,0,0,alpha=0.4)), 
-                                         labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
-                                         title="LISA for level, OA") +
-  tm_legend(text.size = 1) +
-  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
-  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
-  #tm_borders(alpha=.1) +
-  tm_borders(lwd = 0) +
-  tm_shape(gb, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
-  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
-  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
-tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
-}
+# oa.gb.f %>% mutate(fast = ifelse(estimate < 2003, 1, 0)) %>% 
+#   arrange(order) %>% 
+ggplot(data = oa.gb.f %>% arrange(order), aes(x = long, y = lat)) +
+  geom_polygon(aes(group = group, fill = estimate)) + #fill = fast, 
+  #theme_nothing(legend = TRUE) +
+  labs(title = "Clusters") +
+  scale_fill_continuous(type = "viridis") +
+  theme_void()
 ```
 
-Moran I
+## OA classification
+
+[Source](https://data.cdrc.ac.uk/dataset/output-area-classification-2011)
 
 ```{r}
-morani %>% as_tibble() %>% 
-  rename('morani'=V1,
-                   p = V2,
-                   year = V3) %>% 
+oa.class.path <- paste0(path, "/data/raw/Output Area Classification")
+oa.class <- readOGR(dsn = oa.class.path)
+oa.class@data %>% glimpse()
+
+pen.path <- paste0(path, "/data/raw/Output Area Classification/penportraits.csv")
+pen <- read.csv(pen.path, sep = "\t")
+
+oa.class.df <- a %>%  
+  mutate(fast = ifelse(estimate < 2003, "fast", "slow")) %>% 
+  left_join(oa.class@data, by = c("id"="OA_SA")) %>% 
+  select(1:10) %>% 
+  arrange(id)
+
+length(unique(oa.class.df$GRP))
+
+oa.class.df %>% 
+  filter(fast == "slow") %>% 
+  group_by(SUBGRP) %>% 
+  summarise(n=n()) %>% 
+  arrange(-n) %>% 
+  mutate(freq = n / sum(n),
+         freq = round(freq, 2),
+         cum.n = cumsum(n),
+         cum.freq = cumsum(n / sum(n)),
+         cum.freq = round(cum.freq, 2)) %>% 
+  left_join(pen %>% select(code, name),
+            by = c("SUBGRP" = "code")) %>% 
   kable()
 
+# no obvious group / subgroup patterns between slow and fast
 ```
 
-## Gini
+## Regress xmid
 
-```{r}
-
-gini.df <- data.frame()
-
-for (t in 1996:2012){
-    a <- df %>% filter(year==t)
-  b <- as.data.frame(gini(a$n))
-  b$year = t
-  gini.df <- rbind(gini.df, b)
-}
-```
-
-## Getis-Ord
-
-```{r}
-
-t= 1998
-local_g <- localG(OA.Census$Qualification, nb_lw)
-local_g <- localG(df[df$year==t,]$n, listw)
-
-local_g <- cbind(OA.Census, as.matrix(local_g))
-names(local_g)[6] <- "gstat"
-
-```
-
-## Distances to cities and retail centres
+### Distances to cities and retail centres
 
@@
 # Join with distance and area
-df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
-  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
-  relocate(area, .after = n) %>% 
-  mutate(area = area / 1000000,
-         density = n / area) %>% 
-  relocate(density, .after = area)
+oa.class.df <- oa.class.df %>% left_join(dist, by = c("id" = "oa11cd")) 
 sapply(df, function(x) sum(is.na(x)))
@@
 # Join with complete panel
-df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd"))
+oa.class.df <- oa.class.df %>% left_join(dist.retail, by = c("id" = "oa11cd"))
 
-sapply(df, function(x) sum(is.na(x)))
+sapply(oa.class.df, function(x) sum(is.na(x)))
 ```
 
-Website density regressions to London
+Distance regressions
 
 ```{r}
-dens.London <- data.frame()
-out.model <- data.frame()
 
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(London), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.London <- rbind(r2, dens.London)
+oa.class.df <- oa.class.df %>% left_join(df %>% group_by(oa11cd) %>% summarise(max.n = max(n, na.rm = T)), #filter(year==2012), 
+                          by = c("id"="oa11cd"))
 
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model <- rbind(out.model_, out.model)
-}
+model.n <- lm(estimate ~ max.n, data = oa.class.df) #%>%filter(estimate>1996 & estimate <2012)
+summary(model.n)
+nobs(model.n)
 
-names(dens.London)[1] <- "rsquared"
-ggplot(dens.London) +
-  geom_line(aes(x=year, y = rsquared))
+model <- lm(estimate ~ log(London + 1) + log(dist + 1) + log(dist.retail + 1) + 
+              as.factor(SPRGRP) + log(POPULATION) + max.n, data = oa.class.df) #%>%filter(estimate>1996 & estimate <2012)
+summary(model)
+nobs(model)
 
-library(broom)
-out.model <- out.model %>% 
-  mutate(year = 1996:2012)
+oa.class.df <- oa.class.df %>% mutate(fast.f = recode(fast, "fast" = 1, 
+                                          "slow" = 0)) %>% 
+  relocate(fast.f, .after = fast)
 
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-#ggplot(out.model, aes(x=year, y = estimate)) +  
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() 
-```
+model.glm <- glm(fast.f ~ log(London + 1) + log(dist + 1) + log(dist.retail + 1) + 
+              as.factor(SPRGRP) + log(POPULATION) + max.n, family = binomial(link='logit'), data = oa.class.df) #%>%filter(estimate>1996 & estimate <2012)
+summary(model.glm)
+nobs(model.glm)
 
-Website density regressions to the nearest city
-
-```{r}
-dens.distMet <- data.frame()
-out.modelMet <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.distMet <- rbind(r2, dens.distMet)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.modelMet <- rbind(out.model_, out.modelMet)
-}
-
-names(dens.distMet)[1] <- "rsquared"
-ggplot(dens.distMet) +
-  geom_line(aes(x=year, y = rsquared))
-
-out.modelMet <- out.modelMet %>% 
-  mutate(year = 1996:2012)
-
-ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() 
-```
-
-Website density regressions to the nearest retail centre
-
-```{r}
-dens.dist.retail <- data.frame()
-out.model.retail <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model.retail <- rbind(out.model_, out.model.retail)
-}
-
-names(dens.dist.retail)[1] <- "rsquared"
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared))
-
-out.model.retail <- out.model.retail %>% 
-  mutate(year = 1996:2012)
-
-ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip()
-
-# library(pscl)
-# m1 <- zeroinfl((n) ~ log(dist.retail+0.001),
-#   data = df[df$year==t,], dist = "negbin", EM = TRUE)
-# summary(m1)
-# 
-# df %>% filter(year == 1996) %>% 
-#   group_by(group = cut(n, breaks = c(0, 1, 2, 10, 100, 1000, 10000), include.lowest = T, right = F)) %>%  #seq(0, max(n), 100))) %>%
-#   summarise(n = sum(n))
-```
-
-Website density regressions to the nearest retail centre with spatial lag
-
-```{r}
-library(spatialreg)
-
-model <- lagsarlm(log(density+0.1) ~ log(dist.retail+0.1), data = df[df$year==t,], 
-                  listw, tol.solve = 1e-30, zero.policy = T)
-
-dens.dist.retail <- data.frame()
-for (t in 1996:2012){
-  model <- lm(log(density) ~ log(dist.retail+0.1), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-}
-names(dens.dist.retail)[1] <- "rsquared"
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared))
+library(pscl)
+pR2(model.glm)
 
 ```
-
-## HH clusters
-
-```{r}
-hh <- mun_merge_new@data %>% filter(quadrant==1) %>% 
-  select(id, quadrant) %>% 
-  left_join(df[df$year==2000,], by =c("id" = "oa11cd")) %>% 
-  select(id, year, n, area, density, London, dist, dist.retail)
-
-hh.London <- lm(log(density+0.001) ~ log(London), data = hh)
-summary(hh.London)
-hh.dist <- lm(log(density+0.001) ~ log(dist), data = hh)
-summary(hh.dist)
-hh.retail <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = hh)
-summary(hh.retail)
-
-data.folder <- "/hdd/internet_archive/archive/data/"
-data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
-#Created by domain.R, which uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df9610 <- read_csv(data.path9610) %>% 
-  filter(V1.domain == n,
-         year == 1999) %>% 
-  left_join(lookup, by = "pc", suffix =c("",""))
-
-hh %>% left_join(df9610, by =c("id"="oa11cd"))
-
-```
-
-Spatial regression
-
-```{r}
-library(splm)
-m <- spml(log(n)~lag(log(n)), data = df, listw = listw, lag = TRUE, spatial.error = "b", 
-          model = "within", effect = "individual", method = "eigen", na.action = na.fail)
-
-```

### 2022-04-13 src/sppt.Rmd
update

---
title: "diffusion, websites with 1 pc"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(rgdal)
library(spdep)
library(sppt)
library(spatstat)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  select(pcds, msoa11cd, dointr, doterm) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd

path.lookup <- "/hdd/internet_archive/archive/ONSPD_FEB_2017_UK.csv"
lookup <- read_csv(path.lookup) %>% 
  select(pcds, oseast1m, osnrth1m) %>% 
  dplyr::rename(pc = pcds)

```

The internet archive data is saved on /hdd.

```{r}
n = 1 #number of unique postcodes

data.folder9610 <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder9610, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

data.path1112 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610) 
df1112 <- read_csv(data.path1112)

df <- bind_rows(df9610, df1112) %>% 
  filter(V1.domain == n,
         year > 1995) %>% 
  left_join(lookup, by = "pc", suffix =c("",""))

rm(df9610, df1112)
```

## Boundary data

```{r}
# get GB
path.gb <- paste0(path, "/data/raw/Countries_(December_2020)_UK_BUC.geojson")
gb <- readOGR(path.gb) 
# spatial transformations
gb <- spTransform(gb, CRS("+init=epsg:4326"))
gb <- subset(gb, CTRY20NM != "Northern Ireland")
gb.df <- fortify(gb, region = "CTRY20NM")

# get MSOA 
# source: https://geoportal.statistics.gov.uk/datasets/middle-layer-super-output-areas-december-2011-boundaries-generalised-clipped-bgc-ew-v3/explore?location=52.849644%2C-2.489527%2C7.08

path.geo <- paste0(path, "/data/raw/Middle_Layer_Super_Output_Areas_(December_2011)_Boundaries_Generalised_Clipped_(BGC)_EW_V3.geojson")
msoa <- readOGR(path.geo)
# UK BFC: la <- readOGR("https://opendata.arcgis.com/datasets/1d78d47c87df4212b79fe2323aae8e08_0.geojson") 
# source: https://data.gov.uk/dataset/7c387c64-d25f-474a-b07e-b933578caea2/local-authority-districts-april-2019-boundaries-uk-bfe

# cities
cities <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  arrange(pop) %>% tail(10) 
tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)

# get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))

# keep in the data slot only the ONS Output Area id, renaming it as 'id'
oa.ew <- oa.ew[, 'OA11CD']
colnames(oa.ew@data) <- c('id')

# reassign the polygon IDs
oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.ew)

# # get OA for Scotland
path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries

# spatial transformations
oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))

# Scotland (follows same steps as EW, see notes above)
oa.sc <- oa.sc[, 'code']
colnames(oa.sc@data) <- c('id')

# reassign the polygon IDs
oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.sc)

# build oa for GB
oa.gb <- spRbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)

```

## sppt

```{r}

gb <- spTransform(gb, CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
msoa <- spTransform(msoa, CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
oa.gb <- spTransform(oa.gb, CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

results.1 <- data.frame()
results.2 <- data.frame()

for (t in 1996:2012){
  a <- df %>%
    #slice_sample(n = 10000) %>% 
    filter(!is.na(oseast1m), #6k rows with missing east/north for all years
           year == t,
           substr(pc, 1,2) !="BT") %>% #drop Northern Ireland, for 2010 c. 4k obs.
    st_as_sf(coords = c("oseast1m", "osnrth1m"), crs = 27700) %>%
    st_transform(4326) %>%
    st_coordinates() %>%
    as_tibble() %>%
    st_as_sf(coords = c("X", "Y"),
                        crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0") %>% 
  as("Spatial")

  b <- df %>%
    #slice_sample(n = 10000) %>% 
    filter(!is.na(oseast1m), #6k rows with missing east/north for all years
           year == t+1,
           substr(pc, 1,2) !="BT") %>% #drop Northern Ireland, for 2010 c. 4k obs.
    st_as_sf(coords = c("oseast1m", "osnrth1m"), crs = 27700) %>%
    st_transform(4326) %>%
    st_coordinates() %>%
    as_tibble() %>%
    st_as_sf(coords = c("X", "Y"),
                        crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0") %>% 
    as("Spatial")

  a <- spTransform(a, CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
  b <- spTransform(b, CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

  a@data <- a@data %>% mutate(id = 1:nrow(a@data))
  b@data <- b@data %>% mutate(id = 1:nrow(b@data))

  set.seed(93255) # set seed for reproducibility
  #output <- sppt(sp1998, sp1999, msoa)
  #output@data

  #output2 <- sppt_boot(sp1998, sp1999, msoa)
  #output2@data

  #output3 <- sppt_diff(sp1998, sp1999, msoa)
  #output3@data

  output.oa <- sppt(a, b, oa.gb)
  #output.oa@data
  results.1 <- summary_sppt(output.oa) %>% as_tibble() %>%
    mutate(year = t) %>% 
    bind_rows(results.1)

  output.oa2 <- sppt_boot(a, b, oa.gb)
  #output.oa2@data 
  results.2 <- summary_sppt(output.oa2) %>% as_tibble() %>% 
    mutate(year = t) %>% 
    bind_rows(results.2)

  #output.oa3 <- sppt_diff(a, b, oa.gb)
  #output.oa3@data
}

out.path1 <- paste0(path, "/outputs/sppt_oa.csv")
out.path2 <- paste0(path, "/outputs/sppt_boot_oa.csv")
write_csv(results.1, out.path1)
write_csv(results.2, out.path2)

```

## Kernel density estimation

```{r}
xy <- df %>%
    #slice_sample(n = 10000) %>% 
    filter(!is.na(oseast1m), #6k rows with missing east/north for all years
           year == 2010,
           substr(pc, 1,2) !="BT") %>% #drop Northern Ireland, for 2010 c. 4k obs.
    st_as_sf(coords = c("oseast1m", "osnrth1m"), crs = 27700) %>%
    st_transform(27700) %>%
    st_coordinates() %>%
    as_tibble()

gb <- st_as_sf(gb, coords = c("oseast1m", "osnrth1m"), crs = 27700) 
gb <- st_transform(gb, 27700)
gb.win <- as.owin(gb)    
xy.ppp <- ppp(x=xy$X, y = xy$Y, window = gb.win)

plot(density(xy.ppp, sigma = 1000))
```

### 2022-05-26 src/hex_heatmaps.Rmd
new images

@@
 lookup <- read_csv(path.lookup) %>% 
-  select(pcds, msoa11cd, dointr, doterm) %>% 
+  dplyr::select(pcds, msoa11cd, dointr, doterm) %>% 
   dplyr::rename(pc = pcds)
@@
 lookup <- read_csv(path.lookup) %>% 
-  select(pcds, oseast1m, osnrth1m) %>% 
+  dplyr::select(pcds, oseast1m, osnrth1m) %>% 
   dplyr::rename(pc = pcds)
@@
 ```
+## hex heatmaps log transformed
+
+```{r}
+# old code: counts per hexagon based on xy
+#xy point
+
+path.out <- paste0(path, "/outputs/hex_log", n)
+
+for (t in 1996:2012){
+  p <- df %>%
+    #slice_sample(n = 10000) %>% 
+    filter(!is.na(oseast1m), #6k rows with missing east/north for all years
+           year == t,
+           substr(pc, 1,2) !="BT") %>% #drop Northern Ireland, for 2010 c. 4k obs.
+    st_as_sf(coords = c("oseast1m", "osnrth1m"), crs = 27700) %>%
+    st_transform(4326) %>%
+    st_coordinates() %>%
+    as_tibble() %>% 
+    ggplot(aes(x=X,y=Y)) +
+
+    #stat_binhex(bins=150) +
+    stat_binhex(aes(fill=log(..count..)), bins=150) +
+    geom_polygon(data = gb.df, aes(x=long, y = lat, group = group), #gb.df
+                 fill=NA, color = "black", size = 0.1, alpha = 0.8)+
+    scale_fill_gradientn(colours=c('light gray','blue'),
+                         name='Frequency',na.value=NA) + #, trans = "log10") +
+    coord_equal() +
+    labs(x=NULL, y=NULL) +
+    geom_point(data = cities, aes(x=long, y=lat), colour = "black", size = .1) + 
+    geom_text(size = 2, check_overlap = F, nudge_y = .2, data = cities, aes(x=long, y=lat, label = name)) +
+    #theme(legend.key.size = unit(0.05, "cm")) + # legend.key.width=unit(1, "cm"), 
+    ggmap::theme_nothing(legend = T) +
+    theme(legend.title=element_text(size=5.5), legend.text=element_text(size=4)) +
+    #theme(legend.position=c(0.075,0.28))
+    guides(fill = guide_colorbar(barwidth = 0.5, barheight = 2.5)) +
+    ggtitle(t)
+  ggsave(p, filename = paste(path.out, t, ".png", sep = ""))
+}
+```

### 2022-05-26 src/la_data.Rmd
new images

---
title: "LA data"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: FALSE
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

Taken from `broadband.speed.covid` project.

```{r settings, include = FALSE}
library(tidyverse)
library(rprojroot)
# library(lubridate)
# library(knitr)
# library(rprojroot)
# library(nnet)
# library(stargazer)
# library(openair)
# library(rgdal)
# library(corrplot)
# library(DescTools)
library(geosphere)
library(sf)
# library(rgeos)
library(raster)

# This is the project path
path <- find_rstudio_root_file()
```

## Distance to Metro Area

Assumes LA spatial object is already loaded

```{r}

# spatial transformations
ladDist <- la@data[,c("LAD21CD", "LAD21NM")]

############# names(ladDist)[1] <- "lad19cd"

#distance from London and other 
coords_LAD <- la@data[,c("LONG", "LAT")]
London <- la[which(la@data$LAD21NM == "City of London"),]
London <- gCentroid(London)
distLondon <- distm(coords_LAD, London, fun=distCosine) 
#result in meters - change to km
distLondon <- round((distLondon/1000),2)
Birmingham <- la[which(la@data$LAD21NM == "Birmingham"),]
Birmingham <- gCentroid(Birmingham)
distBham <- distm(coords_LAD, Birmingham, fun=distCosine)
distBham <- round((distBham/1000),2)
Manchester <- la[which(la@data$LAD21NM == "Manchester"),]
Manchester <- gCentroid(Manchester)
distManc <- distm(coords_LAD, Manchester, fun=distCosine)
distManc <- round((distManc/1000),2)
Leeds <- la[which(la@data$LAD21NM == "Leeds"),]
Leeds <- gCentroid(Leeds)
distLeeds <- distm(coords_LAD, Leeds, fun=distCosine)
distLeeds <- round((distLeeds/1000),2)
Liverpool <- la[which(la@data$LAD21NM == "Liverpool"),]
Liverpool <- gCentroid(Liverpool)
distLivp <- distm(coords_LAD, Liverpool, fun=distCosine)
distLivp <- round((distLivp/1000),2)
Newcastle <- la[which(la@data$LAD21NM == "Newcastle upon Tyne"),]
Newcastle <- gCentroid(Newcastle)
distNewc <- distm(coords_LAD, Newcastle, fun=distCosine)
distNewc <- round((distNewc/1000),2)
Sheffield <- la[which(la@data$LAD21NM == "Sheffield"),]
Sheffield <- gCentroid(Sheffield)
distShef <- distm(coords_LAD, Sheffield, fun=distCosine)
distShef <- round((distShef/1000),2)
SouthHants <- la[which(la@data$LAD21NM == "Fareham"),]
SouthHants <- gCentroid(SouthHants)
distSHam <- distm(coords_LAD, SouthHants, fun=distCosine)
distSHam <- round((distSHam/1000),2)
Nottingham <- la[which(la@data$LAD21NM == "Nottingham"),]
Nottingham <- gCentroid(Nottingham)
distNotts <- distm(coords_LAD, Nottingham, fun=distCosine)
distNotts <- round((distNotts/1000),2)
Bristol <- la[which(la@data$LAD21NM == "Bristol, City of"),]
Bristol <- gCentroid(Bristol)
distBris <- distm(coords_LAD, Bristol, fun=distCosine)
distBris <- round((distBris/1000),2)
Cardiff <- la[which(la@data$LAD21NM == "Cardiff"),]
Cardiff <- gCentroid(Cardiff)
distCard <- distm(coords_LAD, Cardiff, fun=distCosine)
distCard <- round((distCard/1000),2)
Glasgow <- la[which(la@data$LAD21NM == "Glasgow City"),]
Glasgow <- gCentroid(Glasgow)
distGlas <- distm(coords_LAD, Glasgow, fun=distCosine) 
distGlas <- round((distGlas/1000),2)

ladDist <- cbind(ladDist, distLondon, distBham, distManc, distLeeds,
                 distLivp, distNewc, distShef, distSHam, distNotts, 
                 distBris, distCard, distGlas)

ladDist$distMet <- apply(ladDist[,3:14], 1, min)
ladDist <- ladDist %>%
  dplyr::select(LAD21CD, LAD21NM, distLondon, distMet)

#test <- left_join(lad, ladDist)
```

## Job density 2018

```{r}

job.dens <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_57_1.data.csv?geography=1816133633...1816133848,1820327937...1820328318&date=latest&item=3&measures=20100")

job.dens <- job.dens %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, OBS_VALUE) %>%
  rename(job.dens2018 = OBS_VALUE) %>%
  rename(LAD21CD = GEOGRAPHY_CODE) %>%
  distinct()

# test <- left_join(lad, job.dens, by = "lad19cd")
```

## Population 2018

```{r}
# pop <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_31_1.data.csv?geography=1807745025...1807745028,1807745030...1807745032,1807745034...1807745155,1807745157...1807745164,1807745166...1807745170,1807745172...1807745177,1807745179...1807745194,1807745196,1807745197,1807745199,1807745201...1807745218,1807745221,1807745222,1807745224,1807745226...1807745231,1807745233,1807745234,1807745236...1807745244,1807745271...1807745281,1811939329...1811939332,1811939334...1811939336,1811939338...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS1-latest&sex=7&age=0,22&measures=20100")

pop <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_31_1.data.csv?geography=1820327937...1820328318,1816133633...1816133848&date=latestMINUS1-latest&sex=7&age=0,22&measures=20100")

pop <- pop %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, AGE_NAME, OBS_VALUE) %>%
  distinct(GEOGRAPHY_CODE, AGE_NAME, .keep_all = T) %>% 
  spread(AGE_NAME, OBS_VALUE) %>% 
  rename(pop16_64 = `Aged 16 - 64`) %>%
  rename(pop = `All ages`) %>%
  rename(LAD21CD = GEOGRAPHY_CODE)

# test <- left_join(lad, pop, by = "lad19cd")
# sapply(test, function(x) sum(is.na(x)))
```

## Population density 2018

```{r}

# this is just population
pop.dens <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_31_1.data.csv?geography=1820327937...1820328318,1816133633...1816133848&date=latestMINUS1-latest&sex=7&age=0,22&measures=20100")

pop.dens <- pop.dens %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, AGE_NAME, OBS_VALUE) %>%
  distinct(GEOGRAPHY_CODE, AGE_NAME, .keep_all = T) %>% 
  spread(AGE_NAME, OBS_VALUE) %>% 
  rename(pop16_64 = `Aged 16 - 64`) %>%
  rename(pop = `All ages`) %>%
  rename(LAD21CD = GEOGRAPHY_CODE) %>%
  dplyr::select(-pop16_64)

raster::crs(la)
la$area_sqkm <- raster::area(la) / 1000000

pop.dens <- merge(pop.dens, la@data, by.x = "LAD21CD", by.y = "LAD21CD", all.x = T)

pop.dens <- pop.dens %>%
  mutate(pop.dens = pop/area_sqkm) %>%
  dplyr::select(LAD21CD,pop.dens)

# sapply(pop.dens, function(x) sum(is.na(x)))
# test <- pop.dens[is.na(pop.dens$area_sqkm),]
```

## N/S divide

```{r}

# It only includes only the south

ns <- read_csv("https://opendata.arcgis.com/datasets/3ba3daf9278f47daba0f561889c3521a_0.csv") %>%
  mutate(south = ifelse(RGN19NM=="London" | 
                        RGN19NM=="South West" |
                        RGN19NM=="East of England" |
                        RGN19NM=="South East", 1, 0)) %>%
  dplyr::select(LAD19CD, south) %>% 
  dplyr::filter(south==1) %>%
  rename(LAD21CD = LAD19CD)
```

## Labour supply March 2020

```{r}

labour <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_17_5.data.csv?geography=1816133633...1816133837,1820327937...1820328307&date=latest&variable=18,1532...1540,290,720&measures=20599,21001,21002,21003")

labour <- labour %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, VARIABLE_NAME, OBS_VALUE) %>%
  distinct(GEOGRAPHY_CODE, VARIABLE_NAME, .keep_all = T) %>% 
  spread(VARIABLE_NAME, OBS_VALUE) %>% 
  rename(LAD21CD = GEOGRAPHY_CODE) %>%
  rename(managers =  `% all in employment who are - 1: managers, directors and senior officials (SOC2010)`) %>%
  rename(prof = `% all in employment who are - 2: professional occupations (SOC2010)`) %>%
  rename(tech = `% all in employment who are - 3: associate prof & tech occupations (SOC2010)`) %>%
  rename(admin = `% all in employment who are - 4: administrative and secretarial occupations (SOC2010)`) %>%
  rename(skilled = `% all in employment who are - 5: skilled trades occupations (SOC2010)`) %>%
  rename(caring = `% all in employment who are - 6: caring, leisure and other service occupations (SOC2010)`) %>%
  rename(sales = `% all in employment who are - 7: sales and customer service occupations (SOC2010)`) %>%
  rename(plant = `% all in employment who are - 8: process, plant and machine operatives (SOC2010)`) %>%
  rename(elementary = `% all in employment who are - 9: elementary occupations (SOC2010)`) %>%
  rename(NVQ3 = `% with NVQ3+ - aged 16-64`) %>%
  rename(NVQ4 = `% with NVQ4+ - aged 16-64`) %>%
  rename(econ.act = `Economic activity rate - aged 16-64`) %>% # No NVQ data here as this is for Jul 2019-Jun 2020. NVQ is only  
  dplyr::select(-NVQ3, -NVQ4)                                  # available for Dec - Dec. I download these data seperetely.

# test <- left_join(lad, labour, by = "lad19cd")
# sapply(test, function(x) sum(is.na(x)))
```

## NVQ4 March 2020

```{r}

nvq4 <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_17_5.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1807745025...1807745028,1807745030...1807745032,1807745034...1807745155,1807745157...1807745164,1807745166...1807745170,1807745172...1807745177,1807745179...1807745194,1807745196,1807745197,1807745199,1807745201...1807745218,1807745221,1807745222,1807745224,1807745226...1807745231,1807745233,1807745234,1807745236...1807745244&date=latestMINUS2&variable=290&measures=20599,21001,21002,21003")

nvq4 <- nvq4 %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, MEASURES_NAME, OBS_VALUE) %>%
  filter(MEASURES_NAME=="Variable") %>%
  dplyr::select(-MEASURES_NAME) %>%
  distinct(GEOGRAPHY_CODE, OBS_VALUE, .keep_all = T) %>% 
  rename(LAD21CD = GEOGRAPHY_CODE) %>%
  rename(nvq4 = OBS_VALUE)
```

## Earnings 2019

```{r}

earnings <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_30_1.data.csv?geography=1946157348,2092957698,2013265929,1816133633...1816133837,1820327937...1820328307&date=latest&sex=8&item=2&pay=1&measures=20100,20701")

earnings <- earnings %>%
  filter(MEASURES_NAME == "Value") %>% # dropiing the confedence
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, OBS_VALUE) %>%
  rename(LAD21CD = GEOGRAPHY_CODE) %>%
  rename(earnings =  OBS_VALUE) %>%
  distinct()

# test <- left_join(lad, earnings, by = "lad19cd")
# sapply(test, function(x) sum(is.na(x)))
```

## Business counts 2019

```{r}

# busi <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_141_1.data.csv?geography=1816133633...1816133848,1820327937...1820328318,1870659585...1870659791,1870659801,1870659792...1870659800,1879048193...1879048573,1879048583,1879048574...1879048582&date=latest&industry=163577857...163577874&employment_sizeband=0&legal_status=0&measures=20100")

busi <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_141_1.data.csv?geography=1816133633...1816133848,1820327937...1820328318,1870659585...1870659791,1870659801,1870659792...1870659800,1879048193...1879048573,1879048583,1879048574...1879048582&date=latest&industry=150994945...150994965&employment_sizeband=0&legal_status=0&measures=20100")

busi <- busi %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, INDUSTRY_NAME, OBS_VALUE) %>%
  distinct(GEOGRAPHY_CODE, INDUSTRY_NAME, .keep_all = T) %>% 
  spread(INDUSTRY_NAME, OBS_VALUE) %>% 
  rename(LAD21CD = GEOGRAPHY_CODE) %>%
  rename(A = `A : Agriculture, forestry and fishing`) %>%
  rename(B = `B : Mining and quarrying`) %>%
  rename(C = `C : Manufacturing`) %>%
  rename(D = `D : Electricity, gas, steam and air conditioning supply`) %>%
  rename(E = `E : Water supply; sewerage, waste management and remediation activities`) %>%
  rename(F = `F : Construction`) %>%
  rename(G = `G : Wholesale and retail trade; repair of motor vehicles and motorcycles`) %>%
  rename(H = `H : Transportation and storage`) %>%
  rename(I = `I : Accommodation and food service activities`) %>%
  rename(J = `J : Information and communication`) %>%
  rename(K = `K : Financial and insurance activities`) %>%
  rename(L = `L : Real estate activities`) %>%
  rename(M = `M : Professional, scientific and technical activities`) %>%
  rename(N = `N : Administrative and support service activities`) %>%
  rename(O = `O : Public administration and defence; compulsory social security`) %>%
  rename(P = `P : Education`) %>%
  rename(Q = `Q : Human health and social work activities`) %>%
  rename(R = `R : Arts, entertainment and recreation`) %>%
  rename(S = `S : Other service activities`) %>%
  rename(T = `T : Activities of households as employers;undifferentiated goods-and services-producing activities of households for own use`) %>%
  rename(U = `U : Activities of extraterritorial organisations and bodies`) %>%
  mutate(total.busi = rowSums(.[3:23]))

# test <- left_join(lad, busi, by = "lad19cd")
# sapply(test, function(x) sum(is.na(x)))
```

## Urban/rural

```{r eval=F}
# not complete
url <- "https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/137663/la-class-dataset-post0409.xls"
GET(url, write_disk(tf <- tempfile(fileext = ".xls")))
urban.rural <- read_excel(tf, 1L, range = "A17:CP1877", col_names = T)

```

## merge

```{r}
data.la <- list(ladDist, job.dens, pop, pop.dens, labour, earnings, busi) # ns; nvq4 
sapply(data.la, function(x) dim(x))
sapply(data.la, function(x) names(x))

# merge with reduce
data.la <- data.la %>% 
  reduce(inner_join, by = "LAD21CD") %>%
  #distinct(.keep_all = T)
  unique()
sapply(data.la, function(x) sum(is.na(x)))

data.la <- merge(data.la, ns, all.x = T)
data.la$south <- ifelse(is.na(data.la$south),0, data.la$south)

# NOT RUN
# export file for future reference / backup
# data.out.path <- paste0(path, "/data/temp/data_for_aux.csv")
# write.csv(data, data.out.path)
```

### 2022-05-26 src/s_la.Rmd
new images

@@
 library(rprojroot)
+library(rgdal)
+library(sf)
+library(modelr)
 
@@
 ```
+
+## TODO
+
+- speed, other metrics? 
+- understand rsquare
+- model LAs
+- add SE for the line plots [link](https://stackoverflow.com/questions/26020142/adding-shade-to-r-lineplot-denotes-standard-error)
 
@@
 ```{r}
-df.test <- df %>% filter(ladcd == "S12000033")
+df.test <- df %>% filter(ladcd == "E06000002")
 
@@
 
+rsquare(fit, df.country)
+
 df.country$prednls = predict(fit)
 
-ggplot(df.country, aes(x = year, y = n) ) +
+df.country %>% 
+  ggplot(aes(x = year, y = n) ) +
   geom_point() +
@@
   scale_x_continuous("year", labels = as.character(df.country$year), breaks = df.country$year) +
+  geom_vline(xintercept = 2003) +
   ylab("N. of websites") +
@@
 
+## Loop for S function estimation for Local Authorities keeping the model objects
+
+```{r}
+a.la <- as_tibble()
+all.la <- df %>% filter(ladcd != "NA") %>% distinct(ladcd) %>% simplify2array()
+simplify2array(all.la)
+
+for (i in all.la){
+  data <- df %>% filter(ladcd==i) %>%
+    select(year, ladcd, n) 
+    #lm(n~year, data = .)
+    #tryCatch(tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .)), , error=function(e) NULL)
+  tryCatch(model <- (nls(n ~ SSlogis(year, Asym, xmid, scal), data = data)), error=function(e) NULL)
+  model.name <- paste0("model.", i)
+  assign(model.name, model)
+  tryCatch(b.la <- tidy(model), error=function(e) NULL)
+
+  tryCatch(b.la$r2 <- rsquare(model, data), error=function(e) NULL)
+
+  tryCatch(a.la <-rbind(a.la,b.la), error=function(e) NULL)
+}
+
+a.la <- a.la %>% filter(term=="xmid") %>% 
+  bind_cols(as.tibble(all.la)) %>% 
+  #mutate(estimate = round(estimate, 0)) %>% 
+  #summarise(r = range(r2))
+  filter(p.value < 0.01 & 
+         r2 > 0.9) %>%    # 376 / 376, all sig, 330 r2 > 0.9
+  select(ladcd, estimate, r2, std.error) %>%
+  mutate(fast = ifelse(estimate < 2003, "fast", "slow"))
+
+# predictions 
+pred <- tibble()
+help <- tibble(year = 1996:2012)
+for (i in all.la){
+  model.name <- (paste0("model.", i))
+  #assign(col.name, paste0("pred.", i))
+  pred_ <- df %>% filter(ladcd == i) %>% 
+    mutate(prednls = eval(call("predict", as.name(model.name))))  # "pred.{i}" :=
+  help_ <- help %>% left_join(pred_)
+
+  pred <- pred %>% bind_rows(help_) 
+}
+
+# plots
+path.out <- paste0(path, "/outputs/lad_")
+
+for (i in all.la){
+p <- ggplot(pred %>% left_join(a.la) %>% 
+              filter(ladcd == i), aes(x = year, y = n, color = fast) ) +
+  geom_point() +
+  geom_line(aes(y = prednls), size = 1) +
+  scale_x_continuous("year", labels = as.character(df.test$year), breaks = df.test$year) +
+  geom_vline(xintercept = pred %>% left_join(a.la) %>% filter(ladcd == i) %>% select(estimate) %>% first()) +
+  ylab("N. of websites") + 
+  labs(title = paste0(la@data %>% filter(LAD21CD== i) %>% select(LAD21NM) %>% first(), ", ", i)) +
+  theme(legend.position=c(.9,.15)) +
+  scale_color_manual(values = c("fast" = "purple",
+                                "slow"="orange"))
+
+ggsave(p, filename = paste(path.out, i, ".png", sep = ""))
+
+}
+```
+
+## Xmid frequency graph
+
+```{r}
+ggplot(a.la, aes(as.factor(round(estimate, 0)))) +
+  geom_bar(fill = "#0073C2FF")
+```
+
 write the object with all the models
@@
 write_csv(a.la, path.out)
+```
+
+## Survival
+
+- pop.density
+- income
+- firms
+- employment
+- N/S
+- London
+
+Time series formula will look like this:
+
+`coxph(Surv(tstart, tstop, infect) ~ treat + inherit + steroids, data =newcgd, cluster = id)`
+
+This is what @perkins2005international should be using.
+See p. 7 [here](https://cran.r-project.org/web/packages/survival/vignettes/timedep.pdf).
+
+```{r}
+library(survival)
+
+test.surv.data <- a.la %>% mutate(status = 1) %>% 
+  left_join(data.la, by = c("ladcd" = "LAD21CD")) %>% 
+  left_join(df %>% filter(year == 2012), by = "ladcd")
+
+cf <- coxph(Surv((estimate-1996), status) ~ log(distLondon) + log(distMet) + log(pop16_64) + log(pop.dens) + log(econ.act) + log(earnings) + log(total.busi) + south + log(n), data=test.surv.data, weights = 1/(std.error)^2)
+summary(cf)
+
 ```
@@
 ```
+
+## time series clusters
+
+2 clusters is the preferred solution, but when I plot individual local authorities 
+I cannot interpret the differences between the two clusters
+
+```{r}
+library(dtwclust)
+library(tsbox)
+library(imputeTS)
+
+# turn to a wide format
+ts.wide <- df %>% 
+  filter(!is.na(ladcd)) %>% 
+  rename(time = year) %>%
+  dplyr::select(time, ladcd, n) %>% 
+  #mutate(year = )
+  ts_wide()
+
+# impute with means of each TS
+# TODO check sensitivity
+ts.wide <- na_mean(ts.wide)
+#sapply(ts.wide, function(x) sum(is.na(x)))
+
+# extract names
+ts.wide.names <- names(ts.wide[,-1])
+ts.wide <- ts.wide[,-1]
+
+# standardise
+ts.wide <- zscore(ts.wide)
+
+# convert to list for tsclust
+ts.wide <- split(ts.wide, rep(1:ncol(ts.wide), each = nrow(ts.wide)))
+
+# 5 - 20 cluster sollutions
+pc_k <- tsclust(ts.wide, k = c(2:10), seed = 94L,
+                type="partitional",
+                distance = "dtw_basic", centroid = "pam", trace = T, 
+                args = tsclust_args(dist = list(window.size = 20L)))
+
+# decide k, see RJ-2019-023 for max/min
+#names(pc_k) <- paste0("k_", c("5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15"))
+names(pc_k) <- paste0("k_", c("2", "3", "4", "5", "6", "7", "8", "9", "10"))
+cvis <- sapply(pc_k, cvi, type = "internal")
+cvis
+# max Sil
+# max SF
+# max CH
+# min DB
+# min DBstar
+# max D
+# min COP
+
+# loop for k decision
+pref.cvi <- NULL
+
+max.cvis <- c(1:3,6)
+for (i in max.cvis){
+  a <- which.max(cvis[i,])
+  pref.cvi <- c(pref.cvi, a)
+}
+
+min.cvis <- c(4:5,7)
+for (i in min.cvis){
+  a <- which.min(cvis[i,])
+  pref.cvi <- c(pref.cvi, a)
+}
+
+print(pref.cvi) # just count how many time each k_* appears, this is the preferred solution
+
+pc_k <- tsclust(ts.wide, k = 7, seed = 94L,  # 7 or 2
+                type="partitional", 
+                distance = "dtw_basic", centroid = "pam", trace = T, 
+                args = tsclust_args(dist = list(window.size = 20L)))
+plot(pc_k)
+
+clusters <- data.frame(LAD=ts.wide.names,
+                           cluster=pc_k@cluster)
+
+```
+
+## Mapping the clusters
+
+```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}
+
+la.f <- la.f %>% left_join(clusters, by = c("id" = "LAD")) %>% 
+  arrange(order)  # if i don't order merge.nuts.f loses order and the map has gaps
+
+ggplot(la.f %>% filter(!is.na(r2)), aes(x = long, y = lat)) +
+  geom_polygon(aes( group = group, fill = as.factor(cluster))) +
+  #theme_nothing(legend = TRUE) +
+  labs(title = "Clusters") +
+  scale_fill_viridis_d()+
+  theme_void()
+
+# ggplot(la.f, aes(x = long, y = lat)) +
+#   geom_polygon(aes( group = group, fill = as.factor(cluster.down))) +
+#   theme_nothing(legend = TRUE) +
+#   labs(title = "Download clusters") +
+#   scale_fill_viridis_d()+
+#   theme_void()
+# 
+# cluster.upL <- c(6, 9, 11, 12, 13)
+# cluster.upS <- c(1, 2, 3, 7, 8, 10)
+#  
+# la.f_ <- la.f #plot the background
+# 
+# cluster.upL.plot <- la.f %>% 
+#   filter(cluster.up %in% cluster.upL) %>% 
+#   ggplot(aes(x = long, y = lat)) +
+#   geom_polygon(data = la.f_, color='grey', fill='white', aes(x=long, y=lat, group=group, colour = 'grey30')) +
+#   geom_polygon(aes( group = group, fill = as.factor(cluster.up))) +
+#   labs(title = "Larger clusters, upload speeds") + 
+#   scale_fill_viridis_d() +
+#   guides(fill=guide_legend(title="Clusters")) +
+#   theme_void()
+# 
+# cluster.upS.plot <- la.f %>% 
+#   filter(cluster.up %in% cluster.upS) %>% 
+#   ggplot(aes(x = long, y = lat)) +
+#   geom_polygon(data = la.f_, color='grey', fill='white', aes(x=long, y=lat, group=group, colour = 'grey30')) +
+#   geom_polygon(aes( group = group, fill = as.factor(cluster.up))) +
+#   labs(title = "Smaller clusters, upload speeds") + 
+#   scale_fill_viridis_d() +
+#   guides(fill=guide_legend(title="Clusters")) +
+#   theme_void()
+# 
+# plot_grid(cluster.upL.plot, cluster.upS.plot, nrow = 1,
+#           align = "h", axis = "l", 
+#           #labels = c(".co.uk / pop., 2000", ".co.uk / pop., 2010", "non .co.uk / pop., 2000", "non .co.uk / pop., 2010", ""), 
+#           #label_size = 12,
+#           greedy = T)
+#           #rel_heights = c(1, 1, .3)) # this is the relative position of the 3 lines
+# 
+# #ggsave("all_maps2.png") # this is the colour map for the preprint
+# out.path.map <- paste0(path, "/paper/v1/figures/map.up.clusters.png")
+# ggsave(out.path.map)#,
+#        #width = 210, #A4
+#        #height = 297, #A4
+#        #units = "mm")
+```

### 2022-05-26 paper/paper.Rmd
papers

---
title: Title of submission to PLOS journal
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliation: University of Bristol and The Alan Turing Institute
    corresponding: e.tranos@bristol.ac.uk
address:
  - code: University of Bristol and The Alan Turing Institute
    address: University of Bristol, School of Geographical Sciences, Bristol, UK; The Alan Turing Institute, London, UK
abstract: |
  blah blah

author_summary: |
  blah blah

bibliography: mybibfile.bib
output: rticles::plos_article
csl: plos.csl
header-includes:
- \usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}
---

```{r echo=FALSE, message = FALSE, warning=FALSE}
library(stargazer)
library(tidyverse)
#library(rmarkdown)
```

# Introduction {#sec:1}

@beardsell1999spatial
To characterize evolution of the computer industry, we examine the distribution
of relative employment across cities in 1977 and how that distribution
changes over time to 1992

There is no
tendency of relative size distributions of urban computer employment to collapse,
go bimodal, or fully spread. Overall computers exhibit some turbulence,
with dramatic big winners and losers among cities. In attracting or repelling an
industry, urban heterogeneity is important. Large, well educated cities near San
Jose have much greater chances of attracting high-tech employment and less of
losing it

@bednarz2020pulled
spatial innovation diffusion
By using Bayesian survival models with time-dependent data of wind turbine deployment
and firm foundation for 402 German regions between the years 1970 and 2015, we show that the spatial
evolution of the German wind energy industry was more strongly influenced by local demand–
pull than local supply–push processes.

The industry’s initial locations are distributed relatively arbitrarily
and unpredictably, as their needs in terms of resources and skills are diverse and distinct from the older existing
industries (Boschma and Lambooy, 1999). Consequently, emerging industries are characterized by relatively high
degrees of freedom in terms of location. In later extensions of the concept, the assumption of the randomness of locations
was revised with greater importance assigned to regional conditions (Boschma and Lambooy, 1999; Fornahl
et al., 2012).

larger local technological niches (at least from a demand perspective) are
more likely to be formed in urban areas than in rural ones, which is in line with the work of Ha¨gerstrand (1965a,b).
However, size is not the only regional characteristic that may matter in this context. Regions also differ in accumulated
experiences and the presence of tacit knowledge with respect to products, as well as in actors’ propensity share
this knowledge among producers (Martin et al., 2019)

@bento2015spatial
The concept of absorptive capacity has also been applied in the TIS context in terms of the con-struction of capacity needed to transfer and implement a new technology in the receiving country (cf.Van Alphen, 2011), with repercussions for the nature of system functions

@das2022diffusion [not good paper]
Invention, innovation
and diffusion are considered as the three pillars of
the theory of technological change and Schumpeter is considered
as the father of the study of diffusion

Epidemic approach: The approach emphasizes that the adoption of a technology
spreads like an infection among the potential adopters.
By taking a homogeneous group of potential adopters as
an assumption, it explains the adoption of any innovation.

To summarize
the discussion, first, here the diffusion process is selfperpetuating
as initial use stimulates further use. Second, it
follows a disequilibrium path as the level of users is
always lower than the number of potential users (Stoneman
2002).

The
criticisms are mainly based on the fact that although the
approach gives an idea of aggregate (industry or household)
behaviour, it does not focus on the individual’s
(firm or household) adoption process

Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.) Early adopters: next
13.5%, (III.) Early Majority: following 34%, (IV.) Late Majority: next 34%, and the rest as (V.) Laggards

@morrill2020spatial
Spatial diffusion is the process by which behavior or characteristics of the landscape change as a result of
what happens elsewhere earlier. Spatial diffusion is the spread of the phenomenon, over space and timed,
from limited origins

Contagious diffusion=> spatial autocorelation

expansion-type diffusion

the three geographic diffusion processes are (1) purely contagious where distance or adjacency
is the absolute controlling factor, (2) purely hierarchical where size or urban position in the central place
hierarchy is the absolute controlling factor, or (3) where the location of change is purely random. These are
the pure or ideal forms of the diffusion process; all real diffusion processes are the result of a combination of
these extremes.

The work by the communication sociologist Everett Rogers (1962, 1971, 1983) has emphasized the role of
information, communication, formal and informal media, opinion leaders and social networks, and economic
and psychological constraints on acceptance. Rogers’s work stresses the decision mechanism of the potential
adopter; the work of Rogers partitions the adopter’s process of choosing to accept a new phenomenon or
trait into five stages: stage 1–the potential adopter gains knowledge or awareness of an innovation; stage
2–persuasion is exercised to adopt; stage 3–a decision is made to adopt; stage 4–the decision to adopt is
implemented; stage 5–the adopter confirms the decision to adopt. In Rogers’s stage theory there may be a
substantial time lag between when a potential adopter becomes aware of the new characteristic and when a
decision is actually made to adopt.

Hägerstrand considered diffusion to be a fundamental geographic process: Whatever the phenomenon being
diffused might be, one may consider it in the context of a larger universal process of spatial diffusion.

@zook2022mapping
Because new digital activities are rarely—if
ever—captured in official state data, researchers
must rely on information gathered from alternative
sources. With this in mind, the overarching aim of
this article is outlining an approach for analyzing the
“new spaces and geographies” of digital phenomena
and practices.

@leibowicz2016representing for energy
There is no generally accepted theory that explains diffusion
rate heterogeneity across technologies, but several factors are
considered important. Greater unit scale and larger market size
contribute to slower diffusion. Requirements for interrelated
technologies or complex infrastructures also hinder the diffusion
process (Grubler, 2012).

Mobile phones benefited from early
deployment in recreational boats and automobiles, where the
traditional competitor was not a viable option. In the early
stages of diffusion, performance is a more important driver of
adoption than cost competitiveness. Typically, significant cost
reductions only occur once the technology reaches a deployment
level capable of supporting standardization and mass
production (Wilson, 2012).

Diffusion is a spatial (Hagerstrand, 1967) as well as a
temporal process, and historical evidence confirms that
technologies diffuse at different times, at different rates, and
to different extents in different places, and can be significantly
influenced by policies (Victor, 1993). An illustrative empirically
based theory for technology diffusion acrossmultiple regions is
suggested by “Schmidt's Law” (Grubler, 1990). With respect to a
specific technology, it divides regions into three groups: core,
rim, and periphery. The core iswhere the technology is invented,
subject to experimentation, and first deployed commercially.
To develop the technology, the core must have both sufficient
innovative ability and economic motivation to make the
necessary resource commitment. As the technology spreads
from the core to the rim and eventually to the periphery, later
adopters have the advantage of benefiting from the experiences
of earlier adopters via knowledge spillover effects (Verdolini and
Galeotti, 2011). At this point, the technology has likely been
refined and costs have likely declined significantly. Schmidt's
Law claims that, as a result, diffusion proceeds at a faster rate in
regions which adopt the technology later. However, diffusion
tends to be less pervasive in regions where adoption is delayed,
ultimately saturating at a lower extent than in regions where
adoption occurred earlier. This is often attributed to a lack of
proper institutions and infrastructures in later adopters, markets
whichmay not be as favorable to a technology, or the availability
of yet newer options that invite technological leapfrogging.

Empirical evidence supports the validity of Schmidt's Law
over a wide range of technologies, time periods, and geographical
contexts. A recent meta-analysis of technology up-scaling
found that diffusion accelerated moving from the core to the
rim and periphery for technologies as diverse as natural gas
power, oil refineries, and automobiles (Wilson, 2009). One
historical example that conforms particularly well to Schmidt's
Law is the diffusion of coal power in Europe (Grubler, 2012).
England emerged as the core region for coal power because it
had legal and economic institutions that incentivized scientific
pursuits, domestic coal reserves, and a clear industrial motivation
to replace water power with coal. 
...

Lastly, it can be globally optimal for innovative economies
to deploy advanced technologies more than what is locally
optimal if this enables faster diffusion in less advanced regions
that could benefit from the technology.

@wilson201281
Logistic growth describes an initial period of gradual diffusion as a technology is introduced as a new commercial application, moving then through a rapid, exponential growth phase, before slowing and eventually saturating [@grubler1999dynamics]. The substitution of incumbent technologies by new competitors leads to subsequent decline and eventual obsolescence.

Early on in their lifecycle, new technologies are crude, imperfect,
and expensive (Rosenberg, 1994). New energy technologies
are attractive for their ability to perform a particular task or
deliver a new or improved energy service (Fouquet, 2010). This is
often circumscribed by a particular set of needs in a particular
context: a market ‘niche’. End-users in niche markets are generally
less sensitive to the effective price of the energy service
provided or have a higher willingness to pay for its performance
advantages (Fouquet, 2010). Thus initially, performance dominates
cost competitiveness (Wilson and Grubler, 2011). Market
niches afford some protection from competitive pressures, allowing
technologies to be tested and improved in applied settings,
reducing uncertainties with performance or market demand
(Kemp et al., 1998). Costs may only fall substantively after an
extended period of commercial experimentation, concurrent with
the establishment of an industrial base and characteristic moves
towards standardisation and mass production (Grubler, 1998).
The influence of accumulating production experience on costs is
captured by the concept of learning.

@grubler1990rise
Following the definition given by Rogers, 1983, diffusion is the
process by which an innovation is communicated through certain communication
channels over time among the members of a social system.
An innovation may be an idea, object or practice which is perceived
as new by an individual or another unit of adoption (e.g., an
organization) .
Communication via communication channels refers to how
knowledge about an innovation is transferred between the members of
a social system. Communication takes time, and is therefore an
important aspect of innovation diffusion. In fact, all theoretical and
empirical diffusion studies agree that an innovation does not instantly
diffuse into a social system. Instead a typical time pattern of diffusion
along a S-shaped trajectory seems to be the rule. The S-shaped pattern
of diffusion appears to be a basic anthropologic phenomenon, as
it is also confirmed by a large number of studies· of primitive preindustrial
societies.

The scale of spatial innovation diffusion is similar* to that of
temporal diffusion models: hierarchically decomposed.

Spatial innovation diffusion in Hagerstrand's model consists of
three stages.
Stage 1: Local concentrations of initial acceptances (initial
agglomeration) .
Stage 2: Radial dissemination outward from the initial agglomeration,
accompanied by the rise of secondary agglomerations.
Stage 3: The condensing and saturation stage, in which filling-in
occurs and diffusion eventually ceases.

Thus, three main characteristics define the spatial diffusion process:
(1) As in temporal diffusion models, an S-shaped pattern in the
cumulative level of adoption.
(2) A "hierarchy effect" by which diffusion proceeds from initial
centers to secondary ones. Within the context of urban systems
(central places) this implies a hierarchical spread of an innovation
from larger to smaller centers, and from the centers to the surrounding
hinterland.
(3) A "neighborhood effect" Le., diffusion proceeds outwards from
innovation centers, first "hitting" (Brown, 1981) nearby rather
than far-away locations.

Later Hagerstrand conceptualized
physical "barrier" effects like lakes or uninhabited areas, which, in
addition to distance, act as further retarding effects on diffusion.
These are formalized in the form of "zero" or "half" contact multiplicators
on the (distance decaying) message flows.

With respect to the formalization of the communication flows
Hagerstrand defines a "mean information field" (MIF), in which the
probability of communication is a negative function of distance
between individuals

@ding2010modeling model the spatial diffusion of mobile telecommunications in China
as well as its determinants. regional socioeconomic characteristics play an
important role in determining the timing,
speed, and level of mobile telecommunications
diffusion in China

Rogers
(1995) characterized early adopters as knowledgeable
risk takers. Griliches (1957) looked
at the rate of return for early adopters and
characterized them as profit maximizers. The
geographer Hagerstrand (1967) was focusing
on the character of early adopters in the transmittal
process that produces follow-on participants
in the process. Space was treated as a
contiguity system with embedded characteristics
producing barrier elements to the diffusion
process.

REVIEW ON Technological Diffusion

Perkins and Neumayer (2005)
examined whether latecomers have a higher diffusion
rate of three technologies including digital
telephone mainline

@bento2018time explore 
What determines the duration of
formative phases for energy innovations in different markets? We are interested
both in initial markets (also: core, lead, first mover, early
adopter) where formative phases prepare technologies for mass commercialization,
and in follower markets (also: periphery, lag, late
adopter) where accelerated formative phases may benefit from diffusion
and spillovers.

Hagerstrand (1968) was the first
to demonstrate that diffusion typically starts from innovative centers
(‘core´) and disseminates, through a hierarchy of subcenters, to the
periphery. This process is sequential rather than simultaneous, and
tends to accelerate from core to periphery, where diffusion reaches a
lower intensity than in the core (Hagerstrand, 1968; Grubler, 1990;
Morrill, 2005). A similar effect of spatial acceleration has been identified
in marketing research and is known as the “lead-lag effect” (Peres
et al., 2010).

@perkins2005international analyze whether the rate at which new producer
technologies diffuse is significantly influenced by
(1) latecomer advantage and (2) engagement with the
global economy via trade and foreign investment

Indeed, precisely
because of these latecomer advantages, developing
countries1 are believed to be well placed to catch up with
developed ones (Gerschenkron 1962; Abramovitz
1986).

Geographers, of course, were at the forefront of
quantitative diffusion research. Ha¨gerstrand’s (1967)
early work on understanding innovation diffusion in
Sweden using simulation modeling is particularly noteworthy
in this respect. However, already long before the
‘‘cultural turn’’ in economic geography (Barnes 2001),
geographers’ interest in formally modeling the diffusion
of new innovations waned (Gregory 1985). Indeed,
much of the recent quantitative research into the diffusion
of new technology has been undertaken by
economists, sociologists, and business scholars.

Theoretical models disagree as to why firms adopt
innovations at different times. Epidemic models emphasize
information (Griliches 1957). Certain firms are hypothesized
to adopt earlier because they come into
contact with, and learn from, adopters of the new technology
before others. Economic models, on the other
hand, predominantly emphasize firm heterogeneity (Ireland
and Stoneman 1986). Firms adopt technologies at
different times because they differ with respect to various
organizational and environmental variables influencing
the economic returns from adoption (Blackman 1999).

Either
way, the strong assumption is that developing countries
can acquire modern technology innovated in developed
economies, often at a fraction of the original research
and development (R&D) costs, thereby leapfrogging
many decades of technological progress (Teece 2000).

The first involves a country’s geographical location.
Recent empirical work suggests that diffusion is ‘‘geographically
localized’’ (Globerman, Kokko, and Sjo¨holm
2000; Keller 2002; Milner 2003) in that a technology
diffuses faster in a country where it is already more
widely diffused in neighboring countries. Underlying
these regional effects are contagion and contact with
prior users or producers of technology.

@perkins2011internet
Two main mechanisms are
identified in the literature to explain diffusion:
(1) epidemic-type dynamics whereby contact
with previous adopters stimulates uptake as po
tential
adopters learn about a new innovation;
and (2) economic-type mechanisms whereby
potential users adopt a new innovation as it be
comesmore
profitable, useful, or valuable, with
uptake characteristically spreading as costs be
come
lower, performance improves, or the po
tential
uses of the innovation grow over time.

To this extent,
our results contribute to a growing body of
work that has sought to caution against claims
about the supposed novelty of the Internet and
the suggestion that it is somehow different.

@meade2021modelling
There are numerous agents who would benefit from accurate predictions of introduction times.
These include: suppliers of network equipment for production planning; suppliers of mobile phones who need to have inventories of
appropriately configured handsets; international network planners who need to provide sufficient bandwidth for communications
between different countries; network operators and regulatory authorities.

Classifying adopters of a technology along a time axis was pioneered by
Rogers (1962), who categorized adopters into five groups as innovators, early adopters, early majority, late majority and laggards.

An a priori approach groups similar countries into segments, where similarity is defined using economic, social, or
political covariates. Examples include Hofstede (2001) and Lee (1990). An a posteriori analysis considers segments based on realized
market behaviour, for example see Sood et al. (2009). Some studies investigate changing segment membership over time, see Cannon
and Yaprak (2011).

@fritsch2015new

@fritsch2015new analyze the spatial diffusion of laser technology research in West Germany from
1960, when this technology began, until 2005.

Hägerstrand (1952, 1967) was the first scholar to intensively analyze spatial diffusion pro
cesses.
Based on the empirical observations of the regional spread of process innovations,
he hypothesized that new technology is first implemented in the ‘center’ – the large agglom
erations
– and then moves ‘down’ the spatial hierarchy, the last stop being the ‘periphery’,
i.e. remote and sparsely populated regions. There are at least two factors that may explain
this pattern: first, the transfer of tacit knowledge via face-to-face contact and second, the
absorptive capacity of a region.

# Literature review {#sec2} MAYBE SKIP THIS??

# Materials and Methods {#sec3}

@wilson201281 growth function description in p. 86

from R: Asym/(1+exp((xmid-input)/scal))
using the terms from Wilson k/(1 + e^(t_0 - t)/scal)

Wilson: k(1 + e^-b(t - t_0))
So, scal = - 1/b

The curve is symmetric* around t_0

**minimum R2 = 95%**, see also @grubler1990rise

The literature usually uses the saturation level as the asymptote. I am using the total
number of websites as we cannot compute a rate.

Moran's I as @ding2010modeling
**TODO** for t_0,  diffusion speed

## Multi-country analysis {#sec3.1}

## 3.2	Case study approach {#sec3.2}

# Digital technologies and spatial structure: a global view {#sec4}

# The impact of ICT on the US and the UK spatial structures {#sec5}

# Discussion and conclusions {#sec6}

contrary to results from future studies regarding social media [@lengyel2020role], web technologies 
did not exclusively spread from a central location.

@lengyel2020role
We see that the OSN
spread almost exclusively from the original location (the capital Budapest, with an order of magnitude more inhabitants than
the next size town) to various parts of the country in the early phase of the life-cycle. Later, diffusion became less mono-centric
and other towns also emerged as spreaders. Our findings support the idea that spreading initially happens to large distances and
becomes more local over time. This is illustrated and discussed later in Figure 2I.

# References {#references .unnumbered}

### 2022-05-26 paper/paper.tex
papers

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
% \bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\setcounter{secnumdepth}{0}

\newcommand{\getAff}[1]{
  \getIndex{#1}{University of Bristol and The Alan Turing Institute}
}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Title of submission to PLOS
journal} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Emmanouil Tranos\textsuperscript{\getAff{University of Bristol and The
Alan Turing Institute}}\textsuperscript{*}\\
\bigskip
\textbf{\getAff{University of Bristol and The Alan Turing
Institute}}University of Bristol, School of Geographical Sciences,
Bristol, UK; The Alan Turing Institute, London, UK\\
\bigskip
* Corresponding author: e.tranos@bristol.ac.uk\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
blah blah

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
blah blah

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\hypertarget{sec:1}{%
\section{Introduction}\label{sec:1}}

{[}1{]} To characterize evolution of the computer industry, we examine
the distribution of relative employment across cities in 1977 and how
that distribution changes over time to 1992

There is no tendency of relative size distributions of urban computer
employment to collapse, go bimodal, or fully spread. Overall computers
exhibit some turbulence, with dramatic big winners and losers among
cities. In attracting or repelling an industry, urban heterogeneity is
important. Large, well educated cities near San Jose have much greater
chances of attracting high-tech employment and less of losing it

{[}2{]} spatial innovation diffusion By using Bayesian survival models
with time-dependent data of wind turbine deployment and firm foundation
for 402 German regions between the years 1970 and 2015, we show that the
spatial evolution of the German wind energy industry was more strongly
influenced by local demand-- pull than local supply--push processes.

The industry's initial locations are distributed relatively arbitrarily
and unpredictably, as their needs in terms of resources and skills are
diverse and distinct from the older existing industries (Boschma and
Lambooy, 1999). Consequently, emerging industries are characterized by
relatively high degrees of freedom in terms of location. In later
extensions of the concept, the assumption of the randomness of locations
was revised with greater importance assigned to regional conditions
(Boschma and Lambooy, 1999; Fornahl et al., 2012).

larger local technological niches (at least from a demand perspective)
are more likely to be formed in urban areas than in rural ones, which is
in line with the work of Ha¨gerstrand (1965a,b). However, size is not
the only regional characteristic that may matter in this context.
Regions also differ in accumulated experiences and the presence of tacit
knowledge with respect to products, as well as in actors' propensity
share this knowledge among producers (Martin et al., 2019)

{[}3{]} The concept of absorptive capacity has also been applied in the
TIS context in terms of the con-struction of capacity needed to transfer
and implement a new technology in the receiving country (cf.Van Alphen,
2011), with repercussions for the nature of system functions

{[}4not good paper{]} Invention, innovation and diffusion are considered
as the three pillars of the theory of technological change and
Schumpeter is considered as the father of the study of diffusion

Epidemic approach: The approach emphasizes that the adoption of a
technology spreads like an infection among the potential adopters. By
taking a homogeneous group of potential adopters as an assumption, it
explains the adoption of any innovation.

To summarize the discussion, first, here the diffusion process is
selfperpetuating as initial use stimulates further use. Second, it
follows a disequilibrium path as the level of users is always lower than
the number of potential users (Stoneman 2002).

The criticisms are mainly based on the fact that although the approach
gives an idea of aggregate (industry or household) behaviour, it does
not focus on the individual's (firm or household) adoption process

Rogers3, that divides adopters as: (I.) Innovators: first 2.5\%, (II.)
Early adopters: next 13.5\%, (III.) Early Majority: following 34\%,
(IV.) Late Majority: next 34\%, and the rest as (V.) Laggards

{[}5{]} Spatial diffusion is the process by which behavior or
characteristics of the landscape change as a result of what happens
elsewhere earlier. Spatial diffusion is the spread of the phenomenon,
over space and timed, from limited origins

Contagious diffusion=\textgreater{} spatial autocorelation

expansion-type diffusion

the three geographic diffusion processes are (1) purely contagious where
distance or adjacency is the absolute controlling factor, (2) purely
hierarchical where size or urban position in the central place hierarchy
is the absolute controlling factor, or (3) where the location of change
is purely random. These are the pure or ideal forms of the diffusion
process; all real diffusion processes are the result of a combination of
these extremes.

The work by the communication sociologist Everett Rogers (1962, 1971,
1983) has emphasized the role of information, communication, formal and
informal media, opinion leaders and social networks, and economic and
psychological constraints on acceptance. Rogers's work stresses the
decision mechanism of the potential adopter; the work of Rogers
partitions the adopter's process of choosing to accept a new phenomenon
or trait into five stages: stage 1--the potential adopter gains
knowledge or awareness of an innovation; stage 2--persuasion is
exercised to adopt; stage 3--a decision is made to adopt; stage 4--the
decision to adopt is implemented; stage 5--the adopter confirms the
decision to adopt. In Rogers's stage theory there may be a substantial
time lag between when a potential adopter becomes aware of the new
characteristic and when a decision is actually made to adopt.

Hägerstrand considered diffusion to be a fundamental geographic process:
Whatever the phenomenon being diffused might be, one may consider it in
the context of a larger universal process of spatial diffusion.

{[}6{]} Because new digital activities are rarely---if ever---captured
in official state data, researchers must rely on information gathered
from alternative sources. With this in mind, the overarching aim of this
article is outlining an approach for analyzing the ``new spaces and
geographies'' of digital phenomena and practices.

\hypertarget{literature-review-sec2-maybe-skip-this}{%
\section{Literature review \{\#sec2\} MAYBE SKIP
THIS??}\label{literature-review-sec2-maybe-skip-this}}

\hypertarget{sec3}{%
\section{Materials and Methods}\label{sec3}}

\hypertarget{sec3.1}{%
\subsection{Multi-country analysis}\label{sec3.1}}

\hypertarget{sec3.2}{%
\subsection{3.2 Case study approach}\label{sec3.2}}

\hypertarget{sec4}{%
\section{Digital technologies and spatial structure: a global
view}\label{sec4}}

\hypertarget{sec5}{%
\section{The impact of ICT on the US and the UK spatial
structures}\label{sec5}}

\hypertarget{sec6}{%
\section{Discussion and conclusions}\label{sec6}}

contrary to results from future studies regarding social media {[}7{]},
web technologies did not exclusively spread from a central location.

{[}7{]} We see that the OSN spread almost exclusively from the original
location (the capital Budapest, with an order of magnitude more
inhabitants than the next size town) to various parts of the country in
the early phase of the life-cycle. Later, diffusion became less
mono-centric and other towns also emerged as spreaders. Our findings
support the idea that spreading initially happens to large distances and
becomes more local over time. This is illustrated and discussed later in
Figure 2I.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\hypertarget{ref-beardsell1999spatial}{}%
\CSLLeftMargin{1. }
\CSLRightInline{Beardsell M, Henderson V. Spatial evolution of the
computer industry in the USA. European Economic Review. Elsevier;
1999;43: 431--456. }

\leavevmode\hypertarget{ref-bednarz2020pulled}{}%
\CSLLeftMargin{2. }
\CSLRightInline{Bednarz M, Broekel T. Pulled or pushed? The spatial
diffusion of wind energy between local demand and supply. Industrial and
Corporate Change. Oxford University Press; 2020;29: 893--916. }

\leavevmode\hypertarget{ref-bento2015spatial}{}%
\CSLLeftMargin{3. }
\CSLRightInline{Bento N, Fontes M. Spatial diffusion and the formation
of a technological innovation system in the receiving country: The case
of wind energy in portugal. Environmental Innovation and Societal
Transitions. Elsevier; 2015;15: 158--179. }

\leavevmode\hypertarget{ref-das2022diffusion}{}%
\CSLLeftMargin{4. }
\CSLRightInline{Das B. Diffusion of innovations: Theoretical
perspectives and empirical evidence. African Journal of Science,
Technology, Innovation and Development. Taylor \& Francis; 2022;14:
94--103. }

\leavevmode\hypertarget{ref-morrill2020spatial}{}%
\CSLLeftMargin{5. }
\CSLRightInline{Morrill R, Gaile GL, Thrall GI. Spatial diffusion. Sage
Publications; 1988; }

\leavevmode\hypertarget{ref-zook2022mapping}{}%
\CSLLeftMargin{6. }
\CSLRightInline{Zook M, McCanless M. Mapping the uneven geographies of
digital phenomena: The case of blockchain. The Canadian Geographer/Le
G{é}ographe canadien. Wiley Online Library; 2022;66: 23--36. }

\leavevmode\hypertarget{ref-lengyel2020role}{}%
\CSLLeftMargin{7. }
\CSLRightInline{Lengyel B, Bokányi E, Di Clemente R, Kertész J, González
MC. The role of geography in the complex diffusion of innovations.
Scientific reports. Nature Publishing Group; 2020;10: 1--11. }

\end{CSLReferences}

\nolinenumbers

\end{document}

### 2022-05-26 presentations/gceg.Rmd
papers

---
title: 'Digital economy in the UK: an evolutionary story'
# date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
author: |
  |
  | Emmanouil Tranos
  |
  | <small>University of Bristol, Alan Turing Institute</small>
  | <small>e.tranos@bristol.ac.uk, [\@EmmanouilTranos](https://twitter.com/EmmanouilTranos), [etranos.info](https://etranos.info/)</small>
output: 
  revealjs::revealjs_presentation:
    theme: simple
    self_contained: true
    reveal_options:
      width: 1150
      height: 720
bibliography: mybibfile.bib
biblio-style: "apalike"
css: style.css
---

```{r, echo=FALSE, results=FALSE, include=FALSE}
library(knitr)
library(rprojroot)

knitr::opts_chunk$set(echo = FALSE, message = F, warning = F) # By default, hide code; set to TRUE to see code
#knitr::opts_chunk$set(out.width = '100%', dpi=300) # Figure resolution and size

# This is the project path
path <- find_rstudio_root_file()

#bibliography: ../bibliography.bib
```

## Contents
- Introduction
- Web data and spatial research
- Empirical strategy
- Descriptive statistics
- Results
- Conclusions
<br>
<br>
[etranos.info/post/sad2021](https://etranos.info/post/sad2021)

# Introduction

## Aims

- Map the active engagement with the *digital*
- Over time, early stages of the internet
- Granular spatial level
- Importance: understand how the adoption of technologies evolves
- Predictions of introduction times for future technologies [@meade2021modelling]

  - Network operators

  - Suppliers of network equipment

  - Regulatory authorities

## Diffusion

Large scale adoption studies

Add Figure

## Spatiality of the internet using supply-side measures

- Map internet infrastructure, *both*:

  - backbone (ADD REF), and

  - last mile (ADD REF, Singleton, Hannah)

- Economic effects:

  - Kolko

  - Nathan

## ???

@beardsell1999spatial used the employment in the computer industry to assess
its evolution

# Results

---

```{r results='asis', out.width = "60%"}

library(magick)
library(purrr)
path.images <- paste0(path, "/outputs/")

list.files(path = path.images, pattern = "^hex1", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir

```

---

```{r results='asis', out.width = "60%"}

library(magick)
library(purrr)
path.images <- paste0(path, "/outputs/")

list.files(path = path.images, pattern = "^hex_log1", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir

```

```{r eval=FALSE, fig.show='animate', ffmpeg.format='gif', dev='jpeg', interval=0.01}

hex.images <- list.files(path = path.images, pattern = "^hex", full.names = T)
for (i in hex.images){
  plot(i) # for example
}
```

## References {.allowframebreaks}

### 2022-05-26 abstracts_notes/help.R

@@ -1,0 +1,1801 @@
+# from https://github.com/rsbivand/survival_katrina/blob/master/bivand%2Bgomez-rubio_SMij.R
+# linked to Bivand_Gomez-Rubio_2021.pdf
+
+### R code from vignette source 'bivand+gomez-rubio_SMij.Rnw'
+
+zz <- file("sunk_output.txt")
+sink(zz)
+sink(zz, type="message")
+
+###################################################
+### code chunk number 13: bivand+gomez-rubio_SMij.Rnw:242-254
+###################################################
+library(spatialprobit)
+data(Katrina)
+library(sf)
+sf_katrina <- st_as_sf(Katrina.raw, coords=c("long", "lat"))
+st_crs(sf_katrina) <- "+proj=longlat +datum=WGS84"
+is.na(sf_katrina$days) <- sf_katrina$days == 99999
+sf_katrina$Days <- as.Date("2005-08-29") + sf_katrina$days
+t <- table(sf_katrina$Days)
+diff_days <- diff(as.Date(names(t)))
+names(diff_days) <- names(t)[-length(t)]
+yrs <- format(as.Date(names(diff_days)), "%Y")
+# o
+
+
+###################################################
+### code chunk number 14: bivand+gomez-rubio_SMij.Rnw:265-268
+###################################################
+suppressMessages(library(xtable))
+mat <- table(yrs, diff_days)
+#print(xtable(mat, align=c("l","r","r","r","r","r","r","r"), digits=c(NA, 0, 0, 0, 0, 0, 0, 0), display=c("s", "d", "d", "d", "d", "d", "d", "d")), floating=FALSE, comment=FALSE, sanitize.text.function=function(x){x}, hline.after=c(-1,0,2))
+
+
+###################################################
+### code chunk number 15: bivand+gomez-rubio_SMij.Rnw:283-292
+###################################################
+sf_katrina$days_mod <- sf_katrina$days
+sf_katrina$days[is.na(sf_katrina$days)] <- 99999
+sf_katrina$y0_90 <- as.numeric(!sf_katrina$days < 90)
+sf_katrina$y0_180 <- as.numeric(!sf_katrina$days < 180)
+sf_katrina$y0_360 <- as.numeric(!sf_katrina$days < 360)
+sf_katrina$sliced <- (sf_katrina$y0_90 == 0) + (sf_katrina$y0_180 == 0) + (sf_katrina$y0_360 == 0) + as.numeric(sf_katrina$days < 99999)
[…]

### 2022-05-26 abstracts_notes/help_splm.R

@@ -1,0 +1,5 @@
+library(splm)
+library(Ecdat)
+
+data("Produc", package = "Ecdat")
+data("usaww")

### 2022-05-26 src/help_survival.R

@@ -1,0 +1,21 @@
+install.packages("survival")
+library(survival)
+
+head(lung)
+dim(lung)
+
+lung$status <- lung$status - 0.4
+
+cfit1 <- coxph(Surv((time+.1), status) ~ age + sex + wt.loss, data=lung)
+summary(cfit1, digits=3)
+table(lung$status)
+
+
+head(mgus2)
+
+etime <- with(mgus2, ifelse(pstat==0, futime, ptime))
+event <- with(mgus2, ifelse(pstat==0, 2*death, 1))
+event <- factor(event, 0:2, labels=c("censor", "pcm", "death"))
+table(event)
+
+mfit2 <- survfit(Surv(etime, event) ~ sex, data=mgus2)

### 2022-05-27 src/hex_heatmaps.Rmd
betas

@@
     scale_fill_gradientn(colours=c('light gray','blue'),
-                         name='Frequency',na.value=NA) + #, trans = "log10") +
+                         name='log(Frequency)',na.value=NA) + #, trans = "log10") +
     coord_equal() +

### 2022-05-27 src/level_lisa_oa.Rmd
betas

@@
 library(geosphere)
+library(broom)
 
@@
 lookup <- read_csv(path.lookup) %>% 
-  select(pcds, oa11cd) %>% 
+  dplyr::select(pcds, oa11cd) %>% 
   dplyr::rename(pc = pcds)
@@
   group_by(year, oa11cd) %>%
-  summarise(n = n()) %>%            # for multiple pc  
-  #summarise(n = sum(V1.domain)) %>% # for n == 1 
+  #summarise(n = n()) %>%            # for multiple pc  
+  summarise(n = sum(V1.domain)) %>%  # for n == 1 
   ungroup()
@@
 
-#path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_oa") # for unique PC
-path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_oa")  # for multiple PC
+path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_oa") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_oa")  # for multiple PC
 
@@
 ```{r}
+path.out <- paste0(path, "/outputs/morani.csv")
 morani %>% as_tibble() %>% 
@@
                    year = V3) %>% 
-  kable()
+  kable() #%>% 
+  #write_csv(path.out)
 
@@
 }
+
+path.out <- paste0(path, "/outputs/gini.csv")
+gini.df %>% write_csv(path.out)
 ```
@@
 dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
-dist.retail <- dist.retail %>% select(oa11cd, dist.retail)
+dist.retail <- dist.retail %>% dplyr::select(oa11cd, dist.retail)
 
@@
 names(dens.London)[1] <- "rsquared"
+
+path.out <- paste0(path,"/outputs/rsq_dist_london.png")
 ggplot(dens.London) +
-  geom_line(aes(x=year, y = rsquared))
+  geom_line(aes(x=year, y = rsquared)) +
+  ylab("Rsquared") + 
+  ggtitle("Regression fit, distance to London on website density") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  ggsave(path.out)
 
-library(broom)
 out.model <- out.model %>% 
@@
 
+path.out <- paste0(path,"/outputs/b_dist_london.png")
 ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
@@
   #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() 
+  geom_point() + coord_flip() +
+  ylab("distance coefficient") + xlab("") + 
+  ggtitle("Regression coefficients, distance to London on website density") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  ggsave(path.out)
 ```
@@
 names(dens.distMet)[1] <- "rsquared"
+
+path.out <- paste0(path,"/outputs/rsq_dist_city.png")
 ggplot(dens.distMet) +
-  geom_line(aes(x=year, y = rsquared))
+  geom_line(aes(x=year, y = rsquared)) +
+  ylab("Rsquared") + 
+  ggtitle("Regression fit, distance to nearest city on website density") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  ggsave(path.out)
 
@@
 
+path.out <- paste0(path,"/outputs/b_dist_city.png")
 ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
@@
   #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() 
+  geom_point() + coord_flip() +
+  ylab("distance coefficient") + xlab("") + 
+  ggtitle("Regression coefficients, distance to nearest city on website density") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  ggsave(path.out)
 ```
@@
 names(dens.dist.retail)[1] <- "rsquared"
+
+path.out <- paste0(path,"/outputs/rsq_dist_centre.png")
 ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared))
+  geom_line(aes(x=year, y = rsquared)) +
+  ylab("Rsquared") + 
+  ggtitle("Regression fit, distance to nearest retail centre on website density") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  ggsave(path.out)
 
@@
 
+path.out <- paste0(path,"/outputs/b_dist_centre.png")
 ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
@@
   #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip()
+  geom_point() + coord_flip() +
+  ylab("distance coefficient") + xlab("") + 
+  ggtitle("Regression coefficients, distance to nearest city on website density") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  ggsave(path.out)
 
@@
 ```
+
+## The below are unfinished
 

### 2022-05-30 src/level_lisa_la.Rmd
LAs

--- 2022-04-13 src/level_lisa_oa.Rmd
@@
 lookup <- read_csv(path.lookup) %>% 
-  select(pcds, oa11cd) %>% 
+  dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
   dplyr::rename(pc = pcds)
@@
   filter(
-    V1.domain < m,             # for multiple pc
-    #V1.domain == n,             # for n == 1
+    #V1.domain < m,             # for multiple pc
+    V1.domain == n,             # for n == 1
     year > 1995 & year <2013) %>%   
   left_join(lookup, by = "pc", suffix =c("","")) %>% 
-  group_by(year, oa11cd) %>%
-  summarise(n = n()) %>%            # for multiple pc  
-  #summarise(n = sum(V1.domain)) %>% # for n == 1 
+  group_by(year, ladcd) %>%
+  #summarise(n = n()) %>%            # for multiple pc  
+  summarise(n = sum(V1.domain)) %>% # for n == 1 
   ungroup()
@@
 ```{r}
-# get OA for England and Wales
-path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
-oa.ew <- readOGR(path.geo)
+# get LA for the UK
+path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
+la <- readOGR(path.geo)
 # source: https://geoportal.statistics.gov.uk/
@@
 # spatial transformations
-oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))
+la <- spTransform(la, CRS("+init=epsg:4326"))
 
-# keep in the data slot only the ONS Output Area id, renaming it as 'id'
-oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
-colnames(oa.ew@data) <- c('id', 'area')
+la.f <- fortify(la, region = "LAD21CD")
 
-# reassign the polygon IDs
-oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))
-
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.ew)
-
-# # get OA for Scotland
-# path.geo.sc <- paste0(path, "/data/raw/SG_DataZoneBdry_2011")
-# oa.sc <- readOGR(dsn=path.geo.sc, layer = "SG_DataZone_Bdry_2011")
-# # source: https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011
-# 
-# # spatial transformations
-# oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-# 
-# # Scotland (follows same steps as EW, see notes above)
-# oa.sc <- oa.sc[, 'DataZone']
-# colnames(oa.sc@data) <- c('id')
-# 
-# # reassign the polygon IDs
-# oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-# 
-# # check the CRS has changed correctly, and the data slot has shrink to only the ID
-# summary(oa.sc)
-
-path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
-oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
-# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries
-
-# spatial transformations
-oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-
-# Scotland (follows same steps as EW, see notes above)
-oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
-colnames(oa.sc@data) <- c('id', 'area')
-
-# reassign the polygon IDs
-oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.sc)
-
-# build oa for GB
-oa.gb <- spRbind(oa.ew, oa.sc)
-rm(oa.ew, oa.sc)
-
-# oa.gb$geometry <- oa.gb$geometry %>%
-#   s2::s2_rebuild() %>%
-#   sf::st_as_sfc()
-
-# get GB
-path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
-gb <- readOGR(path.gb)
-# spatial transformations
-gb <- spTransform(gb, CRS("+init=epsg:4326"))
-
+# cities
+cities <- maps::world.cities %>% 
+  filter(country.etc=="UK") %>% 
+  arrange(pop) %>% tail(10) 
+tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
 ```
@@
 
-#path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_oa") # for unique PC
-path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_oa")  # for multiple PC
+path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_la") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_la")  # for multiple PC
 
@@
 sf_use_s2(FALSE)
-nb <- poly2nb(oa.gb)
+nb <- poly2nb(la)
 listw <- nb2listw(nb, style = "W", zero.policy = T)
@@
 for (t in 1996:2012){
-test <- oa.gb
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
+test <- la
+test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
   mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
@@
 
-mun_merge_new<- merge(test, lmoran, by="id")
+mun_merge_new<- merge(test, lmoran, by="LAD21CD")
 
@@
                                          labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
-                                         title="LISA for level, OA") +
+                                         title="LISA for level, LA") +
   tm_legend(text.size = 1) +
@@
   tm_borders(lwd = 0) +
-  tm_shape(gb, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
+  tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
   tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
@@
 ```{r}
+path.out <- paste0(path, "/outputs/morani_la.csv")
 morani %>% as_tibble() %>% 
@@
                    year = V3) %>% 
-  kable()
+  #kable() %>% 
+  write_csv(path.out)
 
@@
 }
+
+path.out <- paste0(path, "/outputs/gini_la.csv")
+gini.df %>% write_csv(path.out)
 ```
@@
 
-```{r}
+```{r eval = F}
 
@@
 
-First set up the destinations. Origins are all the OA.
+First set up the destinations. Origins are all the LA.
 
@@
 ```{r}
-# oa centroids
-# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
-# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
-#   mutate(id = row.names())
-oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
+# LA centroids
+la.c <- gCentroid(la, byid=TRUE)@coords %>% 
   as.data.frame() %>%
-  add_rownames(var = "id")
+  bind_cols(la@data) %>% 
+  dplyr::select(LAD21CD, x, y) %>% 
+  remove_rownames()
 
@@
 ```{r}
-dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist <- distm(cbind(la.c$x, la.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
 dist <- round((dist/1000),2) %>% 
@@
 
-dist <- dist %>% bind_cols(oa.gb.c$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
+dist <- dist %>% bind_cols(la.c$LAD21CD) %>% 
+  rename(ladcd = last_col()) %>% 
+  relocate(ladcd)
 
@@
 
+# calculate area
+la$SHAPE_Area <- area(la) #/ 1000000
+
 # Join with distance and area
-df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
-  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
+df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
+  left_join(la@data, by = c("ladcd" = "LAD21CD")) %>% 
+  rename(area = SHAPE_Area) %>% 
   relocate(area, .after = n) %>% 
@@
 ```{r}
-dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
+dist.retail <- distm(cbind(la.c$x, la.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
 dist.retail <- round((dist.retail/1000),2) %>% 
@@
 
-dist.retail <- dist.retail %>% bind_cols(oa.gb.c$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
+dist.retail <- dist.retail %>% bind_cols(la.c$LAD21CD) %>% 
+  rename(LAD21CD = last_col()) %>% 
+  relocate(LAD21CD)
 
@@
 dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
-dist.retail <- dist.retail %>% select(oa11cd, dist.retail)
+dist.retail <- dist.retail %>% dplyr::select(LAD21CD, dist.retail)
 
 # Join with complete panel
-df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd"))
+df <- df %>% left_join(dist.retail, by = c("ladcd" = "LAD21CD"))
 
@@
 names(dens.London)[1] <- "rsquared"
+
+path.out <- paste0(path,"/outputs/rsq_dist_london_la.png")
 ggplot(dens.London) +
-  geom_line(aes(x=year, y = rsquared))
+  geom_line(aes(x=year, y = rsquared)) +
+  ylab("Rsquared") + 
+  ggtitle("Regression fit, distance to London on website density") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  ggsave(path.out)
 
-library(broom)
 out.model <- out.model %>% 
@@
 
+path.out <- paste0(path,"/outputs/b_dist_london_la.png")
 ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
@@
 names(dens.distMet)[1] <- "rsquared"
+
+path.out <- paste0(path,"/outputs/rsq_dist_city_la.png")
 ggplot(dens.distMet) +
-  geom_line(aes(x=year, y = rsquared))
+  geom_line(aes(x=year, y = rsquared)) +
+  ylab("Rsquared") + 
+  ggtitle("Regression fit, distance to nearest city on website density") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  ggsave(path.out)
 
@@
 
+path.out <- paste0(path,"/outputs/b_dist_city_la.png")
 ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
@@
 names(dens.dist.retail)[1] <- "rsquared"
+
+path.out <- paste0(path,"/outputs/rsq_dist_centre_la.png")
 ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared))
+  geom_line(aes(x=year, y = rsquared)) +
+  ylab("Rsquared") + 
+  ggtitle("Regression fit, distance to nearest retail centre on website density") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  ggsave(path.out)
 
@@
 
+path.out <- paste0(path,"/outputs/b_dist_centre_la.png")
 ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
@@
 ```
-
-Website density regressions to the nearest retail centre with spatial lag
-
-```{r}
-library(spatialreg)
-
-model <- lagsarlm(log(density+0.1) ~ log(dist.retail+0.1), data = df[df$year==t,], 
-                  listw, tol.solve = 1e-30, zero.policy = T)
-
-dens.dist.retail <- data.frame()
-for (t in 1996:2012){
-  model <- lm(log(density) ~ log(dist.retail+0.1), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-}
-names(dens.dist.retail)[1] <- "rsquared"
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared))
-
-```
-
-## HH clusters
-
-```{r}
-hh <- mun_merge_new@data %>% filter(quadrant==1) %>% 
-  select(id, quadrant) %>% 
-  left_join(df[df$year==2000,], by =c("id" = "oa11cd")) %>% 
-  select(id, year, n, area, density, London, dist, dist.retail)
-
-hh.London <- lm(log(density+0.001) ~ log(London), data = hh)
-summary(hh.London)
-hh.dist <- lm(log(density+0.001) ~ log(dist), data = hh)
-summary(hh.dist)
-hh.retail <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = hh)
-summary(hh.retail)
-
-data.folder <- "/hdd/internet_archive/archive/data/"
-data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
-#Created by domain.R, which uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df9610 <- read_csv(data.path9610) %>% 
-  filter(V1.domain == n,
-         year == 1999) %>% 
-  left_join(lookup, by = "pc", suffix =c("",""))
-
-hh %>% left_join(df9610, by =c("id"="oa11cd"))
-
-```
-
-Spatial regression
-
-```{r}
-library(splm)
-m <- spml(log(n)~lag(log(n)), data = df, listw = listw, lag = TRUE, spatial.error = "b", 
-          model = "within", effect = "individual", method = "eigen", na.action = na.fail)
-
-```

### 2022-05-30 src/level_lisa_la.Rmd
edits

@@
 
+# No need to add 0s like OAs as there aren't any LA with 0s
+
 gini.df <- data.frame()
@@
   ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to London on website density") +
+  ggtitle("Regression fit, distance to London on website density, LA") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to London on website density") +
+  ggtitle("Regression coefficients, distance to London on website density, LA") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest city on website density") +
+  ggtitle("Regression fit, distance to nearest city on website density, LA") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density") +
+  ggtitle("Regression coefficients, distance to nearest city on website density, LA") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest retail centre on website density") +
+  ggtitle("Regression fit, distance to nearest retail centre on website density, LA") +
   theme(plot.title = element_text(hjust = 0.5)) +

### 2022-05-30 src/level_lisa_oa.Rmd
edits

@@
 ```{r}
-path.out <- paste0(path, "/outputs/morani.csv")
+path.out <- paste0(path, "/outputs/morani_oa.csv")
 morani %>% as_tibble() %>% 
@@
 
+# add OAs with 0s
+df.with0s <- df %>% mutate(country = substring(oa11cd, 1, 1)) %>%
+  filter(country != "L" & country != "M" & country != "N") %>%
+  filter(!is.na(country)) %>%
+  complete(year, oa11cd, fill = list(n = 0)) %>%
+  filter(year!=1000,
+         !is.na(oa11cd)) %>%
+  dplyr::select(-country)
+
 gini.df <- data.frame()
@@
 for (t in 1996:2012){
-    a <- df %>% filter(year==t)
+    a <- df.with0s %>% filter(year==t)
   b <- as.data.frame(gini(a$n))
@@
 
-path.out <- paste0(path, "/outputs/gini.csv")
+path.out <- paste0(path, "/outputs/gini_oa.csv")
 gini.df %>% write_csv(path.out)
@@
 # Complete panel, N = 3871903, 17 year * 227759 OA
-# df <- df %>% bind_rows(dist %>% select(oa11cd)) %>% 
+# df <- df %>% bind_rows(dist %>% dplyr::select(oa11cd)) %>%
 #   mutate(year = ifelse(is.na(year),1000, year),
@@
 #          !is.na(oa11cd)) %>%
-#   select(-country)
+#   dplyr::select(-country)
 
@@
 # length(unique(test$oa11cd))
+
+# calculate area
+oa.gb$area <- area(oa.gb) #/ 1000000
 
@@
 
-path.out <- paste0(path,"/outputs/rsq_dist_london.png")
+path.out <- paste0(path,"/outputs/rsq_dist_london_oa.png")
 ggplot(dens.London) +
@@
   ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to London on website density") +
+  ggtitle("Regression fit, distance to London on website density, OA") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
 
-path.out <- paste0(path,"/outputs/b_dist_london.png")
+path.out <- paste0(path,"/outputs/b_dist_london_oa.png")
 ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
@@
   ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to London on website density") +
+  ggtitle("Regression coefficients, distance to London on website density, OA") +
   theme(plot.title = element_text(hjust = 0.5)) +
   ggsave(path.out)
+
+r2.all <- dens.London %>% 
+  rename(London = rsquared)
 ```
@@
 
-path.out <- paste0(path,"/outputs/rsq_dist_city.png")
+path.out <- paste0(path,"/outputs/rsq_dist_city_oa.png")
 ggplot(dens.distMet) +
@@
   ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest city on website density") +
+  ggtitle("Regression fit, distance to nearest city on website density, OA") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
 
-path.out <- paste0(path,"/outputs/b_dist_city.png")
+path.out <- paste0(path,"/outputs/b_dist_city_oa.png")
 ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
@@
   ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density") +
+  ggtitle("Regression coefficients, distance to nearest city on website density, OA") +
   theme(plot.title = element_text(hjust = 0.5)) +
   ggsave(path.out)
+
+r2.all <- r2.all %>% left_join(dens.distMet) %>% 
+  rename(cities = rsquared)
 ```
@@
 
-path.out <- paste0(path,"/outputs/rsq_dist_centre.png")
+path.out <- paste0(path,"/outputs/rsq_dist_centre_oa.png")
 ggplot(dens.dist.retail) +
@@
   ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest retail centre on website density") +
+  ggtitle("Regression fit, distance to nearest retail centre on website density, OA") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
 
-path.out <- paste0(path,"/outputs/b_dist_centre.png")
+path.out <- paste0(path,"/outputs/b_dist_centre_oa.png")
 ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
@@
   ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density") +
+  ggtitle("Regression coefficients, distance to nearest city on website density, OA") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
 #   summarise(n = sum(n))
+```
+All R^2 dataframe
+
+```{r}
+path.out <- paste0(path, "/outputs/r2_all.csv")
+
+r2.all <- dens.London %>% 
+  rename(London = rsquared) %>% 
+  left_join(dens.distMet) %>% 
+  rename(cities = rsquared) %>% 
+  left_join(dens.dist.retail) %>% 
+  rename(retail = rsquared) %>% 
+  relocate(year, .before = London) %>% 
+  arrange(year)
+
+write_csv(r2.all, path.out)
 ```

### 2022-06-01 src/level_lisa_la.Rmd
updates

@@
 library(broom)
+library(foreach)
 
@@
   tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
+tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
+}
+```
+
+## lisa maps parallel
+
+```{r}
+
+path.out <- paste0(path, "/outputs/lisa_level_pc_test", n, "_la") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_la")  # for multiple PC
+
+#setup cluster
+#https://www.blasbenito.com/post/02_parallelizing_loops_with_r/
+n.cores <- parallel::detectCores() - 1
+my.cluster <- parallel::makeCluster(
+  n.cores, 
+  type = "PSOCK"
+  )
+print(my.cluster)
+doParallel::registerDoParallel(cl = my.cluster)
+foreach::getDoParWorkers()
+
+morani <- data.frame()
+
+#spatial weight matrix
+sf_use_s2(FALSE)
+nb <- poly2nb(la)
+listw <- nb2listw(nb, style = "W", zero.policy = T)
+
+#for (t in 1996:2012){
+foreach(t = 1996:2012, .packages = c("tidyverse",
+                                     "rgdal",
+                                     "rgeos",
+                                     "spdep",
+                                     "tmap",
+                                     "maptools",
+                                     "knitr",
+                                     "broom")) %dopar% {
+test <- la
+test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
+  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
+         year = ifelse(is.na(year), t, year))
+
+#spatial weight matrix
+#moved out of the loop
+
+#Moran's I
+globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
+a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
+morani <- rbind(a, morani)
+
+#LISA
+lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
+#lmoran
+
+# centers the local Moran's around the mean
+lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
+lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
+
+# centers the variable of interest around its mean
+
+lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
+lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
+
+signif <- 0.05
+#lmoran
+
+lmoran <- lmoran%>% 
+  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
+  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
+  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
+  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
+  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
+  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%
+
+#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
+# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
+#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
+#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
+#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))
+
+mun_merge_new<- merge(test, lmoran, by="LAD21CD")
+
+# R p value map
+breaks = c(0, 1, 2, 3, 4, 5)
+
+cities <- maps::world.cities %>% 
+  filter(country.etc=="UK") %>% 
+  arrange(pop) %>% tail(10) 
+tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
+
+map <- tm_shape(mun_merge_new, projection = "+init=epsg:4326") + tm_fill(col = "quadrant", breaks = breaks, 
+                                         palette = c("white","red","blue",
+                                                     rgb(0,0,1,alpha=0.4),
+                                                     rgb(1,0,0,alpha=0.4)), 
+                                         labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
+                                         title="LISA for level, LA") +
+  tm_legend(text.size = 1) +
+  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
+  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
+  #tm_borders(alpha=.1) +
+  tm_borders(lwd = 0) +
+  tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
+  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
+  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ") +
+tmap_options(check.and.fix = TRUE)
 tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
@@
 ```
+
+All R^2 dataframe
+
+```{r}
+path.out <- paste0(path, "/outputs/r2_all_la.csv")
+
+r2.all <- dens.London %>% 
+  rename(London.LA = rsquared) %>% 
+  left_join(dens.distMet) %>% 
+  rename(cities.LA = rsquared) %>% 
+  left_join(dens.dist.retail) %>% 
+  rename(retail.LA = rsquared) %>% 
+  relocate(year, .before = London.LA) %>% 
+  arrange(year)
+
+write_csv(r2.all, path.out)
+```

### 2022-06-05 src/s_la.Rmd
edits

@@
 library(modelr)
+library(survival)
 
@@
 lookup <- read_csv(path.lookup) %>% 
-  select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
+  dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
   dplyr::rename(pc = pcds)
@@
   group_by(year, ladcd) %>%
-  summarise(n = n()) %>%            # for multiple pc  
-  #summarise(n = sum(V1.domain)) %>% # for n == 1 
+  #summarise(n = n()) %>%            # for multiple pc  
+  summarise(n = sum(V1.domain)) %>% # for n == 1 
   ungroup()
@@
 #               year==2003)
+```
+
+## Spatial data
+
+```{r}
+# get LA for the UK
+path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
+la <- readOGR(path.geo)
+# source: https://geoportal.statistics.gov.uk/
+
+# spatial transformations
+la <- spTransform(la, CRS("+init=epsg:4326"))
+
+la.f <- fortify(la, region = "LAD21CD")
+
+# cities
+cities <- maps::world.cities %>% 
+  filter(country.etc=="UK") %>% 
+  arrange(pop) %>% tail(10) 
+tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
 ```
@@
 
+path.out <- paste0(path, "/outputs/s_uk.png")
+
 df.country %>% 
@@
   ylab("N. of websites") +
-  ggtitle("Country")
+  ggtitle("S-curve, UK") +
+  theme(plot.title = element_text(hjust = 0.5))
+
+ggsave(path.out)
+
   #scale_x_date(breaks = df.test$year) 
@@
   data <- df %>% filter(ladcd==i) %>%
-    select(year, ladcd, n) 
+    dplyr::select(year, ladcd, n) 
     #lm(n~year, data = .)
@@
   mutate(estimate = round(estimate, 0)) %>% 
-  select(estimate) %>% 
+  dplyr::select(estimate) %>% 
   summarise(r = range(estimate),
@@
   filter(p.value < 0.01) %>%    # (227759 - 212782)/227759  # 6% not sig 
-  select(estimate) %>% 
+  dplyr::select(estimate) %>% 
   ggplot(aes(x=estimate)) + 
   geom_histogram(binwidth = 1) +
-  geom_vline(xintercept = 2003, colour = "red")
-  scale_x_continuous("year", labels = round(a$estimate, 0), breaks = round(a$estimate, 0)) 
+  geom_vline(xintercept = 2003, colour = "red") +
+  #scale_x_continuous("year", labels = round(a$estimate, 0), breaks = round(a$estimate, 0)) +
   #scale_x_discrete(labels = as.factor(a$estimate))
+  theme_minimal() +
+  ggtitle("Counts of Xmid per year") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  ylab("LAs") + xlab("year")
+
+path.out <- paste0(path, "/outputs/xmid_f.png")
+ggsave(path.out)
 
@@
   filter(p.value < 0.01) %>%    # 376 / 376, all sig
-  select(ladcd, estimate)
+  dplyr::select(ladcd, estimate)
   #filter(estimate>2011)
@@
          r2 > 0.9) %>%    # 376 / 376, all sig, 330 r2 > 0.9
-  select(ladcd, estimate, r2, std.error) %>%
+  dplyr::select(ladcd, estimate, r2, std.error) %>%
   mutate(fast = ifelse(estimate < 2003, "fast", "slow"))
@@
   geom_line(aes(y = prednls), size = 1) +
-  scale_x_continuous("year", labels = as.character(df.test$year), breaks = df.test$year) +
-  geom_vline(xintercept = pred %>% left_join(a.la) %>% filter(ladcd == i) %>% select(estimate) %>% first()) +
+  scale_x_continuous("year", labels = as.character(df$year), breaks = df$year) +
+  geom_vline(xintercept = pred %>% left_join(a.la) %>% filter(ladcd == i) %>% dplyr::select(estimate) %>% first()) +
   ylab("N. of websites") + 
-  labs(title = paste0(la@data %>% filter(LAD21CD== i) %>% select(LAD21NM) %>% first(), ", ", i)) +
+  labs(color = "Website \ndiffusion speed") +
+  labs(title = paste0(la@data %>% filter(LAD21CD== i) %>% dplyr::select(LAD21NM) %>% first(), ", ", i)) +
   theme(legend.position=c(.9,.15)) +
   scale_color_manual(values = c("fast" = "purple",
-                                "slow"="orange"))
+                                "slow"="orange")) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  geom_vline(xintercept = 2003, colour="grey30", linetype="dashed")
 
@@
 
-```{r}
-ggplot(a.la, aes(as.factor(round(estimate, 0)))) +
-  geom_bar(fill = "#0073C2FF")
+**TO DELETE**, done it above.
+
+```{r, eval=F}
+# path.out <- paste0(path, "/outputs/xmid_f.png")
+# 
+# ggplot(a.la, aes(as.factor(round(estimate, 0)))) +
+#   geom_bar(fill = "#0073C2FF") +
+#   theme_minimal() +
+#   ggtitle("Counts of Xmid per year") +
+#   theme(plot.title = element_text(hjust = 0.5)) +
+#   ylab("LAs") + xlab("year") 
+# 
+# ggsave(path.out)
 ```
@@
 
+using `library(survival)`
+
 ```{r}
-library(survival)
-
 test.surv.data <- a.la %>% mutate(status = 1) %>% 
@@
 ## Map the xmid
-
-Spatial data
-
-```{r}
-# get LA for the UK
-path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-la <- readOGR(path.geo)
-# source: https://geoportal.statistics.gov.uk/
-
-# spatial transformations
-la <- spTransform(la, CRS("+init=epsg:4326"))
-
-la.f <- fortify(la, region = "LAD21CD")
-
-# cities
-cities <- maps::world.cities %>% 
-  filter(country.etc=="UK") %>% 
-  arrange(pop) %>% tail(10) 
-tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
-```
 
@@
 
+path.out <- paste0(path, "/outputs/speed_pam.png")
 ggplot(data = la.f, aes(x = long, y = lat)) +
   geom_polygon(aes(group = group, fill = fast)) + #fill = fast, 
+  scale_fill_manual(values = c("darkorchid3", "orange"), #darkmagenta, purple4
+                    na.value = "grey70") +
+
   #theme_nothing(legend = TRUE) +
-  labs(title = "Clusters") +
+  labs(title = "Website diffusion speed in Local authorities") +
   geom_point(data = cities, aes(x=long, y=lat), colour = "black", size = .1) + 
@@
   #scale_fill_continuous(type = "viridis") +
-  theme_void()
-
+  labs(fill='Diffusion') +
+  theme_void() +
+  theme(plot.title = element_text(hjust = 0.5)) 
+ggsave(path.out)
 ```

### 2022-06-05 src/s_oa.Rmd
edits

@@
 ---
+
+I re-run the below using a `foreach` loop. It created a new object `a_` with c. 40k
+rows. Once I filter the old `a` object derived from the `for` loop I have c. 60k rows.
+I run `lm' and `coxph` using both and the results are not strong not easy to interpret.
+
+The `/data/temp/s_oa.csv` is based on the `for` loop.
 
@@
 ```{r}
-df.test <- df %>% filter(oa11cd == "S00103828")
+df.test <- df %>% filter(oa11cd == "E00138496")
 
@@
 ```{r}
-a <- as_tibble()
-all.oa <- unique(df$oa11cd)
-for (oa in all.oa){
+# a <- as_tibble()
+# all.oa <- unique(df$oa11cd)
+# #for (oa in all.oa){
+# a <- foreach(oa = all.oa, .combine = 'rbind',
+#              #.final =  function(x) setNames(x, names(all.oa)),
+#           .packages = c("tidyverse",
+#                                     "broom")) %dopar% {
+#   data <- df %>% filter(oa11cd==oa) %>%
+#     dplyr::select(year, oa11cd, n) 
+#     #lm(n~year, data = .)
+#     #tryCatch(tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .)), , error=function(e) NULL)
+#   tryCatch(tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = data)), error=function(e) NULL)
+#   #tryCatch(a <-rbind(a,b), error=function(e) NULL)
+# }
+
+# Setup cluster
+# https://www.blasbenito.com/post/02_parallelizing_loops_with_r/
+n.cores <- parallel::detectCores() - 1
+my.cluster <- parallel::makeCluster(
+  n.cores, 
+  type = "PSOCK"
+  )
+print(my.cluster)
+doParallel::registerDoParallel(cl = my.cluster)
+foreach::getDoParWorkers()
+
+a <- foreach(oa = all.oa, .combine = 'rbind', .packages = c("tidyverse",
+                           "broom")) %dopar% 
+  {
     data <- df %>% filter(oa11cd==oa) %>%
-    select(year, oa11cd, n) 
-    #lm(n~year, data = .)
-    #tryCatch(tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .)), , error=function(e) NULL)
-  tryCatch(b <- tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = data)), error=function(e) NULL)
-  tryCatch(a <-rbind(a,b), error=function(e) NULL)
+    dplyr::select(year, oa11cd, n) 
+    tryCatch(model <- tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = data)), error=function(e) NULL)
+
+    tryCatch(model <- model %>% dplyr::filter(term == "xmid"), error=function(e) NULL)
+
+    c(oa, model)
     }
 
-a %>% filter(term=="xmid") %>% 
+stopCluster(my.cluster)
+
+# drop the second duplicates  
+a_ <- as.data.frame(a) %>% 
+  rename(oa11cd = V1) %>% 
+  #mutate(help = duplicated(estimate))
+  distinct(estimate, std.error, statistic, p.value, .keep_all = TRUE) %>% 
+  unnest()
+
+a_ %>% #filter(term=="xmid") %>% 
   mutate(estimate = round(estimate, 0)) %>% 
-  select(estimate) %>% 
+  dplyr::select(estimate) %>% 
   summarise(r = range(estimate))
 
-a %>% filter(term=="xmid") %>% 
+a_ %>% #filter(term=="xmid") %>% 
   mutate(estimate = round(estimate, 0)) %>% 
   filter(p.value < 0.01) %>%    # (227759 - 212782)/227759  # 6% not sig 
-  select(estimate) %>% 
+  dplyr::select(estimate) %>% 
   ggplot(aes(x=estimate)) + 
@@
   #scale_x_discrete(labels = as.factor(a$estimate))
+
+a_ %>% filter(estimate<1996)
 
@@
 ```{r}
-path.out <- paste0(path, "/data/temp/s.csv")
+path.out <- paste0(path, "/data/temp/s_oa.csv")
 write_csv(a, path.out)
@@
 
-oa.gb.f <- oa.gb.f %>% left_join(a, by = "id") %>% 
+oa.gb.f <- oa.gb.f %>% left_join(a_, by = c("id"="oa11cd")) %>% 
   mutate(fast = ifelse(estimate < 2003, "fast", "slow")) %>% 
@@
 #   arrange(order) %>% 
-ggplot(data = oa.gb.f %>% arrange(order), aes(x = long, y = lat)) +
+ggplot(data = oa.gb.f %>% arrange(order) %>% filter(estimate<2020 & estimate >1996), aes(x = long, y = lat)) +
   geom_polygon(aes(group = group, fill = estimate)) + #fill = fast, 
@@
 
-oa.class.df <- a %>%  
+oa.class.df <- a_ %>%  
   mutate(fast = ifelse(estimate < 2003, "fast", "slow")) %>% 
-  left_join(oa.class@data, by = c("id"="OA_SA")) %>% 
-  select(1:10) %>% 
-  arrange(id)
+  left_join(oa.class@data, by = c("oa11cd"="OA_SA")) %>% 
+  dplyr::select(1:14) %>% 
+  arrange(oa11cd)
 
@@
 oa.class.df %>% 
-  filter(fast == "slow") %>% 
+  filter(fast == "fast") %>% 
   group_by(SUBGRP) %>% 
@@
          cum.freq = round(cum.freq, 2)) %>% 
-  left_join(pen %>% select(code, name),
+  left_join(pen %>% dplyr::select(code, name),
             by = c("SUBGRP" = "code")) %>% 
@@
 # Join with distance and area
-oa.class.df <- oa.class.df %>% left_join(dist, by = c("id" = "oa11cd")) 
+oa.class.df <- oa.class.df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) 
 sapply(df, function(x) sum(is.na(x)))
@@
 dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
-dist.retail <- dist.retail %>% select(oa11cd, dist.retail)
+dist.retail <- dist.retail %>% dplyr::select(oa11cd, dist.retail)
 
 # Join with complete panel
-oa.class.df <- oa.class.df %>% left_join(dist.retail, by = c("id" = "oa11cd"))
+oa.class.df <- oa.class.df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd"))
 
@@
 oa.class.df <- oa.class.df %>% left_join(df %>% group_by(oa11cd) %>% summarise(max.n = max(n, na.rm = T)), #filter(year==2012), 
-                          by = c("id"="oa11cd"))
+                          by = c("oa11cd"="oa11cd"))
 
@@
 model <- lm(estimate ~ log(London + 1) + log(dist + 1) + log(dist.retail + 1) + 
-              as.factor(SPRGRP) + log(POPULATION) + max.n, data = oa.class.df) #%>%filter(estimate>1996 & estimate <2012)
+              as.factor(SPRGRP) + log(POPULATION) + max.n, data = oa.class.df %>% filter(estimate>1996 & estimate <2012))
 summary(model)
 nobs(model)
+
+oa.class.df <- oa.class.df %>% mutate(status = 1)
+
+cf <- coxph(Surv((estimate-1996), status) ~ log(London + 1) + log(dist + 1) + log(dist.retail + 1) + 
+              as.factor(SPRGRP) + log(POPULATION) + max.n, data = oa.class.df %>% filter(estimate>1996 & estimate <2012)) #, weights = 1/(std.error)^2)
+summary(cf)
 

### 2022-06-06 src/level_lisa_la.Rmd
edits

@@
   ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density, LA") +
+  ggtitle("Regression coefficients, distance to nearest retail centre on website density, LA") +
   theme(plot.title = element_text(hjust = 0.5)) +

### 2022-06-06 src/level_lisa_oa.Rmd
edits

@@
   ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density, OA") +
+  ggtitle("Regression coefficients, distance to nearest retail centre on website density, OA") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
 ```{r}
-path.out <- paste0(path, "/outputs/r2_all.csv")
+path.out <- paste0(path, "/outputs/r2_all_oa.csv")
 

### 2023-02-16 src/s_la_per_firm.Rmd
per firm

--- 2022-06-05 src/s_la.Rmd
@@
 
+## Firms
+
+```{r}
+firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>%
+  select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
+  rename(ladcd = GEOGRAPHY_CODE,
+         firms = OBS_VALUE)
+
+df <- df %>% left_join(firms) %>% 
+  mutate(n_by_firm = (n*1000)/firms) %>% # '000 websites/firm
+  #filter(!is.na(firms)) %>% 
+  select(-n, -firms) %>% 
+  rename(n = n_by_firm)
+
+# df %>% filter(is.na(n_by_firm))
+# sapply(df %>% left_join(firms), function(x) sum(is.na(x)))
+```
+
 ## Spatial data
@@
 df.country <- df %>% group_by(year) %>% 
-  summarise(n = sum(n))
+  summarise(n = sum(n, na.rm = T))
 
@@
 ggsave(path.out)
-
   #scale_x_date(breaks = df.test$year) 
@@
 
-path.out <- paste0(path, "/outputs/xmid_f.png")
+path.out <- paste0(path, "/outputs/per_firm/xmid_f.png")
 ggsave(path.out)
@@
 # plots
-path.out <- paste0(path, "/outputs/lad_")
+path.out <- paste0(path, "/outputs/per_firm/lad_")
 

### 2023-02-16 src/test_gambling.Rmd
per firm

--- 2022-05-27 src/level_lisa_oa.Rmd
@@
 library(rprojroot)
-library(rgdal)
-library(rgeos)
-library(spdep)
-library(tmap)
-library(maptools)
-library(knitr)
-library(REAT)
-library(tidygeocoder)
-library(geosphere)
-library(broom)
+library(urltools)
 
@@
 path <- find_rstudio_root_file()
+```
+
+all
+
+```{r}
+data.folder <- "/hdd/internet_archive/archive/data/"
+data.path.all <- paste0(data.folder, "all.csv")
+all <- read_csv(data.path.all)
+
+g <- all %>% filter(year==2010) %>% 
+  filter(grepl('gambl', url)) #
+
+#g %>% filter(year==2010)
+
+#all %>% filter(grepl("^www.", host))
+```
+## hyperlinks
+
+zgrep "gambleaware" host-linkage.tsv.gz | sort > gambleaware.tsv
+
+```{r}
+links.path <- "/home/nw19521/Downloads/gambleaware.tsv"
+
+l <- read_delim(links.path, delim = "|", col_names = F) %>% 
+  separate(X3, c('j', 'links'), sep = "\t") %>% 
+  rename(year = X1,
+         i = X2)
+```
+gamble is obtained by:
+zgrep "gamble" host-linkage.tsv.gz | sort > gamble.tsv
+
+It contains the hyperlink between domains that include the word "gamble".
+There are 47795 obs (`dim(l)`) and 46293 unique hyperlinks (`length(unique(l$ij))`).
+
+**Key message** the links are not repeated over time, only 47795 - 46293 = 1202.
+2004 is the link-farm problem. 2008-2010 the number of links drops.
+
+After i do the i join, there 935378 obs and 35751 postcode NA.
+
+```{r}
+links.path <- "/home/nw19521/Downloads/gamble.tsv"
+
+l <- read_delim(links.path, delim = "|", col_names = F) %>% 
+  separate(X3, c('j', 'links'), sep = "\t") %>% 
+  rename(year = X1,
+         i = X2) %>% 
+  mutate(year = as.double(year),
+         ij = paste(i, "_", j))
+
+length(unique(l$ij))
+
+l %>% group_by(year) %>% 
+  summarise(n=n())
+
+l %>% filter(year == 2010)
+
+```
+
+In 2010 492,575 rows. 7,868 without NA -- that is both i and j are geolocated --
+only 11 distinct ij
+
+In 2002 28,729,384 rows.  28,257,752 without NA -- that is both i and j are geolocated --
+only 39 distinct ij
+
+```{r}
+l.pc <- l %>% 
+  filter(year == 2002) %>% 
+  left_join(all, by = c("i" = "host", "year" = "year")) %>% 
+  left_join(all, by = c("j" = "host", "year" = "year"))
+```
+
+```{r}
+l.pc %>% drop_na() %>% distinct(ij)
+```
+
+```{r}
+length(unique(l.pc$ij))
+```
+
+```{r}
+sapply(l.pc, function(x) sum(is.na(x)))
+```
+
+```{r}
+l %>% 
+  #filter(year == 2010) %>% 
+  mutate(i.host = host_extract(i))
+
+host_extract(l$i)
+domain(l$i)
 ```
@@
 ```
-
-## spatial data
-
-```{r}
-# get OA for England and Wales
-path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
-oa.ew <- readOGR(path.geo)
-# source: https://geoportal.statistics.gov.uk/
-
-# spatial transformations
-oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))
-
-# keep in the data slot only the ONS Output Area id, renaming it as 'id'
-oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
-colnames(oa.ew@data) <- c('id', 'area')
-
-# reassign the polygon IDs
-oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))
-
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.ew)
-
-# # get OA for Scotland
-# path.geo.sc <- paste0(path, "/data/raw/SG_DataZoneBdry_2011")
-# oa.sc <- readOGR(dsn=path.geo.sc, layer = "SG_DataZone_Bdry_2011")
-# # source: https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011
-# 
-# # spatial transformations
-# oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-# 
-# # Scotland (follows same steps as EW, see notes above)
-# oa.sc <- oa.sc[, 'DataZone']
-# colnames(oa.sc@data) <- c('id')
-# 
-# # reassign the polygon IDs
-# oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-# 
-# # check the CRS has changed correctly, and the data slot has shrink to only the ID
-# summary(oa.sc)
-
-path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
-oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
-# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries
-
-# spatial transformations
-oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-
-# Scotland (follows same steps as EW, see notes above)
-oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
-colnames(oa.sc@data) <- c('id', 'area')
-
-# reassign the polygon IDs
-oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.sc)
-
-# build oa for GB
-oa.gb <- spRbind(oa.ew, oa.sc)
-rm(oa.ew, oa.sc)
-
-# oa.gb$geometry <- oa.gb$geometry %>%
-#   s2::s2_rebuild() %>%
-#   sf::st_as_sfc()
-
-# get GB
-path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
-gb <- readOGR(path.gb)
-# spatial transformations
-gb <- spTransform(gb, CRS("+init=epsg:4326"))
-
-```
-
-## lisa maps
-
-```{r}
-
-path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_oa") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_oa")  # for multiple PC
-
-morani <- data.frame()
-
-#spatial weight matrix
-sf_use_s2(FALSE)
-nb <- poly2nb(oa.gb)
-listw <- nb2listw(nb, style = "W", zero.policy = T)
-
-for (t in 1996:2012){
-test <- oa.gb
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
-  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
-         year = ifelse(is.na(year), t, year))
-
-#spatial weight matrix
-#moved out of the loop
-
-#Moran's I
-globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
-a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
-morani <- rbind(a, morani)
-
-#LISA
-lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
-#lmoran
-
-# centers the local Moran's around the mean
-lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
-lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
-
-# centers the variable of interest around its mean
-
-lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
-lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
-
-signif <- 0.05
-#lmoran
-
-lmoran <- lmoran%>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
-  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
-  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%
-
-#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
-# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
-#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))
-
-mun_merge_new<- merge(test, lmoran, by="id")
-
-# R p value map
-breaks = c(0, 1, 2, 3, 4, 5)
-
-cities <- maps::world.cities %>% 
-  filter(country.etc=="UK") %>% 
-  arrange(pop) %>% tail(10) 
-tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
-
-map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, 
-                                         palette = c("white","red","blue",
-                                                     rgb(0,0,1,alpha=0.4),
-                                                     rgb(1,0,0,alpha=0.4)), 
-                                         labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
-                                         title="LISA for level, OA") +
-  tm_legend(text.size = 1) +
-  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
-  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
-  #tm_borders(alpha=.1) +
-  tm_borders(lwd = 0) +
-  tm_shape(gb, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
-  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
-  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
-tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
-}
-```
-
-Moran I
-
-```{r}
-path.out <- paste0(path, "/outputs/morani.csv")
-morani %>% as_tibble() %>% 
-  rename('morani'=V1,
-                   p = V2,
-                   year = V3) %>% 
-  kable() #%>% 
-  #write_csv(path.out)
-
-```
-
-## Gini
-
-```{r}
-
-gini.df <- data.frame()
-
-for (t in 1996:2012){
-    a <- df %>% filter(year==t)
-  b <- as.data.frame(gini(a$n))
-  b$year = t
-  gini.df <- rbind(gini.df, b)
-}
-
-path.out <- paste0(path, "/outputs/gini.csv")
-gini.df %>% write_csv(path.out)
-```
-
-## Getis-Ord
-
-```{r}
-
-t= 1998
-local_g <- localG(OA.Census$Qualification, nb_lw)
-local_g <- localG(df[df$year==t,]$n, listw)
-
-local_g <- cbind(OA.Census, as.matrix(local_g))
-names(local_g)[6] <- "gstat"
-
-```
-
-## Distances to cities and retail centres
-
-First set up the destinations. Origins are all the OA.
-
-Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
-Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)
-
-```{r}
-# oa centroids
-# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
-# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
-#   mutate(id = row.names())
-oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id")
-
-# cities
-city.names <- c("London, UK", 
-"Birmingham, UK",
-"Glasgow, UK",
-"Liverpool, UK",
-"Bristol, UK",
-"Manchester, UK",
-"Sheffield, UK",
-"Leeds, UK",
-"Edinburgh, UK",
-"Leicester, UK")
-
-cities <- geo(city.names, no_query = F, method = "arcgis")
-
-# retail centres
-geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
-retail <-readOGR(geo.path) 
-#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
-retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id") #%>%
-
-retail.major.cetres <- retail.major.cetres.help %>% 
-  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
-  st_transform(4326) %>%
-  st_coordinates() %>%
-  as_tibble() %>% 
-  bind_cols(retail.major.cetres.help$id) %>% 
-  rename(id = '...3') %>% 
-  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
-```
-
-Distance to cities
-
-```{r}
-dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist <- round((dist/1000),2) %>% 
-  as_tibble()
-
-city.names <- city.names %>% stringr::str_remove(", UK")
-names(dist) <- city.names
-
-dist <- dist %>% bind_cols(oa.gb.c$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
-# dist$distMet <- apply(dist[,2:11], 1, min)
-
-dist <- transform(dist, dist = do.call(pmin, dist[-1]))
-
-# NOT USED AS IT SCREWS UP THE REGRESSIONS
-# Complete panel, N = 3871903, 17 year * 227759 OA
-# df <- df %>% bind_rows(dist %>% select(oa11cd)) %>% 
-#   mutate(year = ifelse(is.na(year),1000, year),
-#          country = substring(oa11cd, 1, 1)) %>% 
-#   filter(country != "L" & country != "M" & country != "N") %>% 
-#   filter(!is.na(country)) %>% 
-#   complete(year, oa11cd, fill = list(n = 0)) %>% 
-#   filter(year!=1000,
-#          !is.na(oa11cd)) %>% 
-#   select(-country)
-
-# tests
-# test$country <- substring(test$oa11cd, 1, 1)
-# unique(test$country)
-# length(unique(test$oa11cd))
-
-# Join with distance and area
-df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
-  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
-  relocate(area, .after = n) %>% 
-  mutate(area = area / 1000000,
-         density = n / area) %>% 
-  relocate(density, .after = area)
-sapply(df, function(x) sum(is.na(x)))
-```
-
-Distance to retail centres
-
-```{r}
-dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail <- round((dist.retail/1000),2) %>% 
-  as_tibble()
-
-# retail.names <- retail.major.cetres$RC_Name
-# names(dist.retail) <- retail.names
-
-dist.retail <- dist.retail %>% bind_cols(oa.gb.c$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# Minimum distance 
-dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
-dist.retail <- dist.retail %>% dplyr::select(oa11cd, dist.retail)
-
-# Join with complete panel
-df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd"))
-
-sapply(df, function(x) sum(is.na(x)))
-```
-
-Website density regressions to London
-
-```{r}
-dens.London <- data.frame()
-out.model <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(London), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.London <- rbind(r2, dens.London)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model <- rbind(out.model_, out.model)
-}
-
-names(dens.London)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_london.png")
-ggplot(dens.London) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to London on website density") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  ggsave(path.out)
-
-out.model <- out.model %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_london.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-#ggplot(out.model, aes(x=year, y = estimate)) +  
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to London on website density") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  ggsave(path.out)
-```
-
-Website density regressions to the nearest city
-
-```{r}
-dens.distMet <- data.frame()
-out.modelMet <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.distMet <- rbind(r2, dens.distMet)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.modelMet <- rbind(out.model_, out.modelMet)
-}
-
-names(dens.distMet)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_city.png")
-ggplot(dens.distMet) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest city on website density") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  ggsave(path.out)
-
-out.modelMet <- out.modelMet %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_city.png")
-ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  ggsave(path.out)
-```
-
-Website density regressions to the nearest retail centre
-
-```{r}
-dens.dist.retail <- data.frame()
-out.model.retail <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model.retail <- rbind(out.model_, out.model.retail)
-}
-
-names(dens.dist.retail)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_centre.png")
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest retail centre on website density") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  ggsave(path.out)
-
-out.model.retail <- out.model.retail %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_centre.png")
-ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  ggsave(path.out)
-
-# library(pscl)
-# m1 <- zeroinfl((n) ~ log(dist.retail+0.001),
-#   data = df[df$year==t,], dist = "negbin", EM = TRUE)
-# summary(m1)
-# 
-# df %>% filter(year == 1996) %>% 
-#   group_by(group = cut(n, breaks = c(0, 1, 2, 10, 100, 1000, 10000), include.lowest = T, right = F)) %>%  #seq(0, max(n), 100))) %>%
-#   summarise(n = sum(n))
-```
-
-## The below are unfinished
-
-Website density regressions to the nearest retail centre with spatial lag
-
-```{r}
-library(spatialreg)
-
-model <- lagsarlm(log(density+0.1) ~ log(dist.retail+0.1), data = df[df$year==t,], 
-                  listw, tol.solve = 1e-30, zero.policy = T)
-
-dens.dist.retail <- data.frame()
-for (t in 1996:2012){
-  model <- lm(log(density) ~ log(dist.retail+0.1), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-}
-names(dens.dist.retail)[1] <- "rsquared"
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared))
-
-```
-
-## HH clusters
-
-```{r}
-hh <- mun_merge_new@data %>% filter(quadrant==1) %>% 
-  select(id, quadrant) %>% 
-  left_join(df[df$year==2000,], by =c("id" = "oa11cd")) %>% 
-  select(id, year, n, area, density, London, dist, dist.retail)
-
-hh.London <- lm(log(density+0.001) ~ log(London), data = hh)
-summary(hh.London)
-hh.dist <- lm(log(density+0.001) ~ log(dist), data = hh)
-summary(hh.dist)
-hh.retail <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = hh)
-summary(hh.retail)
-
-data.folder <- "/hdd/internet_archive/archive/data/"
-data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
-#Created by domain.R, which uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df9610 <- read_csv(data.path9610) %>% 
-  filter(V1.domain == n,
-         year == 1999) %>% 
-  left_join(lookup, by = "pc", suffix =c("",""))
-
-hh %>% left_join(df9610, by =c("id"="oa11cd"))
-
-```
-
-Spatial regression
-
-```{r}
-library(splm)
-m <- spml(log(n)~lag(log(n)), data = df, listw = listw, lag = TRUE, spatial.error = "b", 
-          model = "within", effect = "individual", method = "eigen", na.action = na.fail)
-
-```

### 2023-02-16 paper/paper.Rmd
presentation

@@
   blah blah
-
 bibliography: mybibfile.bib
@@
 - \usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
@@
 
-@beardsell1999spatial
-To characterize evolution of the computer industry, we examine the distribution
-of relative employment across cities in 1977 and how that distribution
-changes over time to 1992
+@beardsell1999spatial To characterize evolution of the computer
+industry, we examine the distribution of relative employment across
+cities in 1977 and how that distribution changes over time to 1992
 
-There is no
-tendency of relative size distributions of urban computer employment to collapse,
-go bimodal, or fully spread. Overall computers exhibit some turbulence,
-with dramatic big winners and losers among cities. In attracting or repelling an
-industry, urban heterogeneity is important. Large, well educated cities near San
-Jose have much greater chances of attracting high-tech employment and less of
-losing it
+There is no tendency of relative size distributions of urban computer
+employment to collapse, go bimodal, or fully spread. Overall computers
+exhibit some turbulence, with dramatic big winners and losers among
+cities. In attracting or repelling an industry, urban heterogeneity is
+important. Large, well educated cities near San Jose have much greater
+chances of attracting high-tech employment and less of losing it
 
-@bednarz2020pulled
-spatial innovation diffusion
-By using Bayesian survival models with time-dependent data of wind turbine deployment
-and firm foundation for 402 German regions between the years 1970 and 2015, we show that the spatial
-evolution of the German wind energy industry was more strongly influenced by local demand–
-pull than local supply–push processes.
+@bednarz2020pulled spatial innovation diffusion By using Bayesian
+survival models with time-dependent data of wind turbine deployment and
+firm foundation for 402 German regions between the years 1970 and 2015,
+we show that the spatial evolution of the German wind energy industry
+was more strongly influenced by local demand-- pull than local
+supply--push processes.
 
-The industry’s initial locations are distributed relatively arbitrarily
-and unpredictably, as their needs in terms of resources and skills are diverse and distinct from the older existing
-industries (Boschma and Lambooy, 1999). Consequently, emerging industries are characterized by relatively high
-degrees of freedom in terms of location. In later extensions of the concept, the assumption of the randomness of locations
-was revised with greater importance assigned to regional conditions (Boschma and Lambooy, 1999; Fornahl
-et al., 2012).
+The industry's initial locations are distributed relatively arbitrarily
+and unpredictably, as their needs in terms of resources and skills are
+diverse and distinct from the older existing industries (Boschma and
+Lambooy, 1999). Consequently, emerging industries are characterized by
+relatively high degrees of freedom in terms of location. In later
+extensions of the concept, the assumption of the randomness of locations
+was revised with greater importance assigned to regional conditions
+(Boschma and Lambooy, 1999; Fornahl et al., 2012).
 
-larger local technological niches (at least from a demand perspective) are
-more likely to be formed in urban areas than in rural ones, which is in line with the work of Ha¨gerstrand (1965a,b).
-However, size is not the only regional characteristic that may matter in this context. Regions also differ in accumulated
-experiences and the presence of tacit knowledge with respect to products, as well as in actors’ propensity share
-this knowledge among producers (Martin et al., 2019)
+larger local technological niches (at least from a demand perspective)
+are more likely to be formed in urban areas than in rural ones, which is
+in line with the work of Ha¨gerstrand (1965a,b). However, size is not
+the only regional characteristic that may matter in this context.
+Regions also differ in accumulated experiences and the presence of tacit
+knowledge with respect to products, as well as in actors' propensity
+share this knowledge among producers (Martin et al., 2019)
 
-@bento2015spatial
-The concept of absorptive capacity has also been applied in the TIS context in terms of the con-struction of capacity needed to transfer and implement a new technology in the receiving country (cf.Van Alphen, 2011), with repercussions for the nature of system functions
+@bento2015spatial The concept of absorptive capacity has also been
+applied in the TIS context in terms of the con-struction of capacity
+needed to transfer and implement a new technology in the receiving
+country (cf.Van Alphen, 2011), with repercussions for the nature of
+system functions
 
-@das2022diffusion [not good paper]
-Invention, innovation
-and diffusion are considered as the three pillars of
-the theory of technological change and Schumpeter is considered
-as the father of the study of diffusion
+@das2022diffusion [not good paper] Invention, innovation and diffusion
+are considered as the three pillars of the theory of technological
+change and Schumpeter is considered as the father of the study of
+diffusion
 
-Epidemic approach: The approach emphasizes that the adoption of a technology
-spreads like an infection among the potential adopters.
-By taking a homogeneous group of potential adopters as
-an assumption, it explains the adoption of any innovation.
+Epidemic approach: The approach emphasizes that the adoption of a
+technology spreads like an infection among the potential adopters. By
+taking a homogeneous group of potential adopters as an assumption, it
+explains the adoption of any innovation.
 
-To summarize
-the discussion, first, here the diffusion process is selfperpetuating
-as initial use stimulates further use. Second, it
-follows a disequilibrium path as the level of users is
-always lower than the number of potential users (Stoneman
-2002).
+To summarize the discussion, first, here the diffusion process is
+selfperpetuating as initial use stimulates further use. Second, it
+follows a disequilibrium path as the level of users is always lower than
+the number of potential users (Stoneman 2002).
 
-The
-criticisms are mainly based on the fact that although the
-approach gives an idea of aggregate (industry or household)
-behaviour, it does not focus on the individual’s
-(firm or household) adoption process
+The criticisms are mainly based on the fact that although the approach
+gives an idea of aggregate (industry or household) behaviour, it does
+not focus on the individual's (firm or household) adoption process
 
-Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.) Early adopters: next
-13.5%, (III.) Early Majority: following 34%, (IV.) Late Majority: next 34%, and the rest as (V.) Laggards
+Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.)
+Early adopters: next 13.5%, (III.) Early Majority: following 34%, (IV.)
+Late Majority: next 34%, and the rest as (V.) Laggards
 
-@morrill2020spatial
-Spatial diffusion is the process by which behavior or characteristics of the landscape change as a result of
-what happens elsewhere earlier. Spatial diffusion is the spread of the phenomenon, over space and timed,
-from limited origins
+@morrill2020spatial Spatial diffusion is the process by which behavior
+or characteristics of the landscape change as a result of what happens
+elsewhere earlier. Spatial diffusion is the spread of the phenomenon,
+over space and timed, from limited origins
 
-Contagious diffusion=> spatial autocorelation
+Contagious diffusion=\> spatial autocorelation
 
@@
 
-the three geographic diffusion processes are (1) purely contagious where distance or adjacency
-is the absolute controlling factor, (2) purely hierarchical where size or urban position in the central place
-hierarchy is the absolute controlling factor, or (3) where the location of change is purely random. These are
-the pure or ideal forms of the diffusion process; all real diffusion processes are the result of a combination of
+the three geographic diffusion processes are (1) purely contagious where
+distance or adjacency is the absolute controlling factor, (2) purely
+hierarchical where size or urban position in the central place hierarchy
+is the absolute controlling factor, or (3) where the location of change
+is purely random. These are the pure or ideal forms of the diffusion
+process; all real diffusion processes are the result of a combination of
 these extremes.
 
-The work by the communication sociologist Everett Rogers (1962, 1971, 1983) has emphasized the role of
-information, communication, formal and informal media, opinion leaders and social networks, and economic
-and psychological constraints on acceptance. Rogers’s work stresses the decision mechanism of the potential
-adopter; the work of Rogers partitions the adopter’s process of choosing to accept a new phenomenon or
-trait into five stages: stage 1–the potential adopter gains knowledge or awareness of an innovation; stage
-2–persuasion is exercised to adopt; stage 3–a decision is made to adopt; stage 4–the decision to adopt is
-implemented; stage 5–the adopter confirms the decision to adopt. In Rogers’s stage theory there may be a
-substantial time lag between when a potential adopter becomes aware of the new characteristic and when a
-decision is actually made to adopt.
+The work by the communication sociologist Everett Rogers (1962, 1971,
+1983) has emphasized the role of information, communication, formal and
+informal media, opinion leaders and social networks, and economic and
+psychological constraints on acceptance. Rogers's work stresses the
+decision mechanism of the potential adopter; the work of Rogers
+partitions the adopter's process of choosing to accept a new phenomenon
+or trait into five stages: stage 1--the potential adopter gains
+knowledge or awareness of an innovation; stage 2--persuasion is
+exercised to adopt; stage 3--a decision is made to adopt; stage 4--the
+decision to adopt is implemented; stage 5--the adopter confirms the
+decision to adopt. In Rogers's stage theory there may be a substantial
+time lag between when a potential adopter becomes aware of the new
+characteristic and when a decision is actually made to adopt.
 
-Hägerstrand considered diffusion to be a fundamental geographic process: Whatever the phenomenon being
-diffused might be, one may consider it in the context of a larger universal process of spatial diffusion.
+Hägerstrand considered diffusion to be a fundamental geographic process:
+Whatever the phenomenon being diffused might be, one may consider it in
+the context of a larger universal process of spatial diffusion.
 
-@zook2022mapping
-Because new digital activities are rarely—if
-ever—captured in official state data, researchers
-must rely on information gathered from alternative
-sources. With this in mind, the overarching aim of
-this article is outlining an approach for analyzing the
-“new spaces and geographies” of digital phenomena
-and practices.
+@zook2022mapping Because new digital activities are rarely---if
+ever---captured in official state data, researchers must rely on
+information gathered from alternative sources. With this in mind, the
+overarching aim of this article is outlining an approach for analyzing
+the "new spaces and geographies" of digital phenomena and practices.
 
-@leibowicz2016representing for energy
-There is no generally accepted theory that explains diffusion
-rate heterogeneity across technologies, but several factors are
-considered important. Greater unit scale and larger market size
-contribute to slower diffusion. Requirements for interrelated
-technologies or complex infrastructures also hinder the diffusion
-process (Grubler, 2012).
+@leibowicz2016representing for energy There is no generally accepted
+theory that explains diffusion rate heterogeneity across technologies,
+but several factors are considered important. Greater unit scale and
+larger market size contribute to slower diffusion. Requirements for
+interrelated technologies or complex infrastructures also hinder the
+diffusion process (Grubler, 2012).
 
-Mobile phones benefited from early
-deployment in recreational boats and automobiles, where the
-traditional competitor was not a viable option. In the early
-stages of diffusion, performance is a more important driver of
-adoption than cost competitiveness. Typically, significant cost
-reductions only occur once the technology reaches a deployment
-level capable of supporting standardization and mass
-production (Wilson, 2012).
+Mobile phones benefited from early deployment in recreational boats and
+automobiles, where the traditional competitor was not a viable option.
+In the early stages of diffusion, performance is a more important driver
+of adoption than cost competitiveness. Typically, significant cost
+reductions only occur once the technology reaches a deployment level
+capable of supporting standardization and mass production (Wilson,
+2012).
 
-Diffusion is a spatial (Hagerstrand, 1967) as well as a
-temporal process, and historical evidence confirms that
-technologies diffuse at different times, at different rates, and
-to different extents in different places, and can be significantly
-influenced by policies (Victor, 1993). An illustrative empirically
-based theory for technology diffusion acrossmultiple regions is
-suggested by “Schmidt's Law” (Grubler, 1990). With respect to a
-specific technology, it divides regions into three groups: core,
-rim, and periphery. The core iswhere the technology is invented,
-subject to experimentation, and first deployed commercially.
-To develop the technology, the core must have both sufficient
-innovative ability and economic motivation to make the
-necessary resource commitment. As the technology spreads
-from the core to the rim and eventually to the periphery, later
-adopters have the advantage of benefiting from the experiences
-of earlier adopters via knowledge spillover effects (Verdolini and
-Galeotti, 2011). At this point, the technology has likely been
-refined and costs have likely declined significantly. Schmidt's
-Law claims that, as a result, diffusion proceeds at a faster rate in
-regions which adopt the technology later. However, diffusion
-tends to be less pervasive in regions where adoption is delayed,
-ultimately saturating at a lower extent than in regions where
-adoption occurred earlier. This is often attributed to a lack of
-proper institutions and infrastructures in later adopters, markets
-whichmay not be as favorable to a technology, or the availability
-of yet newer options that invite technological leapfrogging.
+Diffusion is a spatial (Hagerstrand, 1967) as well as a temporal
+process, and historical evidence confirms that technologies diffuse at
+different times, at different rates, and to different extents in
+different places, and can be significantly influenced by policies
+(Victor, 1993). An illustrative empirically based theory for technology
+diffusion acrossmultiple regions is suggested by "Schmidt's Law"
+(Grubler, 1990). With respect to a specific technology, it divides
+regions into three groups: core, rim, and periphery. The core iswhere
+the technology is invented, subject to experimentation, and first
+deployed commercially. To develop the technology, the core must have
+both sufficient innovative ability and economic motivation to make the
+necessary resource commitment. As the technology spreads from the core
+to the rim and eventually to the periphery, later adopters have the
+advantage of benefiting from the experiences of earlier adopters via
+knowledge spillover effects (Verdolini and Galeotti, 2011). At this
+point, the technology has likely been refined and costs have likely
+declined significantly. Schmidt's Law claims that, as a result,
+diffusion proceeds at a faster rate in regions which adopt the
+technology later. However, diffusion tends to be less pervasive in
+regions where adoption is delayed, ultimately saturating at a lower
+extent than in regions where adoption occurred earlier. This is often
+attributed to a lack of proper institutions and infrastructures in later
+adopters, markets whichmay not be as favorable to a technology, or the
+availability of yet newer options that invite technological
+leapfrogging.
 
-Empirical evidence supports the validity of Schmidt's Law
-over a wide range of technologies, time periods, and geographical
-contexts. A recent meta-analysis of technology up-scaling
-found that diffusion accelerated moving from the core to the
-rim and periphery for technologies as diverse as natural gas
-power, oil refineries, and automobiles (Wilson, 2009). One
-historical example that conforms particularly well to Schmidt's
-Law is the diffusion of coal power in Europe (Grubler, 2012).
-England emerged as the core region for coal power because it
-had legal and economic institutions that incentivized scientific
-pursuits, domestic coal reserves, and a clear industrial motivation
-to replace water power with coal. 
-...
+Empirical evidence supports the validity of Schmidt's Law over a wide
+range of technologies, time periods, and geographical contexts. A recent
+meta-analysis of technology up-scaling found that diffusion accelerated
+moving from the core to the rim and periphery for technologies as
+diverse as natural gas power, oil refineries, and automobiles (Wilson,
+2009). One historical example that conforms particularly well to
+Schmidt's Law is the diffusion of coal power in Europe (Grubler, 2012).
+England emerged as the core region for coal power because it had legal
+and economic institutions that incentivized scientific pursuits,
+domestic coal reserves, and a clear industrial motivation to replace
+water power with coal. ...
 
-Lastly, it can be globally optimal for innovative economies
-to deploy advanced technologies more than what is locally
-optimal if this enables faster diffusion in less advanced regions
-that could benefit from the technology.
+Lastly, it can be globally optimal for innovative economies to deploy
+advanced technologies more than what is locally optimal if this enables
+faster diffusion in less advanced regions that could benefit from the
+technology.
 
-@wilson201281
-Logistic growth describes an initial period of gradual diffusion as a technology is introduced as a new commercial application, moving then through a rapid, exponential growth phase, before slowing and eventually saturating [@grubler1999dynamics].
-The substitution of incumbent technologies by new competitors leads to subsequent decline and eventual obsolescence.
+@wilson201281 Logistic growth describes an initial period of gradual
+diffusion as a technology is introduced as a new commercial application,
+moving then through a rapid, exponential growth phase, before slowing
+and eventually saturating [@grubler1999dynamics]. The substitution of
+incumbent technologies by new competitors leads to subsequent decline
+and eventual obsolescence.
 
-Early on in their lifecycle, new technologies are crude, imperfect,
-and expensive (Rosenberg, 1994). New energy technologies
-are attractive for their ability to perform a particular task or
-deliver a new or improved energy service (Fouquet, 2010). This is
-often circumscribed by a particular set of needs in a particular
-context: a market ‘niche’. End-users in niche markets are generally
-less sensitive to the effective price of the energy service
-provided or have a higher willingness to pay for its performance
-advantages (Fouquet, 2010). Thus initially, performance dominates
-cost competitiveness (Wilson and Grubler, 2011). Market
-niches afford some protection from competitive pressures, allowing
-technologies to be tested and improved in applied settings,
-reducing uncertainties with performance or market demand
-(Kemp et al., 1998). Costs may only fall substantively after an
-extended period of commercial experimentation, concurrent with
-the establishment of an industrial base and characteristic moves
-towards standardisation and mass production (Grubler, 1998).
-The influence of accumulating production experience on costs is
-captured by the concept of learning.
+Early on in their lifecycle, new technologies are crude, imperfect, and
+expensive (Rosenberg, 1994). New energy technologies are attractive for
+their ability to perform a particular task or deliver a new or improved
+energy service (Fouquet, 2010). This is often circumscribed by a
+particular set of needs in a particular context: a market 'niche'.
+End-users in niche markets are generally less sensitive to the effective
+price of the energy service provided or have a higher willingness to pay
+for its performance advantages (Fouquet, 2010). Thus initially,
+performance dominates cost competitiveness (Wilson and Grubler, 2011).
+Market niches afford some protection from competitive pressures,
+allowing technologies to be tested and improved in applied settings,
+reducing uncertainties with performance or market demand (Kemp et al.,
+1998). Costs may only fall substantively after an extended period of
+commercial experimentation, concurrent with the establishment of an
+industrial base and characteristic moves towards standardisation and
+mass production (Grubler, 1998). The influence of accumulating
+production experience on costs is captured by the concept of learning.
 
-@grubler1990rise
-Following the definition given by Rogers, 1983, diffusion is the
-process by which an innovation is communicated through certain communication
-channels over time among the members of a social system.
-An innovation may be an idea, object or practice which is perceived
-as new by an individual or another unit of adoption (e.g., an
-organization) .
-Communication via communication channels refers to how
-knowledge about an innovation is transferred between the members of
-a social system. Communication takes time, and is therefore an
-important aspect of innovation diffusion. In fact, all theoretical and
-empirical diffusion studies agree that an innovation does not instantly
-diffuse into a social system. Instead a typical time pattern of diffusion
-along a S-shaped trajectory seems to be the rule. The S-shaped pattern
-of diffusion appears to be a basic anthropologic phenomenon, as
-it is also confirmed by a large number of studies· of primitive preindustrial
+@grubler1990rise Following the definition given by Rogers, 1983,
+diffusion is the process by which an innovation is communicated through
+certain communication channels over time among the members of a social
+system. An innovation may be an idea, object or practice which is
+perceived as new by an individual or another unit of adoption (e.g., an
+organization) . Communication via communication channels refers to how
+knowledge about an innovation is transferred between the members of a
+social system. Communication takes time, and is therefore an important
+aspect of innovation diffusion. In fact, all theoretical and empirical
+diffusion studies agree that an innovation does not instantly diffuse
+into a social system. Instead a typical time pattern of diffusion along
+a S-shaped trajectory seems to be the rule. The S-shaped pattern of
+diffusion appears to be a basic anthropologic phenomenon, as it is also
+confirmed by a large number of studies· of primitive preindustrial
 societies.
 
-The scale of spatial innovation diffusion is similar* to that of
+The scale of spatial innovation diffusion is similar\* to that of
 temporal diffusion models: hierarchically decomposed.
 
-Spatial innovation diffusion in Hagerstrand's model consists of
-three stages.
-Stage 1: Local concentrations of initial acceptances (initial
-agglomeration) .
-Stage 2: Radial dissemination outward from the initial agglomeration,
-accompanied by the rise of secondary agglomerations.
-Stage 3: The condensing and saturation stage, in which filling-in
-occurs and diffusion eventually ceases.
+Spatial innovation diffusion in Hagerstrand's model consists of three
+stages. Stage 1: Local concentrations of initial acceptances (initial
+agglomeration) . Stage 2: Radial dissemination outward from the initial
+agglomeration, accompanied by the rise of secondary agglomerations.
+Stage 3: The condensing and saturation stage, in which filling-in occurs
+and diffusion eventually ceases.
 
@@
 (1) As in temporal diffusion models, an S-shaped pattern in the
-cumulative level of adoption.
-(2) A "hierarchy effect" by which diffusion proceeds from initial
-centers to secondary ones. Within the context of urban systems
-(central places) this implies a hierarchical spread of an innovation
-from larger to smaller centers, and from the centers to the surrounding
-hinterland.
-(3) A "neighborhood effect" Le., diffusion proceeds outwards from
-innovation centers, first "hitting" (Brown, 1981) nearby rather
-than far-away locations.
+cumulative level of adoption. (2) A "hierarchy effect" by which
+diffusion proceeds from initial centers to secondary ones. Within the
+context of urban systems (central places) this implies a hierarchical
+spread of an innovation from larger to smaller centers, and from the
+centers to the surrounding hinterland. (3) A "neighborhood effect" Le.,
+diffusion proceeds outwards from innovation centers, first "hitting"
+(Brown, 1981) nearby rather than far-away locations.
 
-Later Hagerstrand conceptualized
-physical "barrier" effects like lakes or uninhabited areas, which, in
-addition to distance, act as further retarding effects on diffusion.
-These are formalized in the form of "zero" or "half" contact multiplicators
-on the (distance decaying) message flows.
+Later Hagerstrand conceptualized physical "barrier" effects like lakes
+or uninhabited areas, which, in addition to distance, act as further
+retarding effects on diffusion. These are formalized in the form of
+"zero" or "half" contact multiplicators on the (distance decaying)
+message flows.
 
-With respect to the formalization of the communication flows
-Hagerstrand defines a "mean information field" (MIF), in which the
-probability of communication is a negative function of distance
-between individuals
+With respect to the formalization of the communication flows Hagerstrand
+defines a "mean information field" (MIF), in which the probability of
+communication is a negative function of distance between individuals
 
-@ding2010modeling model the spatial diffusion of mobile telecommunications in China
-as well as its determinants. regional socioeconomic characteristics play an
-important role in determining the timing,
-speed, and level of mobile telecommunications
-diffusion in China
+@ding2010modeling model the spatial diffusion of mobile
+telecommunications in China as well as its determinants. regional
+socioeconomic characteristics play an important role in determining the
+timing, speed, and level of mobile telecommunications diffusion in China
 
-Rogers
-(1995) characterized early adopters as knowledgeable
-risk takers. Griliches (1957) looked
-at the rate of return for early adopters and
-characterized them as profit maximizers. The
-geographer Hagerstrand (1967) was focusing
-on the character of early adopters in the transmittal
-process that produces follow-on participants
-in the process. Space was treated as a
-contiguity system with embedded characteristics
-producing barrier elements to the diffusion
-process.
+Rogers (1995) characterized early adopters as knowledgeable risk takers.
+Griliches (1957) looked at the rate of return for early adopters and
+characterized them as profit maximizers. The geographer Hagerstrand
+(1967) was focusing on the character of early adopters in the
+transmittal process that produces follow-on participants in the process.
+Space was treated as a contiguity system with embedded characteristics
+producing barrier elements to the diffusion process.
 
@@
 
-Perkins and Neumayer (2005)
-examined whether latecomers have a higher diffusion
-rate of three technologies including digital
-telephone mainline
+Perkins and Neumayer (2005) examined whether latecomers have a higher
+diffusion rate of three technologies including digital telephone
+mainline
 
-@bento2018time explore 
-What determines the duration of
-formative phases for energy innovations in different markets? We are interested
-both in initial markets (also: core, lead, first mover, early
-adopter) where formative phases prepare technologies for mass commercialization,
-and in follower markets (also: periphery, lag, late
-adopter) where accelerated formative phases may benefit from diffusion
-and spillovers.
+@bento2018time explore What determines the duration of formative phases
+for energy innovations in different markets? We are interested both in
+initial markets (also: core, lead, first mover, early adopter) where
+formative phases prepare technologies for mass commercialization, and in
+follower markets (also: periphery, lag, late adopter) where accelerated
+formative phases may benefit from diffusion and spillovers.
 
-Hagerstrand (1968) was the first
-to demonstrate that diffusion typically starts from innovative centers
-(‘core´) and disseminates, through a hierarchy of subcenters, to the
-periphery. This process is sequential rather than simultaneous, and
-tends to accelerate from core to periphery, where diffusion reaches a
-lower intensity than in the core (Hagerstrand, 1968; Grubler, 1990;
-Morrill, 2005). A similar effect of spatial acceleration has been identified
-in marketing research and is known as the “lead-lag effect” (Peres
-et al., 2010).
+Hagerstrand (1968) was the first to demonstrate that diffusion typically
+starts from innovative centers ('core´) and disseminates, through a
+hierarchy of subcenters, to the periphery. This process is sequential
+rather than simultaneous, and tends to accelerate from core to
+periphery, where diffusion reaches a lower intensity than in the core
+(Hagerstrand, 1968; Grubler, 1990; Morrill, 2005). A similar effect of
+spatial acceleration has been identified in marketing research and is
+known as the "lead-lag effect" (Peres et al., 2010).
 
 @perkins2005international analyze whether the rate at which new producer
-technologies diffuse is significantly influenced by
-(1) latecomer advantage and (2) engagement with the
-global economy via trade and foreign investment
+technologies diffuse is significantly influenced by (1) latecomer
+advantage and (2) engagement with the global economy via trade and
+foreign investment
 
-Indeed, precisely
-because of these latecomer advantages, developing
-countries1 are believed to be well placed to catch up with
-developed ones (Gerschenkron 1962; Abramovitz
-1986).
+Indeed, precisely because of these latecomer advantages, developing
+countries1 are believed to be well placed to catch up with developed
+ones (Gerschenkron 1962; Abramovitz 1986).
 
-Geographers, of course, were at the forefront of
-quantitative diffusion research. Ha¨gerstrand’s (1967)
-early work on understanding innovation diffusion in
-Sweden using simulation modeling is particularly noteworthy
-in this respect. However, already long before the
-‘‘cultural turn’’ in economic geography (Barnes 2001),
-geographers’ interest in formally modeling the diffusion
-of new innovations waned (Gregory 1985). Indeed,
-much of the recent quantitative research into the diffusion
-of new technology has been undertaken by
-economists, sociologists, and business scholars.
+Geographers, of course, were at the forefront of quantitative diffusion
+research. Ha¨gerstrand's (1967) early work on understanding innovation
+diffusion in Sweden using simulation modeling is particularly noteworthy
+in this respect. However, already long before the ''cultural turn'' in
+economic geography (Barnes 2001), geographers' interest in formally
+modeling the diffusion of new innovations waned (Gregory 1985). Indeed,
+much of the recent quantitative research into the diffusion of new
+technology has been undertaken by economists, sociologists, and business
+scholars.
 
-Theoretical models disagree as to why firms adopt
-innovations at different times. Epidemic models emphasize
-information (Griliches 1957). Certain firms are hypothesized
-to adopt earlier because they come into
-contact with, and learn from, adopters of the new technology
-before others. Economic models, on the other
-hand, predominantly emphasize firm heterogeneity (Ireland
-and Stoneman 1986). Firms adopt technologies at
+Theoretical models disagree as to why firms adopt innovations at
+different times. Epidemic models emphasize information (Griliches 1957).
+Certain firms are hypothesized to adopt earlier because they come into
+contact with, and learn from, adopters of the new technology before
+others. Economic models, on the other hand, predominantly emphasize firm
+heterogeneity (Ireland and Stoneman 1986). Firms adopt technologies at
 different times because they differ with respect to various
-organizational and environmental variables influencing
-the economic returns from adoption (Blackman 1999).
+organizational and environmental variables influencing the economic
+returns from adoption (Blackman 1999).
 
-Either
-way, the strong assumption is that developing countries
-can acquire modern technology innovated in developed
-economies, often at a fraction of the original research
-and development (R&D) costs, thereby leapfrogging
-many decades of technological progress (Teece 2000).
+Either way, the strong assumption is that developing countries can
+acquire modern technology innovated in developed economies, often at a
+fraction of the original research and development (R&D) costs, thereby
+leapfrogging many decades of technological progress (Teece 2000).
 
-The first involves a country’s geographical location.
-Recent empirical work suggests that diffusion is ‘‘geographically
-localized’’ (Globerman, Kokko, and Sjo¨holm
-2000; Keller 2002; Milner 2003) in that a technology
-diffuses faster in a country where it is already more
-widely diffused in neighboring countries. Underlying
-these regional effects are contagion and contact with
-prior users or producers of technology.
+The first involves a country's geographical location. Recent empirical
+work suggests that diffusion is ''geographically localized'' (Globerman,
+Kokko, and Sjo¨holm 2000; Keller 2002; Milner 2003) in that a technology
+diffuses faster in a country where it is already more widely diffused in
+neighboring countries. Underlying these regional effects are contagion
+and contact with prior users or producers of technology.
 
-@perkins2011internet
-Two main mechanisms are
-identified in the literature to explain diffusion:
-(1) epidemic-type dynamics whereby contact
-with previous adopters stimulates uptake as po
-tential
-adopters learn about a new innovation;
-and (2) economic-type mechanisms whereby
-potential users adopt a new innovation as it be
-comesmore
-profitable, useful, or valuable, with
-uptake characteristically spreading as costs be
-come
-lower, performance improves, or the po
-tential
-uses of the innovation grow over time.
+@perkins2011internet Two main mechanisms are identified in the
+literature to explain diffusion: (1) epidemic-type dynamics whereby
+contact with previous adopters stimulates uptake as po tential adopters
+learn about a new innovation; and (2) economic-type mechanisms whereby
+potential users adopt a new innovation as it be comesmore profitable,
+useful, or valuable, with uptake characteristically spreading as costs
+be come lower, performance improves, or the po tential uses of the
+innovation grow over time.
 
-To this extent,
-our results contribute to a growing body of
-work that has sought to caution against claims
-about the supposed novelty of the Internet and
-the suggestion that it is somehow different.
+To this extent, our results contribute to a growing body of work that
+has sought to caution against claims about the supposed novelty of the
+Internet and the suggestion that it is somehow different.
 
-@meade2021modelling
-There are numerous agents who would benefit from accurate predictions of introduction times.
-These include: suppliers of network equipment for production planning; suppliers of mobile phones who need to have inventories of
-appropriately configured handsets; international network planners who need to provide sufficient bandwidth for communications
-between different countries; network operators and regulatory authorities.
+@meade2021modelling There are numerous agents who would benefit from
+accurate predictions of introduction times. These include: suppliers of
+network equipment for production planning; suppliers of mobile phones
+who need to have inventories of appropriately configured handsets;
+international network planners who need to provide sufficient bandwidth
+for communications between different countries; network operators and
+regulatory authorities.
 
 Classifying adopters of a technology along a time axis was pioneered by
-Rogers (1962), who categorized adopters into five groups as innovators, early adopters, early majority, late majority and laggards.
+Rogers (1962), who categorized adopters into five groups as innovators,
+early adopters, early majority, late majority and laggards.
 
-An a priori approach groups similar countries into segments, where similarity is defined using economic, social, or
-political covariates. Examples include Hofstede (2001) and Lee (1990). An a posteriori analysis considers segments based on realized
-market behaviour, for example see Sood et al. (2009). Some studies investigate changing segment membership over time, see Cannon
-and Yaprak (2011).
+An a priori approach groups similar countries into segments, where
+similarity is defined using economic, social, or political covariates.
+Examples include Hofstede (2001) and Lee (1990). An a posteriori
+analysis considers segments based on realized market behaviour, for
+example see Sood et al. (2009). Some studies investigate changing
+segment membership over time, see Cannon and Yaprak (2011).
 
@@
 
-@fritsch2015new analyze the spatial diffusion of laser technology research in West Germany from
-1960, when this technology began, until 2005.
+@fritsch2015new analyze the spatial diffusion of laser technology
+research in West Germany from 1960, when this technology began, until
+2005.
 
-Hägerstrand (1952, 1967) was the first scholar to intensively analyze spatial diffusion pro
-cesses.
-Based on the empirical observations of the regional spread of process innovations,
-he hypothesized that new technology is first implemented in the ‘center’ – the large agglom
-erations
-– and then moves ‘down’ the spatial hierarchy, the last stop being the ‘periphery’,
-i.e. remote and sparsely populated regions. There are at least two factors that may explain
-this pattern: first, the transfer of tacit knowledge via face-to-face contact and second, the
+Hägerstrand (1952, 1967) was the first scholar to intensively analyze
+spatial diffusion pro cesses. Based on the empirical observations of the
+regional spread of process innovations, he hypothesized that new
+technology is first implemented in the 'center' -- the large agglom
+erations -- and then moves 'down' the spatial hierarchy, the last stop
+being the 'periphery', i.e. remote and sparsely populated regions. There
+are at least two factors that may explain this pattern: first, the
+transfer of tacit knowledge via face-to-face contact and second, the
 absorptive capacity of a region.
 
-# Literature review {#sec2} MAYBE SKIP THIS??
+# Literature review {\#sec2} MAYBE SKIP THIS??
 
@@
 
-from R: Asym/(1+exp((xmid-input)/scal))
-using the terms from Wilson k/(1 + e^(t_0 - t)/scal)
+from R: Asym/(1+exp((xmid-input)/scal)) using the terms from Wilson k/(1
++ e\^(t_0 - t)/scal)
 
-Wilson: k(1 + e^-b(t - t_0))
-So, scal = - 1/b
+Wilson: k(1 + e\^-b(t - t_0)) So, scal = - 1/b
 
-The curve is symmetric* around t_0
+The curve is symmetric\* around t_0
 
@@
 
-The literature usually uses the saturation level as the asymptote. I am using the total
-number of websites as we cannot compute a rate.
+The literature usually uses the saturation level as the asymptote. I am
+using the total number of websites as we cannot compute a rate.
 
-Moran's I as @ding2010modeling
-**TODO** for t_0,  diffusion speed
+Moran's I as @ding2010modeling **TODO** for t_0, diffusion speed
 
@@
 
-contrary to results from future studies regarding social media [@lengyel2020role], web technologies 
-did not exclusively spread from a central location.
+contrary to results from future studies regarding social media
+[@lengyel2020role], web technologies did not exclusively spread from a
+central location.
 
-@lengyel2020role
-We see that the OSN
-spread almost exclusively from the original location (the capital Budapest, with an order of magnitude more inhabitants than
-the next size town) to various parts of the country in the early phase of the life-cycle. Later, diffusion became less mono-centric
-and other towns also emerged as spreaders. Our findings support the idea that spreading initially happens to large distances and
-becomes more local over time. This is illustrated and discussed later in Figure 2I.
+@lengyel2020role We see that the OSN spread almost exclusively from the
+original location (the capital Budapest, with an order of magnitude more
+inhabitants than the next size town) to various parts of the country in
+the early phase of the life-cycle. Later, diffusion became less
+mono-centric and other towns also emerged as spreaders. Our findings
+support the idea that spreading initially happens to large distances and
+becomes more local over time. This is illustrated and discussed later in
+Figure 2I.
 

### 2023-02-16 presentations/gceg.Rmd
presentation

@@
 library(knitr)
+library(tidyverse)
 library(rprojroot)
+library(magick)
+library(purrr)
+library(kableExtra)
 
@@
 path <- find_rstudio_root_file()
-
-#bibliography: ../bibliography.bib
 ```
@@
 ## Contents
+<br>
+<div style="margin-left: 50px !important; text-align:left !important">
 - Introduction
-- Web data and spatial research
-- Empirical strategy
-- Descriptive statistics
-- Results
+- Web data
+- Spatial analysis
+- S-shaped diffusion curves
 - Conclusions
@@
 <br>
-[etranos.info/post/sad2021](https://etranos.info/post/sad2021)
+[etranos.info/post/gceg2022](https://etranos.info/post/gceg2022/)
+</div>
 
@@
 
-- Map the active engagement with the *digital*
-- Over time, early stages of the internet
-- Granular spatial level
-- Importance: understand how the adoption of technologies evolves
-- Predictions of introduction times for future technologies [@meade2021modelling]
+<br>
+
+- Map the **active** engagement with the *digital*
+- Over time, **early stages** of the internet
+- **Granular** and **multi-scale** spatial perspective
+
+## Importance
+
+- Understand how the adoption of new technologies evolves
+
+- Guide policies for deployment of new technologies
+
+- Predictions of introduction times for future technologies [@meade2021modelling]:
 
@@
 
-## Diffusion
+## Technological diffusion
 
-Large scale adoption studies
+<br>
 
-Add Figure
+![](images/S-shaped-logistic-diffusion-curve_W640.jpg)
 
-## Spatiality of the internet using supply-side measures
+## *Spatial* diffusion processes
 
-- Map internet infrastructure, *both*:
+- As in temporal diffusion models, an **S-shaped** pattern in the
+cumulative level of adoption
 
-  - backbone (ADD REF), and
+- A **hierarchy effect**: from main centers to secondary ones -- central places
 
-  - last mile (ADD REF, Singleton, Hannah)
+- A **neighborhood effect**: diffusion proceeds outwards from innovation centers, 
+first "hitting" nearby rather than far-away locations [@grubler1990rise]
 
-- Economic effects:
+@hagerstrand1965monte: from innovative centers (core) through a
+hierarchy of subcenters, to the periphery
 
-  - Kolko
+## Internet geographies
 
-  - Nathan
+- Map internet infrastructure
 
-## ???
+  - backbone [@malecki2002economic; @tranos2013geography]
 
-@beardsell1999spatial used the employment in the computer industry to assess
-its evolution
+  - last mile [@riddlesden2014broadband; @budnitz2022working]
 
-# Results
+- Social media and user generated content
+
+  - Here and now effect [@crampton2013beyond]
+
+> Because new digital activities are rarely---if
+ever---captured in official state data, researchers must rely on
+information gathered from alternative sources [@zook2022mapping].
+
+# Web data
+
+## Web data: The Internet Archive
+- The largest archive of webpages in the world
+- 273 billion webpages from over 361 million websites, 15 petabytes of storage (1996 -)
+- A web crawler starts with a list of URLs (a seed list) to crawl and downloads a copy of their content
+- Using the hyperlinks included in the crawled URLs, new URLs are identified and crawled (snowball sampling)
+- Time-stamp
+
+## Web data: The Internet Archive
+![](C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/figure/Picture1.png){width=70%}
+
+## Web data: The Internet Archive
+![](C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/figure/Picture2.png){width=70%}
+
+## Our web data
+- JISC UK Web Domain Dataset: all archived webpages from the .uk domain 1996-2012
+
+- Curated by the British Library
+
+<small>
+
+- Tranos, E., and C. Stich. 2020. Individual internet usage and the availability of online content of local interest: A multilevel approach. *Computers, Environment and Urban Systems*, 79:101371.
+
+- Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy in the UK: Regional productivity effects of early adoption. *Regional Studies*, 55:12, 1924-1938.
+
+- Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from the ground up: a web data approach. *Environment and Planning B*, in press.
+
+- Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to predict regional trade flows: data extraction, modelling, and validation, *Annals of the AAG*, in press.
+
+</small>
+
+## Our web data
+
+- All .uk archived webpages which contain a UK postcode in the web text
+
+- Circa 0.5 billion URLs with valid UK postcodes
+
+<br>
+<br>
+
+<!-- <small></small> -->
+20080509162138/http://www.website1.co.uk/contact_us IG8 8HD
+
+## Data cleaning
+
+- All the archived .uk webpages
+
+- Archived during 1996-2012
+
+- Commercial webpages (.co.uk)
+
+- From webpages to websites:
+
+    <small>- http://www.website1.co.uk/webpage1 and</small>
+
+    <small>- http://www.website1.co.uk/webpage2 are part of the</small>
+
+    <small>- http://www.website1.co.uk</small>
+
+- 1 *vs.* multuple postcodes in a website
+
+## Unique postcodes frequencies, 2000
+
+```{r eval=TRUE, echo=FALSE, results='asis'}
+
+# load 2000 co.uk
+# setwd("C:/Users/TranosE/DataShare/archive/nuts")
+path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"
+
+all2000.duplicates <- data.table::fread(path.2000) #
+# dim(all2000.duplicates) # 3336162
+# it includes duplicated URLs: if one webpages includes multiple postcodes
+# then it appears multiple times. This is ok for the nuts aggregation, but
+# not for the frequencies
+# This is only .co.uk
+
+# one line for every host
+all2000 <- unique(all2000.duplicates, by = c("host")) # 57897
+
+# unique postcodes per website f table
+f.websites.pc <- DescTools::Freq(all2000$V1, breaks = c(0, 1,2, 10,100,1000,10000,100000), ord = "desc")
+f.websites.pc$level <- factor(f.websites.pc$level, levels = c("[0,1]","(1,2]", "(2,10]", "(10,100]",
+                                                              "(100,1e+03]", "(1e+03,1e+04]",
+                                                              "(1e+04,1e+05]"))
+levels(f.websites.pc$level) <- c("(0,1]","(1,2]", "(2,10]", "(10,100]", "(100,1000]", "(1000,10000]", "(10000,100000]")
+kable(f.websites.pc,
+      format = "html",
+      digits = 3,
+      format.args = list(big.mark = ",")) %>% 
+  kable_classic(full_width = F, html_font = "Cambria")
+```
+<br>
+
+- Websites with a large number of postcodes: e.g. directories, real estate websites
+
+- Focus on websites with **one unique** postcode per year
+
+## Directory website with a lot of postcodes
+![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
+
+## Website with a unique postcode in London
+![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
+
+# Analysis
+
+## Mapping website density
 
@@
 
-```{r results='asis', out.width = "60%"}
-
-library(magick)
-library(purrr)
+```{r heatmaps, results='asis', out.width = "60%"}
 path.images <- paste0(path, "/outputs/")
@@
   #image_write("test.gif") # write to current dir
-
 ```
@@
 
-```{r results='asis', out.width = "60%"}
-
-library(magick)
-library(purrr)
+```{r heatmaps_log, results='asis', out.width = "60%"}
 path.images <- paste0(path, "/outputs/")
@@
   #image_write("test.gif") # write to current dir
+```
+
+## Spatial attributes
+
+<br>
+
+Two scales:
+
+- Output areas (c. 200,000)
+
+- Local authorities (c. 400)
+
+---
+
+<br>
+
+**Neighborhood effect**: diffusion proceeds outwards from innovation centers, 
+first "hitting" nearby rather than far-away locations [@grubler1990rise]
+
+- Moran's I
+
+- LISA maps
+
+- Website density regressions
+
+---
+
+```{r moransi_oa, results='asis'}
+path.in <- paste0(path, "/outputs/morani_oa.csv")
+# read.csv(path.in) %>% dplyr::select(year, morani) %>% 
+#   rename('Moran\'s I' = morani,
+#          Year = year) %>% 
+#   arrange(Year) %>% 
+#   kable(digits = 3, 
+#         label = "Moran\'s I per year, Output Areas") %>% 
+#     kable_classic(full_width = F, html_font = "Cambria")
+
+read.csv(path.in) %>% dplyr::select(year, morani) %>% 
+  ggplot(aes(x=year, y=morani)) +
+  geom_bar(stat = "identity") + 
+  ylab('Moran\'s I') +
+  theme_minimal() +
+  ggtitle("Moran\'s I, number of websites per Output Areas") +
+  theme(plot.title = element_text(hjust = 0.5)) 
+```
+
+---
+
+```{r moransi_la, results='asis'}
+path.in <- paste0(path, "/outputs/morani_la.csv")
+
+read.csv(path.in) %>% dplyr::select(year, morani) %>% 
+  ggplot(aes(x=year, y=morani)) +
+  geom_bar(stat = "identity") + 
+  ylab('Moran\'s I') +
+  theme_minimal() +
+  ggtitle("Moran\'s I, number of websites per Local Authority") +
+  theme(plot.title = element_text(hjust = 0.5)) 
+```
+
+---
+
+```{r lisa_oa, results='asis', out.width = "30%"}
+path.images <- paste0(path, "/outputs/")
+
+list.files(path = path.images, pattern = "^lisa_level_pc1_oa", full.names = T) %>% 
+  map(image_read) %>% # reads each path file
+  image_join() %>% # joins image
+  image_animate(delay=50) #%>% # animates, can opt for number of loops
+  #image_write("test.gif") # write to current dir
+```
+
+---
+
+```{r lisa_la, results='asis', out.width = "30%"}
+path.images <- paste0(path, "/outputs/")
+
+list.files(path = path.images, pattern = "^lisa_level_pc1_la", full.names = T) %>% 
+  map(image_read) %>% # reads each path file
+  image_join() %>% # joins image
+  image_animate(delay=50) #%>% # animates, can opt for number of loops
+  #image_write("test.gif") # write to current dir
+```
+
+## Website density regressions
+
+<br>
+
+$$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$
+
+<br>
+
+<div style="margin-left: 50px !important; text-align:left !important">
+$Place$:
+
+- London, *or*
+
+- Nearest city, *or*
+
+- Nearest retail centre
+
+$i$: Output Areas *or* Local Authorities
+</div>
+
+## Website density regressions
+
+<br>
+
+$\beta$ interpretation:
+
+- The lower the $\beta$ is (or the larger the $|\beta|$ is)...
+
+- ... the larger urban gravitation is for web adoption.
+
+---
+
+```{r b_London_oa, fig.height=5, fig.width=5, message=FALSE}
+#, out.width = "60%"
+path.image <- paste0(path, "/outputs/b_dist_london_oa.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r b_London_la, fig.height=5, fig.width=5, message=FALSE}
+path.image <- paste0(path, "/outputs/b_dist_london_la.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r b_city_oa, fig.height=5, fig.width=5, message=FALSE}
+#, out.width = "60%"
+path.image <- paste0(path, "/outputs/b_dist_city_oa.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r b_city_la, fig.height=5, fig.width=5, message=FALSE}
+#, out.width = "60%"
+path.image <- paste0(path, "/outputs/b_dist_city_la.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r b_centre_oa, message=FALSE, out.height= "60%", out.width="60%"}
+#fig.height=4, fig.width=4, 
+path.image <- paste0(path, "/outputs/b_dist_centre_oa.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r b_centre_la, message=FALSE, out.height= "60%", out.width="60%"}
+path.image <- paste0(path, "/outputs/b_dist_centre_la.png")
+knitr::include_graphics(path.image)
+```
+
+```{r r_squared, eval=F, results='asis'}
+# This is for the RSquared of the different website density regressions.
+# The RSquared does not drop much hence it is not worth reporting.
+
+path.in1 <- paste0(path, "/outputs/r2_all_oa.csv")
+path.in2 <- paste0(path, "/outputs/r2_all_la.csv")
+
+read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
+  ggplot(aes(x=year)) +
+  geom_line(aes(y=London, color = "OA to London"), size=1) +
+  geom_line(aes(y=cities, color = "OA to nearest city"), size=1) +
+  geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) +
+
+  geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) +
+  geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) +
+  geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) +
+
+  scale_color_discrete(name = "Website density regressions") +
+  scale_linetype_manual(name = "Website density regressions") +
+  theme_minimal()
+
+  scale_color_manual(values = c("red", "blue", "green"))
+
+read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
+  pivot_longer(names_to = "destinations",
+               values_to = "Rsquared",
+               cols = 2:7) %>% 
+  arrange(destinations, year) %>% 
+  ggplot(aes(x=year)) +
+  geom_line(aes(y=Rsquared, group=destinations, colour=destinations)) +
+  geom_textline(aes(y=Rsquared, group=destinations, colour=destinations, 
+                    label=destinations),
+                hjust = 1) +
+  theme(legend.position = "none") +
+  theme_minimal()
 
@@
 
-```{r eval=FALSE, fig.show='animate', ffmpeg.format='gif', dev='jpeg', interval=0.01}
+---
 
-hex.images <- list.files(path = path.images, pattern = "^hex", full.names = T)
-for (i in hex.images){
-  plot(i) # for example
-}
+<br>
+
+**Hierarchy effect**: from main centers to secondary ones -- central places
+
+- Gini coefficient
+
+---
+
+```{r gini_oa, results='asis'}
+path.in <- paste0(path, "/outputs/gini_oa.csv")
+
+read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
+  ggplot(aes(x=year, y=gini.a.n.)) +
+  geom_bar(stat = "identity") + 
+  ylab("Gini") +
+  theme_minimal() +
+  ggtitle("Gini coefficient, number of websites per Output Area") +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ```
 
+---
+
+```{r gini_la, results='asis'}
+path.in <- paste0(path, "/outputs/gini_la.csv")
+
+read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
+  ggplot(aes(x=year, y=gini.a.n.)) +
+  geom_bar(stat = "identity") + 
+  ylab("Gini") +
+  theme_minimal() +
+  ggtitle("Gini coefficient, number of websites per Local Auhtority") +
+  theme(plot.title = element_text(hjust = 0.5)) 
+```
+
+## Spatial attributes, a summary
+
+- Spatial dependency relatively small and constant over time / scales
+
+- At local scale, consistent hotspots over time
+
+- More granular analysis reveals hotspots
+
+- Almost perfect polarisation of web adoption in the early stages at a granular level
+
+- More equally diffused at the Local Authority level
+
+- Plateau overtime
+
+- Distance effect: urban gravitation increases over time and then drops
+
+- Consistent across scales and definitions of urban
+
+# S-shaped diffusion curves
+
+---
+
+```{r UK S, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_uk.png")
+knitr::include_graphics(path.image)
+```
+
+```{r LA_S_gif, eval=F, results='asis', out.width = "60%"}
+path.images <- paste0(path, "/outputs/")
+
+library(magick)
+las <- list.files(path = path.images, pattern = "^lad_", full.names = T) #%>% 
+las_list <- lapply(las, image_read)
+
+## join the images together
+las_joined <- image_join(las_list)
+
+## animate at 2 frames per second
+las_animated <- image_animate(las_joined, fps = 2)
+
+## view animated image
+las_animated
+
+## save to disk
+image_write(image = las_animated,
+            path = "las.gif")
+
+  # map(image_read) %>% # reads each path file
+  # image_join() %>% # joins image
+  # image_animate(delay=50) #%>% # animates, can opt for number of loops
+  #image_write("test.gif") # write to current dir
+```
+
+---
+
+```{r Xmid_frequency_graph, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/xmid_f.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r diffusion_speed_map, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/speed_pam.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+<!-- fast  -->
+
+```{r Darlington, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/lad_E06000005.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r Nottingham, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/lad_E06000018.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r Reading, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/lad_E06000038.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r City_of_London, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/lad_E09000001.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+<!-- slow -->
+
+```{r Middlesbrough, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/lad_E06000002.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r Leicester, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/lad_E06000016.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r Bristol, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/lad_E06000023.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r Plymouth, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/lad_E06000026.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r Liverpool, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/lad_E08000012.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r Hackney, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/lad_E09000012.png")
+knitr::include_graphics(path.image)
+```
+
+---
+
+```{r Edinburgh, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/lad_S12000036.png")
+knitr::include_graphics(path.image)
+```
+
+# Conclusions
+
+## What have we learned?
+
+- **Geography matters:** spatial dependency, urban gravitation
+
+- Some indications of a hierarchical diffusion
+
+- Granular analysis reveals patterns otherwise not visible
+
+- Well-established theoretical approaches of diffusion survive even at a granular level
+
+## What have we *yet* to learn?
+
+- Explain the spatial patterns of fast/slow web adoption
+
+  - Applied the same analysis fo OA
+
+  - Survival regressions [@perkins2005international]
+
+- Expand our definition of web to websites with # of postcodes > 1
+
 ## References {.allowframebreaks}

### 2023-02-21 src/s_la_per_firm.Rmd
lads

@@
 - model LAs
-- add SE for the line plots [link](https://stackoverflow.com/questions/26020142/adding-shade-to-r-lineplot-denotes-standard-error)
+- ~~add SE for the line plots~~ [link](https://stackoverflow.com/questions/26020142/adding-shade-to-r-lineplot-denotes-standard-error). No as we have the points in the plots, so SE is not needed
 
@@
 df <- df %>% left_join(firms) %>% 
-  mutate(n_by_firm = (n*1000)/firms) %>% # '000 websites/firm
+  mutate(n_by_firm = n/firms) %>% # N websites/firm
   #filter(!is.na(firms)) %>% 
@@
 df.country <- df %>% group_by(year) %>% 
-  summarise(n = sum(n, na.rm = T))
+  summarise(n = mean(n, na.rm = T))
 
@@
 
-path.out <- paste0(path, "/outputs/s_uk.png")
+path.out <- paste0(path, "/outputs/s_uk_per_firm.png")
 
@@
   geom_vline(xintercept = 2003) +
-  ylab("N. of websites") +
+  ylab("N. of websites / per firm") +
   ggtitle("S-curve, UK") +
@@
 a.la <- as_tibble()
-all.la <- df %>% filter(ladcd != "NA") %>% distinct(ladcd) %>% simplify2array()
+all.la <- df %>% filter(ladcd != "NA") %>% distinct(ladcd) %>% 
+  #top_n(n=2) %>% 
+  simplify2array()
 simplify2array(all.la)
@@
 # plots
-path.out <- paste0(path, "/outputs/per_firm/lad_")
+path.out <- paste0(path, "/outputs/s_lad_per_firm/lad_")
 
 for (i in all.la){
+
+  xmid_la <- pred %>% left_join(a.la) %>% filter(ladcd == i) %>% dplyr::select(estimate) %>% first()
+  lines <- data.frame(
+    intercepts = c(2003, xmid_la[1]),
+    Xmid = c("UK", i)
+  )
+
 p <- ggplot(pred %>% left_join(a.la) %>% 
@@
   scale_x_continuous("year", labels = as.character(df$year), breaks = df$year) +
-  geom_vline(xintercept = pred %>% left_join(a.la) %>% filter(ladcd == i) %>% dplyr::select(estimate) %>% first()) +
-  ylab("N. of websites") + 
+  #geom_vline(xintercept = pred %>% left_join(a.la) %>% filter(ladcd == i) %>% dplyr::select(estimate) %>% first()) +
+  ylab("N. of websites / per firm") + 
   labs(color = "Website \ndiffusion speed") +
@@
   scale_color_manual(values = c("fast" = "purple",
-                                "slow"="orange")) +
+                                "slow"="orange"),
+                     limits = force) +
+  geom_vline(
+    data=lines,
+    aes(xintercept = intercepts, linetype=Xmid),
+    size=.5, key_glyph='path', colour="grey30"
+    ) +
   theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_vline(xintercept = 2003, colour="grey30", linetype="dashed")
+  theme(plot.title = element_text(hjust = 0.5)) #+
+  #geom_vline(xintercept = 2003, colour="grey30", linetype="dashed")
 
-ggsave(p, filename = paste(path.out, i, ".png", sep = ""))
+ggsave(p, filename = paste(path.out, i, ".png", sep = ""), bg="white")
 

### 2023-02-21 src/la_ranks.Rmd

--- 2023-02-16 src/s_la_per_firm.Rmd
@@
 ---
-title: "s.function"
+title: "LA ranks"
 date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 
-- speed, other metrics? 
-- understand rsquare
-- model LAs
-- add SE for the line plots [link](https://stackoverflow.com/questions/26020142/adding-shade-to-r-lineplot-denotes-standard-error)
+- add 0s in the df and then calculate the ranks
 
@@
 ```
-
-## Spatial data
+## Ranks
 
 ```{r}
-# get LA for the UK
-path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-la <- readOGR(path.geo)
-# source: https://geoportal.statistics.gov.uk/
+library(ggbump)
 
-# spatial transformations
-la <- spTransform(la, CRS("+init=epsg:4326"))
+df %>% group_by(year) %>% 
+  mutate(rank = dense_rank(desc(n))) %>%
+  ungroup() %>% 
+  filter(year==1996 | year == 2012) %>% 
+  mutate(diff96_12 = -diff(rank))
 
-la.f <- fortify(la, region = "LAD21CD")
-
-# cities
-cities <- maps::world.cities %>% 
-  filter(country.etc=="UK") %>% 
-  arrange(pop) %>% tail(10) 
-tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
-```
-
-## Help, not really used
-
-```{r, eval=FALSE}
-
-df.test <- df %>% filter(oa11cd == "E00000007" |
-                         oa11cd == "E00000021" |
-                         oa11cd == "E00000025") %>% 
-  mutate(rownum = row_number(),
-         value2 = n/max(n))
-
-# function to select unique groups
-sample_n_groups = function(grouped_df, size, replace = FALSE, weight=NULL) {
-  grp_var <- grouped_df %>% 
-    groups %>%
-    unlist %>% 
-    as.character
-  random_grp <- grouped_df %>% 
-    summarise() %>% 
-    sample_n(size, replace, weight) %>% 
-    mutate(unique_id = 1:NROW(.))
-  grouped_df %>% 
-    right_join(random_grp, by=grp_var) %>% 
-    group_by_(grp_var) 
-}
-
-df.test <- df %>% group_by(oa11cd) %>% 
-  sample_n_groups(10000) %>% 
-  arrange(oa11cd, year) %>% 
-  select(-unique_id)
-
-fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.test)
-
-fd.nls <- df.test %>% 
-  group_by(oa11cd) %>%
-  select(year, oa11cd, n) %>% 
-  do(fitloop = tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .))) %>% 
-  unnest(fitloop)
-
-fd.nls %>% filter(term=="xmid") %>% 
-  summarise(rng = range(estimate))
-```
-
-## Test for S function estimation
-
-To interpret the coefficients, check [SSlogis](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/SSlogis).
-This is a self-starting model, which estimates the starting values.
-
-```{r}
-df.test <- df %>% filter(ladcd == "E06000002")
-
-fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.test)
-summary(fit)
-
-df.test$prednls = predict(fit)
-
-ggplot(df.test, aes(x = year, y = n) ) +
-  geom_point() +
-  geom_line(aes(y = prednls), size = 1) +
-  scale_x_continuous("year", labels = as.character(df.test$year), breaks = df.test$year) +
-  ylab("N. of websites")
-  #scale_x_date(breaks = df.test$year) 
-```
-
-## S function estimation for all the country
-
-```{r}
-df.country <- df %>% group_by(year) %>% 
-  summarise(n = sum(n, na.rm = T))
-
-fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.country)
-summary(fit)
-
-rsquare(fit, df.country)
-
-df.country$prednls = predict(fit)
-
-path.out <- paste0(path, "/outputs/s_uk.png")
-
-df.country %>% 
-  ggplot(aes(x = year, y = n) ) +
-  geom_point() +
-  geom_line(aes(y = prednls), size = 1) +
-  scale_x_continuous("year", labels = as.character(df.country$year), breaks = df.country$year) +
-  geom_vline(xintercept = 2003) +
-  ylab("N. of websites") +
-  ggtitle("S-curve, UK") +
-  theme(plot.title = element_text(hjust = 0.5))
-
-ggsave(path.out)
-  #scale_x_date(breaks = df.test$year) 
-```
-
-## Loop for S function estimation for Local Authorities
-
-```{r}
-a.la <- as_tibble()
-all.la <- df %>% filter(ladcd != "NA") %>% distinct(ladcd) %>% simplify2array()
-simplify2array(all.la)
-
-for (i in all.la){
-  data <- df %>% filter(ladcd==i) %>%
-    dplyr::select(year, ladcd, n) 
-    #lm(n~year, data = .)
-    #tryCatch(tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .)), , error=function(e) NULL)
-  tryCatch(b.la <- tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = data)), error=function(e) NULL)
-  tryCatch(a.la <-rbind(a.la,b.la), error=function(e) NULL)
-}
-
-a.la %>% filter(term=="xmid") %>% 
-  mutate(estimate = round(estimate, 0)) %>% 
-  dplyr::select(estimate) %>% 
-  summarise(r = range(estimate),
-            median = median(estimate))
-
-a.la %>% filter(term=="xmid") %>% 
-  mutate(estimate = round(estimate, 0)) %>% 
-  filter(p.value < 0.01) %>%    # (227759 - 212782)/227759  # 6% not sig 
-  dplyr::select(estimate) %>% 
-  ggplot(aes(x=estimate)) + 
-  geom_histogram(binwidth = 1) +
-  geom_vline(xintercept = 2003, colour = "red") +
-  #scale_x_continuous("year", labels = round(a$estimate, 0), breaks = round(a$estimate, 0)) +
-  #scale_x_discrete(labels = as.factor(a$estimate))
-  theme_minimal() +
-  ggtitle("Counts of Xmid per year") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  ylab("LAs") + xlab("year")
-
-path.out <- paste0(path, "/outputs/per_firm/xmid_f.png")
-ggsave(path.out)
-
-a.la <- a.la %>% filter(term=="xmid") %>% 
-  bind_cols(as.tibble(all.la)) %>% 
-  mutate(estimate = round(estimate, 0)) %>% 
-  filter(p.value < 0.01) %>%    # 376 / 376, all sig
-  dplyr::select(ladcd, estimate)
-  #filter(estimate>2011)
-```
-
-## Loop for S function estimation for Local Authorities keeping the model objects
-
-```{r}
-a.la <- as_tibble()
-all.la <- df %>% filter(ladcd != "NA") %>% distinct(ladcd) %>% simplify2array()
-simplify2array(all.la)
-
-for (i in all.la){
-  data <- df %>% filter(ladcd==i) %>%
-    dplyr::select(year, ladcd, n) 
-    #lm(n~year, data = .)
-    #tryCatch(tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .)), , error=function(e) NULL)
-  tryCatch(model <- (nls(n ~ SSlogis(year, Asym, xmid, scal), data = data)), error=function(e) NULL)
-  model.name <- paste0("model.", i)
-  assign(model.name, model)
-  tryCatch(b.la <- tidy(model), error=function(e) NULL)
-
-  tryCatch(b.la$r2 <- rsquare(model, data), error=function(e) NULL)
-
-  tryCatch(a.la <-rbind(a.la,b.la), error=function(e) NULL)
-}
-
-a.la <- a.la %>% filter(term=="xmid") %>% 
-  bind_cols(as.tibble(all.la)) %>% 
-  #mutate(estimate = round(estimate, 0)) %>% 
-  #summarise(r = range(r2))
-  filter(p.value < 0.01 & 
-         r2 > 0.9) %>%    # 376 / 376, all sig, 330 r2 > 0.9
-  dplyr::select(ladcd, estimate, r2, std.error) %>%
-  mutate(fast = ifelse(estimate < 2003, "fast", "slow"))
-
-# predictions 
-pred <- tibble()
-help <- tibble(year = 1996:2012)
-for (i in all.la){
-  model.name <- (paste0("model.", i))
-  #assign(col.name, paste0("pred.", i))
-  pred_ <- df %>% filter(ladcd == i) %>% 
-    mutate(prednls = eval(call("predict", as.name(model.name))))  # "pred.{i}" :=
-  help_ <- help %>% left_join(pred_)
-
-  pred <- pred %>% bind_rows(help_) 
-}
-
-# plots
-path.out <- paste0(path, "/outputs/per_firm/lad_")
-
-for (i in all.la){
-p <- ggplot(pred %>% left_join(a.la) %>% 
-              filter(ladcd == i), aes(x = year, y = n, color = fast) ) +
-  geom_point() +
-  geom_line(aes(y = prednls), size = 1) +
-  scale_x_continuous("year", labels = as.character(df$year), breaks = df$year) +
-  geom_vline(xintercept = pred %>% left_join(a.la) %>% filter(ladcd == i) %>% dplyr::select(estimate) %>% first()) +
-  ylab("N. of websites") + 
-  labs(color = "Website \ndiffusion speed") +
-  labs(title = paste0(la@data %>% filter(LAD21CD== i) %>% dplyr::select(LAD21NM) %>% first(), ", ", i)) +
-  theme(legend.position=c(.9,.15)) +
-  scale_color_manual(values = c("fast" = "purple",
-                                "slow"="orange")) +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_vline(xintercept = 2003, colour="grey30", linetype="dashed")
-
-ggsave(p, filename = paste(path.out, i, ".png", sep = ""))
-
-}
-```
-
-## Xmid frequency graph
-
-**TO DELETE**, done it above.
-
-```{r, eval=F}
-# path.out <- paste0(path, "/outputs/xmid_f.png")
-# 
-# ggplot(a.la, aes(as.factor(round(estimate, 0)))) +
-#   geom_bar(fill = "#0073C2FF") +
-#   theme_minimal() +
-#   ggtitle("Counts of Xmid per year") +
-#   theme(plot.title = element_text(hjust = 0.5)) +
-#   ylab("LAs") + xlab("year") 
-# 
-# ggsave(path.out)
-```
-
-write the object with all the models
-
-```{r eval=F}
-path.out <- paste0(path, "/data/temp/s_la.csv")
-write_csv(a.la, path.out)
-```
-
-## Survival
-
-- pop.density
-- income
-- firms
-- employment
-- N/S
-- London
-
-Time series formula will look like this:
-
-`coxph(Surv(tstart, tstop, infect) ~ treat + inherit + steroids, data =newcgd, cluster = id)`
-
-This is what @perkins2005international should be using.
-See p. 7 [here](https://cran.r-project.org/web/packages/survival/vignettes/timedep.pdf).
-
-using `library(survival)`
-
-```{r}
-test.surv.data <- a.la %>% mutate(status = 1) %>% 
-  left_join(data.la, by = c("ladcd" = "LAD21CD")) %>% 
-  left_join(df %>% filter(year == 2012), by = "ladcd")
-
-cf <- coxph(Surv((estimate-1996), status) ~ log(distLondon) + log(distMet) + log(pop16_64) + log(pop.dens) + log(econ.act) + log(earnings) + log(total.busi) + south + log(n), data=test.surv.data, weights = 1/(std.error)^2)
-summary(cf)
+  #filter(rank < 50) %>% 
+  ggplot(aes(year, rank, color = ladcd)) +
+  geom_bump() + 
+  theme(legend.position="none")
 
 ```
-
-## Map the xmid
-
-```{r}
-
-la.f <- la.f %>% left_join(a.la, by = c("id" = "ladcd")) %>% 
-  mutate(fast = ifelse(estimate < 2003, "fast", "slow")) %>% 
-  arrange(order)  # if i don't order merge.nuts.f loses order and the map has gaps
-
-# oa.gb.f %>% mutate(fast = ifelse(estimate < 2003, 1, 0)) %>% 
-#   arrange(order) %>% 
-ggplot(data = la.f, aes(x = long, y = lat)) +
-  geom_polygon(aes(group = group, fill = estimate)) + #fill = fast, 
-  #theme_nothing(legend = TRUE) +
-  labs(title = "Clusters") +
-  scale_fill_continuous(type = "viridis") +
-  theme_void()
-
-path.out <- paste0(path, "/outputs/speed_pam.png")
-ggplot(data = la.f, aes(x = long, y = lat)) +
-  geom_polygon(aes(group = group, fill = fast)) + #fill = fast, 
-  scale_fill_manual(values = c("darkorchid3", "orange"), #darkmagenta, purple4
-                    na.value = "grey70") +
-
-  #theme_nothing(legend = TRUE) +
-  labs(title = "Website diffusion speed in Local authorities") +
-  geom_point(data = cities, aes(x=long, y=lat), colour = "black", size = .1) + 
-  geom_text(size = 2, check_overlap = F, nudge_y = .2, data = cities, aes(x=long, y=lat, label = name)) +
-  #scale_fill_continuous(type = "viridis") +
-  labs(fill='Diffusion') +
-  theme_void() +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-```
-
-```{r}
-la.f %>% left_join(df %>% filter(year==2012), by = c("id" = "ladcd")) %>% 
-  arrange(order) %>% # if i don't order merge.nuts.f loses order and the map has gaps
-  ggplot(aes(x = long, y = lat)) +
-  geom_polygon(aes(group = group, fill = n)) + #fill = fast, 
-  #theme_nothing(legend = TRUE) +
-  labs(title = "Clusters") +
-  scale_fill_continuous(type = "viridis") +
-  theme_void()
-
-```
-
-## time series clusters
-
-2 clusters is the preferred solution, but when I plot individual local authorities 
-I cannot interpret the differences between the two clusters
-
-```{r}
-library(dtwclust)
-library(tsbox)
-library(imputeTS)
-
-# turn to a wide format
-ts.wide <- df %>% 
-  filter(!is.na(ladcd)) %>% 
-  rename(time = year) %>%
-  dplyr::select(time, ladcd, n) %>% 
-  #mutate(year = )
-  ts_wide()
-
-# impute with means of each TS
-# TODO check sensitivity
-ts.wide <- na_mean(ts.wide)
-#sapply(ts.wide, function(x) sum(is.na(x)))
-
-# extract names
-ts.wide.names <- names(ts.wide[,-1])
-ts.wide <- ts.wide[,-1]
-
-# standardise
-ts.wide <- zscore(ts.wide)
-
-# convert to list for tsclust
-ts.wide <- split(ts.wide, rep(1:ncol(ts.wide), each = nrow(ts.wide)))
-
-# 5 - 20 cluster sollutions
-pc_k <- tsclust(ts.wide, k = c(2:10), seed = 94L,
-                type="partitional",
-                distance = "dtw_basic", centroid = "pam", trace = T, 
-                args = tsclust_args(dist = list(window.size = 20L)))
-
-# decide k, see RJ-2019-023 for max/min
-#names(pc_k) <- paste0("k_", c("5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15"))
-names(pc_k) <- paste0("k_", c("2", "3", "4", "5", "6", "7", "8", "9", "10"))
-cvis <- sapply(pc_k, cvi, type = "internal")
-cvis
-# max Sil
-# max SF
-# max CH
-# min DB
-# min DBstar
-# max D
-# min COP
-
-# loop for k decision
-pref.cvi <- NULL
-
-max.cvis <- c(1:3,6)
-for (i in max.cvis){
-  a <- which.max(cvis[i,])
-  pref.cvi <- c(pref.cvi, a)
-}
-
-min.cvis <- c(4:5,7)
-for (i in min.cvis){
-  a <- which.min(cvis[i,])
-  pref.cvi <- c(pref.cvi, a)
-}
-
-print(pref.cvi) # just count how many time each k_* appears, this is the preferred solution
-
-pc_k <- tsclust(ts.wide, k = 7, seed = 94L,  # 7 or 2
-                type="partitional", 
-                distance = "dtw_basic", centroid = "pam", trace = T, 
-                args = tsclust_args(dist = list(window.size = 20L)))
-plot(pc_k)
-
-clusters <- data.frame(LAD=ts.wide.names,
-                           cluster=pc_k@cluster)
-
-```
-
-## Mapping the clusters
-
-```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}
-
-la.f <- la.f %>% left_join(clusters, by = c("id" = "LAD")) %>% 
-  arrange(order)  # if i don't order merge.nuts.f loses order and the map has gaps
-
-ggplot(la.f %>% filter(!is.na(r2)), aes(x = long, y = lat)) +
-  geom_polygon(aes( group = group, fill = as.factor(cluster))) +
-  #theme_nothing(legend = TRUE) +
-  labs(title = "Clusters") +
-  scale_fill_viridis_d()+
-  theme_void()
-
-# ggplot(la.f, aes(x = long, y = lat)) +
-#   geom_polygon(aes( group = group, fill = as.factor(cluster.down))) +
-#   theme_nothing(legend = TRUE) +
-#   labs(title = "Download clusters") +
-#   scale_fill_viridis_d()+
-#   theme_void()
-# 
-# cluster.upL <- c(6, 9, 11, 12, 13)
-# cluster.upS <- c(1, 2, 3, 7, 8, 10)
-#  
-# la.f_ <- la.f #plot the background
-# 
-# cluster.upL.plot <- la.f %>% 
-#   filter(cluster.up %in% cluster.upL) %>% 
-#   ggplot(aes(x = long, y = lat)) +
-#   geom_polygon(data = la.f_, color='grey', fill='white', aes(x=long, y=lat, group=group, colour = 'grey30')) +
-#   geom_polygon(aes( group = group, fill = as.factor(cluster.up))) +
-#   labs(title = "Larger clusters, upload speeds") + 
-#   scale_fill_viridis_d() +
-#   guides(fill=guide_legend(title="Clusters")) +
-#   theme_void()
-# 
-# cluster.upS.plot <- la.f %>% 
-#   filter(cluster.up %in% cluster.upS) %>% 
-#   ggplot(aes(x = long, y = lat)) +
-#   geom_polygon(data = la.f_, color='grey', fill='white', aes(x=long, y=lat, group=group, colour = 'grey30')) +
-#   geom_polygon(aes( group = group, fill = as.factor(cluster.up))) +
-#   labs(title = "Smaller clusters, upload speeds") + 
-#   scale_fill_viridis_d() +
-#   guides(fill=guide_legend(title="Clusters")) +
-#   theme_void()
-# 
-# plot_grid(cluster.upL.plot, cluster.upS.plot, nrow = 1,
-#           align = "h", axis = "l", 
-#           #labels = c(".co.uk / pop., 2000", ".co.uk / pop., 2010", "non .co.uk / pop., 2000", "non .co.uk / pop., 2010", ""), 
-#           #label_size = 12,
-#           greedy = T)
-#           #rel_heights = c(1, 1, .3)) # this is the relative position of the 3 lines
-# 
-# #ggsave("all_maps2.png") # this is the colour map for the preprint
-# out.path.map <- paste0(path, "/paper/v1/figures/map.up.clusters.png")
-# ggsave(out.path.map)#,
-#        #width = 210, #A4
-#        #height = 297, #A4
-#        #units = "mm")
-```

### 2023-02-21 src/s_la.Rmd

@@
 # plots
-path.out <- paste0(path, "/outputs/lad_")
+path.out <- paste0(path, "/outputs/s_lad/lad_")
 
 for (i in all.la){
+
+  xmid_la <- pred %>% left_join(a.la) %>% filter(ladcd == i) %>% dplyr::select(estimate) %>% first()
+  lines <- data.frame(
+    intercepts = c(2003, xmid_la[1]),
+    Xmid = c("UK", i)
+  )
+
 p <- ggplot(pred %>% left_join(a.la) %>% 
@@
   scale_x_continuous("year", labels = as.character(df$year), breaks = df$year) +
-  geom_vline(xintercept = pred %>% left_join(a.la) %>% filter(ladcd == i) %>% dplyr::select(estimate) %>% first()) +
+  #geom_vline(xintercept = pred %>% left_join(a.la) %>% filter(ladcd == i) %>% dplyr::select(estimate) %>% first()) +
   ylab("N. of websites") + 
@@
   scale_color_manual(values = c("fast" = "purple",
-                                "slow"="orange")) +
+                                "slow"="orange"),
+                     limits = force) +
+  geom_vline(
+    data=lines,
+    aes(xintercept = intercepts, linetype=Xmid),
+    size=.5, key_glyph='path', colour="grey30"
+    ) +
   theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_vline(xintercept = 2003, colour="grey30", linetype="dashed")
+  theme(plot.title = element_text(hjust = 0.5)) 
+  #geom_vline(xintercept = 2003, colour="grey30", linetype="dashed")
 
-ggsave(p, filename = paste(path.out, i, ".png", sep = ""))
+ggsave(p, filename = paste(path.out, i, ".png", sep = ""), bg="white")
 

### 2023-03-20 src/s_la_per_firm.Rmd
initial commit all outputs

@@
   ungroup()
+
+df <- df %>% filter(!is.na(ladcd)) %>% 
+  complete(ladcd, year) %>% 
+  replace(is.na(.), 0)
 
@@
 
+xmid.uk <- tidy(fit) %>% filter(term=="xmid") %>% select(estimate) %>% as.numeric()
+
 rsquare(fit, df.country)
@@
 
-```{r}
+```{r, eval=F}
 a.la <- as_tibble()
@@
   dplyr::select(ladcd, estimate, r2, std.error) %>%
-  mutate(fast = ifelse(estimate < 2003, "fast", "slow"))
+  mutate(fast = ifelse(estimate < xmid.uk, "fast", "slow")) # it was 2003
 
@@
   lines <- data.frame(
-    intercepts = c(2003, xmid_la[1]),
+    intercepts = c(xmid.uk, xmid_la[1]), # it was 2003
     Xmid = c("UK", i)
@@
 
-## Xmid frequency graph
-
-**TO DELETE**, done it above.
-
-```{r, eval=F}
-# path.out <- paste0(path, "/outputs/xmid_f.png")
-# 
-# ggplot(a.la, aes(as.factor(round(estimate, 0)))) +
-#   geom_bar(fill = "#0073C2FF") +
-#   theme_minimal() +
-#   ggtitle("Counts of Xmid per year") +
-#   theme(plot.title = element_text(hjust = 0.5)) +
-#   ylab("LAs") + xlab("year") 
-# 
-# ggsave(path.out)
-```
-
 write the object with all the models
@@
 ```{r eval=F}
-path.out <- paste0(path, "/data/temp/s_la.csv")
+path.out <- paste0(path, "/data/temp/s_la_per_firm.csv")
 write_csv(a.la, path.out)

### 2023-03-20 src/help_firms.Rmd
all .png files

---
title: "Untitled"
author: "Emmanouil Tranos"
date: "06/03/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
la <- readOGR("https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/LAD_Dec_2009_UGCB_GB_2022/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson")

la.names <- la@data %>% 
  select(lad09cd, lad09nm) %>% 
  rename(ladcd = lad09cd,
         ladnm = lad09nm)

df <- df %>% left_join(la.names)

```
Firms 2010 - 2022
```{r, eval=F}
firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>% 
  select(GEOGRAPHY_CODE, DATE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
  rename(ladcd = GEOGRAPHY_CODE,
         year = DATE,
         firms = OBS_VALUE)

firms %>% ggplot(aes(year, firms, group = ladcd)) +
  geom_line()
```
Firms 2000 - 2012, old LA

```{r}
library(readxl)
library(httr)
url.firms <- "https://www.ons.gov.uk/file?uri=/businessindustryandtrade/business/activitysizeandlocation/adhocs/15580analysisofenterprisesintheukbylocalauthoritydistrict2004to2021/9980223.xls"

GET(url.firms, write_disk(tf <- tempfile(fileext = ".xls")))

test <- read_excel(tf, 11L:19L, range = "A12:B45", col_names = F)

path <- read_excel(tf)
path %>% 
  excel_sheets() %>% 
  set_names() %>% 
  map(read_excel, path = path)

sheets <- excel_sheets('data.xlsx')

path.excel <- paste0(path, "/data/raw/9980223.xls")
all.sheets <- path.excel %>% 
  excel_sheets() %>% 
  set_names() %>% 
  map(read_excel, path = path.excel, range = "A12:B445", col_names = F) #%>%

f <- as_tibble()
for (y in 11:19){
  help <- all.sheets %>%   .[[y]] %>% #11:19
  rename(help = '...1',
         n = '...2') %>% 
  separate(help, c("ladcd", "ladnm"), sep = " : ") %>% 
  mutate(year = 2021-(y-2),
         ladnm = str_squish(ladnm))
  f <- rbind(f, help)
}
f %>%  ggplot(aes(year, n, group = ladnm, color = ladnm)) +
  geom_line() +
  theme(legend.position="none")

f %>% filter(year == 2008) %>% arrange(desc(n))

f %>% filter(ladnm == "Birmingham")

help.f <- f %>% distinct(ladnm, .keep_all=T)

la.names %>% mutate(ladnm = str_squish(ladnm)) %>% 
 left_join(help.f, by = c("ladnm" = "ladnm")) %>% arrange(ladnm) %>% 
  filter(is.na(n))
```

### 2023-03-20 src/la_ranks_m11.Rmd
all .png files

---
title: "LA ranks"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
library(rprojroot)
library(rgdal)
library(sf)
library(modelr)
library(survival)
library(ggbump)
library(ggrepel)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

## TODO

- add 0s in the df and then calculate the ranks

## Load data

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

```{r}
n = 1 #number of unique postcodes. 
m = 12

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610)

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012)

df <- bind_rows(df9610, df1112) %>% 
  filter(
    V1.domain < m,             # for multiple pc
    #V1.domain == n,           # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, ladcd) %>%
  #summarise(n = n()) %>%            # for multiple pc  
  summarise(n = sum(V1.domain)) %>% # for n == 1 
  ungroup()

df <- df %>% filter(!is.na(ladcd)) %>% 
  complete(ladcd, year) %>% 
  replace(is.na(.), 0)
# tests for Scotland
# df %>% filter(substr(oa11cd, 1,1) =="S",
#               year==2010) %>% summarise(mean(n, na.rm = T))
# 
# df %>% filter(oa11cd=="S00090182")
# 
# test@data %>% filter(substr(id, 1,1) =="S",
#                      n!=0)
# 
# df %>% filter(substr(oa11cd, 1,1) =="W",
#               year==2003)
```

## Firms

```{r}
firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>% 
  select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
  rename(ladcd = GEOGRAPHY_CODE,
         firms = OBS_VALUE)

df <- df %>% left_join(firms) %>% 
  mutate(n_by_firm = n/firms) #%>% # N websites/firm
  #filter(!is.na(firms)) %>% 
  #select(-n, -firms) %>% 
  #rename(n = n_by_firm)

# df %>% filter(is.na(n_by_firm))
# sapply(df %>% left_join(firms), function(x) sum(is.na(x)))
```
## Spatial data

```{r}
# get LA for the UK
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- readOGR(path.geo)

la.names <- la@data %>% 
  select(LAD21CD, LAD21NM) %>% 
  rename(ladcd = LAD21CD,
         ladnm = LAD21NM)

df <- df %>% left_join(la.names)
# source: https://geoportal.statistics.gov.uk/
```

## Ranks websites per firm t and 2012 only

```{r}
t <- 1996

out.path <- paste0(path, "/outputs/ranks/web11_per_firm", t,"_2012_only95.png")

df %>% group_by(year) %>% 
  mutate(rank = rank(desc(n_by_firm))) %>% # or n_by_firm
  ungroup() %>% 
  filter(year==t | year == 2012) %>% 
  arrange(ladcd, year) %>% 
  group_by(ladcd) %>% 
  #mutate(diff_1st = -diff(rank))
  mutate(diff_1st = -(rank-lag(rank))) %>% 
  #ungroup() %>% summarise(q = quantile(abs(diff_1st), .95, na.rm = T)) 
  # t= 2000: q.95 = , q.9 = , q.1 = , q.05 = . **abs()**
  # t= 1996: q.95 = 243.625, q.9 = 191.5, q.1 = 11.5, q.05 = 6. **abs()**

  mutate(alpha = ifelse(abs(diff_1st) > 243.625, .8, 0.1)) %>% 
  mutate(alpha = ifelse(is.na(diff_1st), 0.1, alpha)) %>%
  mutate(label = ifelse(alpha==.1, "", ladnm)) %>%

  ggplot(aes(year, rank, group = ladcd, alpha=alpha)) +
  geom_bump(colour = "red") +
  theme(legend.position="none") +
  scale_y_reverse(labels = c(1, 100, 200, 300, 400)) +
  # geom_text(#data = df %>% filter(year == 2012),
  #           aes(x = Inf, label = label), size = 4, hjust = 0)  # x = 2012 + .1
  geom_text_repel(aes(label = label),
                  na.rm = TRUE, 
                  size = 2,
                  nudge_x = 2, nudge_y = 1,
                  direction = "y", hjust = "left")

ggsave(out.path)
```

## Ranks websites 2000 and 2012 only

```{r}
out.path <- paste0(path, "/outputs/ranks/web_2000_2012_only9.png")

df %>% group_by(year) %>% 
  #mutate(rank = dense_rank(desc(n))) %>% # or n_by_firm
  mutate(rank = rank(desc(n))) %>% # or n_by_firm

  ungroup() %>% 
  filter(year==2000 | year == 2012) %>% 
  arrange(ladcd, year) %>% 
  group_by(ladcd) %>% 
  #mutate(diff_1st = -diff(rank))
  mutate(diff_1st = -(rank-lag(rank))) %>% 
  #ungroup() %>% summarise(q = quantile(abs(diff_1st), .9, na.rm = T)) 
  # q.95 = 87.75, q.9 = 71, q.1 = 3.5, q.05 = 2. **abs()**

  mutate(alpha = ifelse(abs(diff_1st) > 71, .8, 0.1)) %>% 
  mutate(alpha = ifelse(is.na(diff_1st), 0.1, alpha)) %>%
  mutate(label = ifelse(alpha==.1, "", ladnm)) %>%

  ggplot(aes(year, rank, group = ladcd, alpha=alpha)) +
  geom_bump(colour = "red") +
  theme(legend.position="none") +
  scale_y_reverse(labels = c(1, 100, 200, 300, 400)) +
  # geom_text(#data = df %>% filter(year == 2012),
  #           aes(x = Inf, label = label), size = 4, hjust = 0)  # x = 2012 + .1
  geom_text_repel(aes(label = label),
                  na.rm = TRUE, 
                  size = 2,
                  nudge_x = 2, nudge_y = 1,
                  direction = "y", hjust = "left")

ggsave(out.path)

```

## Websites per year

```{r}
out.path <- paste0(path, "/outputs/ranks/web11_per_year.png")

df %>% group_by(year) %>% 
  summarise(n = sum(n)) %>% 
  ggplot(aes(x=year, y = n)) +
  geom_bar(stat = "identity") +
  ylab("N. websites")

ggsave(out.path)
```

## Ranks websites per firm all period

```{r}  
out.path <- paste0(path, "/outputs/ranks/web11_per_firm1996_2012_95.png")

df %>% group_by(year) %>% 
  mutate(rank = rank(desc(n_by_firm))) %>% # or n_by_firm
  ungroup() %>% 
  #filter(year==t | year == 2012) %>% 
  arrange(ladcd, year) %>% 
  group_by(ladcd) %>% 
  #mutate(diff_1st = -diff(rank))
  mutate(diff_1st = -(rank-lag(rank, n=16))) %>% 
  #ungroup() %>% summarise(q = quantile(abs(diff_1st), .95, na.rm = T)) 
  # t= 2000: q.95 = , q.9 = , q.1 = , q.05 = . **abs()**
  # t= 1996: q.95 = 243.625, q.9 = 191.5, q.1 = 11.5, q.05 = 6. **abs()**

  mutate(alpha = ifelse(abs(diff_1st) > 243.625, .8, 0.1)) %>% 
  mutate(alpha = ifelse(is.na(diff_1st), 0.1, alpha)) %>%
  mutate(label = ifelse(alpha==.1, "", ladnm)) %>%

  ggplot(aes(year, rank, group = ladcd, alpha=alpha, color = ladnm)) +
  geom_bump() +
  theme(legend.position="none") +
  scale_y_reverse(labels = c(1, 100, 200, 300, 400)) +
  # geom_text(#data = df %>% filter(year == 2012),
  #           aes(x = Inf, label = label), size = 4, hjust = 0)  # x = 2012 + .1
  geom_text_repel(aes(label = label),
                  na.rm = TRUE, 
                  size = 2,
                  nudge_x = 2, nudge_y = 1,
                  direction = "y", hjust = "left")

ggsave(out.path)
```

### 2023-03-20 src/la_ranks_n1.Rmd
all .png files

--- 2023-03-20 src/la_ranks_m11.Rmd
@@
   filter(
-    V1.domain < m,             # for multiple pc
-    #V1.domain == n,           # for n == 1
+    #V1.domain < m,             # for multiple pc
+    V1.domain == n,             # for n == 1
     year > 1995 & year <2013) %>%   
@@
 ```
-
 ## Ranks websites per firm t and 2012 only
@@
 
-out.path <- paste0(path, "/outputs/ranks/web11_per_firm", t,"_2012_only95.png")
+out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only05.png")
 
@@
   #ungroup() %>% summarise(q = quantile(abs(diff_1st), .95, na.rm = T)) 
-  # t= 2000: q.95 = , q.9 = , q.1 = , q.05 = . **abs()**
-  # t= 1996: q.95 = 243.625, q.9 = 191.5, q.1 = 11.5, q.05 = 6. **abs()**
+  # t= 2000: q.95 = 199.5, q.9 = 168, q.1 = 7, q.05 = 3.75. **abs()**
+  # t= 1996: q.95 = 257.125, q.9 = 232.25, q.1 = 17.75, q.05 = 6.875. **abs()**
 
-  mutate(alpha = ifelse(abs(diff_1st) > 243.625, .8, 0.1)) %>% 
+  mutate(alpha = ifelse(abs(diff_1st) < 6.875, .8, 0.1)) %>% 
   mutate(alpha = ifelse(is.na(diff_1st), 0.1, alpha)) %>%
@@
 ```{r}
-out.path <- paste0(path, "/outputs/ranks/web11_per_year.png")
+out.path <- paste0(path, "/outputs/ranks/web_per_year.png")
 
@@
 ggsave(out.path)
+
 ```
@@
 ```{r}  
-out.path <- paste0(path, "/outputs/ranks/web11_per_firm1996_2012_95.png")
+out.path <- paste0(path, "/outputs/ranks/web_per_firm1996_2012_95.png")
 
@@
   # t= 2000: q.95 = , q.9 = , q.1 = , q.05 = . **abs()**
-  # t= 1996: q.95 = 243.625, q.9 = 191.5, q.1 = 11.5, q.05 = 6. **abs()**
+  # t= 1996: q.95 = 257.125, q.9 = 232.25, q.1 = 17.75, q.05 = 6.875. **abs()**
 
-  mutate(alpha = ifelse(abs(diff_1st) > 243.625, .8, 0.1)) %>% 
+  mutate(alpha = ifelse(abs(diff_1st) > 257.125, .8, 0.1)) %>% 
   mutate(alpha = ifelse(is.na(diff_1st), 0.1, alpha)) %>%

### 2023-03-20 src/s_la_per_firm_m11.Rmd
all .png files

--- 2023-03-20 src/s_la_per_firm.Rmd
@@
   filter(
-    #V1.domain < m,             # for multiple pc
-    V1.domain == n,             # for n == 1
+    V1.domain < m,             # for multiple pc
+    #V1.domain == n,           # for n == 1
     year > 1995 & year <2013) %>%   
@@
 
-xmid.uk <- tidy(fit) %>% filter(term=="xmid") %>% select(estimate) %>% as.numeric()
-
 rsquare(fit, df.country)
@@
 
-path.out <- paste0(path, "/outputs/s_uk_per_firm.png")
+path.out <- paste0(path, "/outputs/s_uk_per_firm_11.png")
 
@@
   scale_x_continuous("year", labels = as.character(df.country$year), breaks = df.country$year) +
-  geom_vline(xintercept = 2003) +
+  geom_vline(xintercept = 2002) +
   ylab("N. of websites / per firm") +
@@
   dplyr::select(ladcd, estimate, r2, std.error) %>%
-  mutate(fast = ifelse(estimate < xmid.uk, "fast", "slow")) # it was 2003
+  mutate(fast = ifelse(estimate < 2002, "fast", "slow"))
 
@@
 # plots
-path.out <- paste0(path, "/outputs/s_lad_per_firm/lad_")
+path.out <- paste0(path, "/outputs/s_la_er_firm_m11/lad11_")
 
@@
   lines <- data.frame(
-    intercepts = c(xmid.uk, xmid_la[1]), # it was 2003
+    intercepts = c(2002, xmid_la[1]),
     Xmid = c("UK", i)
@@
 ```{r eval=F}
-path.out <- paste0(path, "/data/temp/s_la_per_firm.csv")
+path.out <- paste0(path, "/data/temp/s_la_per_firm_11.csv")
 write_csv(a.la, path.out)

### 2023-03-20 src/hex_heatmaps.Rmd
lisa per firm

@@
 }
+
+# geo_sf solution. same as lisa, but squeese hexagons
+# gb %>% st_as_sf(crs = 27700) %>% 
+# ggplot() + 
+#     geom_sf(fill=NA, color = "black", size = 0.1, alpha = 0.8) + #geom_point(data = help.df, aes(x=X,y=Y))
+#     stat_binhex(data = help.df, aes(x=X,y=Y),bins=150) + 
+#     # geom_polygon(data = gb.df, aes(x=long, y = lat, group = group), #gb.df
+#     #              fill=NA, color = "black", size = 0.1, alpha = 0.8)+
+#     
+#     #coord_equal() +
+#     scale_fill_gradientn(colours=c('light gray','blue'),
+#                          name='Frequency',na.value=NA) + #, trans = "log10") +
+#     
+#     labs(x=NULL, y=NULL) +
+#     geom_point(data = cities, aes(x=long, y=lat), colour = "black", size = .1) + 
+#     geom_text(size = 2, check_overlap = F, nudge_y = .2, data = cities, aes(x=long, y=lat, label = name)) +
+#     #theme(legend.key.size = unit(0.05, "cm")) + # legend.key.width=unit(1, "cm"), 
+#     ggmap::theme_nothing(legend = T) +
+#     theme(legend.title=element_text(size=5.5), legend.text=element_text(size=4)) +
+#     #theme(legend.position=c(0.075,0.28))
+#     guides(fill = guide_colorbar(barwidth = 0.5, barheight = 2.5)) +
+#     ggtitle(t)
 ```

### 2023-03-20 src/level_lisa_la.Rmd
lisa per firm

--- 2022-05-30 src/level_lisa_la.Rmd
@@
 library(broom)
+library(foreach)
+library(raster)
 
@@
 
+df <- df %>% filter(!is.na(ladcd)) %>% 
+  complete(ladcd, year) %>% 
+  replace(is.na(.), 0)
+
 # tests for Scotland
@@
 #               year==2003)
+```
+
+## Firms
+
+```{r}
+firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>%
+  select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
+  rename(ladcd = GEOGRAPHY_CODE,
+         firms = OBS_VALUE)
+
+df <- df %>% left_join(firms) %>% 
+  mutate(n_by_firm = n/firms) %>% # N websites/firm
+  #filter(!is.na(firms)) %>% 
+  select(-n, -firms) %>% 
+  rename(n = n_by_firm)
+
+# df %>% filter(is.na(n_by_firm))
+# sapply(df %>% left_join(firms), function(x) sum(is.na(x)))
 ```
@@
 
-path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_la") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_la")  # for multiple PC
+path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", n, "_la") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_la")  # for multiple PC
 
@@
 
+## lisa maps parallel
+
+```{r}
+
+path.out <- paste0(path, "/outputs/lisa/lisa_level_pc_test", n, "_la") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_la")  # for multiple PC
+
+#setup cluster
+#https://www.blasbenito.com/post/02_parallelizing_loops_with_r/
+n.cores <- parallel::detectCores() - 1
+my.cluster <- parallel::makeCluster(
+  n.cores, 
+  type = "PSOCK"
+  )
+print(my.cluster)
+doParallel::registerDoParallel(cl = my.cluster)
+foreach::getDoParWorkers()
+
+morani <- data.frame()
+
+#spatial weight matrix
+sf_use_s2(FALSE)
+nb <- poly2nb(la)
+listw <- nb2listw(nb, style = "W", zero.policy = T)
+
+#for (t in 1996:2012){
+foreach(t = 1996:2012, .packages = c("tidyverse",
+                                     "rgdal",
+                                     "rgeos",
+                                     "spdep",
+                                     "tmap",
+                                     "maptools",
+                                     "knitr",
+                                     "broom")) %dopar% {
+test <- la
+test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
+  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
+         year = ifelse(is.na(year), t, year))
+
+#spatial weight matrix
+#moved out of the loop
+
+#Moran's I
+globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
+a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
+morani <- rbind(a, morani)
+
+#LISA
+lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
+#lmoran
+
+# centers the local Moran's around the mean
+lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
+lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
+
+# centers the variable of interest around its mean
+
+lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
+lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
+
+signif <- 0.05
+#lmoran
+
+lmoran <- lmoran%>% 
+  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
+  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
+  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
+  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
+  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
+  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%
+
+#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
+# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
+#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
+#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
+#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))
+
+mun_merge_new<- merge(test, lmoran, by="LAD21CD")
+
+# R p value map
+breaks = c(0, 1, 2, 3, 4, 5)
+
+cities <- maps::world.cities %>% 
+  filter(country.etc=="UK") %>% 
+  arrange(pop) %>% tail(10) 
+tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
+
+map <- tm_shape(mun_merge_new, projection = "+init=epsg:4326") + tm_fill(col = "quadrant", breaks = breaks, 
+                                         palette = c("white","red","blue",
+                                                     rgb(0,0,1,alpha=0.4),
+                                                     rgb(1,0,0,alpha=0.4)), 
+                                         labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
+                                         title="LISA for level, LA") +
+  tm_legend(text.size = 1) +
+  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
+  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
+  #tm_borders(alpha=.1) +
+  tm_borders(lwd = 0) +
+  tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
+  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
+  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ") +
+tmap_options(check.and.fix = TRUE)
+tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
+}
+
+#ggplot map solution.
+# mun_merge_new %>% st_as_sf(crs = 27700) %>% #st_transform(4326) %>%
+#   ggplot()+
+#   geom_sf(aes(fill = as.factor(quadrant)))
+# ggsave("test_ggplot.png", path = path)
+```
+
 Moran I
@@
 ```{r}
-path.out <- paste0(path, "/outputs/morani_la.csv")
+path.out <- paste0(path, "/outputs/lisa/morani_la.csv")
 morani %>% as_tibble() %>% 
@@
   ggtitle("Regression fit, distance to London on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(hjust = 0.5))
 ggsave(path.out)
@@
   ggtitle("Regression coefficients, distance to London on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ggsave(path.out)
@@
   ggtitle("Regression fit, distance to nearest city on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ggsave(path.out)
@@
   ggtitle("Regression coefficients, distance to nearest city on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ggsave(path.out)
@@
 for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = df[df$year==t,])
+  model <- lm(log(n+0.001) ~ log(dist.retail+0.001), data = df[df$year==t,]) #density
   r2 <- as.data.frame(summary(model)$r.squared)
@@
   ggtitle("Regression fit, distance to nearest retail centre on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ggsave(path.out)
@@
   ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  ggtitle("Regression coefficients, distance to nearest retail centre on website density, LA") +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ggsave(path.out)

### 2023-03-20 src/hex_heatmaps.Rmd
regressions per firm

@@
     ggmap::theme_nothing(legend = T) +
-    theme(legend.title=element_text(size=5.5), legend.text=element_text(size=4)) +
+    theme(legend.title=element_text(size=5.5), legend.text=element_text(size=4),
+          plot.title = element_text(hjust = 0.5)) +
     #theme(legend.position=c(0.075,0.28))

### 2023-03-20 src/level_lisa_la.Rmd
regressions per firm

@@
 }
+
+#ggplot map solution.
+# mun_merge_new %>% st_as_sf(crs = 27700) %>% #st_transform(4326) %>%
+#   ggplot()+
+#   geom_sf(aes(fill = as.factor(quadrant)))
+# ggsave("test_ggplot.png", path = path)
+
 ```
@@
 
-#ggplot map solution.
-# mun_merge_new %>% st_as_sf(crs = 27700) %>% #st_transform(4326) %>%
-#   ggplot()+
-#   geom_sf(aes(fill = as.factor(quadrant)))
-# ggsave("test_ggplot.png", path = path)
 ```

### 2023-03-21 src/level_lisa_la.Rmd
level_lisa_la.Rmd

@@
 firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>%
-  select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
+  dplyr::select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
   rename(ladcd = GEOGRAPHY_CODE,
@@
   #filter(!is.na(firms)) %>% 
-  select(-n, -firms) %>% 
+  dplyr::select(-n, -firms) %>% 
   rename(n = n_by_firm)
@@
 for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(London), data = df[df$year==t,])
+  model <- lm(log(n+0.001) ~ log(London), data = df[df$year==t,]) #density
   r2 <- as.data.frame(summary(model)$r.squared)

### 2023-03-21 Untitled/Untitled.Rmd
paper first draft

--- 2023-02-16 paper/paper.Rmd
@@
 ---
-title: Title of submission to PLOS journal
+title: Digital economy in the UK: a multi-scalar story of the diffusion of web technologies
 author:
@@
     affiliation: University of Bristol and The Alan Turing Institute
-    corresponding: e.tranos@bristol.ac.uk
+    correspondingauthor: true
+    footnote: 1
 address:
@@
     address: University of Bristol, School of Geographical Sciences, Bristol, UK; The Alan Turing Institute, London, UK
+footnote:
+  - code: 1
+    text: "This is the first author footnote."
+  - code: 2
+    text: "Another author footnote."
 abstract: |
-  blah blah
+  This paper maps the participation in the digital economy and its evolution in the UK over space and time.
+Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects.
+Useful as these approaches might have been, they cannot capture the micro-processes and the characteristics of the individual online behaviour.
+Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK.
+The data and geolocation strategy allow to capture these processes at a very granular spatial scale.
+The modelling approach, which is based on simple spatial analytical methods and on the estimation of diffusion curves at various scales, enables to depict the role of geography and other cognitive factors which drove the diffusion of web technologies.
+Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.
 
-author_summary: |
-  blah blah
+keywords: 
+  - keyword1
+  - keyword2
+journal: "An awesome journal"
+date: "`r Sys.Date()`"
+classoption: preprint, 3p, authoryear
 bibliography: mybibfile.bib
-output: rticles::plos_article
-csl: plos.csl
-header-includes:
-- \usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}
-editor_options: 
-  markdown: 
-    wrap: 72
+linenumbers: false
+numbersections: true
+# Use a CSL with `citation_package = "default"`
+# csl: https://www.zotero.org/styles/elsevier-harvard
+output: 
+  rticles::elsevier_article:
+    keep_tex: true
+    citation_package: natbib
 ---
-
-```{r echo=FALSE, message = FALSE, warning=FALSE}
-library(stargazer)
-library(tidyverse)
-#library(rmarkdown)
-```
-
 # Introduction {#sec:1}
 
+Geographers were always interested in how new technologies and innovations diffuse across space and, importantly, how such spatio-temporal processes can be modelled.
+The seminal contribution of @hagerstrand1968innovation is illustrative of this early interest.
+However, the torch of exploring and modelling such processes had been passed to other disciplines such as economics, business studies and sociology well before the 'cultural turn' of economic geography [@perkins2005international].
+A potential explanation of lack of studies exploring the diffusion of new and, more specifically, digital technologies across *both* space and time can be attributed to the scarcity of relevant and granular enough data.
+As @zook2022mapping highlight, digital activities are hardly ever captured in official data.
+
+This paper aims to reinstate geographers' interest in how new and digital technologies diffuse over space and time.
+It adopts a geographic standpoint to explore and map the diffusion of web technologies in the UK over space and time.
+It does so by employing a novel source of big data which captures the active engagement with web technologies during the $1996-2013$ period.
+
+The motivation for this paper lies in the fact that there are various stakeholders who are interested in knowing how new digital technologies diffused over space and time use this information to make predictions regarding the diffusion of related future technologies.
+As per @meade2021modelling a variety of actors have a direct interest in gaining such knowledge including network equipment suppliers; network operators, regulatory and local authorities.
+And despite the spatial heterogeneity of such diffusion mechanisms and the policy relevance, there are very limited attempts in the literature to analyse the diffusion of new digital technologies at a granular level.
+
+the paper answers an empirical question: how did a
+
+@zook2022mapping Because new digital activities are rarely---if
+ever---captured in official state data, researchers must rely on
+information gathered from alternative sources. With this in mind, the
+overarching aim of this article is outlining an approach for analyzing
+the "new spaces and geographies" of digital phenomena and practices.
+
+<!-- motivation -->
+
+Classifying adopters of a technology along a time axis was pioneered by
+Rogers (1962), who categorized adopters into five groups as innovators,
+early adopters, early majority, late majority and laggards.
+
+An a priori approach groups similar countries into segments, where
+similarity is defined using economic, social, or political covariates.
+Examples include Hofstede (2001) and Lee (1990). An a posteriori
+analysis considers segments based on realized market behaviour, for
+example see Sood et al. (2009). Some studies investigate changing
+segment membership over time, see Cannon and Yaprak (2011).
+
+@fritsch2015new
+Hägerstrand (1952, 1967) was the first scholar to intensively analyze
+spatial diffusion pro cesses. Based on the empirical observations of the
+regional spread of process innovations, he hypothesized that new
+technology is first implemented in the 'center' -- the large agglom
+erations -- and then moves 'down' the spatial hierarchy, the last stop
+being the 'periphery', i.e. remote and sparsely populated regions. There
+are at least two factors that may explain this pattern: first, the
+transfer of tacit knowledge via face-to-face contact and second, the
+absorptive capacity of a region.
+
+<!-- key concepts -->
+@bento2015spatial The concept of absorptive capacity has also been
+applied in the TIS context in terms of the con-struction of capacity
+needed to transfer and implement a new technology in the receiving
+country (cf.Van Alphen, 2011), with repercussions for the nature of
+system functions
+
+@bednarz2020pulled 
+larger local technological niches (at least from a demand perspective)
+are more likely to be formed in urban areas than in rural ones, which is
+in line with the work of Ha¨gerstrand (1965a,b). However, size is not
+the only regional characteristic that may matter in this context.
+Regions also differ in accumulated experiences and the presence of tacit
+knowledge with respect to products, as well as in actors' propensity
+share this knowledge among producers (Martin et al., 2019)
+
+<!-- @das2022diffusion -->
+<!-- Epidemic approach: The approach emphasizes that the adoption of a -->
+<!-- technology spreads like an infection among the potential adopters. By -->
+<!-- taking a homogeneous group of potential adopters as an assumption, it -->
+<!-- explains the adoption of any innovation. -->
+
+To summarize the discussion, first, here the diffusion process is
+selfperpetuating as initial use stimulates further use. Second, it
+follows a disequilibrium path as the level of users is always lower than
+the number of potential users (Stoneman 2002).
+
+The criticisms are mainly based on the fact that although the approach
+gives an idea of aggregate (industry or household) behaviour, it does
+not focus on the individual's (firm or household) adoption process
+
+<!-- Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.) -->
+<!-- Early adopters: next 13.5%, (III.) Early Majority: following 34%, (IV.) -->
+<!-- Late Majority: next 34%, and the rest as (V.) Laggards -->
+
+@morrill2020spatial Spatial diffusion is the process by which behavior
+or characteristics of the landscape change as a result of what happens
+elsewhere earlier. Spatial diffusion is the spread of the phenomenon,
+over space and timed, from limited origins
+
+Contagious diffusion=> spatial autocorelation
+
+expansion-type diffusion
+
+<!-- the three geographic diffusion processes are (1) purely contagious where -->
+<!-- distance or adjacency is the absolute controlling factor, (2) purely -->
+<!-- hierarchical where size or urban position in the central place hierarchy -->
+<!-- is the absolute controlling factor, or (3) where the location of change -->
+<!-- is purely random. These are the pure or ideal forms of the diffusion -->
+<!-- process; all real diffusion processes are the result of a combination of -->
+<!-- these extremes. -->
+
+The work by the communication sociologist Everett Rogers (1962, 1971,
+1983) has emphasized the role of information, communication, formal and
+informal media, opinion leaders and social networks, and economic and
+psychological constraints on acceptance. Rogers's work stresses the
+decision mechanism of the potential adopter; the work of Rogers
+partitions the adopter's process of choosing to accept a new phenomenon
+or trait into five stages: stage 1--the potential adopter gains
+knowledge or awareness of an innovation; stage 2--persuasion is
+exercised to adopt; stage 3--a decision is made to adopt; stage 4--the
+decision to adopt is implemented; stage 5--the adopter confirms the
+decision to adopt. In Rogers's stage theory there may be a substantial
+time lag between when a potential adopter becomes aware of the new
+characteristic and when a decision is actually made to adopt.
+
+Hägerstrand considered diffusion to be a fundamental geographic process:
+Whatever the phenomenon being diffused might be, one may consider it in
+the context of a larger universal process of spatial diffusion.
+
+<!-- @das2022diffusion [not good paper] Invention, innovation and diffusion -->
+<!-- are considered as the three pillars of the theory of technological -->
+<!-- change and Schumpeter is considered as the father of the study of -->
+<!-- diffusion -->
+
+@leibowicz2016representing
+Diffusion is a spatial (Hagerstrand, 1967) as well as a temporal
+process, and historical evidence confirms that technologies diffuse at
+different times, at different rates, and to different extents in
+different places, and can be significantly influenced by policies
+(Victor, 1993).
+
+<!-- @leibowicz2016representing -->
+<!-- An illustrative empirically based theory for technology -->
+<!-- diffusion across multiple regions is suggested by "Schmidt's Law" -->
+<!-- (Grubler, 1990). With respect to a specific technology, it divides -->
+<!-- regions into three groups: core, rim, and periphery. The core is where -->
+<!-- the technology is invented, subject to experimentation, and first -->
+<!-- deployed commercially. To develop the technology, the core must have -->
+<!-- both sufficient innovative ability and economic motivation to make the -->
+<!-- necessary resource commitment. As the technology spreads from the core -->
+<!-- to the rim and eventually to the periphery, later adopters have the -->
+<!-- advantage of benefiting from the experiences of earlier adopters via -->
+<!-- knowledge spillover effects (Verdolini and Galeotti, 2011). At this -->
+<!-- point, the technology has likely been refined and costs have likely -->
+<!-- declined significantly. Schmidt's Law claims that, as a result, -->
+<!-- diffusion proceeds at a faster rate in regions which adopt the -->
+<!-- technology later. However, diffusion tends to be less pervasive in -->
+<!-- regions where adoption is delayed, ultimately saturating at a lower -->
+<!-- extent than in regions where adoption occurred earlier. This is often -->
+<!-- attributed to a lack of proper institutions and infrastructures in later -->
+<!-- adopters, markets whichmay not be as favorable to a technology, or the -->
+<!-- availability of yet newer options that invite technological -->
+<!-- leapfrogging. -->
+
+Lastly, it can be globally optimal for innovative economies to deploy
+advanced technologies more than what is locally optimal if this enables
+faster diffusion in less advanced regions that could benefit from the
+technology.
+
+wilson201281 Early on in their lifecycle, new technologies are crude, imperfect, and
+expensive (Rosenberg, 1994). New energy technologies are attractive for
+their ability to perform a particular task or deliver a new or improved
+energy service (Fouquet, 2010). This is often circumscribed by a
+particular set of needs in a particular context: a market 'niche'.
+End-users in niche markets are generally less sensitive to the effective
+price of the energy service provided or have a higher willingness to pay
+for its performance advantages (Fouquet, 2010). Thus initially,
+performance dominates cost competitiveness (Wilson and Grubler, 2011).
+Market niches afford some protection from competitive pressures,
+allowing technologies to be tested and improved in applied settings,
+reducing uncertainties with performance or market demand (Kemp et al.,
+1998). Costs may only fall substantively after an extended period of
+commercial experimentation, concurrent with the establishment of an
+industrial base and characteristic moves towards standardisation and
+mass production (Grubler, 1998). The influence of accumulating
+production experience on costs is captured by the concept of learning.
+
+<!-- @grubler1990rise Following the definition given by Rogers, 1983, -->
+<!-- diffusion is the process by which an innovation is communicated through -->
+<!-- certain communication channels over time among the members of a social -->
+<!-- system. An innovation may be an idea, object or practice which is -->
+<!-- perceived as new by an individual or another unit of adoption (e.g., an -->
+<!-- organization) . Communication via communication channels refers to how -->
+<!-- knowledge about an innovation is transferred between the members of a -->
+<!-- social system. Communication takes time, and is therefore an important -->
+<!-- aspect of innovation diffusion. In fact, all theoretical and empirical -->
+<!-- diffusion studies agree that an innovation does not instantly diffuse -->
+<!-- into a social system. Instead a typical time pattern of diffusion along -->
+<!-- a S-shaped trajectory seems to be the rule. The S-shaped pattern of -->
+<!-- diffusion appears to be a basic anthropologic phenomenon, as it is also -->
+<!-- confirmed by a large number of studies· of primitive preindustrial -->
+<!-- societies. -->
+
+grubler1990rise
+The scale of spatial innovation diffusion is similar to that of
+temporal diffusion models: hierarchically decomposed.
+
+<!-- grubler1990rise -->
+<!-- Spatial innovation diffusion in Hagerstrand's model consists of three -->
+<!-- stages. Stage 1: Local concentrations of initial acceptances (initial -->
+<!-- agglomeration) . Stage 2: Radial dissemination outward from the initial -->
+<!-- agglomeration, accompanied by the rise of secondary agglomerations. -->
+<!-- Stage 3: The condensing and saturation stage, in which filling-in occurs -->
+<!-- and diffusion eventually ceases. -->
+
+<!-- grubler1990rise -->
+<!-- Thus, three main characteristics define the spatial diffusion process: -->
+<!-- (1) As in temporal diffusion models, an S-shaped pattern in the -->
+<!-- cumulative level of adoption. (2) A "hierarchy effect" by which -->
+<!-- diffusion proceeds from initial centers to secondary ones. Within the -->
+<!-- context of urban systems (central places) this implies a hierarchical -->
+<!-- spread of an innovation from larger to smaller centers, and from the -->
+<!-- centers to the surrounding hinterland. (3) A "neighborhood effect" Le., -->
+<!-- diffusion proceeds outwards from innovation centers, first "hitting" -->
+<!-- (Brown, 1981) nearby rather than far-away locations. -->
+
+@grubler1990rise
+Later Hagerstrand conceptualized physical "barrier" effects like lakes
+or uninhabited areas, which, in addition to distance, act as further
+retarding effects on diffusion. These are formalized in the form of
+"zero" or "half" contact multiplicators on the (distance decaying)
+message flows.
+
+@grubler1990rise
+With respect to the formalization of the communication flows Hagerstrand
+defines a "mean information field" (MIF), in which the probability of
+communication is a negative function of distance between individuals
+
+@perkins2005international
+Theoretical models disagree as to why firms adopt innovations at
+different times. Epidemic models emphasize information (Griliches 1957).
+Certain firms are hypothesized to adopt earlier because they come into
+contact with, and learn from, adopters of the new technology before
+others. Economic models, on the other hand, predominantly emphasize firm
+heterogeneity (Ireland and Stoneman 1986). Firms adopt technologies at
+different times because they differ with respect to various
+organizational and environmental variables influencing the economic
+returns from adoption (Blackman 1999).
+
+Either way, the strong assumption is that developing countries can
+acquire modern technology innovated in developed economies, often at a
+fraction of the original research and development (R&D) costs, thereby
+leapfrogging many decades of technological progress (Teece 2000).
+
+The first involves a country's geographical location. Recent empirical
+work suggests that diffusion is ''geographically localized'' (Globerman,
+Kokko, and Sjo¨holm 2000; Keller 2002; Milner 2003) in that a technology
+diffuses faster in a country where it is already more widely diffused in
+neighboring countries. Underlying these regional effects are contagion
+and contact with prior users or producers of technology.
+
+@ding2010modeling Rogers (1995) characterized early adopters as knowledgeable risk takers.
+Griliches (1957) looked at the rate of return for early adopters and
+characterized them as profit maximizers. The geographer Hagerstrand
+(1967) was focusing on the character of early adopters in the
+transmittal process that produces follow-on participants in the process.
+Space was treated as a contiguity system with embedded characteristics
+producing barrier elements to the diffusion process.
+
+@ding2010modeling
+REVIEW ON Technological Diffusion
+
+<!-- Hagerstrand (1968) was the first to demonstrate that diffusion typically -->
+<!-- starts from innovative centers ('core´) and disseminates, through a -->
+<!-- hierarchy of subcenters, to the periphery. This process is sequential -->
+<!-- rather than simultaneous, and tends to accelerate from core to -->
+<!-- periphery, where diffusion reaches a lower intensity than in the core -->
+<!-- (Hagerstrand, 1968; Grubler, 1990; Morrill, 2005). A similar effect of -->
+<!-- spatial acceleration has been identified in marketing research and is -->
+<!-- known as the "lead-lag effect" (Peres et al., 2010). -->
+
+@perkins2011internet Two main mechanisms are identified in the
+literature to explain diffusion: (1) epidemic-type dynamics whereby
+contact with previous adopters stimulates uptake as po tential adopters
+learn about a new innovation; and (2) economic-type mechanisms whereby
+potential users adopt a new innovation as it be comesmore profitable,
+useful, or valuable, with uptake characteristically spreading as costs
+be come lower, performance improves, or the po tential uses of the
+innovation grow over time.
+
+To this extent, our results contribute to a growing body of work that
+has sought to caution against claims about the supposed novelty of the
+Internet and the suggestion that it is somehow different.
+
+<!-- methods -->
+@wilson201281 Logistic growth describes an initial period of gradual
+diffusion as a technology is introduced as a new commercial application,
+moving then through a rapid, exponential growth phase, before slowing
+and eventually saturating [@grubler1999dynamics]. The substitution of
+incumbent technologies by new competitors leads to subsequent decline
+and eventual obsolescence.
+
+# Literature review {#sec2} MAYBE SKIP THIS??
+
+<!-- previous studies -->
 @beardsell1999spatial To characterize evolution of the computer
@@
 
-larger local technological niches (at least from a demand perspective)
-are more likely to be formed in urban areas than in rural ones, which is
-in line with the work of Ha¨gerstrand (1965a,b). However, size is not
-the only regional characteristic that may matter in this context.
-Regions also differ in accumulated experiences and the presence of tacit
-knowledge with respect to products, as well as in actors' propensity
-share this knowledge among producers (Martin et al., 2019)
+@perkins2005international analyze whether the rate at which new producer
+technologies diffuse is significantly influenced by (1) latecomer
+advantage and (2) engagement with the global economy via trade and
+foreign investment
 
-@bento2015spatial The concept of absorptive capacity has also been
-applied in the TIS context in terms of the con-struction of capacity
-needed to transfer and implement a new technology in the receiving
-country (cf.Van Alphen, 2011), with repercussions for the nature of
-system functions
+Indeed, precisely because of these latecomer advantages, developing
+countries1 are believed to be well placed to catch up with developed
+ones (Gerschenkron 1962; Abramovitz 1986).
 
-@das2022diffusion [not good paper] Invention, innovation and diffusion
-are considered as the three pillars of the theory of technological
-change and Schumpeter is considered as the father of the study of
-diffusion
+@fritsch2015new analyze the spatial diffusion of laser technology
+research in West Germany from 1960, when this technology began, until
+2005.
 
-Epidemic approach: The approach emphasizes that the adoption of a
-technology spreads like an infection among the potential adopters. By
-taking a homogeneous group of potential adopters as an assumption, it
-explains the adoption of any innovation.
+@perkins2011internet
 
-To summarize the discussion, first, here the diffusion process is
-selfperpetuating as initial use stimulates further use. Second, it
-follows a disequilibrium path as the level of users is always lower than
-the number of potential users (Stoneman 2002).
+@lengyel2020role We see that the OSN spread almost exclusively from the
+original location (the capital Budapest, with an order of magnitude more
+inhabitants than the next size town) to various parts of the country in
+the early phase of the life-cycle. Later, diffusion became less
+mono-centric and other towns also emerged as spreaders. Our findings
+support the idea that spreading initially happens to large distances and
+becomes more local over time. This is illustrated and discussed later in
+Figure 2I.
 
-The criticisms are mainly based on the fact that although the approach
-gives an idea of aggregate (industry or household) behaviour, it does
-not focus on the individual's (firm or household) adoption process
-
-Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.)
-Early adopters: next 13.5%, (III.) Early Majority: following 34%, (IV.)
-Late Majority: next 34%, and the rest as (V.) Laggards
-
-@morrill2020spatial Spatial diffusion is the process by which behavior
-or characteristics of the landscape change as a result of what happens
-elsewhere earlier. Spatial diffusion is the spread of the phenomenon,
-over space and timed, from limited origins
-
-Contagious diffusion=\> spatial autocorelation
-
-expansion-type diffusion
-
-the three geographic diffusion processes are (1) purely contagious where
-distance or adjacency is the absolute controlling factor, (2) purely
-hierarchical where size or urban position in the central place hierarchy
-is the absolute controlling factor, or (3) where the location of change
-is purely random. These are the pure or ideal forms of the diffusion
-process; all real diffusion processes are the result of a combination of
-these extremes.
-
-The work by the communication sociologist Everett Rogers (1962, 1971,
-1983) has emphasized the role of information, communication, formal and
-informal media, opinion leaders and social networks, and economic and
-psychological constraints on acceptance. Rogers's work stresses the
-decision mechanism of the potential adopter; the work of Rogers
-partitions the adopter's process of choosing to accept a new phenomenon
-or trait into five stages: stage 1--the potential adopter gains
-knowledge or awareness of an innovation; stage 2--persuasion is
-exercised to adopt; stage 3--a decision is made to adopt; stage 4--the
-decision to adopt is implemented; stage 5--the adopter confirms the
-decision to adopt. In Rogers's stage theory there may be a substantial
-time lag between when a potential adopter becomes aware of the new
-characteristic and when a decision is actually made to adopt.
-
-Hägerstrand considered diffusion to be a fundamental geographic process:
-Whatever the phenomenon being diffused might be, one may consider it in
-the context of a larger universal process of spatial diffusion.
-
-@zook2022mapping Because new digital activities are rarely---if
-ever---captured in official state data, researchers must rely on
-information gathered from alternative sources. With this in mind, the
-overarching aim of this article is outlining an approach for analyzing
-the "new spaces and geographies" of digital phenomena and practices.
+@PAPAGIANNIDIS2015308
 
@@
 
-Diffusion is a spatial (Hagerstrand, 1967) as well as a temporal
-process, and historical evidence confirms that technologies diffuse at
-different times, at different rates, and to different extents in
-different places, and can be significantly influenced by policies
-(Victor, 1993). An illustrative empirically based theory for technology
-diffusion acrossmultiple regions is suggested by "Schmidt's Law"
-(Grubler, 1990). With respect to a specific technology, it divides
-regions into three groups: core, rim, and periphery. The core iswhere
-the technology is invented, subject to experimentation, and first
-deployed commercially. To develop the technology, the core must have
-both sufficient innovative ability and economic motivation to make the
-necessary resource commitment. As the technology spreads from the core
-to the rim and eventually to the periphery, later adopters have the
-advantage of benefiting from the experiences of earlier adopters via
-knowledge spillover effects (Verdolini and Galeotti, 2011). At this
-point, the technology has likely been refined and costs have likely
-declined significantly. Schmidt's Law claims that, as a result,
-diffusion proceeds at a faster rate in regions which adopt the
-technology later. However, diffusion tends to be less pervasive in
-regions where adoption is delayed, ultimately saturating at a lower
-extent than in regions where adoption occurred earlier. This is often
-attributed to a lack of proper institutions and infrastructures in later
-adopters, markets whichmay not be as favorable to a technology, or the
-availability of yet newer options that invite technological
-leapfrogging.
-
+@leibowicz2016representing
 Empirical evidence supports the validity of Schmidt's Law over a wide
@@
 
-Lastly, it can be globally optimal for innovative economies to deploy
-advanced technologies more than what is locally optimal if this enables
-faster diffusion in less advanced regions that could benefit from the
-technology.
-
-@wilson201281 Logistic growth describes an initial period of gradual
-diffusion as a technology is introduced as a new commercial application,
-moving then through a rapid, exponential growth phase, before slowing
-and eventually saturating [@grubler1999dynamics]. The substitution of
-incumbent technologies by new competitors leads to subsequent decline
-and eventual obsolescence.
-
-Early on in their lifecycle, new technologies are crude, imperfect, and
-expensive (Rosenberg, 1994). New energy technologies are attractive for
-their ability to perform a particular task or deliver a new or improved
-energy service (Fouquet, 2010). This is often circumscribed by a
-particular set of needs in a particular context: a market 'niche'.
-End-users in niche markets are generally less sensitive to the effective
-price of the energy service provided or have a higher willingness to pay
-for its performance advantages (Fouquet, 2010). Thus initially,
-performance dominates cost competitiveness (Wilson and Grubler, 2011).
-Market niches afford some protection from competitive pressures,
-allowing technologies to be tested and improved in applied settings,
-reducing uncertainties with performance or market demand (Kemp et al.,
-1998). Costs may only fall substantively after an extended period of
-commercial experimentation, concurrent with the establishment of an
-industrial base and characteristic moves towards standardisation and
-mass production (Grubler, 1998). The influence of accumulating
-production experience on costs is captured by the concept of learning.
-
-@grubler1990rise Following the definition given by Rogers, 1983,
-diffusion is the process by which an innovation is communicated through
-certain communication channels over time among the members of a social
-system. An innovation may be an idea, object or practice which is
-perceived as new by an individual or another unit of adoption (e.g., an
-organization) . Communication via communication channels refers to how
-knowledge about an innovation is transferred between the members of a
-social system. Communication takes time, and is therefore an important
-aspect of innovation diffusion. In fact, all theoretical and empirical
-diffusion studies agree that an innovation does not instantly diffuse
-into a social system. Instead a typical time pattern of diffusion along
-a S-shaped trajectory seems to be the rule. The S-shaped pattern of
-diffusion appears to be a basic anthropologic phenomenon, as it is also
-confirmed by a large number of studies· of primitive preindustrial
-societies.
-
-The scale of spatial innovation diffusion is similar\* to that of
-temporal diffusion models: hierarchically decomposed.
-
-Spatial innovation diffusion in Hagerstrand's model consists of three
-stages. Stage 1: Local concentrations of initial acceptances (initial
-agglomeration) . Stage 2: Radial dissemination outward from the initial
-agglomeration, accompanied by the rise of secondary agglomerations.
-Stage 3: The condensing and saturation stage, in which filling-in occurs
-and diffusion eventually ceases.
-
-Thus, three main characteristics define the spatial diffusion process:
-(1) As in temporal diffusion models, an S-shaped pattern in the
-cumulative level of adoption. (2) A "hierarchy effect" by which
-diffusion proceeds from initial centers to secondary ones. Within the
-context of urban systems (central places) this implies a hierarchical
-spread of an innovation from larger to smaller centers, and from the
-centers to the surrounding hinterland. (3) A "neighborhood effect" Le.,
-diffusion proceeds outwards from innovation centers, first "hitting"
-(Brown, 1981) nearby rather than far-away locations.
-
-Later Hagerstrand conceptualized physical "barrier" effects like lakes
-or uninhabited areas, which, in addition to distance, act as further
-retarding effects on diffusion. These are formalized in the form of
-"zero" or "half" contact multiplicators on the (distance decaying)
-message flows.
-
-With respect to the formalization of the communication flows Hagerstrand
-defines a "mean information field" (MIF), in which the probability of
-communication is a negative function of distance between individuals
-
 @ding2010modeling model the spatial diffusion of mobile
@@
 timing, speed, and level of mobile telecommunications diffusion in China
-
-Rogers (1995) characterized early adopters as knowledgeable risk takers.
-Griliches (1957) looked at the rate of return for early adopters and
-characterized them as profit maximizers. The geographer Hagerstrand
-(1967) was focusing on the character of early adopters in the
-transmittal process that produces follow-on participants in the process.
-Space was treated as a contiguity system with embedded characteristics
-producing barrier elements to the diffusion process.
-
-REVIEW ON Technological Diffusion
-
-Perkins and Neumayer (2005) examined whether latecomers have a higher
-diffusion rate of three technologies including digital telephone
-mainline
 
@@
 
-Hagerstrand (1968) was the first to demonstrate that diffusion typically
-starts from innovative centers ('core´) and disseminates, through a
-hierarchy of subcenters, to the periphery. This process is sequential
-rather than simultaneous, and tends to accelerate from core to
-periphery, where diffusion reaches a lower intensity than in the core
-(Hagerstrand, 1968; Grubler, 1990; Morrill, 2005). A similar effect of
-spatial acceleration has been identified in marketing research and is
-known as the "lead-lag effect" (Peres et al., 2010).
-
-@perkins2005international analyze whether the rate at which new producer
-technologies diffuse is significantly influenced by (1) latecomer
-advantage and (2) engagement with the global economy via trade and
-foreign investment
-
-Indeed, precisely because of these latecomer advantages, developing
-countries1 are believed to be well placed to catch up with developed
-ones (Gerschenkron 1962; Abramovitz 1986).
-
-Geographers, of course, were at the forefront of quantitative diffusion
-research. Ha¨gerstrand's (1967) early work on understanding innovation
-diffusion in Sweden using simulation modeling is particularly noteworthy
-in this respect. However, already long before the ''cultural turn'' in
-economic geography (Barnes 2001), geographers' interest in formally
-modeling the diffusion of new innovations waned (Gregory 1985). Indeed,
-much of the recent quantitative research into the diffusion of new
-technology has been undertaken by economists, sociologists, and business
-scholars.
-
-Theoretical models disagree as to why firms adopt innovations at
-different times. Epidemic models emphasize information (Griliches 1957).
-Certain firms are hypothesized to adopt earlier because they come into
-contact with, and learn from, adopters of the new technology before
-others. Economic models, on the other hand, predominantly emphasize firm
-heterogeneity (Ireland and Stoneman 1986). Firms adopt technologies at
-different times because they differ with respect to various
-organizational and environmental variables influencing the economic
-returns from adoption (Blackman 1999).
-
-Either way, the strong assumption is that developing countries can
-acquire modern technology innovated in developed economies, often at a
-fraction of the original research and development (R&D) costs, thereby
-leapfrogging many decades of technological progress (Teece 2000).
-
-The first involves a country's geographical location. Recent empirical
-work suggests that diffusion is ''geographically localized'' (Globerman,
-Kokko, and Sjo¨holm 2000; Keller 2002; Milner 2003) in that a technology
-diffuses faster in a country where it is already more widely diffused in
-neighboring countries. Underlying these regional effects are contagion
-and contact with prior users or producers of technology.
-
-@perkins2011internet Two main mechanisms are identified in the
-literature to explain diffusion: (1) epidemic-type dynamics whereby
-contact with previous adopters stimulates uptake as po tential adopters
-learn about a new innovation; and (2) economic-type mechanisms whereby
-potential users adopt a new innovation as it be comesmore profitable,
-useful, or valuable, with uptake characteristically spreading as costs
-be come lower, performance improves, or the po tential uses of the
-innovation grow over time.
-
-To this extent, our results contribute to a growing body of work that
-has sought to caution against claims about the supposed novelty of the
-Internet and the suggestion that it is somehow different.
-
-@meade2021modelling There are numerous agents who would benefit from
-accurate predictions of introduction times. These include: suppliers of
-network equipment for production planning; suppliers of mobile phones
-who need to have inventories of appropriately configured handsets;
-international network planners who need to provide sufficient bandwidth
-for communications between different countries; network operators and
-regulatory authorities.
-
-Classifying adopters of a technology along a time axis was pioneered by
-Rogers (1962), who categorized adopters into five groups as innovators,
-early adopters, early majority, late majority and laggards.
-
-An a priori approach groups similar countries into segments, where
-similarity is defined using economic, social, or political covariates.
-Examples include Hofstede (2001) and Lee (1990). An a posteriori
-analysis considers segments based on realized market behaviour, for
-example see Sood et al. (2009). Some studies investigate changing
-segment membership over time, see Cannon and Yaprak (2011).
-
-@fritsch2015new
-
-@fritsch2015new analyze the spatial diffusion of laser technology
-research in West Germany from 1960, when this technology began, until
-2005.
-
-Hägerstrand (1952, 1967) was the first scholar to intensively analyze
-spatial diffusion pro cesses. Based on the empirical observations of the
-regional spread of process innovations, he hypothesized that new
-technology is first implemented in the 'center' -- the large agglom
-erations -- and then moves 'down' the spatial hierarchy, the last stop
-being the 'periphery', i.e. remote and sparsely populated regions. There
-are at least two factors that may explain this pattern: first, the
-transfer of tacit knowledge via face-to-face contact and second, the
-absorptive capacity of a region.
-
-# Literature review {\#sec2} MAYBE SKIP THIS??
-
 # Materials and Methods {#sec3}
@@
 
-from R: Asym/(1+exp((xmid-input)/scal)) using the terms from Wilson k/(1
-+ e\^(t_0 - t)/scal)
+from R: Asym/(1+exp((xmid-input)/scal)) using the terms from Wilson
+k/(1 + e\^(t_0 - t)/scal)
 
@@
 
-## Multi-country analysis {#sec3.1}
+## Results {#sec3.1}
+
+Hexagon density maps: reflect the spatial structure of Britain. Websites
+are associated in places where people live and work.
+
+**TODO** maps at the local authority level per firms and discuss
+patterns
+
+Neighbourhood effect: diffusion proceeds outwards from innovation
+centers, first "hitting" nearby rather than far-away locations (Grubler
+1990)
+
+-   Moran's I: for OA and LAD over time **TODO** add 0s
+
+-   LISA maps: for OA and LAD over time More and less expected clusters.
+    Different scales show different results
+
+-   Website density regressions: for OA and LAD over time **TODO** add
+    0s Similar pattern
+
+Hierarchy effect: from main centers to secondary ones -- central places
+
+-   Gini coefficient. Almost perfect polarisation of web adoption in the
+    early stages at a granular level More equally diffused at the Local
+    Authority level Plateau overtime
+
+S-shaped diffusion curves
+
+-   S for LADs per firm and OA. **TODO**: OA per firm? fix firms over
+    time.
+
+-   Fast and slow LAs map. There is clustering
+
+-   **TODo** check s_uk_firm_11
+
+-   ranks: there is stability and movement
 
@@
 
-@lengyel2020role We see that the OSN spread almost exclusively from the
-original location (the capital Budapest, with an order of magnitude more
-inhabitants than the next size town) to various parts of the country in
-the early phase of the life-cycle. Later, diffusion became less
-mono-centric and other towns also emerged as spreaders. Our findings
-support the idea that spreading initially happens to large distances and
-becomes more local over time. This is illustrated and discussed later in
-Figure 2I.
-
 # References {#references .unnumbered}

### 2023-03-21 paper/submission1/help.Rmd
paper first draft

---
title: "help"
author: "Emmanouil Tranos"
date: "2023-03-10"
output: html_document
---

## Key concepts

Spatial diffusion processes

Geographical diffusion is a synthesis of different processes. On the one hand, we can identify purely spatial or, in other words, contagious processes. Adjacency and, more broadly, distance are the key drivers of diffusion. 
This perspective draws similarities with epidemics: innovation just like pathogens spreads because of contagion and, consequently, proximity and exposure [@hivner2003facilitating]. On the other hand, we can identify hierarchical processes. Instead of horizontal distance-based diffusion mechanisms, the top-down hierarchy of urban systems drives diffusion. In reality, the synthesis of these two processes represents how new technologies diffuse over space and time [@morrill2020spatial].

These ideas were firstly introduced by Torsten Hägerstrand and his thesis entitled 'Innovation Diffusion as a Spatial Process' [@hagerstrand1968innovation]. Hägerstrand was the first one to identify diffusion as a geographical process. The starting point was the idea that diffusion is based on passing information through social networks, which themselves tend to be defined by geography. Hence, he identified the 'neighbourhood' effect of how information, and consequently, innovation diffuses. He used agricultural innovations to test and model his ideas using Monte Carlo simulations. Hägerstrand also incorporated the role of hierarchy and how some phenomena maybe firstly adopted in larger cities and then diffuse to second tier ones. This is a sequential instead of a simultaneous process, which resembles the 'lead-lag' spatial acceleration effect in market research [@bento2018time; @PERES201091]. Hägerstrand maybe more widely known for highlighting the role time plays in the diffusion of innovations: an early-pioneering period, a middle fast accelerating period and a final saturation period [@morrill2020spatial.

The temporal dimension was further explored by Everett Rogers and his seminal work on 'Diffusion of Innovations' [@rogers2010diffusion]. Rogers being a sociologist  his work focused not on the diffusion of innovations over space and time, but instead on the adoption of new technologies and innovations by individuals and the individual mechanisms that drive the decisions behind adoption. He identified five groups of individuals regarding their adoption speed: innovators, early adopters, early majority, late majority and laggards. Key mechanism of diffusion and adoption is always communication and how knowledge is transferred within a social system. Therefore, all approaches agree on the S-shaped diffusion and cumulative adoption pattern [@grubler1990rise].

Schmidt's Law empirically illustrates a similar pattern. *Core* and ususally highly agglomerated regions is where new technologies are invented and commercially deployed [@grubler1990rise; grubler1990rise]. This is where the first adopters tend to be based. Then, technologies spread to the *rim* and eventually to the *periphery*. Although adoption pace might be higher when new technologies finally arrive to the periphery, the saturation levels there may never reach the ones in the core because of the lack of infrastructure or other necessary institutions [@leibowicz2016representing].

@grubler1990rise effectively summarises the three key characteristics of the spatial diffusion process: (i) the cumulative level of adoption follows an S-shaped pattern just like purely temporal models; (ii) diffusion is shaped by a hierarchy effect in a form of a centrifugal force: from core to periphery; and (iii) diffusion is also shaped by distance and a neighbourhood effect and contaminate nearby locations.

Space: absorptive capacity

Evolutionary economic geography >> path dependency

### 2023-03-21 paper/submission1/tranos2023-IT074305.qmd
paper first draft

--- 2023-03-21 Untitled/Untitled.Rmd
@@
 ---
-title: Digital economy in the UK: a multi-scalar story of the diffusion of web technologies
+title: "Digital economy in the UK: a multi-scalar story of the diffusion of web technologies"
 author:
@@
     email: e.tranos@bristol.ac.uk
-    affiliation: University of Bristol and The Alan Turing Institute
-    correspondingauthor: true
-    footnote: 1
+    affiliations: 
+      - id: some-tech
+        name: University of Bristol and The Alan Turing Institute
+        department: 
         address: 
-  - code: University of Bristol and The Alan Turing Institute
-    address: University of Bristol, School of Geographical Sciences, Bristol, UK; The Alan Turing Institute, London, UK
-footnote:
-  - code: 1
-    text: "This is the first author footnote."
-  - code: 2
-    text: "Another author footnote."
+        city: 
+        state: State
+        postal-code: UK
+    corresponding: e.tranos@bristol.ac.uk
+address:
+  - attributes:
+        corresponding: true
+    note: This is the first author footnote.
 abstract: |
@@
 Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.
-
 keywords: 
@@
   - keyword2
-journal: "An awesome journal"
-date: "`r Sys.Date()`"
-classoption: preprint, 3p, authoryear
-bibliography: mybibfile.bib
-linenumbers: false
-numbersections: true
-# Use a CSL with `citation_package = "default"`
-# csl: https://www.zotero.org/styles/elsevier-harvard
-output: 
-  rticles::elsevier_article:
-    keep_tex: true
-    citation_package: natbib
+date: last-modified
+bibliography: bibliography
+format:
+  elsevier-pdf:
+    keep-tex: true
+    journal:
+      name: Journal Name
+      formatting: preprint
+      model: 3p
+      cite-style: authoryear
+editor: 
+  markdown: 
+    wrap: 72
 ---
+
 # Introduction {#sec:1}
 
-Geographers were always interested in how new technologies and innovations diffuse across space and, importantly, how such spatio-temporal processes can be modelled.
-The seminal contribution of @hagerstrand1968innovation is illustrative of this early interest.
-However, the torch of exploring and modelling such processes had been passed to other disciplines such as economics, business studies and sociology well before the 'cultural turn' of economic geography [@perkins2005international].
-A potential explanation of lack of studies exploring the diffusion of new and, more specifically, digital technologies across *both* space and time can be attributed to the scarcity of relevant and granular enough data.
-As @zook2022mapping highlight, digital activities are hardly ever captured in official data.
+Geographers were always interested in how new technologies and
+innovations diffuse across space and, importantly, how such
+spatio-temporal processes can be modelled. The seminal contribution of
+@hagerstrand1968innovation is illustrative of this early interest.
+However, the torch of exploring and modelling such processes had been
+passed to other disciplines such as economics, business studies and
+sociology well before the 'cultural turn' of economic geography
+[@perkins2005international]. A potential explanation of lack of studies
+exploring the diffusion of new and, more specifically, digital
+technologies across *both* space and time can be attributed to the
+scarcity of relevant and granular enough data. As @zook2022mapping
+highlight, digital activities are hardly ever captured in official data.
 
-This paper aims to reinstate geographers' interest in how new and digital technologies diffuse over space and time.
-It adopts a geographic standpoint to explore and map the diffusion of web technologies in the UK over space and time.
-It does so by employing a novel source of big data which captures the active engagement with web technologies during the $1996-2013$ period.
+This paper aims to reinstate geographers' interest in how new and
+digital technologies diffuse over space and time. It adopts a geographic
+standpoint to explore and map the diffusion of web technologies in the
+UK over space and time. It does so by employing a novel source of big
+data which captures the active engagement with web technologies during
+the $1996-2013$ period.
 
-The motivation for this paper lies in the fact that there are various stakeholders who are interested in knowing how new digital technologies diffused over space and time use this information to make predictions regarding the diffusion of related future technologies.
-As per @meade2021modelling a variety of actors have a direct interest in gaining such knowledge including network equipment suppliers; network operators, regulatory and local authorities.
-And despite the spatial heterogeneity of such diffusion mechanisms and the policy relevance, there are very limited attempts in the literature to analyse the diffusion of new digital technologies at a granular level.
+The motivation for this paper lies in the fact that there are various
+stakeholders who are interested in knowing how new digital technologies
+diffused over space and time use this information to make predictions
+regarding the diffusion of related future technologies. As per
+@meade2021modelling a variety of actors have a direct interest in
+gaining such knowledge including network equipment suppliers; network
+operators, regulatory and local authorities. And despite the spatial
+heterogeneity of such diffusion mechanisms and the policy relevance,
+there are very limited attempts in the literature to analyse the
+diffusion of new digital technologies at a granular level.
 
@@
 
-@fritsch2015new
-Hägerstrand (1952, 1967) was the first scholar to intensively analyze
-spatial diffusion pro cesses. Based on the empirical observations of the
-regional spread of process innovations, he hypothesized that new
-technology is first implemented in the 'center' -- the large agglom
-erations -- and then moves 'down' the spatial hierarchy, the last stop
-being the 'periphery', i.e. remote and sparsely populated regions. There
-are at least two factors that may explain this pattern: first, the
-transfer of tacit knowledge via face-to-face contact and second, the
-absorptive capacity of a region.
+@fritsch2015new Hägerstrand (1952, 1967) was the first scholar to
+intensively analyze spatial diffusion pro cesses. Based on the empirical
+observations of the regional spread of process innovations, he
+hypothesized that new technology is first implemented in the 'center' --
+the large agglom erations -- and then moves 'down' the spatial
+hierarchy, the last stop being the 'periphery', i.e. remote and sparsely
+populated regions. There are at least two factors that may explain this
+pattern: first, the transfer of tacit knowledge via face-to-face contact
+and second, the absorptive capacity of a region.
 
 <!-- key concepts -->
+
 @bento2015spatial The concept of absorptive capacity has also been
@@
 
-@bednarz2020pulled 
-larger local technological niches (at least from a demand perspective)
-are more likely to be formed in urban areas than in rural ones, which is
-in line with the work of Ha¨gerstrand (1965a,b). However, size is not
-the only regional characteristic that may matter in this context.
-Regions also differ in accumulated experiences and the presence of tacit
-knowledge with respect to products, as well as in actors' propensity
-share this knowledge among producers (Martin et al., 2019)
+@bednarz2020pulled larger local technological niches (at least from a
+demand perspective) are more likely to be formed in urban areas than in
+rural ones, which is in line with the work of Ha¨gerstrand (1965a,b).
+However, size is not the only regional characteristic that may matter in
+this context. Regions also differ in accumulated experiences and the
+presence of tacit knowledge with respect to products, as well as in
+actors' propensity share this knowledge among producers (Martin et al.,
+2019)
 
 <!-- @das2022diffusion -->
+
 <!-- Epidemic approach: The approach emphasizes that the adoption of a -->
+
 <!-- technology spreads like an infection among the potential adopters. By -->
+
 <!-- taking a homogeneous group of potential adopters as an assumption, it -->
+
 <!-- explains the adoption of any innovation. -->
@@
 <!-- Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.) -->
+
 <!-- Early adopters: next 13.5%, (III.) Early Majority: following 34%, (IV.) -->
+
 <!-- Late Majority: next 34%, and the rest as (V.) Laggards -->
@@
 
-Contagious diffusion=> spatial autocorelation
+Contagious diffusion=\> spatial autocorelation
 
@@
 <!-- the three geographic diffusion processes are (1) purely contagious where -->
+
 <!-- distance or adjacency is the absolute controlling factor, (2) purely -->
+
 <!-- hierarchical where size or urban position in the central place hierarchy -->
+
 <!-- is the absolute controlling factor, or (3) where the location of change -->
+
 <!-- is purely random. These are the pure or ideal forms of the diffusion -->
+
 <!-- process; all real diffusion processes are the result of a combination of -->
+
 <!-- these extremes. -->
@@
 <!-- @das2022diffusion [not good paper] Invention, innovation and diffusion -->
+
 <!-- are considered as the three pillars of the theory of technological -->
+
 <!-- change and Schumpeter is considered as the father of the study of -->
+
 <!-- diffusion -->
 
-@leibowicz2016representing
-Diffusion is a spatial (Hagerstrand, 1967) as well as a temporal
-process, and historical evidence confirms that technologies diffuse at
-different times, at different rates, and to different extents in
-different places, and can be significantly influenced by policies
-(Victor, 1993).
+@leibowicz2016representing Diffusion is a spatial (Hagerstrand, 1967) as
+well as a temporal process, and historical evidence confirms that
+technologies diffuse at different times, at different rates, and to
+different extents in different places, and can be significantly
+influenced by policies (Victor, 1993).
 
 <!-- @leibowicz2016representing -->
+
 <!-- An illustrative empirically based theory for technology -->
+
 <!-- diffusion across multiple regions is suggested by "Schmidt's Law" -->
+
 <!-- (Grubler, 1990). With respect to a specific technology, it divides -->
+
 <!-- regions into three groups: core, rim, and periphery. The core is where -->
+
 <!-- the technology is invented, subject to experimentation, and first -->
+
 <!-- deployed commercially. To develop the technology, the core must have -->
+
 <!-- both sufficient innovative ability and economic motivation to make the -->
+
 <!-- necessary resource commitment. As the technology spreads from the core -->
+
 <!-- to the rim and eventually to the periphery, later adopters have the -->
+
 <!-- advantage of benefiting from the experiences of earlier adopters via -->
+
 <!-- knowledge spillover effects (Verdolini and Galeotti, 2011). At this -->
+
 <!-- point, the technology has likely been refined and costs have likely -->
+
 <!-- declined significantly. Schmidt's Law claims that, as a result, -->
+
 <!-- diffusion proceeds at a faster rate in regions which adopt the -->
+
 <!-- technology later. However, diffusion tends to be less pervasive in -->
+
 <!-- regions where adoption is delayed, ultimately saturating at a lower -->
+
 <!-- extent than in regions where adoption occurred earlier. This is often -->
+
 <!-- attributed to a lack of proper institutions and infrastructures in later -->
+
 <!-- adopters, markets whichmay not be as favorable to a technology, or the -->
+
 <!-- availability of yet newer options that invite technological -->
+
 <!-- leapfrogging. -->
@@
 
-wilson201281 Early on in their lifecycle, new technologies are crude, imperfect, and
-expensive (Rosenberg, 1994). New energy technologies are attractive for
-their ability to perform a particular task or deliver a new or improved
-energy service (Fouquet, 2010). This is often circumscribed by a
-particular set of needs in a particular context: a market 'niche'.
-End-users in niche markets are generally less sensitive to the effective
-price of the energy service provided or have a higher willingness to pay
-for its performance advantages (Fouquet, 2010). Thus initially,
-performance dominates cost competitiveness (Wilson and Grubler, 2011).
-Market niches afford some protection from competitive pressures,
-allowing technologies to be tested and improved in applied settings,
-reducing uncertainties with performance or market demand (Kemp et al.,
-1998). Costs may only fall substantively after an extended period of
-commercial experimentation, concurrent with the establishment of an
-industrial base and characteristic moves towards standardisation and
-mass production (Grubler, 1998). The influence of accumulating
+wilson201281 Early on in their lifecycle, new technologies are crude,
+imperfect, and expensive (Rosenberg, 1994). New energy technologies are
+attractive for their ability to perform a particular task or deliver a
+new or improved energy service (Fouquet, 2010). This is often
+circumscribed by a particular set of needs in a particular context: a
+market 'niche'. End-users in niche markets are generally less sensitive
+to the effective price of the energy service provided or have a higher
+willingness to pay for its performance advantages (Fouquet, 2010). Thus
+initially, performance dominates cost competitiveness (Wilson and
+Grubler, 2011). Market niches afford some protection from competitive
+pressures, allowing technologies to be tested and improved in applied
+settings, reducing uncertainties with performance or market demand (Kemp
+et al., 1998). Costs may only fall substantively after an extended
+period of commercial experimentation, concurrent with the establishment
+of an industrial base and characteristic moves towards standardisation
+and mass production (Grubler, 1998). The influence of accumulating
 production experience on costs is captured by the concept of learning.
@@
 <!-- @grubler1990rise Following the definition given by Rogers, 1983, -->
+
 <!-- diffusion is the process by which an innovation is communicated through -->
+
 <!-- certain communication channels over time among the members of a social -->
+
 <!-- system. An innovation may be an idea, object or practice which is -->
+
 <!-- perceived as new by an individual or another unit of adoption (e.g., an -->
+
 <!-- organization) . Communication via communication channels refers to how -->
+
 <!-- knowledge about an innovation is transferred between the members of a -->
+
 <!-- social system. Communication takes time, and is therefore an important -->
+
 <!-- aspect of innovation diffusion. In fact, all theoretical and empirical -->
+
 <!-- diffusion studies agree that an innovation does not instantly diffuse -->
+
 <!-- into a social system. Instead a typical time pattern of diffusion along -->
+
 <!-- a S-shaped trajectory seems to be the rule. The S-shaped pattern of -->
+
 <!-- diffusion appears to be a basic anthropologic phenomenon, as it is also -->
+
 <!-- confirmed by a large number of studies· of primitive preindustrial -->
+
 <!-- societies. -->
 
-grubler1990rise
-The scale of spatial innovation diffusion is similar to that of
-temporal diffusion models: hierarchically decomposed.
+grubler1990rise The scale of spatial innovation diffusion is similar to
+that of temporal diffusion models: hierarchically decomposed.
 
 <!-- grubler1990rise -->
+
 <!-- Spatial innovation diffusion in Hagerstrand's model consists of three -->
+
 <!-- stages. Stage 1: Local concentrations of initial acceptances (initial -->
+
 <!-- agglomeration) . Stage 2: Radial dissemination outward from the initial -->
+
 <!-- agglomeration, accompanied by the rise of secondary agglomerations. -->
+
 <!-- Stage 3: The condensing and saturation stage, in which filling-in occurs -->
+
 <!-- and diffusion eventually ceases. -->
@@
 <!-- grubler1990rise -->
+
 <!-- Thus, three main characteristics define the spatial diffusion process: -->
+
 <!-- (1) As in temporal diffusion models, an S-shaped pattern in the -->
+
 <!-- cumulative level of adoption. (2) A "hierarchy effect" by which -->
+
 <!-- diffusion proceeds from initial centers to secondary ones. Within the -->
+
 <!-- context of urban systems (central places) this implies a hierarchical -->
+
 <!-- spread of an innovation from larger to smaller centers, and from the -->
+
 <!-- centers to the surrounding hinterland. (3) A "neighborhood effect" Le., -->
+
 <!-- diffusion proceeds outwards from innovation centers, first "hitting" -->
+
 <!-- (Brown, 1981) nearby rather than far-away locations. -->
 
-@grubler1990rise
-Later Hagerstrand conceptualized physical "barrier" effects like lakes
-or uninhabited areas, which, in addition to distance, act as further
-retarding effects on diffusion. These are formalized in the form of
-"zero" or "half" contact multiplicators on the (distance decaying)
-message flows.
+@grubler1990rise Later Hagerstrand conceptualized physical "barrier"
+effects like lakes or uninhabited areas, which, in addition to distance,
+act as further retarding effects on diffusion. These are formalized in
+the form of "zero" or "half" contact multiplicators on the (distance
+decaying) message flows.
 
-@grubler1990rise
-With respect to the formalization of the communication flows Hagerstrand
-defines a "mean information field" (MIF), in which the probability of
-communication is a negative function of distance between individuals
+@grubler1990rise With respect to the formalization of the communication
+flows Hagerstrand defines a "mean information field" (MIF), in which the
+probability of communication is a negative function of distance between
+individuals
 
-@perkins2005international
-Theoretical models disagree as to why firms adopt innovations at
-different times. Epidemic models emphasize information (Griliches 1957).
-Certain firms are hypothesized to adopt earlier because they come into
-contact with, and learn from, adopters of the new technology before
-others. Economic models, on the other hand, predominantly emphasize firm
-heterogeneity (Ireland and Stoneman 1986). Firms adopt technologies at
-different times because they differ with respect to various
-organizational and environmental variables influencing the economic
-returns from adoption (Blackman 1999).
+@perkins2005international Theoretical models disagree as to why firms
+adopt innovations at different times. Epidemic models emphasize
+information (Griliches 1957). Certain firms are hypothesized to adopt
+earlier because they come into contact with, and learn from, adopters of
+the new technology before others. Economic models, on the other hand,
+predominantly emphasize firm heterogeneity (Ireland and Stoneman 1986).
+Firms adopt technologies at different times because they differ with
+respect to various organizational and environmental variables
+influencing the economic returns from adoption (Blackman 1999).
 
@@
 
-@ding2010modeling Rogers (1995) characterized early adopters as knowledgeable risk takers.
-Griliches (1957) looked at the rate of return for early adopters and
-characterized them as profit maximizers. The geographer Hagerstrand
-(1967) was focusing on the character of early adopters in the
-transmittal process that produces follow-on participants in the process.
-Space was treated as a contiguity system with embedded characteristics
-producing barrier elements to the diffusion process.
+@ding2010modeling Rogers (1995) characterized early adopters as
+knowledgeable risk takers. Griliches (1957) looked at the rate of return
+for early adopters and characterized them as profit maximizers. The
+geographer Hagerstrand (1967) was focusing on the character of early
+adopters in the transmittal process that produces follow-on participants
+in the process. Space was treated as a contiguity system with embedded
+characteristics producing barrier elements to the diffusion process.
 
-@ding2010modeling
-REVIEW ON Technological Diffusion
+@ding2010modeling REVIEW ON Technological Diffusion
 
 <!-- Hagerstrand (1968) was the first to demonstrate that diffusion typically -->
+
 <!-- starts from innovative centers ('core´) and disseminates, through a -->
+
 <!-- hierarchy of subcenters, to the periphery. This process is sequential -->
+
 <!-- rather than simultaneous, and tends to accelerate from core to -->
+
 <!-- periphery, where diffusion reaches a lower intensity than in the core -->
+
 <!-- (Hagerstrand, 1968; Grubler, 1990; Morrill, 2005). A similar effect of -->
+
 <!-- spatial acceleration has been identified in marketing research and is -->
+
 <!-- known as the "lead-lag effect" (Peres et al., 2010). -->
@@
 <!-- methods -->
+
 @wilson201281 Logistic growth describes an initial period of gradual
@@
 <!-- previous studies -->
+
 @beardsell1999spatial To characterize evolution of the computer
@@
 
-@leibowicz2016representing
-Empirical evidence supports the validity of Schmidt's Law over a wide
-range of technologies, time periods, and geographical contexts. A recent
-meta-analysis of technology up-scaling found that diffusion accelerated
-moving from the core to the rim and periphery for technologies as
-diverse as natural gas power, oil refineries, and automobiles (Wilson,
-2009). One historical example that conforms particularly well to
-Schmidt's Law is the diffusion of coal power in Europe (Grubler, 2012).
-England emerged as the core region for coal power because it had legal
-and economic institutions that incentivized scientific pursuits,
-domestic coal reserves, and a clear industrial motivation to replace
-water power with coal. ...
+@leibowicz2016representing Empirical evidence supports the validity of
+Schmidt's Law over a wide range of technologies, time periods, and
+geographical contexts. A recent meta-analysis of technology up-scaling
+found that diffusion accelerated moving from the core to the rim and
+periphery for technologies as diverse as natural gas power, oil
+refineries, and automobiles (Wilson, 2009). One historical example that
+conforms particularly well to Schmidt's Law is the diffusion of coal
+power in Europe (Grubler, 2012). England emerged as the core region for
+coal power because it had legal and economic institutions that
+incentivized scientific pursuits, domestic coal reserves, and a clear
+industrial motivation to replace water power with coal. ...
 

### 2023-03-21 paper/submission1/tranos2023.tex
paper first draft

--- 2023-03-21 paper/submission1/tranos2023-IT074305.qmd
@@
----
-title: "Digital economy in the UK: a multi-scalar story of the diffusion of web technologies"
-author:
-  - name: Emmanouil Tranos
-    email: e.tranos@bristol.ac.uk
-    affiliations: 
-      - id: some-tech
-        name: University of Bristol and The Alan Turing Institute
-        department: 
-        address: 
-        city: 
-        state: State
-        postal-code: UK
-    corresponding: e.tranos@bristol.ac.uk
-address:
-  - attributes:
-        corresponding: true
-    note: This is the first author footnote.
-abstract: |
-  This paper maps the participation in the digital economy and its evolution in the UK over space and time.
-Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects.
-Useful as these approaches might have been, they cannot capture the micro-processes and the characteristics of the individual online behaviour.
-Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK.
-The data and geolocation strategy allow to capture these processes at a very granular spatial scale.
-The modelling approach, which is based on simple spatial analytical methods and on the estimation of diffusion curves at various scales, enables to depict the role of geography and other cognitive factors which drove the diffusion of web technologies.
-Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.
-keywords: 
-  - keyword1
-  - keyword2
-date: last-modified
-bibliography: bibliography
-format:
-  elsevier-pdf:
-    keep-tex: true
-    journal:
-      name: Journal Name
-      formatting: preprint
-      model: 3p
-      cite-style: authoryear
-editor: 
-  markdown: 
-    wrap: 72
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  authoryear,
+  preprint,
+  3p]{elsarticle}
 
-# Introduction {#sec:1}
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{5}
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\journal{Journal Name}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage[]{natbib}
+\bibliographystyle{elsarticle-harv}
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Digital economy in the UK: a multi-scalar story of the diffusion of web technologies},
+  pdfauthor={Emmanouil Tranos},
+  pdfkeywords={keyword1, keyword2},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
+
+\setlength{\parindent}{6pt}
+\begin{document}
+
+\begin{frontmatter}
+\title{Digital economy in the UK: a multi-scalar story of the diffusion
+of web technologies}
+\author[1]{Emmanouil Tranos%
+\corref{cor1}%
+}
+ \ead{e.tranos@bristol.ac.uk}
+
+\affiliation[1]{organization={University of Bristol and The Alan Turing
+Institute},postcode={UK},postcodesep={}}
+
+\cortext[cor1]{Corresponding author}
+
+\begin{abstract}
+This paper maps the participation in the digital economy and its
+evolution in the UK over space and time. Most of the existing economic
+geography literature which dealt with the spatiality of the internet
+employed supply-side measures, such as infrastructural capacity, in
+order to understand the geography of the digital economy and its
+potential spatial economic effects. Useful as these approaches might
+have been, they cannot capture the micro-processes and the
+characteristics of the individual online behaviour. Using large volumes
+of archived and geolocated web content, this paper models the diffusion
+of web technologies over space and time in the UK. The data and
+geolocation strategy allow to capture these processes at a very granular
+spatial scale. The modelling approach, which is based on simple spatial
+analytical methods and on the estimation of diffusion curves at various
+scales, enables to depict the role of geography and other cognitive
+factors which drove the diffusion of web technologies. Although the
+focus is on a recent historical period -- 1996-2012 -- the results of
+the analysis depict diffusion mechanisms which can be very useful in
+understanding the evolutionary patterns of the adoption of other newer
+technologies.
+\end{abstract}
+
+\begin{keyword}
+    keyword1 \sep 
+    keyword2
+\end{keyword}
+\end{frontmatter}
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, frame hidden, interior hidden]}{\end{tcolorbox}}\fi
+
+\hypertarget{sec:1}{%
+\section{Introduction}\label{sec:1}}
 
@@
 spatio-temporal processes can be modelled. The seminal contribution of
-@hagerstrand1968innovation is illustrative of this early interest.
-However, the torch of exploring and modelling such processes had been
-passed to other disciplines such as economics, business studies and
-sociology well before the 'cultural turn' of economic geography
-[@perkins2005international]. A potential explanation of lack of studies
-exploring the diffusion of new and, more specifically, digital
-technologies across *both* space and time can be attributed to the
-scarcity of relevant and granular enough data. As @zook2022mapping
-highlight, digital activities are hardly ever captured in official data.
+\citet{hagerstrand1968innovation} is illustrative of this early
+interest. However, the torch of exploring and modelling such processes
+had been passed to other disciplines such as economics, business studies
+and sociology well before the `cultural turn' of economic geography
+\citep{perkins2005international}. A potential explanation of lack of
+studies exploring the diffusion of new and, more specifically, digital
+technologies across \emph{both} space and time can be attributed to the
+scarcity of relevant and granular enough data. As
+\citet{zook2022mapping} highlight, digital activities are hardly ever
+captured in official data.
 
@@
 data which captures the active engagement with web technologies during
-the $1996-2013$ period.
+the \(1996-2013\) period.
 
@@
 regarding the diffusion of related future technologies. As per
-@meade2021modelling a variety of actors have a direct interest in
+\citet{meade2021modelling} a variety of actors have a direct interest in
 gaining such knowledge including network equipment suppliers; network
@@
 
-@zook2022mapping Because new digital activities are rarely---if
+\citet{zook2022mapping} Because new digital activities are rarely---if
 ever---captured in official state data, researchers must rely on
@@
 overarching aim of this article is outlining an approach for analyzing
-the "new spaces and geographies" of digital phenomena and practices.
-
-<!-- motivation -->
+the ``new spaces and geographies'' of digital phenomena and practices.
 
@@
 analysis considers segments based on realized market behaviour, for
-example see Sood et al. (2009). Some studies investigate changing
+example see Sood et al.~(2009). Some studies investigate changing
 segment membership over time, see Cannon and Yaprak (2011).
 
-@fritsch2015new Hägerstrand (1952, 1967) was the first scholar to
+\citet{fritsch2015new} Hägerstrand (1952, 1967) was the first scholar to
 intensively analyze spatial diffusion pro cesses. Based on the empirical
 observations of the regional spread of process innovations, he
-hypothesized that new technology is first implemented in the 'center' --
-the large agglom erations -- and then moves 'down' the spatial
-hierarchy, the last stop being the 'periphery', i.e. remote and sparsely
+hypothesized that new technology is first implemented in the `center' --
+the large agglom erations -- and then moves `down' the spatial
+hierarchy, the last stop being the `periphery', i.e.~remote and sparsely
 populated regions. There are at least two factors that may explain this
@@
 
-<!-- key concepts -->
+\citet{bento2015spatial} The concept of absorptive capacity has also
+been applied in the TIS context in terms of the con-struction of
+capacity needed to transfer and implement a new technology in the
+receiving country (cf.Van Alphen, 2011), with repercussions for the
+nature of system functions
 
-@bento2015spatial The concept of absorptive capacity has also been
-applied in the TIS context in terms of the con-struction of capacity
-needed to transfer and implement a new technology in the receiving
-country (cf.Van Alphen, 2011), with repercussions for the nature of
-system functions
-
-@bednarz2020pulled larger local technological niches (at least from a
-demand perspective) are more likely to be formed in urban areas than in
-rural ones, which is in line with the work of Ha¨gerstrand (1965a,b).
-However, size is not the only regional characteristic that may matter in
-this context. Regions also differ in accumulated experiences and the
-presence of tacit knowledge with respect to products, as well as in
-actors' propensity share this knowledge among producers (Martin et al.,
-2019)
-
-<!-- @das2022diffusion -->
-
-<!-- Epidemic approach: The approach emphasizes that the adoption of a -->
-
-<!-- technology spreads like an infection among the potential adopters. By -->
-
-<!-- taking a homogeneous group of potential adopters as an assumption, it -->
-
-<!-- explains the adoption of any innovation. -->
+\citet{bednarz2020pulled} larger local technological niches (at least
+from a demand perspective) are more likely to be formed in urban areas
+than in rural ones, which is in line with the work of Ha¨gerstrand
+(1965a,b). However, size is not the only regional characteristic that
+may matter in this context. Regions also differ in accumulated
+experiences and the presence of tacit knowledge with respect to
+products, as well as in actors' propensity share this knowledge among
+producers (Martin et al., 2019)
 
@@
 
-<!-- Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.) -->
+\citet{morrill2020spatial} Spatial diffusion is the process by which
+behavior or characteristics of the landscape change as a result of what
+happens elsewhere earlier. Spatial diffusion is the spread of the
+phenomenon, over space and timed, from limited origins
 
-<!-- Early adopters: next 13.5%, (III.) Early Majority: following 34%, (IV.) -->
-
-<!-- Late Majority: next 34%, and the rest as (V.) Laggards -->
-
-@morrill2020spatial Spatial diffusion is the process by which behavior
-or characteristics of the landscape change as a result of what happens
-elsewhere earlier. Spatial diffusion is the spread of the phenomenon,
-over space and timed, from limited origins
-
-Contagious diffusion=\> spatial autocorelation
+Contagious diffusion=\textgreater{} spatial autocorelation
 
 expansion-type diffusion
-
-<!-- the three geographic diffusion processes are (1) purely contagious where -->
-
-<!-- distance or adjacency is the absolute controlling factor, (2) purely -->
-
-<!-- hierarchical where size or urban position in the central place hierarchy -->
-
-<!-- is the absolute controlling factor, or (3) where the location of change -->
-
-<!-- is purely random. These are the pure or ideal forms of the diffusion -->
-
-<!-- process; all real diffusion processes are the result of a combination of -->
-
-<!-- these extremes. -->
 
@@
 
-<!-- @das2022diffusion [not good paper] Invention, innovation and diffusion -->
-
-<!-- are considered as the three pillars of the theory of technological -->
-
-<!-- change and Schumpeter is considered as the father of the study of -->
-
-<!-- diffusion -->
-
-@leibowicz2016representing Diffusion is a spatial (Hagerstrand, 1967) as
-well as a temporal process, and historical evidence confirms that
-technologies diffuse at different times, at different rates, and to
+\citet{leibowicz2016representing} Diffusion is a spatial (Hagerstrand,
+1967) as well as a temporal process, and historical evidence confirms
+that technologies diffuse at different times, at different rates, and to
 different extents in different places, and can be significantly
 influenced by policies (Victor, 1993).
-
-<!-- @leibowicz2016representing -->
-
-<!-- An illustrative empirically based theory for technology -->
-
-<!-- diffusion across multiple regions is suggested by "Schmidt's Law" -->
-
-<!-- (Grubler, 1990). With respect to a specific technology, it divides -->
-
-<!-- regions into three groups: core, rim, and periphery. The core is where -->
-
-<!-- the technology is invented, subject to experimentation, and first -->
-
-<!-- deployed commercially. To develop the technology, the core must have -->
-
-<!-- both sufficient innovative ability and economic motivation to make the -->
-
-<!-- necessary resource commitment. As the technology spreads from the core -->
-
-<!-- to the rim and eventually to the periphery, later adopters have the -->
-
-<!-- advantage of benefiting from the experiences of earlier adopters via -->
-
-<!-- knowledge spillover effects (Verdolini and Galeotti, 2011). At this -->
-
-<!-- point, the technology has likely been refined and costs have likely -->
-
-<!-- declined significantly. Schmidt's Law claims that, as a result, -->
-
-<!-- diffusion proceeds at a faster rate in regions which adopt the -->
-
-<!-- technology later. However, diffusion tends to be less pervasive in -->
-
-<!-- regions where adoption is delayed, ultimately saturating at a lower -->
-
-<!-- extent than in regions where adoption occurred earlier. This is often -->
-
-<!-- attributed to a lack of proper institutions and infrastructures in later -->
-
-<!-- adopters, markets whichmay not be as favorable to a technology, or the -->
-
-<!-- availability of yet newer options that invite technological -->
-
-<!-- leapfrogging. -->
 
@@
 circumscribed by a particular set of needs in a particular context: a
-market 'niche'. End-users in niche markets are generally less sensitive
+market `niche'. End-users in niche markets are generally less sensitive
 to the effective price of the energy service provided or have a higher
@@
 
-<!-- @grubler1990rise Following the definition given by Rogers, 1983, -->
-
-<!-- diffusion is the process by which an innovation is communicated through -->
-
-<!-- certain communication channels over time among the members of a social -->
-
-<!-- system. An innovation may be an idea, object or practice which is -->
-
-<!-- perceived as new by an individual or another unit of adoption (e.g., an -->
-
-<!-- organization) . Communication via communication channels refers to how -->
-
-<!-- knowledge about an innovation is transferred between the members of a -->
-
-<!-- social system. Communication takes time, and is therefore an important -->
-
-<!-- aspect of innovation diffusion. In fact, all theoretical and empirical -->
-
-<!-- diffusion studies agree that an innovation does not instantly diffuse -->
-
-<!-- into a social system. Instead a typical time pattern of diffusion along -->
-
-<!-- a S-shaped trajectory seems to be the rule. The S-shaped pattern of -->
-
-<!-- diffusion appears to be a basic anthropologic phenomenon, as it is also -->
-
-<!-- confirmed by a large number of studies· of primitive preindustrial -->
-
-<!-- societies. -->
-
 grubler1990rise The scale of spatial innovation diffusion is similar to
@@
 
-<!-- grubler1990rise -->
+\citet{grubler1990rise} Later Hagerstrand conceptualized physical
+``barrier'' effects like lakes or uninhabited areas, which, in addition
+to distance, act as further retarding effects on diffusion. These are
+formalized in the form of ``zero'' or ``half'' contact multiplicators on
+the (distance decaying) message flows.
 
-<!-- Spatial innovation diffusion in Hagerstrand's model consists of three -->
+\citet{grubler1990rise} With respect to the formalization of the
+communication flows Hagerstrand defines a ``mean information field''
+(MIF), in which the probability of communication is a negative function
+of distance between individuals
 
-<!-- stages. Stage 1: Local concentrations of initial acceptances (initial -->
-
-<!-- agglomeration) . Stage 2: Radial dissemination outward from the initial -->
-
-<!-- agglomeration, accompanied by the rise of secondary agglomerations. -->
-
-<!-- Stage 3: The condensing and saturation stage, in which filling-in occurs -->
-
-<!-- and diffusion eventually ceases. -->
-
-<!-- grubler1990rise -->
-
-<!-- Thus, three main characteristics define the spatial diffusion process: -->
-
-<!-- (1) As in temporal diffusion models, an S-shaped pattern in the -->
-
-<!-- cumulative level of adoption. (2) A "hierarchy effect" by which -->
-
-<!-- diffusion proceeds from initial centers to secondary ones. Within the -->
-
-<!-- context of urban systems (central places) this implies a hierarchical -->
-
-<!-- spread of an innovation from larger to smaller centers, and from the -->
-
-<!-- centers to the surrounding hinterland. (3) A "neighborhood effect" Le., -->
-
-<!-- diffusion proceeds outwards from innovation centers, first "hitting" -->
-
-<!-- (Brown, 1981) nearby rather than far-away locations. -->
-
-@grubler1990rise Later Hagerstrand conceptualized physical "barrier"
-effects like lakes or uninhabited areas, which, in addition to distance,
-act as further retarding effects on diffusion. These are formalized in
-the form of "zero" or "half" contact multiplicators on the (distance
-decaying) message flows.
-
-@grubler1990rise With respect to the formalization of the communication
-flows Hagerstrand defines a "mean information field" (MIF), in which the
-probability of communication is a negative function of distance between
-individuals
-
-@perkins2005international Theoretical models disagree as to why firms
-adopt innovations at different times. Epidemic models emphasize
+\citet{perkins2005international} Theoretical models disagree as to why
+firms adopt innovations at different times. Epidemic models emphasize
 information (Griliches 1957). Certain firms are hypothesized to adopt
@@
 acquire modern technology innovated in developed economies, often at a
-fraction of the original research and development (R&D) costs, thereby
+fraction of the original research and development (R\&D) costs, thereby
 leapfrogging many decades of technological progress (Teece 2000).
@@
 The first involves a country's geographical location. Recent empirical
-work suggests that diffusion is ''geographically localized'' (Globerman,
+work suggests that diffusion is `'geographically localized'' (Globerman,
 Kokko, and Sjo¨holm 2000; Keller 2002; Milner 2003) in that a technology
@@
 
-@ding2010modeling Rogers (1995) characterized early adopters as
+\citet{ding2010modeling} Rogers (1995) characterized early adopters as
 knowledgeable risk takers. Griliches (1957) looked at the rate of return
@@
 
-@ding2010modeling REVIEW ON Technological Diffusion
+\citet{ding2010modeling} REVIEW ON Technological Diffusion
 
-<!-- Hagerstrand (1968) was the first to demonstrate that diffusion typically -->
-
-<!-- starts from innovative centers ('core´) and disseminates, through a -->
-
-<!-- hierarchy of subcenters, to the periphery. This process is sequential -->
-
-<!-- rather than simultaneous, and tends to accelerate from core to -->
-
-<!-- periphery, where diffusion reaches a lower intensity than in the core -->
-
-<!-- (Hagerstrand, 1968; Grubler, 1990; Morrill, 2005). A similar effect of -->
-
-<!-- spatial acceleration has been identified in marketing research and is -->
-
-<!-- known as the "lead-lag effect" (Peres et al., 2010). -->
-
-@perkins2011internet Two main mechanisms are identified in the
+\citet{perkins2011internet} Two main mechanisms are identified in the
 literature to explain diffusion: (1) epidemic-type dynamics whereby
@@
 
-<!-- methods -->
+\citet{wilson201281} Logistic growth describes an initial period of
+gradual diffusion as a technology is introduced as a new commercial
+application, moving then through a rapid, exponential growth phase,
+before slowing and eventually saturating \citep{grubler1999dynamics}.
+The substitution of incumbent technologies by new competitors leads to
+subsequent decline and eventual obsolescence.
 
-@wilson201281 Logistic growth describes an initial period of gradual
-diffusion as a technology is introduced as a new commercial application,
-moving then through a rapid, exponential growth phase, before slowing
-and eventually saturating [@grubler1999dynamics]. The substitution of
-incumbent technologies by new competitors leads to subsequent decline
-and eventual obsolescence.
+\hypertarget{literature-review-sec2-maybe-skip-this}{%
+\section{Literature review \{\#sec2\} MAYBE SKIP
+THIS??}\label{literature-review-sec2-maybe-skip-this}}
 
-# Literature review {#sec2} MAYBE SKIP THIS??
-
-<!-- previous studies -->
-
-@beardsell1999spatial To characterize evolution of the computer
+\citet{beardsell1999spatial} To characterize evolution of the computer
 industry, we examine the distribution of relative employment across
@@
 
-@bednarz2020pulled spatial innovation diffusion By using Bayesian
+\citet{bednarz2020pulled} spatial innovation diffusion By using Bayesian
 survival models with time-dependent data of wind turbine deployment and
@@
 
-@perkins2005international analyze whether the rate at which new producer
-technologies diffuse is significantly influenced by (1) latecomer
-advantage and (2) engagement with the global economy via trade and
-foreign investment
+\citet{perkins2005international} analyze whether the rate at which new
+producer technologies diffuse is significantly influenced by (1)
+latecomer advantage and (2) engagement with the global economy via trade
+and foreign investment
 
@@
 
-@fritsch2015new analyze the spatial diffusion of laser technology
+\citet{fritsch2015new} analyze the spatial diffusion of laser technology
 research in West Germany from 1960, when this technology began, until
@@
 
-@perkins2011internet
+\citet{perkins2011internet}
 
-@lengyel2020role We see that the OSN spread almost exclusively from the
-original location (the capital Budapest, with an order of magnitude more
-inhabitants than the next size town) to various parts of the country in
-the early phase of the life-cycle. Later, diffusion became less
-mono-centric and other towns also emerged as spreaders. Our findings
-support the idea that spreading initially happens to large distances and
-becomes more local over time. This is illustrated and discussed later in
-Figure 2I.
+\citet{lengyel2020role} We see that the OSN spread almost exclusively
+from the original location (the capital Budapest, with an order of
+magnitude more inhabitants than the next size town) to various parts of
+the country in the early phase of the life-cycle. Later, diffusion
+became less mono-centric and other towns also emerged as spreaders. Our
+findings support the idea that spreading initially happens to large
+distances and becomes more local over time. This is illustrated and
+discussed later in Figure 2I.
 
-@PAPAGIANNIDIS2015308
+\citet{PAPAGIANNIDIS2015308}
 
-@leibowicz2016representing for energy There is no generally accepted
-theory that explains diffusion rate heterogeneity across technologies,
-but several factors are considered important. Greater unit scale and
-larger market size contribute to slower diffusion. Requirements for
-interrelated technologies or complex infrastructures also hinder the
-diffusion process (Grubler, 2012).
+\citet{leibowicz2016representing} for energy There is no generally
+accepted theory that explains diffusion rate heterogeneity across
+technologies, but several factors are considered important. Greater unit
+scale and larger market size contribute to slower diffusion.
+Requirements for interrelated technologies or complex infrastructures
+also hinder the diffusion process (Grubler, 2012).
 
@@
 
-@leibowicz2016representing Empirical evidence supports the validity of
-Schmidt's Law over a wide range of technologies, time periods, and
-geographical contexts. A recent meta-analysis of technology up-scaling
-found that diffusion accelerated moving from the core to the rim and
-periphery for technologies as diverse as natural gas power, oil
+\citet{leibowicz2016representing} Empirical evidence supports the
+validity of Schmidt's Law over a wide range of technologies, time
+periods, and geographical contexts. A recent meta-analysis of technology
+up-scaling found that diffusion accelerated moving from the core to the
+rim and periphery for technologies as diverse as natural gas power, oil
 refineries, and automobiles (Wilson, 2009). One historical example that
@@
 incentivized scientific pursuits, domestic coal reserves, and a clear
-industrial motivation to replace water power with coal. ...
+industrial motivation to replace water power with coal. \ldots{}
 
-@ding2010modeling model the spatial diffusion of mobile
+\citet{ding2010modeling} model the spatial diffusion of mobile
 telecommunications in China as well as its determinants. regional
@@
 
-@bento2018time explore What determines the duration of formative phases
-for energy innovations in different markets? We are interested both in
-initial markets (also: core, lead, first mover, early adopter) where
-formative phases prepare technologies for mass commercialization, and in
-follower markets (also: periphery, lag, late adopter) where accelerated
-formative phases may benefit from diffusion and spillovers.
+\citet{bento2018time} explore What determines the duration of formative
+phases for energy innovations in different markets? We are interested
+both in initial markets (also: core, lead, first mover, early adopter)
+where formative phases prepare technologies for mass commercialization,
+and in follower markets (also: periphery, lag, late adopter) where
+accelerated formative phases may benefit from diffusion and spillovers.
 
-# Materials and Methods {#sec3}
+\hypertarget{sec3}{%
+\section{Materials and Methods}\label{sec3}}
 
-@wilson201281 growth function description in p. 86
+\citet{wilson201281} growth function description in p.~86
 
-from R: Asym/(1+exp((xmid-input)/scal)) using the terms from Wilson
-k/(1 + e\^(t_0 - t)/scal)
+from R: Asym/(1+exp((xmid-input)/scal)) using the terms from Wilson k/(1
++ e\^{}(t\_0 - t)/scal)
 
-Wilson: k(1 + e\^-b(t - t_0)) So, scal = - 1/b
+Wilson: k(1 + e\^{}-b(t - t\_0)) So, scal = - 1/b
 
-The curve is symmetric\* around t_0
+The curve is symmetric* around t\_0
 
-**minimum R2 = 95%**, see also @grubler1990rise
+\textbf{minimum R2 = 95\%}, see also \citet{grubler1990rise}
 
@@
 
-Moran's I as @ding2010modeling **TODO** for t_0, diffusion speed
+Moran's I as \citet{ding2010modeling} \textbf{TODO} for t\_0, diffusion
+speed
 
-## Results {#sec3.1}
+\hypertarget{sec3.1}{%
+\subsection{Results}\label{sec3.1}}
 
@@
 
-**TODO** maps at the local authority level per firms and discuss
+\textbf{TODO} maps at the local authority level per firms and discuss
 patterns
@@
 Neighbourhood effect: diffusion proceeds outwards from innovation
-centers, first "hitting" nearby rather than far-away locations (Grubler
-1990)
+centers, first ``hitting'' nearby rather than far-away locations
+(Grubler 1990)
 
--   Moran's I: for OA and LAD over time **TODO** add 0s
-
--   LISA maps: for OA and LAD over time More and less expected clusters.
+\begin{itemize}
+\item
+  Moran's I: for OA and LAD over time \textbf{TODO} add 0s
+\item
+  LISA maps: for OA and LAD over time More and less expected clusters.
   Different scales show different results
-
--   Website density regressions: for OA and LAD over time **TODO** add
-    0s Similar pattern
+\item
+  Website density regressions: for OA and LAD over time \textbf{TODO}
+  add 0s Similar pattern
+\end{itemize}
 
@@
 
--   Gini coefficient. Almost perfect polarisation of web adoption in the
+\begin{itemize}
+\tightlist
+\item
+  Gini coefficient. Almost perfect polarisation of web adoption in the
   early stages at a granular level More equally diffused at the Local
   Authority level Plateau overtime
+\end{itemize}
 
@@
 
--   S for LADs per firm and OA. **TODO**: OA per firm? fix firms over
+\begin{itemize}
+\item
+  S for LADs per firm and OA. \textbf{TODO}: OA per firm? fix firms over
   time.
+\item
+  Fast and slow LAs map. There is clustering
+\item
+  \textbf{TODo} check s\_uk\_firm\_11
+\item
+  ranks: there is stability and movement
+\end{itemize}
 
--   Fast and slow LAs map. There is clustering
+\hypertarget{sec3.2}{%
+\subsection{3.2 Case study approach}\label{sec3.2}}
 
--   **TODo** check s_uk_firm_11
+\hypertarget{sec4}{%
+\section{Digital technologies and spatial structure: a global
+view}\label{sec4}}
 
--   ranks: there is stability and movement
+\hypertarget{sec5}{%
+\section{The impact of ICT on the US and the UK spatial
+structures}\label{sec5}}
 
-## 3.2 Case study approach {#sec3.2}
-
-# Digital technologies and spatial structure: a global view {#sec4}
-
-# The impact of ICT on the US and the UK spatial structures {#sec5}
-
-# Discussion and conclusions {#sec6}
+\hypertarget{sec6}{%
+\section{Discussion and conclusions}\label{sec6}}
 
 contrary to results from future studies regarding social media
-[@lengyel2020role], web technologies did not exclusively spread from a
-central location.
+\citep{lengyel2020role}, web technologies did not exclusively spread
+from a central location.
 
-# References {#references .unnumbered}
+\renewcommand\refname{References}
+  \bibliography{bibliography}
+
+\end{document}

### 2023-03-21 presentations/web_diffusion.Rmd
paper first draft

--- 2023-02-16 presentations/gceg.Rmd
@@
 css: style.css
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
@@
 ## Contents
+
 <br>
-<div style="margin-left: 50px !important; text-align:left !important">
+
+::: {style="margin-left: 50px !important; text-align:left !important"}
 -   Introduction
@@
 -   S-shaped diffusion curves
-- Conclusions
-<br>
-<br>
+-   Conclusions <br> <br>
     [etranos.info/post/gceg2022](https://etranos.info/post/gceg2022/)
-</div>
+:::
 
@@
 
-- Predictions of introduction times for future technologies [@meade2021modelling]:
+-   Predictions of introduction times for future technologies
+    [@meade2021modelling]:
 
@@
 
-- A **hierarchy effect**: from main centers to secondary ones -- central places
+-   A **hierarchy effect**: from main centres to secondary ones --
+    central places
 
-- A **neighborhood effect**: diffusion proceeds outwards from innovation centers, 
-first "hitting" nearby rather than far-away locations [@grubler1990rise]
+-   A **neighborhood effect**: diffusion proceeds outwards from
+    innovation centres, first "hitting" nearby rather than far-away
+    locations [@grubler1990rise]
 
-@hagerstrand1965monte: from innovative centers (core) through a
-hierarchy of subcenters, to the periphery
+@hagerstrand1965monte: from innovative centres (core) through a
+hierarchy of sub-centres, to the periphery
 
@@
 
-> Because new digital activities are rarely---if
-ever---captured in official state data, researchers must rely on
-information gathered from alternative sources [@zook2022mapping].
+> Because new digital activities are rarely---if ever---captured in
+> official state data, researchers must rely on information gathered
+> from alternative sources [@zook2022mapping].
 
@@
 ## Web data: The Internet Archive
+
 -   The largest archive of webpages in the world
-- 273 billion webpages from over 361 million websites, 15 petabytes of storage (1996 -)
-- A web crawler starts with a list of URLs (a seed list) to crawl and downloads a copy of their content
-- Using the hyperlinks included in the crawled URLs, new URLs are identified and crawled (snowball sampling)
+-   273 billion webpages from over 361 million websites, 15 petabytes of
+    storage (1996 -)
+-   A web crawler starts with a list of URLs (a seed list) to crawl and
+    downloads a copy of their content
+-   Using the hyperlinks included in the crawled URLs, new URLs are
+    identified and crawled (snowball sampling)
 -   Time-stamp
@@
 ## Web data: The Internet Archive
-![](C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/figure/Picture1.png){width=70%}
+
+![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture1.png){width="70%"}
 
 ## Web data: The Internet Archive
-![](C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/figure/Picture2.png){width=70%}
+
+![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture2.png){width="70%"}
 
 ## Our web data
-- JISC UK Web Domain Dataset: all archived webpages from the .uk domain 1996-2012
+
+-   JISC UK Web Domain Dataset: all archived webpages from the .uk
+    domain 1996-2012
 
@@
 
-- Tranos, E., and C. Stich. 2020. Individual internet usage and the availability of online content of local interest: A multilevel approach. *Computers, Environment and Urban Systems*, 79:101371.
+-   Tranos, E., and C. Stich. 2020. Individual internet usage and the
+    availability of online content of local interest: A multilevel
+    approach. *Computers, Environment and Urban Systems*, 79:101371.
 
-- Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy in the UK: Regional productivity effects of early adoption. *Regional Studies*, 55:12, 1924-1938.
+-   Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy
+    in the UK: Regional productivity effects of early adoption.
+    *Regional Studies*, 55:12, 1924-1938.
 
-- Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from the ground up: a web data approach. *Environment and Planning B*, in press.
+-   Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from
+    the ground up: a web data approach. *Environment and Planning B*, in
+    press.
 
-- Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to predict regional trade flows: data extraction, modelling, and validation, *Annals of the AAG*, in press.
+-   Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to
+    predict regional trade flows: data extraction, modelling, and
+    validation, *Annals of the AAG*, in press.
 
@@
 
-- All .uk archived webpages which contain a UK postcode in the web text
+-   All .uk archived webpages which contain a UK postcode in the web
+    text
 
@@
 
-<br>
-<br>
+<br> <br>
 
 <!-- <small></small> -->
-20080509162138/http://www.website1.co.uk/contact_us IG8 8HD
+
+20080509162138/<http://www.website1.co.uk/contact_us> IG8 8HD
 
@@
 
-    <small>- http://www.website1.co.uk/webpage1 and</small>
+    <small>- <http://www.website1.co.uk/webpage1> and</small>
 
-    <small>- http://www.website1.co.uk/webpage2 are part of the</small>
+    <small>- <http://www.website1.co.uk/webpage2> are part of
+    the</small>
 
-    <small>- http://www.website1.co.uk</small>
+    <small>- <http://www.website1.co.uk></small>
 
@@
 ```
+
 <br>
 
-- Websites with a large number of postcodes: e.g. directories, real estate websites
+-   Websites with a large number of postcodes: e.g. directories, real
+    estate websites
 
@@
 ## Directory website with a lot of postcodes
-![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
+
+![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
 
 ## Website with a unique postcode in London
-![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
+
+![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
 
@@
 
----
+------------------------------------------------------------------------
 
@@
 
----
+<!-- ------------------------------------------------------------------------ -->
 
-```{r heatmaps_log, results='asis', out.width = "60%"}
+```{r eval = F, heatmaps_log, results='asis', out.width = "60%"}
 path.images <- paste0(path, "/outputs/")
@@
 
-<br>
-
-Two scales:
-
-- Output areas (c. 200,000)
-
-- Local authorities (c. 400)
-
----
-
-<br>
-
-**Neighborhood effect**: diffusion proceeds outwards from innovation centers, 
-first "hitting" nearby rather than far-away locations [@grubler1990rise]
+**Neighborhood effect**: diffusion proceeds outwards from innovation
+centers, first "hitting" nearby rather than far-away locations
+[@grubler1990rise]
 
@@
 
----
+-   Websites per firm in Local authorities (c. 400)
 
-```{r moransi_oa, results='asis'}
+<!-- ------------------------------------------------------------------------ -->
+
+```{r, eval=FALSE, moransi_oa, results='asis'}
 path.in <- paste0(path, "/outputs/morani_oa.csv")
@@
 
----
+<!-- ------------------------------------------------------------------------ -->
 
-```{r lisa_oa, results='asis', out.width = "30%"}
+```{r lisa_oa, eval=FALSE,results='asis', out.width = "30%"}
 path.images <- paste0(path, "/outputs/")
@@
 
-<div style="margin-left: 50px !important; text-align:left !important">
+::: {style="margin-left: 50px !important; text-align:left !important"}
 $Place$:
@@
 
-$i$: Output Areas *or* Local Authorities
-</div>
+-   $Website\,Density_{i}$: Websites per firm in a Local Authority $i$
+:::
 
@@
 
----
+<!-- ------------------------------------------------------------------------ -->
 
-```{r b_London_oa, fig.height=5, fig.width=5, message=FALSE}
+```{r b_London_oa, eval=F, fig.height=5, fig.width=5, message=FALSE}
 #, out.width = "60%"
@@
 
----
+<!-- ------------------------------------------------------------------------ -->
 
-```{r b_city_oa, fig.height=5, fig.width=5, message=FALSE}
+```{r b_city_oa, eval = F, fig.height=5, fig.width=5, message=FALSE}
 #, out.width = "60%"
@@
 
----
+<!-- ------------------------------------------------------------------------ -->
 
-```{r b_centre_oa, message=FALSE, out.height= "60%", out.width="60%"}
+```{r b_centre_oa, eval = F, message=FALSE, out.height= "60%", out.width="60%"}
 #fig.height=4, fig.width=4, 
@@
 
----
+------------------------------------------------------------------------
 
-```{r b_centre_la, message=FALSE, out.height= "60%", out.width="60%"}
+```{r b_centre_la, message=FALSE}
+#, out.height= "60%", out.width="60%"
 path.image <- paste0(path, "/outputs/b_dist_centre_la.png")
@@
 
-```{r r_squared, eval=F, results='asis'}
+------------------------------------------------------------------------
+
+```{r r_squared, eval=T, results='asis'}
 # This is for the RSquared of the different website density regressions.
@@
 
-read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
+read.csv(path.in2) %>% #left_join(read.csv(path.in2)) %>% 
   ggplot(aes(x=year)) +
-  geom_line(aes(y=London, color = "OA to London"), size=1) +
-  geom_line(aes(y=cities, color = "OA to nearest city"), size=1) +
-  geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) +
+  # geom_line(aes(y=London, color = "OA to London"), size=1) +
+  # geom_line(aes(y=cities, color = "OA to nearest city"), size=1) +
+  # geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) +
 
@@
 
-  scale_color_discrete(name = "Website density regressions") +
+  scale_color_discrete(name = "R-squared for regression:") +
   scale_linetype_manual(name = "Website density regressions") +
-  theme_minimal()
+  ylab("R-squared") + 
+  ggtitle("Website density regressions") +
+  theme_minimal() +
+  theme(axis.title.x=element_blank(),
+        plot.title = element_text(hjust = 0.5))
 
-  scale_color_manual(values = c("red", "blue", "green"))
+  # scale_color_manual(values = c("red", "blue", "green"))
 
-read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
-  pivot_longer(names_to = "destinations",
-               values_to = "Rsquared",
-               cols = 2:7) %>% 
-  arrange(destinations, year) %>% 
-  ggplot(aes(x=year)) +
-  geom_line(aes(y=Rsquared, group=destinations, colour=destinations)) +
-  geom_textline(aes(y=Rsquared, group=destinations, colour=destinations, 
-                    label=destinations),
-                hjust = 1) +
-  theme(legend.position = "none") +
-  theme_minimal()
+# read.csv(path.in2) #%>% left_join(read.csv(path.in2)) %>% 
+#   pivot_longer(names_to = "destinations",
+#                values_to = "Rsquared",
+#                cols = 2:7) %>% 
+#   arrange(destinations, year) %>% 
+#   ggplot(aes(x=year)) +
+#   geom_line(aes(y=Rsquared, group=destinations, colour=destinations)) +
+#   geom_textline(aes(y=Rsquared, group=destinations, colour=destinations, 
+#                     label=destinations),
+#                 hjust = 1) +
+#   theme(legend.position = "none") +
+#   theme_minimal()
 
@@
 
-**Hierarchy effect**: from main centers to secondary ones -- central places
+**Hierarchy effect**: from main centers to secondary ones -- central
+places
 
@@
 
----
+<!-- ------------------------------------------------------------------------ -->
 
-```{r gini_oa, results='asis'}
+```{r gini_oa, eval=F, results='asis'}
 path.in <- paste0(path, "/outputs/gini_oa.csv")
@@
 
-- Almost perfect polarisation of web adoption in the early stages at a granular level
+-   Almost perfect polarisation of web adoption in the early stages at a
+    granular level
 
@@
 
-- Distance effect: urban gravitation increases over time and then drops
+-   Distance effect: urban gravitation increases over time and then
+    drops
 
@@
 ```{r Darlington, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000005.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000005.png")
 knitr::include_graphics(path.image)
@@
 
----
+------------------------------------------------------------------------
 
 ```{r Nottingham, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000018.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000018.png")
 knitr::include_graphics(path.image)
@@
 
----
+------------------------------------------------------------------------
 
 ```{r Reading, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000038.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000038.png")
 knitr::include_graphics(path.image)
@@
 
----
+------------------------------------------------------------------------
 
 ```{r City_of_London, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E09000001.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000001.png")
 knitr::include_graphics(path.image)
@@
 ```{r Middlesbrough, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000002.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000002.png")
 knitr::include_graphics(path.image)
@@
 
----
+------------------------------------------------------------------------
 
 ```{r Leicester, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000016.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000016.png")
 knitr::include_graphics(path.image)
@@
 
----
+------------------------------------------------------------------------
 
 ```{r Bristol, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000023.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000023.png")
 knitr::include_graphics(path.image)
@@
 
----
+------------------------------------------------------------------------
 
 ```{r Plymouth, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000026.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000026.png")
 knitr::include_graphics(path.image)
@@
 
----
+------------------------------------------------------------------------
 
 ```{r Liverpool, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E08000012.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E08000012.png")
 knitr::include_graphics(path.image)
@@
 
----
+------------------------------------------------------------------------
 
 ```{r Hackney, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E09000012.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000012.png")
 knitr::include_graphics(path.image)
@@
 
----
+------------------------------------------------------------------------
 
 ```{r Edinburgh, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_S12000036.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_S12000036.png")
 knitr::include_graphics(path.image)
@@
 
-- Well-established theoretical approaches of diffusion survive even at a granular level
+-   Well-established theoretical approaches of diffusion survive even at
+    a granular level
 
@@
 
-- Expand our definition of web to websites with # of postcodes > 1
+-   Expand our definition of web to websites with \# of postcodes \> 1
 

### 2023-03-22 paper/submission1/tranos2023_backup.qmd
lisa maps titles

--- 2023-03-21 paper/submission1/tranos2023-IT074305.qmd
@@
 date: last-modified
-bibliography: bibliography
+bibliography: bibliography.bib
 format:
@@
       model: 3p
-      cite-style: authoryear
+      cite-style: super
 editor: 

### 2023-03-22 src/level_lisa_la.Rmd
lisa maps titles

@@
                                          labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
-                                         title="LISA for level, LA") +
+                                         title="LISA, websites per firm in a LA") +
   tm_legend(text.size = 1) +

### 2023-03-22 paper/help/backup_elsevier_template/test.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  super,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Short Paper},
  pdfauthor={Alice Anonymous; Bob Security; Cat Memes; Derek Zoolander},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Short Paper \\\large{A Short Subtitle} }
\author[1]{Alice Anonymous%
\corref{cor1}%
\fnref{fn1}}
 \ead{alice@example.com} 
\author[2]{Bob Security%
%
\fnref{fn2}}
 \ead{bob@example.com} 
\author[2]{Cat Memes%
%
\fnref{fn3}}
 \ead{cat@example.com} 
\author[]{Derek Zoolander%
%
}
 \ead{derek@example.com}

\affiliation[1]{organization={Some Institute of Technology, Department
Name},addressline={Street Address},city={City},postcode={Postal
Code},postcodesep={}}
\affiliation[2]{organization={Another University, Department
Name},addressline={Street Address},city={City},postcode={Postal
Code},postcodesep={}}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{This is the first author footnote.}
\fntext[fn2]{Another author footnote, this is a very long footnote and
it should be a really long footnote. But this footnote is not yet
sufficiently long enough to make two lines of footnote text.}
\fntext[fn3]{Yet another author footnote.}

\begin{abstract}
This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
tempus elit.
\end{abstract}

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi

Please make sure that your manuscript follows the guidelines in the
Guide for Authors of the relevant journal. It is not necessary to
typeset your manuscript in exactly the same way as an article, unless
you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see
\url{https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions}

\hypertarget{bibliography-styles}{%
\section{Bibliography styles}\label{bibliography-styles}}

Here are two sample references: \citet{Feynman1963118}
\citet{Dirac1953888}.

By default, natbib will be used with the \texttt{authoryear} style, set
in \texttt{classoption} variable in YAML. You can sets extra options
with \texttt{natbiboptions} variable in YAML header. Example

\begin{verbatim}
natbiboptions: longnamesfirst,angle,semicolon
\end{verbatim}

There are various more specific bibliography styles available at
\url{https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files}.
To use one of these, add it in the header using, for example,
\texttt{biblio-style:\ model1-num-names}.

\hypertarget{using-csl}{%
\subsection{Using CSL}\label{using-csl}}

If \texttt{cite-method} is set to \texttt{citeproc} in
\texttt{elsevier\_article()}, then pandoc is used for citations instead
of \texttt{natbib}. In this case, the \texttt{csl} option is used to
format the references. By default, this template will provide an
appropriate style, but alternative \texttt{csl} files are available from
\url{https://www.zotero.org/styles?q=elsevier}. These can be downloaded
and stored locally, or the url can be used as in the example header.

\hypertarget{equations}{%
\section{Equations}\label{equations}}

Here is an equation: \[ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
\]

Inline equations work as well: \(\sum_{i = 2}^\infty\{\alpha_i^\beta\}\)

\hypertarget{figures-and-tables}{%
\section{Figures and tables}\label{figures-and-tables}}

Figure~\ref{fig-meaningless} is generated using an R chunk.

\begin{figure}

{\centering \includegraphics[width=0.5\textwidth,height=\textheight]{test_files/figure-pdf/fig-meaningless-1.pdf}

}

\caption{\label{fig-meaningless}A meaningless scatterplot}

\end{figure}

\hypertarget{tables-coming-from-r}{%
\section{Tables coming from R}\label{tables-coming-from-r}}

Tables can also be generated using R chunks, as shown in
Table~\ref{tbl-simple} example.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(}\FunctionTok{head}\NormalTok{(mtcars)[,}\DecValTok{1}\SpecialCharTok{:}\DecValTok{4}\NormalTok{])}
\end{Highlighting}
\end{Shaded}

\hypertarget{tbl-simple}{}
\begin{longtable}[]{@{}lrrrr@{}}
\caption{\label{tbl-simple}Caption centered above table}\tabularnewline
\toprule()
& mpg & cyl & disp & hp \\
\midrule()
\endfirsthead
\toprule()
& mpg & cyl & disp & hp \\
\midrule()
\endhead
Mazda RX4 & 21.0 & 6 & 160 & 110 \\
Mazda RX4 Wag & 21.0 & 6 & 160 & 110 \\
Datsun 710 & 22.8 & 4 & 108 & 93 \\
Hornet 4 Drive & 21.4 & 6 & 258 & 110 \\
Hornet Sportabout & 18.7 & 8 & 360 & 175 \\
Valiant & 18.1 & 6 & 225 & 105 \\
\bottomrule()
\end{longtable}

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}

### 2023-03-22 paper/help/old/paper.Rmd

--- 2023-03-21 Untitled/Untitled.Rmd
@@
 ---
-title: Digital economy in the UK: a multi-scalar story of the diffusion of web technologies
+title: Title of submission to PLOS journal
 author:
@@
     affiliation: University of Bristol and The Alan Turing Institute
-    correspondingauthor: true
-    footnote: 1
+    corresponding: e.tranos@bristol.ac.uk
 address:
@@
     address: University of Bristol, School of Geographical Sciences, Bristol, UK; The Alan Turing Institute, London, UK
-footnote:
-  - code: 1
-    text: "This is the first author footnote."
-  - code: 2
-    text: "Another author footnote."
 abstract: |
-  This paper maps the participation in the digital economy and its evolution in the UK over space and time.
-Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects.
-Useful as these approaches might have been, they cannot capture the micro-processes and the characteristics of the individual online behaviour.
-Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK.
-The data and geolocation strategy allow to capture these processes at a very granular spatial scale.
-The modelling approach, which is based on simple spatial analytical methods and on the estimation of diffusion curves at various scales, enables to depict the role of geography and other cognitive factors which drove the diffusion of web technologies.
-Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.
+  blah blah
 
-keywords: 
-  - keyword1
-  - keyword2
-journal: "An awesome journal"
-date: "`r Sys.Date()`"
-classoption: preprint, 3p, authoryear
+author_summary: |
+  blah blah
 bibliography: mybibfile.bib
-linenumbers: false
-numbersections: true
-# Use a CSL with `citation_package = "default"`
-# csl: https://www.zotero.org/styles/elsevier-harvard
-output: 
-  rticles::elsevier_article:
-    keep_tex: true
-    citation_package: natbib
+output: rticles::plos_article
+csl: plos.csl
+header-includes:
+- \usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
+
+```{r echo=FALSE, message = FALSE, warning=FALSE}
+library(stargazer)
+library(tidyverse)
+#library(rmarkdown)
+```
+
 # Introduction {#sec:1}
@@
 <!-- motivation -->
+@meade2021modelling There are numerous agents who would benefit from
+accurate predictions of introduction times. These include: suppliers of
+network equipment for production planning; suppliers of mobile phones
+who need to have inventories of appropriately configured handsets;
+international network planners who need to provide sufficient bandwidth
+for communications between different countries; network operators and
+regulatory authorities.
 

### 2023-03-22 paper/help/old/paper.tex

--- 2023-03-21 paper/submission1/tranos2023.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+% Template for PLoS
+% Version 3.5 March 2018
 %
-\documentclass[
-  authoryear,
-  preprint,
-  3p]{elsarticle}
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended
+% to minimize problems and delays during our production
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% Once your paper is accepted for publication,
+% PLEASE REMOVE ALL TRACKED CHANGES in this file
+% and leave only the final text of your manuscript.
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file.
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline equations, please be sure to include all portions of an equation in the math environment.
+%
+% Do not include text that is not math in the math environment.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column.
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
 
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% cite package, to clean up citations in the main text. Do not remove.
+% \usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
+\usepackage{array}
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace}
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Use the PLoS provided BiBTeX style
+% \bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
 \makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
+\renewcommand{\@biblabel}[1]{\quad#1.}
 \makeatother
 
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\AtBeginDocument{%
-\ifdefined\contentsname
-  \renewcommand*\contentsname{Table of contents}
-\else
-  \newcommand\contentsname{Table of contents}
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+%\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+%\setlength{\headheight}{27.023pt}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\today}
+
+%% Include all macros below
+
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
+
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
   \fi
-\ifdefined\listfigurename
-  \renewcommand*\listfigurename{List of Figures}
-\else
-  \newcommand\listfigurename{List of Figures}
-\fi
-\ifdefined\listtablename
-  \renewcommand*\listtablename{List of Tables}
-\else
-  \newcommand\listtablename{List of Tables}
-\fi
-\ifdefined\figurename
-  \renewcommand*\figurename{Figure}
-\else
-  \newcommand\figurename{Figure}
-\fi
-\ifdefined\tablename
-  \renewcommand*\tablename{Table}
-\else
-  \newcommand\tablename{Table}
-\fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+\usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}
+
+\usepackage{forarray}
+\usepackage{xstring}
+\newcommand{\getIndex}[2]{
+  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
 }
-\@ifpackageloaded{float}{}{\usepackage{float}}
-\floatstyle{ruled}
-\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
-\floatname{codelisting}{Listing}
-\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\journal{Journal Name}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\usepackage[]{natbib}
-\bibliographystyle{elsarticle-harv}
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  pdftitle={Digital economy in the UK: a multi-scalar story of the diffusion of web technologies},
-  pdfauthor={Emmanouil Tranos},
-  pdfkeywords={keyword1, keyword2},
-  colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
 
-\setlength{\parindent}{6pt}
+\setcounter{secnumdepth}{0}
+
+\newcommand{\getAff}[1]{
+  \getIndex{#1}{University of Bristol and The Alan Turing Institute}
+}
+
 \begin{document}
+\vspace*{0.2in}
 
-\begin{frontmatter}
-\title{Digital economy in the UK: a multi-scalar story of the diffusion
-of web technologies}
-\author[1]{Emmanouil Tranos%
-\corref{cor1}%
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf\newline{Title of submission to PLOS
+journal} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
- \ead{e.tranos@bristol.ac.uk}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Emmanouil Tranos\textsuperscript{\getAff{University of Bristol and The
+Alan Turing Institute}}\textsuperscript{*}\\
+\bigskip
+\textbf{\getAff{University of Bristol and The Alan Turing
+Institute}}University of Bristol, School of Geographical Sciences,
+Bristol, UK; The Alan Turing Institute, London, UK\\
+\bigskip
+* Corresponding author: e.tranos@bristol.ac.uk\\
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
+blah blah
 
-\affiliation[1]{organization={University of Bristol and The Alan Turing
-Institute},postcode={UK},postcodesep={}}
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step.
+% Author Summary not valid for PLOS ONE submissions.
+\section*{Author summary}
+blah blah
 
-\cortext[cor1]{Corresponding author}
+\linenumbers
 
-\begin{abstract}
-This paper maps the participation in the digital economy and its
-evolution in the UK over space and time. Most of the existing economic
-geography literature which dealt with the spatiality of the internet
-employed supply-side measures, such as infrastructural capacity, in
-order to understand the geography of the digital economy and its
-potential spatial economic effects. Useful as these approaches might
-have been, they cannot capture the micro-processes and the
-characteristics of the individual online behaviour. Using large volumes
-of archived and geolocated web content, this paper models the diffusion
-of web technologies over space and time in the UK. The data and
-geolocation strategy allow to capture these processes at a very granular
-spatial scale. The modelling approach, which is based on simple spatial
-analytical methods and on the estimation of diffusion curves at various
-scales, enables to depict the role of geography and other cognitive
-factors which drove the diffusion of web technologies. Although the
-focus is on a recent historical period -- 1996-2012 -- the results of
-the analysis depict diffusion mechanisms which can be very useful in
-understanding the evolutionary patterns of the adoption of other newer
-technologies.
-\end{abstract}
-
-\begin{keyword}
-    keyword1 \sep 
-    keyword2
-\end{keyword}
-\end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, frame hidden, interior hidden]}{\end{tcolorbox}}\fi
-
+% Use "Eq" instead of "Equation" for equation citations.
 \hypertarget{sec:1}{%
@@
 spatio-temporal processes can be modelled. The seminal contribution of
-\citet{hagerstrand1968innovation} is illustrative of this early
-interest. However, the torch of exploring and modelling such processes
-had been passed to other disciplines such as economics, business studies
-and sociology well before the `cultural turn' of economic geography
-\citep{perkins2005international}. A potential explanation of lack of
-studies exploring the diffusion of new and, more specifically, digital
-technologies across \emph{both} space and time can be attributed to the
-scarcity of relevant and granular enough data. As
-\citet{zook2022mapping} highlight, digital activities are hardly ever
-captured in official data.
+{[}1{]} is illustrative of this early interest. However, the torch of
+exploring and modelling such processes had been passed to other
+disciplines such as economics, business studies and sociology well
+before the `cultural turn' of economic geography {[}2{]}. A potential
+explanation of lack of studies exploring the diffusion of new and, more
+specifically, digital technologies across \emph{both} space and time can
+be attributed to the scarcity of relevant and granular enough data. As
+{[}3{]} highlight, digital activities are hardly ever captured in
+official data.
 
@@
 diffused over space and time use this information to make predictions
-regarding the diffusion of related future technologies. As per
-\citet{meade2021modelling} a variety of actors have a direct interest in
-gaining such knowledge including network equipment suppliers; network
-operators, regulatory and local authorities. And despite the spatial
-heterogeneity of such diffusion mechanisms and the policy relevance,
-there are very limited attempts in the literature to analyse the
-diffusion of new digital technologies at a granular level.
+regarding the diffusion of related future technologies. As per {[}4{]} a
+variety of actors have a direct interest in gaining such knowledge
+including network equipment suppliers; network operators, regulatory and
+local authorities. And despite the spatial heterogeneity of such
+diffusion mechanisms and the policy relevance, there are very limited
+attempts in the literature to analyse the diffusion of new digital
+technologies at a granular level.
 
@@
 
-\citet{zook2022mapping} Because new digital activities are rarely---if
-ever---captured in official state data, researchers must rely on
-information gathered from alternative sources. With this in mind, the
-overarching aim of this article is outlining an approach for analyzing
-the ``new spaces and geographies'' of digital phenomena and practices.
+{[}3{]} Because new digital activities are rarely---if ever---captured
+in official state data, researchers must rely on information gathered
+from alternative sources. With this in mind, the overarching aim of this
+article is outlining an approach for analyzing the ``new spaces and
+geographies'' of digital phenomena and practices.
+
+{[}4{]} There are numerous agents who would benefit from accurate
+predictions of introduction times. These include: suppliers of network
+equipment for production planning; suppliers of mobile phones who need
+to have inventories of appropriately configured handsets; international
+network planners who need to provide sufficient bandwidth for
+communications between different countries; network operators and
+regulatory authorities.
 
@@
 
-\citet{fritsch2015new} Hägerstrand (1952, 1967) was the first scholar to
-intensively analyze spatial diffusion pro cesses. Based on the empirical
+{[}5{]} Hägerstrand (1952, 1967) was the first scholar to intensively
+analyze spatial diffusion pro cesses. Based on the empirical
 observations of the regional spread of process innovations, he
@@
 
-\citet{bento2015spatial} The concept of absorptive capacity has also
-been applied in the TIS context in terms of the con-struction of
-capacity needed to transfer and implement a new technology in the
-receiving country (cf.Van Alphen, 2011), with repercussions for the
-nature of system functions
+{[}6{]} The concept of absorptive capacity has also been applied in the
+TIS context in terms of the con-struction of capacity needed to transfer
+and implement a new technology in the receiving country (cf.Van Alphen,
+2011), with repercussions for the nature of system functions
 
-\citet{bednarz2020pulled} larger local technological niches (at least
-from a demand perspective) are more likely to be formed in urban areas
-than in rural ones, which is in line with the work of Ha¨gerstrand
-(1965a,b). However, size is not the only regional characteristic that
-may matter in this context. Regions also differ in accumulated
-experiences and the presence of tacit knowledge with respect to
-products, as well as in actors' propensity share this knowledge among
-producers (Martin et al., 2019)
+{[}7{]} larger local technological niches (at least from a demand
+perspective) are more likely to be formed in urban areas than in rural
+ones, which is in line with the work of Ha¨gerstrand (1965a,b). However,
+size is not the only regional characteristic that may matter in this
+context. Regions also differ in accumulated experiences and the presence
+of tacit knowledge with respect to products, as well as in actors'
+propensity share this knowledge among producers (Martin et al., 2019)
 
@@
 
-\citet{morrill2020spatial} Spatial diffusion is the process by which
-behavior or characteristics of the landscape change as a result of what
-happens elsewhere earlier. Spatial diffusion is the spread of the
-phenomenon, over space and timed, from limited origins
+{[}8{]} Spatial diffusion is the process by which behavior or
+characteristics of the landscape change as a result of what happens
+elsewhere earlier. Spatial diffusion is the spread of the phenomenon,
+over space and timed, from limited origins
 
@@
 
-\citet{leibowicz2016representing} Diffusion is a spatial (Hagerstrand,
-1967) as well as a temporal process, and historical evidence confirms
-that technologies diffuse at different times, at different rates, and to
-different extents in different places, and can be significantly
-influenced by policies (Victor, 1993).
+{[}9{]} Diffusion is a spatial (Hagerstrand, 1967) as well as a temporal
+process, and historical evidence confirms that technologies diffuse at
+different times, at different rates, and to different extents in
+different places, and can be significantly influenced by policies
+(Victor, 1993).
 
@@
 
-\citet{grubler1990rise} Later Hagerstrand conceptualized physical
-``barrier'' effects like lakes or uninhabited areas, which, in addition
-to distance, act as further retarding effects on diffusion. These are
-formalized in the form of ``zero'' or ``half'' contact multiplicators on
-the (distance decaying) message flows.
+{[}10{]} Later Hagerstrand conceptualized physical ``barrier'' effects
+like lakes or uninhabited areas, which, in addition to distance, act as
+further retarding effects on diffusion. These are formalized in the form
+of ``zero'' or ``half'' contact multiplicators on the (distance
+decaying) message flows.
 
-\citet{grubler1990rise} With respect to the formalization of the
-communication flows Hagerstrand defines a ``mean information field''
-(MIF), in which the probability of communication is a negative function
-of distance between individuals
+{[}10{]} With respect to the formalization of the communication flows
+Hagerstrand defines a ``mean information field'' (MIF), in which the
+probability of communication is a negative function of distance between
+individuals
 
-\citet{perkins2005international} Theoretical models disagree as to why
-firms adopt innovations at different times. Epidemic models emphasize
-information (Griliches 1957). Certain firms are hypothesized to adopt
-earlier because they come into contact with, and learn from, adopters of
-the new technology before others. Economic models, on the other hand,
-predominantly emphasize firm heterogeneity (Ireland and Stoneman 1986).
-Firms adopt technologies at different times because they differ with
-respect to various organizational and environmental variables
-influencing the economic returns from adoption (Blackman 1999).
+{[}2{]} Theoretical models disagree as to why firms adopt innovations at
+different times. Epidemic models emphasize information (Griliches 1957).
+Certain firms are hypothesized to adopt earlier because they come into
+contact with, and learn from, adopters of the new technology before
+others. Economic models, on the other hand, predominantly emphasize firm
+heterogeneity (Ireland and Stoneman 1986). Firms adopt technologies at
+different times because they differ with respect to various
+organizational and environmental variables influencing the economic
+returns from adoption (Blackman 1999).
 
@@
 
-\citet{ding2010modeling} Rogers (1995) characterized early adopters as
-knowledgeable risk takers. Griliches (1957) looked at the rate of return
-for early adopters and characterized them as profit maximizers. The
-geographer Hagerstrand (1967) was focusing on the character of early
-adopters in the transmittal process that produces follow-on participants
-in the process. Space was treated as a contiguity system with embedded
+{[}11{]} Rogers (1995) characterized early adopters as knowledgeable
+risk takers. Griliches (1957) looked at the rate of return for early
+adopters and characterized them as profit maximizers. The geographer
+Hagerstrand (1967) was focusing on the character of early adopters in
+the transmittal process that produces follow-on participants in the
+process. Space was treated as a contiguity system with embedded
 characteristics producing barrier elements to the diffusion process.
 
-\citet{ding2010modeling} REVIEW ON Technological Diffusion
+{[}11{]} REVIEW ON Technological Diffusion
 
-\citet{perkins2011internet} Two main mechanisms are identified in the
-literature to explain diffusion: (1) epidemic-type dynamics whereby
-contact with previous adopters stimulates uptake as po tential adopters
-learn about a new innovation; and (2) economic-type mechanisms whereby
-potential users adopt a new innovation as it be comesmore profitable,
-useful, or valuable, with uptake characteristically spreading as costs
-be come lower, performance improves, or the po tential uses of the
-innovation grow over time.
+{[}12{]} Two main mechanisms are identified in the literature to explain
+diffusion: (1) epidemic-type dynamics whereby contact with previous
+adopters stimulates uptake as po tential adopters learn about a new
+innovation; and (2) economic-type mechanisms whereby potential users
+adopt a new innovation as it be comesmore profitable, useful, or
+valuable, with uptake characteristically spreading as costs be come
+lower, performance improves, or the po tential uses of the innovation
+grow over time.
 
@@
 
-\citet{wilson201281} Logistic growth describes an initial period of
-gradual diffusion as a technology is introduced as a new commercial
-application, moving then through a rapid, exponential growth phase,
-before slowing and eventually saturating \citep{grubler1999dynamics}.
-The substitution of incumbent technologies by new competitors leads to
-subsequent decline and eventual obsolescence.
+{[}13{]} Logistic growth describes an initial period of gradual
+diffusion as a technology is introduced as a new commercial application,
+moving then through a rapid, exponential growth phase, before slowing
+and eventually saturating {[}14{]}. The substitution of incumbent
+technologies by new competitors leads to subsequent decline and eventual
+obsolescence.
 
@@
 
-\citet{beardsell1999spatial} To characterize evolution of the computer
-industry, we examine the distribution of relative employment across
-cities in 1977 and how that distribution changes over time to 1992
+{[}15{]} To characterize evolution of the computer industry, we examine
+the distribution of relative employment across cities in 1977 and how
+that distribution changes over time to 1992
 
@@
 
-\citet{bednarz2020pulled} spatial innovation diffusion By using Bayesian
-survival models with time-dependent data of wind turbine deployment and
-firm foundation for 402 German regions between the years 1970 and 2015,
-we show that the spatial evolution of the German wind energy industry
-was more strongly influenced by local demand-- pull than local
-supply--push processes.
+{[}7{]} spatial innovation diffusion By using Bayesian survival models
+with time-dependent data of wind turbine deployment and firm foundation
+for 402 German regions between the years 1970 and 2015, we show that the
+spatial evolution of the German wind energy industry was more strongly
+influenced by local demand-- pull than local supply--push processes.
 
@@
 
-\citet{perkins2005international} analyze whether the rate at which new
-producer technologies diffuse is significantly influenced by (1)
-latecomer advantage and (2) engagement with the global economy via trade
-and foreign investment
+{[}2{]} analyze whether the rate at which new producer technologies
+diffuse is significantly influenced by (1) latecomer advantage and (2)
+engagement with the global economy via trade and foreign investment
 
@@
 
-\citet{fritsch2015new} analyze the spatial diffusion of laser technology
-research in West Germany from 1960, when this technology began, until
-2005.
+{[}5{]} analyze the spatial diffusion of laser technology research in
+West Germany from 1960, when this technology began, until 2005.
 
-\citet{perkins2011internet}
+{[}12{]}
 
-\citet{lengyel2020role} We see that the OSN spread almost exclusively
-from the original location (the capital Budapest, with an order of
-magnitude more inhabitants than the next size town) to various parts of
-the country in the early phase of the life-cycle. Later, diffusion
-became less mono-centric and other towns also emerged as spreaders. Our
-findings support the idea that spreading initially happens to large
-distances and becomes more local over time. This is illustrated and
-discussed later in Figure 2I.
+{[}16{]} We see that the OSN spread almost exclusively from the original
+location (the capital Budapest, with an order of magnitude more
+inhabitants than the next size town) to various parts of the country in
+the early phase of the life-cycle. Later, diffusion became less
+mono-centric and other towns also emerged as spreaders. Our findings
+support the idea that spreading initially happens to large distances and
+becomes more local over time. This is illustrated and discussed later in
+Figure 2I.
 
-\citet{PAPAGIANNIDIS2015308}
+{[}17{]}
 
-\citet{leibowicz2016representing} for energy There is no generally
-accepted theory that explains diffusion rate heterogeneity across
-technologies, but several factors are considered important. Greater unit
-scale and larger market size contribute to slower diffusion.
-Requirements for interrelated technologies or complex infrastructures
-also hinder the diffusion process (Grubler, 2012).
+{[}9{]} for energy There is no generally accepted theory that explains
+diffusion rate heterogeneity across technologies, but several factors
+are considered important. Greater unit scale and larger market size
+contribute to slower diffusion. Requirements for interrelated
+technologies or complex infrastructures also hinder the diffusion
+process (Grubler, 2012).
 
@@
 
-\citet{leibowicz2016representing} Empirical evidence supports the
-validity of Schmidt's Law over a wide range of technologies, time
-periods, and geographical contexts. A recent meta-analysis of technology
-up-scaling found that diffusion accelerated moving from the core to the
-rim and periphery for technologies as diverse as natural gas power, oil
-refineries, and automobiles (Wilson, 2009). One historical example that
-conforms particularly well to Schmidt's Law is the diffusion of coal
-power in Europe (Grubler, 2012). England emerged as the core region for
-coal power because it had legal and economic institutions that
-incentivized scientific pursuits, domestic coal reserves, and a clear
-industrial motivation to replace water power with coal. \ldots{}
+{[}9{]} Empirical evidence supports the validity of Schmidt's Law over a
+wide range of technologies, time periods, and geographical contexts. A
+recent meta-analysis of technology up-scaling found that diffusion
+accelerated moving from the core to the rim and periphery for
+technologies as diverse as natural gas power, oil refineries, and
+automobiles (Wilson, 2009). One historical example that conforms
+particularly well to Schmidt's Law is the diffusion of coal power in
+Europe (Grubler, 2012). England emerged as the core region for coal
+power because it had legal and economic institutions that incentivized
+scientific pursuits, domestic coal reserves, and a clear industrial
+motivation to replace water power with coal. \ldots{}
 
-\citet{ding2010modeling} model the spatial diffusion of mobile
-telecommunications in China as well as its determinants. regional
-socioeconomic characteristics play an important role in determining the
-timing, speed, and level of mobile telecommunications diffusion in China
+{[}11{]} model the spatial diffusion of mobile telecommunications in
+China as well as its determinants. regional socioeconomic
+characteristics play an important role in determining the timing, speed,
+and level of mobile telecommunications diffusion in China
 
-\citet{bento2018time} explore What determines the duration of formative
-phases for energy innovations in different markets? We are interested
-both in initial markets (also: core, lead, first mover, early adopter)
-where formative phases prepare technologies for mass commercialization,
-and in follower markets (also: periphery, lag, late adopter) where
-accelerated formative phases may benefit from diffusion and spillovers.
+{[}18{]} explore What determines the duration of formative phases for
+energy innovations in different markets? We are interested both in
+initial markets (also: core, lead, first mover, early adopter) where
+formative phases prepare technologies for mass commercialization, and in
+follower markets (also: periphery, lag, late adopter) where accelerated
+formative phases may benefit from diffusion and spillovers.
 
@@
 
-\citet{wilson201281} growth function description in p.~86
+{[}13{]} growth function description in p.~86
 
@@
 
-\textbf{minimum R2 = 95\%}, see also \citet{grubler1990rise}
+\textbf{minimum R2 = 95\%}, see also {[}10{]}
 
@@
 
-Moran's I as \citet{ding2010modeling} \textbf{TODO} for t\_0, diffusion
-speed
+Moran's I as {[}11{]} \textbf{TODO} for t\_0, diffusion speed
 
@@
 
-contrary to results from future studies regarding social media
-\citep{lengyel2020role}, web technologies did not exclusively spread
-from a central location.
+contrary to results from future studies regarding social media {[}16{]},
+web technologies did not exclusively spread from a central location.
 
-\renewcommand\refname{References}
-  \bibliography{bibliography}
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{0}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-hagerstrand1968innovation}{}}%
+\CSLLeftMargin{1. }%
+\CSLRightInline{Hagerstrand T et al. Innovation diffusion as a spatial
+process. Innovation diffusion as a spatial process. Chicago, USA: Univ.
+Chicago Press.; 1968; }
+
+\leavevmode\vadjust pre{\hypertarget{ref-perkins2005international}{}}%
+\CSLLeftMargin{2. }%
+\CSLRightInline{Perkins R, Neumayer E. The international diffusion of
+new technologies: A multitechnology analysis of latecomer advantage and
+global economic integration. Annals of the Association of American
+Geographers. Taylor \& Francis; 2005;95: 789--808. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-zook2022mapping}{}}%
+\CSLLeftMargin{3. }%
+\CSLRightInline{Zook M, McCanless M. Mapping the uneven geographies of
+digital phenomena: The case of blockchain. The Canadian Geographer/Le
+G{é}ographe canadien. Wiley Online Library; 2022;66: 23--36. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-meade2021modelling}{}}%
+\CSLLeftMargin{4. }%
+\CSLRightInline{Meade N, Islam T. Modelling and forecasting national
+introduction times for successive generations of mobile telephony.
+Telecommunications Policy. Elsevier; 2021;45: 102088. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-fritsch2015new}{}}%
+\CSLLeftMargin{5. }%
+\CSLRightInline{Fritsch M, Medrano Echalar LF. New technology in the
+region--agglomeration and absorptive capacity effects on laser
+technology research in west germany, 1960--2005. Economics of Innovation
+and New Technology. Taylor \& Francis; 2015;24: 65--94. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-bento2015spatial}{}}%
+\CSLLeftMargin{6. }%
+\CSLRightInline{Bento N, Fontes M. Spatial diffusion and the formation
+of a technological innovation system in the receiving country: The case
+of wind energy in portugal. Environmental Innovation and Societal
+Transitions. Elsevier; 2015;15: 158--179. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-bednarz2020pulled}{}}%
+\CSLLeftMargin{7. }%
+\CSLRightInline{Bednarz M, Broekel T. Pulled or pushed? The spatial
+diffusion of wind energy between local demand and supply. Industrial and
+Corporate Change. Oxford University Press; 2020;29: 893--916. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-morrill2020spatial}{}}%
+\CSLLeftMargin{8. }%
+\CSLRightInline{Morrill R, Gaile GL, Thrall GI. Spatial diffusion. Sage
+Publications; 1988; }
+
+\leavevmode\vadjust pre{\hypertarget{ref-leibowicz2016representing}{}}%
+\CSLLeftMargin{9. }%
+\CSLRightInline{Leibowicz BD, Krey V, Grubler A. Representing spatial
+technology diffusion in an energy system optimization model.
+Technological Forecasting and Social Change. Elsevier; 2016;103:
+350--363. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-grubler1990rise}{}}%
+\CSLLeftMargin{10. }%
+\CSLRightInline{Grubler A. The rise and fall of infrastructures:
+Dynamics of evolution and technological change in transport.
+Physica-Verlag; 1990. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-ding2010modeling}{}}%
+\CSLLeftMargin{11. }%
+\CSLRightInline{Ding L, Haynes KE, Li H. Modeling the spatial diffusion
+of mobile telecommunications in china. The Professional Geographer.
+Taylor \& Francis; 2010;62: 248--263. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-perkins2011internet}{}}%
+\CSLLeftMargin{12. }%
+\CSLRightInline{Perkins R, Neumayer E. Is the internet really new after
+all? The determinants of telecommunications diffusion in historical
+perspective. The Professional Geographer. Taylor \& Francis; 2011;63:
+55--72. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-wilson201281}{}}%
+\CSLLeftMargin{13. }%
+\CSLRightInline{Wilson C. Up-scaling, formative phases, and learning in
+the historical diffusion of energy technologies. Energy Policy. 2012;50:
+81--94. doi:\url{https://doi.org/10.1016/j.enpol.2012.04.077}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-grubler1999dynamics}{}}%
+\CSLLeftMargin{14. }%
+\CSLRightInline{Grübler A, Nakićenović N, Victor DG. Dynamics of energy
+technologies and global change. Energy policy. Elsevier; 1999;27:
+247--280. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-beardsell1999spatial}{}}%
+\CSLLeftMargin{15. }%
+\CSLRightInline{Beardsell M, Henderson V. Spatial evolution of the
+computer industry in the USA. European Economic Review. Elsevier;
+1999;43: 431--456. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-lengyel2020role}{}}%
+\CSLLeftMargin{16. }%
+\CSLRightInline{Lengyel B, Bokányi E, Di Clemente R, Kertész J, González
+MC. The role of geography in the complex diffusion of innovations.
+Scientific reports. Nature Publishing Group; 2020;10: 1--11. }
+
+\leavevmode\vadjust pre{\hypertarget{ref-PAPAGIANNIDIS2015308}{}}%
+\CSLLeftMargin{17. }%
+\CSLRightInline{Papagiannidis S, Gebka B, Gertner D, Stahl F. Diffusion
+of web technologies and practices: A longitudinal study. Technological
+Forecasting and Social Change. 2015;96: 308--321.
+doi:\url{https://doi.org/10.1016/j.techfore.2015.04.011}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bento2018time}{}}%
+\CSLLeftMargin{18. }%
+\CSLRightInline{Bento N, Wilson C, Anadon LD. Time to get ready:
+Conceptualizing the temporal and spatial dynamics of formative phases
+for energy technologies. Energy Policy. Elsevier; 2018;119: 282--293. }
+
+\end{CSLReferences}
+
+\nolinenumbers
 

### 2023-03-22 paper/help/old/test.qmd

--- 2023-03-22 paper/help/old/paper.Rmd
@@
 ---
-title: Title of submission to PLOS journal
+title: "Digital economy in the UK; a multi-scalar story of the diffusion of web technologies"
+format: 
+  elsevier-pdf:
+    journal:
+      name: Journal Name
+      formatting: preprint
+      model: 3p
+      layout: onecolumn
+      cite-style: number
 author:
@@
     address: University of Bristol, School of Geographical Sciences, Bristol, UK; The Alan Turing Institute, London, UK
-abstract: |
-  blah blah
-
-author_summary: |
-  blah blah
-bibliography: mybibfile.bib
-output: rticles::plos_article
-csl: plos.csl
-header-includes:
-- \usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}
-editor_options: 
-  markdown: 
-    wrap: 72
 ---
-
-```{r echo=FALSE, message = FALSE, warning=FALSE}
-library(stargazer)
-library(tidyverse)
-#library(rmarkdown)
-```
 
@@
 <!-- motivation -->
-@meade2021modelling There are numerous agents who would benefit from
-accurate predictions of introduction times. These include: suppliers of
-network equipment for production planning; suppliers of mobile phones
-who need to have inventories of appropriately configured handsets;
-international network planners who need to provide sufficient bandwidth
-for communications between different countries; network operators and
-regulatory authorities.
 

### 2023-03-22 paper/help/old/tranos2023.Rmd

--- 2023-03-21 Untitled/Untitled.Rmd
@@
 ---
-title: Digital economy in the UK: a multi-scalar story of the diffusion of web technologies
+title: Digital economy in the UK; a multi-scalar story of the diffusion of web technologies
 author:
@@
     correspondingauthor: true
-    footnote: 1
 address:
@@
     address: University of Bristol, School of Geographical Sciences, Bristol, UK; The Alan Turing Institute, London, UK
-footnote:
-  - code: 1
-    text: "This is the first author footnote."
-  - code: 2
-    text: "Another author footnote."
 abstract: |

### 2023-03-22 paper/help/old/tranos2023.tex

--- 2023-03-21 paper/submission1/tranos2023.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-  authoryear,
-  preprint,
-  3p]{elsarticle}
+\documentclass[preprint, 3p,
+authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
+%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
 
-\usepackage{amsmath,amssymb}
+\usepackage[hyphens]{url}
+
+  \journal{An awesome journal} % Sets Journal name
+
+\usepackage{lineno} % add
+
+\usepackage{graphicx}
+%%%%%%%%%%%%%%%% end my additions to header
+
+\usepackage[T1]{fontenc}
 \usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
   \fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
+
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Digital economy in the UK; a multi-scalar story of the diffusion of web technologies},
+            colorlinks=false,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+
 \setcounter{secnumdepth}{5}
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+% Pandoc toggle for numbering sections (defaults to be off)
 
+% tightlist command for lists without linebreak
 \providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\AtBeginDocument{%
-\ifdefined\contentsname
-  \renewcommand*\contentsname{Table of contents}
-\else
-  \newcommand\contentsname{Table of contents}
-\fi
-\ifdefined\listfigurename
-  \renewcommand*\listfigurename{List of Figures}
-\else
-  \newcommand\listfigurename{List of Figures}
-\fi
-\ifdefined\listtablename
-  \renewcommand*\listtablename{List of Tables}
-\else
-  \newcommand\listtablename{List of Tables}
-\fi
-\ifdefined\figurename
-  \renewcommand*\figurename{Figure}
-\else
-  \newcommand\figurename{Figure}
-\fi
-\ifdefined\tablename
-  \renewcommand*\tablename{Table}
-\else
-  \newcommand\tablename{Table}
-\fi
-}
-\@ifpackageloaded{float}{}{\usepackage{float}}
-\floatstyle{ruled}
-\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
-\floatname{codelisting}{Listing}
-\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\journal{Journal Name}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\usepackage[]{natbib}
-\bibliographystyle{elsarticle-harv}
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  pdftitle={Digital economy in the UK: a multi-scalar story of the diffusion of web technologies},
-  pdfauthor={Emmanouil Tranos},
-  pdfkeywords={keyword1, keyword2},
-  colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
-
-\setlength{\parindent}{6pt}
 \begin{document}
@@
 \begin{frontmatter}
-\title{Digital economy in the UK: a multi-scalar story of the diffusion
-of web technologies}
-\author[1]{Emmanouil Tranos%
+
+  \title{Digital economy in the UK; a multi-scalar story of the
+diffusion of web technologies}
+    \author[University of Bristol and The Alan Turing
+Institute]{Emmanouil Tranos%
   \corref{cor1}%
@@
    \ead{e.tranos@bristol.ac.uk} 
-
-\affiliation[1]{organization={University of Bristol and The Alan Turing
-Institute},postcode={UK},postcodesep={}}
-
+      \affiliation[University of Bristol and The Alan Turing
+Institute]{University of Bristol, School of Geographical Sciences,
+Bristol, UK; The Alan Turing Institute, London, UK}
     \cortext[cor1]{Corresponding author}
@@
   have been, they cannot capture the micro-processes and the
-characteristics of the individual online behaviour. Using large volumes
-of archived and geolocated web content, this paper models the diffusion
-of web technologies over space and time in the UK. The data and
-geolocation strategy allow to capture these processes at a very granular
-spatial scale. The modelling approach, which is based on simple spatial
-analytical methods and on the estimation of diffusion curves at various
-scales, enables to depict the role of geography and other cognitive
-factors which drove the diffusion of web technologies. Although the
-focus is on a recent historical period -- 1996-2012 -- the results of
-the analysis depict diffusion mechanisms which can be very useful in
-understanding the evolutionary patterns of the adoption of other newer
-technologies.
+  characteristics of the individual online behaviour. Using large
+  volumes of archived and geolocated web content, this paper models the
+  diffusion of web technologies over space and time in the UK. The data
+  and geolocation strategy allow to capture these processes at a very
+  granular spatial scale. The modelling approach, which is based on
+  simple spatial analytical methods and on the estimation of diffusion
+  curves at various scales, enables to depict the role of geography and
+  other cognitive factors which drove the diffusion of web technologies.
+  Although the focus is on a recent historical period -- 1996-2012 --
+  the results of the analysis depict diffusion mechanisms which can be
+  very useful in understanding the evolutionary patterns of the adoption
+  of other newer technologies.
   \end{abstract}
-
     \begin{keyword}
@@
   \end{keyword}
+
  \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, frame hidden, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 \renewcommand\refname{References}
-  \bibliography{bibliography}
+\bibliography{mybibfile.bib}
 

### 2023-03-22 paper/submission1/help.Rmd

@@
 In reality, the synthesis of these two processes represents how new technologies diffuse over space and time [@morrill2020spatial].
+
+grubler1990rise The scale of spatial innovation diffusion is similar to
+that of temporal diffusion models: hierarchically decomposed.
 
@@
 
-Space: absorptive capacity
+<!-- Space: absorptive capacity -->
 
-Evolutionary economic geography >> path dependency
+<!-- Evolutionary economic geography >> path dependency -->
+
+@bento2015spatial The concept of absorptive capacity has also been
+applied in the TIS context in terms of the con-struction of capacity
+needed to transfer and implement a new technology in the receiving
+country (cf.Van Alphen, 2011), with repercussions for the nature of
+system functions
+
+@bednarz2020pulled larger local technological niches (at least from a
+demand perspective) are more likely to be formed in urban areas than in
+rural ones, which is in line with the work of Ha¨gerstrand (1965a,b).
+However, size is not the only regional characteristic that may matter in
+this context. Regions also differ in accumulated experiences and the
+presence of tacit knowledge with respect to products, as well as in
+actors' propensity share this knowledge among producers (Martin et al.,
+2019)

### 2023-03-22 paper/submission1/tranos2023.qmd

--- 2023-03-21 paper/submission1/tranos2023-IT074305.qmd
@@
 
-# Introduction {#sec:1}
+# Introduction {#sec1}
 
@@
 innovations diffuse across space and, importantly, how such
-spatio-temporal processes can be modelled. The seminal contribution of
+spatio-temporal processes can be modelled. After all, diffusion together 
+with invention and innovation are considered the pillars of 
+technological change [@das2022diffusion]. The seminal contribution of
 @hagerstrand1968innovation is illustrative of this early interest.
@@
 sociology well before the 'cultural turn' of economic geography
-[@perkins2005international]. A potential explanation of lack of studies
-exploring the diffusion of new and, more specifically, digital
-technologies across *both* space and time can be attributed to the
-scarcity of relevant and granular enough data. As @zook2022mapping
-highlight, digital activities are hardly ever captured in official data.
+[@perkins2005international]. A potential explanation of the lack of 
+geographical studies exploring the diffusion of new and, more 
+specifically, digital technologies across *both* space and time can be
+attributed to the scarcity of relevant and granular enough data. As 
+@zook2022mapping highlight, digital activities are hardly ever captured 
+in official data.
 
-This paper aims to reinstate geographers' interest in how new and
-digital technologies diffuse over space and time. It adopts a geographic
-standpoint to explore and map the diffusion of web technologies in the
-UK over space and time. It does so by employing a novel source of big
-data which captures the active engagement with web technologies during
-the $1996-2013$ period.
+This paper offers such a contribution: a geographical study illustrating 
+how a new technology that is the web diffused over space and time 
+in the UK at a high level of spatial granularity during the $1996-2012$ 
+period. It does so by employing a novel source of big data which 
+captures the active engagement with web technologies during that period.
+By addressing this empirical question this paper exemplifies how the 
+combination of data sources which escape the traditional social science
+domain and adequate research methods can offer new contributions 
+regarding our understanding of technologies are diffused.
 
@@
 stakeholders who are interested in knowing how new digital technologies
-diffused over space and time use this information to make predictions
-regarding the diffusion of related future technologies. As per
-@meade2021modelling a variety of actors have a direct interest in
-gaining such knowledge including network equipment suppliers; network
-operators, regulatory and local authorities. And despite the spatial
-heterogeneity of such diffusion mechanisms and the policy relevance,
-there are very limited attempts in the literature to analyse the
-diffusion of new digital technologies at a granular level.
+diffused over space and time and use this information to make predictions
+regarding the diffusion of related *future* technologies.
+As per @leibowicz2016representing, historical studies agree on the fact that 
+"technologies diffuse at different times, at different rates, and to
+different extents in different places, and can be significantly
+influenced by policies [@victor1993]". @meade2021modelling highlight that 
+a variety of actors have a direct interest in gaining such knowledge 
+including network equipment suppliers; network operators, regulatory 
+and local authorities. 
+These processes and their effects very a lot between scales: although
+the diffusion of a new technology might not be optimal at a local level,
+it might be beneficial from a global perspective as it might enable 
+faster diffusion to less advantaged places @leibowicz2016representing. 
+Despite the spatial heterogeneity of such diffusion mechanisms and 
+the policy relevance, there are very limited attempts in the literature 
+to analyse the diffusion of new digital technologies at a detailed 
+geographical level.
 
-the paper answers an empirical question: how did a
+<!-- diffusion and adoption -->
 
-@zook2022mapping Because new digital activities are rarely---if
-ever---captured in official state data, researchers must rely on
-information gathered from alternative sources. With this in mind, the
-overarching aim of this article is outlining an approach for analyzing
-the "new spaces and geographies" of digital phenomena and practices.
+Technological diffusion, which is by definition an aggregated process, 
+can be discussed in parallel with individual adoption mechanisms.
+On the one hand, @rogers2010diffusion identifies early adopter of new 
+technologies as 'knowledgeable risk takers' and @griliches1957 as 'profit 
+maximisers' [@ding2010modeling]. Such individual agents are rewarded because 
+of their attitude towards new technologies and innovations. On the other hand,
+@perkins2011internet attribute diffusion to two processes: (i) epidemic-type 
+of mechanisms, which
 
-<!-- motivation -->
-
-Classifying adopters of a technology along a time axis was pioneered by
-Rogers (1962), who categorized adopters into five groups as innovators,
-early adopters, early majority, late majority and laggards.
-
-An a priori approach groups similar countries into segments, where
-similarity is defined using economic, social, or political covariates.
-Examples include Hofstede (2001) and Lee (1990). An a posteriori
-analysis considers segments based on realized market behaviour, for
-example see Sood et al. (2009). Some studies investigate changing
-segment membership over time, see Cannon and Yaprak (2011).
-
-@fritsch2015new Hägerstrand (1952, 1967) was the first scholar to
-intensively analyze spatial diffusion pro cesses. Based on the empirical
-observations of the regional spread of process innovations, he
-hypothesized that new technology is first implemented in the 'center' --
-the large agglom erations -- and then moves 'down' the spatial
-hierarchy, the last stop being the 'periphery', i.e. remote and sparsely
-populated regions. There are at least two factors that may explain this
-pattern: first, the transfer of tacit knowledge via face-to-face contact
-and second, the absorptive capacity of a region.
-
-<!-- key concepts -->
-
-@bento2015spatial The concept of absorptive capacity has also been
-applied in the TIS context in terms of the con-struction of capacity
-needed to transfer and implement a new technology in the receiving
-country (cf.Van Alphen, 2011), with repercussions for the nature of
-system functions
-
-@bednarz2020pulled larger local technological niches (at least from a
-demand perspective) are more likely to be formed in urban areas than in
-rural ones, which is in line with the work of Ha¨gerstrand (1965a,b).
-However, size is not the only regional characteristic that may matter in
-this context. Regions also differ in accumulated experiences and the
-presence of tacit knowledge with respect to products, as well as in
-actors' propensity share this knowledge among producers (Martin et al.,
-2019)
-
-<!-- @das2022diffusion -->
-
-<!-- Epidemic approach: The approach emphasizes that the adoption of a -->
-
-<!-- technology spreads like an infection among the potential adopters. By -->
-
-<!-- taking a homogeneous group of potential adopters as an assumption, it -->
-
-<!-- explains the adoption of any innovation. -->
-
-To summarize the discussion, first, here the diffusion process is
-selfperpetuating as initial use stimulates further use. Second, it
-follows a disequilibrium path as the level of users is always lower than
-the number of potential users (Stoneman 2002).
-
+Two main mechanisms are identified in the
+literature to explain diffusion: (1) epidemic-type dynamics whereby
+contact with previous adopters stimulates uptake as po tential adopters
+learn about a new innovation; and (2) economic-type mechanisms whereby
+potential users adopt a new innovation as it be comesmore profitable,
+useful, or valuable, with uptake characteristically spreading as costs
+be come lower, performance improves, or the po tential uses of the
+innovation grow over time.
 The criticisms are mainly based on the fact that although the approach
@@
 
-<!-- Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.) -->
+@morrill2020spatial understands spatial diffusion as the landscape 
+changes that spread across space because of changes that take place 
+elsewhere.
 
-<!-- Early adopters: next 13.5%, (III.) Early Majority: following 34%, (IV.) -->
-
-<!-- Late Majority: next 34%, and the rest as (V.) Laggards -->
-
-@morrill2020spatial Spatial diffusion is the process by which behavior
-or characteristics of the landscape change as a result of what happens
-elsewhere earlier. Spatial diffusion is the spread of the phenomenon,
-over space and timed, from limited origins
-
-Contagious diffusion=\> spatial autocorelation
-
-expansion-type diffusion
-
-<!-- the three geographic diffusion processes are (1) purely contagious where -->
-
-<!-- distance or adjacency is the absolute controlling factor, (2) purely -->
-
-<!-- hierarchical where size or urban position in the central place hierarchy -->
-
-<!-- is the absolute controlling factor, or (3) where the location of change -->
-
-<!-- is purely random. These are the pure or ideal forms of the diffusion -->
-
-<!-- process; all real diffusion processes are the result of a combination of -->
-
-<!-- these extremes. -->
-
-The work by the communication sociologist Everett Rogers (1962, 1971,
-1983) has emphasized the role of information, communication, formal and
-informal media, opinion leaders and social networks, and economic and
-psychological constraints on acceptance. Rogers's work stresses the
-decision mechanism of the potential adopter; the work of Rogers
-partitions the adopter's process of choosing to accept a new phenomenon
-or trait into five stages: stage 1--the potential adopter gains
-knowledge or awareness of an innovation; stage 2--persuasion is
-exercised to adopt; stage 3--a decision is made to adopt; stage 4--the
-decision to adopt is implemented; stage 5--the adopter confirms the
-decision to adopt. In Rogers's stage theory there may be a substantial
-time lag between when a potential adopter becomes aware of the new
-characteristic and when a decision is actually made to adopt.
-
-Hägerstrand considered diffusion to be a fundamental geographic process:
-Whatever the phenomenon being diffused might be, one may consider it in
-the context of a larger universal process of spatial diffusion.
-
-<!-- @das2022diffusion [not good paper] Invention, innovation and diffusion -->
-
-<!-- are considered as the three pillars of the theory of technological -->
-
-<!-- change and Schumpeter is considered as the father of the study of -->
-
-<!-- diffusion -->
-
-@leibowicz2016representing Diffusion is a spatial (Hagerstrand, 1967) as
-well as a temporal process, and historical evidence confirms that
-technologies diffuse at different times, at different rates, and to
-different extents in different places, and can be significantly
-influenced by policies (Victor, 1993).
-
-<!-- @leibowicz2016representing -->
-
-<!-- An illustrative empirically based theory for technology -->
-
-<!-- diffusion across multiple regions is suggested by "Schmidt's Law" -->
-
-<!-- (Grubler, 1990). With respect to a specific technology, it divides -->
-
-<!-- regions into three groups: core, rim, and periphery. The core is where -->
-
-<!-- the technology is invented, subject to experimentation, and first -->
-
-<!-- deployed commercially. To develop the technology, the core must have -->
-
-<!-- both sufficient innovative ability and economic motivation to make the -->
-
-<!-- necessary resource commitment. As the technology spreads from the core -->
-
-<!-- to the rim and eventually to the periphery, later adopters have the -->
-
-<!-- advantage of benefiting from the experiences of earlier adopters via -->
-
-<!-- knowledge spillover effects (Verdolini and Galeotti, 2011). At this -->
-
-<!-- point, the technology has likely been refined and costs have likely -->
-
-<!-- declined significantly. Schmidt's Law claims that, as a result, -->
-
-<!-- diffusion proceeds at a faster rate in regions which adopt the -->
-
-<!-- technology later. However, diffusion tends to be less pervasive in -->
-
-<!-- regions where adoption is delayed, ultimately saturating at a lower -->
-
-<!-- extent than in regions where adoption occurred earlier. This is often -->
-
-<!-- attributed to a lack of proper institutions and infrastructures in later -->
-
-<!-- adopters, markets whichmay not be as favorable to a technology, or the -->
-
-<!-- availability of yet newer options that invite technological -->
-
-<!-- leapfrogging. -->
-
-Lastly, it can be globally optimal for innovative economies to deploy
-advanced technologies more than what is locally optimal if this enables
-faster diffusion in less advanced regions that could benefit from the
-technology.
+<!-- data and early engagement with the web -->
 
@@
 
-<!-- @grubler1990rise Following the definition given by Rogers, 1983, -->
+@perkins2005international Theoretical models disagree as to why firms
+adopt innovations at different times. Epidemic models emphasize
+information (Griliches 1957). Certain firms are hypothesized to adopt
+earlier because they come into contact with, and learn from, adopters of
+the new technology before others. Economic models, on the other hand,
+predominantly emphasize firm heterogeneity (Ireland and Stoneman 1986).
+Firms adopt technologies at different times because they differ with
+respect to various organizational and environmental variables
+influencing the economic returns from adoption (Blackman 1999).
 
-<!-- diffusion is the process by which an innovation is communicated through -->
-
-<!-- certain communication channels over time among the members of a social -->
-
-<!-- system. An innovation may be an idea, object or practice which is -->
-
-<!-- perceived as new by an individual or another unit of adoption (e.g., an -->
-
-<!-- organization) . Communication via communication channels refers to how -->
-
-<!-- knowledge about an innovation is transferred between the members of a -->
-
-<!-- social system. Communication takes time, and is therefore an important -->
-
-<!-- aspect of innovation diffusion. In fact, all theoretical and empirical -->
-
-<!-- diffusion studies agree that an innovation does not instantly diffuse -->
-
-<!-- into a social system. Instead a typical time pattern of diffusion along -->
-
-<!-- a S-shaped trajectory seems to be the rule. The S-shaped pattern of -->
-
-<!-- diffusion appears to be a basic anthropologic phenomenon, as it is also -->
-
-<!-- confirmed by a large number of studies· of primitive preindustrial -->
-
-<!-- societies. -->
-
-grubler1990rise The scale of spatial innovation diffusion is similar to
-that of temporal diffusion models: hierarchically decomposed.
-
-<!-- grubler1990rise -->
-
-<!-- Spatial innovation diffusion in Hagerstrand's model consists of three -->
-
-<!-- stages. Stage 1: Local concentrations of initial acceptances (initial -->
-
-<!-- agglomeration) . Stage 2: Radial dissemination outward from the initial -->
-
-<!-- agglomeration, accompanied by the rise of secondary agglomerations. -->
-
-<!-- Stage 3: The condensing and saturation stage, in which filling-in occurs -->
-
-<!-- and diffusion eventually ceases. -->
-
-<!-- grubler1990rise -->
-
-<!-- Thus, three main characteristics define the spatial diffusion process: -->
-
-<!-- (1) As in temporal diffusion models, an S-shaped pattern in the -->
-
-<!-- cumulative level of adoption. (2) A "hierarchy effect" by which -->
-
-<!-- diffusion proceeds from initial centers to secondary ones. Within the -->
-
-<!-- context of urban systems (central places) this implies a hierarchical -->
-
-<!-- spread of an innovation from larger to smaller centers, and from the -->
-
-<!-- centers to the surrounding hinterland. (3) A "neighborhood effect" Le., -->
-
-<!-- diffusion proceeds outwards from innovation centers, first "hitting" -->
-
-<!-- (Brown, 1981) nearby rather than far-away locations. -->
+<!-- methods -->
 
@@
 
-@perkins2005international Theoretical models disagree as to why firms
-adopt innovations at different times. Epidemic models emphasize
-information (Griliches 1957). Certain firms are hypothesized to adopt
-earlier because they come into contact with, and learn from, adopters of
-the new technology before others. Economic models, on the other hand,
-predominantly emphasize firm heterogeneity (Ireland and Stoneman 1986).
-Firms adopt technologies at different times because they differ with
-respect to various organizational and environmental variables
-influencing the economic returns from adoption (Blackman 1999).
-
-Either way, the strong assumption is that developing countries can
-acquire modern technology innovated in developed economies, often at a
-fraction of the original research and development (R&D) costs, thereby
-leapfrogging many decades of technological progress (Teece 2000).
-
-The first involves a country's geographical location. Recent empirical
-work suggests that diffusion is ''geographically localized'' (Globerman,
-Kokko, and Sjo¨holm 2000; Keller 2002; Milner 2003) in that a technology
-diffuses faster in a country where it is already more widely diffused in
-neighboring countries. Underlying these regional effects are contagion
-and contact with prior users or producers of technology.
-
-@ding2010modeling Rogers (1995) characterized early adopters as
-knowledgeable risk takers. Griliches (1957) looked at the rate of return
-for early adopters and characterized them as profit maximizers. The
-geographer Hagerstrand (1967) was focusing on the character of early
-adopters in the transmittal process that produces follow-on participants
-in the process. Space was treated as a contiguity system with embedded
-characteristics producing barrier elements to the diffusion process.
-
-@ding2010modeling REVIEW ON Technological Diffusion
-
-<!-- Hagerstrand (1968) was the first to demonstrate that diffusion typically -->
-
-<!-- starts from innovative centers ('core´) and disseminates, through a -->
-
-<!-- hierarchy of subcenters, to the periphery. This process is sequential -->
-
-<!-- rather than simultaneous, and tends to accelerate from core to -->
-
-<!-- periphery, where diffusion reaches a lower intensity than in the core -->
-
-<!-- (Hagerstrand, 1968; Grubler, 1990; Morrill, 2005). A similar effect of -->
-
-<!-- spatial acceleration has been identified in marketing research and is -->
-
-<!-- known as the "lead-lag effect" (Peres et al., 2010). -->
-
-@perkins2011internet Two main mechanisms are identified in the
-literature to explain diffusion: (1) epidemic-type dynamics whereby
-contact with previous adopters stimulates uptake as po tential adopters
-learn about a new innovation; and (2) economic-type mechanisms whereby
-potential users adopt a new innovation as it be comesmore profitable,
-useful, or valuable, with uptake characteristically spreading as costs
-be come lower, performance improves, or the po tential uses of the
-innovation grow over time.
-
-To this extent, our results contribute to a growing body of work that
-has sought to caution against claims about the supposed novelty of the
-Internet and the suggestion that it is somehow different.
-
 <!-- methods -->
@@
 
-# Literature review {#sec2} MAYBE SKIP THIS??
+# Literature review {#sec2}
+
+MAYBE SKIP THIS??
 
@@
 
+@perkins2005internationalEither way, the strong assumption is that developing countries can
+acquire modern technology innovated in developed economies, often at a
+fraction of the original research and development (R&D) costs, thereby
+leapfrogging many decades of technological progress (Teece 2000).
+
+@perkins2005international The first involves a country's geographical location. Recent empirical
+work suggests that diffusion is ''geographically localized'' (Globerman,
+Kokko, and Sjo¨holm 2000; Keller 2002; Milner 2003) in that a technology
+diffuses faster in a country where it is already more widely diffused in
+neighboring countries. Underlying these regional effects are contagion
+and contact with prior users or producers of technology.
+
 # Materials and Methods {#sec3}

### 2023-03-22 paper/submission1/tranos2023.tex

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, frame hidden, interior hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, frame hidden]}{\end{tcolorbox}}\fi
 
-\hypertarget{sec:1}{%
-\section{Introduction}\label{sec:1}}
+\hypertarget{sec1}{%
+\section{Introduction}\label{sec1}}
 
@@
 innovations diffuse across space and, importantly, how such
-spatio-temporal processes can be modelled. The seminal contribution of
-\citet{hagerstrand1968innovation} is illustrative of this early
+spatio-temporal processes can be modelled. After all, diffusion together
+with invention and innovation are considered the pillars of
+technological change \citep{das2022diffusion}. The seminal contribution
+of \citet{hagerstrand1968innovation} is illustrative of this early
 interest. However, the torch of exploring and modelling such processes
@@
 and sociology well before the `cultural turn' of economic geography
-\citep{perkins2005international}. A potential explanation of lack of
-studies exploring the diffusion of new and, more specifically, digital
-technologies across \emph{both} space and time can be attributed to the
-scarcity of relevant and granular enough data. As
+\citep{perkins2005international}. A potential explanation of the lack of
+geographical studies exploring the diffusion of new and, more
+specifically, digital technologies across \emph{both} space and time can
+be attributed to the scarcity of relevant and granular enough data. As
 \citet{zook2022mapping} highlight, digital activities are hardly ever
@@
 
-This paper aims to reinstate geographers' interest in how new and
-digital technologies diffuse over space and time. It adopts a geographic
-standpoint to explore and map the diffusion of web technologies in the
-UK over space and time. It does so by employing a novel source of big
-data which captures the active engagement with web technologies during
-the \(1996-2013\) period.
+This paper offers such a contribution: a geographical study illustrating
+how a new technology that is the web diffused over space and time in the
+UK at a high level of spatial granularity during the \(1996-2012\)
+period. It does so by employing a novel source of big data which
+captures the active engagement with web technologies during that period.
+By addressing this empirical question this paper exemplifies how the
+combination of data sources which escape the traditional social science
+domain and adequate research methods can offer new contributions
+regarding our understanding of technologies are diffused.
 
@@
 stakeholders who are interested in knowing how new digital technologies
-diffused over space and time use this information to make predictions
-regarding the diffusion of related future technologies. As per
-\citet{meade2021modelling} a variety of actors have a direct interest in
-gaining such knowledge including network equipment suppliers; network
-operators, regulatory and local authorities. And despite the spatial
-heterogeneity of such diffusion mechanisms and the policy relevance,
-there are very limited attempts in the literature to analyse the
-diffusion of new digital technologies at a granular level.
-
-the paper answers an empirical question: how did a
-
-\citet{zook2022mapping} Because new digital activities are rarely---if
-ever---captured in official state data, researchers must rely on
-information gathered from alternative sources. With this in mind, the
-overarching aim of this article is outlining an approach for analyzing
-the ``new spaces and geographies'' of digital phenomena and practices.
-
-Classifying adopters of a technology along a time axis was pioneered by
-Rogers (1962), who categorized adopters into five groups as innovators,
-early adopters, early majority, late majority and laggards.
-
-An a priori approach groups similar countries into segments, where
-similarity is defined using economic, social, or political covariates.
-Examples include Hofstede (2001) and Lee (1990). An a posteriori
-analysis considers segments based on realized market behaviour, for
-example see Sood et al.~(2009). Some studies investigate changing
-segment membership over time, see Cannon and Yaprak (2011).
-
-\citet{fritsch2015new} Hägerstrand (1952, 1967) was the first scholar to
-intensively analyze spatial diffusion pro cesses. Based on the empirical
-observations of the regional spread of process innovations, he
-hypothesized that new technology is first implemented in the `center' --
-the large agglom erations -- and then moves `down' the spatial
-hierarchy, the last stop being the `periphery', i.e.~remote and sparsely
-populated regions. There are at least two factors that may explain this
-pattern: first, the transfer of tacit knowledge via face-to-face contact
-and second, the absorptive capacity of a region.
-
-\citet{bento2015spatial} The concept of absorptive capacity has also
-been applied in the TIS context in terms of the con-struction of
-capacity needed to transfer and implement a new technology in the
-receiving country (cf.Van Alphen, 2011), with repercussions for the
-nature of system functions
-
-\citet{bednarz2020pulled} larger local technological niches (at least
-from a demand perspective) are more likely to be formed in urban areas
-than in rural ones, which is in line with the work of Ha¨gerstrand
-(1965a,b). However, size is not the only regional characteristic that
-may matter in this context. Regions also differ in accumulated
-experiences and the presence of tacit knowledge with respect to
-products, as well as in actors' propensity share this knowledge among
-producers (Martin et al., 2019)
+diffused over space and time and use this information to make
+predictions regarding the diffusion of related \emph{future}
+technologies. As per \citet{leibowicz2016representing}, historical
+studies agree on the fact that ``technologies diffuse at different
+times, at different rates, and to different extents in different places,
+and can be significantly influenced by policies
+\citep{victor1993}''.@meade2021modelling highlight that a variety of
+actors have a direct interest in gaining such knowledge including
+network equipment suppliers; network operators, regulatory and local
+authorities. These processes and their effects very a lot between
+scales: although the diffusion of a new technology might not be optimal
+at a local level, it might be beneficial from a global perspective as it
+might enable faster diffusion to less advantaged places
+\citet{leibowicz2016representing}. Despite the spatial heterogeneity of
+such diffusion mechanisms and the policy relevance, there are very
+limited attempts in the literature to analyse the diffusion of new
+digital technologies at a detailed geographical level.
 
@@
 
-Contagious diffusion=\textgreater{} spatial autocorelation
-
-expansion-type diffusion
-
-The work by the communication sociologist Everett Rogers (1962, 1971,
-1983) has emphasized the role of information, communication, formal and
-informal media, opinion leaders and social networks, and economic and
-psychological constraints on acceptance. Rogers's work stresses the
-decision mechanism of the potential adopter; the work of Rogers
-partitions the adopter's process of choosing to accept a new phenomenon
-or trait into five stages: stage 1--the potential adopter gains
-knowledge or awareness of an innovation; stage 2--persuasion is
-exercised to adopt; stage 3--a decision is made to adopt; stage 4--the
-decision to adopt is implemented; stage 5--the adopter confirms the
-decision to adopt. In Rogers's stage theory there may be a substantial
-time lag between when a potential adopter becomes aware of the new
-characteristic and when a decision is actually made to adopt.
-
-Hägerstrand considered diffusion to be a fundamental geographic process:
-Whatever the phenomenon being diffused might be, one may consider it in
-the context of a larger universal process of spatial diffusion.
-
-\citet{leibowicz2016representing} Diffusion is a spatial (Hagerstrand,
-1967) as well as a temporal process, and historical evidence confirms
-that technologies diffuse at different times, at different rates, and to
-different extents in different places, and can be significantly
-influenced by policies (Victor, 1993).
-
-Lastly, it can be globally optimal for innovative economies to deploy
-advanced technologies more than what is locally optimal if this enables
-faster diffusion in less advanced regions that could benefit from the
-technology.
+Compare dissusion with adoption
 
@@
 
-\hypertarget{literature-review-sec2-maybe-skip-this}{%
-\section{Literature review \{\#sec2\} MAYBE SKIP
-THIS??}\label{literature-review-sec2-maybe-skip-this}}
+\hypertarget{sec2}{%
+\section{Literature review}\label{sec2}}
+
+MAYBE SKIP THIS??
 
@@
 \renewcommand\refname{References}
-  \bibliography{bibliography}
+  \bibliography{./paper/submission1/bibliography}
 

### 2023-03-22 presentations/gceg.Rmd

@@
 
-- A **hierarchy effect**: from main centers to secondary ones -- central places
+- A **hierarchy effect**: from main centres to secondary ones -- central places
 
-- A **neighborhood effect**: diffusion proceeds outwards from innovation centers, 
+- A **neighborhood effect**: diffusion proceeds outwards from innovation centres, 
 first "hitting" nearby rather than far-away locations [@grubler1990rise]
 
-@hagerstrand1965monte: from innovative centers (core) through a
-hierarchy of subcenters, to the periphery
+@hagerstrand1965monte: from innovative centres (core) through a
+hierarchy of sub-centres, to the periphery
 
@@
 ```{r Nottingham, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000018.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000018.png")
 knitr::include_graphics(path.image)
@@
 ```{r Reading, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000038.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000038.png")
 knitr::include_graphics(path.image)
@@
 ```{r City_of_London, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E09000001.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000001.png")
 knitr::include_graphics(path.image)
@@
 ```{r Leicester, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000016.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000016.png")
 knitr::include_graphics(path.image)
@@
 ```{r Bristol, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000023.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000023.png")
 knitr::include_graphics(path.image)
@@
 ```{r Plymouth, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E06000026.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000026.png")
 knitr::include_graphics(path.image)
@@
 ```{r Liverpool, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E08000012.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E08000012.png")
 knitr::include_graphics(path.image)
@@
 ```{r Hackney, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_E09000012.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000012.png")
 knitr::include_graphics(path.image)
@@
 ```{r Edinburgh, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/lad_S12000036.png")
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_S12000036.png")
 knitr::include_graphics(path.image)

### 2023-03-22 presentations/web_diffusion.Rmd

@@
       height: 720
-bibliography: mybibfile.bib
+bibliography: bibliography.bib
 biblio-style: "apalike"
@@
 -   S-shaped diffusion curves
+-   Rank dynamics
 -   Conclusions <br> <br>
@@
 -   **Granular** and **multi-scale** spatial perspective
+
+## Aims
+
+<br>
+
+-   Geographers used to be interested in diffusion
+
+-   @hagerstrand1968innovation
+
+-   Passed the torch to economists and sociologists
+
+-   Why? Lack of granular data:
+
+> Because new digital activities are rarely---if ever---captured in
+> official state data, researchers must rely on information gathered
+> from alternative sources [@zook2022mapping].
 
@@
 
-## Internet geographies
+# Web data
 
--   Map internet infrastructure
+## Long story short
 
-    -   backbone [@malecki2002economic; @tranos2013geography]
+-   Archived web data
 
-    -   last mile [@riddlesden2014broadband; @budnitz2022working]
+-   Observe commercial websites 1996 - 2012
 
--   Social media and user generated content
+-   Geolocate to a unique location
 
-    -   Here and now effect [@crampton2013beyond]
-
-> Because new digital activities are rarely---if ever---captured in
-> official state data, researchers must rely on information gathered
-> from alternative sources [@zook2022mapping].
-
-# Web data
+-   Geolocate to multiple locations
 
@@
 
-# Analysis
+# Web diffusion
 
@@
 
-<!-- ------------------------------------------------------------------------ -->
+------------------------------------------------------------------------
 
@@
 
-**Neighborhood effect**: diffusion proceeds outwards from innovation
-centers, first "hitting" nearby rather than far-away locations
-[@grubler1990rise]
+> **Neighborhood effect**: diffusion proceeds outwards from innovation
+> centers, first "hitting" nearby rather than far-away locations
+> [@grubler1990rise]
 
--   Moran's I
-
--   LISA maps
+-   Spatial dependency (Moran's I & LISA maps)
 
@@
 
-<!-- ------------------------------------------------------------------------ -->
+-   Websites in Output Areas (c. 200,000)
 
-```{r, eval=FALSE, moransi_oa, results='asis'}
+------------------------------------------------------------------------
+
+```{r moransi_la, results='asis'}
+path.in <- paste0(path, "/outputs/morani_la.csv")
+
+read.csv(path.in) %>% dplyr::select(year, morani) %>% 
+  ggplot(aes(x=year, y=morani)) +
+  geom_bar(stat = "identity") + 
+  ylab('Moran\'s I') +
+  theme_minimal() +
+  ggtitle("Moran\'s I, websites per firm a Local Authority") +
+  theme(plot.title = element_text(hjust = 0.5)) 
+```
+
+------------------------------------------------------------------------
+
+```{r, eval=T, moransi_oa, results='asis'}
 path.in <- paste0(path, "/outputs/morani_oa.csv")
@@
   theme_minimal() +
-  ggtitle("Moran\'s I, number of websites per Output Areas") +
+  ggtitle("Moran\'s I, websites per Output Area") +
   theme(plot.title = element_text(hjust = 0.5)) 
-```
-
-------------------------------------------------------------------------
-
-```{r moransi_la, results='asis'}
-path.in <- paste0(path, "/outputs/morani_la.csv")
-
-read.csv(path.in) %>% dplyr::select(year, morani) %>% 
-  ggplot(aes(x=year, y=morani)) +
-  geom_bar(stat = "identity") + 
-  ylab('Moran\'s I') +
-  theme_minimal() +
-  ggtitle("Moran\'s I, number of websites per Local Authority") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r lisa_oa, eval=FALSE,results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/")
-
-list.files(path = path.images, pattern = "^lisa_level_pc1_oa", full.names = T) %>% 
-  map(image_read) %>% # reads each path file
-  image_join() %>% # joins image
-  image_animate(delay=50) #%>% # animates, can opt for number of loops
-  #image_write("test.gif") # write to current dir
 ```
@@
 ```
+
+------------------------------------------------------------------------
+
+```{r lisa_oa, eval=T,results='asis', out.width = "30%"}
+path.images <- paste0(path, "/outputs/")
+
+list.files(path = path.images, pattern = "^lisa_level_pc1_oa", full.names = T) %>% 
+  map(image_read) %>% # reads each path file
+  image_join() %>% # joins image
+  image_animate(delay=50) #%>% # animates, can opt for number of loops
+  #image_write("test.gif") # write to current dir
+```
+
+## Website density regressions
+
+<br>
+
+$$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$
+
+<br>
+
+::: {style="margin-left: 50px !important; text-align:left !important"}
+$Website\,Density_{i}$:
+
+-   Websites per firm in a Local Authority $i$, *or*
+
+-   Websites in an Output Area $i$
+
+:::
 
@@
 
--   $Website\,Density_{i}$: Websites per firm in a Local Authority $i$
 :::
@@
 
-<!-- ------------------------------------------------------------------------ -->
-
-```{r b_London_oa, eval=F, fig.height=5, fig.width=5, message=FALSE}
-#, out.width = "60%"
-path.image <- paste0(path, "/outputs/b_dist_london_oa.png")
-knitr::include_graphics(path.image)
-```
-
 ------------------------------------------------------------------------
@@
 
-<!-- ------------------------------------------------------------------------ -->
+------------------------------------------------------------------------
 
-```{r b_city_oa, eval = F, fig.height=5, fig.width=5, message=FALSE}
+```{r b_London_oa, eval=T, fig.height=5, fig.width=5, message=FALSE}
 #, out.width = "60%"
-path.image <- paste0(path, "/outputs/b_dist_city_oa.png")
+path.image <- paste0(path, "/outputs/b_dist_london_oa.png")
 knitr::include_graphics(path.image)
@@
 
-<!-- ------------------------------------------------------------------------ -->
+------------------------------------------------------------------------
 
-```{r b_centre_oa, eval = F, message=FALSE, out.height= "60%", out.width="60%"}
-#fig.height=4, fig.width=4, 
-path.image <- paste0(path, "/outputs/b_dist_centre_oa.png")
+```{r b_city_oa, eval = T, fig.height=5, fig.width=5, message=FALSE}
+#, out.width = "60%"
+path.image <- paste0(path, "/outputs/b_dist_city_oa.png")
 knitr::include_graphics(path.image)
@@
 
-```{r r_squared, eval=T, results='asis'}
+```{r b_centre_oa, eval = T, message=FALSE, out.height= "60%", out.width="60%"}
+#fig.height=4, fig.width=4, 
+path.image <- paste0(path, "/outputs/b_dist_centre_oa.png")
+knitr::include_graphics(path.image)
+```
+
+<!-- ------------------------------------------------------------------------ -->
+
+```{r r_squared, eval=F, results='asis'}
 # This is for the RSquared of the different website density regressions.
@@
 
-**Hierarchy effect**: from main centers to secondary ones -- central
+> **Hierarchy effect**: from main centers to secondary ones -- central
 places
@@
 
-<!-- ------------------------------------------------------------------------ -->
+------------------------------------------------------------------------
 
-```{r gini_oa, eval=F, results='asis'}
+```{r gini_oa, eval=T, results='asis'}
 path.in <- paste0(path, "/outputs/gini_oa.csv")
@@
 
-```{r Darlington, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+```{r Darlington, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
 path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000005.png")
@@
 
-------------------------------------------------------------------------
+<!-- ------------------------------------------------------------------------ -->
 
-```{r Nottingham, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+```{r Nottingham, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
 path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000018.png")
@@
 
-------------------------------------------------------------------------
+<!-- ------------------------------------------------------------------------ -->
 
@@
 
-```{r Leicester, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+```{r Leicester, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
 path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000016.png")
@@
 
-```{r Plymouth, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+```{r Plymouth, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
 path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000026.png")
@@
 
-------------------------------------------------------------------------
+<!-- ------------------------------------------------------------------------ -->
 
-```{r Edinburgh, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+```{r Edinburgh, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
 path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_S12000036.png")
@@
 
+# Rank dynamics: stability vs. deversity
+
+------------------------------------------------------------------------
+
+```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only05.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only95.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+to do: the same for OAs
+
+------------------------------------------------------------------------
 # Conclusions

### 2023-03-22 src/level_lisa_oa.Rmd

@@
                                          labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
-                                         title="LISA for level, OA") +
+                                         title="LISA, websites per OA") + 
   tm_legend(text.size = 1) +

### 2023-03-22 src/level_lisa_la.Rmd
error
lisa maps

--- 2023-03-21 src/level_lisa_la.Rmd
@@
                                                      rgb(1,0,0,alpha=0.4)), 
-                                         labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
-                                         title="LISA for level, LA") +
+                                         labels = c("Not significant",
+                                                    "High-High",
+                                                    "Low-Low",
+                                                    "Low-High",
+                                                    "High-Low"), 
+                                         title="LISA, websites per firm in a LA") +
   tm_legend(text.size = 1) +
@@
 
-```{r}
+```{r, eval=F}
 

### 2023-03-22 src/level_lisa_oa.Rmd
lisa maps

--- 2022-06-06 src/level_lisa_oa.Rmd
@@
                                                      rgb(1,0,0,alpha=0.4)), 
-                                         labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
-                                         title="LISA for level, OA") +
+                                         labels = c("Not significant",
+                                                    "High-High",
+                                                    "Low-Low",
+                                                    "Low-High",
+                                                    "High-Low"), 
+                                         title="LISA, websites per OA") + 
   tm_legend(text.size = 1) +
@@
 
-```{r}
+```{r, eval = F}
 
@@
   ggtitle("Regression fit, distance to London on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ggsave(path.out)
@@
   ggtitle("Regression coefficients, distance to London on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(hjust = 0.5))
 ggsave(path.out)
@@
   rename(London = rsquared)
+
+# Help for OA area size. I double checked the largest: E00102963
+# It is correct, this is its size.
+# ggplot(df, aes(x=area)) +
+#     geom_histogram(binwidth=.5, colour="black", fill="white")
+# 
+# df %>% filter(area>50)
 ```
@@
   ggtitle("Regression fit, distance to nearest city on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ggsave(path.out)
@@
   ggtitle("Regression coefficients, distance to nearest city on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ggsave(path.out)
@@
   ggtitle("Regression fit, distance to nearest retail centre on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ggsave(path.out)
@@
   ggtitle("Regression coefficients, distance to nearest retail centre on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ggsave(path.out)

### 2023-03-22 src/level_lisa_la.Rmd
file system

@@
   write_csv(path.out)
-
 ```

### 2023-03-22 src/level_lisa_oa.Rmd
file system

@@
 
-path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_oa") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_oa")  # for multiple PC
+path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", n, "_oa") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_oa")  # for multiple PC
 
@@
 ```{r}
-path.out <- paste0(path, "/outputs/morani_oa.csv")
+path.out <- paste0(path, "/outputs/lisa/morani_oa.csv")
 morani %>% as_tibble() %>% 
@@
                    year = V3) %>% 
-  kable() #%>% 
-  #write_csv(path.out)
-
+  #kable() %>% 
+  write_csv(path.out)
 ```
@@
 
-path.out <- paste0(path, "/outputs/gini_oa.csv")
+path.out <- paste0(path, "/outputs/lisa/gini_oa.csv")
 gini.df %>% write_csv(path.out)
@@
 
-path.out <- paste0(path,"/outputs/rsq_dist_london_oa.png")
+path.out <- paste0(path,"/outputs/lisa/rsq_dist_london_oa.png")
 ggplot(dens.London) +
@@
 
-path.out <- paste0(path,"/outputs/b_dist_london_oa.png")
+path.out <- paste0(path,"/outputs/lisa/b_dist_london_oa.png")
 ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
@@
 
-path.out <- paste0(path,"/outputs/rsq_dist_city_oa.png")
+path.out <- paste0(path,"/outputs/lisa/rsq_dist_city_oa.png")
 ggplot(dens.distMet) +
@@
 
-path.out <- paste0(path,"/outputs/b_dist_city_oa.png")
+path.out <- paste0(path,"/outputs/lisa/b_dist_city_oa.png")
 ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
@@
 
-path.out <- paste0(path,"/outputs/rsq_dist_centre_oa.png")
+path.out <- paste0(path,"/outputs/lisa/rsq_dist_centre_oa.png")
 ggplot(dens.dist.retail) +
@@
 
-path.out <- paste0(path,"/outputs/b_dist_centre_oa.png")
+path.out <- paste0(path,"/outputs/lisa/b_dist_centre_oa.png")
 ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
@@
 ```{r}
-path.out <- paste0(path, "/outputs/r2_all_oa.csv")
+path.out <- paste0(path, "/outputs/lisa/r2_all_oa.csv")
 

### 2023-03-22 src/la_ranks_n1.Rmd
lisa maps titles
ranks

@@
 
-- add 0s in the df and then calculate the ranks
+-
 
@@
 la.names <- la@data %>% 
-  select(LAD21CD, LAD21NM) %>% 
+  dplyr::select(LAD21CD, LAD21NM) %>% 
   rename(ladcd = LAD21CD,
@@
 ```
+
 ## Ranks websites per firm t and 2012 only
@@
 ```{r}
-t <- 1996
+
+# change percentile in out.path and mutate(alpha = ifelse(abs(diff_1st)
+t <- 2000
 
@@
 
-  mutate(alpha = ifelse(abs(diff_1st) < 6.875, .8, 0.1)) %>% 
+  mutate(alpha = ifelse(abs(diff_1st) < 3.75, .8, 0.1)) %>% 
   mutate(alpha = ifelse(is.na(diff_1st), 0.1, alpha)) %>%
@@
   geom_bump(colour = "red") +
-  theme(legend.position="none") +
   scale_y_reverse(labels = c(1, 100, 200, 300, 400)) +
@@
                   nudge_x = 2, nudge_y = 1,
-                  direction = "y", hjust = "left")
-
+                  direction = "y", hjust = "left") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position="none") 
 ggsave(out.path)
@@
 ```{r}
-out.path <- paste0(path, "/outputs/ranks/web_2000_2012_only9.png")
+out.path <- paste0(path, "/outputs/ranks/web_2000_2012_only95.png")
 
@@
 
-  mutate(alpha = ifelse(abs(diff_1st) > 71, .8, 0.1)) %>% 
+  mutate(alpha = ifelse(abs(diff_1st) > 87.75, .8, 0.1)) %>% 
   mutate(alpha = ifelse(is.na(diff_1st), 0.1, alpha)) %>%
@@
                   nudge_x = 2, nudge_y = 1,
-                  direction = "y", hjust = "left")
+                  direction = "y", hjust = "left") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position="none")
 

### 2023-04-06 src/la_ranks_n1.Rmd
various outputs

@@
 
-out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only05.png")
+out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only95.png")
 
@@
 
-  mutate(alpha = ifelse(abs(diff_1st) < 3.75, .8, 0.1)) %>% 
+  mutate(alpha = ifelse(abs(diff_1st) > 199.5, .8, 0.1)) %>% 
   mutate(alpha = ifelse(is.na(diff_1st), 0.1, alpha)) %>%

### 2023-04-06 src/level_lisa_oa.Rmd
various outputs

@@
 library(broom)
+library(plm)
+library(lmtest)
 
@@
 
-Website density regressions to London
+## Panel regressions
+
+```{r}
+# London
+model <- plm(log(n+0.001) ~ log(London+.001)*as.factor(year), 
+             data = df,
+             index = c("oa11cd", "year"),
+             model = "within",
+             effect = "twoways")
+summary(model)
+
+est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
+int <- confint(est)
+
+out.model <- bind_cols(tidy(est), 
+                       as_tibble(int)) %>% 
+  mutate_if(is.numeric, round, 2) %>% 
+  mutate(year = 1997:2012) %>% 
+  rename(conf.low = '2.5 %',
+         conf.high = '97.5 %')
+
+path.out <- paste0(path,"/outputs/lisa/b_panel_dist_london_oa.png")
+ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
+  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
+                    width = 0.2,size  = 1,
+                    position = "dodge", color="turquoise4") +
+  #geom_hline(yintercept = 0, color = "red", size = 1) +
+  geom_point() + coord_flip() +
+  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
+  ggtitle("Regression coefficients: distance to London on website counts, OA") +
+  theme(plot.title = element_text(hjust = 0.5))
+ggsave(path.out)
+
+# Cities
+model <- plm(log(n+0.001) ~ log(dist+.001)*as.factor(year), 
+             data = df,
+             index = c("oa11cd", "year"),
+             model = "within",
+             effect = "twoways")
+summary(model)
+
+est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
+int <- confint(est)
+
+out.model <- bind_cols(tidy(est), 
+                       as_tibble(int)) %>% 
+  mutate_if(is.numeric, round, 2) %>% 
+  mutate(year = 1997:2012) %>% 
+  rename(conf.low = '2.5 %',
+         conf.high = '97.5 %')
+
+path.out <- paste0(path,"/outputs/lisa/b_panel_dist_cities_oa.png")
+ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
+  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
+                    width = 0.2,size  = 1,
+                    position = "dodge", color="turquoise4") +
+  #geom_hline(yintercept = 0, color = "red", size = 1) +
+  geom_point() + coord_flip() +
+  ylab("Interaction coefficienst: distance*year, robust st. errors") + xlab("") + 
+  ggtitle("Regression coefficients: distance to Cities on website counts, OA") +
+  theme(plot.title = element_text(hjust = 0.5))
+ggsave(path.out)
+
+# Retail centres
+model <- plm(log(n+0.001) ~ log(dist.retail+.001)*as.factor(year), 
+             data = df,
+             index = c("oa11cd", "year"),
+             model = "within",
+             effect = "twoways")
+summary(model)
+
+est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
+int <- confint(est)
+
+out.model <- bind_cols(tidy(est), 
+                       as_tibble(int)) %>% 
+  mutate_if(is.numeric, round, 2) %>% 
+  mutate(year = 1997:2012) %>% 
+  rename(conf.low = '2.5 %',
+         conf.high = '97.5 %')
+
+path.out <- paste0(path,"/outputs/lisa/b_panel_dist_retail_oa.png")
+ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
+  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
+                    width = 0.2,size  = 1,
+                    position = "dodge", color="turquoise4") +
+  #geom_hline(yintercept = 0, color = "red", size = 1) +
+  geom_point() + coord_flip() +
+  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
+  ggtitle("Regression coefficients: distance to Retail Centres on website counts, OA") +
+  theme(plot.title = element_text(hjust = 0.5))
+ggsave(path.out)
+```
+
+## Website density regressions to London, cross-section
 
@@
 
-Website density regressions to the nearest city
+## Website density regressions to the nearest city, cross-section
 
@@
 
-Website density regressions to the nearest retail centre
+## Website density regressions to the nearest retail centre, cross-section
 

### 2023-04-06 src/s_la_per_firm.Rmd
various outputs

@@
 library(survival)
+library(rgeos)
+library(spdep)
+library(rgdal)
+library(tidygeocoder)
+library(geosphere)
+library(raster)
+library(rms)
+library(stargazer)
 
@@
 
+## Explain fast/slow
+
+### load data
+
+```{r}
+a.la <- read_csv(path.out)
+
+library(readxl)
+library(httr)
+url.classes <- "https://www.ons.gov.uk/file?uri=/methodology/geography/geographicalproducts/areaclassifications/2011areaclassifications/datasets/clustermembershipv2.xls"
+
+GET(url.classes, write_disk(tf <- tempfile(fileext = ".xls", tmpdir = paste0(path, "/data/temp"))))
+
+classes.la <- read_excel(tf, sheet = 3, range = "A10:I399") %>% 
+  rename(ladcd = Code,
+         supergroup = 'Supergroup Code',
+         group = 'Group Code',
+         subgroup = 'Subgroup Code')
+
+a.la <- a.la %>% left_join(classes.la) %>% 
+  mutate(fast.binary = if_else(fast == "fast", 1, 0))
+sapply(a.la, function(x) sum(is.na(x))) # 16 mising
+
+```
+
+### Distances
+
+First set up the destinations. Origins are all the LA.
+
+Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
+Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)
+
+```{r}
+# LA centroids
+la.c <- gCentroid(la, byid=TRUE)@coords %>% 
+  as.data.frame() %>%
+  bind_cols(la@data) %>% 
+  dplyr::select(LAD21CD, x, y) %>% 
+  remove_rownames()
+
+# cities
+city.names <- c("London, UK", 
+"Birmingham, UK",
+"Glasgow, UK",
+"Liverpool, UK",
+"Bristol, UK",
+"Manchester, UK",
+"Sheffield, UK",
+"Leeds, UK",
+"Edinburgh, UK",
+"Leicester, UK")
+
+cities <- geo(city.names, no_query = F, method = "arcgis")
+
+# retail centres
+geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
+retail <-readOGR(geo.path) 
+#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
+retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
+  as.data.frame() %>% 
+  add_rownames(var = "id") #%>%
+
+retail.major.cetres <- retail.major.cetres.help %>% 
+  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
+  st_transform(4326) %>%
+  st_coordinates() %>%
+  as_tibble() %>% 
+  bind_cols(retail.major.cetres.help$id) %>% 
+  rename(id = '...3') %>% 
+  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
+```
+
+Distance to cities
+
+```{r}
+dist <- distm(cbind(la.c$x, la.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist <- round((dist/1000),2) %>% 
+  as_tibble()
+
+city.names <- city.names %>% stringr::str_remove(", UK")
+names(dist) <- city.names
+
+dist <- dist %>% bind_cols(la.c$LAD21CD) %>% 
+  rename(ladcd = last_col()) %>% 
+  relocate(ladcd)
+
+# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
+# dist$distMet <- apply(dist[,2:11], 1, min)
+
+dist <- transform(dist, dist = do.call(pmin, dist[-1]))
+
+# NOT USED AS IT SCREWS UP THE REGRESSIONS
+# Complete panel, N = 3871903, 17 year * 227759 OA
+# df <- df %>% bind_rows(dist %>% select(oa11cd)) %>% 
+#   mutate(year = ifelse(is.na(year),1000, year),
+#          country = substring(oa11cd, 1, 1)) %>% 
+#   filter(country != "L" & country != "M" & country != "N") %>% 
+#   filter(!is.na(country)) %>% 
+#   complete(year, oa11cd, fill = list(n = 0)) %>% 
+#   filter(year!=1000,
+#          !is.na(oa11cd)) %>% 
+#   select(-country)
+
+# tests
+# test$country <- substring(test$oa11cd, 1, 1)
+# unique(test$country)
+# length(unique(test$oa11cd))
+
+# calculate area
+la$SHAPE_Area <- area(la) #/ 1000000
+
+# Join with distance and area
+a.la <- a.la %>% left_join(dist, by = c("ladcd" = "ladcd")) 
+sapply(a.la, function(x) sum(is.na(x)))
+```
+
+Distance to retail centres
+
+```{r}
+dist.retail <- distm(cbind(la.c$x, la.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
+dist.retail <- round((dist.retail/1000),2) %>% 
+  as_tibble()
+
+# retail.names <- retail.major.cetres$RC_Name
+# names(dist.retail) <- retail.names
+
+dist.retail <- dist.retail %>% bind_cols(la.c$LAD21CD) %>% 
+  rename(LAD21CD = last_col()) %>% 
+  relocate(LAD21CD)
+
+# Minimum distance 
+dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
+dist.retail <- dist.retail %>% dplyr::select(LAD21CD, dist.retail)
+
+# Join with complete panel
+a.la <- a.la %>% left_join(dist.retail, by = c("ladcd" = "LAD21CD"))
+
+sapply(a.la, function(x) sum(is.na(x)))
+```
+
+### Regressions
+
+**TODO** add distance, see what classes mean, try lm
+
+```{r}
+
+model <- glm(fast.binary ~ log(London + 0.001) + log(dist + 0.001) + log(dist.retail + 0.001) + as.factor(subgroup), family=binomial (link='logit'), data = a.la)
+summary(model)
+stargazer(model, type = 'text')
+```
+
+For Nagelkerke R2
+
+```{r}
+model2 <- lrm(fast.binary ~ log(London + 0.001) + log(dist + 0.001) + log(dist.retail + 0.001) + subgroup, data = a.la)
+print(model2)
+```
+
 ## Survival

### 2023-04-06 src/s_oa.Rmd
various outputs

@@
 library(rprojroot)
+library(tidyverse)
+library(broom)
+library(rprojroot)
+library(rgdal)
+library(sf)
+library(modelr)
+library(rgeos)
+library(spdep)
+library(rgdal)
+library(tidygeocoder)
+library(geosphere)
+library(raster)
+library(Matrix)
+library(rms)
+library(stargazer)
+library(doParallel)
+library(foreach)
+library(maptools)
+#library(survival)
 
@@
 #               year==2003)
+```
+
+## Spatial data
+
+```{r}
+# get OA for England and Wales
+path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
+oa.ew <- readOGR(path.geo)
+# source: https://geoportal.statistics.gov.uk/
+
+# spatial transformations
+oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))
+
+# keep in the data slot only the ONS Output Area id, renaming it as 'id'
+oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
+colnames(oa.ew@data) <- c('id', 'area')
+
+# reassign the polygon IDs
+oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))
+
+# check the CRS has changed correctly, and the data slot has shrink to only the ID
+summary(oa.ew)
+
+# # get OA for Scotland
+# path.geo.sc <- paste0(path, "/data/raw/SG_DataZoneBdry_2011")
+# oa.sc <- readOGR(dsn=path.geo.sc, layer = "SG_DataZone_Bdry_2011")
+# # source: https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011
+# 
+# # spatial transformations
+# oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
+# 
+# # Scotland (follows same steps as EW, see notes above)
+# oa.sc <- oa.sc[, 'DataZone']
+# colnames(oa.sc@data) <- c('id')
+# 
+# # reassign the polygon IDs
+# oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
+# 
+# # check the CRS has changed correctly, and the data slot has shrink to only the ID
+# summary(oa.sc)
+
+path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
+oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
+# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries
+
+# spatial transformations
+oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
+
+# Scotland (follows same steps as EW, see notes above)
+oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
+colnames(oa.sc@data) <- c('id', 'area')
+
+# reassign the polygon IDs
+oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
+
+# check the CRS has changed correctly, and the data slot has shrink to only the ID
+summary(oa.sc)
+
+# build oa for GB
+oa.gb <- spRbind(oa.ew, oa.sc)
+rm(oa.ew, oa.sc)
+
+# oa.gb$geometry <- oa.gb$geometry %>%
+#   s2::s2_rebuild() %>%
+#   sf::st_as_sfc()
+
+# get GB
+path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
+gb <- readOGR(path.gb)
+# spatial transformations
+gb <- spTransform(gb, CRS("+init=epsg:4326"))
 ```
@@
 ```{r}
-# a <- as_tibble()
-# all.oa <- unique(df$oa11cd)
+a <- as_tibble()
+all.oa <- unique(df$oa11cd)
 # #for (oa in all.oa){
@@
 
+# test for a single OA
+help <- df %>% dplyr::filter(oa11cd == "E00000706")
+model <- tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = help))
+model %>% as.data.frame() %>% 
+  filter(term=="xmid") %>% 
+    #bind_cols(all.oa) %>% 
+    mutate(estimate = round(estimate, 0)) %>% 
+    filter(p.value < 0.01) %>%    # (227759 - 212782)/227759  # 6% not sig 
+    rename(id = ...6) %>% 
+    dplyr::select(id, estimate)
+
 # Setup cluster
@@
 
-a <- foreach(oa = all.oa, .combine = 'rbind', .packages = c("tidyverse",
+a <- foreach(oa = all.oa, .combine = 'rbind', .packages = c("tidyverse", #
                            "broom")) %dopar% 
@@
 
-    c(oa, model)
+    #tryCatch(c(oa, model), error=function(e) NULL) #
+    #tryCatch(model$oa <-  oa, error=function(e) NULL) #
+    tryCatch(model %>% mutate(id = oa), error=function(e) NULL)
     }
@@
 a <- a %>% filter(term=="xmid") %>% 
-    bind_cols(all.oa) %>% 
-    mutate(estimate = round(estimate, 0)) %>% 
+    #mutate(estimate = round(estimate, 2)) %>% 
     filter(p.value < 0.01) %>%    # (227759 - 212782)/227759  # 6% not sig 
-    rename(id = ...6) %>% 
-    select(id, estimate)
+    dplyr::select(id, estimate)
 ```
@@
 write_csv(a, path.out)
+```
+
+## Explain fast/slow
+
+### load data
+
+Source: https://www.ons.gov.uk/methodology/geography/geographicalproducts/areaclassifications/2011areaclassifications/datasets
+
+```{r}
+a <- read_csv(path.out)
+
+library(readxl)
+library(httr)
+url.classes <- "https://www.ons.gov.uk/file?uri=/methodology/geography/geographicalproducts/areaclassifications/2011areaclassifications/datasets/2011oacclustersandnamesexcelv2.zips"
+
+path.in <- paste0(path, "/data/raw/2011 OAC Clusters and Names Excel v2.xlsx")
+classes.oa <- read_excel(path.in, sheet = 2) %>% 
+  rename(id = 'Output Area Code',
+         supergroup = 'Supergroup Code',
+         group = 'Group Code',
+         subgroup = 'Subgroup Code')
+
+a <- a %>% left_join(classes.oa) %>% 
+  mutate(fast.binary = if_else(estimate < 2002.926, 1, 0))
+sapply(a, function(x) sum(is.na(x))) # 16 mising
+```
+
+### Distances
+
+First set up the destinations. Origins are all the OA.
+
+Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
+Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)
+
+```{r}
+# oa centroids
+# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
+# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
+#   mutate(id = row.names())
+oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
+  as.data.frame() %>% 
+  add_rownames(var = "id")
+
+# cities
+city.names <- c("London, UK", 
+"Birmingham, UK",
+"Glasgow, UK",
+"Liverpool, UK",
+"Bristol, UK",
+"Manchester, UK",
+"Sheffield, UK",
+"Leeds, UK",
+"Edinburgh, UK",
+"Leicester, UK")
+
+cities <- geo(city.names, no_query = F, method = "arcgis")
+
+# retail centres
+geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
+retail <-readOGR(geo.path) 
+#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
+retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
+  as.data.frame() %>% 
+  add_rownames(var = "id") #%>%
+
+retail.major.cetres <- retail.major.cetres.help %>% 
+  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
+  st_transform(4326) %>%
+  st_coordinates() %>%
+  as_tibble() %>% 
+  bind_cols(retail.major.cetres.help$id) %>% 
+  rename(id = '...3') %>% 
+  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
+```
+
+Distance to cities
+
+```{r}
+dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist <- round((dist/1000),2) %>% 
+  as_tibble()
+
+city.names <- city.names %>% stringr::str_remove(", UK")
+names(dist) <- city.names
+
+dist <- dist %>% bind_cols(oa.gb.c$id) %>% 
+  rename(oa11cd = last_col()) %>% 
+  relocate(oa11cd)
+
+# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
+# dist$distMet <- apply(dist[,2:11], 1, min)
+
+dist <- transform(dist, dist = do.call(pmin, dist[-1]))
+
+# NOT USED AS IT SCREWS UP THE REGRESSIONS
+# Complete panel, N = 3871903, 17 year * 227759 OA
+# df <- df %>% bind_rows(dist %>% dplyr::select(oa11cd)) %>%
+#   mutate(year = ifelse(is.na(year),1000, year),
+#          country = substring(oa11cd, 1, 1)) %>%
+#   filter(country != "L" & country != "M" & country != "N") %>%
+#   filter(!is.na(country)) %>%
+#   complete(year, oa11cd, fill = list(n = 0)) %>%
+#   filter(year!=1000,
+#          !is.na(oa11cd)) %>%
+#   dplyr::select(-country)
+
+# tests
+# test$country <- substring(test$oa11cd, 1, 1)
+# unique(test$country)
+# length(unique(test$oa11cd))
+
+# calculate area
+oa.gb$area <- area(oa.gb) #/ 1000000
+
+# Join with distance and area
+a <- a %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
+  left_join(oa.gb@data, by = c("id" = "id")) %>% 
+  relocate(area, .after = n) %>% 
+  mutate(area = area / 1000000,
+         density = n / area) %>% 
+  relocate(density, .after = area)
+sapply(a, function(x) sum(is.na(x))) # mising NI
+
+```
+
+Distance to retail centres
+
+```{r}
+dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
+dist.retail <- round((dist.retail/1000),2) %>% 
+  as_tibble()
+
+# retail.names <- retail.major.cetres$RC_Name
+# names(dist.retail) <- retail.names
+
+dist.retail <- dist.retail %>% bind_cols(oa.gb.c$id) %>% 
+  rename(oa11cd = last_col()) %>% 
+  relocate(oa11cd)
+
+# Minimum distance 
+dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
+dist.retail <- dist.retail %>% dplyr::select(oa11cd, dist.retail)
+
+# Join with complete panel
+a <- a %>% left_join(dist.retail, by = c("id" = "oa11cd"))
+
+sapply(a, function(x) sum(is.na(x)))
+```
+
+### Regressions
+
+```{r}
+model <- glm(fast.binary ~ log(London + 0.001) + log(dist + 0.001) + log(dist.retail + 0.001) + as.factor(subgroup), family=binomial (link='logit'), data = a)
+summary(model)
+stargazer(model, type = 'text')
+```
+
+For Nagelkerke R2
+
+```{r}
+model2 <- lrm(fast.binary ~ log(London + 0.001) + log(dist + 0.001) + log(dist.retail + 0.001) + group, data = a)
+print(model2)
 ```

### 2023-04-06 src/wpz.Rmd
various outputs

--- 2023-03-22 src/level_lisa_la.Rmd
@@
 ---
-title: "diffusion, websites with 1 pc"
+title: "wpz"
 date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 
+Runs all the analysis for the Workplace Zones.
+
+WPZ sources:
+
+- [info](https://www.ons.gov.uk/methodology/geography/geographicalproducts/areaclassifications/2011workplacebasedareaclassification)
+
+- [lookup OA to WPZ](https://geoportal.statistics.gov.uk/datasets/ons::output-area-to-workplace-zone-to-local-authority-district-december-2011-lookup-in-england-and-wales-1/explore)
+
+- [Ireland](https://www.nisra.gov.uk/support/geography/northern-ireland-workplace-zones)
+
+- [Scotland](https://www.data.gov.uk/dataset/02718f33-a1e7-4447-8e04-d305a950c19e/census-2011-workplace-zones)
+
 ```{r setup, include=FALSE}
@@
 library(tidyverse)
+library(broom)
 library(rprojroot)
 library(rgdal)
-library(rgeos)
-library(spdep)
-library(tmap)
-library(maptools)
-library(knitr)
-library(REAT)
-library(tidygeocoder)
-library(geosphere)
-library(broom)
-library(foreach)
-library(raster)
-
+library(sf)
+library(modelr)
+library(survival)
 options(scipen=10000)
@@
 lookup <- read_csv(path.lookup) %>% 
-  dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
+  dplyr::select(pcds, oa11cd) %>% #, lsoa11cd, msoa11cd, ladcd, ladnm
   dplyr::rename(pc = pcds)
-#glimpse(lookup)
-# The problems refer to Welsh LAD names. Not a problem for the analysis.
-#sapply(lookup, function(x) sum(is.na(x)))
-# 10332 missing msoa11cd
+
+path.lookup <- paste0(path,"/data/raw/Output_Area_to_Workplace_Zone_to_Local_Authority_District_(December_2011)_Lookup_in_England_and_Wales.csv")
+lookup2 <- read_csv(path.lookup) %>% 
+  dplyr::select(OA11CD, WZ11CD) %>% 
+  rename(oa11cd = OA11CD)
+
+lookup <- lookup %>% left_join(lookup2) %>% 
+  filter(!is.na(WZ11CD)) #%>% 
+  #mutate(help = substring(oa11cd, 1, 1)) %>% 
+  #filter(is.na(help))
+  #distinct(WZ11CD)
+
+rm(lookup2)
+
+# 53,578 unique WZ. OK according to the spatial object, but should be 60,709 according the the WPZ description
 ```
@@
   left_join(lookup, by = "pc", suffix =c("","")) %>% 
-  group_by(year, ladcd) %>%
+  group_by(year, WZ11CD) %>%
   #summarise(n = n()) %>%            # for multiple pc  
@@
 
-df <- df %>% filter(!is.na(ladcd)) %>% 
-  complete(ladcd, year) %>% 
+df <- df %>% filter(!is.na(WZ11CD)) %>% 
+  complete(WZ11CD, year) %>% 
   replace(is.na(.), 0)
@@
 ```{r}
-firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>%
-  dplyr::select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
-  rename(ladcd = GEOGRAPHY_CODE,
-         firms = OBS_VALUE)
+# from business census
+path.firms <- paste0(path, "/data/raw/business_census2012.csv")
+firms <- read_csv(path.firms) %>% #distinct(companystatus) 
+  filter(!is.na(postcode),
+         companystatus == "Active") %>% 
+  dplyr::select(postcode) %>% 
+  rename(pc = postcode) %>% 
+  left_join(lookup) %>% 
+  group_by(WZ11CD) %>% 
+  summarise(firms = n())
+
+# firms %>% filter(is.na(oa11cd)) %>% distinct(pc)
+# 200k rows and 73k unique postcodes are missing
 
@@
 # sapply(df %>% left_join(firms), function(x) sum(is.na(x)))
+
 ```
@@
 # get LA for the UK
-path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-la <- readOGR(path.geo)
+path.geo <- paste0(path, "/data/raw/Workplace_Zones_Dec_2011_GCB_in_England_and_Wales_2022_4617806297229452276.geojson")
+wpz <- readOGR(path.geo)
 # source: https://geoportal.statistics.gov.uk/
@@
 # spatial transformations
-la <- spTransform(la, CRS("+init=epsg:4326"))
+wpz <- spTransform(la, CRS("+init=epsg:4326"))
 
-la.f <- fortify(la, region = "LAD21CD")
+wpz.f <- fortify(wpz, region = "LAD21CD")
 
@@
 tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
+
+# get GB
+path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
+gb <- readOGR(path.gb)
+# spatial transformations
+gb <- spTransform(gb, CRS("+init=epsg:4326"))
+
 ```
@@
 
-path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", n, "_la") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_la")  # for multiple PC
+path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", n, "_wpz") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_wpz")  # for multiple PC
 
@@
 sf_use_s2(FALSE)
-nb <- poly2nb(la)
+nb <- poly2nb(wpz)
 listw <- nb2listw(nb, style = "W", zero.policy = T)
@@
 for (t in 1996:2012){
-test <- la
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
+test <- wpz
+test@data <- test@data %>% left_join(df[df$year==t,], by = c("wz11cd"="WZ11CD")) %>% 
   mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
@@
 
-mun_merge_new<- merge(test, lmoran, by="LAD21CD")
+mun_merge_new<- merge(test, lmoran, by="wz11cd")
 
@@
                                                     "High-Low"), 
-                                         title="LISA, websites per firm in a LA") +
+                                         title="LISA, websites per firm in a workplace zone") +
   tm_legend(text.size = 1) +
@@
   tm_borders(lwd = 0) +
-  tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
+  tm_shape(gb, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
   tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
@@
 
-path.out <- paste0(path, "/outputs/lisa/lisa_level_pc_test", n, "_la") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_la")  # for multiple PC
+path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", n, "_wpz") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_wpz")  # for multiple PC
 
@@
                                      "broom")) %dopar% {
-test <- la
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
+test <- wpz
+test@data <- test@data %>% left_join(df[df$year==t,], by = c("wz11cd"="WZ11CD")) %>% 
   mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
@@
   tm_borders(lwd = 0) +
-  tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
+  tm_shape(wpz, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
   tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
@@
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/morani_la.csv")
+path.out <- paste0(path, "/outputs/lisa/morani_wpz.csv")
 morani %>% as_tibble() %>% 
@@
 # LA centroids
-la.c <- gCentroid(la, byid=TRUE)@coords %>% 
+wpz.c <- gCentroid(wpz, byid=TRUE)@coords %>% 
   as.data.frame() %>%
-  bind_cols(la@data) %>% 
-  dplyr::select(LAD21CD, x, y) %>% 
+  bind_cols(wpz@data) %>% 
+  dplyr::select(wz11cd, x, y) %>% 
   remove_rownames()
@@
 ```{r}
-dist <- distm(cbind(la.c$x, la.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist <- distm(cbind(wpz.c$x, wpz.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
 dist <- round((dist/1000),2) %>% 
@@
 
-dist <- dist %>% bind_cols(la.c$LAD21CD) %>% 
-  rename(ladcd = last_col()) %>% 
-  relocate(ladcd)
+dist <- dist %>% bind_cols(wpz.c$wz11cd) %>% 
+  rename(wz11cd = last_col()) %>% 
+  relocate(wz11cd)
 
@@
 # calculate area
-la$SHAPE_Area <- area(la) #/ 1000000
+wpz$SHAPE_Area <- area(wpz) #/ 1000000
 
 # Join with distance and area
-df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
-  left_join(la@data, by = c("ladcd" = "LAD21CD")) %>% 
+df <- df %>% left_join(dist, by = c("WZ11CD" = "wz11cd")) %>% 
+  left_join(wpz@data, by = c("WZ11CD" = "wz11cd")) %>% 
   rename(area = SHAPE_Area) %>% 
@@
 ```{r}
-dist.retail <- distm(cbind(la.c$x, la.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
+dist.retail <- distm(cbind(wpz.c$x, wpz.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
 dist.retail <- round((dist.retail/1000),2) %>% 
@@
 
-dist.retail <- dist.retail %>% bind_cols(la.c$LAD21CD) %>% 
-  rename(LAD21CD = last_col()) %>% 
-  relocate(LAD21CD)
+dist.retail <- dist.retail %>% bind_cols(wpz.c$wz11cd) %>% 
+  rename(wz11cd = last_col()) %>% 
+  relocate(wz11cd)
 
@@
 dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
-dist.retail <- dist.retail %>% dplyr::select(LAD21CD, dist.retail)
+dist.retail <- dist.retail %>% dplyr::select(wz11cd, dist.retail)
 
 # Join with complete panel
-df <- df %>% left_join(dist.retail, by = c("ladcd" = "LAD21CD"))
+df <- df %>% left_join(dist.retail, by = c("WZ11CD" = "wz11cd"))
 
 sapply(df, function(x) sum(is.na(x)))
+```
+
+Classification
+
+```{r}
+
+library(readxl)
+library(httr)
+url.classes <- "https://www.ons.gov.uk/file?uri=/methodology/geography/geographicalproducts/areaclassifications/2011workplacebasedareaclassification/classificationofworkplacezonesfortheukdatasets/wzassignmentsfinal.xlsx"
+
+GET(url.classes, write_disk(tf <- tempfile(fileext = ".xlsx", tmpdir = paste0(path, "/data/temp"))))
+
+classes.wpz <- read_excel(tf, sheet = 1) %>% 
+  rename(WZ11CD= 'Workplace Zone Code',
+         supergroup = 'Supergroup Code',
+         group = 'Group Code')
+
+df <- df %>% left_join(classes.wpz)
 ```
@@
 for (t in 1996:2012){
-  model <- lm(log(n+0.001) ~ log(London), data = df[df$year==t,]) #density
+  model <- lm(log(n+0.001) ~ log(London) + as.factor(group), data = df[df$year==t,]) #density
   r2 <- as.data.frame(summary(model)$r.squared)
@@
 
-path.out <- paste0(path,"/outputs/rsq_dist_london_la.png")
+path.out <- paste0(path,"/outputs/rsq_dist_london_wpz.png")
 ggplot(dens.London) +
@@
   ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to London on website density, LA") +
+  ggtitle("Regression fit, distance to London on website density, Workplace Zone") +
   theme(plot.title = element_text(hjust = 0.5))
@@
 
-path.out <- paste0(path,"/outputs/b_dist_london_la.png")
+path.out <- paste0(path,"/outputs/b_dist_london_wpz.png")
 ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
@@
   ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to London on website density, LA") +
+  ggtitle("Regression coefficients, distance to London on website density, Workplace Zone") +
   theme(plot.title = element_text(hjust = 0.5)) 
@@
 
-path.out <- paste0(path,"/outputs/rsq_dist_city_la.png")
+path.out <- paste0(path,"/outputs/rsq_dist_city_wpz.png")
 ggplot(dens.distMet) +
@@
   ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest city on website density, LA") +
+  ggtitle("Regression fit, distance to nearest city on website density, Workplace Zone") +
   theme(plot.title = element_text(hjust = 0.5)) 
@@
 
-path.out <- paste0(path,"/outputs/b_dist_city_la.png")
+path.out <- paste0(path,"/outputs/b_dist_city_wpz.png")
 ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
@@
   ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density, LA") +
+  ggtitle("Regression coefficients, distance to nearest city on website density, Workplace Zone") +
   theme(plot.title = element_text(hjust = 0.5)) 
@@
 
-path.out <- paste0(path,"/outputs/rsq_dist_centre_la.png")
+path.out <- paste0(path,"/outputs/rsq_dist_centre_wpz.png")
 ggplot(dens.dist.retail) +
@@
   ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest retail centre on website density, LA") +
+  ggtitle("Regression fit, distance to nearest retail centre on website density, Workplace Zone") +
   theme(plot.title = element_text(hjust = 0.5)) 
@@
 
-path.out <- paste0(path,"/outputs/b_dist_centre_la.png")
+path.out <- paste0(path,"/outputs/b_dist_centre_wpz.png")
 ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
@@
   ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest retail centre on website density, LA") +
+  ggtitle("Regression coefficients, distance to nearest retail centre on website density, Workplace Zone") +
   theme(plot.title = element_text(hjust = 0.5)) 
@@
 ```{r}
-path.out <- paste0(path, "/outputs/r2_all_la.csv")
+path.out <- paste0(path, "/outputs/r2_all_wpz.csv")
 
 r2.all <- dens.London %>% 
-  rename(London.LA = rsquared) %>% 
+  rename(London.wpz = rsquared) %>% 
   left_join(dens.distMet) %>% 
-  rename(cities.LA = rsquared) %>% 
+  rename(cities.wpz = rsquared) %>% 
   left_join(dens.dist.retail) %>% 
-  rename(retail.LA = rsquared) %>% 
-  relocate(year, .before = London.LA) %>% 
+  rename(retail.wpz = rsquared) %>% 
+  relocate(year, .before = London.wpz) %>% 
   arrange(year)

### 2023-10-27 paper/submission1/tranos2023.qmd
presentation

@@
 
+<!-- geographers lack of interest + motivation -->
 Geographers were always interested in how new technologies and
@@
 
+<!-- contribution -->
 This paper offers such a contribution: a geographical study illustrating 
@@
 domain and adequate research methods can offer new contributions 
-regarding our understanding of technologies are diffused.
+regarding our understanding of how new technologies diffuse.
 
+<!-- motivation -->
 The motivation for this paper lies in the fact that there are various
 stakeholders who are interested in knowing how new digital technologies
-diffused over space and time and use this information to make predictions
+diffused over space and time and use this knowledge to make predictions
 regarding the diffusion of related *future* technologies.
@@
 the diffusion of a new technology might not be optimal at a local level,
-it might be beneficial from a global perspective as it might enable 
-faster diffusion to less advantaged places @leibowicz2016representing. 
+it might be beneficial from a global perspective as it could lead to 
+faster diffusion to less advantaged places [@leibowicz2016representing]. 
 Despite the spatial heterogeneity of such diffusion mechanisms and 
@@
 <!-- diffusion and adoption -->
-
 Technological diffusion, which is by definition an aggregated process, 
@@
 of their attitude towards new technologies and innovations. On the other hand,
-@perkins2011internet attribute diffusion to two processes: (i) epidemic-type 
-of mechanisms, which
-
-Two main mechanisms are identified in the
-literature to explain diffusion: (1) epidemic-type dynamics whereby
-contact with previous adopters stimulates uptake as po tential adopters
-learn about a new innovation; and (2) economic-type mechanisms whereby
-potential users adopt a new innovation as it be comesmore profitable,
-useful, or valuable, with uptake characteristically spreading as costs
-be come lower, performance improves, or the po tential uses of the
-innovation grow over time.
-The criticisms are mainly based on the fact that although the approach
-gives an idea of aggregate (industry or household) behaviour, it does
-not focus on the individual's (firm or household) adoption process
-
-@morrill2020spatial understands spatial diffusion as the landscape 
-changes that spread across space because of changes that take place 
-elsewhere.
+@perkins2011internet attribute diffusion to two processes: (i) epidemic-like 
+mechanisms, which are governed by distance, proximity and social interactions, 
+and (ii) by economic mechanisms as new innovations are adopted by users as they
+become more profitable, valueable and useful.
+<!-- The criticisms are mainly based on the fact that although the approach -->
+<!-- gives an idea of aggregate (industry or household) behaviour, it does -->
+<!-- not focus on the individual's (firm or household) adoption process -->
 
 <!-- data and early engagement with the web -->
+This paper focuses on the diffusion of the web as new technology during the $1996-2012$
+period. This was an exciting period for digital technologies as it corresponds with
+the commercialisation of the internet and, consequently, its almost universal 
+adoption. The reader is reminded that it was only in 1994 when Netscape Navigator 
+was introduced, a year before Microsoft's Internet Explorer
+^[[https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft](https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft)].
+Also, only $9$ per cent of UK's households had access to the internet in $1998$ 
+[@ons2018], the web included mostly static webpages, there were no social media 
+and web browsing was happening exclusively from desktop PCs as there were no 
+smartphones [@tranosuk]. 
+Hence, it is fair to say that the study period captures the very early stages of 
+the diffusion on a new technology that is the web. 
+This is key point in the lifecycle of a new technology. Firstly, during this period
+new technologies are expensive, crude and imperfect [@rosenberg1994exploring; @wilson201281].
+A simple comparison between Web 1.0 and Web 3.0 applications clearly illustrates this
+[@tranos2020social]. During this period the performance of a new technology
+is the main attraction and not the cost to access and use it [@wilson2011lessons]
+There is a broader theoretical discussion in the literature regarding the motivation
+behind early adoption. As summarised by @perkins2005international, on the one hand,
+epidemic models highlight the role of interpersonal contacts as a way for new 
+technologies to diffuse. On the other hand, economic models underline the importance
+of heterogeneity. Different firms have different structures and business plans,
+which define the potential economic returns of the adoption of a new technology 
+and, therefore, the choice to adopt a new technology becomes an individual option.
+From a broader and evolutionary perspective, initial conditions are essential for 
+the creation and evolution of path-dependent technological development trajectories
+[@neffke2011regions; @simmie2014new]. This argument is even more relevant when the
+focus is on Information and Communication Technologies (ICT) because of the commonly
+found lag between investment and economic returns as reflected in the Solow paradox 
+[@acemoglu2014return; @brynjolfsson2018artificial].
 
-wilson201281 Early on in their lifecycle, new technologies are crude,
-imperfect, and expensive (Rosenberg, 1994). New energy technologies are
-attractive for their ability to perform a particular task or deliver a
-new or improved energy service (Fouquet, 2010). This is often
-circumscribed by a particular set of needs in a particular context: a
-market 'niche'. End-users in niche markets are generally less sensitive
-to the effective price of the energy service provided or have a higher
-willingness to pay for its performance advantages (Fouquet, 2010). Thus
-initially, performance dominates cost competitiveness (Wilson and
-Grubler, 2011). Market niches afford some protection from competitive
-pressures, allowing technologies to be tested and improved in applied
-settings, reducing uncertainties with performance or market demand (Kemp
-et al., 1998). Costs may only fall substantively after an extended
-period of commercial experimentation, concurrent with the establishment
-of an industrial base and characteristic moves towards standardisation
-and mass production (Grubler, 1998). The influence of accumulating
-production experience on costs is captured by the concept of learning.
+Importantly, the data used here depicts the *active* engagement with the web as 
+it contains all the commercial websites archived by the Internet Archive
+^[[https://archive.org/](See https://archive.org/).]
+that include a mention to at least one valid UK postcode in the web text. 
+Websites are identified as commercial ones if they are part of the .co.uk second 
+level domain (SLD). Also, the act of of creating a website is understood here as 
+active engagement with the web vis-à-vis the more passive act of browsing the web 
+or having an internet connection [@tranosuk]. Previous studies have focused mostly
+on more passive notions of engaging with digital technologies such as internet 
+adoption and internet speeds [e.g. @blank2018local; @riddlesden2014broadband].
 
-@perkins2005international Theoretical models disagree as to why firms
-adopt innovations at different times. Epidemic models emphasize
-information (Griliches 1957). Certain firms are hypothesized to adopt
-earlier because they come into contact with, and learn from, adopters of
-the new technology before others. Economic models, on the other hand,
-predominantly emphasize firm heterogeneity (Ireland and Stoneman 1986).
-Firms adopt technologies at different times because they differ with
-respect to various organizational and environmental variables
-influencing the economic returns from adoption (Blackman 1999).
+<!-- description of the methods and results -->
+
+<!-- contents -->
 
 <!-- methods -->
-
 @grubler1990rise Later Hagerstrand conceptualized physical "barrier"
@@
 
-<!-- methods -->
-
 @wilson201281 Logistic growth describes an initial period of gradual
@@
 
-MAYBE SKIP THIS??
+<!-- key concepts -- copied from help.Rmd -->
+Geographical diffusion is a synthesis of different processes. On the one hand, we 
+can identify purely spatial or, in other words, contagious processes. Adjacency 
+and, more broadly, distance are the key drivers of diffusion. 
+This perspective draws similarities with epidemics: innovation just like pathogens 
+spreads because of contagion and, consequently, proximity and exposure 
+[@hivner2003facilitating]. On the other hand, we can identify hierarchical 
+processes. Instead of horizontal distance-based diffusion mechanisms, the top-down 
+hierarchy of urban systems drives diffusion. In reality, the synthesis of these 
+two processes represents how new technologies diffuse over space and time [@morrill2020spatial].
+
+These ideas were firstly introduced by Torsten Hägerstrand and his thesis entitled 
+'Innovation Diffusion as a Spatial Process' [@hagerstrand1968innovation]. 
+Hägerstrand was the first one to identify diffusion as a geographical process. 
+The starting point was the idea that diffusion is based on passing information 
+through social networks, which themselves tend to be defined by geography. Hence, 
+he identified the 'neighbourhood' effect of how information, and consequently, 
+innovation diffuse. He used agricultural innovations to test and model his ideas 
+using Monte Carlo simulations. Hägerstrand also incorporated the role of hierarchy 
+and how some phenomena maybe firstly adopted in larger cities and then diffuse 
+to second tier ones. This is a sequential instead of a simultaneous process, which 
+resembles the 'lead-lag' spatial acceleration effect in market research 
+[@bento2018time; @PERES201091]. Hägerstrand maybe more widely known for highlighting 
+the role time plays in the diffusion of innovations: an early-pioneering 
+period, a middle fast accelerating period and a final saturation one [@morrill2020spatial.
+
+The temporal dimension was further explored by Everett Rogers and his seminal work 
+on 'Diffusion of Innovations' [@rogers2010diffusion]. Rogers being a sociologist 
+his work focused not on the diffusion of innovations over space and time, but 
+instead on the adoption of new technologies and innovations by individuals and the 
+individual mechanisms that drive the decisions behind adoption. He identified five 
+groups of individuals regarding their adoption speed: innovators, early adopters, 
+early majority, late majority and laggards. Key mechanism of diffusion and adoption 
+is always communication and how knowledge is transferred within a social system. 
+Therefore, all approaches agree on the S-shaped diffusion and cumulative adoption 
+pattern [@grubler1990rise].
+
+Schmidt's Law empirically illustrates a similar pattern. *Core* and usually highly 
+agglomerated regions is where new technologies are invented and commercially deployed 
+[@grubler1990rise; grubler1990rise]. This is where the first adopters tend to be 
+based. Then, technologies spread to the *rim* and eventually to the *periphery*. 
+Although adoption pace might be higher when new technologies finally arrive to the 
+periphery, the saturation levels there may never reach the ones in the core because 
+of the lack of infrastructure or other necessary institutions [@leibowicz2016representing].
+@grubler1990rise effectively summarises the three key characteristics of the spatial 
+diffusion process: (i) the cumulative level of adoption follows an S-shaped pattern 
+just like purely temporal models; (ii) diffusion is shaped by a hierarchy effect 
+in a form of a centrifugal force: from core to periphery; and (iii) diffusion is 
+also shaped by distance and a neighbourhood effect and contaminate nearby locations.
 
 <!-- previous studies -->
+The remaining of this section reviews empirical studies which analysed the diffusion 
+on new technologies over space and time. Although the spatial dimension is present
+in most of the following studies, the level of spatial detail is almost always 
+more coarse than the one adopted here. To begin with @beardsell1999spatial studied 
+the evolution of the computer industry in $317$ US metro areas during the $1977-1992$ 
+period using employment data. Their analysis indicated that the relative size distribution
+holds for urban computer employment and also urban heterogeneity is essential in
+explaining this distribution. @bednarz2020pulled focused on wind turbines and modelled
+their spatial diffusion across $402$ German regions during $1970-2015$. Their key
+finding is that local demand than local supply was the main driving factor.
 
-@beardsell1999spatial To characterize evolution of the computer
-industry, we examine the distribution of relative employment across
-cities in 1977 and how that distribution changes over time to 1992
+At a global scale @perkins2005international analyzed if the diffusion rate of new
+technologies is driven by a latecomer advantage and the engagement with the global
+economy via foreign direct investments and trade. Their results illustrate that 
+indeed latecomers and developing countries experience diffusion of new technologies 
+more rapidly than early adopter and developed countries. At the same scale, 
+@perkins2011internet explored whether the adoption of previous communication
+technologies that is mail, telegrams and telephones was shaped by similar socioeconomic 
+factors as the internet. Their results indicated common patterns regarding the 
+drivers behind the adoption of different communication technologies.
 
-There is no tendency of relative size distributions of urban computer
-employment to collapse, go bimodal, or fully spread. Overall computers
-exhibit some turbulence, with dramatic big winners and losers among
-cities. In attracting or repelling an industry, urban heterogeneity is
-important. Large, well educated cities near San Jose have much greater
-chances of attracting high-tech employment and less of losing it
+Turning to studies that share more technological and scalar similarities with this 
+paper, @ding2010modeling modelled the spatial diffusion of mobile telecommunications 
+across regions in China. Their analysis indicated that socioeconomic characteristics 
+are important determinants of the timing, speed and the level of mobile diffusion
+within China. Using data from a Hungarian online social network, @lengyel2020role 
+analysed its adoption and the churn at a very granular spatial level. Their results 
+are in agreement with early theoretical and empirical contributions reviewed here:
+assortativity, urban scaling and distance are the key drivers of spatial diffusion.
+At a global scale @PAPAGIANNIDIS2015308 modelled the diffusion of different web
+technologies technologies and practices. Interestingly, they did so by using similar,
+but less extensive data as the one used here. Their analysis illustrated how the 
+diffusion of different web technologies and practices follow an S-shaped pattern
+as well as the different diffusion rates of the different technologies and practices.
 
-@bednarz2020pulled spatial innovation diffusion By using Bayesian
-survival models with time-dependent data of wind turbine deployment and
-firm foundation for 402 German regions between the years 1970 and 2015,
-we show that the spatial evolution of the German wind energy industry
-was more strongly influenced by local demand-- pull than local
-supply--push processes.
+<!-- @fritsch2015new analyze the spatial diffusion of laser technology -->
+<!-- research in West Germany from 1960, when this technology began, until -->
+<!-- 2005. -->
+<!-- in one region  -->
 
-The industry's initial locations are distributed relatively arbitrarily
-and unpredictably, as their needs in terms of resources and skills are
-diverse and distinct from the older existing industries (Boschma and
-Lambooy, 1999). Consequently, emerging industries are characterized by
-relatively high degrees of freedom in terms of location. In later
-extensions of the concept, the assumption of the randomness of locations
-was revised with greater importance assigned to regional conditions
-(Boschma and Lambooy, 1999; Fornahl et al., 2012).
+<!-- @leibowicz2016representing for energy There is no generally accepted -->
+<!-- theory that explains diffusion rate heterogeneity across technologies, -->
+<!-- but several factors are considered important. Greater unit scale and -->
+<!-- larger market size contribute to slower diffusion. Requirements for -->
+<!-- interrelated technologies or complex infrastructures also hinder the -->
+<!-- diffusion process (Grubler, 2012). -->
 
-@perkins2005international analyze whether the rate at which new producer
-technologies diffuse is significantly influenced by (1) latecomer
-advantage and (2) engagement with the global economy via trade and
-foreign investment
+<!-- Mobile phones benefited from early deployment in recreational boats and -->
+<!-- automobiles, where the traditional competitor was not a viable option. -->
+<!-- In the early stages of diffusion, performance is a more important driver -->
+<!-- of adoption than cost competitiveness. Typically, significant cost -->
+<!-- reductions only occur once the technology reaches a deployment level -->
+<!-- capable of supporting standardization and mass production (Wilson, -->
+<!-- 2012). -->
 
-Indeed, precisely because of these latecomer advantages, developing
-countries1 are believed to be well placed to catch up with developed
-ones (Gerschenkron 1962; Abramovitz 1986).
+<!-- @leibowicz2016representing Empirical evidence supports the validity of -->
+<!-- Schmidt's Law over a wide range of technologies, time periods, and -->
+<!-- geographical contexts. A recent meta-analysis of technology up-scaling -->
+<!-- found that diffusion accelerated moving from the core to the rim and -->
+<!-- periphery for technologies as diverse as natural gas power, oil -->
+<!-- refineries, and automobiles (Wilson, 2009). One historical example that -->
+<!-- conforms particularly well to Schmidt's Law is the diffusion of coal -->
+<!-- power in Europe (Grubler, 2012). England emerged as the core region for -->
+<!-- coal power because it had legal and economic institutions that -->
+<!-- incentivized scientific pursuits, domestic coal reserves, and a clear -->
+<!-- industrial motivation to replace water power with coal. ... -->
 
-@fritsch2015new analyze the spatial diffusion of laser technology
-research in West Germany from 1960, when this technology began, until
-2005.
+<!-- country level, not sure if needed. -->
 
-@perkins2011internet
-
-@lengyel2020role We see that the OSN spread almost exclusively from the
-original location (the capital Budapest, with an order of magnitude more
-inhabitants than the next size town) to various parts of the country in
-the early phase of the life-cycle. Later, diffusion became less
-mono-centric and other towns also emerged as spreaders. Our findings
-support the idea that spreading initially happens to large distances and
-becomes more local over time. This is illustrated and discussed later in
-Figure 2I.
-
-@PAPAGIANNIDIS2015308
-
-@leibowicz2016representing for energy There is no generally accepted
-theory that explains diffusion rate heterogeneity across technologies,
-but several factors are considered important. Greater unit scale and
-larger market size contribute to slower diffusion. Requirements for
-interrelated technologies or complex infrastructures also hinder the
-diffusion process (Grubler, 2012).
-
-Mobile phones benefited from early deployment in recreational boats and
-automobiles, where the traditional competitor was not a viable option.
-In the early stages of diffusion, performance is a more important driver
-of adoption than cost competitiveness. Typically, significant cost
-reductions only occur once the technology reaches a deployment level
-capable of supporting standardization and mass production (Wilson,
-2012).
-
-@leibowicz2016representing Empirical evidence supports the validity of
-Schmidt's Law over a wide range of technologies, time periods, and
-geographical contexts. A recent meta-analysis of technology up-scaling
-found that diffusion accelerated moving from the core to the rim and
-periphery for technologies as diverse as natural gas power, oil
-refineries, and automobiles (Wilson, 2009). One historical example that
-conforms particularly well to Schmidt's Law is the diffusion of coal
-power in Europe (Grubler, 2012). England emerged as the core region for
-coal power because it had legal and economic institutions that
-incentivized scientific pursuits, domestic coal reserves, and a clear
-industrial motivation to replace water power with coal. ...
-
-@ding2010modeling model the spatial diffusion of mobile
-telecommunications in China as well as its determinants. regional
-socioeconomic characteristics play an important role in determining the
-timing, speed, and level of mobile telecommunications diffusion in China
-
-@bento2018time explore What determines the duration of formative phases
-for energy innovations in different markets? We are interested both in
-initial markets (also: core, lead, first mover, early adopter) where
-formative phases prepare technologies for mass commercialization, and in
-follower markets (also: periphery, lag, late adopter) where accelerated
-formative phases may benefit from diffusion and spillovers.
-
-@perkins2005internationalEither way, the strong assumption is that developing countries can
-acquire modern technology innovated in developed economies, often at a
-fraction of the original research and development (R&D) costs, thereby
-leapfrogging many decades of technological progress (Teece 2000).
-
-@perkins2005international The first involves a country's geographical location. Recent empirical
-work suggests that diffusion is ''geographically localized'' (Globerman,
-Kokko, and Sjo¨holm 2000; Keller 2002; Milner 2003) in that a technology
-diffuses faster in a country where it is already more widely diffused in
-neighboring countries. Underlying these regional effects are contagion
-and contact with prior users or producers of technology.
+<!-- @bento2018time explore What determines the duration of formative phases -->
+<!-- for energy innovations in different markets? We are interested both in -->
+<!-- initial markets (also: core, lead, first mover, early adopter) where -->
+<!-- formative phases prepare technologies for mass commercialization, and in -->
+<!-- follower markets (also: periphery, lag, late adopter) where accelerated -->
+<!-- formative phases may benefit from diffusion and spillovers. -->
 
 # Materials and Methods {#sec3}
+
+<!-- from our RS paper -->
+For instance, we can
+approximate digital economic activities in a region by
+measuring the volume of commercial webpages by considering
+only .co.uk websites, which are dedicated to commercial
+activities (Thelwall, 2000). Such commercial
+websites are used to exchange information, support online
+transactions and share opinions (Blazquez & Domenech,
+2018). Although nothing prevents a UK-based company
+from adopting a generic TLD such as.com and, indeed,
+such cases escape our data, we do not expect that such
+omissions could affect our results given the popularity of
+the .uk ccTLD: UK consumers prefer to visit a .uk website
+when they are searching for products or services (Hope,
+2017); and anecdotal evidence indicates that during the
+first half of 2000, three .co.uk domains were registered
+every minute (OECD, 2001). Also, as Table 1 demonstrates,
+.co.uk was the most popular SLD under the .uk
+ccTLD in 2000. Hence, we use the total volume of all
+the archived webpages in a region as a proxy for the level
+of digitization. We start the analysis by focusing on commercial
+webpages (.co.uk).
 

### 2023-10-27 paper/submission1/tranos2023.tex
presentation

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, frame hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 domain and adequate research methods can offer new contributions
-regarding our understanding of technologies are diffused.
+regarding our understanding of how new technologies diffuse.
 
@@
 times, at different rates, and to different extents in different places,
-and can be significantly influenced by policies
-\citep{victor1993}''.@meade2021modelling highlight that a variety of
-actors have a direct interest in gaining such knowledge including
-network equipment suppliers; network operators, regulatory and local
-authorities. These processes and their effects very a lot between
-scales: although the diffusion of a new technology might not be optimal
-at a local level, it might be beneficial from a global perspective as it
-might enable faster diffusion to less advantaged places
-\citet{leibowicz2016representing}. Despite the spatial heterogeneity of
-such diffusion mechanisms and the policy relevance, there are very
-limited attempts in the literature to analyse the diffusion of new
-digital technologies at a detailed geographical level.
+and can be significantly influenced by policies \citep{victor1993}''.
+\citet{meade2021modelling} highlight that a variety of actors have a
+direct interest in gaining such knowledge including network equipment
+suppliers; network operators, regulatory and local authorities. These
+processes and their effects very a lot between scales: although the
+diffusion of a new technology might not be optimal at a local level, it
+might be beneficial from a global perspective as it could lead to faster
+diffusion to less advantaged places \citep{leibowicz2016representing}.
+Despite the spatial heterogeneity of such diffusion mechanisms and the
+policy relevance, there are very limited attempts in the literature to
+analyse the diffusion of new digital technologies at a detailed
+geographical level.
 
-To summarize the discussion, first, here the diffusion process is
-selfperpetuating as initial use stimulates further use. Second, it
-follows a disequilibrium path as the level of users is always lower than
-the number of potential users (Stoneman 2002).
+Technological diffusion, which is by definition an aggregated process,
+can be discussed in parallel with individual adoption mechanisms. On the
+one hand, \citet{rogers2010diffusion} identifies early adopter of new
+technologies as `knowledgeable risk takers' and \citet{griliches1957} as
+`profit maximisers' \citep{ding2010modeling}. Such individual agents are
+rewarded because of their attitude towards new technologies and
+innovations. On the other hand, \citet{perkins2011internet} attribute
+diffusion to two processes: (i) epidemic-like mechanisms, which are
+governed by distance, proximity and social interactions, and (ii) by
+economic mechanisms as new innovations are adopted by users as they
+become more profitable, valueable and useful.
 
-The criticisms are mainly based on the fact that although the approach
-gives an idea of aggregate (industry or household) behaviour, it does
-not focus on the individual's (firm or household) adoption process
+This paper focuses on the diffusion of the web as new technology during
+the \(1996-2012\) period. This was an exciting period for digital
+technologies as it corresponds with the commercialisation of the
+internet and, consequently, its almost universal adoption. The reader is
+reminded that it was only in 1994 when Netscape Navigator was
+introduced, a year before Microsoft's Internet Explorer \footnote{\url{https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft}}.
+Also, only \(9\) per cent of UK's households had access to the internet
+in \(1998\) \citep{ons2018} and during that period that web included
+mostly static webpages, there were no social media and web browsing was
+happening exclusively from desktop PCs as there were no smartphones
+\citep{tranosuk}. Hence, it is fair to say that the study period
+captures the very early stages of the diffusion on a new technology that
+is the web. This is key point in the lifecycle of a new technology.
+Firstly, during this period new technologies are expensive, crude and
+imperfect \citep{rosenberg1994exploring, wilson201281}. A simple
+comparison between Web 1.0 and Web 3.0 applications clearly illustrates
+this \citep{tranos2020social}. During this period the performance of a
+new technology is the main attraction and not the cost to access and use
+it \citep{wilson2011lessons} There is a broader theoretical discussion
+in the literature regarding the motivation behind early adoption. As
+summarised by \citet{perkins2005international}, on the one hand,
+epidemic models highlight the role of interpersonal contacts as a way
+for new technologies to diffuse. On the other hand, economic models
+underline the importance of heterogeneity. Different firms have
+different structures and business plans, which define the potential
+economic returns of the adoption of a new technology and, therefore, the
+choice to adopt a new technology becomes an individual option. From a
+broader and evolutionary perspective, initial conditions are essential
+for the creation and evolution of path-dependent technological
+development trajectories \citep{neffke2011regions, simmie2014new}. This
+argument is even more relevant when the focus is on Information and
+Communication Technologies (ICT) because of the commonly found lag
+between investment and economic returns as reflected in the Solow
+paradox \citep{acemoglu2014return, brynjolfsson2018artificial}.
 
-\citet{morrill2020spatial} Spatial diffusion is the process by which
-behavior or characteristics of the landscape change as a result of what
-happens elsewhere earlier. Spatial diffusion is the spread of the
-phenomenon, over space and timed, from limited origins
-
-Compare dissusion with adoption
-
-wilson201281 Early on in their lifecycle, new technologies are crude,
-imperfect, and expensive (Rosenberg, 1994). New energy technologies are
-attractive for their ability to perform a particular task or deliver a
-new or improved energy service (Fouquet, 2010). This is often
-circumscribed by a particular set of needs in a particular context: a
-market `niche'. End-users in niche markets are generally less sensitive
-to the effective price of the energy service provided or have a higher
-willingness to pay for its performance advantages (Fouquet, 2010). Thus
-initially, performance dominates cost competitiveness (Wilson and
-Grubler, 2011). Market niches afford some protection from competitive
-pressures, allowing technologies to be tested and improved in applied
-settings, reducing uncertainties with performance or market demand (Kemp
-et al., 1998). Costs may only fall substantively after an extended
-period of commercial experimentation, concurrent with the establishment
-of an industrial base and characteristic moves towards standardisation
-and mass production (Grubler, 1998). The influence of accumulating
-production experience on costs is captured by the concept of learning.
-
-grubler1990rise The scale of spatial innovation diffusion is similar to
-that of temporal diffusion models: hierarchically decomposed.
+Importantly, the data used here depicts the \emph{active} engagement
+with the web as it contains all the commercial websites archived by the
+Internet Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
+that include a mention to at least one valid UK postcode in the web
+text. Websites are identified as commercial ones if they are part of the
+.co.uk second level domain (SLD). Also, the act of of creating a website
+is understood here as active engagement with the web vis-à-vis the more
+passive act of browsing the web or having an internet connection
+\citep{tranosuk}. Previous studies have focused mostly on more passive
+notions of engaging with digital technologies such as internet adoption
+and internet speeds
+\citep[e.g.][]{blank2018local, riddlesden2014broadband}.
 
@@
 
-\citet{perkins2005international} Theoretical models disagree as to why
-firms adopt innovations at different times. Epidemic models emphasize
-information (Griliches 1957). Certain firms are hypothesized to adopt
-earlier because they come into contact with, and learn from, adopters of
-the new technology before others. Economic models, on the other hand,
-predominantly emphasize firm heterogeneity (Ireland and Stoneman 1986).
-Firms adopt technologies at different times because they differ with
-respect to various organizational and environmental variables
-influencing the economic returns from adoption (Blackman 1999).
-
-Either way, the strong assumption is that developing countries can
-acquire modern technology innovated in developed economies, often at a
-fraction of the original research and development (R\&D) costs, thereby
-leapfrogging many decades of technological progress (Teece 2000).
-
-The first involves a country's geographical location. Recent empirical
-work suggests that diffusion is `'geographically localized'' (Globerman,
-Kokko, and Sjo¨holm 2000; Keller 2002; Milner 2003) in that a technology
-diffuses faster in a country where it is already more widely diffused in
-neighboring countries. Underlying these regional effects are contagion
-and contact with prior users or producers of technology.
-
-\citet{ding2010modeling} Rogers (1995) characterized early adopters as
-knowledgeable risk takers. Griliches (1957) looked at the rate of return
-for early adopters and characterized them as profit maximizers. The
-geographer Hagerstrand (1967) was focusing on the character of early
-adopters in the transmittal process that produces follow-on participants
-in the process. Space was treated as a contiguity system with embedded
-characteristics producing barrier elements to the diffusion process.
-
-\citet{ding2010modeling} REVIEW ON Technological Diffusion
-
-\citet{perkins2011internet} Two main mechanisms are identified in the
-literature to explain diffusion: (1) epidemic-type dynamics whereby
-contact with previous adopters stimulates uptake as po tential adopters
-learn about a new innovation; and (2) economic-type mechanisms whereby
-potential users adopt a new innovation as it be comesmore profitable,
-useful, or valuable, with uptake characteristically spreading as costs
-be come lower, performance improves, or the po tential uses of the
-innovation grow over time.
-
-To this extent, our results contribute to a growing body of work that
-has sought to caution against claims about the supposed novelty of the
-Internet and the suggestion that it is somehow different.
-
 \citet{wilson201281} Logistic growth describes an initial period of
@@
 
-MAYBE SKIP THIS??
+Geographical diffusion is a synthesis of different processes. On the one
+hand, we can identify purely spatial or, in other words, contagious
+processes. Adjacency and, more broadly, distance are the key drivers of
+diffusion. This perspective draws similarities with epidemics:
+innovation just like pathogens spreads because of contagion and,
+consequently, proximity and exposure \citep{hivner2003facilitating}. On
+the other hand, we can identify hierarchical processes. Instead of
+horizontal distance-based diffusion mechanisms, the top-down hierarchy
+of urban systems drives diffusion. In reality, the synthesis of these
+two processes represents how new technologies diffuse over space and
+time \citep{morrill2020spatial}.
 
-\citet{beardsell1999spatial} To characterize evolution of the computer
-industry, we examine the distribution of relative employment across
-cities in 1977 and how that distribution changes over time to 1992
+These ideas were firstly introduced by Torsten Hägerstrand and his
+thesis entitled `Innovation Diffusion as a Spatial Process'
+\citep{hagerstrand1968innovation}. Hägerstrand was the first one to
+identify diffusion as a geographical process. The starting point was the
+idea that diffusion is based on passing information through social
+networks, which themselves tend to be defined by geography. Hence, he
+identified the `neighbourhood' effect of how information, and
+consequently, innovation diffuse. He used agricultural innovations to
+test and model his ideas using Monte Carlo simulations. Hägerstrand also
+incorporated the role of hierarchy and how some phenomena maybe firstly
+adopted in larger cities and then diffuse to second tier ones. This is a
+sequential instead of a simultaneous process, which resembles the
+`lead-lag' spatial acceleration effect in market research
+\citep{bento2018time, PERES201091}. Hägerstrand maybe more widely known
+for highlighting the role time plays in the diffusion of innovations: an
+early-pioneering period, a middle fast accelerating period and a final
+saturation one {[}\citet{morrill2020spatial}.
 
-There is no tendency of relative size distributions of urban computer
-employment to collapse, go bimodal, or fully spread. Overall computers
-exhibit some turbulence, with dramatic big winners and losers among
-cities. In attracting or repelling an industry, urban heterogeneity is
-important. Large, well educated cities near San Jose have much greater
-chances of attracting high-tech employment and less of losing it
+The temporal dimension was further explored by Everett Rogers and his
+seminal work on `Diffusion of Innovations' \citep{rogers2010diffusion}.
+Rogers being a sociologist his work focused not on the diffusion of
+innovations over space and time, but instead on the adoption of new
+technologies and innovations by individuals and the individual
+mechanisms that drive the decisions behind adoption. He identified five
+groups of individuals regarding their adoption speed: innovators, early
+adopters, early majority, late majority and laggards. Key mechanism of
+diffusion and adoption is always communication and how knowledge is
+transferred within a social system. Therefore, all approaches agree on
+the S-shaped diffusion and cumulative adoption pattern
+\citep{grubler1990rise}.
 
-\citet{bednarz2020pulled} spatial innovation diffusion By using Bayesian
-survival models with time-dependent data of wind turbine deployment and
-firm foundation for 402 German regions between the years 1970 and 2015,
-we show that the spatial evolution of the German wind energy industry
-was more strongly influenced by local demand-- pull than local
-supply--push processes.
+Schmidt's Law empirically illustrates a similar pattern. \emph{Core} and
+usually highly agglomerated regions is where new technologies are
+invented and commercially deployed {[}\citet{grubler1990rise};
+grubler1990rise{]}. This is where the first adopters tend to be based.
+Then, technologies spread to the \emph{rim} and eventually to the
+\emph{periphery}. Although adoption pace might be higher when new
+technologies finally arrive to the periphery, the saturation levels
+there may never reach the ones in the core because of the lack of
+infrastructure or other necessary institutions
+\citep{leibowicz2016representing}. \citet{grubler1990rise} effectively
+summarises the three key characteristics of the spatial diffusion
+process: (i) the cumulative level of adoption follows an S-shaped
+pattern just like purely temporal models; (ii) diffusion is shaped by a
+hierarchy effect in a form of a centrifugal force: from core to
+periphery; and (iii) diffusion is also shaped by distance and a
+neighbourhood effect and contaminate nearby locations.
 
-The industry's initial locations are distributed relatively arbitrarily
-and unpredictably, as their needs in terms of resources and skills are
-diverse and distinct from the older existing industries (Boschma and
-Lambooy, 1999). Consequently, emerging industries are characterized by
-relatively high degrees of freedom in terms of location. In later
-extensions of the concept, the assumption of the randomness of locations
-was revised with greater importance assigned to regional conditions
-(Boschma and Lambooy, 1999; Fornahl et al., 2012).
+The remaining of this section reviews empirical studies which analysed
+the diffusion on new technologies over space and time. Although the
+spatial dimension is present in most of the following studies, the level
+of spatial detail is almost always more coarse than the one adopted
+here. To begin with \citet{beardsell1999spatial} studied the evolution
+of the computer industry in \(317\) US metro areas during the
+\(1977-1992\) period using employment data. Their analysis indicated
+that the relative size distribution holds for urban computer employment
+and also urban heterogeneity is essential in explaining this
+distribution. \citet{bednarz2020pulled} focused on wind turbines and
+modelled their spatial diffusion across \(402\) German regions during
+\(1970-2015\). Their key finding is that local demand than local supply
+was the main driving factor.
 
-\citet{perkins2005international} analyze whether the rate at which new
-producer technologies diffuse is significantly influenced by (1)
-latecomer advantage and (2) engagement with the global economy via trade
-and foreign investment
+At a global scale \citet{perkins2005international} analyzed if the
+diffusion rate of new technologies is driven by a latecomer advantage
+and the engagement with the global economy via foreign direct
+investments and trade. Their results illustrate that indeed latecomers
+and developing countries experience diffusion of new technologies more
+rapidly than early adopter and developed countries. At the same scale,
+\citet{perkins2011internet} explored whether the adoption of previous
+communication technologies that is mail, telegrams and telephones was
+shaped by similar socioeconomic factors as the internet. Their results
+indicated common patterns regarding the drivers behind the adoption of
+different communication technologies.
 
-Indeed, precisely because of these latecomer advantages, developing
-countries1 are believed to be well placed to catch up with developed
-ones (Gerschenkron 1962; Abramovitz 1986).
-
-\citet{fritsch2015new} analyze the spatial diffusion of laser technology
-research in West Germany from 1960, when this technology began, until
-2005.
-
-\citet{perkins2011internet}
-
-\citet{lengyel2020role} We see that the OSN spread almost exclusively
-from the original location (the capital Budapest, with an order of
-magnitude more inhabitants than the next size town) to various parts of
-the country in the early phase of the life-cycle. Later, diffusion
-became less mono-centric and other towns also emerged as spreaders. Our
-findings support the idea that spreading initially happens to large
-distances and becomes more local over time. This is illustrated and
-discussed later in Figure 2I.
-
-\citet{PAPAGIANNIDIS2015308}
-
-\citet{leibowicz2016representing} for energy There is no generally
-accepted theory that explains diffusion rate heterogeneity across
-technologies, but several factors are considered important. Greater unit
-scale and larger market size contribute to slower diffusion.
-Requirements for interrelated technologies or complex infrastructures
-also hinder the diffusion process (Grubler, 2012).
-
-Mobile phones benefited from early deployment in recreational boats and
-automobiles, where the traditional competitor was not a viable option.
-In the early stages of diffusion, performance is a more important driver
-of adoption than cost competitiveness. Typically, significant cost
-reductions only occur once the technology reaches a deployment level
-capable of supporting standardization and mass production (Wilson,
-2012).
-
-\citet{leibowicz2016representing} Empirical evidence supports the
-validity of Schmidt's Law over a wide range of technologies, time
-periods, and geographical contexts. A recent meta-analysis of technology
-up-scaling found that diffusion accelerated moving from the core to the
-rim and periphery for technologies as diverse as natural gas power, oil
-refineries, and automobiles (Wilson, 2009). One historical example that
-conforms particularly well to Schmidt's Law is the diffusion of coal
-power in Europe (Grubler, 2012). England emerged as the core region for
-coal power because it had legal and economic institutions that
-incentivized scientific pursuits, domestic coal reserves, and a clear
-industrial motivation to replace water power with coal. \ldots{}
-
-\citet{ding2010modeling} model the spatial diffusion of mobile
-telecommunications in China as well as its determinants. regional
-socioeconomic characteristics play an important role in determining the
-timing, speed, and level of mobile telecommunications diffusion in China
-
-\citet{bento2018time} explore What determines the duration of formative
-phases for energy innovations in different markets? We are interested
-both in initial markets (also: core, lead, first mover, early adopter)
-where formative phases prepare technologies for mass commercialization,
-and in follower markets (also: periphery, lag, late adopter) where
-accelerated formative phases may benefit from diffusion and spillovers.
+Turning to studies that share more technological and scalar similarities
+with this paper, \citet{ding2010modeling} modelled the spatial diffusion
+of mobile telecommunications across regions in China. Their analysis
+indicated that socioeconomic characteristics are important determinants
+of the timing, speed and the level of mobile diffusion within China.
+Using data from a Hungarian online social network,
+\citet{lengyel2020role} analysed its adoption and the churn at a very
+granular spatial level. Their results are in agreement with early
+theoretical and empirical contributions reviewed here: assortativity,
+urban scaling and distance are the key drivers of spatial diffusion. At
+a global scale \citet{PAPAGIANNIDIS2015308} modelled the diffusion of
+different web technologies technologies and practices. Interestingly,
+they did so by using similar, but less extensive data as the one used
+here. Their analysis illustrated how the diffusion of different web
+technologies and practices follow an S-shaped pattern as well as the
+different diffusion rates of the different technologies and practices.
 
@@
 \section{Materials and Methods}\label{sec3}}
+
+For instance, we can approximate digital economic activities in a region
+by measuring the volume of commercial webpages by considering only
+.co.uk websites, which are dedicated to commercial activities (Thelwall,
+2000). Such commercial websites are used to exchange information,
+support online transactions and share opinions (Blazquez \& Domenech,
+2018). Although nothing prevents a UK-based company from adopting a
+generic TLD such as.com and, indeed, such cases escape our data, we do
+not expect that such omissions could affect our results given the
+popularity of the .uk ccTLD: UK consumers prefer to visit a .uk website
+when they are searching for products or services (Hope, 2017); and
+anecdotal evidence indicates that during the first half of 2000, three
+.co.uk domains were registered every minute (OECD, 2001). Also, as Table
+1 demonstrates, .co.uk was the most popular SLD under the .uk ccTLD in
+2000. Hence, we use the total volume of all the archived webpages in a
+region as a proxy for the level of digitization. We start the analysis
+by focusing on commercial webpages (.co.uk).
 
@@
 \renewcommand\refname{References}
-  \bibliography{./paper/submission1/bibliography}
+  \bibliography{bibliography}
 

### 2023-10-27 presentations/web_diffusion.Rmd
presentation

@@
 -   Conclusions <br> <br>
-    [etranos.info/post/gceg2022](https://etranos.info/post/gceg2022/)
+    <!-- [etranos.info/post/gceg2022](https://etranos.info/post/gceg2022/) -->
 :::
@@
 
-<br>
-
+-   Diffusion of a new technology: the *web*
 -   Geographers used to be interested in diffusion
-
 -   @hagerstrand1968innovation
-
 -   Passed the torch to economists and sociologists 
-
 -   Why? Lack of granular data:
@@
 
+<br>
+
 -   Archived web data
 
--   Observe commercial websites 1996 - 2012
+-   Observe commercial websites 1996 - 2012 in the UK (*.co.uk*)
 
--   Geolocate to a unique location
+-   Geolocate:
 
--   Geolocate to multiple locations
+    - to a unique postcode
+
+    - to multiple postcodes
+
+-------------------------------------------
+
+```{r , eval=T, message=FALSE, out.width = "1200px"}
+path.image <- paste0(path, "/presentations/images/past_publications.png")
+knitr::include_graphics(path.image)
+```
 
@@
 
-![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
+![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
 
@@
 
-![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
+![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
 
@@
 
-## Spatial attributes
+---
 
-> **Neighborhood effect**: diffusion proceeds outwards from innovation
+> **Neighbourhood effect**: diffusion proceeds outwards from innovation
 > centers, first "hitting" nearby rather than far-away locations
@@
 
--   Website density regressions
+-   Website density regressions -- distance effect
 
@@
 ```{r moransi_la, results='asis'}
-path.in <- paste0(path, "/outputs/morani_la.csv")
+path.in <- paste0(path, "/outputs/lisa/morani_la.csv")
 
@@
 ```{r, eval=T, moransi_oa, results='asis'}
-path.in <- paste0(path, "/outputs/morani_oa.csv")
+path.in <- paste0(path, "/outputs/lisa/morani_oa.csv")
 # read.csv(path.in) %>% dplyr::select(year, morani) %>% 
@@
 ```{r lisa_la, results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/")
+path.images <- paste0(path, "/outputs/lisa/")
 
@@
 ```{r lisa_oa, eval=T,results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/")
+path.images <- paste0(path, "/outputs/lisa/")
 
@@
 ```{r b_London_la, fig.height=5, fig.width=5, message=FALSE}
-path.image <- paste0(path, "/outputs/b_dist_london_la.png")
+path.image <- paste0(path, "/outputs/lisa/b_dist_london_la.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r b_city_la, fig.height=5, fig.width=5, message=FALSE}
+#, out.width = "60%"
+path.image <- paste0(path, "/outputs/lisa/b_dist_city_la.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r b_centre_la,  fig.height=5, fig.width=5,message=FALSE}
+#, out.height= "60%", out.width="60%"
+path.image <- paste0(path, "/outputs/lisa/b_dist_centre_la.png")
 knitr::include_graphics(path.image)
@@
 #, out.width = "60%"
-path.image <- paste0(path, "/outputs/b_dist_london_oa.png")
+path.image <- paste0(path, "/outputs/lisa/b_dist_london_oa.png")
 knitr::include_graphics(path.image)
@@
 
-```{r b_city_la, fig.height=5, fig.width=5, message=FALSE}
-#, out.width = "60%"
-path.image <- paste0(path, "/outputs/b_dist_city_la.png")
+```{r b_centre_oa, eval = T,  fig.height=5, fig.width=5,message=FALSE}
+#fig.height=4, fig.width=4, 
+#, out.height= "60%", out.width="60%"
+path.image <- paste0(path, "/outputs/lisa/b_dist_city_oa.png")
 knitr::include_graphics(path.image)
@@
 #, out.width = "60%"
-path.image <- paste0(path, "/outputs/b_dist_city_oa.png")
+path.image <- paste0(path, "/outputs/lisa/b_dist_centre_oa.png")
 knitr::include_graphics(path.image)
@@
 
-```{r b_centre_la, message=FALSE}
-#, out.height= "60%", out.width="60%"
-path.image <- paste0(path, "/outputs/b_dist_centre_la.png")
-knitr::include_graphics(path.image)
-```
+```{r r_squared, eval=T, results='asis'}
+# This is for the RSquared of the different website density regressions.
 
-------------------------------------------------------------------------
+path.in1 <- paste0(path, "/outputs/lisa/r2_all_oa.csv")
+path.in2 <- paste0(path, "/outputs/lisa/r2_all_la.csv")
 
-```{r b_centre_oa, eval = T, message=FALSE, out.height= "60%", out.width="60%"}
-#fig.height=4, fig.width=4, 
-path.image <- paste0(path, "/outputs/b_dist_centre_oa.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r r_squared, eval=F, results='asis'}
-# This is for the RSquared of the different website density regressions.
-# The RSquared does not drop much hence it is not worth reporting.
-
-path.in1 <- paste0(path, "/outputs/r2_all_oa.csv")
-path.in2 <- paste0(path, "/outputs/r2_all_la.csv")
-
-read.csv(path.in2) %>% #left_join(read.csv(path.in2)) %>% 
-  ggplot(aes(x=year)) +
+# read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
+#   ggplot(aes(x=year)) +
 #   geom_line(aes(y=London, color = "OA to London"), size=1) +
@@
 #   geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) +
+# 
+#   geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) +
+#   geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) +
+#   geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) +
+# 
+#   scale_color_discrete(name = "R-squared for regression:") +
+#   #scale_linetype_manual(name = "Website density regressions111") 
+#   ylab("R-squared") + 
+#   ggtitle("Website density regressions") +
+#   theme_minimal() +
+#   theme(axis.title.x=element_blank(),
+#         plot.title = element_text(hjust = 0.5))
 
-  geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) +
-  geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) +
-  geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) +
-
+read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
+  pivot_longer(!year, names_to = "regression", values_to = "Rsquared") %>% 
+  mutate(LA = if_else(str_ends(regression, ".LA"), " to LA", "to OA"),
+         regression = str_remove_all(regression, ".LA"),
+         line = paste0(regression, " ", LA)) %>% 
+  ggplot() +
+  geom_line(aes(x=year, y = Rsquared, color = line),
+            size = 1) +
   scale_color_discrete(name = "R-squared for regression:") +
-  scale_linetype_manual(name = "Website density regressions") +
+  #scale_linetype_manual(name = "Website density regressions111") 
   ylab("R-squared") + 
@@
 
-# read.csv(path.in2) #%>% left_join(read.csv(path.in2)) %>% 
-#   pivot_longer(names_to = "destinations",
-#                values_to = "Rsquared",
-#                cols = 2:7) %>% 
-#   arrange(destinations, year) %>% 
-#   ggplot(aes(x=year)) +
-#   geom_line(aes(y=Rsquared, group=destinations, colour=destinations)) +
-#   geom_textline(aes(y=Rsquared, group=destinations, colour=destinations, 
-#                     label=destinations),
-#                 hjust = 1) +
-#   theme(legend.position = "none") +
-#   theme_minimal()
+```
 
-```
+## Neighbourhood effect
+
+- Spatial dependency
+	- Relatively small and constant over time / scales
+	- London hot spot early on
+	- At local scale, consistent hotspots over time
+	- Granular analysis reveals other hotspots
+
+- Distance effect
+	- Urban gravitation increases over time and then drops
+	- Granular analysis: gravitation
+	- Lost explanatory power over time
 
@@
 
-```{r gini_oa, eval=T, results='asis'}
-path.in <- paste0(path, "/outputs/gini_oa.csv")
+```{r gini_la, results='asis'}
+path.in <- paste0(path, "/outputs/lisa/gini_la.csv")
 
@@
   theme_minimal() +
-  ggtitle("Gini coefficient, number of websites per Output Area") +
+  ggtitle("Gini coefficient, websites per firm in Local Auhtority") +
   theme(plot.title = element_text(hjust = 0.5)) 
@@
 
-```{r gini_la, results='asis'}
-path.in <- paste0(path, "/outputs/gini_la.csv")
+```{r gini_oa, eval=T, results='asis'}
+path.in <- paste0(path, "/outputs/lisa/gini_oa.csv")
 
@@
   theme_minimal() +
-  ggtitle("Gini coefficient, number of websites per Local Auhtority") +
+  ggtitle("Gini coefficient, websites per Output Area") +
   theme(plot.title = element_text(hjust = 0.5)) 
@@
 
-## Spatial attributes, a summary
-
--   Spatial dependency relatively small and constant over time / scales
-
--   At local scale, consistent hotspots over time
-
--   More granular analysis reveals hotspots
+## Hieararchy
 
@@
 
+-   Polarisation decreases over time
+
 -   More equally diffused at the Local Authority level
-
--   Plateau overtime
-
--   Distance effect: urban gravitation increases over time and then
-    drops
-
--   Consistent across scales and definitions of urban
 
@@
 
-```{r UK S, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+```{r UK S, eval=T}
+#, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"
+
 path.image <- paste0(path, "/outputs/s_uk.png")
@@
 
-# Rank dynamics: stability vs. deversity
+## Diffusion speed
+
+<br>
+
+- Spatial heterogeneity
+- Not clear pattern
+
+# Rank dynamics: stability vs. volatility
+
+------------------------------------------------------------------------
+
+<br>
+
+- Adoption heterogeneity 
+- Perceptions of risk and economic returns from new technologies
+- Early adopters vs. laggards, leapfrogging
 
@@
 
-------------------------------------------------------------------------
-
-to do: the same for OAs
-
-------------------------------------------------------------------------
 # Conclusions
 
-## What have we learned?
+----
+
+<br>
 
@@
 
-## What have we *yet* to learn?
-
--   Explain the spatial patterns of fast/slow web adoption
-
-    -   Applied the same analysis fo OA
-
-    -   Survival regressions [@perkins2005international]
-
--   Expand our definition of web to websites with \# of postcodes \> 1
+-   Stability and volatility: leapfrogging, early adopters dropping, but also maintaining
 

### 2023-10-27 src/level_lisa_la.Rmd
presentation

@@
 library(rprojroot)
-library(rgdal)
-library(rgeos)
+# library(rgdal)
+# library(rgeos)
+library(sf)
 library(spdep)
 library(tmap)
-library(maptools)
+#library(maptools)
 library(knitr)
@@
 library(raster)
+library(plm)
+library(lmtest)
 
@@
 
-```{r}
+```{r, eval=F}
 firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>%
@@
 
-Moran I
+## Moran I
 
@@
 
-Website density regressions to London
+## Panel regressions
+
+```{r}
+# London
+model <- plm(log(n+0.001) ~ log(London+.001)*as.factor(year), 
+             data = df,
+             index = c("ladcd", "year"),
+             model = "within",
+             effect = "twoways")
+summary(model)
+
+est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
+int <- confint(est)
+
+out.model <- bind_cols(tidy(est), 
+                       as_tibble(int)) %>% 
+  mutate_if(is.numeric, round, 2) %>% 
+  mutate(year = 1997:2012) %>% 
+  rename(conf.low = '2.5 %',
+         conf.high = '97.5 %')
+
+path.out <- paste0(path,"/outputs/lisa/b_panel_dist_london_la.png")
+ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
+  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
+                    width = 0.2,size  = 1,
+                    position = "dodge", color="turquoise4") +
+  #geom_hline(yintercept = 0, color = "red", size = 1) +
+  geom_point() + coord_flip() +
+  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
+  ggtitle("Regression coefficients: distance to London on website counts, LA") +
+  theme(plot.title = element_text(hjust = 0.5))
+ggsave(path.out)
+
+# Cities
+model <- plm(log(n+0.001) ~ log(dist+.001)*as.factor(year), 
+             data = df,
+             index = c("ladcd", "year"),
+             model = "within",
+             effect = "twoways")
+summary(model)
+
+est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
+int <- confint(est)
+
+out.model <- bind_cols(tidy(est), 
+                       as_tibble(int)) %>% 
+  mutate_if(is.numeric, round, 2) %>% 
+  mutate(year = 1997:2012) %>% 
+  rename(conf.low = '2.5 %',
+         conf.high = '97.5 %')
+
+path.out <- paste0(path,"/outputs/lisa/b_panel_dist_cities_la.png")
+ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
+  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
+                    width = 0.2,size  = 1,
+                    position = "dodge", color="turquoise4") +
+  #geom_hline(yintercept = 0, color = "red", size = 1) +
+  geom_point() + coord_flip() +
+  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
+  ggtitle("Regression coefficients: distance to Cities on website counts, LA") +
+  theme(plot.title = element_text(hjust = 0.5))
+ggsave(path.out)
+
+# Retail centres
+model <- plm(log(n+0.001) ~ log(dist.retail+.001)*as.factor(year), 
+             data = df,
+             index = c("ladcd", "year"),
+             model = "within",
+             effect = "twoways")
+summary(model)
+
+est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
+int <- confint(est)
+
+out.model <- bind_cols(tidy(est), 
+                       as_tibble(int)) %>% 
+  mutate_if(is.numeric, round, 2) %>% 
+  mutate(year = 1997:2012) %>% 
+  rename(conf.low = '2.5 %',
+         conf.high = '97.5 %')
+
+path.out <- paste0(path,"/outputs/lisa/b_panel_dist_retail_la.png")
+ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
+  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
+                    width = 0.2,size  = 1,
+                    position = "dodge", color="turquoise4") +
+  #geom_hline(yintercept = 0, color = "red", size = 1) +
+  geom_point() + coord_flip() +
+  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
+  ggtitle("Regression coefficients: distance to Retail Centres on website counts, LA") +
+  theme(plot.title = element_text(hjust = 0.5))
+ggsave(path.out)
+```
+
+## Website density regressions to London
 
@@
 
-Website density regressions to the nearest city
+## Website density regressions to the nearest city
 
@@
 
-Website density regressions to the nearest retail centre
+## Website density regressions to the nearest retail centre
 

### 2023-11-05 src/help_predict.Rmd
rf

---
title: "Untitled"
author: "Emmanouil Tranos"
date: "17/04/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(caret)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

```{r}
df.model <- df %>% complete(year, oa11cd, fill = list(n = 0)) %>% 
  arrange(oa11cd, year) %>% 
  left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
  left_join(dist.retail, by = c("oa11cd" = "oa11cd")) %>% 
  filter(!is.na(London)) %>% 
  dplyr::select(year, oa11cd, n, London, dist, dist.retail) %>% 
  left_join(classes.oa %>% dplyr::select(id, subgroup), by = c("oa11cd" = "id")) %>% 
  mutate(subgroup == as.factor(subgroup))

sapply(df.model, function(x) sum(is.na(x))) # mising NI
```

```{r}

train <- df.model %>% filter(year>2011)# | year==2009)
test  <- df.model %>% filter(year == 2012)

train_x <- train %>% dplyr::select(London, dist, dist.retail) #subgroup
train_x = scale(train_x)[,]
train_y = train %>% dplyr::select(n) %>% as.matrix()

test_x <- test %>% dplyr::select(London, dist, dist.retail) #subgroup
test_x = scale(test_x)[,]
test_y = test %>% dplyr::select(n) %>% as.matrix()

knnmodel = knnreg(train_x, train_y)

str(knnmodel)

pred_y = predict(knnmodel, data.frame(test_x))

print(data.frame(test_y, pred_y))

mse = mean((test_y - pred_y)^2)
mae = caret::MAE(test_y, pred_y)
rmse = caret::RMSE(test_y, pred_y)
#rsqr =

x = 1:length(test_y)

# plot(x, test_y, col = "red", type = "l", lwd=2,
#      main = "Boston housing test data prediction")
# lines(x, pred_y, col = "blue", lwd=2)
# legend("topright",  legend = c("original-medv", "predicted-medv"), 
#        fill = c("red", "blue"), col = 2:3,  adj = c(0, 0.6))
# grid()

test_y %>% as_tibble() %>% ggplot() +
  geom_line(aes(x = x, y = test_y), col = "red", lwd=.5) +
  geom_line(aes(x = x, y = pred_y), col = "blue", lwd=.5)

test_y %>% as_tibble() %>% ggplot() +
  geom_point(aes(x = test_y, y = pred_y))

```

### 2023-11-05 src/la_RF.Rmd
rf

---
title: "Random Forests"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(rgdal)
library(rgeos)
library(sf)
library(spdep)
#library(tmap)
#library(maptools)
library(knitr)
library(REAT)
library(tidygeocoder)
library(geosphere)
library(broom)
#library(foreach)
library(doParallel)
library(raster)
library(plm)
#library(lmtest)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

## Load data

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

```{r}
n = 1 #number of unique postcodes. 
m = 12

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610)

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012)

df <- bind_rows(df9610, df1112) %>% 
  filter(
    #V1.domain < m,             # for multiple pc
    V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, ladcd) %>%
  #summarise(n = n()) %>%            # for multiple pc  
  summarise(n = sum(V1.domain)) %>% # for n == 1 
  ungroup()

df <- df %>% filter(!is.na(ladcd)) %>% 
  complete(ladcd, year) %>% 
  replace(is.na(.), 0)

# tests for Scotland
# df %>% filter(substr(oa11cd, 1,1) =="S",
#               year==2010) %>% summarise(mean(n, na.rm = T))
# 
# df %>% filter(oa11cd=="S00090182")
# 
# test@data %>% filter(substr(id, 1,1) =="S",
#                      n!=0)
# 
# df %>% filter(substr(oa11cd, 1,1) =="W",
#               year==2003)
```

## Firms

```{r, eval=F}
firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>% 
  dplyr::select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
  rename(ladcd = GEOGRAPHY_CODE,
         firms = OBS_VALUE)

df <- df %>% left_join(firms) %>% 
  mutate(n_by_firm = n/firms) %>% # N websites/firm
  #filter(!is.na(firms)) %>% 
  dplyr::select(-n, -firms) %>% 
  rename(n = n_by_firm)

# df %>% filter(is.na(n_by_firm))
# sapply(df %>% left_join(firms), function(x) sum(is.na(x)))
```

## spatial data

```{r}
# get LA for the UK
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
la <- spTransform(la, CRS("+init=epsg:4326"))

#la.f <- fortify(la, region = "LAD21CD")

# cities
cities <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  arrange(pop) %>% tail(10) 
cities.sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
```

## Distances to cities and retail centres

First set up the destinations. Origins are all the LA.

Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)

```{r}
# LA centroids
la.c <- gCentroid(la, byid=TRUE)@coords %>% 
  as.data.frame() %>%
  bind_cols(la@data) %>% 
  dplyr::select(LAD21CD, x, y) %>% 
  remove_rownames()

# cities
city.names <- c("London, UK", 
"Birmingham, UK",
"Glasgow, UK",
"Liverpool, UK",
"Bristol, UK",
"Manchester, UK",
"Sheffield, UK",
"Leeds, UK",
"Edinburgh, UK",
"Leicester, UK")

cities <- geo(city.names, no_query = F, method = "arcgis")

# retail centres
geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
retail <-readOGR(geo.path) 
#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
  as.data.frame() %>% 
  add_rownames(var = "id") #%>%

retail.major.cetres <- retail.major.cetres.help %>% 
  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
  st_transform(4326) %>%
  st_coordinates() %>%
  as_tibble() %>% 
  bind_cols(retail.major.cetres.help$id) %>% 
  rename(id = '...3') %>% 
  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
```

Distance to cities

```{r}
dist <- distm(cbind(la.c$x, la.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist <- round((dist/1000),2) %>% 
  as_tibble()

city.names <- city.names %>% stringr::str_remove(", UK")
names(dist) <- city.names

dist <- dist %>% bind_cols(la.c$LAD21CD) %>% 
  rename(ladcd = last_col()) %>% 
  relocate(ladcd)

# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
# dist$distMet <- apply(dist[,2:11], 1, min)

dist <- transform(dist, dist = do.call(pmin, dist[-1]))

# NOT USED AS IT SCREWS UP THE REGRESSIONS
# Complete panel, N = 3871903, 17 year * 227759 OA
# df <- df %>% bind_rows(dist %>% select(oa11cd)) %>% 
#   mutate(year = ifelse(is.na(year),1000, year),
#          country = substring(oa11cd, 1, 1)) %>% 
#   filter(country != "L" & country != "M" & country != "N") %>% 
#   filter(!is.na(country)) %>% 
#   complete(year, oa11cd, fill = list(n = 0)) %>% 
#   filter(year!=1000,
#          !is.na(oa11cd)) %>% 
#   select(-country)

# tests
# test$country <- substring(test$oa11cd, 1, 1)
# unique(test$country)
# length(unique(test$oa11cd))

# calculate area
la$SHAPE_Area <- area(la) #/ 1000000

# Join with distance and area
df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
  left_join(la@data, by = c("ladcd" = "LAD21CD")) %>% 
  rename(area = SHAPE_Area) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,
         density = n / area) %>% 
  relocate(density, .after = area)
sapply(df, function(x) sum(is.na(x)))
```

Distance to retail centres

```{r}
dist.retail <- distm(cbind(la.c$x, la.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail <- round((dist.retail/1000),2) %>% 
  as_tibble()

# retail.names <- retail.major.cetres$RC_Name
# names(dist.retail) <- retail.names

dist.retail <- dist.retail %>% bind_cols(la.c$LAD21CD) %>% 
  rename(LAD21CD = last_col()) %>% 
  relocate(LAD21CD)

# Minimum distance 
dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
dist.retail <- dist.retail %>% dplyr::select(LAD21CD, dist.retail)

# Join with complete panel
df <- df %>% left_join(dist.retail, by = c("ladcd" = "LAD21CD"))

sapply(df, function(x) sum(is.na(x)))
```

## Spatial panel

```{r}
#spatial weight matrix
sf_use_s2(FALSE)
nb <- poly2nb(la, queen = T, snap = 1)
listw <- nb2listw(nb, style = "W", zero.policy = T)

test <- la
test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
         year = ifelse(is.na(year), t, year))

# Load required libraries
library(spdep)
library(splm)

# Assuming you have a dataset named 'data' containing your variables: Y, W, Wc, and X

# Create spatial lag matrices
W <- mat2listw(W, style="W")
Wc <- mat2listw(Wc, style="W")

# Estimate spatial panel model
spatial_panel_model <- spml(n ~ density + London + dist + dist.retail, 
                            listw=listw,
                            data=df %>% filter(!is.na(n)), 
                            index = c("ladcd", "year"),
                            effect = "individual",
                            #method = "BFGS", # eigen
                            model= "random", 
                            lag=TRUE, 
                            na.action = na.omit) #na.fail, 
                            #zero.policy = NULL)

m.spreml <- spreml(n ~ density + London + dist + dist.retail, 
                            w=listw,
                            data=df %>% filter(!is.na(n)), 
                            index = c("ladcd", "year"),
                            effect = "individual",
                            #method = "BFGS", # eigen
                            lag=TRUE, 
                            na.action = na.omit) #na.fail, 
                            #zero.policy = NULL)

# Print summary of the model
summary(m.spreml)
```

calculate k neighbours and find out why can't have spatial FE. Hierarchical?

```{r}
#install.packages('SDPDmod')
library(SDPDmod)

#spatial weights matrix
sf_use_s2(FALSE)
nb <- poly2nb(la, queen = T, snap = 1)
listw <- nb2listw(nb, style = "W", zero.policy = T)

#spatial weight matrix
binary_matrix <- nb2mat(nb, style = "B", zero.policy = TRUE)
class(binary_matrix)
w.binary <- rownor(binary_matrix)
isrownor(w.binary)
w.binary <- rownor(binary_matrix)
isrownor(w.binary)

#spatial weights matrix
library("sf")
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la.sf <- st_read(path.geo)
W_2n <- mOrdNbr(sf_pol = la.sf, m = 2) ## second order neighbors
W_knn <- mNearestN(distMat = gN3dist, m = 5) ## 5 nearest neighbors

#spatial weights matrix
sf_use_s2(FALSE)
la.c <- st_centroid(la.sf)
kn <- knearneigh(la.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn_binary_matrix <- nb2mat(knn, style = "W", zero.policy = TRUE)
isrownor(knn_binary_matrix)

mod2<-SDPDm(formula = n ~ density,# + London,# + dist + dist.retail, 
            #W = w.binary,
            #W = binary_matrix,
            #W = knn_binary_matrix,
            W = knn_binary_matrix,
            data=df %>% drop_na(),
            index = c("ladcd", "year"),
            model = "sar", 
            effect = "twoways",
            dynamic = T,
            LYtrans = TRUE, #it affects coefficients, need to read about it
            tlaginfo = list(ind = NULL, tl = T, stl = T),
            )
summary(mod2)

# if distances are included, effect = year. 
# Rho is positive and sig, AR is positive and sig, BUT Wn(t-1) is negative and sig
```

## df for RF

```{r}
# London = Islington 
sf_use_s2(FALSE)
help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)

help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
  mutate(cities = c("Leicester",
                    "Bristol",
                    "Manchester",
                    "Liverpool",
                    "Sheffield",
                    "Birmingham",
                    "Leeds",
                    "London",
                    "Edinburgh",
                    "Glasgow"))

dist$dist.city.name <- names(dist)[apply(dist[,1:9], MARGIN = 1, FUN = which.min)]
help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
  rename(nearest.city.LAD21CD = LAD21CD) %>% 
  dplyr::select(ladcd, nearest.city.LAD21CD)

knn.l <- nb2listw(knn)

df <- df %>% drop_na() %>% group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(ladcd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
  left_join(help2, by = c("ladcd" = "ladcd"))

df <- df %>% left_join(df %>% dplyr::select(ladcd,year, n) %>% 
                         rename(n.nearest.city = n), 
                       by = c("nearest.city.LAD21CD" = "ladcd",
                              "year" = "year")) %>% 
  dplyr::select(-c(LAD21NMW, OBJECTID, BNG_E, BNG_N, LONG, LAT, SHAPE_Length)) %>% 
  relocate(LAD21NM, .after = ladcd)

path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup)%>% dplyr::select(LAD21CD, RGN21CD)

df <- df %>% left_join(lookup.region, by = c("ladcd" = "LAD21CD")) %>% 
  relocate(RGN21CD, .after=year)
```

## RF

### train on all but one regions

```{r}
library(caret)
library(randomForest)

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   savePredictions = 'final')

# London region: E12000007
# training dataset
tr.df <- df %>% ungroup() %>% filter(RGN21CD!="E12000007"|
                                     RGN21CD!="E12000008")

set.seed(71)
m.test <- train(n ~ density + area + dist + dist.retail + London +
                  n.slag + year + n.nearest.city, #n.l.slag
                data = tr.df,
                trControl = tc,
                method = "rf", 
                importance = TRUE,
                na.action = na.omit)
summary(m.test)
plot(m.test)
m.test
print(m.test)

varimp_mars <- varImp(m)
plot(varimp_mars, main="Variable Importance")
```

### test on the hold out regions

Wales region is missing!!!

```{r}
pred <- predict(m, df %>% ungroup() %>% filter(RGN21CD=="E12000008"), na.action = na.pass)
length(pred)

df %>% ungroup() %>% filter(RGN21CD=="E12000008") %>% filter(if_any(everything(), is.na))

pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN21CD=="E12000008"))
postResample(pred = pred$value, obs = pred$n) ## 0.99
```

## time slices

```{r}
help <- df %>% createTimeSlices(initialWindow = 5,
                        horizon = 1,
                        fixedWindow = F)

#Adding t-1 lag for n

set.seed(71)
m.test.time <- train(n ~ area + dist + dist.retail + London + #density
             n.lag + n.slag + n.l.slag + n.nearest.city + year,
           data = df %>% filter(year==1997 | year == 1998),
           trControl = tc,
           method = "rf", 
           importance = TRUE,
           na.action = na.omit)
summary(m.test.time)
plot(m.test.time)
m.test.time

varimp_mars <- varImp(m.test)
plot(varimp_mars, main="Variable Importance")

pred <- predict(m.test, df %>% ungroup() %>% filter(year==1999), na.action = na.pass)
length(pred)

pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(year==1999))
postResample(pred = pred$value, obs = pred$n) ## 0.60 
```

## RF rolling

### train the model in a loop

TODO:

- [x] read more about trainColntrol
- [ ] other variables, check regressions, N/S?
- [x] reproducible
- [ ] n vs. log(n/n.lag), %>% mutate(n.lag = n.lag + 0.00001, n = n + 0.00001, way to bring S?
- [ ] ceter or scale with preProcess??
- [ ] direction of relationship, or cor()
- [ ]importance over time
- [ ] spatial sampling
- [ ] OA

```{r}
# Train on t & t + 1

# Make the parallelisation reproducible using https://stackoverflow.com/questions/13403427/fully-reproducible-parallel-models-using-caret. 
# See also https://stackoverflow.com/questions/27944558/set-seed-parallel-random-forest-in-caret-for-reproducible-result

#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8)

#for the last model
seeds[[11]]<-sample.int(1000, 1)

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   savePredictions = 'final')

detectCores()
cl <- makePSOCKcluster(14)
registerDoParallel(cl)

# run the model for every 2 years
start.time <- Sys.time()
years <- 1997:2012

for (t in years){
  tplus1 <- t + 1
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + 
                     n.lag + n.slag + n.l.slag + n.nearest.city + year,
                     data = df %>% filter(year==t | year==tplus1),
                     trControl = tc,
                     method = "rf", 
                     importance = TRUE)
  assign(paste0("model.all",as.character(t + 1)), model.all)
}

stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken

# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
```

### Resamples for plots

```{r}
# create a list of the above model outputs
fit.model.all <- mget(ls(pattern = "^model.all\\d"))

# resamples
fit.model.all.res <- resamples(fit.model.all)
summary(fit.model.all.res)

# plot metrics

# TODO: fix the year labels

scales <- list(x=list(relation="free"), y=list(relation="free"))
bwplot(fit.model.all.res, scales=scales)

# TODO: fix the year labels
```

### Variable importance

```{r , include=TRUE, results= 'markup', message=FALSE}
lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
```

### test

```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}

# for a reference point
df %>% group_by(year) %>%
  summarise(min = min(n), max=max(n),
            mean = mean(n), median = median(n)) %>%
  round(2) %>% kable()
```

```{r}
## Test on t + 2

# It will create: 2001 (train on 2000-2001) - 2011 (train on 2010)
# I need: 2001 (train on 2000-2001) - 2009 (train on 2008-2009)

# use the model trainned in year t to predict io of year t + 1
years <- (1998:2011)
pred.by.year.all <- data.frame(matrix(NA,nrow=374, ncol = 0))
for (t in years){
  t_plus_1 <- t + 1
  pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",t)], df[df$year==t_plus_1,])
  pred <- as.data.frame(pred)
  rownames(pred) <- c()
  pred.by.year.all <- cbind(pred.by.year.all, pred)
}

#change column names to match with total
names(pred.by.year.all) <- 1999:2012

# wide to long
pred.by.year.all <- gather(pred.by.year.all,key = year,value = predictions,1:14)

# drop 2000 and select variables form total
df.no1997_1 <- df %>% 
  filter(year!=1996 & year!=1997 & year!=1998) %>% 
  dplyr::select(ladcd, year, n)

# column bind prediction and data  
pred.by.year.all <- cbind(pred.by.year.all, df.no1997_1 %>% arrange(year, ladcd)) #DOUBLE CHECK ARRANGE

# pred.by.year to list by year
pred.by.year.all.list <- split(pred.by.year.all, pred.by.year.all$year)

# calculate metrics for every year
rf.year.all.metrics <- lapply(pred.by.year.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))
rf.year.all.metrics %>% as.data.frame() %>%
  round(2) %>%
  kable()

# there are 2 columns named year, so I drop one
pred.by.year.all$year <- NULL

ggplot(data = pred.by.year.all,aes(x = n, y = predictions)) +
  geom_point(colour = "blue") +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
  facet_wrap(~ year,ncol = 2) +
  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
  ggtitle("Predicted vs. actual trade by year")
```

### 2023-11-05 src/level_lisa_la.Rmd
rf

@@
   arrange(pop) %>% tail(10) 
-tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
+cities.sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
 ```
@@
   arrange(pop) %>% tail(10) 
-tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
+cities.sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
 
@@
   tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
-  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
+  tm_shape(cities.sf) + tm_dots() + tm_text("name", size = .5) +
   tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
@@
   tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
-  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
+  tm_shape(cities.sf) + tm_dots() + tm_text("name", size = .5) +
   tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ") +
@@
 ```
+
+## Spatial panel
+
+```{r}
+#spatial weight matrix
+sf_use_s2(FALSE)
+nb <- poly2nb(la, queen = T, snap = 1)
+listw <- nb2listw(nb, style = "W", zero.policy = T)
+
+test <- la
+test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
+  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
+         year = ifelse(is.na(year), t, year))
+
+# Load required libraries
+library(spdep)
+library(splm)
+
+# Assuming you have a dataset named 'data' containing your variables: Y, W, Wc, and X
+
+# Create spatial lag matrices
+W <- mat2listw(W, style="W")
+Wc <- mat2listw(Wc, style="W")
+
+# Estimate spatial panel model
+spatial_panel_model <- spml(n ~ density + London + dist + dist.retail, 
+                            listw=listw,
+                            data=df %>% filter(!is.na(n)), 
+                            index = c("ladcd", "year"),
+                            effect = "individual",
+                            #method = "BFGS", # eigen
+                            model= "random", 
+                            lag=TRUE, 
+                            na.action = na.omit) #na.fail, 
+                            #zero.policy = NULL)
+
+m.spreml <- spreml(n ~ density + London + dist + dist.retail, 
+                            w=listw,
+                            data=df %>% filter(!is.na(n)), 
+                            index = c("ladcd", "year"),
+                            effect = "individual",
+                            #method = "BFGS", # eigen
+                            lag=TRUE, 
+                            na.action = na.omit) #na.fail, 
+                            #zero.policy = NULL)
+
+# Print summary of the model
+summary(m.spreml)
+```
+
+calculate k neighbours and find out why can't have spatial FE. Hierarchical?
+
+```{r}
+#install.packages('SDPDmod')
+library(SDPDmod)
+
+#spatial weights matrix
+sf_use_s2(FALSE)
+nb <- poly2nb(la, queen = T, snap = 1)
+listw <- nb2listw(nb, style = "W", zero.policy = T)
+
+#spatial weight matrix
+binary_matrix <- nb2mat(nb, style = "B", zero.policy = TRUE)
+class(binary_matrix)
+w.binary <- rownor(binary_matrix)
+isrownor(w.binary)
+w.binary <- rownor(binary_matrix)
+isrownor(w.binary)
+
+#spatial weights matrix
+library("sf")
+path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
+la.sf <- st_read(path.geo)
+W_2n <- mOrdNbr(sf_pol = la.sf, m = 2) ## second order neighbors
+W_knn <- mNearestN(distMat = gN3dist, m = 5) ## 5 nearest neighbors
+
+#spatial weights matrix
+sf_use_s2(FALSE)
+la.c <- st_centroid(la.sf)
+kn <- knearneigh(la.c, k = 5)
+knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
+knn_binary_matrix <- nb2mat(knn, style = "W", zero.policy = TRUE)
+isrownor(knn_binary_matrix)
+
+mod2<-SDPDm(formula = n ~ density,# + London,# + dist + dist.retail, 
+            #W = w.binary,
+            #W = binary_matrix,
+            #W = knn_binary_matrix,
+            W = knn_binary_matrix,
+            data=df %>% drop_na(),
+            index = c("ladcd", "year"),
+            model = "sar", 
+            effect = "twoways",
+            dynamic = T,
+            LYtrans = TRUE, #it affects coefficients, need to read about it
+            tlaginfo = list(ind = NULL, tl = T, stl = T),
+            )
+summary(mod2)
+
+# if distances are included, effect = year. 
+# Rho is positive and sig, AR is positive and sig, BUT Wn(t-1) is negative and sig
+```
+## df for RF
+
+```{r}
+
+# London = Islington 
+help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
+  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)
+
+help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
+  mutate(cities = c("Leicester",
+                    "Bristol",
+                    "Manchester",
+                    "Liverpool",
+                    "Sheffield",
+                    "Birmingham",
+                    "Leeds",
+                    "London",
+                    "Edinburgh",
+                    "Glasgow"))
+
+dist$dist.city.name <- names(dist)[apply(dist[,1:9], MARGIN = 1, FUN = which.min)]
+help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
+  rename(nearest.city.LAD21CD = LAD21CD) %>% 
+  dplyr::select(ladcd, nearest.city.LAD21CD)
+
+knn.l <- nb2listw(knn)
+
+df <- df %>% drop_na() %>% group_by(year) %>% 
+  mutate(n.slag = lag.listw(knn.l, n)) %>% 
+  group_by(ladcd) %>% 
+  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year)) %>% 
+  left_join(help2, by = c("ladcd" = "ladcd"))
+
+df <- df %>% left_join(df %>% dplyr::select(ladcd,year, n) %>% 
+                         rename(n.nearest.city = n), 
+                       by = c("nearest.city.LAD21CD" = "ladcd",
+                              "year" = "year")) %>% 
+  dplyr::select(-c(LAD21NMW, OBJECTID, BNG_E, BNG_N, LONG, LAT, SHAPE_Length)) %>% 
+  relocate(LAD21NM, .after = ladcd)
+
+path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup)%>% dplyr::select(LAD21CD, RGN21CD)
+
+df <- df %>% left_join(lookup.region, by = c("ladcd" = "LAD21CD")) %>% 
+  relocate(RGN21CD, .after=year)
+```
+
+## RF
+
+### train
+
+```{r}
+library(caret)
+library(randomForest)
+
+# CV
+tc <- trainControl(method = "cv",
+                   number = 10,
+                   savePredictions = 'final')
+
+# London region: E12000007
+# training dataset
+tr.df <- df %>% ungroup() %>% filter(RGN21CD!="E12000007"|
+                                     RGN21CD!="E12000008")
+
+set.seed(71)
+m <- train(n ~ density + area + dist + dist.retail + London +
+             n.slag + year + n.nearest.city, #n.l.slag
+           data = tr.df,
+           trControl = tc,
+           method = "rf", 
+           importance = TRUE,
+           na.action = na.omit)
+summary(m)
+plot(m)
+m
+
+varimp_mars <- varImp(m)
+plot(varimp_mars, main="Variable Importance")
+```
+
+### test
+
+```{r}
+pred <- predict(m, df %>% ungroup() %>% filter(RGN21CD=="E12000008"), na.action = na.pass)
+length(pred)
+
+df %>% ungroup() %>% filter(RGN21CD=="E12000008") %>% filter(if_any(everything(), is.na))
+
+pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN21CD=="E12000008"))
+postResample(pred = pred$value, obs = pred$n) ## 0.99
+```
+## Firms
+
+```{r}
+firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>%
+  dplyr::select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
+  rename(ladcd = GEOGRAPHY_CODE,
+         firms = OBS_VALUE)
+
+df <- df %>% left_join(firms) %>% 
+  mutate(n_by_firm = n/firms) %>% # N websites/firm
+  #filter(!is.na(firms)) %>% 
+  dplyr::select(-n, -firms) %>% 
+  rename(n = n_by_firm)
+
+# df %>% filter(is.na(n_by_firm))
+# sapply(df %>% left_join(firms), function(x) sum(is.na(x)))
+```
+
+## time slices
+
+```{r}
+hepl <- df %>% createTimeSlices(initialWindow = 5,
+                        horizon = 1,
+                        fixedWindow = F)
+
+#Adding t-1 lag for n
+
+set.seed(71)
+m.test <- train(n ~ area + dist + dist.retail + London + #density
+             n.slag + year + n.nearest.city + n.l.slag,
+           data = df %>% filter(year < 1998),
+           trControl = tc,
+           method = "rf", 
+           importance = TRUE,
+           na.action = na.omit)
+summary(m.test)
+plot(m.test)
+m.test
+
+varimp_mars <- varImp(m.test)
+plot(varimp_mars, main="Variable Importance")
+
+pred <- predict(m.test, df %>% ungroup() %>% filter(year==1998), na.action = na.pass)
+length(pred)
+
+pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(year==1998))
+postResample(pred = pred$value, obs = pred$n) ## 0.99
+```

### 2023-11-08 src/la_RF.Rmd
rf

@@
 
+## REPLACE WITH 0.1 INSTEAD OF OF 0 FOR THE GROWTH
 df <- df %>% filter(!is.na(ladcd)) %>% 
@@
 
-```{r, eval=F}
+```{r, eval=T}
 firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>%
@@
 path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup)%>% dplyr::select(LAD21CD, RGN21CD)
+lookup.region <- read_csv(path.lookup) %>% dplyr::select(LAD21CD, RGN21CD)
 
@@
   relocate(RGN21CD, .after=year)
+
+# North/South
+#lookup.region %>% distinct(RGN21NM, .keep_all = T)
+south <- c("E12000007", "E12000008", "E12000009" )
+df <- df %>% mutate(south = ifelse(RGN21CD %in% south, 1, 0))
+
+# For growth:
+#df <- df %>% mutate(growth = log(n/n.lag))
+
+sapply(df, function(x) sum(is.na(x)))
 ```
@@
 - [x] read more about trainColntrol
-- [ ] other variables, check regressions, N/S?
+- [ ] other variables, LA Classification, check regressions, 
+- [x] N/S
 - [x] reproducible
 - [ ] n vs. log(n/n.lag), %>% mutate(n.lag = n.lag + 0.00001, n = n + 0.00001, way to bring S?
-- [ ] ceter or scale with preProcess??
+- [1] center or scale with preProcess??
 - [ ] direction of relationship, or cor()
-- [ ]importance over time
-- [ ] spatial sampling
+- [x] importance over time
+- [2] spatial sampling
 - [ ] OA
@@
 #(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South
 
@@
   set.seed(71)
-  model.all <- train(n ~ area + dist + dist.retail + London + 
-                     n.lag + n.slag + n.l.slag + n.nearest.city + year,
-                     data = df %>% filter(year==t | year==tplus1),
+  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
+                      n.slag + n.l.slag + n.nearest.city + year, #n.lag +
+                     data = df %>% filter(year==t | year==tplus1), #%>% 
+                       # mutate(growth = ifelse(growth == -Inf | growth == Inf, NA, growth)) %>% 
+                       # mutate(growth = ifelse(is.na(growth), 0, growth)), 
+                     #%>% mutate(growth = ifelse(growth== -Inf | growth == Inf, NA, growth))
                      trControl = tc,
-                     method = "rf", 
+                     method = "rf", #"rpart", # rf lm
+                     na.action = na.omit,
                      importance = TRUE)
@@
 time.taken
+
+# df %>% filter(growth == Inf) 
+# log(n/n.lag)
 
@@
 
+```{r}
+help <- lapply(fit.model.all, function(x) varImp(x)) 
+help <- lapply(help, function(x) x[[1]])
+help <- do.call("rbind", help)
+
+importance <- help %>% rownames_to_column() %>% 
+  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
+         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
+  rename(importance = Overall) %>% 
+  relocate(importance, .after = year) %>% 
+  dplyr::select(-rowname)
+
+importance %>% ggplot() +
+  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
+  scale_color_discrete(name = "Features") +
+  #scale_linetype_manual(name = "Website density regressions111") 
+  ylab("Importance") + 
+  ggtitle("Feature importance") +
+  theme_minimal() +
+  theme(axis.title.x=element_blank(),
+        plot.title = element_text(hjust = 0.5))
+
+```
+
 ### test
@@
 ```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
+
+# growth or n
 
@@
   filter(year!=1996 & year!=1997 & year!=1998) %>% 
-  dplyr::select(ladcd, year, n)
+  dplyr::select(ladcd, year, n) # CHANGE n WITH growth
 
@@
 rf.year.all.metrics <- lapply(pred.by.year.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))
+                                                                     obs = x$n))  # CHANGE n WITH growth
 rf.year.all.metrics %>% as.data.frame() %>%

### 2023-11-12 src/la_RF.Rmd
rf

@@
 - [ ] n vs. log(n/n.lag), %>% mutate(n.lag = n.lag + 0.00001, n = n + 0.00001, way to bring S?
-- [1] center or scale with preProcess??
+- [x] center or scale with preProcess?? preProc = c("center", "scale") makes no difference
 - [ ] direction of relationship, or cor()
 - [x] importance over time
-- [2] spatial sampling
+- [x] spatial sampling
+- [ ] hierarchy: ranking correlation and line graphs 
 - [ ] OA
@@
                      na.action = na.omit,
+                     preProc = c("center", "scale"),
                      importance = TRUE)
-  assign(paste0("model.all",as.character(t + 1)), model.all)
 }
@@
 ```
+
+## RF spatial
+
+```{r}
+df <- df %>% mutate(help = substr(ladcd, start = 1, stop = 2)) %>% 
+  mutate(RGN21CD = ifelse(is.na(RGN21CD), help, RGN21CD)) %>% 
+  ungroup() %>% 
+  dplyr::select(-help)
+
+df <- df %>% dplyr::select(-n.lag)
+
+```
+
+### train
+
+```{r}
+#length is = (n_repeats*nresampling)+1
+seeds <- vector(mode = "list", length = 11)
+
+#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South
+
+#for the last model
+seeds[[11]]<-sample.int(1000, 1)
+
+# CV
+tc <- trainControl(method = "cv",
+                   number = 10,
+                   #seeds = seeds,
+                   allowParallel = T,
+                   savePredictions = 'final')
+
+detectCores()
+cl <- makePSOCKcluster(14)
+registerDoParallel(cl)
+
+# run the model for every 2 years
+start.time <- Sys.time()
+regions <- df %>% distinct(RGN21CD)
+regions <- as.vector(regions$RGN21CD)
+
+for (i in regions){
+  set.seed(71)
+  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
+                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
+                     data = df %>% filter(RGN21CD==i), trControl = tc,
+                     method = "xgbTree", # rf "rpart", # rf lm
+                     na.action = na.omit,
+                     #preProc = c("center", "scale"),
+                     importance = TRUE)
+  assign(paste0("model.all",as.character(i)), model.all)
+}
+
+stopCluster(cl)
+end.time <- Sys.time()
+time.taken <- round(end.time - start.time,2)
+time.taken
+
+# df %>% filter(growth == Inf) 
+# log(n/n.lag)
+
+# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
+# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
+```
+
+### Resamples for plots
+
+```{r}
+# create a list of the above model outputs
+fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
+
+# resamples
+fit.model.all.res <- resamples(fit.model.all)
+summary(fit.model.all.res)
+
+# plot metrics
+
+# TODO: fix the year labels
+
+scales <- list(x=list(relation="free"), y=list(relation="free"))
+bwplot(fit.model.all.res, scales=scales)
+
+# TODO: fix the year labels
+```
+
+### Variable importance
+
+```{r , include=TRUE, results= 'markup', message=FALSE}
+lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
+```
+
+```{r}
+help <- lapply(fit.model.all, function(x) varImp(x)) 
+help <- lapply(help, function(x) x[[1]])
+help <- do.call("rbind", help)
+
+importance <- help %>% rownames_to_column() %>% 
+  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
+         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
+  rename(importance = Overall) %>% 
+  relocate(importance, .after = year) %>% 
+  dplyr::select(-rowname)
+
+importance %>% ggplot() +
+  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
+  scale_color_discrete(name = "Features") +
+  #scale_linetype_manual(name = "Website density regressions111") 
+  ylab("Importance") + 
+  ggtitle("Feature importance") +
+  theme_minimal() +
+  theme(axis.title.x=element_blank(),
+        plot.title = element_text(hjust = 0.5))
+
+```
+
+### test
+
+```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
+
+# growth or n
+
+# for a reference point
+df %>% group_by(year) %>%
+  summarise(min = min(n), max=max(n),
+            mean = mean(n), median = median(n)) %>%
+  round(2) %>% kable()
+```
+
+```{r}
+## Test on t + 2
+
+# It will create: 2001 (train on 2000-2001) - 2011 (train on 2010)
+# I need: 2001 (train on 2000-2001) - 2009 (train on 2008-2009)
+
+# use the model trainned in year t to predict io of year t + 1
+#pred.by.year.all <- data.frame(matrix(NA,nrow=336, ncol = 0))
+for (i in regions){
+  for (j in regions[!regions == i]){
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
+    pred <- as.data.frame(pred)
+    rownames(pred) <- c()
+    assign(paste0("region.predict.model.", i, ".on.", j), pred)
+    #pred.by.year.all <- rbind(pred.by.year.all, pred)
+  }
+}
+
+df.predictions <- data.frame()
+for (i in regions){
+  for (j in regions[!regions == i]){
+    d <- paste0("region.predict.model.", i, ".on.", j)
+    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
+    d <- d[,c(5, 2, 4, 6, 1)]
+    colnames(d)[5] <- "predictions"
+    d <- d %>% mutate(trained.on = i)
+    df.predictions <- rbind(d, df.predictions)
+  }
+}
+
+df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))
+
+# pred.by.year to list by year
+pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
+
+# calculate metrics for every year
+rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
+                                                                     obs = x$n))  # CHANGE n WITH growth
+rf.year.all.metrics %>% as.data.frame() %>%
+  round(2) %>%
+  kable()
+
+# there are 2 columns named year, so I drop one
+pred.by.region.all$year <- NULL
+
+ggplot(data = df.predictions, aes(x = n, y = predictions)) +
+  geom_point(colour = "blue") +
+  geom_abline(intercept = 0, slope = 1, colour = "red") +
+  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
+  facet_wrap(~ test.train, ncol = 2) +
+  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
+  ggtitle("Predicted vs. actual trade by year")
+```

### 2023-11-13 src/la_RF.Rmd
rf

@@
 ---
-title: "Random Forests"
+title: "Random Forests, LAs"
 date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 
-```{r}
+```{r, eval=F}
 #spatial weight matrix
@@
 
-```{r}
+```{r, eval=F}
 #install.packages('SDPDmod')
@@
 sf_use_s2(FALSE)
+path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
+la.sf <- st_read(path.geo)
+
 help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
@@
 
+sf_use_s2(FALSE)
+la.c <- st_centroid(la.sf)
+kn <- knearneigh(la.c, k = 5)
+knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
 knn.l <- nb2listw(knn)
@@
                    number = 10,
-                   #seeds = seeds,
+                   seeds = seeds,
                    allowParallel = T,
@@
                      data = df %>% filter(RGN21CD==i), trControl = tc,
-                     method = "xgbTree", # rf "rpart", # rf lm
+                     method = "rf", # xgbTree "rpart", # rf lm
                      na.action = na.omit,
@@
 
-# It will create: 2001 (train on 2000-2001) - 2011 (train on 2010)
-# I need: 2001 (train on 2000-2001) - 2009 (train on 2008-2009)
+# train on one region, test on all the rest
 
-# use the model trainned in year t to predict io of year t + 1
-#pred.by.year.all <- data.frame(matrix(NA,nrow=336, ncol = 0))
+df.predictions <- data.frame()
 for (i in regions){
@@
     assign(paste0("region.predict.model.", i, ".on.", j), pred)
-    #pred.by.year.all <- rbind(pred.by.year.all, pred)
-  }
-}
 
-df.predictions <- data.frame()
-for (i in regions){
-  for (j in regions[!regions == i]){
     d <- paste0("region.predict.model.", i, ".on.", j)
@@
 
-# pred.by.year to list by year
+# split to list by region pair
 pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
 
-# calculate metrics for every year
+# calculate metrics for every region pair 
 rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                      obs = x$n))  # CHANGE n WITH growth
-rf.year.all.metrics %>% as.data.frame() %>%
-  round(2) %>%
-  kable()
 
-# there are 2 columns named year, so I drop one
-pred.by.region.all$year <- NULL
+path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup) %>% 
+  dplyr::select(RGN21CD, RGN21NM) %>% 
+  distinct() %>% 
+  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
 
+rf.year.all.metrics %>% 
+  as.data.frame() %>%
+  add_rownames(var = "metrics") %>% 
+  pivot_longer(!metrics, names_to = "train.test") %>% 
+  pivot_wider(names_from = metrics, values_from = value) %>% 
+  separate(train.test, c("train", "test"), remove = F) %>% 
+  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
+  rename(train.region=RGN21NM) %>% 
+  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
+  rename(test.region=RGN21NM) %>% 
+  arrange(Rsquared)
+  #summarise(r = range(Rsquared))
+  #round(2) %>%
+  #kable()
+```
+
+### test plots
+
+```{r}
+#plots
 ggplot(data = df.predictions, aes(x = n, y = predictions)) +

### 2023-11-13 src/oa_RF.Rmd
rf

--- 2023-04-06 src/level_lisa_oa.Rmd
@@
 ---
-title: "diffusion, websites with 1 pc"
+title: "Random Forests", OAs
 date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 library(rgeos)
+library(sf)
 library(spdep)
-library(tmap)
-library(maptools)
+#library(tmap)
+#library(maptools)
 library(knitr)
@@
 library(broom)
+#library(foreach)
+library(doParallel)
+library(raster)
 library(plm)
-library(lmtest)
+#library(lmtest)
 
@@
 # build oa for GB
-oa.gb <- spRbind(oa.ew, oa.sc)
+oa.gb <- maptools::spRbind(oa.ew, oa.sc)
 rm(oa.ew, oa.sc)
@@
 gb <- spTransform(gb, CRS("+init=epsg:4326"))
+```
 
+## df for RF
+
+```{r}
+# London = Islington 
+oa.sf <- st_as_sf(oa.gb)
+
+help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
+  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)
+
+help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
+  mutate(cities = c("Leicester",
+                    "Bristol",
+                    "Manchester",
+                    "Liverpool",
+                    "Sheffield",
+                    "Birmingham",
+                    "Leeds",
+                    "London",
+                    "Edinburgh",
+                    "Glasgow"))
+
+dist$dist.city.name <- names(dist)[apply(dist[,1:9], MARGIN = 1, FUN = which.min)]
+help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
+  rename(nearest.city.LAD21CD = LAD21CD) %>% 
+  dplyr::select(ladcd, nearest.city.LAD21CD)
+
+sf_use_s2(FALSE)
+la.c <- st_centroid(la.sf)
+kn <- knearneigh(la.c, k = 5)
+knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
+knn.l <- nb2listw(knn)
+
+df <- df %>% drop_na() %>% group_by(year) %>% 
+  mutate(n.slag = lag.listw(knn.l, n)) %>% 
+  group_by(ladcd) %>% 
+  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
+         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
+  left_join(help2, by = c("ladcd" = "ladcd"))
+
+df <- df %>% left_join(df %>% dplyr::select(ladcd,year, n) %>% 
+                         rename(n.nearest.city = n), 
+                       by = c("nearest.city.LAD21CD" = "ladcd",
+                              "year" = "year")) %>% 
+  dplyr::select(-c(LAD21NMW, OBJECTID, BNG_E, BNG_N, LONG, LAT, SHAPE_Length)) %>% 
+  relocate(LAD21NM, .after = ladcd)
+
+path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup) %>% dplyr::select(LAD21CD, RGN21CD)
+
+df <- df %>% left_join(lookup.region, by = c("ladcd" = "LAD21CD")) %>% 
+  relocate(RGN21CD, .after=year)
+
+# North/South
+#lookup.region %>% distinct(RGN21NM, .keep_all = T)
+south <- c("E12000007", "E12000008", "E12000009" )
+df <- df %>% mutate(south = ifelse(RGN21CD %in% south, 1, 0))
+
+# For growth:
+#df <- df %>% mutate(growth = log(n/n.lag))
+
+sapply(df, function(x) sum(is.na(x)))
 ```
@@
 ```{r}
+
+sf_use_s2(FALSE)
+la.c <- st_centroid(la.sf)
+kn <- knearneigh(la.c, k = 5)
+knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
+knn.l <- nb2listw(knn)
 
@@
 ```
-
-## The below are unfinished
-
-Website density regressions to the nearest retail centre with spatial lag
-
-```{r}
-library(spatialreg)
-
-model <- lagsarlm(log(density+0.1) ~ log(dist.retail+0.1), data = df[df$year==t,], 
-                  listw, tol.solve = 1e-30, zero.policy = T)
-
-dens.dist.retail <- data.frame()
-for (t in 1996:2012){
-  model <- lm(log(density) ~ log(dist.retail+0.1), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-}
-names(dens.dist.retail)[1] <- "rsquared"
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared))
-
-```
-
-## HH clusters
-
-```{r}
-hh <- mun_merge_new@data %>% filter(quadrant==1) %>% 
-  select(id, quadrant) %>% 
-  left_join(df[df$year==2000,], by =c("id" = "oa11cd")) %>% 
-  select(id, year, n, area, density, London, dist, dist.retail)
-
-hh.London <- lm(log(density+0.001) ~ log(London), data = hh)
-summary(hh.London)
-hh.dist <- lm(log(density+0.001) ~ log(dist), data = hh)
-summary(hh.dist)
-hh.retail <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = hh)
-summary(hh.retail)
-
-data.folder <- "/hdd/internet_archive/archive/data/"
-data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
-#Created by domain.R, which uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df9610 <- read_csv(data.path9610) %>% 
-  filter(V1.domain == n,
-         year == 1999) %>% 
-  left_join(lookup, by = "pc", suffix =c("",""))
-
-hh %>% left_join(df9610, by =c("id"="oa11cd"))
-
-```
-
-Spatial regression
-
-```{r}
-library(splm)
-m <- spml(log(n)~lag(log(n)), data = df, listw = listw, lag = TRUE, spatial.error = "b", 
-          model = "within", effect = "individual", method = "eigen", na.action = na.fail)
-
-```

### 2023-11-14 src/oa_RF.Rmd
oa.rf

@@
 
+# Complete panel
+df <- df %>% filter(!is.na(oa11cd)) %>% 
+  complete(oa11cd, year) %>% 
+  replace(is.na(.), 0)
+
 # tests for Scotland
@@
 
+## Distances to cities and retail centres
+
+First set up the destinations. Origins are all the OA.
+
+Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
+Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)
+
+```{r}
+# oa centroids
+# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
+# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
+#   mutate(id = row.names())
+oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
+  as.data.frame() %>% 
+  add_rownames(var = "id")
+
+# cities
+city.names <- c("London, UK", 
+"Birmingham, UK",
+"Glasgow, UK",
+"Liverpool, UK",
+"Bristol, UK",
+"Manchester, UK",
+"Sheffield, UK",
+"Leeds, UK",
+"Edinburgh, UK",
+"Leicester, UK")
+
+cities <- geo(city.names, no_query = F, method = "arcgis")
+
+# retail centres
+geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
+retail <-readOGR(geo.path) 
+#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
+retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
+  as.data.frame() %>% 
+  add_rownames(var = "id") #%>%
+
+retail.major.cetres <- retail.major.cetres.help %>% 
+  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
+  st_transform(4326) %>%
+  st_coordinates() %>%
+  as_tibble() %>% 
+  bind_cols(retail.major.cetres.help$id) %>% 
+  rename(id = '...3') %>% 
+  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
+```
+
+Distance to cities
+
+```{r}
+dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist <- round((dist/1000),2) %>% 
+  as_tibble()
+
+city.names <- city.names %>% stringr::str_remove(", UK")
+names(dist) <- city.names
+
+dist <- dist %>% bind_cols(oa.gb.c$id) %>% 
+  rename(oa11cd = last_col()) %>% 
+  relocate(oa11cd)
+
+# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
+# dist$distMet <- apply(dist[,2:11], 1, min)
+
+dist <- transform(dist, dist = do.call(pmin, dist[-1]))
+
+# NOT USED AS IT SCREWS UP THE REGRESSIONS
+# Complete panel, N = 3871903, 17 year * 227759 OA
+# df <- df %>% bind_rows(dist %>% dplyr::select(oa11cd)) %>%
+#   mutate(year = ifelse(is.na(year),1000, year),
+#          country = substring(oa11cd, 1, 1)) %>%
+#   filter(country != "L" & country != "M" & country != "N") %>%
+#   filter(!is.na(country)) %>%
+#   complete(year, oa11cd, fill = list(n = 0)) %>%
+#   filter(year!=1000,
+#          !is.na(oa11cd)) %>%
+#   dplyr::select(-country)
+
+# tests
+# test$country <- substring(test$oa11cd, 1, 1)
+# unique(test$country)
+# length(unique(test$oa11cd))
+
+# calculate area
+oa.gb$area <- area(oa.gb) #/ 1000000
+
+# Join with distance and area
+df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
+  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
+  relocate(area, .after = n) %>% 
+  mutate(area = area / 1000000,
+         density = n / area) %>% 
+  relocate(density, .after = area)
+
+sapply(df, function(x) sum(is.na(x)))
+```
+
+Distance to retail centres
+
+```{r}
+dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
+dist.retail <- round((dist.retail/1000),2) %>% 
+  as_tibble()
+
+# retail.names <- retail.major.cetres$RC_Name
+# names(dist.retail) <- retail.names
+
+dist.retail <- dist.retail %>% bind_cols(oa.gb.c$id) %>% 
+  rename(oa11cd = last_col()) %>% 
+  relocate(oa11cd)
+
+# Minimum distance 
+dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
+dist.retail <- dist.retail %>% dplyr::select(oa11cd, dist.retail)
+
+# Join with complete panel
+df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd"))
+
+sapply(df, function(x) sum(is.na(x)))
+```
+
 ## df for RF
@@
 ```{r}
+## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
+df <- df %>% filter(!is.na(area))
+
 # London = Islington 
 oa.sf <- st_as_sf(oa.gb)
+cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
+cities.sf <- st_set_crs(cities.sf, 4326)
 
-help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
-  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)
+help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
+  left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
 
-help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
+sf_use_s2(FALSE)
+help1 <- st_filter( oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
   mutate(cities = c("Leicester",
@@
 
-dist$dist.city.name <- names(dist)[apply(dist[,1:9], MARGIN = 1, FUN = which.min)]
+# too slow
+#dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
+
+dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
+
 help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
-  rename(nearest.city.LAD21CD = LAD21CD) %>% 
-  dplyr::select(ladcd, nearest.city.LAD21CD)
+  rename(nearest.city.oa11cd = id) %>% 
+  dplyr::select(oa11cd, nearest.city.oa11cd)
 
 sf_use_s2(FALSE)
-la.c <- st_centroid(la.sf)
-kn <- knearneigh(la.c, k = 5)
+oa.c <- st_centroid(oa.sf)
+kn <- knearneigh(oa.c, k = 5)
 knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
@@
 
-df <- df %>% drop_na() %>% group_by(year) %>% 
+df <- df %>% drop_na() %>% 
+  group_by(year) %>% 
   mutate(n.slag = lag.listw(knn.l, n)) %>% 
-  group_by(ladcd) %>% 
+  group_by(oa11cd) %>% 
   mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
          n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
-  left_join(help2, by = c("ladcd" = "ladcd"))
+  left_join(help2, by = c("oa11cd" = "oa11cd"))
 
@@
 
-```
-
-## Distances to cities and retail centres
-
-First set up the destinations. Origins are all the OA.
-
-Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
-Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)
-
-```{r}
-# oa centroids
-# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
-# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
-#   mutate(id = row.names())
-oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id")
-
-# cities
-city.names <- c("London, UK", 
-"Birmingham, UK",
-"Glasgow, UK",
-"Liverpool, UK",
-"Bristol, UK",
-"Manchester, UK",
-"Sheffield, UK",
-"Leeds, UK",
-"Edinburgh, UK",
-"Leicester, UK")
-
-cities <- geo(city.names, no_query = F, method = "arcgis")
-
-# retail centres
-geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
-retail <-readOGR(geo.path) 
-#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
-retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id") #%>%
-
-retail.major.cetres <- retail.major.cetres.help %>% 
-  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
-  st_transform(4326) %>%
-  st_coordinates() %>%
-  as_tibble() %>% 
-  bind_cols(retail.major.cetres.help$id) %>% 
-  rename(id = '...3') %>% 
-  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
-```
-
-Distance to cities
-
-```{r}
-dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist <- round((dist/1000),2) %>% 
-  as_tibble()
-
-city.names <- city.names %>% stringr::str_remove(", UK")
-names(dist) <- city.names
-
-dist <- dist %>% bind_cols(oa.gb.c$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
-# dist$distMet <- apply(dist[,2:11], 1, min)
-
-dist <- transform(dist, dist = do.call(pmin, dist[-1]))
-
-# NOT USED AS IT SCREWS UP THE REGRESSIONS
-# Complete panel, N = 3871903, 17 year * 227759 OA
-# df <- df %>% bind_rows(dist %>% dplyr::select(oa11cd)) %>%
-#   mutate(year = ifelse(is.na(year),1000, year),
-#          country = substring(oa11cd, 1, 1)) %>%
-#   filter(country != "L" & country != "M" & country != "N") %>%
-#   filter(!is.na(country)) %>%
-#   complete(year, oa11cd, fill = list(n = 0)) %>%
-#   filter(year!=1000,
-#          !is.na(oa11cd)) %>%
-#   dplyr::select(-country)
-
-# tests
-# test$country <- substring(test$oa11cd, 1, 1)
-# unique(test$country)
-# length(unique(test$oa11cd))
-
-# calculate area
-oa.gb$area <- area(oa.gb) #/ 1000000
-
-# Join with distance and area
-df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
-  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
-  relocate(area, .after = n) %>% 
-  mutate(area = area / 1000000,
-         density = n / area) %>% 
-  relocate(density, .after = area)
-sapply(df, function(x) sum(is.na(x)))
-```
-
-Distance to retail centres
-
-```{r}
-dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail <- round((dist.retail/1000),2) %>% 
-  as_tibble()
-
-# retail.names <- retail.major.cetres$RC_Name
-# names(dist.retail) <- retail.names
-
-dist.retail <- dist.retail %>% bind_cols(oa.gb.c$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# Minimum distance 
-dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
-dist.retail <- dist.retail %>% dplyr::select(oa11cd, dist.retail)
-
-# Join with complete panel
-df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd"))
-
-sapply(df, function(x) sum(is.na(x)))
 ```

### 2023-11-14 src/oa_RF.Rmd
oa.rf

@@
 
-df <- df %>% drop_na() %>% 
+df <- df %>% #drop_na() %>% 
   group_by(year) %>% 

### 2023-11-17 src/la_RF.Rmd
oa.rf
ranger

@@
 #library(lmtest)
+library(caret)
+library(randomForest)
 
@@
 
+# Complete panel
 ## REPLACE WITH 0.1 INSTEAD OF OF 0 FOR THE GROWTH
@@
 
+#########CHECK 1:9################
+
 dist$dist.city.name <- names(dist)[apply(dist[,1:9], MARGIN = 1, FUN = which.min)]
@@
 
+##############ADD EAST, 4 REGIONS IN TOTAL###########
+
 # North/South
@@
 ```{r}
-library(caret)
-library(randomForest)
-
 # CV
@@
   model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
-                      n.slag + n.l.slag + n.nearest.city + year, #n.lag +
+                      n.slag + n.l.slag + n.nearest.city + year + n.lag,
                      data = df %>% filter(year==t | year==tplus1), #%>% 
@@
                      trControl = tc,
-                     method = "rf", #"rpart", # rf lm
+                     method = "ranger", #"rpart", # rf lm
                      na.action = na.omit,
-                     preProc = c("center", "scale"),
-                     importance = TRUE)
+                     #preProc = c("center", "scale"),
+                     importance = 'impurity') # TRUE for rf
+    assign(paste0("model.all",as.character(t + 1)), model.all)
 }

### 2023-11-17 src/oa_RF.Rmd
ranger

--- 2023-11-14 src/oa_RF.Rmd
@@
 
-# Complete panel
+# Partially complete panel as not all OAs have a website in at least ine year
 df <- df %>% filter(!is.na(oa11cd)) %>% 
@@
 gb <- spTransform(gb, CRS("+init=epsg:4326"))
+```
+
+## Full panel
+
+```{r}
+oa.sf <- st_as_sf(oa.gb)
+
+oa.full <- replicate(17, oa.sf %>% as_tibble(), simplify = FALSE) %>%
+  bind_rows(.id = 'id_') %>% 
+  rename(year = id_) %>% 
+  mutate(year = as.numeric(year) + 1995) %>% 
+  dplyr::select(id, year)
+
+df <- oa.full %>% left_join(df, by = c("id" = "oa11cd", "year" = "year")) %>% 
+  replace(is.na(.), 0)
+
+sapply(df, function(x) sum(is.na(x)))
+rm(oa.full)
 ```
@@
 # Join with distance and area
-df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
+df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
+  rename(oa11cd=id) %>% 
   left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
@@
 ## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
-df <- df %>% filter(!is.na(area))
+#df <- df %>% filter(!is.na(area))
 
 # London = Islington 
-oa.sf <- st_as_sf(oa.gb)
 cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
@@
 
-help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
-  left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
+# help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
+#   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
 
@@
 
+# This efficiently returns the name of the column with the shortest distance
 dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
@@
 
+#####save this#####
 sf_use_s2(FALSE)
@@
 knn.l <- nb2listw(knn)
+##################
 
-df <- df %>% drop_na() %>% 
+df <- df %>% 
   group_by(year) %>% 
@@
 
-df <- df %>% left_join(df %>% dplyr::select(ladcd,year, n) %>% 
+df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
                          rename(n.nearest.city = n), 
-                       by = c("nearest.city.LAD21CD" = "ladcd",
-                              "year" = "year")) %>% 
-  dplyr::select(-c(LAD21NMW, OBJECTID, BNG_E, BNG_N, LONG, LAT, SHAPE_Length)) %>% 
-  relocate(LAD21NM, .after = ladcd)
+                       by = c("nearest.city.oa11cd" = "oa11cd",
+                              "year" = "year"))
 
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% dplyr::select(LAD21CD, RGN21CD)
+# Lookup: OA to regions
+path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup) %>% dplyr::select(OA11CD, RGN11CD, RGN11NM)
 
-df <- df %>% left_join(lookup.region, by = c("ladcd" = "LAD21CD")) %>% 
-  relocate(RGN21CD, .after=year)
+df <- df %>% left_join(lookup.region, by = c("oa11cd" = "OA11CD")) %>% 
+  relocate(c(RGN11NM, RGN11CD), .after=year) %>% 
+  mutate(help = substring(oa11cd, 1, 1),
+         RGN11CD = ifelse(help=="E", RGN11CD,
+                         ifelse(help=="W", "Wales",
+                                ifelse(help == "S", "Scotland", NA))),
+         RGN11NM = ifelse(help=="E", RGN11NM,
+                         ifelse(help=="W", "Wales",
+                                ifelse(help == "S", "Scotland", NA)))) %>% 
+  dplyr::select(-help)
+
+# df %>% ungroup() %>% distinct(RGN11NM)
+# sapply(df, function(x) sum(is.na(x)))
 
@@
 #lookup.region %>% distinct(RGN21NM, .keep_all = T)
-south <- c("E12000007", "E12000008", "E12000009" )
-df <- df %>% mutate(south = ifelse(RGN21CD %in% south, 1, 0))
+south <- c("London", "South West", "South West", "East of England")
+df <- df %>% mutate(south = ifelse(RGN11NM %in% south, 1, 0))
 
@@
 
-## lisa maps
+## RF
+
+### train on all but one regions
 
 ```{r}
+out.path <- paste0(path, "/data/temp/oa_rf.RData")
+# save.image(file = out.path)
+start.time <- Sys.time()
+load(out.path)
+end.time <- Sys.time()
+time.taken <- round(end.time - start.time,2)
+time.taken
 
-sf_use_s2(FALSE)
-la.c <- st_centroid(la.sf)
-kn <- knearneigh(la.c, k = 5)
-knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
-knn.l <- nb2listw(knn)
+unregister_dopar <- function() {
+  env <- foreach:::.foreachGlobals
+  rm(list=ls(name=env), pos=env)
+}
+unregister_dopar()
 
-path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", n, "_oa") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_oa")  # for multiple PC
+# CV
+tc <- trainControl(method = "cv",
+                   number = 10, #10,
+                   allowParallel = T,
+                   savePredictions = 'final')
 
-morani <- data.frame()
+# London region: E12000007
 
-#spatial weight matrix
-sf_use_s2(FALSE)
-nb <- poly2nb(oa.gb)
-listw <- nb2listw(nb, style = "W", zero.policy = T)
+# detectCores()
+# cl <- makePSOCKcluster(14)
+# registerDoParallel(cl)
 
-for (t in 1996:2012){
-test <- oa.gb
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
-  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
-         year = ifelse(is.na(year), t, year))
+# train
+start.time <- Sys.time()
 
-#spatial weight matrix
-#moved out of the loop
+set.seed(71)
+m.test <- train(n ~ density + area + dist + dist.retail + London +
+                  n.slag + year + n.nearest.city, #n.l.slag
+                data = df %>% filter(RGN11CD=="E12000008"),
+                trControl = tc,
+                method = "ranger", #"rf", 
+                importance = "impurity",  #TRUE,
+                na.action = na.omit)
 
-#Moran's I
-globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
-a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
-morani <- rbind(a, morani)
+#stopCluster(cl)
+end.time <- Sys.time()
+time.taken <- round(end.time - start.time,2)
+time.taken
 
-#LISA
-lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
-#lmoran
+summary(m.test)
+plot(m.test)
+m.test
+print(m.test)
 
-# centers the local Moran's around the mean
-lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
-lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
-
-# centers the variable of interest around its mean
-
-lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
-lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
-
-signif <- 0.05
-#lmoran
-
-lmoran <- lmoran%>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
-  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
-  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%
-
-#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
-# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
-#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))
-
-mun_merge_new<- merge(test, lmoran, by="id")
-
-# R p value map
-breaks = c(0, 1, 2, 3, 4, 5)
-
-cities <- maps::world.cities %>% 
-  filter(country.etc=="UK") %>% 
-  arrange(pop) %>% tail(10) 
-tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
-
-map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, 
-                                         palette = c("white","red","blue",
-                                                     rgb(0,0,1,alpha=0.4),
-                                                     rgb(1,0,0,alpha=0.4)), 
-                                         labels = c("Not significant",
-                                                    "High-High",
-                                                    "Low-Low",
-                                                    "Low-High",
-                                                    "High-Low"), 
-                                         title="LISA, websites per OA") + 
-  tm_legend(text.size = 1) +
-  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
-  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
-  #tm_borders(alpha=.1) +
-  tm_borders(lwd = 0) +
-  tm_shape(gb, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
-  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
-  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
-tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
-}
+varimp_mars <- varImp(m)
+plot(varimp_mars, main="Variable Importance")
 ```
 
-Moran I
+### test on the hold out regions
+
+Wales region is missing!!!
 
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/morani_oa.csv")
-morani %>% as_tibble() %>% 
-  rename('morani'=V1,
-                   p = V2,
-                   year = V3) %>% 
-  #kable() %>% 
-  write_csv(path.out)
+pred <- predict(m, df %>% ungroup() %>% filter(RGN21CD=="E12000008"), na.action = na.pass)
+length(pred)
+
+df %>% ungroup() %>% filter(RGN21CD=="E12000008") %>% filter(if_any(everything(), is.na))
+
+pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN21CD=="E12000008"))
+postResample(pred = pred$value, obs = pred$n) ## 0.99
 ```
 
-## Gini
+## RF rolling
+
+### train the model in a loop
 
 ```{r}
+# Train on t & t + 1
 
-# add OAs with 0s
-df.with0s <- df %>% mutate(country = substring(oa11cd, 1, 1)) %>%
-  filter(country != "L" & country != "M" & country != "N") %>%
-  filter(!is.na(country)) %>%
-  complete(year, oa11cd, fill = list(n = 0)) %>%
-  filter(year!=1000,
-         !is.na(oa11cd)) %>%
-  dplyr::select(-country)
+# Make the parallelisation reproducible using https://stackoverflow.com/questions/13403427/fully-reproducible-parallel-models-using-caret. 
+# See also https://stackoverflow.com/questions/27944558/set-seed-parallel-random-forest-in-caret-for-reproducible-result
 
-gini.df <- data.frame()
+#length is = (n_repeats*nresampling)+1
+seeds <- vector(mode = "list", length = 11)
 
-for (t in 1996:2012){
-    a <- df.with0s %>% filter(year==t)
-  b <- as.data.frame(gini(a$n))
-  b$year = t
-  gini.df <- rbind(gini.df, b)
+#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South
+
+#for the last model
+seeds[[11]]<-sample.int(1000, 1)
+
+# CV
+tc <- trainControl(method = "cv",
+                   number = 10,
+                   seeds = seeds,
+                   allowParallel = T,
+                   savePredictions = 'final')
+
+detectCores()
+cl <- makePSOCKcluster(14)
+registerDoParallel(cl)
+
+# run the model for every 2 years
+start.time <- Sys.time()
+years <- 1997:2012
+
+for (t in years){
+  tplus1 <- t + 1
+  set.seed(71)
+  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
+                      n.slag + n.l.slag + n.nearest.city + year, #n.lag +
+                     data = df %>% filter(year==t | year==tplus1), #%>% 
+                       # mutate(growth = ifelse(growth == -Inf | growth == Inf, NA, growth)) %>% 
+                       # mutate(growth = ifelse(is.na(growth), 0, growth)), 
+                     #%>% mutate(growth = ifelse(growth== -Inf | growth == Inf, NA, growth))
+                     trControl = tc,
+                     method = "rf", #"rpart", # rf lm
+                     na.action = na.omit,
+                     preProc = c("center", "scale"),
+                     importance = TRUE)
 }
 
-path.out <- paste0(path, "/outputs/lisa/gini_oa.csv")
-gini.df %>% write_csv(path.out)
+stopCluster(cl)
+end.time <- Sys.time()
+time.taken <- round(end.time - start.time,2)
+time.taken
+
+# df %>% filter(growth == Inf) 
+# log(n/n.lag)
+
+# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
+# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
 ```
 
-## Getis-Ord
+### Resamples for plots
 
-```{r, eval = F}
+```{r}
+# create a list of the above model outputs
+fit.model.all <- mget(ls(pattern = "^model.all\\d"))
 
-t= 1998
-local_g <- localG(OA.Census$Qualification, nb_lw)
-local_g <- localG(df[df$year==t,]$n, listw)
+# resamples
+fit.model.all.res <- resamples(fit.model.all)
+summary(fit.model.all.res)
 
-local_g <- cbind(OA.Census, as.matrix(local_g))
-names(local_g)[6] <- "gstat"
+# plot metrics
+
+# TODO: fix the year labels
+
+scales <- list(x=list(relation="free"), y=list(relation="free"))
+bwplot(fit.model.all.res, scales=scales)
+
+# TODO: fix the year labels
+```
+
+### Variable importance
+
+```{r , include=TRUE, results= 'markup', message=FALSE}
+lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
+```
+
+```{r}
+help <- lapply(fit.model.all, function(x) varImp(x)) 
+help <- lapply(help, function(x) x[[1]])
+help <- do.call("rbind", help)
+
+importance <- help %>% rownames_to_column() %>% 
+  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
+         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
+  rename(importance = Overall) %>% 
+  relocate(importance, .after = year) %>% 
+  dplyr::select(-rowname)
+
+importance %>% ggplot() +
+  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
+  scale_color_discrete(name = "Features") +
+  #scale_linetype_manual(name = "Website density regressions111") 
+  ylab("Importance") + 
+  ggtitle("Feature importance") +
+  theme_minimal() +
+  theme(axis.title.x=element_blank(),
+        plot.title = element_text(hjust = 0.5))
 
@@
 
-## Panel regressions
+### test
+
+```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
+
+# growth or n
+
+# for a reference point
+df %>% group_by(year) %>%
+  summarise(min = min(n), max=max(n),
+            mean = mean(n), median = median(n)) %>%
+  round(2) %>% kable()
+```
 
 ```{r}
-# London
-model <- plm(log(n+0.001) ~ log(London+.001)*as.factor(year), 
-             data = df,
-             index = c("oa11cd", "year"),
-             model = "within",
-             effect = "twoways")
-summary(model)
+## Test on t + 2
 
-est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
-int <- confint(est)
+# It will create: 2001 (train on 2000-2001) - 2011 (train on 2010)
+# I need: 2001 (train on 2000-2001) - 2009 (train on 2008-2009)
 
-out.model <- bind_cols(tidy(est), 
-                       as_tibble(int)) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(year = 1997:2012) %>% 
-  rename(conf.low = '2.5 %',
-         conf.high = '97.5 %')
-
-path.out <- paste0(path,"/outputs/lisa/b_panel_dist_london_oa.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                    width = 0.2,size  = 1,
-                    position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
-  ggtitle("Regression coefficients: distance to London on website counts, OA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-
-# Cities
-model <- plm(log(n+0.001) ~ log(dist+.001)*as.factor(year), 
-             data = df,
-             index = c("oa11cd", "year"),
-             model = "within",
-             effect = "twoways")
-summary(model)
-
-est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
-int <- confint(est)
-
-out.model <- bind_cols(tidy(est), 
-                       as_tibble(int)) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(year = 1997:2012) %>% 
-  rename(conf.low = '2.5 %',
-         conf.high = '97.5 %')
-
-path.out <- paste0(path,"/outputs/lisa/b_panel_dist_cities_oa.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                    width = 0.2,size  = 1,
-                    position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("Interaction coefficienst: distance*year, robust st. errors") + xlab("") + 
-  ggtitle("Regression coefficients: distance to Cities on website counts, OA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-
-# Retail centres
-model <- plm(log(n+0.001) ~ log(dist.retail+.001)*as.factor(year), 
-             data = df,
-             index = c("oa11cd", "year"),
-             model = "within",
-             effect = "twoways")
-summary(model)
-
-est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
-int <- confint(est)
-
-out.model <- bind_cols(tidy(est), 
-                       as_tibble(int)) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(year = 1997:2012) %>% 
-  rename(conf.low = '2.5 %',
-         conf.high = '97.5 %')
-
-path.out <- paste0(path,"/outputs/lisa/b_panel_dist_retail_oa.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                    width = 0.2,size  = 1,
-                    position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
-  ggtitle("Regression coefficients: distance to Retail Centres on website counts, OA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-```
-
-## Website density regressions to London, cross-section
-
-```{r}
-dens.London <- data.frame()
-out.model <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(London), data = df[df$year==t,]) 
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.London <- rbind(r2, dens.London)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model <- rbind(out.model_, out.model)
+# use the model trainned in year t to predict io of year t + 1
+years <- (1998:2011)
+pred.by.year.all <- data.frame(matrix(NA,nrow=374, ncol = 0))
+for (t in years){
+  t_plus_1 <- t + 1
+  pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",t)], df[df$year==t_plus_1,])
+  pred <- as.data.frame(pred)
+  rownames(pred) <- c()
+  pred.by.year.all <- cbind(pred.by.year.all, pred)
 }
 
-names(dens.London)[1] <- "rsquared"
+#change column names to match with total
+names(pred.by.year.all) <- 1999:2012
 
-path.out <- paste0(path,"/outputs/lisa/rsq_dist_london_oa.png")
-ggplot(dens.London) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to London on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
+# wide to long
+pred.by.year.all <- gather(pred.by.year.all,key = year,value = predictions,1:14)
 
-out.model <- out.model %>% 
-  mutate(year = 1996:2012)
+# drop 2000 and select variables form total
+df.no1997_1 <- df %>% 
+  filter(year!=1996 & year!=1997 & year!=1998) %>% 
+  dplyr::select(ladcd, year, n) # CHANGE n WITH growth
 
-path.out <- paste0(path,"/outputs/lisa/b_dist_london_oa.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-#ggplot(out.model, aes(x=year, y = estimate)) +  
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to London on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
+# column bind prediction and data  
+pred.by.year.all <- cbind(pred.by.year.all, df.no1997_1 %>% arrange(year, ladcd)) #DOUBLE CHECK ARRANGE
 
-r2.all <- dens.London %>% 
-  rename(London = rsquared)
+# pred.by.year to list by year
+pred.by.year.all.list <- split(pred.by.year.all, pred.by.year.all$year)
 
-# Help for OA area size. I double checked the largest: E00102963
-# It is correct, this is its size.
-# ggplot(df, aes(x=area)) +
-#     geom_histogram(binwidth=.5, colour="black", fill="white")
-# 
-# df %>% filter(area>50)
+# calculate metrics for every year
+rf.year.all.metrics <- lapply(pred.by.year.all.list, function(x) postResample(pred = x$predictions,
+                                                                     obs = x$n))  # CHANGE n WITH growth
+rf.year.all.metrics %>% as.data.frame() %>%
+  round(2) %>%
+  kable()
+
+# there are 2 columns named year, so I drop one
+pred.by.year.all$year <- NULL
+
+ggplot(data = pred.by.year.all,aes(x = n, y = predictions)) +
+  geom_point(colour = "blue") +
+  geom_abline(intercept = 0, slope = 1, colour = "red") +
+  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
+  facet_wrap(~ year,ncol = 2) +
+  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
+  ggtitle("Predicted vs. actual trade by year")
 ```
-
-## Website density regressions to the nearest city, cross-section
-
-```{r}
-dens.distMet <- data.frame()
-out.modelMet <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.distMet <- rbind(r2, dens.distMet)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.modelMet <- rbind(out.model_, out.modelMet)
-}
-
-names(dens.distMet)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/lisa/rsq_dist_city_oa.png")
-ggplot(dens.distMet) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest city on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-out.modelMet <- out.modelMet %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/lisa/b_dist_city_oa.png")
-ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-r2.all <- r2.all %>% left_join(dens.distMet) %>% 
-  rename(cities = rsquared)
-```
-
-## Website density regressions to the nearest retail centre, cross-section
-
-```{r}
-dens.dist.retail <- data.frame()
-out.model.retail <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = df[df$year==t,]) 
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model.retail <- rbind(out.model_, out.model.retail)
-}
-
-names(dens.dist.retail)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/lisa/rsq_dist_centre_oa.png")
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest retail centre on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-out.model.retail <- out.model.retail %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/lisa/b_dist_centre_oa.png")
-ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest retail centre on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-# library(pscl)
-# m1 <- zeroinfl((n) ~ log(dist.retail+0.001),
-#   data = df[df$year==t,], dist = "negbin", EM = TRUE)
-# summary(m1)
-# 
-# df %>% filter(year == 1996) %>% 
-#   group_by(group = cut(n, breaks = c(0, 1, 2, 10, 100, 1000, 10000), include.lowest = T, right = F)) %>%  #seq(0, max(n), 100))) %>%
-#   summarise(n = sum(n))
-```
-All R^2 dataframe
-
-```{r}
-path.out <- paste0(path, "/outputs/lisa/r2_all_oa.csv")
-
-r2.all <- dens.London %>% 
-  rename(London = rsquared) %>% 
-  left_join(dens.distMet) %>% 
-  rename(cities = rsquared) %>% 
-  left_join(dens.dist.retail) %>% 
-  rename(retail = rsquared) %>% 
-  relocate(year, .before = London) %>% 
-  arrange(year)
-
-write_csv(r2.all, path.out)
-```

### 2023-11-20 src/la_RF.Rmd
CAST

@@
 library(randomForest)
+library(CAST)
 
@@
 
-#########CHECK 1:9################
-
-dist$dist.city.name <- names(dist)[apply(dist[,1:9], MARGIN = 1, FUN = which.min)]
+# dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
+# help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
+#   rename(nearest.city.LAD21CD = LAD21CD) %>% 
+#   dplyr::select(ladcd, nearest.city.LAD21CD)
+
+# This efficiently returns the name of the column with the shortest distance
+dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
 help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
@@
 
-##############ADD EAST, 4 REGIONS IN TOTAL###########
-
 # North/South
 #lookup.region %>% distinct(RGN21NM, .keep_all = T)
-south <- c("E12000007", "E12000008", "E12000009" )
+south <- c("E12000006", "E12000007", "E12000008", "E12000009" )
 df <- df %>% mutate(south = ifelse(RGN21CD %in% south, 1, 0))
@@
 
-### train
+### train the model in a loop
 
@@
                      data = df %>% filter(RGN21CD==i), trControl = tc,
-                     method = "rf", # xgbTree "rpart", # rf lm
+                     method = "ranger", 
                      na.action = na.omit,
@@
 ```
+
+## RF spatial with CAST
+
+```{r}
+df <- df %>% mutate(help = substr(ladcd, start = 1, stop = 2)) %>% 
+  mutate(RGN21CD = ifelse(is.na(RGN21CD), help, RGN21CD)) %>% 
+  ungroup() %>% 
+  dplyr::select(-help)
+
+df <- df %>% filter(year > 1996)
+
+#df <- df %>% dplyr::select(-n.lag)
+```
+
+### train the model in all data
+
+```{r}
+indices <- CreateSpacetimeFolds(df, 
+                                spacevar = "ladcd",
+                                timevar = "year", 
+                                k = 10)
+
+#length is = (n_repeats*nresampling)+1
+seeds <- vector(mode = "list", length = 11)
+
+#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South
+
+#for the last model
+seeds[[11]]<-sample.int(1000, 1)
+
+# CV
+tc <- trainControl(method = "cv",
+                   number = 10,
+                   seeds = seeds,
+                   allowParallel = T,
+                   index = indices$index,
+                   savePredictions = 'final')
+
+detectCores()
+cl <- makePSOCKcluster(14)
+registerDoParallel(cl)
+
+start.time <- Sys.time()
+model.all <- train(n ~ area + dist + dist.retail + London + south + 
+                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
+                   data = df, 
+                   trControl = tc,
+                   method = "ranger",
+                   na.action = na.omit,
+                   #preProc = c("center", "scale"),
+                   importance = "impurity")
+
+stopCluster(cl)
+end.time <- Sys.time()
+time.taken <- round(end.time - start.time,2)
+time.taken
+
+# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
+#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
+#                                                         "year")], 
+#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
+#                     trControl = tc,
+#                     method = "ranger",
+#                     #na.action = na.omit,
+#                     #importance = "impurity",
+#                     metric="Rsquared",
+#                     verbose=FALSE)
+# ffsmodel_LLO
+# ffsmodel_LLO$selectedvars
+# plot_ffs(ffsmodel_LLO)
+```
+
+### train the model in a loop
+
+```{r}
+
+#length is = (n_repeats*nresampling)+1
+seeds <- vector(mode = "list", length = 11)
+
+#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South
+
+#for the last model
+seeds[[11]]<-sample.int(1000, 1)
+
+# CV
+tc <- trainControl(method = "cv",
+                   number = 10,
+                   seeds = seeds,
+                   allowParallel = T,
+                   index = indices$index,
+                   savePredictions = 'final')
+
+detectCores()
+cl <- makePSOCKcluster(14)
+registerDoParallel(cl)
+
+# train in every region
+start.time <- Sys.time()
+
+regions <- df %>% distinct(RGN21CD)
+regions <- as.vector(regions$RGN21CD)
+
+for (i in regions){
+
+  set.seed(71)
+  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD==i), 
+                                spacevar = "ladcd",
+                                timevar = "year", 
+                                k = 10)
+
+  set.seed(71)
+  model.all <- train(n ~ area + dist + dist.retail + London + south + 
+                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
+                   data = df %>% filter(RGN21CD==i), 
+                   trControl = tc,
+                   method = "ranger",
+                   na.action = na.omit,
+                   #preProc = c("center", "scale"),
+                   importance = "impurity")
+  assign(paste0("model.all",as.character(i)), model.all)
+}
+
+stopCluster(cl)
+end.time <- Sys.time()
+time.taken <- round(end.time - start.time,2)
+time.taken
+
+# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
+# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
+```
+
+### Resamples for plots
+
+```{r}
+# create a list of the above model outputs
+fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
+
+# resamples
+fit.model.all.res <- resamples(fit.model.all)
+summary(fit.model.all.res)
+
+# plot metrics
+
+# TODO: fix the year labels
+
+scales <- list(x=list(relation="free"), y=list(relation="free"))
+bwplot(fit.model.all.res, scales=scales)
+
+# TODO: fix the year labels
+```
+
+### Variable importance
+
+```{r , include=TRUE, results= 'markup', message=FALSE}
+lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
+```
+
+```{r}
+help <- lapply(fit.model.all, function(x) varImp(x)) 
+help <- lapply(help, function(x) x[[1]])
+help <- do.call("rbind", help)
+
+importance <- help %>% rownames_to_column() %>% 
+  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
+         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
+  rename(importance = Overall) %>% 
+  relocate(importance, .after = year) %>% 
+  dplyr::select(-rowname)
+
+importance %>% ggplot() +
+  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
+  scale_color_discrete(name = "Features") +
+  #scale_linetype_manual(name = "Website density regressions111") 
+  ylab("Importance") + 
+  ggtitle("Feature importance") +
+  theme_minimal() +
+  theme(axis.title.x=element_blank(),
+        plot.title = element_text(hjust = 0.5))
+
+```
+
+### test
+
+```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
+
+# growth or n
+
+# for a reference point
+df %>% group_by(year) %>%
+  summarise(min = min(n), max=max(n),
+            mean = mean(n), median = median(n)) %>%
+  round(2) %>% kable()
+```
+
+```{r}
+## Test on t + 2
+
+# train on one region, test on all the rest
+
+df.predictions <- data.frame()
+for (i in regions){
+  for (j in regions[!regions == i]){
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
+    pred <- as.data.frame(pred)
+    rownames(pred) <- c()
+    assign(paste0("region.predict.model.", i, ".on.", j), pred)
+
+    d <- paste0("region.predict.model.", i, ".on.", j)
+    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
+    d <- d[,c(5, 2, 4, 6, 1)]
+    colnames(d)[5] <- "predictions"
+    d <- d %>% mutate(trained.on = i)
+    df.predictions <- rbind(d, df.predictions)
+  }
+}
+
+df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))
+
+# split to list by region pair
+pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
+
+# calculate metrics for every region pair 
+rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
+                                                                     obs = x$n))  # CHANGE n WITH growth
+
+path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup) %>% 
+  dplyr::select(RGN21CD, RGN21NM) %>% 
+  distinct() %>% 
+  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
+
+rf.year.all.metrics %>% 
+  as.data.frame() %>%
+  rownames_to_column(var = "metrics") %>% 
+  pivot_longer(!metrics, names_to = "train.test") %>% 
+  pivot_wider(names_from = metrics, values_from = value) %>% 
+  separate(train.test, c("train", "test"), remove = F) %>% 
+  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
+  rename(train.region=RGN21NM) %>% 
+  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
+  rename(test.region=RGN21NM) %>% 
+  arrange(Rsquared) %>% 
+  dplyr::select(train.region, test.region, Rsquared) %>% 
+  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
+  geom_tile() +
+  coord_fixed() +
+  theme(legend.position = "none")
+  #summarise(r = range(Rsquared))
+  #round(2) %>%
+  #kable()
+```
+
+### test plots
+
+```{r}
+#plots
+ggplot(data = df.predictions, aes(x = n, y = predictions)) +
+  geom_point(colour = "blue") +
+  geom_abline(intercept = 0, slope = 1, colour = "red") +
+  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
+  facet_wrap(~ test.train, ncol = 2) +
+  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
+  ggtitle("Predicted vs. actual trade by year")
+```

### 2023-11-20 src/oa_RF.Rmd
CAST

@@
 sapply(df, function(x) sum(is.na(x)))
+
+# out.path <- paste0(path, "/data/temp/oa_rf.RData")
+# save.image(file = out.path)
+# load(out.path)
+
+# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
+# df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol", 
+#                         "Manchester", "Sheffield", "Leeds", "Edinburgh", 
+#                         "Leicester", "nearest.city.oa11cd")) %>% 
+#   write_csv(file = out.path)
 ```
@@
 ```{r}
-out.path <- paste0(path, "/data/temp/oa_rf.RData")
-# save.image(file = out.path)
-start.time <- Sys.time()
-load(out.path)
-end.time <- Sys.time()
-time.taken <- round(end.time - start.time,2)
-time.taken
+# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
+# df <- read_csv(out.path)
 
@@
 time.taken <- round(end.time - start.time,2)
-time.taken
+time.taken #1h
 
@@
 
-varimp_mars <- varImp(m)
+varimp_mars <- varImp(m.test)
 plot(varimp_mars, main="Variable Importance")
@@
 ```{r}
-pred <- predict(m, df %>% ungroup() %>% filter(RGN21CD=="E12000008"), na.action = na.pass)
+pred <- predict(m.test, df %>% ungroup() %>% filter(RGN11CD=="E12000006"), na.action = na.pass)
 length(pred)
 
-df %>% ungroup() %>% filter(RGN21CD=="E12000008") %>% filter(if_any(everything(), is.na))
+df %>% ungroup() %>% filter(RGN11CD=="E12000006") %>% filter(if_any(everything(), is.na))
 
-pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN21CD=="E12000008"))
+pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN11CD=="E12000006"))
 postResample(pred = pred$value, obs = pred$n) ## 0.99
@@
 detectCores()
-cl <- makePSOCKcluster(14)
+cl <- makePSOCKcluster(6)#14
 registerDoParallel(cl)
@@
 start.time <- Sys.time()
-years <- 1997:2012
+years <- 2001:2012 #1997
 
 for (t in years){
+
   tplus1 <- t + 1
   set.seed(71)
+  print(t)
+
   model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
@@
                      trControl = tc,
-                     method = "rf", #"rpart", # rf lm
+                     method = "ranger", 
                      na.action = na.omit,
-                     preProc = c("center", "scale"),
-                     importance = TRUE)
+                     #preProc = c("center", "scale"),
+                     importance = 'impurity') #TRUE
+
+  #assign(paste0("model.all",as.character(t + 1)), model.all)
+  file <- paste0(path, "/data/temp/oa_rf/timeseries", as.character(t + 1), ".RData")
+  save(model.all, file = file)
+  rm(model.all)
 }
@@
 # create a list of the above model outputs
-fit.model.all <- mget(ls(pattern = "^model.all\\d"))
+
+load(file = (paste0(path, "/data/temp/oa_rf/timeseries2001.RData")))
+
+filenames <- list.files(paste0(path, "/data/temp/oa_rf"), pattern = "*.RData", full.names = T)
+
+t <- 1997
+for (i in filenames){
+  load(file = i)
+  t <- t + 1
+  assign(paste0("timeseries",t), model.all)  
+}
+
+fit.model.all <- mget(ls(pattern = "^timeseries\\d"))
+rm(list = ls(pattern = "^timeseries\\d"))
 
@@
 years <- (1998:2011)
-pred.by.year.all <- data.frame(matrix(NA,nrow=374, ncol = 0))
+pred.by.year.all <- data.frame(matrix(NA,nrow=227759, ncol = 0))
 for (t in years){
   t_plus_1 <- t + 1
-  pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",t)], df[df$year==t_plus_1,])
+  pred <- predict(fit.model.all[names(fit.model.all)==paste0("timeseries",t)], df[df$year==t_plus_1,])
   pred <- as.data.frame(pred)
@@
   filter(year!=1996 & year!=1997 & year!=1998) %>% 
-  dplyr::select(ladcd, year, n) # CHANGE n WITH growth
+  dplyr::select(oa11cd, year, n) # CHANGE n WITH growth
 
 # column bind prediction and data  
-pred.by.year.all <- cbind(pred.by.year.all, df.no1997_1 %>% arrange(year, ladcd)) #DOUBLE CHECK ARRANGE
+pred.by.year.all <- cbind(pred.by.year.all, df.no1997_1 %>% arrange(year, oa11cd)) #DOUBLE CHECK ARRANGE
 
@@
 ```
+
+## RF spatial
+
+```{r}
+df <- df %>% mutate(help = substr(ladcd, start = 1, stop = 2)) %>% 
+  mutate(RGN21CD = ifelse(is.na(RGN21CD), help, RGN21CD)) %>% 
+  ungroup() %>% 
+  dplyr::select(-help)
+
+df <- df %>% dplyr::select(-n.lag)
+
+```
+
+### train the model in a loop
+
+```{r}
+#length is = (n_repeats*nresampling)+1
+seeds <- vector(mode = "list", length = 11)
+
+#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South
+
+#for the last model
+seeds[[11]]<-sample.int(1000, 1)
+
+# CV
+tc <- trainControl(method = "cv",
+                   number = 10,
+                   seeds = seeds,
+                   allowParallel = T,
+                   savePredictions = 'final')
+
+detectCores()
+cl <- makePSOCKcluster(14)
+registerDoParallel(cl)
+
+# run the model for every 2 years
+start.time <- Sys.time()
+regions <- df %>% distinct(RGN21CD)
+regions <- as.vector(regions$RGN21CD)
+
+for (i in regions){
+  set.seed(71)
+  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
+                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
+                     data = df %>% filter(RGN21CD==i), trControl = tc,
+                     method = "ranger",
+                     na.action = na.omit,
+                     #preProc = c("center", "scale"),
+                     importance = 'impurity') #TRUE
+  assign(paste0("model.all",as.character(i)), model.all)
+}
+
+stopCluster(cl)
+end.time <- Sys.time()
+time.taken <- round(end.time - start.time,2)
+time.taken
+
+# df %>% filter(growth == Inf) 
+# log(n/n.lag)
+
+# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
+# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
+```
+
+### Resamples for plots
+
+```{r}
+# create a list of the above model outputs
+fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
+
+# resamples
+fit.model.all.res <- resamples(fit.model.all)
+summary(fit.model.all.res)
+
+# plot metrics
+
+# TODO: fix the year labels
+
+scales <- list(x=list(relation="free"), y=list(relation="free"))
+bwplot(fit.model.all.res, scales=scales)
+
+# TODO: fix the year labels
+```
+
+### Variable importance
+
+```{r , include=TRUE, results= 'markup', message=FALSE}
+lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
+```
+
+```{r}
+help <- lapply(fit.model.all, function(x) varImp(x)) 
+help <- lapply(help, function(x) x[[1]])
+help <- do.call("rbind", help)
+
+importance <- help %>% rownames_to_column() %>% 
+  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
+         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
+  rename(importance = Overall) %>% 
+  relocate(importance, .after = year) %>% 
+  dplyr::select(-rowname)
+
+importance %>% ggplot() +
+  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
+  scale_color_discrete(name = "Features") +
+  #scale_linetype_manual(name = "Website density regressions111") 
+  ylab("Importance") + 
+  ggtitle("Feature importance") +
+  theme_minimal() +
+  theme(axis.title.x=element_blank(),
+        plot.title = element_text(hjust = 0.5))
+
+```
+
+### test
+
+```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
+
+# growth or n
+
+# for a reference point
+df %>% group_by(year) %>%
+  summarise(min = min(n), max=max(n),
+            mean = mean(n), median = median(n)) %>%
+  round(2) %>% kable()
+```
+
+```{r}
+## Test on t + 2
+
+# train on one region, test on all the rest
+
+df.predictions <- data.frame()
+for (i in regions){
+  for (j in regions[!regions == i]){
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
+    pred <- as.data.frame(pred)
+    rownames(pred) <- c()
+    assign(paste0("region.predict.model.", i, ".on.", j), pred)
+
+    d <- paste0("region.predict.model.", i, ".on.", j)
+    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
+    d <- d[,c(5, 2, 4, 6, 1)]
+    colnames(d)[5] <- "predictions"
+    d <- d %>% mutate(trained.on = i)
+    df.predictions <- rbind(d, df.predictions)
+  }
+}
+
+df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))
+
+# split to list by region pair
+pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
+
+# calculate metrics for every region pair 
+rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
+                                                                     obs = x$n))  # CHANGE n WITH growth
+
+path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup) %>% 
+  dplyr::select(RGN21CD, RGN21NM) %>% 
+  distinct() %>% 
+  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
+
+rf.year.all.metrics %>% 
+  as.data.frame() %>%
+  add_rownames(var = "metrics") %>% 
+  pivot_longer(!metrics, names_to = "train.test") %>% 
+  pivot_wider(names_from = metrics, values_from = value) %>% 
+  separate(train.test, c("train", "test"), remove = F) %>% 
+  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
+  rename(train.region=RGN21NM) %>% 
+  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
+  rename(test.region=RGN21NM) %>% 
+  arrange(Rsquared)
+  #summarise(r = range(Rsquared))
+  #round(2) %>%
+  #kable()
+```
+
+### test plots
+
+```{r}
+#plots
+ggplot(data = df.predictions, aes(x = n, y = predictions)) +
+  geom_point(colour = "blue") +
+  geom_abline(intercept = 0, slope = 1, colour = "red") +
+  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
+  facet_wrap(~ test.train, ncol = 2) +
+  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
+  ggtitle("Predicted vs. actual trade by year")
+```

### 2023-11-21 src/la_RF.Rmd
CAST

@@
 
-## RF spatial with CAST
+## RF with CAST
 
@@
 ```{r}
+set.seed(123)
 indices <- CreateSpacetimeFolds(df, 
@@
 #(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 11) # 9 after adding South
 
@@
 time.taken
+
+print(model.all)
+# Random Forest 
+# 
+# 5984 samples
+#   10 predictor
+# 
+# No pre-processing
+# Resampling: Cross-Validated (10 fold) 
+# Summary of sample sizes: 4718, 4690, 5040, 4718, 5055, 4704, ... 
+# Resampling results across tuning parameters:
+# 
+#   mtry  splitrule   RMSE        Rsquared   MAE       
+#    2    variance    0.02194750  0.9015001  0.01162254
+#    2    extratrees  0.02245492  0.8984940  0.01229071
+#    6    variance    0.02208770  0.9025087  0.01175267
+#    6    extratrees  0.02165214  0.9081808  0.01161281
+#   10    variance    0.02447416  0.8859265  0.01326348
+#   10    extratrees  0.02204464  0.9064730  0.01184945
+# 
+# Tuning parameter 'min.node.size' was held constant at a value of 5
+# RMSE was used to select the optimal model using the smallest value.
+# The final values used for the model were mtry = 6, splitrule = extratrees and min.node.size = 5.
+
+varimp_mars <- varImp(model.all) 
+varimp_mars$importance <- varimp_mars$importance %>% 
+  rownames_to_column() %>% 
+  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
+                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
+                                 ifelse(rowname == "London", "distance to London",
+                                        ifelse(rowname == "n.slag", "spatial lag of web density",
+                                               ifelse(rowname == "n.lag", "temporal lag of web density",
+                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag  of web density",
+                                                             ifelse(rowname == "n.nearest.city", "web density in nearest city", rowname)))))))) %>% 
+  column_to_rownames()
+
+path.out <- paste0(path, "/outputs/rf/varimp_LA.png")
+ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
+ggsave(path.out)
 
@@
 
-### train the model in a loop
+### train the model in a loop per region
 
@@
 time.taken
-
-# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
-# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
 ```
@@
 
-### test
+### Test on all other regions
 
 ```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
-
-# growth or n
 
@@
 ```{r}
-## Test on t + 2
-
 # train on one region, test on all the rest
@@
 
+path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
 rf.year.all.metrics %>% 
@@
   coord_fixed() +
-  theme(legend.position = "none")
+  labs(y = "", x = "") +
+  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
+  theme(#legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust=1),
+        panel.border = element_blank(), 
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.background = element_rect(fill = "white"))
+ggsave(path.out)
+
   #summarise(r = range(Rsquared))
@@
   #kable()
-```
-
-### test plots
-
-```{r}
-#plots
-ggplot(data = df.predictions, aes(x = n, y = predictions)) +
-  geom_point(colour = "blue") +
-  geom_abline(intercept = 0, slope = 1, colour = "red") +
-  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
-  facet_wrap(~ test.train, ncol = 2) +
-  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
-  ggtitle("Predicted vs. actual trade by year")
-```
+
+```

### 2023-11-21 src/la_RF.Rmd
CAST

@@
 sapply(df, function(x) sum(is.na(x)))
+```
+
+## unregister_dopar function
+
+This is helpful when the parallel loop collapses.
+
+```{r}
+unregister_dopar <- function() {
+  env <- foreach:::.foreachGlobals
+  rm(list=ls(name=env), pos=env)
+}
+unregister_dopar()
 ```
@@
   indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD==i), 
-                                spacevar = "ladcd",
+                                #spacevar = "ladcd",
                                 timevar = "year", 
@@
 
-  set.seed(71)
   model.all <- train(n ~ area + dist + dist.retail + London + south + 
@@
   coord_fixed() +
-  labs(y = "", x = "") +
+  labs(y = "", x = "") + #y: test.region, x: train.region
   scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +

### 2023-11-21 src/oa_RF.Rmd
CAST

@@
 
-## spatial data
+## spatial data OLD
 
@@
 
+## spatial data NEW
+
+```{r}
+library(sf)
+
+# Get OA for England and Wales
+path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
+oa.ew <- st_read(path.geo)
+
+# Spatial transformations
+oa.ew <- st_transform(oa.ew, 4326)  # EPSG code for WGS84
+
+# Rename columns and keep only necessary columns
+oa.ew <- oa.ew %>% dplyr::select(OA11CD, Shape__Area) %>% 
+  rename(id = OA11CD,
+         area = Shape__Area)
+# oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
+# colnames(oa.ew) <- c('id', 'area')
+
+# Check the data structure
+print(summary(oa.ew))
+
+# Get OA for Scotland
+path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
+oa.sc <- st_read(dsn = path.geo.sc, layer = "OutputArea2011_MHW")
+
+# Spatial transformations
+oa.sc <- st_transform(oa.sc, 4326)  # EPSG code for WGS84
+
+# Rename columns and keep only necessary columns
+oa.sc <- oa.sc %>% dplyr::select(code, SHAPE_1_Ar) %>% 
+  rename(id = code,
+         area = SHAPE_1_Ar)
+# oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
+# colnames(oa.sc) <- c('id', 'area')
+
+# Check the data structure
+print(summary(oa.sc))
+
+# Build OA for GB
+oa.gb <- rbind(oa.ew, oa.sc)
+rm(oa.ew, oa.sc)
+
+# Get GB
+path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
+gb <- st_read(path.gb)
+
+# Spatial transformations
+gb <- st_transform(gb, 4326)  # EPSG code for WGS84
+```
+
 ## Full panel
@@
 ```{r}
-oa.sf <- st_as_sf(oa.gb)
+#oa.sf <- st_as_sf(oa.gb)
 
-oa.full <- replicate(17, oa.sf %>% as_tibble(), simplify = FALSE) %>%
+oa.full <- replicate(17, oa.gb %>% as_tibble(), simplify = FALSE) %>%
   bind_rows(.id = 'id_') %>% 
@@
 #   mutate(id = row.names())
-oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
+
+# oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
+#   as.data.frame() %>% 
+#   add_rownames(var = "id")
+sf_use_s2(FALSE)
+oa.gb.c <- st_centroid(oa.gb) %>%
+  st_coordinates() %>%
   as.data.frame() %>%
-  add_rownames(var = "id")
+  #add_rownames(var = "id") %>% 
+  bind_cols(oa.gb %>% dplyr::select(id))
 
@@
 geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
-retail <-readOGR(geo.path) 
+#retail <-readOGR(geo.path) 
+retail <- st_read(geo.path)
+
 #retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
-retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id") #%>%
-
-retail.major.cetres <- retail.major.cetres.help %>% 
-  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
+# retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
+#   as.data.frame() %>% 
+#   add_rownames(var = "id") #%>%
+retail.major.cetres.help <- st_centroid(retail) %>% 
   st_transform(4326) %>%
   st_coordinates() %>% 
-  as_tibble() %>% 
-  bind_cols(retail.major.cetres.help$id) %>% 
-  rename(id = '...3') %>% 
-  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
+  as.data.frame() %>% 
+  add_rownames(var = "id")
+
+# retail.major.cetres <- retail.major.cetres.help %>% 
+#   st_as_sf(coords = c("x", "y"), crs = 27700) %>%
+#   st_transform(4326) %>%
+#   st_coordinates() %>%
+#   as_tibble() %>% 
+#   bind_cols(retail.major.cetres.help$id) %>% 
+#   rename(id = '...3') %>% 
+#   left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
+retail.major.cetres <- retail.major.cetres.help %>% 
+  # as_tibble() %>% 
+  # bind_cols(retail.major.cetres.help$id) %>% 
+  # #rename(id = '...3') %>% 
+  left_join(retail %>% rownames_to_column(var = "id"), by = 'id')
 ```
@@
 ```{r}
-dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist <- distm(cbind(oa.gb.c$Y, oa.gb.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
 dist <- round((dist/1000),2) %>% 
@@
 # calculate area
-oa.gb$area <- area(oa.gb) #/ 1000000
+oa.gb$area <- st_area(oa.gb) #/ 1000000
 
@@
   rename(oa11cd=id) %>% 
-  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
+  left_join(oa.gb, by = c("oa11cd" = "id")) %>% 
   relocate(area, .after = n) %>% 
@@
          density = n / area) %>% 
-  relocate(density, .after = area)
+  relocate(density, .after = area) %>% 
+  dplyr::select(-geometry)
 
@@
 ```{r}
-dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
+dist.retail <- distm(cbind(oa.gb.c$X, oa.gb.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
 dist.retail <- round((dist.retail/1000),2) %>% 
@@
 sf_use_s2(FALSE)
-help1 <- st_filter( oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
+help1 <- st_filter( oa.gb, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
   mutate(cities = c("Leicester",
@@
 sf_use_s2(FALSE)
-oa.c <- st_centroid(oa.sf)
+oa.c <- st_centroid(oa.gb)
 kn <- knearneigh(oa.c, k = 5)

### 2023-11-23 src/oa_RF.Rmd
CAST

@@
 
-# Partially complete panel as not all OAs have a website in at least ine year
+# Partially complete panel as not all OAs have a website in at least one year
 df <- df %>% filter(!is.na(oa11cd)) %>% 
@@
 
-```{r}
+```{r eval=F}
 # get OA for England and Wales
@@
 
+# NI
+path.geo.ni <- paste0(path, "/data/raw/ni_small_area/")
+oa.ni <- st_read(dsn = path.geo.ni, layer = "SA2011")
+oa.ni <- st_transform(oa.ni, 4326)  # EPSG code for WGS84
+oa.ni <- oa.ni %>% dplyr::select(SA2011, Hectares) %>% 
+  rename(id= SA2011,
+         area = Hectares)
+
 # Build OA for GB
-oa.gb <- rbind(oa.ew, oa.sc)
-rm(oa.ew, oa.sc)
+oa.gb <- rbind(oa.ew, oa.sc, oa.ni)
+rm(oa.ew, oa.sc, oa.ni)
 
@@
 ```{r}
-dist <- distm(cbind(oa.gb.c$Y, oa.gb.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist <- distm(cbind(oa.gb.c$X, oa.gb.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
 dist <- round((dist/1000),2) %>% 
@@
 
+# in.path <- paste0(path, "/data/temp/df_for_oa_rf_old.csv")
+# df.compare <- read_csv(in.path)
+
 # out.path <- paste0(path, "/data/temp/oa_rf.RData")
@@
 
+df <- df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol",
+                        "Manchester", "Sheffield", "Leeds", "Edinburgh",
+                        "Leicester", "nearest.city.oa11cd"))
+
 # out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
-# df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol", 
-#                         "Manchester", "Sheffield", "Leeds", "Edinburgh", 
-#                         "Leicester", "nearest.city.oa11cd")) %>% 
-#   write_csv(file = out.path)
+# df <- read_csv(out.path)
+# df %>% write_csv(file = out.path)
+```
+
+## RF with CAST
+
+```{r}
+df <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
+  mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
+  ungroup() %>% 
+  dplyr::select(-help)
+
+df <- df %>% filter(year > 1996)
+
+#df <- df %>% dplyr::select(-n.lag)
+```
+
+### unregister_dopar function
+
+This is helpful when the parallel loop collapses.
+
+```{r}
+unregister_dopar <- function() {
+  env <- foreach:::.foreachGlobals
+  rm(list=ls(name=env), pos=env)
+}
+unregister_dopar()
+```
+
+### train the model in all data
+
+```{r}
+set.seed(123)
+indices <- CreateSpacetimeFolds(df, 
+                                spacevar = "oa11cd",
+                                timevar = "year", 
+                                k = 10)
+
+#length is = (n_repeats*nresampling)+1
+seeds <- vector(mode = "list", length = 11)
+
+#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 11) # 9 after adding South
+
+#for the last model
+seeds[[11]]<-sample.int(1000, 1)
+
+# CV
+tc <- trainControl(method = "cv",
+                   number = 10,
+                   seeds = seeds,
+                   allowParallel = T,
+                   index = indices$index,
+                   savePredictions = 'final')
+
+# detectCores()
+# cl <- makePSOCKcluster(2)
+# registerDoParallel(cl)
+
+start.time <- Sys.time()
+model.all <- train(n ~ area + dist + dist.retail + London + south + 
+                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
+                   data = df, 
+                   trControl = tc,
+                   method = "ranger",
+                   na.action = na.omit,
+                   #preProc = c("center", "scale"),
+                   importance = "impurity")
+
+# stopCluster(cl)
+end.time <- Sys.time()
+time.taken <- round(end.time - start.time,2)
+time.taken
+
+print(model.all)
+
+varimp_mars <- varImp(model.all) 
+varimp_mars$importance <- varimp_mars$importance %>% 
+  rownames_to_column() %>% 
+  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
+                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
+                                 ifelse(rowname == "London", "distance to London",
+                                        ifelse(rowname == "n.slag", "spatial lag of web density",
+                                               ifelse(rowname == "n.lag", "temporal lag of web density",
+                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag  of web density",
+                                                             ifelse(rowname == "n.nearest.city", "web density in nearest city", rowname)))))))) %>% 
+  column_to_rownames()
+
+path.out <- paste0(path, "/outputs/rf/varimp_OA.png")
+ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
+ggsave(path.out)
+
+# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
+#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
+#                                                         "year")], 
+#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
+#                     trControl = tc,
+#                     method = "ranger",
+#                     #na.action = na.omit,
+#                     #importance = "impurity",
+#                     metric="Rsquared",
+#                     verbose=FALSE)
+# ffsmodel_LLO
+# ffsmodel_LLO$selectedvars
+# plot_ffs(ffsmodel_LLO)
+```
+
+### train the model in a loop per region
+
+```{r}
+
+#length is = (n_repeats*nresampling)+1
+seeds <- vector(mode = "list", length = 11)
+
+#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 11) # 9 after adding South
+
+#for the last model
+seeds[[11]]<-sample.int(1000, 1)
+
+# CV
+tc <- trainControl(method = "cv",
+                   number = 10,
+                   seeds = seeds,
+                   allowParallel = T,
+                   index = indices$index,
+                   savePredictions = 'final')
+
+detectCores()
+cl <- makePSOCKcluster(14)
+registerDoParallel(cl)
+
+# train in every region
+start.time <- Sys.time()
+
+regions <- df %>% distinct(RGN21CD)
+regions <- as.vector(regions$RGN21CD)
+
+for (i in regions){
+
+  set.seed(71)
+  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD==i), 
+                                spacevar = "oa11cd",
+                                timevar = "year", 
+                                k = 10)
+
+  model.all <- train(n ~ area + dist + dist.retail + London + south + 
+                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
+                   data = df %>% filter(RGN21CD==i), 
+                   trControl = tc,
+                   method = "ranger",
+                   na.action = na.omit,
+                   #preProc = c("center", "scale"),
+                   importance = "impurity")
+  assign(paste0("model.all",as.character(i)), model.all)
+}
+
+stopCluster(cl)
+end.time <- Sys.time()
+time.taken <- round(end.time - start.time,2)
+time.taken
+```
+
+### Resamples for plots
+
+```{r}
+# create a list of the above model outputs
+fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
+
+# resamples
+fit.model.all.res <- resamples(fit.model.all)
+summary(fit.model.all.res)
+
+# plot metrics
+
+# TODO: fix the year labels
+
+scales <- list(x=list(relation="free"), y=list(relation="free"))
+bwplot(fit.model.all.res, scales=scales)
+
+# TODO: fix the year labels
+```
+
+### Variable importance
+
+```{r , include=TRUE, results= 'markup', message=FALSE}
+lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
+```
+
+```{r}
+help <- lapply(fit.model.all, function(x) varImp(x)) 
+help <- lapply(help, function(x) x[[1]])
+help <- do.call("rbind", help)
+
+importance <- help %>% rownames_to_column() %>% 
+  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
+         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
+  rename(importance = Overall) %>% 
+  relocate(importance, .after = year) %>% 
+  dplyr::select(-rowname)
+
+importance %>% ggplot() +
+  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
+  scale_color_discrete(name = "Features") +
+  #scale_linetype_manual(name = "Website density regressions111") 
+  ylab("Importance") + 
+  ggtitle("Feature importance") +
+  theme_minimal() +
+  theme(axis.title.x=element_blank(),
+        plot.title = element_text(hjust = 0.5))
+
+```
+
+### Test on all other regions
+
+```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
+
+# for a reference point
+df %>% group_by(year) %>%
+  summarise(min = min(n), max=max(n),
+            mean = mean(n), median = median(n)) %>%
+  round(2) %>% kable()
+```
+
+```{r}
+# train on one region, test on all the rest
+
+df.predictions <- data.frame()
+for (i in regions){
+  for (j in regions[!regions == i]){
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
+    pred <- as.data.frame(pred)
+    rownames(pred) <- c()
+    assign(paste0("region.predict.model.", i, ".on.", j), pred)
+
+    d <- paste0("region.predict.model.", i, ".on.", j)
+    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
+    d <- d[,c(5, 2, 4, 6, 1)]
+    colnames(d)[5] <- "predictions"
+    d <- d %>% mutate(trained.on = i)
+    df.predictions <- rbind(d, df.predictions)
+  }
+}
+
+df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))
+
+# split to list by region pair
+pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
+
+# calculate metrics for every region pair 
+rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
+                                                                     obs = x$n))  # CHANGE n WITH growth
+
+path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup) %>% 
+  dplyr::select(RGN21CD, RGN21NM) %>% 
+  distinct() %>% 
+  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
+
+path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
+rf.year.all.metrics %>% 
+  as.data.frame() %>%
+  rownames_to_column(var = "metrics") %>% 
+  pivot_longer(!metrics, names_to = "train.test") %>% 
+  pivot_wider(names_from = metrics, values_from = value) %>% 
+  separate(train.test, c("train", "test"), remove = F) %>% 
+  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
+  rename(train.region=RGN21NM) %>% 
+  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
+  rename(test.region=RGN21NM) %>% 
+  arrange(Rsquared) %>% 
+  dplyr::select(train.region, test.region, Rsquared) %>% 
+  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
+  geom_tile() +
+  coord_fixed() +
+  labs(y = "", x = "") + #y: test.region, x: train.region
+  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
+  theme(#legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust=1),
+        panel.border = element_blank(), 
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.background = element_rect(fill = "white"))
+ggsave(path.out)
+
+  #summarise(r = range(Rsquared))
+  #round(2) %>%
+  #kable()
+
 ```

### 2023-11-24 abstracts_notes/abstract_GfR_2024/Tranos_GfR_2024.tex
CAT OA regions

\documentclass[11pt]{article}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage[demo]{graphicx}
%\usepackage{subfig}
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

%\usepackage[square,numbers]{natbib}
\usepackage[round]{natbib} % remove number, use ()
\bibliographystyle{abbrvnat}

\newcommand{\numpy}{{\tt numpy}}    % tt font for numpy

\topmargin -.5in
\textheight 9in
\oddsidemargin -.25in
\evensidemargin -.25in
\textwidth 7in

\begin{document}

% remove date
%\date{}

% ========== Edit your name here
\author[]{Professor Emmanouil Tranos}
\affil[]{University of Bristol and The Alan Turing Institute}

\title{Digital economy in the UK: a multi-scalar story of the diffusion of web technologies}
\maketitle

\begin{center}
\href{mailto:e.tranos@bristol.ac.uk}{e.tranos@bristol.ac.uk},
\href{https://etranos.info/}{etranos.info},
\href{https://twitter.com/EmmanouilTranos}{@EmmanouilTranos} \&
\href{https://www.linkedin.com/in/emmanouil-tranos-42947812a/}{LinkedIn}
\end{center}

\medskip

% general, summary
This paper maps the participation in the digital economy and its evolution in the UK over space and time. Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects. Useful as these approaches might have been, they cannot capture the micro-processes and the characteristics of the individual online behaviour. Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK. Instead of metrics capturing the passive engagement with digital technologies, for instance internet subscription metrics, this paper targets the active engagement with the digital as reflected in website creation.

Importantly, the data and geolocation strategy allow to capture these processes at small spatial scales. This level of granularity differentiate this paper with previous approaches in the literature, which were only able to capture technological adoption at more coarse level. Thus, this paper tests how well established theoretical approaches regarding the diffusion of technologies are still applicable when the focus is on local scales. Although we know that spatial contagion and urban hierarchies are key drivers of technological diffusion, such theoretical concepts have not been tested at small geographical scales. Apart from an empirical interest, understanding such processes at small scales can support designing strategies for the adoption of new technologies and the development of relevant infrastructure.

The paper begins with simple analytical tools which reveal the scale-free character of the 'S curve' that defines the diffusion of new technologies and innovations. Indeed, such an 'S curve' describes the adoption of web technologies even at the level of local authorities in the UK. Then the paper continues with other exploratory approaches, which signal the spatial dependence of the diffusion on web technologies. In the last section, the paper employs state of the art machine learning techniques to predict the diffusion of web technologies over space and, importantly, highlight how different or not UK regions have been in terms of the adoption of web technologies.

Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.

\begin{figure}%
    \centering
    \subfloat[\centering label 1]{{\includegraphics[width=5cm]{./lad_E07000031.png} }}%
    \qquad
    \subfloat[\centering label 2]{{\includegraphics[width=5cm]{./lad_E07000073} }}%
    \caption{2 Figures side by side}%
    \label{fig:example}%
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}{.3\textwidth}
		\includegraphics[width=\textwidth]{lad_E07000031.png}
		\caption{Atom 1}
	\end{subfigure}
%%%%%%%%%%%%%%
	\begin{subfigure}{.3\textwidth}
		\includegraphics[width=\textwidth]{lad_E07000031.png}
		\caption{Atom 2}
	\end{subfigure}
%%%%%%%%%%%%%%
	\begin{subfigure}{.3\textwidth}
		\includegraphics[width=\textwidth]{lad_E07000031.png}
		\caption{Atom 3}
	\end{subfigure}
	\caption{Atoms are fun!}
\end{figure}

%\newpage
%\bibliography{bibliography}

\end{document}
\grid
\grid

### 2023-11-24 paper/submission1/tranos2023.qmd
CAT OA regions

@@
 Geographers were always interested in how new technologies and
-innovations diffuse across space and, importantly, how such
+innovations diffuse across space and time and, importantly, how such
 spatio-temporal processes can be modelled. After all, diffusion together 
@@
 geographical studies exploring the diffusion of new and, more 
-specifically, digital technologies across *both* space and time can be
-attributed to the scarcity of relevant and granular enough data. As 
-@zook2022mapping highlight, digital activities are hardly ever captured 
-in official data.
+specifically for this paper, digital technologies across *both* space 
+and time can be attributed to the scarcity of relevant and granular 
+enough data. As @zook2022mapping highlight, digital activities are 
+hardly ever captured in official data.
 
@@
 combination of data sources which escape the traditional social science
-domain and adequate research methods can offer new contributions 
-regarding our understanding of how new technologies diffuse.
+domain and adequate research methods can offer new lenses to geographical
+research regarding the understanding of technological diffusion.
 
@@
 and local authorities. 
-These processes and their effects very a lot between scales: although
+These processes and their effects very a lot across scales: although
 the diffusion of a new technology might not be optimal at a local level,
@@
 
-Importantly, the data used here depicts the *active* engagement with the web as 
-it contains all the commercial websites archived by the Internet Archive
+Importantly, the data used here depicts the *active* engagement with the web in 
+the UK as it contains all the commercial websites archived by the Internet Archive
 ^[[https://archive.org/](See https://archive.org/).]
@@
 adoption and internet speeds [e.g. @blank2018local; @riddlesden2014broadband]. 
+More details about the data and the data generation process can be found in Section
+[XX](#sec3). **ADD DATA DESCRIPTION, WHY .CO.UK**
 
@@
 resembles the 'lead-lag' spatial acceleration effect in market research 
-[@bento2018time; @PERES201091]. Hägerstrand maybe more widely known for highlighting 
-the role time plays in the diffusion of innovations: an early-pioneering 
-period, a middle fast accelerating period and a final saturation one [@morrill2020spatial.
+[@bento2018time; @PERES201091]. Hägerstrand is more widely known though for highlighting 
+the role time plays in the diffusion of innovations: an early-pioneering period, 
+a middle fast accelerating period and a final saturation one [@morrill2020spatial.
 
 The temporal dimension was further explored by Everett Rogers and his seminal work 
-on 'Diffusion of Innovations' [@rogers2010diffusion]. Rogers being a sociologist 
+on 'Diffusion of Innovations' [@rogers2010diffusion]. Rogers being a sociologist, 
 his work focused not on the diffusion of innovations over space and time, but 
@@
 holds for urban computer employment and also urban heterogeneity is essential in
-explaining this distribution. @bednarz2020pulled focused on wind turbines and modelled
-their spatial diffusion across $402$ German regions during $1970-2015$. Their key
-finding is that local demand than local supply was the main driving factor.
+explaining this distribution. in a recent study, @bednarz2020pulled focused on 
+wind turbines and modelled their spatial diffusion across $402$ German regions 
+during $1970-2015$. Their key finding is that local demand than local supply was 
+the main driving factor.
 
-At a global scale @perkins2005international analyzed if the diffusion rate of new
+At a global scale @perkins2005international explored whther the diffusion rate of new
 technologies is driven by a latecomer advantage and the engagement with the global
@@
 
-## Results {#sec3.1}
+# Results {#sec3}
+
+## Old ideas
 
@@
 
--   Moran's I: for OA and LAD over time **TODO** add 0s
+-   Moran's I: for OA and LAD over time
 
@@
 
--   S for LADs per firm and OA. **TODO**: OA per firm? fix firms over
-    time.
+-   S for LADs per firm and OA.
 
@@
 
--   **TODo** check s_uk_firm_11
+-   s_uk_firm_11
 
@@
 
-## 3.2 Case study approach {#sec3.2}
+## New structure
 
-# Digital technologies and spatial structure: a global view {#sec4}
+1. S-shaped diffusion curves: S for LADs per firm. UK, fast/slow/examples
 
-# The impact of ICT on the US and the UK spatial structures {#sec5}
+2. ranks: there is stability and movement
 
-# Discussion and conclusions {#sec6}
+3. Neighbourhood effect: diffusion proceeds outwards from innovation
+centers, first "hitting" nearby rather than far-away locations (Grubler
+1990)
+
+  -   Moran's I: for OA and LAD over time
+
+  -   LISA maps: for OA and LAD over time More and less expected clusters.
+    Different scales show different results
+
+  -   Website density regressions: for OA and LAD over time
+
+4. Hierarchy effect: from main centers to secondary ones -- central places
+
+  -   Gini coefficient. Almost perfect polarisation of web adoption in the
+    early stages at a granular level More equally diffused at the Local
+    Authority level Plateau overtime
+
+5. RF
+
+  - ideal: (i) train RF for all years and all (1) LADs and (2) OAs with CAST
+  and report metrics. (ii) train for all years and per region for (1) LADs and (2)
+  OAs, and predict to the other regions. Reports predictions as region similarities
+
+# Discussion and conclusions {#sec4}
 

### 2023-11-24 paper/submission1/tranos2023.tex
CAT OA regions

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
 \fi
@@
 \makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
 \makeatother
@@
 \@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
 \makeatother
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, interior hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, interior hidden, sharp corners, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 Geographers were always interested in how new technologies and
-innovations diffuse across space and, importantly, how such
+innovations diffuse across space and time and, importantly, how such
 spatio-temporal processes can be modelled. After all, diffusion together
@@
 geographical studies exploring the diffusion of new and, more
-specifically, digital technologies across \emph{both} space and time can
-be attributed to the scarcity of relevant and granular enough data. As
-\citet{zook2022mapping} highlight, digital activities are hardly ever
-captured in official data.
+specifically for this paper, digital technologies across \emph{both}
+space and time can be attributed to the scarcity of relevant and
+granular enough data. As \citet{zook2022mapping} highlight, digital
+activities are hardly ever captured in official data.
 
@@
 combination of data sources which escape the traditional social science
-domain and adequate research methods can offer new contributions
-regarding our understanding of how new technologies diffuse.
+domain and adequate research methods can offer new lenses to
+geographical research regarding the understanding of technological
+diffusion.
 
@@
 stakeholders who are interested in knowing how new digital technologies
-diffused over space and time and use this information to make
-predictions regarding the diffusion of related \emph{future}
-technologies. As per \citet{leibowicz2016representing}, historical
-studies agree on the fact that ``technologies diffuse at different
-times, at different rates, and to different extents in different places,
-and can be significantly influenced by policies \citep{victor1993}''.
-\citet{meade2021modelling} highlight that a variety of actors have a
-direct interest in gaining such knowledge including network equipment
-suppliers; network operators, regulatory and local authorities. These
-processes and their effects very a lot between scales: although the
-diffusion of a new technology might not be optimal at a local level, it
-might be beneficial from a global perspective as it could lead to faster
-diffusion to less advantaged places \citep{leibowicz2016representing}.
-Despite the spatial heterogeneity of such diffusion mechanisms and the
-policy relevance, there are very limited attempts in the literature to
-analyse the diffusion of new digital technologies at a detailed
-geographical level.
+diffused over space and time and use this knowledge to make predictions
+regarding the diffusion of related \emph{future} technologies. As per
+\citet{leibowicz2016representing}, historical studies agree on the fact
+that ``technologies diffuse at different times, at different rates, and
+to different extents in different places, and can be significantly
+influenced by policies \citep{victor1993}''. \citet{meade2021modelling}
+highlight that a variety of actors have a direct interest in gaining
+such knowledge including network equipment suppliers; network operators,
+regulatory and local authorities. These processes and their effects very
+a lot across scales: although the diffusion of a new technology might
+not be optimal at a local level, it might be beneficial from a global
+perspective as it could lead to faster diffusion to less advantaged
+places \citep{leibowicz2016representing}. Despite the spatial
+heterogeneity of such diffusion mechanisms and the policy relevance,
+there are very limited attempts in the literature to analyse the
+diffusion of new digital technologies at a detailed geographical level.
 
@@
 Also, only \(9\) per cent of UK's households had access to the internet
-in \(1998\) \citep{ons2018} and during that period that web included
-mostly static webpages, there were no social media and web browsing was
-happening exclusively from desktop PCs as there were no smartphones
-\citep{tranosuk}. Hence, it is fair to say that the study period
-captures the very early stages of the diffusion on a new technology that
-is the web. This is key point in the lifecycle of a new technology.
-Firstly, during this period new technologies are expensive, crude and
-imperfect \citep{rosenberg1994exploring, wilson201281}. A simple
-comparison between Web 1.0 and Web 3.0 applications clearly illustrates
-this \citep{tranos2020social}. During this period the performance of a
-new technology is the main attraction and not the cost to access and use
-it \citep{wilson2011lessons} There is a broader theoretical discussion
-in the literature regarding the motivation behind early adoption. As
+in \(1998\) \citep{ons2018}, the web included mostly static webpages,
+there were no social media and web browsing was happening exclusively
+from desktop PCs as there were no smartphones \citep{tranosuk}. Hence,
+it is fair to say that the study period captures the very early stages
+of the diffusion on a new technology that is the web. This is key point
+in the lifecycle of a new technology. Firstly, during this period new
+technologies are expensive, crude and imperfect
+\citep{rosenberg1994exploring, wilson201281}. A simple comparison
+between Web 1.0 and Web 3.0 applications clearly illustrates this
+\citep{tranos2020social}. During this period the performance of a new
+technology is the main attraction and not the cost to access and use it
+\citep{wilson2011lessons} There is a broader theoretical discussion in
+the literature regarding the motivation behind early adoption. As
 summarised by \citet{perkins2005international}, on the one hand,
@@
 Importantly, the data used here depicts the \emph{active} engagement
-with the web as it contains all the commercial websites archived by the
-Internet Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
+with the web in the UK as it contains all the commercial websites
+archived by the Internet Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
 that include a mention to at least one valid UK postcode in the web
@@
 and internet speeds
-\citep[e.g.][]{blank2018local, riddlesden2014broadband}.
+\citep[e.g.][]{blank2018local, riddlesden2014broadband}. More details
+about the data and the data generation process can be found in Section
+\protect\hyperlink{sec3}{XX}. \textbf{ADD DATA DESCRIPTION, WHY .CO.UK}
 

### 2023-11-24 src/oa_RF.Rmd
cast

@@
 library(rprojroot)
-library(rgdal)
-library(rgeos)
+#library(rgdal)
+#library(rgeos)
 library(sf)
@@
 
-# out.path <- paste0(path, "/data/temp/oa_rf.RData")
+# out.path <- paste0(path, "/data/temp/for_oa_rf.RData")
 # save.image(file = out.path)
@@
 print(model.all)
+# Random Forest 
+# 
+# 3716736 samples
+#      10 predictor
+# 
+# No pre-processing
+# Resampling: Cross-Validated (10 fold) 
+# Summary of sample sizes: 2926910, 2926924, 3136005, 2926924, 3136005, 2926924, ... 
+# Resampling results across tuning parameters:
+# 
+#   mtry  splitrule   RMSE      Rsquared   MAE      
+#    2    variance    3.865809  0.5809699  0.5313035
+#    2    extratrees  3.730542  0.6168954  0.5884731
+#    6    variance    3.761356  0.6063340  0.5366004
+#    6    extratrees  3.612670  0.6273041  0.5331196
+#   10    variance    3.929176  0.5775066  0.5449132
+#   10    extratrees  3.617598  0.6278302  0.5403551
+# 
+# Tuning parameter 'min.node.size' was held constant at a value of 5
+# RMSE was used to select the optimal model using the smallest value.
+# The final values used for the model were mtry = 6, splitrule =
+#  extratrees and min.node.size = 5.
 
@@
 ggsave(path.out)
+
+path.out <- paste0(path, "/data/temp/oa_rf/all_OAs/all_OAs.RData")
+save(model.all, file = path.out)
 
@@
 
+#folds
+k <- 3 # 7.09h for k=3
+
 #length is = (n_repeats*nresampling)+1
@@
 
+detectCores()
+cl <- makePSOCKcluster(2)
+registerDoParallel(cl)
+
+# train in every region
+start.time <- Sys.time()
+
+regions <- df %>% distinct(RGN11CD)
+regions <- as.vector(regions$RGN11CD)
+
+for (i in regions){
+
+  set.seed(71)
+  indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD==i), 
+                                spacevar = "oa11cd",
+                                timevar = "year", 
+                                k = k)
+
   # CV
   tc <- trainControl(method = "cv",
-                   number = 10,
+                   number = k,
                    seeds = seeds,
@@
 
-detectCores()
-cl <- makePSOCKcluster(14)
-registerDoParallel(cl)
-
-# train in every region
-start.time <- Sys.time()
-
-regions <- df %>% distinct(RGN21CD)
-regions <- as.vector(regions$RGN21CD)
-
-for (i in regions){
-
-  set.seed(71)
-  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD==i), 
-                                spacevar = "oa11cd",
-                                timevar = "year", 
-                                k = 10)
-
   model.all <- train(n ~ area + dist + dist.retail + London + south + 
                      n.slag + n.lag + n.l.slag + n.nearest.city + year,
-                   data = df %>% filter(RGN21CD==i), 
+                   data = df %>% filter(RGN11CD==i), 
                    trControl = tc,
@@
                    importance = "impurity")
-  assign(paste0("model.all",as.character(i)), model.all)
+  #assign(paste0("model.all",as.character(i)), model.all)
+
+  file <- paste0(path, "/data/temp/oa_rf/regions/region_", as.character(i), ".RData")
+  save(model.all, file = file)
+  rm(model.all)
 }
@@
 # create a list of the above model outputs
-fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
+filenames <- list.files(paste0(path, "/data/temp/oa_rf/regions"), pattern = "*.RData", full.names = T)
+
+for (i in filenames){
+  load(file = i)
+  i <- gsub("/home/nw19521/projects/web.diffusion/data/temp/oa_rf/regions/|.RData", "", i)
+  assign(i, model.all)  
+}
+
+fit.model.all <- mget(ls(pattern = "^region_"))
+
+# clear up
+rm(model.all) # this is the last item from the loop
+rm(list = ls(pattern = "^region_"))
 
@@
   for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], df[df$RGN11CD==j,])
     pred <- as.data.frame(pred)
@@
     d <- paste0("region.predict.model.", i, ".on.", j)
-    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
+    d <- cbind(get(d), df %>% filter(RGN11CD==j)) #%>% arrange(year, ladcd)
     d <- d[,c(5, 2, 4, 6, 1)]
@@
 
-df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))
+df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN11CD, ".", trained.on))
 
@@
   #assign(paste0("model.all",as.character(t + 1)), model.all)
-  file <- paste0(path, "/data/temp/oa_rf/timeseries", as.character(t + 1), ".RData")
+  file <- paste0(path, "/data/temp/oa_rf/timeseries/timeseries", as.character(t + 1), ".RData")
   save(model.all, file = file)
@@
 
-load(file = (paste0(path, "/data/temp/oa_rf/timeseries2001.RData")))
-
-filenames <- list.files(paste0(path, "/data/temp/oa_rf"), pattern = "*.RData", full.names = T)
+#load(file = (paste0(path, "/data/temp/oa_rf/timeseries/timeseries2001.RData")))
+
+filenames <- list.files(paste0(path, "/data/temp/oa_rf/timeseries/"), pattern = "*.RData", full.names = T)
 

### 2023-11-24 src/oa_RF.Rmd
cast

@@
   distinct() %>% 
-  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
-
-path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
+  rename(RGN11CD = RGN21CD) %>% 
+  add_row(RGN11CD = c("Scotland", "N0", "Wales"), 
+          RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
+
+path.out <- paste0(path, "/outputs/rf/heatmap_OA.png")
 rf.year.all.metrics %>% 
@@
   separate(train.test, c("train", "test"), remove = F) %>% 
-  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
+  left_join(lookup.region, by = c("train" = "RGN11CD")) %>% 
   rename(train.region=RGN21NM) %>% 
-  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
+  left_join(lookup.region, by = c("test" = "RGN11CD")) %>% 
   rename(test.region=RGN21NM) %>% 

### 2023-12-12 src/la_RF.Rmd
CAST
refined model LA

@@
 
-## spatial data
+## spatial data NEW
 
@@
 path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-la <- readOGR(path.geo)
+la <- st_read(path.geo)
 # source: https://geoportal.statistics.gov.uk/
@@
 # spatial transformations
-la <- spTransform(la, CRS("+init=epsg:4326"))
+la <- st_transform(la, 4326)
 
@@
 
+## spatial data
+
+```{r, eval=FALSE}
+# get LA for the UK
+path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
+la <- readOGR(path.geo)
+# source: https://geoportal.statistics.gov.uk/
+
+# spatial transformations
+la <- spTransform(la, CRS("+init=epsg:4326"))
+
+#la.f <- fortify(la, region = "LAD21CD")
+
+# cities
+cities <- maps::world.cities %>% 
+  filter(country.etc=="UK") %>% 
+  arrange(pop) %>% tail(10) 
+cities.sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
+```
+
 ## Distances to cities and retail centres
@@
 # LA centroids
-la.c <- gCentroid(la, byid=TRUE)@coords %>% 
+# la.c <- gCentroid(la, byid=TRUE)@coords %>% 
+#   as.data.frame() %>%
+#   bind_cols(la@data) %>% 
+#   dplyr::select(LAD21CD, x, y) %>% 
+#   remove_rownames()
+
+sf_use_s2(FALSE)
+la.c <- st_centroid(la) %>% 
+  st_coordinates() %>%
   as.data.frame() %>% 
-  bind_cols(la@data) %>% 
-  dplyr::select(LAD21CD, x, y) %>% 
-  remove_rownames()
+  bind_cols(la %>% dplyr::select(LAD21CD))
 
@@
 ```{r}
-dist <- distm(cbind(la.c$x, la.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist <- distm(cbind(la.c$X, la.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
 dist <- round((dist/1000),2) %>% 
@@
 # calculate area
-la$SHAPE_Area <- area(la) #/ 1000000
+#la$SHAPE_Area <- area(la) #/ 1000000
+la$area <- st_area(la) #/ 1000000
 
@@
 df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
-  left_join(la@data, by = c("ladcd" = "LAD21CD")) %>% 
-  rename(area = SHAPE_Area) %>% 
+  left_join(la, by = c("ladcd" = "LAD21CD")) %>% 
   relocate(area, .after = n) %>% 
-  mutate(area = area / 1000000,
+  mutate(area = area / 1000000,     # UNITS APPEAR WRONGLY IN M AND NOT IN KM
          density = n / area) %>% 
@@
 ```{r}
-dist.retail <- distm(cbind(la.c$x, la.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
+dist.retail <- distm(cbind(la.c$X, la.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
 dist.retail <- round((dist.retail/1000),2) %>% 
@@
 # London = Islington 
-sf_use_s2(FALSE)
-path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-la.sf <- st_read(path.geo)
-
-help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
-  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)
-
-help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
+# sf_use_s2(FALSE)
+# path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
+# la.sf <- st_read(path.geo)
+
+#help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
+#  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)
+
+# nearest city
+help1 <- st_filter( la, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
   mutate(cities = c("Leicester",
@@
 sf_use_s2(FALSE)
-la.c <- st_centroid(la.sf)
+la.c <- st_centroid(la)
 kn <- knearneigh(la.c, k = 5)
@@
 
-df <- df %>% drop_na() %>% group_by(year) %>% 
+df <- df %>% drop_na() %>% # THIS DROPS L999999
+  group_by(year) %>% 
   mutate(n.slag = lag.listw(knn.l, n)) %>% 
@@
 # For growth:
-#df <- df %>% mutate(growth = log(n/n.lag))
+df <- df %>% mutate(growth = log(n/n.lag))
 
@@
 
+df <- df %>% left_join(df %>% ungroup() %>% filter(LAD21NM == "Islington") %>% dplyr::select(year, n) %>% rename(n.London = n)) %>% 
+  group_by(ladcd) %>% 
+  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
+  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city))
+
 df <- df %>% filter(year > 1996)
-
-#df <- df %>% dplyr::select(-n.lag)
+sapply(df, function(x) sum(is.na(x)))
 ```
@@
 set.seed(123)
-indices <- CreateSpacetimeFolds(df, 
+indices <- CreateSpacetimeFolds(df %>% filter(year > 1999), 
                                 spacevar = "ladcd",
@@
 #(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 11) # 9 after adding South
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # , 11) 9 after adding South
 
@@
 start.time <- Sys.time()
-model.all <- train(n ~ area + dist + dist.retail + London + south + 
-                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
-                   data = df, 
+model.all <- train(growth ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
+                     I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
+                   data = df %>% filter(year > 1999), 
                    trControl = tc,
@@
 
-### train the model in a loop per region
+### train the model in a loop for all but one regions
 
@@
 # CV
-tc <- trainControl(method = "cv",
+rm(indices)
+
+detectCores()
+cl <- makePSOCKcluster(14)
+registerDoParallel(cl)
+
+# train in every region
+start.time <- Sys.time()
+
+regions <- df %>% distinct(RGN21CD)
+regions <- as.vector(regions$RGN21CD)
+
+for (i in regions){
+
+  set.seed(71)
+  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i), 
+                                spacevar = "ladcd",
+                                timevar = "year", 
+                                k = 10)
+
+  tc <- trainControl(method = "cv", #MOVED HERE< DOUBLE CHECK
                    number = 10,
@@
 
-detectCores()
-cl <- makePSOCKcluster(14)
-registerDoParallel(cl)
-
-# train in every region
-start.time <- Sys.time()
-
-regions <- df %>% distinct(RGN21CD)
-regions <- as.vector(regions$RGN21CD)
-
-for (i in regions){
-
-  set.seed(71)
-  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD==i), 
-                                #spacevar = "ladcd",
-                                timevar = "year", 
-                                k = 10)
-
-  model.all <- train(n ~ area + dist + dist.retail + London + south + 
-                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
-                   data = df %>% filter(RGN21CD==i), 
+  model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
+                       I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
+                   data = df %>% filter(RGN21CD!=i), 
                    trControl = tc,
@@
                    importance = "impurity")
-  assign(paste0("model.all",as.character(i)), model.all)
+  assign(paste0("model.all",as.character(i)), model.all) # name = region not in the training data
 }
@@
 stopCluster(cl)
-end.time <- Sys.time()
+end.time <- Sys.time() # ca. 2h
 time.taken <- round(end.time - start.time,2)
@@
 
-### Test on all other regions
+### Test on one region
 
@@
 ```{r}
+# train on all but one region, test on that region
+
+df.predictions <- data.frame()
+for (i in regions){
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==i,])
+    pred <- as.data.frame(pred)
+    rownames(pred) <- c()
+    assign(paste0("region.predict.model.", ".on.", i), pred)
+
+    d <- paste0("region.predict.model.", ".on.", i)
+    d <- cbind(get(d), df %>% filter(RGN21CD==i)) #%>% arrange(year, ladcd)
+    d <- d[,c(5, 2, 4, 6, 1)]
+    colnames(d)[5] <- "predictions"
+    d <- d %>% mutate(tested.on = i)
+    df.predictions <- rbind(d, df.predictions)
+}
+
+df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", tested.on))
+
+# split to list by region pair
+pred.by.region.all.list <- split(df.predictions, df.predictions$RGN21CD)
+
+# calculate metrics for every region pair 
+rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
+                                                                     obs = x$n))  # CHANGE n WITH growth
+
+path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup) %>% 
+  dplyr::select(RGN21CD, RGN21NM) %>% 
+  distinct() %>% 
+  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
+
+path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
+rf.year.all.metrics %>% 
+  as.data.frame() %>%
+  rownames_to_column(var = "metrics") %>% 
+  pivot_longer(!metrics, names_to = "test.on") %>% 
+  pivot_wider(names_from = metrics, values_from = value) %>% 
+  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
+  arrange(Rsquared)
+
+ggplot(aes(train.region, test.region, fill= Rsquared)) + 
+  geom_tile() +
+  coord_fixed() +
+  labs(y = "", x = "") + #y: test.region, x: train.region
+  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
+  theme(#legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust=1),
+        panel.border = element_blank(), 
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.background = element_rect(fill = "white"))
+ggsave(path.out)
+
+  #summarise(r = range(Rsquared))
+  #round(2) %>%
+  #kable()
+
+```
+
+#### OLD
+
+```{r, eval=FALSE}
 # train on one region, test on all the rest

### 2023-12-12 src/s_la_per_firm.Rmd
refined model LA

@@
 ```{r}
-a.la <- as_tibble()
+a.la <- tibble()
 all.la <- df %>% filter(ladcd != "NA") %>% distinct(ladcd) %>% 
@@
 a.la <- a.la %>% filter(term=="xmid") %>% 
-  bind_cols(as.tibble(all.la)) %>% 
+  bind_cols(as_tibble(all.la)) %>% 
   #mutate(estimate = round(estimate, 0)) %>% 

### 2023-12-14 src/la_RF.Rmd
refined model LA

@@
 # For growth:
-df <- df %>% mutate(growth = log(n/n.lag))
+#df <- df %>% mutate(growth = log(n/n.lag))                          
+df <- df %>% mutate(growth = log((n+0.001)/(n.lag+0.01)))                       # CONSTANT FOR GROWTH
+#df <- df %>% mutate(growth = n- n.lag)                                         # ABS. GROWTH
 
@@
 
-df <- df %>% left_join(df %>% ungroup() %>% filter(LAD21NM == "Islington") %>% dplyr::select(year, n) %>% rename(n.London = n)) %>% 
+df <- df %>% left_join(df %>% 
+                         ungroup() %>% 
+                         filter(LAD21NM == "Islington") %>% 
+                         dplyr::select(year, n) %>% 
+                         rename(n.London = n)) %>% 
   group_by(ladcd) %>% 
   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city))
+  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
+  ungroup()
 
@@
 set.seed(123)
-indices <- CreateSpacetimeFolds(df %>% filter(year > 1999), 
+indices <- CreateSpacetimeFolds(df %>% filter(year > 1999),                     # FILTER BY YEAR FOR GROWTH
                                 spacevar = "ladcd",
@@
 start.time <- Sys.time()
-model.all <- train(growth ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
+model.all <- train(growth ~                                                     # REPLACE n WITH growth
+                     #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
                      I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
-                   data = df %>% filter(year > 1999), 
+                   data = df %>% filter(year > 1999),                           # FILTER BY YEAR FOR GROWTH
                    trControl = tc,
@@
 ```{r}
+rm(indices)
 
@@
 #(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # , 11) 9 after adding South
+set.seed(831)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # 4 feature + 2
 
 #for the last model
+set.seed(832)
 seeds[[11]]<-sample.int(1000, 1)
 
-# CV
-rm(indices)
-
-detectCores()
+# detectCores()
 cl <- makePSOCKcluster(14)
@@
 
-regions <- df %>% distinct(RGN21CD)
+regions <- df %>% ungroup() %>% distinct(RGN21CD)
 regions <- as.vector(regions$RGN21CD)
@@
   set.seed(71)
-  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i), 
+  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i), #& year > 1999),   # FILTER BY YEAR FOR GROWTH 
                                 spacevar = "ladcd",
@@
                                 k = 10)
-
+  # CV
   tc <- trainControl(method = "cv", #MOVED HERE< DOUBLE CHECK
@@
 
-  model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
+  model.all <- train(growth ~                                                   # REPLACE n WITH growth
+                       #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                        I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
-                   data = df %>% filter(RGN21CD!=i), 
+                   data = df %>% filter(RGN21CD!=i), #& year > 1999),           # FILTER BY YEAR FOR GROWTH
                    trControl = tc,
@@
                    importance = "impurity")
-  assign(paste0("model.all",as.character(i)), model.all) # name = region not in the training data
+  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data
 }
@@
 for (i in regions){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==i,])
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], 
+                    df[df$RGN21CD==i,]) #& df$year > 1999,])                            # FILTER BY YEAR
     pred <- as.data.frame(pred)
@@
     d <- paste0("region.predict.model.", ".on.", i)
-    d <- cbind(get(d), df %>% filter(RGN21CD==i)) #%>% arrange(year, ladcd)
-    d <- d[,c(5, 2, 4, 6, 1)]
+    d <- cbind(get(d), df %>% filter(RGN21CD==i)) #& year > 1999))                      # FILTER BY YEAR
+    d <- d %>% dplyr::select(RGN21CD, ladcd, year, growth, 1)                           # REPLACE n WITH growth
     colnames(d)[5] <- "predictions"
@@
 rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))  # CHANGE n WITH growth
+                                                                     obs = x$growth))    # REPLACE n WITH growth
 
@@
 ```
+
+## Granger
+
+```{r}
+library(plm)
+p.df <- pdata.frame(df, index= c("ladcd","year"))
+
+test <- pgrangertest(n ~ n.l.slag, data = p.df, order = 1L) # *** reject the null that x does NOT cause y
+# n ~ Lags(n, 1) + Lags(n.l.slag, 1), or
+# n ~ Lags(n, 1) + Lags(n.slag, 2)
+# n ~ Lags(n, 1)
+test$indgranger %>% as_tibble() %>% arrange(`p-value`) #filter(`p-value` < 0.01)
+
+test <- pgrangertest(n ~ n.slag, data = p.df, order = 1L)
+# n ~ Lags(n, 1:1) + Lags(n.slag, 1:1)
+# n ~ Lags(n, 1:1)
+test$indgranger %>% as_tibble() %>% arrange(`p-value`) #filter(`p-value` < 0.01)
+
+test <- pgrangertest(n.slag ~ n, data = p.df, order = 1L) # ** reject the null that x does NOT cause y
+test$indgranger %>% as_tibble() %>% arrange(`p-value`)
+
+p.df$n.London_dist = p.df$n.London/p.df$London
+p.df$n.nearest.city_dist = p.df$n.nearest.city/p.df$dist
+
+pgrangertest(n ~ n.London_dist, data = p.df, order = 1L)
+pgrangertest(n ~ n.nearest.city_dist, data = p.df, order = 1L)
+
+cor(p.df$n.London_dist, dplyr::lag(p.df$n.London_dist), use = "complete.obs")
+```

### 2023-12-14 src/level_lisa_la.Rmd
local minor changes

@@
 ```
-## df for RF
-
-```{r}
-
-# London = Islington 
-help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
-  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)
-
-help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
-  mutate(cities = c("Leicester",
-                    "Bristol",
-                    "Manchester",
-                    "Liverpool",
-                    "Sheffield",
-                    "Birmingham",
-                    "Leeds",
-                    "London",
-                    "Edinburgh",
-                    "Glasgow"))
-
-dist$dist.city.name <- names(dist)[apply(dist[,1:9], MARGIN = 1, FUN = which.min)]
-help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
-  rename(nearest.city.LAD21CD = LAD21CD) %>% 
-  dplyr::select(ladcd, nearest.city.LAD21CD)
-
-knn.l <- nb2listw(knn)
-
-df <- df %>% drop_na() %>% group_by(year) %>% 
-  mutate(n.slag = lag.listw(knn.l, n)) %>% 
-  group_by(ladcd) %>% 
-  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year)) %>% 
-  left_join(help2, by = c("ladcd" = "ladcd"))
-
-df <- df %>% left_join(df %>% dplyr::select(ladcd,year, n) %>% 
-                         rename(n.nearest.city = n), 
-                       by = c("nearest.city.LAD21CD" = "ladcd",
-                              "year" = "year")) %>% 
-  dplyr::select(-c(LAD21NMW, OBJECTID, BNG_E, BNG_N, LONG, LAT, SHAPE_Length)) %>% 
-  relocate(LAD21NM, .after = ladcd)
-
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup)%>% dplyr::select(LAD21CD, RGN21CD)
-
-df <- df %>% left_join(lookup.region, by = c("ladcd" = "LAD21CD")) %>% 
-  relocate(RGN21CD, .after=year)
-```
-
-## RF
-
-### train
-
-```{r}
-library(caret)
-library(randomForest)
-
-# CV
-tc <- trainControl(method = "cv",
-                   number = 10,
-                   savePredictions = 'final')
-
-# London region: E12000007
-# training dataset
-tr.df <- df %>% ungroup() %>% filter(RGN21CD!="E12000007"|
-                                     RGN21CD!="E12000008")
-
-set.seed(71)
-m <- train(n ~ density + area + dist + dist.retail + London +
-             n.slag + year + n.nearest.city, #n.l.slag
-           data = tr.df,
-           trControl = tc,
-           method = "rf", 
-           importance = TRUE,
-           na.action = na.omit)
-summary(m)
-plot(m)
-m
-
-varimp_mars <- varImp(m)
-plot(varimp_mars, main="Variable Importance")
-```
-
-### test
-
-```{r}
-pred <- predict(m, df %>% ungroup() %>% filter(RGN21CD=="E12000008"), na.action = na.pass)
-length(pred)
-
-df %>% ungroup() %>% filter(RGN21CD=="E12000008") %>% filter(if_any(everything(), is.na))
-
-pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN21CD=="E12000008"))
-postResample(pred = pred$value, obs = pred$n) ## 0.99
-```
-## Firms
-
-```{r}
-firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>%
-  dplyr::select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
-  rename(ladcd = GEOGRAPHY_CODE,
-         firms = OBS_VALUE)
-
-df <- df %>% left_join(firms) %>% 
-  mutate(n_by_firm = n/firms) %>% # N websites/firm
-  #filter(!is.na(firms)) %>% 
-  dplyr::select(-n, -firms) %>% 
-  rename(n = n_by_firm)
-
-# df %>% filter(is.na(n_by_firm))
-# sapply(df %>% left_join(firms), function(x) sum(is.na(x)))
-```
-
-## time slices
-
-```{r}
-hepl <- df %>% createTimeSlices(initialWindow = 5,
-                        horizon = 1,
-                        fixedWindow = F)
-
-#Adding t-1 lag for n
-
-set.seed(71)
-m.test <- train(n ~ area + dist + dist.retail + London + #density
-             n.slag + year + n.nearest.city + n.l.slag,
-           data = df %>% filter(year < 1998),
-           trControl = tc,
-           method = "rf", 
-           importance = TRUE,
-           na.action = na.omit)
-summary(m.test)
-plot(m.test)
-m.test
-
-varimp_mars <- varImp(m.test)
-plot(varimp_mars, main="Variable Importance")
-
-pred <- predict(m.test, df %>% ungroup() %>% filter(year==1998), na.action = na.pass)
-length(pred)
-
-pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(year==1998))
-postResample(pred = pred$value, obs = pred$n) ## 0.99
-```

### 2023-12-14 abstracts_notes/abstract_GfR_2024/Tranos_GfR_2024.tex
local changes

@@
 \usepackage{hyperref}
-\usepackage[demo]{graphicx}
-%\usepackage{subfig}
-\usepackage{float}
 \usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
+\graphicspath{ {./} }
 
-%\usepackage[square,numbers]{natbib}
-\usepackage[round]{natbib} % remove number, use ()
+\usepackage[square,numbers]{natbib}
 \bibliographystyle{abbrvnat}
@@
 
-% remove date
-%\date{}
-
 % ========== Edit your name here
-\author[]{Professor Emmanouil Tranos}
-\affil[]{University of Bristol and The Alan Turing Institute}
+\author[1]{Prof Emmanouil Tranos}
+\affil[1]{School of Geographical Sciences, University of Bristol and The Alan Turing Institute}
 
@@
 \href{mailto:e.tranos@bristol.ac.uk}{e.tranos@bristol.ac.uk},
-\href{https://etranos.info/}{etranos.info},
-\href{https://twitter.com/EmmanouilTranos}{@EmmanouilTranos} \&
-\href{https://www.linkedin.com/in/emmanouil-tranos-42947812a/}{LinkedIn}
+\href{https://etranos.info/}{etranos.info}
+%\href{https://twitter.com/EmmanouilTranos}{@EmmanouilTranos} \&
+%\href{https://www.linkedin.com/in/emmanouil-tranos-42947812a/}{LinkedIn}
 \end{center}
@@
 
-% general, summary
-This paper maps the participation in the digital economy and its evolution in the UK over space and time.
+This paper maps and models the participation in the digital economy and its evolution in the UK over space and time.
 Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects.
@@
 Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK.
-Instead of metrics capturing the passive engagement with digital technologies, for instance internet subscription metrics, this paper targets the active engagement with the digital as reflected in website creation.
+Instead of using metrics capturing the passive engagement with digital technologies, for instance internet subscription metrics, this paper targets the active engagement with the digital as reflected in website creation.
 
@@
 This level of granularity differentiate this paper with previous approaches in the literature, which were only able to capture technological adoption at more coarse level.
-Thus, this paper tests how well established theoretical approaches regarding the diffusion of technologies are still applicable when the focus is on local scales.
+Thus, this paper tests how well established theoretical approaches regarding the diffusion of new technologies are still applicable when the focus is on local scales.
 Although we know that spatial contagion and urban hierarchies are key drivers of technological diffusion, such theoretical concepts have not been tested at small geographical scales.
@@
 The paper begins with simple analytical tools which reveal the scale-free character of the 'S curve' that defines the diffusion of new technologies and innovations.
-Indeed, such an 'S curve' describes the adoption of web technologies even at the level of local authorities in the UK.
-Then the paper continues with other exploratory approaches, which signal the spatial dependence of the diffusion on web technologies.
-In the last section, the paper employs state of the art machine learning techniques to predict the diffusion of web technologies over space and, importantly, highlight how different or not UK regions have been in terms of the adoption of web technologies.
+Indeed, such an 'S curve' describes the adoption of web technologies across scales and even at the level of local authorities in the UK (Figure \ref{fig:s}).
+Then, the paper continues with exploratory approaches, which signal the spatial dependence of the diffusion on web technologies.
+All the above provide useful insights for the last section, when the paper employs state of the art machine learning techniques to predict the diffusion of web technologies over space.
+Specifically, Random Forest models are trained at two different scales: local authorities (c.
+400 in the UK) and census output areas (c.
+230,000).
+The models are firstly trained and tested in all data using space and time sensitive \emph{k}-folds.
+This modelling exercise highlights the importance of the key spatial predictors as described in the literature in predicting technological diffusion.
+Then, these model are trained for each UK region separately and tested iteratively to all other regions in order obtain an understanding of how differently these spatial drivers affect the diffusion of web technologies across different places (Figures \ref{fig:la} and \ref{fig:oa}).
+In other words, this modelling strategy reveals how different or not UK regions have been in terms of the adoption of web technologies across scales.
+The granularity of the data used here allows to expose spatial differences that otherwise would have gone unnoticed.
+The results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.
 
-Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.
-
-\begin{figure}%
+\begin{figure}[h]
     \centering
-    \subfloat[\centering label 1]{{\includegraphics[width=5cm]{./lad_E07000031.png} }}%
-    \qquad
-    \subfloat[\centering label 2]{{\includegraphics[width=5cm]{./lad_E07000073} }}%
-    \caption{2 Figures side by side}%
-    \label{fig:example}%
+    \includegraphics[width=0.75\textwidth]{lad_E09000001}
+    \caption{Example of an S-curve for a local authority}
+    \label{fig:s}
 \end{figure}
 
-\begin{figure}[H]
+\begin{figure}[h]
     \centering
-	\begin{subfigure}{.3\textwidth}
-		\includegraphics[width=\textwidth]{lad_E07000031.png}
-		\caption{Atom 1}
-	\end{subfigure}
-%%%%%%%%%%%%%%
-	\begin{subfigure}{.3\textwidth}
-		\includegraphics[width=\textwidth]{lad_E07000031.png}
-		\caption{Atom 2}
-	\end{subfigure}
-%%%%%%%%%%%%%%
-	\begin{subfigure}{.3\textwidth}
-		\includegraphics[width=\textwidth]{lad_E07000031.png}
-		\caption{Atom 3}
-	\end{subfigure}
-	\caption{Atoms are fun!}
+    \includegraphics[width=0.9\textwidth]{heatmap_LA}
+    \caption{Regional similarities based on cross-region training/predictions at the Local Authority level}
+    \label{fig:la}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.75\textwidth]{heatmap_OA}
+    \caption{Regional similarities based on cross-region training/predictions at the Local Authority level}
+    \label{fig:oa}
 \end{figure}

### 2023-12-14 presentations/web_diffusion.Rmd
local changes

@@
 ---
-title: 'Digital economy in the UK: an evolutionary story'
+title: 'A multi-scalar story of the diffusion of a new technology: the web'
 # date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 
-<br>
-
 ::: {style="margin-left: 50px !important; text-align:left !important"}
 -   Introduction
+
 -   Web data
--   Spatial analysis
+
+-   Methods
+
+-   Results
+
     -   S-shaped diffusion curves
     -   Rank dynamics
+    -   Spatial analysis
+    -   Modelling
+
 -   Conclusions <br> <br>
@@
 
-## Aims
-
-<br>
-
--   Map the **active** engagement with the *digital*
--   Over time, **early stages** of the internet
--   **Granular** and **multi-scale** spatial perspective
-
-## Aims
+## Aim
 
@@
 ## Importance
-
--   Understand how the adoption of new technologies evolves
 
@@
 
+## Diffusion of a new *digital* technology
+
+<br>
+
+-   Diffusion of an intangible/digital technology
+
+-   Map the **active** engagement with the *digital*
+
+-   Over time, **early stages** of the internet
+
+-   **Granular** and **multi-scale** spatial perspective
+
 # Web data
@@
 
-<br>
-
--   Archived web data
+-   Data from the Internet Archive, the oldest web archive
 
@@
 
--   Geolocate:
+-   Geolocation: postcode references in the text
 
-    - to a unique postcode
+-   Timestamp: archival year
 
-    - to multiple postcodes
+-   Counts
 
--------------------------------------------
+## Web data: The Internet Archive
+
+![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture1.png){width="70%"}
+
+## Web data: The Internet Archive
+
+![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture2.png){width="70%"}
+
+## Long story short
+
+-   Data from a Web Archive -- The Internet Archive
+
+-   Observe commercial websites 1996 - 2012 in the UK (*.co.uk*)
+
+-   Geolocation: postcode references in the text
+
+-   Timestamp: archival year
+
+-   Counts
+
+------------------------------------------------------------------------
 
@@
 -   Time-stamp
-
-## Web data: The Internet Archive
-
-![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture1.png){width="70%"}
-
-## Web data: The Internet Archive
-
-![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture2.png){width="70%"}
 
@@
 
-![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
+![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
 
@@
 
-![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
+![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
 
-# Web diffusion
+# Methods
 
-## Mapping website density
+## Reminder: diffusion mechanisms
+
+-   **S-shaped** pattern in the cumulative level of adoption
+
+-   A **hierarchy effect**: from main centres to secondary ones
+
+-   A **neighborhood effect**: first "hitting" nearby locations
+
+## Methods
+
+-   Model cumulative adoption
+
+-   Descriptive statistics, ESDA & density regressions
+
+-   Modelling framework
+
+-   Two scales:
+
+    -   websites per firm in a Local Authority
+    -   websites in an Output Area
+
+# S-shaped diffusion curves
 
@@
 
-```{r heatmaps, results='asis', out.width = "60%"}
+```{r UK S, eval=T, out.width = "60%", out.height="60%"}
+#, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"
+
+path.image <- paste0(path, "/outputs/s_uk_per_firm.png")
+knitr::include_graphics(path.image)
+```
+
+```{r LA_S_gif, eval=F, results='asis', out.width = "60%"}
+path.images <- paste0(path, "/outputs/")
+
+library(magick)
+las <- list.files(path = path.images, pattern = "^lad_", full.names = T) #%>% 
+las_list <- lapply(las, image_read)
+
+## join the images together
+las_joined <- image_join(las_list)
+
+## animate at 2 frames per second
+las_animated <- image_animate(las_joined, fps = 2)
+
+## view animated image
+las_animated
+
+## save to disk
+image_write(image = las_animated,
+            path = "las.gif")
+
+  # map(image_read) %>% # reads each path file
+  # image_join() %>% # joins image
+  # image_animate(delay=50) #%>% # animates, can opt for number of loops
+  #image_write("test.gif") # write to current dir
+```
+
+------------------------------------------------------------------------
+
+```{r Xmid_frequency_graph, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/xmid_f.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r diffusion_speed_map, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/speed_pam.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+<!-- fast  -->
+
+```{r Darlington, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000005.png")
+knitr::include_graphics(path.image)
+```
+
+<!-- ------------------------------------------------------------------------ -->
+
+```{r Nottingham, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000018.png")
+knitr::include_graphics(path.image)
+```
+
+<!-- ------------------------------------------------------------------------ -->
+
+```{r Reading, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000038.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r City_of_London, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000001.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+<!-- slow -->
+
+```{r Middlesbrough, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000002.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r Leicester, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000016.png")
+knitr::include_graphics(path.image)
+```
+
+<!-- ------------------------------------------------------------------------ -->
+
+```{r Bristol, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000023.png")
+knitr::include_graphics(path.image)
+```
+
+<!-- ------------------------------------------------------------------------ -->
+
+```{r Plymouth, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000026.png")
+knitr::include_graphics(path.image)
+```
+
+<!-- ------------------------------------------------------------------------ -->
+
+```{r Liverpool, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E08000012.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r Hackney, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000012.png")
+knitr::include_graphics(path.image)
+```
+
+<!-- ------------------------------------------------------------------------ -->
+
+```{r Edinburgh, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_S12000036.png")
+knitr::include_graphics(path.image)
+```
+
+## Diffusion speed
+
+<br>
+
+-   Spatial heterogeneity
+
+-   Not a clear, easy to explain pattern
+
+# Rank dynamics: stability vs. volatility
+
+------------------------------------------------------------------------
+
+<br>
+
+-   Adoption heterogeneity
+
+-   Different perceptions of risk and economic returns from new
+    technologies
+
+-   Early adopters vs. laggards, leapfrogging
+
+------------------------------------------------------------------------
+
+```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only05.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only95.png")
+knitr::include_graphics(path.image)
+```
+
+## Rank dynamics and diffusion speed
+
+-   Spatial heterogeneity
+
+-   Expected volatility
+
+# Spatial mechanisms
+
+<!-- ------------------------------------------------------------------------ -->
+
+```{r, eval=FALSE, heatmaps, results='asis', out.width = "60%"}
 path.images <- paste0(path, "/outputs/")
@@
 -   Websites in an Output Area $i$
-
 :::
@@
 -   Nearest retail centre
-
 :::
@@
 -   Spatial dependency
-	- Relatively small and constant over time / scales
+    -   Relatively small, consistent over time / scales
     -   London hot spot early on
@@
     -   Granular analysis reveals other hotspots
-
 -   Distance effect
@@
 > **Hierarchy effect**: from main centers to secondary ones -- central
-places
+> places
 
@@
 
-# S-shaped diffusion curves
+# Putting all of these together
 
-------------------------------------------------------------------------
+## Modelling framework
 
-```{r UK S, eval=T}
-#, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"
+-   Random forests to predict $Website\,Density_{i,t}$
 
-path.image <- paste0(path, "/outputs/s_uk.png")
+-   4 sets of models:
+
+    -   All the data ⟹ *variable importance*
+    -   Train on one region, test on the rest ⟹ *spatial differences and
+        similarities* of diffusion mechanisms
+    -   For Local Authorities and Output Areas
+
+-   Space-time sensitive 10-fold CV (`CAST`)
+
+## Models trained on all data
+<br>
+
+|                   | RMSE  | RSquared | MAE   |
+|-------------------|-------|----------|-------|
+| Local Authorities | 0.022 | 0.908    | 0.012 |
+| Output Areas      | 3.613 | 0.627    | 0.533 |
+
+## Variable importance for Local Authorities
+
+```{r varimpLA, eval = T, out.height= "58%", out.width="58%", message=FALSE}
+
+path.image <- paste0(path, "/outputs/rf/varimp_LA.png")
 knitr::include_graphics(path.image)
@@
 
-```{r LA_S_gif, eval=F, results='asis', out.width = "60%"}
-path.images <- paste0(path, "/outputs/")
+## Variable importance for Output Areas
 
-library(magick)
-las <- list.files(path = path.images, pattern = "^lad_", full.names = T) #%>% 
-las_list <- lapply(las, image_read)
+```{r varimpOA, eval = T, out.height= "58%", out.width="58%", message=FALSE}
+#,  fig.height=5, fig.width=5,message=FALSE
 
-## join the images together
-las_joined <- image_join(las_list)
-
-## animate at 2 frames per second
-las_animated <- image_animate(las_joined, fps = 2)
-
-## view animated image
-las_animated
-
-## save to disk
-image_write(image = las_animated,
-            path = "las.gif")
-
-  # map(image_read) %>% # reads each path file
-  # image_join() %>% # joins image
-  # image_animate(delay=50) #%>% # animates, can opt for number of loops
-  #image_write("test.gif") # write to current dir
-```
-
-------------------------------------------------------------------------
-
-```{r Xmid_frequency_graph, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/xmid_f.png")
+path.image <- paste0(path, "/outputs/rf/varimp_OA.png")
 knitr::include_graphics(path.image)
@@
 
-------------------------------------------------------------------------
+## Regional similarities for Local Authorities
 
-```{r diffusion_speed_map, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/speed_pam.png")
+```{r similarity_LA, eval = T, message=FALSE}
+#fig.height=5, fig.width=5 
+#, out.height= "60%", out.width="60%"
+path.image <- paste0(path, "/outputs/rf/heatmap_LA.png")
 knitr::include_graphics(path.image)
@@
 
-------------------------------------------------------------------------
+## Regional similarities for Output Areas
 
-<!-- fast  -->
+```{r similarity_OA, eval = T, out.height= "70%", out.width="70%", message=FALSE}
 
-```{r Darlington, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000005.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r Nottingham, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000018.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r Reading, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000038.png")
-knitr::include_graphics(path.image)
-```
-
-------------------------------------------------------------------------
-
-```{r City_of_London, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000001.png")
-knitr::include_graphics(path.image)
-```
-
-------------------------------------------------------------------------
-
-<!-- slow -->
-
-```{r Middlesbrough, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000002.png")
-knitr::include_graphics(path.image)
-```
-
-------------------------------------------------------------------------
-
-```{r Leicester, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000016.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r Bristol, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000023.png")
-knitr::include_graphics(path.image)
-```
-
-------------------------------------------------------------------------
-
-```{r Plymouth, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000026.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r Liverpool, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E08000012.png")
-knitr::include_graphics(path.image)
-```
-
-------------------------------------------------------------------------
-
-```{r Hackney, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000012.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r Edinburgh, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_S12000036.png")
-knitr::include_graphics(path.image)
-```
-
-## Diffusion speed
-
-<br>
-
-- Spatial heterogeneity
-- Not clear pattern
-
-# Rank dynamics: stability vs. volatility
-
-------------------------------------------------------------------------
-
-<br>
-
-- Adoption heterogeneity 
-- Perceptions of risk and economic returns from new technologies
-- Early adopters vs. laggards, leapfrogging
-
-------------------------------------------------------------------------
-
-```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
-path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only05.png")
-knitr::include_graphics(path.image)
-```
-
-------------------------------------------------------------------------
-
-```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
-path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only95.png")
+path.image <- paste0(path, "/outputs/rf/heatmap_OA.png")
 knitr::include_graphics(path.image)
@@
 
-----
+------------------------------------------------------------------------
 
-<br>
+-   Established technological diffusion drivers still apply
 
--   **Geography matters:** spatial dependency, urban gravitation
+    -   for a digital technology
+    -   at local scales
+
+-   Geography matters: spatial dependency, urban gravitation
 
@@
 
--   Well-established theoretical approaches of diffusion survive even at
-    a granular level
+-   Stability and volatility: leapfrogging, early adopters dropping, but
+    also stable positions
 
--   Stability and volatility: leapfrogging, early adopters dropping, but also maintaining
+-   Spatially consistent mechanisms at local scale
+
+-   Heterogeneity increases with resolution
 

### 2023-12-21 src/oa_RF.Rmd
refined model LA
local minor changes
refined model OA

@@
 dist.retail <- distm(cbind(oa.gb.c$X, oa.gb.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail <- round((dist.retail/1000),2) %>% 
+dist.retail<- round((dist.retail/1000),2) %>% 
   as_tibble()
@@
 
+# This efficiently returns the name of the column with the shortest distance
+dist.retail <- dist.retail %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])
+
+# dist.retail_ <- dist.retail %>% slice_sample(n=100)
+# dist.retail_[,6425:6426]
+# dist.retail_ <- transform(dist.retail_, dist.retail = do.call(pmin, dist.retail_[2:6424])) 
+# sum(is.na(dist.retail_$dist.retail.name))
+
 # Minimum distance 
-dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
-dist.retail <- dist.retail %>% dplyr::select(oa11cd, dist.retail)
+dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[2:6424]))
+
+dist.retail <- dist.retail  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)
+# 232,296 
+#sapply(dist.retail, function(x) sum(is.na(x))) #0
+
+#spatial join retail centroids into OA
+retail.major.cetres.help <- st_centroid(retail) %>% 
+  st_transform(4326)
+
+# lookup obj with retail id and OA id
+help <- st_join(retail.major.cetres.help, oa.gb) %>% 
+  as_tibble %>% 
+  mutate(dist.retail.name = row_number(),
+         dist.retail.name = paste0("V", dist.retail.name)) %>% 
+  dplyr::select(id, dist.retail.name) %>% 
+  rename(oa11cd.retail = id)
+
+dist.retail <- dist.retail %>% 
+  left_join(help, by = c('dist.retail.name' = 'dist.retail.name')) %>% 
+  dplyr::select(-dist.retail.name)
+sapply(dist.retail, function(x) sum(is.na(x))) 
+# 1 retail centre is missing, because the centroid falls in the water
+
+dist.retail %>% filter(is.na(oa11cd.retail))
+missing.retail.centre <- retail.major.cetres.help %>% filter(RC_ID=="RC_EW_24")
+st_nearest_feature(missing.retail.centre, oa.gb)
+oa.gb[91986,] # missing OA: E00094765
+
+dist.retail <- dist.retail %>%
+  mutate(oa11cd.retail = ifelse(is.na(oa11cd.retail), "E00094765", oa11cd.retail))
+
+# dist.retail %>% distinct(dist.retail.name) #6405
+# help %>% distinct(dist.retail.name) # 6423 
+# dist.retail %>% distinct(oa11cd.retail) #6217 as there are multiple centres in an OA
 
@@
 sapply(df, function(x) sum(is.na(x)))
+df %>% filter(is.na(oa11cd.retail)) %>% distinct(oa11cd) #493 rows or 29 OA don't have oa11cd.retail
+
+# df <- df %>% dplyr::select(-dist.retail.x, -oa11cd.retail.x) %>% 
+#   rename(dist.retail = dist.retail.y,
+#          oa11cd.retail = oa11cd.retail.y)
 ```
@@
 
+help.retail <- dist.retail
+
 #####save this#####
@@
 
+# lags
 df <- df %>% 
@@
 
+# n nearest city
 df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
@@
                               "year" = "year")) 
+# n London 
+df <- df %>% left_join(df %>% 
+                         ungroup() %>% 
+                         filter(oa11cd == "E00176659") %>% 
+                         dplyr::select(year, n) %>% 
+                         rename(n.London = n)) %>% 
+  group_by(oa11cd) %>% 
+  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
+  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
+  ungroup()
+
+# n nearest retail
+df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
+                         rename(n.nearest.retail = n), 
+                       by = c("oa11cd.retail" = "oa11cd",
+                              "year" = "year")) %>% 
+  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
 
 # Lookup: OA to regions
+# NEED TO ADD NI. NOT A PROBLEM FOR NOW AS IT IS FIXED IN THE TEST SECTION
 path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
@@
 
-# df %>% ungroup() %>% distinct(RGN11NM)
-# sapply(df, function(x) sum(is.na(x)))
-
 # North/South
@@
 # For growth:
-#df <- df %>% mutate(growth = log(n/n.lag))
-
-sapply(df, function(x) sum(is.na(x)))
-
-# in.path <- paste0(path, "/data/temp/df_for_oa_rf_old.csv")
-# df.compare <- read_csv(in.path)
-
-# out.path <- paste0(path, "/data/temp/for_oa_rf.RData")
-# save.image(file = out.path)
-# load(out.path)
-
-df <- df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol",
-                        "Manchester", "Sheffield", "Leeds", "Edinburgh",
-                        "Leicester", "nearest.city.oa11cd"))
-
-# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
-# df <- read_csv(out.path)
-# df %>% write_csv(file = out.path)
-```
-
-## RF with CAST
-
-```{r}
+df <- df %>% mutate(growth = log(n/n.lag),
+                    abs.growth = n - n.lag)
+
+# regional identifier
 df <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
@@
 
-df <- df %>% filter(year > 1996)
-
-#df <- df %>% dplyr::select(-n.lag)
+sapply(df, function(x) sum(is.na(x)))
+
+# remove not used columns
+df <- df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol",
+                        "Manchester", "Sheffield", "Leeds", "Edinburgh",
+                        "Leicester", "nearest.city.oa11cd"))
+
+# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
+# df <- read_csv(out.path)
+# df %>% write_csv(file = out.path)
+
+```
+
+## RF with CAST
+
+```{r}
+london.help <- df %>% group_by(RGN11NM, year) %>% 
+  summarise(n.London = sum(n)) %>% 
+  filter(RGN11NM == "London") %>%
+  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
+  ungroup() %>% 
+  dplyr::select(-RGN11NM)
+
+#n.nearest.retail IS WRONG. FIX UPSTREAM
+df <- df %>% 
+  rename(oa11cd.retail = n.nearest.retail) %>% 
+  dplyr::select(-n.nearest.retail.lag) %>% 
+  left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
+                         rename(n.nearest.retail = n), 
+                       by = c("oa11cd.retail" = "oa11cd",
+                              "year" = "year")) %>% 
+  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
+
+df <- df %>% dplyr::select(-n.London, -n.London.lag) %>% 
+  left_join(london.help, by = c('year' = 'year'))
+
+# There are no lags for 1996, so it is removed
+df <- df %>% filter(year > 1996) %>% 
+  dplyr::select(oa11cd, year, RGN11CD, n, n.l.slag, n.nearest.city.lag, n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) #growth, abs.growth, RGN11NM,
+
+df %>% group_by(year) %>% 
+  filter(n == 0) %>% 
+  count(n)
+# in 2012, 120848 0s
+
+# df %>% dplyr::select(area, dist, dist.retail, London, south, n.slag, n.lag, n.l.slag, n.nearest.city, year) %>% 
+#   #sapply(function(x) sum(is.na(x)))
+#   summarise(across(where(is.numeric), .fns = 
+#                      list(#n = is.na,
+#                           min = min,
+#                           median = median,
+#                           mean = mean,
+#                           stdev = sd,
+#                           # q25 = ~quantile(., 0.25),
+#                           # q75 = ~quantile(., 0.75),
+#                           max = max))) %>%
+#   pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))
+# 
+# df %>% dplyr::select(n.London.lag, n.nearest.city.lag, n.nearest.retail.lag, n.l.slag, year, London, dist, dist.retail) %>% 
+#   #sapply(function(x) sum(is.na(x)))
+#   summarise(across(where(is.numeric), .fns = 
+#                      list(#n = is.na,
+#                           min = min,
+#                           median = median,
+#                           mean = mean,
+#                           stdev = sd,
+#                           # q25 = ~quantile(., 0.25),
+#                           # q75 = ~quantile(., 0.75),
+#                           max = max))) %>%
+#   pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))
+
 ```
@@
 #(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 11) # 9 after adding South
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # 9 after adding South. It should be RHS variables + 2
 
@@
 start.time <- Sys.time()
-model.all <- train(n ~ area + dist + dist.retail + London + south + 
-                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
+model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
+                     n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
                    data = df, 
@@
 time.taken <- round(end.time - start.time,2)
-time.taken
+time.taken #1.3h
 
@@
 #folds
-k <- 3 # 7.09h for k=3
+k <- 10 # 7.09h for k=3
 
@@
 detectCores()
-cl <- makePSOCKcluster(2)
+cl <- makePSOCKcluster(8)
 registerDoParallel(cl)

### 2024-01-05 src/oa_RF.Rmd
refined model OA

@@
 df <- df %>% filter(year > 1996) %>% 
-  dplyr::select(oa11cd, year, RGN11CD, n, n.l.slag, n.nearest.city.lag, n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) #growth, abs.growth, RGN11NM,
+  dplyr::select(oa11cd, year, RGN11CD, n, abs.growth, n.l.slag, n.nearest.city.lag, n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) #growth, abs.growth, RGN11NM,
 
@@
 # 
-# df %>% dplyr::select(n.London.lag, n.nearest.city.lag, n.nearest.retail.lag, n.l.slag, year, London, dist, dist.retail) %>% 
+# df %>% dplyr::select(n, n.London.lag, n.nearest.city.lag, n.nearest.retail.lag, n.l.slag, year, London, dist, dist.retail) %>%
 #   #sapply(function(x) sum(is.na(x)))
@@
 #   pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))
-
+# 
+# df %>% slice_max(order_by = n, n = 5, by = year)
 ```

### 2024-01-08 src/oa_RF.Rmd
refined model OA

@@
 
+### Cities as points
+
 First set up the destinations. Origins are all the OA.
@@
 Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)
+
+I use the centroids from arcgis API as they better reflect the city centres than 
+than the geometric centres of the boundaries, which I use for `sum(n)`
 
@@
 
-cities <- geo(city.names, no_query = F, method = "arcgis")
-
-# retail centres
+cities <- geo(city.names, no_query = F, method = "arcgis") %>% 
+  mutate(address = stringr::str_remove(address, ", UK"))
+
+# # Replace London with Inner London centroid
+# # Source: https://data.london.gov.uk/dataset/sub-regions-london-plan-consultation-2009
+# london.path <- paste0(path, "/data/raw/London_shape")
+# london.sf <- st_read(dsn = london.path, layer = "lp-consultation-oct-2009-subregions")
+# # Spatial transformations
+# london.sf <- st_transform(london.sf, 4326)  # EPSG code for WGS84
+# 
+# london.c <- london.sf %>% 
+#   filter(Name == "Central") %>% 
+#   st_centroid() %>% 
+#   st_coordinates() %>%
+#   as.data.frame() %>% 
+#   rename(lat = Y, long = X) %>% 
+#   mutate(address = "London, UK") %>% 
+#   dplyr::select(address, lat, long)
+# 
+# cities <- cities %>% bind_rows(london.c) %>% 
+#   slice(2:n()) %>% # to remove the old London centroid based on the cities geometry
+#   mutate(address = stringr::str_remove(address, ", UK"))
+```
+
+### City boundaries
+
+```{r}
+
+# cities
+city.names <- c(
+  "London", 
+  "Birmingham",
+  "Glasgow",
+  "Liverpool",
+  "Bristol",
+  "Manchester",
+  "Sheffield",
+  "Leeds",
+  "Edinburgh",
+  "Leicester"
+  )
+
+# City boundaries
+# Source: https://geoportal.statistics.gov.uk/search?collection=Dataset&sort=name&tags=all(BDY_TCITY%2CDEC_2015)
+path.in <- paste0(path, "/data/raw/Major_Towns_and_Cities_Dec_2015_Boundaries_V2_2022_121402779136482658.geojson")
+
+city.names <- city.names %>% stringr::str_remove(", UK")
+
+city.boundaries.ew <- st_read(path.in) %>% 
+  st_transform(4326) %>% 
+  filter(TCITY15NM %in% city.names) %>% 
+  dplyr::select(OBJECTID, TCITY15CD, TCITY15NM) %>% 
+  rename(code = TCITY15CD,
+         name = TCITY15NM)
+
+# Source: https://geoportal.statistics.gov.uk/
+path.in <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
+city.boundaries.sc <- st_read(path.geo) %>% 
+  st_transform(4326) %>% 
+  filter(grepl('Edinburgh|Glasgow', LAD21NM)) %>% 
+  dplyr::select(OBJECTID, LAD21CD, LAD21NM) %>% 
+  rename(code = LAD21CD,
+         name = LAD21NM)
+
+# merge
+city.boundaries <- rbind(city.boundaries.ew, city.boundaries.sc)
+
+```
+
+### Retail centres
+
+```{r}
+# retail centres <- 
 geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
 #retail <-readOGR(geo.path) 
-retail <- st_read(geo.path)
+retail <- st_read(geo.path) %>% st_transform(4326)  # EPSG code for WGS84
 
@@
 retail.major.cetres.help <- st_centroid(retail) %>% 
-  st_transform(4326) %>%
+  #st_transform(4326) %>%
   st_coordinates() %>% 
@@
   left_join(retail %>% rownames_to_column(var = "id"), by = 'id')
-```
-
-Distance to cities
+
+```
+
+### Distance to cities
 
@@
 
-city.names <- city.names %>% stringr::str_remove(", UK")
-names(dist) <- city.names
+names(dist) <- cities$address #city.names
 
@@
 
+# min distance
 dist <- transform(dist, dist = do.call(pmin, dist[-1]))
+
+# City name for minimum distance 
+dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
 
@@
 # calculate area
-oa.gb$area <- st_area(oa.gb) #/ 1000000
+#oa.gb$area <- st_area(oa.gb) #/ 1000000
 
@@
 df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
-  rename(oa11cd=id) %>% 
-  left_join(oa.gb, by = c("oa11cd" = "id")) %>% 
-  relocate(area, .after = n) %>% 
-  mutate(area = area / 1000000,
-         density = n / area) %>% 
-  relocate(density, .after = area) %>% 
-  dplyr::select(-geometry)
+  rename(oa11cd=id)  
+  # left_join(oa.gb, by = c("oa11cd" = "id")) %>% 
+  # relocate(area, .after = n) %>% 
+  # mutate(area = area / 1000000,
+  #        density = n / area) %>% 
+  # relocate(density, .after = area) %>% 
+  # dplyr::select(-geometry)
 
@@
 
-Distance to retail centres
+### Distance to retail centres
 
@@
 
-# retail.names <- retail.major.cetres$RC_Name
-# names(dist.retail) <- retail.names
+#retail.names <- retail.major.cetres$RC_ID
+names(dist.retail) <- retail.major.cetres$RC_ID #city.names
 
@@
 
-#spatial join retail centroids into OA
-retail.major.cetres.help <- st_centroid(retail) %>% 
-  st_transform(4326)
-
-# lookup obj with retail id and OA id
-help <- st_join(retail.major.cetres.help, oa.gb) %>% 
-  as_tibble %>% 
-  mutate(dist.retail.name = row_number(),
-         dist.retail.name = paste0("V", dist.retail.name)) %>% 
-  dplyr::select(id, dist.retail.name) %>% 
-  rename(oa11cd.retail = id)
-
-dist.retail <- dist.retail %>% 
-  left_join(help, by = c('dist.retail.name' = 'dist.retail.name')) %>% 
-  dplyr::select(-dist.retail.name)
-sapply(dist.retail, function(x) sum(is.na(x))) 
-# 1 retail centre is missing, because the centroid falls in the water
-
-dist.retail %>% filter(is.na(oa11cd.retail))
-missing.retail.centre <- retail.major.cetres.help %>% filter(RC_ID=="RC_EW_24")
-st_nearest_feature(missing.retail.centre, oa.gb)
-oa.gb[91986,] # missing OA: E00094765
-
-dist.retail <- dist.retail %>%
-  mutate(oa11cd.retail = ifelse(is.na(oa11cd.retail), "E00094765", oa11cd.retail))
-
-# dist.retail %>% distinct(dist.retail.name) #6405
-# help %>% distinct(dist.retail.name) # 6423 
-# dist.retail %>% distinct(oa11cd.retail) #6217 as there are multiple centres in an OA
+### TO DELETE
+
+# # spatial join retail centroids into OA
+# retail.major.cetres.help <- st_centroid(retail) %>% 
+#   st_transform(4326)
+# 
+# # lookup obj with retail id and OA id
+# help <- st_join(retail.major.cetres.help, oa.gb) %>% 
+#   as_tibble %>% 
+#   mutate(dist.retail.name = row_number(),
+#          dist.retail.name = paste0("V", dist.retail.name)) %>% 
+#   dplyr::select(id, dist.retail.name) %>% 
+#   rename(oa11cd.retail = id)
+# 
+# dist.retail <- dist.retail %>% 
+#   left_join(help, by = c('dist.retail.name' = 'dist.retail.name')) %>% 
+#   dplyr::select(-dist.retail.name)
+# sapply(dist.retail, function(x) sum(is.na(x))) 
+# # 1 retail centre is missing, because the centroid falls in the water
+# 
+# dist.retail %>% filter(is.na(oa11cd.retail))
+# missing.retail.centre <- retail.major.cetres.help %>% filter(RC_ID=="RC_EW_24")
+# st_nearest_feature(missing.retail.centre, oa.gb)
+# oa.gb[91986,] # missing OA: E00094765.
+# 
+# dist.retail <- dist.retail %>%
+#   mutate(oa11cd.retail = ifelse(is.na(oa11cd.retail), "E00094765", oa11cd.retail))
+# 
+# # dist.retail %>% distinct(dist.retail.name) #6405
+# # help %>% distinct(dist.retail.name) # 6423 
+# # dist.retail %>% distinct(oa11cd.retail) #6217 as there are multiple centres in an OA
 
@@
 df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd")) 
-
-sapply(df, function(x) sum(is.na(x)))
-df %>% filter(is.na(oa11cd.retail)) %>% distinct(oa11cd) #493 rows or 29 OA don't have oa11cd.retail
-
-# df <- df %>% dplyr::select(-dist.retail.x, -oa11cd.retail.x) %>% 
-#   rename(dist.retail = dist.retail.y,
-#          oa11cd.retail = oa11cd.retail.y)
-```
-
-## df for RF
-
-```{r}
-## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
-#df <- df %>% filter(!is.na(area))
-
-# London = Islington 
-cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
-cities.sf <- st_set_crs(cities.sf, 4326)
-
-# help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
-#   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
+#sapply(df, function(x) sum(is.na(x)))
+```
+
+## n for London, nearest city and retail
+
+### TODELETE
+
+```{r, eval=FALSE}
+# ## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
+# #df <- df %>% filter(!is.na(area)) 
+# 
+# cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
+# cities.sf <- st_set_crs(cities.sf, 4326)
+# 
+# # help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
+# #   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
+# 
+# sf_use_s2(FALSE)
+# help1 <- st_filter( oa.gb, cities.sf) %>% 
+#   #st_coordinates() %>% 
+#   as_tibble() %>% 
+#   dplyr::select(id) %>% 
+#   mutate(cities = c("London",     # The city names have been manually checked and assigned
+#                     "Leicester",
+#                     "Leeds",
+#                     "Sheffield",
+#                     "Bristol",
+#                     "Birmingham",
+#                     "Manchester",
+#                     "Liverpool",
+#                     "Edinburgh",
+#                     "Glasgow"))
+# 
+# # too slow
+# #dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
+# 
+# # This efficiently returns the name of the column with the shortest distance
+# dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
+# 
+# help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
+#   #rename(nearest.city.oa11cd = id) %>% 
+#   dplyr::select(oa11cd, dist.city.name)
+# 
+# df <- df %>%  left_join(help2, by = c("oa11cd" = "oa11cd")) 
+```
+
+### n for cities
+
+```{R}
+#help.retail <- dist.retail
+
+# n nearest city
+# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
+#                          rename(n.nearest.city = n), 
+#                        by = c("nearest.city.oa11cd" = "oa11cd",
+#                               "year" = "year"))
 
 sf_use_s2(FALSE)
-help1 <- st_filter( oa.gb, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
-  mutate(cities = c("Leicester",
-                    "Bristol",
-                    "Manchester",
-                    "Liverpool",
-                    "Sheffield",
-                    "Birmingham",
-                    "Leeds",
-                    "London",
-                    "Edinburgh",
-                    "Glasgow"))
-
-# too slow
-#dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
-
-# This efficiently returns the name of the column with the shortest distance
-dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
-
-help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
-  rename(nearest.city.oa11cd = id) %>% 
-  dplyr::select(oa11cd, nearest.city.oa11cd)
-
-help.retail <- dist.retail
-
-#####save this#####
-sf_use_s2(FALSE)
-oa.c <- st_centroid(oa.gb)
-kn <- knearneigh(oa.c, k = 5)
-knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
-knn.l <- nb2listw(knn)
-##################
-
-# lags
+city.boundaries.oa <- st_join(oa.gb, city.boundaries, join = st_intersects) %>% 
+  as_tibble() %>% 
+  filter(!is.na(name)) %>% 
+  dplyr::select(id, name) %>% 
+  mutate(name = ifelse(name == "City of Edinburgh", "Edinburgh", name)) %>% 
+  mutate(name = ifelse(name == "Glasgow City", "Glasgow", name))
+
+city.boundaries.help <- df %>% inner_join(city.boundaries.oa, by = c("oa11cd" = "id")) %>% 
+  group_by(name, year) %>% 
+  summarise(n.nearest.city = sum(n)) %>% 
+  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
+  ungroup()
+
+london.boundaries.help <- city.boundaries.help %>% 
+  filter(name == "London") %>% 
+  rename(n.London = n.nearest.city,
+         n.London.lag = n.nearest.city.lag)
+
+df <- df %>% left_join(city.boundaries.help, by = c("dist.city.name" = "name", "year" = "year"))
+
+df <- df %>% left_join(london.boundaries.help, by = c("year" = "year"))
+#sapply(df, function(x) sum(is.na(x)))
+```
+
+### DELETE n for London
+
+```{r, eval=FALSE}
+
+# # OLD London definitions
+# # df <- df %>% left_join(df %>% 
+# #                          ungroup() %>% 
+# #                          filter(oa11cd == "E00176659") %>% 
+# #                          dplyr::select(year, n) %>% 
+# #                          rename(n.London = n)) %>% 
+# #   group_by(oa11cd) %>% 
+# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
+# #   mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
+# #   ungroup() 
+# # 
+# # london.help <- df %>% group_by(RGN11NM, year) %>% 
+# #   summarise(n.London = sum(n)) %>% 
+# #   filter(RGN11NM == "London") %>%
+# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
+# #   ungroup() %>% 
+# #   dplyr::select(-RGN11NM) 
+# 
+# london.oa <- st_filter(oa.gb, london.sf %>% 
+#                          filter(Name == "Central")) %>% 
+#   as_tibble() %>% 
+#   dplyr::select(id)
+# 
+# london.help <- df %>% inner_join(london.oa, by = c("oa11cd" = "id")) %>% 
+#   group_by(year) %>% 
+#   summarise(n.London = sum(n)) %>% 
+#   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
+#   ungroup()
+# 
+# df <- df %>% left_join(london.help, by = c("year" = "year"))
+```
+
+### n for nearest retail
+
+```{r}
+# Old definition for retail centre
+# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
+#                          rename(n.nearest.retail = n), 
+#                        by = c("oa11cd.retail" = "oa11cd",
+#                               "year" = "year")) %>% 
+#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
+# 
+# 
+# #n.nearest.retail IS WRONG. FIX UPSTREAM
+# df <- df %>% 
+#   rename(oa11cd.retail = n.nearest.retail) %>% 
+#   dplyr::select(-n.nearest.retail.lag) %>% 
+#   left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
+#                          rename(n.nearest.retail = n), 
+#                        by = c("oa11cd.retail" = "oa11cd",
+#                               "year" = "year")) %>% 
+#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
+# 
+# 
+# df <- df %>% dplyr::select(-n.London, -n.London.lag) %>% 
+#   left_join(london.help, by = c('year' = 'year'))
+
+retail.oa <- st_join(oa.gb, retail, join = st_intersects) %>% 
+  as_tibble() %>% 
+  filter(!is.na(RC_ID)) %>% 
+  dplyr::select(id, RC_ID)
+
+# st_join(oa.gb, retail, join = st_is_within_distance, dist = 500) %>% 
+#   as_tibble() %>% 
+#   filter(!is.na(RC_ID)) %>% 
+#   dplyr::select(id, RC_ID)
+# 
+# st_filter(oa.gb, retail,  .pred = st_is_within_distance, dist = 100) %>% 
+#   as_tibble() %>% 
+#   filter(!is.na(RC_ID)) %>% 
+#   dplyr::select(id, RC_ID)
+
+# h <- oa.gb %>% st_intersects(retail)
+
+# retail.oa_ %>% filter(id == "E00004828"|id == "E00004829"|id == "E00004827")
+# # RC_EW_5285
+# retail.oa <- retail.oa %>% add_row(id = "E00004828", RC_ID = "RC_EW_5285")
+# retail.oa <- retail.oa %>% arrange(id)
+
+retail.help <- df %>% inner_join(retail.oa, by = c("oa11cd" = "id")) %>% 
+  group_by(RC_ID, year) %>% 
+  summarise(n.nearest.retail = sum(n)) %>% 
+  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail)) %>% 
+  ungroup()
+
+# retail.help2 <- retail.oa %>% left_join(retail.help, by = c("RC_ID" = "RC_ID")) %>% 
+#   dplyr::select(-RC_ID)
+
+df <- df %>% left_join(retail.help, by = c("dist.retail.name" = "RC_ID", "year" = "year"))
+#sapply(df_, function(x) sum(is.na(x)))
+```
+
+## TODO
+
+```{r}
+# Lookup: OA to regions
+# **FIXED< DELETE IN TEST** NEED TO ADD NI. NOT A PROBLEM FOR NOW AS IT IS FIXED IN THE TEST SECTION
+path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup) %>% dplyr::select(OA11CD, RGN11CD, RGN11NM)
+
+df <- df %>% left_join(lookup.region, by = c("oa11cd" = "OA11CD")) %>% 
+  relocate(c(RGN11NM, RGN11CD), .after=year) %>% 
+  mutate(help = substring(oa11cd, 1, 1),
+         RGN11CD = ifelse(help=="E", RGN11CD,
+                         ifelse(help=="W", "Wales",
+                                ifelse(help == "S", "Scotland", "NI"))),
+         RGN11NM = ifelse(help=="E", RGN11NM,
+                         ifelse(help=="W", "Wales",
+                                ifelse(help == "S", "Scotland", "NI")))) %>% 
+  dplyr::select(-help)
+
+# # North/South
+# #lookup.region %>% distinct(RGN21NM, .keep_all = T)
+# south <- c("London", "South West", "South West", "East of England")
+# df <- df %>% mutate(south = ifelse(RGN11NM %in% south, 1, 0))
+
+# For growth:
+df <- df %>% group_by(oa11cd) %>% 
+  mutate(n.lag = dplyr::lag(n),
+         growth = log(n/n.lag),
+                    abs.growth = n - n.lag) %>% 
+  relocate(n.lag, .after = n)
+
+# # regional identifier
+# df_ <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
+#   mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
+#   ungroup() %>% 
+#   dplyr::select(-help)
+
+#sapply(df, function(x) sum(is.na(x)))
+```
+
+## Spatial and spatio-temporal lags
+
+```{r}
+# sf_use_s2(FALSE)
+# oa.c <- st_centroid(oa.gb)
+# kn <- knearneigh(oa.c, k = 5)
+# knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
+# knn.l <- nb2listw(knn)
+
+path.out <- paste0(path, "/data/temp/oa_rf/all_OAs/knnl.RData")
+# save(knn.l, file = path.out)
+load(path.out)
+
 df <- df %>% 
@@
   mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
-         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
-  left_join(help2, by = c("oa11cd" = "oa11cd"))
-
-# n nearest city
-df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-                         rename(n.nearest.city = n), 
-                       by = c("nearest.city.oa11cd" = "oa11cd",
-                              "year" = "year")) 
-# n London 
-df <- df %>% left_join(df %>% 
-                         ungroup() %>% 
-                         filter(oa11cd == "E00176659") %>% 
-                         dplyr::select(year, n) %>% 
-                         rename(n.London = n)) %>% 
-  group_by(oa11cd) %>% 
-  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
-  ungroup()
-
-# n nearest retail
-df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-                         rename(n.nearest.retail = n), 
-                       by = c("oa11cd.retail" = "oa11cd",
-                              "year" = "year")) %>% 
-  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
-
-# Lookup: OA to regions
-# NEED TO ADD NI. NOT A PROBLEM FOR NOW AS IT IS FIXED IN THE TEST SECTION
-path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% dplyr::select(OA11CD, RGN11CD, RGN11NM)
-
-df <- df %>% left_join(lookup.region, by = c("oa11cd" = "OA11CD")) %>% 
-  relocate(c(RGN11NM, RGN11CD), .after=year) %>% 
-  mutate(help = substring(oa11cd, 1, 1),
-         RGN11CD = ifelse(help=="E", RGN11CD,
-                         ifelse(help=="W", "Wales",
-                                ifelse(help == "S", "Scotland", NA))),
-         RGN11NM = ifelse(help=="E", RGN11NM,
-                         ifelse(help=="W", "Wales",
-                                ifelse(help == "S", "Scotland", NA)))) %>% 
-  dplyr::select(-help)
-
-# North/South
-#lookup.region %>% distinct(RGN21NM, .keep_all = T)
-south <- c("London", "South West", "South West", "East of England")
-df <- df %>% mutate(south = ifelse(RGN11NM %in% south, 1, 0))
-
-# For growth:
-df <- df %>% mutate(growth = log(n/n.lag),
-                    abs.growth = n - n.lag)
-
-# regional identifier
-df <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
-  mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
-  ungroup() %>% 
-  dplyr::select(-help)
-
-sapply(df, function(x) sum(is.na(x)))
+         n.lag = dplyr::lag(n, n=1, order_by = year))        # duplicate line
+```
+
+## RF with CAST
+
+```{r}
+# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
+# df <- read_csv(out.path)
+# df %>% write_csv(file = out.path)
 
@@
                         "Manchester", "Sheffield", "Leeds", "Edinburgh",
-                        "Leicester", "nearest.city.oa11cd"))
-
-# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
-# df <- read_csv(out.path)
-# df %>% write_csv(file = out.path)
-
-```
-
-## RF with CAST
-
-```{r}
-london.help <- df %>% group_by(RGN11NM, year) %>% 
-  summarise(n.London = sum(n)) %>% 
-  filter(RGN11NM == "London") %>%
-  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-  ungroup() %>% 
-  dplyr::select(-RGN11NM)
-
-#n.nearest.retail IS WRONG. FIX UPSTREAM
-df <- df %>% 
-  rename(oa11cd.retail = n.nearest.retail) %>% 
-  dplyr::select(-n.nearest.retail.lag) %>% 
-  left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-                         rename(n.nearest.retail = n), 
-                       by = c("oa11cd.retail" = "oa11cd",
-                              "year" = "year")) %>% 
-  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
-
-df <- df %>% dplyr::select(-n.London, -n.London.lag) %>% 
-  left_join(london.help, by = c('year' = 'year'))
+                        "Leicester", "dist.city.name", "dist.retail.name", "name"))
 
@@
 df <- df %>% filter(year > 1996) %>% 
-  dplyr::select(oa11cd, year, RGN11CD, n, abs.growth, n.l.slag, n.nearest.city.lag, n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) #growth, abs.growth, RGN11NM,
+  dplyr::select(oa11cd, year, RGN11CD, n, abs.growth, n.l.slag, n.nearest.city.lag, 
+                n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) %>% #growth, abs.growth, RGN11NM, 
+  mutate(London.v = n.London.lag / London,
+         city.v = n.nearest.city.lag / dist,
+         retail.v = n.nearest.retail.lag / (dist.retail + .1)) %>% 
+  dplyr::select(-n.London.lag, -London, -n.nearest.city.lag, -dist, -n.nearest.retail.lag, -dist.retail)
 
@@
 #(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # 9 after adding South. It should be RHS variables + 2
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # It should be RHS variables + 2
 
@@
                      n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
+                     #London.v + city.v + retail.v + n.l.slag + year,
                    data = df, 
@@
 print(model.all)
+
 # Random Forest 
@@
 
-path.out <- paste0(path, "/outputs/rf/varimp_OA.png")
+path.out <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
 ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
@@
 
-path.out <- paste0(path, "/data/temp/oa_rf/all_OAs/all_OAs.RData")
+path.out <- paste0(path, "/outputs/rf/models/oa_rf/all_OAs/all_OAs.RData")
 save(model.all, file = path.out)
@@
 #(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 11) # 9 after adding South
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # It should be RHS variables + 2
 
@@
 
-  model.all <- train(n ~ area + dist + dist.retail + London + south + 
-                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
+  model.all <- train(n ~ n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
                    data = df %>% filter(RGN11CD==i), 
@@
 
-  file <- paste0(path, "/data/temp/oa_rf/regions/region_", as.character(i), ".RData")
+  file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
   save(model.all, file = file)

### 2024-01-13 src/oa_RF.Rmd
refined model OA

@@
 
-# Build OA for GB
-oa.gb <- rbind(oa.ew, oa.sc, oa.ni)
+# Build OA for UK
+oa.uk <- rbind(oa.ew, oa.sc, oa.ni)
 rm(oa.ew, oa.sc, oa.ni)
@@
 ```{r}
-#oa.sf <- st_as_sf(oa.gb)
-
-oa.full <- replicate(17, oa.gb %>% as_tibble(), simplify = FALSE) %>%
+#oa.sf <- st_as_sf(oa.uk)
+
+oa.full <- replicate(17, oa.uk %>% as_tibble(), simplify = FALSE) %>%
   bind_rows(.id = 'id_') %>% 
@@
 # oa centroids
-# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
-# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
+# oa.uk.c <- (rgeos::gCentroid(oa.uk, byid=TRUE))@coords %>% as.data.frame()
+# oa.uk.c$id <- row.names(oa.uk.c)  #as_tibble(rownames = NA) %>% 
 #   mutate(id = row.names())
 
-# oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
+# oa.uk.c <- gCentroid(oa.uk, byid=TRUE)@coords %>% 
 #   as.data.frame() %>% 
@@
 sf_use_s2(FALSE)
-oa.gb.c <- st_centroid(oa.gb) %>%
+oa.uk.c <- st_centroid(oa.uk) %>%
   st_coordinates() %>%
@@
   #add_rownames(var = "id") %>% 
-  bind_cols(oa.gb %>% dplyr::select(id))
+  bind_cols(oa.uk %>% dplyr::select(id))
 
@@
 "Edinburgh, UK",
-"Leicester, UK")
+"Leicester, UK",
+"Belfast, UK")
 
@@
 
-# cities
-city.names <- c(
-  "London", 
-  "Birmingham",
-  "Glasgow",
-  "Liverpool",
-  "Bristol",
-  "Manchester",
-  "Sheffield",
-  "Leeds",
-  "Edinburgh",
-  "Leicester"
-  )
+# # cities
+# city.names <- c(
+#   "London", 
+#   "Birmingham",
+#   "Glasgow",
+#   "Liverpool",
+#   "Bristol",
+#   "Manchester",
+#   "Sheffield",
+#   "Leeds",
+#   "Edinburgh",
+#   "Leicester",
+#   )
 
@@
 path.in <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-city.boundaries.sc <- st_read(path.geo) %>% 
+city.boundaries.sc.ni <- st_read(path.in) %>% 
   st_transform(4326) %>% 
-  filter(grepl('Edinburgh|Glasgow', LAD21NM)) %>% 
+  filter(grepl('Edinburgh|Glasgow|Belfast', LAD21NM)) %>% 
   dplyr::select(OBJECTID, LAD21CD, LAD21NM) %>% 
@@
 # merge
-city.boundaries <- rbind(city.boundaries.ew, city.boundaries.sc)
-
+city.boundaries <- rbind(city.boundaries.ew, city.boundaries.sc.ni)
 ```
@@
   left_join(retail %>% rownames_to_column(var = "id"), by = 'id')
-
 ```
@@
 ```{r}
-dist <- distm(cbind(oa.gb.c$X, oa.gb.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist <- round((dist/1000),2) %>% 
+# NI
+oa.uk.c.ni <- oa.uk.c %>% filter(str_detect(id, "^N"))
+
+dist.ni <- distm(cbind(oa.uk.c.ni$X, oa.uk.c.ni$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist.ni <- round((dist.ni/1000),2) %>% 
   as_tibble()
 
-names(dist) <- cities$address #city.names
-
-dist <- dist %>% bind_cols(oa.gb.c$id) %>% 
+names(dist.ni) <- cities$address #city.names
+
+dist.ni <- dist.ni %>% bind_cols(oa.uk.c.ni$id) %>% 
   rename(oa11cd = last_col()) %>% 
@@
 # min distance
-dist <- transform(dist, dist = do.call(pmin, dist[-1]))
+dist.ni <- transform(dist.ni, dist = do.call(pmin, dist.ni[-1]))
 
 # City name for minimum distance 
-dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
-
-# NOT USED AS IT SCREWS UP THE REGRESSIONS
-# Complete panel, N = 3871903, 17 year * 227759 OA
-# df <- df %>% bind_rows(dist %>% dplyr::select(oa11cd)) %>%
-#   mutate(year = ifelse(is.na(year),1000, year),
-#          country = substring(oa11cd, 1, 1)) %>%
-#   filter(country != "L" & country != "M" & country != "N") %>%
-#   filter(!is.na(country)) %>%
-#   complete(year, oa11cd, fill = list(n = 0)) %>%
-#   filter(year!=1000,
-#          !is.na(oa11cd)) %>%
-#   dplyr::select(-country)
-
-# tests
-# test$country <- substring(test$oa11cd, 1, 1)
-# unique(test$country)
-# length(unique(test$oa11cd))
+dist.ni <- dist.ni %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
+
+# GB
+oa.uk.c.gb <- oa.uk.c %>% filter(!str_detect(id, "^N"))
+
+dist.gb <- distm(cbind(oa.uk.c.gb$X, oa.uk.c.gb$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist.gb <- round((dist.gb/1000),2) %>% 
+  as_tibble()
+
+names(dist.gb) <- cities$address #city.names
+
+dist.gb <- dist.gb %>% bind_cols(oa.uk.c.gb$id) %>% 
+  rename(oa11cd = last_col()) %>% 
+  relocate(oa11cd)
+
+# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
+# dist$distMet <- apply(dist[,2:11], 1, min)
+
+# min distance
+dist.gb <- transform(dist.gb, dist = do.call(pmin, dist.gb[-1]))
+
+# City name for minimum distance 
+dist.gb <- dist.gb %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
+
+# UK
+dist <- rbind(dist.ni, dist.gb)
 
 # calculate area
-#oa.gb$area <- st_area(oa.gb) #/ 1000000
+#oa.uk$area <- st_area(oa.uk) #/ 1000000
 
@@
   rename(oa11cd=id)
-  # left_join(oa.gb, by = c("oa11cd" = "id")) %>% 
-  # relocate(area, .after = n) %>% 
-  # mutate(area = area / 1000000,
-  #        density = n / area) %>% 
-  # relocate(density, .after = area) %>% 
-  # dplyr::select(-geometry)
 
@@
 ```{r}
-dist.retail <- distm(cbind(oa.gb.c$X, oa.gb.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail<- round((dist.retail/1000),2) %>% 
+# dist.retail <- distm(cbind(oa.uk.c$X, oa.uk.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
+# dist.retail <- round((dist.retail/1000),2) %>% 
+#   as_tibble()
+
+# NI
+oa.uk.c.ni <- oa.uk.c %>% filter(str_detect(id, "^N"))
+dist.retail.ni <- distm(cbind(oa.uk.c.ni$X, oa.uk.c.ni$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
+dist.retail.ni <- round((dist.retail.ni/1000),2) %>% 
   as_tibble()
 
-#retail.names <- retail.major.cetres$RC_ID
-names(dist.retail) <- retail.major.cetres$RC_ID #city.names
-
-dist.retail <- dist.retail %>% bind_cols(oa.gb.c$id) %>% 
+names(dist.retail.ni) <- retail.major.cetres$RC_ID
+
+dist.retail.ni <- dist.retail.ni %>% bind_cols(oa.uk.c.ni$id) %>% 
   rename(oa11cd = last_col()) %>% 
@@
 # This efficiently returns the name of the column with the shortest distance
-dist.retail <- dist.retail %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])
+dist.retail.ni <- dist.retail.ni %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])
 
@@
 # Minimum distance 
-dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[2:6424]))
-
-dist.retail <- dist.retail  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)
-# 232,296 
-#sapply(dist.retail, function(x) sum(is.na(x))) #0
-
-### TO DELETE
-
-# # spatial join retail centroids into OA
-# retail.major.cetres.help <- st_centroid(retail) %>% 
-#   st_transform(4326)
-# 
-# # lookup obj with retail id and OA id
-# help <- st_join(retail.major.cetres.help, oa.gb) %>% 
-#   as_tibble %>% 
-#   mutate(dist.retail.name = row_number(),
-#          dist.retail.name = paste0("V", dist.retail.name)) %>% 
-#   dplyr::select(id, dist.retail.name) %>% 
-#   rename(oa11cd.retail = id)
-# 
-# dist.retail <- dist.retail %>% 
-#   left_join(help, by = c('dist.retail.name' = 'dist.retail.name')) %>% 
-#   dplyr::select(-dist.retail.name)
-# sapply(dist.retail, function(x) sum(is.na(x))) 
-# # 1 retail centre is missing, because the centroid falls in the water
-# 
-# dist.retail %>% filter(is.na(oa11cd.retail))
-# missing.retail.centre <- retail.major.cetres.help %>% filter(RC_ID=="RC_EW_24")
-# st_nearest_feature(missing.retail.centre, oa.gb)
-# oa.gb[91986,] # missing OA: E00094765.
-# 
-# dist.retail <- dist.retail %>%
-#   mutate(oa11cd.retail = ifelse(is.na(oa11cd.retail), "E00094765", oa11cd.retail))
-# 
-# # dist.retail %>% distinct(dist.retail.name) #6405
-# # help %>% distinct(dist.retail.name) # 6423 
-# # dist.retail %>% distinct(oa11cd.retail) #6217 as there are multiple centres in an OA
+dist.retail.ni <- transform(dist.retail.ni, dist.retail = do.call(pmin, dist.retail.ni[2:6424])) 
+dist.retail.ni <- dist.retail.ni  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)
+
+# GB
+oa.uk.c.gb <- oa.uk.c %>% filter(!str_detect(id, "^N"))
+dist.retail.gb <- distm(cbind(oa.uk.c.gb$X, oa.uk.c.gb$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
+dist.retail.gb <- round((dist.retail.gb/1000),2) %>% 
+  as_tibble()
+
+names(dist.retail.gb) <- retail.major.cetres$RC_ID
+
+dist.retail.gb <- dist.retail.gb %>% bind_cols(oa.uk.c.gb$id) %>% 
+  rename(oa11cd = last_col()) %>% 
+  relocate(oa11cd)
+
+# This efficiently returns the name of the column with the shortest distance
+dist.retail.gb <- dist.retail.gb %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])
+
+# dist.retail_ <- dist.retail %>% slice_sample(n=100)
+# dist.retail_[,6425:6426]
+# dist.retail_ <- transform(dist.retail_, dist.retail = do.call(pmin, dist.retail_[2:6424])) 
+# sum(is.na(dist.retail_$dist.retail.name))
+
+# Minimum distance 
+dist.retail.gb <- transform(dist.retail.gb, dist.retail = do.call(pmin, dist.retail.gb[2:6424])) 
+dist.retail.gb <- dist.retail.gb  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)
+
+# UK
+dist.retail <- rbind(dist.retail.ni, dist.retail.gb)
 
@@
 sf_use_s2(FALSE)
-city.boundaries.oa <- st_join(oa.gb, city.boundaries, join = st_intersects) %>% 
+city.boundaries.oa <- st_join(oa.uk, city.boundaries, join = st_intersects) %>% 
   as_tibble() %>% 
@@
 df <- df %>% left_join(london.boundaries.help, by = c("year" = "year"))
+
 #sapply(df, function(x) sum(is.na(x)))
@@
 
-retail.oa <- st_join(oa.gb, retail, join = st_intersects) %>% 
+retail.oa <- st_join(oa.uk, retail, 
+                     join = st_intersects) %>% #, 
+                    #largest = TRUE) %>% # returns the x features augmented with 
+                                         # the fields of y that have the largest 
+                                         # overlap with each of the features of x
   as_tibble() %>% 
   filter(!is.na(RC_ID)) %>% 
-  dplyr::select(id, RC_ID)
-
-# st_join(oa.gb, retail, join = st_is_within_distance, dist = 500) %>% 
+  dplyr::select(id, RC_ID) %>% 
+  add_row(id = "E00019779", RC_ID = "RC_EW_5951")
+
+# If `larget = FALSE`, we have multiple RC_IDs for OAs.
+# Also, RC_ID="RC_EW_5951" would fall in Thames as it is the London Bridge Tube Station.
+# To avoid NAs for for n.nearest.retail, I would have assigned dist.retail.name = RC_EW_1935 for E00019779
+
+# st_join(oa.uk, retail, join = st_is_within_distance, dist = 500) %>% 
 #   as_tibble() %>% 
@@
 # 
-# st_filter(oa.gb, retail,  .pred = st_is_within_distance, dist = 100) %>% 
+# st_filter(oa.uk, retail,  .pred = st_is_within_distance, dist = 100) %>% 
 #   as_tibble() %>% 
@@
 
-# h <- oa.gb %>% st_intersects(retail)
+# h <- oa.uk %>% st_intersects(retail)
 
@@
 
-retail.help <- df %>% inner_join(retail.oa, by = c("oa11cd" = "id")) %>% 
+retail.help <- df %>% full_join(retail.oa, by = c("oa11cd" = "id")) %>% 
   group_by(RC_ID, year) %>% 
@@
 df <- df %>% left_join(retail.help, by = c("dist.retail.name" = "RC_ID", "year" = "year"))
-#sapply(df_, function(x) sum(is.na(x)))
-```
+#sapply(df, function(x) sum(is.na(x)))
+```
+
+### Urban Rural TODO
 
@@
 # **FIXED< DELETE IN TEST** NEED TO ADD NI. NOT A PROBLEM FOR NOW AS IT IS FIXED IN THE TEST SECTION
+
+# # regional identifier
+# df_ <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
+#   mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
+#   ungroup() %>% 
+#   dplyr::select(-help)
+
+#sapply(df, function(x) sum(is.na(x)))
+```
+
+## Spatial and spatio-temporal lags
+
+```{r}
+# sf_use_s2(FALSE)
+# oa.c <- st_centroid(oa.uk)
+# kn <- knearneigh(oa.c, k = 5)
+# knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
+# knn.l <- nb2listw(knn)
+
+path.out <- paste0(path, "/data/temp/knnl.RData")
+# save(knn.l, file = path.out)
+load(path.out)
+
+df <- df %>% 
+  group_by(year) %>% 
+  mutate(n.slag = lag.listw(knn.l, n)) %>% 
+  group_by(oa11cd) %>% 
+  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
+         n.lag = dplyr::lag(n, n=1, order_by = year))        # duplicate line
+```
+
+## RF with CAST
+
+```{r}
+
+# Lookup: OA to regions
 path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
@@
 df <- df %>% group_by(oa11cd) %>% 
-  mutate(n.lag = dplyr::lag(n),
+  mutate(#n.lag = dplyr::lag(n),
          growth = log(n/n.lag),
@@
 
-# # regional identifier
-# df_ <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
-#   mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
-#   ungroup() %>% 
-#   dplyr::select(-help)
-
-#sapply(df, function(x) sum(is.na(x)))
-```
-
-## Spatial and spatio-temporal lags
-
-```{r}
-# sf_use_s2(FALSE)
-# oa.c <- st_centroid(oa.gb)
-# kn <- knearneigh(oa.c, k = 5)
-# knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
-# knn.l <- nb2listw(knn)
-
-path.out <- paste0(path, "/data/temp/oa_rf/all_OAs/knnl.RData")
-# save(knn.l, file = path.out)
-load(path.out)
-
-df <- df %>% 
-  group_by(year) %>% 
-  mutate(n.slag = lag.listw(knn.l, n)) %>% 
-  group_by(oa11cd) %>% 
-  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
-         n.lag = dplyr::lag(n, n=1, order_by = year))        # duplicate line
-```
-
-## RF with CAST
-
-```{r}
 # out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
@@
 # df %>% write_csv(file = out.path)
+
+# rm(list=setdiff(ls(), "df"))
+
+# # This is the project path
+# path <- find_rstudio_root_file()
 
@@
   dplyr::select(oa11cd, year, RGN11CD, n, abs.growth, n.l.slag, n.nearest.city.lag, 
-                n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) %>% #growth, abs.growth, RGN11NM, 
-  mutate(London.v = n.London.lag / London,
-         city.v = n.nearest.city.lag / dist,
-         retail.v = n.nearest.retail.lag / (dist.retail + .1)) %>% 
-  dplyr::select(-n.London.lag, -London, -n.nearest.city.lag, -dist, -n.nearest.retail.lag, -dist.retail)
+                n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) #%>% #growth, abs.growth, RGN11NM, 
+  # mutate(London.v = n.London.lag / London,
+  #        city.v = n.nearest.city.lag / dist,
+  #        retail.v = n.nearest.retail.lag / (dist.retail + .1)) %>% 
+  # dplyr::select(-n.London.lag, -London, -n.nearest.city.lag, -dist, -n.nearest.retail.lag, -dist.retail)
 
@@
 time.taken <- round(end.time - start.time,2)
-time.taken #1.3h
+time.taken #20.6h
 
@@
 #folds
-k <- 10 # 7.09h for k=3
+k <- 10 #
 
@@
 
-detectCores()
-cl <- makePSOCKcluster(8)
-registerDoParallel(cl)
+# detectCores()
+# cl <- makePSOCKcluster(2)
+# registerDoParallel(cl)
 
@@
 
-regions <- df %>% distinct(RGN11CD)
+regions <- df %>% ungroup() %>% distinct(RGN11CD) 
 regions <- as.vector(regions$RGN11CD)
@@
 for (i in regions){
+
+  print(paste0("start loop for ", i))
 
@@
   model.all <- train(n ~ n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
-                   data = df %>% filter(RGN11CD==i), 
+                   data = df %>% filter(RGN11CD!=i), 
                    trControl = tc,
@@
   file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
-  save(model.all, file = file)
+  save(model.all, file = file)  # file name = region not in the training data
   rm(model.all)
+
+  print(paste0("end loop for ", i))
+
 }
 
-stopCluster(cl)
+# stopCluster(cl)
 end.time <- Sys.time()
@@
 # create a list of the above model outputs
-filenames <- list.files(paste0(path, "/data/temp/oa_rf/regions"), pattern = "*.RData", full.names = T)
+filenames <- list.files(paste0(path, "/outputs/rf/models/oa_rf/regions"), pattern = "*.RData", full.names = T)
 
@@
   load(file = i)
-  i <- gsub("/home/nw19521/projects/web.diffusion/data/temp/oa_rf/regions/|.RData", "", i)
+  i <- gsub("/home/nw19521/projects/web.diffusion/outputs/rf/models/oa_rf/regions/|.RData", "", i)
   assign(i, model.all)  
@@
 ```{r}
-# train on one region, test on all the rest
+# train on all but one region, test on that region
 
@@
 for (i in regions){
-  for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], df[df$RGN11CD==j,])
+  #for (j in regions[!regions == i]){
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], 
+                    df[df$RGN11CD==i,])
     pred <- as.data.frame(pred)
     rownames(pred) <- c()
-    assign(paste0("region.predict.model.", i, ".on.", j), pred)
-
-    d <- paste0("region.predict.model.", i, ".on.", j)
-    d <- cbind(get(d), df %>% filter(RGN11CD==j)) #%>% arrange(year, ladcd)
+    assign(paste0("region.predict.model.", ".on.", i), pred)
+
+    d <- paste0("region.predict.model.", ".on.", i)
+    d <- cbind(get(d), df %>% filter(RGN11CD==i)) #%>% arrange(year, ladcd)
+
+    #***dplyr::select(RGN21CD, ladcd, year, growth, 1)***
+
     d <- d[,c(5, 2, 4, 6, 1)]     
@@
     df.predictions <- rbind(d, df.predictions)
-  }
+  #}
 }
@@
 
+### OLD
+
+```{r, eval=FALSE}
+# train on all but one region, test on that region
+
+df.predictions <- data.frame()
+for (i in regions){
+  for (j in regions[!regions == i]){
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], df[df$RGN11CD==j,])
+    pred <- as.data.frame(pred)
+    rownames(pred) <- c()
+    assign(paste0("region.predict.model.", i, ".on.", j), pred)
+
+    d <- paste0("region.predict.model.", i, ".on.", j)
+    d <- cbind(get(d), df %>% filter(RGN11CD==j)) #%>% arrange(year, ladcd)
+    d <- d[,c(5, 2, 4, 6, 1)]
+    colnames(d)[5] <- "predictions"
+    d <- d %>% mutate(trained.on = i)
+    df.predictions <- rbind(d, df.predictions)
+  }
+}
+
+df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN11CD, ".", trained.on))
+
+# split to list by region pair
+pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
+
+# calculate metrics for every region pair 
+rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
+                                                                     obs = x$n))  # CHANGE n WITH growth
+
+path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup) %>% 
+  dplyr::select(RGN21CD, RGN21NM) %>% 
+  distinct() %>% 
+  rename(RGN11CD = RGN21CD) %>% 
+  add_row(RGN11CD = c("Scotland", "N0", "Wales"), 
+          RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
+
+path.out <- paste0(path, "/outputs/rf/heatmap_OA.png")
+rf.year.all.metrics %>% 
+  as.data.frame() %>%
+  rownames_to_column(var = "metrics") %>% 
+  pivot_longer(!metrics, names_to = "train.test") %>% 
+  pivot_wider(names_from = metrics, values_from = value) %>% 
+  separate(train.test, c("train", "test"), remove = F) %>% 
+  left_join(lookup.region, by = c("train" = "RGN11CD")) %>% 
+  rename(train.region=RGN21NM) %>% 
+  left_join(lookup.region, by = c("test" = "RGN11CD")) %>% 
+  rename(test.region=RGN21NM) %>% 
+  arrange(Rsquared) %>% 
+  dplyr::select(train.region, test.region, Rsquared) %>% 
+  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
+  geom_tile() +
+  coord_fixed() +
+  labs(y = "", x = "") + #y: test.region, x: train.region
+  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
+  theme(#legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust=1),
+        panel.border = element_blank(), 
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.background = element_rect(fill = "white"))
+ggsave(path.out)
+
+  #summarise(r = range(Rsquared))
+  #round(2) %>%
+  #kable()
+
+```
+
 ## RF

### 2024-01-27 src/la_RF.Rmd
local minor changes
refined model OA
for bham presentation

@@
 ```{r, eval=T}
+
+# also saved locally in /data/raw/firm_data_backup.csv
+
 firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>%
@@
 # cities
-cities <- maps::world.cities %>% 
+cities.1 <- maps::world.cities %>% 
   filter(country.etc=="UK") %>% 
-  arrange(pop) %>% tail(10) 
-cities.sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
+  slice_max(pop, n = 10)
+cities.2 <- maps::world.cities %>% 
+  filter(country.etc=="UK") %>% 
+  filter(name == "Belfast")
+
+cities.sf <- bind_rows(cities.1, cities.2) %>% 
+  st_as_sf(coords = c("long", "lat"), crs = 4326)
 ```
@@
   relocate(density, .after = area)
+
+df <- df %>% filter(!is.na(London))  # to drop 34 for L99999999 and M99999999
+
 sapply(df, function(x) sum(is.na(x)))
@@
 
+#************** DOUBLE CHECK THE ORDER OF CITY NAMES BELOW **************
+
 # nearest city
@@
                     "London",
+                    "Belfast",
                     "Edinburgh",
@@
 # This efficiently returns the name of the column with the shortest distance
-dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
+dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
 help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
@@
 
-df <- df %>% drop_na() %>% # THIS DROPS L999999
+df <- df %>% 
   group_by(year) %>% 
@@
 
+# There are no lags for 1996, so it is removed
 df <- df %>% filter(year > 1996)
@@
 set.seed(123)
-indices <- CreateSpacetimeFolds(df %>% filter(year > 1999),                     # FILTER BY YEAR FOR GROWTH
+indices <- CreateSpacetimeFolds(df,# %>% filter(year > 1999),                   # FILTER BY YEAR FOR GROWTH
                                 spacevar = "ladcd",
@@
 start.time <- Sys.time()
-model.all <- train(growth ~                                                     # REPLACE n WITH growth
+model.all <- train(n ~                                                          # REPLACE n WITH growth
                      #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
-                     I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
-                   data = df %>% filter(year > 1999),                           # FILTER BY YEAR FOR GROWTH
+                     #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
+                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist, #n.nearest.retail.lag + dist.retail,
+                   data = df,# %>% filter(year > 1999),                         # FILTER BY YEAR FOR GROWTH
                    trControl = tc,
@@
 print(model.all)
+
+# [reproducible results]
 # Random Forest 
@@
 # 5984 samples
-#   10 predictor
+#    6 predictor
 # 
@@
 # Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 4718, 4690, 5040, 4718, 5055, 4704, ... 
+# Summary of sample sizes: 5055, 4690, 4732, 5070, 5040, 4718, ... 
 # Resampling results across tuning parameters:
@@
 #   mtry  splitrule   RMSE        Rsquared   MAE       
-#    2    variance    0.02194750  0.9015001  0.01162254
-#    2    extratrees  0.02245492  0.8984940  0.01229071
-#    6    variance    0.02208770  0.9025087  0.01175267
-#    6    extratrees  0.02165214  0.9081808  0.01161281
-#   10    variance    0.02447416  0.8859265  0.01326348
-#   10    extratrees  0.02204464  0.9064730  0.01184945
+#   2     variance    0.03227897  0.8023185  0.01991724
+#   2     extratrees  0.03150004  0.8057960  0.01912348
+#   4     variance    0.03271887  0.7985855  0.02025685
+#   4     extratrees  0.03174791  0.8047090  0.01932473
+#   6     variance    0.03358260  0.7879953  0.02080616
+#   6     extratrees  0.03207532  0.8018495  0.01964516
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 6, splitrule = extratrees and min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule = extratrees and min.node.size = 5.
 
@@
   mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
-                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
+                          ifelse(rowname == "n.London.lag", "London's web density, t-1", 
                                  ifelse(rowname == "London", "distance to London",
-                                        ifelse(rowname == "n.slag", "spatial lag of web density",
-                                               ifelse(rowname == "n.lag", "temporal lag of web density",
                                         ifelse(rowname == "n.l.slag", "spatial and temporal lag of web density",
-                                                             ifelse(rowname == "n.nearest.city", "web density in nearest city", rowname)))))))) %>% 
+                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's web density, t-1", rowname)))))) %>% 
   column_to_rownames()
 
-path.out <- paste0(path, "/outputs/rf/varimp_LA.png")
+path.out <- paste0(path, "/outputs/rf/figures/varimp_LA.png")
 ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
 ggsave(path.out)
-
-# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
-#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
-#                                                         "year")], 
-#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
-#                     trControl = tc,
-#                     method = "ranger",
-#                     #na.action = na.omit,
-#                     #importance = "impurity",
-#                     metric="Rsquared",
-#                     verbose=FALSE)
-# ffsmodel_LLO
-# ffsmodel_LLO$selectedvars
-# plot_ffs(ffsmodel_LLO)
 ```
@@
 set.seed(831)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # 4 feature + 2
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # 6 feature + 2
 
@@
 
-  model.all <- train(growth ~                                                   # REPLACE n WITH growth
+  model.all <- train(n ~                                                        # REPLACE n WITH growth
                        #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
-                       I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
+                       #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
+                       n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                    data = df %>% filter(RGN21CD!=i), #& year > 1999),           # FILTER BY YEAR FOR GROWTH
@@
 stopCluster(cl)
-end.time <- Sys.time() # ca. 2h
+end.time <- Sys.time() 
 time.taken <- round(end.time - start.time,2)
-time.taken
+time.taken # ca. 4 minutes 
 ```
@@
     d <- cbind(get(d), df %>% filter(RGN21CD==i)) #& year > 1999))                      # FILTER BY YEAR
-    d <- d %>% dplyr::select(RGN21CD, ladcd, year, growth, 1)                           # REPLACE n WITH growth
+    d <- d %>% dplyr::select(RGN21CD, ladcd, year, n, 1)                                # REPLACE n WITH growth
     colnames(d)[5] <- "predictions"
@@
 rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$growth))    # REPLACE n WITH growth
+                                                                     obs = x$n))       # REPLACE n WITH growth
 
@@
 
-path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
+path.out <- paste0(path, "/outputs/rf/figures/test_regions.csv")
 rf.year.all.metrics %>% 
@@
   left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
-  arrange(Rsquared)
-
-ggplot(aes(train.region, test.region, fill= Rsquared)) + 
-  geom_tile() +
-  coord_fixed() +
-  labs(y = "", x = "") + #y: test.region, x: train.region
-  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
-  theme(#legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust=1),
-        panel.border = element_blank(), 
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.background = element_rect(fill = "white"))
-ggsave(path.out)
+  arrange(Rsquared) %>% 
+  dplyr::select(RGN21NM, Rsquared) %>% 
+  rename(Region = RGN21NM) %>% 
+  write_csv(path.out)
+  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()
+
+# ggplot(aes(train.region, test.region, fill= Rsquared)) + 
+#   geom_tile() +
+#   coord_fixed() +
+#   labs(y = "", x = "") + #y: test.region, x: train.region
+#   scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
+#   theme(#legend.position = "none",
+#         axis.text.x = element_text(angle = 45, hjust=1),
+#         panel.border = element_blank(), 
+#         panel.grid.major = element_blank(),
+#         panel.grid.minor = element_blank(),
+#         panel.background = element_rect(fill = "white"))
+# ggsave(path.out)
 
@@
   #kable()
-
 ```

### 2024-01-27 src/oa_RF.Rmd
for bham presentation

@@
 # 3716736 samples
-#      10 predictor
+#       8 predictor
 # 
@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#    2    variance    3.865809  0.5809699  0.5313035
-#    2    extratrees  3.730542  0.6168954  0.5884731
-#    6    variance    3.761356  0.6063340  0.5366004
-#    6    extratrees  3.612670  0.6273041  0.5331196
-#   10    variance    3.929176  0.5775066  0.5449132
-#   10    extratrees  3.617598  0.6278302  0.5403551
+#   2     variance    4.999920  0.2053697  1.047293
+#   2     extratrees  5.167655  0.1540213  1.103048
+#   5     variance    5.187039  0.1710128  1.034382
+#   5     extratrees  5.169570  0.1557868  1.122845
+#   8     variance    5.633427  0.1247189  1.035171
+#   8     extratrees  5.182860  0.1563555  1.137332
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 6, splitrule =
-#  extratrees and min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule =
+#  variance and min.node.size = 5.
 
@@
                                  ifelse(rowname == "London", "distance to London",
-                                        ifelse(rowname == "n.slag", "spatial lag of web density",
-                                               ifelse(rowname == "n.lag", "temporal lag of web density",
+                                        ifelse(rowname == "n.nearest.retail.lag", "Nearest retail centre's web density, t-1",
+                                               ifelse(rowname == "n.London.lag", "London's web density, t-1",
                                                       ifelse(rowname == "n.l.slag", "spatial and temporal lag of web density",
-                                                             ifelse(rowname == "n.nearest.city", "web density in nearest city", rowname)))))))) %>% 
+                                                             ifelse(rowname == "n.nearest.city", "Nearest city's web density, t-1", rowname)))))))) %>% 
   column_to_rownames()
@@
   set.seed(71)
-  indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD==i), 
+  indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD!=i), 
                                 spacevar = "oa11cd",
@@
 
-  file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
+  #file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
+  file <- paste0("/hdd/tmp/regions/region_", as.character(i), ".RData")  
   save(model.all, file = file)  # file name = region not in the training data

### 2024-01-27 src/la_RF.Rmd
for bham presentation

@@
 
-path.out <- paste0(path, "/outputs/rf/figures/test_regions.csv")
+path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
 rf.year.all.metrics %>% 

### 2024-01-27 src/oa_RF.Rmd
for bham presentation

@@
                                  ifelse(rowname == "London", "distance to London",
-                                        ifelse(rowname == "n.nearest.retail.lag", "Nearest retail centre's web density, t-1",
-                                               ifelse(rowname == "n.London.lag", "London's web density, t-1",
-                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag of web density",
-                                                             ifelse(rowname == "n.nearest.city", "Nearest city's web density, t-1", rowname)))))))) %>% 
+                                        ifelse(rowname == "n.nearest.retail.lag", "Nearest retail centre's wensite density, t-1",
+                                               ifelse(rowname == "n.London.lag", "London's wensite density, t-1",
+                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag of wensite density",
+                                                             ifelse(rowname == "n.nearest.city", "Nearest city's wensite density, t-1", rowname)))))))) %>% 
   column_to_rownames()

### 2024-02-08 paper/submission1/tranos2023.qmd
bua

@@
 different extents in different places, and can be significantly
-influenced by policies [@victor1993]". @meade2021modelling highlight that 
+influenced by policies" [@victor1993]. @meade2021modelling highlight that 
 a variety of actors have a direct interest in gaining such knowledge 
@@
 and local authorities. 
-These processes and their effects very a lot across scales: although
+These processes and their effects vary a lot across scales: although
 the diffusion of a new technology might not be optimal at a local level,
@@
 in most of the following studies, the level of spatial detail is almost always 
-more coarse than the one adopted here. To begin with @beardsell1999spatial studied 
+more coarse than the one adopted here. @beardsell1999spatial studied 
 the evolution of the computer industry in $317$ US metro areas during the $1977-1992$ 

### 2024-02-08 presentations/web_diffusion.Rmd
bua

@@
     -   Rank dynamics
-    -   Spatial analysis
+    -   Exploratory spatial analysis
     -   Modelling
@@
 
--   Diffusion of an intangible/digital technology
+-   Diffusion of an **intangible**, digital technology [*web*]
 
@@
 
--   Over time, **early stages** of the internet
+-   Over time, **early stages** of the internet [*1996-2012*]
 
@@
 
-<small>
+<!-- <small> -->
 
--   Tranos, E., and C. Stich. 2020. Individual internet usage and the
-    availability of online content of local interest: A multilevel
-    approach. *Computers, Environment and Urban Systems*, 79:101371.
+<!-- -   Tranos, E., and C. Stich. 2020. Individual internet usage and the -->
 
--   Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy
-    in the UK: Regional productivity effects of early adoption.
-    *Regional Studies*, 55:12, 1924-1938.
+<!--     availability of online content of local interest: A multilevel -->
 
--   Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from
-    the ground up: a web data approach. *Environment and Planning B*, in
-    press.
+<!--     approach. *Computers, Environment and Urban Systems*, 79:101371. -->
 
--   Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to
-    predict regional trade flows: data extraction, modelling, and
-    validation, *Annals of the AAG*, in press.
+<!-- -   Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy -->
 
-</small>
+<!--     in the UK: Regional productivity effects of early adoption. -->
+
+<!--     *Regional Studies*, 55:12, 1924-1938. -->
+
+<!-- -   Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from -->
+
+<!--     the ground up: a web data approach. *Environment and Planning B*, in -->
+
+<!--     press. -->
+
+<!-- -   Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to -->
+
+<!--     predict regional trade flows: data extraction, modelling, and -->
+
+<!--     validation, *Annals of the AAG*, in press. -->
+
+<!-- </small> -->
 
@@
 
-20080509162138/<http://www.website1.co.uk/contact_us> IG8 8HD
+20080509162138 \| <http://www.website1.co.uk/contact_us> \| IG8 8HD
 
@@
 
--   Model cumulative adoption
+-   Cumulative adoption: Self-starting logistic growth model [`nls` and
+    `SSlogis`]
 
@@
 
--   Modelling framework
+-   Machine learning framework [random forests]
 
@@
 ```{r LA_S_gif, eval=F, results='asis', out.width = "60%"}
-path.images <- paste0(path, "/outputs/")
+path.images <- paste0(path, "/outputs/s_lad_per_firm/")
 
 library(magick)
-las <- list.files(path = path.images, pattern = "^lad_", full.names = T) #%>% 
+las <- list.files(path = path.images, pattern = "^lad_", full.names = T) 
+las <- sample(las, 50)
 las_list <- lapply(las, image_read)
@@
 ## animate at 2 frames per second
-las_animated <- image_animate(las_joined, fps = 2)
+las_animated <- image_animate(las_joined, fps = 4)
 
 ## view animated image
-las_animated
+#las_animated
 
 ## save to disk
-image_write(image = las_animated,
-            path = "las.gif")
+# image_write(image = las_animated,
+#             path = "las.gif")
+image_write_gif(las_animated, path = "las.gif")
 
@@
   #image_write("test.gif") # write to current dir
+```
+
+------------------------------------------------------------------------
+
+```{r gif S, eval=T, out.width = "60%", out.height="60%"}
+#, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"
+
+path.image <- paste0(path, "/presentations/las.gif")
+knitr::include_graphics(path.image)
 ```
@@
 
--   Website density regressions -- distance effect
+<!-- -   Website density regressions -- distance effect -->
 
@@
 
-## Website density regressions
+<!-- ## Website density regressions -->
 
-<br>
+<!-- <br> -->
 
-$$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$
+<!-- $$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$ -->
 
-<br>
+<!-- <br> -->
 
-::: {style="margin-left: 50px !important; text-align:left !important"}
-$Website\,Density_{i}$:
+<!-- ::: {style="margin-left: 50px !important; text-align:left !important"} -->
 
--   Websites per firm in a Local Authority $i$, *or*
+<!-- $Website\,Density_{i}$: -->
 
--   Websites in an Output Area $i$
-:::
+<!-- -   Websites per firm in a Local Authority $i$, *or* -->
 
-## Website density regressions
+<!-- -   Websites in an Output Area $i$ -->
 
-<br>
+<!-- ::: -->
 
-$$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$
+<!-- ## Website density regressions -->
 
-<br>
+<!-- <br> -->
 
-::: {style="margin-left: 50px !important; text-align:left !important"}
-$Place$:
+<!-- $$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$ -->
 
--   London, *or*
+<!-- <br> -->
 
--   Nearest city, *or*
+<!-- ::: {style="margin-left: 50px !important; text-align:left !important"} -->
 
--   Nearest retail centre
-:::
+<!-- $Place$: -->
 
-## Website density regressions
+<!-- -   London, *or* -->
 
-<br>
+<!-- -   Nearest city, *or* -->
 
-$\beta$ interpretation:
+<!-- -   Nearest retail centre -->
 
--   The lower the $\beta$ is (or the larger the $|\beta|$ is)...
+<!-- ::: -->
 
--   ... the larger urban gravitation is for web adoption.
+<!-- ## Website density regressions -->
 
-------------------------------------------------------------------------
+<!-- <br> -->
 
-```{r b_London_la, fig.height=5, fig.width=5, message=FALSE}
-path.image <- paste0(path, "/outputs/lisa/b_dist_london_la.png")
-knitr::include_graphics(path.image)
-```
+<!-- $\beta$ interpretation: -->
 
-------------------------------------------------------------------------
+<!-- -   The lower the $\beta$ is (or the larger the $|\beta|$ is)... -->
 
-```{r b_city_la, fig.height=5, fig.width=5, message=FALSE}
-#, out.width = "60%"
-path.image <- paste0(path, "/outputs/lisa/b_dist_city_la.png")
-knitr::include_graphics(path.image)
-```
+<!-- -   ... the larger urban gravitation is for web adoption. -->
 
-------------------------------------------------------------------------
+<!-- ------------------------------------------------------------------------ -->
 
-```{r b_centre_la,  fig.height=5, fig.width=5,message=FALSE}
-#, out.height= "60%", out.width="60%"
-path.image <- paste0(path, "/outputs/lisa/b_dist_centre_la.png")
-knitr::include_graphics(path.image)
-```
+<!-- ```{r b_London_la, fig.height=5, fig.width=5, message=FALSE} -->
 
-------------------------------------------------------------------------
+<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_london_la.png") -->
 
-```{r b_London_oa, eval=T, fig.height=5, fig.width=5, message=FALSE}
-#, out.width = "60%"
-path.image <- paste0(path, "/outputs/lisa/b_dist_london_oa.png")
-knitr::include_graphics(path.image)
-```
+<!-- knitr::include_graphics(path.image) -->
 
-------------------------------------------------------------------------
+<!-- ``` -->
 
-```{r b_centre_oa, eval = T,  fig.height=5, fig.width=5,message=FALSE}
-#fig.height=4, fig.width=4, 
-#, out.height= "60%", out.width="60%"
-path.image <- paste0(path, "/outputs/lisa/b_dist_city_oa.png")
-knitr::include_graphics(path.image)
-```
+<!-- ------------------------------------------------------------------------ -->
 
-------------------------------------------------------------------------
+<!-- ```{r b_city_la, fig.height=5, fig.width=5, message=FALSE} -->
 
-```{r b_city_oa, eval = T, fig.height=5, fig.width=5, message=FALSE}
-#, out.width = "60%"
-path.image <- paste0(path, "/outputs/lisa/b_dist_centre_oa.png")
-knitr::include_graphics(path.image)
-```
+<!-- #, out.width = "60%" -->
 
-------------------------------------------------------------------------
+<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_city_la.png") -->
 
-```{r r_squared, eval=T, results='asis'}
-# This is for the RSquared of the different website density regressions.
+<!-- knitr::include_graphics(path.image) -->
 
-path.in1 <- paste0(path, "/outputs/lisa/r2_all_oa.csv")
-path.in2 <- paste0(path, "/outputs/lisa/r2_all_la.csv")
+<!-- ``` -->
 
-# read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
-#   ggplot(aes(x=year)) +
-#   geom_line(aes(y=London, color = "OA to London"), size=1) +
-#   geom_line(aes(y=cities, color = "OA to nearest city"), size=1) +
-#   geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) +
-# 
-#   geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) +
-#   geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) +
-#   geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) +
-# 
-#   scale_color_discrete(name = "R-squared for regression:") +
-#   #scale_linetype_manual(name = "Website density regressions111") 
-#   ylab("R-squared") + 
-#   ggtitle("Website density regressions") +
-#   theme_minimal() +
-#   theme(axis.title.x=element_blank(),
-#         plot.title = element_text(hjust = 0.5))
+<!-- ------------------------------------------------------------------------ -->
 
-read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
-  pivot_longer(!year, names_to = "regression", values_to = "Rsquared") %>% 
-  mutate(LA = if_else(str_ends(regression, ".LA"), " to LA", "to OA"),
-         regression = str_remove_all(regression, ".LA"),
-         line = paste0(regression, " ", LA)) %>% 
-  ggplot() +
-  geom_line(aes(x=year, y = Rsquared, color = line),
-            size = 1) +
-  scale_color_discrete(name = "R-squared for regression:") +
-  #scale_linetype_manual(name = "Website density regressions111") 
-  ylab("R-squared") + 
-  ggtitle("Website density regressions") +
-  theme_minimal() +
-  theme(axis.title.x=element_blank(),
-        plot.title = element_text(hjust = 0.5))
+<!-- ```{r b_centre_la,  fig.height=5, fig.width=5,message=FALSE} -->
 
-#   scale_color_manual(values = c("red", "blue", "green"))
+<!-- #, out.height= "60%", out.width="60%" -->
 
-```
+<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_centre_la.png") -->
+
+<!-- knitr::include_graphics(path.image) -->
+
+<!-- ``` -->
+
+<!-- ------------------------------------------------------------------------ -->
+
+<!-- ```{r b_London_oa, eval=T, fig.height=5, fig.width=5, message=FALSE} -->
+
+<!-- #, out.width = "60%" -->
+
+<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_london_oa.png") -->
+
+<!-- knitr::include_graphics(path.image) -->
+
+<!-- ``` -->
+
+<!-- ------------------------------------------------------------------------ -->
+
+<!-- ```{r b_centre_oa, eval = T,  fig.height=5, fig.width=5,message=FALSE} -->
+
+<!-- #fig.height=4, fig.width=4,  -->
+
+<!-- #, out.height= "60%", out.width="60%" -->
+
+<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_city_oa.png") -->
+
+<!-- knitr::include_graphics(path.image) -->
+
+<!-- ``` -->
+
+<!-- ------------------------------------------------------------------------ -->
+
+<!-- ```{r b_city_oa, eval = T, fig.height=5, fig.width=5, message=FALSE} -->
+
+<!-- #, out.width = "60%" -->
+
+<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_centre_oa.png") -->
+
+<!-- knitr::include_graphics(path.image) -->
+
+<!-- ``` -->
+
+<!-- ------------------------------------------------------------------------ -->
+
+<!-- ```{r r_squared, eval=T, results='asis'} -->
+
+<!-- # This is for the RSquared of the different website density regressions. -->
+
+<!-- path.in1 <- paste0(path, "/outputs/lisa/r2_all_oa.csv") -->
+
+<!-- path.in2 <- paste0(path, "/outputs/lisa/r2_all_la.csv") -->
+
+<!-- # read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>%  -->
+
+<!-- #   ggplot(aes(x=year)) + -->
+
+<!-- #   geom_line(aes(y=London, color = "OA to London"), size=1) + -->
+
+<!-- #   geom_line(aes(y=cities, color = "OA to nearest city"), size=1) + -->
+
+<!-- #   geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) + -->
+
+<!-- #  -->
+
+<!-- #   geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) + -->
+
+<!-- #   geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) + -->
+
+<!-- #   geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) + -->
+
+<!-- #  -->
+
+<!-- #   scale_color_discrete(name = "R-squared for regression:") + -->
+
+<!-- #   #scale_linetype_manual(name = "Website density regressions111")  -->
+
+<!-- #   ylab("R-squared") +  -->
+
+<!-- #   ggtitle("Website density regressions") + -->
+
+<!-- #   theme_minimal() + -->
+
+<!-- #   theme(axis.title.x=element_blank(), -->
+
+<!-- #         plot.title = element_text(hjust = 0.5)) -->
+
+<!-- read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>%  -->
+
+<!--   pivot_longer(!year, names_to = "regression", values_to = "Rsquared") %>%  -->
+
+<!--   mutate(LA = if_else(str_ends(regression, ".LA"), " to LA", "to OA"), -->
+
+<!--          regression = str_remove_all(regression, ".LA"), -->
+
+<!--          line = paste0(regression, " ", LA)) %>%  -->
+
+<!--   ggplot() + -->
+
+<!--   geom_line(aes(x=year, y = Rsquared, color = line), -->
+
+<!--             size = 1) + -->
+
+<!--   scale_color_discrete(name = "R-squared for regression:") + -->
+
+<!--   #scale_linetype_manual(name = "Website density regressions111")  -->
+
+<!--   ylab("R-squared") +  -->
+
+<!--   ggtitle("Website density regressions") + -->
+
+<!--   theme_minimal() + -->
+
+<!--   theme(axis.title.x=element_blank(), -->
+
+<!--         plot.title = element_text(hjust = 0.5)) -->
+
+<!-- #   scale_color_manual(values = c("red", "blue", "green")) -->
+
+<!-- ``` -->
 
@@
     -   Granular analysis reveals other hotspots
--   Distance effect
-    -   Urban gravitation increases over time and then drops
-    -   Granular analysis: gravitation
-    -   Lost explanatory power over time
+        <!-- -   Distance effect -->
+        <!--     -   Urban gravitation increases over time and then drops -->
+        <!--     -   Granular analysis: gravitation -->
+        <!--     -   Lost explanatory power over time -->
 
@@
 
+-   [A **hierarchy effect**: from main centres to secondary ones]{style="color:orange;"}
+
+-   [A **neighborhood effect**: first "hitting" nearby locations]{style="color:olive;"}
+
+-   [**S-shaped** pattern in the cumulative level of adoption]{style="color:violet;"}
+
+<br> <small>
+$$Website\,Density_{t} \sim \color{orange}{Distance\,London} + 
+\color{orange}{Website\,density\,London_{t-1}} + \\
+\color{orange}{Distance\,Nearest\,City} + 
+\color{orange}{Website\,density\,Nearest\,City_{t-1}} + \\
+\color{orange}{Distance\,Nearest\,Retail_{i}} +
+\color{orange}{Website\,density\,Nearest\,Retail_{t-1}} +\\
+\color{olive}{W*\, Website\,density_{t-1}} +\\ 
+\color{violet}{year_{t}}$$ </small>
+
+## Modelling framework
+
 -   Random forests to predict $Website\,Density_{i,t}$
 
--   4 sets of models:
+-   2 spatial resolutions:
+
+    -   Local Authorities (websites per firm)
+    -   Output Areas (websites)
+
+-   2 sets of models:
 
     -   All the data ⟹ *variable importance*
-    -   Train on one region, test on the rest ⟹ *spatial differences and
-        similarities* of diffusion mechanisms
-    -   For Local Authorities and Output Areas
+    -   Train on all but one region, test on the holdout region ⟹
+        *spatial differences and similarities* of diffusion mechanisms
 
@@
 ## Models trained on all data
+
 <br>
@@
 |-------------------|-------|----------|-------|
-| Local Authorities | 0.022 | 0.908    | 0.012 |
-| Output Areas      | 3.613 | 0.627    | 0.533 |
+| Local Authorities | 0.032 | 0.810    | 0.019 |
+| Output Areas*      | 5.000 | 0.205    | 1.047 |
+
+<small>
+*No retail centre variables for the Local Authority models
+</small>
 
@@
 
-path.image <- paste0(path, "/outputs/rf/varimp_LA.png")
+path.image <- paste0(path, "/outputs/rf/figures/varimp_LA.png")
 knitr::include_graphics(path.image)
@@
 
-path.image <- paste0(path, "/outputs/rf/varimp_OA.png")
+path.image <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
 knitr::include_graphics(path.image)
@@
 ```{r similarity_LA, eval = T, message=FALSE}
-#fig.height=5, fig.width=5 
-#, out.height= "60%", out.width="60%"
-path.image <- paste0(path, "/outputs/rf/heatmap_LA.png")
-knitr::include_graphics(path.image)
+path.table <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
+read.csv(path.table) %>% 
+  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
+  kable()
 ```
@@
 
-```{r similarity_OA, eval = T, out.height= "70%", out.width="70%", message=FALSE}
+```{r similarity_OA, eval = F, out.height= "70%", out.width="70%", message=FALSE}
 
-path.image <- paste0(path, "/outputs/rf/heatmap_OA.png")
+path.image <- paste0(path, "/outputs/rf/figures/heatmap_OA.png")
 knitr::include_graphics(path.image)
+```
+
+## Alternative specification: growth, trained on all data
+
+<br> <small>
+$$Website\,Density\,Growth_{t} \sim \color{orange}{Distance\,London} + 
+\color{orange}{Website\,density\,London_{t-1}} + \\
+\color{orange}{Distance\,Nearest\,City} + 
+\color{orange}{Website\,density\,Nearest\,City_{t-1}} + \\
+\color{orange}{Distance\,Nearest\,Retail_{i}} +
+\color{orange}{Website\,density\,Nearest\,Retail_{t-1}} +\\
+\color{olive}{W*\, Website\,density_{t-1}} +\\ 
+\color{violet}{year_{t}}$$ </small>
+
+<br>
+
+|                   | RMSE  | RSquared | MAE   |
+|-------------------|-------|----------|-------|
+| Local Authorities* | 0.200 | 0.634    | 0.148 |
+
+<small>
+*No retail centre variables for the Local Authority models
+</small>
+
+## Alternative specification: growth, trained on all data
+
+```{r growth LA}
+path.table <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv")
+read.csv(path.table) %>% 
+  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
+  kable()
 ```
@@
 
--   Some indications of a hierarchical diffusion
+-   Hierarchical diffusion
 
@@
 
+-   Things to do: OA?
+
 ## References {.allowframebreaks}

### 2024-02-08 src/bua_RF.Rmd
bua

--- 2024-01-27 src/oa_RF.Rmd
@@
 ---
-title: "Random Forests", OAs
+title: "Random Forests", BUAs
 date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 ---
+
+TODO:
+
+- check missing bua
+- check out
+- regions
+- growth
 
@@
 
-# tests for Scotland
-# df %>% filter(substr(oa11cd, 1,1) =="S",
-#               year==2010) %>% summarise(mean(n, na.rm = T))
-# 
-# df %>% filter(oa11cd=="S00090182")
-# 
-# test@data %>% filter(substr(id, 1,1) =="S",
-#                      n!=0)
-# 
-# df %>% filter(substr(oa11cd, 1,1) =="W",
-#               year==2003)
-```
-
-## spatial data OLD
-
-```{r eval=F}
-# get OA for England and Wales
-path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
-oa.ew <- readOGR(path.geo)
-# source: https://geoportal.statistics.gov.uk/
-
-# spatial transformations
-oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))
-
-# keep in the data slot only the ONS Output Area id, renaming it as 'id'
-oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
-colnames(oa.ew@data) <- c('id', 'area')
-
-# reassign the polygon IDs
-oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))
-
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.ew)
-
-# # get OA for Scotland
-# path.geo.sc <- paste0(path, "/data/raw/SG_DataZoneBdry_2011")
-# oa.sc <- readOGR(dsn=path.geo.sc, layer = "SG_DataZone_Bdry_2011")
-# # source: https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011
-# 
-# # spatial transformations
-# oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-# 
-# # Scotland (follows same steps as EW, see notes above)
-# oa.sc <- oa.sc[, 'DataZone']
-# colnames(oa.sc@data) <- c('id')
-# 
-# # reassign the polygon IDs
-# oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-# 
-# # check the CRS has changed correctly, and the data slot has shrink to only the ID
-# summary(oa.sc)
-
-path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
-oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
-# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries
-
-# spatial transformations
-oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-
-# Scotland (follows same steps as EW, see notes above)
-oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
-colnames(oa.sc@data) <- c('id', 'area')
-
-# reassign the polygon IDs
-oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.sc)
-
-# build oa for GB
-oa.gb <- maptools::spRbind(oa.ew, oa.sc)
-rm(oa.ew, oa.sc)
-
-# oa.gb$geometry <- oa.gb$geometry %>%
-#   s2::s2_rebuild() %>%
-#   sf::st_as_sfc()
-
-# get GB
-path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
-gb <- readOGR(path.gb)
-# spatial transformations
-gb <- spTransform(gb, CRS("+init=epsg:4326"))
-```
-
-## spatial data NEW
-
-```{r}
-library(sf)
+# temp for working on laptop
+# path.in <- paste0(path, "/data/temp/help.oa.csv")
+# df <- read_csv(path.in)
+
+# https://geoportal.statistics.gov.uk/datasets/df834d6d654d4f4a9a211fdbbb8893bd_0/explore
+# path.lookup.rg <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")
+# lookup.rg <- read_csv(path.lookup.rg) %>% 
+#   dplyr::select(-ObjectId)
+```
+
+## spatial data
+
+```{r}
 
@@
 
-# NI
-path.geo.ni <- paste0(path, "/data/raw/ni_small_area/")
-oa.ni <- st_read(dsn = path.geo.ni, layer = "SA2011")
-oa.ni <- st_transform(oa.ni, 4326)  # EPSG code for WGS84
-oa.ni <- oa.ni %>% dplyr::select(SA2011, Hectares) %>% 
-  rename(id= SA2011,
-         area = Hectares)
-
-# Build OA for UK
-oa.uk <- rbind(oa.ew, oa.sc, oa.ni)
-rm(oa.ew, oa.sc, oa.ni)
-
-# Get GB
-path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
-gb <- st_read(path.gb)
-
-# Spatial transformations
-gb <- st_transform(gb, 4326)  # EPSG code for WGS84
+# Build OA for GB
+oa.gb <- rbind(oa.ew, oa.sc)
+rm(oa.ew, oa.sc)
+
+# # Get GB
+# path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
+# gb <- st_read(path.gb)
+# 
+# # Spatial transformations
+# gb <- st_transform(gb, 4326)  # EPSG code for WGS84
 ```
@@
 
-oa.full <- replicate(17, oa.uk %>% as_tibble(), simplify = FALSE) %>%
+oa.full <- replicate(17, oa.gb %>% as_tibble(), simplify = FALSE) %>%
   bind_rows(.id = 'id_') %>% 
@@
 
-## Distances to cities and retail centres
-
-### Cities as points
-
-First set up the destinations. Origins are all the OA.
-
-Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
-Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)
-
-I use the centroids from arcgis API as they better reflect the city centres than 
-than the geometric centres of the boundaries, which I use for `sum(n)`
-
-```{r}
-# oa centroids
-# oa.uk.c <- (rgeos::gCentroid(oa.uk, byid=TRUE))@coords %>% as.data.frame()
-# oa.uk.c$id <- row.names(oa.uk.c)  #as_tibble(rownames = NA) %>% 
-#   mutate(id = row.names())
-
-# oa.uk.c <- gCentroid(oa.uk, byid=TRUE)@coords %>% 
-#   as.data.frame() %>% 
-#   add_rownames(var = "id")
+## OA to BUA
+
+```{r}
+# long to wide
+df.w <- df %>% pivot_wider(names_from = year, values_from = n)
+
 sf_use_s2(FALSE)
-oa.uk.c <- st_centroid(oa.uk) %>%
+oa.gb.c <- st_centroid(oa.gb) %>%
+  left_join(df.w)
+
+# BUA
+# source: https://geoportal.statistics.gov.uk/datasets/ad30b234308f4b02b4bb9b0f4766f7bb_0/explore
+bua.path <- paste0(path, "/data/raw/BUA_2022_GB_-8042218937152150708.geojson")
+bua <- st_read(bua.path)
+bua <- st_transform(bua, 4326)  # EPSG code for WGS84
+
+# OA fall within a BUA
+df.within <- st_join(oa.gb.c, bua, join = st_intersects) %>% #, join = st_intersects
+  as_tibble()
+
+df.out <- df.within %>% filter(is.na(OBJECTID))
+df.within <- df.within %>% filter(!is.na(OBJECTID))
+
+# help <- st_is_within_distance(oa.gb.c %>% filter(id %in% df.out$id), bua, dist=500) %>% #, join = st_intersects
+#   as_tibble()
+df.out <- st_join(oa.gb.c %>% filter(id %in% df.out$id), bua, join = st_nearest_feature) %>% #, join = st_intersects
+  as_tibble()
+
+df <- bind_rows(df.within, df.out) %>% # 142953
+  dplyr::select(BUA22CD, BUA22NM, 3:19) %>% 
+  pivot_longer(!c(BUA22CD, BUA22NM), names_to = "year", values_to = "n") %>% 
+  group_by(BUA22CD, year) %>% 
+  summarise(n = sum(n)) %>% 
+  ungroup() %>% 
+  left_join(bua %>% as_tibble() %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
+  relocate(BUA22NM, .after = BUA22CD) %>% 
+  mutate(year = as.numeric(year))
+
+# # WE ARE MISSING 42855 websites were in OA not in a BUA
+# # The above might help but takes a long time
+# df <- st_join(oa.gb.c, bua, join = st_intersects) %>% # 94265
+#   as_tibble() %>% 
+#   dplyr::select(BUA22CD, BUA22NM, 3:19) %>% 
+#   pivot_longer(!c(BUA22CD, BUA22NM), names_to = "year", values_to = "n") %>% 
+#   group_by(BUA22CD, year) %>% 
+#   summarise(n = sum(n)) %>% 
+#   ungroup() %>% 
+#   left_join(bua %>% as_tibble() %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
+#   relocate(BUA22NM, .after = BUA22CD) %>% 
+#   mutate(year = as.numeric(year))
+
+sapply(df, function(x) sum(is.na(x)))  
+# df <- df %>% filter(is.na(BUA22CD))
+# 42855 websites were in OA not in a BUA (3,536,635 - 3,493,780)
+```
+
+## Distances to cities
+
+### City BUA
+
+```{r}
+
+city.names <- c("City of Westminster",
+"Birmingham",
+"Glasgow",
+"Liverpool",
+"Bristol",
+"Manchester",
+"Sheffield",
+"Leeds",
+"Edinburgh",
+"Leicester")
+
+#cities.bua <- df %>% filter(BUA22NM %in% city.names) %>% distinct(BUA22CD)
+cities.bua <- bua %>% filter(BUA22NM %in% city.names)
+
+cities.bua.c <- cities.bua %>% filter(BUA22NM %in% city.names) %>% 
+  st_centroid() %>%
   st_coordinates() %>%
@@
   #add_rownames(var = "id") %>% 
-  bind_cols(oa.uk %>% dplyr::select(id))
-
-# cities
-city.names <- c("London, UK", 
-"Birmingham, UK",
-"Glasgow, UK",
-"Liverpool, UK",
-"Bristol, UK",
-"Manchester, UK",
-"Sheffield, UK",
-"Leeds, UK",
-"Edinburgh, UK",
-"Leicester, UK",
-"Belfast, UK")
-
-cities <- geo(city.names, no_query = F, method = "arcgis") %>% 
-  mutate(address = stringr::str_remove(address, ", UK"))
-
-# # Replace London with Inner London centroid
-# # Source: https://data.london.gov.uk/dataset/sub-regions-london-plan-consultation-2009
-# london.path <- paste0(path, "/data/raw/London_shape")
-# london.sf <- st_read(dsn = london.path, layer = "lp-consultation-oct-2009-subregions")
-# # Spatial transformations
-# london.sf <- st_transform(london.sf, 4326)  # EPSG code for WGS84
-# 
-# london.c <- london.sf %>% 
-#   filter(Name == "Central") %>% 
-#   st_centroid() %>% 
-#   st_coordinates() %>%
-#   as.data.frame() %>% 
-#   rename(lat = Y, long = X) %>% 
-#   mutate(address = "London, UK") %>% 
-#   dplyr::select(address, lat, long)
-# 
-# cities <- cities %>% bind_rows(london.c) %>% 
-#   slice(2:n()) %>% # to remove the old London centroid based on the cities geometry
-#   mutate(address = stringr::str_remove(address, ", UK"))
-```
-
-### City boundaries
-
-```{r}
-
-# # cities
-# city.names <- c(
-#   "London", 
-#   "Birmingham",
-#   "Glasgow",
-#   "Liverpool",
-#   "Bristol",
-#   "Manchester",
-#   "Sheffield",
-#   "Leeds",
-#   "Edinburgh",
-#   "Leicester",
-#   )
-
-# City boundaries
-# Source: https://geoportal.statistics.gov.uk/search?collection=Dataset&sort=name&tags=all(BDY_TCITY%2CDEC_2015)
-path.in <- paste0(path, "/data/raw/Major_Towns_and_Cities_Dec_2015_Boundaries_V2_2022_121402779136482658.geojson")
-
-city.names <- city.names %>% stringr::str_remove(", UK")
-
-city.boundaries.ew <- st_read(path.in) %>% 
-  st_transform(4326) %>% 
-  filter(TCITY15NM %in% city.names) %>% 
-  dplyr::select(OBJECTID, TCITY15CD, TCITY15NM) %>% 
-  rename(code = TCITY15CD,
-         name = TCITY15NM)
-
-# Source: https://geoportal.statistics.gov.uk/
-path.in <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-city.boundaries.sc.ni <- st_read(path.in) %>% 
-  st_transform(4326) %>% 
-  filter(grepl('Edinburgh|Glasgow|Belfast', LAD21NM)) %>% 
-  dplyr::select(OBJECTID, LAD21CD, LAD21NM) %>% 
-  rename(code = LAD21CD,
-         name = LAD21NM)
-
-# merge
-city.boundaries <- rbind(city.boundaries.ew, city.boundaries.sc.ni)
-```
-
-### Retail centres
-
-```{r}
-# retail centres <- 
-geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
-#retail <-readOGR(geo.path) 
-retail <- st_read(geo.path) %>% st_transform(4326)  # EPSG code for WGS84
-
-#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
-# retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
-#   as.data.frame() %>% 
-#   add_rownames(var = "id") #%>%
-retail.major.cetres.help <- st_centroid(retail) %>% 
-  #st_transform(4326) %>%
+  bind_cols(cities.bua %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
+  dplyr::select(BUA22CD, BUA22NM, X, Y)
+
+bua.c <- bua %>% 
+  st_centroid() %>%
   st_coordinates() %>%
   as.data.frame() %>%
-  add_rownames(var = "id")
-
-# retail.major.cetres <- retail.major.cetres.help %>% 
-#   st_as_sf(coords = c("x", "y"), crs = 27700) %>%
-#   st_transform(4326) %>%
-#   st_coordinates() %>%
-#   as_tibble() %>% 
-#   bind_cols(retail.major.cetres.help$id) %>% 
-#   rename(id = '...3') %>% 
-#   left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
-retail.major.cetres <- retail.major.cetres.help %>% 
-  # as_tibble() %>% 
-  # bind_cols(retail.major.cetres.help$id) %>% 
-  # #rename(id = '...3') %>% 
-  left_join(retail %>% rownames_to_column(var = "id"), by = 'id')
+  #add_rownames(var = "id") %>% 
+  bind_cols(bua %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
+  dplyr::select(BUA22CD, BUA22NM, X, Y)
 ```
@@
 ```{r}
-# NI
-oa.uk.c.ni <- oa.uk.c %>% filter(str_detect(id, "^N"))
-
-dist.ni <- distm(cbind(oa.uk.c.ni$X, oa.uk.c.ni$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist.ni <- round((dist.ni/1000),2) %>% 
+
+dist.gb <- distm(cbind(bua.c$X, bua.c$Y), cbind(cities.bua.c$X, cities.bua.c$Y), fun=distHaversine) 
+dist.gb <- round((dist.gb/1000),2) %>% 
   as_tibble()
 
-names(dist.ni) <- cities$address #city.names
-
-dist.ni <- dist.ni %>% bind_cols(oa.uk.c.ni$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
+names(dist.gb) <- cities.bua.c$BUA22NM #city.names
+
+dist.gb <- dist.gb %>% 
+  mutate(BUA22CD = bua.c$BUA22CD,
+         BUA22NM = bua.c$BUA22NM) %>% 
+  relocate(c(BUA22CD, BUA22NM))
 
@@
 # min distance
-dist.ni <- transform(dist.ni, dist = do.call(pmin, dist.ni[-1]))
+dist.gb <- transform(dist.gb, dist = do.call(pmin, dist.gb[-(1:2)]))
 
 # City name for minimum distance 
-dist.ni <- dist.ni %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
-
-# GB
-oa.uk.c.gb <- oa.uk.c %>% filter(!str_detect(id, "^N"))
-
-dist.gb <- distm(cbind(oa.uk.c.gb$X, oa.uk.c.gb$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist.gb <- round((dist.gb/1000),2) %>% 
-  as_tibble()
-
-names(dist.gb) <- cities$address #city.names
-
-dist.gb <- dist.gb %>% bind_cols(oa.uk.c.gb$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
-# dist$distMet <- apply(dist[,2:11], 1, min)
-
-# min distance
-dist.gb <- transform(dist.gb, dist = do.call(pmin, dist.gb[-1]))
-
-# City name for minimum distance 
-dist.gb <- dist.gb %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
-
-# UK
-dist <- rbind(dist.ni, dist.gb)
-
-# calculate area
-#oa.uk$area <- st_area(oa.uk) #/ 1000000
+# +2L to accommodate for the first two columns: BUA22CD and BUA22NM
+dist.gb <- dist.gb %>% mutate(dist.city.name = names(.)[max.col(.[3:12]*-1)+2L])
 
 # Join with distance and area
-df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
-  rename(oa11cd=id)
+df <- df %>% left_join(dist.gb %>% dplyr::select(-BUA22NM))
 
@@
 
-### Distance to retail centres
-
-```{r}
-# dist.retail <- distm(cbind(oa.uk.c$X, oa.uk.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-# dist.retail <- round((dist.retail/1000),2) %>% 
-#   as_tibble()
-
-# NI
-oa.uk.c.ni <- oa.uk.c %>% filter(str_detect(id, "^N"))
-dist.retail.ni <- distm(cbind(oa.uk.c.ni$X, oa.uk.c.ni$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail.ni <- round((dist.retail.ni/1000),2) %>% 
-  as_tibble()
-
-names(dist.retail.ni) <- retail.major.cetres$RC_ID
-
-dist.retail.ni <- dist.retail.ni %>% bind_cols(oa.uk.c.ni$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# This efficiently returns the name of the column with the shortest distance
-dist.retail.ni <- dist.retail.ni %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])
-
-# dist.retail_ <- dist.retail %>% slice_sample(n=100)
-# dist.retail_[,6425:6426]
-# dist.retail_ <- transform(dist.retail_, dist.retail = do.call(pmin, dist.retail_[2:6424])) 
-# sum(is.na(dist.retail_$dist.retail.name))
-
-# Minimum distance 
-dist.retail.ni <- transform(dist.retail.ni, dist.retail = do.call(pmin, dist.retail.ni[2:6424])) 
-dist.retail.ni <- dist.retail.ni  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)
-
-# GB
-oa.uk.c.gb <- oa.uk.c %>% filter(!str_detect(id, "^N"))
-dist.retail.gb <- distm(cbind(oa.uk.c.gb$X, oa.uk.c.gb$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail.gb <- round((dist.retail.gb/1000),2) %>% 
-  as_tibble()
-
-names(dist.retail.gb) <- retail.major.cetres$RC_ID
-
-dist.retail.gb <- dist.retail.gb %>% bind_cols(oa.uk.c.gb$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# This efficiently returns the name of the column with the shortest distance
-dist.retail.gb <- dist.retail.gb %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])
-
-# dist.retail_ <- dist.retail %>% slice_sample(n=100)
-# dist.retail_[,6425:6426]
-# dist.retail_ <- transform(dist.retail_, dist.retail = do.call(pmin, dist.retail_[2:6424])) 
-# sum(is.na(dist.retail_$dist.retail.name))
-
-# Minimum distance 
-dist.retail.gb <- transform(dist.retail.gb, dist.retail = do.call(pmin, dist.retail.gb[2:6424])) 
-dist.retail.gb <- dist.retail.gb  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)
-
-# UK
-dist.retail <- rbind(dist.retail.ni, dist.retail.gb)
-
-# Join with complete panel
-df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd")) 
-#sapply(df, function(x) sum(is.na(x)))
-```
-
-## n for London, nearest city and retail
-
-### TODELETE
-
-```{r, eval=FALSE}
-# ## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
-# #df <- df %>% filter(!is.na(area)) 
-# 
-# cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
-# cities.sf <- st_set_crs(cities.sf, 4326)
-# 
-# # help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
-# #   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
-# 
-# sf_use_s2(FALSE)
-# help1 <- st_filter( oa.gb, cities.sf) %>% 
-#   #st_coordinates() %>% 
-#   as_tibble() %>% 
-#   dplyr::select(id) %>% 
-#   mutate(cities = c("London",     # The city names have been manually checked and assigned
-#                     "Leicester",
-#                     "Leeds",
-#                     "Sheffield",
-#                     "Bristol",
-#                     "Birmingham",
-#                     "Manchester",
-#                     "Liverpool",
-#                     "Edinburgh",
-#                     "Glasgow"))
-# 
-# # too slow
-# #dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
-# 
-# # This efficiently returns the name of the column with the shortest distance
-# dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
-# 
-# help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
-#   #rename(nearest.city.oa11cd = id) %>% 
-#   dplyr::select(oa11cd, dist.city.name)
-# 
-# df <- df %>%  left_join(help2, by = c("oa11cd" = "oa11cd")) 
-```
-
-### n for cities
+## n for London and nearest city
 
 ```{R}
-#help.retail <- dist.retail
-
-# n nearest city
-# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-#                          rename(n.nearest.city = n), 
-#                        by = c("nearest.city.oa11cd" = "oa11cd",
-#                               "year" = "year"))
-
+df <- df %>% mutate(dist.city.name = ifelse(dist.city.name == "City.of.Westminster",
+                                             "City of Westminster", dist.city.name)) %>% 
+  left_join(cities.bua.c %>%
+              dplyr::select(BUA22CD, BUA22NM) %>% 
+              rename(BUA22CD.city = BUA22CD),
+            by = c("dist.city.name" = "BUA22NM")) %>% 
+  left_join(df %>% dplyr::select(BUA22CD, year, n) %>% 
+              rename(n.nearest.city = n),
+            by = c("BUA22CD.city" = "BUA22CD", "year" = "year")) %>% 
+  left_join(df %>% dplyr::select(BUA22NM, year, n) %>% 
+              rename(n.London = n) %>% 
+              filter(BUA22NM == "City of Westminster") %>% 
+              rename(help = BUA22NM),
+            by = c("year" = "year")) %>% 
+  dplyr::select(-c(help, Leeds, Liverpool, Manchester, Sheffield, Leicester, Birmingham, 
+                   Bristol, Edinburgh, Glasgow))
+```
+
+## Spatial and spatio-temporal lags
+
+```{r}
+
+bua.used <- df %>% distinct(BUA22CD) %>% as.vector()
 sf_use_s2(FALSE)
-city.boundaries.oa <- st_join(oa.uk, city.boundaries, join = st_intersects) %>% 
-  as_tibble() %>% 
-  filter(!is.na(name)) %>% 
-  dplyr::select(id, name) %>% 
-  mutate(name = ifelse(name == "City of Edinburgh", "Edinburgh", name)) %>% 
-  mutate(name = ifelse(name == "Glasgow City", "Glasgow", name))
-
-city.boundaries.help <- df %>% inner_join(city.boundaries.oa, by = c("oa11cd" = "id")) %>% 
-  group_by(name, year) %>% 
-  summarise(n.nearest.city = sum(n)) %>% 
-  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
-  ungroup()
-
-london.boundaries.help <- city.boundaries.help %>% 
-  filter(name == "London") %>% 
-  rename(n.London = n.nearest.city,
-         n.London.lag = n.nearest.city.lag)
-
-df <- df %>% left_join(city.boundaries.help, by = c("dist.city.name" = "name", "year" = "year"))
-
-df <- df %>% left_join(london.boundaries.help, by = c("year" = "year"))
-
-#sapply(df, function(x) sum(is.na(x)))
-```
-
-### DELETE n for London
-
-```{r, eval=FALSE}
-
-# # OLD London definitions
-# # df <- df %>% left_join(df %>% 
-# #                          ungroup() %>% 
-# #                          filter(oa11cd == "E00176659") %>% 
-# #                          dplyr::select(year, n) %>% 
-# #                          rename(n.London = n)) %>% 
-# #   group_by(oa11cd) %>% 
-# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-# #   mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
-# #   ungroup() 
-# # 
-# # london.help <- df %>% group_by(RGN11NM, year) %>% 
-# #   summarise(n.London = sum(n)) %>% 
-# #   filter(RGN11NM == "London") %>%
-# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-# #   ungroup() %>% 
-# #   dplyr::select(-RGN11NM) 
-# 
-# london.oa <- st_filter(oa.gb, london.sf %>% 
-#                          filter(Name == "Central")) %>% 
-#   as_tibble() %>% 
-#   dplyr::select(id)
-# 
-# london.help <- df %>% inner_join(london.oa, by = c("oa11cd" = "id")) %>% 
-#   group_by(year) %>% 
-#   summarise(n.London = sum(n)) %>% 
-#   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-#   ungroup()
-# 
-# df <- df %>% left_join(london.help, by = c("year" = "year"))
-```
-
-### n for nearest retail
-
-```{r}
-# Old definition for retail centre
-# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-#                          rename(n.nearest.retail = n), 
-#                        by = c("oa11cd.retail" = "oa11cd",
-#                               "year" = "year")) %>% 
-#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
-# 
-# 
-# #n.nearest.retail IS WRONG. FIX UPSTREAM
-# df <- df %>% 
-#   rename(oa11cd.retail = n.nearest.retail) %>% 
-#   dplyr::select(-n.nearest.retail.lag) %>% 
-#   left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-#                          rename(n.nearest.retail = n), 
-#                        by = c("oa11cd.retail" = "oa11cd",
-#                               "year" = "year")) %>% 
-#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
-# 
-# 
-# df <- df %>% dplyr::select(-n.London, -n.London.lag) %>% 
-#   left_join(london.help, by = c('year' = 'year'))
-
-retail.oa <- st_join(oa.uk, retail, 
-                     join = st_intersects) %>% #, 
-                    #largest = TRUE) %>% # returns the x features augmented with 
-                                         # the fields of y that have the largest 
-                                         # overlap with each of the features of x
-  as_tibble() %>% 
-  filter(!is.na(RC_ID)) %>% 
-  dplyr::select(id, RC_ID) %>% 
-  add_row(id = "E00019779", RC_ID = "RC_EW_5951")
-
-# If `larget = FALSE`, we have multiple RC_IDs for OAs.
-# Also, RC_ID="RC_EW_5951" would fall in Thames as it is the London Bridge Tube Station.
-# To avoid NAs for for n.nearest.retail, I would have assigned dist.retail.name = RC_EW_1935 for E00019779
-
-# st_join(oa.uk, retail, join = st_is_within_distance, dist = 500) %>% 
-#   as_tibble() %>% 
-#   filter(!is.na(RC_ID)) %>% 
-#   dplyr::select(id, RC_ID)
-# 
-# st_filter(oa.uk, retail,  .pred = st_is_within_distance, dist = 100) %>% 
-#   as_tibble() %>% 
-#   filter(!is.na(RC_ID)) %>% 
-#   dplyr::select(id, RC_ID)
-
-# h <- oa.uk %>% st_intersects(retail)
-
-# retail.oa_ %>% filter(id == "E00004828"|id == "E00004829"|id == "E00004827")
-# # RC_EW_5285
-# retail.oa <- retail.oa %>% add_row(id = "E00004828", RC_ID = "RC_EW_5285")
-# retail.oa <- retail.oa %>% arrange(id)
-
-retail.help <- df %>% full_join(retail.oa, by = c("oa11cd" = "id")) %>% 
-  group_by(RC_ID, year) %>% 
-  summarise(n.nearest.retail = sum(n)) %>% 
-  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail)) %>% 
-  ungroup()
-
-# retail.help2 <- retail.oa %>% left_join(retail.help, by = c("RC_ID" = "RC_ID")) %>% 
-#   dplyr::select(-RC_ID)
-
-df <- df %>% left_join(retail.help, by = c("dist.retail.name" = "RC_ID", "year" = "year"))
-#sapply(df, function(x) sum(is.na(x)))
-```
-
-### Urban Rural TODO
-
-## TODO
-
-```{r}
-# Lookup: OA to regions
-# **FIXED< DELETE IN TEST** NEED TO ADD NI. NOT A PROBLEM FOR NOW AS IT IS FIXED IN THE TEST SECTION
-
-# # regional identifier
-# df_ <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
-#   mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
-#   ungroup() %>% 
-#   dplyr::select(-help)
-
-#sapply(df, function(x) sum(is.na(x)))
-```
-
-## Spatial and spatio-temporal lags
-
-```{r}
-# sf_use_s2(FALSE)
-# oa.c <- st_centroid(oa.uk)
-# kn <- knearneigh(oa.c, k = 5)
-# knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
-# knn.l <- nb2listw(knn)
-
-path.out <- paste0(path, "/data/temp/knnl.RData")
-# save(knn.l, file = path.out)
-load(path.out)
+bua.c <- bua %>% filter(BUA22CD %in% bua.used$BUA22CD) %>% 
+  st_centroid()
+kn <- knearneigh(bua.c, k = 5)
+knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
+knn.l <- nb2listw(knn)
 
@@
   mutate(n.slag = lag.listw(knn.l, n)) %>% 
-  group_by(oa11cd) %>% 
+  group_by(BUA22CD) %>% 
   mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
-         n.lag = dplyr::lag(n, n=1, order_by = year))        # duplicate line
+         n.London.lag = dplyr::lag(n.London, n=1, order_by = year),
+         n.nearest.city.lag = dplyr::lag(n.nearest.city, n=1, order_by = year))
 ```
@@
 ## RF with CAST
+
+NOT DONE
 
@@
 df <- df %>% filter(year > 1996) %>% 
-  dplyr::select(oa11cd, year, RGN11CD, n, abs.growth, n.l.slag, n.nearest.city.lag, 
-                n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) #%>% #growth, abs.growth, RGN11NM, 
-  # mutate(London.v = n.London.lag / London,
-  #        city.v = n.nearest.city.lag / dist,
-  #        retail.v = n.nearest.retail.lag / (dist.retail + .1)) %>% 
-  # dplyr::select(-n.London.lag, -London, -n.nearest.city.lag, -dist, -n.nearest.retail.lag, -dist.retail)
+  dplyr::select(BUA22CD, BUA22NM, year, n, n.l.slag, n.nearest.city.lag, 
+                n.London.lag, City.of.Westminster, dist,) %>% 
+  rename(London = City.of.Westminster)
 
@@
 
-# df %>% dplyr::select(area, dist, dist.retail, London, south, n.slag, n.lag, n.l.slag, n.nearest.city, year) %>% 
-#   #sapply(function(x) sum(is.na(x)))
-#   summarise(across(where(is.numeric), .fns = 
-#                      list(#n = is.na,
-#                           min = min,
-#                           median = median,
-#                           mean = mean,
-#                           stdev = sd,
-#                           # q25 = ~quantile(., 0.25),
-#                           # q75 = ~quantile(., 0.75),
-#                           max = max))) %>%
-#   pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))
-# 
-# df %>% dplyr::select(n, n.London.lag, n.nearest.city.lag, n.nearest.retail.lag, n.l.slag, year, London, dist, dist.retail) %>%
-#   #sapply(function(x) sum(is.na(x)))
-#   summarise(across(where(is.numeric), .fns =
-#                      list(#n = is.na,
-#                           min = min,
-#                           median = median,
-#                           mean = mean,
-#                           stdev = sd,
-#                           # q25 = ~quantile(., 0.25),
-#                           # q75 = ~quantile(., 0.75),
-#                           max = max))) %>%
-#   pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))
-# 
-# df %>% slice_max(order_by = n, n = 5, by = year)
 ```
@@
 indices <- CreateSpacetimeFolds(df, 
-                                spacevar = "oa11cd",
+                                spacevar = "BUA22CD",
                                 timevar = "year", 
@@
 #(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # It should be RHS variables + 2
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # It should be RHS variables + 2
 
@@
 
-# detectCores()
-# cl <- makePSOCKcluster(2)
-# registerDoParallel(cl)
+detectCores()
+cl <- makePSOCKcluster(6)
+registerDoParallel(cl)
 
@@
 model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
-                     n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
+                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist ,
                      #London.v + city.v + retail.v + n.l.slag + year,
@@
 time.taken <- round(end.time - start.time,2)
-time.taken #20.6h
+time.taken #7.3m
 
@@
 # 
-# 3716736 samples
-#       8 predictor
+# 134528 samples
+#      6 predictor
 # 
@@
 # Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 2926910, 2926924, 3136005, 2926924, 3136005, 2926924, ... 
+# Summary of sample sizes: 113505, 105938, 105938, 113520, 113490, 105938, ... 
 # Resampling results across tuning parameters:
@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    4.999920  0.2053697  1.047293
-#   2     extratrees  5.167655  0.1540213  1.103048
-#   5     variance    5.187039  0.1710128  1.034382
-#   5     extratrees  5.169570  0.1557868  1.122845
-#   8     variance    5.633427  0.1247189  1.035171
-#   8     extratrees  5.182860  0.1563555  1.137332
+#   2     variance    79.06941  0.6270099  24.31033
+#   2     extratrees  93.85626  0.5098337  26.64545
+#   4     variance    75.36044  0.6507424  21.62872
+#   4     extratrees  85.91194  0.5748039  25.88575
+#   6     variance    75.32465  0.6476521  21.19409
+#   6     extratrees  83.85574  0.5904021  26.05747
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule =
-#  variance and min.node.size = 5.
+# The final values used for the model were mtry = 6, splitrule = variance and min.node.size = 5.
 
@@
 ```
-
-### OLD
-
-```{r, eval=FALSE}
-# train on all but one region, test on that region
-
-df.predictions <- data.frame()
-for (i in regions){
-  for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], df[df$RGN11CD==j,])
-    pred <- as.data.frame(pred)
-    rownames(pred) <- c()
-    assign(paste0("region.predict.model.", i, ".on.", j), pred)
-
-    d <- paste0("region.predict.model.", i, ".on.", j)
-    d <- cbind(get(d), df %>% filter(RGN11CD==j)) #%>% arrange(year, ladcd)
-    d <- d[,c(5, 2, 4, 6, 1)]
-    colnames(d)[5] <- "predictions"
-    d <- d %>% mutate(trained.on = i)
-    df.predictions <- rbind(d, df.predictions)
-  }
-}
-
-df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN11CD, ".", trained.on))
-
-# split to list by region pair
-pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
-
-# calculate metrics for every region pair 
-rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))  # CHANGE n WITH growth
-
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% 
-  dplyr::select(RGN21CD, RGN21NM) %>% 
-  distinct() %>% 
-  rename(RGN11CD = RGN21CD) %>% 
-  add_row(RGN11CD = c("Scotland", "N0", "Wales"), 
-          RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
-
-path.out <- paste0(path, "/outputs/rf/heatmap_OA.png")
-rf.year.all.metrics %>% 
-  as.data.frame() %>%
-  rownames_to_column(var = "metrics") %>% 
-  pivot_longer(!metrics, names_to = "train.test") %>% 
-  pivot_wider(names_from = metrics, values_from = value) %>% 
-  separate(train.test, c("train", "test"), remove = F) %>% 
-  left_join(lookup.region, by = c("train" = "RGN11CD")) %>% 
-  rename(train.region=RGN21NM) %>% 
-  left_join(lookup.region, by = c("test" = "RGN11CD")) %>% 
-  rename(test.region=RGN21NM) %>% 
-  arrange(Rsquared) %>% 
-  dplyr::select(train.region, test.region, Rsquared) %>% 
-  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
-  geom_tile() +
-  coord_fixed() +
-  labs(y = "", x = "") + #y: test.region, x: train.region
-  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
-  theme(#legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust=1),
-        panel.border = element_blank(), 
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.background = element_rect(fill = "white"))
-ggsave(path.out)
-
-  #summarise(r = range(Rsquared))
-  #round(2) %>%
-  #kable()
-
-```
-
-## RF
-
-### train on all but one regions
-
-```{r}
-# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
-# df <- read_csv(out.path)
-
-unregister_dopar <- function() {
-  env <- foreach:::.foreachGlobals
-  rm(list=ls(name=env), pos=env)
-}
-unregister_dopar()
-
-# CV
-tc <- trainControl(method = "cv",
-                   number = 10, #10,
-                   allowParallel = T,
-                   savePredictions = 'final')
-
-# London region: E12000007
-
-# detectCores()
-# cl <- makePSOCKcluster(14)
-# registerDoParallel(cl)
-
-# train
-start.time <- Sys.time()
-
-set.seed(71)
-m.test <- train(n ~ density + area + dist + dist.retail + London +
-                  n.slag + year + n.nearest.city, #n.l.slag
-                data = df %>% filter(RGN11CD=="E12000008"),
-                trControl = tc,
-                method = "ranger", #"rf", 
-                importance = "impurity",  #TRUE,
-                na.action = na.omit)
-
-#stopCluster(cl)
-end.time <- Sys.time()
-time.taken <- round(end.time - start.time,2)
-time.taken #1h
-
-summary(m.test)
-plot(m.test)
-m.test
-print(m.test)
-
-varimp_mars <- varImp(m.test)
-plot(varimp_mars, main="Variable Importance")
-```
-
-### test on the hold out regions
-
-Wales region is missing!!!
-
-```{r}
-pred <- predict(m.test, df %>% ungroup() %>% filter(RGN11CD=="E12000006"), na.action = na.pass)
-length(pred)
-
-df %>% ungroup() %>% filter(RGN11CD=="E12000006") %>% filter(if_any(everything(), is.na))
-
-pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN11CD=="E12000006"))
-postResample(pred = pred$value, obs = pred$n) ## 0.99
-```
-
-## RF rolling
-
-### train the model in a loop
-
-```{r}
-# Train on t & t + 1
-
-# Make the parallelisation reproducible using https://stackoverflow.com/questions/13403427/fully-reproducible-parallel-models-using-caret. 
-# See also https://stackoverflow.com/questions/27944558/set-seed-parallel-random-forest-in-caret-for-reproducible-result
-
-#length is = (n_repeats*nresampling)+1
-seeds <- vector(mode = "list", length = 11)
-
-#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South
-
-#for the last model
-seeds[[11]]<-sample.int(1000, 1)
-
-# CV
-tc <- trainControl(method = "cv",
-                   number = 10,
-                   seeds = seeds,
-                   allowParallel = T,
-                   savePredictions = 'final')
-
-detectCores()
-cl <- makePSOCKcluster(6)#14
-registerDoParallel(cl)
-
-# run the model for every 2 years
-start.time <- Sys.time()
-years <- 2001:2012 #1997
-
-for (t in years){
-
-  tplus1 <- t + 1
-  set.seed(71)
-  print(t)
-
-  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
-                      n.slag + n.l.slag + n.nearest.city + year, #n.lag +
-                     data = df %>% filter(year==t | year==tplus1), #%>% 
-                       # mutate(growth = ifelse(growth == -Inf | growth == Inf, NA, growth)) %>% 
-                       # mutate(growth = ifelse(is.na(growth), 0, growth)), 
-                     #%>% mutate(growth = ifelse(growth== -Inf | growth == Inf, NA, growth))
-                     trControl = tc,
-                     method = "ranger", 
-                     na.action = na.omit,
-                     #preProc = c("center", "scale"),
-                     importance = 'impurity') #TRUE
-
-  #assign(paste0("model.all",as.character(t + 1)), model.all)
-  file <- paste0(path, "/data/temp/oa_rf/timeseries/timeseries", as.character(t + 1), ".RData")
-  save(model.all, file = file)
-  rm(model.all)
-}
-
-stopCluster(cl)
-end.time <- Sys.time()
-time.taken <- round(end.time - start.time,2)
-time.taken
-
-# df %>% filter(growth == Inf) 
-# log(n/n.lag)
-
-# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
-# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
-```
-
-### Resamples for plots
-
-```{r}
-# create a list of the above model outputs
-
-#load(file = (paste0(path, "/data/temp/oa_rf/timeseries/timeseries2001.RData")))
-
-filenames <- list.files(paste0(path, "/data/temp/oa_rf/timeseries/"), pattern = "*.RData", full.names = T)
-
-t <- 1997
-for (i in filenames){
-  load(file = i)
-  t <- t + 1
-  assign(paste0("timeseries",t), model.all)  
-}
-
-fit.model.all <- mget(ls(pattern = "^timeseries\\d"))
-rm(list = ls(pattern = "^timeseries\\d"))
-
-# resamples
-fit.model.all.res <- resamples(fit.model.all)
-summary(fit.model.all.res)
-
-# plot metrics
-
-# TODO: fix the year labels
-
-scales <- list(x=list(relation="free"), y=list(relation="free"))
-bwplot(fit.model.all.res, scales=scales)
-
-# TODO: fix the year labels
-```
-
-### Variable importance
-
-```{r , include=TRUE, results= 'markup', message=FALSE}
-lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
-```
-
-```{r}
-help <- lapply(fit.model.all, function(x) varImp(x)) 
-help <- lapply(help, function(x) x[[1]])
-help <- do.call("rbind", help)
-
-importance <- help %>% rownames_to_column() %>% 
-  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
-         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
-  rename(importance = Overall) %>% 
-  relocate(importance, .after = year) %>% 
-  dplyr::select(-rowname)
-
-importance %>% ggplot() +
-  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
-  scale_color_discrete(name = "Features") +
-  #scale_linetype_manual(name = "Website density regressions111") 
-  ylab("Importance") + 
-  ggtitle("Feature importance") +
-  theme_minimal() +
-  theme(axis.title.x=element_blank(),
-        plot.title = element_text(hjust = 0.5))
-
-```
-
-### test
-
-```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
-
-# growth or n
-
-# for a reference point
-df %>% group_by(year) %>%
-  summarise(min = min(n), max=max(n),
-            mean = mean(n), median = median(n)) %>%
-  round(2) %>% kable()
-```
-
-```{r}
-## Test on t + 2
-
-# It will create: 2001 (train on 2000-2001) - 2011 (train on 2010)
-# I need: 2001 (train on 2000-2001) - 2009 (train on 2008-2009)
-
-# use the model trainned in year t to predict io of year t + 1
-years <- (1998:2011)
-pred.by.year.all <- data.frame(matrix(NA,nrow=227759, ncol = 0))
-for (t in years){
-  t_plus_1 <- t + 1
-  pred <- predict(fit.model.all[names(fit.model.all)==paste0("timeseries",t)], df[df$year==t_plus_1,])
-  pred <- as.data.frame(pred)
-  rownames(pred) <- c()
-  pred.by.year.all <- cbind(pred.by.year.all, pred)
-}
-
-#change column names to match with total
-names(pred.by.year.all) <- 1999:2012
-
-# wide to long
-pred.by.year.all <- gather(pred.by.year.all,key = year,value = predictions,1:14)
-
-# drop 2000 and select variables form total
-df.no1997_1 <- df %>% 
-  filter(year!=1996 & year!=1997 & year!=1998) %>% 
-  dplyr::select(oa11cd, year, n) # CHANGE n WITH growth
-
-# column bind prediction and data  
-pred.by.year.all <- cbind(pred.by.year.all, df.no1997_1 %>% arrange(year, oa11cd)) #DOUBLE CHECK ARRANGE
-
-# pred.by.year to list by year
-pred.by.year.all.list <- split(pred.by.year.all, pred.by.year.all$year)
-
-# calculate metrics for every year
-rf.year.all.metrics <- lapply(pred.by.year.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))  # CHANGE n WITH growth
-rf.year.all.metrics %>% as.data.frame() %>%
-  round(2) %>%
-  kable()
-
-# there are 2 columns named year, so I drop one
-pred.by.year.all$year <- NULL
-
-ggplot(data = pred.by.year.all,aes(x = n, y = predictions)) +
-  geom_point(colour = "blue") +
-  geom_abline(intercept = 0, slope = 1, colour = "red") +
-  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
-  facet_wrap(~ year,ncol = 2) +
-  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
-  ggtitle("Predicted vs. actual trade by year")
-```
-
-## RF spatial
-
-```{r}
-df <- df %>% mutate(help = substr(ladcd, start = 1, stop = 2)) %>% 
-  mutate(RGN21CD = ifelse(is.na(RGN21CD), help, RGN21CD)) %>% 
-  ungroup() %>% 
-  dplyr::select(-help)
-
-df <- df %>% dplyr::select(-n.lag)
-
-```
-
-### train the model in a loop
-
-```{r}
-#length is = (n_repeats*nresampling)+1
-seeds <- vector(mode = "list", length = 11)
-
-#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South
-
-#for the last model
-seeds[[11]]<-sample.int(1000, 1)
-
-# CV
-tc <- trainControl(method = "cv",
-                   number = 10,
-                   seeds = seeds,
-                   allowParallel = T,
-                   savePredictions = 'final')
-
-detectCores()
-cl <- makePSOCKcluster(14)
-registerDoParallel(cl)
-
-# run the model for every 2 years
-start.time <- Sys.time()
-regions <- df %>% distinct(RGN21CD)
-regions <- as.vector(regions$RGN21CD)
-
-for (i in regions){
-  set.seed(71)
-  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
-                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
-                     data = df %>% filter(RGN21CD==i), trControl = tc,
-                     method = "ranger",
-                     na.action = na.omit,
-                     #preProc = c("center", "scale"),
-                     importance = 'impurity') #TRUE
-  assign(paste0("model.all",as.character(i)), model.all)
-}
-
-stopCluster(cl)
-end.time <- Sys.time()
-time.taken <- round(end.time - start.time,2)
-time.taken
-
-# df %>% filter(growth == Inf) 
-# log(n/n.lag)
-
-# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
-# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
-```
-
-### Resamples for plots
-
-```{r}
-# create a list of the above model outputs
-fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
-
-# resamples
-fit.model.all.res <- resamples(fit.model.all)
-summary(fit.model.all.res)
-
-# plot metrics
-
-# TODO: fix the year labels
-
-scales <- list(x=list(relation="free"), y=list(relation="free"))
-bwplot(fit.model.all.res, scales=scales)
-
-# TODO: fix the year labels
-```
-
-### Variable importance
-
-```{r , include=TRUE, results= 'markup', message=FALSE}
-lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
-```
-
-```{r}
-help <- lapply(fit.model.all, function(x) varImp(x)) 
-help <- lapply(help, function(x) x[[1]])
-help <- do.call("rbind", help)
-
-importance <- help %>% rownames_to_column() %>% 
-  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
-         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
-  rename(importance = Overall) %>% 
-  relocate(importance, .after = year) %>% 
-  dplyr::select(-rowname)
-
-importance %>% ggplot() +
-  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
-  scale_color_discrete(name = "Features") +
-  #scale_linetype_manual(name = "Website density regressions111") 
-  ylab("Importance") + 
-  ggtitle("Feature importance") +
-  theme_minimal() +
-  theme(axis.title.x=element_blank(),
-        plot.title = element_text(hjust = 0.5))
-
-```
-
-### test
-
-```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
-
-# growth or n
-
-# for a reference point
-df %>% group_by(year) %>%
-  summarise(min = min(n), max=max(n),
-            mean = mean(n), median = median(n)) %>%
-  round(2) %>% kable()
-```
-
-```{r}
-## Test on t + 2
-
-# train on one region, test on all the rest
-
-df.predictions <- data.frame()
-for (i in regions){
-  for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
-    pred <- as.data.frame(pred)
-    rownames(pred) <- c()
-    assign(paste0("region.predict.model.", i, ".on.", j), pred)
-
-    d <- paste0("region.predict.model.", i, ".on.", j)
-    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
-    d <- d[,c(5, 2, 4, 6, 1)]
-    colnames(d)[5] <- "predictions"
-    d <- d %>% mutate(trained.on = i)
-    df.predictions <- rbind(d, df.predictions)
-  }
-}
-
-df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))
-
-# split to list by region pair
-pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
-
-# calculate metrics for every region pair 
-rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))  # CHANGE n WITH growth
-
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% 
-  dplyr::select(RGN21CD, RGN21NM) %>% 
-  distinct() %>% 
-  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
-
-rf.year.all.metrics %>% 
-  as.data.frame() %>%
-  add_rownames(var = "metrics") %>% 
-  pivot_longer(!metrics, names_to = "train.test") %>% 
-  pivot_wider(names_from = metrics, values_from = value) %>% 
-  separate(train.test, c("train", "test"), remove = F) %>% 
-  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
-  rename(train.region=RGN21NM) %>% 
-  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
-  rename(test.region=RGN21NM) %>% 
-  arrange(Rsquared)
-  #summarise(r = range(Rsquared))
-  #round(2) %>%
-  #kable()
-```
-
-### test plots
-
-```{r}
-#plots
-ggplot(data = df.predictions, aes(x = n, y = predictions)) +
-  geom_point(colour = "blue") +
-  geom_abline(intercept = 0, slope = 1, colour = "red") +
-  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
-  facet_wrap(~ test.train, ncol = 2) +
-  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
-  ggtitle("Predicted vs. actual trade by year")
-```

### 2024-02-08 src/oa_RF.Rmd
oa rf

@@
 
-```{r}
+```{r, eval=F}
 # create a list of the above model outputs
@@
 
-```{r , include=TRUE, results= 'markup', message=FALSE}
+```{r, eval=F, include=TRUE, results= 'markup', message=FALSE}
 lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
@@
 
-```{r}
+```{r, eval=F}
 help <- lapply(fit.model.all, function(x) varImp(x)) 
@@
 
+regions <- df %>% ungroup() %>% distinct(RGN11CD) 
+regions <- as.vector(regions$RGN11CD)
+
+new.path <- "/hdd/tmp/regions/"
+#filenames <- list.files(paste0(path, "/outputs/rf/models/oa_rf/regions"), pattern = "*.RData", full.names = T)
+
+df.predictions <- data.frame()
+
+for (i in regions){
+  load(file = paste0(new.path, "region_", as.character(i), ".RData"))
+  #model.name <- gsub("/home/nw19521/projects/web.diffusion/outputs/rf/models/oa_rf/regions/|.RData", "", i)
+  #assign(as.character(i), model.all)
+
+  pred <- predict(model.all, df[df$RGN11CD==i,])
+  pred <- as.data.frame(pred)
+
+  rownames(pred) <- c()
+  assign(paste0("region.predict.model.", ".on.", i), pred)
+
+  d <- paste0("region.predict.model.", ".on.", i)
+  d <- cbind(get(d), df %>% filter(RGN11CD==i)) #%>% arrange(year, ladcd)
+
+  d <- d %>% dplyr::select(RGN11CD, oa11cd, year, n, pred)
+
+  colnames(d)[5] <- "predictions"
+  d <- d %>% mutate(tested.on = i)
+  df.predictions <- rbind(d, df.predictions)
+
+  rm(model.all)
+}
+
+# df.predictions <- data.frame()
+# for (i in regions){
+#   #for (j in regions[!regions == i]){
+#     pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], 
+#                     df[df$RGN11CD==i,])
+#     pred <- as.data.frame(pred)
+#     rownames(pred) <- c()
+#     assign(paste0("region.predict.model.", ".on.", i), pred)
+#     
+#     d <- paste0("region.predict.model.", ".on.", i)
+#     d <- cbind(get(d), df %>% filter(RGN11CD==i)) #%>% arrange(year, ladcd)
+#     
+#     #***dplyr::select(RGN21CD, ladcd, year, growth, 1)***
+#     
+#     d <- d[,c(5, 2, 4, 6, 1)]     
+#     colnames(d)[5] <- "predictions"
+#     d <- d %>% mutate(trained.on = i)
+#     df.predictions <- rbind(d, df.predictions)
+#   #}
+# }
+
+df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN11CD, ".", tested.on))
+
+# split to list by region pair
+pred.by.region.all.list <- split(df.predictions, df.predictions$test.train)
+
+# calculate metrics for every region pair 
+rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
+                                                                     obs = x$n))  # CHANGE n WITH growth
+
+path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup) %>% 
+  dplyr::select(RGN21CD, RGN21NM) %>% 
+  distinct() %>% 
+  rename(RGN11CD = RGN21CD) %>% 
+  add_row(RGN11CD = c("Scotland", "NI", "Wales"), 
+          RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
+
+path.out <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
+rf.year.all.metrics %>% 
+  as.data.frame() %>%
+  rownames_to_column(var = "metrics") %>% 
+  pivot_longer(!metrics, names_to = "train.test") %>% 
+  pivot_wider(names_from = metrics, values_from = value) %>% 
+  separate(train.test, c("train", "test"), remove = F) %>% 
+  # left_join(lookup.region, by = c("train" = "RGN11CD")) %>% 
+  # rename(train.region=RGN21NM) %>% 
+  left_join(lookup.region, by = c("test" = "RGN11CD")) %>% 
+  rename(test.region=RGN21NM) %>% 
+  arrange(Rsquared) %>% 
+  dplyr::select(test.region, Rsquared) %>% 
+  write_csv(path.out)
+```
+
+### OLD
+
+```{r, eval=FALSE}
+# train on all but one region, test on that region
+
 df.predictions <- data.frame()
 for (i in regions){
-  #for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], 
-                    df[df$RGN11CD==i,])
+  for (j in regions[!regions == i]){
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], df[df$RGN11CD==j,])
     pred <- as.data.frame(pred)
     rownames(pred) <- c()
-    assign(paste0("region.predict.model.", ".on.", i), pred)
-
-    d <- paste0("region.predict.model.", ".on.", i)
-    d <- cbind(get(d), df %>% filter(RGN11CD==i)) #%>% arrange(year, ladcd)
-
-    #***dplyr::select(RGN21CD, ladcd, year, growth, 1)***
-
+    assign(paste0("region.predict.model.", i, ".on.", j), pred)
+
+    d <- paste0("region.predict.model.", i, ".on.", j)
+    d <- cbind(get(d), df %>% filter(RGN11CD==j)) #%>% arrange(year, ladcd)
     d <- d[,c(5, 2, 4, 6, 1)]
@@
     df.predictions <- rbind(d, df.predictions)
-  #}
+  }
 }
@@
 
-### OLD
-
-```{r, eval=FALSE}
-# train on all but one region, test on that region
-
-df.predictions <- data.frame()
-for (i in regions){
-  for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], df[df$RGN11CD==j,])
-    pred <- as.data.frame(pred)
-    rownames(pred) <- c()
-    assign(paste0("region.predict.model.", i, ".on.", j), pred)
-
-    d <- paste0("region.predict.model.", i, ".on.", j)
-    d <- cbind(get(d), df %>% filter(RGN11CD==j)) #%>% arrange(year, ladcd)
-    d <- d[,c(5, 2, 4, 6, 1)]
-    colnames(d)[5] <- "predictions"
-    d <- d %>% mutate(trained.on = i)
-    df.predictions <- rbind(d, df.predictions)
-  }
-}
-
-df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN11CD, ".", trained.on))
-
-# split to list by region pair
-pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
-
-# calculate metrics for every region pair 
-rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))  # CHANGE n WITH growth
-
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% 
-  dplyr::select(RGN21CD, RGN21NM) %>% 
-  distinct() %>% 
-  rename(RGN11CD = RGN21CD) %>% 
-  add_row(RGN11CD = c("Scotland", "N0", "Wales"), 
-          RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
-
-path.out <- paste0(path, "/outputs/rf/heatmap_OA.png")
-rf.year.all.metrics %>% 
-  as.data.frame() %>%
-  rownames_to_column(var = "metrics") %>% 
-  pivot_longer(!metrics, names_to = "train.test") %>% 
-  pivot_wider(names_from = metrics, values_from = value) %>% 
-  separate(train.test, c("train", "test"), remove = F) %>% 
-  left_join(lookup.region, by = c("train" = "RGN11CD")) %>% 
-  rename(train.region=RGN21NM) %>% 
-  left_join(lookup.region, by = c("test" = "RGN11CD")) %>% 
-  rename(test.region=RGN21NM) %>% 
-  arrange(Rsquared) %>% 
-  dplyr::select(train.region, test.region, Rsquared) %>% 
-  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
-  geom_tile() +
-  coord_fixed() +
-  labs(y = "", x = "") + #y: test.region, x: train.region
-  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
-  theme(#legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust=1),
-        panel.border = element_blank(), 
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.background = element_rect(fill = "white"))
-ggsave(path.out)
-
-  #summarise(r = range(Rsquared))
-  #round(2) %>%
-  #kable()
-
-```
-
 ## RF

### 2024-02-12 src/oa_RF.Rmd
figure

@@
                                                       ifelse(rowname == "n.l.slag", "spatial and temporal lag of wensite density",
-                                                             ifelse(rowname == "n.nearest.city", "Nearest city's wensite density, t-1", rowname)))))))) %>% 
+                                                             ifelse(rowname == "n.nearest.city.lag", "Nearest city's wensite density, t-1", rowname)))))))) %>% 
   column_to_rownames()

### 2024-02-12 src/bua_RF.Rmd
bua

@@
   group_by(BUA22CD) %>% 
-  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
+  mutate(n.lag = dplyr::lag(n, n=1, order_by=year),
+         n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
          n.London.lag = dplyr::lag(n.London, n=1, order_by = year),
-         n.nearest.city.lag = dplyr::lag(n.nearest.city, n=1, order_by = year))
+         n.nearest.city.lag = dplyr::lag(n.nearest.city, n=1, order_by = year)) %>%
+  ungroup()
 ```
@@
 
-NOT DONE
-
 ```{r}
 
-# Lookup: OA to regions
-path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% dplyr::select(OA11CD, RGN11CD, RGN11NM)
+# Lookup: BUA to regions
 
-df <- df %>% left_join(lookup.region, by = c("oa11cd" = "OA11CD")) %>% 
-  relocate(c(RGN11NM, RGN11CD), .after=year) %>% 
-  mutate(help = substring(oa11cd, 1, 1),
-         RGN11CD = ifelse(help=="E", RGN11CD,
-                         ifelse(help=="W", "Wales",
-                                ifelse(help == "S", "Scotland", "NI"))),
-         RGN11NM = ifelse(help=="E", RGN11NM,
-                         ifelse(help=="W", "Wales",
-                                ifelse(help == "S", "Scotland", "NI")))) %>% 
-  dplyr::select(-help)
+# source: https://www.data.gov.uk/dataset/aca8408d-e422-4ec2-a22f-9bd8beb1d549/built-up-area-to-region-december-2022-lookup-in-great-britain
+path.lookup <- paste0(path, "/data/raw/Built_Up_Area_to_Region_(December_2022)_Lookup_in_Great_Britain.csv")
+
+# 142 BUA in more than one regions
+lookup.region <- read_csv(path.lookup)
+
+df <- df %>% left_join(lookup.region %>% 
+                         filter(WHOLE_PART=="Whole") %>% 
+                         dplyr::select(BUA22CD, RGN22CD, RGN22NM), 
+                       by = c("BUA22CD" = "BUA22CD")) %>% 
+  #filter(is.na(RGN22CD))                                      #142 BUA in more than one regions dropped
+  relocate(c(RGN22NM, RGN22CD), .after=BUA22NM) %>% 
+  ungroup()
 
@@
 # For growth:
-df <- df %>% group_by(oa11cd) %>% 
+df <- df %>% group_by(BUA22CD) %>% 
   mutate(#n.lag = dplyr::lag(n),
@@
 
-# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
+# out.path <- paste0(path, "/data/temp/df_for_bua_rf.csv")
 # df <- read_csv(out.path)
@@
 
-# remove not used columns
-df <- df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol",
-                        "Manchester", "Sheffield", "Leeds", "Edinburgh",
-                        "Leicester", "dist.city.name", "dist.retail.name", "name"))
-
 # There are no lags for 1996, so it is removed
 df <- df %>% filter(year > 1996) %>% 
-  dplyr::select(BUA22CD, BUA22NM, year, n, n.l.slag, n.nearest.city.lag, 
+  dplyr::select(BUA22CD, BUA22NM, RGN22NM, RGN22CD, year, n, n.l.slag, n.nearest.city.lag, 
                 n.London.lag, City.of.Westminster, dist,) %>% 
   rename(London = City.of.Westminster)
-
-df %>% group_by(year) %>% 
-  filter(n == 0) %>% 
-  count(n)
-# in 2012, 120848 0s
-
 ```
@@
 ```{r}
+#RNGversion("4.0.2")
+
 set.seed(123)
@@
                                 timevar = "year", 
-                                k = 10)
+                                k = 10,
+                                seed = 123)
 
@@
 model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
-                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist ,
+                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                      #London.v + city.v + retail.v + n.l.slag + year,
@@
 
-# stopCluster(cl)
+stopCluster(cl)
+
 end.time <- Sys.time()
@@
 # Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 113505, 105938, 105938, 113520, 113490, 105938, ... 
+# Summary of sample sizes: 113520, 105938, 105938, 113505, 113505, 105938, ... 
 # Resampling results across tuning parameters:
@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    79.06941  0.6270099  24.31033
-#   2     extratrees  93.85626  0.5098337  26.64545
-#   4     variance    75.36044  0.6507424  21.62872
-#   4     extratrees  85.91194  0.5748039  25.88575
-#   6     variance    75.32465  0.6476521  21.19409
-#   6     extratrees  83.85574  0.5904021  26.05747
+#   2     variance    81.30945  0.6421819  23.39696
+#   2     extratrees  96.94266  0.5158416  26.31072
+#   4     variance    78.74613  0.6597365  21.00403
+#   4     extratrees  88.39690  0.5849950  25.41510
+#   6     variance    79.57023  0.6514796  20.86435
+#   6     extratrees  86.15949  0.6003589  25.47657
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 6, splitrule = variance and min.node.size = 5.
+# The final values used for the model were mtry = 4, splitrule = variance and min.node.size = 5.
 
@@
                                                       ifelse(rowname == "n.l.slag", "spatial and temporal lag of wensite density",
-                                                             ifelse(rowname == "n.nearest.city", "Nearest city's wensite density, t-1", rowname)))))))) %>% 
+                                                             ifelse(rowname == "n.nearest.city.lag", "Nearest city's wensite density, t-1", rowname)))))))) %>% 
   column_to_rownames()
 
-path.out <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
+path.out <- paste0(path, "/outputs/rf/figures/varimp_BUA.png")
 ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
 ggsave(path.out)
-
-path.out <- paste0(path, "/outputs/rf/models/oa_rf/all_OAs/all_OAs.RData")
-save(model.all, file = path.out)
-
-# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
-#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
-#                                                         "year")], 
-#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
-#                     trControl = tc,
-#                     method = "ranger",
-#                     #na.action = na.omit,
-#                     #importance = "impurity",
-#                     metric="Rsquared",
-#                     verbose=FALSE)
-# ffsmodel_LLO
-# ffsmodel_LLO$selectedvars
-# plot_ffs(ffsmodel_LLO)
 ```
@@
 ```{r}
+# Drop the 142 BUA in more than one regions
+df <- df %>% filter(!is.na(RGN22CD))
+
+#rm(list = setdiff(ls(), c("df", "indices")))
 
@@
 # detectCores()
-# cl <- makePSOCKcluster(2)
-# registerDoParallel(cl)
+cl <- makePSOCKcluster(4)
+registerDoParallel(cl)
 
@@
 
-regions <- df %>% ungroup() %>% distinct(RGN11CD) 
-regions <- as.vector(regions$RGN11CD)
+regions <- df %>% ungroup() %>% distinct(RGN22CD) 
+regions <- as.vector(regions$RGN22CD)
 
@@
   set.seed(71)
-  indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD!=i), 
-                                spacevar = "oa11cd",
+  indices <- CreateSpacetimeFolds(df %>% filter(RGN22CD!=i), 
+                                spacevar = "BUA22CD",
                                 timevar = "year", 
@@
 
-  model.all <- train(n ~ n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
-                   data = df %>% filter(RGN11CD!=i), 
+  model.all <- train(n ~ n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
+                   data = df %>% filter(RGN22CD!=i), 
                    trControl = tc,
@@
                    importance = "impurity")
-  #assign(paste0("model.all",as.character(i)), model.all)
+  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data
 
-  #file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
-  file <- paste0("/hdd/tmp/regions/region_", as.character(i), ".RData")  
-  save(model.all, file = file)  # file name = region not in the training data
+  # file <- paste0("/hdd/tmp/regions/region_", as.character(i), ".RData")  
+  # save(model.all, file = file)  # file name = region not in the training data
   rm(model.all)
@@
 
+# create a list of the above model outputs
+fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
+
 # clear up
-rm(model.all) # this is the last item from the loop
-rm(list = ls(pattern = "^region_"))
+rm(list = ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
 
@@
   #for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], 
-                    df[df$RGN11CD==i,])
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all", i)], 
+                    df[df$RGN22CD==i,])
     pred <- as.data.frame(pred)
@@
     d <- paste0("region.predict.model.", ".on.", i)
-    d <- cbind(get(d), df %>% filter(RGN11CD==i)) #%>% arrange(year, ladcd)
+    d <- cbind(get(d), df %>% filter(RGN22CD==i)) #%>% arrange(year, ladcd)
 
-    #***dplyr::select(RGN21CD, ladcd, year, growth, 1)***
+    d <- d %>% dplyr::select(RGN22CD, RGN22NM, BUA22CD, year, n, 1)
 
-    d <- d[,c(5, 2, 4, 6, 1)]     
-    colnames(d)[5] <- "predictions"
+    colnames(d)[6] <- "predictions"
     d <- d %>% mutate(trained.on = i)
@@
 
-df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN11CD, ".", trained.on))
+df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN22CD, ".", trained.on))
 
@@
 
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
+# source: https://www.data.gov.uk/dataset/aca8408d-e422-4ec2-a22f-9bd8beb1d549/built-up-area-to-region-december-2022-lookup-in-great-britain
+path.lookup <- paste0(path, "/data/raw/Built_Up_Area_to_Region_(December_2022)_Lookup_in_Great_Britain.csv")
+
+# 142 BUA in more than one regions
 lookup.region <- read_csv(path.lookup) %>% 
-  dplyr::select(RGN21CD, RGN21NM) %>% 
-  distinct() %>% 
-  rename(RGN11CD = RGN21CD) %>% 
-  add_row(RGN11CD = c("Scotland", "N0", "Wales"), 
-          RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
+  dplyr::select(RGN22CD, RGN22NM) %>% 
+  distinct()
 
-path.out <- paste0(path, "/outputs/rf/heatmap_OA.png")
+path.out <- paste0(path, "/outputs/rf/test_regions_BUA.csv")
+
 rf.year.all.metrics %>% 
@@
   separate(train.test, c("train", "test"), remove = F) %>% 
-  left_join(lookup.region, by = c("train" = "RGN11CD")) %>% 
-  rename(train.region=RGN21NM) %>% 
-  left_join(lookup.region, by = c("test" = "RGN11CD")) %>% 
-  rename(test.region=RGN21NM) %>% 
+  left_join(lookup.region, by = c("train" = "RGN22CD")) %>% 
+  rename(Region = RGN22NM) %>% 
   arrange(Rsquared) %>% 
-  dplyr::select(train.region, test.region, Rsquared) %>% 
-  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
-  geom_tile() +
-  coord_fixed() +
-  labs(y = "", x = "") + #y: test.region, x: train.region
-  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
-  theme(#legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust=1),
-        panel.border = element_blank(), 
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.background = element_rect(fill = "white"))
-ggsave(path.out)
-
-  #summarise(r = range(Rsquared))
-  #round(2) %>%
-  #kable()
-
+  dplyr::select(Region, Rsquared) %>% 
+  write_csv(path.out)
 ```

### 2024-02-12 src/bua_RF.Rmd
bua

@@
 
+```{r}
+# out.path <- paste0(path, "/data/temp/df_for_bua_rf.csv")
+# df <- read_csv(out.path)
+# df %>% write_csv(file = out.path)
+
+# rm(list=setdiff(ls(), "df"))
+
+# # This is the project path
+# path <- find_rstudio_root_file()
+
+```
+
 ## RF with CAST
@@
 df <- df %>% left_join(lookup.region %>% 
-                         filter(WHOLE_PART=="Whole") %>% 
+                         #filter(WHOLE_PART=="Whole") %>% 
                          dplyr::select(BUA22CD, RGN22CD, RGN22NM), 
@@
   relocate(n.lag, .after = n)
-
-# out.path <- paste0(path, "/data/temp/df_for_bua_rf.csv")
-# df <- read_csv(out.path)
-# df %>% write_csv(file = out.path)
-
-# rm(list=setdiff(ls(), "df"))
-
-# # This is the project path
-# path <- find_rstudio_root_file()
 
@@
 # detectCores()
-cl <- makePSOCKcluster(4)
+cl <- makePSOCKcluster(6)
 registerDoParallel(cl)
@@
 ```{r}
-# create a list of the above model outputs
-filenames <- list.files(paste0(path, "/outputs/rf/models/oa_rf/regions"), pattern = "*.RData", full.names = T)
-
-for (i in filenames){
-  load(file = i)
-  i <- gsub("/home/nw19521/projects/web.diffusion/outputs/rf/models/oa_rf/regions/|.RData", "", i)
-  assign(i, model.all)  
-}
-
- fit.model.all <- mget(ls(pattern = "^region_"))
-
 # create a list of the above model outputs
@@
 for (i in regions){
-  #for (j in regions[!regions == i]){
     pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all", i)], 
@@
 
-path.out <- paste0(path, "/outputs/rf/test_regions_BUA.csv")
+path.out <- paste0(path, "/outputs/rf/figures/test_regions_BUA.csv")
 

### 2024-02-15 presentations/web_diffusion.Rmd
mac

@@
 
-![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture1.png){width="70%"}
+<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture1.png){width="70%"} -->
+![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture1.png){width="70%"}
 
@@
 
-![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture2.png){width="70%"}
+<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture2.png){width="70%"} -->
+![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture2.png){width="70%"}
 
@@
 # load 2000 co.uk
-# setwd("C:/Users/TranosE/DataShare/archive/nuts")
-path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"
+# path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"
+path.2000 <- "/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/all2000couk.csv"
 
@@
 
-![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
+<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png) -->
+![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
 
@@
 
-![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
+<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png) -->
+![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
 
@@
 
--   [A **hierarchy effect**: from main centres to secondary ones]{style="color:orange;"}
+-   [A hierarchy effect: from main centres to secondary ones]{style="color:orange;"}
 
--   [A **neighborhood effect**: first "hitting" nearby locations]{style="color:olive;"}
+-   [A neighborhood effect: first "hitting" nearby locations]{style="color:olive;"}
 
--   [**S-shaped** pattern in the cumulative level of adoption]{style="color:violet;"}
+-   [S-shaped pattern in the cumulative level of adoption]{style="color:violet;"}
 
@@
 
-```{r similarity_OA, eval = F, out.height= "70%", out.width="70%", message=FALSE}
-
-path.image <- paste0(path, "/outputs/rf/figures/heatmap_OA.png")
-knitr::include_graphics(path.image)
+```{r similarity_OA, out.height= "70%", out.width="70%", message=FALSE}
+path.table <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
+read.csv(path.table) %>% 
+  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
+  kable()
 ```

### 2024-02-15 src/la_RF.Rmd
mac

@@
 
-## spatial data NEW
+## Spatial data
 
@@
   st_as_sf(coords = c("long", "lat"), crs = 4326)
-```
-
-## spatial data
-
-```{r, eval=FALSE}
-# get LA for the UK
-path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-la <- readOGR(path.geo)
-# source: https://geoportal.statistics.gov.uk/
-
-# spatial transformations
-la <- spTransform(la, CRS("+init=epsg:4326"))
-
-#la.f <- fortify(la, region = "LAD21CD")
-
-# cities
-cities <- maps::world.cities %>% 
-  filter(country.etc=="UK") %>% 
-  arrange(pop) %>% tail(10) 
-cities.sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
 ```
@@
 
-## Spatial panel
-
-```{r, eval=F}
-#spatial weight matrix
-sf_use_s2(FALSE)
-nb <- poly2nb(la, queen = T, snap = 1)
-listw <- nb2listw(nb, style = "W", zero.policy = T)
-
-test <- la
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
-  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
-         year = ifelse(is.na(year), t, year))
-
-# Load required libraries
-library(spdep)
-library(splm)
-
-# Assuming you have a dataset named 'data' containing your variables: Y, W, Wc, and X
-
-# Create spatial lag matrices
-W <- mat2listw(W, style="W")
-Wc <- mat2listw(Wc, style="W")
-
-# Estimate spatial panel model
-spatial_panel_model <- spml(n ~ density + London + dist + dist.retail, 
-                            listw=listw,
-                            data=df %>% filter(!is.na(n)), 
-                            index = c("ladcd", "year"),
-                            effect = "individual",
-                            #method = "BFGS", # eigen
-                            model= "random", 
-                            lag=TRUE, 
-                            na.action = na.omit) #na.fail, 
-                            #zero.policy = NULL)
-
-m.spreml <- spreml(n ~ density + London + dist + dist.retail, 
-                            w=listw,
-                            data=df %>% filter(!is.na(n)), 
-                            index = c("ladcd", "year"),
-                            effect = "individual",
-                            #method = "BFGS", # eigen
-                            lag=TRUE, 
-                            na.action = na.omit) #na.fail, 
-                            #zero.policy = NULL)
-
-# Print summary of the model
-summary(m.spreml)
-```
-
-calculate k neighbours and find out why can't have spatial FE. Hierarchical?
-
-```{r, eval=F}
-#install.packages('SDPDmod')
-library(SDPDmod)
-
-#spatial weights matrix
-sf_use_s2(FALSE)
-nb <- poly2nb(la, queen = T, snap = 1)
-listw <- nb2listw(nb, style = "W", zero.policy = T)
-
-#spatial weight matrix
-binary_matrix <- nb2mat(nb, style = "B", zero.policy = TRUE)
-class(binary_matrix)
-w.binary <- rownor(binary_matrix)
-isrownor(w.binary)
-w.binary <- rownor(binary_matrix)
-isrownor(w.binary)
-
-#spatial weights matrix
-library("sf")
-path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-la.sf <- st_read(path.geo)
-W_2n <- mOrdNbr(sf_pol = la.sf, m = 2) ## second order neighbors
-W_knn <- mNearestN(distMat = gN3dist, m = 5) ## 5 nearest neighbors
-
-#spatial weights matrix
-sf_use_s2(FALSE)
-la.c <- st_centroid(la.sf)
-kn <- knearneigh(la.c, k = 5)
-knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
-knn_binary_matrix <- nb2mat(knn, style = "W", zero.policy = TRUE)
-isrownor(knn_binary_matrix)
-
-mod2<-SDPDm(formula = n ~ density,# + London,# + dist + dist.retail, 
-            #W = w.binary,
-            #W = binary_matrix,
-            #W = knn_binary_matrix,
-            W = knn_binary_matrix,
-            data=df %>% drop_na(),
-            index = c("ladcd", "year"),
-            model = "sar", 
-            effect = "twoways",
-            dynamic = T,
-            LYtrans = TRUE, #it affects coefficients, need to read about it
-            tlaginfo = list(ind = NULL, tl = T, stl = T),
-            )
-summary(mod2)
-
-# if distances are included, effect = year. 
-# Rho is positive and sig, AR is positive and sig, BUT Wn(t-1) is negative and sig
-```
-
 ## df for RF
@@
 
-## RF
-
-### train on all but one regions
-
-```{r}
-# CV
-tc <- trainControl(method = "cv",
-                   number = 10,
-                   savePredictions = 'final')
-
-# London region: E12000007
-# training dataset
-tr.df <- df %>% ungroup() %>% filter(RGN21CD!="E12000007"|
-                                     RGN21CD!="E12000008")
-
-set.seed(71)
-m.test <- train(n ~ density + area + dist + dist.retail + London +
-                  n.slag + year + n.nearest.city, #n.l.slag
-                data = tr.df,
-                trControl = tc,
-                method = "rf", 
-                importance = TRUE,
-                na.action = na.omit)
-summary(m.test)
-plot(m.test)
-m.test
-print(m.test)
-
-varimp_mars <- varImp(m)
-plot(varimp_mars, main="Variable Importance")
-```
-
-### test on the hold out regions
-
-Wales region is missing!!!
-
-```{r}
-pred <- predict(m, df %>% ungroup() %>% filter(RGN21CD=="E12000008"), na.action = na.pass)
-length(pred)
-
-df %>% ungroup() %>% filter(RGN21CD=="E12000008") %>% filter(if_any(everything(), is.na))
-
-pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN21CD=="E12000008"))
-postResample(pred = pred$value, obs = pred$n) ## 0.99
-```
-
-## time slices
-
-```{r}
-help <- df %>% createTimeSlices(initialWindow = 5,
-                        horizon = 1,
-                        fixedWindow = F)
-
-#Adding t-1 lag for n
-
-set.seed(71)
-m.test.time <- train(n ~ area + dist + dist.retail + London + #density
-             n.lag + n.slag + n.l.slag + n.nearest.city + year,
-           data = df %>% filter(year==1997 | year == 1998),
-           trControl = tc,
-           method = "rf", 
-           importance = TRUE,
-           na.action = na.omit)
-summary(m.test.time)
-plot(m.test.time)
-m.test.time
-
-varimp_mars <- varImp(m.test)
-plot(varimp_mars, main="Variable Importance")
-
-pred <- predict(m.test, df %>% ungroup() %>% filter(year==1999), na.action = na.pass)
-length(pred)
-
-pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(year==1999))
-postResample(pred = pred$value, obs = pred$n) ## 0.60 
-```
-
-## RF rolling
-
-### train the model in a loop
-
-TODO:
-
-- [x] read more about trainColntrol
-- [ ] other variables, LA Classification, check regressions, 
-- [x] N/S
-- [x] reproducible
-- [ ] n vs. log(n/n.lag), %>% mutate(n.lag = n.lag + 0.00001, n = n + 0.00001, way to bring S?
-- [x] center or scale with preProcess?? preProc = c("center", "scale") makes no difference
-- [ ] direction of relationship, or cor()
-- [x] importance over time
-- [x] spatial sampling
-- [ ] hierarchy: ranking correlation and line graphs 
-- [ ] OA
-
-```{r}
-# Train on t & t + 1
-
-# Make the parallelisation reproducible using https://stackoverflow.com/questions/13403427/fully-reproducible-parallel-models-using-caret. 
-# See also https://stackoverflow.com/questions/27944558/set-seed-parallel-random-forest-in-caret-for-reproducible-result
-
-#length is = (n_repeats*nresampling)+1
-seeds <- vector(mode = "list", length = 11)
-
-#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South
-
-#for the last model
-seeds[[11]]<-sample.int(1000, 1)
-
-# CV
-tc <- trainControl(method = "cv",
-                   number = 10,
-                   seeds = seeds,
-                   allowParallel = T,
-                   savePredictions = 'final')
-
-detectCores()
-cl <- makePSOCKcluster(14)
-registerDoParallel(cl)
-
-# run the model for every 2 years
-start.time <- Sys.time()
-years <- 1997:2012
-
-for (t in years){
-  tplus1 <- t + 1
-  set.seed(71)
-  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
-                      n.slag + n.l.slag + n.nearest.city + year + n.lag,
-                     data = df %>% filter(year==t | year==tplus1), #%>% 
-                       # mutate(growth = ifelse(growth == -Inf | growth == Inf, NA, growth)) %>% 
-                       # mutate(growth = ifelse(is.na(growth), 0, growth)), 
-                     #%>% mutate(growth = ifelse(growth== -Inf | growth == Inf, NA, growth))
-                     trControl = tc,
-                     method = "ranger", #"rpart", # rf lm
-                     na.action = na.omit,
-                     #preProc = c("center", "scale"),
-                     importance = 'impurity') # TRUE for rf
-    assign(paste0("model.all",as.character(t + 1)), model.all)
-}
-
-stopCluster(cl)
-end.time <- Sys.time()
-time.taken <- round(end.time - start.time,2)
-time.taken
-
-# df %>% filter(growth == Inf) 
-# log(n/n.lag)
-
-# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
-# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
-```
-
-### Resamples for plots
-
-```{r}
-# create a list of the above model outputs
-fit.model.all <- mget(ls(pattern = "^model.all\\d"))
-
-# resamples
-fit.model.all.res <- resamples(fit.model.all)
-summary(fit.model.all.res)
-
-# plot metrics
-
-# TODO: fix the year labels
-
-scales <- list(x=list(relation="free"), y=list(relation="free"))
-bwplot(fit.model.all.res, scales=scales)
-
-# TODO: fix the year labels
-```
-
-### Variable importance
-
-```{r , include=TRUE, results= 'markup', message=FALSE}
-lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
-```
-
-```{r}
-help <- lapply(fit.model.all, function(x) varImp(x)) 
-help <- lapply(help, function(x) x[[1]])
-help <- do.call("rbind", help)
-
-importance <- help %>% rownames_to_column() %>% 
-  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
-         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
-  rename(importance = Overall) %>% 
-  relocate(importance, .after = year) %>% 
-  dplyr::select(-rowname)
-
-importance %>% ggplot() +
-  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
-  scale_color_discrete(name = "Features") +
-  #scale_linetype_manual(name = "Website density regressions111") 
-  ylab("Importance") + 
-  ggtitle("Feature importance") +
-  theme_minimal() +
-  theme(axis.title.x=element_blank(),
-        plot.title = element_text(hjust = 0.5))
-
-```
-
-### test
-
-```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
-
-# growth or n
-
-# for a reference point
-df %>% group_by(year) %>%
-  summarise(min = min(n), max=max(n),
-            mean = mean(n), median = median(n)) %>%
-  round(2) %>% kable()
-```
-
-```{r}
-## Test on t + 2
-
-# It will create: 2001 (train on 2000-2001) - 2011 (train on 2010)
-# I need: 2001 (train on 2000-2001) - 2009 (train on 2008-2009)
-
-# use the model trainned in year t to predict io of year t + 1
-years <- (1998:2011)
-pred.by.year.all <- data.frame(matrix(NA,nrow=374, ncol = 0))
-for (t in years){
-  t_plus_1 <- t + 1
-  pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",t)], df[df$year==t_plus_1,])
-  pred <- as.data.frame(pred)
-  rownames(pred) <- c()
-  pred.by.year.all <- cbind(pred.by.year.all, pred)
-}
-
-#change column names to match with total
-names(pred.by.year.all) <- 1999:2012
-
-# wide to long
-pred.by.year.all <- gather(pred.by.year.all,key = year,value = predictions,1:14)
-
-# drop 2000 and select variables form total
-df.no1997_1 <- df %>% 
-  filter(year!=1996 & year!=1997 & year!=1998) %>% 
-  dplyr::select(ladcd, year, n) # CHANGE n WITH growth
-
-# column bind prediction and data  
-pred.by.year.all <- cbind(pred.by.year.all, df.no1997_1 %>% arrange(year, ladcd)) #DOUBLE CHECK ARRANGE
-
-# pred.by.year to list by year
-pred.by.year.all.list <- split(pred.by.year.all, pred.by.year.all$year)
-
-# calculate metrics for every year
-rf.year.all.metrics <- lapply(pred.by.year.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))  # CHANGE n WITH growth
-rf.year.all.metrics %>% as.data.frame() %>%
-  round(2) %>%
-  kable()
-
-# there are 2 columns named year, so I drop one
-pred.by.year.all$year <- NULL
-
-ggplot(data = pred.by.year.all,aes(x = n, y = predictions)) +
-  geom_point(colour = "blue") +
-  geom_abline(intercept = 0, slope = 1, colour = "red") +
-  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
-  facet_wrap(~ year,ncol = 2) +
-  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
-  ggtitle("Predicted vs. actual trade by year")
-```
-
-## RF spatial
-
-```{r}
-df <- df %>% mutate(help = substr(ladcd, start = 1, stop = 2)) %>% 
-  mutate(RGN21CD = ifelse(is.na(RGN21CD), help, RGN21CD)) %>% 
-  ungroup() %>% 
-  dplyr::select(-help)
-
-df <- df %>% dplyr::select(-n.lag)
-
-```
-
-### train the model in a loop
-
-```{r}
-#length is = (n_repeats*nresampling)+1
-seeds <- vector(mode = "list", length = 11)
-
-#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South
-
-#for the last model
-seeds[[11]]<-sample.int(1000, 1)
-
-# CV
-tc <- trainControl(method = "cv",
-                   number = 10,
-                   seeds = seeds,
-                   allowParallel = T,
-                   savePredictions = 'final')
-
-detectCores()
-cl <- makePSOCKcluster(14)
-registerDoParallel(cl)
-
-# run the model for every 2 years
-start.time <- Sys.time()
-regions <- df %>% distinct(RGN21CD)
-regions <- as.vector(regions$RGN21CD)
-
-for (i in regions){
-  set.seed(71)
-  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
-                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
-                     data = df %>% filter(RGN21CD==i), trControl = tc,
-                     method = "ranger", 
-                     na.action = na.omit,
-                     #preProc = c("center", "scale"),
-                     importance = TRUE)
-  assign(paste0("model.all",as.character(i)), model.all)
-}
-
-stopCluster(cl)
-end.time <- Sys.time()
-time.taken <- round(end.time - start.time,2)
-time.taken
-
-# df %>% filter(growth == Inf) 
-# log(n/n.lag)
-
-# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
-# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
-```
-
-### Resamples for plots
-
-```{r}
-# create a list of the above model outputs
-fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
-
-# resamples
-fit.model.all.res <- resamples(fit.model.all)
-summary(fit.model.all.res)
-
-# plot metrics
-
-# TODO: fix the year labels
-
-scales <- list(x=list(relation="free"), y=list(relation="free"))
-bwplot(fit.model.all.res, scales=scales)
-
-# TODO: fix the year labels
-```
-
-### Variable importance
-
-```{r , include=TRUE, results= 'markup', message=FALSE}
-lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
-```
-
-```{r}
-help <- lapply(fit.model.all, function(x) varImp(x)) 
-help <- lapply(help, function(x) x[[1]])
-help <- do.call("rbind", help)
-
-importance <- help %>% rownames_to_column() %>% 
-  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
-         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
-  rename(importance = Overall) %>% 
-  relocate(importance, .after = year) %>% 
-  dplyr::select(-rowname)
-
-importance %>% ggplot() +
-  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
-  scale_color_discrete(name = "Features") +
-  #scale_linetype_manual(name = "Website density regressions111") 
-  ylab("Importance") + 
-  ggtitle("Feature importance") +
-  theme_minimal() +
-  theme(axis.title.x=element_blank(),
-        plot.title = element_text(hjust = 0.5))
-
-```
-
-### test
-
-```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
-
-# growth or n
-
-# for a reference point
-df %>% group_by(year) %>%
-  summarise(min = min(n), max=max(n),
-            mean = mean(n), median = median(n)) %>%
-  round(2) %>% kable()
-```
-
-```{r}
-## Test on t + 2
-
-# train on one region, test on all the rest
-
-df.predictions <- data.frame()
-for (i in regions){
-  for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
-    pred <- as.data.frame(pred)
-    rownames(pred) <- c()
-    assign(paste0("region.predict.model.", i, ".on.", j), pred)
-
-    d <- paste0("region.predict.model.", i, ".on.", j)
-    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
-    d <- d[,c(5, 2, 4, 6, 1)]
-    colnames(d)[5] <- "predictions"
-    d <- d %>% mutate(trained.on = i)
-    df.predictions <- rbind(d, df.predictions)
-  }
-}
-
-df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))
-
-# split to list by region pair
-pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
-
-# calculate metrics for every region pair 
-rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))  # CHANGE n WITH growth
-
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% 
-  dplyr::select(RGN21CD, RGN21NM) %>% 
-  distinct() %>% 
-  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
-
-rf.year.all.metrics %>% 
-  as.data.frame() %>%
-  add_rownames(var = "metrics") %>% 
-  pivot_longer(!metrics, names_to = "train.test") %>% 
-  pivot_wider(names_from = metrics, values_from = value) %>% 
-  separate(train.test, c("train", "test"), remove = F) %>% 
-  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
-  rename(train.region=RGN21NM) %>% 
-  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
-  rename(test.region=RGN21NM) %>% 
-  arrange(Rsquared)
-  #summarise(r = range(Rsquared))
-  #round(2) %>%
-  #kable()
-```
-
-### test plots
-
-```{r}
-#plots
-ggplot(data = df.predictions, aes(x = n, y = predictions)) +
-  geom_point(colour = "blue") +
-  geom_abline(intercept = 0, slope = 1, colour = "red") +
-  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
-  facet_wrap(~ test.train, ncol = 2) +
-  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
-  ggtitle("Predicted vs. actual trade by year")
-```
-
 ## RF with CAST
@@
   mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
-                          ifelse(rowname == "n.London.lag", "London's web density, t-1", 
+                          ifelse(rowname == "n.London.lag", "London's website density, t-1", 
                                  ifelse(rowname == "London", "distance to London",
-                                        ifelse(rowname == "n.l.slag", "spatial and temporal lag of web density",
-                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's web density, t-1", rowname)))))) %>% 
+                                        ifelse(rowname == "n.l.slag", "spatial and temporal lag of website density",
+                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's website density, t-1", rowname)))))) %>% 
   column_to_rownames()
@@
 
-#### OLD
-
-```{r, eval=FALSE}
-# train on one region, test on all the rest
+## RF growth with CAST
+
+### train the model in all data, growth
+
+```{r}
+set.seed(123)
+indices <- CreateSpacetimeFolds(df %>% filter(year > 1999),                     # FILTER BY YEAR FOR GROWTH
+                                spacevar = "ladcd",
+                                timevar = "year", 
+                                k = 10)
+
+#length is = (n_repeats*nresampling)+1
+seeds <- vector(mode = "list", length = 11)
+
+#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # , 11) 9 after adding South
+
+#for the last model
+seeds[[11]]<-sample.int(1000, 1)
+
+# CV
+tc <- trainControl(method = "cv",
+                   number = 10,
+                   seeds = seeds,
+                   allowParallel = T,
+                   index = indices$index,
+                   savePredictions = 'final')
+
+detectCores()
+cl <- makePSOCKcluster(14)
+registerDoParallel(cl)
+
+start.time <- Sys.time()
+model.all <- train(growth ~                                                     # REPLACE n WITH growth
+                     #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
+                     #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
+                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist, #n.nearest.retail.lag + dist.retail,
+                   data = df %>% filter(year > 1999),                           # FILTER BY YEAR FOR GROWTH
+                   trControl = tc,
+                   method = "ranger",
+                   na.action = na.omit,
+                   #preProc = c("center", "scale"),
+                   importance = "impurity")
+
+stopCluster(cl)
+end.time <- Sys.time()
+time.taken <- round(end.time - start.time,2)
+time.taken
+
+print(model.all)
+
+# [reproducible results]
+# Random Forest 
+# 
+# 4862 samples
+#    6 predictor
+# 
+# No pre-processing
+# Resampling: Cross-Validated (10 fold) 
+# Summary of sample sizes: 4044, 3685, 4056, 4056, 4032, 3707, ... 
+# Resampling results across tuning parameters:
+# 
+#   mtry  splitrule   RMSE       Rsquared   MAE      
+#   2     variance    0.2004565  0.6285613  0.1460545
+#   2     extratrees  0.2001702  0.6342523  0.1480880
+#   4     variance    0.2080782  0.6066952  0.1503476
+#   4     extratrees  0.2065532  0.6220336  0.1527747
+#   6     variance    0.2099617  0.5996347  0.1507628
+#   6     extratrees  0.2084788  0.6180880  0.1544252
+# 
+# Tuning parameter 'min.node.size' was held constant at a value of 5
+# RMSE was used to select the optimal model using the smallest value.
+# The final values used for the model were mtry = 2, splitrule = extratrees and min.node.size = 5.
+
+varimp_mars <- varImp(model.all) 
+varimp_mars$importance <- varimp_mars$importance %>% 
+  rownames_to_column() %>% 
+  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
+                          ifelse(rowname == "n.London.lag", "London's website density, t-1", 
+                                 ifelse(rowname == "London", "distance to London",
+                                        ifelse(rowname == "n.l.slag", "spatial and temporal lag of website density",
+                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's website density, t-1", rowname)))))) %>% 
+  column_to_rownames()
+
+path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_growth.png")
+ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
+ggsave(path.out)
+```
+
+### train the model in a loop for all but one regions, growth
+
+```{r}
+rm(indices)
+
+#length is = (n_repeats*nresampling)+1
+seeds <- vector(mode = "list", length = 11)
+
+#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
+set.seed(831)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # 6 feature + 2
+
+#for the last model
+set.seed(832)
+seeds[[11]]<-sample.int(1000, 1)
+
+# detectCores()
+cl <- makePSOCKcluster(14)
+registerDoParallel(cl)
+
+# train in every region
+start.time <- Sys.time()
+
+regions <- df %>% ungroup() %>% distinct(RGN21CD)
+regions <- as.vector(regions$RGN21CD)
+
+for (i in regions){
+
+  set.seed(71)
+  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i & year > 1999),      # FILTER BY YEAR FOR GROWTH 
+                                spacevar = "ladcd",
+                                timevar = "year", 
+                                k = 10)
+  # CV
+  tc <- trainControl(method = "cv", #MOVED HERE< DOUBLE CHECK
+                   number = 10,
+                   seeds = seeds,
+                   allowParallel = T,
+                   index = indices$index,
+                   savePredictions = 'final')
+
+  model.all <- train(growth ~                                                   # REPLACE n WITH growth
+                       #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
+                       #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
+                       n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
+                   data = df %>% filter(RGN21CD!=i & year > 1999),              # FILTER BY YEAR FOR GROWTH
+                   trControl = tc,
+                   method = "ranger",
+                   na.action = na.omit,
+                   #preProc = c("center", "scale"),
+                   importance = "impurity")
+  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data
+}
+
+stopCluster(cl)
+end.time <- Sys.time() 
+time.taken <- round(end.time - start.time,2)
+time.taken # ca. 4 minutes 
+```
+
+### Resamples for plots, growth
+
+```{r}
+# create a list of the above model outputs
+fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
+
+# resamples
+fit.model.all.res <- resamples(fit.model.all)
+summary(fit.model.all.res)
+
+# plot metrics
+
+# TODO: fix the year labels
+
+scales <- list(x=list(relation="free"), y=list(relation="free"))
+bwplot(fit.model.all.res, scales=scales)
+
+# TODO: fix the year labels
+```
+
+### Variable importance, growth
+
+```{r , include=TRUE, results= 'markup', message=FALSE}
+lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
+```
+
+```{r}
+help <- lapply(fit.model.all, function(x) varImp(x)) 
+help <- lapply(help, function(x) x[[1]])
+help <- do.call("rbind", help)
+
+importance <- help %>% rownames_to_column() %>% 
+  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
+         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
+  rename(importance = Overall) %>% 
+  relocate(importance, .after = year) %>% 
+  dplyr::select(-rowname)
+
+importance %>% ggplot() +
+  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
+  scale_color_discrete(name = "Features") +
+  #scale_linetype_manual(name = "Website density regressions111") 
+  ylab("Importance") + 
+  ggtitle("Feature importance") +
+  theme_minimal() +
+  theme(axis.title.x=element_blank(),
+        plot.title = element_text(hjust = 0.5))
+
+```
+
+### Test on one region, growth
+
+```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
+
+# for a reference point
+df %>% group_by(year) %>%
+  summarise(min = min(n), max=max(n),
+            mean = mean(n), median = median(n)) %>%
+  round(2) %>% kable()
+```
+
+```{r}
+# train on all but one region, test on that region
 
@@
 for (i in regions){
-  for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], 
+                    df[df$RGN21CD==i & df$year > 1999,])                                # FILTER BY YEAR
     pred <- as.data.frame(pred)
     rownames(pred) <- c()
-    assign(paste0("region.predict.model.", i, ".on.", j), pred)
-
-    d <- paste0("region.predict.model.", i, ".on.", j)
-    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
-    d <- d[,c(5, 2, 4, 6, 1)]
+    assign(paste0("region.predict.model.", ".on.", i), pred)
+
+    d <- paste0("region.predict.model.", ".on.", i)
+    d <- cbind(get(d), df %>% filter(RGN21CD==i & year > 1999))                         # FILTER BY YEAR
+    d <- d %>% dplyr::select(RGN21CD, ladcd, year, growth, 1)                           # REPLACE n WITH growth
     colnames(d)[5] <- "predictions"
-    d <- d %>% mutate(trained.on = i)
+    d <- d %>% mutate(tested.on = i)
     df.predictions <- rbind(d, df.predictions)
 }
-}
-
-df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))
+
+df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", tested.on))
 
 # split to list by region pair
-pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
+pred.by.region.all.list <- split(df.predictions, df.predictions$RGN21CD)
 
@@
 rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))  # CHANGE n WITH growth
+                                                                     obs = x$growth))   # REPLACE n WITH growth
 
@@
 
-path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
+path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv")
 rf.year.all.metrics %>% 
@@
   rownames_to_column(var = "metrics") %>% 
-  pivot_longer(!metrics, names_to = "train.test") %>% 
+  pivot_longer(!metrics, names_to = "test.on") %>% 
   pivot_wider(names_from = metrics, values_from = value) %>% 
-  separate(train.test, c("train", "test"), remove = F) %>% 
-  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
-  rename(train.region=RGN21NM) %>% 
-  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
-  rename(test.region=RGN21NM) %>% 
+  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
   arrange(Rsquared) %>% 
-  dplyr::select(train.region, test.region, Rsquared) %>% 
-  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
-  geom_tile() +
-  coord_fixed() +
-  labs(y = "", x = "") + #y: test.region, x: train.region
-  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
-  theme(#legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust=1),
-        panel.border = element_blank(), 
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.background = element_rect(fill = "white"))
-ggsave(path.out)
+  dplyr::select(RGN21NM, Rsquared) %>% 
+  rename(Region = RGN21NM) %>% 
+  write_csv(path.out)
+  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()
+
+# ggplot(aes(train.region, test.region, fill= Rsquared)) + 
+#   geom_tile() +
+#   coord_fixed() +
+#   labs(y = "", x = "") + #y: test.region, x: train.region
+#   scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
+#   theme(#legend.position = "none",
+#         axis.text.x = element_text(angle = 45, hjust=1),
+#         panel.border = element_blank(), 
+#         panel.grid.major = element_blank(),
+#         panel.grid.minor = element_blank(),
+#         panel.background = element_rect(fill = "white"))
+# ggsave(path.out)
 

### 2024-06-11 presentations/web_diffusion.Rmd
ggplot

@@
 ```{r moransi_la, results='asis'}
-path.in <- paste0(path, "/outputs/lisa/morani_la.csv")
+path.in <- paste0(path, "/outputs/lisa/oa_ggplot/morani_la.csv")
 
@@
 ```{r, eval=T, moransi_oa, results='asis'}
-path.in <- paste0(path, "/outputs/lisa/morani_oa.csv")
+path.in <- paste0(path, "/outputs/lisa/oa_ggplot/morani_oa.csv")
 # read.csv(path.in) %>% dplyr::select(year, morani) %>% 
@@
 ```{r lisa_la, results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/lisa/")
+path.images <- paste0(path, "/outputs/oa_ggplot/lisa/")
 
@@
 ```{r lisa_oa, eval=T,results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/lisa/")
+path.images <- paste0(path, "/outputs/oa_ggplot/lisa/")
 

### 2024-06-11 src/level_lisa_la_sf.Rmd
ggplot

--- 2023-10-27 src/level_lisa_la.Rmd
@@
 library(rprojroot)
-# library(rgdal)
-# library(rgeos)
+library(plm)
 library(sf)
-library(spdep)
-library(tmap)
-#library(maptools)
-library(knitr)
-library(REAT)
-library(tidygeocoder)
-library(geosphere)
-library(broom)
-library(foreach)
-library(raster)
-library(plm)
-library(lmtest)
+library(magick)
 
@@
 # spatial transformations
-la <- spTransform(la, CRS("+init=epsg:4326"))
-
-la.f <- fortify(la, region = "LAD21CD")
+la <- st_transform(la, 4326)
 
@@
 ```{r}
-
-path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", n, "_la") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_la")  # for multiple PC
+path.out <- paste0(path, "/outputs/lisa/la_ggplot/lisa_level_pc", n, "_la") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa/la_ggplot/lisa_level_pc", m, "_la")  # for multiple PC
 
@@
   test <- la
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
+  test <- test %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
     mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
@@
   #LISA
-lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
-#lmoran
+  #lmoran <- cbind(test, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
+  lmoran <- localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude)
 
-# centers the local Moran's around the mean
-lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
-lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
+  quadr <- attr(lmoran, "quadr")
+  #quadr[lmoran[,5] > 0.05 | is.na(lmoran[,5]), ] <- NA
+  quadr[p.adjust(lmoran[,5], method = "fdr") > 0.05 | is.na(lmoran[,5]), ] <- NA
 
-# centers the variable of interest around its mean
+  palette_new <- grDevices::colorRampPalette(
+            colors = c("blue", "lightsalmon", "lightblue", "red", "white"))(5)
 
-lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
-lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
+  ggplot(data = la, aes(fill = quadr$median)) +
+    geom_sf() +
+    scale_fill_manual(values = palette_new, na.value="white", drop = F) +
+    theme_minimal() +
+    theme(panel.grid.major = element_line(colour = "transparent"),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          legend.title=element_blank()) +
+          #legend.position="none") +
+    guides(fill = guide_legend(reverse = TRUE)) +
+    labs(title = t,
+        subtitle = "LISA, websites per firm in a LA")
 
-signif <- 0.05
-#lmoran
-
-lmoran <- lmoran%>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
-  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
-  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%
-
-#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
-# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
-#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))
-
-mun_merge_new<- merge(test, lmoran, by="LAD21CD")
-
-# R p value map
-breaks = c(0, 1, 2, 3, 4, 5)
-
-cities <- maps::world.cities %>% 
-  filter(country.etc=="UK") %>% 
-  arrange(pop) %>% tail(10) 
-tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
-
-map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, 
-                                         palette = c("white","red","blue",
-                                                     rgb(0,0,1,alpha=0.4),
-                                                     rgb(1,0,0,alpha=0.4)), 
-                                         labels = c("Not significant",
-                                                    "High-High",
-                                                    "Low-Low",
-                                                    "Low-High",
-                                                    "High-Low"), 
-                                         title="LISA, websites per firm in a LA") +
-  tm_legend(text.size = 1) +
-  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
-  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
-  #tm_borders(alpha=.1) +
-  tm_borders(lwd = 0) +
-  tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
-  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
-  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
-tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
+  filename = paste(path.out, t, ".png", sep = "")
+  ggsave(filename, bg='#ffffff')
 }
-
-#ggplot map solution.
-# mun_merge_new %>% st_as_sf(crs = 27700) %>% #st_transform(4326) %>%
-#   ggplot()+
-#   geom_sf(aes(fill = as.factor(quadrant)))
-# ggsave("test_ggplot.png", path = path)
-
 ```
 
-## lisa maps parallel
-
-```{r, eval=F}
-
-path.out <- paste0(path, "/outputs/lisa/lisa_level_pc_test", n, "_la") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_la")  # for multiple PC
-
-#setup cluster
-#https://www.blasbenito.com/post/02_parallelizing_loops_with_r/
-n.cores <- parallel::detectCores() - 1
-my.cluster <- parallel::makeCluster(
-  n.cores, 
-  type = "PSOCK"
-  )
-print(my.cluster)
-doParallel::registerDoParallel(cl = my.cluster)
-foreach::getDoParWorkers()
-
-morani <- data.frame()
-
-#spatial weight matrix
-sf_use_s2(FALSE)
-nb <- poly2nb(la)
-listw <- nb2listw(nb, style = "W", zero.policy = T)
-
-#for (t in 1996:2012){
-foreach(t = 1996:2012, .packages = c("tidyverse",
-                                     "rgdal",
-                                     "rgeos",
-                                     "spdep",
-                                     "tmap",
-                                     "maptools",
-                                     "knitr",
-                                     "broom")) %dopar% {
-test <- la
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
-  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
-         year = ifelse(is.na(year), t, year))
-
-#spatial weight matrix
-#moved out of the loop
-
-#Moran's I
-globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
-a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
-morani <- rbind(a, morani)
-
-#LISA
-lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
-#lmoran
-
-# centers the local Moran's around the mean
-lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
-lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
-
-# centers the variable of interest around its mean
-
-lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
-lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
-
-signif <- 0.05
-#lmoran
-
-lmoran <- lmoran%>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
-  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
-  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%
-
-#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
-# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
-#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))
-
-mun_merge_new<- merge(test, lmoran, by="LAD21CD")
-
-# R p value map
-breaks = c(0, 1, 2, 3, 4, 5)
-
-cities <- maps::world.cities %>% 
-  filter(country.etc=="UK") %>% 
-  arrange(pop) %>% tail(10) 
-tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
-
-map <- tm_shape(mun_merge_new, projection = "+init=epsg:4326") + tm_fill(col = "quadrant", breaks = breaks, 
-                                         palette = c("white","red","blue",
-                                                     rgb(0,0,1,alpha=0.4),
-                                                     rgb(1,0,0,alpha=0.4)), 
-                                         labels = c("Not significant",
-                                                    "High-High",
-                                                    "Low-Low",
-                                                    "Low-High",
-                                                    "High-Low"), 
-                                         title="LISA, websites per firm in a LA") +
-  tm_legend(text.size = 1) +
-  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
-  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
-  #tm_borders(alpha=.1) +
-  tm_borders(lwd = 0) +
-  tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
-  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
-  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ") +
-tmap_options(check.and.fix = TRUE)
-tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
-                                     }
-
-```
-
-## Moran I
+# Moran I
 
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/morani_la.csv")
+path.out <- paste0(path, "/outputs/lisa/la_ggplot/morani_la.csv")
 morani %>% as_tibble() %>% 
@@
 
-## Gini
+Both LISA and Moran's I are not the same as before. Double-check the process and 
+if okay no need to do the correction as Moran's I is not dropping a lot.
+Add legend up left, check grey background, create gif.
 
-```{r}
-
-# No need to add 0s like OAs as there aren't any LA with 0s
-
-gini.df <- data.frame()
-
-for (t in 1996:2012){
-    a <- df %>% filter(year==t)
-  b <- as.data.frame(gini(a$n))
-  b$year = t
-  gini.df <- rbind(gini.df, b)
-}
-
-path.out <- paste0(path, "/outputs/gini_la.csv")
-gini.df %>% write_csv(path.out)
-```
-
-## Getis-Ord
-
-```{r eval = F}
-
-t= 1998
-local_g <- localG(OA.Census$Qualification, nb_lw)
-local_g <- localG(df[df$year==t,]$n, listw)
-
-local_g <- cbind(OA.Census, as.matrix(local_g))
-names(local_g)[6] <- "gstat"
-
-```
-
-## Distances to cities and retail centres
-
-First set up the destinations. Origins are all the LA.
-
-Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
-Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)
-
-```{r}
-# LA centroids
-la.c <- gCentroid(la, byid=TRUE)@coords %>% 
-  as.data.frame() %>%
-  bind_cols(la@data) %>% 
-  dplyr::select(LAD21CD, x, y) %>% 
-  remove_rownames()
-
-# cities
-city.names <- c("London, UK", 
-"Birmingham, UK",
-"Glasgow, UK",
-"Liverpool, UK",
-"Bristol, UK",
-"Manchester, UK",
-"Sheffield, UK",
-"Leeds, UK",
-"Edinburgh, UK",
-"Leicester, UK")
-
-cities <- geo(city.names, no_query = F, method = "arcgis")
-
-# retail centres
-geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
-retail <-readOGR(geo.path) 
-#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
-retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id") #%>%
-
-retail.major.cetres <- retail.major.cetres.help %>% 
-  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
-  st_transform(4326) %>%
-  st_coordinates() %>%
-  as_tibble() %>% 
-  bind_cols(retail.major.cetres.help$id) %>% 
-  rename(id = '...3') %>% 
-  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
-```
-
-Distance to cities
-
-```{r}
-dist <- distm(cbind(la.c$x, la.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist <- round((dist/1000),2) %>% 
-  as_tibble()
-
-city.names <- city.names %>% stringr::str_remove(", UK")
-names(dist) <- city.names
-
-dist <- dist %>% bind_cols(la.c$LAD21CD) %>% 
-  rename(ladcd = last_col()) %>% 
-  relocate(ladcd)
-
-# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
-# dist$distMet <- apply(dist[,2:11], 1, min)
-
-dist <- transform(dist, dist = do.call(pmin, dist[-1]))
-
-# NOT USED AS IT SCREWS UP THE REGRESSIONS
-# Complete panel, N = 3871903, 17 year * 227759 OA
-# df <- df %>% bind_rows(dist %>% select(oa11cd)) %>% 
-#   mutate(year = ifelse(is.na(year),1000, year),
-#          country = substring(oa11cd, 1, 1)) %>% 
-#   filter(country != "L" & country != "M" & country != "N") %>% 
-#   filter(!is.na(country)) %>% 
-#   complete(year, oa11cd, fill = list(n = 0)) %>% 
-#   filter(year!=1000,
-#          !is.na(oa11cd)) %>% 
-#   select(-country)
-
-# tests
-# test$country <- substring(test$oa11cd, 1, 1)
-# unique(test$country)
-# length(unique(test$oa11cd))
-
-# calculate area
-la$SHAPE_Area <- area(la) #/ 1000000
-
-# Join with distance and area
-df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
-  left_join(la@data, by = c("ladcd" = "LAD21CD")) %>% 
-  rename(area = SHAPE_Area) %>% 
-  relocate(area, .after = n) %>% 
-  mutate(area = area / 1000000,
-         density = n / area) %>% 
-  relocate(density, .after = area)
-sapply(df, function(x) sum(is.na(x)))
-```
-
-Distance to retail centres
-
-```{r}
-dist.retail <- distm(cbind(la.c$x, la.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail <- round((dist.retail/1000),2) %>% 
-  as_tibble()
-
-# retail.names <- retail.major.cetres$RC_Name
-# names(dist.retail) <- retail.names
-
-dist.retail <- dist.retail %>% bind_cols(la.c$LAD21CD) %>% 
-  rename(LAD21CD = last_col()) %>% 
-  relocate(LAD21CD)
-
-# Minimum distance 
-dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
-dist.retail <- dist.retail %>% dplyr::select(LAD21CD, dist.retail)
-
-# Join with complete panel
-df <- df %>% left_join(dist.retail, by = c("ladcd" = "LAD21CD"))
-
-sapply(df, function(x) sum(is.na(x)))
-```
-
-## Panel regressions
-
-```{r}
-# London
-model <- plm(log(n+0.001) ~ log(London+.001)*as.factor(year), 
-             data = df,
-             index = c("ladcd", "year"),
-             model = "within",
-             effect = "twoways")
-summary(model)
-
-est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
-int <- confint(est)
-
-out.model <- bind_cols(tidy(est), 
-                       as_tibble(int)) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(year = 1997:2012) %>% 
-  rename(conf.low = '2.5 %',
-         conf.high = '97.5 %')
-
-path.out <- paste0(path,"/outputs/lisa/b_panel_dist_london_la.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                    width = 0.2,size  = 1,
-                    position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
-  ggtitle("Regression coefficients: distance to London on website counts, LA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-
-# Cities
-model <- plm(log(n+0.001) ~ log(dist+.001)*as.factor(year), 
-             data = df,
-             index = c("ladcd", "year"),
-             model = "within",
-             effect = "twoways")
-summary(model)
-
-est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
-int <- confint(est)
-
-out.model <- bind_cols(tidy(est), 
-                       as_tibble(int)) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(year = 1997:2012) %>% 
-  rename(conf.low = '2.5 %',
-         conf.high = '97.5 %')
-
-path.out <- paste0(path,"/outputs/lisa/b_panel_dist_cities_la.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                    width = 0.2,size  = 1,
-                    position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
-  ggtitle("Regression coefficients: distance to Cities on website counts, LA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-
-# Retail centres
-model <- plm(log(n+0.001) ~ log(dist.retail+.001)*as.factor(year), 
-             data = df,
-             index = c("ladcd", "year"),
-             model = "within",
-             effect = "twoways")
-summary(model)
-
-est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
-int <- confint(est)
-
-out.model <- bind_cols(tidy(est), 
-                       as_tibble(int)) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(year = 1997:2012) %>% 
-  rename(conf.low = '2.5 %',
-         conf.high = '97.5 %')
-
-path.out <- paste0(path,"/outputs/lisa/b_panel_dist_retail_la.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                    width = 0.2,size  = 1,
-                    position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
-  ggtitle("Regression coefficients: distance to Retail Centres on website counts, LA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-```
-
-## Website density regressions to London
-
-```{r}
-dens.London <- data.frame()
-out.model <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(n+0.001) ~ log(London), data = df[df$year==t,]) #density
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.London <- rbind(r2, dens.London)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model <- rbind(out.model_, out.model)
-}
-
-names(dens.London)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_london_la.png")
-ggplot(dens.London) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to London on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-
-out.model <- out.model %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_london_la.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-#ggplot(out.model, aes(x=year, y = estimate)) +  
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to London on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-```
-
-## Website density regressions to the nearest city
-
-```{r}
-dens.distMet <- data.frame()
-out.modelMet <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.distMet <- rbind(r2, dens.distMet)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.modelMet <- rbind(out.model_, out.modelMet)
-}
-
-names(dens.distMet)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_city_la.png")
-ggplot(dens.distMet) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest city on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-out.modelMet <- out.modelMet %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_city_la.png")
-ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-```
-
-## Website density regressions to the nearest retail centre
-
-```{r}
-dens.dist.retail <- data.frame()
-out.model.retail <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(n+0.001) ~ log(dist.retail+0.001), data = df[df$year==t,]) #density
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model.retail <- rbind(out.model_, out.model.retail)
-}
-
-names(dens.dist.retail)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_centre_la.png")
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest retail centre on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-out.model.retail <- out.model.retail %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_centre_la.png")
-ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest retail centre on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-# library(pscl)
-# m1 <- zeroinfl((n) ~ log(dist.retail+0.001),
-#   data = df[df$year==t,], dist = "negbin", EM = TRUE)
-# summary(m1)
-# 
-# df %>% filter(year == 1996) %>% 
-#   group_by(group = cut(n, breaks = c(0, 1, 2, 10, 100, 1000, 10000), include.lowest = T, right = F)) %>%  #seq(0, max(n), 100))) %>%
-#   summarise(n = sum(n))
-```
-
-All R^2 dataframe
-
-```{r}
-path.out <- paste0(path, "/outputs/r2_all_la.csv")
-
-r2.all <- dens.London %>% 
-  rename(London.LA = rsquared) %>% 
-  left_join(dens.distMet) %>% 
-  rename(cities.LA = rsquared) %>% 
-  left_join(dens.dist.retail) %>% 
-  rename(retail.LA = rsquared) %>% 
-  relocate(year, .before = London.LA) %>% 
-  arrange(year)
-
-write_csv(r2.all, path.out)
-```
+https://profrichharris.github.io/MandM/autocorrelation.html#moran-plot-and-local-moran-values

### 2024-06-11 src/level_lisa_oa_sf.Rmd
ggplot

--- 2023-03-22 src/level_lisa_oa.Rmd
@@
 library(rprojroot)
-library(rgdal)
-library(rgeos)
+library(sf)
+library(magick)
 library(spdep)
-library(tmap)
-library(maptools)
-library(knitr)
-library(REAT)
-library(tidygeocoder)
-library(geosphere)
-library(broom)
+library(kableExtra)
 
@@
 path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
-oa.ew <- readOGR(path.geo)
+oa.ew <- st_read(path.geo)
 # source: https://geoportal.statistics.gov.uk/
@@
 # spatial transformations
-oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))
+oa.ew <- st_transform(oa.ew, 4326)  # EPSG code for WGS84
 
 # keep in the data slot only the ONS Output Area id, renaming it as 'id'
-oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
-colnames(oa.ew@data) <- c('id', 'area')
-
-# reassign the polygon IDs
-oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))
-
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.ew)
-
-# # get OA for Scotland
-# path.geo.sc <- paste0(path, "/data/raw/SG_DataZoneBdry_2011")
-# oa.sc <- readOGR(dsn=path.geo.sc, layer = "SG_DataZone_Bdry_2011")
-# # source: https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011
-# 
-# # spatial transformations
-# oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-# 
-# # Scotland (follows same steps as EW, see notes above)
-# oa.sc <- oa.sc[, 'DataZone']
-# colnames(oa.sc@data) <- c('id')
-# 
-# # reassign the polygon IDs
-# oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-# 
-# # check the CRS has changed correctly, and the data slot has shrink to only the ID
-# summary(oa.sc)
+oa.ew <- oa.ew %>% 
+  dplyr::select(OA11CD, Shape__Area) %>% 
+  rename(id = OA11CD,
+         area = Shape__Area)
 
 path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
-oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
+oa.sc <- st_read(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
 # source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries
@@
 # spatial transformations
-oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
+oa.sc <- st_transform(oa.sc, 4326)  # EPSG code for WGS84
 
 # Scotland (follows same steps as EW, see notes above)
-oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
-colnames(oa.sc@data) <- c('id', 'area')
+oa.sc <- oa.sc %>% 
+  dplyr::select(code, SHAPE_1_Ar) %>% 
+  rename(id = code,
+         area = SHAPE_1_Ar)
 
-# reassign the polygon IDs
-oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
+# NI
+path.geo.ni <- paste0(path, "/data/raw/ni_small_area/")
+oa.ni <- st_read(dsn = path.geo.ni, layer = "SA2011")
+oa.ni <- st_transform(oa.ni, 4326)  # EPSG code for WGS84
+oa.ni <- oa.ni %>% dplyr::select(SA2011, Hectares) %>% 
+  rename(id= SA2011,
+         area = Hectares)
 
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.sc)
+# Build OA for UK
+oa.uk <- rbind(oa.ew, oa.sc, oa.ni)
+rm(oa.ew, oa.sc, oa.ni)
 
-# build oa for GB
-oa.gb <- spRbind(oa.ew, oa.sc)
-rm(oa.ew, oa.sc)
-
-# oa.gb$geometry <- oa.gb$geometry %>%
-#   s2::s2_rebuild() %>%
-#   sf::st_as_sfc()
-
-# get GB
-path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
-gb <- readOGR(path.gb)
+# get UK
+path.uk <- "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/Countries_December_2023_Boundaries_UK_BUC/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson"
+uk <- st_read(path.uk)
 # spatial transformations
-gb <- spTransform(gb, CRS("+init=epsg:4326"))
-
+uk <- st_transform(uk, 4326)  # EPSG code for WGS84
 ```
@@
 ```{r}
-
-path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_oa") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_oa")  # for multiple PC
+path.out <- paste0(path, "/outputs/lisa/oa_ggplot/lisa_level_pc", n, "_oa") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa/oa_ggplot/lisa_level_pc", m, "_oa")  # for multiple PC
 
@@
 sf_use_s2(FALSE)
-nb <- poly2nb(oa.gb)
+nb <- poly2nb(oa.uk)
 listw <- nb2listw(nb, style = "W", zero.policy = T)
@@
 for (t in 1996:2012){
-test <- oa.gb
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
+  test <- oa.uk
+  test <- test %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
     mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
@@
   #LISA
-lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
-#lmoran
+  lmoran <- localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude)
 
-# centers the local Moran's around the mean
-lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
-lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
+  quadr <- attr(lmoran, "quadr")
+  #quadr[lmoran[,5] > 0.05 | is.na(lmoran[,5]), ] <- NA
+  quadr[p.adjust(lmoran[,5], method = "fdr") > 0.05 | is.na(lmoran[,5]), ] <- NA
 
-# centers the variable of interest around its mean
-
-lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
-lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
-
-signif <- 0.05
-#lmoran
-
-lmoran <- lmoran%>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
-  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
-  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%
-
-#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
-# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
-#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))
-
-mun_merge_new<- merge(test, lmoran, by="id")
-
-# R p value map
-breaks = c(0, 1, 2, 3, 4, 5)
+  palette_new <- grDevices::colorRampPalette(
+            colors = c("blue", "lightsalmon", "lightblue", "red", "white"))(5)
 
@@
 
-map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, 
-                                         palette = c("white","red","blue",
-                                                     rgb(0,0,1,alpha=0.4),
-                                                     rgb(1,0,0,alpha=0.4)), 
-                                         labels = c("Not significant",
-                                                    "High-High",
-                                                    "Low-Low",
-                                                    "Low-High",
-                                                    "High-Low"), 
-                                         title="LISA, websites per OA") + 
-  tm_legend(text.size = 1) +
-  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
-  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
-  #tm_borders(alpha=.1) +
-  tm_borders(lwd = 0) +
-  tm_shape(gb, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
-  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
-  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
-tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
+  ggplot(data = oa.uk, aes(fill = quadr$median)) +
+    geom_sf(data = uk, fill = "white", linewidth = .5) +
+    geom_sf(linewidth = 0) +
+    geom_sf(data = tif_sf, fill = "black", size = .3) +
+    #geom_sf_text(data = tif_sf, aes(label = name, fill = "black"), size = 4) +
+    scale_fill_manual(values = palette_new, na.value="white", drop = F) +
+    theme_minimal() +
+    theme(panel.grid.major = element_line(colour = "transparent"),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          legend.title=element_blank()) +
+          #legend.position="none") +
+    guides(fill = guide_legend(reverse = TRUE)) +
+    labs(title = t,
+        subtitle = "LISA, websites per OA")
+
+  filename = paste(path.out, t, ".png", sep = "")
+  ggsave(filename, bg='#ffffff')
 }
@@
 ```{r}
-path.out <- paste0(path, "/outputs/morani_oa.csv")
+path.out <- paste0(path, "/outputs/lisa/oa_ggplot/morani_oa.csv")
 morani %>% as_tibble() %>% 
@@
                    year = V3) %>% 
-  kable() #%>% 
-  #write_csv(path.out)
-
+  #kable() %>% 
+  write_csv(path.out)
 ```
-
-## Gini
-
-```{r}
-
-# add OAs with 0s
-df.with0s <- df %>% mutate(country = substring(oa11cd, 1, 1)) %>%
-  filter(country != "L" & country != "M" & country != "N") %>%
-  filter(!is.na(country)) %>%
-  complete(year, oa11cd, fill = list(n = 0)) %>%
-  filter(year!=1000,
-         !is.na(oa11cd)) %>%
-  dplyr::select(-country)
-
-gini.df <- data.frame()
-
-for (t in 1996:2012){
-    a <- df.with0s %>% filter(year==t)
-  b <- as.data.frame(gini(a$n))
-  b$year = t
-  gini.df <- rbind(gini.df, b)
-}
-
-path.out <- paste0(path, "/outputs/gini_oa.csv")
-gini.df %>% write_csv(path.out)
-```
-
-## Getis-Ord
-
-```{r, eval = F}
-
-t= 1998
-local_g <- localG(OA.Census$Qualification, nb_lw)
-local_g <- localG(df[df$year==t,]$n, listw)
-
-local_g <- cbind(OA.Census, as.matrix(local_g))
-names(local_g)[6] <- "gstat"
-
-```
-
-## Distances to cities and retail centres
-
-First set up the destinations. Origins are all the OA.
-
-Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
-Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)
-
-```{r}
-# oa centroids
-# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
-# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
-#   mutate(id = row.names())
-oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id")
-
-# cities
-city.names <- c("London, UK", 
-"Birmingham, UK",
-"Glasgow, UK",
-"Liverpool, UK",
-"Bristol, UK",
-"Manchester, UK",
-"Sheffield, UK",
-"Leeds, UK",
-"Edinburgh, UK",
-"Leicester, UK")
-
-cities <- geo(city.names, no_query = F, method = "arcgis")
-
-# retail centres
-geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
-retail <-readOGR(geo.path) 
-#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
-retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id") #%>%
-
-retail.major.cetres <- retail.major.cetres.help %>% 
-  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
-  st_transform(4326) %>%
-  st_coordinates() %>%
-  as_tibble() %>% 
-  bind_cols(retail.major.cetres.help$id) %>% 
-  rename(id = '...3') %>% 
-  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
-```
-
-Distance to cities
-
-```{r}
-dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist <- round((dist/1000),2) %>% 
-  as_tibble()
-
-city.names <- city.names %>% stringr::str_remove(", UK")
-names(dist) <- city.names
-
-dist <- dist %>% bind_cols(oa.gb.c$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
-# dist$distMet <- apply(dist[,2:11], 1, min)
-
-dist <- transform(dist, dist = do.call(pmin, dist[-1]))
-
-# NOT USED AS IT SCREWS UP THE REGRESSIONS
-# Complete panel, N = 3871903, 17 year * 227759 OA
-# df <- df %>% bind_rows(dist %>% dplyr::select(oa11cd)) %>%
-#   mutate(year = ifelse(is.na(year),1000, year),
-#          country = substring(oa11cd, 1, 1)) %>%
-#   filter(country != "L" & country != "M" & country != "N") %>%
-#   filter(!is.na(country)) %>%
-#   complete(year, oa11cd, fill = list(n = 0)) %>%
-#   filter(year!=1000,
-#          !is.na(oa11cd)) %>%
-#   dplyr::select(-country)
-
-# tests
-# test$country <- substring(test$oa11cd, 1, 1)
-# unique(test$country)
-# length(unique(test$oa11cd))
-
-# calculate area
-oa.gb$area <- area(oa.gb) #/ 1000000
-
-# Join with distance and area
-df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
-  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
-  relocate(area, .after = n) %>% 
-  mutate(area = area / 1000000,
-         density = n / area) %>% 
-  relocate(density, .after = area)
-sapply(df, function(x) sum(is.na(x)))
-```
-
-Distance to retail centres
-
-```{r}
-dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail <- round((dist.retail/1000),2) %>% 
-  as_tibble()
-
-# retail.names <- retail.major.cetres$RC_Name
-# names(dist.retail) <- retail.names
-
-dist.retail <- dist.retail %>% bind_cols(oa.gb.c$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# Minimum distance 
-dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
-dist.retail <- dist.retail %>% dplyr::select(oa11cd, dist.retail)
-
-# Join with complete panel
-df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd"))
-
-sapply(df, function(x) sum(is.na(x)))
-```
-
-Website density regressions to London
-
-```{r}
-dens.London <- data.frame()
-out.model <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(London), data = df[df$year==t,]) 
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.London <- rbind(r2, dens.London)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model <- rbind(out.model_, out.model)
-}
-
-names(dens.London)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_london_oa.png")
-ggplot(dens.London) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to London on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-out.model <- out.model %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_london_oa.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-#ggplot(out.model, aes(x=year, y = estimate)) +  
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to London on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-
-r2.all <- dens.London %>% 
-  rename(London = rsquared)
-
-# Help for OA area size. I double checked the largest: E00102963
-# It is correct, this is its size.
-# ggplot(df, aes(x=area)) +
-#     geom_histogram(binwidth=.5, colour="black", fill="white")
-# 
-# df %>% filter(area>50)
-```
-
-Website density regressions to the nearest city
-
-```{r}
-dens.distMet <- data.frame()
-out.modelMet <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.distMet <- rbind(r2, dens.distMet)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.modelMet <- rbind(out.model_, out.modelMet)
-}
-
-names(dens.distMet)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_city_oa.png")
-ggplot(dens.distMet) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest city on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-out.modelMet <- out.modelMet %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_city_oa.png")
-ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-r2.all <- r2.all %>% left_join(dens.distMet) %>% 
-  rename(cities = rsquared)
-```
-
-Website density regressions to the nearest retail centre
-
-```{r}
-dens.dist.retail <- data.frame()
-out.model.retail <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = df[df$year==t,]) 
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model.retail <- rbind(out.model_, out.model.retail)
-}
-
-names(dens.dist.retail)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_centre_oa.png")
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest retail centre on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-out.model.retail <- out.model.retail %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_centre_oa.png")
-ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest retail centre on website density, OA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-# library(pscl)
-# m1 <- zeroinfl((n) ~ log(dist.retail+0.001),
-#   data = df[df$year==t,], dist = "negbin", EM = TRUE)
-# summary(m1)
-# 
-# df %>% filter(year == 1996) %>% 
-#   group_by(group = cut(n, breaks = c(0, 1, 2, 10, 100, 1000, 10000), include.lowest = T, right = F)) %>%  #seq(0, max(n), 100))) %>%
-#   summarise(n = sum(n))
-```
-All R^2 dataframe
-
-```{r}
-path.out <- paste0(path, "/outputs/r2_all_oa.csv")
-
-r2.all <- dens.London %>% 
-  rename(London = rsquared) %>% 
-  left_join(dens.distMet) %>% 
-  rename(cities = rsquared) %>% 
-  left_join(dens.dist.retail) %>% 
-  rename(retail = rsquared) %>% 
-  relocate(year, .before = London) %>% 
-  arrange(year)
-
-write_csv(r2.all, path.out)
-```
-
-## The below are unfinished
-
-Website density regressions to the nearest retail centre with spatial lag
-
-```{r}
-library(spatialreg)
-
-model <- lagsarlm(log(density+0.1) ~ log(dist.retail+0.1), data = df[df$year==t,], 
-                  listw, tol.solve = 1e-30, zero.policy = T)
-
-dens.dist.retail <- data.frame()
-for (t in 1996:2012){
-  model <- lm(log(density) ~ log(dist.retail+0.1), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-}
-names(dens.dist.retail)[1] <- "rsquared"
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared))
-
-```
-
-## HH clusters
-
-```{r}
-hh <- mun_merge_new@data %>% filter(quadrant==1) %>% 
-  select(id, quadrant) %>% 
-  left_join(df[df$year==2000,], by =c("id" = "oa11cd")) %>% 
-  select(id, year, n, area, density, London, dist, dist.retail)
-
-hh.London <- lm(log(density+0.001) ~ log(London), data = hh)
-summary(hh.London)
-hh.dist <- lm(log(density+0.001) ~ log(dist), data = hh)
-summary(hh.dist)
-hh.retail <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = hh)
-summary(hh.retail)
-
-data.folder <- "/hdd/internet_archive/archive/data/"
-data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
-#Created by domain.R, which uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df9610 <- read_csv(data.path9610) %>% 
-  filter(V1.domain == n,
-         year == 1999) %>% 
-  left_join(lookup, by = "pc", suffix =c("",""))
-
-hh %>% left_join(df9610, by =c("id"="oa11cd"))
-
-```
-
-Spatial regression
-
-```{r}
-library(splm)
-m <- spml(log(n)~lag(log(n)), data = df, listw = listw, lag = TRUE, spatial.error = "b", 
-          model = "within", effect = "individual", method = "eigen", na.action = na.fail)
-
-```

### 2024-06-11 src/problem_years.Rmd
ggplot

--- 2023-10-27 src/level_lisa_la.Rmd
@@
 ---
-title: "diffusion, websites with 1 pc"
-date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
-output: 
-  html_document:
-    df_print: paged
-    toc: true
-    toc_float: true
-knit: (function(inputFile, encoding) {
-    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
-  })
+title: "problem years"
+author: "Emmanouil Tranos"
+date: '2024-05-17'
+output: html_document
 ---
@@
 ```{r}
-n = 1 #number of unique postcodes. 
-m = 12
-
-data.folder <- "/hdd/internet_archive/archive/data/"
-data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
-#Created by domain.R, which uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df9610 <- read_csv(data.path9610)
 
@@
 df <- df %>% filter(!is.na(ladcd)) %>% 
+  complete(ladcd, year) %>% 
+  replace(is.na(.), 0)
+
+# tests for Scotland
+# df %>% filter(substr(oa11cd, 1,1) =="S",
+#               year==2010) %>% summarise(mean(n, na.rm = T))
+# 
+# df %>% filter(oa11cd=="S00090182")
+# 
+# test@data %>% filter(substr(id, 1,1) =="S",
+#                      n!=0)
+# 
+# df %>% filter(substr(oa11cd, 1,1) =="W",
+#               year==2003)
+```
+
+### Problem postcodes
+
+```{r}
+n = 1 #number of unique postcodes. 
+m = 12
+
+data.folder <- "/hdd/internet_archive/archive/data/"
+data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
+#Created by domain.R, which uses domain instead of host.
+#This is what we use for the hyperlinks paper as per George's script
+
+df9610 <- read_csv(data.path9610) %>% 
+  filter(
+      #V1.domain < m,             # for multiple pc
+      V1.domain == n,             # for n == 1
+      year > 1995)
+
+# df9610 %>% filter(V1.domain == 1) %>% 
+#   left_join(lookup, by = "pc", suffix =c("","")) %>% 
+#   filter(ladcd == "E08000006" & year == 2006) %>% 
+#   group_by(pc) %>% 
+#   summarise(n = n()) %>% 
+#   arrange(desc(n))
+# 
+# df9610 %>% filter(pc == "M28 2SL" & year == 2006 & V1.domain == 1)
+
+problem.pcs <- df9610 %>% filter(year > 1999 & year < 2011) %>% 
+  group_by(pc, year) %>% 
+  summarise(n = n()) %>%
+  ungroup() %>% 
+  # complete(pc, year) %>% 
+  # replace(is.na(n), 0) %>% 
+  arrange(desc(n)) %>% 
+  pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
+  dplyr::select(pc, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010) %>% 
+  slice_head(n=10) %>% 
+  dplyr::select(pc)
+
+# [1] "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"  "EC1V 4PY" "DE21 7BF" "E14 9SR"  "GL16 7YA" "WC1N 3XX"
+# [10] "CB2 1AB" 
+```
+
+### Correct 1996-2010
+
+The above exploratory analysis indicated that there is a huge increase in 2004-2006
+in the above postcodes. I turn them to NAs and then use a regression to impute these
+gaps.
+
+```{r}
+df9610 <- df9610 %>% 
+  group_by(pc, year) %>% 
+  summarise(n = n()) %>% 
+  ungroup() %>% 
+  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2003 & year <2007, NA, n))
+
+df9610 <- pdata.frame(df9610)
+predictions <- round(predict(plm(n ~ as.factor(year), effect = "individual", 
+                                 model = "within", data = df9610), newdata = df9610), 2)
+
+df9610$n[is.na(df9610$n)] <- predictions[is.na(df9610$n)]
+df9610 %>% filter(pc == "M28 2SL") %>% distinct(year)
+
+df9610 <- as_tibble(df9610) %>% 
+  mutate(year = as.double(as.character(year))) %>% 
+  left_join(lookup, by = "pc", suffix =c("","")) %>%
+  group_by(year, ladcd) %>%
+  summarise(n = n()) %>%            
+  ungroup()
+```
+
+### Load remaining data
+
+```{r}
+data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
+#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
+#This is what we use for the hyperlinks paper as per George's script
+
+df1112 <- read_csv(data.path2011_2012) %>% 
+  filter(
+    #V1.domain < m,             # for multiple pc
+    V1.domain == n,             # for n == 1
+    year > 1995 & year <2013) %>%   
+  left_join(lookup, by = "pc", suffix =c("","")) %>% 
+  group_by(year, ladcd) %>%
+  summarise(n = n()) %>%             # for n == 1  
+  #summarise(n = sum(V1.domain)) %>% # for multiple pc  
+  ungroup()
+
+df <- bind_rows(df9610, df1112) %>% 
+  filter(!is.na(ladcd)) %>% 
   complete(ladcd, year) %>% 
@@
 
-path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", n, "_la") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_la")  # for multiple PC
+path.out <- paste0(path, "/outputs/lisa/corrected/lisa_level_pc", n, "_la") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa/corrected/lisa_level_pc", m, "_la")  # for multiple PC
 
@@
 test <- la
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
+test <- test %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
   mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
@@
 #LISA
-lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
-#lmoran
+#lmoran <- cbind(test, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
+lmoran <- localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude)
 
+quadr <- attr(lmoran, "quadr")
+quadr[lmoran[,5] > 0.05 | is.na(lmoran[,5]), ] <- NA
+
+palette_new <- grDevices::colorRampPalette(
+          colors = c("blue", "lightsalmon", "lightblue", "red", "white"))(5)
+
+ggplot(data = la, aes(fill = quadr$median)) +
+  geom_sf() +
+  # scale_fill_discrete_c4a_cat(name = "%", palette = "parks.charmonix",
+  #                             na.translate = FALSE) +
+  scale_fill_manual(values = palette_new, na.value="white", breaks = NULL, drop = F) +
+  theme_minimal() +
+  theme(panel.grid.major = element_line(colour = "transparent"),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks = element_blank()) +
+  #guides(fill = guide_legend(reverse = TRUE)) +
+  labs(title = "LISA, websites per firm in a LA") + 
+  theme(legend.title=element_blank())
+
+theme(axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks = element_blank(),
+        rect = element_blank())
+
+filename = paste(path.out, t, ".png", sep = "")
+ggsave(filename)
+}
+```
+
+# Moran I
+
+```{r}
+path.out <- paste0(path, "/outputs/lisa/corrected/morani_la.csv")
+morani %>% as_tibble() %>% 
+  rename('morani'=V1,
+                   p = V2,
+                   year = V3) %>% 
+  #kable() %>% 
+  write_csv(path.out)
+```
+
+```{r}
+path.images <- paste0(path, "/outputs/lisa/corrected/")
+
+list.files(path = path.images, pattern = "^lisa_level_pc1_la", full.names = T) %>% 
+  map(image_read) %>% # reads each path file
+  image_join() %>% # joins image
+  image_animate(delay=50) %>% # animates, can opt for number of loops
+  image_write("corrected_lisa.gif", format = "png") # write to current dir
+
+```
+
+```{r}
 # centers the local Moran's around the mean
@@
   #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
-  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%
+  mutate(quadrant= ifelse(lmoran$`Pr.z....E.Ii..` > signif, 0, quadrant)) #%>%
 
@@
 mun_merge_new<- merge(test, lmoran, by="LAD21CD")
+mun_merge_new<- test %>% 
+  left_join(lmoran, by="LAD21CD")
 
@@
 ```
-
-## lisa maps parallel
-
-```{r, eval=F}
-
-path.out <- paste0(path, "/outputs/lisa/lisa_level_pc_test", n, "_la") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa/lisa_level_pc", m, "_la")  # for multiple PC
-
-#setup cluster
-#https://www.blasbenito.com/post/02_parallelizing_loops_with_r/
-n.cores <- parallel::detectCores() - 1
-my.cluster <- parallel::makeCluster(
-  n.cores, 
-  type = "PSOCK"
-  )
-print(my.cluster)
-doParallel::registerDoParallel(cl = my.cluster)
-foreach::getDoParWorkers()
-
-morani <- data.frame()
-
-#spatial weight matrix
-sf_use_s2(FALSE)
-nb <- poly2nb(la)
-listw <- nb2listw(nb, style = "W", zero.policy = T)
-
-#for (t in 1996:2012){
-foreach(t = 1996:2012, .packages = c("tidyverse",
-                                     "rgdal",
-                                     "rgeos",
-                                     "spdep",
-                                     "tmap",
-                                     "maptools",
-                                     "knitr",
-                                     "broom")) %dopar% {
-test <- la
-test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
-  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
-         year = ifelse(is.na(year), t, year))
-
-#spatial weight matrix
-#moved out of the loop
-
-#Moran's I
-globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
-a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
-morani <- rbind(a, morani)
-
-#LISA
-lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
-#lmoran
-
-# centers the local Moran's around the mean
-lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
-lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
-
-# centers the variable of interest around its mean
-
-lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
-lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
-
-signif <- 0.05
-#lmoran
-
-lmoran <- lmoran%>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
-  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
-  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%
-
-#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
-# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
-#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))
-
-mun_merge_new<- merge(test, lmoran, by="LAD21CD")
-
-# R p value map
-breaks = c(0, 1, 2, 3, 4, 5)
-
-cities <- maps::world.cities %>% 
-  filter(country.etc=="UK") %>% 
-  arrange(pop) %>% tail(10) 
-tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
-
-map <- tm_shape(mun_merge_new, projection = "+init=epsg:4326") + tm_fill(col = "quadrant", breaks = breaks, 
-                                         palette = c("white","red","blue",
-                                                     rgb(0,0,1,alpha=0.4),
-                                                     rgb(1,0,0,alpha=0.4)), 
-                                         labels = c("Not significant",
-                                                    "High-High",
-                                                    "Low-Low",
-                                                    "Low-High",
-                                                    "High-Low"), 
-                                         title="LISA, websites per firm in a LA") +
-  tm_legend(text.size = 1) +
-  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
-  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
-  #tm_borders(alpha=.1) +
-  tm_borders(lwd = 0) +
-  tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
-  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
-  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ") +
-tmap_options(check.and.fix = TRUE)
-tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
-                                     }
-
-```
-
-## Moran I
-
-```{r}
-path.out <- paste0(path, "/outputs/lisa/morani_la.csv")
-morani %>% as_tibble() %>% 
-  rename('morani'=V1,
-                   p = V2,
-                   year = V3) %>% 
-  #kable() %>% 
-  write_csv(path.out)
-```
-
-## Gini
-
-```{r}
-
-# No need to add 0s like OAs as there aren't any LA with 0s
-
-gini.df <- data.frame()
-
-for (t in 1996:2012){
-    a <- df %>% filter(year==t)
-  b <- as.data.frame(gini(a$n))
-  b$year = t
-  gini.df <- rbind(gini.df, b)
-}
-
-path.out <- paste0(path, "/outputs/gini_la.csv")
-gini.df %>% write_csv(path.out)
-```
-
-## Getis-Ord
-
-```{r eval = F}
-
-t= 1998
-local_g <- localG(OA.Census$Qualification, nb_lw)
-local_g <- localG(df[df$year==t,]$n, listw)
-
-local_g <- cbind(OA.Census, as.matrix(local_g))
-names(local_g)[6] <- "gstat"
-
-```
-
-## Distances to cities and retail centres
-
-First set up the destinations. Origins are all the LA.
-
-Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
-Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)
-
-```{r}
-# LA centroids
-la.c <- gCentroid(la, byid=TRUE)@coords %>% 
-  as.data.frame() %>%
-  bind_cols(la@data) %>% 
-  dplyr::select(LAD21CD, x, y) %>% 
-  remove_rownames()
-
-# cities
-city.names <- c("London, UK", 
-"Birmingham, UK",
-"Glasgow, UK",
-"Liverpool, UK",
-"Bristol, UK",
-"Manchester, UK",
-"Sheffield, UK",
-"Leeds, UK",
-"Edinburgh, UK",
-"Leicester, UK")
-
-cities <- geo(city.names, no_query = F, method = "arcgis")
-
-# retail centres
-geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
-retail <-readOGR(geo.path) 
-#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
-retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id") #%>%
-
-retail.major.cetres <- retail.major.cetres.help %>% 
-  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
-  st_transform(4326) %>%
-  st_coordinates() %>%
-  as_tibble() %>% 
-  bind_cols(retail.major.cetres.help$id) %>% 
-  rename(id = '...3') %>% 
-  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
-```
-
-Distance to cities
-
-```{r}
-dist <- distm(cbind(la.c$x, la.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist <- round((dist/1000),2) %>% 
-  as_tibble()
-
-city.names <- city.names %>% stringr::str_remove(", UK")
-names(dist) <- city.names
-
-dist <- dist %>% bind_cols(la.c$LAD21CD) %>% 
-  rename(ladcd = last_col()) %>% 
-  relocate(ladcd)
-
-# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
-# dist$distMet <- apply(dist[,2:11], 1, min)
-
-dist <- transform(dist, dist = do.call(pmin, dist[-1]))
-
-# NOT USED AS IT SCREWS UP THE REGRESSIONS
-# Complete panel, N = 3871903, 17 year * 227759 OA
-# df <- df %>% bind_rows(dist %>% select(oa11cd)) %>% 
-#   mutate(year = ifelse(is.na(year),1000, year),
-#          country = substring(oa11cd, 1, 1)) %>% 
-#   filter(country != "L" & country != "M" & country != "N") %>% 
-#   filter(!is.na(country)) %>% 
-#   complete(year, oa11cd, fill = list(n = 0)) %>% 
-#   filter(year!=1000,
-#          !is.na(oa11cd)) %>% 
-#   select(-country)
-
-# tests
-# test$country <- substring(test$oa11cd, 1, 1)
-# unique(test$country)
-# length(unique(test$oa11cd))
-
-# calculate area
-la$SHAPE_Area <- area(la) #/ 1000000
-
-# Join with distance and area
-df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
-  left_join(la@data, by = c("ladcd" = "LAD21CD")) %>% 
-  rename(area = SHAPE_Area) %>% 
-  relocate(area, .after = n) %>% 
-  mutate(area = area / 1000000,
-         density = n / area) %>% 
-  relocate(density, .after = area)
-sapply(df, function(x) sum(is.na(x)))
-```
-
-Distance to retail centres
-
-```{r}
-dist.retail <- distm(cbind(la.c$x, la.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail <- round((dist.retail/1000),2) %>% 
-  as_tibble()
-
-# retail.names <- retail.major.cetres$RC_Name
-# names(dist.retail) <- retail.names
-
-dist.retail <- dist.retail %>% bind_cols(la.c$LAD21CD) %>% 
-  rename(LAD21CD = last_col()) %>% 
-  relocate(LAD21CD)
-
-# Minimum distance 
-dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
-dist.retail <- dist.retail %>% dplyr::select(LAD21CD, dist.retail)
-
-# Join with complete panel
-df <- df %>% left_join(dist.retail, by = c("ladcd" = "LAD21CD"))
-
-sapply(df, function(x) sum(is.na(x)))
-```
-
-## Panel regressions
-
-```{r}
-# London
-model <- plm(log(n+0.001) ~ log(London+.001)*as.factor(year), 
-             data = df,
-             index = c("ladcd", "year"),
-             model = "within",
-             effect = "twoways")
-summary(model)
-
-est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
-int <- confint(est)
-
-out.model <- bind_cols(tidy(est), 
-                       as_tibble(int)) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(year = 1997:2012) %>% 
-  rename(conf.low = '2.5 %',
-         conf.high = '97.5 %')
-
-path.out <- paste0(path,"/outputs/lisa/b_panel_dist_london_la.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                    width = 0.2,size  = 1,
-                    position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
-  ggtitle("Regression coefficients: distance to London on website counts, LA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-
-# Cities
-model <- plm(log(n+0.001) ~ log(dist+.001)*as.factor(year), 
-             data = df,
-             index = c("ladcd", "year"),
-             model = "within",
-             effect = "twoways")
-summary(model)
-
-est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
-int <- confint(est)
-
-out.model <- bind_cols(tidy(est), 
-                       as_tibble(int)) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(year = 1997:2012) %>% 
-  rename(conf.low = '2.5 %',
-         conf.high = '97.5 %')
-
-path.out <- paste0(path,"/outputs/lisa/b_panel_dist_cities_la.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                    width = 0.2,size  = 1,
-                    position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
-  ggtitle("Regression coefficients: distance to Cities on website counts, LA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-
-# Retail centres
-model <- plm(log(n+0.001) ~ log(dist.retail+.001)*as.factor(year), 
-             data = df,
-             index = c("ladcd", "year"),
-             model = "within",
-             effect = "twoways")
-summary(model)
-
-est <- coeftest(model, vcov = vcovHC, type = "HC1") # same
-int <- confint(est)
-
-out.model <- bind_cols(tidy(est), 
-                       as_tibble(int)) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(year = 1997:2012) %>% 
-  rename(conf.low = '2.5 %',
-         conf.high = '97.5 %')
-
-path.out <- paste0(path,"/outputs/lisa/b_panel_dist_retail_la.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                    width = 0.2,size  = 1,
-                    position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("Interaction coefficients: distance*year, robust st. errors") + xlab("") + 
-  ggtitle("Regression coefficients: distance to Retail Centres on website counts, LA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-```
-
-## Website density regressions to London
-
-```{r}
-dens.London <- data.frame()
-out.model <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(n+0.001) ~ log(London), data = df[df$year==t,]) #density
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.London <- rbind(r2, dens.London)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model <- rbind(out.model_, out.model)
-}
-
-names(dens.London)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_london_la.png")
-ggplot(dens.London) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to London on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5))
-ggsave(path.out)
-
-out.model <- out.model %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_london_la.png")
-ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
-#ggplot(out.model, aes(x=year, y = estimate)) +  
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to London on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-```
-
-## Website density regressions to the nearest city
-
-```{r}
-dens.distMet <- data.frame()
-out.modelMet <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(density+0.001) ~ log(dist), data = df[df$year==t,])
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.distMet <- rbind(r2, dens.distMet)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.modelMet <- rbind(out.model_, out.modelMet)
-}
-
-names(dens.distMet)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_city_la.png")
-ggplot(dens.distMet) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest city on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-out.modelMet <- out.modelMet %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_city_la.png")
-ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest city on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-```
-
-## Website density regressions to the nearest retail centre
-
-```{r}
-dens.dist.retail <- data.frame()
-out.model.retail <- data.frame()
-
-for (t in 1996:2012){
-  model <- lm(log(n+0.001) ~ log(dist.retail+0.001), data = df[df$year==t,]) #density
-  r2 <- as.data.frame(summary(model)$r.squared)
-  r2$year = t
-  dens.dist.retail <- rbind(r2, dens.dist.retail)
-
-  out.model_ <- tidy(model, conf.int = TRUE) %>% 
-    slice(2) %>% 
-    mutate_if(is.numeric, round, 2)
-  out.model.retail <- rbind(out.model_, out.model.retail)
-}
-
-names(dens.dist.retail)[1] <- "rsquared"
-
-path.out <- paste0(path,"/outputs/rsq_dist_centre_la.png")
-ggplot(dens.dist.retail) +
-  geom_line(aes(x=year, y = rsquared)) +
-  ylab("Rsquared") + 
-  ggtitle("Regression fit, distance to nearest retail centre on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-out.model.retail <- out.model.retail %>% 
-  mutate(year = 1996:2012)
-
-path.out <- paste0(path,"/outputs/b_dist_centre_la.png")
-ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
-           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
-                       width = 0.2,size  = 1,
-                       position = "dodge", color="turquoise4") +
-  #geom_hline(yintercept = 0, color = "red", size = 1) +
-  geom_point() + coord_flip() +
-  ylab("distance coefficient") + xlab("") + 
-  ggtitle("Regression coefficients, distance to nearest retail centre on website density, LA") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-ggsave(path.out)
-
-# library(pscl)
-# m1 <- zeroinfl((n) ~ log(dist.retail+0.001),
-#   data = df[df$year==t,], dist = "negbin", EM = TRUE)
-# summary(m1)
-# 
-# df %>% filter(year == 1996) %>% 
-#   group_by(group = cut(n, breaks = c(0, 1, 2, 10, 100, 1000, 10000), include.lowest = T, right = F)) %>%  #seq(0, max(n), 100))) %>%
-#   summarise(n = sum(n))
-```
-
-All R^2 dataframe
-
-```{r}
-path.out <- paste0(path, "/outputs/r2_all_la.csv")
-
-r2.all <- dens.London %>% 
-  rename(London.LA = rsquared) %>% 
-  left_join(dens.distMet) %>% 
-  rename(cities.LA = rsquared) %>% 
-  left_join(dens.dist.retail) %>% 
-  rename(retail.LA = rsquared) %>% 
-  relocate(year, .before = London.LA) %>% 
-  arrange(year)
-
-write_csv(r2.all, path.out)
-```

### 2024-06-11 abstracts_notes/abstract_UiS_2024/Tranos_2024.tex
pre-ggplot

--- 2023-12-14 abstracts_notes/abstract_GfR_2024/Tranos_GfR_2024.tex
@@
 
-\title{Digital economy in the UK: a multi-scalar story of the diffusion of web technologies}
+\title{A multi-scale story of the diffusion of a new technology: the web}
 \maketitle
@@
 Importantly, the data and geolocation strategy allow to capture these processes at small spatial scales.
-This level of granularity differentiate this paper with previous approaches in the literature, which were only able to capture technological adoption at more coarse level.
+This level of granularity differentiates this paper from previous approaches in the literature, which were only able to capture technological adoption at more coarse level.
 Thus, this paper tests how well established theoretical approaches regarding the diffusion of new technologies are still applicable when the focus is on local scales.
@@
 This modelling exercise highlights the importance of the key spatial predictors as described in the literature in predicting technological diffusion.
-Then, these model are trained for each UK region separately and tested iteratively to all other regions in order obtain an understanding of how differently these spatial drivers affect the diffusion of web technologies across different places (Figures \ref{fig:la} and \ref{fig:oa}).
+Then, these model are trained for each UK region separately and tested iteratively to all other regions in order obtain an understanding of how differently these spatial drivers affect the diffusion of web technologies across different places.
 In other words, this modelling strategy reveals how different or not UK regions have been in terms of the adoption of web technologies across scales.
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.9\textwidth]{heatmap_LA}
-    \caption{Regional similarities based on cross-region training/predictions at the Local Authority level}
-    \label{fig:la}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.75\textwidth]{heatmap_OA}
-    \caption{Regional similarities based on cross-region training/predictions at the Local Authority level}
-    \label{fig:oa}
-\end{figure}
-
 %\newpage

### 2024-06-11 paper/submission1/tranos2023.qmd
mac
pre-ggplot

@@
 ---
-title: "Digital economy in the UK: a multi-scalar story of the diffusion of web technologies"
+title: "A multi-scale story of the diffusion of a new technology: the web"
 author:
@@
 
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, message = FALSE) # By default, hide code; set to TRUE to see code
+#knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
+knitr::opts_chunk$set(out.width = '100%', dpi=300) # Figure resolution and size
+knitr::opts_chunk$set(fig.env="figure") # Latex figure environment
+
+#insert libraries here
+library(rprojroot)
+library(tidyverse)
+library(knitr)
+
+# This is the project path
+path <- find_rstudio_root_file()
+```
+
 # Introduction {#sec1}
@@
 <!-- geographers lack of interest + motivation -->
+
 Geographers were always interested in how new technologies and
@@
 geographical studies exploring the diffusion of new and, more
-specifically for this paper, digital technologies across *both* space 
+specifically for this paper, digital technologies across both space
 and time can be attributed to the scarcity of relevant and granular
@@
 <!-- contribution -->
+
 This paper offers such a contribution: a geographical study illustrating
-how a new technology that is the web diffused over space and time 
-in the UK at a high level of spatial granularity during the $1996-2012$ 
-period. It does so by employing a novel source of big data which 
-captures the active engagement with web technologies during that period.
-By addressing this empirical question this paper exemplifies how the 
+how a new technology that is the web diffused over space and time in the
+UK at a high level of spatial granularity during the $1996-2012$ period.
+It does so by employing a novel source of big data which captures the
+active engagement with web technologies during that period. By
+addressing this empirical question this paper exemplifies how the
 combination of data sources which escape the traditional social science
-domain and adequate research methods can offer new lenses to geographical
-research regarding the understanding of technological diffusion.
+domain and adequate research methods can offer new lenses to
+geographical research regarding the understanding of technological
+diffusion.
+
+**SAY SOMETHING ABOUT SPATIAL ANALYTICS, DIST PAPER BY ARP, NOT SPECIFIC EXPLANATORY VARIABLES, INTANGIBLE**
 
 <!-- motivation -->
+
 The motivation for this paper lies in the fact that there are various
@@
 diffused over space and time and use this knowledge to make predictions
-regarding the diffusion of related *future* technologies.
-As per @leibowicz2016representing, historical studies agree on the fact that 
+regarding the diffusion of related *future* technologies. As per
+@leibowicz2016representing, historical studies agree on the fact that
 "technologies diffuse at different times, at different rates, and to
 different extents in different places, and can be significantly
-influenced by policies" [@victor1993]. @meade2021modelling highlight that 
-a variety of actors have a direct interest in gaining such knowledge 
-including network equipment suppliers; network operators, regulatory 
-and local authorities. 
-These processes and their effects vary a lot across scales: although
-the diffusion of a new technology might not be optimal at a local level,
-it might be beneficial from a global perspective as it could lead to 
-faster diffusion to less advantaged places [@leibowicz2016representing]. 
-Despite the spatial heterogeneity of such diffusion mechanisms and 
-the policy relevance, there are very limited attempts in the literature 
-to analyse the diffusion of new digital technologies at a detailed 
-geographical level.
+influenced by policies" [@victor1993]. @meade2021modelling highlight
+that a variety of actors have a direct interest in gaining such
+knowledge including network equipment suppliers; network operators,
+regulatory and local authorities. These processes and their effects vary
+a lot across scales: although the diffusion of a new technology might
+not be optimal at a local level, it might be beneficial from a global
+perspective as it could lead to faster diffusion to less advantaged
+places [@leibowicz2016representing]. Despite the spatial heterogeneity
+of such diffusion mechanisms and the policy relevance, there are very
+limited attempts in the literature to analyse the diffusion of new
+digital technologies at a detailed geographical level.
 
 <!-- diffusion and adoption -->
+
 Technological diffusion, which is by definition an aggregated process,
-can be discussed in parallel with individual adoption mechanisms.
-On the one hand, @rogers2010diffusion identifies early adopter of new 
-technologies as 'knowledgeable risk takers' and @griliches1957 as 'profit 
-maximisers' [@ding2010modeling]. Such individual agents are rewarded because 
-of their attitude towards new technologies and innovations. On the other hand,
-@perkins2011internet attribute diffusion to two processes: (i) epidemic-like 
-mechanisms, which are governed by distance, proximity and social interactions, 
-and (ii) by economic mechanisms as new innovations are adopted by users as they
-become more profitable, valueable and useful.
+can be discussed in parallel with individual adoption mechanisms. On the
+one hand, @rogers2010diffusion identifies early adopter of new
+technologies as 'knowledgeable risk takers' and @griliches1957 as
+'profit maximisers' [@ding2010modeling]. Such individual agents are
+rewarded because of their attitude towards new technologies and
+innovations. On the other hand, @perkins2011internet attribute diffusion
+to two processes: (i) epidemic-like mechanisms, which are governed by
+distance, proximity and social interactions, and (ii) by economic
+mechanisms as new innovations are adopted by users as they become more
+profitable, valueable and useful.
 <!-- The criticisms are mainly based on the fact that although the approach -->
@@
 <!-- data and early engagement with the web -->
-This paper focuses on the diffusion of the web as new technology during the $1996-2012$
-period. This was an exciting period for digital technologies as it corresponds with
-the commercialisation of the internet and, consequently, its almost universal 
-adoption. The reader is reminded that it was only in 1994 when Netscape Navigator 
-was introduced, a year before Microsoft's Internet Explorer
-^[[https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft](https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft)].
-Also, only $9$ per cent of UK's households had access to the internet in $1998$ 
-[@ons2018], the web included mostly static webpages, there were no social media 
-and web browsing was happening exclusively from desktop PCs as there were no 
-smartphones [@tranosuk]. 
-Hence, it is fair to say that the study period captures the very early stages of 
-the diffusion on a new technology that is the web. 
-This is key point in the lifecycle of a new technology. Firstly, during this period
-new technologies are expensive, crude and imperfect [@rosenberg1994exploring; @wilson201281].
-A simple comparison between Web 1.0 and Web 3.0 applications clearly illustrates this
-[@tranos2020social]. During this period the performance of a new technology
-is the main attraction and not the cost to access and use it [@wilson2011lessons]
-There is a broader theoretical discussion in the literature regarding the motivation
-behind early adoption. As summarised by @perkins2005international, on the one hand,
-epidemic models highlight the role of interpersonal contacts as a way for new 
-technologies to diffuse. On the other hand, economic models underline the importance
-of heterogeneity. Different firms have different structures and business plans,
-which define the potential economic returns of the adoption of a new technology 
-and, therefore, the choice to adopt a new technology becomes an individual option.
-From a broader and evolutionary perspective, initial conditions are essential for 
-the creation and evolution of path-dependent technological development trajectories
-[@neffke2011regions; @simmie2014new]. This argument is even more relevant when the
-focus is on Information and Communication Technologies (ICT) because of the commonly
-found lag between investment and economic returns as reflected in the Solow paradox 
-[@acemoglu2014return; @brynjolfsson2018artificial].
 
-Importantly, the data used here depicts the *active* engagement with the web in 
-the UK as it contains all the commercial websites archived by the Internet Archive
-^[[https://archive.org/](See https://archive.org/).]
-that include a mention to at least one valid UK postcode in the web text. 
-Websites are identified as commercial ones if they are part of the .co.uk second 
-level domain (SLD). Also, the act of of creating a website is understood here as 
-active engagement with the web vis-à-vis the more passive act of browsing the web 
-or having an internet connection [@tranosuk]. Previous studies have focused mostly
-on more passive notions of engaging with digital technologies such as internet 
-adoption and internet speeds [e.g. @blank2018local; @riddlesden2014broadband]. 
-More details about the data and the data generation process can be found in Section
-[XX](#sec3). **ADD DATA DESCRIPTION, WHY .CO.UK**
+This paper focuses on the diffusion of the web as new technology during
+the $1996-2012$ period. This was an exciting period for digital
+technologies as it corresponds with the commercialisation of the
+internet and, consequently, its almost universal adoption. The reader is
+reminded that it was only in 1994 when Netscape Navigator was
+introduced, a year before Microsoft's Internet Explorer.[^1] Also, only
+$9$ per cent of UK's households had access to the internet in $1998$
+[@ons2018], the web included mostly static webpages, there were no
+social media and web browsing was happening exclusively from desktop PCs
+as there were no smartphones [@tranosuk]. Hence, it is fair to say that
+the study period captures the very early stages of the diffusion on a
+new technology that is the web. This is key point in the lifecycle of a
+new technology. Firstly, during this period new technologies are
+expensive, crude and imperfect [@rosenberg1994exploring; @wilson201281].
+A simple comparison between Web 1.0 and Web 3.0 applications clearly
+illustrates this [@tranos2020social]. During this period the performance
+of a new technology is the main attraction and not the cost to access
+and use it [@wilson2011lessons] There is a broader theoretical
+discussion in the literature regarding the motivation behind early
+adoption. As summarised by @perkins2005international, on the one hand,
+epidemic models highlight the role of interpersonal contacts as a way
+for new technologies to diffuse. On the other hand, economic models
+underline the importance of heterogeneity. Different firms have
+different structures and business plans, which define the potential
+economic returns of the adoption of a new technology and, therefore, the
+choice to adopt a new technology becomes an individual option. From a
+broader and evolutionary perspective, initial conditions are essential
+for the creation and evolution of path-dependent technological
+development trajectories [@neffke2011regions; @simmie2014new]. This
+argument is even more relevant when the focus is on digital technologies because 
+of the commonly found lag between investment and economic returns as reflected in 
+the Solow paradox [@acemoglu2014return; @brynjolfsson2018artificial].
+
+[^1]: <https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft>
+
+Importantly, the data used here depicts the *active* engagement with the
+web in the UK as it contains all the commercial websites archived by the
+Internet Archive [^2] that include a mention to at least one valid UK
+postcode in the web text. Websites are identified as commercial ones if
+they are part of the .co.uk second level domain (SLD). Also, the act of
+of creating a website is understood here as active engagement with the
+web vis-à-vis the more passive act of browsing the web or having an
+internet connection [@tranosuk]. Previous studies have focused mostly on
+more passive notions of engaging with digital technologies such as
+internet adoption and internet speeds [e.g. @blank2018local;
+@riddlesden2014broadband]. More details about the data and the data
+generation process can be found in Section [XX](#sec3).
 
@@
 
-<!-- methods -->
 @grubler1990rise Later Hagerstrand conceptualized physical "barrier"
@@
 <!-- key concepts -- copied from help.Rmd -->
-Geographical diffusion is a synthesis of different processes. On the one hand, we 
-can identify purely spatial or, in other words, contagious processes. Adjacency 
-and, more broadly, distance are the key drivers of diffusion. 
-This perspective draws similarities with epidemics: innovation just like pathogens 
-spreads because of contagion and, consequently, proximity and exposure 
-[@hivner2003facilitating]. On the other hand, we can identify hierarchical 
-processes. Instead of horizontal distance-based diffusion mechanisms, the top-down 
-hierarchy of urban systems drives diffusion. In reality, the synthesis of these 
-two processes represents how new technologies diffuse over space and time [@morrill2020spatial].
 
-These ideas were firstly introduced by Torsten Hägerstrand and his thesis entitled 
-'Innovation Diffusion as a Spatial Process' [@hagerstrand1968innovation]. 
-Hägerstrand was the first one to identify diffusion as a geographical process. 
-The starting point was the idea that diffusion is based on passing information 
-through social networks, which themselves tend to be defined by geography. Hence, 
-he identified the 'neighbourhood' effect of how information, and consequently, 
-innovation diffuse. He used agricultural innovations to test and model his ideas 
-using Monte Carlo simulations. Hägerstrand also incorporated the role of hierarchy 
-and how some phenomena maybe firstly adopted in larger cities and then diffuse 
-to second tier ones. This is a sequential instead of a simultaneous process, which 
-resembles the 'lead-lag' spatial acceleration effect in market research 
-[@bento2018time; @PERES201091]. Hägerstrand is more widely known though for highlighting 
-the role time plays in the diffusion of innovations: an early-pioneering period, 
-a middle fast accelerating period and a final saturation one [@morrill2020spatial.
+Geographical diffusion is a synthesis of different processes. On the one
+hand, we can identify purely spatial or, in other words, contagious
+processes. Adjacency and, more broadly, distance are the key drivers of
+diffusion. This perspective draws similarities with epidemics:
+innovation just like pathogens spreads because of contagion and,
+consequently, proximity and exposure [@hivner2003facilitating]. On the
+other hand, we can identify hierarchical processes. Instead of
+horizontal distance-based diffusion mechanisms, the top-down hierarchy
+of urban systems drives diffusion. In reality, the synthesis of these
+two processes represents how new technologies diffuse over space and
+time [@morrill2020spatial].
 
-The temporal dimension was further explored by Everett Rogers and his seminal work 
-on 'Diffusion of Innovations' [@rogers2010diffusion]. Rogers being a sociologist, 
-his work focused not on the diffusion of innovations over space and time, but 
-instead on the adoption of new technologies and innovations by individuals and the 
-individual mechanisms that drive the decisions behind adoption. He identified five 
-groups of individuals regarding their adoption speed: innovators, early adopters, 
-early majority, late majority and laggards. Key mechanism of diffusion and adoption 
-is always communication and how knowledge is transferred within a social system. 
-Therefore, all approaches agree on the S-shaped diffusion and cumulative adoption 
-pattern [@grubler1990rise].
+These ideas were firstly introduced by Torsten Hägerstrand and his
+thesis entitled 'Innovation Diffusion as a Spatial Process'
+[@hagerstrand1968innovation]. Hägerstrand was the first one to identify
+diffusion as a geographical process. The starting point was the idea
+that diffusion is based on passing information through social networks,
+which themselves tend to be defined by geography. Hence, he identified
+the 'neighbourhood' effect of how information, and consequently,
+innovation diffuse. He used agricultural innovations to test and model
+his ideas using Monte Carlo simulations. Hägerstrand also incorporated
+the role of hierarchy and how some phenomena maybe firstly adopted in
+larger cities and then diffuse to second tier ones. This is a sequential
+instead of a simultaneous process, which resembles the 'lead-lag'
+spatial acceleration effect in market research [@bento2018time;
+@PERES201091]. Hägerstrand is more widely known though for highlighting
+the role time plays in the diffusion of innovations: an early-pioneering
+period, a middle fast accelerating period and a final saturation one
+[@morrill2020spatial].
 
-Schmidt's Law empirically illustrates a similar pattern. *Core* and usually highly 
-agglomerated regions is where new technologies are invented and commercially deployed 
-[@grubler1990rise; grubler1990rise]. This is where the first adopters tend to be 
-based. Then, technologies spread to the *rim* and eventually to the *periphery*. 
-Although adoption pace might be higher when new technologies finally arrive to the 
-periphery, the saturation levels there may never reach the ones in the core because 
-of the lack of infrastructure or other necessary institutions [@leibowicz2016representing].
-@grubler1990rise effectively summarises the three key characteristics of the spatial 
-diffusion process: (i) the cumulative level of adoption follows an S-shaped pattern 
-just like purely temporal models; (ii) diffusion is shaped by a hierarchy effect 
-in a form of a centrifugal force: from core to periphery; and (iii) diffusion is 
-also shaped by distance and a neighbourhood effect and contaminate nearby locations.
+The temporal dimension was further explored by Everett Rogers and his
+seminal work on 'Diffusion of Innovations' [@rogers2010diffusion].
+Rogers being a sociologist, his work focused not on the diffusion of
+innovations over space and time, but instead on the adoption of new
+technologies and innovations by individuals and the individual
+mechanisms that drive the decisions behind adoption. He identified five
+groups of individuals regarding their adoption speed: innovators, early
+adopters, early majority, late majority and laggards. The key mechanism of
+diffusion and adoption is communication and how knowledge is
+transferred within a social system. Therefore, all approaches agree on
+the S-shaped diffusion and cumulative adoption pattern
+[@grubler1990rise].
+
+Schmidt's Law empirically illustrates a similar pattern. *Core* and
+usually highly agglomerated regions is where new technologies are
+invented and commercially deployed [@grubler1990rise; @grubler1990rise].
+This is where the first adopters tend to be based. Then, technologies
+spread to the *rim* and eventually to the *periphery*. Although adoption
+pace might be higher when new technologies finally arrive to the
+periphery, the saturation levels there may never reach the ones in the
+core because of the lack of infrastructure or other necessary
+institutions [@leibowicz2016representing]. @grubler1990rise effectively
+summarises the three key characteristics of the spatial diffusion
+process: (i) the cumulative level of adoption follows an S-shaped
+pattern just like purely temporal models; (ii) diffusion is shaped by a
+hierarchy effect in a form of a centrifugal force: from core to
+periphery; and (iii) diffusion is also shaped by distance and a
+neighbourhood effect and contaminate nearby locations. These are the three
+mechanisms that the empirical analysis in Section [4]{#sec4} investigates.
 
 <!-- previous studies -->
-The remaining of this section reviews empirical studies which analysed the diffusion 
-on new technologies over space and time. Although the spatial dimension is present
-in most of the following studies, the level of spatial detail is almost always 
-more coarse than the one adopted here. @beardsell1999spatial studied 
-the evolution of the computer industry in $317$ US metro areas during the $1977-1992$ 
-period using employment data. Their analysis indicated that the relative size distribution
-holds for urban computer employment and also urban heterogeneity is essential in
-explaining this distribution. in a recent study, @bednarz2020pulled focused on 
-wind turbines and modelled their spatial diffusion across $402$ German regions 
-during $1970-2015$. Their key finding is that local demand than local supply was 
-the main driving factor.
 
-At a global scale @perkins2005international explored whther the diffusion rate of new
-technologies is driven by a latecomer advantage and the engagement with the global
-economy via foreign direct investments and trade. Their results illustrate that 
-indeed latecomers and developing countries experience diffusion of new technologies 
-more rapidly than early adopter and developed countries. At the same scale, 
-@perkins2011internet explored whether the adoption of previous communication
-technologies that is mail, telegrams and telephones was shaped by similar socioeconomic 
-factors as the internet. Their results indicated common patterns regarding the 
-drivers behind the adoption of different communication technologies.
+The remaining of this section reviews empirical studies which analysed
+the diffusion on new technologies over space and time. Although the
+spatial dimension is present in most of the following studies, the level
+of spatial detail is always more coarse than the one adopted
+here. @beardsell1999spatial studied the evolution of the computer
+industry in $317$ US metro areas during the $1977-1992$ period using
+employment data. Their analysis indicated that the relative size
+distribution holds for urban computer employment and also urban
+heterogeneity is essential in explaining this distribution. In a recent
+study, @bednarz2020pulled focused on wind turbines and modelled their
+spatial diffusion across $402$ German regions during $1970-2015$. Their
+key finding is that local demand than local supply was the main driving
+factor.
 
-Turning to studies that share more technological and scalar similarities with this 
-paper, @ding2010modeling modelled the spatial diffusion of mobile telecommunications 
-across regions in China. Their analysis indicated that socioeconomic characteristics 
-are important determinants of the timing, speed and the level of mobile diffusion
-within China. Using data from a Hungarian online social network, @lengyel2020role 
-analysed its adoption and the churn at a very granular spatial level. Their results 
-are in agreement with early theoretical and empirical contributions reviewed here:
-assortativity, urban scaling and distance are the key drivers of spatial diffusion.
-At a global scale @PAPAGIANNIDIS2015308 modelled the diffusion of different web
-technologies technologies and practices. Interestingly, they did so by using similar,
-but less extensive data as the one used here. Their analysis illustrated how the 
-diffusion of different web technologies and practices follow an S-shaped pattern
-as well as the different diffusion rates of the different technologies and practices.
+At a global scale @perkins2005international explored whther the
+diffusion rate of new technologies is driven by a latecomer advantage
+and the engagement with the global economy via foreign direct
+investments and trade. Their results illustrate that indeed latecomers
+and developing countries experience diffusion of new technologies more
+rapidly than early adopters and developed countries. At the same scale,
+@perkins2011internet explored whether the adoption of previous
+communication technologies that is mail, telegrams and telephones was
+shaped by similar socioeconomic factors as the internet. Their results
+indicated common patterns regarding the drivers behind the adoption of
+different communication technologies.
+
+Turning to studies that share more technological and scalar similarities
+with this paper, @ding2010modeling modelled the spatial diffusion of
+mobile telecommunications across regions in China. Their analysis
+indicated that socioeconomic characteristics are important determinants
+of the timing, speed and the level of mobile diffusion within China.
+Using data from a Hungarian online social network, @lengyel2020role
+analysed its adoption and the churn at a very granular spatial level.
+Their results are in agreement with early theoretical and empirical
+contributions reviewed here: assortativity, urban scaling and distance
+are the key drivers of spatial diffusion. At a global scale
+@PAPAGIANNIDIS2015308 modelled the diffusion of different web
+technologies technologies and practices. Interestingly, they did so by
+using similar, but less extensive data as the one used here. Their
+analysis illustrated how the diffusion of different web technologies and
+practices follow an S-shaped pattern as well as the different diffusion
+rates of the different technologies and practices.
+
+All in all, ...
 
@@
 <!-- in one region  -->
-
 <!-- @leibowicz2016representing for energy There is no generally accepted -->
@@
 <!-- incentivized scientific pursuits, domestic coal reserves, and a clear -->
+
 <!-- industrial motivation to replace water power with coal. ... -->
@@
 <!-- @bento2018time explore What determines the duration of formative phases -->
+
 <!-- for energy innovations in different markets? We are interested both in -->
+
 <!-- initial markets (also: core, lead, first mover, early adopter) where -->
+
 <!-- formative phases prepare technologies for mass commercialization, and in -->
+
 <!-- follower markets (also: periphery, lag, late adopter) where accelerated -->
+
 <!-- formative phases may benefit from diffusion and spillovers. -->
@@
 
+**ADD DATA
+DESCRIPTION, WHY .CO.UK, FAMILY OF WEB TECHNOLOGIES**
+
+[^2]: [https://archive.org/](See%20https://archive.org/).
+
 <!-- from our RS paper -->
-For instance, we can
-approximate digital economic activities in a region by
-measuring the volume of commercial webpages by considering
-only .co.uk websites, which are dedicated to commercial
-activities (Thelwall, 2000). Such commercial
-websites are used to exchange information, support online
-transactions and share opinions (Blazquez & Domenech,
-2018). Although nothing prevents a UK-based company
-from adopting a generic TLD such as.com and, indeed,
-such cases escape our data, we do not expect that such
-omissions could affect our results given the popularity of
-the .uk ccTLD: UK consumers prefer to visit a .uk website
-when they are searching for products or services (Hope,
-2017); and anecdotal evidence indicates that during the
-first half of 2000, three .co.uk domains were registered
-every minute (OECD, 2001). Also, as Table 1 demonstrates,
-.co.uk was the most popular SLD under the .uk
-ccTLD in 2000. Hence, we use the total volume of all
-the archived webpages in a region as a proxy for the level
-of digitization. We start the analysis by focusing on commercial
-webpages (.co.uk).
+
+For instance, we can approximate digital economic activities in a region
+by measuring the volume of commercial webpages by considering only
+.co.uk websites, which are dedicated to commercial activities (Thelwall,
+2000). Such commercial websites are used to exchange information,
+support online transactions and share opinions (Blazquez & Domenech,
+2018). Although nothing prevents a UK-based company from adopting a
+generic TLD such as.com and, indeed, such cases escape our data, we do
+not expect that such omissions could affect our results given the
+popularity of the .uk ccTLD: UK consumers prefer to visit a .uk website
+when they are searching for products or services (Hope, 2017); and
+anecdotal evidence indicates that during the first half of 2000, three
+.co.uk domains were registered every minute (OECD, 2001). Also, as Table
+1 demonstrates, .co.uk was the most popular SLD under the .uk ccTLD in
+2000. Hence, we use the total volume of all the archived webpages in a
+region as a proxy for the level of digitization. We start the analysis
+by focusing on commercial webpages (.co.uk).
 
@@
 
-1. S-shaped diffusion curves: S for LADs per firm. UK, fast/slow/examples
+1.  S-shaped diffusion curves: S for LADs per firm. UK,
+    fast/slow/examples
 
@@
 3.  Neighbourhood effect: diffusion proceeds outwards from innovation
-centers, first "hitting" nearby rather than far-away locations (Grubler
-1990)
+    centers, first "hitting" nearby rather than far-away locations
+    (Grubler 1990)
 
@@
 
-  -   Website density regressions: for OA and LAD over time
+-   ~~Website density regressions: for OA and LAD over time~~
 
-4. Hierarchy effect: from main centers to secondary ones -- central places
+4.  Hierarchy effect: from main centers to secondary ones -- central
+    places
 
@@
 
-  - ideal: (i) train RF for all years and all (1) LADs and (2) OAs with CAST
-  and report metrics. (ii) train for all years and per region for (1) LADs and (2)
-  OAs, and predict to the other regions. Reports predictions as region similarities
+-   ideal: (i) train RF for all years and all (1) LADs and (2) OAs with
+    CAST and report metrics. (ii) train for all years and all but one
+    region for (1) LADs and (2) OAs to predict to the holdout region.
+    Reports predictions as region similarities.
 
-# Discussion and conclusions {#sec4}
+### RF results
+
+The next section incorporates the above discussed spatial processes of the
+diffusion of web technologies into the same modelling framework. The aim
+is to use variables depicting these spatial processes in order to
+predict the diffusion of web technologies in the UK over space and time
+and across different scales. Specifically, four different models are
+estimated. Firstly, all the data points at the OA and LAD are utilised
+in order to build RF models and assess their capacity to predict the
+adoption of web technologies. These two models will reveal the
+predictive capacity of the spatial processes and also allow to see how
+the importance of such variables changes across scales. The next two
+sets of models will be again trained on web diffusion at the two working scales:
+OA and LAD. However, instead of using all the data points, the OA and
+the LAD from one of the twelve UK regions are held out and then the trained
+model is used to predict web adoption in the OA or the LAD of the
+held-out region. This process takes place recursively for all UK
+regions. The difference in the predictive capacity of the different
+samples will reveal how dissimilar are these spatial process across
+regions and, importantly, at different scales.
+
+It needs to be highlighted here that the cross-validation for all models
+is spatially and temporally sensitive. Instead of using
+10 random folds, we employ the `CAST` package which allows holding back
+data points from specific years and spatial units and use them for
+testing in order to estimate the model performance
+[@meyer2018improving].
+
+The models need to include variables that capture the three processes
+that the relevant literature and the descriptive analysis presented in
+the Section **XX** highlighted. Namely, the models capture: (i) a
+hierarchy effect with diffusion running from main centres to secondary
+ones, (ii) a neighborhood effect according to which diffusion first hits
+nearby locations, and (iii) the rather canonical pattern of 
+diffusion over time as reflected in the S-shaped pattern in the
+cumulative level of adoption.
+
+To capture the hierarchy effect the models include as predictors a one
+year lag of website density in London -- the largest city in the UK, a
+one year lag of the website density in the nearest city and the same for
+the nearest retail centre. Due to the small sizes of the retail centres,
+the latter is only relevant for the OA-level models. In addition, the
+models include the distance to London, the nearest city and the nearest
+retail centre. The underlying logic is that the level of website
+adoption in a spatial unit depends on the level of the adoption in
+places further up in the urban hierarchy the previous year. To depict
+the neighbourhood effect, the web density of the neighbouring spatial
+units in the previous year is employed. Again, the underpinning
+rationale is that the level of web adoption within a spatial unit
+depends on the level of web adoption in the neighbouring spatial units
+the year before. This is the 'hitting nearby locations first' argument.
+Therefore, the spatial and temporal lag of the website density in LAD
+and OA is calculated. Lastly, the time effect which is reflected on the
+S curve for the cumulative adoption is captured by trend variable.
+Hence, all four model will follow the following generic form 
+(Eq. \ref{model}):
+
+```{=tex}
+\begin{align} \label{model}
+Website\,Density_{t} \sim Distance\,London +
+Website\,density\,London_{t-1} +\notag\\
+Distance\,Nearest\,City +
+Website\,density\,Nearest\,City_{t-1} +\notag\\
+Distance\,Nearest\,Retail_{i} +
+Website\,density\,Nearest\,Retail_{t-1} +\notag\\
+W*\, Website\,density_{t-1} +\notag\\ 
+year_{t}
+\end{align}
+```
+
+To assess the predictive capability of the model, three broadly utilised
+metrics are employed: the coefficient of determination (R squared), mean
+absolute error (MAE) and root mean square error (RMSE):
+
+```{=tex}
+\begin{align}
+R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
+\end{align}
+```
+```{=tex}
+\begin{align}
+MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
+\end{align}
+```
+```{=tex}
+\begin{align}
+RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
+\end{align}
+```
+
+$y_{k}$ is the $k^{th}$ observation of the dataset, which consists of
+$N$ observations in total. $\hat{y_{k}}$ is the $k_{th}$ predicted value
+for the dependent variable and $\overline{y_{k}}$ is the average value
+of $y$. The last two metrics are expressed in the same units as the
+dependent variable -- websites per firm for the LAD modes and the number
+of websites for the OA models -- while the first one is the coefficient
+of determination between the observed and the predicted values of
+website adoption. Regarding $MAE$, it is the absolute difference between
+the observed and the predicted website adoption. While $MAE$ does not
+penalise for large errors, $RSME$ does so as it is proportional to the
+squared difference between the observed and the predicted trade flows.
+Hence larger errors weigh more for $RMSE$ [@pontius2008components].
+
+Table \ref{table.metrics.all} presents the model performance for the
+first set on models, for which all data points are employed for training
+and testing via cross validation. The first one is trained and tested on 374 LAD
+and the second on 232,296 OA, both for a 16 year period (1997-2012). The results are 
+remarkably good considering that the are the outcome of space and time sensitive 
+CV, so the the model does not suffer from overfitting. At the LAD level the model 
+predicts 81% of the variation of website density. Both error metrics indicate that
+the model error is $(\frac{1}{20}, \frac{1}{30})$ of a website per firm. At 
+the OA the RSquared drops down to 21%. Considering its granularity, this is still 
+a remarkable performance. To contextualise it, the model results in a MAE of one 
+website for areas small enough to host less than 140 households.^[According to the Office for
+National Statistics, 80% of OA in England and Wales host 110-139 households, [www.ons.gov.uk](https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas).] 
+Because of the small size of the spatial units, the distribution is highly skewed
+and a significant part of them is not linked to any websites. In 1997 only 1% of 
+the UK OA were associated with a website. This should not come as a surprise as 
+this was the very beginning of the commercial internet and any activities with a 
+digital footprint were concentrated in a handful of areas. This was clearly illustrated
+in Section **ADD**. At the end of the study period almost half of the UK OA were
+not associated with a website. Again, given the granularity of the data this should
+not come as a surprise.
+
+|                   | RMSE  | RSquared | MAE   |
+|-------------------|-------|----------|-------|
+| Local Authorities | 0.032 | 0.810    | 0.019 |
+| Output Areas      | 5.000 | 0.205    | 1.047 |
+
+: Model metrics {#table.metrics.all}
+
+Figures \ref{var.imp.LAD} and \ref{var.imp.OA} plot the importance of the different 
+predictors. When the focus is on the LAD, the website density in the nearest city,
+in London and in the neighbouring LADs the year before are the most important predictors.
+They are followed by the yearly trend, while the spatial configuration as reflected
+in distances to London or the nearest city only play a minor role. This can be attributed
+to the rather coarse spatial scale of analysis. Nevertheless, all previously discussed
+spatial processes are at play in the diffusion of web technologies at the LAD level: 
+the first two predictors depict the hierarchical effect, the spatial and temporal lag of website
+density depict the neighbourhood effect and the yearly trend the time-sensitive 
+cumulative adoption pattern.
+
+When the much more granular scale of OA is adopted, the picture is reversed. The 
+most important predictors are the three distance variables to London, the nearest 
+city and the nearest retail centre. They stil depict the hierarchical effect, but
+proximity to the different population centres is more important than their lagged 
+web densities in predicting website diffusion. The neighbouring effect is less 
+important at this scale. What is interesting is the almost negligible role of the 
+yearly trend and London's website density. While the former probably illustrates 
+the large heterogeneity in how web technologies have been adopted at this very fine 
+scale, the later highlights that the importance of past web adoption rates in large 
+population centres is surpassed by proximity to them and soatial configuration 
+at this scale.
+
+```{r varimpLA, eval = T, echo=FALSE, message = F, fig.cap="\\label{var.imp.LAD}Variable importance, LAD"}
+path.image <- paste0(path, "/outputs/rf/figures/varimp_LA.png")
+knitr::include_graphics(path.image)
+```
+
+```{r varimpOA, eval = T, echo=FALSE, message = F, fig.cap="\\label{var.imp.OA}Variable importance, OA"}
+path.image <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
+knitr::include_graphics(path.image)
+```
+
+Table \ref{table.regions} presents the results of the recursive hold out models,
+which aim to highlight the potential regional heterogeneity of the spatial processes
+behind the diffusion of web technologies. To begin with, as highlighted before, 
+there is a difference of magnitude 
+of one order between the LAD and the OA prediction errors, which is aligned with 
+previous results that employed all data points. What is of interest here is the 
+regional comparison. Table \ref{table.regions} illustrates some striking similarities, 
+but also a few significant differences. The regions the web diffusion of which is 
+better predicted using models trained in the rest of the country are the same 
+despite the scale of analysis: South East, Wales, Yorkshire and The Humber and 
+the North East of England. In other words, these are the regions whose spatial 
+diffusion mechanisms of web technologies is closer to the country's average. Despite the 
+consistency across scales, this is a diverse set of regions: **ADD CHARACTERISTICS**.
+
+At the other end of the spectrum, Scotland's and the North West's web diffusion
+mechanisms are consistently diverging from the country's average. This should not
+come as a surprise as these regions are characterised of high levels of rurality
+and remoteness. Similarly, London diffusion mechanisms diverge from the country's 
+average and this is consistent across scales. London's uniqueness in UK's urban 
+system and economy is also reflected in the spatial diffusion mechanisms of web 
+technologies within its LAD and OA. It needs to be highlighted though that the 
+difference between the RSquared of LAD and OA is more that an order of magnitude 
+signaling how difficult is to predict diffusion at such a small spatial scale.
+Northern Ireland is an interesting case. While it ranks at the bottom of the scale
+when the models are trained and tested on LAD data, when the modelling adopts the
+more granular OA scale, the spatial mechanisms that shape the web diffusion within
+this region appear to be closer to the country's average. At this scale, proximity,
+or lack of, relative to the rest of the country become less important and the 
+internal to the region spatial structure predictors start playing a more
+important role **CHECK NI OA model**.
+
+```{r regions}
+path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
+path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
+
+lad <- read.csv(path.lad) %>% 
+  rename('RSquared LAD' = Rsquared)
+
+oa <- read.csv(path.oa) %>% rename(Region = test.region,
+                                   'RSquared OA' = Rsquared)
+
+lad %>% left_join(oa) %>% 
+  mutate('Rank OA' = rank(desc(`RSquared OA`)),
+         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
+  relocate(5, .after = 2) %>% 
+  arrange(`Rank LAD`) %>% 
+  kable(caption= "Regional differences\\label{table.regions}", digits=3)
+```
+
+# Discussion and conclusions
 

### 2024-06-11 paper/submission1/tranos2023.tex
pre-ggplot

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, interior hidden, sharp corners, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 to different extents in different places, and can be significantly
-influenced by policies \citep{victor1993}''. \citet{meade2021modelling}
+influenced by policies'' \citep{victor1993}. \citet{meade2021modelling}
 highlight that a variety of actors have a direct interest in gaining
 such knowledge including network equipment suppliers; network operators,
-regulatory and local authorities. These processes and their effects very
+regulatory and local authorities. These processes and their effects vary
 a lot across scales: although the diffusion of a new technology might
@@
 `lead-lag' spatial acceleration effect in market research
-\citep{bento2018time, PERES201091}. Hägerstrand maybe more widely known
-for highlighting the role time plays in the diffusion of innovations: an
-early-pioneering period, a middle fast accelerating period and a final
-saturation one {[}\citet{morrill2020spatial}.
+\citep{bento2018time, PERES201091}. Hägerstrand is more widely known
+though for highlighting the role time plays in the diffusion of
+innovations: an early-pioneering period, a middle fast accelerating
+period and a final saturation one \citep{morrill2020spatial}.
 
@@
 seminal work on `Diffusion of Innovations' \citep{rogers2010diffusion}.
-Rogers being a sociologist his work focused not on the diffusion of
+Rogers being a sociologist, his work focused not on the diffusion of
 innovations over space and time, but instead on the adoption of new
@@
 of spatial detail is almost always more coarse than the one adopted
-here. To begin with \citet{beardsell1999spatial} studied the evolution
-of the computer industry in \(317\) US metro areas during the
-\(1977-1992\) period using employment data. Their analysis indicated
-that the relative size distribution holds for urban computer employment
-and also urban heterogeneity is essential in explaining this
-distribution. \citet{bednarz2020pulled} focused on wind turbines and
-modelled their spatial diffusion across \(402\) German regions during
+here. \citet{beardsell1999spatial} studied the evolution of the computer
+industry in \(317\) US metro areas during the \(1977-1992\) period using
+employment data. Their analysis indicated that the relative size
+distribution holds for urban computer employment and also urban
+heterogeneity is essential in explaining this distribution. in a recent
+study, \citet{bednarz2020pulled} focused on wind turbines and modelled
+their spatial diffusion across \(402\) German regions during
 \(1970-2015\). Their key finding is that local demand than local supply
@@
 
-At a global scale \citet{perkins2005international} analyzed if the
+At a global scale \citet{perkins2005international} explored whther the
 diffusion rate of new technologies is driven by a latecomer advantage
@@
 
-\hypertarget{sec3.1}{%
-\subsection{Results}\label{sec3.1}}
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{old-ideas}{%
+\subsection{Old ideas}\label{old-ideas}}
 
@@
 \item
-  Moran's I: for OA and LAD over time \textbf{TODO} add 0s
+  Moran's I: for OA and LAD over time
 \item
@@
 \item
-  S for LADs per firm and OA. \textbf{TODO}: OA per firm? fix firms over
-  time.
+  S for LADs per firm and OA.
 \item
@@
 \item
-  \textbf{TODo} check s\_uk\_firm\_11
+  s\_uk\_firm\_11
 \item
@@
 
-\hypertarget{sec3.2}{%
-\subsection{3.2 Case study approach}\label{sec3.2}}
+\hypertarget{new-structure}{%
+\subsection{New structure}\label{new-structure}}
 
-\hypertarget{sec4}{%
-\section{Digital technologies and spatial structure: a global
-view}\label{sec4}}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  S-shaped diffusion curves: S for LADs per firm. UK, fast/slow/examples
+\item
+  ranks: there is stability and movement
+\item
+  Neighbourhood effect: diffusion proceeds outwards from innovation
+  centers, first ``hitting'' nearby rather than far-away locations
+  (Grubler
+\end{enumerate}
 
-\hypertarget{sec5}{%
-\section{The impact of ICT on the US and the UK spatial
-structures}\label{sec5}}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\setcounter{enumi}{1989}
+\tightlist
+\item
+\end{enumerate}
 
-\hypertarget{sec6}{%
-\section{Discussion and conclusions}\label{sec6}}
+\begin{itemize}
+\item
+  Moran's I: for OA and LAD over time
+\item
+  LISA maps: for OA and LAD over time More and less expected clusters.
+  Different scales show different results
+\item
+  Website density regressions: for OA and LAD over time
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Hierarchy effect: from main centers to secondary ones -- central
+  places
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  Gini coefficient. Almost perfect polarisation of web adoption in the
+  early stages at a granular level More equally diffused at the Local
+  Authority level Plateau overtime
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{4}
+\tightlist
+\item
+  RF
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  ideal: (i) train RF for all years and all (1) LADs and (2) OAs with
+  CAST and report metrics. (ii) train for all years and all but one
+  region for (1) LADs and (2) OAs to predict to the holdout region.
+  Reports predictions as region similarities.
+\end{itemize}
+
+\hypertarget{rf-results}{%
+\subsubsection{RF results}\label{rf-results}}
+
+The next section incorporates the above discussed spatial drivers of the
+diffusion of web technologies into the same modelling framework. The aim
+is to use variables depicting these spatial processes in order to
+predict the diffusion of web technologies in the UK over space and time
+and across different scales. Specifically, four different models are
+estimated. Firstly, all the data points at the OA and LAD are utilised
+in order to build RF models and assess their capacity to predict the
+adoption of web technologies. These two models will reveal the
+predictive capacity of the spatial processes and also allow to see how
+the importance of such variables changes across scales. The next two
+models will be again trained on web difusion at the two working scales:
+OA and LAD. However, instead of using all the data points, the OA and
+the LAD from one of the 12 UK regions are hold out and then the trained
+model is used to predict web adoption in the OA or the LAD of the
+held-out region. This process takes place recursively for all UK
+regions. The difference in the predictive capacity of the different
+samples will reveal how different are these spatial process across
+regions and, importantly, at different scales.
+
+It needs to be highlighted here that the cross-validation for all models
+presented here is spatially and temporally sensitive. Instead of using
+10 random folds, we employ the \texttt{CAST} package which allows holds
+back data points from specific years and spatial units and use them for
+testing in order to estimate the model performance
+\citep{meyer2018improving}.
+
+The models need to include variables that capture the three processes
+that the relevant literature and the descriptive analysis presented in
+the previous section highlighted. Namely, the models capture: (i) a
+hierarchy effect with diffusion running from main centres to secondary
+ones, (ii) a neighborhood effect according to which diffusion first hits
+nearby locations, and (iii) the rather canonical pattern of the
+diffusion over time as reflected in the S-shaped pattern in the
+cumulative level of adoption.
+
+To capture the hierarchy effect the models include as predictors a one
+year lag of website density in London -- the largest city in the UK, a
+one year lag of the website density in the nearest city and the same for
+the nearest retail centre. Due to the small sizes of the retail centres,
+the latter is only relevant for the OA-level models. In addition, the
+models include the distance to London, the nearest city and the nearest
+retail centre. The underlying logic is that the level of website
+adoption in a spatial unit depends on the level of the adoption in
+places further up in the urban hierarchy the previous year. To depict
+the neighbourhood effect, the web density of the neighbouring spatial
+units in the previous year is calculated. Again, the underpinning
+rationale is that the level of web adoption within a spatial unit
+depends on the level of web adoption in the neighbouring spatial units
+the year before. This is the `hitting nearby locations first' argument.
+Therefore, the spatial and temporal lag of the website density in LAD
+and OA is calculated. Lastly, the time effect which is reflected on the
+S curve for the cumulative adoption is captured by trend variable.
+Hence, all four model will follow the following generic form
+\ref{model}:
+
+\begin{align} \label{model}
+Website\,Density_{t} \sim Distance\,London +
+Website\,density\,London_{t-1} +\notag\\
+Distance\,Nearest\,City +
+Website\,density\,Nearest\,City_{t-1} +\notag\\
+Distance\,Nearest\,Retail_{i} +
+Website\,density\,Nearest\,Retail_{t-1} +\notag\\
+W*\, Website\,density_{t-1} +\notag\\ 
+year_{t}
+\end{align}
+
+To assess the predictive capability of the model, three broadly utilised
+metrics are employed: the coefficient of determination (R squared), mean
+absolute error (MAE) and root mean square error (RMSE):
+
+\begin{align}
+R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
+\end{align}
+
+\begin{align}
+MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
+\end{align}
+
+\begin{align}
+RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
+\end{align}
+
+\(y_{k}\) is the \(k^{th}\) observation of the dataset, which consists
+of \(N\) observations in total. \(\hat{y_{k}}\) is the \(k_{th}\)
+predicted value for the dependent variable and \(\overline{y_{k}}\) is
+the average value of \(y\). The last two metrics are expressed in the
+same units as the dependent variable -- websites per firm for the LAD
+modes and the number of websites for the OA models -- while the first
+one is the coefficient of determination between the observed and the
+predicted values of website adoption. Regarding \(MAE\), it is the
+absolute difference between the observed and the predicted website
+adoption. While \(MAE\) does not penalise for large errors, \(RSME\)
+does so as it is proportional to the squared difference between the
+observed and the predicted trade flows. Hence larger errors weigh more
+for \(RMSE\) \citep{pontius2008components}.
+
+Table \ref{table.metrics.all} presents the model performance for the
+first set on models, for which all data points are employed for training
+and testing via cross validation. The first one is trained and tested on
+374 LAD and the second on 232,296 OA for a 16 year period (1997-2012).
+The results are remarkably good considering that the are the outcome of
+space and time sensitive CV, so the the model does not suffer from
+overfit. At the LAD level we the model predicts 81\% of the variation of
+website density. Both error metrics indicate that the model error is
+\((\frac{1}{20} - \frac{1}{30})\) of a website per firm. At the OA this
+score drops down to 21\%. Considering its granularity, this is still a
+remarkable performance. To contextualise it, the model results in a MAE
+of one website for areas small enough to host less than 140
+households.\footnote{According to the Office for National Statistics,
+  80\% of OA in England and Wales host 110-139 households,
+  \href{https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas}{www.ons.gov.uk}.}
+Because of the small size of the spatial units, the distribution is
+highly skewed and a significant part of them is not linked to any
+websites. In 1997 only 1\% of the UK OA were associated with a website.
+This should not come as a surprise as this was the very beginning of the
+commercial internet and any activities with a digital footprint were
+concentrated in a handful of areas. This was clearly illustrated in
+Section \textbf{ADD}. At the end of the study period almost half of the
+UK OA were not associated with a website. Again, given the granularity
+of the data this should not come as a surprise. Despite the large
+heterogeneity in website adoption, the model reached an R-Squared of
+0.2.
+
+\begin{longtable}[]{@{}llll@{}}
+\caption{Model metrics \{\#table.metrics.all\}}\tabularnewline
+\toprule\noalign{}
+& RMSE & RSquared & MAE \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+& RMSE & RSquared & MAE \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Local Authorities & 0.032 & 0.810 & 0.019 \\
+Output Areas & 5.000 & 0.205 & 1.047 \\
+\end{longtable}
+
+Figures \ref{var.imp.LAD} and \ref{var.imp.OA} plot the importance of
+the difrerent predictors. When the focus is on the LAD, the website
+density in the nearest city, in London and on the neighbouring LADs the
+year before are the most important predictors. They are followed by the
+yearly trend, while the spatial configuration as reflected on distances
+to London or the nearest city only play a minor role. This can be
+attributed to the rather coarse spatial scale of analysis. Nevertheless,
+all previously discussed forces are at play in the diffusion of web
+technologies at the LAD: the first two predictors depict the
+hierarchical effect, the spatial and temporal lag of website density the
+neighbourhood effect and the yearly trend the time-sensitive cumulative
+adoption pattern.
+
+When the much more granular scale of OA is adopted, the picture is
+reversed. The most important predictors are the three distance variables
+to London, the nearest city and the nearest retail centre. They stil
+depict the hierarchical effect, but proximity to the different
+population centres is more important than their lagged web densities in
+predicting website diffusion. The neighbouring effect is less important
+at this scale. What is interesting is the almost negligible role of the
+yearly trend and London's website density. While the former probably
+illustrates the large heterogeneity in how web technologies have been
+adopted at this very fine scale, the later highlights that the
+importance of past web adoption rates in large population centres is
+surpassed by proximity to them and soatial configuration at this scale.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/rf/figures/varimp_LA.png}
+
+}
+
+\caption{\label{var.imp.LAD}Variable importance, LAD}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/rf/figures/varimp_OA.png}
+
+}
+
+\caption{\label{var.imp.OA}Variable importance, OA}
+
+\end{figure}
+
+Table \ref{table.regions} presents the results of the recursive hold out
+models which aim to highlight potential regional heteorigeinty of the
+web diffusion spatial mechanisms. To begin with, as highlighted before,
+there is a difference of magnitude of one order between the LAD and the
+OA prediction errors, which is aligned with previous results that
+employed all data points. What is of interest here is the regional
+comparison. Table \ref{table.regions} illustrates some striking
+similarities, but also a few significant differences. The regions the
+web diffusion of which is better predicted using models trained in the
+rest of the country are the same despite the scale of analysis: South
+East, Wales, Yorkshire and The Humber and the North East of England. In
+other words, these are the regions whose spatial diffusion of web
+technologies is closer to the country's average. Despite the consistency
+across scales, this is a diverse set of regions: \textbf{ADD
+CHARACTERISTICS}.
+
+At the other end of the spectrum, Scotland's and the North West's web
+diffusion mechanisms are consistently diverging from the country's
+average. This should not come as a surpise as these regions are
+characterised of high levels of rurality and remoteness. Similarly,
+London diffusion mechanisms diverge from the country's average and this
+is consistent across scales. London's uniqueness in UK's urban system
+and economy is also reflected in the spatial diffusion mechanisms of web
+technologies within its LAD and OA. It needs to be highlighted though
+that the difference between the RSquared of LAD and OA is more that an
+order of magnitude signaling how difficult is to predict diffusion at
+such a small spatial scale. Northern Ireland is an interesting case.
+While it ranks at the bottom of the scale when the models are trained
+and tested on LAD data, when the modelling adopts the more granular OA
+scale, the spatial mechanisms that shape the web diffusion within this
+region appear to be closer to the country's average. At this scale,
+proximity or lack of in relative terms to the rest of the country become
+a less important and the internal to the region spatial structure
+predictors start playing a more important role \textbf{CHECK NI OA}.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3731}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1940}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1343}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1791}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1194}}@{}}
+\caption{Regional differences\label{table.regions}}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Region
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+RSquared LAD
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Rank LAD
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+RSquared OA
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Rank OA
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Region
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+RSquared LAD
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Rank LAD
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+RSquared OA
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Rank OA
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+South East & 0.947 & 1 & 0.134 & 2 \\
+Wales & 0.916 & 2 & 0.131 & 3 \\
+Yorkshire and The Humber & 0.906 & 3 & 0.144 & 1 \\
+North East & 0.895 & 4 & 0.128 & 4 \\
+West Midlands & 0.883 & 5 & 0.070 & 9 \\
+East Midlands & 0.882 & 6 & 0.088 & 8 \\
+East of England & 0.876 & 7 & 0.106 & 6 \\
+South West & 0.864 & 8 & 0.117 & 5 \\
+London & 0.805 & 9 & 0.055 & 10 \\
+Scotland & 0.770 & 10 & 0.035 & 11 \\
+North West & 0.664 & 11 & 0.017 & 12 \\
+Nortern Ireland & 0.576 & 12 & 0.101 & 7 \\
+\end{longtable}
+
+\hypertarget{discussion-and-conclusions}{%
+\section{Discussion and conclusions}\label{discussion-and-conclusions}}
 

### 2024-06-11 presentations/web_diffusion.Rmd
pre-ggplot

--- 2024-02-15 presentations/web_diffusion.Rmd
@@
 ---
-title: 'A multi-scalar story of the diffusion of a new technology: the web'
+title: 'A multi-scale story of the diffusion of a new technology: the web'
 # date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 
--   **Granular** and **multi-scale** spatial perspective
+-   Granular and multi-scale **spatial processes**
 
@@
 
--   Descriptive statistics, ESDA & density regressions
+-   Descriptive statistics & ESDA
 
@@
 
--   Websites in Output Areas (c. 200,000)
+-   Websites in Output Areas (c. 230,000)
 
@@
 |-------------------|-------|----------|-------|
-| Local Authorities | 0.032 | 0.810    | 0.019 |
-| Output Areas*      | 5.000 | 0.205    | 1.047 |
+| Local Authorities* | 0.032 | 0.810    | 0.019 |
+| Output Areas       | 5.000 | 0.205    | 1.047 |
 
 <small>
-*No retail centre variables for the Local Authority models
+*No retail centre predictors for the Local Authority models
 </small>
@@
 
-## Regional similarities for Local Authorities
+## Regional similarities
 
-```{r similarity_LA, eval = T, message=FALSE}
-path.table <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
-read.csv(path.table) %>% 
-  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
-  kable()
-```
+```{r similarity, eval = T, message=FALSE}
+path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
+path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
 
-## Regional similarities for Output Areas
+lad <- read.csv(path.lad) %>% 
+  rename('RSquared LAD' = Rsquared)
 
-```{r similarity_OA, out.height= "70%", out.width="70%", message=FALSE}
-path.table <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
-read.csv(path.table) %>% 
-  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
-  kable()
+oa <- read.csv(path.oa) %>% rename(Region = test.region,
+                                   'RSquared OA' = Rsquared)
+
+lad %>% left_join(oa) %>% 
+  mutate('Rank OA' = rank(desc(`RSquared OA`)),
+         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
+  relocate(5, .after = 2) %>% 
+  arrange(`Rank LAD`) %>% 
+  kable(digits=3)
 ```
@@
 <small>
-*No retail centre variables for the Local Authority models
+*No retail centre predictors for the Local Authority models
 </small>
@@
   mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
+  arrange(desc(Rsquared)) %>% 
   kable()
@@
 
--   Established technological diffusion drivers still apply
+-   Established technological diffusion spatial processes still apply
 
@@
 
--   Heterogeneity increases with resolution
-
--   Things to do: OA?
+-   Regional heterogeneity
 

### 2024-06-11 presentations/web_diffusion.Rmd
path

@@
 <!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture1.png){width="70%"} -->
+
 ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture1.png){width="70%"}
@@
 <!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture2.png){width="70%"} -->
+
 ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture2.png){width="70%"}
@@
 <!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png) -->
+
 ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
@@
 <!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png) -->
+
 ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
@@
 ```{r moransi_la, results='asis'}
-path.in <- paste0(path, "/outputs/lisa/oa_ggplot/morani_la.csv")
+path.in <- paste0(path, "/outputs/lisa/la_ggplot/morani_la.csv")
 
@@
 ```{r lisa_la, results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/oa_ggplot/lisa/")
+path.images <- paste0(path, "/outputs/la_ggplot/lisa/")
 
@@
 
--   [A hierarchy effect: from main centres to secondary ones]{style="color:orange;"}
+-   [A hierarchy effect: from main centres to secondary
+    ones]{style="color:orange;"}
 
--   [A neighborhood effect: first "hitting" nearby locations]{style="color:olive;"}
+-   [A neighborhood effect: first "hitting" nearby
+    locations]{style="color:olive;"}
 
--   [S-shaped pattern in the cumulative level of adoption]{style="color:violet;"}
+-   [S-shaped pattern in the cumulative level of
+    adoption]{style="color:violet;"}
 
@@
 |                     | RMSE  | RSquared | MAE   |
-|-------------------|-------|----------|-------|
-| Local Authorities* | 0.032 | 0.810    | 0.019 |
+|---------------------|-------|----------|-------|
+| Local Authorities\* | 0.032 | 0.810    | 0.019 |
 | Output Areas        | 5.000 | 0.205    | 1.047 |
 
-<small>
-*No retail centre predictors for the Local Authority models
+<small> \*No retail centre predictors for the Local Authority models
 </small>
@@
 |                     | RMSE  | RSquared | MAE   |
-|-------------------|-------|----------|-------|
-| Local Authorities* | 0.200 | 0.634    | 0.148 |
+|---------------------|-------|----------|-------|
+| Local Authorities\* | 0.200 | 0.634    | 0.148 |
 
-<small>
-*No retail centre predictors for the Local Authority models
+<small> \*No retail centre predictors for the Local Authority models
 </small>

### 2024-06-11 presentations/web_diffusion.Rmd
path

@@
 # path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"
-path.2000 <- "/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/all2000couk.csv"
+# path.2000 <- "/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/all2000couk.csv"
+path.2000 <- paste0(path, "/data/temp/all2000couk.csv")
 

### 2024-06-11 presentations/web_diffusion.Rmd
path

@@
 <!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture1.png){width="70%"} -->
-
-![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture1.png){width="70%"}
+<!-- ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture1.png){width="70%"} -->
+![](images/Picture1.png){width="70%"}
 
@@
 <!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture2.png){width="70%"} -->
-
-![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture2.png){width="70%"}
+<!-- ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture2.png){width="70%"} -->
+![](images/Picture2.png){width="70%"}
 

### 2024-06-11 presentations/web_diffusion.Rmd
path

@@
 <!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png) -->
-
-![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
+<!-- ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png) -->
+![](images/screenshot_directory_website_2010.png)
 
@@
 <!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png) -->
-
-![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
+<!-- ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png) -->
+![](images/screenshot_2010.png)
 

### 2024-06-12 presentations/web_diffusion.Rmd
presentation

@@
 ```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
-path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only05.png")
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only05_av.png")
 knitr::include_graphics(path.image)
@@
 ```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
-path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only95.png")
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only95_av.png")
 knitr::include_graphics(path.image)
@@
 
-```{r lisa_la, results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/la_ggplot/lisa/")
+```{r lisa_la, results='asis', out.width = "60%"}
+path.images <- paste0(path, "/outputs/lisa/la_ggplot/")
 
@@
 
-```{r lisa_oa, eval=T,results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/oa_ggplot/lisa/")
+```{r lisa_oa, eval=T,results='asis', out.width = "60%"}
+path.images <- paste0(path, "/outputs/lisa/oa_ggplot/")
 

### 2024-06-12 src/la_ranks_n1.Rmd
mac

@@
 path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-la <- readOGR(path.geo)
+la <- st_read(path.geo)
+# source: https://geoportal.statistics.gov.uk/
 
-la.names <- la@data %>% 
+la.names <- la %>% 
   dplyr::select(LAD21CD, LAD21NM) %>% 
@@
 
-df <- df %>% left_join(la.names)
+df <- df %>% left_join(la.names) %>% 
+  dplyr::select(-geometry)
 # source: https://geoportal.statistics.gov.uk/
@@
   mutate(diff_1st = -(rank-lag(rank))) %>% 
-  #ungroup() %>% summarise(q = quantile(abs(diff_1st), .95, na.rm = T)) 
-  # t= 2000: q.95 = 199.5, q.9 = 168, q.1 = 7, q.05 = 3.75. **abs()**
-  # t= 1996: q.95 = 257.125, q.9 = 232.25, q.1 = 17.75, q.05 = 6.875. **abs()**
+  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .1, na.rm = T)) 
+  # t = 2000: q.95 = 194.5, q.90 = 168.5, q.1 = 9, q.05 = 3.75. **abs()**, 12/6/2024
+  # t = 2000: q.95 = 199.5, q.9 = 168, q.1 = 7, q.05 = 3.75. **abs()**
+  # t = 1996: q.95 = 257.125, q.9 = 232.25, q.1 = 17.75, q.05 = 6.875. **abs()**
 
-  mutate(alpha = ifelse(abs(diff_1st) > 199.5, .8, 0.1)) %>% 
+  mutate(alpha = ifelse(abs(diff_1st) > 194.5, .8, 0.1)) %>% 
+  #mutate(alpha = ifelse(abs(diff_1st) <= 3.75, .8, 0.1)) %>%
+
   mutate(alpha = ifelse(is.na(diff_1st), 0.1, alpha)) %>%
@@
 
-## Ranks websites 2000 and 2012 only
+## Average ranks websites per firm t and 2012 only
+
+```{r}
+
+# change percentile in out.path and mutate(alpha = ifelse(abs(diff_1st)
+
+out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only95_av.png")
+
+df %>% 
+  mutate(period = ifelse(year < 1999, 1, 
+                         ifelse(year > 2009, 3, 2))) %>%
+  group_by(ladcd, period) %>%
+  mutate(n_by_firm = ifelse(period != 2, mean(n_by_firm), n_by_firm)) %>% 
+  filter(year > 1997 & year < 2011) %>% 
+  group_by(year) %>% 
+  mutate(rank = rank(desc(n_by_firm))) %>% # or n_by_firm
+  ungroup() %>% 
+  filter(year==1998 | year == 2010) %>% 
+  arrange(ladcd, year) %>% 
+  group_by(ladcd) %>% 
+  #mutate(diff_1st = -diff(rank))
+  mutate(diff_1st = -(rank-lag(rank))) %>%
+
+  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .1, na.rm = T)) 
+  # q.95 = 253, q.9 = 214, q.1 = 11, q.05 = 3.75. **abs()**
+
+  mutate(alpha = ifelse(abs(diff_1st) > 253, .8, 0.1)) %>% 
+  #mutate(alpha = ifelse(abs(diff_1st) <= 3.75, .8, 0.1)) %>% 
+  mutate(alpha = ifelse(is.na(diff_1st), 0.1, alpha)) %>%
+  mutate(label = ifelse(alpha==.1, "", ladnm)) %>%
+
+  ggplot(aes(year, rank, group = ladcd, alpha=alpha)) +
+  geom_bump(colour = "red") +
+  scale_y_reverse(labels = c(1, 100, 200, 300, 400)) +
+  # geom_text(#data = df %>% filter(year == 2012),
+  #           aes(x = Inf, label = label), size = 4, hjust = 0)  # x = 2012 + .1
+  geom_text_repel(aes(label = label),
+                  na.rm = TRUE, 
+                  size = 2,
+                  nudge_x = 2, nudge_y = 1,
+                  direction = "y", hjust = "left") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position="none") +
+  scale_x_continuous("year", 
+                     #limits = c(1998, 2010), 
+                     breaks=c(1998, 2002, 2006, 2010),
+                     labels = c("1996-98","2002", "2006", "2010-12"))
+ggsave(out.path)
+```
+
+## Not used
+
+### Ranks websites 2000 and 2012 only
 
@@
 
-## Websites per year
+### Websites per year
 
@@
 
-## Ranks websites per firm all period
+### Ranks websites per firm all period
 

### 2024-06-18 paper/submission1/tranos2023.qmd
postpresentation

@@
 library(knitr)
+library(kableExtra)
 
@@
 innovations diffuse across space and time and, importantly, how such
-spatio-temporal processes can be modelled. After all, diffusion together
+spatio-temporal processes can be modeled. After all, diffusion together
 with invention and innovation are considered the pillars of
@@
 the study period captures the very early stages of the diffusion on a
-new technology that is the web. This is key point in the lifecycle of a
+new technology that is the web as well as its maturity. 
+The former is a key point in the lifecycle of a
 new technology. Firstly, during this period new technologies are
@@
 A simple comparison between Web 1.0 and Web 3.0 applications clearly
-illustrates this [@tranos2020social]. During this period the performance
+illustrates this [@tranos2020social]; for instance a static website compared 
+with a platform like `github`, which enables cooperation between users and 
+the creation of new information, meaning, and knowledge 
+[@faraj2016special; @barassi2012does]. 
+During this period the performance
 of a new technology is the main attraction and not the cost to access
-and use it [@wilson2011lessons] There is a broader theoretical
+and use it [@wilson2011lessons]. There is a broader theoretical
 discussion in the literature regarding the motivation behind early
@@
 Importantly, the data used here depicts the *active* engagement with the
-web in the UK as it contains all the commercial websites archived by the
-Internet Archive [^2] that include a mention to at least one valid UK
-postcode in the web text. Websites are identified as commercial ones if
-they are part of the .co.uk second level domain (SLD). Also, the act of
-of creating a website is understood here as active engagement with the
+web in the UK as it contains all commercial websites that 
+(i) are part of the UK's relevant second level domain (SLD, .co.uk), 
+(ii) have been archived by the Internet Archive[^2], and 
+(iii) include a mention to at least one valid UK
+postcode in the web text. 
+The act of creating a website is understood here as active engagement with the
 web vis-à-vis the more passive act of browsing the web or having an
@@
 usually highly agglomerated regions is where new technologies are
-invented and commercially deployed [@grubler1990rise; @grubler1990rise].
+invented and commercially deployed [@grubler1990rise].
 This is where the first adopters tend to be based. Then, technologies
@@
 
-**ADD DATA
-DESCRIPTION, WHY .CO.UK, FAMILY OF WEB TECHNOLOGIES**
+**ADD FAMILY OF WEB TECHNOLOGIES**
+
+To capture the diffusion of web technologies, a website density metric is developed 
+for two different geographical scales: the Local Authority Districts (LAD) and the 
+Output Areas (OA). The former is an administrative unit and there are 
+\\textcolor{red}c. 380\\textcolor{black} such units in the UK. The latter is a census-based
+geographical unit, which is very small as there are c. 230,000 of them in the UK.
+This methodological choice will allow the mapping of the diffusion 
+of web technologies and the assessment of the diffusion mechanisms 
+at these two very different spatial scales.
+
+The population of websites is calculated using data from the 
+Internet Archive [^2] and, specifically, the JISC UK Web Domain Dataset [@ukwebarchive]. 
+The Internet Archive is one of the most complete and oldest archive of webpages 
+in the world operating since 1996 [@ainsworth2011much; @holzmann2016dawn].
+It is a web crawler, which discovers webpages by following the hyperlinks
+of every webpage its archives.
+This dataset, which is curated by the British Library, contains all
+the archived webpages from the UK ccTLD (.uk) from the 1996–2012 period. 
+In essence, this is a long list of 2.5 billion URLs of archived webpages including 
+also the archival timestamp.
 
@@
 
-<!-- from our RS paper -->
+Instead of using the whole .uk country code top-level domain (ccTLD), this 
+paper focuses on its commercial subset, the .co.uk SLD. 
+This choice decrease the heterogeneity of the web data as 
+such commercial websites have rather specific aims: they are used to diffuse information,
+support online transactions and share opinions [@THELWALL2000441; @blazquez2018big]. 
+Although a UK company can adopt a generic TLD such as.com and such cases escape 
+the data used here, such omissions should not affect our results given the
+popularity of the .uk ccTLD [@tranosuk]: UK consumers prefer to visit a .uk website
+when they are searching for products or services [@hope]; and
+anecdotal evidence indicates that during the first half of 2000, three
+.co.uk domains were registered every minute [@oecd_coms]. 
+Importantly, previous studies illustrated that .co.uk is the most popular UK SLD
+[@tranosuk].
 
-For instance, we can approximate digital economic activities in a region
-by measuring the volume of commercial webpages by considering only
-.co.uk websites, which are dedicated to commercial activities (Thelwall,
-2000). Such commercial websites are used to exchange information,
-support online transactions and share opinions (Blazquez & Domenech,
-2018). Although nothing prevents a UK-based company from adopting a
-generic TLD such as.com and, indeed, such cases escape our data, we do
-not expect that such omissions could affect our results given the
-popularity of the .uk ccTLD: UK consumers prefer to visit a .uk website
-when they are searching for products or services (Hope, 2017); and
-anecdotal evidence indicates that during the first half of 2000, three
-.co.uk domains were registered every minute (OECD, 2001). Also, as Table
-1 demonstrates, .co.uk was the most popular SLD under the .uk ccTLD in
-2000. Hence, we use the total volume of all the archived webpages in a
-region as a proxy for the level of digitization. We start the analysis
-by focusing on commercial webpages (.co.uk).
+The text from these webpages was scanned using regular expressions (regex) to
+identify strings of text which resemble UK postcodes and one fifth of them included
+a mention to a postcode [@BL2013geo]. This information allows the geolocation of the 
+data and the creation of the LAD and OA counts.
+
+Before that, the data 
+cleaning process included a data aggregation step, through which
+the webpages were aggregated to the relevant websites. 
+Based on the following example, all three webpages listed below are part of the 
+same website ([http://www.website.co.uk](http://www.website.co.uk)) and at the end 
+only websites and not the nested webpages were considered as otherwise the metric 
+would have been biased towards large websites.
+
+  - [http://www.website.co.uk/webpage_a](http://www.website.co.uk/webpage_a) B152TT
+
+  - [http://www.website.co.uk/webpage_b](http://www.website.co.uk/webpage_b) BS81TH
+
+  - [http://www.website.co.uk/webpage_c](http://www.website.co.uk/webpage_c) B152TT
+
+\noindent What is challenging is that this aggregation approach,
+which has been used elsewhere [@tranosuk; @shoreditch] may lead to websites with 
+multiple postcodes. As per the above example, ([http://www.website.co.uk](http://www.website.co.uk))
+includes two unique postcodes: B152TT and BS81SS. 
+The distribution of postcodes for 2000 is presented in Table \ref{f2000}, which 
+clearly illustrates the wide range. On one hand, we
+have websites anchored to a unique location (72% of all
+the reconstructed websites), which may represent
+a small company with a single trading location. On the other hand,
+we have websites with thousands of different postcodes.
+Considering the timeperiod of the analysis, such cases can represent directories
+which used to be popular in the pre-search engines early times of internet [@tranosuk].[^3]
+
+The analysis presented here is based on two subsets of these data.
+Firstly, the analysis is based on websites which only contain only one unique postcode
+and, therefore, the geolocation process might suffer from less noise. 
+As a robustness check, the analysis is replicated the analysis for an extended subset 
+of websites, which include up to ten unique postcodes. These websites are geolocated 
+by equally attaching them to multiple locations. This extended sample includes 
+94 percent of all the archived websites in 2000.
+
+[^3]: See Figure A1 in **Appendix A in the supplemental data online for examples**.
+
+```{r eval=T, echo=FALSE, results='asis'}
+
+# load 2000 co.uk
+# path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"
+# path.2000 <- "/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/all2000couk.csv"
+path.2000 <- paste0(path, "/data/temp/all2000couk.csv")
+
+all2000.duplicates <- data.table::fread(path.2000) #
+# dim(all2000.duplicates) # 3336162
+# it includes duplicated URLs: if one webpages includes multiple postcodes
+# then it appears multiple times. This is ok for the nuts aggregation, but
+# not for the frequencies
+# This is only .co.uk
+
+# one line for every host
+all2000 <- unique(all2000.duplicates, by = c("host")) # 57897
+
+# unique postcodes per website f table
+f.websites.pc <- DescTools::Freq(all2000$V1, breaks = c(0, 1,2, 10,100,1000,10000,100000), ord = "desc")
+f.websites.pc$level <- factor(f.websites.pc$level, levels = c("[0,1]","(1,2]", "(2,10]", "(10,100]",
+                                                              "(100,1e+03]", "(1e+03,1e+04]",
+                                                              "(1e+04,1e+05]"))
+levels(f.websites.pc$level) <- c("(0,1]","(1,2]", "(2,10]", "(10,100]", "(100,1000]", "(1000,10000]", "(10000,100000]")
+
+f.websites.pc <- f.websites.pc %>% 
+  dplyr::select(-cumfreq) %>% 
+  rename(Postcodes = level,
+         F = freq,
+         'F (%)' = perc,
+         'Cummulative F' = cumperc)
+
+kable(f.websites.pc,
+      format = "latex",
+      digits = 3,
+      booktabs = T,
+      format.args = list(big.mark = ","),
+      caption = "Number of unique postcodes per .co.uk website, 2000.\\label{f2000}") %>%
+  kableExtra::footnote(general = "Tranos et al. 2021",
+                       general_title = "Source: ",
+                       footnote_as_chunk = T)  
+   # kable_classic(full_width = F, html_font = "Cambria")
+```
+
+To create the web density metric, the yearly website counts at the LAD level are
+standardised by the number of firms in LAD to avoid biases associated with LADs 
+hosting a large number of firms. Given that there is no such statistic for the OA,
+the actual OA level counts are used. However, because of the the rather
+consistent spatial definition of OA (they host 40-250 households),[^4] 
+website counts in OA are interpreted as a density metric too.
+
+[^4]: [https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies](https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies)
+
+This website metric...
 
@@
 
-# Results {#sec3}
-
-## Old ideas
-
-Hexagon density maps: reflect the spatial structure of Britain. Websites
-are associated in places where people live and work.
-
-**TODO** maps at the local authority level per firms and discuss
-patterns
-
-Neighbourhood effect: diffusion proceeds outwards from innovation
-centers, first "hitting" nearby rather than far-away locations (Grubler
-1990)
-
--   Moran's I: for OA and LAD over time
-
--   LISA maps: for OA and LAD over time More and less expected clusters.
-    Different scales show different results
-
--   Website density regressions: for OA and LAD over time **TODO** add
-    0s Similar pattern
-
-Hierarchy effect: from main centers to secondary ones -- central places
-
--   Gini coefficient. Almost perfect polarisation of web adoption in the
-    early stages at a granular level More equally diffused at the Local
-    Authority level Plateau overtime
-
-S-shaped diffusion curves
-
--   S for LADs per firm and OA.
-
--   Fast and slow LAs map. There is clustering
-
--   s_uk_firm_11
-
--   ranks: there is stability and movement
-
-## New structure
+# Results {#sec4}
 
@@
     Different scales show different results
-
--   ~~Website density regressions: for OA and LAD over time~~
 
@@
 scale, the later highlights that the importance of past web adoption rates in large 
-population centres is surpassed by proximity to them and soatial configuration 
+population centres is surpassed by proximity to them and spatial configuration 
 at this scale.

### 2024-06-18 paper/submission1/tranos2023.tex
postpresentation

@@
 
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \makeatletter
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
 \@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
 \makeatother
@@
 \bibliographystyle{elsarticle-harv}
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \hypersetup{
-  pdftitle={Digital economy in the UK: a multi-scalar story of the diffusion of web technologies},
+  pdftitle={A multi-scale story of the diffusion of a new technology: the web},
   pdfauthor={Emmanouil Tranos},
@@
 \begin{frontmatter}
-\title{Digital economy in the UK: a multi-scalar story of the diffusion
-of web technologies}
+\title{A multi-scale story of the diffusion of a new technology: the
+web}
 \author[1]{Emmanouil Tranos%
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 
-\hypertarget{sec1}{%
-\section{Introduction}\label{sec1}}
+\section{Introduction}\label{sec1}
 
@@
 geographical studies exploring the diffusion of new and, more
-specifically for this paper, digital technologies across \emph{both}
-space and time can be attributed to the scarcity of relevant and
-granular enough data. As \citet{zook2022mapping} highlight, digital
-activities are hardly ever captured in official data.
+specifically for this paper, digital technologies across both space and
+time can be attributed to the scarcity of relevant and granular enough
+data. As \citet{zook2022mapping} highlight, digital activities are
+hardly ever captured in official data.
 
@@
 diffusion.
+
+\textbf{SAY SOMETHING ABOUT SPATIAL ANALYTICS, DIST PAPER BY ARP, NOT
+SPECIFIC EXPLANATORY VARIABLES, INTANGIBLE}
 
@@
 reminded that it was only in 1994 when Netscape Navigator was
-introduced, a year before Microsoft's Internet Explorer \footnote{\url{https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft}}.
+introduced, a year before Microsoft's Internet Explorer.\footnote{\url{https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft}}
 Also, only \(9\) per cent of UK's households had access to the internet
@@
 it is fair to say that the study period captures the very early stages
-of the diffusion on a new technology that is the web. This is key point
-in the lifecycle of a new technology. Firstly, during this period new
-technologies are expensive, crude and imperfect
-\citep{rosenberg1994exploring, wilson201281}. A simple comparison
-between Web 1.0 and Web 3.0 applications clearly illustrates this
-\citep{tranos2020social}. During this period the performance of a new
-technology is the main attraction and not the cost to access and use it
-\citep{wilson2011lessons} There is a broader theoretical discussion in
-the literature regarding the motivation behind early adoption. As
-summarised by \citet{perkins2005international}, on the one hand,
-epidemic models highlight the role of interpersonal contacts as a way
-for new technologies to diffuse. On the other hand, economic models
-underline the importance of heterogeneity. Different firms have
-different structures and business plans, which define the potential
-economic returns of the adoption of a new technology and, therefore, the
-choice to adopt a new technology becomes an individual option. From a
-broader and evolutionary perspective, initial conditions are essential
-for the creation and evolution of path-dependent technological
-development trajectories \citep{neffke2011regions, simmie2014new}. This
-argument is even more relevant when the focus is on Information and
-Communication Technologies (ICT) because of the commonly found lag
-between investment and economic returns as reflected in the Solow
-paradox \citep{acemoglu2014return, brynjolfsson2018artificial}.
+of the diffusion on a new technology that is the web as well as its
+maturity. The former is a key point in the lifecycle of a new
+technology. Firstly, during this period new technologies are expensive,
+crude and imperfect \citep{rosenberg1994exploring, wilson201281}. A
+simple comparison between Web 1.0 and Web 3.0 applications clearly
+illustrates this \citep{tranos2020social}; for instance a static website
+compared with a platform like \texttt{github}, which enables cooperation
+between users and the creation of new information, meaning, and
+knowledge \citep{faraj2016special, barassi2012does}. During this period
+the performance of a new technology is the main attraction and not the
+cost to access and use it \citep{wilson2011lessons}. There is a broader
+theoretical discussion in the literature regarding the motivation behind
+early adoption. As summarised by \citet{perkins2005international}, on
+the one hand, epidemic models highlight the role of interpersonal
+contacts as a way for new technologies to diffuse. On the other hand,
+economic models underline the importance of heterogeneity. Different
+firms have different structures and business plans, which define the
+potential economic returns of the adoption of a new technology and,
+therefore, the choice to adopt a new technology becomes an individual
+option. From a broader and evolutionary perspective, initial conditions
+are essential for the creation and evolution of path-dependent
+technological development trajectories
+\citep{neffke2011regions, simmie2014new}. This argument is even more
+relevant when the focus is on digital technologies because of the
+commonly found lag between investment and economic returns as reflected
+in the Solow paradox
+\citep{acemoglu2014return, brynjolfsson2018artificial}.
 
 Importantly, the data used here depicts the \emph{active} engagement
-with the web in the UK as it contains all the commercial websites
-archived by the Internet Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
-that include a mention to at least one valid UK postcode in the web
-text. Websites are identified as commercial ones if they are part of the
-.co.uk second level domain (SLD). Also, the act of of creating a website
-is understood here as active engagement with the web vis-à-vis the more
-passive act of browsing the web or having an internet connection
-\citep{tranosuk}. Previous studies have focused mostly on more passive
-notions of engaging with digital technologies such as internet adoption
-and internet speeds
+with the web in the UK as it contains all commercial websites that (i)
+are part of the UK's relevant second level domain (SLD, .co.uk), (ii)
+have been archived by the Internet Archive\footnote{\href{See\%20https://archive.org/}{https://archive.org/}.},
+and (iii) include a mention to at least one valid UK postcode in the web
+text. The act of creating a website is understood here as active
+engagement with the web vis-à-vis the more passive act of browsing the
+web or having an internet connection \citep{tranosuk}. Previous studies
+have focused mostly on more passive notions of engaging with digital
+technologies such as internet adoption and internet speeds
 \citep[e.g.][]{blank2018local, riddlesden2014broadband}. More details
 about the data and the data generation process can be found in Section
-\protect\hyperlink{sec3}{XX}. \textbf{ADD DATA DESCRIPTION, WHY .CO.UK}
+\hyperref[sec3]{XX}.
 
@@
 
-\hypertarget{sec2}{%
-\section{Literature review}\label{sec2}}
+\section{Literature review}\label{sec2}
 
@@
 groups of individuals regarding their adoption speed: innovators, early
-adopters, early majority, late majority and laggards. Key mechanism of
-diffusion and adoption is always communication and how knowledge is
+adopters, early majority, late majority and laggards. The key mechanism
+of diffusion and adoption is communication and how knowledge is
 transferred within a social system. Therefore, all approaches agree on
@@
 usually highly agglomerated regions is where new technologies are
-invented and commercially deployed {[}\citet{grubler1990rise};
-grubler1990rise{]}. This is where the first adopters tend to be based.
-Then, technologies spread to the \emph{rim} and eventually to the
-\emph{periphery}. Although adoption pace might be higher when new
-technologies finally arrive to the periphery, the saturation levels
-there may never reach the ones in the core because of the lack of
-infrastructure or other necessary institutions
-\citep{leibowicz2016representing}. \citet{grubler1990rise} effectively
-summarises the three key characteristics of the spatial diffusion
-process: (i) the cumulative level of adoption follows an S-shaped
-pattern just like purely temporal models; (ii) diffusion is shaped by a
-hierarchy effect in a form of a centrifugal force: from core to
-periphery; and (iii) diffusion is also shaped by distance and a
-neighbourhood effect and contaminate nearby locations.
+invented and commercially deployed \citep{grubler1990rise}. This is
+where the first adopters tend to be based. Then, technologies spread to
+the \emph{rim} and eventually to the \emph{periphery}. Although adoption
+pace might be higher when new technologies finally arrive to the
+periphery, the saturation levels there may never reach the ones in the
+core because of the lack of infrastructure or other necessary
+institutions \citep{leibowicz2016representing}. \citet{grubler1990rise}
+effectively summarises the three key characteristics of the spatial
+diffusion process: (i) the cumulative level of adoption follows an
+S-shaped pattern just like purely temporal models; (ii) diffusion is
+shaped by a hierarchy effect in a form of a centrifugal force: from core
+to periphery; and (iii) diffusion is also shaped by distance and a
+neighbourhood effect and contaminate nearby locations. These are the
+three mechanisms that the empirical analysis in Section
+\phantomsection\label{sec4}{4} investigates.
 
@@
 spatial dimension is present in most of the following studies, the level
-of spatial detail is almost always more coarse than the one adopted
-here. \citet{beardsell1999spatial} studied the evolution of the computer
+of spatial detail is always more coarse than the one adopted here.
+\citet{beardsell1999spatial} studied the evolution of the computer
 industry in \(317\) US metro areas during the \(1977-1992\) period using
@@
 distribution holds for urban computer employment and also urban
-heterogeneity is essential in explaining this distribution. in a recent
+heterogeneity is essential in explaining this distribution. In a recent
 study, \citet{bednarz2020pulled} focused on wind turbines and modelled
@@
 and developing countries experience diffusion of new technologies more
-rapidly than early adopter and developed countries. At the same scale,
+rapidly than early adopters and developed countries. At the same scale,
 \citet{perkins2011internet} explored whether the adoption of previous
@@
 
-\hypertarget{sec3}{%
-\section{Materials and Methods}\label{sec3}}
+All in all, \ldots{}
 
-For instance, we can approximate digital economic activities in a region
-by measuring the volume of commercial webpages by considering only
-.co.uk websites, which are dedicated to commercial activities (Thelwall,
-2000). Such commercial websites are used to exchange information,
-support online transactions and share opinions (Blazquez \& Domenech,
-2018). Although nothing prevents a UK-based company from adopting a
-generic TLD such as.com and, indeed, such cases escape our data, we do
-not expect that such omissions could affect our results given the
-popularity of the .uk ccTLD: UK consumers prefer to visit a .uk website
-when they are searching for products or services (Hope, 2017); and
-anecdotal evidence indicates that during the first half of 2000, three
-.co.uk domains were registered every minute (OECD, 2001). Also, as Table
-1 demonstrates, .co.uk was the most popular SLD under the .uk ccTLD in
-2000. Hence, we use the total volume of all the archived webpages in a
-region as a proxy for the level of digitization. We start the analysis
-by focusing on commercial webpages (.co.uk).
+\section{Materials and Methods}\label{sec3}
+
+\textbf{ADD FAMILY OF WEB TECHNOLOGIES}
+
+To capture the diffusion of web technologies, a website density metric
+is developed for two different geographical scales: the Local Authority
+Districts (LAD) and the Output Areas (OA). The former is an
+administrative unit and there are
+\textbackslash textcolor\{red\}c.~380\textbackslash textcolor\{black\}
+such units in the UK. The latter is a census-based geographical unit,
+which is very small as there are c.~230,000 of them in the UK. This
+methodological choice will allow the mapping of the diffusion of web
+technologies and the assessment of the diffusion mechanisms at these two
+very different spatial scales.
+
+The population of websites is calculated using data from the Internet
+Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
+and, specifically, the JISC UK Web Domain Dataset \citep{ukwebarchive}.
+The Internet Archive is one of the most complete and oldest archive of
+webpages in the world operating since 1996
+\citep{ainsworth2011much, holzmann2016dawn}. It is a web crawler, which
+discovers webpages by following the hyperlinks of every webpage its
+archives. This dataset, which is curated by the British Library,
+contains all the archived webpages from the UK ccTLD (.uk) from the
+1996--2012 period. In essence, this is a long list of 2.5 billion URLs
+of archived webpages including also the archival timestamp.
+
+Instead of using the whole .uk country code top-level domain (ccTLD),
+this paper focuses on its commercial subset, the .co.uk SLD. This choice
+decrease the heterogeneity of the web data as such commercial websites
+have rather specific aims: they are used to diffuse information, support
+online transactions and share opinions
+\citep{THELWALL2000441, blazquez2018big}. Although a UK company can
+adopt a generic TLD such as.com and such cases escape the data used
+here, such omissions should not affect our results given the popularity
+of the .uk ccTLD \citep{tranosuk}: UK consumers prefer to visit a .uk
+website when they are searching for products or services \citep{hope};
+and anecdotal evidence indicates that during the first half of 2000,
+three .co.uk domains were registered every minute \citep{oecd_coms}.
+Importantly, previous studies illustrated that .co.uk is the most
+popular UK SLD \citep{tranosuk}.
+
+The text from these webpages was scanned using regular expressions
+(regex) to identify strings of text which resemble UK postcodes and one
+fifth of them included a mention to a postcode \citep{BL2013geo}. This
+information allows the geolocation of the data and the creation of the
+LAD and OA counts.
+
+Before that, the data cleaning process included a data aggregation step,
+through which the webpages were aggregated to the relevant websites.
+Based on the following example, all three webpages listed below are part
+of the same website (\url{http://www.website.co.uk}) and at the end only
+websites and not the nested webpages were considered as otherwise the
+metric would have been biased towards large websites.
+
+\begin{itemize}
+\item
+  \url{http://www.website.co.uk/webpage_a} B152TT
+\item
+  \url{http://www.website.co.uk/webpage_b} BS81TH
+\item
+  \url{http://www.website.co.uk/webpage_c} B152TT
+\end{itemize}
+
+\noindent What is challenging is that this aggregation approach, which
+has been used elsewhere \citep{tranosuk, shoreditch} may lead to
+websites with multiple postcodes. As per the above example,
+(\url{http://www.website.co.uk}) includes two unique postcodes: B152TT
+and BS81SS. The distribution of postcodes for 2000 is presented in Table
+\ref{f2000}, which clearly illustrates the wide range. On one hand, we
+have websites anchored to a unique location (72\% of all the
+reconstructed websites), which may represent a small company with a
+single trading location. On the other hand, we have websites with
+thousands of different postcodes. Considering the timeperiod of the
+analysis, such cases can represent directories which used to be popular
+in the pre-search engines early times of internet
+\citep{tranosuk}.\footnote{See Figure A1 in \textbf{Appendix A in the
+  supplemental data online for examples}.}
+
+The analysis presented here is based on two subsets of these data.
+Firstly, the analysis is based on websites which only contain only one
+unique postcode and, therefore, the geolocation process might suffer
+from less noise. As a robustness check, the analysis is replicated the
+analysis for an extended subset of websites, which include up to ten
+unique postcodes. These websites are geolocated by equally attaching
+them to multiple locations. This extended sample includes 94 percent of
+all the archived websites in 2000.
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-1}Number of unique postcodes per .co.uk website, 2000.\label{f2000}}
+\centering
+\begin{tabular}[t]{lrrr}
+\toprule
+Postcodes & F & F (\%) & Cummulative F\\
+\midrule
+(0,1] & 41,596 & 0.718 & 0.718\\
+(1,2] & 6,451 & 0.111 & 0.830\\
+(2,10] & 6,163 & 0.106 & 0.936\\
+(10,100] & 2,975 & 0.051 & 0.988\\
+(100,1000] & 646 & 0.011 & 0.999\\
+\addlinespace
+(1000,10000] & 62 & 0.001 & 1.000\\
+(10000,100000] & 4 & 0.000 & 1.000\\
+\bottomrule
+\multicolumn{4}{l}{\rule{0pt}{1em}\textit{Source: } Tranos et al. 2021}\\
+\end{tabular}
+\end{table}
+
+To create the web density metric, the yearly website counts at the LAD
+level are standardised by the number of firms in LAD to avoid biases
+associated with LADs hosting a large number of firms. Given that there
+is no such statistic for the OA, the actual OA level counts are used.
+However, because of the the rather consistent spatial definition of OA
+(they host 40-250 households),\footnote{\url{https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies}}
+website counts in OA are interpreted as a density metric too.
+
+This website metric\ldots{}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{old-ideas}{%
-\subsection{Old ideas}\label{old-ideas}}
-
-Hexagon density maps: reflect the spatial structure of Britain. Websites
-are associated in places where people live and work.
-
-\textbf{TODO} maps at the local authority level per firms and discuss
-patterns
-
-Neighbourhood effect: diffusion proceeds outwards from innovation
-centers, first ``hitting'' nearby rather than far-away locations
-(Grubler 1990)
-
-\begin{itemize}
-\item
-  Moran's I: for OA and LAD over time
-\item
-  LISA maps: for OA and LAD over time More and less expected clusters.
-  Different scales show different results
-\item
-  Website density regressions: for OA and LAD over time \textbf{TODO}
-  add 0s Similar pattern
-\end{itemize}
-
-Hierarchy effect: from main centers to secondary ones -- central places
-
-\begin{itemize}
-\tightlist
-\item
-  Gini coefficient. Almost perfect polarisation of web adoption in the
-  early stages at a granular level More equally diffused at the Local
-  Authority level Plateau overtime
-\end{itemize}
-
-S-shaped diffusion curves
-
-\begin{itemize}
-\item
-  S for LADs per firm and OA.
-\item
-  Fast and slow LAs map. There is clustering
-\item
-  s\_uk\_firm\_11
-\item
-  ranks: there is stability and movement
-\end{itemize}
-
-\hypertarget{new-structure}{%
-\subsection{New structure}\label{new-structure}}
+\section{Results}\label{sec4}
 
@@
   centers, first ``hitting'' nearby rather than far-away locations
-  (Grubler
-\end{enumerate}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\setcounter{enumi}{1989}
-\tightlist
-\item
+  (Grubler 1990)
 \end{enumerate}
@@
   Different scales show different results
-\item
-  Website density regressions: for OA and LAD over time
 \end{itemize}
@@
 
-\hypertarget{rf-results}{%
-\subsubsection{RF results}\label{rf-results}}
+\subsubsection{RF results}\label{rf-results}
 
-The next section incorporates the above discussed spatial drivers of the
-diffusion of web technologies into the same modelling framework. The aim
-is to use variables depicting these spatial processes in order to
+The next section incorporates the above discussed spatial processes of
+the diffusion of web technologies into the same modelling framework. The
+aim is to use variables depicting these spatial processes in order to
 predict the diffusion of web technologies in the UK over space and time
@@
 the importance of such variables changes across scales. The next two
-models will be again trained on web difusion at the two working scales:
-OA and LAD. However, instead of using all the data points, the OA and
-the LAD from one of the 12 UK regions are hold out and then the trained
-model is used to predict web adoption in the OA or the LAD of the
-held-out region. This process takes place recursively for all UK
+sets of models will be again trained on web diffusion at the two working
+scales: OA and LAD. However, instead of using all the data points, the
+OA and the LAD from one of the twelve UK regions are held out and then
+the trained model is used to predict web adoption in the OA or the LAD
+of the held-out region. This process takes place recursively for all UK
 regions. The difference in the predictive capacity of the different
-samples will reveal how different are these spatial process across
+samples will reveal how dissimilar are these spatial process across
 regions and, importantly, at different scales.
@@
 It needs to be highlighted here that the cross-validation for all models
-presented here is spatially and temporally sensitive. Instead of using
-10 random folds, we employ the \texttt{CAST} package which allows holds
-back data points from specific years and spatial units and use them for
-testing in order to estimate the model performance
-\citep{meyer2018improving}.
+is spatially and temporally sensitive. Instead of using 10 random folds,
+we employ the \texttt{CAST} package which allows holding back data
+points from specific years and spatial units and use them for testing in
+order to estimate the model performance \citep{meyer2018improving}.
 
@@
 that the relevant literature and the descriptive analysis presented in
-the previous section highlighted. Namely, the models capture: (i) a
+the Section \textbf{XX} highlighted. Namely, the models capture: (i) a
 hierarchy effect with diffusion running from main centres to secondary
 ones, (ii) a neighborhood effect according to which diffusion first hits
-nearby locations, and (iii) the rather canonical pattern of the
-diffusion over time as reflected in the S-shaped pattern in the
-cumulative level of adoption.
+nearby locations, and (iii) the rather canonical pattern of diffusion
+over time as reflected in the S-shaped pattern in the cumulative level
+of adoption.
 
@@
 the neighbourhood effect, the web density of the neighbouring spatial
-units in the previous year is calculated. Again, the underpinning
+units in the previous year is employed. Again, the underpinning
 rationale is that the level of web adoption within a spatial unit
@@
 S curve for the cumulative adoption is captured by trend variable.
-Hence, all four model will follow the following generic form
-\ref{model}:
+Hence, all four model will follow the following generic form (Eq.
+\ref{model}):
 
@@
 and testing via cross validation. The first one is trained and tested on
-374 LAD and the second on 232,296 OA for a 16 year period (1997-2012).
-The results are remarkably good considering that the are the outcome of
-space and time sensitive CV, so the the model does not suffer from
-overfit. At the LAD level we the model predicts 81\% of the variation of
-website density. Both error metrics indicate that the model error is
-\((\frac{1}{20} - \frac{1}{30})\) of a website per firm. At the OA this
-score drops down to 21\%. Considering its granularity, this is still a
-remarkable performance. To contextualise it, the model results in a MAE
-of one website for areas small enough to host less than 140
+374 LAD and the second on 232,296 OA, both for a 16 year period
+(1997-2012). The results are remarkably good considering that the are
+the outcome of space and time sensitive CV, so the the model does not
+suffer from overfitting. At the LAD level the model predicts 81\% of the
+variation of website density. Both error metrics indicate that the model
+error is \((\frac{1}{20}, \frac{1}{30})\) of a website per firm. At the
+OA the RSquared drops down to 21\%. Considering its granularity, this is
+still a remarkable performance. To contextualise it, the model results
+in a MAE of one website for areas small enough to host less than 140
 households.\footnote{According to the Office for National Statistics,
@@
 UK OA were not associated with a website. Again, given the granularity
-of the data this should not come as a surprise. Despite the large
-heterogeneity in website adoption, the model reached an R-Squared of
-0.2.
+of the data this should not come as a surprise.
 
@@
 Figures \ref{var.imp.LAD} and \ref{var.imp.OA} plot the importance of
-the difrerent predictors. When the focus is on the LAD, the website
-density in the nearest city, in London and on the neighbouring LADs the
+the different predictors. When the focus is on the LAD, the website
+density in the nearest city, in London and in the neighbouring LADs the
 year before are the most important predictors. They are followed by the
-yearly trend, while the spatial configuration as reflected on distances
+yearly trend, while the spatial configuration as reflected in distances
 to London or the nearest city only play a minor role. This can be
 attributed to the rather coarse spatial scale of analysis. Nevertheless,
-all previously discussed forces are at play in the diffusion of web
-technologies at the LAD: the first two predictors depict the
-hierarchical effect, the spatial and temporal lag of website density the
-neighbourhood effect and the yearly trend the time-sensitive cumulative
-adoption pattern.
+all previously discussed spatial processes are at play in the diffusion
+of web technologies at the LAD level: the first two predictors depict
+the hierarchical effect, the spatial and temporal lag of website density
+depict the neighbourhood effect and the yearly trend the time-sensitive
+cumulative adoption pattern.
 
@@
 importance of past web adoption rates in large population centres is
-surpassed by proximity to them and soatial configuration at this scale.
+surpassed by proximity to them and spatial configuration at this scale.
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 
-\end{figure}
+\end{figure}%
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 
-\end{figure}
+\end{figure}%
 
 Table \ref{table.regions} presents the results of the recursive hold out
-models which aim to highlight potential regional heteorigeinty of the
-web diffusion spatial mechanisms. To begin with, as highlighted before,
-there is a difference of magnitude of one order between the LAD and the
-OA prediction errors, which is aligned with previous results that
-employed all data points. What is of interest here is the regional
-comparison. Table \ref{table.regions} illustrates some striking
-similarities, but also a few significant differences. The regions the
-web diffusion of which is better predicted using models trained in the
-rest of the country are the same despite the scale of analysis: South
-East, Wales, Yorkshire and The Humber and the North East of England. In
-other words, these are the regions whose spatial diffusion of web
-technologies is closer to the country's average. Despite the consistency
-across scales, this is a diverse set of regions: \textbf{ADD
-CHARACTERISTICS}.
+models, which aim to highlight the potential regional heterogeneity of
+the spatial processes behind the diffusion of web technologies. To begin
+with, as highlighted before, there is a difference of magnitude of one
+order between the LAD and the OA prediction errors, which is aligned
+with previous results that employed all data points. What is of interest
+here is the regional comparison. Table \ref{table.regions} illustrates
+some striking similarities, but also a few significant differences. The
+regions the web diffusion of which is better predicted using models
+trained in the rest of the country are the same despite the scale of
+analysis: South East, Wales, Yorkshire and The Humber and the North East
+of England. In other words, these are the regions whose spatial
+diffusion mechanisms of web technologies is closer to the country's
+average. Despite the consistency across scales, this is a diverse set of
+regions: \textbf{ADD CHARACTERISTICS}.
 
@@
 diffusion mechanisms are consistently diverging from the country's
-average. This should not come as a surpise as these regions are
+average. This should not come as a surprise as these regions are
 characterised of high levels of rurality and remoteness. Similarly,
@@
 region appear to be closer to the country's average. At this scale,
-proximity or lack of in relative terms to the rest of the country become
-a less important and the internal to the region spatial structure
-predictors start playing a more important role \textbf{CHECK NI OA}.
+proximity, or lack of, relative to the rest of the country become less
+important and the internal to the region spatial structure predictors
+start playing a more important role \textbf{CHECK NI OA model}.
 
@@
 
-\hypertarget{discussion-and-conclusions}{%
-\section{Discussion and conclusions}\label{discussion-and-conclusions}}
+\section{Discussion and conclusions}\label{discussion-and-conclusions}
 

### 2024-06-18 paper/submission1/help.R

@@ -1,0 +1,35 @@
+## Old ideas
+
+Hexagon density maps: reflect the spatial structure of Britain. Websites
+are associated in places where people live and work.
+
+**TODO** maps at the local authority level per firms and discuss
+patterns
+
+Neighbourhood effect: diffusion proceeds outwards from innovation
+centers, first "hitting" nearby rather than far-away locations (Grubler
+                                                                1990)
+
+-   Moran's I: for OA and LAD over time
+
+-   LISA maps: for OA and LAD over time More and less expected clusters.
+    Different scales show different results
+
+-   Website density regressions: for OA and LAD over time **TODO** add
+    0s Similar pattern
+
+Hierarchy effect: from main centers to secondary ones -- central places
+
+-   Gini coefficient. Almost perfect polarisation of web adoption in the
+    early stages at a granular level More equally diffused at the Local
+    Authority level Plateau overtime
+
+S-shaped diffusion curves
+
+-   S for LADs per firm and OA.
+
+-   Fast and slow LAs map. There is clustering
+
+-   s_uk_firm_11
+
+-   ranks: there is stability and movement

### 2024-06-18 src/problem_years.Rmd
corrected OAs

@@
 
-```{r}
+```{r, eval=FALSE}
 
@@
 
-### Problem postcodes
+## Problem postcodes
 
@@
 
-### Correct 1996-2010
+## Correct 1996-2010
 
@@
 df9610 %>% filter(pc == "M28 2SL") %>% distinct(year)
+```
 
+## LA
+
+### Load remaining data
+
+```{r}
 df9610 <- as_tibble(df9610) %>% 
@@
   ungroup()
-```
 
-### Load remaining data
-
-```{r}
 data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
@@
 
-## Firms
+### Firms
 
@@
 
-## spatial data
+### spatial data
 
@@
 
-## lisa maps
+### lisa maps
 
 ```{r}
-
-path.out <- paste0(path, "/outputs/lisa/corrected/lisa_level_pc", n, "_la") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa/corrected/lisa_level_pc", m, "_la")  # for multiple PC
+path.out <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc", n, "_la") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc", m, "_la")  # for multiple PC
 
@@
 
-# Moran I
+### Moran I
 
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/corrected/morani_la.csv")
+path.out <- paste0(path, "/outputs/lisa/corrected/LA/morani_la.csv")
 morani %>% as_tibble() %>% 
@@
 
+## OA
+
+### Load remaining data
+
 ```{r}
-path.images <- paste0(path, "/outputs/lisa/corrected/")
+df9610 <- as_tibble(df9610) %>% 
+  mutate(year = as.double(as.character(year))) %>% 
+  left_join(lookup, by = "pc", suffix =c("","")) %>%
+  group_by(year, oa11cd) %>%
+  summarise(n = n()) %>%            
+  ungroup()
 
-list.files(path = path.images, pattern = "^lisa_level_pc1_la", full.names = T) %>% 
-  map(image_read) %>% # reads each path file
-  image_join() %>% # joins image
-  image_animate(delay=50) %>% # animates, can opt for number of loops
-  image_write("corrected_lisa.gif", format = "png") # write to current dir
+data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
+#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
+#This is what we use for the hyperlinks paper as per George's script
 
+df1112 <- read_csv(data.path2011_2012) %>% 
+  filter(
+    #V1.domain < m,             # for multiple pc
+    V1.domain == n,             # for n == 1
+    year > 1995 & year <2013) %>%   
+  left_join(lookup, by = "pc", suffix =c("","")) %>% 
+  group_by(year, oa11cd) %>%
+  summarise(n = n()) %>%             # for n == 1  
+  #summarise(n = sum(V1.domain)) %>% # for multiple pc  
+  ungroup()
+
+df <- bind_rows(df9610, df1112) %>% 
+  filter(!is.na(oa11cd)) %>% 
+  complete(oa11cd, year) %>% 
+  replace(is.na(.), 0)
+
+# tests for Scotland
+# df %>% filter(substr(oa11cd, 1,1) =="S",
+#               year==2010) %>% summarise(mean(n, na.rm = T))
+# 
+# df %>% filter(oa11cd=="S00090182")
+# 
+# test@data %>% filter(substr(id, 1,1) =="S",
+#                      n!=0)
+# 
+# df %>% filter(substr(oa11cd, 1,1) =="W",
+#               year==2003)
 ```
 
+### spatial data
+
 ```{r}
-# centers the local Moran's around the mean
-lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
-lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)
+# get OA for England and Wales
+path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
+oa.ew <- st_read(path.geo)
+# source: https://geoportal.statistics.gov.uk/
 
-# centers the variable of interest around its mean
+# spatial transformations
+oa.ew <- st_transform(oa.ew, 4326)  # EPSG code for WGS84
 
-lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
-lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)
+# keep in the data slot only the ONS Output Area id, renaming it as 'id'
+oa.ew <- oa.ew %>% 
+  dplyr::select(OA11CD, Shape__Area) %>% 
+  rename(id = OA11CD,
+         area = Shape__Area)
 
-signif <- 0.05
-#lmoran
+path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
+oa.sc <- st_read(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
+# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries
 
-lmoran <- lmoran%>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
-  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
-  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
-  mutate(quadrant= ifelse(lmoran$`Pr.z....E.Ii..` > signif, 0, quadrant)) #%>%
+# spatial transformations
+oa.sc <- st_transform(oa.sc, 4326)  # EPSG code for WGS84
 
-#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
-# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
-#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
-#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))
+# Scotland (follows same steps as EW, see notes above)
+oa.sc <- oa.sc %>% 
+  dplyr::select(code, SHAPE_1_Ar) %>% 
+  rename(id = code,
+         area = SHAPE_1_Ar)
 
-mun_merge_new<- merge(test, lmoran, by="LAD21CD")
-mun_merge_new<- test %>% 
-  left_join(lmoran, by="LAD21CD")
+# NI
+path.geo.ni <- paste0(path, "/data/raw/ni_small_area/")
+oa.ni <- st_read(dsn = path.geo.ni, layer = "SA2011")
+oa.ni <- st_transform(oa.ni, 4326)  # EPSG code for WGS84
+oa.ni <- oa.ni %>% dplyr::select(SA2011, Hectares) %>% 
+  rename(id= SA2011,
+         area = Hectares)
 
-# R p value map
-breaks = c(0, 1, 2, 3, 4, 5)
+# Build OA for UK
+oa.uk <- rbind(oa.ew, oa.sc, oa.ni)
+rm(oa.ew, oa.sc, oa.ni)
+
+# get UK
+path.uk <- "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/Countries_December_2023_Boundaries_UK_BUC/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson"
+uk <- st_read(path.uk)
+# spatial transformations
+uk <- st_transform(uk, 4326)  # EPSG code for WGS84
+```
+
+### lisa maps
+
+```{r}
+path.out <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc", n, "_oa") # for unique PC
+#path.out <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc", m, "_oa")  # for multiple PC
+
+morani <- data.frame()
+
+#spatial weight matrix
+sf_use_s2(FALSE)
+nb <- poly2nb(oa.uk)
+listw <- nb2listw(nb, style = "W", zero.policy = T)
+
+for (t in 1996:2012){
+  test <- oa.uk
+  test <- test %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
+    mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
+           year = ifelse(is.na(year), t, year))
+
+  #spatial weight matrix
+  #moved out of the loop
+
+  #Moran's I
+  globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
+  a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
+  morani <- rbind(a, morani)
+
+  #LISA
+  lmoran <- localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude)
+
+  quadr <- attr(lmoran, "quadr")
+  #quadr[lmoran[,5] > 0.05 | is.na(lmoran[,5]), ] <- NA
+  quadr[p.adjust(lmoran[,5], method = "fdr") > 0.05 | is.na(lmoran[,5]), ] <- NA
+
+  palette_new <- grDevices::colorRampPalette(
+            colors = c("blue", "lightsalmon", "lightblue", "red", "white"))(5)
 
@@
     arrange(pop) %>% tail(10) 
-cities.sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
+  tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
 
-map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, 
-                                         palette = c("white","red","blue",
-                                                     rgb(0,0,1,alpha=0.4),
-                                                     rgb(1,0,0,alpha=0.4)), 
-                                         labels = c("Not significant",
-                                                    "High-High",
-                                                    "Low-Low",
-                                                    "Low-High",
-                                                    "High-Low"), 
-                                         title="LISA, websites per firm in a LA") +
-  tm_legend(text.size = 1) +
-  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
-  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
-  #tm_borders(alpha=.1) +
-  tm_borders(lwd = 0) +
-  tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
-  tm_shape(cities.sf) + tm_dots() + tm_text("name", size = .5) +
-  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
-tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
+  ggplot(data = oa.uk, aes(fill = quadr$median)) +
+    geom_sf(data = uk, fill = "white", linewidth = .5) +
+    geom_sf(linewidth = 0) +
+    geom_sf(data = tif_sf, fill = "black", size = .3) +
+    #geom_sf_text(data = tif_sf, aes(label = name, fill = "black"), size = 4) +
+    scale_fill_manual(values = palette_new, na.value="white", drop = F) +
+    theme_minimal() +
+    theme(panel.grid.major = element_line(colour = "transparent"),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          legend.title=element_blank()) +
+          #legend.position="none") +
+    guides(fill = guide_legend(reverse = TRUE)) +
+    labs(title = t,
+        subtitle = "LISA, websites per OA")
+
+  filename = paste(path.out, t, ".png", sep = "")
+  ggsave(filename, bg='#ffffff')
 }
+```
 
-#ggplot map solution.
-# mun_merge_new %>% st_as_sf(crs = 27700) %>% #st_transform(4326) %>%
-#   ggplot()+
-#   geom_sf(aes(fill = as.factor(quadrant)))
-# ggsave("test_ggplot.png", path = path)
+### Moran I
 
+```{r}
+path.out <- paste0(path, "/outputs/lisa/corrected/OA/morani_oa.csv")
+morani %>% as_tibble() %>% 
+  rename('morani'=V1,
+                   p = V2,
+                   year = V3) %>% 
+  kable() %>% 
+  write_csv(path.out)
 ```

### 2024-06-18 src/la_RF_corrected.Rmd
corrected RF

--- 2024-02-15 src/la_RF.Rmd
@@
 
+## Problem postcodes
+
 ```{r}
@@
 
-df9610 <- read_csv(data.path9610)
+df9610 <- read_csv(data.path9610) %>% 
+  filter(
+      #V1.domain < m,             # for multiple pc
+      V1.domain == n,             # for n == 1
+      year > 1995)
+
+# df9610 %>% filter(V1.domain == 1) %>% 
+#   left_join(lookup, by = "pc", suffix =c("","")) %>% 
+#   filter(ladcd == "E08000006" & year == 2006) %>% 
+#   group_by(pc) %>% 
+#   summarise(n = n()) %>% 
+#   arrange(desc(n))
+# 
+# df9610 %>% filter(pc == "M28 2SL" & year == 2006 & V1.domain == 1)
+
+problem.pcs <- df9610 %>% filter(year > 1999 & year < 2011) %>% 
+  group_by(pc, year) %>% 
+  summarise(n = n()) %>%
+  ungroup() %>% 
+  # complete(pc, year) %>% 
+  # replace(is.na(n), 0) %>% 
+  arrange(desc(n)) %>% 
+  pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
+  dplyr::select(pc, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010) %>% 
+  slice_head(n=10) %>% 
+  dplyr::select(pc)
+
+# [1] "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"  "EC1V 4PY" "DE21 7BF" "E14 9SR"  "GL16 7YA" "WC1N 3XX"
+# [10] "CB2 1AB" 
+```
+
+## Correct 1996-2010
+
+The above exploratory analysis indicated that there is a huge increase in 2004-2006
+in the above postcodes. I turn them to NAs and then use a regression to impute these
+gaps.
+
+```{r}
+df9610 <- df9610 %>% 
+  group_by(pc, year) %>% 
+  summarise(n = n()) %>% 
+  ungroup() %>% 
+  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2003 & year <2007, NA, n))
+
+df9610 <- pdata.frame(df9610)
+predictions <- round(predict(plm(n ~ as.factor(year), effect = "individual", 
+                                 model = "within", data = df9610), newdata = df9610), 2)
+
+df9610$n[is.na(df9610$n)] <- predictions[is.na(df9610$n)]
+df9610 %>% filter(pc == "M28 2SL") %>% distinct(year)
+```
+
+## Load remaining data
+
+```{r}
+df9610 <- as_tibble(df9610) %>% 
+  mutate(year = as.double(as.character(year))) %>% 
+  left_join(lookup, by = "pc", suffix =c("","")) %>%
+  group_by(year, ladcd) %>%
+  summarise(n = n()) %>%            
+  ungroup()
 
@@
 
-df1112 <- read_csv(data.path2011_2012)
-
-df <- bind_rows(df9610, df1112) %>% 
+df1112 <- read_csv(data.path2011_2012) %>% 
   filter(
@@
   group_by(year, ladcd) %>%
-  #summarise(n = n()) %>%            # for multiple pc  
-  summarise(n = sum(V1.domain)) %>% # for n == 1 
+  summarise(n = n()) %>%             # for n == 1  
+  #summarise(n = sum(V1.domain)) %>% # for multiple pc  
   ungroup()
 
-# Complete panel
-## REPLACE WITH 0.1 INSTEAD OF OF 0 FOR THE GROWTH
-df <- df %>% filter(!is.na(ladcd)) %>% 
+df <- bind_rows(df9610, df1112) %>% 
+  filter(!is.na(ladcd)) %>% 
   complete(ladcd, year) %>% 
@@
 # Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 5055, 4690, 4732, 5070, 5040, 4718, ... 
+# Summary of sample sizes: 4718, 4690, 5040, 4718, 5055, 4704, ... 
 # Resampling results across tuning parameters:
@@
 #   mtry  splitrule   RMSE        Rsquared   MAE       
-#   2     variance    0.03227897  0.8023185  0.01991724
-#   2     extratrees  0.03150004  0.8057960  0.01912348
-#   4     variance    0.03271887  0.7985855  0.02025685
-#   4     extratrees  0.03174791  0.8047090  0.01932473
-#   6     variance    0.03358260  0.7879953  0.02080616
-#   6     extratrees  0.03207532  0.8018495  0.01964516
+#   2     variance    0.02336229  0.8713835  0.01536402
+#   2     extratrees  0.02351745  0.8698132  0.01553849
+#   4     variance    0.02405073  0.8667793  0.01599751
+#   4     extratrees  0.02369748  0.8716644  0.01577096
+#   6     variance    0.02504045  0.8591307  0.01680305
+#   6     extratrees  0.02386677  0.8708794  0.01596582
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule = extratrees and min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule =
+#  variance and min.node.size = 5.
 
@@
 
-path.out <- paste0(path, "/outputs/rf/figures/varimp_LA.png")
+path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.png")
 ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
@@
 
-path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
+path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected.csv")
 rf.year.all.metrics %>% 
@@
 
-path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_growth.png")
+path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_growth_corrected.png")
 ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
@@
 time.taken <- round(end.time - start.time,2)
-time.taken # ca. 4 minutes 
+time.taken # ca. 3 minutes 
 ```
@@
 
-path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv")
+path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth_corrected.csv")
 rf.year.all.metrics %>% 
@@
 ```
-
-## Granger
-
-```{r}
-library(plm)
-p.df <- pdata.frame(df, index= c("ladcd","year"))
-
-test <- pgrangertest(n ~ n.l.slag, data = p.df, order = 1L) # *** reject the null that x does NOT cause y
-# n ~ Lags(n, 1) + Lags(n.l.slag, 1), or
-# n ~ Lags(n, 1) + Lags(n.slag, 2)
-# n ~ Lags(n, 1)
-test$indgranger %>% as_tibble() %>% arrange(`p-value`) #filter(`p-value` < 0.01)
-
-test <- pgrangertest(n ~ n.slag, data = p.df, order = 1L)
-# n ~ Lags(n, 1:1) + Lags(n.slag, 1:1)
-# n ~ Lags(n, 1:1)
-test$indgranger %>% as_tibble() %>% arrange(`p-value`) #filter(`p-value` < 0.01)
-
-test <- pgrangertest(n.slag ~ n, data = p.df, order = 1L) # ** reject the null that x does NOT cause y
-test$indgranger %>% as_tibble() %>% arrange(`p-value`)
-
-p.df$n.London_dist = p.df$n.London/p.df$London
-p.df$n.nearest.city_dist = p.df$n.nearest.city/p.df$dist
-
-pgrangertest(n ~ n.London_dist, data = p.df, order = 1L)
-pgrangertest(n ~ n.nearest.city_dist, data = p.df, order = 1L)
-
-cor(p.df$n.London_dist, dplyr::lag(p.df$n.London_dist), use = "complete.obs")
-```

### 2024-06-18 src/oa_RF_corrected.Rmd
corrected RF

--- 2024-02-12 src/oa_RF.Rmd
@@
 lookup <- read_csv(path.lookup) %>% 
-  dplyr::select(pcds, oa11cd) %>% 
+  dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
   dplyr::rename(pc = pcds)
@@
 
-df9610 <- read_csv(data.path9610)
+df9610 <- read_csv(data.path9610) %>% 
+  filter(
+      #V1.domain < m,             # for multiple pc
+      V1.domain == n,             # for n == 1
+      year > 1995)
+
+# df9610 %>% filter(V1.domain == 1) %>% 
+#   left_join(lookup, by = "pc", suffix =c("","")) %>% 
+#   filter(ladcd == "E08000006" & year == 2006) %>% 
+#   group_by(pc) %>% 
+#   summarise(n = n()) %>% 
+#   arrange(desc(n))
+# 
+# df9610 %>% filter(pc == "M28 2SL" & year == 2006 & V1.domain == 1)
+
+problem.pcs <- df9610 %>% filter(year > 1999 & year < 2011) %>% 
+  group_by(pc, year) %>% 
+  summarise(n = n()) %>%
+  ungroup() %>% 
+  # complete(pc, year) %>% 
+  # replace(is.na(n), 0) %>% 
+  arrange(desc(n)) %>% 
+  pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
+  dplyr::select(pc, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010) %>% 
+  slice_head(n=10) %>% 
+  dplyr::select(pc)
+
+# [1] "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"  "EC1V 4PY" "DE21 7BF" "E14 9SR"  "GL16 7YA" "WC1N 3XX"
+# [10] "CB2 1AB" 
+```
+
+## Correct 1996-2010
+
+The above exploratory analysis indicated that there is a huge increase in 2004-2006
+in the above postcodes. I turn them to NAs and then use a regression to impute these
+gaps.
+
+```{r}
+df9610 <- df9610 %>% 
+  group_by(pc, year) %>% 
+  summarise(n = n()) %>% 
+  ungroup() %>% 
+  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2003 & year <2007, NA, n))
+
+df9610 <- pdata.frame(df9610)
+predictions <- round(predict(plm(n ~ as.factor(year), effect = "individual", 
+                                 model = "within", data = df9610), newdata = df9610), 2)
+
+df9610$n[is.na(df9610$n)] <- predictions[is.na(df9610$n)]
+df9610 %>% filter(pc == "M28 2SL") %>% distinct(year)
+```
+
+## Load remaining data
+
+```{r}
+df9610 <- as_tibble(df9610) %>% 
+  mutate(year = as.double(as.character(year))) %>% 
+  left_join(lookup, by = "pc", suffix =c("","")) %>%
+  group_by(year, oa11cd) %>%
+  summarise(n = n()) %>%            
+  ungroup()
 
@@
   group_by(year, oa11cd) %>%
-  #summarise(n = n()) %>%            # for multiple pc  
-  summarise(n = sum(V1.domain)) %>%  # for n == 1 
+  summarise(n = n()) %>%             # for n == 1  
+  #summarise(n = sum(V1.domain)) %>% # for multiple pc  
   ungroup()
 
-# Partially complete panel as not all OAs have a website in at least one year
-df <- df %>% filter(!is.na(oa11cd)) %>% 
+df <- bind_rows(df9610, df1112) %>% 
+  filter(!is.na(oa11cd)) %>% 
   complete(oa11cd, year) %>% 
@@
 
-## spatial data OLD
-
-```{r eval=F}
+## spatial data
+
+```{r}
 # get OA for England and Wales
 path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
-oa.ew <- readOGR(path.geo)
+oa.ew <- st_read(path.geo)
 # source: https://geoportal.statistics.gov.uk/
@@
 # spatial transformations
-oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))
+oa.ew <- st_transform(oa.ew, 4326)  # EPSG code for WGS84
 
 # keep in the data slot only the ONS Output Area id, renaming it as 'id'
-oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
-colnames(oa.ew@data) <- c('id', 'area')
-
-# reassign the polygon IDs
-oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))
-
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.ew)
-
-# # get OA for Scotland
-# path.geo.sc <- paste0(path, "/data/raw/SG_DataZoneBdry_2011")
-# oa.sc <- readOGR(dsn=path.geo.sc, layer = "SG_DataZone_Bdry_2011")
-# # source: https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011
-# 
-# # spatial transformations
-# oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-# 
-# # Scotland (follows same steps as EW, see notes above)
-# oa.sc <- oa.sc[, 'DataZone']
-# colnames(oa.sc@data) <- c('id')
-# 
-# # reassign the polygon IDs
-# oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-# 
-# # check the CRS has changed correctly, and the data slot has shrink to only the ID
-# summary(oa.sc)
-
-path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
-oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
-# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries
-
-# spatial transformations
-oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-
-# Scotland (follows same steps as EW, see notes above)
-oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
-colnames(oa.sc@data) <- c('id', 'area')
-
-# reassign the polygon IDs
-oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.sc)
-
-# build oa for GB
-oa.gb <- maptools::spRbind(oa.ew, oa.sc)
-rm(oa.ew, oa.sc)
-
-# oa.gb$geometry <- oa.gb$geometry %>%
-#   s2::s2_rebuild() %>%
-#   sf::st_as_sfc()
-
-# get GB
-path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
-gb <- readOGR(path.gb)
-# spatial transformations
-gb <- spTransform(gb, CRS("+init=epsg:4326"))
-```
-
-## spatial data NEW
-
-```{r}
-library(sf)
-
-# Get OA for England and Wales
-path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
-oa.ew <- st_read(path.geo)
-
-# Spatial transformations
-oa.ew <- st_transform(oa.ew, 4326)  # EPSG code for WGS84
-
-# Rename columns and keep only necessary columns
-oa.ew <- oa.ew %>% dplyr::select(OA11CD, Shape__Area) %>% 
+oa.ew <- oa.ew %>% 
+  dplyr::select(OA11CD, Shape__Area) %>% 
   rename(id = OA11CD,
          area = Shape__Area)
-# oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
-# colnames(oa.ew) <- c('id', 'area')
-
-# Check the data structure
-print(summary(oa.ew))
-
-# Get OA for Scotland
+
 path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
-oa.sc <- st_read(dsn = path.geo.sc, layer = "OutputArea2011_MHW")
-
-# Spatial transformations
+oa.sc <- st_read(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
+# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries
+
+# spatial transformations
 oa.sc <- st_transform(oa.sc, 4326)  # EPSG code for WGS84
 
-# Rename columns and keep only necessary columns
-oa.sc <- oa.sc %>% dplyr::select(code, SHAPE_1_Ar) %>% 
+# Scotland (follows same steps as EW, see notes above)
+oa.sc <- oa.sc %>% 
+  dplyr::select(code, SHAPE_1_Ar) %>% 
   rename(id = code,
          area = SHAPE_1_Ar)
-# oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
-# colnames(oa.sc) <- c('id', 'area')
-
-# Check the data structure
-print(summary(oa.sc))
 
@@
 
-# Get GB
-path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
-gb <- st_read(path.gb)
-
-# Spatial transformations
-gb <- st_transform(gb, 4326)  # EPSG code for WGS84
+# get UK
+path.uk <- "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/Countries_December_2023_Boundaries_UK_BUC/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson"
+uk <- st_read(path.uk)
+# spatial transformations
+uk <- st_transform(uk, 4326)  # EPSG code for WGS84
 ```
@@
 
-### TODELETE
-
-```{r, eval=FALSE}
-# ## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
-# #df <- df %>% filter(!is.na(area)) 
-# 
-# cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
-# cities.sf <- st_set_crs(cities.sf, 4326)
-# 
-# # help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
-# #   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
-# 
-# sf_use_s2(FALSE)
-# help1 <- st_filter( oa.gb, cities.sf) %>% 
-#   #st_coordinates() %>% 
-#   as_tibble() %>% 
-#   dplyr::select(id) %>% 
-#   mutate(cities = c("London",     # The city names have been manually checked and assigned
-#                     "Leicester",
-#                     "Leeds",
-#                     "Sheffield",
-#                     "Bristol",
-#                     "Birmingham",
-#                     "Manchester",
-#                     "Liverpool",
-#                     "Edinburgh",
-#                     "Glasgow"))
-# 
-# # too slow
-# #dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
-# 
-# # This efficiently returns the name of the column with the shortest distance
-# dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
-# 
-# help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
-#   #rename(nearest.city.oa11cd = id) %>% 
-#   dplyr::select(oa11cd, dist.city.name)
-# 
-# df <- df %>%  left_join(help2, by = c("oa11cd" = "oa11cd")) 
-```
-
 ### n for cities
@@
 #sapply(df, function(x) sum(is.na(x)))
-```
-
-### DELETE n for London
-
-```{r, eval=FALSE}
-
-# # OLD London definitions
-# # df <- df %>% left_join(df %>% 
-# #                          ungroup() %>% 
-# #                          filter(oa11cd == "E00176659") %>% 
-# #                          dplyr::select(year, n) %>% 
-# #                          rename(n.London = n)) %>% 
-# #   group_by(oa11cd) %>% 
-# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-# #   mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
-# #   ungroup() 
-# # 
-# # london.help <- df %>% group_by(RGN11NM, year) %>% 
-# #   summarise(n.London = sum(n)) %>% 
-# #   filter(RGN11NM == "London") %>%
-# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-# #   ungroup() %>% 
-# #   dplyr::select(-RGN11NM) 
-# 
-# london.oa <- st_filter(oa.gb, london.sf %>% 
-#                          filter(Name == "Central")) %>% 
-#   as_tibble() %>% 
-#   dplyr::select(id)
-# 
-# london.help <- df %>% inner_join(london.oa, by = c("oa11cd" = "id")) %>% 
-#   group_by(year) %>% 
-#   summarise(n.London = sum(n)) %>% 
-#   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-#   ungroup()
-# 
-# df <- df %>% left_join(london.help, by = c("year" = "year"))
 ```
@@
 
-path.out <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
+path.out <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.png")
 ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
@@
 
-path.out <- paste0(path, "/outputs/rf/models/oa_rf/all_OAs/all_OAs.RData")
+path.out <- paste0(path, "/outputs/rf/models/oa_rf/all_OAs/all_OAs_corrected.RData")
 save(model.all, file = path.out)
+
+path.out <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.csv")
+write_csv(varimp_mars, path.out)
 

### 2024-06-18 src/problem_years.Rmd
corrected RF

@@
 The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.
-
-```{r, eval=FALSE}
-
-data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
-#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df1112 <- read_csv(data.path2011_2012)
-
-df <- bind_rows(df9610, df1112) %>% 
-  filter(
-    #V1.domain < m,             # for multiple pc
-    V1.domain == n,             # for n == 1
-    year > 1995 & year <2013) %>%   
-  left_join(lookup, by = "pc", suffix =c("","")) %>% 
-  group_by(year, ladcd) %>%
-  #summarise(n = n()) %>%            # for multiple pc  
-  summarise(n = sum(V1.domain)) %>% # for n == 1 
-  ungroup()
-
-df <- df %>% filter(!is.na(ladcd)) %>% 
-  complete(ladcd, year) %>% 
-  replace(is.na(.), 0)
-
-# tests for Scotland
-# df %>% filter(substr(oa11cd, 1,1) =="S",
-#               year==2010) %>% summarise(mean(n, na.rm = T))
-# 
-# df %>% filter(oa11cd=="S00090182")
-# 
-# test@data %>% filter(substr(id, 1,1) =="S",
-#                      n!=0)
-# 
-# df %>% filter(substr(oa11cd, 1,1) =="W",
-#               year==2003)
-```
 

### 2024-07-03 src/la_RF_corrected.Rmd
corrected RF v2 LA

@@
 n = 1 #number of unique postcodes. 
-m = 12
+m = 11
 
@@
 
-# df9610 %>% filter(V1.domain == 1) %>% 
-#   left_join(lookup, by = "pc", suffix =c("","")) %>% 
-#   filter(ladcd == "E08000006" & year == 2006) %>% 
-#   group_by(pc) %>% 
-#   summarise(n = n()) %>% 
-#   arrange(desc(n))
-# 
-# df9610 %>% filter(pc == "M28 2SL" & year == 2006 & V1.domain == 1)
-
-problem.pcs <- df9610 %>% filter(year > 1999 & year < 2011) %>% 
+data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
+#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
+#This is what we use for the hyperlinks paper as per George's script
+
+df1112 <- read_csv(data.path2011_2012) %>% 
+  filter(
+    #V1.domain < m,             # for multiple pc
+    V1.domain == n)             # for n == 1
+
+df.long <- bind_rows(df9610, df1112) %>% 
+  filter(year > 1995 & year < 2013) %>% 
   group_by(pc, year) %>% 
@@
   ungroup() %>% 
-  # complete(pc, year) %>% 
-  # replace(is.na(n), 0) %>% 
-  arrange(desc(n)) %>% 
+  #arrange(desc(n)) %>% 
   pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
-  dplyr::select(pc, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010) %>% 
-  slice_head(n=10) %>% 
-  dplyr::select(pc)
-
-# [1] "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"  "EC1V 4PY" "DE21 7BF" "E14 9SR"  "GL16 7YA" "WC1N 3XX"
-# [10] "CB2 1AB" 
-```
-
-## Correct 1996-2010
+  dplyr::select(pc, n_1996, n_1999, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010, n_2011, n_2012) %>%
+  arrange(desc(n_2004), desc(n_2005)) %>% 
+  slice_head(n=1000) %>% 
+  pivot_longer(!pc, names_to = "year", values_to = "n") %>% 
+  mutate(year = as.integer(sub("n_", "", year, fixed = TRUE))) %>% 
+  replace(is.na(.), 0) %>% 
+  mutate(outlier = ifelse(n > 1000 & (year==2004 | year == 2005), pc, ""))
+
+problem.pcs <- df.long %>% filter(year==2004 | year == 2005, 
+                   n > 1000) %>% 
+  distinct(pc)
+
+# n > 1500: "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"   
+# n > 1000: SE24 9HP, CV8 2ED, GL16 7YA, CW1 6GL, M28 2SL, DE21 7BF
+```
+
+## Corrected
 
@@
 ```{r}
-df9610 <- df9610 %>% 
+df.corrected <- bind_rows(df9610, df1112) %>% 
+  filter(year > 1995 & year < 2013) %>% 
   group_by(pc, year) %>% 
@@
 
-df9610 <- pdata.frame(df9610)
+df.corrected <- pdata.frame(df.corrected)
 predictions <- round(predict(plm(n ~ as.factor(year), effect = "individual", 
-                                 model = "within", data = df9610), newdata = df9610), 2)
-
-df9610$n[is.na(df9610$n)] <- predictions[is.na(df9610$n)]
-df9610 %>% filter(pc == "M28 2SL") %>% distinct(year)
-```
-
-## Load remaining data
-
-```{r}
-df9610 <- as_tibble(df9610) %>% 
-  mutate(year = as.double(as.character(year))) %>% 
+                                 model = "within", data = df.corrected), newdata = df.corrected), 2)
+
+df.corrected$n[is.na(df.corrected$n)] <- predictions[is.na(df.corrected$n)]
+```
+
+## LA df
+
+```{r}
+df <- df.corrected %>% 
+  mutate(year = as.integer(as.character(year))) %>% 
+  filter(year > 1995 & year < 2013) %>% 
   left_join(lookup, by = "pc", suffix =c("","")) %>% 
   group_by(year, ladcd) %>%
-  summarise(n = n()) %>%            
-  ungroup()
-
-data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
-#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df1112 <- read_csv(data.path2011_2012) %>% 
-  filter(
-    #V1.domain < m,             # for multiple pc
-    V1.domain == n,             # for n == 1
-    year > 1995 & year <2013) %>%   
-  left_join(lookup, by = "pc", suffix =c("","")) %>% 
-  group_by(year, ladcd) %>%
-  summarise(n = n()) %>%             # for n == 1  
-  #summarise(n = sum(V1.domain)) %>% # for multiple pc  
-  ungroup()
-
-df <- bind_rows(df9610, df1112) %>% 
-  filter(!is.na(ladcd)) %>% 
+  summarise(n = n()) %>% # for multiple pc  
+  ungroup() %>% 
   complete(ladcd, year) %>% 
+  filter(!is.na(ladcd)) %>% # drop c. 10 - 200 per year, which are not assigned to a LAD
   replace(is.na(.), 0) 
-
-# tests for Scotland
-# df %>% filter(substr(oa11cd, 1,1) =="S",
-#               year==2010) %>% summarise(mean(n, na.rm = T))
-# 
-# df %>% filter(oa11cd=="S00090182")
-# 
-# test@data %>% filter(substr(id, 1,1) =="S",
-#                      n!=0)
-# 
-# df %>% filter(substr(oa11cd, 1,1) =="W",
-#               year==2003)
 ```
@@
 #   mtry  splitrule   RMSE        Rsquared   MAE        
-#   2     variance    0.02336229  0.8713835  0.01536402
-#   2     extratrees  0.02351745  0.8698132  0.01553849
-#   4     variance    0.02405073  0.8667793  0.01599751
-#   4     extratrees  0.02369748  0.8716644  0.01577096
-#   6     variance    0.02504045  0.8591307  0.01680305
-#   6     extratrees  0.02386677  0.8708794  0.01596582
+#   2     variance    0.01366816  0.8895425  0.009767759
+#   2     extratrees  0.01350182  0.8900993  0.009603882
+#   4     variance    0.01408821  0.8841051  0.010134321
+#   4     extratrees  0.01354075  0.8914524  0.009664646
+#   6     variance    0.01445541  0.8772260  0.010417962
+#   6     extratrees  0.01369444  0.8897055  0.009772824
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule =
-#  variance and min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule = extratrees
+#  and min.node.size = 5.
 
@@
 # Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 4044, 3685, 4056, 4056, 4032, 3707, ... 
+# Summary of sample sizes: 4044, 3685, 4032, 4044, 4044, 3696, ... 
 # Resampling results across tuning parameters:
@@
 #   mtry  splitrule   RMSE       Rsquared   MAE      
-#   2     variance    0.2004565  0.6285613  0.1460545
-#   2     extratrees  0.2001702  0.6342523  0.1480880
-#   4     variance    0.2080782  0.6066952  0.1503476
-#   4     extratrees  0.2065532  0.6220336  0.1527747
-#   6     variance    0.2099617  0.5996347  0.1507628
-#   6     extratrees  0.2084788  0.6180880  0.1544252
+#   2     variance    0.2182933  0.5361604  0.1560099
+#   2     extratrees  0.2099702  0.5660477  0.1501194
+#   4     variance    0.2168118  0.5238461  0.1550630
+#   4     extratrees  0.2074722  0.5590058  0.1489474
+#   6     variance    0.2268891  0.5014882  0.1631466
+#   6     extratrees  0.2082266  0.5503986  0.1490911
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule = extratrees and min.node.size = 5.
+# The final values used for the model were mtry = 4, splitrule = extratrees
+#  and min.node.size = 5.
 

### 2024-07-03 src/la_RF_corrected_10.Rmd
corrected RF v2 LA

--- 2024-07-03 src/la_RF_corrected.Rmd
@@
 
+I use the problem postcodes from n = 1 as these are the true outliers as per the
+line plots.
+
 ```{r}
@@
   filter(
-      #V1.domain < m,             # for multiple pc
-      V1.domain == n,             # for n == 1
+      V1.domain < m,               # for multiple pc
+      #V1.domain == n,             # for n == 1
       year > 1995)
@@
   filter(
-    #V1.domain < m,             # for multiple pc
-    V1.domain == n)             # for n == 1
+    V1.domain < m)               # for multiple pc
+    #V1.domain == n)             # for n == 1
 
@@
   group_by(pc, year) %>% 
-  summarise(n = n()) %>%
+  summarise(n = sum(V1.domain)) %>%
   ungroup() %>% 
@@
 
-problem.pcs <- df.long %>% filter(year==2004 | year == 2005, 
-                   n > 1000) %>% 
-  distinct(pc)
-
-# n > 1500: "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"   
-# n > 1000: SE24 9HP, CV8 2ED, GL16 7YA, CW1 6GL, M28 2SL, DE21 7BF
+# From n = 1 as these are the 'true' outliers
+problem.pcs <- c("SE24 9HP", "CV8 2ED", "GL16 7YA", "CW1 6GL", "M28 2SL", "DE21 7BF")
 ```
@@
   group_by(pc, year) %>% 
-  summarise(n = n()) %>% 
+  summarise(n = sum(V1.domain)) %>% 
   ungroup() %>% 
-  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2003 & year <2007, NA, n))
+  # The next line is different than the same for  n = 1: problem.pcs$pc
+  mutate(n = ifelse(pc %in% as.character(problem.pcs) & year > 2003 & year <2007, NA, n))
 
@@
 df.corrected$n[is.na(df.corrected$n)] <- predictions[is.na(df.corrected$n)]
+
 ```
@@
   group_by(year, ladcd) %>%
-  summarise(n = n()) %>% # for multiple pc  
+  summarise(n = sum(n)) %>% # for multiple pc  
   ungroup() %>% 
@@
 
-detectCores()
-cl <- makePSOCKcluster(14)
-registerDoParallel(cl)
+# detectCores()
+# cl <- makePSOCKcluster(14)
+# registerDoParallel(cl)
 
@@
 
-stopCluster(cl)
+#stopCluster(cl)
 end.time <- Sys.time()
@@
 #   mtry  splitrule   RMSE       Rsquared   MAE      
-#   2     variance    0.01366816  0.8895425  0.009767759
-#   2     extratrees  0.01350182  0.8900993  0.009603882
-#   4     variance    0.01408821  0.8841051  0.010134321
-#   4     extratrees  0.01354075  0.8914524  0.009664646
-#   6     variance    0.01445541  0.8772260  0.010417962
-#   6     extratrees  0.01369444  0.8897055  0.009772824
+#   2     variance    0.1881989  0.7750790  0.1223893
+#   2     extratrees  0.1944376  0.7543149  0.1274344
+#   4     variance    0.1906884  0.7730657  0.1256452
+#   4     extratrees  0.1930998  0.7596818  0.1276155
+#   6     variance    0.1937399  0.7677967  0.1282843
+#   6     extratrees  0.1927437  0.7608415  0.1277447
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule = extratrees
-#  and min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule = variance and
+#  min.node.size = 5.
 
@@
 
-path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.png")
+path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected_10.png")
 ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
@@
 # detectCores()
-cl <- makePSOCKcluster(14)
-registerDoParallel(cl)
+# cl <- makePSOCKcluster(14)
+# registerDoParallel(cl)
 
@@
 
-stopCluster(cl)
+# stopCluster(cl)
 end.time <- Sys.time() 
@@
 
-path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected.csv")
+path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected_10.csv")
 rf.year.all.metrics %>% 
@@
 #   mtry  splitrule   RMSE       Rsquared   MAE      
-#   2     variance    0.2182933  0.5361604  0.1560099
-#   2     extratrees  0.2099702  0.5660477  0.1501194
-#   4     variance    0.2168118  0.5238461  0.1550630
-#   4     extratrees  0.2074722  0.5590058  0.1489474
-#   6     variance    0.2268891  0.5014882  0.1631466
-#   6     extratrees  0.2082266  0.5503986  0.1490911
+#   2     variance    0.2773324  0.6169393  0.2025296
+#   2     extratrees  0.2692856  0.6437281  0.1982852
+#   4     variance    0.2836790  0.5992575  0.2060827
+#   4     extratrees  0.2717096  0.6339425  0.2004439
+#   6     variance    0.2899917  0.5806674  0.2100017
+#   6     extratrees  0.2755481  0.6223759  0.2037709
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 4, splitrule = extratrees
+# The final values used for the model were mtry = 2, splitrule = extratrees
 #  and min.node.size = 5.
@@
 
-path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_growth_corrected.png")
+path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_growth_corrected_10.png")
 ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
@@
 
-path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth_corrected.csv")
+path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth_corrected_10.csv")
 rf.year.all.metrics %>% 

### 2024-07-03 src/oa_RF_corrected.Rmd
corrected RF v2 LA

@@
 
-df9610 <- pdata.frame(df9610)
+df.corrected <- pdata.frame(df.corrected)
 predictions <- round(predict(plm(n ~ as.factor(year), effect = "individual", 
-                                 model = "within", data = df9610), newdata = df9610), 2)
-
-df9610$n[is.na(df9610$n)] <- predictions[is.na(df9610$n)]
-df9610 %>% filter(pc == "M28 2SL") %>% distinct(year)
-```
-
-## Load remaining data
-
-```{r}
-df9610 <- as_tibble(df9610) %>% 
-  mutate(year = as.double(as.character(year))) %>% 
+                                 model = "within", data = df.corrected), newdata = df.corrected), 2)
+
+df.corrected$n[is.na(df.corrected$n)] <- predictions[is.na(df.corrected$n)]
+```
+
+## OA df
+
+```{r}
+df <- df.corrected %>% 
+  mutate(year = as.integer(as.character(year))) %>% 
+  filter(year > 1995 & year < 2013) %>% 
   left_join(lookup, by = "pc", suffix =c("","")) %>% 
   group_by(year, oa11cd) %>%
-  summarise(n = n()) %>%            
-  ungroup()
-
-data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
-#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df1112 <- read_csv(data.path2011_2012) %>% 
-  filter(
-    #V1.domain < m,             # for multiple pc
-    V1.domain == n,             # for n == 1
-    year > 1995 & year <2013) %>%   
-  left_join(lookup, by = "pc", suffix =c("","")) %>% 
-  group_by(year, oa11cd) %>%
-  summarise(n = n()) %>%             # for n == 1  
-  #summarise(n = sum(V1.domain)) %>% # for multiple pc  
-  ungroup()
-
-df <- bind_rows(df9610, df1112) %>% 
-  filter(!is.na(oa11cd)) %>% 
+  summarise(n = n()) %>% # for multiple pc  
+  ungroup() %>% 
   complete(oa11cd, year) %>% 
+  filter(!is.na(oa11cd)) %>% # drop c. 10 - 200 per year, which are not assigned to a LAD
   replace(is.na(.), 0) 
-
-# tests for Scotland
-# df %>% filter(substr(oa11cd, 1,1) =="S",
-#               year==2010) %>% summarise(mean(n, na.rm = T))
-# 
-# df %>% filter(oa11cd=="S00090182")
-# 
-# test@data %>% filter(substr(id, 1,1) =="S",
-#                      n!=0)
-# 
-# df %>% filter(substr(oa11cd, 1,1) =="W",
-#               year==2003)
 ```
@@
 time.taken <- round(end.time - start.time,2)
-time.taken #20.6h
+time.taken #20.16h
 
@@
 #   mtry  splitrule   RMSE      Rsquared   MAE      
-#   2     variance    4.999920  0.2053697  1.047293
-#   2     extratrees  5.167655  0.1540213  1.103048
-#   5     variance    5.187039  0.1710128  1.034382
-#   5     extratrees  5.169570  0.1557868  1.122845
-#   8     variance    5.633427  0.1247189  1.035171
-#   8     extratrees  5.182860  0.1563555  1.137332
+#   2     variance    2.307827  0.2770946  0.7497850
+#   2     extratrees  2.329867  0.2306295  0.7719781
+#   5     variance    2.432873  0.2744482  0.7596487
+#   5     extratrees  2.409776  0.2344573  0.7960409
+#   8     variance    2.486179  0.2616929  0.7571522
+#   8     extratrees  2.452737  0.2318285  0.8086679
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule =
-#  variance and min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule = variance and min.node.size = 5.
 
@@
 path.out <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.csv")
-write_csv(varimp_mars, path.out)
-
-# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
-#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
-#                                                         "year")], 
-#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
-#                     trControl = tc,
-#                     method = "ranger",
-#                     #na.action = na.omit,
-#                     #importance = "impurity",
-#                     metric="Rsquared",
-#                     verbose=FALSE)
-# ffsmodel_LLO
-# ffsmodel_LLO$selectedvars
-# plot_ffs(ffsmodel_LLO)
-```
+varimp_mars$importance %>% 
+  rownames_to_column() %>%
+  rename(variable = rowname,
+         importance = Overall) %>% 
+  write_csv(path.out)
+```
+
+THE BELLOW HAS NOT BEEN RUN NEITHER THE FILENAMES HAS BEEN CHANGED (add: _corrected)
 

### 2024-07-03 src/oa_RF_corrected_10.Rmd
corrected RF v2 LA

--- 2024-06-18 src/oa_RF_corrected.Rmd
@@
   filter(
-    #V1.domain < m,             # for multiple pc
-    V1.domain == n,             # for n == 1
+    V1.domain < m,               # for multiple pc
+    #V1.domain == n,             # for n == 1
     year > 1995 & year <2013) %>%   
@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    4.999920  0.2053697  1.047293
-#   2     extratrees  5.167655  0.1540213  1.103048
-#   5     variance    5.187039  0.1710128  1.034382
-#   5     extratrees  5.169570  0.1557868  1.122845
-#   8     variance    5.633427  0.1247189  1.035171
-#   8     extratrees  5.182860  0.1563555  1.137332
+#   2     variance    5.341123  0.3211637  1.294570
+#   2     extratrees  5.114541  0.2727356  1.288928
+#   5     variance    6.248341  0.3136499  1.380152
+#   5     extratrees  5.584957  0.2834177  1.397949
+#   8     variance    6.801851  0.2977672  1.395883
+#   8     extratrees  5.957788  0.2811840  1.444036
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule =
-#  variance and min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule
+#  = extratrees and min.node.size = 5.
 
@@
 
-path.out <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.png")
+path.out <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected_10.png")
 ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
@@
 
-path.out <- paste0(path, "/outputs/rf/models/oa_rf/all_OAs/all_OAs_corrected.RData")
+path.out <- paste0(path, "/outputs/rf/models/oa_rf/all_OAs/all_OAs_corrected_10.RData")
 save(model.all, file = path.out)
 
-path.out <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.csv")
-write_csv(varimp_mars, path.out)
-
-# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
-#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
-#                                                         "year")], 
-#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
-#                     trControl = tc,
-#                     method = "ranger",
-#                     #na.action = na.omit,
-#                     #importance = "impurity",
-#                     metric="Rsquared",
-#                     verbose=FALSE)
-# ffsmodel_LLO
-# ffsmodel_LLO$selectedvars
-# plot_ffs(ffsmodel_LLO)
-```
+path.out <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected_10.csv")
+varimp_mars$importance %>% 
+  rownames_to_column() %>%
+  rename(variable = rowname,
+         importance = Overall) %>% 
+  write_csv(path.out)
+```
+
+THE BELLOW HAS NOT BEEN RUN NEITHER THE FILENAMES HAS BEEN CHANGED (add: _corrected)
 

### 2024-07-03 src/problem_years.Rmd
corrected RF v2 LA

@@
 library(magick)
+library(ggrepel)
+library(scales)
 
@@
 
-# df9610 %>% filter(V1.domain == 1) %>% 
-#   left_join(lookup, by = "pc", suffix =c("","")) %>% 
-#   filter(ladcd == "E08000006" & year == 2006) %>% 
-#   group_by(pc) %>% 
-#   summarise(n = n()) %>% 
-#   arrange(desc(n))
-# 
-# df9610 %>% filter(pc == "M28 2SL" & year == 2006 & V1.domain == 1)
-
-problem.pcs <- df9610 %>% filter(year > 1999 & year < 2011) %>% 
-  group_by(pc, year) %>% 
-  summarise(n = n()) %>%
-  ungroup() %>% 
-  # complete(pc, year) %>% 
-  # replace(is.na(n), 0) %>% 
-  arrange(desc(n)) %>% 
-  pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
-  dplyr::select(pc, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010) %>% 
-  slice_head(n=10) %>% 
-  dplyr::select(pc)
-
-# [1] "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"  "EC1V 4PY" "DE21 7BF" "E14 9SR"  "GL16 7YA" "WC1N 3XX"
-# [10] "CB2 1AB" 
-```
-
-## Correct 1996-2010
-
-The above exploratory analysis indicated that there is a huge increase in 2004-2006
-in the above postcodes. I turn them to NAs and then use a regression to impute these
-gaps.
-
-```{r}
-df9610 <- df9610 %>% 
-  group_by(pc, year) %>% 
-  summarise(n = n()) %>% 
-  ungroup() %>% 
-  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2003 & year <2007, NA, n))
-
-df9610 <- pdata.frame(df9610)
-predictions <- round(predict(plm(n ~ as.factor(year), effect = "individual", 
-                                 model = "within", data = df9610), newdata = df9610), 2)
-
-df9610$n[is.na(df9610$n)] <- predictions[is.na(df9610$n)]
-df9610 %>% filter(pc == "M28 2SL") %>% distinct(year)
-```
-
-## LA
-
-### Load remaining data
-
-```{r}
-df9610 <- as_tibble(df9610) %>% 
-  mutate(year = as.double(as.character(year))) %>% 
-  left_join(lookup, by = "pc", suffix =c("","")) %>%
-  group_by(year, ladcd) %>%
-  summarise(n = n()) %>%            
-  ungroup()
-
 data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
@@
     #V1.domain < m,             # for multiple pc
-    V1.domain == n,             # for n == 1
-    year > 1995 & year <2013) %>%   
+    V1.domain == n)             # for n == 1
+
+df.long <- bind_rows(df9610, df1112) %>% 
+  filter(year > 1995 & year < 2013) %>% 
+  group_by(pc, year) %>% 
+  summarise(n = n()) %>%
+  ungroup() %>% 
+  #arrange(desc(n)) %>% 
+  pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
+  dplyr::select(pc, n_1996, n_1999, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010, n_2011, n_2012) %>%
+  arrange(desc(n_2004), desc(n_2005)) %>% 
+  slice_head(n=1000) %>% 
+  pivot_longer(!pc, names_to = "year", values_to = "n") %>% 
+  mutate(year = as.integer(sub("n_", "", year, fixed = TRUE))) %>% 
+  replace(is.na(.), 0) %>% 
+  mutate(outlier = ifelse(n > 1000 & (year==2004 | year == 2005), pc, ""))
+
+problem.pcs <- df.long %>% filter(year==2004 | year == 2005, 
+                   n > 1000) %>% 
+  distinct(pc)
+
+# n > 1500: "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"   
+# n > 1000: SE24 9HP, CV8 2ED, GL16 7YA, CW1 6GL, M28 2SL, DE21 7BF
+
+df.long %>% 
+  mutate(year = as.double(year)) %>% 
+  ggplot() +
+  geom_line(mapping = aes(x = year, y = n, group = pc), alpha = .3, size = .2) +
+  ylab("N. of websites per postcode") +
+  scale_x_continuous(breaks=seq(1996, 2012, 1)) +
+  scale_y_continuous(labels=scales::comma) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  geom_text_repel(aes(x = year, y = n, label=outlier), cex = 3)
+
+path.out <- paste0(path, "/outputs/pc_year_1.png")
+ggsave(path.out, bg='#ffffff')  
+```
+
+## Corrected
+
+The above exploratory analysis indicated that there is a huge increase in 2004-2006
+in the above postcodes. I turn them to NAs and then use a regression to impute these
+gaps.
+
+```{r}
+df.corrected <- bind_rows(df9610, df1112) %>% 
+  filter(year > 1995 & year < 2013) %>% 
+  group_by(pc, year) %>% 
+  summarise(n = n()) %>% 
+  ungroup() %>% 
+  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2003 & year <2007, NA, n))
+
+df.corrected <- pdata.frame(df.corrected)
+predictions <- round(predict(plm(n ~ as.factor(year), effect = "individual", 
+                                 model = "within", data = df.corrected), newdata = df.corrected), 2)
+
+df.corrected$n[is.na(df.corrected$n)] <- predictions[is.na(df.corrected$n)]
+
+df.corrected %>% 
+  as_tibble() %>% 
+  mutate(year = as.double(as.character(year))) %>% 
+  ggplot() +
+  geom_line(mapping = aes(x = year, y = n, group = pc), alpha = .3, size = .2) +
+  ylab("N. of websites per postcode") +
+  scale_x_continuous(breaks=seq(1996, 2012, 1)) +
+  scale_y_continuous(labels=scales::comma) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+path.out <- paste0(path, "/outputs/pc_year_1_corrected.png")
+ggsave(path.out, bg='#ffffff')  
+```
+
+## LA
+
+### LA df
+
+```{r}
+df <- df.corrected %>% 
+  mutate(year = as.integer(as.character(year))) %>% 
+  filter(year > 1995 & year < 2013) %>% 
   left_join(lookup, by = "pc", suffix =c("","")) %>% 
   group_by(year, ladcd) %>%
-  summarise(n = n()) %>%             # for n == 1  
-  #summarise(n = sum(V1.domain)) %>% # for multiple pc  
-  ungroup()
-
-df <- bind_rows(df9610, df1112) %>% 
-  filter(!is.na(ladcd)) %>% 
+  summarise(n = n()) %>% # for multiple pc  
+  ungroup() %>% 
   complete(ladcd, year) %>% 
+  filter(!is.na(ladcd)) %>% # drop c. 10 - 200 per year, which are not assigned to a LAD
   replace(is.na(.), 0) 
-
-# tests for Scotland
-# df %>% filter(substr(oa11cd, 1,1) =="S",
-#               year==2010) %>% summarise(mean(n, na.rm = T))
-# 
-# df %>% filter(oa11cd=="S00090182")
-# 
-# test@data %>% filter(substr(id, 1,1) =="S",
-#                      n!=0)
-# 
-# df %>% filter(substr(oa11cd, 1,1) =="W",
-#               year==2003)
 ```
@@
 filename = paste(path.out, t, ".png", sep = "")
-ggsave(filename)
+ggsave(filename, bg='#ffffff')
 }
@@
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/corrected/LA/morani_la.csv")
+path.out <- paste0(path, "/outputs/lisa/corrected/LA/morani_la_1.csv")
 morani %>% as_tibble() %>% 
@@
 
-### Load remaining data
+### OA df
 
 ```{r}
-df9610 <- as_tibble(df9610) %>% 
-  mutate(year = as.double(as.character(year))) %>% 
+df <- df.corrected %>% 
+  mutate(year = as.integer(as.character(year))) %>% 
+  filter(year > 1995 & year < 2013) %>% 
   left_join(lookup, by = "pc", suffix =c("","")) %>% 
   group_by(year, oa11cd) %>%
-  summarise(n = n()) %>%            
-  ungroup()
-
-data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
-#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df1112 <- read_csv(data.path2011_2012) %>% 
-  filter(
-    #V1.domain < m,             # for multiple pc
-    V1.domain == n,             # for n == 1
-    year > 1995 & year <2013) %>%   
-  left_join(lookup, by = "pc", suffix =c("","")) %>% 
-  group_by(year, oa11cd) %>%
-  summarise(n = n()) %>%             # for n == 1  
-  #summarise(n = sum(V1.domain)) %>% # for multiple pc  
-  ungroup()
-
-df <- bind_rows(df9610, df1112) %>% 
-  filter(!is.na(oa11cd)) %>% 
+  summarise(n = n()) %>% # for multiple pc  
+  ungroup() %>% 
   complete(oa11cd, year) %>% 
+  filter(!is.na(oa11cd)) %>% # drop c. 10 - 200 per year, which are not assigned to a LAD
   replace(is.na(.), 0) 
-
-# tests for Scotland
-# df %>% filter(substr(oa11cd, 1,1) =="S",
-#               year==2010) %>% summarise(mean(n, na.rm = T))
-# 
-# df %>% filter(oa11cd=="S00090182")
-# 
-# test@data %>% filter(substr(id, 1,1) =="S",
-#                      n!=0)
-# 
-# df %>% filter(substr(oa11cd, 1,1) =="W",
-#               year==2003)
 ```
@@
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/corrected/OA/morani_oa.csv")
+path.out <- paste0(path, "/outputs/lisa/corrected/OA/morani_oa_1.csv")
 morani %>% as_tibble() %>% 
@@
                    year = V3) %>% 
-  kable() %>% 
+  #kable() %>% 
   write_csv(path.out)

### 2024-07-03 src/problem_years_10.Rmd
corrected RF v2 LA

--- 2024-07-03 src/problem_years.Rmd
@@
 
-problem.pcs <- df.long %>% filter(year==2004 | year == 2005, 
-                   n > 1000) %>% 
-  distinct(pc)
-
-# n > 1500: "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"   
-# n > 1000: SE24 9HP, CV8 2ED, GL16 7YA, CW1 6GL, M28 2SL, DE21 7BF
+# From n = 1 as these are the 'true' outliers
+problem.pcs <- c("SE24 9HP", "CV8 2ED", "GL16 7YA", "CW1 6GL", "M28 2SL", "DE21 7BF")
 
@@
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc", n, "_la") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc", m, "_la")  # for multiple PC
+#path.out <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc", n, "_la") # for unique PC
+path.out <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc", m, "_la")  # for multiple PC
 
@@
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/corrected/LA/morani_la_1.csv")
+path.out <- paste0(path, "/outputs/lisa/corrected/LA/morani_la_10.csv")
 morani %>% as_tibble() %>% 
@@
   group_by(year, oa11cd) %>%
-  summarise(n = n()) %>% # for multiple pc  
+  summarise(n = sum(n)) %>% # for multiple pc  
   ungroup() %>% 
@@
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc", n, "_oa") # for unique PC
-#path.out <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc", m, "_oa")  # for multiple PC
+#path.out <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc", n, "_oa") # for unique PC
+path.out <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc", m, "_oa")  # for multiple PC
 
@@
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/corrected/OA/morani_oa_1.csv")
+path.out <- paste0(path, "/outputs/lisa/corrected/OA/morani_oa_10.csv")
 morani %>% as_tibble() %>% 

### 2024-07-03 src/problem_years_old.Rmd
corrected RF v2 LA

--- 2024-06-18 src/problem_years.Rmd
@@
 library(magick)
+library(ggrepel)
 
@@
 # [10] "CB2 1AB" 
+```
+
+```{r}
+df %>%  pivot_longer(!LSOA11CD, names_to = "time", values_to = "premises") %>% 
+  left_join(look.up) %>%
+  group_by(LAD22CD, time) %>% 
+  summarise(premises = sum(premises)) %>% #ungroup() %>% filter(LAD22CD == 'E08000025')
+  # group_by(LAD22CD) %>% 
+  # mutate(help = mean(premises)) %>% 
+  # filter(help<70) %>% 
+  ggplot() +
+  geom_line(mapping = aes(x = time, y = premises, group = LAD22CD), alpha = .3, size = .2) + #, color = 'Premises'
+  #ggtitle("Median experienced internet speeds for Local Authorities") + 
+  xlab("time") + ylab("Gambling premises") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+a <- df9610 %>% #filter(year > 1999 & year < 2011) %>% 
+  group_by(pc, year) %>% 
+  summarise(n = n()) %>%
+  ungroup() %>% 
+  #arrange(desc(n)) %>% 
+  pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
+  dplyr::select(pc, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010) %>%
+  arrange(desc(n_2004), desc(n_2005)) %>% 
+  slice_head(n=1000) %>% 
+  pivot_longer(!pc, names_to = "year", values_to = "n") %>% 
+  mutate(year = as.integer(sub("n_", "", year, fixed = TRUE)))  
+a %>% ggplot() +
+  geom_line(mapping = aes(x = year, y =n, group = pc), alpha = .3, size = .2)
 ```

### 2024-07-07 src/oa_RF_corrected.Rmd
all corrected but regions

@@
 #   mtry  splitrule   RMSE      Rsquared   MAE      
-#   2     variance    2.307827  0.2770946  0.7497850
-#   2     extratrees  2.329867  0.2306295  0.7719781
-#   5     variance    2.432873  0.2744482  0.7596487
-#   5     extratrees  2.409776  0.2344573  0.7960409
-#   8     variance    2.486179  0.2616929  0.7571522
-#   8     extratrees  2.452737  0.2318285  0.8086679
+#   2     variance    1.221299  0.4082199  0.5829598
+#   2     extratrees  1.308769  0.3297824  0.6136490
+#   5     variance    1.203082  0.4241359  0.5703508
+#   5     extratrees  1.280883  0.3513153  0.6124031
+#   8     variance    1.214268  0.4155308  0.5674246
+#   8     extratrees  1.284107  0.3502621  0.6180390
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule = variance and min.node.size = 5.
+# The final values used for the model were mtry = 5, splitrule = variance and
+#  min.node.size = 5.
 

### 2024-07-07 src/oa_RF_corrected_10.Rmd
all corrected but regions

@@
 
-# df9610 %>% filter(V1.domain == 1) %>% 
-#   left_join(lookup, by = "pc", suffix =c("","")) %>% 
-#   filter(ladcd == "E08000006" & year == 2006) %>% 
-#   group_by(pc) %>% 
-#   summarise(n = n()) %>% 
-#   arrange(desc(n))
-# 
-# df9610 %>% filter(pc == "M28 2SL" & year == 2006 & V1.domain == 1)
-
-problem.pcs <- df9610 %>% filter(year > 1999 & year < 2011) %>% 
+data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
+#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
+#This is what we use for the hyperlinks paper as per George's script
+
+df1112 <- read_csv(data.path2011_2012) %>% 
+  filter(
+    V1.domain < m)               # for multiple pc
+    #V1.domain == n)             # for n == 1
+
+df.long <- bind_rows(df9610, df1112) %>% 
+  filter(year > 1995 & year < 2013) %>% 
   group_by(pc, year) %>% 
-  summarise(n = n()) %>%
+  summarise(n = sum(V1.domain)) %>%
   ungroup() %>% 
-  # complete(pc, year) %>% 
-  # replace(is.na(n), 0) %>% 
-  arrange(desc(n)) %>% 
+  #arrange(desc(n)) %>% 
   pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
-  dplyr::select(pc, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010) %>% 
-  slice_head(n=10) %>% 
-  dplyr::select(pc)
-
-# [1] "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"  "EC1V 4PY" "DE21 7BF" "E14 9SR"  "GL16 7YA" "WC1N 3XX"
-# [10] "CB2 1AB" 
-```
-
-## Correct 1996-2010
+  dplyr::select(pc, n_1996, n_1999, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010, n_2011, n_2012) %>%
+  arrange(desc(n_2004), desc(n_2005)) %>% 
+  slice_head(n=1000) %>% 
+  pivot_longer(!pc, names_to = "year", values_to = "n") %>% 
+  mutate(year = as.integer(sub("n_", "", year, fixed = TRUE))) %>% 
+  replace(is.na(.), 0) %>% 
+  mutate(outlier = ifelse(n > 1000 & (year==2004 | year == 2005), pc, ""))
+
+# From n = 1 as these are the 'true' outliers
+problem.pcs <- c("SE24 9HP", "CV8 2ED", "GL16 7YA", "CW1 6GL", "M28 2SL", "DE21 7BF")
+```
+
+## Corrected
 
@@
 ```{r}
-df9610 <- df9610 %>% 
+df.corrected <- bind_rows(df9610, df1112) %>% 
+  filter(year > 1995 & year < 2013) %>% 
   group_by(pc, year) %>% 
-  summarise(n = n()) %>% 
+  summarise(n = sum(V1.domain)) %>% 
   ungroup() %>% 
-  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2003 & year <2007, NA, n))
-
-df9610 <- pdata.frame(df9610)
+  # The next line is different than the same for  n = 1: problem.pcs$pc
+  mutate(n = ifelse(pc %in% as.character(problem.pcs) & year > 2003 & year <2007, NA, n))
+
+df.corrected <- pdata.frame(df.corrected)
 predictions <- round(predict(plm(n ~ as.factor(year), effect = "individual", 
-                                 model = "within", data = df9610), newdata = df9610), 2)
-
-df9610$n[is.na(df9610$n)] <- predictions[is.na(df9610$n)]
-df9610 %>% filter(pc == "M28 2SL") %>% distinct(year)
-```
-
-## Load remaining data
-
-```{r}
-df9610 <- as_tibble(df9610) %>% 
-  mutate(year = as.double(as.character(year))) %>% 
+                                 model = "within", data = df.corrected), newdata = df.corrected), 2)
+
+df.corrected$n[is.na(df.corrected$n)] <- predictions[is.na(df.corrected$n)]
+```
+
+## OA df
+
+```{r}
+df <- df.corrected %>% 
+  mutate(year = as.integer(as.character(year))) %>% 
+  filter(year > 1995 & year < 2013) %>% 
   left_join(lookup, by = "pc", suffix =c("","")) %>% 
   group_by(year, oa11cd) %>%
-  summarise(n = n()) %>%            
-  ungroup()
-
-data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
-#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
-#This is what we use for the hyperlinks paper as per George's script
-
-df1112 <- read_csv(data.path2011_2012) %>% 
-  filter(
-    V1.domain < m,               # for multiple pc
-    #V1.domain == n,             # for n == 1
-    year > 1995 & year <2013) %>%   
-  left_join(lookup, by = "pc", suffix =c("","")) %>% 
-  group_by(year, oa11cd) %>%
-  summarise(n = n()) %>%             # for n == 1  
-  #summarise(n = sum(V1.domain)) %>% # for multiple pc  
-  ungroup()
-
-df <- bind_rows(df9610, df1112) %>% 
-  filter(!is.na(oa11cd)) %>% 
+  summarise(n = sum(n)) %>% # for multiple pc  
+  ungroup() %>% 
   complete(oa11cd, year) %>% 
+  filter(!is.na(oa11cd)) %>% # drop c. 10 - 200 per year, which are not assigned to a LAD
   replace(is.na(.), 0) 
-
-# tests for Scotland
-# df %>% filter(substr(oa11cd, 1,1) =="S",
-#               year==2010) %>% summarise(mean(n, na.rm = T))
-# 
-# df %>% filter(oa11cd=="S00090182")
-# 
-# test@data %>% filter(substr(id, 1,1) =="S",
-#                      n!=0)
-# 
-# df %>% filter(substr(oa11cd, 1,1) =="W",
-#               year==2003)
 ```
@@
 #   mtry  splitrule   RMSE      Rsquared   MAE      
-#   2     variance    5.341123  0.3211637  1.294570
-#   2     extratrees  5.114541  0.2727356  1.288928
-#   5     variance    6.248341  0.3136499  1.380152
-#   5     extratrees  5.584957  0.2834177  1.397949
-#   8     variance    6.801851  0.2977672  1.395883
-#   8     extratrees  5.957788  0.2811840  1.444036
+#   2     variance    1.221299  0.4082199  0.5829598
+#   2     extratrees  1.308769  0.3297824  0.6136490
+#   5     variance    1.203082  0.4241359  0.5703508
+#   5     extratrees  1.280883  0.3513153  0.6124031
+#   8     variance    1.214268  0.4155308  0.5674246
+#   8     extratrees  1.284107  0.3502621  0.6180390
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule
-#  = extratrees and min.node.size = 5.
+# The final values used for the model were mtry = 5, splitrule = variance
+#  and min.node.size = 5.
 

### 2024-08-12 src/oa_RF_corrected.Rmd
corrected regions

@@
 
-THE BELLOW HAS NOT BEEN RUN NEITHER THE FILENAMES HAS BEEN CHANGED (add: _corrected)
-
 ### train the model in a loop for all but one regions
+
+64 GB of swap memory to run the below
 
@@
   #file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
-  file <- paste0("/hdd/tmp/regions/region_", as.character(i), ".RData")  
+  file <- paste0("/hdd/tmp/regions/corrected/regions_corrected_", as.character(i), ".RData")  
   save(model.all, file = file)  # file name = region not in the training data
@@
 
-### Resamples for plots
-
-```{r, eval=F}
-# create a list of the above model outputs
-filenames <- list.files(paste0(path, "/outputs/rf/models/oa_rf/regions"), pattern = "*.RData", full.names = T)
-
-for (i in filenames){
-  load(file = i)
-  i <- gsub("/home/nw19521/projects/web.diffusion/outputs/rf/models/oa_rf/regions/|.RData", "", i)
-  assign(i, model.all)  
-}
-
- fit.model.all <- mget(ls(pattern = "^region_"))
-
-# clear up
-rm(model.all) # this is the last item from the loop
-rm(list = ls(pattern = "^region_"))
-
-# resamples
-fit.model.all.res <- resamples(fit.model.all)
-summary(fit.model.all.res)
-
-# plot metrics
-
-# TODO: fix the year labels
-
-scales <- list(x=list(relation="free"), y=list(relation="free"))
-bwplot(fit.model.all.res, scales=scales)
-
-# TODO: fix the year labels
-```
-
-### Variable importance
-
-```{r, eval=F, include=TRUE, results= 'markup', message=FALSE}
-lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
-```
-
-```{r, eval=F}
-help <- lapply(fit.model.all, function(x) varImp(x)) 
-help <- lapply(help, function(x) x[[1]])
-help <- do.call("rbind", help)
-
-importance <- help %>% rownames_to_column() %>% 
-  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
-         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
-  rename(importance = Overall) %>% 
-  relocate(importance, .after = year) %>% 
-  dplyr::select(-rowname)
-
-importance %>% ggplot() +
-  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
-  scale_color_discrete(name = "Features") +
-  #scale_linetype_manual(name = "Website density regressions111") 
-  ylab("Importance") + 
-  ggtitle("Feature importance") +
-  theme_minimal() +
-  theme(axis.title.x=element_blank(),
-        plot.title = element_text(hjust = 0.5))
-
-```
+THE BELLOW HAS NOT BEEN RUN NEITHER THE FILENAMES HAS BEEN CHANGED (add: _corrected)
 
@@
 
-new.path <- "/hdd/tmp/regions/"
+new.path <- "/hdd/tmp/regions/corrected/"
 #filenames <- list.files(paste0(path, "/outputs/rf/models/oa_rf/regions"), pattern = "*.RData", full.names = T)
@@
 for (i in regions){
-  load(file = paste0(new.path, "region_", as.character(i), ".RData"))
+
+  print(i)
+
+  load(file = paste0(new.path, "regions_corrected_", as.character(i), ".RData"))
   #model.name <- gsub("/home/nw19521/projects/web.diffusion/outputs/rf/models/oa_rf/regions/|.RData", "", i)
@@
 
-path.out <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
+path.out <- paste0(path, "/outputs/rf/figures/test_regions_OA_corrected.csv")
 rf.year.all.metrics %>% 

### 2024-08-12 src/oa_RF_corrected_10.Rmd
corrected regions

@@
 
-THE BELLOW HAS NOT BEEN RUN NEITHER THE FILENAMES HAS BEEN CHANGED (add: _corrected)
-
 ### train the model in a loop for all but one regions
+
+```{r}
+rm(list=setdiff(ls(), c("df", "path")))
+```
+
+64 GB of swap memory to run the below
 
@@
 
-  #file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
-  file <- paste0("/hdd/tmp/regions/region_", as.character(i), ".RData")  
+  #file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_corrected_", as.character(i), ".RData")
+  file <- paste0("/hdd/tmp/regions/corrected_10/regions_corrected_10", as.character(i), ".RData")  
   save(model.all, file = file)  # file name = region not in the training data
@@
 
-### Resamples for plots
-
-```{r, eval=F}
-# create a list of the above model outputs
-filenames <- list.files(paste0(path, "/outputs/rf/models/oa_rf/regions"), pattern = "*.RData", full.names = T)
-
-for (i in filenames){
-  load(file = i)
-  i <- gsub("/home/nw19521/projects/web.diffusion/outputs/rf/models/oa_rf/regions/|.RData", "", i)
-  assign(i, model.all)  
-}
-
- fit.model.all <- mget(ls(pattern = "^region_"))
-
-# clear up
-rm(model.all) # this is the last item from the loop
-rm(list = ls(pattern = "^region_"))
-
-# resamples
-fit.model.all.res <- resamples(fit.model.all)
-summary(fit.model.all.res)
-
-# plot metrics
-
-# TODO: fix the year labels
-
-scales <- list(x=list(relation="free"), y=list(relation="free"))
-bwplot(fit.model.all.res, scales=scales)
-
-# TODO: fix the year labels
-```
-
-### Variable importance
-
-```{r, eval=F, include=TRUE, results= 'markup', message=FALSE}
-lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
-```
-
-```{r, eval=F}
-help <- lapply(fit.model.all, function(x) varImp(x)) 
-help <- lapply(help, function(x) x[[1]])
-help <- do.call("rbind", help)
-
-importance <- help %>% rownames_to_column() %>% 
-  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
-         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
-  rename(importance = Overall) %>% 
-  relocate(importance, .after = year) %>% 
-  dplyr::select(-rowname)
-
-importance %>% ggplot() +
-  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
-  scale_color_discrete(name = "Features") +
-  #scale_linetype_manual(name = "Website density regressions111") 
-  ylab("Importance") + 
-  ggtitle("Feature importance") +
-  theme_minimal() +
-  theme(axis.title.x=element_blank(),
-        plot.title = element_text(hjust = 0.5))
-
-```
-
 ### Test on one region
+
+THE RESULTS ARE WEIRD. REVISIT.
 
@@
 
-regions <- df %>% ungroup() %>% distinct(RGN11CD) 
+# correct for non-England file names
+df <- df %>% 
+  mutate(RGN11CD = ifelse(RGN11CD == "Wales", "W0",
+                          ifelse(RGN11CD == "Scotland", "S1",
+                                 ifelse(RGN11CD == "NI", "N0", RGN11CD))))
+
+regions <- df %>% 
+  ungroup() %>% 
+  distinct(RGN11CD)  
 regions <- as.vector(regions$RGN11CD)
 
-new.path <- "/hdd/tmp/regions/"
+new.path <- "/hdd/tmp/regions/corrected_10/"
 #filenames <- list.files(paste0(path, "/outputs/rf/models/oa_rf/regions"), pattern = "*.RData", full.names = T)
@@
 for (i in regions){
-  load(file = paste0(new.path, "region_", as.character(i), ".RData"))
+
+  print(i)
+
+  load(file = paste0(new.path, "regions_corrected_10", as.character(i), ".RData"))
   #model.name <- gsub("/home/nw19521/projects/web.diffusion/outputs/rf/models/oa_rf/regions/|.RData", "", i)
@@
   rm(model.all)
+
+  print(i)
+
 }
@@
   rename(RGN11CD = RGN21CD) %>% 
-  add_row(RGN11CD = c("Scotland", "NI", "Wales"), 
+  add_row(RGN11CD = c("S1", "N0", "W0"), 
           RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
 
-path.out <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
+path.out <- paste0(path, "/outputs/rf/figures/test_regions_OA_corrected_10.csv")
 rf.year.all.metrics %>% 

### 2024-08-12 src/problem_years_10.Rmd
corrected regions

@@
 
-path.out <- paste0(path, "/outputs/pc_year_1.png")
+path.out <- paste0(path, "/outputs/pc_year_10.png")
 ggsave(path.out, bg='#ffffff')  
@@
 
-path.out <- paste0(path, "/outputs/pc_year_1_corrected.png")
+path.out <- paste0(path, "/outputs/pc_year_10_corrected.png")
 ggsave(path.out, bg='#ffffff')  

### 2024-08-12 src/problem_years.Rmd
LAD lisa

@@
 n = 1 #number of unique postcodes. 
-m = 11
+#m = 11
 
@@
 
+path.out <- paste0(path, "/outputs/pc_year_1.png")
+
 df.long %>% 
@@
   ylab("N. of websites per postcode") +
-  scale_x_continuous(breaks=seq(1996, 2012, 1)) +
+  scale_x_continuous(NULL, breaks=seq(1996, 2012, 1)) +
   scale_y_continuous(labels=scales::comma) +
   theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_text_repel(aes(x = year, y = n, label=outlier), cex = 3)
+  theme(plot.title = element_text(hjust = 0.5),
+          panel.grid.minor.x = element_blank()) 
+  #geom_text_repel(aes(x = year, y = n, label=outlier), cex = 3)
 
-path.out <- paste0(path, "/outputs/pc_year_1.png")
 ggsave(path.out, bg='#ffffff')  
@@
 
+path.out <- paste0(path, "/outputs/pc_year_1_corrected.png")
+
 df.corrected %>% 
@@
   ylab("N. of websites per postcode") +
-  scale_x_continuous(breaks=seq(1996, 2012, 1)) +
+  scale_x_continuous(NULL, breaks=seq(1996, 2012, 1)) +
   scale_y_continuous(labels=scales::comma) +
   theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5),
+          panel.grid.minor.x = element_blank())
 
-path.out <- paste0(path, "/outputs/pc_year_1_corrected.png")
 ggsave(path.out, bg='#ffffff')  
@@
 
-#spatial weight matrix
-#moved out of the loop
+#spatial weight matrix moved out of the loop
 
@@
       axis.ticks = element_blank(),
-        rect = element_blank())
+      rect = element_blank(),
+      plot.margin = unit(c(0, 0, 0, 0), "cm"))
 

### 2024-08-12 src/problem_years.Rmd
LAD lisa

@@
   #guides(fill = guide_legend(reverse = TRUE)) +
-  labs(title = "LISA, websites per firm in a LA") + 
-  theme(legend.title=element_blank())
+  labs(title = paste0("LAD, ", t)) + #"LISA, websites per firm in a LA") + 
+  theme(legend.title=element_blank(),
+        plot.margin = unit(c(0, 0, 0, 0), "cm"),
+        plot.title = element_text(hjust = 0.5))
 
-theme(axis.text.x = element_blank(),
-      axis.text.y = element_blank(),
-      axis.ticks = element_blank(),
-      rect = element_blank(),
-      plot.margin = unit(c(0, 0, 0, 0), "cm"))
+# theme(axis.text.x = element_blank(),
+#       axis.text.y = element_blank(),
+#       axis.ticks = element_blank(),
+#       rect = element_blank())
 
@@
 }
+
+# legend
+path.out <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc", n, "_la_legend.png") 
+legend <- ggplot(data = la, aes(fill = quadr$median)) +
+  geom_sf() +
+  # scale_fill_discrete_c4a_cat(name = "%", palette = "parks.charmonix",
+  #                             na.translate = FALSE) +
+  scale_fill_manual(values = palette_new, na.value="white", drop = F, name = NULL) +
+  theme_minimal() +
+  theme(legend.position="bottom")
+
+legend <- cowplot::get_legend(for.legend)
+plot(legend)  
+ggsave(path.out, bg='#ffffff')
 ```

### 2024-08-13 src/problem_years.Rmd
legend

@@
 library(scales)
+library(ggpubr)
 
@@
 }
+```
 
-# legend
+### lisa map legend
+```{r}
 path.out <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc", n, "_la_legend.png") 
@@
   geom_sf() +
-  # scale_fill_discrete_c4a_cat(name = "%", palette = "parks.charmonix",
-  #                             na.translate = FALSE) +
   scale_fill_manual(values = palette_new, na.value="white", drop = F, name = NULL) +
   theme_minimal() +
-  theme(legend.position="bottom")
+  theme(legend.position="bottom",
+        legend.margin=ggplot2::margin(c(1,1,1,1)))
 
 legend <- cowplot::get_legend(for.legend)
-plot(legend)  
-ggsave(path.out, bg='#ffffff')
+
+leg <- as_ggplot(legend)
+leg <- leg + theme(
+  legend.margin=ggplot2::margin(c(0,0,0,0)))
+ggsave(path.out, width = 5, height = 0.4, bg='#ffffff')
 ```

### 2024-08-13 src/problem_years.Rmd
legend

--- 2024-08-12 src/problem_years.Rmd
@@
   #guides(fill = guide_legend(reverse = TRUE)) +
-  labs(title = "LISA, websites per firm in a LA") + 
-  theme(legend.title=element_blank())
+  labs(title = paste0("LAD, ", t)) + #"LISA, websites per firm in a LA") + 
+  theme(legend.title=element_blank(),
+        plot.margin = unit(c(0, 0, 0, 0), "cm"),
+        plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(size=20, face = "bold"))
 
-theme(axis.text.x = element_blank(),
-      axis.text.y = element_blank(),
-      axis.ticks = element_blank(),
-      rect = element_blank(),
-      plot.margin = unit(c(0, 0, 0, 0), "cm"))
+# theme(axis.text.x = element_blank(),
+#       axis.text.y = element_blank(),
+#       axis.ticks = element_blank(),
+#       rect = element_blank())
 
@@
 }
+```
+
+### lisa map legend
+```{r}
+path.out <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc", n, "_la_legend.png") 
+legend <- ggplot(data = la, aes(fill = quadr$median)) +
+  geom_sf() +
+  scale_fill_manual(values = palette_new, na.value="white", drop = F, name = NULL) +
+  theme_minimal() +
+  theme(legend.position="right",
+        legend.margin=ggplot2::margin(c(1,1,1,1)))
+
+legend <- cowplot::get_legend(legend)
+
+leg <- as_ggplot(legend)
+leg <- leg + theme(
+  legend.margin=ggplot2::margin(c(0,0,0,0)))
+#ggsave(path.out, width = 5, height = 0.4, bg='#ffffff')
+ggsave(path.out, width = 1, height = 1.4, bg='#ffffff')
 ```

### 2024-08-13 src/oa_RF_corrected.Rmd
LAD lisa
oa lisa

@@
                    importance = "impurity")
-  #assign(paste0("model.all",as.character(i)), model.all)
 
-  #file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
   file <- paste0("/hdd/tmp/regions/corrected/regions_corrected_", as.character(i), ".RData")  
@@
 ```
-
-THE BELLOW HAS NOT BEEN RUN NEITHER THE FILENAMES HAS BEEN CHANGED (add: _corrected)
 

### 2024-08-13 src/oa_RF_corrected_10.Rmd
LAD lisa
oa lisa

@@
                    importance = "impurity")
-  #assign(paste0("model.all",as.character(i)), model.all)
-
-  #file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_corrected_", as.character(i), ".RData")
+
   file <- paste0("/hdd/tmp/regions/corrected_10/regions_corrected_10", as.character(i), ".RData")  
@@
 THE RESULTS ARE WEIRD. REVISIT.
+
+The corrected for n = 1 makes sense. Check the name changes below.
 

### 2024-08-13 src/problem_years.Rmd
oa lisa

@@
 
-# theme(axis.text.x = element_blank(),
-#       axis.text.y = element_blank(),
-#       axis.ticks = element_blank(),
-#       rect = element_blank())
-
 filename = paste(path.out, t, ".png", sep = "")
@@
 ### lisa map legend
+
 ```{r}
@@
 
-  #spatial weight matrix
-  #moved out of the loop
+  #spatial weight matrix moved out of the loop
 
@@
 
-  cities <- maps::world.cities %>%
-    filter(country.etc=="UK") %>% 
-    arrange(pop) %>% tail(10) 
-  tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
+  # cities <- maps::world.cities %>%
+  #   filter(country.etc=="UK") %>% 
+  #   arrange(pop) %>% tail(10) 
+  # cities <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
 
@@
     geom_sf(linewidth = 0) +
-    geom_sf(data = tif_sf, fill = "black", size = .3) +
-    #geom_sf_text(data = tif_sf, aes(label = name, fill = "black"), size = 4) +
+    #geom_sf(data = cities, fill = "black", size = .3) +
+    #geom_sf_text(data = cities, aes(label = name, fill = "black"), size = 4) +
     scale_fill_manual(values = palette_new, na.value="white", drop = F) +
@@
           axis.ticks = element_blank(),
-          legend.title=element_blank()) +
-          #legend.position="none") +
-    guides(fill = guide_legend(reverse = TRUE)) +
-    labs(title = t,
-        subtitle = "LISA, websites per OA")
+          legend.position="none") +
+    #guides(fill = guide_legend(reverse = TRUE)) +
+    labs(title = paste0("OA, ", t)) +
+    theme(legend.title=element_blank(),
+        plot.margin = unit(c(0, 0, 0, 0), "cm"),
+        plot.title = element_text(hjust = 0.5)) +
+    theme(plot.title = element_text(size=20, face = "bold"))
 

### 2024-08-14 src/la_RF_corrected.Rmd
sum corrected

@@
   group_by(year, ladcd) %>%
-  summarise(n = n()) %>% # for multiple pc  
+  summarise(n = sum(n)) %>% # **********ATTENTION*********    
   ungroup() %>% 

### 2024-08-14 src/old/msoa_RF.Rmd
sum corrected

--- 2024-01-27 src/oa_RF.Rmd
@@
 ---
-title: "Random Forests", OAs
+title: "Random Forests, MSOAs
 date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 library(rprojroot)
-#library(rgdal)
-#library(rgeos)
 library(sf)
 library(spdep)
-#library(tmap)
-#library(maptools)
 library(knitr)
@@
 library(doParallel)
-library(raster)
-library(plm)
-#library(lmtest)
+#library(raster)
 library(caret)
@@
 lookup <- read_csv(path.lookup) %>% 
-  dplyr::select(pcds, oa11cd) %>% 
+  dplyr::select(pcds, msoa11cd) %>% 
   dplyr::rename(pc = pcds)
@@
   left_join(lookup, by = "pc", suffix =c("","")) %>% 
-  group_by(year, oa11cd) %>%
+  group_by(year, msoa11cd) %>%
   #summarise(n = n()) %>%            # for multiple pc  
@@
 
-# Partially complete panel as not all OAs have a website in at least one year
-df <- df %>% filter(!is.na(oa11cd)) %>% 
-  complete(oa11cd, year) %>% 
+# Partially complete panel as not all MSOAs have a website in at least one year
+df <- df %>% filter(!is.na(msoa11cd)) %>% 
+  complete(msoa11cd, year) %>% 
   replace(is.na(.), 0)
@@
 
-## spatial data OLD
-
-```{r eval=F}
-# get OA for England and Wales
-path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
-oa.ew <- readOGR(path.geo)
-# source: https://geoportal.statistics.gov.uk/
-
-# spatial transformations
-oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))
-
-# keep in the data slot only the ONS Output Area id, renaming it as 'id'
-oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
-colnames(oa.ew@data) <- c('id', 'area')
-
-# reassign the polygon IDs
-oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))
-
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.ew)
-
-# # get OA for Scotland
-# path.geo.sc <- paste0(path, "/data/raw/SG_DataZoneBdry_2011")
-# oa.sc <- readOGR(dsn=path.geo.sc, layer = "SG_DataZone_Bdry_2011")
-# # source: https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011
-# 
-# # spatial transformations
-# oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-# 
-# # Scotland (follows same steps as EW, see notes above)
-# oa.sc <- oa.sc[, 'DataZone']
-# colnames(oa.sc@data) <- c('id')
-# 
-# # reassign the polygon IDs
-# oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-# 
-# # check the CRS has changed correctly, and the data slot has shrink to only the ID
-# summary(oa.sc)
-
-path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
-oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
-# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries
-
-# spatial transformations
-oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
-
-# Scotland (follows same steps as EW, see notes above)
-oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
-colnames(oa.sc@data) <- c('id', 'area')
-
-# reassign the polygon IDs
-oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
-
-# check the CRS has changed correctly, and the data slot has shrink to only the ID
-summary(oa.sc)
-
-# build oa for GB
-oa.gb <- maptools::spRbind(oa.ew, oa.sc)
-rm(oa.ew, oa.sc)
-
-# oa.gb$geometry <- oa.gb$geometry %>%
-#   s2::s2_rebuild() %>%
-#   sf::st_as_sfc()
-
-# get GB
-path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
-gb <- readOGR(path.gb)
-# spatial transformations
-gb <- spTransform(gb, CRS("+init=epsg:4326"))
-```
-
-## spatial data NEW
-
-```{r}
-library(sf)
-
+## spatial data
+
+```{r}
 # Get OA for England and Wales
-path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
-oa.ew <- st_read(path.geo)
+path.geo <- paste0(path, "/data/raw/infuse_msoa_lyr_2011_clipped/infuse_msoa_lyr_2011_clipped.shp")
+msoa <- st_read(path.geo) %>% 
+  dplyr::select(geo_code, geo_label) %>% 
+  rename(id = geo_code)
 
 # Spatial transformations
-oa.ew <- st_transform(oa.ew, 4326)  # EPSG code for WGS84
-
-# Rename columns and keep only necessary columns
-oa.ew <- oa.ew %>% dplyr::select(OA11CD, Shape__Area) %>% 
-  rename(id = OA11CD,
-         area = Shape__Area)
-# oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
-# colnames(oa.ew) <- c('id', 'area')
-
-# Check the data structure
-print(summary(oa.ew))
-
-# Get OA for Scotland
-path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
-oa.sc <- st_read(dsn = path.geo.sc, layer = "OutputArea2011_MHW")
-
-# Spatial transformations
-oa.sc <- st_transform(oa.sc, 4326)  # EPSG code for WGS84
-
-# Rename columns and keep only necessary columns
-oa.sc <- oa.sc %>% dplyr::select(code, SHAPE_1_Ar) %>% 
-  rename(id = code,
-         area = SHAPE_1_Ar)
-# oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
-# colnames(oa.sc) <- c('id', 'area')
-
-# Check the data structure
-print(summary(oa.sc))
-
-# NI
-path.geo.ni <- paste0(path, "/data/raw/ni_small_area/")
-oa.ni <- st_read(dsn = path.geo.ni, layer = "SA2011")
-oa.ni <- st_transform(oa.ni, 4326)  # EPSG code for WGS84
-oa.ni <- oa.ni %>% dplyr::select(SA2011, Hectares) %>% 
-  rename(id= SA2011,
-         area = Hectares)
-
-# Build OA for UK
-oa.uk <- rbind(oa.ew, oa.sc, oa.ni)
-rm(oa.ew, oa.sc, oa.ni)
+msoa <- st_transform(msoa, 4326)  # EPSG code for WGS84
 
@@
 
-oa.full <- replicate(17, oa.uk %>% as_tibble(), simplify = FALSE) %>%
+msoa.full <- replicate(17, msoa %>% as_tibble(), simplify = FALSE) %>%
   bind_rows(.id = 'id_') %>% 
@@
 
-df <- oa.full %>% left_join(df, by = c("id" = "oa11cd", "year" = "year")) %>% 
+df <- msoa.full %>% left_join(df, by = c("id" = "msoa11cd", "year" = "year")) %>% 
   replace(is.na(.), 0)
@@
 sapply(df, function(x) sum(is.na(x)))
-rm(oa.full)
+rm(msoa.full)
 ```
@@
 ```{r}
-# oa centroids
-# oa.uk.c <- (rgeos::gCentroid(oa.uk, byid=TRUE))@coords %>% as.data.frame()
-# oa.uk.c$id <- row.names(oa.uk.c)  #as_tibble(rownames = NA) %>% 
-#   mutate(id = row.names())
-
-# oa.uk.c <- gCentroid(oa.uk, byid=TRUE)@coords %>% 
-#   as.data.frame() %>% 
-#   add_rownames(var = "id")
 sf_use_s2(FALSE)
-oa.uk.c <- st_centroid(oa.uk) %>%
+msoa.c <- st_centroid(msoa) %>%
   st_coordinates() %>%
   as.data.frame() %>%
-  #add_rownames(var = "id") %>% 
-  bind_cols(oa.uk %>% dplyr::select(id))
+  bind_cols(msoa %>% dplyr::select(id))
 
@@
 "Edinburgh, UK",
-"Leicester, UK",
-"Belfast, UK")
+"Leicester, UK")
 
@@
 #   "Edinburgh",
-#   "Leicester",
+#   "Leicester"
 #   )
@@
   st_transform(4326) %>% 
-  filter(grepl('Edinburgh|Glasgow|Belfast', LAD21NM)) %>% 
+  filter(grepl('Edinburgh|Glasgow', LAD21NM)) %>% 
   dplyr::select(OBJECTID, LAD21CD, LAD21NM) %>% 
@@
 
-### Retail centres
-
-```{r}
-# retail centres <- 
-geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
-#retail <-readOGR(geo.path) 
-retail <- st_read(geo.path) %>% st_transform(4326)  # EPSG code for WGS84
-
-#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
-# retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
-#   as.data.frame() %>% 
-#   add_rownames(var = "id") #%>%
-retail.major.cetres.help <- st_centroid(retail) %>% 
-  #st_transform(4326) %>%
-  st_coordinates() %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id")
-
-# retail.major.cetres <- retail.major.cetres.help %>% 
-#   st_as_sf(coords = c("x", "y"), crs = 27700) %>%
-#   st_transform(4326) %>%
-#   st_coordinates() %>%
-#   as_tibble() %>% 
-#   bind_cols(retail.major.cetres.help$id) %>% 
-#   rename(id = '...3') %>% 
-#   left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
-retail.major.cetres <- retail.major.cetres.help %>% 
-  # as_tibble() %>% 
-  # bind_cols(retail.major.cetres.help$id) %>% 
-  # #rename(id = '...3') %>% 
-  left_join(retail %>% rownames_to_column(var = "id"), by = 'id')
-```
-
 ### Distance to cities
@@
 ```{r}
-# NI
-oa.uk.c.ni <- oa.uk.c %>% filter(str_detect(id, "^N"))
-
-dist.ni <- distm(cbind(oa.uk.c.ni$X, oa.uk.c.ni$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist.ni <- round((dist.ni/1000),2) %>% 
+# GB
+dist <- distm(cbind(msoa.c$X, msoa.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist <- round((dist/1000),2) %>% 
   as_tibble()
 
-names(dist.ni) <- cities$address #city.names
-
-dist.ni <- dist.ni %>% bind_cols(oa.uk.c.ni$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
+names(dist) <- cities$address #city.names
+
+dist <- dist %>% bind_cols(msoa.c$id) %>% 
+  rename(msoa11cd = last_col()) %>% 
+  relocate(msoa11cd)
 
@@
 # min distance
-dist.ni <- transform(dist.ni, dist = do.call(pmin, dist.ni[-1]))
+dist <- transform(dist, dist = do.call(pmin, dist[-1]))
 
 # City name for minimum distance 
-dist.ni <- dist.ni %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
-
-# GB
-oa.uk.c.gb <- oa.uk.c %>% filter(!str_detect(id, "^N"))
-
-dist.gb <- distm(cbind(oa.uk.c.gb$X, oa.uk.c.gb$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist.gb <- round((dist.gb/1000),2) %>% 
-  as_tibble()
-
-names(dist.gb) <- cities$address #city.names
-
-dist.gb <- dist.gb %>% bind_cols(oa.uk.c.gb$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
-# dist$distMet <- apply(dist[,2:11], 1, min)
-
-# min distance
-dist.gb <- transform(dist.gb, dist = do.call(pmin, dist.gb[-1]))
-
-# City name for minimum distance 
-dist.gb <- dist.gb %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
-
-# UK
-dist <- rbind(dist.ni, dist.gb)
-
-# calculate area
-#oa.uk$area <- st_area(oa.uk) #/ 1000000
+dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
 
 # Join with distance and area
-df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
-  rename(oa11cd=id)
+df <- df %>% left_join(dist, by = c("id" = "msoa11cd")) %>% 
+  rename(msoa11cd=id)
 
@@
 
-### Distance to retail centres
-
-```{r}
-# dist.retail <- distm(cbind(oa.uk.c$X, oa.uk.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-# dist.retail <- round((dist.retail/1000),2) %>% 
-#   as_tibble()
-
-# NI
-oa.uk.c.ni <- oa.uk.c %>% filter(str_detect(id, "^N"))
-dist.retail.ni <- distm(cbind(oa.uk.c.ni$X, oa.uk.c.ni$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail.ni <- round((dist.retail.ni/1000),2) %>% 
-  as_tibble()
-
-names(dist.retail.ni) <- retail.major.cetres$RC_ID
-
-dist.retail.ni <- dist.retail.ni %>% bind_cols(oa.uk.c.ni$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# This efficiently returns the name of the column with the shortest distance
-dist.retail.ni <- dist.retail.ni %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])
-
-# dist.retail_ <- dist.retail %>% slice_sample(n=100)
-# dist.retail_[,6425:6426]
-# dist.retail_ <- transform(dist.retail_, dist.retail = do.call(pmin, dist.retail_[2:6424])) 
-# sum(is.na(dist.retail_$dist.retail.name))
-
-# Minimum distance 
-dist.retail.ni <- transform(dist.retail.ni, dist.retail = do.call(pmin, dist.retail.ni[2:6424])) 
-dist.retail.ni <- dist.retail.ni  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)
-
-# GB
-oa.uk.c.gb <- oa.uk.c %>% filter(!str_detect(id, "^N"))
-dist.retail.gb <- distm(cbind(oa.uk.c.gb$X, oa.uk.c.gb$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail.gb <- round((dist.retail.gb/1000),2) %>% 
-  as_tibble()
-
-names(dist.retail.gb) <- retail.major.cetres$RC_ID
-
-dist.retail.gb <- dist.retail.gb %>% bind_cols(oa.uk.c.gb$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# This efficiently returns the name of the column with the shortest distance
-dist.retail.gb <- dist.retail.gb %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])
-
-# dist.retail_ <- dist.retail %>% slice_sample(n=100)
-# dist.retail_[,6425:6426]
-# dist.retail_ <- transform(dist.retail_, dist.retail = do.call(pmin, dist.retail_[2:6424])) 
-# sum(is.na(dist.retail_$dist.retail.name))
-
-# Minimum distance 
-dist.retail.gb <- transform(dist.retail.gb, dist.retail = do.call(pmin, dist.retail.gb[2:6424])) 
-dist.retail.gb <- dist.retail.gb  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)
-
-# UK
-dist.retail <- rbind(dist.retail.ni, dist.retail.gb)
-
-# Join with complete panel
-df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd")) 
-#sapply(df, function(x) sum(is.na(x)))
-```
-
 ## n for London, nearest city and retail
 
-### TODELETE
-
-```{r, eval=FALSE}
-# ## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
-# #df <- df %>% filter(!is.na(area)) 
-# 
-# cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
-# cities.sf <- st_set_crs(cities.sf, 4326)
-# 
-# # help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
-# #   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
-# 
-# sf_use_s2(FALSE)
-# help1 <- st_filter( oa.gb, cities.sf) %>% 
-#   #st_coordinates() %>% 
-#   as_tibble() %>% 
-#   dplyr::select(id) %>% 
-#   mutate(cities = c("London",     # The city names have been manually checked and assigned
-#                     "Leicester",
-#                     "Leeds",
-#                     "Sheffield",
-#                     "Bristol",
-#                     "Birmingham",
-#                     "Manchester",
-#                     "Liverpool",
-#                     "Edinburgh",
-#                     "Glasgow"))
-# 
-# # too slow
-# #dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
-# 
-# # This efficiently returns the name of the column with the shortest distance
-# dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
-# 
-# help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
-#   #rename(nearest.city.oa11cd = id) %>% 
-#   dplyr::select(oa11cd, dist.city.name)
-# 
-# df <- df %>%  left_join(help2, by = c("oa11cd" = "oa11cd")) 
-```
-
 ### n for cities
 
-```{R}
-#help.retail <- dist.retail
-
-# n nearest city
-# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-#                          rename(n.nearest.city = n), 
-#                        by = c("nearest.city.oa11cd" = "oa11cd",
-#                               "year" = "year"))
-
+```{r}
 sf_use_s2(FALSE)
-city.boundaries.oa <- st_join(oa.uk, city.boundaries, join = st_intersects) %>% 
+city.boundaries.msoa <- st_join(msoa, city.boundaries, join = st_intersects) %>% 
   as_tibble() %>% 
@@
 
-city.boundaries.help <- df %>% inner_join(city.boundaries.oa, by = c("oa11cd" = "id")) %>% 
+city.boundaries.help <- df %>% inner_join(city.boundaries.msoa, by = c("msoa11cd" = "id")) %>% 
   group_by(name, year) %>% 
@@
 
-#sapply(df, function(x) sum(is.na(x)))
-```
-
-### DELETE n for London
-
-```{r, eval=FALSE}
-
-# # OLD London definitions
-# # df <- df %>% left_join(df %>% 
-# #                          ungroup() %>% 
-# #                          filter(oa11cd == "E00176659") %>% 
-# #                          dplyr::select(year, n) %>% 
-# #                          rename(n.London = n)) %>% 
-# #   group_by(oa11cd) %>% 
-# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-# #   mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
-# #   ungroup() 
-# # 
-# # london.help <- df %>% group_by(RGN11NM, year) %>% 
-# #   summarise(n.London = sum(n)) %>% 
-# #   filter(RGN11NM == "London") %>%
-# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-# #   ungroup() %>% 
-# #   dplyr::select(-RGN11NM) 
-# 
-# london.oa <- st_filter(oa.gb, london.sf %>% 
-#                          filter(Name == "Central")) %>% 
-#   as_tibble() %>% 
-#   dplyr::select(id)
-# 
-# london.help <- df %>% inner_join(london.oa, by = c("oa11cd" = "id")) %>% 
-#   group_by(year) %>% 
-#   summarise(n.London = sum(n)) %>% 
-#   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-#   ungroup()
-# 
-# df <- df %>% left_join(london.help, by = c("year" = "year"))
-```
-
-### n for nearest retail
-
-```{r}
-# Old definition for retail centre
-# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-#                          rename(n.nearest.retail = n), 
-#                        by = c("oa11cd.retail" = "oa11cd",
-#                               "year" = "year")) %>% 
-#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
-# 
-# 
-# #n.nearest.retail IS WRONG. FIX UPSTREAM
-# df <- df %>% 
-#   rename(oa11cd.retail = n.nearest.retail) %>% 
-#   dplyr::select(-n.nearest.retail.lag) %>% 
-#   left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-#                          rename(n.nearest.retail = n), 
-#                        by = c("oa11cd.retail" = "oa11cd",
-#                               "year" = "year")) %>% 
-#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
-# 
-# 
-# df <- df %>% dplyr::select(-n.London, -n.London.lag) %>% 
-#   left_join(london.help, by = c('year' = 'year'))
-
-retail.oa <- st_join(oa.uk, retail, 
-                     join = st_intersects) %>% #, 
-                    #largest = TRUE) %>% # returns the x features augmented with 
-                                         # the fields of y that have the largest 
-                                         # overlap with each of the features of x
-  as_tibble() %>% 
-  filter(!is.na(RC_ID)) %>% 
-  dplyr::select(id, RC_ID) %>% 
-  add_row(id = "E00019779", RC_ID = "RC_EW_5951")
-
-# If `larget = FALSE`, we have multiple RC_IDs for OAs.
-# Also, RC_ID="RC_EW_5951" would fall in Thames as it is the London Bridge Tube Station.
-# To avoid NAs for for n.nearest.retail, I would have assigned dist.retail.name = RC_EW_1935 for E00019779
-
-# st_join(oa.uk, retail, join = st_is_within_distance, dist = 500) %>% 
-#   as_tibble() %>% 
-#   filter(!is.na(RC_ID)) %>% 
-#   dplyr::select(id, RC_ID)
-# 
-# st_filter(oa.uk, retail,  .pred = st_is_within_distance, dist = 100) %>% 
-#   as_tibble() %>% 
-#   filter(!is.na(RC_ID)) %>% 
-#   dplyr::select(id, RC_ID)
-
-# h <- oa.uk %>% st_intersects(retail)
-
-# retail.oa_ %>% filter(id == "E00004828"|id == "E00004829"|id == "E00004827")
-# # RC_EW_5285
-# retail.oa <- retail.oa %>% add_row(id = "E00004828", RC_ID = "RC_EW_5285")
-# retail.oa <- retail.oa %>% arrange(id)
-
-retail.help <- df %>% full_join(retail.oa, by = c("oa11cd" = "id")) %>% 
-  group_by(RC_ID, year) %>% 
-  summarise(n.nearest.retail = sum(n)) %>% 
-  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail)) %>% 
-  ungroup()
-
-# retail.help2 <- retail.oa %>% left_join(retail.help, by = c("RC_ID" = "RC_ID")) %>% 
-#   dplyr::select(-RC_ID)
-
-df <- df %>% left_join(retail.help, by = c("dist.retail.name" = "RC_ID", "year" = "year"))
-#sapply(df, function(x) sum(is.na(x)))
-```
-
-### Urban Rural TODO
-
-## TODO
-
-```{r}
-# Lookup: OA to regions
-# **FIXED< DELETE IN TEST** NEED TO ADD NI. NOT A PROBLEM FOR NOW AS IT IS FIXED IN THE TEST SECTION
-
-# # regional identifier
-# df_ <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
-#   mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
-#   ungroup() %>% 
-#   dplyr::select(-help)
-
-#sapply(df, function(x) sum(is.na(x)))
+sapply(df, function(x) sum(is.na(x)))
 ```
@@
 ```{r}
-# sf_use_s2(FALSE)
-# oa.c <- st_centroid(oa.uk)
-# kn <- knearneigh(oa.c, k = 5)
-# knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
-# knn.l <- nb2listw(knn)
-
-path.out <- paste0(path, "/data/temp/knnl.RData")
-# save(knn.l, file = path.out)
-load(path.out)
+sf_use_s2(FALSE)
+msoa.c <- st_centroid(msoa)
+kn <- knearneigh(msoa.c, k = 5)
+knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
+knn.l <- nb2listw(knn)
 
@@
   mutate(n.slag = lag.listw(knn.l, n)) %>% 
-  group_by(oa11cd) %>% 
+  group_by(msoa11cd) %>% 
   mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
          n.lag = dplyr::lag(n, n=1, order_by = year))        # duplicate line
+```
+
+```{r}
+# out.path <- paste0(path, "/data/temp/df_for_msoa_rf.csv")
+# df <- read_csv(out.path)
+# df %>% write_csv(file = out.path)
+# 
+# rm(list=setdiff(ls(), "df"))
+
+# This is the project path
+# path <- find_rstudio_root_file()
 ```
@@
 
-df <- df %>% left_join(lookup.region, by = c("oa11cd" = "OA11CD")) %>% 
+path.lookup2 <- paste0(path, "/data/raw/Output_Area_to_Lower_layer_Super_Output_Area_to_Middle_layer_Super_Output_Area_to_Local_Authority_District_(December_2011)_Lookup_in_England_and_Wales.csv")
+lookup.region2 <- read_csv(path.lookup2) %>% 
+  dplyr::select(OA11CD, MSOA11CD)
+
+lookup.region <- lookup.region %>% left_join(lookup.region2) %>% 
+  relocate(MSOA11CD, .after = OA11CD) %>% 
+  dplyr::select(-OA11CD) %>% 
+  distinct()
+
+df <- df %>% left_join(lookup.region, by = c("msoa11cd" = "MSOA11CD")) %>% 
   relocate(c(RGN11NM, RGN11CD), .after=year) %>% 
-  mutate(help = substring(oa11cd, 1, 1),
+  mutate(help = substring(msoa11cd, 1, 1),
          RGN11CD = ifelse(help=="E", RGN11CD,
-                         ifelse(help=="W", "Wales",
-                                ifelse(help == "S", "Scotland", "NI"))),
+                         ifelse(help=="W", "Wales", "Scotland")),
          RGN11NM = ifelse(help=="E", RGN11NM,
-                         ifelse(help=="W", "Wales",
-                                ifelse(help == "S", "Scotland", "NI")))) %>% 
+                         ifelse(help=="W", "Wales", "Scotland"))) %>% 
   dplyr::select(-help)
@@
 # For growth:
-df <- df %>% group_by(oa11cd) %>% 
+df <- df %>% group_by(msoa11cd) %>% 
   mutate(#n.lag = dplyr::lag(n),
@@
 
-# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
-# df <- read_csv(out.path)
-# df %>% write_csv(file = out.path)
-
-# rm(list=setdiff(ls(), "df"))
-
-# # This is the project path
-# path <- find_rstudio_root_file()
-
 # remove not used columns
@@
                         "Manchester", "Sheffield", "Leeds", "Edinburgh",
-                        "Leicester", "dist.city.name", "dist.retail.name", "name"))
+                        "Leicester", "dist.city.name", "name"))
 
@@
 df <- df %>% filter(year > 1996) %>% 
-  dplyr::select(oa11cd, year, RGN11CD, n, abs.growth, n.l.slag, n.nearest.city.lag, 
-                n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) #%>% #growth, abs.growth, RGN11NM, 
+  dplyr::select(msoa11cd, year, RGN11CD, n, abs.growth, n.l.slag, n.nearest.city.lag, 
+                n.London.lag, London, dist) #%>% #growth, abs.growth, RGN11NM, 
   # mutate(London.v = n.London.lag / London,
@@
   count(n)
-# in 2012, 120848 0s
+# in 2012, 14 0s
 
@@
 indices <- CreateSpacetimeFolds(df, 
-                                spacevar = "oa11cd",
+                                spacevar = "msoa11cd",
                                 timevar = "year", 
@@
 model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
-                     n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
+                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                      #London.v + city.v + retail.v + n.l.slag + year,
@@
 time.taken <- round(end.time - start.time,2)
-time.taken #20.6h
+time.taken #14m
 
@@
 # 
-# 3716736 samples
-#       8 predictor
+# 135680 samples
+#      6 predictor
 # 
@@
 # Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 2926910, 2926924, 3136005, 2926924, 3136005, 2926924, ... 
+# Summary of sample sizes: 114480, 106848, 106848, 114480, 114480, 106848, ... 
 # Resampling results across tuning parameters:
@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    4.999920  0.2053697  1.047293
-#   2     extratrees  5.167655  0.1540213  1.103048
-#   5     variance    5.187039  0.1710128  1.034382
-#   5     extratrees  5.169570  0.1557868  1.122845
-#   8     variance    5.633427  0.1247189  1.035171
-#   8     extratrees  5.182860  0.1563555  1.137332
+#   2     variance    39.87146  0.4371416  14.83116
+#   2     extratrees  40.70997  0.3962037  15.49807
+#   4     variance    40.12686  0.4400353  14.73566
+#   4     extratrees  40.23814  0.4073727  15.39737
+#   6     variance    40.66090  0.4248514  14.81426
+#   6     extratrees  40.46715  0.4016897  15.63003
 # 
@@
   mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
-                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
                           ifelse(rowname == "London", "distance to London",
-                                        ifelse(rowname == "n.nearest.retail.lag", "Nearest retail centre's wensite density, t-1",
                                  ifelse(rowname == "n.London.lag", "London's wensite density, t-1",
                                         ifelse(rowname == "n.l.slag", "spatial and temporal lag of wensite density",
-                                                             ifelse(rowname == "n.nearest.city", "Nearest city's wensite density, t-1", rowname)))))))) %>% 
+                                               ifelse(rowname == "n.nearest.city.lag", "Nearest city's wensite density, t-1", rowname)))))) %>% 
   column_to_rownames()
 
-path.out <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
+path.out <- paste0(path, "/outputs/rf/figures/varimp_MSOA.png")
 ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
 ggsave(path.out)
-
-path.out <- paste0(path, "/outputs/rf/models/oa_rf/all_OAs/all_OAs.RData")
-save(model.all, file = path.out)
-
-# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
-#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
-#                                                         "year")], 
-#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
-#                     trControl = tc,
-#                     method = "ranger",
-#                     #na.action = na.omit,
-#                     #importance = "impurity",
-#                     metric="Rsquared",
-#                     verbose=FALSE)
-# ffsmodel_LLO
-# ffsmodel_LLO$selectedvars
-# plot_ffs(ffsmodel_LLO)
 ```
@@
 # detectCores()
-# cl <- makePSOCKcluster(2)
-# registerDoParallel(cl)
+cl <- makePSOCKcluster(6)
+registerDoParallel(cl)
 
@@
   indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD!=i), 
-                                spacevar = "oa11cd",
+                                spacevar = "msoa11cd",
                                 timevar = "year", 
@@
 
-  model.all <- train(n ~ n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
+  model.all <- train(n ~ n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                    data = df %>% filter(RGN11CD!=i), 
@@
   #file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
-  file <- paste0("/hdd/tmp/regions/region_", as.character(i), ".RData")  
+  file <- paste0("/hdd/tmp/regions/msoa/region_", as.character(i), ".RData")  
   save(model.all, file = file)  # file name = region not in the training data
@@
   print(paste0("end loop for ", i))
-
 }
@@
 time.taken
-```
-
-### Resamples for plots
-
-```{r}
-# create a list of the above model outputs
-filenames <- list.files(paste0(path, "/outputs/rf/models/oa_rf/regions"), pattern = "*.RData", full.names = T)
-
-for (i in filenames){
-  load(file = i)
-  i <- gsub("/home/nw19521/projects/web.diffusion/outputs/rf/models/oa_rf/regions/|.RData", "", i)
-  assign(i, model.all)  
-}
-
- fit.model.all <- mget(ls(pattern = "^region_"))
-
-# clear up
-rm(model.all) # this is the last item from the loop
-rm(list = ls(pattern = "^region_"))
-
-# resamples
-fit.model.all.res <- resamples(fit.model.all)
-summary(fit.model.all.res)
-
-# plot metrics
-
-# TODO: fix the year labels
-
-scales <- list(x=list(relation="free"), y=list(relation="free"))
-bwplot(fit.model.all.res, scales=scales)
-
-# TODO: fix the year labels
-```
-
-### Variable importance
-
-```{r , include=TRUE, results= 'markup', message=FALSE}
-lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
-```
-
-```{r}
-help <- lapply(fit.model.all, function(x) varImp(x)) 
-help <- lapply(help, function(x) x[[1]])
-help <- do.call("rbind", help)
-
-importance <- help %>% rownames_to_column() %>% 
-  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
-         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
-  rename(importance = Overall) %>% 
-  relocate(importance, .after = year) %>% 
-  dplyr::select(-rowname)
-
-importance %>% ggplot() +
-  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
-  scale_color_discrete(name = "Features") +
-  #scale_linetype_manual(name = "Website density regressions111") 
-  ylab("Importance") + 
-  ggtitle("Feature importance") +
-  theme_minimal() +
-  theme(axis.title.x=element_blank(),
-        plot.title = element_text(hjust = 0.5))
-
 ```
@@
 
+regions <- df %>% ungroup() %>% distinct(RGN11CD) 
+regions <- as.vector(regions$RGN11CD)
+
+new.path <- "/hdd/tmp/regions/msoa/"
+
+df.predictions <- data.frame()
+
+for (i in regions){
+  load(file = paste0(new.path, "region_", as.character(i), ".RData"))
+  #model.name <- gsub("/home/nw19521/projects/web.diffusion/outputs/rf/models/oa_rf/regions/|.RData", "", i)
+  #assign(as.character(i), model.all)
+
+  pred <- predict(model.all, df[df$RGN11CD==i,])
+  pred <- as.data.frame(pred)
+
+  rownames(pred) <- c()
+  assign(paste0("region.predict.model.", ".on.", i), pred)
+
+  d <- paste0("region.predict.model.", ".on.", i)
+  d <- cbind(get(d), df %>% filter(RGN11CD==i)) #%>% arrange(year, ladcd)
+
+  d <- d %>% dplyr::select(RGN11CD, msoa11cd, year, n, pred)
+
+  colnames(d)[5] <- "predictions"
+  d <- d %>% mutate(tested.on = i)
+  df.predictions <- rbind(d, df.predictions)
+
+  rm(model.all)
+}
+
+df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN11CD, ".", tested.on))
+
+# split to list by region pair
+pred.by.region.all.list <- split(df.predictions, df.predictions$test.train)
+
+# calculate metrics for every region pair 
+rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
+                                                                     obs = x$n))  # CHANGE n WITH growth
+
+path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
+lookup.region <- read_csv(path.lookup) %>% 
+  dplyr::select(RGN21CD, RGN21NM) %>% 
+  distinct()
+
+path.out <- paste0(path, "/outputs/rf/figures/test_regions_MSOA.csv")
+rf.year.all.metrics %>% 
+  as.data.frame() %>%
+  rownames_to_column(var = "metrics") %>% 
+  pivot_longer(!metrics, names_to = "train.test") %>% 
+  pivot_wider(names_from = metrics, values_from = value) %>% 
+  separate(train.test, c("train", "test"), remove = F) %>% 
+  left_join(lookup.region, by = c("test" = "RGN21CD")) %>%
+  mutate(RGN21NM = ifelse(train == "Scotland", "Scotland",
+                          ifelse(train == "Wales", "Wales", RGN21NM))) %>% 
+  rename(region = RGN21NM) %>% 
+  arrange(Rsquared) %>% 
+  dplyr::select(region, Rsquared) %>% 
+  write_csv(path.out)
+```
+
+### OLD
+
+```{r, eval=FALSE}
+# train on all but one region, test on that region
+
 df.predictions <- data.frame()
 for (i in regions){
-  #for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], 
-                    df[df$RGN11CD==i,])
+  for (j in regions[!regions == i]){
+    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], df[df$RGN11CD==j,])
     pred <- as.data.frame(pred)
     rownames(pred) <- c()
-    assign(paste0("region.predict.model.", ".on.", i), pred)
-
-    d <- paste0("region.predict.model.", ".on.", i)
-    d <- cbind(get(d), df %>% filter(RGN11CD==i)) #%>% arrange(year, ladcd)
-
-    #***dplyr::select(RGN21CD, ladcd, year, growth, 1)***
-
+    assign(paste0("region.predict.model.", i, ".on.", j), pred)
+
+    d <- paste0("region.predict.model.", i, ".on.", j)
+    d <- cbind(get(d), df %>% filter(RGN11CD==j)) #%>% arrange(year, ladcd)
     d <- d[,c(5, 2, 4, 6, 1)]

### 2024-08-14 src/problem_years.Rmd
sum corrected

@@
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/corrected/LA/morani_la_1.csv")
+path.out <- paste0(path, "/outputs/lisa/corrected/LA/morani_la.csv")
 morani %>% as_tibble() %>% 
@@
   group_by(year, oa11cd) %>%
-  summarise(n = n()) %>% # for multiple pc  
+  summarise(n = sum(n)) %>% # **********ATTENTION*********    
   ungroup() %>% 
@@
 ```{r}
-path.out <- paste0(path, "/outputs/lisa/corrected/OA/morani_oa_1.csv")
+path.out <- paste0(path, "/outputs/lisa/corrected/OA/morani_oa.csv")
 morani %>% as_tibble() %>% 

### 2024-08-14 src/s_la_per_firm_corrected.Rmd
sum corrected

--- 2024-08-14 src/la_RF_corrected.Rmd
@@
 ---
-title: "Random Forests, LAs"
+title: "s.function"
 date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 library(tidyverse)
+library(broom)
 library(rprojroot)
-library(rgdal)
-library(rgeos)
+#library(rgdal)
 library(sf)
+library(modelr)
+#library(survival)
+#library(rgeos)
 library(spdep)
-#library(tmap)
-#library(maptools)
-library(knitr)
-library(REAT)
 library(tidygeocoder)
-library(geosphere)
-library(broom)
-#library(foreach)
-library(doParallel)
-library(raster)
-library(plm)
-#library(lmtest)
-library(caret)
-library(randomForest)
-library(CAST)
+# library(geosphere)
+# library(raster)
+library(rms)
+library(stargazer)
 
@@
 ```
+## S function estimation for all the country
 
-## Distances to cities and retail centres
-
-First set up the destinations. Origins are all the LA.
-
-Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
-Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)
+To interpret the coefficients, check [SSlogis](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/SSlogis).
+This is a self-starting model, which estimates the starting values.
 
 ```{r}
-# LA centroids
-# la.c <- gCentroid(la, byid=TRUE)@coords %>% 
-#   as.data.frame() %>%
-#   bind_cols(la@data) %>% 
-#   dplyr::select(LAD21CD, x, y) %>% 
-#   remove_rownames()
+df.country <- df.corrected %>% 
+  mutate(year = as.integer(as.character(year))) %>% 
+  filter(year > 1995 & year < 2013) %>% 
+  left_join(lookup, by = "pc", suffix =c("","")) %>% 
+  group_by(year, ladcd) %>%
+  summarise(n = sum(n)) %>% # it was n=n() **********ATTENTION*********  
+  ungroup() %>% 
+  complete(ladcd, year) %>% 
+  filter(!is.na(ladcd)) %>% # drop c. 10 - 200 per year, which are not assigned to a LAD
+  replace(is.na(.), 0) %>% 
+  left_join(firms) %>% 
+  group_by(year) %>% 
+  summarise(n = sum(n, na.rm = T),
+            firms = sum(firms, na.rm = T)) %>% 
+  mutate(n = n/firms) %>% 
+  dplyr::select(-firms)
 
-sf_use_s2(FALSE)
-la.c <- st_centroid(la) %>% 
-  st_coordinates() %>%
-  as.data.frame() %>% 
-  bind_cols(la %>% dplyr::select(LAD21CD))
+fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.country)
+summary(fit)
 
-# cities
-city.names <- c("London, UK", 
-"Birmingham, UK",
-"Glasgow, UK",
-"Liverpool, UK",
-"Bristol, UK",
-"Manchester, UK",
-"Sheffield, UK",
-"Leeds, UK",
-"Edinburgh, UK",
-"Leicester, UK",
-"Belfast, UK")
+# xmid = 2003.118
+xmid.uk <- tidy(fit) %>% 
+  filter(term=="xmid") %>% 
+  dplyr::select(estimate) %>% as.numeric()
 
-cities <- geo(city.names, no_query = F, method = "arcgis")
+rsquare(fit, df.country)
 
-# retail centres
-geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
-#retail <-readOGR(geo.path) 
-retail <- st_read(geo.path)
+df.country$prednls = predict(fit)
 
-#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
-# retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
-#   as.data.frame() %>% 
-#   add_rownames(var = "id") #%>%
-retail.major.cetres.help <- st_centroid(retail) %>% 
-  st_transform(4326) %>%
-  st_coordinates() %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id")
+path.out <- paste0(path, "/outputs/s/s_uk_per_firm.png")
 
-# retail.major.cetres <- retail.major.cetres.help %>% 
-#   st_as_sf(coords = c("x", "y"), crs = 27700) %>%
-#   st_transform(4326) %>%
-#   st_coordinates() %>%
-#   as_tibble() %>% 
-#   bind_cols(retail.major.cetres.help$id) %>% 
-#   rename(id = '...3') %>% 
-#   left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
-retail.major.cetres <- retail.major.cetres.help %>% 
-  # as_tibble() %>% 
-  # bind_cols(retail.major.cetres.help$id) %>% 
-  # #rename(id = '...3') %>% 
-  left_join(retail %>% rownames_to_column(var = "id"), by = 'id')
+df.country %>% 
+  ggplot(aes(x = year, y = n) ) +
+  geom_point() +
+  geom_line(aes(y = prednls), size = 1) +
+  scale_x_continuous(NULL, labels = as.character(df.country$year), breaks = df.country$year) +
+  geom_vline(xintercept = xmid.uk) +
+  ylab("N. of websites per firm") +
+  #ggtitle("S-curve, UK") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5),
+          panel.grid.minor.x = element_blank())
+
+ggsave(path.out)
+  #scale_x_date(breaks = df.test$year) 
 ```
 
-Distance to cities
+## Loop for S function estimation for Local Authorities keeping the model objects
 
 ```{r}
-dist <- distm(cbind(la.c$X, la.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist <- round((dist/1000),2) %>% 
-  as_tibble()
+a.la <- tibble()
+all.la <- df %>% filter(ladcd != "NA") %>% distinct(ladcd) %>% 
+  #top_n(n=2) %>% 
+  simplify2array()
+simplify2array(all.la)
 
-city.names <- city.names %>% stringr::str_remove(", UK")
-names(dist) <- city.names
+for (i in all.la){
+  data <- df %>% filter(ladcd==i) %>%
+    dplyr::select(year, ladcd, n) 
+    #lm(n~year, data = .)
+    #tryCatch(tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .)), , error=function(e) NULL)
+  tryCatch(model <- (nls(n ~ SSlogis(year, Asym, xmid, scal), 
+                         data = data, 
+                         control = nls.control(maxiter = 1000))), error=function(e) NULL)
+  model.name <- paste0("model.", i)
+  assign(model.name, model)
+  tryCatch(b.la <- tidy(model), error=function(e) NULL)
 
-dist <- dist %>% bind_cols(la.c$LAD21CD) %>% 
-  rename(ladcd = last_col()) %>% 
-  relocate(ladcd)
+  tryCatch(b.la$r2 <- rsquare(model, data), error=function(e) NULL)
 
-# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
-# dist$distMet <- apply(dist[,2:11], 1, min)
+  tryCatch(a.la <-rbind(a.la,b.la), error=function(e) NULL)
+}
 
-dist <- transform(dist, dist = do.call(pmin, dist[-1]))
+a.la <- a.la %>% filter(term=="xmid") %>% 
+  bind_cols(as_tibble(all.la)) %>% 
+  #mutate(estimate = round(estimate, 0)) %>% 
+  #summarise(r = range(r2))
+  filter(p.value < 0.01 & 
+         r2 > 0.9) %>%    # 376 / 376, all sig, 330 r2 > 0.9
+  dplyr::select(ladcd, estimate, r2, std.error) %>%
+  mutate(fast = ifelse(estimate < xmid.uk, "fast", "slow")) # it was 2003
 
-# NOT USED AS IT SCREWS UP THE REGRESSIONS
-# Complete panel, N = 3871903, 17 year * 227759 OA
-# df <- df %>% bind_rows(dist %>% select(oa11cd)) %>% 
-#   mutate(year = ifelse(is.na(year),1000, year),
-#          country = substring(oa11cd, 1, 1)) %>% 
-#   filter(country != "L" & country != "M" & country != "N") %>% 
-#   filter(!is.na(country)) %>% 
-#   complete(year, oa11cd, fill = list(n = 0)) %>% 
-#   filter(year!=1000,
-#          !is.na(oa11cd)) %>% 
-#   select(-country)
+# predictions 
+pred <- tibble()
+help <- tibble(year = 1996:2012)
+for (i in all.la){
+  model.name <- (paste0("model.", i))
+  #assign(col.name, paste0("pred.", i))
+  pred_ <- df %>% filter(ladcd == i) %>% 
+    mutate(prednls = eval(call("predict", as.name(model.name))))  # "pred.{i}" :=
+  help_ <- help %>% left_join(pred_)
 
-# tests
-# test$country <- substring(test$oa11cd, 1, 1)
-# unique(test$country)
-# length(unique(test$oa11cd))
+  pred <- pred %>% bind_rows(help_) 
+}
 
-# calculate area
-#la$SHAPE_Area <- area(la) #/ 1000000
-la$area <- st_area(la) #/ 1000000
+# plots
+path.out <- paste0(path, "/outputs/s/s_lad_per_firm/lad_")
 
-# Join with distance and area
-df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
-  left_join(la, by = c("ladcd" = "LAD21CD")) %>% 
-  relocate(area, .after = n) %>% 
-  mutate(area = area / 1000000,     # UNITS APPEAR WRONGLY IN M AND NOT IN KM
-         density = n / area) %>% 
-  relocate(density, .after = area)
+for (i in all.la){
 
-df <- df %>% filter(!is.na(London))  # to drop 34 for L99999999 and M99999999
+  xmid_la <- pred %>% left_join(a.la) %>% filter(ladcd == i) %>% dplyr::select(estimate) %>% first() #******
+  lines <- data.frame(
+    intercepts = c(xmid.uk, xmid_la[[1]]), # ******************
+    Xmid = c("UK", i)
+  )
 
-sapply(df, function(x) sum(is.na(x)))
+p <- ggplot(pred %>% left_join(a.la) %>% 
+              filter(ladcd == i), aes(x = year, y = n, color = fast) ) +
+  geom_point() +
+  geom_line(aes(y = prednls), size = 1) +
+  scale_x_continuous(NULL, labels = as.character(df$year), breaks = df$year) +
+  #geom_vline(xintercept = pred %>% left_join(a.la) %>% filter(ladcd == i) %>% dplyr::select(estimate) %>% first()) +
+  ylab("N. of websites / per firm") + 
+  labs(color = "Website \ndiffusion speed") +
+  labs(title = paste0(la  %>% as_tibble() %>% filter(LAD21CD== i) %>% dplyr::select(LAD21NM) %>% first(), ", ", i)) +
+  theme(legend.position=c(.9,.15)) +
+  scale_color_manual(values = c("fast" = "purple",
+                                "slow"="orange"),
+                     limits = force) +
+  geom_vline(
+    data=lines,
+    aes(xintercept = intercepts, linetype=Xmid),
+    size=.5, key_glyph='path', colour="grey30"
+    ) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5)) #+
+  #geom_vline(xintercept = 2003, colour="grey30", linetype="dashed")
+
+ggsave(p, filename = paste(path.out, i, ".png", sep = ""), bg="white")
+}
 ```
 
-Distance to retail centres
+Write the object with all the models. 
+NAs are for LAD with R^2 < 0.9.
+
+```{r eval=F}
+path.out <- paste0(path, "/outputs/s/s_la_per_firm.csv")
+
+la %>% as_tibble() %>% 
+  dplyr::select(LAD21CD, LAD21NM) %>% 
+  left_join(a.la, by = c("LAD21CD" = "ladcd")) %>% 
+  write_csv(path.out)
+```
+
+## Map the xmid
 
 ```{r}
-dist.retail <- distm(cbind(la.c$X, la.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail <- round((dist.retail/1000),2) %>% 
-  as_tibble()
+path.out <- paste0(path, "/outputs/s/speed_map.png")
 
-# retail.names <- retail.major.cetres$RC_Name
-# names(dist.retail) <- retail.names
+la %>% left_join(a.la, by = c("LAD21CD" = "ladcd")) %>% 
+  ggplot(aes(fill = fast)) +
+  geom_sf() +  
+  #theme_nothing(legend = TRUE) +
+  theme_void() +
+  labs(fill='Diffusion speed') +
+  scale_fill_manual(values = c("darkorchid3", "orange"), #darkmagenta, purple4
+                    na.value = "grey70")
 
-dist.retail <- dist.retail %>% bind_cols(la.c$LAD21CD) %>% 
-  rename(LAD21CD = last_col()) %>% 
-  relocate(LAD21CD)
-
-# Minimum distance 
-dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
-dist.retail <- dist.retail %>% dplyr::select(LAD21CD, dist.retail)
-
-# Join with complete panel
-df <- df %>% left_join(dist.retail, by = c("ladcd" = "LAD21CD"))
-
-sapply(df, function(x) sum(is.na(x)))
-```
-
-## df for RF
-
-```{r}
-# London = Islington 
-# sf_use_s2(FALSE)
-# path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-# la.sf <- st_read(path.geo)
-
-#help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
-#  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)
-
-#************** DOUBLE CHECK THE ORDER OF CITY NAMES BELOW **************
-
-# nearest city
-help1 <- st_filter( la, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
-  mutate(cities = c("Leicester",
-                    "Bristol",
-                    "Manchester",
-                    "Liverpool",
-                    "Sheffield",
-                    "Birmingham",
-                    "Leeds",
-                    "London",
-                    "Belfast",
-                    "Edinburgh",
-                    "Glasgow"))
-
-# dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
-# help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
-#   rename(nearest.city.LAD21CD = LAD21CD) %>% 
-#   dplyr::select(ladcd, nearest.city.LAD21CD)
-
-# This efficiently returns the name of the column with the shortest distance
-dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
-help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
-  rename(nearest.city.LAD21CD = LAD21CD) %>% 
-  dplyr::select(ladcd, nearest.city.LAD21CD)
-
-sf_use_s2(FALSE)
-la.c <- st_centroid(la)
-kn <- knearneigh(la.c, k = 5)
-knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
-knn.l <- nb2listw(knn)
-
-df <- df %>% 
-  group_by(year) %>% 
-  mutate(n.slag = lag.listw(knn.l, n)) %>% 
-  group_by(ladcd) %>% 
-  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
-         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
-  left_join(help2, by = c("ladcd" = "ladcd"))
-
-df <- df %>% left_join(df %>% dplyr::select(ladcd,year, n) %>% 
-                         rename(n.nearest.city = n), 
-                       by = c("nearest.city.LAD21CD" = "ladcd",
-                              "year" = "year")) %>% 
-  dplyr::select(-c(LAD21NMW, OBJECTID, BNG_E, BNG_N, LONG, LAT, SHAPE_Length)) %>% 
-  relocate(LAD21NM, .after = ladcd)
-
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% dplyr::select(LAD21CD, RGN21CD)
-
-df <- df %>% left_join(lookup.region, by = c("ladcd" = "LAD21CD")) %>% 
-  relocate(RGN21CD, .after=year)
-
-# North/South
-#lookup.region %>% distinct(RGN21NM, .keep_all = T)
-south <- c("E12000006", "E12000007", "E12000008", "E12000009" )
-df <- df %>% mutate(south = ifelse(RGN21CD %in% south, 1, 0))
-
-# For growth:
-#df <- df %>% mutate(growth = log(n/n.lag))                          
-df <- df %>% mutate(growth = log((n+0.001)/(n.lag+0.01)))                       # CONSTANT FOR GROWTH
-#df <- df %>% mutate(growth = n- n.lag)                                         # ABS. GROWTH
-
-sapply(df, function(x) sum(is.na(x)))
-```
-
-## unregister_dopar function
-
-This is helpful when the parallel loop collapses.
-
-```{r}
-unregister_dopar <- function() {
-  env <- foreach:::.foreachGlobals
-  rm(list=ls(name=env), pos=env)
-}
-unregister_dopar()
-```
-
-## RF with CAST
-
-```{r}
-df <- df %>% mutate(help = substr(ladcd, start = 1, stop = 2)) %>% 
-  mutate(RGN21CD = ifelse(is.na(RGN21CD), help, RGN21CD)) %>% 
-  ungroup() %>% 
-  dplyr::select(-help)
-
-df <- df %>% left_join(df %>% 
-                         ungroup() %>% 
-                         filter(LAD21NM == "Islington") %>% 
-                         dplyr::select(year, n) %>% 
-                         rename(n.London = n)) %>% 
-  group_by(ladcd) %>% 
-  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
-  ungroup()
-
-# There are no lags for 1996, so it is removed
-df <- df %>% filter(year > 1996)
-sapply(df, function(x) sum(is.na(x)))
-```
-
-### train the model in all data
-
-```{r}
-set.seed(123)
-indices <- CreateSpacetimeFolds(df,# %>% filter(year > 1999),                   # FILTER BY YEAR FOR GROWTH
-                                spacevar = "ladcd",
-                                timevar = "year", 
-                                k = 10)
-
-#length is = (n_repeats*nresampling)+1
-seeds <- vector(mode = "list", length = 11)
-
-#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # , 11) 9 after adding South
-
-#for the last model
-seeds[[11]]<-sample.int(1000, 1)
-
-# CV
-tc <- trainControl(method = "cv",
-                   number = 10,
-                   seeds = seeds,
-                   allowParallel = T,
-                   index = indices$index,
-                   savePredictions = 'final')
-
-detectCores()
-cl <- makePSOCKcluster(14)
-registerDoParallel(cl)
-
-start.time <- Sys.time()
-model.all <- train(n ~                                                          # REPLACE n WITH growth
-                     #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
-                     #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
-                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist, #n.nearest.retail.lag + dist.retail,
-                   data = df,# %>% filter(year > 1999),                         # FILTER BY YEAR FOR GROWTH
-                   trControl = tc,
-                   method = "ranger",
-                   na.action = na.omit,
-                   #preProc = c("center", "scale"),
-                   importance = "impurity")
-
-stopCluster(cl)
-end.time <- Sys.time()
-time.taken <- round(end.time - start.time,2)
-time.taken
-
-print(model.all)
-
-# [reproducible results]
-# Random Forest 
-# 
-# 5984 samples
-#    6 predictor
-# 
-# No pre-processing
-# Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 4718, 4690, 5040, 4718, 5055, 4704, ... 
-# Resampling results across tuning parameters:
-# 
-#   mtry  splitrule   RMSE        Rsquared   MAE        
-#   2     variance    0.01366816  0.8895425  0.009767759
-#   2     extratrees  0.01350182  0.8900993  0.009603882
-#   4     variance    0.01408821  0.8841051  0.010134321
-#   4     extratrees  0.01354075  0.8914524  0.009664646
-#   6     variance    0.01445541  0.8772260  0.010417962
-#   6     extratrees  0.01369444  0.8897055  0.009772824
-# 
-# Tuning parameter 'min.node.size' was held constant at a value of 5
-# RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule = extratrees
-#  and min.node.size = 5.
-
-varimp_mars <- varImp(model.all) 
-varimp_mars$importance <- varimp_mars$importance %>% 
-  rownames_to_column() %>% 
-  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
-                          ifelse(rowname == "n.London.lag", "London's website density, t-1", 
-                                 ifelse(rowname == "London", "distance to London",
-                                        ifelse(rowname == "n.l.slag", "spatial and temporal lag of website density",
-                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's website density, t-1", rowname)))))) %>% 
-  column_to_rownames()
-
-path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.png")
-ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
 ggsave(path.out)
@@
 
-### train the model in a loop for all but one regions
+## Average ranks websites per firm t and 2012 only
+
+Change percentile in out.path and uncomment `mutate(alpha = ifelse(abs(diff_1st) > ...`
 
 ```{r}
-rm(indices)
 
-#length is = (n_repeats*nresampling)+1
-seeds <- vector(mode = "list", length = 11)
+t <- 2000
 
-#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-set.seed(831)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # 6 feature + 2
+out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only95_av.png")
 
-#for the last model
-set.seed(832)
-seeds[[11]]<-sample.int(1000, 1)
+df %>%
 
-# detectCores()
-cl <- makePSOCKcluster(14)
-registerDoParallel(cl)
+  left_join(la %>% as_tibble() %>% dplyr::select(LAD21CD, LAD21NM),
+            by = c("ladcd" = "LAD21CD")) %>%
 
-# train in every region
-start.time <- Sys.time()
+  mutate(period = ifelse(year < 1999, 1, 
+                         ifelse(year > 2009, 3, 2))) %>%
+  group_by(ladcd, period) %>%
+  mutate(n = ifelse(period != 2, mean(n), n)) %>% 
+  filter(year > 1997 & year < 2011) %>% 
+  group_by(year) %>% 
+  mutate(rank = rank(desc(n))) %>% 
+  ungroup() %>% 
+  filter(year==1998 | year == 2010) %>% 
+  arrange(ladcd, year) %>% 
+  group_by(ladcd) %>% 
+  #mutate(diff_1st = -diff(rank))
+  mutate(diff_1st = -(rank-dplyr::lag(rank))) %>%
 
-regions <- df %>% ungroup() %>% distinct(RGN21CD)
-regions <- as.vector(regions$RGN21CD)
+  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .05, na.rm = T))
+  # q.95 = 253, q.9 = 214, q.1 = 11, q.05 = 3.75. **abs()**
 
-for (i in regions){
+  mutate(alpha = ifelse(abs(diff_1st) > 253, .8, 0.1)) %>% 
+  #mutate(alpha = ifelse(abs(diff_1st) <= 3.75, .8, 0.1)) %>%
 
-  set.seed(71)
-  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i), #& year > 1999),   # FILTER BY YEAR FOR GROWTH 
-                                spacevar = "ladcd",
-                                timevar = "year", 
-                                k = 10)
-  # CV
-  tc <- trainControl(method = "cv", #MOVED HERE< DOUBLE CHECK
-                   number = 10,
-                   seeds = seeds,
-                   allowParallel = T,
-                   index = indices$index,
-                   savePredictions = 'final')
+  mutate(alpha = ifelse(is.na(diff_1st), 0.1, alpha)) %>%
+  mutate(label = ifelse(alpha==.1, "", LAD21NM)) %>%
 
-  model.all <- train(n ~                                                        # REPLACE n WITH growth
-                       #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
-                       #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
-                       n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
-                   data = df %>% filter(RGN21CD!=i), #& year > 1999),           # FILTER BY YEAR FOR GROWTH
-                   trControl = tc,
-                   method = "ranger",
-                   na.action = na.omit,
-                   #preProc = c("center", "scale"),
-                   importance = "impurity")
-  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data
-}
-
-stopCluster(cl)
-end.time <- Sys.time() 
-time.taken <- round(end.time - start.time,2)
-time.taken # ca. 4 minutes 
+  ggplot(aes(year, rank, group = ladcd, alpha=alpha)) +
+  geom_bump(colour = "red") +
+  scale_y_reverse(labels = c(1, 100, 200, 300, 400)) +
+  # geom_text(#data = df %>% filter(year == 2012),
+  #           aes(x = Inf, label = label), size = 4, hjust = 0)  # x = 2012 + .1
+  geom_text_repel(aes(label = label),
+                  na.rm = TRUE, 
+                  size = 2,
+                  nudge_x = 2, nudge_y = 1,
+                  direction = "y", hjust = "left") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position="none") +
+  scale_x_continuous(NULL, 
+                     #limits = c(1998, 2010), 
+                     breaks=c(1998, 2010),
+                     labels = c("1996-98","2010-12"))
+ggsave(out.path)
 ```
-
-### Resamples for plots
-
-```{r}
-# create a list of the above model outputs
-fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
-
-# resamples
-fit.model.all.res <- resamples(fit.model.all)
-summary(fit.model.all.res)
-
-# plot metrics
-
-# TODO: fix the year labels
-
-scales <- list(x=list(relation="free"), y=list(relation="free"))
-bwplot(fit.model.all.res, scales=scales)
-
-# TODO: fix the year labels
-```
-
-### Variable importance
-
-```{r , include=TRUE, results= 'markup', message=FALSE}
-lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
-```
-
-```{r}
-help <- lapply(fit.model.all, function(x) varImp(x)) 
-help <- lapply(help, function(x) x[[1]])
-help <- do.call("rbind", help)
-
-importance <- help %>% rownames_to_column() %>% 
-  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
-         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
-  rename(importance = Overall) %>% 
-  relocate(importance, .after = year) %>% 
-  dplyr::select(-rowname)
-
-importance %>% ggplot() +
-  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
-  scale_color_discrete(name = "Features") +
-  #scale_linetype_manual(name = "Website density regressions111") 
-  ylab("Importance") + 
-  ggtitle("Feature importance") +
-  theme_minimal() +
-  theme(axis.title.x=element_blank(),
-        plot.title = element_text(hjust = 0.5))
-
-```
-
-### Test on one region
-
-```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
-
-# for a reference point
-df %>% group_by(year) %>%
-  summarise(min = min(n), max=max(n),
-            mean = mean(n), median = median(n)) %>%
-  round(2) %>% kable()
-```
-
-```{r}
-# train on all but one region, test on that region
-
-df.predictions <- data.frame()
-for (i in regions){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], 
-                    df[df$RGN21CD==i,]) #& df$year > 1999,])                            # FILTER BY YEAR
-    pred <- as.data.frame(pred)
-    rownames(pred) <- c()
-    assign(paste0("region.predict.model.", ".on.", i), pred)
-
-    d <- paste0("region.predict.model.", ".on.", i)
-    d <- cbind(get(d), df %>% filter(RGN21CD==i)) #& year > 1999))                      # FILTER BY YEAR
-    d <- d %>% dplyr::select(RGN21CD, ladcd, year, n, 1)                                # REPLACE n WITH growth
-    colnames(d)[5] <- "predictions"
-    d <- d %>% mutate(tested.on = i)
-    df.predictions <- rbind(d, df.predictions)
-}
-
-df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", tested.on))
-
-# split to list by region pair
-pred.by.region.all.list <- split(df.predictions, df.predictions$RGN21CD)
-
-# calculate metrics for every region pair 
-rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))       # REPLACE n WITH growth
-
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% 
-  dplyr::select(RGN21CD, RGN21NM) %>% 
-  distinct() %>% 
-  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
-
-path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected.csv")
-rf.year.all.metrics %>% 
-  as.data.frame() %>%
-  rownames_to_column(var = "metrics") %>% 
-  pivot_longer(!metrics, names_to = "test.on") %>% 
-  pivot_wider(names_from = metrics, values_from = value) %>% 
-  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
-  arrange(Rsquared) %>% 
-  dplyr::select(RGN21NM, Rsquared) %>% 
-  rename(Region = RGN21NM) %>% 
-  write_csv(path.out)
-  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()
-
-# ggplot(aes(train.region, test.region, fill= Rsquared)) + 
-#   geom_tile() +
-#   coord_fixed() +
-#   labs(y = "", x = "") + #y: test.region, x: train.region
-#   scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
-#   theme(#legend.position = "none",
-#         axis.text.x = element_text(angle = 45, hjust=1),
-#         panel.border = element_blank(), 
-#         panel.grid.major = element_blank(),
-#         panel.grid.minor = element_blank(),
-#         panel.background = element_rect(fill = "white"))
-# ggsave(path.out)
-
-  #summarise(r = range(Rsquared))
-  #round(2) %>%
-  #kable()
-```
-
-## RF growth with CAST
-
-### train the model in all data, growth
-
-```{r}
-set.seed(123)
-indices <- CreateSpacetimeFolds(df %>% filter(year > 1999),                     # FILTER BY YEAR FOR GROWTH
-                                spacevar = "ladcd",
-                                timevar = "year", 
-                                k = 10)
-
-#length is = (n_repeats*nresampling)+1
-seeds <- vector(mode = "list", length = 11)
-
-#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # , 11) 9 after adding South
-
-#for the last model
-seeds[[11]]<-sample.int(1000, 1)
-
-# CV
-tc <- trainControl(method = "cv",
-                   number = 10,
-                   seeds = seeds,
-                   allowParallel = T,
-                   index = indices$index,
-                   savePredictions = 'final')
-
-detectCores()
-cl <- makePSOCKcluster(14)
-registerDoParallel(cl)
-
-start.time <- Sys.time()
-model.all <- train(growth ~                                                     # REPLACE n WITH growth
-                     #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
-                     #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
-                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist, #n.nearest.retail.lag + dist.retail,
-                   data = df %>% filter(year > 1999),                           # FILTER BY YEAR FOR GROWTH
-                   trControl = tc,
-                   method = "ranger",
-                   na.action = na.omit,
-                   #preProc = c("center", "scale"),
-                   importance = "impurity")
-
-stopCluster(cl)
-end.time <- Sys.time()
-time.taken <- round(end.time - start.time,2)
-time.taken
-
-print(model.all)
-
-# [reproducible results]
-# Random Forest 
-# 
-# 4862 samples
-#    6 predictor
-# 
-# No pre-processing
-# Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 4044, 3685, 4032, 4044, 4044, 3696, ... 
-# Resampling results across tuning parameters:
-# 
-#   mtry  splitrule   RMSE       Rsquared   MAE      
-#   2     variance    0.2182933  0.5361604  0.1560099
-#   2     extratrees  0.2099702  0.5660477  0.1501194
-#   4     variance    0.2168118  0.5238461  0.1550630
-#   4     extratrees  0.2074722  0.5590058  0.1489474
-#   6     variance    0.2268891  0.5014882  0.1631466
-#   6     extratrees  0.2082266  0.5503986  0.1490911
-# 
-# Tuning parameter 'min.node.size' was held constant at a value of 5
-# RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 4, splitrule = extratrees
-#  and min.node.size = 5.
-
-varimp_mars <- varImp(model.all) 
-varimp_mars$importance <- varimp_mars$importance %>% 
-  rownames_to_column() %>% 
-  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
-                          ifelse(rowname == "n.London.lag", "London's website density, t-1", 
-                                 ifelse(rowname == "London", "distance to London",
-                                        ifelse(rowname == "n.l.slag", "spatial and temporal lag of website density",
-                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's website density, t-1", rowname)))))) %>% 
-  column_to_rownames()
-
-path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_growth_corrected.png")
-ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
-ggsave(path.out)
-```
-
-### train the model in a loop for all but one regions, growth
-
-```{r}
-rm(indices)
-
-#length is = (n_repeats*nresampling)+1
-seeds <- vector(mode = "list", length = 11)
-
-#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-set.seed(831)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # 6 feature + 2
-
-#for the last model
-set.seed(832)
-seeds[[11]]<-sample.int(1000, 1)
-
-# detectCores()
-cl <- makePSOCKcluster(14)
-registerDoParallel(cl)
-
-# train in every region
-start.time <- Sys.time()
-
-regions <- df %>% ungroup() %>% distinct(RGN21CD)
-regions <- as.vector(regions$RGN21CD)
-
-for (i in regions){
-
-  set.seed(71)
-  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i & year > 1999),      # FILTER BY YEAR FOR GROWTH 
-                                spacevar = "ladcd",
-                                timevar = "year", 
-                                k = 10)
-  # CV
-  tc <- trainControl(method = "cv", #MOVED HERE< DOUBLE CHECK
-                   number = 10,
-                   seeds = seeds,
-                   allowParallel = T,
-                   index = indices$index,
-                   savePredictions = 'final')
-
-  model.all <- train(growth ~                                                   # REPLACE n WITH growth
-                       #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
-                       #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
-                       n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
-                   data = df %>% filter(RGN21CD!=i & year > 1999),              # FILTER BY YEAR FOR GROWTH
-                   trControl = tc,
-                   method = "ranger",
-                   na.action = na.omit,
-                   #preProc = c("center", "scale"),
-                   importance = "impurity")
-  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data
-}
-
-stopCluster(cl)
-end.time <- Sys.time() 
-time.taken <- round(end.time - start.time,2)
-time.taken # ca. 3 minutes 
-```
-
-### Resamples for plots, growth
-
-```{r}
-# create a list of the above model outputs
-fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))
-
-# resamples
-fit.model.all.res <- resamples(fit.model.all)
-summary(fit.model.all.res)
-
-# plot metrics
-
-# TODO: fix the year labels
-
-scales <- list(x=list(relation="free"), y=list(relation="free"))
-bwplot(fit.model.all.res, scales=scales)
-
-# TODO: fix the year labels
-```
-
-### Variable importance, growth
-
-```{r , include=TRUE, results= 'markup', message=FALSE}
-lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
-```
-
-```{r}
-help <- lapply(fit.model.all, function(x) varImp(x)) 
-help <- lapply(help, function(x) x[[1]])
-help <- do.call("rbind", help)
-
-importance <- help %>% rownames_to_column() %>% 
-  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
-         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
-  rename(importance = Overall) %>% 
-  relocate(importance, .after = year) %>% 
-  dplyr::select(-rowname)
-
-importance %>% ggplot() +
-  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
-  scale_color_discrete(name = "Features") +
-  #scale_linetype_manual(name = "Website density regressions111") 
-  ylab("Importance") + 
-  ggtitle("Feature importance") +
-  theme_minimal() +
-  theme(axis.title.x=element_blank(),
-        plot.title = element_text(hjust = 0.5))
-
-```
-
-### Test on one region, growth
-
-```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
-
-# for a reference point
-df %>% group_by(year) %>%
-  summarise(min = min(n), max=max(n),
-            mean = mean(n), median = median(n)) %>%
-  round(2) %>% kable()
-```
-
-```{r}
-# train on all but one region, test on that region
-
-df.predictions <- data.frame()
-for (i in regions){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], 
-                    df[df$RGN21CD==i & df$year > 1999,])                                # FILTER BY YEAR
-    pred <- as.data.frame(pred)
-    rownames(pred) <- c()
-    assign(paste0("region.predict.model.", ".on.", i), pred)
-
-    d <- paste0("region.predict.model.", ".on.", i)
-    d <- cbind(get(d), df %>% filter(RGN21CD==i & year > 1999))                         # FILTER BY YEAR
-    d <- d %>% dplyr::select(RGN21CD, ladcd, year, growth, 1)                           # REPLACE n WITH growth
-    colnames(d)[5] <- "predictions"
-    d <- d %>% mutate(tested.on = i)
-    df.predictions <- rbind(d, df.predictions)
-}
-
-df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", tested.on))
-
-# split to list by region pair
-pred.by.region.all.list <- split(df.predictions, df.predictions$RGN21CD)
-
-# calculate metrics for every region pair 
-rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$growth))   # REPLACE n WITH growth
-
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% 
-  dplyr::select(RGN21CD, RGN21NM) %>% 
-  distinct() %>% 
-  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
-
-path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth_corrected.csv")
-rf.year.all.metrics %>% 
-  as.data.frame() %>%
-  rownames_to_column(var = "metrics") %>% 
-  pivot_longer(!metrics, names_to = "test.on") %>% 
-  pivot_wider(names_from = metrics, values_from = value) %>% 
-  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
-  arrange(Rsquared) %>% 
-  dplyr::select(RGN21NM, Rsquared) %>% 
-  rename(Region = RGN21NM) %>% 
-  write_csv(path.out)
-  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()
-
-# ggplot(aes(train.region, test.region, fill= Rsquared)) + 
-#   geom_tile() +
-#   coord_fixed() +
-#   labs(y = "", x = "") + #y: test.region, x: train.region
-#   scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
-#   theme(#legend.position = "none",
-#         axis.text.x = element_text(angle = 45, hjust=1),
-#         panel.border = element_blank(), 
-#         panel.grid.major = element_blank(),
-#         panel.grid.minor = element_blank(),
-#         panel.background = element_rect(fill = "white"))
-# ggsave(path.out)
-
-  #summarise(r = range(Rsquared))
-  #round(2) %>%
-  #kable()
-```

### 2024-08-14 src/s_la_per_firm_corrected_10.Rmd
sum corrected

--- 2024-08-14 src/s_la_per_firm_corrected.Rmd
@@
 ---
-title: "s.function"
+title: "s.function for 10 postcodes"
 date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 library(rprojroot)
-#library(rgdal)
+library(rgdal)
 library(sf)
 library(modelr)
-#library(survival)
-#library(rgeos)
-library(spdep)
-library(tidygeocoder)
-# library(geosphere)
-# library(raster)
-library(rms)
-library(stargazer)
+library(ggbump)
+library(ggrepel)
 
@@
   group_by(year, ladcd) %>%
-  summarise(n = sum(n)) %>% # **********ATTENTION*********  
+  summarise(n = sum(n)) %>% # for multiple pc  **********ATTENTION*********
   ungroup() %>% 
@@
 ```
+
 ## S function estimation for all the country
@@
   group_by(year, ladcd) %>%
-  summarise(n = sum(n)) %>% # it was n=n() **********ATTENTION*********  
+  summarise(n = sum(n)) %>% # for multiple pc  **********ATTENTION*********
   ungroup() %>% 
@@
 
-# xmid = 2003.118
+# xmid = 2002.303
 xmid.uk <- tidy(fit) %>% 
@@
 
-path.out <- paste0(path, "/outputs/s/s_uk_per_firm.png")
+path.out <- paste0(path, "/outputs/s/s_uk_per_firm_10.png")
 
@@
 # plots
-path.out <- paste0(path, "/outputs/s/s_lad_per_firm/lad_")
+path.out <- paste0(path, "/outputs/s/s_lad_per_firm_10/lad_10_")
 
@@
 ```{r eval=F}
-path.out <- paste0(path, "/outputs/s/s_la_per_firm.csv")
+path.out <- paste0(path, "/outputs/s/s_la_per_firm_10.csv")
 
@@
 ```{r}
-path.out <- paste0(path, "/outputs/s/speed_map.png")
+path.out <- paste0(path, "/outputs/s/speed_map_10.png")
 
@@
 
-Change percentile in out.path and uncomment `mutate(alpha = ifelse(abs(diff_1st) > ...`
+Change percentile in out.path and mutate(alpha = ifelse(abs(diff_1st)
 
@@
 
-out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only95_av.png")
+out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only95_av_10.png")
 
@@
 
-  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .05, na.rm = T))
-  # q.95 = 253, q.9 = 214, q.1 = 11, q.05 = 3.75. **abs()**
+  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .95, na.rm = T))
+  # q.95 = 194.5, q.9 = 162.5, q.1 = 10.5, q.05 = 5. **abs()**
 
-  mutate(alpha = ifelse(abs(diff_1st) > 253, .8, 0.1)) %>% 
-  #mutate(alpha = ifelse(abs(diff_1st) <= 3.75, .8, 0.1)) %>%
+  mutate(alpha = ifelse(abs(diff_1st) > 194.5, .8, 0.1)) %>% 
+  #mutate(alpha = ifelse(abs(diff_1st) <= 5, .8, 0.1)) %>%
 

### 2024-08-14 paper/submission1/tranos2023.tex
postpresentation

@@
 
-\section{Introduction}\label{sec1}
+\section{Introduction}\label{sec-introduction}
 
@@
 regarding the diffusion of related \emph{future} technologies. As per
-\citet{leibowicz2016representing}, historical studies agree on the fact
-that ``technologies diffuse at different times, at different rates, and
-to different extents in different places, and can be significantly
-influenced by policies'' \citep{victor1993}. \citet{meade2021modelling}
-highlight that a variety of actors have a direct interest in gaining
-such knowledge including network equipment suppliers; network operators,
-regulatory and local authorities. These processes and their effects vary
-a lot across scales: although the diffusion of a new technology might
-not be optimal at a local level, it might be beneficial from a global
-perspective as it could lead to faster diffusion to less advantaged
-places \citep{leibowicz2016representing}. Despite the spatial
-heterogeneity of such diffusion mechanisms and the policy relevance,
-there are very limited attempts in the literature to analyse the
-diffusion of new digital technologies at a detailed geographical level.
+\citet{leibowicz2016representing}, historical studies agree that
+technologies diffuse differently in terms of times, rates, and
+geographies and can be driven by related policies \citep{victor1993}.
+\citet{meade2021modelling} highlight that a variety of actors have a
+direct interest in gaining such knowledge including network equipment
+suppliers; network operators, regulatory and local authorities. These
+processes and their effects vary a lot across scales: although the
+diffusion of a new technology might not be optimal at a local level, it
+might be beneficial from a global perspective as it could lead to faster
+diffusion to less advantaged places \citep{leibowicz2016representing}.
+Despite the spatial heterogeneity of such diffusion mechanisms and the
+policy relevance, there are very limited attempts in the literature to
+analyse the diffusion of new digital technologies at a detailed
+geographical level.
 
@@
 diffusion to two processes: (i) epidemic-like mechanisms, which are
-governed by distance, proximity and social interactions, and (ii) by
+governed by distance, proximity and social interactions, and (ii)
 economic mechanisms as new innovations are adopted by users as they
@@
 \citep[e.g.][]{blank2018local, riddlesden2014broadband}. More details
-about the data and the data generation process can be found in Section
-\hyperref[sec3]{XX}.
+about the data and the data generation process can be found in
+Section~\ref{sec-datamethods}.
+
+\textbf{S curve??}
 
@@
 
-\section{Literature review}\label{sec2}
+\section{Literature review}\label{sec-litreview}
 
 Geographical diffusion is a synthesis of different processes. On the one
-hand, we can identify purely spatial or, in other words, contagious
-processes. Adjacency and, more broadly, distance are the key drivers of
+hand, a purely spatial or, in other words, contagious processes can be
+identified. Adjacency and, more broadly, distance are the key drivers of
 diffusion. This perspective draws similarities with epidemics:
@@
 consequently, proximity and exposure \citep{hivner2003facilitating}. On
-the other hand, we can identify hierarchical processes. Instead of
-horizontal distance-based diffusion mechanisms, the top-down hierarchy
-of urban systems drives diffusion. In reality, the synthesis of these
-two processes represents how new technologies diffuse over space and
-time \citep{morrill2020spatial}.
+the other hand, there is a hierarchical processes. Instead of horizontal
+distance-based diffusion mechanisms, the top-down hierarchy of urban
+systems shapes technological diffusion. In reality, the synthesis of
+these two processes represents how new technologies diffuse over space
+and time \citep{morrill2020spatial}.
 
@@
 seminal work on `Diffusion of Innovations' \citep{rogers2010diffusion}.
-Rogers being a sociologist, his work focused not on the diffusion of
-innovations over space and time, but instead on the adoption of new
-technologies and innovations by individuals and the individual
-mechanisms that drive the decisions behind adoption. He identified five
-groups of individuals regarding their adoption speed: innovators, early
-adopters, early majority, late majority and laggards. The key mechanism
-of diffusion and adoption is communication and how knowledge is
-transferred within a social system. Therefore, all approaches agree on
-the S-shaped diffusion and cumulative adoption pattern
-\citep{grubler1990rise}.
+Rogers being a sociologist focused not on the diffusion of innovations
+over space and time, but instead on the adoption of new technologies and
+innovations by individuals and the individual mechanisms that drive the
+decisions behind adoption. He identified five groups of individuals
+regarding their adoption speed: innovators, early adopters, early
+majority, late majority and laggards. The key mechanism of diffusion and
+adoption is communication and how knowledge is transferred within a
+social system. Therefore, all approaches agree on the S-shaped diffusion
+and cumulative adoption pattern \citep{grubler1990rise}.
 
@@
 neighbourhood effect and contaminate nearby locations. These are the
-three mechanisms that the empirical analysis in Section
-\phantomsection\label{sec4}{4} investigates.
+three mechanisms that the empirical analysis in
+Section~\ref{sec-results} investigates.
 
@@
 
-\section{Materials and Methods}\label{sec3}
+\section{Data and Methods}\label{sec-datamethods}
 
@@
 Districts (LAD) and the Output Areas (OA). The former is an
-administrative unit and there are
-\textbackslash textcolor\{red\}c.~380\textbackslash textcolor\{black\}
-such units in the UK. The latter is a census-based geographical unit,
-which is very small as there are c.~230,000 of them in the UK. This
-methodological choice will allow the mapping of the diffusion of web
-technologies and the assessment of the diffusion mechanisms at these two
-very different spatial scales.
-
-The population of websites is calculated using data from the Internet
-Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
+administrative unit and there are c.~374 such units in the UK. The
+latter is a census-based geographical unit, which is very small as there
+are c.~230,000 of them in the UK. This methodological choice will allow
+the mapping of the diffusion of web technologies and the assessment of
+the diffusion mechanisms at these two very different spatial scales.
+
+The counts of websites at these scales are calculated using data from
+the Internet Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
 and, specifically, the JISC UK Web Domain Dataset \citep{ukwebarchive}.
@@
 \citep{THELWALL2000441, blazquez2018big}. Although a UK company can
-adopt a generic TLD such as.com and such cases escape the data used
+adopt a generic TLD such as .com and these cases escape the data used
 here, such omissions should not affect our results given the popularity
@@
 
-The text from these webpages was scanned using regular expressions
+The text from these webpages was scanned using a regular expression
 (regex) to identify strings of text which resemble UK postcodes and one
@@
 
-Before that, the data cleaning process included a data aggregation step,
-through which the webpages were aggregated to the relevant websites.
-Based on the following example, all three webpages listed below are part
-of the same website (\url{http://www.website.co.uk}) and at the end only
-websites and not the nested webpages were considered as otherwise the
-metric would have been biased towards large websites.
+The data cleaning process included an aggregation step, through which
+the archived webpages were aggregated to the parent websites. This
+website reconstruction allows the creation of a \emph{website} instead
+of a \emph{webpage} density metric. Websites, which are hierarchical
+structures, tend to represent specific orgnanisations or entities, and,
+arguably, are more meaningful observational units than webpages, which
+ignore the upstream dependencies. Based on the following example, all
+three webpages are part of the same website
+(\url{http://www.website.co.uk}) and at the end only websites and not
+the nested webpages were considered as otherwise the metric would have
+been biased towards large, place specific, websites.
 
@@
 \item
-  \url{http://www.website.co.uk/webpage_a} B152TT
+  \url{http://www.website.co.uk/webpage_a} B15 2TT
 \item
-  \url{http://www.website.co.uk/webpage_b} BS81TH
+  \url{http://www.website.co.uk/webpage_b} BS8 1TH
 \item
-  \url{http://www.website.co.uk/webpage_c} B152TT
+  \url{http://www.website.co.uk/webpage_c} B15 2TT
 \end{itemize}
@@
 websites with multiple postcodes. As per the above example,
-(\url{http://www.website.co.uk}) includes two unique postcodes: B152TT
-and BS81SS. The distribution of postcodes for 2000 is presented in Table
-\ref{f2000}, which clearly illustrates the wide range. On one hand, we
-have websites anchored to a unique location (72\% of all the
-reconstructed websites), which may represent a small company with a
-single trading location. On the other hand, we have websites with
-thousands of different postcodes. Considering the timeperiod of the
-analysis, such cases can represent directories which used to be popular
-in the pre-search engines early times of internet
-\citep{tranosuk}.\footnote{See Figure A1 in \textbf{Appendix A in the
-  supplemental data online for examples}.}
+\href{http://www.website.co.uk}{www.website.co.uk} includes two unique
+postcodes: B15 2TT and BS8 1SS. The distribution of postcodes for 2000
+is presented in Table \ref{f2000}, which clearly illustrates the wide
+range. At the left end of the distribution, there are websites anchored
+to a unique location (72\% of all the reconstructed websites in 2000),
+which may represent a small company with a single trading location. At
+the right end, we have websites with thousands of different postcodes.
+Considering the time period of the analysis, such cases can represent
+directories which used to be popular in the pre-search engines early
+times of the commercial internet \citep{tranosuk}.\footnote{See Figure
+  A1 in \textbf{Appendix A in the supplemental data online for
+  examples}.}
 
 The analysis presented here is based on two subsets of these data.
-Firstly, the analysis is based on websites which only contain only one
-unique postcode and, therefore, the geolocation process might suffer
-from less noise. As a robustness check, the analysis is replicated the
-analysis for an extended subset of websites, which include up to ten
-unique postcodes. These websites are geolocated by equally attaching
-them to multiple locations. This extended sample includes 94 percent of
-all the archived websites in 2000.
+Firstly, on websites, which only contain only one unique postcode and,
+therefore, the geolocation process does not suffer from noise. As a
+robustness check, the analysis is replicated for an extended subset of
+websites, which include up to ten unique postcodes. These websites are
+geolocated by equally attaching them to multiple locations. This
+extended sample includes 94 percent of all the archived websites in
+2000.
 
@@
 
-To create the web density metric, the yearly website counts at the LAD
-level are standardised by the number of firms in LAD to avoid biases
-associated with LADs hosting a large number of firms. Given that there
-is no such statistic for the OA, the actual OA level counts are used.
-However, because of the the rather consistent spatial definition of OA
-(they host 40-250 households),\footnote{\url{https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies}}
+Another data cleaning step dealt with some extreme outliers. Figure
+\ref{pc1} plots the website counts for the top 1000 postcodes. Some
+obvious outliers can be observed for the 2004-2006 period, which can be
+attributed to a link farm \citep{BL2013links}.\\
+The website counts for these six postcodes (SE24 9HP, CV8 2ED, GL16 7YA,
+CW1 6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1000
+websites pointing to them, were replaced with predicted values based on
+a simple panel regression model with postcode fixed effects and yearly
+dummy variables. As the Section~\ref{sec-results} illustrates, this led
+to a small increase of the model predictive capability. To put the
+magnitude of the data imputation into perspective, this process affected
+6 out of the 557,808 postcodes present in the data.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{../../outputs/pc_year_1.png}
+
+}
+
+\caption{\label{pc1}Yearly website counts per postcode (top 1000)}
+
+\end{figure}%
+
+To create the website density metric, the yearly website counts at the
+LAD level are standardised by the number of firms in LAD to avoid biases
+associated with LAD hosting a large number of firms. Given that there is
+no such statistic for the OA, the actual OA level counts are used.
+However, because of the the consistent spatial definition of OA (they
+host 40-250 households),\footnote{\url{https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies}}
 website counts in OA are interpreted as a density metric too.
 
-This website metric\ldots{}
-
-\citet{wilson201281} growth function description in p.~86
-
-from R: Asym/(1+exp((xmid-input)/scal)) using the terms from Wilson k/(1
-+ e\^{}(t\_0 - t)/scal)
-
-Wilson: k(1 + e\^{}-b(t - t\_0)) So, scal = - 1/b
-
-The curve is symmetric* around t\_0
-
-\textbf{minimum R2 = 95\%}, see also \citet{grubler1990rise}
-
-The literature usually uses the saturation level as the asymptote. I am
-using the total number of websites as we cannot compute a rate.
-
-Moran's I as \citet{ding2010modeling} \textbf{TODO} for t\_0, diffusion
-speed
-
-\section{Results}\label{sec4}
+These website density metrics are used for the three different steps of
+the analysis. Firstly, a system-level analysis explores how the
+diffusion of web technologies in the UK fits with the well-established
+S-curve. To do so, the following logistic function (Equation \ref{eq:s})
+is estimated for the whole of the UK and for each LAD separately:
+
+\begin{align}
+y = k /(1 + e^{-(t-t_{o})})\label{eq:s}
+\end{align}
+
+\noindent \(k\) is the asymptote or, in other words, the saturation
+level, \(b\) the overall growth rate, and \(t_{0}\) the \emph{inflection
+point} of maximum growth at \(k/2\), where the logistic function is
+symmetrical \citep{wilson201281}. We compare the LAD \(t_0\) of each LAD
+against the \(t_0\) of the UK to delineate whether a LAD reached that
+point faster or slower than the country average. Importantly, an
+accuracy criterion was imposed and only S-curves with \(R^2 > 0.9\) were
+included in the analysis. To estimate Equation \ref{eq:s} a
+self-starting logistic growth model was employed using the \texttt{nls}
+and \texttt{SSlogis} functions in \texttt{R}.
+
+The system-level analysis also focuses on the volatility of the web
+adoption to depict places with high concentration of early adopters and,
+equally, latecomers. To do so, the change of the ranking of the UK LAD
+over time is plotted and discussed. Both the S-curve and the volatility
+analysis focus only on LAD as the very large number of OA would have
+made such analysis diffucult to visualise and interpret.
+
+Next, exploratory spatial data analysis depicts whether the two main
+drivers of spatial diffusion -- namely neighbourhood and hierarchy --
+underpin the diffusion of web technologies in the UK. To capture the
+former and following \citet{ding2010modeling} the Moran's I and the
+Local Index of Spatial Association (LISA) are estimated for website
+density. To address the hierarchy mechanism, the Gini coefficient -- a
+well established metric of inequality -- is calculated. All the above
+are computed and plotted longitudinally both for the LAD and the OA.
+
+Lastly, a modelling framework is developed to test the above diffusion
+mechanisms. The overarching aim is to build a model that can test the
+relationship between the website density and these mechanisms:
+
+\begin{align}
+Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S-curve_{t}\label{eq:rf.generic}
+\end{align}
+
+To estimate Equation \ref{eq:rf.generic}, Random Forest (RF) models were
+built. This is a popular machine learning algorith for both regression
+and classification problems \citep{biau2012analysis}. It was introduced
+by \citet{breiman2001random} and has\\
+gained popularity, becoming a go-to data science tool. RF can
+effectively handle skewed distributions and outliers, model non-linear
+relationships, require minimal hyperparameter tuning, exhibit low
+sensitivity to these parameters, and have relatively short training
+times \citep{Caruana2008, liaw2002classification, yan2020using}. These
+attributes match well with the website density data characteristics
+including skewness especially for OA. Also, the large data size (c.~230k
+data points for each of the 17 years) call for fast training times.
+Importantly, RF predictions tend to be more accurate than those from
+single regression trees and outperform Ordinary Least Squares in
+out-of-sample predictions, even with moderate-sized training data and a
+small number of predictors
+\citep{mullainathan2017machine, athey2019machine, sulaiman2011intelligent, pourebrahim2019trip, biau2012analysis}.
+
+RF is a tree-based ensemble learning algorithm
+\citep{breiman2001random}. It begins by generating random samples of the
+training data, which are then used to grow regression trees to predict
+the dependent variable. Data points and predictors are randomly sampled
+for the different trees. The trees are trained in parallel using their
+own bootstrapped samples of the training data. A crucial feature of RF
+is their ability not to overfit, meaning they can generalize well to
+unseen test data. While each tree may overfit individually, the ensemble
+of trees does not because the errors of individual trees are averaged,
+reducing the overall variance and preventing overfitting
+\citep{last2002improving}. For regression problems, RF predictions are
+made by averaging the predictions of all decision trees.
+
+RF have been widely employed to address regression research problems.
+\citet{pourebrahim2019trip} combined a spatial interaction modeling
+framework with ML algorithms including RF to predict commuting flows in
+New York City. \citet{sinha2019assessing} advocated for adopting spatial
+ensemble learning approaches, such as RF, to model spatial data with
+high autocorrelation and heterogeneity. \citet{creditspatial} predicted
+employment density in Los Angeles using spatially explicit RF.
+\citet{guns2014recommending} used RF to build a recommendation system
+for research collaborations. \citet{ren2019predicting} trained RF to
+predict the socio-economic status of cities using various online and
+mobility predictors. \citet{tranos2023using} utilised hyperlinks data
+and RF to make out-of-sample predictions of interregional trade.
+\citet{zhou2023geography} employed such a framework to assess whether
+key predictors of obesity differ across English cities.
+
+\section{Results}\label{sec-results}
 
@@
 \item
-  S-shaped diffusion curves: S for LADs per firm. UK, fast/slow/examples
+  S-shaped diffusion curves: S for LAD per firm. UK, fast/slow/examples
 \item
@@
 \item
-  ideal: (i) train RF for all years and all (1) LADs and (2) OAs with
-  CAST and report metrics. (ii) train for all years and all but one
-  region for (1) LADs and (2) OAs to predict to the holdout region.
-  Reports predictions as region similarities.
+  ideal: (i) train RF for all years and all (1) LAD and (2) OA with CAST
+  and report metrics. (ii) train for all years and all but one region
+  for (1) LAD and (2) OA to predict to the holdout region. Reports
+  predictions as region similarities.
 \end{itemize}
 
-\subsubsection{RF results}\label{rf-results}
+Figure \ref{s_uk} plots the S curve for the cumulative adoption of
+website technologies in the UK during the 1996-2012 period. It
+demonstrates a pattern well aligned with previous studies discussed in
+Section~\ref{sec-litreview}. The vertical line for year 2003 illustrates
+the point where the modeled cumulative adoption was equal to 50\% of the
+maximum. This \(t_0\) \emph{inflection} point signals the maximum
+adoption speed and is used here to determine whether the UK LAD reached
+that inflection point earlier or later than the UK average.
+Specifically, the S curve and the inflection point are estimated for
+every LAD individually and then compared with the UK average. LAD that
+reached their inflection point earlier than the country average are
+labelled as \emph{fast} and the rest as \emph{slow}. Figure \ref{s_map}
+maps this pattern and the picture is not entirely clear. On the one
+hand, there are some expected examples of LAD with relevant industrial
+backgrounds delineated as fast: the City of London, a world-renowned
+cluster of finance industries \citep{cook2007role}, and Reading, a town
+with high-tech service industries in proximity to London and its main
+airport, Heathrow \citep{england2005polynet}. On the other hand, LAD
+which were expected to appear as fast -- e.g.~Hackney in central London
+and Bristol, a well-established creative cluster
+\citep{oatley1999cultural, bassett2002cultural} -- were delineated as
+slow. Nevertheless, out of the 10 fastest LAD, eight were located in
+Greater London and one in Cambridge (see \textbf{Appendix}), but in
+overall there is a spatially heterogenous and not easy to explain
+spatial pattern.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/s_uk_per_firm.png}
+
+}
+
+\caption{\label{s_uk}Grwoth curve, UK}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/speed_map.png}
+
+}
+
+\caption{\label{s_map}LAD adoption rates}
+
+\end{figure}%
+
+The next step is to assess the stability and volatility of the LAD in
+terms of their adoption of web technologies. As we know from the
+literature \citep{risk_perceptions}, different agents have different
+perceptions about and levels of acceptance of the risks and the
+potential economic returns associated with the adoption of new
+technologies -- see for instance the seminal work of
+\citet{venkatesh2000theoretical}. To reveal such aggregated patterns,
+Figures \ref{rank_stable} and \ref{rank_unstable} plot the ranking of UK
+LAD based on website density. To decrease noise, the average ranking of
+1996-1998 and 2010-2012 is plotted instead of the individual years.
+While there are quite a few obvious cases of LAD that maintained their
+position between the beginning and the end of the study period either at
+the top or at the bottom of the hierarchy (Figure \ref{rank_stable}),
+there are also quite a few LAD that changed drastically their position.
+Some of these LAD enjoyed a process that at the first instance looks
+like leapfrogging since they managed to jump at the top of the hierarchy
+despite their slow start (top of Figure \ref{unstable}). There is
+extensive literature regarding the potential benefits of technological
+leapfrogging. The underpinning argument is that latecomers can adopt and
+benefit from new technologies that have been developed elsewhere without
+incurring the hefty initial R\&D costs \citep{teece2008firm}. Although
+the leapfrogging literature does not pay much attention to cities and
+regions \citep{yu2018sustainability}, previous research highlighted the
+long term and sustained productivity benefit of the early adopters of
+web technologies in the UK \citep{tranosuk}. So, although the LAD system
+is volatile, it is not clear whether the LAD with high concentration of
+late adopters will gain any latecomer benefits in a way similar to
+countries experiencing such technological leapfrogging.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/ranks/web_per_firm2000_2012_only05_av.png}
+
+}
+
+\caption{\label{rank_stable}Dynamics of wed diffusion: stability of LAD}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/ranks/web_per_firm2000_2012_only95_av.png}
+
+}
+
+\caption{\label{rank_unstable}Dynamics of wed diffusion: volatility LAD}
+
+\end{figure}%
+
+Then, figures \ref{morani} and \ref{lisa} offer a first insight into
+whether a neighbourhood effect underpins the diffusion of web
+technologies in the UK as they plot the Moran's I and the LISA maps of
+website density respectively for LAD and OA. Starting from the former,
+spatial autocorrelation was higher in the beginning and then around year
+2000 dropped slightly and stabled at 0.2. This reflects the early
+concentration of high website density around London, which over time
+diffused as high-high clusters can be seen in other parts of the country
+away from London as seen in Figure \ref{lisa}. An almost reverse pattern
+can be observed for OA. At the beginning of the study period Moran's I
+was around 0.1 and it plateaued after 2000-2001 around 0.2. Because of
+the very small size of OA, at the early stages of the diffusion of web
+technologies their adoption was spatially scattered. This is reflected
+in the lack of any significant clusters in 1996. Eventually, as the
+adoption rate increased, more such clusters of high website density were
+formed and this is reflected both in the Moran's I and the LISA maps in
+Figures \ref{morani} and \ref{lisa}. All in all, the exploratory spatial
+data analysis advocates towards an underpinning neighbourhood effect and
+the different scales of analysis illustrate how it evolved differently
+over time.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2023_files/figure-pdf/morani-1.pdf}
+
+}
+
+\caption{\label{morani}Website density Moran's I}
+
+\end{figure}%
+
+Figure \ref{lisa}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2023_files/figure-pdf/unnamed-chunk-7-1.pdf}
+
+}
+
+\caption{\label{lisa}\centering Website density LISA maps}
+
+\end{figure}%
+
+To illustrate whether a hierarchical process underpins the diffusion of
+web technologies, the Gini coefficient is calculated yearly both for LAD
+and OA. As a metric of inequality, the Gini coefficient demonstrates
+whether website density is concentrated in a small number of LAD/OA, or
+whether it is more equally spread across the country. Both scales of
+analysis in Figure \ref{gini} illustrate the same picture. At the
+beginning of the commercial internet website density was extremely
+unequal, or in other words, only a few places had websites associated
+with them. Inequality dropped and plateaued after 2000 for both scales
+illustrating that at the first stages of the commercial internet,
+website density was concentrated in a limited number of places. This is
+an indication of a hierarchical diffusion mechanism that led over time
+to a more equal spread. Interestingly, the year 2000 is again a period
+of change for this diffusion mechanisms as it was for the neighbourhood
+process. There is a substantial difference between the Gini coefficient
+magnitude for LAD and OA, but this is well expected as the very small
+size of OA equates to a lot polygons without any websites pointing to
+them -- for instance residential OA. LAD/OA
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2023_files/figure-pdf/gini-1.pdf}
+
+}
+
+\caption{\label{gini}Website density Gini coefficient}
+
+\end{figure}%
 
 The next section incorporates the above discussed spatial processes of
-the diffusion of web technologies into the same modelling framework. The
-aim is to use variables depicting these spatial processes in order to
-predict the diffusion of web technologies in the UK over space and time
-and across different scales. Specifically, four different models are
-estimated. Firstly, all the data points at the OA and LAD are utilised
-in order to build RF models and assess their capacity to predict the
-adoption of web technologies. These two models will reveal the
-predictive capacity of the spatial processes and also allow to see how
-the importance of such variables changes across scales. The next two
-sets of models will be again trained on web diffusion at the two working
-scales: OA and LAD. However, instead of using all the data points, the
-OA and the LAD from one of the twelve UK regions are held out and then
-the trained model is used to predict web adoption in the OA or the LAD
-of the held-out region. This process takes place recursively for all UK
-regions. The difference in the predictive capacity of the different
-samples will reveal how dissimilar are these spatial process across
-regions and, importantly, at different scales.
+the diffusion of web technologies into a modelling framework. The aim is
+to use variables depicting these spatial processes in order to predict
+the diffusion of web technologies in the UK over space and time and
+across different scales. Specifically, four different models are
+estimated. Firstly, all the data points for the OA and LAD are utilised
+in order to build two respective RF models and assess their capacity to
+predict the adoption of web technologies. These two models will reveal
+the predictive capacity of the diffusion mechanisms and also allow to
+see how the importance of such variables changes across scales. The next
+two sets of models will be again trained on web diffusion at the two
+working scales: OA and LAD. However, instead of using all the data
+points, the OA and the LAD from one of the twelve UK regions are held
+out and then the trained model is used to predict website density in the
+OA or the LAD of the held-out region. This process takes place
+recursively for all UK regions. The difference in the predictive
+capacity of the different samples will reveal how dissimilar are these
+spatial process across regions and, importantly, at different scales.
 
@@
 
-The models need to include variables that capture the three processes
-that the relevant literature and the descriptive analysis presented in
-the Section \textbf{XX} highlighted. Namely, the models capture: (i) a
-hierarchy effect with diffusion running from main centres to secondary
-ones, (ii) a neighborhood effect according to which diffusion first hits
-nearby locations, and (iii) the rather canonical pattern of diffusion
-over time as reflected in the S-shaped pattern in the cumulative level
-of adoption.
+The models include variables that capture the three processes that the
+relevant literature and the descriptive analysis presented above
+highlighted. Namely, the models capture: (i) a hierarchy effect with
+diffusion running from main centres to secondary ones, (ii) a
+neighborhood effect according to which diffusion first hits nearby
+locations, and (iii) the rather canonical pattern of diffusion over time
+as reflected in the S-shaped pattern in the cumulative level of
+adoption.
 
 To capture the hierarchy effect the models include as predictors a one
-year lag of website density in London -- the largest city in the UK, a
-one year lag of the website density in the nearest city and the same for
-the nearest retail centre. Due to the small sizes of the retail centres,
-the latter is only relevant for the OA-level models. In addition, the
-models include the distance to London, the nearest city and the nearest
-retail centre. The underlying logic is that the level of website
-adoption in a spatial unit depends on the level of the adoption in
-places further up in the urban hierarchy the previous year. To depict
-the neighbourhood effect, the web density of the neighbouring spatial
-units in the previous year is employed. Again, the underpinning
+year lag of website density in London, the largest city in the UK, a one
+year lag of the website density in the nearest city and the same for the
+nearest retail centre. Due to the small sizes of the retail centres, the
+latter is only relevant for the OA-level models. In addition, the models
+include the distance to London, the nearest city and the nearest retail
+centre. The underlying logic is that the level of website adoption in a
+spatial unit depends on the level of the adoption in places further up
+in the urban hierarchy the previous year. The inclusion of the distance
+variables incorporates spatial structure into the hierarchy argument. To
+depict the neighbourhood effect, the website density of the neighbouring
+spatial units in the previous year is employed. Again, the underpinning
 rationale is that the level of web adoption within a spatial unit
 depends on the level of web adoption in the neighbouring spatial units
-the year before. This is the `hitting nearby locations first' argument.
-Therefore, the spatial and temporal lag of the website density in LAD
-and OA is calculated. Lastly, the time effect which is reflected on the
-S curve for the cumulative adoption is captured by trend variable.
-Hence, all four model will follow the following generic form (Eq.
-\ref{model}):
+the year before. This represents the `hitting nearby locations first'
+argument. Therefore, the spatial and temporal lag of the website density
+in LAD and OA is calculated. Lastly, the time effect which is reflected
+on the S-curve for the cumulative adoption is captured by a time trend
+variable. Hence, all four model will follow the following generic form
+(Eq. \ref{model}):
 
@@
 To assess the predictive capability of the model, three broadly utilised
-metrics are employed: the coefficient of determination (R squared), mean
+metrics are employed: the coefficient of determination (\(R^2\)), mean
 absolute error (MAE) and root mean square error (RMSE):
@@
 
-Table \ref{table.metrics.all} presents the model performance for the
-first set on models, for which all data points are employed for training
+Table~\ref{tbl-model-metrics} presents the model performance for the
+first set of models, for which all data points are employed for training
 and testing via cross validation. The first one is trained and tested on
@@
 error is \((\frac{1}{20}, \frac{1}{30})\) of a website per firm. At the
-OA the RSquared drops down to 21\%. Considering its granularity, this is
+OA the \(R^2\) drops down to 21\%. Considering its granularity, this is
 still a remarkable performance. To contextualise it, the model results
@@
 highly skewed and a significant part of them is not linked to any
-websites. In 1997 only 1\% of the UK OA were associated with a website.
-This should not come as a surprise as this was the very beginning of the
-commercial internet and any activities with a digital footprint were
-concentrated in a handful of areas. This was clearly illustrated in
-Section \textbf{ADD}. At the end of the study period almost half of the
-UK OA were not associated with a website. Again, given the granularity
-of the data this should not come as a surprise.
+websites. In 1997 only 1\% of the UK OA were associated with at least
+one website. This should not come as a surprise as this was the very
+beginning of the commercial internet and any activities with a digital
+footprint were concentrated in a handful of areas. This was clearly
+illustrated in Figure \ref{lisa}. At the end of the study period almost
+half of the UK OA were not associated with a website. Again, given the
+granularity of the data this should not come as a surprise.
 
 \begin{longtable}[]{@{}llll@{}}
-\caption{Model metrics \{\#table.metrics.all\}}\tabularnewline
+\caption{Model metrics}\label{tbl-model-metrics}\tabularnewline
 \toprule\noalign{}
-& RMSE & RSquared & MAE \\
+& RMSE & \(R^{2}\) & MAE \\
 \midrule\noalign{}
@@
 \toprule\noalign{}
-& RMSE & RSquared & MAE \\
+& RMSE & \(R^{2}\) & MAE \\
 \midrule\noalign{}
@@
 the different predictors. When the focus is on the LAD, the website
-density in the nearest city, in London and in the neighbouring LADs the
+density in the nearest city, in London and in the neighbouring LAD the
 year before are the most important predictors. They are followed by the
@@
 \caption{\label{var.imp.OA}Variable importance, OA}
+
+\end{figure}%
+
+Figure \ref{var.imp} to replace Figures \ref{var.imp.LAD} and
+\ref{var.imp.OA} once I have the variable importance for LAD.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2023_files/figure-pdf/varimp-1.pdf}
+
+}
+
+\caption{\label{var.imp}Variable importance}
 
@@
 technologies within its LAD and OA. It needs to be highlighted though
-that the difference between the RSquared of LAD and OA is more that an
+that the difference between the \(R^2\) of LAD and OA is more that an
 order of magnitude signaling how difficult is to predict diffusion at
@@
 
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3731}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1940}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1343}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1791}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1194}}@{}}
+\begin{longtable}[]{@{}lrrrr@{}}
 \caption{Regional differences\label{table.regions}}\tabularnewline
 \toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Region
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-RSquared LAD
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-Rank LAD
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-RSquared OA
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-Rank OA
-\end{minipage} \\
+Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
 \midrule\noalign{}
@@
 \toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Region
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-RSquared LAD
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-Rank LAD
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-RSquared OA
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-Rank OA
-\end{minipage} \\
+Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
 \midrule\noalign{}
@@
 
-\section{Discussion and conclusions}\label{discussion-and-conclusions}
+\section{Discussion and conclusions}\label{sec-conclusions}
 
@@
 
+\setcounter{section}{0}
+\renewcommand{\thesection}{\Alph{section}}
+
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{A\arabic{figure}}
+
+\section{Appendices}\label{appendices}
+
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
+
+\subsection{\texorpdfstring{\(t_0\) estunates for all
+LAD}{t\_0 estunates for all LAD}}\label{t_0-estunates-for-all-lad}
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4091}}
+  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1818}}
+  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1364}}
+  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0795}}
+  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1932}}@{}}
+\caption{S-curve estiamtes for LAD\label{table.s.lads}}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+LAD
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(t_0\) estimate
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Std. error
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(R^2\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Diffusion speed
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+LAD
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(t_0\) estimate
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Std. error
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(R^2\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Diffusion speed
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Horsham & 2001.423 & 0.254 & 0.950 & fast \\
+Fareham & 2001.443 & 0.346 & 0.925 & fast \\
+Kingston upon Thames & 2001.513 & 0.354 & 0.933 & fast \\
+Kensington and Chelsea & 2001.581 & 0.341 & 0.943 & fast \\
+Runnymede & 2001.582 & 0.266 & 0.943 & fast \\
+Bracknell Forest & 2001.586 & 0.324 & 0.928 & fast \\
+Elmbridge & 2001.656 & 0.333 & 0.932 & fast \\
+Reigate and Banstead & 2001.665 & 0.253 & 0.964 & fast \\
+South Cambridgeshire & 2001.680 & 0.433 & 0.905 & fast \\
+Walsall & 2001.715 & 0.413 & 0.901 & fast \\
+Surrey Heath & 2001.718 & 0.333 & 0.933 & fast \\
+Woking & 2001.759 & 0.292 & 0.951 & fast \\
+South Norfolk & 2001.874 & 0.358 & 0.931 & fast \\
+City of London & 2001.874 & 0.499 & 0.930 & fast \\
+Wokingham & 2001.891 & 0.287 & 0.958 & fast \\
+Reading & 2001.900 & 0.318 & 0.948 & fast \\
+Sevenoaks & 2001.904 & 0.407 & 0.926 & fast \\
+Huntingdonshire & 2001.912 & 0.402 & 0.925 & fast \\
+Pendle & 2001.924 & 0.407 & 0.923 & fast \\
+St Albans & 2001.933 & 0.396 & 0.927 & fast \\
+Perth and Kinross & 2001.943 & 0.228 & 0.970 & fast \\
+Bromley & 2001.945 & 0.396 & 0.919 & fast \\
+Rushmoor & 2001.962 & 0.518 & 0.912 & fast \\
+Powys & 2001.969 & 0.278 & 0.963 & fast \\
+Swindon & 2001.996 & 0.390 & 0.916 & fast \\
+Chelmsford & 2002.008 & 0.545 & 0.911 & fast \\
+Crawley & 2002.008 & 0.379 & 0.949 & fast \\
+Dumfries and Galloway & 2002.012 & 0.251 & 0.967 & fast \\
+Inverclyde & 2002.030 & 0.330 & 0.915 & fast \\
+North Hertfordshire & 2002.031 & 0.475 & 0.920 & fast \\
+Moray & 2002.050 & 0.358 & 0.947 & fast \\
+Orkney Islands & 2002.070 & 0.283 & 0.958 & fast \\
+Fife & 2002.070 & 0.363 & 0.945 & fast \\
+Mole Valley & 2002.071 & 0.460 & 0.925 & fast \\
+Guildford & 2002.113 & 0.413 & 0.941 & fast \\
+Buckinghamshire & 2002.114 & 0.390 & 0.942 & fast \\
+Cotswold & 2002.116 & 0.375 & 0.946 & fast \\
+Shetland Islands & 2002.120 & 0.332 & 0.949 & fast \\
+South Oxfordshire & 2002.146 & 0.367 & 0.950 & fast \\
+Watford & 2002.157 & 0.451 & 0.929 & fast \\
+Aberdeenshire & 2002.168 & 0.416 & 0.931 & fast \\
+Southampton & 2002.189 & 0.472 & 0.922 & fast \\
+Mid Suffolk & 2002.196 & 0.442 & 0.922 & fast \\
+Eden & 2002.196 & 0.334 & 0.949 & fast \\
+South Lakeland & 2002.197 & 0.238 & 0.974 & fast \\
+Stoke-on-Trent & 2002.203 & 0.419 & 0.927 & fast \\
+Babergh & 2002.209 & 0.449 & 0.927 & fast \\
+Bexley & 2002.213 & 0.493 & 0.909 & fast \\
+Torbay & 2002.215 & 0.374 & 0.942 & fast \\
+Brentwood & 2002.219 & 0.336 & 0.936 & fast \\
+Ceredigion & 2002.227 & 0.241 & 0.971 & fast \\
+Spelthorne & 2002.232 & 0.451 & 0.933 & fast \\
+Bath and North East Somerset & 2002.238 & 0.402 & 0.941 & fast \\
+Falkirk & 2002.245 & 0.367 & 0.938 & fast \\
+Broxtowe & 2002.258 & 0.435 & 0.932 & fast \\
+East Hertfordshire & 2002.261 & 0.453 & 0.928 & fast \\
+Stirling & 2002.268 & 0.251 & 0.975 & fast \\
+Ealing & 2002.294 & 0.523 & 0.922 & fast \\
+Mid Sussex & 2002.295 & 0.453 & 0.936 & fast \\
+Barrow-in-Furness & 2002.303 & 0.284 & 0.959 & fast \\
+West Oxfordshire & 2002.322 & 0.467 & 0.929 & fast \\
+Sutton & 2002.326 & 0.544 & 0.911 & fast \\
+Portsmouth & 2002.356 & 0.479 & 0.921 & fast \\
+Great Yarmouth & 2002.369 & 0.265 & 0.968 & fast \\
+East Hampshire & 2002.370 & 0.474 & 0.931 & fast \\
+Wyre Forest & 2002.376 & 0.561 & 0.916 & fast \\
+Angus & 2002.378 & 0.396 & 0.939 & fast \\
+Argyll and Bute & 2002.427 & 0.321 & 0.962 & fast \\
+Scottish Borders & 2002.428 & 0.298 & 0.962 & fast \\
+East Suffolk & 2002.431 & 0.533 & 0.919 & fast \\
+Blackpool & 2002.443 & 0.294 & 0.962 & fast \\
+Cherwell & 2002.448 & 0.533 & 0.923 & fast \\
+Wandsworth & 2002.449 & 0.570 & 0.921 & fast \\
+Tendring & 2002.456 & 0.456 & 0.935 & fast \\
+Westminster & 2002.459 & 0.437 & 0.952 & fast \\
+Richmond upon Thames & 2002.462 & 0.389 & 0.956 & fast \\
+Test Valley & 2002.463 & 0.553 & 0.920 & fast \\
+Isle of Wight & 2002.478 & 0.377 & 0.953 & fast \\
+Vale of White Horse & 2002.496 & 0.408 & 0.949 & fast \\
+Croydon & 2002.510 & 0.504 & 0.927 & fast \\
+Hounslow & 2002.521 & 0.426 & 0.952 & fast \\
+Copeland & 2002.524 & 0.237 & 0.971 & fast \\
+Wiltshire & 2002.528 & 0.450 & 0.940 & fast \\
+Gwynedd & 2002.536 & 0.372 & 0.954 & fast \\
+Hillingdon & 2002.560 & 0.470 & 0.940 & fast \\
+Conwy & 2002.564 & 0.372 & 0.954 & fast \\
+Highland & 2002.566 & 0.270 & 0.972 & fast \\
+Basingstoke and Deane & 2002.572 & 0.461 & 0.947 & fast \\
+Harrow & 2002.587 & 0.509 & 0.936 & fast \\
+West Berkshire & 2002.595 & 0.696 & 0.902 & fast \\
+North Lincolnshire & 2002.596 & 0.501 & 0.933 & fast \\
+East Dunbartonshire & 2002.598 & 0.528 & 0.925 & fast \\
+Bradford & 2002.598 & 0.613 & 0.916 & fast \\
+Waverley & 2002.606 & 0.512 & 0.928 & fast \\
+Stroud & 2002.608 & 0.559 & 0.916 & fast \\
+Herefordshire, County of & 2002.622 & 0.388 & 0.956 & fast \\
+North Norfolk & 2002.624 & 0.422 & 0.945 & fast \\
+Nottingham & 2002.630 & 0.531 & 0.936 & fast \\
+Castle Point & 2002.645 & 0.470 & 0.936 & fast \\
+North Warwickshire & 2002.657 & 0.380 & 0.955 & fast \\
+Gravesham & 2002.680 & 0.473 & 0.933 & fast \\
+Lancaster & 2002.692 & 0.336 & 0.963 & fast \\
+West Suffolk & 2002.699 & 0.536 & 0.930 & fast \\
+Stafford & 2002.719 & 0.499 & 0.933 & fast \\
+Tunbridge Wells & 2002.720 & 0.401 & 0.958 & fast \\
+Oxford & 2002.723 & 0.583 & 0.934 & fast \\
+East Lothian & 2002.731 & 0.537 & 0.927 & fast \\
+Redcar and Cleveland & 2002.750 & 0.509 & 0.929 & fast \\
+Tamworth & 2002.755 & 0.551 & 0.929 & fast \\
+South Hams & 2002.773 & 0.422 & 0.952 & fast \\
+Allerdale & 2002.784 & 0.376 & 0.958 & fast \\
+Central Bedfordshire & 2002.825 & 0.620 & 0.921 & fast \\
+East Cambridgeshire & 2002.843 & 0.636 & 0.927 & fast \\
+Havant & 2002.853 & 0.541 & 0.932 & fast \\
+South Lanarkshire & 2002.874 & 0.486 & 0.938 & fast \\
+Stratford-on-Avon & 2002.879 & 0.470 & 0.945 & fast \\
+Arun & 2002.882 & 0.526 & 0.938 & fast \\
+Fenland & 2002.891 & 0.632 & 0.918 & fast \\
+Monmouthshire & 2002.893 & 0.501 & 0.935 & fast \\
+Denbighshire & 2002.901 & 0.457 & 0.947 & fast \\
+Southend-on-Sea & 2002.903 & 0.551 & 0.932 & fast \\
+High Peak & 2002.903 & 0.488 & 0.946 & fast \\
+Pembrokeshire & 2002.905 & 0.387 & 0.958 & fast \\
+Malvern Hills & 2002.907 & 0.524 & 0.940 & fast \\
+Camden & 2002.915 & 0.599 & 0.941 & fast \\
+North Devon & 2002.926 & 0.426 & 0.950 & fast \\
+Wyre & 2002.927 & 0.553 & 0.927 & fast \\
+Darlington & 2002.933 & 0.615 & 0.929 & fast \\
+Swale & 2002.940 & 0.447 & 0.948 & fast \\
+Welwyn Hatfield & 2002.951 & 0.628 & 0.922 & fast \\
+Mid Devon & 2002.960 & 0.485 & 0.946 & fast \\
+Lewes & 2002.964 & 0.507 & 0.945 & fast \\
+Dorset & 2002.968 & 0.407 & 0.958 & fast \\
+Brighton and Hove & 2002.976 & 0.578 & 0.934 & fast \\
+Braintree & 2002.978 & 0.534 & 0.938 & fast \\
+Epsom and Ewell & 2002.979 & 0.746 & 0.908 & fast \\
+Sefton & 2002.980 & 0.616 & 0.928 & fast \\
+Eastleigh & 2002.992 & 0.595 & 0.929 & fast \\
+Craven & 2002.993 & 0.379 & 0.965 & fast \\
+Hammersmith and Fulham & 2002.997 & 0.585 & 0.944 & fast \\
+Greenwich & 2003.014 & 0.571 & 0.930 & slow \\
+Cheltenham & 2003.025 & 0.602 & 0.937 & slow \\
+Worcester & 2003.041 & 0.659 & 0.920 & slow \\
+Tower Hamlets & 2003.043 & 0.660 & 0.926 & slow \\
+New Forest & 2003.046 & 0.554 & 0.941 & slow \\
+East Renfrewshire & 2003.046 & 0.739 & 0.900 & slow \\
+Fylde & 2003.056 & 0.501 & 0.946 & slow \\
+Carmarthenshire & 2003.067 & 0.390 & 0.958 & slow \\
+West Lancashire & 2003.076 & 0.547 & 0.937 & slow \\
+Hastings & 2003.086 & 0.685 & 0.911 & slow \\
+Carlisle & 2003.111 & 0.541 & 0.934 & slow \\
+Wychavon & 2003.115 & 0.570 & 0.935 & slow \\
+Isle of Anglesey & 2003.144 & 0.403 & 0.960 & slow \\
+Dudley & 2003.150 & 0.678 & 0.925 & slow \\
+Ryedale & 2003.152 & 0.380 & 0.961 & slow \\
+Rother & 2003.162 & 0.366 & 0.966 & slow \\
+Leicester & 2003.167 & 0.686 & 0.924 & slow \\
+Breckland & 2003.182 & 0.567 & 0.938 & slow \\
+Tandridge & 2003.190 & 0.747 & 0.918 & slow \\
+Shropshire & 2003.206 & 0.621 & 0.928 & slow \\
+West Devon & 2003.228 & 0.489 & 0.949 & slow \\
+Milton Keynes & 2003.231 & 0.700 & 0.929 & slow \\
+Sandwell & 2003.234 & 0.760 & 0.916 & slow \\
+Winchester & 2003.234 & 0.443 & 0.956 & slow \\
+Redditch & 2003.238 & 0.706 & 0.923 & slow \\
+Rugby & 2003.265 & 0.709 & 0.922 & slow \\
+North Somerset & 2003.278 & 0.558 & 0.939 & slow \\
+West Lindsey & 2003.285 & 0.674 & 0.924 & slow \\
+Richmondshire & 2003.286 & 0.425 & 0.958 & slow \\
+Melton & 2003.292 & 0.534 & 0.946 & slow \\
+Chichester & 2003.295 & 0.582 & 0.942 & slow \\
+Torridge & 2003.301 & 0.532 & 0.942 & slow \\
+Dundee City & 2003.303 & 0.568 & 0.939 & slow \\
+South Ayrshire & 2003.303 & 0.626 & 0.931 & slow \\
+East Devon & 2003.308 & 0.492 & 0.947 & slow \\
+Bolton & 2003.323 & 0.774 & 0.913 & slow \\
+Hertsmere & 2003.326 & 0.726 & 0.924 & slow \\
+Bournemouth, Christchurch and Poole & 2003.347 & 0.662 & 0.936 & slow \\
+Maldon & 2003.349 & 0.495 & 0.951 & slow \\
+Barnet & 2003.352 & 0.766 & 0.919 & slow \\
+Brent & 2003.389 & 0.762 & 0.925 & slow \\
+Midlothian & 2003.413 & 0.732 & 0.921 & slow \\
+Boston & 2003.416 & 0.593 & 0.933 & slow \\
+Scarborough & 2003.421 & 0.431 & 0.963 & slow \\
+Kirklees & 2003.435 & 0.725 & 0.925 & slow \\
+Tameside & 2003.441 & 0.819 & 0.907 & slow \\
+Newport & 2003.449 & 0.635 & 0.930 & slow \\
+Glasgow City & 2003.451 & 0.759 & 0.922 & slow \\
+Clackmannanshire & 2003.454 & 0.585 & 0.936 & slow \\
+Three Rivers & 2003.458 & 0.628 & 0.947 & slow \\
+Harborough & 2003.488 & 0.549 & 0.950 & slow \\
+Rochdale & 2003.496 & 0.685 & 0.937 & slow \\
+Tonbridge and Malling & 2003.497 & 0.710 & 0.927 & slow \\
+Somerset West and Taunton & 2003.502 & 0.491 & 0.955 & slow \\
+Wrexham & 2003.512 & 0.553 & 0.942 & slow \\
+Worthing & 2003.522 & 0.622 & 0.938 & slow \\
+Tewkesbury & 2003.540 & 0.580 & 0.946 & slow \\
+City of Edinburgh & 2003.541 & 0.685 & 0.941 & slow \\
+Coventry & 2003.592 & 0.564 & 0.948 & slow \\
+Northumberland & 2003.607 & 0.508 & 0.954 & slow \\
+Mid and East Antrim & 2003.609 & 0.893 & 0.900 & slow \\
+Wigan & 2003.617 & 0.755 & 0.929 & slow \\
+Lichfield & 2003.629 & 0.625 & 0.939 & slow \\
+Ashford & 2003.633 & 0.535 & 0.944 & slow \\
+Cornwall & 2003.638 & 0.443 & 0.963 & slow \\
+Mendip & 2003.652 & 0.505 & 0.955 & slow \\
+Stevenage & 2003.665 & 0.525 & 0.958 & slow \\
+Blaenau Gwent & 2003.666 & 0.643 & 0.933 & slow \\
+Lincoln & 2003.666 & 0.634 & 0.940 & slow \\
+South Somerset & 2003.676 & 0.553 & 0.948 & slow \\
+Charnwood & 2003.677 & 0.699 & 0.934 & slow \\
+North Ayrshire & 2003.682 & 0.607 & 0.942 & slow \\
+Folkestone and Hythe & 2003.684 & 0.792 & 0.905 & slow \\
+Eastbourne & 2003.684 & 0.697 & 0.930 & slow \\
+Blackburn with Darwen & 2003.685 & 0.680 & 0.938 & slow \\
+Dover & 2003.699 & 0.574 & 0.948 & slow \\
+Adur & 2003.735 & 0.551 & 0.954 & slow \\
+Haringey & 2003.739 & 0.667 & 0.943 & slow \\
+Gosport & 2003.751 & 0.600 & 0.934 & slow \\
+Trafford & 2003.752 & 0.640 & 0.949 & slow \\
+Norwich & 2003.756 & 0.742 & 0.939 & slow \\
+West Lothian & 2003.758 & 0.652 & 0.943 & slow \\
+North East Lincolnshire & 2003.811 & 0.858 & 0.920 & slow \\
+Bristol, City of & 2003.828 & 0.831 & 0.928 & slow \\
+Rushcliffe & 2003.839 & 0.591 & 0.954 & slow \\
+Hinckley and Bosworth & 2003.857 & 0.749 & 0.934 & slow \\
+Torfaen & 2003.861 & 0.841 & 0.922 & slow \\
+Swansea & 2003.875 & 0.650 & 0.947 & slow \\
+Wolverhampton & 2003.882 & 0.793 & 0.931 & slow \\
+Kingston upon Hull, City of & 2003.885 & 0.757 & 0.936 & slow \\
+Calderdale & 2003.889 & 0.714 & 0.939 & slow \\
+North Northamptonshire & 2003.902 & 0.746 & 0.935 & slow \\
+Stockton-on-Tees & 2003.902 & 0.955 & 0.916 & slow \\
+Wealden & 2003.906 & 0.736 & 0.938 & slow \\
+Thanet & 2003.953 & 0.667 & 0.947 & slow \\
+Broadland & 2003.962 & 0.727 & 0.938 & slow \\
+West Northamptonshire & 2003.972 & 0.736 & 0.940 & slow \\
+North Kesteven & 2003.976 & 0.633 & 0.942 & slow \\
+Neath Port Talbot & 2003.981 & 0.767 & 0.933 & slow \\
+Rhondda Cynon Taf & 2003.991 & 0.769 & 0.925 & slow \\
+Enfield & 2003.992 & 0.588 & 0.954 & slow \\
+East Lindsey & 2004.010 & 0.577 & 0.952 & slow \\
+Teignbridge & 2004.015 & 0.681 & 0.941 & slow \\
+York & 2004.025 & 0.739 & 0.944 & slow \\
+Harlow & 2004.030 & 0.803 & 0.936 & slow \\
+Vale of Glamorgan & 2004.033 & 0.592 & 0.951 & slow \\
+Newark and Sherwood & 2004.034 & 0.668 & 0.946 & slow \\
+Merton & 2004.041 & 0.932 & 0.922 & slow \\
+Hambleton & 2004.065 & 0.636 & 0.953 & slow \\
+Bedford & 2004.068 & 0.852 & 0.933 & slow \\
+Stockport & 2004.075 & 0.757 & 0.942 & slow \\
+South Ribble & 2004.126 & 0.674 & 0.948 & slow \\
+Uttlesford & 2004.144 & 0.905 & 0.926 & slow \\
+Rossendale & 2004.160 & 0.976 & 0.923 & slow \\
+Oldham & 2004.162 & 0.801 & 0.938 & slow \\
+Sheffield & 2004.189 & 0.770 & 0.944 & slow \\
+Ribble Valley & 2004.204 & 0.718 & 0.945 & slow \\
+Maidstone & 2004.222 & 0.854 & 0.932 & slow \\
+Cheshire West and Chester & 2004.234 & 0.742 & 0.947 & slow \\
+Thurrock & 2004.241 & 0.841 & 0.924 & slow \\
+Telford and Wrekin & 2004.241 & 1.024 & 0.917 & slow \\
+Cardiff & 2004.246 & 0.823 & 0.937 & slow \\
+County Durham & 2004.250 & 0.626 & 0.953 & slow \\
+Plymouth & 2004.252 & 0.785 & 0.934 & slow \\
+Bridgend & 2004.256 & 0.547 & 0.956 & slow \\
+Rotherham & 2004.273 & 0.722 & 0.938 & slow \\
+Hyndburn & 2004.298 & 0.881 & 0.932 & slow \\
+Newcastle upon Tyne & 2004.305 & 0.906 & 0.932 & slow \\
+Wakefield & 2004.314 & 0.905 & 0.923 & slow \\
+Havering & 2004.345 & 0.822 & 0.938 & slow \\
+South Kesteven & 2004.353 & 1.026 & 0.915 & slow \\
+Canterbury & 2004.354 & 0.602 & 0.960 & slow \\
+Broxbourne & 2004.363 & 0.817 & 0.944 & slow \\
+Peterborough & 2004.382 & 0.937 & 0.932 & slow \\
+Redbridge & 2004.411 & 1.028 & 0.925 & slow \\
+Bury & 2004.431 & 0.735 & 0.953 & slow \\
+Harrogate & 2004.433 & 0.971 & 0.927 & slow \\
+Birmingham & 2004.435 & 1.182 & 0.912 & slow \\
+South Tyneside & 2004.446 & 0.636 & 0.948 & slow \\
+Chorley & 2004.462 & 1.093 & 0.913 & slow \\
+North East Derbyshire & 2004.483 & 0.703 & 0.953 & slow \\
+Lambeth & 2004.485 & 0.958 & 0.937 & slow \\
+Doncaster & 2004.496 & 0.672 & 0.953 & slow \\
+South Gloucestershire & 2004.499 & 0.881 & 0.937 & slow \\
+Caerphilly & 2004.499 & 0.819 & 0.938 & slow \\
+East Staffordshire & 2004.501 & 1.132 & 0.910 & slow \\
+North Tyneside & 2004.509 & 0.864 & 0.933 & slow \\
+South Holland & 2004.534 & 0.851 & 0.941 & slow \\
+Gateshead & 2004.535 & 0.869 & 0.934 & slow \\
+St.~Helens & 2004.542 & 1.097 & 0.923 & slow \\
+Flintshire & 2004.549 & 0.764 & 0.951 & slow \\
+Selby & 2004.549 & 0.716 & 0.947 & slow \\
+Belfast & 2004.556 & 1.332 & 0.906 & slow \\
+Sedgemoor & 2004.613 & 0.540 & 0.966 & slow \\
+Exeter & 2004.617 & 0.964 & 0.936 & slow \\
+South Derbyshire & 2004.629 & 0.934 & 0.938 & slow \\
+Derbyshire Dales & 2004.668 & 0.648 & 0.963 & slow \\
+East Riding of Yorkshire & 2004.683 & 0.728 & 0.955 & slow \\
+Basildon & 2004.691 & 1.067 & 0.925 & slow \\
+Barking and Dagenham & 2004.800 & 1.198 & 0.916 & slow \\
+Causeway Coast and Glens & 2004.834 & 1.024 & 0.928 & slow \\
+North Lanarkshire & 2004.839 & 0.878 & 0.945 & slow \\
+Chesterfield & 2004.887 & 1.173 & 0.927 & slow \\
+Newcastle-under-Lyme & 2004.933 & 1.172 & 0.925 & slow \\
+East Ayrshire & 2004.934 & 1.280 & 0.903 & slow \\
+Nuneaton and Bedworth & 2004.956 & 0.950 & 0.941 & slow \\
+Blaby & 2004.959 & 1.059 & 0.929 & slow \\
+Leeds & 2004.979 & 0.836 & 0.956 & slow \\
+Erewash & 2004.987 & 0.939 & 0.944 & slow \\
+Bassetlaw & 2004.998 & 0.982 & 0.944 & slow \\
+Hackney & 2005.026 & 1.157 & 0.934 & slow \\
+Barnsley & 2005.055 & 0.912 & 0.942 & slow \\
+Amber Valley & 2005.105 & 1.017 & 0.941 & slow \\
+Colchester & 2005.144 & 1.250 & 0.927 & slow \\
+Ashfield & 2005.176 & 1.073 & 0.942 & slow \\
+Wirral & 2005.184 & 1.084 & 0.939 & slow \\
+Knowsley & 2005.186 & 1.160 & 0.923 & slow \\
+Windsor and Maidenhead & 2005.195 & 1.746 & 0.902 & slow \\
+Bromsgrove & 2005.494 & 1.144 & 0.941 & slow \\
+Liverpool & 2005.526 & 1.276 & 0.936 & slow \\
+Staffordshire Moorlands & 2005.583 & 1.426 & 0.924 & slow \\
+Islington & 2005.970 & 1.223 & 0.955 & slow \\
+Warrington & 2006.015 & 1.387 & 0.942 & slow \\
+Newham & 2006.078 & 1.908 & 0.907 & slow \\
+Bolsover & 2006.188 & 1.143 & 0.957 & slow \\
+Lisburn and Castlereagh & 2006.193 & 1.970 & 0.903 & slow \\
+Manchester & 2006.449 & 1.434 & 0.950 & slow \\
+Sunderland & 2006.479 & 1.679 & 0.928 & slow \\
+South Staffordshire & 2006.490 & 1.837 & 0.922 & slow \\
+Rutland & 2006.529 & 0.986 & 0.967 & slow \\
+Cannock Chase & 2006.700 & 2.231 & 0.917 & slow \\
+Mansfield & 2006.889 & 1.579 & 0.946 & slow \\
+Ards and North Down & 2006.947 & 2.485 & 0.907 & slow \\
+Halton & 2007.069 & 0.932 & 0.967 & slow \\
+Burnley & 2007.372 & 2.286 & 0.919 & slow \\
+Middlesbrough & 2010.074 & 5.297 & 0.900 & slow \\
+Isles of Scilly & 2012.163 & 4.257 & 0.953 & slow \\
+\end{longtable}
+
 \renewcommand\refname{References}

### 2024-08-14 src/problem_years.Rmd
gini

@@
 
+### Gini
+
+```{r}
+
+# No need to add 0s like OAs as there aren't any LA with 0s
+
+gini.df <- data.frame()
+
+for (t in 1996:2012){
+    a <- df %>% filter(year==t)
+  b <- as.data.frame(gini(a$n))
+  b$year = t
+  gini.df <- rbind(gini.df, b)
+}
+
+path.out <- paste0(path, "/outputs/gini/gini_la.csv")
+gini.df %>% write_csv(path.out)
+```
+
 ## OA
@@
 ```
+
+### Gini
+
+```{r}
+
+# No need to add 0s like OAs as there aren't any LA with 0s
+
+gini.df <- data.frame()
+
+for (t in 1996:2012){
+    a <- df %>% filter(year==t)
+  b <- as.data.frame(gini(a$n))
+  b$year = t
+  gini.df <- rbind(gini.df, b)
+}
+
+path.out <- paste0(path, "/outputs/gini/gini_oa.csv")
+gini.df %>% write_csv(path.out)
+```

### 2024-08-14 src/problem_years_10.Rmd
gini

@@
 
+### Gini
+
+```{r}
+
+# No need to add 0s like OAs as there aren't any LA with 0s
+
+gini.df <- data.frame()
+
+for (t in 1996:2012){
+    a <- df %>% filter(year==t)
+  b <- as.data.frame(gini(a$n))
+  b$year = t
+  gini.df <- rbind(gini.df, b)
+}
+
+path.out <- paste0(path, "/outputs/gini/gini_la_10.csv")
+gini.df %>% write_csv(path.out)
+```
+
 ## OA
@@
 ```
+
+### Gini
+
+```{r}
+
+# No need to add 0s like OAs as there aren't any LA with 0s
+
+gini.df <- data.frame()
+
+for (t in 1996:2012){
+    a <- df %>% filter(year==t)
+  b <- as.data.frame(gini(a$n))
+  b$year = t
+  gini.df <- rbind(gini.df, b)
+}
+
+path.out <- paste0(path, "/outputs/gini/gini_oa_10.csv")
+gini.df %>% write_csv(path.out)
+```

### 2024-08-14 paper/submission1/tranos2023.qmd

@@
     keep-tex: true
+    text: |
+      \usepackage[demo]{graphicx} % "demo" option just for this example
+      \usepackage{subcaption}
     journal:
@@
 library(kableExtra)
+library(sf)
+library(cowplot)
 
@@
 
-# Introduction {#sec1}
+# Introduction {#sec-introduction}
 
@@
 regarding the diffusion of related *future* technologies. As per
-@leibowicz2016representing, historical studies agree on the fact that
-"technologies diffuse at different times, at different rates, and to
-different extents in different places, and can be significantly
-influenced by policies" [@victor1993]. @meade2021modelling highlight
+@leibowicz2016representing, historical studies agree that technologies diffuse 
+differently in terms of times, rates, and geographies and can be driven
+by related policies [@victor1993]. @meade2021modelling highlight
 that a variety of actors have a direct interest in gaining such
@@
 to two processes: (i) epidemic-like mechanisms, which are governed by
-distance, proximity and social interactions, and (ii) by economic
+distance, proximity and social interactions, and (ii) economic
 mechanisms as new innovations are adopted by users as they become more
@@
 @riddlesden2014broadband]. More details about the data and the data
-generation process can be found in Section [XX](#sec3).
+generation process can be found in @sec-datamethods.
+
+**S curve??**
 
@@
 
-# Literature review {#sec2}
+# Literature review {#sec-litreview}
 
@@
 Geographical diffusion is a synthesis of different processes. On the one
-hand, we can identify purely spatial or, in other words, contagious
-processes. Adjacency and, more broadly, distance are the key drivers of
+hand, a purely spatial or, in other words, contagious processes can be identified. 
+Adjacency and, more broadly, distance are the key drivers of
 diffusion. This perspective draws similarities with epidemics:
@@
 consequently, proximity and exposure [@hivner2003facilitating]. On the
-other hand, we can identify hierarchical processes. Instead of
+other hand, there is a hierarchical processes. Instead of
 horizontal distance-based diffusion mechanisms, the top-down hierarchy
-of urban systems drives diffusion. In reality, the synthesis of these
+of urban systems shapes technological diffusion. In reality, the synthesis of these
 two processes represents how new technologies diffuse over space and
@@
 seminal work on 'Diffusion of Innovations' [@rogers2010diffusion].
-Rogers being a sociologist, his work focused not on the diffusion of
+Rogers being a sociologist focused not on the diffusion of
 innovations over space and time, but instead on the adoption of new
@@
 neighbourhood effect and contaminate nearby locations. These are the three
-mechanisms that the empirical analysis in Section [4]{#sec4} investigates.
+mechanisms that the empirical analysis in @sec-results investigates.
 
@@
 
-# Materials and Methods {#sec3}
+# Data and Methods {#sec-datamethods}
 
 **ADD FAMILY OF WEB TECHNOLOGIES**
+
+<!-- main metric -->
 
@@
 Output Areas (OA). The former is an administrative unit and there are 
-\\textcolor{red}c. 380\\textcolor{black} such units in the UK. The latter is a census-based
-geographical unit, which is very small as there are c. 230,000 of them in the UK.
-This methodological choice will allow the mapping of the diffusion 
-of web technologies and the assessment of the diffusion mechanisms 
-at these two very different spatial scales.
+c. 374 such units in the UK. The latter is a census-based geographical unit, which 
+is very small as there are c. 230,000 of them in the UK. This methodological choice 
+will allow the mapping of the diffusion of web technologies and the assessment of 
+the diffusion mechanisms at these two very different spatial scales.
 
-The population of websites is calculated using data from the 
+<!-- web data -->
+
+The counts of websites at these scales are calculated using data from the 
 Internet Archive [^2] and, specifically, the JISC UK Web Domain Dataset [@ukwebarchive]. 
@@
 support online transactions and share opinions [@THELWALL2000441; @blazquez2018big]. 
-Although a UK company can adopt a generic TLD such as.com and such cases escape 
+Although a UK company can adopt a generic TLD such as .com and these cases escape 
 the data used here, such omissions should not affect our results given the
@@
 
-The text from these webpages was scanned using regular expressions (regex) to
+The text from these webpages was scanned using a regular expression (regex) to
 identify strings of text which resemble UK postcodes and one fifth of them included
@@
 
-Before that, the data 
-cleaning process included a data aggregation step, through which
-the webpages were aggregated to the relevant websites. 
-Based on the following example, all three webpages listed below are part of the 
+The data cleaning process included an aggregation step, through which
+the archived webpages were aggregated to the parent websites. This website 
+reconstruction allows the creation of a *website* instead of a *webpage* density 
+metric. Websites, which are hierarchical structures, tend to represent specific 
+orgnanisations or entities, and, arguably, are more meaningful observational units 
+than webpages, which ignore the upstream dependencies. 
+Based on the following example, all three webpages are part of the 
 same website ([http://www.website.co.uk](http://www.website.co.uk)) and at the end 
 only websites and not the nested webpages were considered as otherwise the metric 
-would have been biased towards large websites.
+would have been biased towards large, place specific, websites.
 
-  - [http://www.website.co.uk/webpage_a](http://www.website.co.uk/webpage_a) B152TT
+  - [http://www.website.co.uk/webpage_a](http://www.website.co.uk/webpage_a) B15 2TT
 
-  - [http://www.website.co.uk/webpage_b](http://www.website.co.uk/webpage_b) BS81TH
+  - [http://www.website.co.uk/webpage_b](http://www.website.co.uk/webpage_b) BS8 1TH
 
-  - [http://www.website.co.uk/webpage_c](http://www.website.co.uk/webpage_c) B152TT
+  - [http://www.website.co.uk/webpage_c](http://www.website.co.uk/webpage_c) B15 2TT
 
@@
 which has been used elsewhere [@tranosuk; @shoreditch] may lead to websites with 
-multiple postcodes. As per the above example, ([http://www.website.co.uk](http://www.website.co.uk))
-includes two unique postcodes: B152TT and BS81SS. 
+multiple postcodes. As per the above example, [www.website.co.uk](http://www.website.co.uk)
+includes two unique postcodes: B15 2TT and BS8 1SS. 
 The distribution of postcodes for 2000 is presented in Table \ref{f2000}, which 
-clearly illustrates the wide range. On one hand, we
-have websites anchored to a unique location (72% of all
-the reconstructed websites), which may represent
-a small company with a single trading location. On the other hand,
-we have websites with thousands of different postcodes.
-Considering the timeperiod of the analysis, such cases can represent directories
-which used to be popular in the pre-search engines early times of internet [@tranosuk].[^3]
+clearly illustrates the wide range. At the left end of the distribution, there are 
+websites anchored to a unique location (72% of all the reconstructed websites in
+2000), which may represent a small company with a single trading location. At the 
+right end, we have websites with thousands of different postcodes.
+Considering the time period of the analysis, such cases can represent directories
+which used to be popular in the pre-search engines early times of the commercial 
+internet [@tranosuk].[^3]
 
-The analysis presented here is based on two subsets of these data.
-Firstly, the analysis is based on websites which only contain only one unique postcode
-and, therefore, the geolocation process might suffer from less noise. 
-As a robustness check, the analysis is replicated the analysis for an extended subset 
-of websites, which include up to ten unique postcodes. These websites are geolocated 
-by equally attaching them to multiple locations. This extended sample includes 
-94 percent of all the archived websites in 2000.
+The analysis presented here is based on two subsets of these data. Firstly, on websites, 
+which only contain only one unique postcode and, therefore, the geolocation process 
+does not suffer from noise. As a robustness check, the analysis is replicated 
+for an extended subset of websites, which include up to ten unique postcodes. These 
+websites are geolocated by equally attaching them to multiple locations. This extended 
+sample includes 94 percent of all the archived websites in 2000.
 
@@
 
-To create the web density metric, the yearly website counts at the LAD level are
-standardised by the number of firms in LAD to avoid biases associated with LADs 
+Another data cleaning step dealt with some extreme outliers. Figure \ref{pc1} plots
+the website counts for the top 1000 postcodes. Some obvious outliers can be observed 
+for the 2004-2006 period, which can be attributed to a link farm [@BL2013links].  
+The website counts for these six postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1 6GL, 
+M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1000 websites pointing to 
+them, were replaced with predicted values based on a simple panel regression model 
+with postcode fixed effects and yearly dummy variables. As the @sec-results illustrates,
+this led to a small increase of the model predictive capability. To put the magnitude 
+of the data imputation into perspective, this process affected 6 out of the 557,808
+postcodes present in the data.
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{pc1}Yearly website counts per postcode (top 1000)", out.height="60%"}
+path.image <- paste0(path, "/outputs/pc_year_1.png")
+knitr::include_graphics(path.image)
+```
+
+To create the website density metric, the yearly website counts at the LAD level are
+standardised by the number of firms in LAD to avoid biases associated with LAD 
 hosting a large number of firms. Given that there is no such statistic for the OA,
-the actual OA level counts are used. However, because of the the rather
-consistent spatial definition of OA (they host 40-250 households),[^4] 
-website counts in OA are interpreted as a density metric too.
+the actual OA level counts are used. However, because of the the consistent spatial 
+definition of OA (they host 40-250 households),[^4] website counts in OA are 
+interpreted as a density metric too.
 
@@
 
-This website metric...
+<!-- system-level analysis -->
 
-@wilson201281 growth function description in p. 86
+These website density metrics are used for the three different steps of the analysis. 
+Firstly, a system-level analysis explores how the diffusion of web technologies in
+the UK fits with the well-established S-curve. To do so, the following logistic 
+function (Equation \ref{eq:s}) is estimated for the whole of the UK and for each 
+LAD separately:
 
-from R: Asym/(1+exp((xmid-input)/scal)) using the terms from Wilson
-k/(1 + e\^(t_0 - t)/scal)
+\begin{align}
+y = k /(1 + e^{-(t-t_{o})})\label{eq:s}
+\end{align}
 
-Wilson: k(1 + e\^-b(t - t_0)) So, scal = - 1/b
+\noindent $k$ is the asymptote or, in other words, the saturation level, $b$ the overall 
+growth rate, and $t_{0}$ the *inflection point*
+of maximum growth at $k/2$, where the logistic function is symmetrical [@wilson201281]. 
+We compare  the LAD $t_0$ of each LAD against the $t_0$ of the UK to delineate 
+whether a LAD reached that point faster or slower than the country average. Importantly, 
+an accuracy criterion was imposed and only S-curves with $R^2 > 0.9$ were included 
+in the analysis. To estimate Equation \ref{eq:s} a self-starting logistic growth 
+model was employed using the `nls` and `SSlogis` functions in `R`.
 
-The curve is symmetric\* around t_0
+<!-- **minimum R2 = 95%** in @wilson201281, see also @grubler1990rise -->
+<!-- The literature usually uses the saturation level as the asymptote. I am -->
+<!-- using the total number of websites as we cannot compute a rate. -->
 
-**minimum R2 = 95%**, see also @grubler1990rise
+The system-level analysis also focuses on the volatility of the web adoption to 
+depict places with high concentration of early adopters and, equally, latecomers. 
+To do so, the change of the ranking of the UK LAD over time is plotted and discussed. 
+Both the S-curve and the volatility analysis focus only on LAD as the very large 
+number of OA would have made such analysis diffucult to visualise and interpret.
 
-The literature usually uses the saturation level as the asymptote. I am
-using the total number of websites as we cannot compute a rate.
+<!-- ESDA -->
 
-Moran's I as @ding2010modeling **TODO** for t_0, diffusion speed
+Next, exploratory spatial data analysis depicts whether the two main 
+drivers of spatial diffusion -- namely neighbourhood and hierarchy -- underpin the
+diffusion of web technologies in the UK. To capture the former and following 
+@ding2010modeling the Moran's I and the Local Index of Spatial Association (LISA) 
+are estimated for website density. To address the hierarchy mechanism, the Gini 
+coefficient -- a well established metric of inequality -- is calculated. All the 
+above are computed and plotted longitudinally both for the LAD and the OA.
 
-# Results {#sec4}
+<!-- RF -->
 
-1.  S-shaped diffusion curves: S for LADs per firm. UK,
+Lastly, a modelling framework is developed to test the above diffusion mechanisms.
+The overarching aim is to build a model that can test the relationship between the
+website density and these mechanisms:
+
+\begin{align}
+Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S-curve_{t}\label{eq:rf.generic}
+\end{align}
+
+To estimate Equation \ref{eq:rf.generic}, Random Forest (RF) models were built. 
+This is a popular machine learning algorith for both regression and classification 
+problems [@biau2012analysis]. It was introduced by @breiman2001random and has  
+gained popularity, becoming a go-to data science tool. RF can effectively 
+handle skewed distributions and outliers, model non-linear relationships, require 
+minimal hyperparameter tuning, exhibit low sensitivity to these parameters, and 
+have relatively short training times [@Caruana2008; @liaw2002classification; @yan2020using]. 
+These attributes match well with the website density data characteristics including 
+skewness especially for OA. Also, the large data size (c. 230k data points for 
+each of the 17 years) call for fast training times. 
+Importantly, RF predictions tend to be more accurate than those from single regression 
+trees and outperform Ordinary Least Squares in out-of-sample predictions, even with 
+moderate-sized training data and a small number of predictors [@mullainathan2017machine; 
+@athey2019machine; @sulaiman2011intelligent; @pourebrahim2019trip; @biau2012analysis].
+
+RF is a tree-based ensemble learning algorithm [@breiman2001random]. It begins by 
+generating random samples of the training data, which are then used to grow 
+regression trees to predict the dependent variable. Data points and predictors are 
+randomly sampled for the different trees. The trees are trained in parallel using their 
+own bootstrapped samples of the training data. A crucial feature of RF is their 
+ability not to overfit, meaning they can generalize well to unseen test data. While 
+each tree may overfit individually, the ensemble of trees does not because the errors 
+of individual trees are averaged, reducing the overall variance and preventing overfitting 
+[@last2002improving]. For regression problems, RF predictions are made by averaging 
+the predictions of all decision trees.
+
+RF have been widely employed to address regression research problems. @pourebrahim2019trip 
+combined a spatial interaction modeling framework with ML algorithms including RF
+to predict commuting flows in New York City. @sinha2019assessing advocated for adopting
+spatial ensemble learning approaches, such as RF, to model spatial data with high
+autocorrelation and heterogeneity. @creditspatial predicted employment density in 
+Los Angeles using spatially explicit RF. @guns2014recommending used RF to build a 
+recommendation system for research collaborations. @ren2019predicting trained RF 
+to predict the socio-economic status of cities using various online and mobility 
+predictors. @tranos2023using utilised hyperlinks data and RF to make out-of-sample
+predictions of interregional trade. @zhou2023geography employed such a framework
+to assess whether key predictors of obesity differ across English cities.
+
+# Results {#sec-results}
+
+1.  S-shaped diffusion curves: S for LAD per firm. UK,
     fast/slow/examples
@@
 
--   ideal: (i) train RF for all years and all (1) LADs and (2) OAs with
+-   ideal: (i) train RF for all years and all (1) LAD and (2) OA with
     CAST and report metrics. (ii) train for all years and all but one
-    region for (1) LADs and (2) OAs to predict to the holdout region.
+    region for (1) LAD and (2) OA to predict to the holdout region.
     Reports predictions as region similarities.
 
-### RF results
+<!-- S-shaped -->
+
+Figure \ref{s_uk} plots the S curve for the cumulative adoption of website technologies
+in the UK during the 1996-2012 period. It demonstrates a pattern well aligned with
+previous studies discussed in @sec-litreview. The vertical line for year 
+2003 illustrates the point where the modeled cumulative adoption was equal to 50% 
+of the maximum. This $t_0$ *inflection* point signals the maximum adoption speed 
+and is used here to determine whether the UK LAD reached that inflection point 
+earlier or later than the UK average. Specifically, the S curve and the inflection 
+point are estimated for every LAD individually and then compared with the UK average. 
+LAD that reached their inflection point earlier than the country average are labelled 
+as *fast* and the rest as *slow*. Figure \ref{s_map} maps this pattern and the picture 
+is not entirely clear. On the one hand, there are some expected examples of LAD
+with relevant industrial backgrounds delineated as fast: the City of London, a 
+world-renowned cluster of finance industries [@cook2007role], and Reading, a town 
+with high-tech service industries in proximity to London and its main airport, Heathrow 
+[@england2005polynet]. On the other hand, LAD which were expected to appear as fast
+-- e.g. Hackney in central London and Bristol, a well-established creative cluster 
+[@oatley1999cultural; @bassett2002cultural] -- were delineated as slow. Nevertheless, 
+out of the 10 fastest LAD, eight were located in Greater London and one in 
+Cambridge (see **Appendix**), but in overall there is a spatially heterogenous and
+not easy to explain spatial pattern.
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk}Grwoth curve, UK"}
+path.image <- paste0(path, "/outputs/s/s_uk_per_firm.png")
+knitr::include_graphics(path.image)
+```
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map}LAD adoption rates"}
+path.image <- paste0(path, "/outputs/s/speed_map.png")
+knitr::include_graphics(path.image)
+```
+
+<!-- rankings -->
+The next step is to assess the stability and volatility of the LAD in terms of 
+their adoption of web technologies. As we know from the literature [@risk_perceptions],
+different agents have different perceptions about and levels of acceptance of the 
+risks and the potential economic returns associated with the adoption of new technologies 
+-- see for instance the seminal work of @venkatesh2000theoretical. To reveal such 
+aggregated patterns, Figures \ref{rank_stable} and \ref{rank_unstable} plot the ranking
+of UK LAD based on website density. To decrease noise, the average ranking 
+of 1996-1998 and 2010-2012 is plotted instead of the individual years. 
+While there are quite a few obvious cases of LAD that maintained their position 
+between the beginning and the end of the study period either at the top or at the 
+bottom of the hierarchy (Figure \ref{rank_stable}), there are also 
+quite a few LAD that changed drastically their position. Some of these LAD enjoyed
+a process that at the first instance looks like leapfrogging since 
+they managed to jump at the top of the hierarchy despite their slow start (top of 
+Figure \ref{unstable}). There is extensive literature regarding the potential benefits of
+technological leapfrogging. The underpinning argument is that latecomers can adopt
+and benefit from new technologies that have been developed elsewhere without incurring 
+the hefty initial R&D costs [@teece2008firm]. Although the leapfrogging literature
+does not pay much attention to cities and regions [@yu2018sustainability], previous 
+research highlighted the long term and sustained productivity benefit of the early
+adopters of web technologies in the UK [@tranosuk]. So, although the LAD system 
+is volatile, it is not clear whether the LAD with high concentration of late adopters 
+will gain any latecomer benefits in a way similar to countries experiencing such
+technological leapfrogging.
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank_stable}Dynamics of wed diffusion: stability of LAD"}
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only05_av.png")
+knitr::include_graphics(path.image)
+```
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank_unstable}Dynamics of wed diffusion: volatility LAD"}
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only95_av.png")
+knitr::include_graphics(path.image)
+```
+
+<!-- Neighbourhood effect -->
+
+Then, figures \ref{morani} and \ref{lisa} offer a first insight into whether a neighbourhood 
+effect underpins the diffusion of web technologies in the UK as they plot the Moran's I
+and the LISA maps of website density respectively for LAD and OA. Starting from 
+the former, spatial autocorrelation was higher in the beginning and then around 
+year 2000 dropped slightly and stabled at 0.2. This reflects the early concentration 
+of high website density around London, which over time diffused as high-high clusters
+can be seen in other parts of the country away from London as seen in Figure \ref{lisa}. An 
+almost reverse pattern can be observed for OA. At the beginning of the study period
+Moran's I was around 0.1 and it plateaued after 2000-2001 around 0.2. Because of the 
+very small size of OA, at the early stages of the diffusion of web technologies 
+their adoption was spatially scattered. This is reflected in the lack of any significant
+clusters in 1996. Eventually, as the adoption rate increased, more such clusters 
+of high website density were formed and this is reflected both in the Moran's I 
+and the LISA maps in Figures \ref{morani} and \ref{lisa}. All in all, the exploratory 
+spatial data analysis advocates towards an underpinning neighbourhood effect and 
+the different scales of analysis illustrate how it evolved differently over time.
+
+```{r, morani, echo=FALSE, message=FALSE, fig.cap="\\label{morani}Website density Moran's I"}
+
+path.in.la <- paste0(path, "/outputs/lisa/corrected/LA/morani_la.csv")
+path.in.oa <- paste0(path, "/outputs/lisa/corrected/oa/morani_oa.csv")
+
+la.m <- read_csv(path.in.la) %>% 
+  filter(p < 0.001) %>% # all values were significant
+  mutate(spatial.unit = 'LAD')
+oa.m <- read_csv(path.in.oa) %>% 
+  filter(p < 0.001) %>% # all values were significant
+  mutate(spatial.unit = 'OA')
+
+bind_rows(la.m, oa.m) %>% 
+  ggplot(aes(x=year, y=morani, fill= spatial.unit)) +
+  geom_bar(stat = 'identity', position = 'dodge') +
+  ylab('Moran\'s I') +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_fill_manual('Spatial unit', 
+                    values = c("LAD" = "grey80",
+                                "OA"="grey48"),
+                     limits = force) +
+  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) 
+```
+
+Figure \ref{lisa}
+
+```{r, echo=FALSE, message=FALSE, fig.cap="\\label{lisa}\\centering Website density LISA maps", out.height="80%"}
+
+path.image1 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la1996.png")
+path.image2 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la2000.png")
+path.image3 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la2012.png")
+path.image4 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa1996.png")
+path.image5 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa2000.png")
+path.image6 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa2012.png")
+path.image7 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la_legend.png")
+
+p1 <- ggdraw() + draw_image(path.image1)  
+p2 <- ggdraw() + draw_image(path.image2)  
+p3 <- ggdraw() + draw_image(path.image3)  
+p4 <- ggdraw() + draw_image(path.image4) 
+p5 <- ggdraw() + draw_image(path.image5)
+p6 <- ggdraw() + draw_image(path.image6)  
+p7 <- ggdraw() + draw_image(path.image7)
+
+par(mar=rep(0,4)) # no margins
+
+library(patchwork)
+# (p1 | p4)/(p2 | p5)/(p3 | p6)/ p7 +
+#   #plot_layout(widths = 1)
+#     plot_layout(widths = c(1, -0.5))
+
+# p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 / p7 +
+#   plot_layout(ncol = 3, widths = c(1, -0.85, 1), heights = c(1,1,1,2))
+
+p <- p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 +
+  plot_layout(ncol = 3, widths = c(1, -0.8, 1), heights = c(1,1,1))
+
+# p + inset_element(p7, left = 0.45, bottom = -0.5, right = 0.6, top = 0.1)
+p + inset_element(p7, left = 0.29, bottom = 0.8, right = 0.49, top = 1.2)
+
+```
+
+To illustrate whether a hierarchical process underpins the diffusion of web technologies,
+the Gini coefficient is calculated yearly both for LAD and OA. As a metric of 
+inequality, the Gini coefficient demonstrates whether website density is concentrated
+in a small number of LAD/OA, or whether it is more equally spread across the country.
+Both scales of analysis in Figure \ref{gini} illustrate the same picture. At the 
+beginning of the commercial internet website density was extremely unequal, or in
+other words, only a few places had websites associated with them. Inequality dropped 
+and plateaued after 2000 for both scales illustrating that at the first stages of 
+the commercial internet, website density was concentrated in a limited number of 
+places. This is an indication of a hierarchical diffusion mechanism that led over 
+time to a more equal spread. Interestingly, the year 2000 is again a period of change
+for this diffusion mechanisms as it was for the neighbourhood process. 
+There is a substantial difference between the Gini coefficient magnitude for LAD
+and OA, but this is well expected as the very small size of OA equates to a lot 
+polygons without any websites pointing to them -- for instance residential OA. 
+LAD/OA
+
+```{r, gini, echo=FALSE, message=FALSE, fig.cap="\\label{gini}Website density Gini coefficient"}
+
+path.in.la <- paste0(path, "/outputs/lisa/gini_la.csv")
+path.in.oa <- paste0(path, "/outputs/lisa/gini_oa.csv")
+
+# path.in <- paste0(path, "/outputs/lisa/gini_la.csv")
+# path.in <- paste0(path, "/outputs/lisa/gini_oa.csv")
+
+la.gini <- read_csv(path.in.la) %>% 
+  mutate(spatial.unit = 'LAD')
+oa.gini <- read_csv(path.in.oa) %>% 
+  mutate(spatial.unit = 'OA')
+
+bind_rows(la.gini, oa.gini) %>%
+  rename(Gini = 'gini(a$n)') %>% 
+  ggplot(aes(x=year, y=Gini, fill= spatial.unit)) +
+  geom_bar(stat = 'identity', position = 'dodge') +
+  ylab('Gini') +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_fill_manual('Spatial unit', 
+                    values = c("LAD" = "grey80",
+                                "OA"="grey48"),
+                     limits = force) +
+  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) 
+```
+
+<!-- RF results -->
 
 The next section incorporates the above discussed spatial processes of the
-diffusion of web technologies into the same modelling framework. The aim
+diffusion of web technologies into a modelling framework. The aim
 is to use variables depicting these spatial processes in order to
@@
 and across different scales. Specifically, four different models are
-estimated. Firstly, all the data points at the OA and LAD are utilised
-in order to build RF models and assess their capacity to predict the
+estimated. Firstly, all the data points for the OA and LAD are utilised
+in order to build two respective RF models and assess their capacity to predict the
 adoption of web technologies. These two models will reveal the
-predictive capacity of the spatial processes and also allow to see how
+predictive capacity of the diffusion mechanisms and also allow to see how
 the importance of such variables changes across scales. The next two
@@
 the LAD from one of the twelve UK regions are held out and then the trained
-model is used to predict web adoption in the OA or the LAD of the
+model is used to predict website density in the OA or the LAD of the
 held-out region. This process takes place recursively for all UK
@@
 
-The models need to include variables that capture the three processes
-that the relevant literature and the descriptive analysis presented in
-the Section **XX** highlighted. Namely, the models capture: (i) a
-hierarchy effect with diffusion running from main centres to secondary
-ones, (ii) a neighborhood effect according to which diffusion first hits
-nearby locations, and (iii) the rather canonical pattern of 
-diffusion over time as reflected in the S-shaped pattern in the
-cumulative level of adoption.
+The models include variables that capture the three processes
+that the relevant literature and the descriptive analysis presented above highlighted. 
+Namely, the models capture: (i) a hierarchy effect with diffusion running from main 
+centres to secondary ones, (ii) a neighborhood effect according to which diffusion 
+first hits nearby locations, and (iii) the rather canonical pattern of diffusion 
+over time as reflected in the S-shaped pattern in the cumulative level of adoption.
 
 To capture the hierarchy effect the models include as predictors a one
-year lag of website density in London -- the largest city in the UK, a
+year lag of website density in London, the largest city in the UK, a
 one year lag of the website density in the nearest city and the same for
@@
 adoption in a spatial unit depends on the level of the adoption in
-places further up in the urban hierarchy the previous year. To depict
-the neighbourhood effect, the web density of the neighbouring spatial
-units in the previous year is employed. Again, the underpinning
-rationale is that the level of web adoption within a spatial unit
-depends on the level of web adoption in the neighbouring spatial units
-the year before. This is the 'hitting nearby locations first' argument.
-Therefore, the spatial and temporal lag of the website density in LAD
-and OA is calculated. Lastly, the time effect which is reflected on the
-S curve for the cumulative adoption is captured by trend variable.
+places further up in the urban hierarchy the previous year. The inclusion of the
+distance variables incorporates spatial structure into the hierarchy argument.
+To depict the neighbourhood effect, the website density of the neighbouring spatial
+units in the previous year is employed. Again, the underpinning rationale is that 
+the level of web adoption within a spatial unit depends on the level of web adoption 
+in the neighbouring spatial units the year before. This represents the 'hitting 
+nearby locations first' argument. Therefore, the spatial and temporal lag of the 
+website density in LAD and OA is calculated. Lastly, the time effect which is reflected 
+on the S-curve for the cumulative adoption is captured by a time trend variable.
 Hence, all four model will follow the following generic form 
@@
 To assess the predictive capability of the model, three broadly utilised
-metrics are employed: the coefficient of determination (R squared), mean
+metrics are employed: the coefficient of determination ($R^2$), mean
 absolute error (MAE) and root mean square error (RMSE):
@@
 
-Table \ref{table.metrics.all} presents the model performance for the
-first set on models, for which all data points are employed for training
+@tbl-model-metrics presents the model performance for the
+first set of models, for which all data points are employed for training
 and testing via cross validation. The first one is trained and tested on 374 LAD
@@
 the model error is $(\frac{1}{20}, \frac{1}{30})$ of a website per firm. At 
-the OA the RSquared drops down to 21%. Considering its granularity, this is still 
+the OA the $R^2$ drops down to 21%. Considering its granularity, this is still 
 a remarkable performance. To contextualise it, the model results in a MAE of one 
@@
 and a significant part of them is not linked to any websites. In 1997 only 1% of 
-the UK OA were associated with a website. This should not come as a surprise as 
+the UK OA were associated with at least one website. This should not come as a surprise as 
 this was the very beginning of the commercial internet and any activities with a 
 digital footprint were concentrated in a handful of areas. This was clearly illustrated
-in Section **ADD**. At the end of the study period almost half of the UK OA were
+in Figure \ref{lisa}. At the end of the study period almost half of the UK OA were
 not associated with a website. Again, given the granularity of the data this should
@@
 
-|                   | RMSE  | RSquared | MAE   |
-|-------------------|-------|----------|-------|
+|                   | RMSE  | $R^{2}$   | MAE   |
+|-------------------|-------|-----------|-------|
 | Local Authorities | 0.032 | 0.810     | 0.019 |
@@
 
-: Model metrics {#table.metrics.all}
+: Model metrics {#tbl-model-metrics}
 
@@
 predictors. When the focus is on the LAD, the website density in the nearest city,
-in London and in the neighbouring LADs the year before are the most important predictors.
+in London and in the neighbouring LAD the year before are the most important predictors.
 They are followed by the yearly trend, while the spatial configuration as reflected
@@
 
+Figure \ref{var.imp} to replace Figures \ref{var.imp.LAD} and \ref{var.imp.OA} once
+I have the variable importance for LAD.
+
+```{r varimp, eval = T, echo=FALSE, message = F, fig.cap="\\label{var.imp}Variable importance"}
+path.importance.lad <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.csv")
+path.importance.oa <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.csv")
+
+importance.oa <- read_csv(path.importance.oa)
+importance.lad <- read_csv(path.importance.oa) %>% 
+  mutate(importance = ifelse(variable == "year", 10, 
+                             ifelse(variable == "distance to London", 75, importance)))
+
+importance.oa %>% 
+  rename(OA = importance) %>% 
+  left_join(importance.lad %>% rename(LAD = importance), by = "variable") %>%
+  pivot_longer(!variable, names_to = "spatial.unit", values_to = "importance") %>% 
+  # reorder with fct_reorder
+  ggplot(aes(x=importance, y=forcats::fct_reorder(variable, importance), 
+         fill = spatial.unit)) +
+  geom_col(position = 'dodge') +
+  ylab('') +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text(vjust = 0.5, hjust=1)) +
+  scale_fill_manual('Spatial unit', 
+                    values = c("LAD" = "grey80",
+                                "OA"="grey48"),
+                     limits = force) +
+  scale_x_continuous('Variable importance')
+
+```
+
 Table \ref{table.regions} presents the results of the recursive hold out models,
@@
 technologies within its LAD and OA. It needs to be highlighted though that the 
-difference between the RSquared of LAD and OA is more that an order of magnitude 
+difference between the $R^2$ of LAD and OA is more that an order of magnitude 
 signaling how difficult is to predict diffusion at such a small spatial scale.
@@
   arrange(`Rank LAD`) %>% 
-  kable(caption= "Regional differences\\label{table.regions}", digits=3)
+  kable(caption= "Regional differences\\label{table.regions}", 
+        digits=3,
+        col.names = c("Region", "$R^2$ LAD", "Rank LAD", "$R^2$ OA", "Rank OA")) 
 ```
 
-# Discussion and conclusions
+# Discussion and conclusions {#sec-conclusions}
 
@@
 
+<!-- reset the figure, table and section numbering -->
+\setcounter{section}{0}
+\renewcommand{\thesection}{\Alph{section}}
+
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{A\arabic{figure}}
+
+# Appendices {.appendix}
+
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
+
+## $t_0$ estunates for all LAD
+
+```{r, echo=FALSE, message = F, error=FALSE}
+
+path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
+la <- st_read(path.geo, quiet = T)
+# source: https://geoportal.statistics.gov.uk/
+
+# spatial transformations
+la <- st_transform(la, 4326)
+
+s.la.path <- paste0(path, "/data/temp/s_la_per_firm.csv")
+read.csv(s.la.path) %>% 
+  left_join(la %>% as_tibble() %>% select(LAD21CD, LAD21NM),
+            by = c("ladcd" = "LAD21CD")) %>% 
+  arrange(estimate) %>% 
+  select(LAD21NM, estimate, std.error, r2, fast) %>% 
+  rename(LAD = LAD21NM,
+         "t_0 estimate" = "estimate",
+         "R-squared" = "r2",
+         "Std. error" = "std.error",
+         "Diffusion speed" = "fast") %>% 
+  kable(caption= "S-curve estiamtes for LAD\\label{table.s.lads}", 
+        digits=3,
+        align = c("l", "c", "c", "c", "c"),
+        col.names = c("LAD", "$t_0$ estimate", "Std. error", "$R^2$", "Diffusion speed"))
+```
+
 # References {#references .unnumbered}

### 2024-08-15 src/problem_years.Rmd
extended correction

@@
   ungroup() %>% 
-  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2003 & year <2007, NA, n))
+  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2001 & year <2007, NA, n))
 

### 2024-08-15 src/problem_years_10.Rmd
extended correction

@@
   ylab("N. of websites per postcode") +
-  scale_x_continuous(breaks=seq(1996, 2012, 1)) +
+  scale_x_continuous(NULL, breaks=seq(1996, 2012, 1)) +
   scale_y_continuous(labels=scales::comma) +
   theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_text_repel(aes(x = year, y = n, label=outlier), cex = 3)
+  theme(plot.title = element_text(hjust = 0.5),
+        panel.grid.minor.x = element_blank()) 
+  #geom_text_repel(aes(x = year, y = n, label=outlier), cex = 3)
 
@@
   # The next line is different than the same for  n = 1: problem.pcs$pc
-  mutate(n = ifelse(pc %in% as.character(problem.pcs) & year > 2003 & year <2007, NA, n))
+  mutate(n = ifelse(pc %in% as.character(problem.pcs) & year > 2001 & year <2007, NA, n))
 

### 2024-08-15 src/s_la_per_firm_corrected.Rmd
extended correction

@@
 
-out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only95_av.png")
+out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only05_av.png")
 
@@
 
-  mutate(alpha = ifelse(abs(diff_1st) > 253, .8, 0.1)) %>% 
-  #mutate(alpha = ifelse(abs(diff_1st) <= 3.75, .8, 0.1)) %>%
+  #mutate(alpha = ifelse(abs(diff_1st) > 253, .8, 0.1)) %>% 
+  mutate(alpha = ifelse(abs(diff_1st) <= 3.75, .8, 0.1)) %>%
 

### 2024-08-15 src/s_la_per_firm_corrected_10.Rmd
extended correction

@@
 
-out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only95_av_10.png")
+out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only05_av_10.png")
 
@@
 
-  mutate(alpha = ifelse(abs(diff_1st) > 194.5, .8, 0.1)) %>% 
-  #mutate(alpha = ifelse(abs(diff_1st) <= 5, .8, 0.1)) %>%
+  #mutate(alpha = ifelse(abs(diff_1st) > 194.5, .8, 0.1)) %>% 
+  mutate(alpha = ifelse(abs(diff_1st) <= 5, .8, 0.1)) %>%
 

### 2024-08-16 src/la_RF_corrected.Rmd
all but OA

@@
 #   mtry  splitrule   RMSE        Rsquared   MAE       
-#   2     variance    0.01366816  0.8895425  0.009767759
-#   2     extratrees  0.01350182  0.8900993  0.009603882
-#   4     variance    0.01408821  0.8841051  0.010134321
-#   4     extratrees  0.01354075  0.8914524  0.009664646
-#   6     variance    0.01445541  0.8772260  0.010417962
-#   6     extratrees  0.01369444  0.8897055  0.009772824
+#   2     variance    0.02811777  0.8509273  0.01912432
+#   2     extratrees  0.02782836  0.8508316  0.01901921
+#   4     variance    0.02867344  0.8466191  0.01970102
+#   4     extratrees  0.02795250  0.8509000  0.01911956
+#   6     variance    0.02974784  0.8364575  0.02066007
+#   6     extratrees  0.02830277  0.8482189  0.01944133
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule = extratrees
-#  and min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule = extratrees and min.node.size = 5.
 
@@
 ggsave(path.out)
+
+path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.csv")
+varimp_mars$importance %>% 
+  rownames_to_column() %>%
+  rename(variable = rowname,
+         importance = Overall) %>% 
+  write_csv(path.out)
 ```
@@
 #   mtry  splitrule   RMSE       Rsquared   MAE      
-#   2     variance    0.2182933  0.5361604  0.1560099
-#   2     extratrees  0.2099702  0.5660477  0.1501194
-#   4     variance    0.2168118  0.5238461  0.1550630
-#   4     extratrees  0.2074722  0.5590058  0.1489474
-#   6     variance    0.2268891  0.5014882  0.1631466
-#   6     extratrees  0.2082266  0.5503986  0.1490911
+#   2     variance    0.2658416  0.5284505  0.1934243
+#   2     extratrees  0.2612133  0.5292689  0.1900297
+#   4     variance    0.2621564  0.5194092  0.1912807
+#   4     extratrees  0.2611575  0.5183202  0.1910029
+#   6     variance    0.2609983  0.5140347  0.1891971
+#   6     extratrees  0.2607463  0.5124544  0.1907650
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 4, splitrule = extratrees
-#  and min.node.size = 5.
+# The final values used for the model were mtry = 6, splitrule = extratrees and min.node.size = 5.
 

### 2024-08-16 src/la_RF_corrected_10.Rmd
all but OA

@@
 #   mtry  splitrule   RMSE       Rsquared   MAE      
-#   2     variance    0.1881989  0.7750790  0.1223893
-#   2     extratrees  0.1944376  0.7543149  0.1274344
-#   4     variance    0.1906884  0.7730657  0.1256452
-#   4     extratrees  0.1930998  0.7596818  0.1276155
-#   6     variance    0.1937399  0.7677967  0.1282843
-#   6     extratrees  0.1927437  0.7608415  0.1277447
+#   2     variance    0.1877289  0.7762104  0.1222690
+#   2     extratrees  0.1945882  0.7537303  0.1274455
+#   4     variance    0.1899769  0.7754308  0.1253349
+#   4     extratrees  0.1925299  0.7604062  0.1270665
+#   6     variance    0.1926697  0.7707287  0.1277624
+#   6     extratrees  0.1930151  0.7601929  0.1280934
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule = variance and
-#  min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule = variance and min.node.size = 5.
 
@@
 ggsave(path.out)
+
+path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected_10.csv")
+varimp_mars$importance %>% 
+  rownames_to_column() %>%
+  rename(variable = rowname,
+         importance = Overall) %>% 
+  write_csv(path.out)
 ```
@@
 #   mtry  splitrule   RMSE       Rsquared   MAE      
-#   2     variance    0.2773324  0.6169393  0.2025296
-#   2     extratrees  0.2692856  0.6437281  0.1982852
-#   4     variance    0.2836790  0.5992575  0.2060827
-#   4     extratrees  0.2717096  0.6339425  0.2004439
-#   6     variance    0.2899917  0.5806674  0.2100017
-#   6     extratrees  0.2755481  0.6223759  0.2037709
+#   2     variance    0.2773784  0.6171406  0.2025901
+#   2     extratrees  0.2696423  0.6424804  0.1985126
+#   4     variance    0.2834811  0.5991036  0.2058774
+#   4     extratrees  0.2720828  0.6324695  0.2007287
+#   6     variance    0.2902072  0.5801010  0.2101899
+#   6     extratrees  0.2755322  0.6215775  0.2037045
 # 

### 2024-08-16 src/oa_RF_corrected.Rmd
all but OA

@@
   group_by(year, oa11cd) %>%
-  summarise(n = n()) %>% # for multiple pc  
+  summarise(n = sum(n)) %>% # **********ATTENTION*********  
   ungroup() %>% 

### 2024-08-19 src/oa_RF_corrected.Rmd
0595A

@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    1.221299  0.4082199  0.5829598
-#   2     extratrees  1.308769  0.3297824  0.6136490
-#   5     variance    1.203082  0.4241359  0.5703508
-#   5     extratrees  1.280883  0.3513153  0.6124031
-#   8     variance    1.214268  0.4155308  0.5674246
-#   8     extratrees  1.284107  0.3502621  0.6180390
+#   2     variance    3.283708  0.3202280  1.034153
+#   2     extratrees  3.516978  0.2323646  1.087689
+#   5     variance    3.240309  0.3354508  1.019184
+#   5     extratrees  3.443571  0.2565672  1.107110
+#   8     variance    3.278693  0.3251487  1.017672
+#   8     extratrees  3.449965  0.2562403  1.124713
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 5, splitrule = variance and
-#  min.node.size = 5.
+# The final values used for the model were mtry = 5, splitrule = variance and min.node.size = 5.
 

### 2024-08-19 src/oa_RF_corrected_10.Rmd
0595A

@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    1.221299  0.4082199  0.5829598
-#   2     extratrees  1.308769  0.3297824  0.6136490
-#   5     variance    1.203082  0.4241359  0.5703508
-#   5     extratrees  1.280883  0.3513153  0.6124031
-#   8     variance    1.214268  0.4155308  0.5674246
-#   8     extratrees  1.284107  0.3502621  0.6180390
+#   2     variance    22.46734  0.4784290  5.681843
+#   2     extratrees  24.69943  0.3875775  6.049473
+#   5     variance    22.50225  0.4803563  5.664478
+#   5     extratrees  23.95316  0.4143508  6.102471
+#   8     variance    23.14682  0.4644807  5.678126
+#   8     extratrees  24.14426  0.4095256  6.213216
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 5, splitrule = variance
-#  and min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule = variance and min.node.size = 5.
 

### 2024-08-19 src/s_la_per_firm_corrected.Rmd
0595A

@@
 library(stargazer)
+library(ggbump)
 
@@
 ```
+
+## Average ranks websites per firm t and 2012 only
+
+```{r}
+
+t <- 2000
+
+out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only0595_av.png")
+
+df %>%
+
+  left_join(la %>% as_tibble() %>% dplyr::select(LAD21CD, LAD21NM),
+            by = c("ladcd" = "LAD21CD")) %>%
+
+  mutate(period = ifelse(year < 1999, 1, 
+                         ifelse(year > 2009, 3, 2))) %>%
+  group_by(ladcd, period) %>%
+  mutate(n = ifelse(period != 2, mean(n), n)) %>% 
+  filter(year > 1997 & year < 2011) %>% 
+  group_by(year) %>% 
+  mutate(rank = rank(desc(n))) %>% 
+  ungroup() %>% 
+  filter(year==1998 | year == 2010) %>% 
+  arrange(ladcd, year) %>% 
+  group_by(ladcd) %>% 
+  #mutate(diff_1st = -diff(rank))
+  mutate(diff_1st = -(rank-dplyr::lag(rank))) %>%
+
+  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .05, na.rm = T))
+  # q.95 = 253, q.9 = 214, q.1 = 11, q.05 = 3.75. **abs()**
+
+  mutate(alpha = ifelse(abs(diff_1st) <= 3.75 | abs(diff_1st) > 253, 1, 0.5)) %>% 
+  mutate(alpha = ifelse(is.na(diff_1st), 0.5, alpha)) %>%
+
+  mutate(q = ifelse(abs(diff_1st) <= 3.75, "05",
+                    ifelse(abs(diff_1st) > 253, "95", "m"))) %>% 
+  mutate(q = ifelse(is.na(diff_1st), "m", q)) %>%
+
+  ggplot(aes(year, rank, group = ladcd, alpha=alpha)) +
+  geom_bump(aes(colour = q)) + #colour = "red"
+  scale_color_manual(values = c("darkorchid3", "orange", "grey70"),
+                     breaks = c("05", "95"),
+                     labels = c("Stability","Volatility")) + 
+  labs(color = "LAD") +
+  scale_y_reverse(labels = c(1, 100, 200, 300, 400)) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  #theme(legend.position="none") +
+  scale_x_continuous(NULL, 
+                     #limits = c(1998, 2010), 
+                     breaks=c(1998, 2010),
+                     labels = c("1996-98","2010-12")) +
+  guides(alpha = "none")
+
+ggsave(out.path)
+```

### 2024-08-19 src/problem_years_10.Rmd
lisa10

@@
   geom_sf() +
-  # scale_fill_discrete_c4a_cat(name = "%", palette = "parks.charmonix",
-  #                             na.translate = FALSE) +
   scale_fill_manual(values = palette_new, na.value="white", breaks = NULL, drop = F) +
@@
   #guides(fill = guide_legend(reverse = TRUE)) +
-  labs(title = "LISA, websites per firm in a LA") + 
-  theme(legend.title=element_blank())
-
-theme(axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks = element_blank(),
-        rect = element_blank())
+  labs(title = "LAD", t) + 
+  theme(legend.title=element_blank(),
+        plot.margin = unit(c(0, 0, 0, 0), "cm"),
+        plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(size=20, face = "bold"))
 
@@
           #legend.position="none") +
-    guides(fill = guide_legend(reverse = TRUE)) +
-    labs(title = t,
-        subtitle = "LISA, websites per OA")
+    #guides(fill = guide_legend(reverse = TRUE)) +
+    labs(title = "OA", t) + 
+    theme(legend.title=element_blank(),
+          plot.margin = unit(c(0, 0, 0, 0), "cm"),
+          plot.title = element_text(hjust = 0.5)) +
+    theme(plot.title = element_text(size=20, face = "bold"))
 

### 2024-08-19 src/problem_years_10.Rmd
lisa10

--- 2024-08-15 src/problem_years_10.Rmd
@@
   #guides(fill = guide_legend(reverse = TRUE)) +
-  labs(title = "LISA, websites per firm in a LA") + 
-  theme(legend.title=element_blank())
-
-theme(axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks = element_blank(),
-        rect = element_blank())
+  labs(title = "LAD, ", t) + 
+  theme(legend.title=element_blank(),
+        plot.margin = unit(c(0, 0, 0, 0), "cm"),
+        plot.title = element_text(hjust = 0.5)) +
+  theme(plot.title = element_text(size=20, face = "bold"))
 
@@
           #legend.position="none") +
-    guides(fill = guide_legend(reverse = TRUE)) +
-    labs(title = t,
-        subtitle = "LISA, websites per OA")
+    #guides(fill = guide_legend(reverse = TRUE)) +
+    labs(title = paste0("OA, ", t)) + 
+    theme(legend.title=element_blank(),
+          plot.margin = unit(c(0, 0, 0, 0), "cm"),
+          plot.title = element_text(hjust = 0.5)) +
+    theme(plot.title = element_text(size=20, face = "bold"))
 

### 2024-09-11 src/oa_RF_corrected_10.Rmd
oa corrected 10

@@
 
-# correct for non-England file names
-df <- df %>% 
-  mutate(RGN11CD = ifelse(RGN11CD == "Wales", "W0",
-                          ifelse(RGN11CD == "Scotland", "S1",
-                                 ifelse(RGN11CD == "NI", "N0", RGN11CD))))
-
-regions <- df %>% 
-  ungroup() %>% 
-  distinct(RGN11CD)  
+regions <- df %>% ungroup() %>% distinct(RGN11CD) 
 regions <- as.vector(regions$RGN11CD)
@@
   rename(RGN11CD = RGN21CD) %>% 
-  add_row(RGN11CD = c("S1", "N0", "W0"), 
+  add_row(RGN11CD = c("Scotland", "NI", "Wales"),
           RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))

### 2024-09-11 src/la_RF_corrected.Rmd
oa correlations

@@
 ```
+
+## Correlations
+
+```{r}
+cor <- df %>% dplyr::select(n, n.London.lag, n.nearest.city.lag, n.l.slag, year, London, dist) 
+cor <- rcorr(as.matrix(cor))
+r <- cor$r
+p <- cor$P
+
+path.out <- paste0(path, "/outputs/rf/correlations/corr_LAD_corrected.csv")
+r %>% as_tibble(rownames = "rowname") %>% 
+  write_csv(path.out)
+
+path.out <- paste0(path, "/outputs/rf/correlations/corr_LAD_corrected_p.csv")
+p %>% as_tibble(rownames = "rowname") %>% 
+  write_csv(path.out)
+```

### 2024-09-11 src/la_RF_corrected_10.Rmd
oa correlations

@@
 ```
+
+## Correlations
+
+```{r}
+cor <- df %>% dplyr::select(n, n.London.lag, n.nearest.city.lag, n.l.slag, year, London, dist) 
+cor <- rcorr(as.matrix(cor))
+r <- cor$r
+p <- cor$P
+
+path.out <- paste0(path, "/outputs/rf/correlations/corr_LAD_corrected_10.csv")
+r %>% as_tibble(rownames = "rowname") %>% 
+  write_csv(path.out)
+
+path.out <- paste0(path, "/outputs/rf/correlations/corr_LAD_corrected_10p.csv")
+p %>% as_tibble(rownames = "rowname") %>% 
+  write_csv(path.out)
+```

### 2024-09-11 src/oa_RF_corrected.Rmd
oa correlations

@@
 ```
+
+## Correlations
+
+```{r}
+cor <- rcorr(as.matrix(df[-(1:3)]))
+r <- cor$r
+p <- cor$P
+
+path.out <- paste0(path, "/outputs/rf/correlations/corr_OA_corrected.csv")
+r %>% as_tibble(rownames = "rowname") %>% 
+  write_csv(path.out)
+
+path.out <- paste0(path, "/outputs/rf/correlations/corr_OA_corrected_p.csv")
+p %>% as_tibble(rownames = "rowname") %>% 
+  write_csv(path.out)
+```

### 2024-09-11 src/oa_RF_corrected_10.Rmd
oa correlations

@@
 library(CAST)
+library(Hmisc)
 
@@
 ```
+
+## Correlations
+
+```{r}
+cor <- rcorr(as.matrix(df[-(1:3)]))
+r <- cor$r
+p <- cor$P
+
+path.out <- paste0(path, "/outputs/rf/correlations/corr_OA_corrected_10.csv")
+r %>% as_tibble(rownames = "rowname") %>% 
+  write_csv(path.out)
+
+path.out <- paste0(path, "/outputs/rf/correlations/corr_OA_corrected_10p.csv")
+p %>% as_tibble(rownames = "rowname") %>% 
+  write_csv(path.out)
+```

### 2024-09-12 paper/submission1/tranos2023.qmd
postpresentation2

@@
       - id: some-tech
-        name: University of Bristol and The Alan Turing Institute
+        name: University of Bristol
         department: 
@@
 library(cowplot)
+library(patchwork)
 
@@
 
-<!-- geographers lack of interest + motivation -->
-
-Geographers were always interested in how new technologies and
-innovations diffuse across space and time and, importantly, how such
-spatio-temporal processes can be modeled. After all, diffusion together
-with invention and innovation are considered the pillars of
-technological change [@das2022diffusion]. The seminal contribution of
-@hagerstrand1968innovation is illustrative of this early interest.
-However, the torch of exploring and modelling such processes had been
+<!-- geographers lack of interest for diffusion + motivation -->
+Economic geographers were always interested in how new technologies and
+innovations diffuse across space and time with the work of @hagerstrand1968innovation,
+which demonstrated the importance of communications as a mechanism for the 
+diffusion of innovations, being the most impactful. It was followed by various 
+empirical studies -- see for instance @ormrod1990 and @iso2005. 
+However, the torch of exploring and modelling such diffusion processes had been
 passed to other disciplines such as economics, business studies and
 sociology well before the 'cultural turn' of economic geography
-[@perkins2005international]. A potential explanation of the lack of
-geographical studies exploring the diffusion of new and, more
-specifically for this paper, digital technologies across both space
-and time can be attributed to the scarcity of relevant and granular
-enough data. As @zook2022mapping highlight, digital activities are
-hardly ever captured in official data.
+[@perkins2005international]. A potential explanation of the recent lack of
+geographical studies exploring the diffusion of new and, more specifically for 
+this paper, digital technologies across both space and time can be attributed to 
+the scarcity of relevant and granular enough data, a problem also highlighted 
+by @iso2005 and @kemeny2011international. As @zook2022mapping pointed, digital 
+activities are hardly ever captured in official data.
 
 <!-- contribution -->
-
-This paper offers such a contribution: a geographical study illustrating
-how a new technology that is the web diffused over space and time in the
-UK at a high level of spatial granularity during the $1996-2012$ period.
-It does so by employing a novel source of big data which captures the
-active engagement with web technologies during that period. By
-addressing this empirical question this paper exemplifies how the
-combination of data sources which escape the traditional social science
-domain and adequate research methods can offer new lenses to
-geographical research regarding the understanding of technological
-diffusion.
-
-**SAY SOMETHING ABOUT SPATIAL ANALYTICS, DIST PAPER BY ARP, NOT SPECIFIC EXPLANATORY VARIABLES, INTANGIBLE**
-
-<!-- motivation -->
-
+This paper offers such a contribution: a geographical study demonstrating
+how a new technology that is the Web diffused over space and time in the
+UK at a high level of spatial granularity during the 1996-2012 period.
+It does so by employing a novel source of big data and machine 
+learning algorithms to model the *active* engagement with Web during that 
+period and tests how different spatial and temporal mechanisms shaped its 
+diffusion. This paper also exemplifies how the combination of data sources 
+which escape the traditional social science domain and adequate research 
+methods can offer new lenses to geographical research regarding the 
+understanding of technological diffusion.
+
+<!-- motivation: stakeholders -->
 The motivation for this paper lies in the fact that there are various
@@
 diffused over space and time and use this knowledge to make predictions
-regarding the diffusion of related *future* technologies. As per
+about the diffusion of related *future* technologies. As per
 @leibowicz2016representing, historical studies agree that technologies diffuse 
@@
 that a variety of actors have a direct interest in gaining such
-knowledge including network equipment suppliers; network operators,
+knowledge including network equipment suppliers, network operators,
 regulatory and local authorities. These processes and their effects vary
@@
 places [@leibowicz2016representing]. Despite the spatial heterogeneity
-of such diffusion mechanisms and the policy relevance, there are very
-limited attempts in the literature to analyse the diffusion of new
-digital technologies at a detailed geographical level.
+of such diffusion mechanisms and the policy relevance, there is a scarcity 
+of empirical studies analysing the diffusion of new digital technologies 
+at a detailed geographical level.
+
+<!-- motivation: regional development -->
+This gap becomes even more prominent if we consider the importance of 
+technology and its adoption in economic geography and macroeconomic growth 
+theories. Simply put, differences in technological adoption and sophistication 
+may lead to uneven development as new technologies can increase productivity 
+and allow firms to introduce new products [@solow1957technical; 
+@aghion1990model; @kemeny2011international]. Linked to the subject of this 
+paper, the Web -- and the active engagement with this technology that is 
+creating instead of just browsing a website -- can be considered as the 
+underpinning technology for the transformative process known as the
+*digital service economy* [@capello2024nexus].
+After all, diffusion together with invention and innovation are considered 
+the pillars of technological change [@das2022diffusion].
+Findings from previous empirical work illustrated that early engagement with 
+digital technologies such as the Web led to longstanding positive productivity 
+gains for these regions that hosted early adopters [@tranosuk].
+
+<!-- motivation: intangible/digital -->
+Another line of motivation has to do with the intangible nature of the Web 
+and the debate about the footloose nature of digital technologies, which
+contrary to several earlier general purpose technologies are light in weight 
+[@coyle1999weightless]. Following the seminal work of @gaspar1998information, 
+which demonstrated that technological improvements in telecommunications lead 
+to increased demand for face-to-face communications and, consequently, 
+enhance the importance of cities, empirical studies from economic geography 
+and urban economics explored whether digital technologies substitute or 
+complement agglomeration externalities and spatial structure 
+[@kolko_death_2000; @ioannides2008effect; @tranos2021ubiquitous].
+With the exception of @sinai2004geography, who focused, among other things, 
+on web content and demonstrated that more such content is available in larger
+markets, this strand of literature has been mostly concerned with the
+spatiality of the internet's hard and tangible infrastructure
+and shied away from exploring the spatial dimensions of one of the 
+most distinct software layers of the internet infrastructure, the Web.
 
 <!-- diffusion and adoption -->
-
 Technological diffusion, which is by definition an aggregated process,
@@
 
-<!-- data and early engagement with the web -->
-
-This paper focuses on the diffusion of the web as new technology during
-the $1996-2012$ period. This was an exciting period for digital
+<!-- early engagement with the web -->
+This paper focuses on the diffusion of the Web as new technology during
+the 1996-2012 period. This was an exciting period for digital
 technologies as it corresponds with the commercialisation of the
@@
 introduced, a year before Microsoft's Internet Explorer.[^1] Also, only
-$9$ per cent of UK's households had access to the internet in $1998$
+9 per cent of UK's households had access to the internet in 1998
 [@ons2018], the web included mostly static webpages, there were no
-social media and web browsing was happening exclusively from desktop PCs
+social media and web browsing involved exclusively a desktop PCs
 as there were no smartphones [@tranosuk]. Hence, it is fair to say that
-the study period captures the very early stages of the diffusion on a
-new technology that is the web as well as its maturity. 
-The former is a key point in the lifecycle of a
-new technology. Firstly, during this period new technologies are
-expensive, crude and imperfect [@rosenberg1994exploring; @wilson201281].
+the study period captures the very early stages of the diffusion of a
+new technology that is the Web as well until its maturity. 
+The former is a key point in the lifecycle of a new technology. 
+Firstly, during this period new technologies are expensive, crude and 
+imperfect [@rosenberg1994exploring; @wilson201281].
 A simple comparison between Web 1.0 and Web 3.0 applications clearly
-illustrates this [@tranos2020social]; for instance a static website compared 
-with a platform like `github`, which enables cooperation between users and 
-the creation of new information, meaning, and knowledge 
-[@faraj2016special; @barassi2012does]. 
-During this period the performance
-of a new technology is the main attraction and not the cost to access
-and use it [@wilson2011lessons]. There is a broader theoretical
-discussion in the literature regarding the motivation behind early
-adoption. As summarised by @perkins2005international, on the one hand,
-epidemic models highlight the role of interpersonal contacts as a way
-for new technologies to diffuse. On the other hand, economic models
-underline the importance of heterogeneity. Different firms have
+illustrates this [@tranos2020social]: while a static website facilitates
+one-way information dissemination, a platform like `github` enables 
+cooperation between users and the creation of new information, meaning, 
+and even knowledge [@faraj2016special; @barassi2012does]. 
+During this period the performance of a new technology is the main 
+attraction and not the cost to access and use it [@wilson2011lessons]. 
+There is a broader theoretical discussion in the literature about 
+the early adoption motives. As summarised by @perkins2005international, 
+on one hand, epidemic models highlight the role of interpersonal 
+contacts as a way for new technologies to diffuse. On the other hand, 
+economic models underline the importance of heterogeneity. Different firms have
 different structures and business plans, which define the potential
@@
 
-Importantly, the data used here depicts the *active* engagement with the
-web in the UK as it contains all commercial websites that 
-(i) are part of the UK's relevant second level domain (SLD, .co.uk), 
-(ii) have been archived by the Internet Archive[^2], and 
-(iii) include a mention to at least one valid UK
-postcode in the web text. 
-The act of creating a website is understood here as active engagement with the
-web vis-à-vis the more passive act of browsing the web or having an
-internet connection [@tranosuk]. Previous studies have focused mostly on
-more passive notions of engaging with digital technologies such as
-internet adoption and internet speeds [e.g. @blank2018local;
-@riddlesden2014broadband]. More details about the data and the data
-generation process can be found in @sec-datamethods.
-
-**S curve??**
+<!-- data -->
+The data used here depict the active engagement with the Web as they contain 
+geolocated and time-stamped counts of commercial websites from the UK. Instead 
+of adopting metrics reflecting the passive engagement with the internet such as 
+internet adoption and speeds [e.g. @blank2018local; @destefano2022fuzzy], this 
+paper observes yearly counts of geolocated websites, which reflect the creation 
+and maintenance of websites. It needs to be highlighted here that although 
+there are various types of underpinning web technologies -- namely markup, 
+stylesheet, client- and server-side coding languages as well as different
+web browsers and other related programming language -- the focus here
+is on the diffusion of the Web in its totality. For a more detailed and 
+cross-country technological diffusion study, see @PAPAGIANNIDIS2015308.
 
 <!-- description of the methods and results -->
+To analyse these data, a diverse set of methods are employed. Firstly, 
+cumulative adoption curves are estimated for the UK as a whole and for the Local 
+Authority Districts (LAD), illustrating, for the first time, such S-shaped 
+diffusion patterns for local areas. This system level analysis also reveals 
+insights about the volatility of the adoption of the Web. Then, exploratory 
+spatial data analysis offers some first indications of the importance of the 
+different spatial diffusion mechanisms in shaping the diffusion of the Web in 
+the UK. Importantly, the analysis adopts two distinct geographical scales: 
+the LAD and the much more detailed scale of the smallest census areas in the UK,
+the Output Areas. Lastly, a novel methodological framework based on 
+well-established Machine Learning (ML) algorithms is developed to test the role 
+of such mechanisms at these two scales. The results reveal for the first time 
+the importance and relevance of different spatial diffusion mechanisms for the 
+diffusion of a tangible, digital technology at very granular geographical scales.
 
 <!-- contents -->
-
-@grubler1990rise Later Hagerstrand conceptualized physical "barrier"
-effects like lakes or uninhabited areas, which, in addition to distance,
-act as further retarding effects on diffusion. These are formalized in
-the form of "zero" or "half" contact multiplicators on the (distance
-decaying) message flows.
-
-@grubler1990rise With respect to the formalization of the communication
-flows Hagerstrand defines a "mean information field" (MIF), in which the
-probability of communication is a negative function of distance between
-individuals
-
-@wilson201281 Logistic growth describes an initial period of gradual
-diffusion as a technology is introduced as a new commercial application,
-moving then through a rapid, exponential growth phase, before slowing
-and eventually saturating [@grubler1999dynamics]. The substitution of
-incumbent technologies by new competitors leads to subsequent decline
-and eventual obsolescence.
+The structure of the paper goes as follows. @sec-litreview reviews the key
+concepts in the diffusion literature as well as relevant empirical studies. 
+Then, @sec-datamethods describes the data and methods used in this paper. 
+The results of the different stages of the analysis are presented in 
+@sec-results. The discussion of the results and the conlcusions of the paper
+are offered in @sec-conclusions.
+
+<!-- @grubler1990rise Later Hagerstrand conceptualized physical "barrier" -->
+<!-- effects like lakes or uninhabited areas, which, in addition to distance, -->
+<!-- act as further retarding effects on diffusion. These are formalized in -->
+<!-- the form of "zero" or "half" contact multiplicators on the (distance -->
+<!-- decaying) message flows. -->
+
+<!-- @grubler1990rise With respect to the formalization of the communication -->
+<!-- flows Hagerstrand defines a "mean information field" (MIF), in which the -->
+<!-- probability of communication is a negative function of distance between -->
+<!-- individuals -->
+
+**SAY SOMETHING ABOUT DIST PAPER BY ARP, NOT SPECIFIC EXPLANATORY VARIABLES**
 
@@
 <!-- key concepts -- copied from help.Rmd -->
-
 Geographical diffusion is a synthesis of different processes. On the one
-hand, a purely spatial or, in other words, contagious processes can be identified. 
-Adjacency and, more broadly, distance are the key drivers of
+hand, a purely spatial or, in other words, contagious processes can be 
+identified. Adjacency and, more broadly, distance are the key drivers of
 diffusion. This perspective draws similarities with epidemics:
@@
 consequently, proximity and exposure [@hivner2003facilitating]. On the
-other hand, there is a hierarchical processes. Instead of
-horizontal distance-based diffusion mechanisms, the top-down hierarchy
-of urban systems shapes technological diffusion. In reality, the synthesis of these
-two processes represents how new technologies diffuse over space and
-time [@morrill2020spatial].
+other hand, there is a hierarchical processes. Instead of a
+horizontal distance-based diffusion mechanism, the top-down hierarchy
+of urban systems shapes technological diffusion. In reality, the synthesis 
+of these two processes better reflects how new technologies diffuse over 
+space and time [@morrill2020spatial].
 
@@
 which themselves tend to be defined by geography. Hence, he identified
-the 'neighbourhood' effect of how information, and consequently,
+a 'neighbourhood' mechanism of how information, and consequently,
 innovation diffuse. He used agricultural innovations to test and model
@@
 adopters, early majority, late majority and laggards. The key mechanism of
-diffusion and adoption is communication and how knowledge is
-transferred within a social system. Therefore, all approaches agree on
-the S-shaped diffusion and cumulative adoption pattern
-[@grubler1990rise].
+diffusion and adoption is communication and how knowledge is transferred 
+within a social system. All above approaches agree that diffusion and the
+cumulative adoption of a new technology are represented by an S-shaped 
+line plot [@grubler1990rise].
 
@@
 core because of the lack of infrastructure or other necessary
-institutions [@leibowicz2016representing]. @grubler1990rise effectively
-summarises the three key characteristics of the spatial diffusion
-process: (i) the cumulative level of adoption follows an S-shaped
+institutions [@leibowicz2016representing].
+
+@grubler1990rise effectively summarises the three key spatial diffusion 
+mechanisms: (i) the cumulative level of adoption follows an S-shaped
 pattern just like purely temporal models; (ii) diffusion is shaped by a
-hierarchy effect in a form of a centrifugal force: from core to
-periphery; and (iii) diffusion is also shaped by distance and a
-neighbourhood effect and contaminate nearby locations. These are the three
-mechanisms that the empirical analysis in @sec-results investigates.
+hierarchy mechanisms in a form of a centrifugal force: from core to
+periphery; and (iii) diffusion is also shaped by distance and proximity. 
+These are the three mechanisms that the empirical analysis in @sec-results 
+investigates.
 
 <!-- previous studies -->
-
 The remaining of this section reviews empirical studies which analysed
@@
 spatial dimension is present in most of the following studies, the level
-of spatial detail is always more coarse than the one adopted
-here. @beardsell1999spatial studied the evolution of the computer
-industry in $317$ US metro areas during the $1977-1992$ period using
+of spatial detail is always more coarse than the one adopted in this
+paper. @beardsell1999spatial studied the evolution of the computer
+industry in 317 US metro areas during the 1977-1992 period using
 employment data. Their analysis indicated that the relative size
@@
 heterogeneity is essential in explaining this distribution. In a recent
-study, @bednarz2020pulled focused on wind turbines and modelled their
-spatial diffusion across $402$ German regions during $1970-2015$. Their
+study, @bednarz2020pulled focused on wind turbines and modeled their
+spatial diffusion across 402 German regions during 1970-2015. Their
 key finding is that local demand than local supply was the main driving
-factor.
-
-At a global scale @perkins2005international explored whther the
+factor. @haller2011determinants employed firm panel data for Ireland
+during 2003-2005 and illustrated that when firms are located in a
+region or industry with high share of firms having a website, they are
+more likely to have a website too.
+
+At a global scale @perkins2005international explored whether the
 diffusion rate of new technologies is driven by a latecomer advantage
@@
 @perkins2011internet explored whether the adoption of previous
-communication technologies that is mail, telegrams and telephones was
+communication technologies that is mail, telegrams and telephones were
 shaped by similar socioeconomic factors as the internet. Their results
@@
 Turning to studies that share more technological and scalar similarities
-with this paper, @ding2010modeling modelled the spatial diffusion of
+with this paper, @ding2010modeling modeled the spatial diffusion of
 mobile telecommunications across regions in China. Their analysis
@@
 contributions reviewed here: assortativity, urban scaling and distance
-are the key drivers of spatial diffusion. At a global scale
-@PAPAGIANNIDIS2015308 modelled the diffusion of different web
-technologies technologies and practices. Interestingly, they did so by
-using similar, but less extensive data as the one used here. Their
-analysis illustrated how the diffusion of different web technologies and
+are the key drivers of spatial diffusion. @bakher2013diffusion compared 
+different the growth curves as a means to model diffusion of online 
+shopping in Australia during the 1998-2009 period. At a global scale
+@PAPAGIANNIDIS2015308 modeled the diffusion of different web
+technologies and practices across countries. Interestingly, they used 
+similar, but less extensive data as the ones used here. Their analysis 
+illustrated how the diffusion of different web technologies and
 practices follow an S-shaped pattern as well as the different diffusion
@@
 
-All in all, ...
+All in all, this paper tests whether the three distinct spatial diffusion 
+mechanisms shaped the diffusion of the Web in the UK at very granular
+spatial scales from the Web's early stages onwards.
 
@@
 <!-- @bento2018time explore What determines the duration of formative phases -->
-
 <!-- for energy innovations in different markets? We are interested both in -->
-
 <!-- initial markets (also: core, lead, first mover, early adopter) where -->
-
 <!-- formative phases prepare technologies for mass commercialization, and in -->
-
 <!-- follower markets (also: periphery, lag, late adopter) where accelerated -->
-
 <!-- formative phases may benefit from diffusion and spillovers. -->
@@
 
-**ADD FAMILY OF WEB TECHNOLOGIES**
-
 <!-- main metric -->
-
-To capture the diffusion of web technologies, a website density metric is developed 
+To capture the diffusion of the Web, a website density metric is developed 
 for two different geographical scales: the Local Authority Districts (LAD) and the 
 Output Areas (OA). The former is an administrative unit and there are 
-c. 374 such units in the UK. The latter is a census-based geographical unit, which 
-is very small as there are c. 230,000 of them in the UK. This methodological choice 
+374 LAD in the UK. The latter is the smallest census-based geographical unit and 
+there are c. 230,000 of them in the UK. This methodological choice 
 will allow the mapping of the diffusion of web technologies and the assessment of 
@@
 <!-- web data -->
-
 The counts of websites at these scales are calculated using data from the 
-Internet Archive [^2] and, specifically, the JISC UK Web Domain Dataset [@ukwebarchive]. 
-The Internet Archive is one of the most complete and oldest archive of webpages 
-in the world operating since 1996 [@ainsworth2011much; @holzmann2016dawn].
-It is a web crawler, which discovers webpages by following the hyperlinks
-of every webpage its archives.
+Internet Archive[^2] and, specifically, the JISC UK Web Domain Dataset 
+[@ukwebarchive]. The Internet Archive is one of the most complete and oldest 
+archive of webpages in the world operating since 1996 
+[@ainsworth2011much; @holzmann2016dawn]. It is a web crawler, which discovers 
+webpages by following the hyperlinks of every webpage it archives.
 This dataset, which is curated by the British Library, contains all
@@
 
-Instead of using the whole .uk country code top-level domain (ccTLD), this 
-paper focuses on its commercial subset, the .co.uk SLD. 
-This choice decrease the heterogeneity of the web data as 
-such commercial websites have rather specific aims: they are used to diffuse information,
-support online transactions and share opinions [@THELWALL2000441; @blazquez2018big]. 
-Although a UK company can adopt a generic TLD such as .com and these cases escape 
-the data used here, such omissions should not affect our results given the
-popularity of the .uk ccTLD [@tranosuk]: UK consumers prefer to visit a .uk website
-when they are searching for products or services [@hope]; and
-anecdotal evidence indicates that during the first half of 2000, three
-.co.uk domains were registered every minute [@oecd_coms]. 
-Importantly, previous studies illustrated that .co.uk is the most popular UK SLD
-[@tranosuk].
+Instead of using the whole .uk ccTLD, this paper focuses on its commercial 
+subset, the .co.uk second level domain (SLD). This choices decrease the 
+heterogeneity of the web data as such commercial websites have specific 
+aims: they are used to diffuse information, support online transactions 
+and share opinions [@THELWALL2000441; @blazquez2018big]. 
+Although a UK company can adopt a generic TLD such as .com and these 
+cases escape the data used here, such omissions should not affect the 
+validity of the results given the popularity of the .uk ccTLD [@tranosuk]: 
+UK consumers prefer to visit a .uk website when they are searching for 
+products or services [@hope]; and canecdotal evidence indicates that during 
+the first half of 2000, threec .co.uk domains were registered every minute 
+[@oecd_coms]. Importantly, previous studies illustrated that .co.uk is the most 
+popular UK SLD [@tranosuk].
 
 The text from these webpages was scanned using a regular expression (regex) to
-identify strings of text which resemble UK postcodes and one fifth of them included
-a mention to a postcode [@BL2013geo]. This information allows the geolocation of the 
-data and the creation of the LAD and OA counts.
-
-The data cleaning process included an aggregation step, through which
-the archived webpages were aggregated to the parent websites. This website 
-reconstruction allows the creation of a *website* instead of a *webpage* density 
-metric. Websites, which are hierarchical structures, tend to represent specific 
-orgnanisations or entities, and, arguably, are more meaningful observational units 
-than webpages, which ignore the upstream dependencies. 
+identify strings of text which resemble UK postcodes and one fifth of them 
+included a mention to a postcode [@BL2013geo]. This information allows the 
+geolocation of the data and the creation of the LAD and OA counts.
+
+The data cleaning process includes an aggregation step, through which
+the archived webpages are aggregated to the parent websites. This website 
+reconstruction process lend itself to the consteruction of *website* instead 
+of *webpage* density metrics. Websites tend to represent specific orgnanisations 
+or entities, and, arguably, are more meaningful observational units than 
+webpages, which can ignore the upstream dependency of the website they belong to. 
 Based on the following example, all three webpages are part of the 
-same website ([http://www.website.co.uk](http://www.website.co.uk)) and at the end 
-only websites and not the nested webpages were considered as otherwise the metric 
-would have been biased towards large, place specific, websites.
+same website ([http://www.website.co.uk](http://www.website.co.uk)). Only 
+the overall website and not the nested webpages are considered for the counts 
+as otherwise the density metrics would have been biased towards large, place 
+specific, websites.
 
@@
 which has been used elsewhere [@tranosuk; @shoreditch] may lead to websites with 
-multiple postcodes. As per the above example, [www.website.co.uk](http://www.website.co.uk)
-includes two unique postcodes: B15 2TT and BS8 1SS. 
-The distribution of postcodes for 2000 is presented in Table \ref{f2000}, which 
-clearly illustrates the wide range. At the left end of the distribution, there are 
-websites anchored to a unique location (72% of all the reconstructed websites in
-2000), which may represent a small company with a single trading location. At the 
-right end, we have websites with thousands of different postcodes.
-Considering the time period of the analysis, such cases can represent directories
-which used to be popular in the pre-search engines early times of the commercial 
-internet [@tranosuk].[^3]
-
-The analysis presented here is based on two subsets of these data. Firstly, on websites, 
-which only contain only one unique postcode and, therefore, the geolocation process 
-does not suffer from noise. As a robustness check, the analysis is replicated 
-for an extended subset of websites, which include up to ten unique postcodes. These 
-websites are geolocated by equally attaching them to multiple locations. This extended 
-sample includes 94 percent of all the archived websites in 2000.
-
-[^3]: See Figure A1 in **Appendix A in the supplemental data online for examples**.
+multiple postcodes. As per the above example, 
+[www.website.co.uk](http://www.website.co.uk) includes two unique postcodes: 
+B15 2TT and BS8 1SS. The distribution of the number of postcodes per website for 
+2000 is presented in Table \ref{f2000}, which  clearly illustrates a wide range. 
+At the left end of the distribution, there are websites anchored to a unique 
+location (72% of all the reconstructed websites in 2000), which may represent a 
+small company with a single trading location. At the right end, there are 
+websites with thousands of different postcodes. Considering the time period of 
+the analysis, such cases can represent directories which used to be popular 
+in the pre-search engines early times of the commercial internet as well as real 
+estate websites [@tranosuk].
+
+The analysis presented here is based on two subsets of these data. Firstly, on 
+websites, which only contain one unique postcode. As a robustness check, the 
+analysis is replicated for an extended subset of websites, which include up to 
+10 unique postcodes to capture commercial websites which point to multiple 
+locations. They are geolocated by equally attaching them to these locations. 
+This extended sample includes 94 percent of all the archived websites in 2000.
 
@@
 
-Another data cleaning step dealt with some extreme outliers. Figure \ref{pc1} plots
-the website counts for the top 1000 postcodes. Some obvious outliers can be observed 
-for the 2004-2006 period, which can be attributed to a link farm [@BL2013links].  
-The website counts for these six postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1 6GL, 
-M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1000 websites pointing to 
-them, were replaced with predicted values based on a simple panel regression model 
-with postcode fixed effects and yearly dummy variables. As the @sec-results illustrates,
-this led to a small increase of the model predictive capability. To put the magnitude 
-of the data imputation into perspective, this process affected 6 out of the 557,808
-postcodes present in the data.
-
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{pc1}Yearly website counts per postcode (top 1000)", out.height="60%"}
-path.image <- paste0(path, "/outputs/pc_year_1.png")
-knitr::include_graphics(path.image)
-```
-
-To create the website density metric, the yearly website counts at the LAD level are
-standardised by the number of firms in LAD to avoid biases associated with LAD 
-hosting a large number of firms. Given that there is no such statistic for the OA,
-the actual OA level counts are used. However, because of the the consistent spatial 
-definition of OA (they host 40-250 households),[^4] website counts in OA are 
-interpreted as a density metric too.
+Another data cleaning step deals with some extreme outliers. Figure 
+\ref{correct} plots the website counts for the top 1000 postcodes. Some obvious 
+outliers can be observed for the 2002-2006 period for a handful of postcodes, 
+which can be attributed to link farms [@BL2013links]. The website counts for 
+these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1 6GL, M28 2SL, DE21 7BF), which 
+in 2004 or 2005 had more than 1000 websites pointing to them, were replaced 
+with predicted values based on a simple panel regression model with postcode 
+fixed effects and yearly dummy variables. These postcodes refer to a residential 
+area with a small park in South London, residential areas outside Warwick and 
+Manchester, a rural area in the south border of England and Wales and small 
+business parks in Crew and Derby. Hence, it is difficult to believe that these
+postcodes genuinely hosted such a large amounts of websites, which were later 
+extensively declined. To put the magnitude of the data imputation into 
+perspective, this process affected 6 out of the 557,808 postcodes included in 
+the data. As the @sec-results illustrates, this led to a small increase of the 
+model predictive capability.
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{correct}Yearly website counts per postcode (top 1000) before (top) and after (down) data imputation", out.height="70%"}
+path.image1 <- paste0(path, "/outputs/pc_year_1.png")
+path.image2 <- paste0(path, "/outputs/pc_year_1_corrected.png")
+
+p1 <- ggdraw() + draw_image(path.image1)  
+p2 <- ggdraw() + draw_image(path.image2)
+
+par(mar=rep(0,4)) # no margins
+
+p1 + plot_spacer() + p2 + plot_layout(ncol = 2)#, widths = c(1, -0.8, 1))
+```
+
+To create the website density metric, the yearly website counts at the LAD level 
+are standardised by the number of firms in LAD to avoid biases associated with 
+LAD hosting a large number of firms. Given that there is no such statistic for 
+the OA, the actual OA level counts are used. However, because of the the 
+consistent spatial definition of OA (they host 40-250 households),[^4] website 
+counts in OA are interpreted as a density metric too.
 
@@
 <!-- system-level analysis -->
-
-These website density metrics are used for the three different steps of the analysis. 
-Firstly, a system-level analysis explores how the diffusion of web technologies in
-the UK fits with the well-established S-curve. To do so, the following logistic 
-function (Equation \ref{eq:s}) is estimated for the whole of the UK and for each 
-LAD separately:
+These website density metrics are used for the three different stages of the 
+analysis. Firstly, a system-level analysis explores how the diffusion of web 
+technologies in the UK fits with the well-established S-curve. To do so, the 
+following logistic function (Equation \ref{eq:s}) is estimated for the whole 
+of the UK and for each LAD separately:
 
@@
 
-\noindent $k$ is the asymptote or, in other words, the saturation level, $b$ the overall 
-growth rate, and $t_{0}$ the *inflection point*
-of maximum growth at $k/2$, where the logistic function is symmetrical [@wilson201281]. 
-We compare  the LAD $t_0$ of each LAD against the $t_0$ of the UK to delineate 
-whether a LAD reached that point faster or slower than the country average. Importantly, 
-an accuracy criterion was imposed and only S-curves with $R^2 > 0.9$ were included 
-in the analysis. To estimate Equation \ref{eq:s} a self-starting logistic growth 
-model was employed using the `nls` and `SSlogis` functions in `R`.
+\noindent $k$ is the asymptote or, in other words, the saturation level, $b$ 
+the overall growth rate, and $t_{0}$ the *inflection point* of maximum growth 
+at $k/2$, where the logistic function is symmetrical [@wilson201281]. 
+The $t_0$ of each LAD is compared against the $t_0$ of the UK to delineate 
+whether a LAD reached that point faster or slower than the country average. 
+Importantly, an accuracy criterion was imposed and only LAD with $R^2 > 0.9$ 
+were included in this analysis. To estimate Equation \ref{eq:s} a self-starting 
+logistic growth model was employed using the `nls` and `SSlogis` functions in 
+`R`. This logistic growth model depicts the initial period of slow diffusion of 
+a new technology, which is followed by the fast, exponential growth stage to
+then eventually slow down and saturate [@wilson201281; @grubler1999dynamics].
 
@@
 The system-level analysis also focuses on the volatility of the web adoption to 
-depict places with high concentration of early adopters and, equally, latecomers. 
-To do so, the change of the ranking of the UK LAD over time is plotted and discussed. 
-Both the S-curve and the volatility analysis focus only on LAD as the very large 
-number of OA would have made such analysis diffucult to visualise and interpret.
+depict places with high concentration of early adopters and, equally, 
+latecomers. To do so, the change of the ranking of the UK LAD over time is 
+plotted and discussed. Both the S-curve and the volatility analysis focus only 
+on LAD as the very large number of OA would have made such analysis difficult 
+to visualise and interpret.
 
 <!-- ESDA -->
-
 Next, exploratory spatial data analysis depicts whether the two main 
-drivers of spatial diffusion -- namely neighbourhood and hierarchy -- underpin the
-diffusion of web technologies in the UK. To capture the former and following 
-@ding2010modeling the Moran's I and the Local Index of Spatial Association (LISA) 
-are estimated for website density. To address the hierarchy mechanism, the Gini 
-coefficient -- a well established metric of inequality -- is calculated. All the 
-above are computed and plotted longitudinally both for the LAD and the OA.
+drivers of spatial diffusion -- namely neighbourhood and hierarchy -- underpin 
+thediffusion of web technologies in the UK. To capture the former and following 
+@ding2010modeling the Moran's I and the Local Index of Spatial Association 
+(LISA) are estimated for the website density. To address the hierarchy mechanism, 
+the Gini coefficient -- a well established metric of inequality -- is 
+calculated. All the above are computed and plotted longitudinally both for the 
+LAD and the OA.
 
 <!-- RF -->
-
-Lastly, a modelling framework is developed to test the above diffusion mechanisms.
-The overarching aim is to build a model that can test the relationship between the
-website density and these mechanisms:
+Lastly, a modelling framework is developed to test the above diffusion 
+mechanisms. The overarching aim is to build a model that can test the 
+relationship between website density and these mechanisms:
 
@@
 
-To estimate Equation \ref{eq:rf.generic}, Random Forest (RF) models were built. 
-This is a popular machine learning algorith for both regression and classification 
+\noindent Equation \ref{eq:rf.generic} is estimated using Random Forest (RF).
+This is a popular ML algorithm for both regression and classification
 problems [@biau2012analysis]. It was introduced by @breiman2001random and has 
-gained popularity, becoming a go-to data science tool. RF can effectively 
-handle skewed distributions and outliers, model non-linear relationships, require 
-minimal hyperparameter tuning, exhibit low sensitivity to these parameters, and 
-have relatively short training times [@Caruana2008; @liaw2002classification; @yan2020using]. 
-These attributes match well with the website density data characteristics including 
-skewness especially for OA. Also, the large data size (c. 230k data points for 
-each of the 17 years) call for fast training times. 
-Importantly, RF predictions tend to be more accurate than those from single regression 
-trees and outperform Ordinary Least Squares in out-of-sample predictions, even with 
-moderate-sized training data and a small number of predictors [@mullainathan2017machine; 
-@athey2019machine; @sulaiman2011intelligent; @pourebrahim2019trip; @biau2012analysis].
-
-RF is a tree-based ensemble learning algorithm [@breiman2001random]. It begins by 
-generating random samples of the training data, which are then used to grow 
-regression trees to predict the dependent variable. Data points and predictors are 
-randomly sampled for the different trees. The trees are trained in parallel using their 
-own bootstrapped samples of the training data. A crucial feature of RF is their 
-ability not to overfit, meaning they can generalize well to unseen test data. While 
-each tree may overfit individually, the ensemble of trees does not because the errors 
-of individual trees are averaged, reducing the overall variance and preventing overfitting 
-[@last2002improving]. For regression problems, RF predictions are made by averaging 
-the predictions of all decision trees.
-
-RF have been widely employed to address regression research problems. @pourebrahim2019trip 
-combined a spatial interaction modeling framework with ML algorithms including RF
-to predict commuting flows in New York City. @sinha2019assessing advocated for adopting
-spatial ensemble learning approaches, such as RF, to model spatial data with high
-autocorrelation and heterogeneity. @creditspatial predicted employment density in 
-Los Angeles using spatially explicit RF. @guns2014recommending used RF to build a 
-recommendation system for research collaborations. @ren2019predicting trained RF 
-to predict the socio-economic status of cities using various online and mobility 
-predictors. @tranos2023using utilised hyperlinks data and RF to make out-of-sample
-predictions of interregional trade. @zhou2023geography employed such a framework
-to assess whether key predictors of obesity differ across English cities.
+become a go-to data science tool. RF can effectively handle skewed distributions 
+and outliers, model non-linear relationships, require minimal hyperparameter 
+tuning, exhibit low sensitivity to these parameters, and have relatively short 
+training times [@Caruana2008; @liaw2002classification; @yan2020using]. These 
+attributes match well with the website density data characteristics including 
+skewness especially for the OA. Also, the large data size (c. 230k data points 
+for each of the 17 years) calls for fast training times. Importantly, RF 
+predictions tend to be more accurate than those from single regression trees and 
+outperform Ordinary Least Squares in out-of-sample predictions, even with 
+moderate-sized training data and a small number of predictors 
+[@mullainathan2017machine; @athey2019machine; @sulaiman2011intelligent; 
+@pourebrahim2019trip; @biau2012analysis].
+
+RF is a tree-based ensemble learning algorithm [@breiman2001random]. It begins 
+by generating random samples of the training data, which are then used to grow 
+regression trees to predict the dependent variable. Data points and predictors 
+are randomly sampled for the different trees. The trees are trained in parallel 
+using their own bootstrapped samples of the training data. A crucial feature of 
+RF is their ability not to overfit, meaning they can generalize well to unseen 
+test data. While each tree may overfit individually, the ensemble of trees does 
+not because the errors of individual trees are averaged, reducing the overall 
+variance and preventing overfitting [@last2002improving]. For regression 
+problems, RF predictions are made by averaging the predictions of all decision 
+trees.
+
+RF have been widely employed to address regression research problems. 
+@pourebrahim2019trip combined a spatial interaction modeling framework with ML 
+algorithms including RF to predict commuting flows in New York City. 
+@sinha2019assessing advocated for adopting spatial ensemble learning approaches, 
+such as RF, to model spatial data with high autocorrelation and heterogeneity. 
+@creditspatial predicted employment density in Los Angeles using spatially 
+explicit RF. @guns2014recommending used RF to build a recommendation system for 
+research collaborations. @ren2019predicting trained RF to predict the 
+socio-economic status of cities using various online and mobility predictors. 
+@tranos2023using utilised hyperlinks data and RF to make out-of-sample 
+predictions of interregional trade. @zhou2023geography employed such a 
+framework to assess whether key predictors of obesity differ across English 
+cities.
 
@@
 
-1.  S-shaped diffusion curves: S for LAD per firm. UK,
-    fast/slow/examples
-
-2.  ranks: there is stability and movement
-
-3.  Neighbourhood effect: diffusion proceeds outwards from innovation
-    centers, first "hitting" nearby rather than far-away locations
-    (Grubler 1990)
-
--   Moran's I: for OA and LAD over time
-
--   LISA maps: for OA and LAD over time More and less expected clusters.
-    Different scales show different results
-
-4.  Hierarchy effect: from main centers to secondary ones -- central
-    places
-
--   Gini coefficient. Almost perfect polarisation of web adoption in the
-    early stages at a granular level More equally diffused at the Local
-    Authority level Plateau overtime
-
-5.  RF
-
--   ideal: (i) train RF for all years and all (1) LAD and (2) OA with
-    CAST and report metrics. (ii) train for all years and all but one
-    region for (1) LAD and (2) OA to predict to the holdout region.
-    Reports predictions as region similarities.
+## System-level analysis
 
 <!-- S-shaped -->
-
-Figure \ref{s_uk} plots the S curve for the cumulative adoption of website technologies
-in the UK during the 1996-2012 period. It demonstrates a pattern well aligned with
-previous studies discussed in @sec-litreview. The vertical line for year 
-2003 illustrates the point where the modeled cumulative adoption was equal to 50% 
-of the maximum. This $t_0$ *inflection* point signals the maximum adoption speed 
-and is used here to determine whether the UK LAD reached that inflection point 
-earlier or later than the UK average. Specifically, the S curve and the inflection 
-point are estimated for every LAD individually and then compared with the UK average. 
-LAD that reached their inflection point earlier than the country average are labelled 
-as *fast* and the rest as *slow*. Figure \ref{s_map} maps this pattern and the picture 
-is not entirely clear. On the one hand, there are some expected examples of LAD
-with relevant industrial backgrounds delineated as fast: the City of London, a 
-world-renowned cluster of finance industries [@cook2007role], and Reading, a town 
-with high-tech service industries in proximity to London and its main airport, Heathrow 
-[@england2005polynet]. On the other hand, LAD which were expected to appear as fast
--- e.g. Hackney in central London and Bristol, a well-established creative cluster 
-[@oatley1999cultural; @bassett2002cultural] -- were delineated as slow. Nevertheless, 
-out of the 10 fastest LAD, eight were located in Greater London and one in 
-Cambridge (see **Appendix**), but in overall there is a spatially heterogenous and
-not easy to explain spatial pattern.
-
 ```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk}Grwoth curve, UK"}
@@
 
+Figure \ref{s_uk} plots the cumulative adoption of the Web in total in the UK 
+during the 1996-2012 period. Not surprisingly, the cumulative adoption is well 
+represented by curve that resembles the figure S. The vertical 
+line around year 2003 illustrates the point where the modeled cumulative 
+adoption is equal to 50% of the maximum. This $t_0$ *inflection* point signals 
+the maximum adoption speed and is used here to determine whether a UK LAD 
+reached that inflection point earlier or later than the UK average. 
+Specifically, the S curve and the inflection point are estimated for every LAD 
+individually and then compared against the UK average. LAD that reached their 
+inflection point earlier than the country average are labelled as *fast* and 
+the rest as *slow*. Figure \ref{s_map} maps this pattern. There is a clear high 
+concentration of fast adopting LAD in the South and East of England as expected. 
+However, there is also such a concentration in Wales, Scotland and the North 
+West of England. At a more detailed level, there are some expected 
+examples of LAD with relevant industrial backgrounds delineated as fast: the 
+City of London, a world-renowned cluster of finance industries [@cook2007role], 
+and Reading, a town with high-tech service industries in proximity to London 
+and its main airport, Heathrow [@england2005polynet]. However, there are also 
+LAD which were expected to appear as fast -- e.g. Hackney in central London and 
+Bristol, a well-established creative cluster [@oatley1999cultural; 
+@bassett2002cultural] -- but were delineated as slow. Nevertheless, there is 
+obvious spatial clustering of fast and slow LAD. Importantly, out of the ten 
+fastest LAD, nine were located in the South East of England and London and one 
+in Cambridge (see Table \ref{table.s.lads} in the Appendix). The above 
+observations remain almost unchanged when the analysis includes websites 
+with up to 10 unique postcodes -- see Figures \ref{s_uk10} and \ref{s_map10} in 
+the Appendix -- advocating towards the robustness of the analysis.
+
 ```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map}LAD adoption rates"}
@@
 
+<!-- discussion -->
+The S-curves of the cumulative adoption of Web in the UK in total and in the 
+UK LAD demonstrate a pattern well aligned with all previous studies discussed 
+in @sec-litreview. Importantly, the above findings are in line with previous 
+studies exploring the diffusion of web technologies [@PAPAGIANNIDIS2015308], 
+social media [@lengyel2020role] and online shopping [@bakher2013diffusion]. 
+Similar to @beardsell1999spatial, who analysed the spatial evolution of 
+computer industry employment across 317 US metro areas, spatial heterogeneity
+is evident in the diffusion speed when the scale of analysis is as detailed
+as the LAD. Importantly, no previous studies estimated such cumulative 
+adoption curves for areal units as small as the UK LADS.
+
 <!-- rankings -->
-The next step is to assess the stability and volatility of the LAD in terms of 
-their adoption of web technologies. As we know from the literature [@risk_perceptions],
-different agents have different perceptions about and levels of acceptance of the 
-risks and the potential economic returns associated with the adoption of new technologies 
--- see for instance the seminal work of @venkatesh2000theoretical. To reveal such 
-aggregated patterns, Figures \ref{rank_stable} and \ref{rank_unstable} plot the ranking
-of UK LAD based on website density. To decrease noise, the average ranking 
-of 1996-1998 and 2010-2012 is plotted instead of the individual years. 
-While there are quite a few obvious cases of LAD that maintained their position 
-between the beginning and the end of the study period either at the top or at the 
-bottom of the hierarchy (Figure \ref{rank_stable}), there are also 
-quite a few LAD that changed drastically their position. Some of these LAD enjoyed
-a process that at the first instance looks like leapfrogging since 
-they managed to jump at the top of the hierarchy despite their slow start (top of 
-Figure \ref{unstable}). There is extensive literature regarding the potential benefits of
-technological leapfrogging. The underpinning argument is that latecomers can adopt
-and benefit from new technologies that have been developed elsewhere without incurring 
-the hefty initial R&D costs [@teece2008firm]. Although the leapfrogging literature
-does not pay much attention to cities and regions [@yu2018sustainability], previous 
-research highlighted the long term and sustained productivity benefit of the early
-adopters of web technologies in the UK [@tranosuk]. So, although the LAD system 
-is volatile, it is not clear whether the LAD with high concentration of late adopters 
-will gain any latecomer benefits in a way similar to countries experiencing such
-technological leapfrogging.
-
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank_stable}Dynamics of wed diffusion: stability of LAD"}
-path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only05_av.png")
+The next step is assesses the stability and volatility of the LAD in terms of 
+their adoption of web technologies. As highlighted in the literature 
+[@risk_perceptions], different agents have different perceptions about and 
+levels of acceptance of the risks and the potential economic returns associated 
+with the adoption of new technologies -- see for instance the seminal work of
+@venkatesh2000theoretical. To reveal such aggregated patterns, Figure \ref{rank} 
+plots the ranking of UK LAD based on website density. To decrease noise, the 
+average ranking of 1996-1998 and 2010-2012 is plotted instead of the individual 
+years. The colours depict the 5th and 95th percentile of the absolute 
+difference of the LAD ranking between 1996-1998 and 2010-2012. While there are 
+quite a few obvious cases of LAD that maintained their position between the 
+beginning and the end of the study period both at the top or at the bottom of 
+the hierarchy, there are also quite a few LAD that changed drastically 
+their position. Some of these LAD enjoyed a process that at the first instance 
+looks like *leapfrogging* since they managed to jump at the top of the hierarchy 
+despite their slow start. There is extensive literature regarding the potential 
+benefits of technological leapfrogging. The underpinning argument is that 
+latecomers can adopt and benefit from new technologies that have been developed 
+elsewhere without incurring the hefty initial R&D costs [@teece2008firm]. 
+Although the leapfrogging literature does not pay much attention to cities and 
+regions [@yu2018sustainability], previous research highlighted the long term 
+and sustained productivity benefit of the early adopters of web technologies 
+in the UK [@tranosuk]. So, although the LAD system is volatile, it is not clear 
+whether the LAD with high concentration of late adopters will gain any latecomer 
+benefits in a way similar to countries experiencing such technological 
+leapfrogging. Still, the opportunity to catch up that @perkins2005international
+illustrated at the country level, is also visible at the much more detail LAD 
+scale. As previously, this is consistent with including websites with up 
+to 10 postcodes (see Figure \ref{rank10}).
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank}Dynamics of wed diffusion (LAD)"}
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av.png")
 knitr::include_graphics(path.image)
@@
 
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank_unstable}Dynamics of wed diffusion: volatility LAD"}
-path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only95_av.png")
-knitr::include_graphics(path.image)
-```
+## Exploratory analysis
 
 <!-- Neighbourhood effect -->
-
-Then, figures \ref{morani} and \ref{lisa} offer a first insight into whether a neighbourhood 
-effect underpins the diffusion of web technologies in the UK as they plot the Moran's I
-and the LISA maps of website density respectively for LAD and OA. Starting from 
-the former, spatial autocorrelation was higher in the beginning and then around 
-year 2000 dropped slightly and stabled at 0.2. This reflects the early concentration 
-of high website density around London, which over time diffused as high-high clusters
-can be seen in other parts of the country away from London as seen in Figure \ref{lisa}. An 
-almost reverse pattern can be observed for OA. At the beginning of the study period
-Moran's I was around 0.1 and it plateaued after 2000-2001 around 0.2. Because of the 
-very small size of OA, at the early stages of the diffusion of web technologies 
-their adoption was spatially scattered. This is reflected in the lack of any significant
-clusters in 1996. Eventually, as the adoption rate increased, more such clusters 
-of high website density were formed and this is reflected both in the Moran's I 
-and the LISA maps in Figures \ref{morani} and \ref{lisa}. All in all, the exploratory 
-spatial data analysis advocates towards an underpinning neighbourhood effect and 
-the different scales of analysis illustrate how it evolved differently over time.
+Figures \ref{morani} and \ref{lisa} offer a first insight into whether a 
+neighbourhood mechanism underpins the diffusion of the Web in the UK as they 
+plot the Moran's I and the LISA maps of website density respectively for LAD 
+and OA. Starting from the former, spatial autocorrelation was higher in the 
+beginning of the study period and then after 2000 dropped slightly and 
+stabilised around 0.2. This reflects the early concentration of high website 
+density around London, which over time diffused as high-high clusters can be 
+seen in other parts of the country away from London (Figure \ref{lisa}). An 
+almost reverse pattern can be observed for OA. At the beginning of the study 
+period Moran's I was around 0.5 and it plateaued after 2000 around 0.2. 
+Because of the very small size of OA, at the early stages of the diffusion 
+of web technologies their adoption was spatially scattered. This is reflected in 
+the lack of any significant clusters in 1996 (Figure \ref{lisa}). Eventually, as 
+the adoption rate increased, more such clusters of high website density were 
+formed and this is reflected both in the Moran's I and the LISA maps in Figures 
+\ref{morani} and \ref{lisa}. A similar pattern is observed for the extended 
+dataset (see Figures\ref{morani10} and \ref{lisa10}). In this case, the 
+magnitude of spatial autocorrelation is slightly higher illustrating a stronger 
+spatial clustering mechanism. All in all, the exploratory spatial data analysis 
+advocates towards an underpinning neighbourhood mechanism and the different 
+scales of analysis illustrate how it evolved differently over time. 
+<!-- discussion -->
+Interestingly, the magnitude of Moran's I is quite similar to the ones revealed
+by @ding2010modeling about mobile phone adoption in Chine albeit the much more 
+detailed spatial scale of this paper.
 
@@
                      limits = force) +
-  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) 
-```
-
-Figure \ref{lisa}
+  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
+  theme(legend.position="bottom",
+        text = element_text(size = 7),
+        legend.text=element_text(size=7),
+        legend.title=element_text(size=7),
+        axis.title=element_text(size=7),
+        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
+```
 
@@
 p + inset_element(p7, left = 0.29, bottom = 0.8, right = 0.49, top = 1.2)
-
-```
-
-To illustrate whether a hierarchical process underpins the diffusion of web technologies,
-the Gini coefficient is calculated yearly both for LAD and OA. As a metric of 
-inequality, the Gini coefficient demonstrates whether website density is concentrated
-in a small number of LAD/OA, or whether it is more equally spread across the country.
-Both scales of analysis in Figure \ref{gini} illustrate the same picture. At the 
-beginning of the commercial internet website density was extremely unequal, or in
-other words, only a few places had websites associated with them. Inequality dropped 
-and plateaued after 2000 for both scales illustrating that at the first stages of 
-the commercial internet, website density was concentrated in a limited number of 
-places. This is an indication of a hierarchical diffusion mechanism that led over 
-time to a more equal spread. Interestingly, the year 2000 is again a period of change
-for this diffusion mechanisms as it was for the neighbourhood process. 
-There is a substantial difference between the Gini coefficient magnitude for LAD
-and OA, but this is well expected as the very small size of OA equates to a lot 
-polygons without any websites pointing to them -- for instance residential OA. 
-LAD/OA
+```
+
+To illustrate whether a hierarchical process also underpins the diffusion of web 
+technologies, the Gini coefficient is calculated yearly both for LAD and OA. As 
+a metric of inequality, the Gini coefficient demonstrates whether website 
+density is concentrated in a small number of LAD or OA, or whether it is more 
+equally spread across the country. Both scales of analysis in Figure \ref{gini} 
+illustrate the same picture. At the beginning of the commercial internet website 
+density was extremely unequal, or, in other words, only a few places had 
+websites associated with them. Inequality dropped and plateaued after 2000 for 
+both scales. This is illustrative of a hierarchical diffusion mechanism that 
+led over time to a more equal spread. Interestingly, the year 2000 is again a 
+period of change for this diffusion mechanism as it was for the neighbourhood 
+process. There is a substantial difference between the Gini coefficient 
+magnitude for LAD and OA, but this is expected as the very small size of OA 
+equates to a lot of polygons without any websites pointing to them -- for 
+example residential OA.
 
@@
 
-path.in.la <- paste0(path, "/outputs/lisa/gini_la.csv")
-path.in.oa <- paste0(path, "/outputs/lisa/gini_oa.csv")
-
-# path.in <- paste0(path, "/outputs/lisa/gini_la.csv")
-# path.in <- paste0(path, "/outputs/lisa/gini_oa.csv")
+path.in.la <- paste0(path, "/outputs/gini/gini_la.csv")
+path.in.oa <- paste0(path, "/outputs/gini/gini_oa.csv")
 
@@
                      limits = force) +
-  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) 
-```
-
-<!-- RF results -->
-
-The next section incorporates the above discussed spatial processes of the
-diffusion of web technologies into a modelling framework. The aim
-is to use variables depicting these spatial processes in order to
-predict the diffusion of web technologies in the UK over space and time
-and across different scales. Specifically, four different models are
-estimated. Firstly, all the data points for the OA and LAD are utilised
-in order to build two respective RF models and assess their capacity to predict the
-adoption of web technologies. These two models will reveal the
-predictive capacity of the diffusion mechanisms and also allow to see how
-the importance of such variables changes across scales. The next two
-sets of models will be again trained on web diffusion at the two working scales:
+  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
+  theme(legend.position="bottom",
+        text = element_text(size = 7),
+        legend.text=element_text(size=7),
+        legend.title=element_text(size=7),
+        axis.title=element_text(size=7),
+        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
+```
+
+## Modelling results
+
+The next section incorporates the above mechanisms of the spatial
+diffusion of the Web into a modelling framework. The aim is to use variables 
+depicting these spatial processes in order to predict the diffusion of the Web 
+as a new technology in the UK over space and time and across different scales. 
+Specifically, four different models are estimated. Firstly, all the data points 
+for the OA and LAD are utilised in order to build two RF models and assess their 
+capacity to predict the adoption of web technologies at the two different scales. 
+These two models will reveal the predictive capacity of the diffusion mechanisms 
+and also demonstrate how the importance of such variables changes across different 
+scales. The next two sets of models are again trained at these two scales: 
 OA and LAD. However, instead of using all the data points, the OA and
-the LAD from one of the twelve UK regions are held out and then the trained
-model is used to predict website density in the OA or the LAD of the
-held-out region. This process takes place recursively for all UK
-regions. The difference in the predictive capacity of the different
-samples will reveal how dissimilar are these spatial process across
-regions and, importantly, at different scales.
+the LAD from one of the twelve UK regions are held out for the model training.
+Then the trained model is used to predict website density for the OA or the LAD 
+of the held-out region. This process takes place recursively for all twelve UK
+regions. The regional differences in the predictive capacity of the different
+samples will reveal how dissimilar these spatial process are across regions 
+and, importantly, at different scales.
 
@@
 is spatially and temporally sensitive. Instead of using
-10 random folds, we employ the `CAST` package which allows holding back
-data points from specific years and spatial units and use them for
-testing in order to estimate the model performance
+10 random and space- and time-agnostic folds,the `CAST` package is employed 
+which allows holding back data points from specific years and spatial units 
+and use them for testing in order to estimate the model performance 
 [@meyer2018improving].
 
-The models include variables that capture the three processes
-that the relevant literature and the descriptive analysis presented above highlighted. 
-Namely, the models capture: (i) a hierarchy effect with diffusion running from main 
-centres to secondary ones, (ii) a neighborhood effect according to which diffusion 
-first hits nearby locations, and (iii) the rather canonical pattern of diffusion 
-over time as reflected in the S-shaped pattern in the cumulative level of adoption.
+The models include variables that capture the three processes that the relevant 
+literature and the descriptive analysis highlighted. Namely, the models capture: 
+(i) a hierarchy mechanism with diffusion running from main centres to secondary 
+ones, (ii) a neighborhood mechanism according to which diffusion first hits 
+nearby locations, and (iii) the rather canonical pattern of diffusion over time 
+as reflected in the S-shaped pattern in the cumulative level of adoption.
 
@@
 adoption in a spatial unit depends on the level of the adoption in
-places further up in the urban hierarchy the previous year. The inclusion of the
-distance variables incorporates spatial structure into the hierarchy argument.
-To depict the neighbourhood effect, the website density of the neighbouring spatial
-units in the previous year is employed. Again, the underpinning rationale is that 
-the level of web adoption within a spatial unit depends on the level of web adoption 
-in the neighbouring spatial units the year before. This represents the 'hitting 
-nearby locations first' argument. Therefore, the spatial and temporal lag of the 
-website density in LAD and OA is calculated. Lastly, the time effect which is reflected 
-on the S-curve for the cumulative adoption is captured by a time trend variable.
-Hence, all four model will follow the following generic form 
-(Eq. \ref{model}):
+places further up in the urban hierarchy the previous year. The inclusion of
+the distance variables incorporates spatial structure into the hierarchy 
+argument. To depict the neighbourhood effect, the website density of the 
+neighbouring spatial units in the previous year is employed (rook continuity). 
+Again, the underpinning rationale is that the level of web adoption within a 
+spatial unit depends on the level of web adoption in the neighbouring spatial 
+units the year before. This represents the 'hitting nearby locations first' 
+argument. Therefore, the spatial and temporal lag of the website density in 
+LAD and OA is calculated. Lastly, the temporal pattern of the cumulative 
+adoption is captured by a time trend variable. Hence, all four models will 
+follow the following generic form (Eq. \ref{model}):
 
@@
 
-$y_{k}$ is the $k^{th}$ observation of the dataset, which consists of
+\noindent $y_{k}$ is the $k^{th}$ observation of the dataset, which consists of
 $N$ observations in total. $\hat{y_{k}}$ is the $k_{th}$ predicted value
@@
 
-@tbl-model-metrics presents the model performance for the
-first set of models, for which all data points are employed for training
-and testing via cross validation. The first one is trained and tested on 374 LAD
-and the second on 232,296 OA, both for a 16 year period (1997-2012). The results are 
-remarkably good considering that the are the outcome of space and time sensitive 
-CV, so the the model does not suffer from overfitting. At the LAD level the model 
-predicts 81% of the variation of website density. Both error metrics indicate that
-the model error is $(\frac{1}{20}, \frac{1}{30})$ of a website per firm. At 
-the OA the $R^2$ drops down to 21%. Considering its granularity, this is still 
-a remarkable performance. To contextualise it, the model results in a MAE of one 
-website for areas small enough to host less than 140 households.^[According to the Office for
-National Statistics, 80% of OA in England and Wales host 110-139 households, [www.ons.gov.uk](https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas).] 
+@tbl-model-metrics presents the model performance for the first set of models, 
+for which all data points are employed for training and testing via cross 
+validation. The first one is trained and tested on 374 LAD and the second on 
+232,296 OA, both for a 16 year period (1997-2012). The results are remarkably 
+good considering that the are the outcome of space and time sensitive 
+CV, so the the model does not suffer from overfitting. At the LAD level the 
+model predicts 85% of the variation of website density. Both error metrics 
+indicate that the model error is $(\frac{1}{20}, \frac{1}{30})$ of a website per 
+firm. At the OA the $R^2$ drops down to 32%. Considering its granularity, this 
+is still a remarkable performance. To contextualise it, the model results in a 
+$MAE$ equal to one website for areas small enough to host less than 140 
+households.
+^[According to the Office for National Statistics, 80% of OA in England and 
+Wales host 110-139 households,
+[www.ons.gov.uk](https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas).] 
 Because of the small size of the spatial units, the distribution is highly skewed
 and a significant part of them is not linked to any websites. In 1997 only 1% of 
-the UK OA were associated with at least one website. This should not come as a surprise as 
-this was the very beginning of the commercial internet and any activities with a 
-digital footprint were concentrated in a handful of areas. This was clearly illustrated
-in Figure \ref{lisa}. At the end of the study period almost half of the UK OA were
-not associated with a website. Again, given the granularity of the data this should
-not come as a surprise.
-
-|                   | RMSE  | $R^{2}$   | MAE   |
-|-------------------|-------|-----------|-------|
+the UK OA were associated with at least one website. This should not come as a 
+surprise as this was the very beginning of the commercial internet and any 
+activities with a digital footprint were concentrated in a handful of areas. 
+This was clearly illustrated in Figure \ref{lisa}. At the end of the study 
+period almost half of the UK OA were not associated with a website. Again, 
+given the granularity of the data this should not come as a surprise.
+
+|                   | $RMSE$ | $R^{2}$ | $MAE$ |
+|-------------------|--------|---------|-------|
+| Local Authorities | 0.028 | 0.850    | 0.019 |
+| Output Areas      | 3.284 | 0.320    | 1.034 |
+
+: Model metrics {#tbl-model-metrics}
+
+The results in @tbl-model-metrics are robust against different specification. 
+As noted earlier, the data imputation process resulted in a modest improvement 
+of the predictive capacity of the model. As per
+@tbl-model-metrics-no-correction all accuracy metrics are only slightly worst 
+when the data imputation is not applied. As expected, the improvement is 
+higher for the more detailed scale of analysis. Then, @tbl-model-metrics-10 
+presents the model results for the extended dataset, which includes websites 
+with up to 10 unique postcodes. While the results for the LAD are slightly 
+less accurate, the $R^{2}$ of the OA model is more than double in comparison 
+to the base results in @tbl-model-metrics. All in all, the modelling exercise
+revealed the how well can the three established diffusion mechanisms predict 
+the diffusion of the Web as a new technology. Importantly, this predictive
+capacity is highly robust against against different scales and delineations
+of the diffusion of the web.
+
+|                   | $RMSE$ | $R^{2}$ | $MAE$ |
+|-------------------|--------|---------|-------|
 | Local Authorities | 0.032 | 0.810    | 0.019 |
@@
 
-: Model metrics {#tbl-model-metrics}
-
-Figures \ref{var.imp.LAD} and \ref{var.imp.OA} plot the importance of the different 
-predictors. When the focus is on the LAD, the website density in the nearest city,
-in London and in the neighbouring LAD the year before are the most important predictors.
-They are followed by the yearly trend, while the spatial configuration as reflected
-in distances to London or the nearest city only play a minor role. This can be attributed
-to the rather coarse spatial scale of analysis. Nevertheless, all previously discussed
-spatial processes are at play in the diffusion of web technologies at the LAD level: 
-the first two predictors depict the hierarchical effect, the spatial and temporal lag of website
-density depict the neighbourhood effect and the yearly trend the time-sensitive 
-cumulative adoption pattern.
-
-When the much more granular scale of OA is adopted, the picture is reversed. The 
-most important predictors are the three distance variables to London, the nearest 
-city and the nearest retail centre. They stil depict the hierarchical effect, but
-proximity to the different population centres is more important than their lagged 
-web densities in predicting website diffusion. The neighbouring effect is less 
-important at this scale. What is interesting is the almost negligible role of the 
-yearly trend and London's website density. While the former probably illustrates 
-the large heterogeneity in how web technologies have been adopted at this very fine 
-scale, the later highlights that the importance of past web adoption rates in large 
-population centres is surpassed by proximity to them and spatial configuration 
-at this scale.
-
-```{r varimpLA, eval = T, echo=FALSE, message = F, fig.cap="\\label{var.imp.LAD}Variable importance, LAD"}
-path.image <- paste0(path, "/outputs/rf/figures/varimp_LA.png")
+: Model metrics without data imputation {#tbl-model-metrics-no-correction}
+
+|                   | $RMSE$ | $R^{2}$ | $MAE$ |
+|-------------------|--------|---------|-------|
+| Local Authorities | 0.188  | 0.77    | 0.122 |
+| Output Areas      | 22.467 | 0.478   | 5.681 |
+
+: Model metrics for the extended dataset {#tbl-model-metrics-10}
+
+Figure \ref{var.imp} plots the importance of the different predictors. When the 
+focus is on LAD, the website density in the nearest city, in London and in the 
+neighbouring LAD the year before are the most important predictors. They are 
+followed by the yearly trend, while the spatial configuration as reflected
+in distances to London or the nearest city only plays a minor role. This can be 
+attributed to the relatively coarse spatial scale of LAD. Nevertheless, all 
+previously discussed spatial processes are at play in the diffusion of web 
+technologies at the LAD level: the first two predictors depict the hierarchy 
+mechanism, the spatial and temporal lag of website density the neighbourhood 
+mechanism, and the yearly trend the time-sensitive cumulative adoption pattern.
+
+When the much more granular scale of OA is adopted, the picture is almost 
+reversed. The two most important predictors are the distance to London and to 
+the nearest city followed by the spatial and temporal lag of website density 
+and the distance to the nearest retail centre. They still depict a hierarchical 
+mechanism, but proximity to the different population centres is more important 
+than their lagged web densities in predicting website diffusion. The 
+neighbourhood mechanism is still strong but less important at this scale. What 
+is interesting is the almost negligible role of the yearly trend and London's 
+website density. While the former probably illustrates the large heterogeneity 
+in how the Web has been adopted at this very fine scale, the latter highlights 
+that the importance of past web adoption rates in large population centres is 
+surpassed by proximity to them and spatial configuration at this scale.
+
+The above pattern of variable importance is largely in accordance with the models
+estimated for the extended dataset. A comparison with Figure \ref{var.imp10} 
+illustrates that the main difference is the importance of the neighbourhood
+mechanism for the OA as the space and time lag of the website density is by far 
+the most important predictor when the extended dataset is employed. But still, 
+the variable importance patterns for the two datasets are not dissimilar.
+
+```{r varimpLA, eval = F, echo=FALSE, message = F, fig.cap="\\label{var.imp.LAD}Variable importance, LAD"}
+path.image <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.png")
 knitr::include_graphics(path.image)
@@
 
-```{r varimpOA, eval = T, echo=FALSE, message = F, fig.cap="\\label{var.imp.OA}Variable importance, OA"}
-path.image <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
+```{r varimpOA, eval = F, echo=FALSE, message = F, fig.cap="\\label{var.imp.OA}Variable importance, OA"}
+path.image <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.png")
 knitr::include_graphics(path.image)
@@
 
-Figure \ref{var.imp} to replace Figures \ref{var.imp.LAD} and \ref{var.imp.OA} once
-I have the variable importance for LAD.
-
-```{r varimp, eval = T, echo=FALSE, message = F, fig.cap="\\label{var.imp}Variable importance"}
+```{r varimp, eval = T, echo=FALSE, message = F, warning=FALSE, fig.cap="\\label{var.imp}Variable importance", out.height="60%"}
 path.importance.lad <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.csv")
@@
 
-importance.oa <- read_csv(path.importance.oa)
-importance.lad <- read_csv(path.importance.oa) %>% 
-  mutate(importance = ifelse(variable == "year", 10, 
-                             ifelse(variable == "distance to London", 75, importance)))
+importance.oa <- read_csv(path.importance.oa) %>% 
+  mutate(id = c("LD", "NC", "NR", "STL", "Y", "DL", "DNC", "DNR")) %>% 
+  mutate(variable = str_to_sentence(variable))
+importance.lad <- read_csv(path.importance.lad) %>% 
+  mutate(id = c("LD", "NC", "STL", "Y", "DL", "DNC")) %>% 
+  mutate(variable = str_to_sentence(variable))
 
@@
   rename(OA = importance) %>% 
-  left_join(importance.lad %>% rename(LAD = importance), by = "variable") %>%
+  left_join(importance.lad %>% rename(LAD = importance), by = "id") %>%
+  dplyr::select(-variable.y, -id) %>% 
+  rename(variable = variable.x) %>% 
   pivot_longer(!variable, names_to = "spatial.unit", values_to = "importance") %>% 
-  # reorder with fct_reorder
-  ggplot(aes(x=importance, y=forcats::fct_reorder(variable, importance), 
+  mutate(order = ifelse(variable == "Distance to london", 1,
+                        ifelse(variable == "London's wensite density, t-1", 2,
+                               ifelse(variable == "Distance to the nearest city", 3,
+                                      ifelse(variable == "Nearest city's wensite density, t-1", 4,
+                                             ifelse(variable == "Distance to the nearest retail centre", 5,
+                                                    ifelse(variable == "Nearest retail centre's wensite density, t-1", 6,
+                                                           ifelse(variable == "Spatial and temporal lag of wensite density", 7, 8)))))))) %>% 
+  ggplot(aes(x=importance, y=reorder(variable, desc(order)), #forcats::fct_reorder(variable, importance), 
          fill = spatial.unit)) +
@@
         axis.text.x = element_text(vjust = 0.5, hjust=1)) +
-  scale_fill_manual('Spatial unit', 
+  scale_fill_manual('Spatial unit: ', 
                     values = c("LAD" = "grey80",
@@
                      limits = force) +
-  scale_x_continuous('Variable importance')
-
+  scale_x_continuous('Variable importance') +
+  theme(legend.position="bottom",
+        text = element_text(size = 7),
+        legend.text=element_text(size=7),
+        legend.title=element_text(size=7),
+        axis.title=element_text(size=7))
 ```
@@
 ```{r regions}
-path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
-path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
+path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected.csv")
+path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA_corrected.csv")
 
@@
 
-# Appendices {.appendix}
+# Appendix {#appendix}
+
+<!-- #sec-appendix -->
+<!-- .appendix -->
 
@@
 
-## $t_0$ estunates for all LAD
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk10}Grwoth curve, UK; up to 10 postcodes per website"}
+path.image <- paste0(path, "/outputs/s/s_uk_per_firm.png")
+knitr::include_graphics(path.image)
+```
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map10}LAD adoption rates; up to 10 postcodes per website"}
+path.image <- paste0(path, "/outputs/s/speed_map_10.png")
+knitr::include_graphics(path.image)
+```
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank10}Dynamics of wed diffusion (LAD); up to 10 postcodes per website"}
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av_10.png")
+knitr::include_graphics(path.image)
+```
+
+```{r, morani10, echo=FALSE, message=FALSE, fig.cap="\\label{morani10}Website density Moran's I; up to 10 postcodes per website"}
+
+path.in.la <- paste0(path, "/outputs/lisa/corrected/LA/morani_la_10.csv")
+path.in.oa <- paste0(path, "/outputs/lisa/corrected/oa/morani_oa_10.csv")
+
+la.m <- read_csv(path.in.la) %>% 
+  filter(p < 0.001) %>% # all values were significant
+  mutate(spatial.unit = 'LAD')
+oa.m <- read_csv(path.in.oa) %>% 
+  filter(p < 0.001) %>% # all values were significant
+  mutate(spatial.unit = 'OA')
+
+bind_rows(la.m, oa.m) %>% 
+  ggplot(aes(x=year, y=morani, fill= spatial.unit)) +
+  geom_bar(stat = 'identity', position = 'dodge') +
+  ylab('Moran\'s I') +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_fill_manual('Spatial unit', 
+                    values = c("LAD" = "grey80",
+                                "OA"="grey48"),
+                     limits = force) +
+  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
+  theme(legend.position="bottom",
+        text = element_text(size = 7),
+        legend.text=element_text(size=7),
+        legend.title=element_text(size=7),
+        axis.title=element_text(size=7),
+        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
+```
+
+```{r, echo=FALSE, message=FALSE, fig.cap="\\label{lisa10}\\centering Website density LISA maps; up to 10 postcodes per website", out.height="80%"}
+
+path.image1 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la1996.png")
+path.image2 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la2000.png")
+path.image3 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la2012.png")
+path.image4 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa1996.png")
+path.image5 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa2000.png")
+path.image6 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa2012.png")
+path.image7 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la_legend.png")
+
+p1 <- ggdraw() + draw_image(path.image1)  
+p2 <- ggdraw() + draw_image(path.image2)  
+p3 <- ggdraw() + draw_image(path.image3)  
+p4 <- ggdraw() + draw_image(path.image4) 
+p5 <- ggdraw() + draw_image(path.image5)
+p6 <- ggdraw() + draw_image(path.image6)  
+p7 <- ggdraw() + draw_image(path.image7)
+
+par(mar=rep(0,4)) # no margins
+
+p <- p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 +
+  plot_layout(ncol = 3, widths = c(1, -0.8, 1), heights = c(1,1,1))
+
+# p + inset_element(p7, left = 0.45, bottom = -0.5, right = 0.6, top = 0.1)
+p + inset_element(p7, left = 0.29, bottom = 0.8, right = 0.49, top = 1.2)
+
+```
+
+```{r, gini10, echo=FALSE, message=FALSE, fig.cap="\\label{gini10}Website density Gini coefficient; up to 10 postcodes per website"}
+
+path.in.la <- paste0(path, "/outputs/gini/gini_la_10.csv")
+path.in.oa <- paste0(path, "/outputs/gini/gini_oa_10.csv")
+
+la.gini <- read_csv(path.in.la) %>% 
+  mutate(spatial.unit = 'LAD')
+oa.gini <- read_csv(path.in.oa) %>% 
+  mutate(spatial.unit = 'OA')
+
+bind_rows(la.gini, oa.gini) %>%
+  rename(Gini = 'gini(a$n)') %>% 
+  ggplot(aes(x=year, y=Gini, fill= spatial.unit)) +
+  geom_bar(stat = 'identity', position = 'dodge') +
+  ylab('Gini') +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_fill_manual('Spatial unit', 
+                    values = c("LAD" = "grey80",
+                                "OA"="grey48"),
+                     limits = force) +
+  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
+  theme(legend.position="bottom",
+        text = element_text(size = 7),
+        legend.text=element_text(size=7),
+        legend.title=element_text(size=7),
+        axis.title=element_text(size=7),
+        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
+```
+
+```{r varimp10, eval = T, echo=FALSE, message = F, warning=FALSE, fig.cap="\\label{var.imp10}Variable importance; up to 10 postcodes per website", out.height="60%"}
+path.importance.lad <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected_10.csv")
+path.importance.oa <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected_10.csv")
+
+importance.oa <- read_csv(path.importance.oa) %>% 
+  mutate(id = c("LD", "NC", "NR", "STL", "Y", "DL", "DNC", "DNR")) %>% 
+  mutate(variable = str_to_sentence(variable))
+importance.lad <- read_csv(path.importance.lad) %>% 
+  mutate(id = c("LD", "NC", "STL", "Y", "DL", "DNC")) %>% 
+  mutate(variable = str_to_sentence(variable))
+
+importance.oa %>% 
+  rename(OA = importance) %>% 
+  left_join(importance.lad %>% rename(LAD = importance), by = "id") %>%
+  dplyr::select(-variable.y, -id) %>% 
+  rename(variable = variable.x) %>% 
+  pivot_longer(!variable, names_to = "spatial.unit", values_to = "importance") %>% 
+  mutate(order = ifelse(variable == "Distance to london", 1,
+                        ifelse(variable == "London's wensite density, t-1", 2,
+                               ifelse(variable == "Distance to the nearest city", 3,
+                                      ifelse(variable == "Nearest city's wensite density, t-1", 4,
+                                             ifelse(variable == "Distance to the nearest retail centre", 5,
+                                                    ifelse(variable == "Nearest retail centre's wensite density, t-1", 6,
+                                                           ifelse(variable == "Spatial and temporal lag of wensite density", 7, 8)))))))) %>% 
+  ggplot(aes(x=importance, y=reorder(variable, desc(order)), #forcats::fct_reorder(variable, importance), 
+         fill = spatial.unit)) +
+  geom_col(position = 'dodge') +
+  ylab('') +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text(vjust = 0.5, hjust=1)) +
+  scale_fill_manual('Spatial unit: ', 
+                    values = c("LAD" = "grey80",
+                                "OA"="grey48"),
+                     limits = force) +
+  scale_x_continuous('Variable importance') +
+  theme(legend.position="bottom",
+        text = element_text(size = 7),
+        legend.text=element_text(size=7),
+        legend.title=element_text(size=7),
+        axis.title=element_text(size=7))
+```
+
+```{r regions_no_imp}
+path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
+path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
+
+lad <- read.csv(path.lad) %>% 
+  rename('RSquared LAD' = Rsquared)
+
+oa <- read.csv(path.oa) %>% rename(Region = test.region,
+                                   'RSquared OA' = Rsquared)
+
+lad %>% left_join(oa) %>% 
+  mutate('Rank OA' = rank(desc(`RSquared OA`)),
+         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
+  relocate(5, .after = 2) %>% 
+  arrange(`Rank LAD`) %>% 
+  kable(caption= "Regional differences; up to 10 postcodes per website\\label{table.regions.no.imp}", 
+        digits=3,
+        col.names = c("Region", "$R^2$ LAD", "Rank LAD", "$R^2$ OA", "Rank OA")) 
+```
 
@@
 
+# regions
+path.regions <- paste0(path, "/data/raw/LAD21_RGN21_EN_LU.csv")
+regions <- read_csv(path.regions)
+
 s.la.path <- paste0(path, "/data/temp/s_la_per_firm.csv")
+
 read.csv(s.la.path) %>% 
@@
   arrange(estimate) %>%
-  select(LAD21NM, estimate, std.error, r2, fast) %>% 
+  left_join(regions %>% select(LAD21CD, RGN21NM), by = c("ladcd" = "LAD21CD")) %>% 
+  select(LAD21NM, RGN21NM, estimate, std.error, r2, fast) %>% 
   rename(LAD = LAD21NM,
+         Region = RGN21NM,
          "t_0 estimate" = "estimate",
@@
          "Diffusion speed" = "fast") %>% 
-  kable(caption= "S-curve estiamtes for LAD\\label{table.s.lads}", 
+  kable(caption= "S-curve estiamtes for LAD; up to 10 postcodes per website\\label{table.s.lads}", 
         digits=3,
-        align = c("l", "c", "c", "c", "c"),
-        col.names = c("LAD", "$t_0$ estimate", "Std. error", "$R^2$", "Diffusion speed"))
+        align = c("l", "l", "c", "c", "c", "c"),
+        col.names = c("LAD", "Region", "$t_0$ estimate", "Std. error", "$R^2$", "Diffusion speed"))
 ```

### 2024-09-12 paper/submission1/tranos2023.tex
postpresentation2

--- 2024-06-18 paper/submission1/tranos2023.tex
@@
 
-\affiliation[1]{organization={University of Bristol and The Alan Turing
-Institute},postcode={UK},postcodesep={}}
+\affiliation[1]{organization={University of
+Bristol},postcode={UK},postcodesep={}}
 
@@
 
-\section{Introduction}\label{sec1}
-
-Geographers were always interested in how new technologies and
-innovations diffuse across space and time and, importantly, how such
-spatio-temporal processes can be modeled. After all, diffusion together
-with invention and innovation are considered the pillars of
-technological change \citep{das2022diffusion}. The seminal contribution
-of \citet{hagerstrand1968innovation} is illustrative of this early
-interest. However, the torch of exploring and modelling such processes
-had been passed to other disciplines such as economics, business studies
-and sociology well before the `cultural turn' of economic geography
-\citep{perkins2005international}. A potential explanation of the lack of
-geographical studies exploring the diffusion of new and, more
+\section{Introduction}\label{sec-introduction}
+
+Economic geographers were always interested in how new technologies and
+innovations diffuse across space and time with the work of
+\citet{hagerstrand1968innovation}, which demonstrated the importance of
+communications as a mechanism for the diffusion of innovations, being
+the most impactful. It was followed by various empirical studies -- see
+for instance \citet{ormrod1990} and \citet{iso2005}. However, the torch
+of exploring and modelling such diffusion processes had been passed to
+other disciplines such as economics, business studies and sociology well
+before the `cultural turn' of economic geography
+\citep{perkins2005international}. A potential explanation of the recent
+lack of geographical studies exploring the diffusion of new and, more
 specifically for this paper, digital technologies across both space and
 time can be attributed to the scarcity of relevant and granular enough
-data. As \citet{zook2022mapping} highlight, digital activities are
-hardly ever captured in official data.
+data, a problem also highlighted by \citet{iso2005} and
+\citet{kemeny2011international}. As \citet{zook2022mapping} pointed,
+digital activities are hardly ever captured in official data.
 
 This paper offers such a contribution: a geographical study illustrating
-how a new technology that is the web diffused over space and time in the
-UK at a high level of spatial granularity during the \(1996-2012\)
-period. It does so by employing a novel source of big data which
-captures the active engagement with web technologies during that period.
-By addressing this empirical question this paper exemplifies how the
-combination of data sources which escape the traditional social science
-domain and adequate research methods can offer new lenses to
-geographical research regarding the understanding of technological
-diffusion.
-
-\textbf{SAY SOMETHING ABOUT SPATIAL ANALYTICS, DIST PAPER BY ARP, NOT
-SPECIFIC EXPLANATORY VARIABLES, INTANGIBLE}
+how a new technology that is the Web diffused over space and time in the
+UK at a high level of spatial granularity during the 1996-2012 period.
+It does so by employing a novel source of big data and machine learning
+algorithms to model the \emph{active} engagement with Web during that
+period and tests how different spatial and temporal mechanisms shaped
+its diffusion. This paper exemplifies how the combination of data
+sources which escape the traditional social science domain and adequate
+research methods can offer new lenses to geographical research regarding
+the understanding of technological diffusion.
 
@@
 diffused over space and time and use this knowledge to make predictions
-regarding the diffusion of related \emph{future} technologies. As per
-\citet{leibowicz2016representing}, historical studies agree on the fact
-that ``technologies diffuse at different times, at different rates, and
-to different extents in different places, and can be significantly
-influenced by policies'' \citep{victor1993}. \citet{meade2021modelling}
-highlight that a variety of actors have a direct interest in gaining
-such knowledge including network equipment suppliers; network operators,
-regulatory and local authorities. These processes and their effects vary
-a lot across scales: although the diffusion of a new technology might
-not be optimal at a local level, it might be beneficial from a global
-perspective as it could lead to faster diffusion to less advantaged
-places \citep{leibowicz2016representing}. Despite the spatial
-heterogeneity of such diffusion mechanisms and the policy relevance,
-there are very limited attempts in the literature to analyse the
+about the diffusion of related \emph{future} technologies. As per
+\citet{leibowicz2016representing}, historical studies agree that
+technologies diffuse differently in terms of times, rates, and
+geographies and can be driven by related policies \citep{victor1993}.
+\citet{meade2021modelling} highlight that a variety of actors have a
+direct interest in gaining such knowledge including network equipment
+suppliers, network operators, regulatory and local authorities. These
+processes and their effects vary a lot across scales: although the
+diffusion of a new technology might not be optimal at a local level, it
+might be beneficial from a global perspective as it could lead to faster
+diffusion to less advantaged places \citep{leibowicz2016representing}.
+Despite the spatial heterogeneity of such diffusion mechanisms and the
+policy relevance, there is a scarcity of empirical studies analysing the
 diffusion of new digital technologies at a detailed geographical level.
+
+This gap becomes even more prominent if we consider the importance of
+technology and its adoption in economic geography and macroeconomic
+growth theories. Simply put, differences in technological adoption and
+sophistication may lead to uneven development as new technologies can
+increase productivity and allow firms to introduce new products
+\citep{solow1957technical, aghion1990model, kemeny2011international}.
+Linked to the subject of this paper, the Web -- and the active
+engagement with this technology that is creating instead of just
+browsing a website -- can be considered as the underpinning technology
+for the transformative process known as the \emph{digital service
+economy} \citep{capello2024nexus}. After all, diffusion together with
+invention and innovation are considered the pillars of technological
+change \citep{das2022diffusion}. Findings from previous empirical work
+illustrated that early engagement with digital technologies such as the
+Web led to longstanding positive productivity gains for these regions
+that hosted early adopters \citep{tranosuk}.
+
+Another line of motivation has to do with the intangible nature of the
+Web and the debate about the footloose nature of digital technologies,
+which contrary to several earlier general purpose technologies are light
+in weight \citep{coyle1999weightless}. Following the seminal work of
+\citet{gaspar1998information}, which demonstrated that technological
+improvements in telecommunications lead to increased demand for
+face-to-face communications and, consequently, enhance the importance of
+cities, empirical studies from economic geography and urban economics
+explored whether digital technologies substitute or complement
+agglomeration externalities and spatial structure
+\citep{kolko_death_2000, ioannides2008effect, tranos2021ubiquitous}.
+With the exception of \citet{sinai2004geography}, who focused, among
+other things, on web content and demonstrated that more such content is
+available in larger markets, this strand of literature has been mostly
+concerned with the spatiality of the internet's hard and tangible
+infrastructure and shied away from exploring the spatial dimensions of
+one of the most distinct software layers of the internet infrastructure,
+the Web.
 
@@
 
-This paper focuses on the diffusion of the web as new technology during
-the \(1996-2012\) period. This was an exciting period for digital
+This paper focuses on the diffusion of the Web as new technology during
+the 1996-2012 period. This was an exciting period for digital
 technologies as it corresponds with the commercialisation of the
@@
 introduced, a year before Microsoft's Internet Explorer.\footnote{\url{https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft}}
-Also, only \(9\) per cent of UK's households had access to the internet
-in \(1998\) \citep{ons2018}, the web included mostly static webpages,
-there were no social media and web browsing was happening exclusively
-from desktop PCs as there were no smartphones \citep{tranosuk}. Hence,
-it is fair to say that the study period captures the very early stages
-of the diffusion on a new technology that is the web as well as its
-maturity. The former is a key point in the lifecycle of a new
-technology. Firstly, during this period new technologies are expensive,
-crude and imperfect \citep{rosenberg1994exploring, wilson201281}. A
-simple comparison between Web 1.0 and Web 3.0 applications clearly
-illustrates this \citep{tranos2020social}; for instance a static website
-compared with a platform like \texttt{github}, which enables cooperation
-between users and the creation of new information, meaning, and
-knowledge \citep{faraj2016special, barassi2012does}. During this period
-the performance of a new technology is the main attraction and not the
-cost to access and use it \citep{wilson2011lessons}. There is a broader
-theoretical discussion in the literature regarding the motivation behind
-early adoption. As summarised by \citet{perkins2005international}, on
-the one hand, epidemic models highlight the role of interpersonal
-contacts as a way for new technologies to diffuse. On the other hand,
-economic models underline the importance of heterogeneity. Different
-firms have different structures and business plans, which define the
-potential economic returns of the adoption of a new technology and,
-therefore, the choice to adopt a new technology becomes an individual
-option. From a broader and evolutionary perspective, initial conditions
-are essential for the creation and evolution of path-dependent
-technological development trajectories
-\citep{neffke2011regions, simmie2014new}. This argument is even more
-relevant when the focus is on digital technologies because of the
-commonly found lag between investment and economic returns as reflected
-in the Solow paradox
+Also, only 9 per cent of UK's households had access to the internet in
+1998 \citep{ons2018}, the web included mostly static webpages, there
+were no social media and web browsing involved exclusively a desktop PCs
+as there were no smartphones \citep{tranosuk}. Hence, it is fair to say
+that the study period captures the very early stages of the diffusion of
+a new technology that is the Web as well until its maturity. The former
+is a key point in the lifecycle of a new technology. Firstly, during
+this period new technologies are expensive, crude and imperfect
+\citep{rosenberg1994exploring, wilson201281}. A simple comparison
+between Web 1.0 and Web 3.0 applications clearly illustrates this
+\citep{tranos2020social}: while a static website facilitates one-way
+information dissemination, a platform like \texttt{github} enables
+cooperation between users and the creation of new information, meaning,
+and even knowledge \citep{faraj2016special, barassi2012does}. During
+this period the performance of a new technology is the main attraction
+and not the cost to access and use it \citep{wilson2011lessons}. There
+is a broader theoretical discussion in the literature about the early
+adoption motives. As summarised by \citet{perkins2005international}, on
+one hand, epidemic models highlight the role of interpersonal contacts
+as a way for new technologies to diffuse. On the other hand, economic
+models underline the importance of heterogeneity. Different firms have
+different structures and business plans, which define the potential
+economic returns of the adoption of a new technology and, therefore, the
+choice to adopt a new technology becomes an individual option. From a
+broader and evolutionary perspective, initial conditions are essential
+for the creation and evolution of path-dependent technological
+development trajectories \citep{neffke2011regions, simmie2014new}. This
+argument is even more relevant when the focus is on digital technologies
+because of the commonly found lag between investment and economic
+returns as reflected in the Solow paradox
 \citep{acemoglu2014return, brynjolfsson2018artificial}.
 
-Importantly, the data used here depicts the \emph{active} engagement
-with the web in the UK as it contains all commercial websites that (i)
-are part of the UK's relevant second level domain (SLD, .co.uk), (ii)
-have been archived by the Internet Archive\footnote{\href{See\%20https://archive.org/}{https://archive.org/}.},
-and (iii) include a mention to at least one valid UK postcode in the web
-text. The act of creating a website is understood here as active
-engagement with the web vis-à-vis the more passive act of browsing the
-web or having an internet connection \citep{tranosuk}. Previous studies
-have focused mostly on more passive notions of engaging with digital
-technologies such as internet adoption and internet speeds
-\citep[e.g.][]{blank2018local, riddlesden2014broadband}. More details
-about the data and the data generation process can be found in Section
-\hyperref[sec3]{XX}.
-
-\citet{grubler1990rise} Later Hagerstrand conceptualized physical
-``barrier'' effects like lakes or uninhabited areas, which, in addition
-to distance, act as further retarding effects on diffusion. These are
-formalized in the form of ``zero'' or ``half'' contact multiplicators on
-the (distance decaying) message flows.
-
-\citet{grubler1990rise} With respect to the formalization of the
-communication flows Hagerstrand defines a ``mean information field''
-(MIF), in which the probability of communication is a negative function
-of distance between individuals
-
-\citet{wilson201281} Logistic growth describes an initial period of
-gradual diffusion as a technology is introduced as a new commercial
-application, moving then through a rapid, exponential growth phase,
-before slowing and eventually saturating \citep{grubler1999dynamics}.
-The substitution of incumbent technologies by new competitors leads to
-subsequent decline and eventual obsolescence.
-
-\section{Literature review}\label{sec2}
+The data used here depict the active engagement with the Web as they
+contain geolocated and time-stamped counts of commercial websites from
+the UK. Instead of adopting metrics reflecting the passive engagement
+with the internet such as internet adoption and speeds
+\citep[e.g.][]{blank2018local, destefano2022fuzzy}, this paper observes
+yearly counts of geolocated websites, which reflect the creation and
+maintenance of websites. It needs to be highlighted here that although
+there are various types of underpinning web technologies -- namely
+markup, stylesheet, client- and server-side coding languages as well as
+different web browsers and other related programming language -- the
+focus here is on the diffusion of the Web in its totality. For a more
+detailed and cross-country technological diffusion study, see
+\citet{PAPAGIANNIDIS2015308}.
+
+To analyse these data, a diverse set of methods are employed. Firstly,
+cumulative adoption curves are estimated for the UK as a whole and for
+the Local Authority Districts (LAD), illustrating, for the first time,
+such S-shaped diffusion patterns for local areas. This system level
+analysis also reveals insights about the volatility of the adoption of
+the Web. Then, exploratory spatial data analysis offers some first
+indications of the importance of the different spatial diffusion
+mechanisms in shaping the diffusion of the Web in the UK. Importantly,
+the analysis adopts two distinct geographical scales: the LAD and the
+much more detailed scale of the smallest census areas in the UK, the
+Output Areas. Lastly, a novel methodological framework based on
+well-established Machine Learning (ML) algorithms is developed to test
+the role of such mechanisms at these two scales. The results reveal for
+the first time the importance and relevance of different spatial
+diffusion mechanisms for the diffusion of a tangible, digital technology
+at very granular geographical scales.
+
+The structure of the paper goes as follows. Section~\ref{sec-litreview}
+reviews the key concepts in the diffusion literature as well as relevant
+empirical studies. Then, Section~\ref{sec-datamethods} describes the
+data and methods used in this paper. The results of the different stages
+of the analysis are presented in Section~\ref{sec-results}. The
+discussion of the results and the conlcusions of the paper are offered
+in Section~\ref{sec-conclusions}.
+
+\textbf{SAY SOMETHING ABOUT DIST PAPER BY ARP, NOT SPECIFIC EXPLANATORY
+VARIABLES}
+
+\section{Literature review}\label{sec-litreview}
 
 Geographical diffusion is a synthesis of different processes. On the one
-hand, we can identify purely spatial or, in other words, contagious
-processes. Adjacency and, more broadly, distance are the key drivers of
+hand, a purely spatial or, in other words, contagious processes can be
+identified. Adjacency and, more broadly, distance are the key drivers of
 diffusion. This perspective draws similarities with epidemics:
@@
 consequently, proximity and exposure \citep{hivner2003facilitating}. On
-the other hand, we can identify hierarchical processes. Instead of
-horizontal distance-based diffusion mechanisms, the top-down hierarchy
-of urban systems drives diffusion. In reality, the synthesis of these
-two processes represents how new technologies diffuse over space and
-time \citep{morrill2020spatial}.
+the other hand, there is a hierarchical processes. Instead of a
+horizontal distance-based diffusion mechanism, the top-down hierarchy of
+urban systems shapes technological diffusion. In reality, the synthesis
+of these two processes better reflects how new technologies diffuse over
+space and time \citep{morrill2020spatial}.
 
@@
 networks, which themselves tend to be defined by geography. Hence, he
-identified the `neighbourhood' effect of how information, and
+identified a `neighbourhood' mechanism of how information, and
 consequently, innovation diffuse. He used agricultural innovations to
@@
 seminal work on `Diffusion of Innovations' \citep{rogers2010diffusion}.
-Rogers being a sociologist, his work focused not on the diffusion of
-innovations over space and time, but instead on the adoption of new
-technologies and innovations by individuals and the individual
-mechanisms that drive the decisions behind adoption. He identified five
-groups of individuals regarding their adoption speed: innovators, early
-adopters, early majority, late majority and laggards. The key mechanism
-of diffusion and adoption is communication and how knowledge is
-transferred within a social system. Therefore, all approaches agree on
-the S-shaped diffusion and cumulative adoption pattern
-\citep{grubler1990rise}.
+Rogers being a sociologist focused not on the diffusion of innovations
+over space and time, but instead on the adoption of new technologies and
+innovations by individuals and the individual mechanisms that drive the
+decisions behind adoption. He identified five groups of individuals
+regarding their adoption speed: innovators, early adopters, early
+majority, late majority and laggards. The key mechanism of diffusion and
+adoption is communication and how knowledge is transferred within a
+social system. All above approaches agree that diffusion and the
+cumulative adoption of a new technology are represented by an S-shaped
+line plot \citep{grubler1990rise}.
 
@@
 core because of the lack of infrastructure or other necessary
-institutions \citep{leibowicz2016representing}. \citet{grubler1990rise}
-effectively summarises the three key characteristics of the spatial
-diffusion process: (i) the cumulative level of adoption follows an
+institutions \citep{leibowicz2016representing}.
+
+\citet{grubler1990rise} effectively summarises the three key spatial
+diffusion mechanisms: (i) the cumulative level of adoption follows an
 S-shaped pattern just like purely temporal models; (ii) diffusion is
-shaped by a hierarchy effect in a form of a centrifugal force: from core
-to periphery; and (iii) diffusion is also shaped by distance and a
-neighbourhood effect and contaminate nearby locations. These are the
-three mechanisms that the empirical analysis in Section
-\phantomsection\label{sec4}{4} investigates.
+shaped by a hierarchy mechanisms in a form of a centrifugal force: from
+core to periphery; and (iii) diffusion is also shaped by distance and
+proximity. These are the three mechanisms that the empirical analysis in
+Section~\ref{sec-results} investigates.
 
@@
 spatial dimension is present in most of the following studies, the level
-of spatial detail is always more coarse than the one adopted here.
-\citet{beardsell1999spatial} studied the evolution of the computer
-industry in \(317\) US metro areas during the \(1977-1992\) period using
-employment data. Their analysis indicated that the relative size
+of spatial detail is always more coarse than the one adopted in this
+paper. \citet{beardsell1999spatial} studied the evolution of the
+computer industry in 317 US metro areas during the 1977-1992 period
+using employment data. Their analysis indicated that the relative size
 distribution holds for urban computer employment and also urban
 heterogeneity is essential in explaining this distribution. In a recent
-study, \citet{bednarz2020pulled} focused on wind turbines and modelled
-their spatial diffusion across \(402\) German regions during
-\(1970-2015\). Their key finding is that local demand than local supply
-was the main driving factor.
-
-At a global scale \citet{perkins2005international} explored whther the
+study, \citet{bednarz2020pulled} focused on wind turbines and modeled
+their spatial diffusion across 402 German regions during 1970-2015.
+Their key finding is that local demand than local supply was the main
+driving factor. \citet{haller2011determinants} employed firm panel data
+for Ireland during 2003-2005 and illustrated that when firms are located
+in a region or industry with high share of firms having a website, they
+are more likely to have a website too.
+
+At a global scale \citet{perkins2005international} explored whether the
 diffusion rate of new technologies is driven by a latecomer advantage
@@
 \citet{perkins2011internet} explored whether the adoption of previous
-communication technologies that is mail, telegrams and telephones was
+communication technologies that is mail, telegrams and telephones were
 shaped by similar socioeconomic factors as the internet. Their results
@@
 Turning to studies that share more technological and scalar similarities
-with this paper, \citet{ding2010modeling} modelled the spatial diffusion
+with this paper, \citet{ding2010modeling} modeled the spatial diffusion
 of mobile telecommunications across regions in China. Their analysis
@@
 urban scaling and distance are the key drivers of spatial diffusion. At
-a global scale \citet{PAPAGIANNIDIS2015308} modelled the diffusion of
-different web technologies technologies and practices. Interestingly,
-they did so by using similar, but less extensive data as the one used
-here. Their analysis illustrated how the diffusion of different web
+a global scale \citet{PAPAGIANNIDIS2015308} modeled the diffusion of
+different web technologies and practices across countries.
+Interestingly, they used similar, but less extensive data as the ones
+used here. Their analysis illustrated how the diffusion of different web
 technologies and practices follow an S-shaped pattern as well as the
@@
 
-All in all, \ldots{}
-
-\section{Materials and Methods}\label{sec3}
-
-\textbf{ADD FAMILY OF WEB TECHNOLOGIES}
-
-To capture the diffusion of web technologies, a website density metric
-is developed for two different geographical scales: the Local Authority
+All in all, this paper tests whether the three distinct spatial
+diffusion mechanisms shaped the diffusion of the Web in the UK at very
+granular spatial scales from the Web's early stages onwards.
+
+\section{Data and Methods}\label{sec-datamethods}
+
+To capture the diffusion of the Web, a website density metric is
+developed for two different geographical scales: the Local Authority
 Districts (LAD) and the Output Areas (OA). The former is an
-administrative unit and there are
-\textbackslash textcolor\{red\}c.~380\textbackslash textcolor\{black\}
-such units in the UK. The latter is a census-based geographical unit,
-which is very small as there are c.~230,000 of them in the UK. This
-methodological choice will allow the mapping of the diffusion of web
-technologies and the assessment of the diffusion mechanisms at these two
-very different spatial scales.
-
-The population of websites is calculated using data from the Internet
-Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
+administrative unit and there are 374 LAD in the UK. The latter is the
+smallest census-based geographical unit and there are c.~230,000 of them
+in the UK. This methodological choice will allow the mapping of the
+diffusion of web technologies and the assessment of the diffusion
+mechanisms at these two very different spatial scales.
+
+The counts of websites at these scales are calculated using data from
+the Internet Archive\footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
 and, specifically, the JISC UK Web Domain Dataset \citep{ukwebarchive}.
@@
 \citep{ainsworth2011much, holzmann2016dawn}. It is a web crawler, which
-discovers webpages by following the hyperlinks of every webpage its
+discovers webpages by following the hyperlinks of every webpage it
 archives. This dataset, which is curated by the British Library,
@@
 
-Instead of using the whole .uk country code top-level domain (ccTLD),
-this paper focuses on its commercial subset, the .co.uk SLD. This choice
+Instead of using the whole .uk ccTLD, this paper focuses on its
+commercial subset, the .co.uk second level domain (SLD). This choices
 decrease the heterogeneity of the web data as such commercial websites
-have rather specific aims: they are used to diffuse information, support
-online transactions and share opinions
+have specific aims: they are used to diffuse information, support online
+transactions and share opinions
 \citep{THELWALL2000441, blazquez2018big}. Although a UK company can
-adopt a generic TLD such as.com and such cases escape the data used
-here, such omissions should not affect our results given the popularity
-of the .uk ccTLD \citep{tranosuk}: UK consumers prefer to visit a .uk
-website when they are searching for products or services \citep{hope};
-and anecdotal evidence indicates that during the first half of 2000,
-three .co.uk domains were registered every minute \citep{oecd_coms}.
-Importantly, previous studies illustrated that .co.uk is the most
-popular UK SLD \citep{tranosuk}.
-
-The text from these webpages was scanned using regular expressions
+adopt a generic TLD such as .com and these cases escape the data used
+here, such omissions should not affect the validity of the results given
+the popularity of the .uk ccTLD \citep{tranosuk}: UK consumers prefer to
+visit a .uk website when they are searching for products or services
+\citep{hope}; and canecdotal evidence indicates that during the first
+half of 2000, threec .co.uk domains were registered every minute
+\citep{oecd_coms}. Importantly, previous studies illustrated that .co.uk
+is the most popular UK SLD \citep{tranosuk}.
+
+The text from these webpages was scanned using a regular expression
 (regex) to identify strings of text which resemble UK postcodes and one
@@
 
-Before that, the data cleaning process included a data aggregation step,
-through which the webpages were aggregated to the relevant websites.
-Based on the following example, all three webpages listed below are part
-of the same website (\url{http://www.website.co.uk}) and at the end only
-websites and not the nested webpages were considered as otherwise the
-metric would have been biased towards large websites.
+The data cleaning process includes an aggregation step, through which
+the archived webpages are aggregated to the parent websites. This
+website reconstruction process lend itself to the consteruction of
+\emph{website} instead of \emph{webpage} density metrics. Websites tend
+to represent specific orgnanisations or entities, and, arguably, are
+more meaningful observational units than webpages, which can ignore the
+upstream dependency of the website they belong to. Based on the
+following example, all three webpages are part of the same website
+(\url{http://www.website.co.uk}). Only the overall website and not the
+nested webpages are considered for the counts as otherwise the density
+metrics would have been biased towards large, place specific, websites.
 
@@
 websites with multiple postcodes. As per the above example,
-(\url{http://www.website.co.uk}) includes two unique postcodes: B152TT
-and BS81SS. The distribution of postcodes for 2000 is presented in Table
-\ref{f2000}, which clearly illustrates the wide range. On one hand, we
-have websites anchored to a unique location (72\% of all the
-reconstructed websites), which may represent a small company with a
-single trading location. On the other hand, we have websites with
-thousands of different postcodes. Considering the timeperiod of the
-analysis, such cases can represent directories which used to be popular
-in the pre-search engines early times of internet
-\citep{tranosuk}.\footnote{See Figure A1 in \textbf{Appendix A in the
-  supplemental data online for examples}.}
+\href{http://www.website.co.uk}{www.website.co.uk} includes two unique
+postcodes: B15 2TT and BS8 1SS. The distribution of the number of
+postcodes per website for 2000 is presented in Table \ref{f2000}, which
+clearly illustrates a wide range. At the left end of the distribution,
+there are websites anchored to a unique location (72\% of all the
+reconstructed websites in 2000), which may represent a small company
+with a single trading location. At the right end, there are websites
+with thousands of different postcodes. Considering the time period of
+the analysis, such cases can represent directories which used to be
+popular in the pre-search engines early times of the commercial internet
+as well as real estate websites \citep{tranosuk}.
 
 The analysis presented here is based on two subsets of these data.
-Firstly, the analysis is based on websites which only contain only one
-unique postcode and, therefore, the geolocation process might suffer
-from less noise. As a robustness check, the analysis is replicated the
-analysis for an extended subset of websites, which include up to ten
-unique postcodes. These websites are geolocated by equally attaching
-them to multiple locations. This extended sample includes 94 percent of
-all the archived websites in 2000.
+Firstly, on websites, which only contain one unique postcode. As a
+robustness check, the analysis is replicated for an extended subset of
+websites, which include up to 10 unique postcodes to capture commercial
+websites which point to multiple locations. They are geolocated by
+equally attaching them to these locations. This extended sample includes
+94 percent of all the archived websites in 2000.
 
@@
 
-To create the web density metric, the yearly website counts at the LAD
-level are standardised by the number of firms in LAD to avoid biases
-associated with LADs hosting a large number of firms. Given that there
-is no such statistic for the OA, the actual OA level counts are used.
-However, because of the the rather consistent spatial definition of OA
-(they host 40-250 households),\footnote{\url{https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies}}
+Another data cleaning step deals with some extreme outliers. Figure
+\ref{correct} plots the website counts for the top 1000 postcodes. Some
+obvious outliers can be observed for the 2002-2006 period for a handful
+of postcodes, which can be attributed to link farms \citep{BL2013links}.
+The website counts for these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1
+6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1000
+websites pointing to them, were replaced with predicted values based on
+a simple panel regression model with postcode fixed effects and yearly
+dummy variables. These postcodes refer to a residential area with a
+small park in South London, residential areas outside Warwick and
+Manchester, a rural area in the south border of England and Wales and
+small business parks in Crew and Derby. Hence, it is difficult to
+believe that these postcodes genuinely hosted such a large amounts of
+websites, which were later extensively declined. To put the magnitude of
+the data imputation into perspective, this process affected 6 out of the
+557,808 postcodes included in the data. As the Section~\ref{sec-results}
+illustrates, this led to a small increase of the model predictive
+capability.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{tranos2023_files/figure-pdf/unnamed-chunk-2-1.pdf}
+
+}
+
+\caption{\label{correct}Yearly website counts per postcode (top 1000)
+before (top) and after (down) data imputation}
+
+\end{figure}%
+
+To create the website density metric, the yearly website counts at the
+LAD level are standardised by the number of firms in LAD to avoid biases
+associated with LAD hosting a large number of firms. Given that there is
+no such statistic for the OA, the actual OA level counts are used.
+However, because of the the consistent spatial definition of OA (they
+host 40-250 households),\footnote{\url{https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies}}
 website counts in OA are interpreted as a density metric too.
 
-This website metric\ldots{}
-
-\citet{wilson201281} growth function description in p.~86
-
-from R: Asym/(1+exp((xmid-input)/scal)) using the terms from Wilson k/(1
-+ e\^{}(t\_0 - t)/scal)
-
-Wilson: k(1 + e\^{}-b(t - t\_0)) So, scal = - 1/b
-
-The curve is symmetric* around t\_0
-
-\textbf{minimum R2 = 95\%}, see also \citet{grubler1990rise}
-
-The literature usually uses the saturation level as the asymptote. I am
-using the total number of websites as we cannot compute a rate.
-
-Moran's I as \citet{ding2010modeling} \textbf{TODO} for t\_0, diffusion
-speed
-
-\section{Results}\label{sec4}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  S-shaped diffusion curves: S for LADs per firm. UK, fast/slow/examples
-\item
-  ranks: there is stability and movement
-\item
-  Neighbourhood effect: diffusion proceeds outwards from innovation
-  centers, first ``hitting'' nearby rather than far-away locations
-  (Grubler 1990)
-\end{enumerate}
-
-\begin{itemize}
-\item
-  Moran's I: for OA and LAD over time
-\item
-  LISA maps: for OA and LAD over time More and less expected clusters.
-  Different scales show different results
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{3}
-\tightlist
-\item
-  Hierarchy effect: from main centers to secondary ones -- central
-  places
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  Gini coefficient. Almost perfect polarisation of web adoption in the
-  early stages at a granular level More equally diffused at the Local
-  Authority level Plateau overtime
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{4}
-\tightlist
-\item
-  RF
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  ideal: (i) train RF for all years and all (1) LADs and (2) OAs with
-  CAST and report metrics. (ii) train for all years and all but one
-  region for (1) LADs and (2) OAs to predict to the holdout region.
-  Reports predictions as region similarities.
-\end{itemize}
-
-\subsubsection{RF results}\label{rf-results}
-
-The next section incorporates the above discussed spatial processes of
-the diffusion of web technologies into the same modelling framework. The
-aim is to use variables depicting these spatial processes in order to
-predict the diffusion of web technologies in the UK over space and time
-and across different scales. Specifically, four different models are
-estimated. Firstly, all the data points at the OA and LAD are utilised
-in order to build RF models and assess their capacity to predict the
-adoption of web technologies. These two models will reveal the
-predictive capacity of the spatial processes and also allow to see how
-the importance of such variables changes across scales. The next two
-sets of models will be again trained on web diffusion at the two working
-scales: OA and LAD. However, instead of using all the data points, the
-OA and the LAD from one of the twelve UK regions are held out and then
-the trained model is used to predict web adoption in the OA or the LAD
-of the held-out region. This process takes place recursively for all UK
-regions. The difference in the predictive capacity of the different
-samples will reveal how dissimilar are these spatial process across
-regions and, importantly, at different scales.
+These website density metrics are used for the three different stages of
+the analysis. Firstly, a system-level analysis explores how the
+diffusion of web technologies in the UK fits with the well-established
+S-curve. To do so, the following logistic function (Equation \ref{eq:s})
+is estimated for the whole of the UK and for each LAD separately:
+
+\begin{align}
+y = k /(1 + e^{-(t-t_{o})})\label{eq:s}
+\end{align}
+
+\noindent \(k\) is the asymptote or, in other words, the saturation
+level, \(b\) the overall growth rate, and \(t_{0}\) the \emph{inflection
+point} of maximum growth at \(k/2\), where the logistic function is
+symmetrical \citep{wilson201281}. The \(t_0\) of each LAD is compared
+against the \(t_0\) of the UK to delineate whether a LAD reached that
+point faster or slower than the country average. Importantly, an
+accuracy criterion was imposed and only LAD with \(R^2 > 0.9\) were
+included in this analysis. To estimate Equation \ref{eq:s} a
+self-starting logistic growth model was employed using the \texttt{nls}
+and \texttt{SSlogis} functions in \texttt{R}. This logistic growth model
+depicts the initial period of slow diffusion of a new technology, which
+is followed by the fast, exponential growth stage to then eventually
+slow down and saturate \citep{wilson201281, grubler1999dynamics}.
+
+The system-level analysis also focuses on the volatility of the web
+adoption to depict places with high concentration of early adopters and,
+equally, latecomers. To do so, the change of the ranking of the UK LAD
+over time is plotted and discussed. Both the S-curve and the volatility
+analysis focus only on LAD as the very large number of OA would have
+made such analysis difficult to visualise and interpret.
+
+Next, exploratory spatial data analysis depicts whether the two main
+drivers of spatial diffusion -- namely neighbourhood and hierarchy --
+underpin thediffusion of web technologies in the UK. To capture the
+former and following \citet{ding2010modeling} the Moran's I and the
+Local Index of Spatial Association (LISA) are estimated for the website
+density. To address the hierarchy mechanism, the Gini coefficient -- a
+well established metric of inequality -- is calculated. All the above
+are computed and plotted longitudinally both for the LAD and the OA.
+
+Lastly, a modelling framework is developed to test the above diffusion
+mechanisms. The overarching aim is to build a model that can test the
+relationship between website density and these mechanisms:
+
+\begin{align}
+Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S-curve_{t}\label{eq:rf.generic}
+\end{align}
+
+\noindent Equation \ref{eq:rf.generic} is estimated using Random Forest
+(RF). This is a popular ML algorithm for both regression and
+classification problems \citep{biau2012analysis}. It was introduced by
+\citet{breiman2001random} and has become a go-to data science tool. RF
+can effectively handle skewed distributions and outliers, model
+non-linear relationships, require minimal hyperparameter tuning, exhibit
+low sensitivity to these parameters, and have relatively short training
+times \citep{Caruana2008, liaw2002classification, yan2020using}. These
+attributes match well with the website density data characteristics
+including skewness especially for the OA. Also, the large data size
+(c.~230k data points for each of the 17 years) calls for fast training
+times. Importantly, RF predictions tend to be more accurate than those
+from single regression trees and outperform Ordinary Least Squares in
+out-of-sample predictions, even with moderate-sized training data and a
+small number of predictors
+\citep{mullainathan2017machine, athey2019machine, sulaiman2011intelligent, pourebrahim2019trip, biau2012analysis}.
+
+RF is a tree-based ensemble learning algorithm
+\citep{breiman2001random}. It begins by generating random samples of the
+training data, which are then used to grow regression trees to predict
+the dependent variable. Data points and predictors are randomly sampled
+for the different trees. The trees are trained in parallel using their
+own bootstrapped samples of the training data. A crucial feature of RF
+is their ability not to overfit, meaning they can generalize well to
+unseen test data. While each tree may overfit individually, the ensemble
+of trees does not because the errors of individual trees are averaged,
+reducing the overall variance and preventing overfitting
+\citep{last2002improving}. For regression problems, RF predictions are
+made by averaging the predictions of all decision trees.
+
+RF have been widely employed to address regression research problems.
+\citet{pourebrahim2019trip} combined a spatial interaction modeling
+framework with ML algorithms including RF to predict commuting flows in
+New York City. \citet{sinha2019assessing} advocated for adopting spatial
+ensemble learning approaches, such as RF, to model spatial data with
+high autocorrelation and heterogeneity. \citet{creditspatial} predicted
+employment density in Los Angeles using spatially explicit RF.
+\citet{guns2014recommending} used RF to build a recommendation system
+for research collaborations. \citet{ren2019predicting} trained RF to
+predict the socio-economic status of cities using various online and
+mobility predictors. \citet{tranos2023using} utilised hyperlinks data
+and RF to make out-of-sample predictions of interregional trade.
+\citet{zhou2023geography} employed such a framework to assess whether
+key predictors of obesity differ across English cities.
+
+\section{Results}\label{sec-results}
+
+\subsection{System-level analysis}\label{system-level-analysis}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/s_uk_per_firm.png}
+
+}
+
+\caption{\label{s_uk}Grwoth curve, UK}
+
+\end{figure}%
+
+Figure \ref{s_uk} plots the S curve for the cumulative adoption of
+website technologies in the UK during the 1996-2012 period. It
+demonstrates a pattern well aligned with previous studies discussed in
+Section~\ref{sec-litreview}. The vertical line around year 2003
+illustrates the point where the modeled cumulative adoption is equal to
+50\% of the maximum. This \(t_0\) \emph{inflection} point signals the
+maximum adoption speed and is used here to determine whether a UK LAD
+reached that inflection point earlier or later than the UK average.
+Specifically, the S curve and the inflection point are estimated for
+every LAD individually and then compared against the UK average. LAD
+that reached their inflection point earlier than the country average are
+labelled as \emph{fast} and the rest as \emph{slow}. Figure \ref{s_map}
+maps this pattern. There is a clear high concentration of fast adopting
+LAD in the South and East of England as expected. However, there is also
+such a concentration in Wales, Scotland and the North West of England.
+At a more detailed level, there are some expected examples of LAD with
+relevant industrial backgrounds delineated as fast: the City of London,
+a world-renowned cluster of finance industries \citep{cook2007role}, and
+Reading, a town with high-tech service industries in proximity to London
+and its main airport, Heathrow \citep{england2005polynet}. However,
+there are also LAD which were expected to appear as fast -- e.g.~Hackney
+in central London and Bristol, a well-established creative cluster
+\citep{oatley1999cultural, bassett2002cultural} -- but were delineated
+as slow. Nevertheless, there is obvious spatial clustering of fast and
+slow LAD. Importantly, out of the ten fastest LAD, nine were located in
+the South East of England and London and one in Cambridge (see Table
+\ref{table.s.lads} in the Appendix). The above observations remain
+almost unchanged when the analysis includes websites with up to 10
+unique postcodes -- see Figures \ref{s_uk10} and \ref{s_map10} in the
+Appendix -- advocating towards the robustness of the analysis.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/speed_map.png}
+
+}
+
+\caption{\label{s_map}LAD adoption rates}
+
+\end{figure}%
+
+The next step is assesses the stability and volatility of the LAD in
+terms of their adoption of web technologies. As we know from the
+literature \citep{risk_perceptions}, different agents have different
+perceptions about and levels of acceptance of the risks and the
+potential economic returns associated with the adoption of new
+technologies -- see for instance the seminal work of
+\citet{venkatesh2000theoretical}. To reveal such aggregated patterns,
+Figure \ref{rank} plots the ranking of UK LAD based on website density.
+To decrease noise, the average ranking of 1996-1998 and 2010-2012 is
+plotted instead of the individual years. The colours depict the 5th and
+95th percentile of the absolute difference of the LAD ranking between
+1996-1998 and 2010-2012. While there are quite a few obvious cases of
+LAD that maintained their position between the beginning and the end of
+the study period both at the top or at the bottom of the hierarchy,
+there are also quite a few LAD that changed drastically their position.
+Some of these LAD enjoyed a process that at the first instance looks
+like \emph{leapfrogging} since they managed to jump at the top of the
+hierarchy despite their slow start. There is extensive literature
+regarding the potential benefits of technological leapfrogging. The
+underpinning argument is that latecomers can adopt and benefit from new
+technologies that have been developed elsewhere without incurring the
+hefty initial R\&D costs \citep{teece2008firm}. Although the
+leapfrogging literature does not pay much attention to cities and
+regions \citep{yu2018sustainability}, previous research highlighted the
+long term and sustained productivity benefit of the early adopters of
+web technologies in the UK \citep{tranosuk}. So, although the LAD system
+is volatile, it is not clear whether the LAD with high concentration of
+late adopters will gain any latecomer benefits in a way similar to
+countries experiencing such technological leapfrogging. As previously,
+this is consistent with including websites with up to 10 postcodes (see
+Figure \ref{rank10}).
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/ranks/web_per_firm2000_2012_only0595_av.png}
+
+}
+
+\caption{\label{rank}Dynamics of wed diffusion (LAD)}
+
+\end{figure}%
+
+\subsection{Exploratory analysis}\label{exploratory-analysis}
+
+Figures \ref{morani} and \ref{lisa} offer a first insight into whether a
+neighbourhood mechanism underpins the diffusion of the Web in the UK as
+they plot the Moran's I and the LISA maps of website density
+respectively for LAD and OA. Starting from the former, spatial
+autocorrelation was higher in the beginning of the study period and then
+after 2000 dropped slightly and stabilised around 0.2. This reflects the
+early concentration of high website density around London, which over
+time diffused as high-high clusters can be seen in other parts of the
+country away from London (Figure \ref{lisa}). An almost reverse pattern
+can be observed for OA. At the beginning of the study period Moran's I
+was around 0.5 and it plateaued after 2000 around 0.2. Because of the
+very small size of OA, at the early stages of the diffusion of web
+technologies their adoption was spatially scattered. This is reflected
+in the lack of any significant clusters in 1996 (Figure \ref{lisa}).
+Eventually, as the adoption rate increased, more such clusters of high
+website density were formed and this is reflected both in the Moran's I
+and the LISA maps in Figures \ref{morani} and \ref{lisa}. A similar
+pattern is observed for the extended dataset (see Figures\ref{morani10}
+and \ref{lisa10}). In this case, the magnitude of spatial
+autocorrelation is slightly higher illustrating a stronger spatial
+clustering mechanism. All in all, the exploratory spatial data analysis
+advocates towards an underpinning neighbourhood mechanism and the
+different scales of analysis illustrate how it evolved differently over
+time.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2023_files/figure-pdf/morani-1.pdf}
+
+}
+
+\caption{\label{morani}Website density Moran's I}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2023_files/figure-pdf/unnamed-chunk-6-1.pdf}
+
+}
+
+\caption{\label{lisa}\centering Website density LISA maps}
+
+\end{figure}%
+
+To illustrate whether a hierarchical process also underpins the
+diffusion of web technologies, the Gini coefficient is calculated yearly
+both for LAD and OA. As a metric of inequality, the Gini coefficient
+demonstrates whether website density is concentrated in a small number
+of LAD or OA, or whether it is more equally spread across the country.
+Both scales of analysis in Figure \ref{gini} illustrate the same
+picture. At the beginning of the commercial internet website density was
+extremely unequal, or, in other words, only a few places had websites
+associated with them. Inequality dropped and plateaued after 2000 for
+both scales. This is illustrative of a hierarchical diffusion mechanism
+that led over time to a more equal spread. Interestingly, the year 2000
+is again a period of change for this diffusion mechanism as it was for
+the neighbourhood process. There is a substantial difference between the
+Gini coefficient magnitude for LAD and OA, but this is expected as the
+very small size of OA equates to a lot of polygons without any websites
+pointing to them -- for example residential OA.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2023_files/figure-pdf/gini-1.pdf}
+
+}
+
+\caption{\label{gini}Website density Gini coefficient}
+
+\end{figure}%
+
+\subsection{Modelling results}\label{modelling-results}
+
+The next section incorporates the above mechanisms of the spatial
+diffusion of the Web into a modelling framework. The aim is to use
+variables depicting these spatial processes in order to predict the
+diffusion of web technologies in the UK over space and time and across
+different scales. Specifically, four different models are estimated.
+Firstly, all the data points for the OA and LAD are utilised in order to
+build two RF models and assess their capacity to predict the adoption of
+web technologies at the two different scales. These two models will
+reveal the predictive capacity of the diffusion mechanisms and also
+demonstrate how the importance of such variables changes across scales.
+The next two sets of models are again trained at the two working scales:
+OA and LAD. However, instead of using all the data points, the OA and
+the LAD from one of the twelve UK regions are held out for the model
+training. Then the trained model is used to predict website density for
+the OA or the LAD of the held-out region. This process takes place
+recursively for all twelve UK regions. The regional difference in the
+predictive capacity of the different samples will reveal how dissimilar
+these spatial process are across regions and, importantly, at different
+scales.
 
 It needs to be highlighted here that the cross-validation for all models
-is spatially and temporally sensitive. Instead of using 10 random folds,
-we employ the \texttt{CAST} package which allows holding back data
-points from specific years and spatial units and use them for testing in
-order to estimate the model performance \citep{meyer2018improving}.
-
-The models need to include variables that capture the three processes
-that the relevant literature and the descriptive analysis presented in
-the Section \textbf{XX} highlighted. Namely, the models capture: (i) a
-hierarchy effect with diffusion running from main centres to secondary
-ones, (ii) a neighborhood effect according to which diffusion first hits
-nearby locations, and (iii) the rather canonical pattern of diffusion
-over time as reflected in the S-shaped pattern in the cumulative level
-of adoption.
+is spatially and temporally sensitive. Instead of using 10 random
+folds,the \texttt{CAST} package is employed which allows holding back
+data points from specific years and spatial units and use them for
+testing in order to estimate the model performance
+\citep{meyer2018improving}.
+
+The models include variables that capture the three processes that the
+relevant literature and the descriptive analysis highlighted. Namely,
+the models capture: (i) a hierarchy mechanism with diffusion running
+from main centres to secondary ones, (ii) a neighborhood mechanism
+according to which diffusion first hits nearby locations, and (iii) the
+rather canonical pattern of diffusion over time as reflected in the
+S-shaped pattern in the cumulative level of adoption.
 
 To capture the hierarchy effect the models include as predictors a one
-year lag of website density in London -- the largest city in the UK, a
-one year lag of the website density in the nearest city and the same for
-the nearest retail centre. Due to the small sizes of the retail centres,
-the latter is only relevant for the OA-level models. In addition, the
-models include the distance to London, the nearest city and the nearest
-retail centre. The underlying logic is that the level of website
-adoption in a spatial unit depends on the level of the adoption in
-places further up in the urban hierarchy the previous year. To depict
-the neighbourhood effect, the web density of the neighbouring spatial
-units in the previous year is employed. Again, the underpinning
-rationale is that the level of web adoption within a spatial unit
-depends on the level of web adoption in the neighbouring spatial units
-the year before. This is the `hitting nearby locations first' argument.
-Therefore, the spatial and temporal lag of the website density in LAD
-and OA is calculated. Lastly, the time effect which is reflected on the
-S curve for the cumulative adoption is captured by trend variable.
-Hence, all four model will follow the following generic form (Eq.
+year lag of website density in London, the largest city in the UK, a one
+year lag of the website density in the nearest city and the same for the
+nearest retail centre. Due to the small sizes of the retail centres, the
+latter is only relevant for the OA-level models. In addition, the models
+include the distance to London, the nearest city and the nearest retail
+centre. The underlying logic is that the level of website adoption in a
+spatial unit depends on the level of the adoption in places further up
+in the urban hierarchy the previous year. The inclusion of the distance
+variables incorporates spatial structure into the hierarchy argument. To
+depict the neighbourhood effect, the website density of the neighbouring
+spatial units in the previous year is employed (rook continuity). Again,
+the underpinning rationale is that the level of web adoption within a
+spatial unit depends on the level of web adoption in the neighbouring
+spatial units the year before. This represents the `hitting nearby
+locations first' argument. Therefore, the spatial and temporal lag of
+the website density in LAD and OA is calculated. Lastly, the temporal
+pattern of the cumulative adoption is captured by a time trend variable.
+Hence, all four models will follow the following generic form (Eq.
 \ref{model}):
@@
 
-\(y_{k}\) is the \(k^{th}\) observation of the dataset, which consists
-of \(N\) observations in total. \(\hat{y_{k}}\) is the \(k_{th}\)
-predicted value for the dependent variable and \(\overline{y_{k}}\) is
-the average value of \(y\). The last two metrics are expressed in the
-same units as the dependent variable -- websites per firm for the LAD
-modes and the number of websites for the OA models -- while the first
-one is the coefficient of determination between the observed and the
-predicted values of website adoption. Regarding \(MAE\), it is the
-absolute difference between the observed and the predicted website
-adoption. While \(MAE\) does not penalise for large errors, \(RSME\)
-does so as it is proportional to the squared difference between the
-observed and the predicted trade flows. Hence larger errors weigh more
-for \(RMSE\) \citep{pontius2008components}.
-
-Table \ref{table.metrics.all} presents the model performance for the
-first set on models, for which all data points are employed for training
+\noindent \(y_{k}\) is the \(k^{th}\) observation of the dataset, which
+consists of \(N\) observations in total. \(\hat{y_{k}}\) is the
+\(k_{th}\) predicted value for the dependent variable and
+\(\overline{y_{k}}\) is the average value of \(y\). The last two metrics
+are expressed in the same units as the dependent variable -- websites
+per firm for the LAD modes and the number of websites for the OA models
+-- while the first one is the coefficient of determination between the
+observed and the predicted values of website adoption. Regarding
+\(MAE\), it is the absolute difference between the observed and the
+predicted website adoption. While \(MAE\) does not penalise for large
+errors, \(RSME\) does so as it is proportional to the squared difference
+between the observed and the predicted trade flows. Hence larger errors
+weigh more for \(RMSE\) \citep{pontius2008components}.
+
+Table~\ref{tbl-model-metrics} presents the model performance for the
+first set of models, for which all data points are employed for training
 and testing via cross validation. The first one is trained and tested on
@@
 the outcome of space and time sensitive CV, so the the model does not
-suffer from overfitting. At the LAD level the model predicts 81\% of the
+suffer from overfitting. At the LAD level the model predicts 85\% of the
 variation of website density. Both error metrics indicate that the model
 error is \((\frac{1}{20}, \frac{1}{30})\) of a website per firm. At the
-OA the RSquared drops down to 21\%. Considering its granularity, this is
+OA the \(R^2\) drops down to 32\%. Considering its granularity, this is
 still a remarkable performance. To contextualise it, the model results
 in a MAE of one website for areas small enough to host less than 140
-households.\footnote{According to the Office for National Statistics,
+households. \footnote{According to the Office for National Statistics,
   80\% of OA in England and Wales host 110-139 households,
@@
 highly skewed and a significant part of them is not linked to any
-websites. In 1997 only 1\% of the UK OA were associated with a website.
-This should not come as a surprise as this was the very beginning of the
-commercial internet and any activities with a digital footprint were
-concentrated in a handful of areas. This was clearly illustrated in
-Section \textbf{ADD}. At the end of the study period almost half of the
-UK OA were not associated with a website. Again, given the granularity
-of the data this should not come as a surprise.
+websites. In 1997 only 1\% of the UK OA were associated with at least
+one website. This should not come as a surprise as this was the very
+beginning of the commercial internet and any activities with a digital
+footprint were concentrated in a handful of areas. This was clearly
+illustrated in Figure \ref{lisa}. At the end of the study period almost
+half of the UK OA were not associated with a website. Again, given the
+granularity of the data this should not come as a surprise.
 
 \begin{longtable}[]{@{}llll@{}}
-\caption{Model metrics \{\#table.metrics.all\}}\tabularnewline
+\caption{Model metrics}\label{tbl-model-metrics}\tabularnewline
 \toprule\noalign{}
-& RMSE & RSquared & MAE \\
+& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
 \midrule\noalign{}
@@
 \toprule\noalign{}
-& RMSE & RSquared & MAE \\
+& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Local Authorities & 0.028 & 0.850 & 0.019 \\
+Output Areas & 3.284 & 0.320 & 1.034 \\
+\end{longtable}
+
+The results in Table~\ref{tbl-model-metrics} are robust against
+different specification. As noted earlier, the data imputation process
+resulted in a modest improvement of the predictive capacity of the
+model. As per Table~\ref{tbl-model-metrics-no-correction} all accuracy
+metrics are slightly worst when the data imputation is applied. Then,
+Table~\ref{tbl-model-metrics-10} presents the model results for the
+extended dataset, which includes websites with up to 10 unique
+postcodes. While the results for the LAD are slightly less accurate, the
+\(R^{2}\) of the OA model is more than double in comparison to the base
+results in Table~\ref{tbl-model-metrics}.
+
+\begin{longtable}[]{@{}llll@{}}
+\caption{Model metrics without data
+imputation}\label{tbl-model-metrics-no-correction}\tabularnewline
+\toprule\noalign{}
+& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
 \midrule\noalign{}
@@
 
-Figures \ref{var.imp.LAD} and \ref{var.imp.OA} plot the importance of
-the different predictors. When the focus is on the LAD, the website
-density in the nearest city, in London and in the neighbouring LADs the
-year before are the most important predictors. They are followed by the
-yearly trend, while the spatial configuration as reflected in distances
-to London or the nearest city only play a minor role. This can be
-attributed to the rather coarse spatial scale of analysis. Nevertheless,
-all previously discussed spatial processes are at play in the diffusion
-of web technologies at the LAD level: the first two predictors depict
-the hierarchical effect, the spatial and temporal lag of website density
-depict the neighbourhood effect and the yearly trend the time-sensitive
-cumulative adoption pattern.
+\begin{longtable}[]{@{}llll@{}}
+\caption{Model metrics for the extended
+dataset}\label{tbl-model-metrics-10}\tabularnewline
+\toprule\noalign{}
+& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Local Authorities & 0.188 & 0.77 & 0.122 \\
+Output Areas & 22.467 & 0.478 & 5.681 \\
+\end{longtable}
+
+Figure \ref{var.imp} plots the importance of the different predictors.
+When the focus is on LAD, the website density in the nearest city, in
+London and in the neighbouring LAD the year before are the most
+important predictors. They are followed by the yearly trend, while the
+spatial configuration as reflected in distances to London or the nearest
+city only plays a minor role. This can be attributed to the relatively
+coarse spatial scale of LAD. Nevertheless, all previously discussed
+spatial processes are at play in the diffusion of web technologies at
+the LAD level: the first two predictors depict the hierarchy mechanism,
+the spatial and temporal lag of website density the neighbourhood
+mechanism, and the yearly trend the time-sensitive cumulative adoption
+pattern.
 
 When the much more granular scale of OA is adopted, the picture is
-reversed. The most important predictors are the three distance variables
-to London, the nearest city and the nearest retail centre. They stil
-depict the hierarchical effect, but proximity to the different
+almost reversed. The two most important predictors are the distance to
+London and to the nearest city followed by the spatial and temporal lag
+of website density and the distance to the nearest retail centre. They
+still depict a hierarchical mechanism, but proximity to the different
 population centres is more important than their lagged web densities in
-predicting website diffusion. The neighbouring effect is less important
-at this scale. What is interesting is the almost negligible role of the
-yearly trend and London's website density. While the former probably
-illustrates the large heterogeneity in how web technologies have been
-adopted at this very fine scale, the later highlights that the
-importance of past web adoption rates in large population centres is
+predicting website diffusion. The neighbourhood mechanism is still
+strong but less important at this scale. What is interesting is the
+almost negligible role of the yearly trend and London's website density.
+While the former probably illustrates the large heterogeneity in how the
+Web has been adopted at this very fine scale, the latter highlights that
+the importance of past web adoption rates in large population centres is
 surpassed by proximity to them and spatial configuration at this scale.
 
+The above pattern of variable importance is largely in accordance with
+the models estimated for the extended dataset. A comparison with Figure
+\ref{var.imp10} illustrates that the main difference is the importance
+of the neighbourhood mechanism for the OA as the space and time lag of
+the website density is by far the most important predictor when the
+extended dataset is employed. But still, the variable importance
+patterns for the two datasets are not dissimilar.
+
 \begin{figure}[H]
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/rf/figures/varimp_LA.png}
+{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{tranos2023_files/figure-pdf/varimp-1.pdf}
 
@@
 
-\caption{\label{var.imp.LAD}Variable importance, LAD}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/rf/figures/varimp_OA.png}
-
-}
-
-\caption{\label{var.imp.OA}Variable importance, OA}
+\caption{\label{var.imp}Variable importance}
 
@@
 \toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Region
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-RSquared LAD
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-Rank LAD
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-RSquared OA
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-Rank OA
-\end{minipage} \\
+Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+South East & 0.945 & 1 & 0.183 & 7 \\
+West Midlands & 0.922 & 2 & 0.190 & 5 \\
+Wales & 0.915 & 3 & 0.193 & 4 \\
+Yorkshire and The Humber & 0.909 & 4 & 0.200 & 2 \\
+East Midlands & 0.903 & 5 & 0.187 & 6 \\
+North East & 0.896 & 6 & 0.156 & 9 \\
+South West & 0.894 & 7 & 0.221 & 1 \\
+London & 0.881 & 8 & 0.104 & 10 \\
+East of England & 0.874 & 9 & 0.170 & 8 \\
+North West & 0.837 & 10 & 0.194 & 3 \\
+Scotland & 0.780 & 11 & 0.091 & 12 \\
+Nortern Ireland & 0.579 & 12 & 0.099 & 11 \\
+\end{longtable}
+
+\section{Discussion and conclusions}\label{sec-conclusions}
+
+contrary to results from future studies regarding social media
+\citep{lengyel2020role}, web technologies did not exclusively spread
+from a central location.
+
+\setcounter{section}{0}
+\renewcommand{\thesection}{\Alph{section}}
+
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{A\arabic{figure}}
+
+\section{Appendix}\label{appendix}
+
+\counterwithin{figure}{section}
+\counterwithin{table}{section}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/s_uk_per_firm.png}
+
+}
+
+\caption{\label{s_uk10}Grwoth curve, UK; up to 10 postcodes per website}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/speed_map_10.png}
+
+}
+
+\caption{\label{s_map10}LAD adoption rates; up to 10 postcodes per
+website}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/ranks/web_per_firm2000_2012_only0595_av_10.png}
+
+}
+
+\caption{\label{rank10}Dynamics of wed diffusion (LAD); up to 10
+postcodes per website}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2023_files/figure-pdf/morani10-1.pdf}
+
+}
+
+\caption{\label{morani10}Website density Moran's I; up to 10 postcodes
+per website}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2023_files/figure-pdf/unnamed-chunk-10-1.pdf}
+
+}
+
+\caption{\label{lisa10}\centering Website density LISA maps; up to 10
+postcodes per website}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2023_files/figure-pdf/gini10-1.pdf}
+
+}
+
+\caption{\label{gini10}Website density Gini coefficient; up to 10
+postcodes per website}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{tranos2023_files/figure-pdf/varimp10-1.pdf}
+
+}
+
+\caption{\label{var.imp10}Variable importance; up to 10 postcodes per
+website}
+
+\end{figure}%
+
+\begin{longtable}[]{@{}lrrrr@{}}
+\caption{Regional differences; up to 10 postcodes per
+website\label{table.regions.no.imp}}\tabularnewline
+\toprule\noalign{}
+Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
 \midrule\noalign{}
@@
 
-\section{Discussion and conclusions}\label{discussion-and-conclusions}
-
-contrary to results from future studies regarding social media
-\citep{lengyel2020role}, web technologies did not exclusively spread
-from a central location.
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.3186}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2212}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1416}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1062}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0619}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1504}}@{}}
+\caption{S-curve estiamtes for LAD; up to 10 postcodes per
+website\label{table.s.lads}}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+LAD
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Region
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(t_0\) estimate
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Std. error
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(R^2\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Diffusion speed
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+LAD
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Region
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(t_0\) estimate
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Std. error
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(R^2\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Diffusion speed
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Horsham & South East & 2001.423 & 0.254 & 0.950 & fast \\
+Fareham & South East & 2001.443 & 0.346 & 0.925 & fast \\
+Kingston upon Thames & London & 2001.513 & 0.354 & 0.933 & fast \\
+Kensington and Chelsea & London & 2001.581 & 0.341 & 0.943 & fast \\
+Runnymede & South East & 2001.582 & 0.266 & 0.943 & fast \\
+Bracknell Forest & South East & 2001.586 & 0.324 & 0.928 & fast \\
+Elmbridge & South East & 2001.656 & 0.333 & 0.932 & fast \\
+Reigate and Banstead & South East & 2001.665 & 0.253 & 0.964 & fast \\
+South Cambridgeshire & East of England & 2001.680 & 0.433 & 0.905 &
+fast \\
+Walsall & West Midlands & 2001.715 & 0.413 & 0.901 & fast \\
+Surrey Heath & South East & 2001.718 & 0.333 & 0.933 & fast \\
+Woking & South East & 2001.759 & 0.292 & 0.951 & fast \\
+South Norfolk & East of England & 2001.874 & 0.358 & 0.931 & fast \\
+City of London & London & 2001.874 & 0.499 & 0.930 & fast \\
+Wokingham & South East & 2001.891 & 0.287 & 0.958 & fast \\
+Reading & South East & 2001.900 & 0.318 & 0.948 & fast \\
+Sevenoaks & South East & 2001.904 & 0.407 & 0.926 & fast \\
+Huntingdonshire & East of England & 2001.912 & 0.402 & 0.925 & fast \\
+Pendle & North West & 2001.924 & 0.407 & 0.923 & fast \\
+St Albans & East of England & 2001.933 & 0.396 & 0.927 & fast \\
+Perth and Kinross & NA & 2001.943 & 0.228 & 0.970 & fast \\
+Bromley & London & 2001.945 & 0.396 & 0.919 & fast \\
+Rushmoor & South East & 2001.962 & 0.518 & 0.912 & fast \\
+Powys & NA & 2001.969 & 0.278 & 0.963 & fast \\
+Swindon & South West & 2001.996 & 0.390 & 0.916 & fast \\
+Chelmsford & East of England & 2002.008 & 0.545 & 0.911 & fast \\
+Crawley & South East & 2002.008 & 0.379 & 0.949 & fast \\
+Dumfries and Galloway & NA & 2002.012 & 0.251 & 0.967 & fast \\
+Inverclyde & NA & 2002.030 & 0.330 & 0.915 & fast \\
+North Hertfordshire & East of England & 2002.031 & 0.475 & 0.920 &
+fast \\
+Moray & NA & 2002.050 & 0.358 & 0.947 & fast \\
+Orkney Islands & NA & 2002.070 & 0.283 & 0.958 & fast \\
+Fife & NA & 2002.070 & 0.363 & 0.945 & fast \\
+Mole Valley & South East & 2002.071 & 0.460 & 0.925 & fast \\
+Guildford & South East & 2002.113 & 0.413 & 0.941 & fast \\
+Buckinghamshire & South East & 2002.114 & 0.390 & 0.942 & fast \\
+Cotswold & South West & 2002.116 & 0.375 & 0.946 & fast \\
+Shetland Islands & NA & 2002.120 & 0.332 & 0.949 & fast \\
+South Oxfordshire & South East & 2002.146 & 0.367 & 0.950 & fast \\
+Watford & East of England & 2002.157 & 0.451 & 0.929 & fast \\
+Aberdeenshire & NA & 2002.168 & 0.416 & 0.931 & fast \\
+Southampton & South East & 2002.189 & 0.472 & 0.922 & fast \\
+Mid Suffolk & East of England & 2002.196 & 0.442 & 0.922 & fast \\
+Eden & North West & 2002.196 & 0.334 & 0.949 & fast \\
+South Lakeland & North West & 2002.197 & 0.238 & 0.974 & fast \\
+Stoke-on-Trent & West Midlands & 2002.203 & 0.419 & 0.927 & fast \\
+Babergh & East of England & 2002.209 & 0.449 & 0.927 & fast \\
+Bexley & London & 2002.213 & 0.493 & 0.909 & fast \\
+Torbay & South West & 2002.215 & 0.374 & 0.942 & fast \\
+Brentwood & East of England & 2002.219 & 0.336 & 0.936 & fast \\
+Ceredigion & NA & 2002.227 & 0.241 & 0.971 & fast \\
+Spelthorne & South East & 2002.232 & 0.451 & 0.933 & fast \\
+Bath and North East Somerset & South West & 2002.238 & 0.402 & 0.941 &
+fast \\
+Falkirk & NA & 2002.245 & 0.367 & 0.938 & fast \\
+Broxtowe & East Midlands & 2002.258 & 0.435 & 0.932 & fast \\
+East Hertfordshire & East of England & 2002.261 & 0.453 & 0.928 &
+fast \\
+Stirling & NA & 2002.268 & 0.251 & 0.975 & fast \\
+Ealing & London & 2002.294 & 0.523 & 0.922 & fast \\
+Mid Sussex & South East & 2002.295 & 0.453 & 0.936 & fast \\
+Barrow-in-Furness & North West & 2002.303 & 0.284 & 0.959 & fast \\
+West Oxfordshire & South East & 2002.322 & 0.467 & 0.929 & fast \\
+Sutton & London & 2002.326 & 0.544 & 0.911 & fast \\
+Portsmouth & South East & 2002.356 & 0.479 & 0.921 & fast \\
+Great Yarmouth & East of England & 2002.369 & 0.265 & 0.968 & fast \\
+East Hampshire & South East & 2002.370 & 0.474 & 0.931 & fast \\
+Wyre Forest & West Midlands & 2002.376 & 0.561 & 0.916 & fast \\
+Angus & NA & 2002.378 & 0.396 & 0.939 & fast \\
+Argyll and Bute & NA & 2002.427 & 0.321 & 0.962 & fast \\
+Scottish Borders & NA & 2002.428 & 0.298 & 0.962 & fast \\
+East Suffolk & East of England & 2002.431 & 0.533 & 0.919 & fast \\
+Blackpool & North West & 2002.443 & 0.294 & 0.962 & fast \\
+Cherwell & South East & 2002.448 & 0.533 & 0.923 & fast \\
+Wandsworth & London & 2002.449 & 0.570 & 0.921 & fast \\
+Tendring & East of England & 2002.456 & 0.456 & 0.935 & fast \\
+Westminster & London & 2002.459 & 0.437 & 0.952 & fast \\
+Richmond upon Thames & London & 2002.462 & 0.389 & 0.956 & fast \\
+Test Valley & South East & 2002.463 & 0.553 & 0.920 & fast \\
+Isle of Wight & South East & 2002.478 & 0.377 & 0.953 & fast \\
+Vale of White Horse & South East & 2002.496 & 0.408 & 0.949 & fast \\
+Croydon & London & 2002.510 & 0.504 & 0.927 & fast \\
+Hounslow & London & 2002.521 & 0.426 & 0.952 & fast \\
+Copeland & North West & 2002.524 & 0.237 & 0.971 & fast \\
+Wiltshire & South West & 2002.528 & 0.450 & 0.940 & fast \\
+Gwynedd & NA & 2002.536 & 0.372 & 0.954 & fast \\
+Hillingdon & London & 2002.560 & 0.470 & 0.940 & fast \\
+Conwy & NA & 2002.564 & 0.372 & 0.954 & fast \\
+Highland & NA & 2002.566 & 0.270 & 0.972 & fast \\
+Basingstoke and Deane & South East & 2002.572 & 0.461 & 0.947 & fast \\
+Harrow & London & 2002.587 & 0.509 & 0.936 & fast \\
+West Berkshire & South East & 2002.595 & 0.696 & 0.902 & fast \\
+North Lincolnshire & Yorkshire and The Humber & 2002.596 & 0.501 & 0.933
+& fast \\
+East Dunbartonshire & NA & 2002.598 & 0.528 & 0.925 & fast \\
+Bradford & Yorkshire and The Humber & 2002.598 & 0.613 & 0.916 & fast \\
+Waverley & South East & 2002.606 & 0.512 & 0.928 & fast \\
+Stroud & South West & 2002.608 & 0.559 & 0.916 & fast \\
+Herefordshire, County of & West Midlands & 2002.622 & 0.388 & 0.956 &
+fast \\
+North Norfolk & East of England & 2002.624 & 0.422 & 0.945 & fast \\
+Nottingham & East Midlands & 2002.630 & 0.531 & 0.936 & fast \\
+Castle Point & East of England & 2002.645 & 0.470 & 0.936 & fast \\
+North Warwickshire & West Midlands & 2002.657 & 0.380 & 0.955 & fast \\
+Gravesham & South East & 2002.680 & 0.473 & 0.933 & fast \\
+Lancaster & North West & 2002.692 & 0.336 & 0.963 & fast \\
+West Suffolk & East of England & 2002.699 & 0.536 & 0.930 & fast \\
+Stafford & West Midlands & 2002.719 & 0.499 & 0.933 & fast \\
+Tunbridge Wells & South East & 2002.720 & 0.401 & 0.958 & fast \\
+Oxford & South East & 2002.723 & 0.583 & 0.934 & fast \\
+East Lothian & NA & 2002.731 & 0.537 & 0.927 & fast \\
+Redcar and Cleveland & North East & 2002.750 & 0.509 & 0.929 & fast \\
+Tamworth & West Midlands & 2002.755 & 0.551 & 0.929 & fast \\
+South Hams & South West & 2002.773 & 0.422 & 0.952 & fast \\
+Allerdale & North West & 2002.784 & 0.376 & 0.958 & fast \\
+Central Bedfordshire & East of England & 2002.825 & 0.620 & 0.921 &
+fast \\
+East Cambridgeshire & East of England & 2002.843 & 0.636 & 0.927 &
+fast \\
+Havant & South East & 2002.853 & 0.541 & 0.932 & fast \\
+South Lanarkshire & NA & 2002.874 & 0.486 & 0.938 & fast \\
+Stratford-on-Avon & West Midlands & 2002.879 & 0.470 & 0.945 & fast \\
+Arun & South East & 2002.882 & 0.526 & 0.938 & fast \\
+Fenland & East of England & 2002.891 & 0.632 & 0.918 & fast \\
+Monmouthshire & NA & 2002.893 & 0.501 & 0.935 & fast \\
+Denbighshire & NA & 2002.901 & 0.457 & 0.947 & fast \\
+Southend-on-Sea & East of England & 2002.903 & 0.551 & 0.932 & fast \\
+High Peak & East Midlands & 2002.903 & 0.488 & 0.946 & fast \\
+Pembrokeshire & NA & 2002.905 & 0.387 & 0.958 & fast \\
+Malvern Hills & West Midlands & 2002.907 & 0.524 & 0.940 & fast \\
+Camden & London & 2002.915 & 0.599 & 0.941 & fast \\
+North Devon & South West & 2002.926 & 0.426 & 0.950 & fast \\
+Wyre & North West & 2002.927 & 0.553 & 0.927 & fast \\
+Darlington & North East & 2002.933 & 0.615 & 0.929 & fast \\
+Swale & South East & 2002.940 & 0.447 & 0.948 & fast \\
+Welwyn Hatfield & East of England & 2002.951 & 0.628 & 0.922 & fast \\
+Mid Devon & South West & 2002.960 & 0.485 & 0.946 & fast \\
+Lewes & South East & 2002.964 & 0.507 & 0.945 & fast \\
+Dorset & South West & 2002.968 & 0.407 & 0.958 & fast \\
+Brighton and Hove & South East & 2002.976 & 0.578 & 0.934 & fast \\
+Braintree & East of England & 2002.978 & 0.534 & 0.938 & fast \\
+Epsom and Ewell & South East & 2002.979 & 0.746 & 0.908 & fast \\
+Sefton & North West & 2002.980 & 0.616 & 0.928 & fast \\
+Eastleigh & South East & 2002.992 & 0.595 & 0.929 & fast \\
+Craven & Yorkshire and The Humber & 2002.993 & 0.379 & 0.965 & fast \\
+Hammersmith and Fulham & London & 2002.997 & 0.585 & 0.944 & fast \\
+Greenwich & London & 2003.014 & 0.571 & 0.930 & slow \\
+Cheltenham & South West & 2003.025 & 0.602 & 0.937 & slow \\
+Worcester & West Midlands & 2003.041 & 0.659 & 0.920 & slow \\
+Tower Hamlets & London & 2003.043 & 0.660 & 0.926 & slow \\
+New Forest & South East & 2003.046 & 0.554 & 0.941 & slow \\
+East Renfrewshire & NA & 2003.046 & 0.739 & 0.900 & slow \\
+Fylde & North West & 2003.056 & 0.501 & 0.946 & slow \\
+Carmarthenshire & NA & 2003.067 & 0.390 & 0.958 & slow \\
+West Lancashire & North West & 2003.076 & 0.547 & 0.937 & slow \\
+Hastings & South East & 2003.086 & 0.685 & 0.911 & slow \\
+Carlisle & North West & 2003.111 & 0.541 & 0.934 & slow \\
+Wychavon & West Midlands & 2003.115 & 0.570 & 0.935 & slow \\
+Isle of Anglesey & NA & 2003.144 & 0.403 & 0.960 & slow \\
+Dudley & West Midlands & 2003.150 & 0.678 & 0.925 & slow \\
+Ryedale & Yorkshire and The Humber & 2003.152 & 0.380 & 0.961 & slow \\
+Rother & South East & 2003.162 & 0.366 & 0.966 & slow \\
+Leicester & East Midlands & 2003.167 & 0.686 & 0.924 & slow \\
+Breckland & East of England & 2003.182 & 0.567 & 0.938 & slow \\
+Tandridge & South East & 2003.190 & 0.747 & 0.918 & slow \\
+Shropshire & West Midlands & 2003.206 & 0.621 & 0.928 & slow \\
+West Devon & South West & 2003.228 & 0.489 & 0.949 & slow \\
+Milton Keynes & South East & 2003.231 & 0.700 & 0.929 & slow \\
+Sandwell & West Midlands & 2003.234 & 0.760 & 0.916 & slow \\
+Winchester & South East & 2003.234 & 0.443 & 0.956 & slow \\
+Redditch & West Midlands & 2003.238 & 0.706 & 0.923 & slow \\
+Rugby & West Midlands & 2003.265 & 0.709 & 0.922 & slow \\
+North Somerset & South West & 2003.278 & 0.558 & 0.939 & slow \\
+West Lindsey & East Midlands & 2003.285 & 0.674 & 0.924 & slow \\
+Richmondshire & Yorkshire and The Humber & 2003.286 & 0.425 & 0.958 &
+slow \\
+Melton & East Midlands & 2003.292 & 0.534 & 0.946 & slow \\
+Chichester & South East & 2003.295 & 0.582 & 0.942 & slow \\
+Torridge & South West & 2003.301 & 0.532 & 0.942 & slow \\
+Dundee City & NA & 2003.303 & 0.568 & 0.939 & slow \\
+South Ayrshire & NA & 2003.303 & 0.626 & 0.931 & slow \\
+East Devon & South West & 2003.308 & 0.492 & 0.947 & slow \\
+Bolton & North West & 2003.323 & 0.774 & 0.913 & slow \\
+Hertsmere & East of England & 2003.326 & 0.726 & 0.924 & slow \\
+Bournemouth, Christchurch and Poole & South West & 2003.347 & 0.662 &
+0.936 & slow \\
+Maldon & East of England & 2003.349 & 0.495 & 0.951 & slow \\
+Barnet & London & 2003.352 & 0.766 & 0.919 & slow \\
+Brent & London & 2003.389 & 0.762 & 0.925 & slow \\
+Midlothian & NA & 2003.413 & 0.732 & 0.921 & slow \\
+Boston & East Midlands & 2003.416 & 0.593 & 0.933 & slow \\
+Scarborough & Yorkshire and The Humber & 2003.421 & 0.431 & 0.963 &
+slow \\
+Kirklees & Yorkshire and The Humber & 2003.435 & 0.725 & 0.925 & slow \\
+Tameside & North West & 2003.441 & 0.819 & 0.907 & slow \\
+Newport & NA & 2003.449 & 0.635 & 0.930 & slow \\
+Glasgow City & NA & 2003.451 & 0.759 & 0.922 & slow \\
+Clackmannanshire & NA & 2003.454 & 0.585 & 0.936 & slow \\
+Three Rivers & East of England & 2003.458 & 0.628 & 0.947 & slow \\
+Harborough & East Midlands & 2003.488 & 0.549 & 0.950 & slow \\
+Rochdale & North West & 2003.496 & 0.685 & 0.937 & slow \\
+Tonbridge and Malling & South East & 2003.497 & 0.710 & 0.927 & slow \\
+Somerset West and Taunton & South West & 2003.502 & 0.491 & 0.955 &
+slow \\
+Wrexham & NA & 2003.512 & 0.553 & 0.942 & slow \\
+Worthing & South East & 2003.522 & 0.622 & 0.938 & slow \\
+Tewkesbury & South West & 2003.540 & 0.580 & 0.946 & slow \\
+City of Edinburgh & NA & 2003.541 & 0.685 & 0.941 & slow \\
+Coventry & West Midlands & 2003.592 & 0.564 & 0.948 & slow \\
+Northumberland & North East & 2003.607 & 0.508 & 0.954 & slow \\
+Mid and East Antrim & NA & 2003.609 & 0.893 & 0.900 & slow \\
+Wigan & North West & 2003.617 & 0.755 & 0.929 & slow \\
+Lichfield & West Midlands & 2003.629 & 0.625 & 0.939 & slow \\
+Ashford & South East & 2003.633 & 0.535 & 0.944 & slow \\
+Cornwall & South West & 2003.638 & 0.443 & 0.963 & slow \\
+Mendip & South West & 2003.652 & 0.505 & 0.955 & slow \\
+Stevenage & East of England & 2003.665 & 0.525 & 0.958 & slow \\
+Blaenau Gwent & NA & 2003.666 & 0.643 & 0.933 & slow \\
+Lincoln & East Midlands & 2003.666 & 0.634 & 0.940 & slow \\
+South Somerset & South West & 2003.676 & 0.553 & 0.948 & slow \\
+Charnwood & East Midlands & 2003.677 & 0.699 & 0.934 & slow \\
+North Ayrshire & NA & 2003.682 & 0.607 & 0.942 & slow \\
+Folkestone and Hythe & South East & 2003.684 & 0.792 & 0.905 & slow \\
+Eastbourne & South East & 2003.684 & 0.697 & 0.930 & slow \\
+Blackburn with Darwen & North West & 2003.685 & 0.680 & 0.938 & slow \\
+Dover & South East & 2003.699 & 0.574 & 0.948 & slow \\
+Adur & South East & 2003.735 & 0.551 & 0.954 & slow \\
+Haringey & London & 2003.739 & 0.667 & 0.943 & slow \\
+Gosport & South East & 2003.751 & 0.600 & 0.934 & slow \\
+Trafford & North West & 2003.752 & 0.640 & 0.949 & slow \\
+Norwich & East of England & 2003.756 & 0.742 & 0.939 & slow \\
+West Lothian & NA & 2003.758 & 0.652 & 0.943 & slow \\
+North East Lincolnshire & Yorkshire and The Humber & 2003.811 & 0.858 &
+0.920 & slow \\
+Bristol, City of & South West & 2003.828 & 0.831 & 0.928 & slow \\
+Rushcliffe & East Midlands & 2003.839 & 0.591 & 0.954 & slow \\
+Hinckley and Bosworth & East Midlands & 2003.857 & 0.749 & 0.934 &
+slow \\
+Torfaen & NA & 2003.861 & 0.841 & 0.922 & slow \\
+Swansea & NA & 2003.875 & 0.650 & 0.947 & slow \\
+Wolverhampton & West Midlands & 2003.882 & 0.793 & 0.931 & slow \\
+Kingston upon Hull, City of & Yorkshire and The Humber & 2003.885 &
+0.757 & 0.936 & slow \\
+Calderdale & Yorkshire and The Humber & 2003.889 & 0.714 & 0.939 &
+slow \\
+North Northamptonshire & East Midlands & 2003.902 & 0.746 & 0.935 &
+slow \\
+Stockton-on-Tees & North East & 2003.902 & 0.955 & 0.916 & slow \\
+Wealden & South East & 2003.906 & 0.736 & 0.938 & slow \\
+Thanet & South East & 2003.953 & 0.667 & 0.947 & slow \\
+Broadland & East of England & 2003.962 & 0.727 & 0.938 & slow \\
+West Northamptonshire & East Midlands & 2003.972 & 0.736 & 0.940 &
+slow \\
+North Kesteven & East Midlands & 2003.976 & 0.633 & 0.942 & slow \\
+Neath Port Talbot & NA & 2003.981 & 0.767 & 0.933 & slow \\
+Rhondda Cynon Taf & NA & 2003.991 & 0.769 & 0.925 & slow \\
+Enfield & London & 2003.992 & 0.588 & 0.954 & slow \\
+East Lindsey & East Midlands & 2004.010 & 0.577 & 0.952 & slow \\
+Teignbridge & South West & 2004.015 & 0.681 & 0.941 & slow \\
+York & Yorkshire and The Humber & 2004.025 & 0.739 & 0.944 & slow \\
+Harlow & East of England & 2004.030 & 0.803 & 0.936 & slow \\
+Vale of Glamorgan & NA & 2004.033 & 0.592 & 0.951 & slow \\
+Newark and Sherwood & East Midlands & 2004.034 & 0.668 & 0.946 & slow \\
+Merton & London & 2004.041 & 0.932 & 0.922 & slow \\
+Hambleton & Yorkshire and The Humber & 2004.065 & 0.636 & 0.953 &
+slow \\
+Bedford & East of England & 2004.068 & 0.852 & 0.933 & slow \\
+Stockport & North West & 2004.075 & 0.757 & 0.942 & slow \\
+South Ribble & North West & 2004.126 & 0.674 & 0.948 & slow \\
+Uttlesford & East of England & 2004.144 & 0.905 & 0.926 & slow \\
+Rossendale & North West & 2004.160 & 0.976 & 0.923 & slow \\
+Oldham & North West & 2004.162 & 0.801 & 0.938 & slow \\
+Sheffield & Yorkshire and The Humber & 2004.189 & 0.770 & 0.944 &
+slow \\
+Ribble Valley & North West & 2004.204 & 0.718 & 0.945 & slow \\
+Maidstone & South East & 2004.222 & 0.854 & 0.932 & slow \\
+Cheshire West and Chester & North West & 2004.234 & 0.742 & 0.947 &
+slow \\
+Thurrock & East of England & 2004.241 & 0.841 & 0.924 & slow \\
+Telford and Wrekin & West Midlands & 2004.241 & 1.024 & 0.917 & slow \\
+Cardiff & NA & 2004.246 & 0.823 & 0.937 & slow \\
+County Durham & North East & 2004.250 & 0.626 & 0.953 & slow \\
+Plymouth & South West & 2004.252 & 0.785 & 0.934 & slow \\
+Bridgend & NA & 2004.256 & 0.547 & 0.956 & slow \\
+Rotherham & Yorkshire and The Humber & 2004.273 & 0.722 & 0.938 &
+slow \\
+Hyndburn & North West & 2004.298 & 0.881 & 0.932 & slow \\
+Newcastle upon Tyne & North East & 2004.305 & 0.906 & 0.932 & slow \\
+Wakefield & Yorkshire and The Humber & 2004.314 & 0.905 & 0.923 &
+slow \\
+Havering & London & 2004.345 & 0.822 & 0.938 & slow \\
+South Kesteven & East Midlands & 2004.353 & 1.026 & 0.915 & slow \\
+Canterbury & South East & 2004.354 & 0.602 & 0.960 & slow \\
+Broxbourne & East of England & 2004.363 & 0.817 & 0.944 & slow \\
+Peterborough & East of England & 2004.382 & 0.937 & 0.932 & slow \\
+Redbridge & London & 2004.411 & 1.028 & 0.925 & slow \\
+Bury & North West & 2004.431 & 0.735 & 0.953 & slow \\
+Harrogate & Yorkshire and The Humber & 2004.433 & 0.971 & 0.927 &
+slow \\
+Birmingham & West Midlands & 2004.435 & 1.182 & 0.912 & slow \\
+South Tyneside & North East & 2004.446 & 0.636 & 0.948 & slow \\
+Chorley & North West & 2004.462 & 1.093 & 0.913 & slow \\
+North East Derbyshire & East Midlands & 2004.483 & 0.703 & 0.953 &
+slow \\
+Lambeth & London & 2004.485 & 0.958 & 0.937 & slow \\
+Doncaster & Yorkshire and The Humber & 2004.496 & 0.672 & 0.953 &
+slow \\
+South Gloucestershire & South West & 2004.499 & 0.881 & 0.937 & slow \\
+Caerphilly & NA & 2004.499 & 0.819 & 0.938 & slow \\
+East Staffordshire & West Midlands & 2004.501 & 1.132 & 0.910 & slow \\
+North Tyneside & North East & 2004.509 & 0.864 & 0.933 & slow \\
+South Holland & East Midlands & 2004.534 & 0.851 & 0.941 & slow \\
+Gateshead & North East & 2004.535 & 0.869 & 0.934 & slow \\
+St.~Helens & North West & 2004.542 & 1.097 & 0.923 & slow \\
+Flintshire & NA & 2004.549 & 0.764 & 0.951 & slow \\
+Selby & Yorkshire and The Humber & 2004.549 & 0.716 & 0.947 & slow \\
+Belfast & NA & 2004.556 & 1.332 & 0.906 & slow \\
+Sedgemoor & South West & 2004.613 & 0.540 & 0.966 & slow \\
+Exeter & South West & 2004.617 & 0.964 & 0.936 & slow \\
+South Derbyshire & East Midlands & 2004.629 & 0.934 & 0.938 & slow \\
+Derbyshire Dales & East Midlands & 2004.668 & 0.648 & 0.963 & slow \\
+East Riding of Yorkshire & Yorkshire and The Humber & 2004.683 & 0.728 &
+0.955 & slow \\
+Basildon & East of England & 2004.691 & 1.067 & 0.925 & slow \\
+Barking and Dagenham & London & 2004.800 & 1.198 & 0.916 & slow \\
+Causeway Coast and Glens & NA & 2004.834 & 1.024 & 0.928 & slow \\
+North Lanarkshire & NA & 2004.839 & 0.878 & 0.945 & slow \\
+Chesterfield & East Midlands & 2004.887 & 1.173 & 0.927 & slow \\
+Newcastle-under-Lyme & West Midlands & 2004.933 & 1.172 & 0.925 &
+slow \\
+East Ayrshire & NA & 2004.934 & 1.280 & 0.903 & slow \\
+Nuneaton and Bedworth & West Midlands & 2004.956 & 0.950 & 0.941 &
+slow \\
+Blaby & East Midlands & 2004.959 & 1.059 & 0.929 & slow \\
+Leeds & Yorkshire and The Humber & 2004.979 & 0.836 & 0.956 & slow \\
+Erewash & East Midlands & 2004.987 & 0.939 & 0.944 & slow \\
+Bassetlaw & East Midlands & 2004.998 & 0.982 & 0.944 & slow \\
+Hackney & London & 2005.026 & 1.157 & 0.934 & slow \\
+Barnsley & Yorkshire and The Humber & 2005.055 & 0.912 & 0.942 & slow \\
+Amber Valley & East Midlands & 2005.105 & 1.017 & 0.941 & slow \\
+Colchester & East of England & 2005.144 & 1.250 & 0.927 & slow \\
+Ashfield & East Midlands & 2005.176 & 1.073 & 0.942 & slow \\
+Wirral & North West & 2005.184 & 1.084 & 0.939 & slow \\
+Knowsley & North West & 2005.186 & 1.160 & 0.923 & slow \\
+Windsor and Maidenhead & South East & 2005.195 & 1.746 & 0.902 & slow \\
+Bromsgrove & West Midlands & 2005.494 & 1.144 & 0.941 & slow \\
+Liverpool & North West & 2005.526 & 1.276 & 0.936 & slow \\
+Staffordshire Moorlands & West Midlands & 2005.583 & 1.426 & 0.924 &
+slow \\
+Islington & London & 2005.970 & 1.223 & 0.955 & slow \\
+Warrington & North West & 2006.015 & 1.387 & 0.942 & slow \\
+Newham & London & 2006.078 & 1.908 & 0.907 & slow \\
+Bolsover & East Midlands & 2006.188 & 1.143 & 0.957 & slow \\
+Lisburn and Castlereagh & NA & 2006.193 & 1.970 & 0.903 & slow \\
+Manchester & North West & 2006.449 & 1.434 & 0.950 & slow \\
+Sunderland & North East & 2006.479 & 1.679 & 0.928 & slow \\
+South Staffordshire & West Midlands & 2006.490 & 1.837 & 0.922 & slow \\
+Rutland & East Midlands & 2006.529 & 0.986 & 0.967 & slow \\
+Cannock Chase & West Midlands & 2006.700 & 2.231 & 0.917 & slow \\
+Mansfield & East Midlands & 2006.889 & 1.579 & 0.946 & slow \\
+Ards and North Down & NA & 2006.947 & 2.485 & 0.907 & slow \\
+Halton & North West & 2007.069 & 0.932 & 0.967 & slow \\
+Burnley & North West & 2007.372 & 2.286 & 0.919 & slow \\
+Middlesbrough & North East & 2010.074 & 5.297 & 0.900 & slow \\
+Isles of Scilly & South West & 2012.163 & 4.257 & 0.953 & slow \\
+\end{longtable}
 

### 2025-01-03 paper/submission2/response.qmd
ssh problem

---
title: "Response to editor and reviewer comments"
format: pdf
bibliography: bibliography.bib
---

18-Dec-2024

Dear Emmanouil

Thank you for submitting your manuscript to Journal of Economic Geography.

I have now received comments from reviewers. Their comments are at the end of this email. The reviewers raise a range of concerns and have questions about the suitability of the paper for publication. The overall feeling is that the paper has probably been submitted too early and needs major work on most sections if it is to become publishable. I concur with the reviewers and see a number of significant issues that need to be addressed. This left me with a dilemma in terms of the decision. I have decided to invite you to revise and resubmit the paper so that it can be considered further for publication. However, this is a very high-risk revision, and I was bordering on rejection given the scale of the work needed and the limitations that the reviewers highlight. I am saying this to be open about the risk that after the next round the paper could be rejected unless reviewers believe you have addressed the range of issues identified. You may want to factor this into your decision about whether to revise the paper, given how much work is needed.

Having said this, let me offer some suggestions about the main areas to focus on if you decide to revise the paper. Reviewer 1 is more positive but very brief in their report. I am, therefore, putting more weight on reviewers 2 and 3 when identifying priorities for the revision.

## Editor; theory and justification 
\color{teal}
First, you will need to develop a more convincing theoretical argument and contribution. You will see that both reviewers 2 and 3 raise questions about your claims in relation to the gap in the literature and suggest you fail to build on existing work appropriately and do not convincingly identify the motivation for this analysis. In general, claims that an issue has never been addressed are hard to sustain and it is better to build on existing work and extend knowledge relating to a particular theoretical issue. The reviewers challenge you to do this using your data and suggest both the argument and structure of the paper will need to change.
\color{black}

## Editor; data
\color{teal}
Second, you need to justify your data and analysis better. Reviewer 2 fairly questions the end point of the analysis – which is not explained and given the 10+ years gap to your submission raises questions about the timeliness of the analysis. The questions about coverage of the UK website population also need to be addressed convincingly. More generally, the reviewers ask for more careful presentation of the data to make clear the significance of each figure and finding. As it stands, the paper feels rather overloaded with data, rather than acting as a focused analysis.
\color{black}

## Editor; conclusions
\color{teal}
The problems identified above result in the paper feeling imprecise, confusing and lacking a contribution. The end point is a conclusion that the reviewers do not find justified or original. The reviewers do see potential. As I noted, the feeling is a paper that is underdeveloped and submitted prematurely. There is, then, a route to publication, but it will require significant changes to convince the reviewers.
\color{black}

## Referee: 1

I have read your paper with interest and I appreciate its quality and clarity. The work is both well-executed and well-written, making it a valuable contribution to the field.

However, I would like to offer a few minor suggestions that could further enhance the readability and overall coherence of the paper:

> Simplify the Introduction: I recommend streamlining the introduction by reducing the inclusion of overly technical or highly specific details. Instead, consider expanding on the broader conceptual framework that situates the paper within its scholarly context. This approach would help orient readers and provide a more accessible foundation for understanding the study.
*Simplify the introduction*

Strengthen the Connection Between Hypotheses and Results: It may be beneficial to establish a clearer and more explicit link between the study's hypotheses and its findings. One way to achieve this could be by using section titles that are more conceptually informative, which would guide the reader through the logical progression of the argument and findings more effectively.
*change section titles?*

## Referee: 2

\color{teal}
OVERVIEW\
- focused on innovation diffusion, using the case of the web in the UK
- “this paper tests whether the three distinct spatial diffusion mechanisms shaped the diffusion of the Web in the UK at very granular spatial scales from the Web’s early stages onwards”; presumably “S-shaped pattern”/temporal models, core to periphery; distance/proximity.

Figure 2 shows the diffusion of domains per company at the scale of the UK matches the S growth model. Figure 3 shows some spatial clustering. I think this is for all years in the dataset but this could be better labeled
\color{black}

**explain, change title**

\color{teal}
Figure 4, not sure what this demonstrates
\color{black}

\color{teal}
Figure 5&6, shows early clustering and then diffusion (I think)
\color{black}

**yes**

\color{teal}
DATA\
Focusing just on .co.uk website is OK but needs to be better justified given that it only represented about 50 percent of domains in the UK at that time. See my back of the envelop calculations: As per Zook (2001, Table 1) .uk represented 66% of all domains in the UK, .co.uk is a subset of that, so this is perhaps 50% of all domains registered in the UK at that time. If 1/5 of .co.uk domains had findable postal codes (p. 5), this study is based on 10% of the domains in the UK. (perhaps this is too low, aggregating from webpages to websites brought the number higher, not entirely clear to me)
\color{black}

Thanks for going the extra mile to check the data used here! I agree with the 
above calculation up until the findable postcode point as 1/5 of the archived 
*webpages* -- not *websites*, i.e. domains -- included a postcode in the web text. 
Think of the website of a small company with one trading address. I think it is 
fair to expect that website to have multiple webpages to present the different
products and services and then one webpage to showcase the address -- the 
'Contact us' page. So, the 1/5 share of *webpages* containing a postcode should
translate to a much higher share of *websites* -- i.e. domains -- containing 
a postcode. Previous estimates indicated that around 70% of all websites contain 
place reference [@hill2009georeferencing] and this reference has no been added 
in the paper.

So, based on the above 35% of all domains with a registered address in the UK
were represented in the data. Most importantly though the paper does not model 
the adoption of the Web from the general population but instead from firms. So, 
although the data only represents 35% of the population of websites with a 
registered address in the UK, it represents almost all the *relevant* archived
websites -- that is websites using the commercial .co.uk SLD. As stated in the 
paper, this has been the primary choice for commercial websites in the UK.

The third paragraph in the Data and Methods section includes the following text:

> Instead of using the whole .uk ccTLD, this paper focuses on its
commercial subset, the .co.uk second level domain (SLD). This choice
is aligned with the topic of the paper -- the diffusion of the Web for 
commercial functions and decreases the heterogeneity of the web data as such 
commercial websites have specific aims: they are used to diffuse information, 
support online transactions and share opinions [@THELWALL2000441; 
@blazquez2018big].

<!-- Active/inactive, don;t suffer from multiple registrations -->

\color{teal}
overall the time period 1996-2012 is a good time to study the diffusion of the commercial internet. But why stop at 2012? Is it when the Library stopped its data series?
\color{black}

Yes, the British Library had obtained a copy of all of the .uk contents of the 
Internet Archive for that period. From a data perspective, it wouldn't be correct
to extend this dataset with data from other open archives such as the Common 
Crawl as they use different crawlers and there will be inconsistencies. Most
importantly though, this period covers the diffusion of the web from the very
early stages until its maturity. This is also illustrated by the figure 
depicting the cumulative adoption and the flattening right end of the line.

There is a specific paragraph justifying how the time period of the data fits
the aim of the paper and another line is now included in the second paragraph 
of the introductions ("from the very early days until its maturity (1996-2012)").

\color{teal}
Framing
“With the exception of Sinai and Waldfogel (2004), who focused, among other things, on web content and demonstrated that more such content is available in larger markets, this strand of literature has been mostly concerned with the spatiality of the Internet’s hard and tangible infrastructure and shied away from exploring the spatial dimensions of one of the most distinct software layers of the Internet infrastructure, the Web.” This statement is simply not true, the literature reviews misses the entirety of web content diffusion studies: Zook, 2000, Malecki, 2002. Townsend, 2000, Gorman, 2002, Feldman, 2002, lots more
\color{black}

Point taken. Despite being aware of these studies, I omitted citing them as I 
was primarily searching for studies testing the diffusion mechanisms over longer
time periods. This error has now been corrected and the above paragraph has been 
replaced with the following one, which emphasises the long time series motivation. 
Regarding the actual references, @zook2000web, @zook2000web and @moss1997tracking 
used relevant, domain name registration data, but the rest (e.g. @malecki2002economic 
and @moss2000internet) focused on hard infrastructure. @gorman2002web examined 
the head and branch locations of the top 40 e-business integration firms in the 
USA, which although interesting is not directly relevant to this paper. I also
went through of Feldman's work around 2002, but I did not find papers
adopting a similar empirical perspective.

> Another line of motivation has to do with the very nature of the Web as a 
digital technology and how it evolved over time. The digital geographies field 
has debunked premature claims about the Internet's placeless-ness by mapping 
various of its facets and technological layers [@tranos2013death], but the lack 
of relevant long time-series data prevented researchers from observing digital
phenomena overtime. Most of these studies focused on its hard infrastructure 
[e.g. @malecki2002economic; @moss2000internet], on Internet subscribers 
[e.g. @blank2018local], or on social media users [e.g. @crampton2013beyond].
Exceptions include early work from @zook2000web, @zook2000web and 
@moss1997tracking, which utilised domain name registration data and use it to
map and illustrate the spatial footprint of the web content. Pivotal as these 
contributions have been in understanding the economic geographies of the 
Internet, they were based on cross-sectional, often spatially detailed, mapping 
and on shallow time comparisons (e.g. 1993-1998) and did not observe the 
evolution and diffusion of the Web until its maturity. As we know from the 
relevant literature, different technologies diffuse at different rates and 
there is no theory that can explain these differences 
[@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
Web among firms can offer useful insights.

\color{teal}
“for a digital technology as weightless as the Web” – this runs a bit counter to lots of work by digital geographers who have shown again and again how the digital is not weightless or placeless
\color{black}

\color{teal}
Conclusion
p. 18 “This is the first time that the importance of granular spatial mechanisms for the diffusion of an intangible, digital technology is exposed.” Honestly, I would not expect the diffusion of the web to be any different than any other kind of innovation diffusion. The spatiality of digital life, and its connection to material spaces is a constant refrain within digital geography. So I’m not convinced it makes much sense to emphasize this, if a granular spatial analysis of innovation diffusion is already novel.
\color{black}

**for the last two comments, I need to soften the weightless argument throughout.**
I have added a long time series argument and also said that there is heterogeneity
[@leibowicz2016representing] among diffusion rates of different technologies.

As per the 'framing' point above, the digital motivation para in the 
introduction has been amended acordilngly.

## Referee: 3

\color{teal}
This paper addresses a gap in Economic Geography by examining the spatial diffusion of digital technologies, specifically the adoption of commercial websites, from 1996 to 2012. Recognizing the potential of digital technologies to shape local development trajectories through increased productivity and innovation, the study employs novel data sources and machine learning techniques to analyze diffusion patterns at highly granular geographical scales. The authors explore the roles of various spatial mechanisms driving this process, offering insights into how digital technology adoption unfolds across space and time.
The paper has the potential to make an interesting contribution to the literature as it combines a fascinating topic with novel data and methodologies. However, in its current form, it is not ready for publication, as it never becomes clear what it actually delivers in terms of new knowledge and what knowledge gap it seeks to fill. Despite some analytical elements, it is rather descriptive in nature, which doesn’t fit well with this journal.

Most importantly, I suggest developing some clear hypotheses to structure and motivate the paper better. In light of this, I recommend major revisions.
\color{black}

This is very helpful. The first two sections have been restructured around two 
hypotheses. I believe the structure and the flow of the paper is much clearer
now.

\color{teal}
\noindent 1. Abstract:\
\noindent 1. The abstract needs significant improvement. It needs to adequately report the main findings, relevance, or implications of the study. Additionally, the assertion that geographers have "shied away" from analyzing the diffusion of digital technologies is misleading; the issue was primarily due to the lack of available data.
\color{black}

**fix the abstract**

"shied away" has been replaced in the opening paragraph.

\color{teal}
\noindent 2. Introduction\
\noindent 2. General Diffusion Studies: The statement that there are few contemporary general diffusion studies in Economic Geography is not accurate, although it may apply specifically to digital technologies. Consider the following recent examples:
o Peris, A., Meijers, E., & van Ham, M. (2021). Information diffusion between Dutch cities: Revisiting Zipf and Pred using a computational social science approach. Computers, Environment and Urban Systems, 85(November), 101565. https://doi.org/10.1016/j.compenvurbsys.2020.101565
o Bednarz, M., & Broekel, T. (2020). Pulled or pushed? The spatial diffusion of wind energy between local demand and supply. Industrial and Corporate Change, 29(4), 893–916. https://doi.org/10.1093/icc/dtaa012
o Feldman, M. P., Kogler, D. F., & Rigby, D. L. (2015). rKnowledge: The Spatial Diffusion and Adoption of rDNA Methods. Regional Studies, 49(5), 798–817. https://doi.org/10.1080/00343404.2014.980799
\color{black}

@PERIS2021101565 is indeed an interesting paper, but it is not directly relevant 
as it focuses on information diffusion rather than diffusion of new 
technologies. Information diffuses through social and, consequently, urban 
networks, but it is not adopted like new technologies. @bednarz2020pulled was 
already cited and @feldman2015rknowledge has now been added. These references 
are included in the second section.

Regarding the statement about economic geography studies on diffusion of 
new technologies and innovation, I do believe that this is a fair statement. 
The references listed below are the ones reviewed in the second section and none 
of them are published in one of the flagship economic geography journals 
(*Journal of Economic Geography* and *Economic Geography*). There are two papers
published in *The Professional Geographer*, which is a broad-scope geography 
journal and then only one paper in *Regional Studies*, which is the closest to
economic geography. Importantly though, the *Regional Studies* paper 
[@feldman2015rknowledge] paper is focused on inventions as observed by patent 
data contrary to the broader 'adoption of a new technology' approach of this 
paper. All other papers listed below are published in business and economics 
journals.

1. @beardsell1999spatial: European Economic Review
2. @bednarz2020pulled: Industrial and Corporate Change
3. @haller2011determinants: Applied Economics
4. @perkins2011internet: The Professional Geographer
5. @feldman2015rknowledge: Regional Studies
6. @ding2010modeling: The Professional Geographer
7. @bakher2013diffusion: Journal of Marketing Analytics
8. @PAPAGIANNIDIS2015308: Technological Forecasting and Social Change
9. @lengyel2020role: Scientific reports

Also, the lack of spatial granularity argument has been highlighted in various 
instances throughout the paper as the main line of justification.

**ADD SPATIAL GRANULARITY IN CONCLUSIONS**

\color{teal}
\noindent 3. Novelty: The study needs to be more explicit about its novel contributions to understanding spatial diffusion mechanisms. Currently, the discussion is too general.
\color{black}

**granularity**

**HAVE A THINK, intro has been restructured**

\color{teal}
\noindent 4. Motivation: The motivation for the study is unclear. The text raises several issues, but it is unclear which are specific to digital technologies and which pertain to technological diffusion more broadly.
\color{black}

**HAVE A THINK, intro has been restructured**

\color{teal}
\noindent 5. Structure and Focus: The discussion on motivations should be streamlined and clarified. For example, after specifying that the study focuses on commercial websites rather than the web, subsequent references revert to the general term “web.” Additionally, the discussion confuses diffusion patterns, firm perspectives, and early adopter advantages. Consider restructuring this section to clearly differentiate between literature gaps (motivations) and expected outcomes (mechanisms).
\color{black}

The revised paper consistently illustrates that the focus is on commercial 
websites.

In the relevant paragrapg in the Data and Methods section, the following line
has been added: "Henceforward any reference to websites will refer to commercial 
-- .co.uk -- websites".

**fixed for intro, lit rev  and data -- check the rest of the paper.**

The introduction and literature review sections have been restructured, so 
the firm perspectives and early adopter advantages discussions have been moved.

\color{teal}
\noindent 6. Data Consistency: The terminology for the data should be consistent. At one point, the text refers to the “totality of the web,” while elsewhere, it specifies “commercial websites.” Clarify and maintain consistency throughout.
\color{black}

The terminology is now consistent across the paper. The "totality" point 
referred to different web technologies and not to the totality of the web. It 
has now been replaced with "an overarching technology".

\color{teal}
\noindent 7. Recommended Structure: To align with the common structure in Journal of Economic Geography (JoEG), consider splitting the introduction into a shorter introduction and a dedicated theory section merged with the literature review.
\color{black}

The introduction is now shorter and the literature review section has been 
re-structured and re-named as "The mechanisms that shape the diffusion of a new 
technology".

\color{teal}
\noindent 3. Literature Review / Theory\
\noindent 1. Synthesis and Motivation: The literature review should be transformed into a theory section. Rather than presenting a historical review, synthesize the arguments to clearly motivate and inform your hypotheses.
\color{black}

Done. Restructured: theory, gaps, hypotheses.

\color{teal}
\noindent 2. Knowledge Gaps: The current review does not clearly identify the knowledge gaps that this study aims to address. Clarify whether the goal is to confirm known processes with new data and methods or to explore unique diffusion patterns related to websites.
\color{black}

Known processes at very detailed scale: explanatory power of these mechanisms at granular, local
scales

\color{teal}
\noindent 3. Lack of Hypotheses: To clarify what the paper delivers and what knowledge gaps it fills, I strongly recommend developing a number of hypotheses that will be tested.
\color{black}

Added.

\color{teal}
\noindent 4. Fit to title: The term “multi-scalar” appears in the title but is never really discussed in the paper.
\color{black}

This was discussed from the data section onwards, but indeed not before that. 
Section 2 includes a new paragraph describing the second hypothesis (H~2~), 
which explicitely discusses the value of the multiscale approach: 
"Understanding the diffusion of a new technology in multiple, detailed spatial
scales is of interest to economic geography as ..."

\color{teal}
\noindent 4. Data & Methods\
\noindent 1. Order of Presentation: Introduce the data first, followed by the geographical units used in the study.
\color{black}

The order of presentation has been amended accordingly and the two scales of 
analyses are presented after the data.

\color{teal}
\noindent 2. Postcode Identification: The approach of identifying postcodes on websites is innovative, but potential errors need to be discussed. Specifically:
- How do you ensure that a postcode on a website accurately represents the 
company’s location or its subsidiary?
- How do you handle numbers that resemble postcodes but are not actual postcodes?
- What about references to events or other organizations unrelated to the website’s owner?
Include a brief discussion of these potential issues.
\color{black}

A number of different strategies have been developed to address the potential 
issues described above.

\noindent 1. To prevent false positives, the resulted postcodes
were matched against official postcode lookup files [@ons]. This was not 
explicitly mentioned in the first submission on the paper and the previous 
sentence has now been added in the relevant paragraph:

> The text from these webpages was scanned using a regular expression
(regex) to identify strings of text which resemble UK postcodes and one
fifth of them included a mention to a postcode [@BL2013geo]. This
information allows the geolocation of the data and the creation of the
LAD and OA counts. To prevent false positives, the resulted postcodes
were matched against official postcode lookup files [@ons].

\noindent 2. Then, to remove the bias from link farms associated with a handful of postcodes
having irregular high numbers of websites pointing to them, a targeted data 
imputation procedure was developed. This is detailed in the data section -- see
also Figure 1.

\noindent 3. Regarding the points about subsidiary company location and event postings, 
these cannot be directly observed and distinguished. However, a process that has 
been successfully used in previous studies has been adopted here. Firstly, given 
that the focus of the paper is commercial websites, the relevant literature 
illustrates that such websites serve specific purposes:

> they are used to diffuse information, support online transactions and share 
opinions [@THELWALL2000441; @blazquez2018big].

Secondly, the webpage-to-website aggregation process leads to websites with 
different number of postcodes as presented in Table 1. To decrease heterogeneity, 
the analysis firstly focuses on websites with one unique postcode. Given the the
commercial nature of these websites (.co.uk), it is fair to expect that such 
websites with one unique postcode in their web text represent companies with a 
single trading location. The analysis included in the main body of the paper is
based on this subset of the website population. For year 2000, this includes 
70% of all the archived, geolocated commercial websites. Then, as a robustness
check all the different steps of the analysis are repeated for websites with up 
to 10 postcodes and the results are presented in the appendix. This includes 94%
of all the archived, geolocated commercial websites. For this step, websites 
with multiple postcodes were counted multiple times, one for each different 
postcode. All the results are qualitatively the same and only marginally 
different, which is indication of the robustness of the website geolocation 
strategy.

As a further argument for the robustness of the above process, the results of a 
previous study which used the same approach for a small subset of these data 
from a specific neighbourhood in London matched in terms of accuracy previous 
research that was based on extensive qualitative work, interviews and 
observational studies [@shoreditch]. This last point has now been added in the 
relevant section:

> Previous research which used the same methodology for a small subset of these 
data illustrates the robustness of the above approach: its results matched in 
terms of accuracy past rounds of research based on extensive qualitative work, 
interviews and observational studies for the same study area [@shoreditch].

\color{teal}
\noindent 3. Equation 2: The rationale behind Equation 2 is unclear. Shouldn’t the primary predictor of website density at time tbe the location density at t-1? Why was this approach chosen over a growth model? Additionally, clarify how spatial dependencies are accounted for in the model.
\color{black}

**growth as an appendix, spatial dependencies have been included**

\color{teal}
\noindent 5. Results\
\noindent 1. Diffusion Speed: The decision to binarize diffusion speed is questionable. Continuous measures, such as growth rates or the time required to reach 50% saturation, would provide more informative insights.
\color{black}

**This is only for the map, this is a critical point in diffusion studies**

\color{teal}
\noindent 2. Geographical Scale: While the effort to model diffusion curves at a fine geographical scale is commendable, the paper does not sufficiently explain the benefits or additional knowledge gained from this approach.
\color{black}

**reveal that theories are valid for lcoal scales and inform local tech policies**

\color{teal}
\noindent 3. Early Adopters and Latecomers: The discussion of early-adopter and latecomer advantages is underdeveloped. How do these advantages manifest locally? What specific benefits arise if a company registers its website six months earlier than competitors in other locations? If this aspect is retained, it should be elaborated in the theory section, including how it relates to local capabilities or diffusion hierarchies.
\color{black}

**relates to local capabilities or diffusion hierarchies**

\color{teal}
\noindent 4. Descriptive Statistics: The value of the descriptive Moran’s I and Gini coefficient analyses in Section 4.2 is unclear. Instead, calculating growth rates of adoption and examining their distribution over time would offer a better sense of adoption speed and magnitude.
\color{black}

**think, growth rates will not work for OA**

\color{teal}
\noindent 5. Retail Centers: The sudden shift to retail centers on page 15 is confusing, given that the data pertains to all commercial sites. Clarify the rationale for this focus.
\color{black}

**hierarchy**

\color{teal}
\noindent 6. Modeling Details: The methodological details in Section 4.3 should be moved to the methods section to maintain clarity and coherence. Including these details in the results section is confusing.
\color{black}

\color{teal}
\noindent 7. Clarity and Structure: Section 4.3 is difficult to follow. Clearly state the main message and highlight how diffusion mechanisms vary spatially. Develop hypotheses to structure the results and provide potential explanations for observed heterogeneity. Without this, the section feels like a descriptive summary without clear takeaways for the reader.
\color{black}

\color{teal}
\noindent 
Overall Recommendations
\noindent 1. Consistency: Ensure consistent terminology and focus throughout the paper, particularly when referring to the web and commercial websites.
\noindent 2. Structure: Adopt a clearer structure by separating the introduction, theory/literature review, methods, and results sections.
\noindent 3. Clarity: Develop hypotheses to frame the results and provide clearer explanations for methodological choices and findings.
\color{black}

## References

### 2025-01-03 paper/submission1/anonymised.qmd
ssh problem

--- 2024-09-12 paper/submission1/tranos2023.qmd
@@
 title: "A multi-scale story of the diffusion of a new technology: the web"
-author:
-  - name: Emmanouil Tranos
-    email: e.tranos@bristol.ac.uk
-    affiliations: 
-      - id: some-tech
-        name: University of Bristol
-        department: 
-        address: 
-        city: 
-        state: State
-        postal-code: UK
-    corresponding: e.tranos@bristol.ac.uk
-address:
-  - attributes:
-        corresponding: true
-    note: This is the first author footnote.
 abstract: |
-  This paper maps the participation in the digital economy and its evolution in the UK over space and time.
-Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects.
-Useful as these approaches might have been, they cannot capture the micro-processes and the characteristics of the individual online behaviour.
-Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK.
-The data and geolocation strategy allow to capture these processes at a very granular spatial scale.
-The modelling approach, which is based on simple spatial analytical methods and on the estimation of diffusion curves at various scales, enables to depict the role of geography and other cognitive factors which drove the diffusion of web technologies.
-Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.
+    Despite the broad consensus that differences in technological adoption and sophistication can lead to uneven trajectories of local development as new technologies can increase productivity and allow firms to introduce new products and services, economic geographers have shied away from exploring the diffusion of digital technologies across space and time.
+This paper fills this gap by exposing the roles of different spatial mechanisms in the diffusion of an intangible, digital technology at very granular geographical scales.
+It does so by employing a novel source of data and machine learning algorithms to model the active engagement with Web during 1996-2012.
 keywords: 
-  - keyword1
-  - keyword2
+  - diffusion
+  - technologies
+  - Web
+  - digital
 date: last-modified
@@
 Economic geographers were always interested in how new technologies and
-innovations diffuse across space and time with the work of @hagerstrand1968innovation,
-which demonstrated the importance of communications as a mechanism for the 
-diffusion of innovations, being the most impactful. It was followed by various 
-empirical studies -- see for instance @ormrod1990 and @iso2005. 
-However, the torch of exploring and modelling such diffusion processes had been
-passed to other disciplines such as economics, business studies and
-sociology well before the 'cultural turn' of economic geography
-[@perkins2005international]. A potential explanation of the recent lack of
+innovations diffuse across space and time with the work of 
+@hagerstrand1968innovation, which demonstrated the importance of communications 
+as a mechanism for the diffusion of innovations, being the most impactful. 
+It was followed by various empirical studies -- see for instance @ormrod1990 
+and @iso2005. However, the torch of exploring and modelling such diffusion 
+processes had been passed to other disciplines such as economics, business 
+studies and sociology well before the 'cultural turn' of economic geography
+[@perkins2005international]. A potential explanation of the lack of
 geographical studies exploring the diffusion of new and, more specifically for 
@@
 <!-- contribution -->
-This paper offers such a contribution: a geographical study demonstrating
-how a new technology that is the Web diffused over space and time in the
-UK at a high level of spatial granularity during the 1996-2012 period.
+This paper offers such a contribution: an economic geography study revealing
+the importance of spatial mechanisms for the diffusion of an intangible, 
+digital technology that is the Web in the UK at a high level of spatial 
+granularity during the 1996-2012 period.
+<!-- how a new technology that is the Web diffused over space and time  -->
 It does so by employing a novel source of big data and machine 
@@
 diffusion. This paper also exemplifies how the combination of data sources 
-which escape the traditional social science domain and adequate research 
-methods can offer new lenses to geographical research regarding the 
+and state-of-the-art research methods which escape the traditional social 
+science domain can offer new lenses to geographical research regarding the 
 understanding of technological diffusion.
@@
 stakeholders who are interested in knowing how new digital technologies
-diffused over space and time and use this knowledge to make predictions
-about the diffusion of related *future* technologies. As per
+diffuse over space and time and use this knowledge to make predictions
+about the diffusion of related future technologies. As per
 @leibowicz2016representing, historical studies agree that technologies diffuse 
@@
 of such diffusion mechanisms and the policy relevance, there is a scarcity 
-of empirical studies analysing the diffusion of new digital technologies 
-at a detailed geographical level.
+of studies analysing the diffusion of new digital technologies at a detailed
+geographical level because of the lack of detailed enough data and adequate 
+research methods.
 
@@
 paper, the Web -- and the active engagement with this technology that is 
-creating instead of just browsing a website -- can be considered as the 
-underpinning technology for the transformative process known as the
-*digital service economy* [@capello2024nexus].
-After all, diffusion together with invention and innovation are considered 
-the pillars of technological change [@das2022diffusion].
-Findings from previous empirical work illustrated that early engagement with 
-digital technologies such as the Web led to longstanding positive productivity 
-gains for these regions that hosted early adopters [@tranosuk].
+creating and maintaining instead of just browsing a website -- can be 
+considered as the underpinning technology for the transformative process 
+identified by @capello2024nexus as the *digital service economy*. Importantly, 
+the focus here is not the Web in general, but instead commercial websites.
+After all, diffusion together with invention and innovation are considered the 
+pillars of technological change [@das2022diffusion]. Findings from previous 
+studies illustrated that early engagement with digital technologies such as 
+the Web led to longstanding positive productivity gains for these regions that 
+hosted early adopters [@tranosuk]. Hence, obtaining a robust understanding of 
+the spatial diffusion mechanisms is crucial in mapping future development 
+trajectories.
 
@@
 and the debate about the footloose nature of digital technologies, which
-contrary to several earlier general purpose technologies are light in weight 
+contrary to several earlier general purpose technologies are 'light in weight' 
 [@coyle1999weightless]. Following the seminal work of @gaspar1998information, 
@@
 markets, this strand of literature has been mostly concerned with the
-spatiality of the internet's hard and tangible infrastructure
+spatiality of the Internet's hard and tangible infrastructure
 and shied away from exploring the spatial dimensions of one of the 
-most distinct software layers of the internet infrastructure, the Web.
+most distinct software layers of the Internet infrastructure, the Web.
 
@@
 technologies as it corresponds with the commercialisation of the
-internet and, consequently, its almost universal adoption. The reader is
+Internet and, consequently, its almost universal adoption. The reader is
 reminded that it was only in 1994 when Netscape Navigator was
 introduced, a year before Microsoft's Internet Explorer.[^1] Also, only
-9 per cent of UK's households had access to the internet in 1998
+9 per cent of UK's households had access to the Internet in 1998
 [@ons2018], the web included mostly static webpages, there were no
@@
 the study period captures the very early stages of the diffusion of a
-new technology that is the Web as well until its maturity. 
-The former is a key point in the lifecycle of a new technology. 
-Firstly, during this period new technologies are expensive, crude and 
-imperfect [@rosenberg1994exploring; @wilson201281].
-A simple comparison between Web 1.0 and Web 3.0 applications clearly
-illustrates this [@tranos2020social]: while a static website facilitates
-one-way information dissemination, a platform like `github` enables 
-cooperation between users and the creation of new information, meaning, 
-and even knowledge [@faraj2016special; @barassi2012does]. 
+new technology that is the Web until its maturity. The former is a key 
+point in the lifecycle of a new technology. Firstly, during this period new 
+technologies are expensive, crude and imperfect [@rosenberg1994exploring; 
+@wilson201281]. A simple comparison between Web 1.0 and Web 3.0 applications 
+clearly illustrates this [@tranos2020social]: while a static website 
+facilitates one-way information dissemination, a platform like `github` 
+enables cooperation between users and the creation of new information, 
+meaning, and even knowledge [@faraj2016special; @barassi2012does]. 
 During this period the performance of a new technology is the main 
@@
 contacts as a way for new technologies to diffuse. On the other hand, 
-economic models underline the importance of heterogeneity. Different firms have
-different structures and business plans, which define the potential
+economic models underline the importance of heterogeneity. Different firms 
+have different structures and business plans, which define the potential
 economic returns of the adoption of a new technology and, therefore, the
@@
 development trajectories [@neffke2011regions; @simmie2014new]. This
-argument is even more relevant when the focus is on digital technologies because 
-of the commonly found lag between investment and economic returns as reflected in 
-the Solow paradox [@acemoglu2014return; @brynjolfsson2018artificial].
+argument is even more relevant when the focus is on digital technologies 
+because of the commonly found lag between investment and economic returns as 
+reflected in the Solow paradox [@acemoglu2014return; 
+@brynjolfsson2018artificial].
 
@@
 geolocated and time-stamped counts of commercial websites from the UK. Instead 
-of adopting metrics reflecting the passive engagement with the internet such as 
-internet adoption and speeds [e.g. @blank2018local; @destefano2022fuzzy], this 
+of adopting metrics reflecting the passive engagement with the Internet such as 
+Internet adoption and speeds [e.g. @blank2018local; @destefano2022fuzzy], this 
 paper observes yearly counts of geolocated websites, which reflect the creation 
@@
 To analyse these data, a diverse set of methods are employed. Firstly, 
-cumulative adoption curves are estimated for the UK as a whole and for the Local 
-Authority Districts (LAD), illustrating, for the first time, such S-shaped 
-diffusion patterns for local areas. This system level analysis also reveals 
-insights about the volatility of the adoption of the Web. Then, exploratory 
-spatial data analysis offers some first indications of the importance of the 
-different spatial diffusion mechanisms in shaping the diffusion of the Web in 
-the UK. Importantly, the analysis adopts two distinct geographical scales: 
-the LAD and the much more detailed scale of the smallest census areas in the UK,
-the Output Areas. Lastly, a novel methodological framework based on 
-well-established Machine Learning (ML) algorithms is developed to test the role 
-of such mechanisms at these two scales. The results reveal for the first time 
-the importance and relevance of different spatial diffusion mechanisms for the 
-diffusion of a tangible, digital technology at very granular geographical scales.
+cumulative adoption curves are estimated for the UK as a whole and for the 
+Local Authority Districts (LAD), illustrating, for the first time, such 
+S-shaped diffusion patterns for local areas. This system level analysis also 
+reveals insights about the volatility of the adoption of the Web. Then, 
+exploratory spatial data analysis offers some first indications of the 
+importance of the different spatial diffusion mechanisms in shaping the 
+diffusion of the Web in the UK. Importantly, the analysis adopts two distinct 
+geographical scales: the LAD and the most detailed scale of UK census areas, 
+the Output Areas (OA). Lastly, a novel methodological framework based on 
+Machine Learning (ML) algorithms is developed to test the 
+role of such mechanisms at these two scales. The results reveal, for the first 
+time, the importance of different spatial diffusion mechanisms in the diffusion 
+of an intangible, digital technology at very granular geographical scales.
 
@@
 The results of the different stages of the analysis are presented in 
-@sec-results. The discussion of the results and the conlcusions of the paper
+@sec-results. The discussion of the results and the conclusions of the paper
 are offered in @sec-conclusions.
@@
 
-**SAY SOMETHING ABOUT DIST PAPER BY ARP, NOT SPECIFIC EXPLANATORY VARIABLES**
-
 # Literature review {#sec-litreview}
@@
 diffusion. This perspective draws similarities with epidemics:
-innovation just like pathogens spreads because of contagion and,
+innovations just like pathogens spread because of contagion and,
 consequently, proximity and exposure [@hivner2003facilitating]. On the
-other hand, there is a hierarchical processes. Instead of a
-horizontal distance-based diffusion mechanism, the top-down hierarchy
+other hand, there is a hierarchical processes. According to this, instead of 
+a horizontal distance-based diffusion mechanism, the top-down hierarchy
 of urban systems shapes technological diffusion. In reality, the synthesis 
 of these two processes better reflects how new technologies diffuse over 
-space and time [@morrill2020spatial].
+space [@morrill2020spatial].
 
@@
 thesis entitled 'Innovation Diffusion as a Spatial Process'
-[@hagerstrand1968innovation]. Hägerstrand was the first one to identify
+[@hagerstrand1968innovation]. Hägerstrand was the first to identify
 diffusion as a geographical process. The starting point was the idea
@@
 pattern just like purely temporal models; (ii) diffusion is shaped by a
-hierarchy mechanisms in a form of a centrifugal force: from core to
+hierarchical mechanism in a form of a centrifugal force: from core to
 periphery; and (iii) diffusion is also shaped by distance and proximity. 
@@
 heterogeneity is essential in explaining this distribution. In a recent
-study, @bednarz2020pulled focused on wind turbines and modeled their
+study, @bednarz2020pulled focused on wind turbines and modelled their
 spatial diffusion across 402 German regions during 1970-2015. Their
@@
 communication technologies that is mail, telegrams and telephones were
-shaped by similar socioeconomic factors as the internet. Their results
+shaped by similar socioeconomic factors as the Internet. Their results
 indicated common patterns regarding the drivers behind the adoption of
@@
 Turning to studies that share more technological and scalar similarities
-with this paper, @ding2010modeling modeled the spatial diffusion of
+with this paper, @ding2010modeling modelled the spatial diffusion of
 mobile telecommunications across regions in China. Their analysis
@@
 analysed its adoption and the churn at a very granular spatial level.
-Their results are in agreement with early theoretical and empirical
+Their results were in agreement with the early theoretical and empirical
 contributions reviewed here: assortativity, urban scaling and distance
@@
 shopping in Australia during the 1998-2009 period. At a global scale
-@PAPAGIANNIDIS2015308 modeled the diffusion of different web
-technologies and practices across countries. Interestingly, they used 
-similar, but less extensive data as the ones used here. Their analysis 
-illustrated how the diffusion of different web technologies and
-practices follow an S-shaped pattern as well as the different diffusion
-rates of the different technologies and practices.
+@PAPAGIANNIDIS2015308 modelled the diffusion of different web
+technologies and practices across countries. Using similar but less extensive 
+data as the one used here, their analysis illustrated how the diffusion 
+of different web technologies and practices follow an S-shaped pattern 
+as well as the different diffusion rates of the different technologies 
+and practices.
 
@@
 mechanisms shaped the diffusion of the Web in the UK at very granular
-spatial scales from the Web's early stages onwards.
+spatial scales from the Web's early stages onwards. The next section discusses
+the empirical framework.
 
@@
 To capture the diffusion of the Web, a website density metric is developed 
-for two different geographical scales: the Local Authority Districts (LAD) and the 
-Output Areas (OA). The former is an administrative unit and there are 
-374 LAD in the UK. The latter is the smallest census-based geographical unit and 
-there are c. 230,000 of them in the UK. This methodological choice 
-will allow the mapping of the diffusion of web technologies and the assessment of 
-the diffusion mechanisms at these two very different spatial scales.
+for two different geographical scales: the Local Authority Districts (LAD) 
+and the Output Areas (OA). The former is an administrative unit and there 
+are 374 LAD in the UK. The latter is the smallest census-based geographical 
+unit and there are c. 230,000 of them in the UK. This methodological choice 
+will allow the mapping of the diffusion of web technologies and the assessment 
+of the diffusion mechanisms at these two very different spatial scales.
 
@@
 This dataset, which is curated by the British Library, contains all
-the archived webpages from the UK ccTLD (.uk) from the 1996–2012 period. 
-In essence, this is a long list of 2.5 billion URLs of archived webpages including 
-also the archival timestamp.
+the archived webpages from the UK country code top level domain (ccTLD - .uk) 
+from the 1996–2012 period. In essence, this is a long list of 2.5 billion URLs 
+of archived webpages including also the archival timestamp.
 
@@
 UK consumers prefer to visit a .uk website when they are searching for 
-products or services [@hope]; and canecdotal evidence indicates that during 
-the first half of 2000, threec .co.uk domains were registered every minute 
-[@oecd_coms]. Importantly, previous studies illustrated that .co.uk is the most 
-popular UK SLD [@tranosuk].
+products or services [@hope]; and anecdotal evidence indicates that during 
+the first half of 2000, three .co.uk domains were registered every minute 
+[@oecd_coms]. Importantly, previous studies illustrated that .co.uk is the 
+most popular UK SLD [@tranosuk].
 
@@
 the archived webpages are aggregated to the parent websites. This website 
-reconstruction process lend itself to the consteruction of *website* instead 
-of *webpage* density metrics. Websites tend to represent specific orgnanisations 
+reconstruction process lend itself to the construction of *website* instead 
+of *webpage* density metrics. Websites tend to represent specific organisations 
 or entities, and, arguably, are more meaningful observational units than 
-webpages, which can ignore the upstream dependency of the website they belong to. 
-Based on the following example, all three webpages are part of the 
+webpages, which can ignore the upstream dependency of the website they belong 
+to. Based on the following example, all three webpages are part of the 
 same website ([http://www.website.co.uk](http://www.website.co.uk)). Only 
@@
 B15 2TT and BS8 1SS. The distribution of the number of postcodes per website for 
-2000 is presented in Table \ref{f2000}, which  clearly illustrates a wide range. 
+2000 is presented in Table \ref{f2000}, which illustrates a wide range. 
 At the left end of the distribution, there are websites anchored to a unique 
@@
 the analysis, such cases can represent directories which used to be popular 
-in the pre-search engines early times of the commercial internet as well as real 
+in the pre-search engines early times of the commercial Internet as well as real 
 estate websites [@tranosuk].
 
-The analysis presented here is based on two subsets of these data. Firstly, on 
-websites, which only contain one unique postcode. As a robustness check, the 
-analysis is replicated for an extended subset of websites, which include up to 
-10 unique postcodes to capture commercial websites which point to multiple 
-locations. They are geolocated by equally attaching them to these locations. 
-This extended sample includes 94 percent of all the archived websites in 2000.
+The analysis is firstly conducted on websites, which only contain one unique 
+postcode. As a robustness check, the analysis is replicated for an extended 
+subset of websites, which include up to 10 unique postcodes to capture 
+commercial websites which point to multiple locations. They are geolocated 
+by equally attaching them to these locations. This extended sample includes 94 
+percent of all the archived websites in 2000.
 
@@
 the data. As the @sec-results illustrates, this led to a small increase of the 
-model predictive capability.
+model predictive power.
 
@@
 are standardised by the number of firms in LAD to avoid biases associated with 
-LAD hosting a large number of firms. Given that there is no such statistic for 
-the OA, the actual OA level counts are used. However, because of the the 
-consistent spatial definition of OA (they host 40-250 households),[^4] website 
-counts in OA are interpreted as a density metric too.
+LAD hosting a large number of firms. Given that there is no such readily 
+available statistic for the OA, the actual OA level counts are used. However, 
+because of the the consistent spatial definition of OA (they host 40-250 
+households),[^4] website counts in OA are interpreted as a density metric too.
 
@@
 analysis. Firstly, a system-level analysis explores how the diffusion of web 
-technologies in the UK fits with the well-established S-curve. To do so, the 
-following logistic function (Equation \ref{eq:s}) is estimated for the whole 
-of the UK and for each LAD separately:
+technologies in the UK fits with the well-established S-curve of cumulative 
+adoption. To do so, the following logistic function (Equation \ref{eq:s}) is 
+estimated for the whole of the UK and for each LAD separately:
 
@@
 drivers of spatial diffusion -- namely neighbourhood and hierarchy -- underpin 
-thediffusion of web technologies in the UK. To capture the former and following 
+the diffusion of the Web in the UK. To capture the former and following 
 @ding2010modeling the Moran's I and the Local Index of Spatial Association 
 (LISA) are estimated for the website density. To address the hierarchy mechanism, 
-the Gini coefficient -- a well established metric of inequality -- is 
+the Gini coefficient -- an established metric of inequality -- is 
 calculated. All the above are computed and plotted longitudinally both for the 
@@
 Lastly, a modelling framework is developed to test the above diffusion 
-mechanisms. The overarching aim is to build a model that can test the 
-relationship between website density and these mechanisms:
+mechanisms. The overarching aim is to build a model that can test how well
+these mechanisms can predict website density:
 
@@
 RF have been widely employed to address regression research problems. 
-@pourebrahim2019trip combined a spatial interaction modeling framework with ML 
+@pourebrahim2019trip combined a spatial interaction modelling framework with ML 
 algorithms including RF to predict commuting flows in New York City. 
@@
 during the 1996-2012 period. Not surprisingly, the cumulative adoption is well 
-represented by curve that resembles the figure S. The vertical 
-line around year 2003 illustrates the point where the modeled cumulative 
+represented by a curve that resembles the figure S. The vertical 
+line around year 2003 illustrates the point where the modelled cumulative 
 adoption is equal to 50% of the maximum. This $t_0$ *inflection* point signals 
@@
 <!-- rankings -->
-The next step is assesses the stability and volatility of the LAD in terms of 
+The next step assesses the stability and volatility of the LAD in terms of 
 their adoption of web technologies. As highlighted in the literature 
@@
 in the UK [@tranosuk]. So, although the LAD system is volatile, it is not clear 
-whether the LAD with high concentration of late adopters will gain any latecomer 
-benefits in a way similar to countries experiencing such technological 
-leapfrogging. Still, the opportunity to catch up that @perkins2005international
-illustrated at the country level, is also visible at the much more detail LAD 
-scale. As previously, this is consistent with including websites with up 
-to 10 postcodes (see Figure \ref{rank10}).
+whether the LAD with high concentration of late adopters will gain any 
+latecomer-related benefits in a way similar to countries experiencing 
+technological leapfrogging. Still, the opportunity to catch up that 
+@perkins2005international illustrated at the country level, is also visible at 
+the much more detailed LAD scale. As previously, this is consistent with 
+including websites with up to 10 postcodes (see Figure \ref{rank10}).
 
@@
 neighbourhood mechanism underpins the diffusion of the Web in the UK as they 
-plot the Moran's I and the LISA maps of website density respectively for LAD 
+present the Moran's I and the LISA maps of website density respectively for LAD 
 and OA. Starting from the former, spatial autocorrelation was higher in the 
@@
 stabilised around 0.2. This reflects the early concentration of high website 
-density around London, which over time diffused as high-high clusters can be 
-seen in other parts of the country away from London (Figure \ref{lisa}). An 
-almost reverse pattern can be observed for OA. At the beginning of the study 
+density around London, which over time diffused as high website density clusters 
+can be seen in other parts of the country away from London (Figure \ref{lisa}). 
+An almost reverse pattern can be observed for OA. At the beginning of the study 
 period Moran's I was around 0.5 and it plateaued after 2000 around 0.2. 
@@
 <!-- discussion -->
-Interestingly, the magnitude of Moran's I is quite similar to the ones revealed
-by @ding2010modeling about mobile phone adoption in Chine albeit the much more 
+Interestingly, the magnitude of Moran's I is similar to the ones revealed
+by @ding2010modeling about mobile phone adoption in China albeit the much more 
 detailed spatial scale of this paper.
@@
 equally spread across the country. Both scales of analysis in Figure \ref{gini} 
-illustrate the same picture. At the beginning of the commercial internet website 
-density was extremely unequal, or, in other words, only a few places had 
+illustrate the same picture. At the beginning of the commercial Internet, 
+website density was extremely unequal, or, in other words, only a few places had 
 websites associated with them. Inequality dropped and plateaued after 2000 for 
@@
 
-## Modelling results
+## Modelling outputs
 
@@
 for the OA and LAD are utilised in order to build two RF models and assess their 
-capacity to predict the adoption of web technologies at the two different scales. 
+capacity to predict the adoption of the Web at the two different scales. 
 These two models will reveal the predictive capacity of the diffusion mechanisms 
@@
 the LAD from one of the twelve UK regions are held out for the model training.
-Then the trained model is used to predict website density for the OA or the LAD 
+Then, the trained model is used to predict website density for the OA or the LAD 
 of the held-out region. This process takes place recursively for all twelve UK
@@
 
-It needs to be highlighted here that the cross-validation for all models
+It needs to be highlighted here that the cross validation (CV) for all models
 is spatially and temporally sensitive. Instead of using
@@
 literature and the descriptive analysis highlighted. Namely, the models capture: 
-(i) a hierarchy mechanism with diffusion running from main centres to secondary 
-ones, (ii) a neighborhood mechanism according to which diffusion first hits 
-nearby locations, and (iii) the rather canonical pattern of diffusion over time 
-as reflected in the S-shaped pattern in the cumulative level of adoption.
-
-To capture the hierarchy effect the models include as predictors a one
+(i) a hierarchical mechanism with diffusion running from main centres to 
+secondary ones, (ii) a neighborhood mechanism according to which diffusion 
+first hits nearby locations, and (iii) the rather canonical pattern of 
+diffusion over time as reflected in the S-shaped pattern in the cumulative level 
+of adoption.
+
+To capture the hierarchical mechanism the models include as predictors a one
 year lag of website density in London, the largest city in the UK, a
@@
 the nearest retail centre. Due to the small sizes of the retail centres,
-the latter is only relevant for the OA-level models. In addition, the
-models include the distance to London, the nearest city and the nearest
-retail centre. The underlying logic is that the level of website
+the latter is only relevant for and included in the OA-level models. In 
+addition, the models include the distance to London, the nearest city and the 
+nearest retail centre. The underlying logic is that the level of website
 adoption in a spatial unit depends on the level of the adoption in
@@
 the distance variables incorporates spatial structure into the hierarchy 
-argument. To depict the neighbourhood effect, the website density of the 
-neighbouring spatial units in the previous year is employed (rook continuity). 
-Again, the underpinning rationale is that the level of web adoption within a 
-spatial unit depends on the level of web adoption in the neighbouring spatial 
-units the year before. This represents the 'hitting nearby locations first' 
-argument. Therefore, the spatial and temporal lag of the website density in 
-LAD and OA is calculated. Lastly, the temporal pattern of the cumulative 
-adoption is captured by a time trend variable. Hence, all four models will 
-follow the following generic form (Eq. \ref{model}):
+argument. To depict the neighbourhood mechanism, the website density of the 
+neighbouring spatial units in the previous year is employed. This is calculated 
+using rook continuity. Again, the underpinning rationale is that the level of 
+web adoption within a spatial unit depends on the level of web adoption in the 
+neighbouring spatial units the year before. This represents the 'hitting nearby 
+locations first' argument. Therefore, the spatial and temporal lag of the 
+website density in LAD and OA is calculated. Lastly, the temporal pattern of 
+the cumulative adoption is captured by a time trend variable. Hence, all four 
+models will follow the following generic form (Eq. \ref{model}):
 
@@
 
-To assess the predictive capability of the model, three broadly utilised
+To assess the predictive power of the model, three broadly utilised
 metrics are employed: the coefficient of determination ($R^2$), mean
@@
 the observed and the predicted website adoption. While $MAE$ does not
-penalise for large errors, $RSME$ does so as it is proportional to the
+penalise for large errors, $RSME$ does, as it is proportional to the
 squared difference between the observed and the predicted trade flows.
@@
 @tbl-model-metrics presents the model performance for the first set of models, 
-for which all data points are employed for training and testing via cross 
-validation. The first one is trained and tested on 374 LAD and the second on 
-232,296 OA, both for a 16 year period (1997-2012). The results are remarkably 
+for which all data points are employed for training and testing via CV. The 
+first one is trained and tested on 374 LAD and the second on 
+232,296 OA, both for the 16 year period (1997-2012). The results are remarkably 
 good considering that the are the outcome of space and time sensitive 
@@
 the UK OA were associated with at least one website. This should not come as a 
-surprise as this was the very beginning of the commercial internet and any 
+surprise as this was the very beginning of the commercial Internet and any 
 activities with a digital footprint were concentrated in a handful of areas. 
@@
 to the base results in @tbl-model-metrics. All in all, the modelling exercise
-revealed the how well can the three established diffusion mechanisms predict 
+revealed how well can the three diffusion mechanisms predict 
 the diffusion of the Web as a new technology. Importantly, this predictive
-capacity is highly robust against against different scales and delineations
+power is highly robust against against different scales and delineations
 of the diffusion of the web.
@@
 previously discussed spatial processes are at play in the diffusion of web 
-technologies at the LAD level: the first two predictors depict the hierarchy 
+technologies at the LAD level: the first two predictors depict the hierarchical 
 mechanism, the spatial and temporal lag of website density the neighbourhood 
@@
 
-The above pattern of variable importance is largely in accordance with the models
-estimated for the extended dataset. A comparison with Figure \ref{var.imp10} 
-illustrates that the main difference is the importance of the neighbourhood
-mechanism for the OA as the space and time lag of the website density is by far 
-the most important predictor when the extended dataset is employed. But still, 
-the variable importance patterns for the two datasets are not dissimilar.
+The above pattern of variable importance is largely in accordance with the 
+models estimated for the extended dataset. A comparison with Figure 
+\ref{var.imp10} illustrates that the main difference is the importance of the 
+neighbourhood mechanism for the OA as the space and time lag of the website 
+density is by far the most important predictor when the extended dataset is 
+employed. But still, the variable importance patterns for the two datasets are 
+not dissimilar.
 
@@
 
-Table \ref{table.regions} presents the results of the recursive hold out models,
-which aim to highlight the potential regional heterogeneity of the spatial processes
-behind the diffusion of web technologies. To begin with, as highlighted before, 
-there is a difference of magnitude 
-of one order between the LAD and the OA prediction errors, which is aligned with 
-previous results that employed all data points. What is of interest here is the 
-regional comparison. Table \ref{table.regions} illustrates some striking similarities, 
-but also a few significant differences. The regions the web diffusion of which is 
-better predicted using models trained in the rest of the country are the same 
-despite the scale of analysis: South East, Wales, Yorkshire and The Humber and 
-the North East of England. In other words, these are the regions whose spatial 
-diffusion mechanisms of web technologies is closer to the country's average. Despite the 
-consistency across scales, this is a diverse set of regions: **ADD CHARACTERISTICS**.
-
-At the other end of the spectrum, Scotland's and the North West's web diffusion
-mechanisms are consistently diverging from the country's average. This should not
-come as a surprise as these regions are characterised of high levels of rurality
-and remoteness. Similarly, London diffusion mechanisms diverge from the country's 
-average and this is consistent across scales. London's uniqueness in UK's urban 
-system and economy is also reflected in the spatial diffusion mechanisms of web 
-technologies within its LAD and OA. It needs to be highlighted though that the 
-difference between the $R^2$ of LAD and OA is more that an order of magnitude 
-signaling how difficult is to predict diffusion at such a small spatial scale.
-Northern Ireland is an interesting case. While it ranks at the bottom of the scale
-when the models are trained and tested on LAD data, when the modelling adopts the
-more granular OA scale, the spatial mechanisms that shape the web diffusion within
-this region appear to be closer to the country's average. At this scale, proximity,
-or lack of, relative to the rest of the country become less important and the 
-internal to the region spatial structure predictors start playing a more
-important role **CHECK NI OA model**.
-
 ```{r regions}
@@
 
-# Discussion and conclusions {#sec-conclusions}
-
-contrary to results from future studies regarding social media
-[@lengyel2020role], web technologies did not exclusively spread from a
-central location.
+Table \ref{table.regions} presents the results of the recursive hold out models
+and allows to (i) highlight the regional heterogeneity of the spatial 
+processes behind the diffusion of the Web, and (ii) compare this heterogeneity
+across the two different scales of analysis. It illustrates some striking 
+similarities, but also a few significant differences. At the top of the 
+similarity with the rest of the country we see West Midlands and Yorkshire
+and the Humber for both scales. The web density for the LAD and OA of these 
+regions can be very well predicted by models trained on data for the LAD 
+and OA of the rest of the country. In other words, the spatial diffusion 
+mechanisms for these regions represent quite well the UK's average. This
+should not come as as surprise given their urban structure: they both contain
+a large urban centre (Birmingham and Leeds), smaller commuter towns as well 
+as more rural settlements. The bottom of the similarity ranking is populated 
+by Scotland, Northern Ireland, East of England and London for both scales. 
+The diffusion mechanisms for these regions diverge significantly from the 
+country's average and this could be attributed to rurality and remoteness 
+or to high levels of urbanisation and sophistication of economic activities 
+for the case of London and for East of England, which contains the Cambridge 
+area.
+
+Still, there are three regions the spatial diffusion mechanisms of which 
+differ a lot between the two scales. North West and South West are very
+close to the UK's average at the more granular OA scale, while the opposite
+applies for South East. Interestingly, the first two regions are well-known 
+tourist destinations (Lake District and Cornwall-Devon) and the exploratory
+analysis revealed clusters of high web density only visible at the OA scale 
+(Figure \ref{lisa}. Then, the South East is very well integrated with the 
+Greater London, hosts the broader Oxford area -- a known hub of research 
+excellence -- and Figure \ref{lisa} revealed high web density clusters at 
+the LAD level.
+
+# Conclusions {#sec-conclusions}
+
+This paper assessed the importance of spatial mechanisms in the diffusion 
+of a new technology -- the Web -- at a high level of spatial granularity.
+Specifically, it assessed the capacity of the following three spatial 
+diffusion mechanisms in predicting website density in the UK over space and 
+time: (i) the hierarchical mechanism according to which diffusion runs 
+from main urban centres to secondary ones, (ii) the neighborhood mechanism 
+which highlights that diffusion first hits nearby locations, and (iii) the 
+rather canonical pattern of diffusion over time as reflected in the S-shaped 
+pattern of the cumulative level of adoption. This is the first time that 
+the importance of granular spatial mechanisms for the diffusion of an
+intangible, digital technology is exposed. The time period of the study is 
+also important as it captures the very early stages of this technology until 
+its maturity. Importantly, the empirical framework and the data employed here 
+allows to observe the active engagement with the Web that is creating and 
+maintaining a commercial website instead of the more passive action of 
+browsing webpages and accessing the Internet.
+
+The analysis firstly illustrated how well the S-pattern of cumulative adoption 
+fits at the local scale of LAD. It highlighted the heterogeneity of adoption 
+rates between LAD -- with both more and less expected examples -- as well as 
+the spatial clustering of LAD with similar adoption rates. The exploratory 
+analysis also illustrated instances of remarkable stability as well as 
+volatility of the web diffusion system that  is indicative of leapfrogging. 
+Previous research had already indicated the long-term productivity gains of 
+the early adoption of such technologies [@tranosuk]. Then, the paper developed 
+a robust and novel ML modelling framework to predict website diffusion over 
+space and time using the three spatial diffusion mechanisms based on space- 
+and time-sensitive CV without suffering from overfitting. The model revealed 
+the remarkably good predictive capacity of these mechanisms at both scales. 
+For the first time, the power of these mechanisms in predicting the diffusion 
+of a new technology was empirically tested, importantly, at very fine scales.
+
+The modelling framework also revealed that the importance of these spatial 
+mechanisms differ across different scales. While spatial structure as 
+reflected in distances to different urban centres is highly important at the 
+OA scale, it hardly plays any role at the more coarse LAD scale. And while 
+the almost canonical pattern of cumulative adoption is a key predictor at the 
+LAD level, it only plays a small role at the detailed OA level. Then, the 
+analysis revealed another level of spatial heterogeneity. While for some UK 
+regions these spatial diffusion mechanisms are very similar to the rest of 
+the country, other regions diverge significantly from the country average.
+Importantly, the above results are robust against different specifications 
+and subsets of the data.
+
+The paper offers a number of different contributions. Firstly, it assesses, 
+for the first time, the role of these spatial diffusion mechanisms at 
+such granular spatial scales. While previous studies had 
+highlighted these mechanisms [@hagerstrand1968innovation; @rogers2010diffusion; 
+@grubler1990rise], their role had never been tested before at such granular 
+scale and, importantly, for a digital technology as weightless as the Web. 
+The findings talk directly to this literature  [e.g. @beardsell1999spatial; 
+@haller2011determinants; @lengyel2020role]. More broadly, the paper offers 
+a robust methodological framework, which utilises novel data and modern ML 
+algorithms and techniques to model the diffusion of a new technology across 
+space and time. It directly addresses the data problem which prevented economic
+geographers in engaging  with diffusion studies [@iso2005; 
+@kemeny2011international; @zook2022mapping].
+Thus, the paper seeks to renew economic geography's interest in the diffusion 
+of new technologies. Current data science advances including the availability 
+of granular enough data and adequate methods can enable economic 
+geographers to reveal interesting spatial patterns of diffusion of new 
+technologies as this paper has demonstrated. This is important as (i) 
+various stakeholders are interested in predicting localised demand for new 
+technologies for planning purposes [@leibowicz2016representing; 
+@meade2021modelling], and (ii) the adoption of such digital technologies 
+is associated with economic gains [@solow1957technical; @aghion1990model; 
+@kemeny2011international; @tranosuk; @capello2024nexus]. In addition, while 
+most related studies focused on tangible technologies, this paper sheds 
+light on the diffusion of a truly weightless digital technology that is the 
+Web. Thus, it adds much more granular findings to the limited related studies
+[@sinai2004geography; @haller2011determinants]. Importantly, although the 
+geographical focus of this paper is the UK, the empirical framework developed 
+here is easily transferable to other geographical contexts.
+
+<!-- Future research can further investigate other -->
+<!-- determinants of the diffusion of the Web. For instance, the archived web data -->
+<!-- can reveal important information about the industrial base of different places -->
+<!-- as well as the industries which were faster in adopting the Web. -->
 

### 2025-01-03 paper/submission1/anonymised.tex
ssh problem

--- 2024-09-12 paper/submission1/tranos2023.tex
@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
@@
 \journal{Journal Name}
+
 \ifLuaTeX
@@
   pdftitle={A multi-scale story of the diffusion of a new technology: the web},
-  pdfauthor={Emmanouil Tranos},
-  pdfkeywords={keyword1, keyword2},
+  pdfkeywords={diffusion, technologies, Web, digital},
   colorlinks=true,
@@
 web}
-\author[1]{Emmanouil Tranos%
-\corref{cor1}%
-}
- \ead{e.tranos@bristol.ac.uk}
-
-\affiliation[1]{organization={University of
-Bristol},postcode={UK},postcodesep={}}
 
@@
 \begin{abstract}
-This paper maps the participation in the digital economy and its
-evolution in the UK over space and time. Most of the existing economic
-geography literature which dealt with the spatiality of the internet
-employed supply-side measures, such as infrastructural capacity, in
-order to understand the geography of the digital economy and its
-potential spatial economic effects. Useful as these approaches might
-have been, they cannot capture the micro-processes and the
-characteristics of the individual online behaviour. Using large volumes
-of archived and geolocated web content, this paper models the diffusion
-of web technologies over space and time in the UK. The data and
-geolocation strategy allow to capture these processes at a very granular
-spatial scale. The modelling approach, which is based on simple spatial
-analytical methods and on the estimation of diffusion curves at various
-scales, enables to depict the role of geography and other cognitive
-factors which drove the diffusion of web technologies. Although the
-focus is on a recent historical period -- 1996-2012 -- the results of
-the analysis depict diffusion mechanisms which can be very useful in
-understanding the evolutionary patterns of the adoption of other newer
-technologies.
+Despite the broad consensus that differences in technological adoption
+and sophistication can lead to uneven trajectories of local development
+as new technologies can increase productivity and allow firms to
+introduce new products and services, economic geographers have shied
+away from exploring the diffusion of digital technologies across space
+and time. This paper fills this gap by exposing the roles of different
+spatial mechanisms in the diffusion of an intangible, digital technology
+at very granular geographical scales. It does so by employing a novel
+source of data and machine learning algorithms to model the active
+engagement with Web during 1996-2012.
 \end{abstract}
@@
 \begin{keyword}
-    keyword1 \sep 
-    keyword2
+    diffusion \sep technologies \sep Web \sep 
+    digital
 \end{keyword}
@@
 before the `cultural turn' of economic geography
-\citep{perkins2005international}. A potential explanation of the recent
-lack of geographical studies exploring the diffusion of new and, more
+\citep{perkins2005international}. A potential explanation of the lack of
+geographical studies exploring the diffusion of new and, more
 specifically for this paper, digital technologies across both space and
@@
 
-This paper offers such a contribution: a geographical study illustrating
-how a new technology that is the Web diffused over space and time in the
-UK at a high level of spatial granularity during the 1996-2012 period.
-It does so by employing a novel source of big data and machine learning
-algorithms to model the \emph{active} engagement with Web during that
-period and tests how different spatial and temporal mechanisms shaped
-its diffusion. This paper exemplifies how the combination of data
-sources which escape the traditional social science domain and adequate
-research methods can offer new lenses to geographical research regarding
+This paper offers such a contribution: an economic geography study
+revealing the importance of spatial mechanisms for the diffusion of an
+intangible, digital technology that is the Web in the UK at a high level
+of spatial granularity during the 1996-2012 period. It does so by
+employing a novel source of big data and machine learning algorithms to
+model the \emph{active} engagement with Web during that period and tests
+how different spatial and temporal mechanisms shaped its diffusion. This
+paper also exemplifies how the combination of data sources and
+state-of-the-art research methods which escape the traditional social
+science domain can offer new lenses to geographical research regarding
 the understanding of technological diffusion.
@@
 stakeholders who are interested in knowing how new digital technologies
-diffused over space and time and use this knowledge to make predictions
-about the diffusion of related \emph{future} technologies. As per
+diffuse over space and time and use this knowledge to make predictions
+about the diffusion of related future technologies. As per
 \citet{leibowicz2016representing}, historical studies agree that
@@
 Despite the spatial heterogeneity of such diffusion mechanisms and the
-policy relevance, there is a scarcity of empirical studies analysing the
-diffusion of new digital technologies at a detailed geographical level.
+policy relevance, there is a scarcity of studies analysing the diffusion
+of new digital technologies at a detailed geographical level because of
+the lack of detailed enough data and adequate research methods.
 
@@
 Linked to the subject of this paper, the Web -- and the active
-engagement with this technology that is creating instead of just
-browsing a website -- can be considered as the underpinning technology
-for the transformative process known as the \emph{digital service
-economy} \citep{capello2024nexus}. After all, diffusion together with
-invention and innovation are considered the pillars of technological
-change \citep{das2022diffusion}. Findings from previous empirical work
-illustrated that early engagement with digital technologies such as the
-Web led to longstanding positive productivity gains for these regions
-that hosted early adopters \citep{tranosuk}.
+engagement with this technology that is creating and maintaining instead
+of just browsing a website -- can be considered as the underpinning
+technology for the transformative process identified by
+\citet{capello2024nexus} as the \emph{digital service economy}.
+Importantly, the focus here is not the Web in general, but instead
+commercial websites. After all, diffusion together with invention and
+innovation are considered the pillars of technological change
+\citep{das2022diffusion}. Findings from previous studies illustrated
+that early engagement with digital technologies such as the Web led to
+longstanding positive productivity gains for these regions that hosted
+early adopters \citep{tranosuk}. Hence, obtaining a robust understanding
+of the spatial diffusion mechanisms is crucial in mapping future
+development trajectories.
 
@@
 Web and the debate about the footloose nature of digital technologies,
-which contrary to several earlier general purpose technologies are light
-in weight \citep{coyle1999weightless}. Following the seminal work of
-\citet{gaspar1998information}, which demonstrated that technological
-improvements in telecommunications lead to increased demand for
-face-to-face communications and, consequently, enhance the importance of
-cities, empirical studies from economic geography and urban economics
-explored whether digital technologies substitute or complement
-agglomeration externalities and spatial structure
+which contrary to several earlier general purpose technologies are
+`light in weight' \citep{coyle1999weightless}. Following the seminal
+work of \citet{gaspar1998information}, which demonstrated that
+technological improvements in telecommunications lead to increased
+demand for face-to-face communications and, consequently, enhance the
+importance of cities, empirical studies from economic geography and
+urban economics explored whether digital technologies substitute or
+complement agglomeration externalities and spatial structure
 \citep{kolko_death_2000, ioannides2008effect, tranos2021ubiquitous}.
@@
 available in larger markets, this strand of literature has been mostly
-concerned with the spatiality of the internet's hard and tangible
+concerned with the spatiality of the Internet's hard and tangible
 infrastructure and shied away from exploring the spatial dimensions of
-one of the most distinct software layers of the internet infrastructure,
+one of the most distinct software layers of the Internet infrastructure,
 the Web.
@@
 technologies as it corresponds with the commercialisation of the
-internet and, consequently, its almost universal adoption. The reader is
+Internet and, consequently, its almost universal adoption. The reader is
 reminded that it was only in 1994 when Netscape Navigator was
 introduced, a year before Microsoft's Internet Explorer.\footnote{\url{https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft}}
-Also, only 9 per cent of UK's households had access to the internet in
+Also, only 9 per cent of UK's households had access to the Internet in
 1998 \citep{ons2018}, the web included mostly static webpages, there
@@
 that the study period captures the very early stages of the diffusion of
-a new technology that is the Web as well until its maturity. The former
-is a key point in the lifecycle of a new technology. Firstly, during
-this period new technologies are expensive, crude and imperfect
+a new technology that is the Web until its maturity. The former is a key
+point in the lifecycle of a new technology. Firstly, during this period
+new technologies are expensive, crude and imperfect
 \citep{rosenberg1994exploring, wilson201281}. A simple comparison
@@
 the UK. Instead of adopting metrics reflecting the passive engagement
-with the internet such as internet adoption and speeds
+with the Internet such as Internet adoption and speeds
 \citep[e.g.][]{blank2018local, destefano2022fuzzy}, this paper observes
@@
 the analysis adopts two distinct geographical scales: the LAD and the
-much more detailed scale of the smallest census areas in the UK, the
-Output Areas. Lastly, a novel methodological framework based on
-well-established Machine Learning (ML) algorithms is developed to test
-the role of such mechanisms at these two scales. The results reveal for
-the first time the importance and relevance of different spatial
-diffusion mechanisms for the diffusion of a tangible, digital technology
-at very granular geographical scales.
+most detailed scale of UK census areas, the Output Areas (OA). Lastly, a
+novel methodological framework based on Machine Learning (ML) algorithms
+is developed to test the role of such mechanisms at these two scales.
+The results reveal, for the first time, the importance of different
+spatial diffusion mechanisms in the diffusion of an intangible, digital
+technology at very granular geographical scales.
 
@@
 of the analysis are presented in Section~\ref{sec-results}. The
-discussion of the results and the conlcusions of the paper are offered
+discussion of the results and the conclusions of the paper are offered
 in Section~\ref{sec-conclusions}.
-
-\textbf{SAY SOMETHING ABOUT DIST PAPER BY ARP, NOT SPECIFIC EXPLANATORY
-VARIABLES}
 
@@
 diffusion. This perspective draws similarities with epidemics:
-innovation just like pathogens spreads because of contagion and,
+innovations just like pathogens spread because of contagion and,
 consequently, proximity and exposure \citep{hivner2003facilitating}. On
-the other hand, there is a hierarchical processes. Instead of a
-horizontal distance-based diffusion mechanism, the top-down hierarchy of
-urban systems shapes technological diffusion. In reality, the synthesis
-of these two processes better reflects how new technologies diffuse over
-space and time \citep{morrill2020spatial}.
+the other hand, there is a hierarchical processes. According to this,
+instead of a horizontal distance-based diffusion mechanism, the top-down
+hierarchy of urban systems shapes technological diffusion. In reality,
+the synthesis of these two processes better reflects how new
+technologies diffuse over space \citep{morrill2020spatial}.
 
@@
 thesis entitled `Innovation Diffusion as a Spatial Process'
-\citep{hagerstrand1968innovation}. Hägerstrand was the first one to
-identify diffusion as a geographical process. The starting point was the
-idea that diffusion is based on passing information through social
-networks, which themselves tend to be defined by geography. Hence, he
-identified a `neighbourhood' mechanism of how information, and
-consequently, innovation diffuse. He used agricultural innovations to
-test and model his ideas using Monte Carlo simulations. Hägerstrand also
-incorporated the role of hierarchy and how some phenomena maybe firstly
-adopted in larger cities and then diffuse to second tier ones. This is a
-sequential instead of a simultaneous process, which resembles the
-`lead-lag' spatial acceleration effect in market research
+\citep{hagerstrand1968innovation}. Hägerstrand was the first to identify
+diffusion as a geographical process. The starting point was the idea
+that diffusion is based on passing information through social networks,
+which themselves tend to be defined by geography. Hence, he identified a
+`neighbourhood' mechanism of how information, and consequently,
+innovation diffuse. He used agricultural innovations to test and model
+his ideas using Monte Carlo simulations. Hägerstrand also incorporated
+the role of hierarchy and how some phenomena maybe firstly adopted in
+larger cities and then diffuse to second tier ones. This is a sequential
+instead of a simultaneous process, which resembles the `lead-lag'
+spatial acceleration effect in market research
 \citep{bento2018time, PERES201091}. Hägerstrand is more widely known
@@
 S-shaped pattern just like purely temporal models; (ii) diffusion is
-shaped by a hierarchy mechanisms in a form of a centrifugal force: from
-core to periphery; and (iii) diffusion is also shaped by distance and
-proximity. These are the three mechanisms that the empirical analysis in
-Section~\ref{sec-results} investigates.
+shaped by a hierarchical mechanism in a form of a centrifugal force:
+from core to periphery; and (iii) diffusion is also shaped by distance
+and proximity. These are the three mechanisms that the empirical
+analysis in Section~\ref{sec-results} investigates.
 
@@
 heterogeneity is essential in explaining this distribution. In a recent
-study, \citet{bednarz2020pulled} focused on wind turbines and modeled
+study, \citet{bednarz2020pulled} focused on wind turbines and modelled
 their spatial diffusion across 402 German regions during 1970-2015.
@@
 Turning to studies that share more technological and scalar similarities
-with this paper, \citet{ding2010modeling} modeled the spatial diffusion
+with this paper, \citet{ding2010modeling} modelled the spatial diffusion
 of mobile telecommunications across regions in China. Their analysis
@@
 \citet{lengyel2020role} analysed its adoption and the churn at a very
-granular spatial level. Their results are in agreement with early
+granular spatial level. Their results were in agreement with the early
 theoretical and empirical contributions reviewed here: assortativity,
-urban scaling and distance are the key drivers of spatial diffusion. At
-a global scale \citet{PAPAGIANNIDIS2015308} modeled the diffusion of
-different web technologies and practices across countries.
-Interestingly, they used similar, but less extensive data as the ones
-used here. Their analysis illustrated how the diffusion of different web
+urban scaling and distance are the key drivers of spatial diffusion.
+\citet{bakher2013diffusion} compared different the growth curves as a
+means to model diffusion of online shopping in Australia during the
+1998-2009 period. At a global scale \citet{PAPAGIANNIDIS2015308}
+modelled the diffusion of different web technologies and practices
+across countries. Using similar but less extensive data as the one used
+here, their analysis illustrated how the diffusion of different web
 technologies and practices follow an S-shaped pattern as well as the
@@
 diffusion mechanisms shaped the diffusion of the Web in the UK at very
-granular spatial scales from the Web's early stages onwards.
+granular spatial scales from the Web's early stages onwards. The next
+section discusses the empirical framework.
 
@@
 archives. This dataset, which is curated by the British Library,
-contains all the archived webpages from the UK ccTLD (.uk) from the
-1996--2012 period. In essence, this is a long list of 2.5 billion URLs
-of archived webpages including also the archival timestamp.
+contains all the archived webpages from the UK country code top level
+domain (ccTLD - .uk) from the 1996--2012 period. In essence, this is a
+long list of 2.5 billion URLs of archived webpages including also the
+archival timestamp.
 
@@
 visit a .uk website when they are searching for products or services
-\citep{hope}; and canecdotal evidence indicates that during the first
-half of 2000, threec .co.uk domains were registered every minute
+\citep{hope}; and anecdotal evidence indicates that during the first
+half of 2000, three .co.uk domains were registered every minute
 \citep{oecd_coms}. Importantly, previous studies illustrated that .co.uk
@@
 the archived webpages are aggregated to the parent websites. This
-website reconstruction process lend itself to the consteruction of
+website reconstruction process lend itself to the construction of
 \emph{website} instead of \emph{webpage} density metrics. Websites tend
-to represent specific orgnanisations or entities, and, arguably, are
-more meaningful observational units than webpages, which can ignore the
+to represent specific organisations or entities, and, arguably, are more
+meaningful observational units than webpages, which can ignore the
 upstream dependency of the website they belong to. Based on the
@@
 postcodes per website for 2000 is presented in Table \ref{f2000}, which
-clearly illustrates a wide range. At the left end of the distribution,
-there are websites anchored to a unique location (72\% of all the
-reconstructed websites in 2000), which may represent a small company
-with a single trading location. At the right end, there are websites
-with thousands of different postcodes. Considering the time period of
-the analysis, such cases can represent directories which used to be
-popular in the pre-search engines early times of the commercial internet
-as well as real estate websites \citep{tranosuk}.
-
-The analysis presented here is based on two subsets of these data.
-Firstly, on websites, which only contain one unique postcode. As a
-robustness check, the analysis is replicated for an extended subset of
-websites, which include up to 10 unique postcodes to capture commercial
-websites which point to multiple locations. They are geolocated by
-equally attaching them to these locations. This extended sample includes
-94 percent of all the archived websites in 2000.
+illustrates a wide range. At the left end of the distribution, there are
+websites anchored to a unique location (72\% of all the reconstructed
+websites in 2000), which may represent a small company with a single
+trading location. At the right end, there are websites with thousands of
+different postcodes. Considering the time period of the analysis, such
+cases can represent directories which used to be popular in the
+pre-search engines early times of the commercial Internet as well as
+real estate websites \citep{tranosuk}.
+
+The analysis is firstly conducted on websites, which only contain one
+unique postcode. As a robustness check, the analysis is replicated for
+an extended subset of websites, which include up to 10 unique postcodes
+to capture commercial websites which point to multiple locations. They
+are geolocated by equally attaching them to these locations. This
+extended sample includes 94 percent of all the archived websites in
+2000.
 
@@
 
-\caption{\label{tab:unnamed-chunk-1}Number of unique postcodes per .co.uk website, 2000.\label{f2000}}
+\caption{Number of unique postcodes per .co.uk website, 2000.\label{f2000}}
 \centering
@@
 557,808 postcodes included in the data. As the Section~\ref{sec-results}
-illustrates, this led to a small increase of the model predictive
-capability.
+illustrates, this led to a small increase of the model predictive power.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{tranos2023_files/figure-pdf/unnamed-chunk-2-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{anonymised_files/figure-pdf/unnamed-chunk-2-1.pdf}
 
@@
 associated with LAD hosting a large number of firms. Given that there is
-no such statistic for the OA, the actual OA level counts are used.
-However, because of the the consistent spatial definition of OA (they
-host 40-250 households),\footnote{\url{https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies}}
+no such readily available statistic for the OA, the actual OA level
+counts are used. However, because of the the consistent spatial
+definition of OA (they host 40-250 households),\footnote{\url{https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies}}
 website counts in OA are interpreted as a density metric too.
@@
 diffusion of web technologies in the UK fits with the well-established
-S-curve. To do so, the following logistic function (Equation \ref{eq:s})
-is estimated for the whole of the UK and for each LAD separately:
+S-curve of cumulative adoption. To do so, the following logistic
+function (Equation \ref{eq:s}) is estimated for the whole of the UK and
+for each LAD separately:
 
@@
 drivers of spatial diffusion -- namely neighbourhood and hierarchy --
-underpin thediffusion of web technologies in the UK. To capture the
-former and following \citet{ding2010modeling} the Moran's I and the
-Local Index of Spatial Association (LISA) are estimated for the website
-density. To address the hierarchy mechanism, the Gini coefficient -- a
-well established metric of inequality -- is calculated. All the above
-are computed and plotted longitudinally both for the LAD and the OA.
+underpin the diffusion of the Web in the UK. To capture the former and
+following \citet{ding2010modeling} the Moran's I and the Local Index of
+Spatial Association (LISA) are estimated for the website density. To
+address the hierarchy mechanism, the Gini coefficient -- an established
+metric of inequality -- is calculated. All the above are computed and
+plotted longitudinally both for the LAD and the OA.
 
 Lastly, a modelling framework is developed to test the above diffusion
-mechanisms. The overarching aim is to build a model that can test the
-relationship between website density and these mechanisms:
+mechanisms. The overarching aim is to build a model that can test how
+well these mechanisms can predict website density:
 
@@
 RF have been widely employed to address regression research problems.
-\citet{pourebrahim2019trip} combined a spatial interaction modeling
+\citet{pourebrahim2019trip} combined a spatial interaction modelling
 framework with ML algorithms including RF to predict commuting flows in
@@
 
-Figure \ref{s_uk} plots the S curve for the cumulative adoption of
-website technologies in the UK during the 1996-2012 period. It
-demonstrates a pattern well aligned with previous studies discussed in
-Section~\ref{sec-litreview}. The vertical line around year 2003
-illustrates the point where the modeled cumulative adoption is equal to
-50\% of the maximum. This \(t_0\) \emph{inflection} point signals the
-maximum adoption speed and is used here to determine whether a UK LAD
-reached that inflection point earlier or later than the UK average.
-Specifically, the S curve and the inflection point are estimated for
-every LAD individually and then compared against the UK average. LAD
-that reached their inflection point earlier than the country average are
-labelled as \emph{fast} and the rest as \emph{slow}. Figure \ref{s_map}
-maps this pattern. There is a clear high concentration of fast adopting
-LAD in the South and East of England as expected. However, there is also
-such a concentration in Wales, Scotland and the North West of England.
-At a more detailed level, there are some expected examples of LAD with
-relevant industrial backgrounds delineated as fast: the City of London,
-a world-renowned cluster of finance industries \citep{cook2007role}, and
-Reading, a town with high-tech service industries in proximity to London
-and its main airport, Heathrow \citep{england2005polynet}. However,
-there are also LAD which were expected to appear as fast -- e.g.~Hackney
-in central London and Bristol, a well-established creative cluster
+Figure \ref{s_uk} plots the cumulative adoption of the Web in total in
+the UK during the 1996-2012 period. Not surprisingly, the cumulative
+adoption is well represented by a curve that resembles the figure S. The
+vertical line around year 2003 illustrates the point where the modelled
+cumulative adoption is equal to 50\% of the maximum. This \(t_0\)
+\emph{inflection} point signals the maximum adoption speed and is used
+here to determine whether a UK LAD reached that inflection point earlier
+or later than the UK average. Specifically, the S curve and the
+inflection point are estimated for every LAD individually and then
+compared against the UK average. LAD that reached their inflection point
+earlier than the country average are labelled as \emph{fast} and the
+rest as \emph{slow}. Figure \ref{s_map} maps this pattern. There is a
+clear high concentration of fast adopting LAD in the South and East of
+England as expected. However, there is also such a concentration in
+Wales, Scotland and the North West of England. At a more detailed level,
+there are some expected examples of LAD with relevant industrial
+backgrounds delineated as fast: the City of London, a world-renowned
+cluster of finance industries \citep{cook2007role}, and Reading, a town
+with high-tech service industries in proximity to London and its main
+airport, Heathrow \citep{england2005polynet}. However, there are also
+LAD which were expected to appear as fast -- e.g.~Hackney in central
+London and Bristol, a well-established creative cluster
 \citep{oatley1999cultural, bassett2002cultural} -- but were delineated
@@
 
-The next step is assesses the stability and volatility of the LAD in
-terms of their adoption of web technologies. As we know from the
-literature \citep{risk_perceptions}, different agents have different
-perceptions about and levels of acceptance of the risks and the
-potential economic returns associated with the adoption of new
-technologies -- see for instance the seminal work of
-\citet{venkatesh2000theoretical}. To reveal such aggregated patterns,
-Figure \ref{rank} plots the ranking of UK LAD based on website density.
-To decrease noise, the average ranking of 1996-1998 and 2010-2012 is
-plotted instead of the individual years. The colours depict the 5th and
-95th percentile of the absolute difference of the LAD ranking between
-1996-1998 and 2010-2012. While there are quite a few obvious cases of
-LAD that maintained their position between the beginning and the end of
-the study period both at the top or at the bottom of the hierarchy,
-there are also quite a few LAD that changed drastically their position.
-Some of these LAD enjoyed a process that at the first instance looks
-like \emph{leapfrogging} since they managed to jump at the top of the
-hierarchy despite their slow start. There is extensive literature
-regarding the potential benefits of technological leapfrogging. The
-underpinning argument is that latecomers can adopt and benefit from new
-technologies that have been developed elsewhere without incurring the
-hefty initial R\&D costs \citep{teece2008firm}. Although the
-leapfrogging literature does not pay much attention to cities and
+The S-curves of the cumulative adoption of Web in the UK in total and in
+the UK LAD demonstrate a pattern well aligned with all previous studies
+discussed in Section~\ref{sec-litreview}. Importantly, the above
+findings are in line with previous studies exploring the diffusion of
+web technologies \citep{PAPAGIANNIDIS2015308}, social media
+\citep{lengyel2020role} and online shopping \citep{bakher2013diffusion}.
+Similar to \citet{beardsell1999spatial}, who analysed the spatial
+evolution of computer industry employment across 317 US metro areas,
+spatial heterogeneity is evident in the diffusion speed when the scale
+of analysis is as detailed as the LAD. Importantly, no previous studies
+estimated such cumulative adoption curves for areal units as small as
+the UK LADS.
+
+The next step assesses the stability and volatility of the LAD in terms
+of their adoption of web technologies. As highlighted in the literature
+\citep{risk_perceptions}, different agents have different perceptions
+about and levels of acceptance of the risks and the potential economic
+returns associated with the adoption of new technologies -- see for
+instance the seminal work of \citet{venkatesh2000theoretical}. To reveal
+such aggregated patterns, Figure \ref{rank} plots the ranking of UK LAD
+based on website density. To decrease noise, the average ranking of
+1996-1998 and 2010-2012 is plotted instead of the individual years. The
+colours depict the 5th and 95th percentile of the absolute difference of
+the LAD ranking between 1996-1998 and 2010-2012. While there are quite a
+few obvious cases of LAD that maintained their position between the
+beginning and the end of the study period both at the top or at the
+bottom of the hierarchy, there are also quite a few LAD that changed
+drastically their position. Some of these LAD enjoyed a process that at
+the first instance looks like \emph{leapfrogging} since they managed to
+jump at the top of the hierarchy despite their slow start. There is
+extensive literature regarding the potential benefits of technological
+leapfrogging. The underpinning argument is that latecomers can adopt and
+benefit from new technologies that have been developed elsewhere without
+incurring the hefty initial R\&D costs \citep{teece2008firm}. Although
+the leapfrogging literature does not pay much attention to cities and
 regions \citep{yu2018sustainability}, previous research highlighted the
@@
 is volatile, it is not clear whether the LAD with high concentration of
-late adopters will gain any latecomer benefits in a way similar to
-countries experiencing such technological leapfrogging. As previously,
-this is consistent with including websites with up to 10 postcodes (see
-Figure \ref{rank10}).
+late adopters will gain any latecomer-related benefits in a way similar
+to countries experiencing technological leapfrogging. Still, the
+opportunity to catch up that \citet{perkins2005international}
+illustrated at the country level, is also visible at the much more
+detailed LAD scale. As previously, this is consistent with including
+websites with up to 10 postcodes (see Figure \ref{rank10}).
 
@@
 neighbourhood mechanism underpins the diffusion of the Web in the UK as
-they plot the Moran's I and the LISA maps of website density
+they present the Moran's I and the LISA maps of website density
 respectively for LAD and OA. Starting from the former, spatial
@@
 early concentration of high website density around London, which over
-time diffused as high-high clusters can be seen in other parts of the
-country away from London (Figure \ref{lisa}). An almost reverse pattern
-can be observed for OA. At the beginning of the study period Moran's I
-was around 0.5 and it plateaued after 2000 around 0.2. Because of the
-very small size of OA, at the early stages of the diffusion of web
-technologies their adoption was spatially scattered. This is reflected
-in the lack of any significant clusters in 1996 (Figure \ref{lisa}).
-Eventually, as the adoption rate increased, more such clusters of high
-website density were formed and this is reflected both in the Moran's I
-and the LISA maps in Figures \ref{morani} and \ref{lisa}. A similar
-pattern is observed for the extended dataset (see Figures\ref{morani10}
-and \ref{lisa10}). In this case, the magnitude of spatial
-autocorrelation is slightly higher illustrating a stronger spatial
-clustering mechanism. All in all, the exploratory spatial data analysis
-advocates towards an underpinning neighbourhood mechanism and the
-different scales of analysis illustrate how it evolved differently over
-time.
+time diffused as high website density clusters can be seen in other
+parts of the country away from London (Figure \ref{lisa}). An almost
+reverse pattern can be observed for OA. At the beginning of the study
+period Moran's I was around 0.5 and it plateaued after 2000 around 0.2.
+Because of the very small size of OA, at the early stages of the
+diffusion of web technologies their adoption was spatially scattered.
+This is reflected in the lack of any significant clusters in 1996
+(Figure \ref{lisa}). Eventually, as the adoption rate increased, more
+such clusters of high website density were formed and this is reflected
+both in the Moran's I and the LISA maps in Figures \ref{morani} and
+\ref{lisa}. A similar pattern is observed for the extended dataset (see
+Figures\ref{morani10} and \ref{lisa10}). In this case, the magnitude of
+spatial autocorrelation is slightly higher illustrating a stronger
+spatial clustering mechanism. All in all, the exploratory spatial data
+analysis advocates towards an underpinning neighbourhood mechanism and
+the different scales of analysis illustrate how it evolved differently
+over time. Interestingly, the magnitude of Moran's I is similar to the
+ones revealed by \citet{ding2010modeling} about mobile phone adoption in
+China albeit the much more detailed spatial scale of this paper.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2023_files/figure-pdf/morani-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/morani-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2023_files/figure-pdf/unnamed-chunk-6-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-6-1.pdf}
 
@@
 Both scales of analysis in Figure \ref{gini} illustrate the same
-picture. At the beginning of the commercial internet website density was
-extremely unequal, or, in other words, only a few places had websites
-associated with them. Inequality dropped and plateaued after 2000 for
-both scales. This is illustrative of a hierarchical diffusion mechanism
-that led over time to a more equal spread. Interestingly, the year 2000
-is again a period of change for this diffusion mechanism as it was for
-the neighbourhood process. There is a substantial difference between the
-Gini coefficient magnitude for LAD and OA, but this is expected as the
-very small size of OA equates to a lot of polygons without any websites
-pointing to them -- for example residential OA.
+picture. At the beginning of the commercial Internet, website density
+was extremely unequal, or, in other words, only a few places had
+websites associated with them. Inequality dropped and plateaued after
+2000 for both scales. This is illustrative of a hierarchical diffusion
+mechanism that led over time to a more equal spread. Interestingly, the
+year 2000 is again a period of change for this diffusion mechanism as it
+was for the neighbourhood process. There is a substantial difference
+between the Gini coefficient magnitude for LAD and OA, but this is
+expected as the very small size of OA equates to a lot of polygons
+without any websites pointing to them -- for example residential OA.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2023_files/figure-pdf/gini-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/gini-1.pdf}
 
@@
 
-\subsection{Modelling results}\label{modelling-results}
+\subsection{Modelling outputs}\label{modelling-outputs}
 
@@
 variables depicting these spatial processes in order to predict the
-diffusion of web technologies in the UK over space and time and across
-different scales. Specifically, four different models are estimated.
-Firstly, all the data points for the OA and LAD are utilised in order to
-build two RF models and assess their capacity to predict the adoption of
-web technologies at the two different scales. These two models will
+diffusion of the Web as a new technology in the UK over space and time
+and across different scales. Specifically, four different models are
+estimated. Firstly, all the data points for the OA and LAD are utilised
+in order to build two RF models and assess their capacity to predict the
+adoption of the Web at the two different scales. These two models will
 reveal the predictive capacity of the diffusion mechanisms and also
-demonstrate how the importance of such variables changes across scales.
-The next two sets of models are again trained at the two working scales:
-OA and LAD. However, instead of using all the data points, the OA and
-the LAD from one of the twelve UK regions are held out for the model
-training. Then the trained model is used to predict website density for
-the OA or the LAD of the held-out region. This process takes place
-recursively for all twelve UK regions. The regional difference in the
-predictive capacity of the different samples will reveal how dissimilar
-these spatial process are across regions and, importantly, at different
-scales.
-
-It needs to be highlighted here that the cross-validation for all models
-is spatially and temporally sensitive. Instead of using 10 random
-folds,the \texttt{CAST} package is employed which allows holding back
-data points from specific years and spatial units and use them for
-testing in order to estimate the model performance
-\citep{meyer2018improving}.
+demonstrate how the importance of such variables changes across
+different scales. The next two sets of models are again trained at these
+two scales: OA and LAD. However, instead of using all the data points,
+the OA and the LAD from one of the twelve UK regions are held out for
+the model training. Then, the trained model is used to predict website
+density for the OA or the LAD of the held-out region. This process takes
+place recursively for all twelve UK regions. The regional differences in
+the predictive capacity of the different samples will reveal how
+dissimilar these spatial process are across regions and, importantly, at
+different scales.
+
+It needs to be highlighted here that the cross validation (CV) for all
+models is spatially and temporally sensitive. Instead of using 10 random
+and space- and time-agnostic folds,the \texttt{CAST} package is employed
+which allows holding back data points from specific years and spatial
+units and use them for testing in order to estimate the model
+performance \citep{meyer2018improving}.
 
@@
 relevant literature and the descriptive analysis highlighted. Namely,
-the models capture: (i) a hierarchy mechanism with diffusion running
+the models capture: (i) a hierarchical mechanism with diffusion running
 from main centres to secondary ones, (ii) a neighborhood mechanism
@@
 
-To capture the hierarchy effect the models include as predictors a one
-year lag of website density in London, the largest city in the UK, a one
-year lag of the website density in the nearest city and the same for the
-nearest retail centre. Due to the small sizes of the retail centres, the
-latter is only relevant for the OA-level models. In addition, the models
-include the distance to London, the nearest city and the nearest retail
-centre. The underlying logic is that the level of website adoption in a
-spatial unit depends on the level of the adoption in places further up
-in the urban hierarchy the previous year. The inclusion of the distance
-variables incorporates spatial structure into the hierarchy argument. To
-depict the neighbourhood effect, the website density of the neighbouring
-spatial units in the previous year is employed (rook continuity). Again,
-the underpinning rationale is that the level of web adoption within a
+To capture the hierarchical mechanism the models include as predictors a
+one year lag of website density in London, the largest city in the UK, a
+one year lag of the website density in the nearest city and the same for
+the nearest retail centre. Due to the small sizes of the retail centres,
+the latter is only relevant for and included in the OA-level models. In
+addition, the models include the distance to London, the nearest city
+and the nearest retail centre. The underlying logic is that the level of
+website adoption in a spatial unit depends on the level of the adoption
+in places further up in the urban hierarchy the previous year. The
+inclusion of the distance variables incorporates spatial structure into
+the hierarchy argument. To depict the neighbourhood mechanism, the
+website density of the neighbouring spatial units in the previous year
+is employed. This is calculated using rook continuity. Again, the
+underpinning rationale is that the level of web adoption within a
 spatial unit depends on the level of web adoption in the neighbouring
@@
 
-To assess the predictive capability of the model, three broadly utilised
+To assess the predictive power of the model, three broadly utilised
 metrics are employed: the coefficient of determination (\(R^2\)), mean
@@
 predicted website adoption. While \(MAE\) does not penalise for large
-errors, \(RSME\) does so as it is proportional to the squared difference
+errors, \(RSME\) does, as it is proportional to the squared difference
 between the observed and the predicted trade flows. Hence larger errors
@@
 first set of models, for which all data points are employed for training
-and testing via cross validation. The first one is trained and tested on
-374 LAD and the second on 232,296 OA, both for a 16 year period
-(1997-2012). The results are remarkably good considering that the are
-the outcome of space and time sensitive CV, so the the model does not
-suffer from overfitting. At the LAD level the model predicts 85\% of the
-variation of website density. Both error metrics indicate that the model
-error is \((\frac{1}{20}, \frac{1}{30})\) of a website per firm. At the
-OA the \(R^2\) drops down to 32\%. Considering its granularity, this is
-still a remarkable performance. To contextualise it, the model results
-in a MAE of one website for areas small enough to host less than 140
-households. \footnote{According to the Office for National Statistics,
-  80\% of OA in England and Wales host 110-139 households,
+and testing via CV. The first one is trained and tested on 374 LAD and
+the second on 232,296 OA, both for the 16 year period (1997-2012). The
+results are remarkably good considering that the are the outcome of
+space and time sensitive CV, so the the model does not suffer from
+overfitting. At the LAD level the model predicts 85\% of the variation
+of website density. Both error metrics indicate that the model error is
+\((\frac{1}{20}, \frac{1}{30})\) of a website per firm. At the OA the
+\(R^2\) drops down to 32\%. Considering its granularity, this is still a
+remarkable performance. To contextualise it, the model results in a
+\(MAE\) equal to one website for areas small enough to host less than
+140 households. \footnote{According to the Office for National
+  Statistics, 80\% of OA in England and Wales host 110-139 households,
   \href{https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas}{www.ons.gov.uk}.}
@@
 one website. This should not come as a surprise as this was the very
-beginning of the commercial internet and any activities with a digital
+beginning of the commercial Internet and any activities with a digital
 footprint were concentrated in a handful of areas. This was clearly
@@
 model. As per Table~\ref{tbl-model-metrics-no-correction} all accuracy
-metrics are slightly worst when the data imputation is applied. Then,
-Table~\ref{tbl-model-metrics-10} presents the model results for the
-extended dataset, which includes websites with up to 10 unique
-postcodes. While the results for the LAD are slightly less accurate, the
-\(R^{2}\) of the OA model is more than double in comparison to the base
-results in Table~\ref{tbl-model-metrics}.
+metrics are only slightly worst when the data imputation is not applied.
+As expected, the improvement is higher for the more detailed scale of
+analysis. Then, Table~\ref{tbl-model-metrics-10} presents the model
+results for the extended dataset, which includes websites with up to 10
+unique postcodes. While the results for the LAD are slightly less
+accurate, the \(R^{2}\) of the OA model is more than double in
+comparison to the base results in Table~\ref{tbl-model-metrics}. All in
+all, the modelling exercise revealed how well can the three diffusion
+mechanisms predict the diffusion of the Web as a new technology.
+Importantly, this predictive power is highly robust against against
+different scales and delineations of the diffusion of the web.
 
@@
 spatial processes are at play in the diffusion of web technologies at
-the LAD level: the first two predictors depict the hierarchy mechanism,
-the spatial and temporal lag of website density the neighbourhood
-mechanism, and the yearly trend the time-sensitive cumulative adoption
-pattern.
+the LAD level: the first two predictors depict the hierarchical
+mechanism, the spatial and temporal lag of website density the
+neighbourhood mechanism, and the yearly trend the time-sensitive
+cumulative adoption pattern.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{tranos2023_files/figure-pdf/varimp-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{anonymised_files/figure-pdf/varimp-1.pdf}
 
@@
 \end{figure}%
-
-Table \ref{table.regions} presents the results of the recursive hold out
-models, which aim to highlight the potential regional heterogeneity of
-the spatial processes behind the diffusion of web technologies. To begin
-with, as highlighted before, there is a difference of magnitude of one
-order between the LAD and the OA prediction errors, which is aligned
-with previous results that employed all data points. What is of interest
-here is the regional comparison. Table \ref{table.regions} illustrates
-some striking similarities, but also a few significant differences. The
-regions the web diffusion of which is better predicted using models
-trained in the rest of the country are the same despite the scale of
-analysis: South East, Wales, Yorkshire and The Humber and the North East
-of England. In other words, these are the regions whose spatial
-diffusion mechanisms of web technologies is closer to the country's
-average. Despite the consistency across scales, this is a diverse set of
-regions: \textbf{ADD CHARACTERISTICS}.
-
-At the other end of the spectrum, Scotland's and the North West's web
-diffusion mechanisms are consistently diverging from the country's
-average. This should not come as a surprise as these regions are
-characterised of high levels of rurality and remoteness. Similarly,
-London diffusion mechanisms diverge from the country's average and this
-is consistent across scales. London's uniqueness in UK's urban system
-and economy is also reflected in the spatial diffusion mechanisms of web
-technologies within its LAD and OA. It needs to be highlighted though
-that the difference between the \(R^2\) of LAD and OA is more that an
-order of magnitude signaling how difficult is to predict diffusion at
-such a small spatial scale. Northern Ireland is an interesting case.
-While it ranks at the bottom of the scale when the models are trained
-and tested on LAD data, when the modelling adopts the more granular OA
-scale, the spatial mechanisms that shape the web diffusion within this
-region appear to be closer to the country's average. At this scale,
-proximity, or lack of, relative to the rest of the country become less
-important and the internal to the region spatial structure predictors
-start playing a more important role \textbf{CHECK NI OA model}.
 
@@
 \endlastfoot
-South East & 0.945 & 1 & 0.183 & 7 \\
-West Midlands & 0.922 & 2 & 0.190 & 5 \\
-Wales & 0.915 & 3 & 0.193 & 4 \\
-Yorkshire and The Humber & 0.909 & 4 & 0.200 & 2 \\
-East Midlands & 0.903 & 5 & 0.187 & 6 \\
-North East & 0.896 & 6 & 0.156 & 9 \\
-South West & 0.894 & 7 & 0.221 & 1 \\
-London & 0.881 & 8 & 0.104 & 10 \\
-East of England & 0.874 & 9 & 0.170 & 8 \\
-North West & 0.837 & 10 & 0.194 & 3 \\
-Scotland & 0.780 & 11 & 0.091 & 12 \\
-Nortern Ireland & 0.579 & 12 & 0.099 & 11 \\
+South East & 0.945 & 1 & 0.116 & 8 \\
+West Midlands & 0.922 & 2 & 0.125 & 4 \\
+Wales & 0.915 & 3 & 0.125 & 5 \\
+Yorkshire and The Humber & 0.909 & 4 & 0.160 & 2 \\
+East Midlands & 0.903 & 5 & 0.120 & 7 \\
+North East & 0.896 & 6 & 0.124 & 6 \\
+South West & 0.894 & 7 & 0.184 & 1 \\
+London & 0.881 & 8 & 0.093 & 9 \\
+East of England & 0.874 & 9 & 0.092 & 10 \\
+North West & 0.837 & 10 & 0.143 & 3 \\
+Scotland & 0.780 & 11 & 0.055 & 12 \\
+Nortern Ireland & 0.579 & 12 & 0.060 & 11 \\
 \end{longtable}
 
-\section{Discussion and conclusions}\label{sec-conclusions}
-
-contrary to results from future studies regarding social media
-\citep{lengyel2020role}, web technologies did not exclusively spread
-from a central location.
+Table \ref{table.regions} presents the results of the recursive hold out
+models and allows to (i) highlight the regional heterogeneity of the
+spatial processes behind the diffusion of the Web, and (ii) compare this
+heterogeneity across the two different scales of analysis. It
+illustrates some striking similarities, but also a few significant
+differences. At the top of the similarity with the rest of the country
+we see West Midlands and Yorkshire and the Humber for both scales. The
+web density for the LAD and OA of these regions can be very well
+predicted by models trained on data for the LAD and OA of the rest of
+the country. In other words, the spatial diffusion mechanisms for these
+regions represent quite well the UK's average. This should not come as
+as surprise given their urban structure: they both contain a large urban
+centre (Birmingham and Leeds), smaller commuter towns as well as more
+rural settlements. The bottom of the similarity ranking is populated by
+Scotland, Northern Ireland, East of England and London for both scales.
+The diffusion mechanisms for these regions diverge significantly from
+the country's average and this could be attributed to rurality and
+remoteness or to high levels of urbanisation and sophistication of
+economic activities for the case of London and for East of England,
+which contains the Cambridge area.
+
+Still, there are three regions the spatial diffusion mechanisms of which
+differ a lot between the two scales. North West and South West are very
+close to the UK's average at the more granular OA scale, while the
+opposite applies for South East. Interestingly, the first two regions
+are well-known tourist destinations (Lake District and Cornwall-Devon)
+and the exploratory analysis revealed clusters of high web density only
+visible at the OA scale (Figure \ref{lisa}. Then, the South East is very
+well integrated with the Greater London, hosts the broader Oxford area
+-- a known hub of research excellence -- and Figure \ref{lisa} revealed
+high web density clusters at the LAD level.
+
+\section{Conclusions}\label{sec-conclusions}
+
+This paper assessed the importance of spatial mechanisms in the
+diffusion of a new technology -- the Web -- at a high level of spatial
+granularity. Specifically, it assessed the capacity of the following
+three spatial diffusion mechanisms in predicting website density in the
+UK over space and time: (i) the hierarchical mechanism according to
+which diffusion runs from main urban centres to secondary ones, (ii) the
+neighborhood mechanism which highlights that diffusion first hits nearby
+locations, and (iii) the rather canonical pattern of diffusion over time
+as reflected in the S-shaped pattern of the cumulative level of
+adoption. This is the first time that the importance of granular spatial
+mechanisms for the diffusion of an intangible, digital technology is
+exposed. The time period of the study is also important as it captures
+the very early stages of this technology until its maturity.
+Importantly, the empirical framework and the data employed here allows
+to observe the active engagement with the Web that is creating and
+maintaining a commercial website instead of the more passive action of
+browsing webpages and accessing the Internet.
+
+The analysis firstly illustrated how well the S-pattern of cumulative
+adoption fits at the local scale of LAD. It highlighted the
+heterogeneity of adoption rates between LAD -- with both more and less
+expected examples -- as well as the spatial clustering of LAD with
+similar adoption rates. The exploratory analysis also illustrated
+instances of remarkable stability as well as volatility of the web
+diffusion system that is indicative of leapfrogging. Previous research
+had already indicated the long-term productivity gains of the early
+adoption of such technologies \citep{tranosuk}. Then, the paper
+developed a robust and novel ML modelling framework to predict website
+diffusion over space and time using the three spatial diffusion
+mechanisms based on space- and time-sensitive CV without suffering from
+overfitting. The model revealed the remarkably good predictive capacity
+of these mechanisms at both scales. For the first time, the power of
+these mechanisms in predicting the diffusion of a new technology was
+empirically tested, importantly, at very fine scales.
+
+The modelling framework also revealed that the importance of these
+spatial mechanisms differ across different scales. While spatial
+structure as reflected in distances to different urban centres is highly
+important at the OA scale, it hardly plays any role at the more coarse
+LAD scale. And while the almost canonical pattern of cumulative adoption
+is a key predictor at the LAD level, it only plays a small role at the
+detailed OA level. Then, the analysis revealed another level of spatial
+heterogeneity. While for some UK regions these spatial diffusion
+mechanisms are very similar to the rest of the country, other regions
+diverge significantly from the country average. Importantly, the above
+results are robust against different specifications and subsets of the
+data.
+
+The paper offers a number of different contributions. Firstly, it
+assesses, for the first time, the role of these spatial diffusion
+mechanisms at such granular spatial scales. While previous studies had
+highlighted these mechanisms
+\citep{hagerstrand1968innovation, rogers2010diffusion, grubler1990rise},
+their role had never been tested before at such granular scale and,
+importantly, for a digital technology as weightless as the Web. The
+findings talk directly to this literature
+\citep[e.g.][]{beardsell1999spatial, haller2011determinants, lengyel2020role}.
+More broadly, the paper offers a robust methodological framework, which
+utilises novel data and modern ML algorithms and techniques to model the
+diffusion of a new technology across space and time. It directly
+addresses the data problem which prevented economic geographers in
+engaging with diffusion studies
+\citep{iso2005, kemeny2011international, zook2022mapping}. Thus, the
+paper seeks to renew economic geography's interest in the diffusion of
+new technologies. Current data science advances including the
+availability of granular enough data and adequate methods can enable
+economic geographers to reveal interesting spatial patterns of diffusion
+of new technologies as this paper has demonstrated. This is important as
+(i) various stakeholders are interested in predicting localised demand
+for new technologies for planning purposes
+\citep{leibowicz2016representing, meade2021modelling}, and (ii) the
+adoption of such digital technologies is associated with economic gains
+\citep{solow1957technical, aghion1990model, kemeny2011international, tranosuk, capello2024nexus}.
+In addition, while most related studies focused on tangible
+technologies, this paper sheds light on the diffusion of a truly
+weightless digital technology that is the Web. Thus, it adds much more
+granular findings to the limited related studies
+\citep{sinai2004geography, haller2011determinants}. Importantly,
+although the geographical focus of this paper is the UK, the empirical
+framework developed here is easily transferable to other geographical
+contexts.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2023_files/figure-pdf/morani10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/morani10-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2023_files/figure-pdf/unnamed-chunk-10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-10-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2023_files/figure-pdf/gini10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/gini10-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{tranos2023_files/figure-pdf/varimp10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{anonymised_files/figure-pdf/varimp10-1.pdf}
 

### 2025-01-03 paper/submission1/tranos2023.qmd
ssh problem

@@
 abstract: |
-  This paper maps the participation in the digital economy and its evolution in the UK over space and time.
-Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects.
-Useful as these approaches might have been, they cannot capture the micro-processes and the characteristics of the individual online behaviour.
-Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK.
-The data and geolocation strategy allow to capture these processes at a very granular spatial scale.
-The modelling approach, which is based on simple spatial analytical methods and on the estimation of diffusion curves at various scales, enables to depict the role of geography and other cognitive factors which drove the diffusion of web technologies.
-Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.
+    Despite the broad consensus that differences in technological adoption and sophistication can lead to uneven trajectories of local development as new technologies can increase productivity and allow firms to introduce new products and services, economic geographers have shied away from exploring the diffusion of digital technologies across space and time.
+This paper fills this gap by exposing the roles of different spatial mechanisms in the diffusion of an intangible, digital technology at very granular geographical scales.
+It does so by employing a novel source of data and machine learning algorithms to model the active engagement with Web during 1996-2012.
 keywords: 
-  - keyword1
-  - keyword2
+  - diffusion
+  - technologies
+  - Web
+  - digital
 date: last-modified

### 2025-01-03 paper/submission2/tranos2025.tex
ssh problem

--- 2025-01-03 paper/submission1/anonymised.tex
@@
   pdftitle={A multi-scale story of the diffusion of a new technology: the web},
+  pdfauthor={Emmanouil Tranos},
   pdfkeywords={diffusion, technologies, Web, digital},
@@
 web}
+\author[1]{Emmanouil Tranos%
+\corref{cor1}%
+}
+ \ead{e.tranos@bristol.ac.uk}
+
+\affiliation[1]{organization={University of
+Bristol},postcode={UK},postcodesep={}}
 
@@
 before the `cultural turn' of economic geography
-\citep{perkins2005international}. A potential explanation of the lack of
-geographical studies exploring the diffusion of new and, more
-specifically for this paper, digital technologies across both space and
-time can be attributed to the scarcity of relevant and granular enough
-data, a problem also highlighted by \citet{iso2005} and
-\citet{kemeny2011international}. As \citet{zook2022mapping} pointed,
-digital activities are hardly ever captured in official data.
+\citep{perkins2005international, ding2010modeling}. As a consequence,
+relevant studies did not investigate diffusion mechanisms at granular
+geographical scales to provide local policy insights, for instance,
+regarding infrastructural d evelopment and local technology policies. A
+potential explanation of the lack of geographical studies exploring the
+diffusion of new and, more specifically for this paper, digital
+technologies across both space and time can be attributed to the
+scarcity of relevant and granular enough data, a problem also
+highlighted by \citet{iso2005} and \citet{kemeny2011international}. As
+\citet{zook2022mapping} pointed, digital activities are hardly ever
+captured in official data.
 
@@
 intangible, digital technology that is the Web in the UK at a high level
-of spatial granularity during the 1996-2012 period. It does so by
-employing a novel source of big data and machine learning algorithms to
-model the \emph{active} engagement with Web during that period and tests
-how different spatial and temporal mechanisms shaped its diffusion. This
-paper also exemplifies how the combination of data sources and
-state-of-the-art research methods which escape the traditional social
-science domain can offer new lenses to geographical research regarding
-the understanding of technological diffusion.
+of spatial granularity from the very early days until its maturity
+(1996-2012). It specifically focuses on the diffusion of commercial
+websites. It does so by employing a novel source of big data and machine
+learning algorithms to model the \emph{active} engagement with Web --
+that is creating and maintaining instead of just browsing a website --
+during that period and tests how different spatial and temporal
+mechanisms shaped its diffusion. This paper also exemplifies how the
+combination of data sources and state-of-the-art research methods which
+escape the traditional social science domain can offer new lenses to
+geographical research regarding the understanding of technological
+diffusion.
 
@@
 \citep{solow1957technical, aghion1990model, kemeny2011international}.
-Linked to the subject of this paper, the Web -- and the active
-engagement with this technology that is creating and maintaining instead
-of just browsing a website -- can be considered as the underpinning
-technology for the transformative process identified by
-\citet{capello2024nexus} as the \emph{digital service economy}.
-Importantly, the focus here is not the Web in general, but instead
-commercial websites. After all, diffusion together with invention and
-innovation are considered the pillars of technological change
-\citep{das2022diffusion}. Findings from previous studies illustrated
-that early engagement with digital technologies such as the Web led to
-longstanding positive productivity gains for these regions that hosted
-early adopters \citep{tranosuk}. Hence, obtaining a robust understanding
-of the spatial diffusion mechanisms is crucial in mapping future
-development trajectories.
-
-Another line of motivation has to do with the intangible nature of the
-Web and the debate about the footloose nature of digital technologies,
-which contrary to several earlier general purpose technologies are
-`light in weight' \citep{coyle1999weightless}. Following the seminal
-work of \citet{gaspar1998information}, which demonstrated that
-technological improvements in telecommunications lead to increased
-demand for face-to-face communications and, consequently, enhance the
-importance of cities, empirical studies from economic geography and
-urban economics explored whether digital technologies substitute or
-complement agglomeration externalities and spatial structure
-\citep{kolko_death_2000, ioannides2008effect, tranos2021ubiquitous}.
-With the exception of \citet{sinai2004geography}, who focused, among
-other things, on web content and demonstrated that more such content is
-available in larger markets, this strand of literature has been mostly
-concerned with the spatiality of the Internet's hard and tangible
-infrastructure and shied away from exploring the spatial dimensions of
-one of the most distinct software layers of the Internet infrastructure,
-the Web.
-
-Technological diffusion, which is by definition an aggregated process,
-can be discussed in parallel with individual adoption mechanisms. On the
-one hand, \citet{rogers2010diffusion} identifies early adopter of new
-technologies as `knowledgeable risk takers' and \citet{griliches1957} as
-`profit maximisers' \citep{ding2010modeling}. Such individual agents are
-rewarded because of their attitude towards new technologies and
-innovations. On the other hand, \citet{perkins2011internet} attribute
-diffusion to two processes: (i) epidemic-like mechanisms, which are
-governed by distance, proximity and social interactions, and (ii)
-economic mechanisms as new innovations are adopted by users as they
-become more profitable, valueable and useful.
+Linked to the subject of this paper, the Web and, more specifically,
+commercial websites can be considered as the underpinning technology for
+the transformative process identified by \citet{capello2024nexus} as the
+\emph{digital service economy}. Importantly, the focus here is not the
+Web in general, but instead commercial websites. After all, diffusion
+together with invention and innovation are considered the pillars of
+technological change \citep{das2022diffusion}. Findings from previous
+studies illustrated that early engagement with digital technologies such
+as the Web led to longstanding positive productivity gains for these
+regions that hosted early adopters \citep{tranosuk}. Hence, obtaining a
+robust understanding of the spatial diffusion mechanisms is crucial in
+mapping future development trajectories.
+
+Another line of motivation has to do with the very nature of the Web as
+a digital technology and how it evolved over time. The digital
+geographies field has debunked premature claims about the Internet's
+placeless-ness by mapping various of its facets and technological layers
+\citep{tranos2013death}, but the lack of relevant long time-series data
+prevented researchers from observing digital phenomena overtime. Most of
+these studies focused on its hard infrastructure
+\citep[e.g.][]{malecki2002economic, moss2000internet}, on Internet
+subscribers \citep[e.g.][]{blank2018local}, or on social media users
+\citep[e.g.][]{crampton2013beyond}. Exceptions include early work from
+\citet{zook2000web}, \citet{zook2000web} and \citet{moss1997tracking},
+which utilised domain name registration data and use it to map and
+illustrate the spatial footprint of the web content. Pivotal as these
+contributions have been in understanding the economic geographies of the
+Internet, they were based on cross-sectional, often spatially detailed,
+mapping and on shallow time comparisons (e.g.~1993-1998) and did not
+observe the evolution and diffusion of the Web until its maturity. As we
+know from the relevant literature, different technologies diffuse at
+different rates and there is no theory that can explain these
+differences \citep{leibowicz2016representing}. Hence, mapping and
+modelling the diffusion of the Web among firms can offer useful
+insights.
 
@@
 \citep[e.g.][]{blank2018local, destefano2022fuzzy}, this paper observes
-yearly counts of geolocated websites, which reflect the creation and
-maintenance of websites. It needs to be highlighted here that although
-there are various types of underpinning web technologies -- namely
-markup, stylesheet, client- and server-side coding languages as well as
-different web browsers and other related programming language -- the
-focus here is on the diffusion of the Web in its totality. For a more
-detailed and cross-country technological diffusion study, see
+yearly counts of geolocated, commercial websites, which reflect the
+creation and maintenance of such websites. It needs to be highlighted
+here that although there are various types of underpinning web
+technologies -- namely markup, stylesheet, client- and server-side
+coding languages as well as different web browsers and other related
+programming language -- the focus here is on the diffusion of the Web as
+an overarching technology instead of more specific web technologies. For
+a such a detailed and cross-country technological diffusion study, see
 \citet{PAPAGIANNIDIS2015308}.
@@
 
-\section{Literature review}\label{sec-litreview}
-
-Geographical diffusion is a synthesis of different processes. On the one
-hand, a purely spatial or, in other words, contagious processes can be
-identified. Adjacency and, more broadly, distance are the key drivers of
-diffusion. This perspective draws similarities with epidemics:
-innovations just like pathogens spread because of contagion and,
-consequently, proximity and exposure \citep{hivner2003facilitating}. On
-the other hand, there is a hierarchical processes. According to this,
-instead of a horizontal distance-based diffusion mechanism, the top-down
-hierarchy of urban systems shapes technological diffusion. In reality,
-the synthesis of these two processes better reflects how new
-technologies diffuse over space \citep{morrill2020spatial}.
-
-These ideas were firstly introduced by Torsten Hägerstrand and his
-thesis entitled `Innovation Diffusion as a Spatial Process'
-\citep{hagerstrand1968innovation}. Hägerstrand was the first to identify
+\section{The mechanisms that shape the diffusion of a new
+technology}\label{sec-litreview}
+
+The diffusion of new technologies and innovations is driven by three
+different mechanisms \citep{grubler1990rise, morrill2020spatial}: (i)
+distance and proximity, (ii) urban hierarchies in the form of a
+centrifugal forces as new technologies diffuse from core to periphery,
+and (iii) over time the cumulative level of adoption follows an S-shaped
+pattern. \citet{hagerstrand1968innovation} was the first to identify
 diffusion as a geographical process. The starting point was the idea
 that diffusion is based on passing information through social networks,
-which themselves tend to be defined by geography. Hence, he identified a
-`neighbourhood' mechanism of how information, and consequently,
-innovation diffuse. He used agricultural innovations to test and model
-his ideas using Monte Carlo simulations. Hägerstrand also incorporated
-the role of hierarchy and how some phenomena maybe firstly adopted in
-larger cities and then diffuse to second tier ones. This is a sequential
-instead of a simultaneous process, which resembles the `lead-lag'
-spatial acceleration effect in market research
-\citep{bento2018time, PERES201091}. Hägerstrand is more widely known
-though for highlighting the role time plays in the diffusion of
-innovations: an early-pioneering period, a middle fast accelerating
-period and a final saturation one \citep{morrill2020spatial}.
-
-The temporal dimension was further explored by Everett Rogers and his
-seminal work on `Diffusion of Innovations' \citep{rogers2010diffusion}.
-Rogers being a sociologist focused not on the diffusion of innovations
-over space and time, but instead on the adoption of new technologies and
-innovations by individuals and the individual mechanisms that drive the
-decisions behind adoption. He identified five groups of individuals
-regarding their adoption speed: innovators, early adopters, early
-majority, late majority and laggards. The key mechanism of diffusion and
-adoption is communication and how knowledge is transferred within a
-social system. All above approaches agree that diffusion and the
-cumulative adoption of a new technology are represented by an S-shaped
-line plot \citep{grubler1990rise}.
-
-Schmidt's Law empirically illustrates a similar pattern. \emph{Core} and
-usually highly agglomerated regions is where new technologies are
-invented and commercially deployed \citep{grubler1990rise}. This is
-where the first adopters tend to be based. Then, technologies spread to
-the \emph{rim} and eventually to the \emph{periphery}. Although adoption
-pace might be higher when new technologies finally arrive to the
-periphery, the saturation levels there may never reach the ones in the
-core because of the lack of infrastructure or other necessary
-institutions \citep{leibowicz2016representing}.
-
-\citet{grubler1990rise} effectively summarises the three key spatial
-diffusion mechanisms: (i) the cumulative level of adoption follows an
-S-shaped pattern just like purely temporal models; (ii) diffusion is
-shaped by a hierarchical mechanism in a form of a centrifugal force:
-from core to periphery; and (iii) diffusion is also shaped by distance
-and proximity. These are the three mechanisms that the empirical
-analysis in Section~\ref{sec-results} investigates.
-
-The remaining of this section reviews empirical studies which analysed
-the diffusion on new technologies over space and time. Although the
-spatial dimension is present in most of the following studies, the level
-of spatial detail is always more coarse than the one adopted in this
-paper. \citet{beardsell1999spatial} studied the evolution of the
-computer industry in 317 US metro areas during the 1977-1992 period
-using employment data. Their analysis indicated that the relative size
-distribution holds for urban computer employment and also urban
-heterogeneity is essential in explaining this distribution. In a recent
-study, \citet{bednarz2020pulled} focused on wind turbines and modelled
-their spatial diffusion across 402 German regions during 1970-2015.
-Their key finding is that local demand than local supply was the main
-driving factor. \citet{haller2011determinants} employed firm panel data
-for Ireland during 2003-2005 and illustrated that when firms are located
-in a region or industry with high share of firms having a website, they
-are more likely to have a website too.
-
-At a global scale \citet{perkins2005international} explored whether the
-diffusion rate of new technologies is driven by a latecomer advantage
-and the engagement with the global economy via foreign direct
-investments and trade. Their results illustrate that indeed latecomers
-and developing countries experience diffusion of new technologies more
-rapidly than early adopters and developed countries. At the same scale,
-\citet{perkins2011internet} explored whether the adoption of previous
-communication technologies that is mail, telegrams and telephones were
-shaped by similar socioeconomic factors as the Internet. Their results
-indicated common patterns regarding the drivers behind the adoption of
-different communication technologies.
-
-Turning to studies that share more technological and scalar similarities
-with this paper, \citet{ding2010modeling} modelled the spatial diffusion
-of mobile telecommunications across regions in China. Their analysis
+which themselves tend to be geographically bounded. Hence, he identified
+a `neighbourhood' mechanism of how information, and consequently,
+innovations and new technologies diffuse. This perspective draws
+similarities with epidemics: new technologies and innovations, just like
+pathogens, spread because of contagion and, consequently, proximity and
+exposure \citep{hivner2003facilitating}.
+
+\citet{hagerstrand1968innovation} also emphasised the role of the
+top-down hierarchy of urban systems and highlighted that new
+technologies and innovations are firstly adopted in larger cities and
+then diffuse to second tier ones. This is a sequential instead of a
+simultaneous process, which resembles the `lead-lag' spatial
+acceleration effect in market research
+\citep{bento2018time, PERES201091}. Schmidt's Law empirically
+illustrates a similar pattern \citep{grubler1990rise}. Core and highly
+agglomerated regions is where new technologies are invented and
+commercially deployed. This is where the first adopters tend to be
+based. Then, technologies spread to the rim and eventually to the
+periphery. Although adoption pace might be higher when new technologies
+finally arrive to the periphery, the saturation levels there may never
+reach the ones in the core because of the lack of infrastructure or
+other necessary institutions \citep{leibowicz2016representing}.
+
+Diffusion is naturally a dynamic process and Hägerstrand recognised the
+role time plays in the diffusion of innovations and new technologies: an
+early-pioneering period, a middle fast accelerating period and a final
+saturation one \citep{morrill2020spatial}. This was further explored by
+\citet{rogers2010diffusion}, who identified five types of adopters of
+new technologies and innovations: innovators, early adopters, early
+majority, late majority and laggards. Hence, plotting the cumulative
+adoption of a new technology over time often leads to an S-shaped line
+plot \citep{grubler1990rise}.
+
+Although the above processes have been studied in innovation studies,
+the distancing of economic geographers from research questions about the
+diffusion of new technologies and innovations resulted to a knowledge
+gap regarding the explanatory power of these mechanisms at granular,
+local scales \citep{ding2010modeling}. As discussed in
+Section~\ref{sec-introduction}, such spatially explicit knowledge can
+directly inform local technology and infrastructure policies. Indeed,
+while the spatial dimension is present in the following studies, the
+geographical level of analysis is often too coarse to ascertain the role
+of the above mechanisms at the local scale. \citet{beardsell1999spatial}
+studied the evolution of the computer industry in 317 US metropolitan
+areas during the 1977-1992 period using employment data. Their analysis
+indicated that the relative size distribution holds for urban computer
+employment and also urban heterogeneity is essential in explaining this
+distribution. In a recent study, \citet{bednarz2020pulled} focused on
+wind turbines and modelled their spatial diffusion across 402 German
+regions during 1970-2015. Their key finding is that local demand than
+local supply was the main driving factor. \citet{haller2011determinants}
+employed firm panel data for Ireland during 2003-2005 and illustrated
+that when firms are located in a region or industry with high share of
+firms having a website, they are more likely to have a website too. At a
+global scale \citet{perkins2011internet} explored whether the adoption
+of previous communication technologies that is mail, telegrams and
+telephones were shaped by similar socioeconomic factors as the Internet.
+Their results indicated common patterns regarding the drivers behind the
+adoption of different communication technologies.
+\citet{feldman2015rknowledge} focused on inventions and used patent data
+to map how the the rDNA technology diffused among inventors across 366
+US metropolitan areas. \citet{ding2010modeling} modelled the spatial
+diffusion of mobile telecommunications across the highest level of
+administrative division in China (29 regions in total). Their analysis
 indicated that socioeconomic characteristics are important determinants
 of the timing, speed and the level of mobile diffusion within China.
-Using data from a Hungarian online social network,
-\citet{lengyel2020role} analysed its adoption and the churn at a very
-granular spatial level. Their results were in agreement with the early
-theoretical and empirical contributions reviewed here: assortativity,
-urban scaling and distance are the key drivers of spatial diffusion.
 \citet{bakher2013diffusion} compared different the growth curves as a
-means to model diffusion of online shopping in Australia during the
-1998-2009 period. At a global scale \citet{PAPAGIANNIDIS2015308}
-modelled the diffusion of different web technologies and practices
-across countries. Using similar but less extensive data as the one used
-here, their analysis illustrated how the diffusion of different web
-technologies and practices follow an S-shaped pattern as well as the
-different diffusion rates of the different technologies and practices.
-
-All in all, this paper tests whether the three distinct spatial
-diffusion mechanisms shaped the diffusion of the Web in the UK at very
-granular spatial scales from the Web's early stages onwards. The next
-section discusses the empirical framework.
+means to model the diffusion of online shopping in Australia during the
+1998-2009 period, but they did not consider any spatial mechanisms. At a
+global scale \citet{PAPAGIANNIDIS2015308} modelled the diffusion of
+different web technologies and practices across countries. Using similar
+but less extensive data as the one used here, their analysis illustrated
+how the diffusion of different web technologies and practices follows an
+S-shaped pattern as well as the different diffusion rates of the
+different technologies and practices. The most detailed consideration of
+the spatial footprint of the diffusion of a new technology comes from
+\citet{lengyel2020role}, who analysed the adoption and churn of a
+Hungarian online social network. Their results were in agreement with
+the early theoretical and empirical contributions reviewed here:
+assortativity, urban scaling and distance are the key drivers of spatial
+diffusion. Based on the above, the main empirical hypothesis states:
+
+\begin{quote}
+H\textsubscript{1}: The three diffusion mechanisms namely distance,
+urban hierarchy and the the S-shaped pattern of the cummulative level of
+adoption shaped the diffusion of the commercial Web in the UK even at
+granular geographical scales.
+\end{quote}
+
+Understanding the diffusion of a new technology in multiple, detailed
+spatial scales is of interest to economic geography as potential local
+economic benefits may occur because of concentration of early adopters
+or through technological leapfrogging. On the one hand, early adopters
+are often rewarded because of their attitudes towards new technologies
+and innovations. \citet{rogers2010diffusion} identifies them as
+`knowledgeable risk takers' and \citet{griliches1957} as `profit
+maximisers' \citep{ding2010modeling}. On the other hand, latecomers can
+adopt and benefit from new technologies that have been developed
+elsewhere without incurring the hefty initial R\&D costs
+\citep{teece2008firm}. At a global scale
+\citet{perkins2005international} explored whether the diffusion rate of
+new technologies is driven by a latecomer advantage and the engagement
+with the global economy via foreign direct investments and trade. Their
+results illustrate that indeed latecomers and developing countries
+experience diffusion of new technologies more rapidly than early
+adopters and developed countries. At a regional scale, previous research
+highlighted the long term and sustained productivity benefit of the
+early adopters of web technologies in the UK \citep{tranosuk}. Although
+the early adoption and leapfrogging literature has not paid much
+attention on cities and regions \citep{yu2018sustainability}, the
+potential local economic benefits justify the need to investigate the
+dynamics of the web diffusion in different scales in the UK. Hence, the
+second empirical hypothesis states:
+
+\begin{quote}
+H\textsubscript{2}: The diffusion of the commecrial Web is shaped by
+both early adoption and late comer effects.
+\end{quote}
+
+These are the two hypotheses are empirically tested in
+Section~\ref{sec-results} using the data and methods described in the
+next section.
 
@@
 
-To capture the diffusion of the Web, a website density metric is
-developed for two different geographical scales: the Local Authority
-Districts (LAD) and the Output Areas (OA). The former is an
-administrative unit and there are 374 LAD in the UK. The latter is the
-smallest census-based geographical unit and there are c.~230,000 of them
-in the UK. This methodological choice will allow the mapping of the
-diffusion of web technologies and the assessment of the diffusion
-mechanisms at these two very different spatial scales.
-
-The counts of websites at these scales are calculated using data from
-the Internet Archive\footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
+To capture the diffusion of the commercial Web, a website density metric
+is developed for two different geographical scales that are discussed
+later in this section. The main data are counts of commercial websites
+that are calculated using data from the Internet Archive\footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
 and, specifically, the JISC UK Web Domain Dataset \citep{ukwebarchive}.
@@
 Instead of using the whole .uk ccTLD, this paper focuses on its
-commercial subset, the .co.uk second level domain (SLD). This choices
-decrease the heterogeneity of the web data as such commercial websites
-have specific aims: they are used to diffuse information, support online
-transactions and share opinions
+commercial subset, the .co.uk second level domain (SLD). This choice is
+aligned with the topic of the paper -- the diffusion of the Web for
+commercial functions -- and decreases the heterogeneity of the web data
+as such commercial websites have specific aims: they are used to diffuse
+information, support online transactions and share opinions
 \citep{THELWALL2000441, blazquez2018big}. Although a UK company can
@@
 \citep{oecd_coms}. Importantly, previous studies illustrated that .co.uk
-is the most popular UK SLD \citep{tranosuk}.
+is the most popular UK SLD \citep{tranosuk}. Henceforward any reference
+to websites will refer to commercial -- .co.uk -- websites.
 
@@
 information allows the geolocation of the data and the creation of the
-LAD and OA counts.
+LAD and OA counts. To prevent false positives, the resulted postcodes
+were matched against official postcode lookup files \citep{ons}.
 
@@
 metrics would have been biased towards large, place specific, websites.
+It is important to note here that websites commonly contain place
+references, with previous research estimating this share to be around
+70\% \citep{hill2009georeferencing}.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{anonymised_files/figure-pdf/unnamed-chunk-2-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-2-1.pdf}
 
@@
 \end{figure}%
+
+the Local Authority Districts (LAD) and the Output Areas (OA). The
+former is an administrative unit and there are 374 LAD in the UK. The
+latter is the smallest census-based geographical unit and there are
+c.~230,000 of them in the UK. This methodological choice will allow the
+mapping of the diffusion of web technologies and the assessment of the
+diffusion mechanisms at these two very different spatial scales.
 
@@
 attributes match well with the website density data characteristics
-including skewness especially for the OA. Also, the large data size
-(c.~230k data points for each of the 17 years) calls for fast training
+including skewness especially for the OA. Also, the large data size (c.
+230k data points for each of the 17 years) calls for fast training
 times. Importantly, RF predictions tend to be more accurate than those
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/morani-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2025_files/figure-pdf/morani-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-6-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-6-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/gini-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2025_files/figure-pdf/gini-1.pdf}
 
@@
 R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
-\end{align}
-
-\begin{align}
+\end{align} \begin{align}
 MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
-\end{align}
-
-\begin{align}
+\end{align} \begin{align}
 RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
@@
 \begin{longtable}[]{@{}llll@{}}
-\caption{Model metrics without data
-imputation}\label{tbl-model-metrics-no-correction}\tabularnewline
+\caption{Model metrics without data imputation
+}\label{tbl-model-metrics-no-correction}\tabularnewline
 \toprule\noalign{}
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{anonymised_files/figure-pdf/varimp-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{tranos2025_files/figure-pdf/varimp-1.pdf}
 
@@
 \renewcommand{\thesection}{\Alph{section}}
-
 \setcounter{table}{0}
 \renewcommand{\thetable}{A\arabic{table}}
-
 \setcounter{figure}{0}
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/morani10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2025_files/figure-pdf/morani10-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-10-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/gini10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2025_files/figure-pdf/gini10-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{anonymised_files/figure-pdf/varimp10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{tranos2025_files/figure-pdf/varimp10-1.pdf}
 

### 2025-01-21 presentations/old/web_diffusion.Rmd
early/late

--- 2024-06-11 presentations/web_diffusion.Rmd
@@
 ---
-title: 'A multi-scale story of the diffusion of a new technology: the web'
+title: 'Digital economy in the UK: an evolutionary story'
 # date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
@@
 
+<br>
+
 ::: {style="margin-left: 50px !important; text-align:left !important"}
 -   Introduction
-
 -   Web data
-
--   Methods
-
--   Results
-
+-   Spatial analysis
 -   S-shaped diffusion curves
 -   Rank dynamics
-    -   Exploratory spatial analysis
-    -   Modelling
-
 -   Conclusions <br> <br>
@@
 
-## Aim
+## Aims
+
+<br>
+
+-   Map the **active** engagement with the *digital*
+-   Over time, **early stages** of the internet
+-   **Granular** and **multi-scale** spatial perspective
+
+## Aims
 
@@
 ## Importance
+
+-   Understand how the adoption of new technologies evolves
 
@@
 
-## Diffusion of a new *digital* technology
-
-<br>
-
--   Diffusion of an **intangible**, digital technology [*web*]
-
--   Map the **active** engagement with the *digital*
-
--   Over time, **early stages** of the internet [*1996-2012*]
-
--   Granular and multi-scale **spatial processes**
-
 # Web data
@@
 
--   Data from the Internet Archive, the oldest web archive
+<br>
+
+-   Archived web data
 
@@
 
--   Geolocation: postcode references in the text
+-   Geolocate:
 
--   Timestamp: archival year
+    - to a unique postcode
 
--   Counts
+    - to multiple postcodes
 
-## Web data: The Internet Archive
-
-<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture1.png){width="70%"} -->
-<!-- ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture1.png){width="70%"} -->
-![](images/Picture1.png){width="70%"}
-
-## Web data: The Internet Archive
-
-<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture2.png){width="70%"} -->
-<!-- ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture2.png){width="70%"} -->
-![](images/Picture2.png){width="70%"}
-
-## Long story short
-
--   Data from a Web Archive -- The Internet Archive
-
--   Observe commercial websites 1996 - 2012 in the UK (*.co.uk*)
-
--   Geolocation: postcode references in the text
-
--   Timestamp: archival year
-
--   Counts
-
-------------------------------------------------------------------------
+-------------------------------------------
 
@@
 
+## Web data: The Internet Archive
+
+![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture1.png){width="70%"}
+
+## Web data: The Internet Archive
+
+![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture2.png){width="70%"}
+
 ## Our web data
@@
 
-<!-- <small> -->
+<small>
 
-<!-- -   Tranos, E., and C. Stich. 2020. Individual internet usage and the -->
+-   Tranos, E., and C. Stich. 2020. Individual internet usage and the
+    availability of online content of local interest: A multilevel
+    approach. *Computers, Environment and Urban Systems*, 79:101371.
 
-<!--     availability of online content of local interest: A multilevel -->
+-   Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy
+    in the UK: Regional productivity effects of early adoption.
+    *Regional Studies*, 55:12, 1924-1938.
 
-<!--     approach. *Computers, Environment and Urban Systems*, 79:101371. -->
+-   Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from
+    the ground up: a web data approach. *Environment and Planning B*, in
+    press.
 
-<!-- -   Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy -->
+-   Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to
+    predict regional trade flows: data extraction, modelling, and
+    validation, *Annals of the AAG*, in press.
 
-<!--     in the UK: Regional productivity effects of early adoption. -->
-
-<!--     *Regional Studies*, 55:12, 1924-1938. -->
-
-<!-- -   Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from -->
-
-<!--     the ground up: a web data approach. *Environment and Planning B*, in -->
-
-<!--     press. -->
-
-<!-- -   Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to -->
-
-<!--     predict regional trade flows: data extraction, modelling, and -->
-
-<!--     validation, *Annals of the AAG*, in press. -->
-
-<!-- </small> -->
+</small>
 
@@
 
-20080509162138 \| <http://www.website1.co.uk/contact_us> \| IG8 8HD
+20080509162138/<http://www.website1.co.uk/contact_us> IG8 8HD
 
@@
 # load 2000 co.uk
-# path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"
-# path.2000 <- "/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/all2000couk.csv"
-path.2000 <- paste0(path, "/data/temp/all2000couk.csv")
+# setwd("C:/Users/TranosE/DataShare/archive/nuts")
+path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"
 
@@
 
-<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png) -->
-
-![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
+![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
 
@@
 
-<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png) -->
+![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
 
-![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
+# Web diffusion
 
-# Methods
+## Mapping website density
 
-## Reminder: diffusion mechanisms
+------------------------------------------------------------------------
 
--   **S-shaped** pattern in the cumulative level of adoption
+```{r heatmaps, results='asis', out.width = "60%"}
+path.images <- paste0(path, "/outputs/")
 
--   A **hierarchy effect**: from main centres to secondary ones
+list.files(path = path.images, pattern = "^hex1", full.names = T) %>% 
+  map(image_read) %>% # reads each path file
+  image_join() %>% # joins image
+  image_animate(delay=50) #%>% # animates, can opt for number of loops
+  #image_write("test.gif") # write to current dir
+```
 
--   A **neighborhood effect**: first "hitting" nearby locations
+<!-- ------------------------------------------------------------------------ -->
 
-## Methods
+```{r eval = F, heatmaps_log, results='asis', out.width = "60%"}
+path.images <- paste0(path, "/outputs/")
 
--   Cumulative adoption: Self-starting logistic growth model [`nls` and
-    `SSlogis`]
+list.files(path = path.images, pattern = "^hex_log1", full.names = T) %>% 
+  map(image_read) %>% # reads each path file
+  image_join() %>% # joins image
+  image_animate(delay=50) #%>% # animates, can opt for number of loops
+  #image_write("test.gif") # write to current dir
+```
 
--   Descriptive statistics & ESDA
+---
 
--   Machine learning framework [random forests]
+> **Neighbourhood effect**: diffusion proceeds outwards from innovation
+> centers, first "hitting" nearby rather than far-away locations
+> [@grubler1990rise]
 
--   Two scales:
+-   Spatial dependency (Moran's I & LISA maps)
 
-    -   websites per firm in a Local Authority
-    -   websites in an Output Area
+-   Website density regressions -- distance effect
+
+-   Websites per firm in Local authorities (c. 400)
+
+-   Websites in Output Areas (c. 200,000)
+
+------------------------------------------------------------------------
+
+```{r moransi_la, results='asis'}
+path.in <- paste0(path, "/outputs/lisa/morani_la.csv")
+
+read.csv(path.in) %>% dplyr::select(year, morani) %>% 
+  ggplot(aes(x=year, y=morani)) +
+  geom_bar(stat = "identity") + 
+  ylab('Moran\'s I') +
+  theme_minimal() +
+  ggtitle("Moran\'s I, websites per firm a Local Authority") +
+  theme(plot.title = element_text(hjust = 0.5)) 
+```
+
+------------------------------------------------------------------------
+
+```{r, eval=T, moransi_oa, results='asis'}
+path.in <- paste0(path, "/outputs/lisa/morani_oa.csv")
+# read.csv(path.in) %>% dplyr::select(year, morani) %>% 
+#   rename('Moran\'s I' = morani,
+#          Year = year) %>% 
+#   arrange(Year) %>% 
+#   kable(digits = 3, 
+#         label = "Moran\'s I per year, Output Areas") %>% 
+#     kable_classic(full_width = F, html_font = "Cambria")
+
+read.csv(path.in) %>% dplyr::select(year, morani) %>% 
+  ggplot(aes(x=year, y=morani)) +
+  geom_bar(stat = "identity") + 
+  ylab('Moran\'s I') +
+  theme_minimal() +
+  ggtitle("Moran\'s I, websites per Output Area") +
+  theme(plot.title = element_text(hjust = 0.5)) 
+```
+
+------------------------------------------------------------------------
+
+```{r lisa_la, results='asis', out.width = "30%"}
+path.images <- paste0(path, "/outputs/lisa/")
+
+list.files(path = path.images, pattern = "^lisa_level_pc1_la", full.names = T) %>% 
+  map(image_read) %>% # reads each path file
+  image_join() %>% # joins image
+  image_animate(delay=50) #%>% # animates, can opt for number of loops
+  #image_write("test.gif") # write to current dir
+```
+
+------------------------------------------------------------------------
+
+```{r lisa_oa, eval=T,results='asis', out.width = "30%"}
+path.images <- paste0(path, "/outputs/lisa/")
+
+list.files(path = path.images, pattern = "^lisa_level_pc1_oa", full.names = T) %>% 
+  map(image_read) %>% # reads each path file
+  image_join() %>% # joins image
+  image_animate(delay=50) #%>% # animates, can opt for number of loops
+  #image_write("test.gif") # write to current dir
+```
+
+## Website density regressions
+
+<br>
+
+$$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$
+
+<br>
+
+::: {style="margin-left: 50px !important; text-align:left !important"}
+$Website\,Density_{i}$:
+
+-   Websites per firm in a Local Authority $i$, *or*
+
+-   Websites in an Output Area $i$
+
+:::
+
+## Website density regressions
+
+<br>
+
+$$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$
+
+<br>
+
+::: {style="margin-left: 50px !important; text-align:left !important"}
+$Place$:
+
+-   London, *or*
+
+-   Nearest city, *or*
+
+-   Nearest retail centre
+
+:::
+
+## Website density regressions
+
+<br>
+
+$\beta$ interpretation:
+
+-   The lower the $\beta$ is (or the larger the $|\beta|$ is)...
+
+-   ... the larger urban gravitation is for web adoption.
+
+------------------------------------------------------------------------
+
+```{r b_London_la, fig.height=5, fig.width=5, message=FALSE}
+path.image <- paste0(path, "/outputs/lisa/b_dist_london_la.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r b_city_la, fig.height=5, fig.width=5, message=FALSE}
+#, out.width = "60%"
+path.image <- paste0(path, "/outputs/lisa/b_dist_city_la.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r b_centre_la,  fig.height=5, fig.width=5,message=FALSE}
+#, out.height= "60%", out.width="60%"
+path.image <- paste0(path, "/outputs/lisa/b_dist_centre_la.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r b_London_oa, eval=T, fig.height=5, fig.width=5, message=FALSE}
+#, out.width = "60%"
+path.image <- paste0(path, "/outputs/lisa/b_dist_london_oa.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r b_centre_oa, eval = T,  fig.height=5, fig.width=5,message=FALSE}
+#fig.height=4, fig.width=4, 
+#, out.height= "60%", out.width="60%"
+path.image <- paste0(path, "/outputs/lisa/b_dist_city_oa.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r b_city_oa, eval = T, fig.height=5, fig.width=5, message=FALSE}
+#, out.width = "60%"
+path.image <- paste0(path, "/outputs/lisa/b_dist_centre_oa.png")
+knitr::include_graphics(path.image)
+```
+
+------------------------------------------------------------------------
+
+```{r r_squared, eval=T, results='asis'}
+# This is for the RSquared of the different website density regressions.
+
+path.in1 <- paste0(path, "/outputs/lisa/r2_all_oa.csv")
+path.in2 <- paste0(path, "/outputs/lisa/r2_all_la.csv")
+
+# read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
+#   ggplot(aes(x=year)) +
+#   geom_line(aes(y=London, color = "OA to London"), size=1) +
+#   geom_line(aes(y=cities, color = "OA to nearest city"), size=1) +
+#   geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) +
+# 
+#   geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) +
+#   geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) +
+#   geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) +
+# 
+#   scale_color_discrete(name = "R-squared for regression:") +
+#   #scale_linetype_manual(name = "Website density regressions111") 
+#   ylab("R-squared") + 
+#   ggtitle("Website density regressions") +
+#   theme_minimal() +
+#   theme(axis.title.x=element_blank(),
+#         plot.title = element_text(hjust = 0.5))
+
+read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
+  pivot_longer(!year, names_to = "regression", values_to = "Rsquared") %>% 
+  mutate(LA = if_else(str_ends(regression, ".LA"), " to LA", "to OA"),
+         regression = str_remove_all(regression, ".LA"),
+         line = paste0(regression, " ", LA)) %>% 
+  ggplot() +
+  geom_line(aes(x=year, y = Rsquared, color = line),
+            size = 1) +
+  scale_color_discrete(name = "R-squared for regression:") +
+  #scale_linetype_manual(name = "Website density regressions111") 
+  ylab("R-squared") + 
+  ggtitle("Website density regressions") +
+  theme_minimal() +
+  theme(axis.title.x=element_blank(),
+        plot.title = element_text(hjust = 0.5))
+
+#   scale_color_manual(values = c("red", "blue", "green"))
+
+```
+
+## Neighbourhood effect
+
+- Spatial dependency
+	- Relatively small and constant over time / scales
+	- London hot spot early on
+	- At local scale, consistent hotspots over time
+	- Granular analysis reveals other hotspots
+
+- Distance effect
+	- Urban gravitation increases over time and then drops
+	- Granular analysis: gravitation
+	- Lost explanatory power over time
+
+------------------------------------------------------------------------
+
+<br>
+
+> **Hierarchy effect**: from main centers to secondary ones -- central
+places
+
+-   Gini coefficient
+
+------------------------------------------------------------------------
+
+```{r gini_la, results='asis'}
+path.in <- paste0(path, "/outputs/lisa/gini_la.csv")
+
+read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
+  ggplot(aes(x=year, y=gini.a.n.)) +
+  geom_bar(stat = "identity") + 
+  ylab("Gini") +
+  theme_minimal() +
+  ggtitle("Gini coefficient, websites per firm in Local Auhtority") +
+  theme(plot.title = element_text(hjust = 0.5)) 
+```
+
+------------------------------------------------------------------------
+
+```{r gini_oa, eval=T, results='asis'}
+path.in <- paste0(path, "/outputs/lisa/gini_oa.csv")
+
+read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
+  ggplot(aes(x=year, y=gini.a.n.)) +
+  geom_bar(stat = "identity") + 
+  ylab("Gini") +
+  theme_minimal() +
+  ggtitle("Gini coefficient, websites per Output Area") +
+  theme(plot.title = element_text(hjust = 0.5)) 
+```
+
+## Hieararchy
+
+-   Almost perfect polarisation of web adoption in the early stages at a
+    granular level
+
+-   Polarisation decreases over time
+
+-   More equally diffused at the Local Authority level
 
@@
 
-```{r UK S, eval=T, out.width = "60%", out.height="60%"}
+```{r UK S, eval=T}
 #, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"
 
-path.image <- paste0(path, "/outputs/s_uk_per_firm.png")
+path.image <- paste0(path, "/outputs/s_uk.png")
 knitr::include_graphics(path.image)
@@
 ```{r LA_S_gif, eval=F, results='asis', out.width = "60%"}
-path.images <- paste0(path, "/outputs/s_lad_per_firm/")
+path.images <- paste0(path, "/outputs/")
 
 library(magick)
-las <- list.files(path = path.images, pattern = "^lad_", full.names = T) 
-las <- sample(las, 50)
+las <- list.files(path = path.images, pattern = "^lad_", full.names = T) #%>% 
 las_list <- lapply(las, image_read)
@@
 ## animate at 2 frames per second
-las_animated <- image_animate(las_joined, fps = 4)
+las_animated <- image_animate(las_joined, fps = 2)
 
 ## view animated image
-#las_animated
+las_animated
 
 ## save to disk
-# image_write(image = las_animated,
-#             path = "las.gif")
-image_write_gif(las_animated, path = "las.gif")
+image_write(image = las_animated,
+            path = "las.gif")
 
@@
   #image_write("test.gif") # write to current dir
-```
-
-------------------------------------------------------------------------
-
-```{r gif S, eval=T, out.width = "60%", out.height="60%"}
-#, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"
-
-path.image <- paste0(path, "/presentations/las.gif")
-knitr::include_graphics(path.image)
 ```
@@
 
-```{r Liverpool, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
+```{r Liverpool, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
 path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E08000012.png")
@@
 - Spatial heterogeneity
-
--   Not a clear, easy to explain pattern
+- Not clear pattern
 
@@
 - Adoption heterogeneity 
-
--   Different perceptions of risk and economic returns from new
-    technologies
-
+- Perceptions of risk and economic returns from new technologies
 - Early adopters vs. laggards, leapfrogging
@@
 
-## Rank dynamics and diffusion speed
+# Conclusions
 
--   Spatial heterogeneity
-
--   Expected volatility
-
-# Spatial mechanisms
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r, eval=FALSE, heatmaps, results='asis', out.width = "60%"}
-path.images <- paste0(path, "/outputs/")
-
-list.files(path = path.images, pattern = "^hex1", full.names = T) %>% 
-  map(image_read) %>% # reads each path file
-  image_join() %>% # joins image
-  image_animate(delay=50) #%>% # animates, can opt for number of loops
-  #image_write("test.gif") # write to current dir
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r eval = F, heatmaps_log, results='asis', out.width = "60%"}
-path.images <- paste0(path, "/outputs/")
-
-list.files(path = path.images, pattern = "^hex_log1", full.names = T) %>% 
-  map(image_read) %>% # reads each path file
-  image_join() %>% # joins image
-  image_animate(delay=50) #%>% # animates, can opt for number of loops
-  #image_write("test.gif") # write to current dir
-```
-
-------------------------------------------------------------------------
-
-> **Neighbourhood effect**: diffusion proceeds outwards from innovation
-> centers, first "hitting" nearby rather than far-away locations
-> [@grubler1990rise]
-
--   Spatial dependency (Moran's I & LISA maps)
-
-<!-- -   Website density regressions -- distance effect -->
-
--   Websites per firm in Local authorities (c. 400)
-
--   Websites in Output Areas (c. 230,000)
-
-------------------------------------------------------------------------
-
-```{r moransi_la, results='asis'}
-path.in <- paste0(path, "/outputs/lisa/la_ggplot/morani_la.csv")
-
-read.csv(path.in) %>% dplyr::select(year, morani) %>% 
-  ggplot(aes(x=year, y=morani)) +
-  geom_bar(stat = "identity") + 
-  ylab('Moran\'s I') +
-  theme_minimal() +
-  ggtitle("Moran\'s I, websites per firm a Local Authority") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-```
-
-------------------------------------------------------------------------
-
-```{r, eval=T, moransi_oa, results='asis'}
-path.in <- paste0(path, "/outputs/lisa/oa_ggplot/morani_oa.csv")
-# read.csv(path.in) %>% dplyr::select(year, morani) %>% 
-#   rename('Moran\'s I' = morani,
-#          Year = year) %>% 
-#   arrange(Year) %>% 
-#   kable(digits = 3, 
-#         label = "Moran\'s I per year, Output Areas") %>% 
-#     kable_classic(full_width = F, html_font = "Cambria")
-
-read.csv(path.in) %>% dplyr::select(year, morani) %>% 
-  ggplot(aes(x=year, y=morani)) +
-  geom_bar(stat = "identity") + 
-  ylab('Moran\'s I') +
-  theme_minimal() +
-  ggtitle("Moran\'s I, websites per Output Area") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-```
-
-------------------------------------------------------------------------
-
-```{r lisa_la, results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/la_ggplot/lisa/")
-
-list.files(path = path.images, pattern = "^lisa_level_pc1_la", full.names = T) %>% 
-  map(image_read) %>% # reads each path file
-  image_join() %>% # joins image
-  image_animate(delay=50) #%>% # animates, can opt for number of loops
-  #image_write("test.gif") # write to current dir
-```
-
-------------------------------------------------------------------------
-
-```{r lisa_oa, eval=T,results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/oa_ggplot/lisa/")
-
-list.files(path = path.images, pattern = "^lisa_level_pc1_oa", full.names = T) %>% 
-  map(image_read) %>% # reads each path file
-  image_join() %>% # joins image
-  image_animate(delay=50) #%>% # animates, can opt for number of loops
-  #image_write("test.gif") # write to current dir
-```
-
-<!-- ## Website density regressions -->
-
-<!-- <br> -->
-
-<!-- $$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$ -->
-
-<!-- <br> -->
-
-<!-- ::: {style="margin-left: 50px !important; text-align:left !important"} -->
-
-<!-- $Website\,Density_{i}$: -->
-
-<!-- -   Websites per firm in a Local Authority $i$, *or* -->
-
-<!-- -   Websites in an Output Area $i$ -->
-
-<!-- ::: -->
-
-<!-- ## Website density regressions -->
-
-<!-- <br> -->
-
-<!-- $$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$ -->
-
-<!-- <br> -->
-
-<!-- ::: {style="margin-left: 50px !important; text-align:left !important"} -->
-
-<!-- $Place$: -->
-
-<!-- -   London, *or* -->
-
-<!-- -   Nearest city, *or* -->
-
-<!-- -   Nearest retail centre -->
-
-<!-- ::: -->
-
-<!-- ## Website density regressions -->
-
-<!-- <br> -->
-
-<!-- $\beta$ interpretation: -->
-
-<!-- -   The lower the $\beta$ is (or the larger the $|\beta|$ is)... -->
-
-<!-- -   ... the larger urban gravitation is for web adoption. -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_London_la, fig.height=5, fig.width=5, message=FALSE} -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_london_la.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_city_la, fig.height=5, fig.width=5, message=FALSE} -->
-
-<!-- #, out.width = "60%" -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_city_la.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_centre_la,  fig.height=5, fig.width=5,message=FALSE} -->
-
-<!-- #, out.height= "60%", out.width="60%" -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_centre_la.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_London_oa, eval=T, fig.height=5, fig.width=5, message=FALSE} -->
-
-<!-- #, out.width = "60%" -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_london_oa.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_centre_oa, eval = T,  fig.height=5, fig.width=5,message=FALSE} -->
-
-<!-- #fig.height=4, fig.width=4,  -->
-
-<!-- #, out.height= "60%", out.width="60%" -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_city_oa.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_city_oa, eval = T, fig.height=5, fig.width=5, message=FALSE} -->
-
-<!-- #, out.width = "60%" -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_centre_oa.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r r_squared, eval=T, results='asis'} -->
-
-<!-- # This is for the RSquared of the different website density regressions. -->
-
-<!-- path.in1 <- paste0(path, "/outputs/lisa/r2_all_oa.csv") -->
-
-<!-- path.in2 <- paste0(path, "/outputs/lisa/r2_all_la.csv") -->
-
-<!-- # read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>%  -->
-
-<!-- #   ggplot(aes(x=year)) + -->
-
-<!-- #   geom_line(aes(y=London, color = "OA to London"), size=1) + -->
-
-<!-- #   geom_line(aes(y=cities, color = "OA to nearest city"), size=1) + -->
-
-<!-- #   geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) + -->
-
-<!-- #  -->
-
-<!-- #   geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) + -->
-
-<!-- #   geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) + -->
-
-<!-- #   geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) + -->
-
-<!-- #  -->
-
-<!-- #   scale_color_discrete(name = "R-squared for regression:") + -->
-
-<!-- #   #scale_linetype_manual(name = "Website density regressions111")  -->
-
-<!-- #   ylab("R-squared") +  -->
-
-<!-- #   ggtitle("Website density regressions") + -->
-
-<!-- #   theme_minimal() + -->
-
-<!-- #   theme(axis.title.x=element_blank(), -->
-
-<!-- #         plot.title = element_text(hjust = 0.5)) -->
-
-<!-- read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>%  -->
-
-<!--   pivot_longer(!year, names_to = "regression", values_to = "Rsquared") %>%  -->
-
-<!--   mutate(LA = if_else(str_ends(regression, ".LA"), " to LA", "to OA"), -->
-
-<!--          regression = str_remove_all(regression, ".LA"), -->
-
-<!--          line = paste0(regression, " ", LA)) %>%  -->
-
-<!--   ggplot() + -->
-
-<!--   geom_line(aes(x=year, y = Rsquared, color = line), -->
-
-<!--             size = 1) + -->
-
-<!--   scale_color_discrete(name = "R-squared for regression:") + -->
-
-<!--   #scale_linetype_manual(name = "Website density regressions111")  -->
-
-<!--   ylab("R-squared") +  -->
-
-<!--   ggtitle("Website density regressions") + -->
-
-<!--   theme_minimal() + -->
-
-<!--   theme(axis.title.x=element_blank(), -->
-
-<!--         plot.title = element_text(hjust = 0.5)) -->
-
-<!-- #   scale_color_manual(values = c("red", "blue", "green")) -->
-
-<!-- ``` -->
-
-## Neighbourhood effect
-
--   Spatial dependency
-    -   Relatively small, consistent over time / scales
-    -   London hot spot early on
-    -   At local scale, consistent hotspots over time
-    -   Granular analysis reveals other hotspots
-        <!-- -   Distance effect -->
-        <!--     -   Urban gravitation increases over time and then drops -->
-        <!--     -   Granular analysis: gravitation -->
-        <!--     -   Lost explanatory power over time -->
-
-------------------------------------------------------------------------
+----
 
@@
 
-> **Hierarchy effect**: from main centers to secondary ones -- central
-> places
+-   **Geography matters:** spatial dependency, urban gravitation
 
--   Gini coefficient
-
-------------------------------------------------------------------------
-
-```{r gini_la, results='asis'}
-path.in <- paste0(path, "/outputs/lisa/gini_la.csv")
-
-read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
-  ggplot(aes(x=year, y=gini.a.n.)) +
-  geom_bar(stat = "identity") + 
-  ylab("Gini") +
-  theme_minimal() +
-  ggtitle("Gini coefficient, websites per firm in Local Auhtority") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-```
-
-------------------------------------------------------------------------
-
-```{r gini_oa, eval=T, results='asis'}
-path.in <- paste0(path, "/outputs/lisa/gini_oa.csv")
-
-read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
-  ggplot(aes(x=year, y=gini.a.n.)) +
-  geom_bar(stat = "identity") + 
-  ylab("Gini") +
-  theme_minimal() +
-  ggtitle("Gini coefficient, websites per Output Area") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-```
-
-## Hieararchy
-
--   Almost perfect polarisation of web adoption in the early stages at a
-    granular level
-
--   Polarisation decreases over time
-
--   More equally diffused at the Local Authority level
-
-# Putting all of these together
-
-## Modelling framework
-
--   [A hierarchy effect: from main centres to secondary
-    ones]{style="color:orange;"}
-
--   [A neighborhood effect: first "hitting" nearby
-    locations]{style="color:olive;"}
-
--   [S-shaped pattern in the cumulative level of
-    adoption]{style="color:violet;"}
-
-<br> <small>
-$$Website\,Density_{t} \sim \color{orange}{Distance\,London} + 
-\color{orange}{Website\,density\,London_{t-1}} + \\
-\color{orange}{Distance\,Nearest\,City} + 
-\color{orange}{Website\,density\,Nearest\,City_{t-1}} + \\
-\color{orange}{Distance\,Nearest\,Retail_{i}} +
-\color{orange}{Website\,density\,Nearest\,Retail_{t-1}} +\\
-\color{olive}{W*\, Website\,density_{t-1}} +\\ 
-\color{violet}{year_{t}}$$ </small>
-
-## Modelling framework
-
--   Random forests to predict $Website\,Density_{i,t}$
-
--   2 spatial resolutions:
-
-    -   Local Authorities (websites per firm)
-    -   Output Areas (websites)
-
--   2 sets of models:
-
-    -   All the data ⟹ *variable importance*
-    -   Train on all but one region, test on the holdout region ⟹
-        *spatial differences and similarities* of diffusion mechanisms
-
--   Space-time sensitive 10-fold CV (`CAST`)
-
-## Models trained on all data
-
-<br>
-
-|                     | RMSE  | RSquared | MAE   |
-|---------------------|-------|----------|-------|
-| Local Authorities\* | 0.032 | 0.810    | 0.019 |
-| Output Areas        | 5.000 | 0.205    | 1.047 |
-
-<small> \*No retail centre predictors for the Local Authority models
-</small>
-
-## Variable importance for Local Authorities
-
-```{r varimpLA, eval = T, out.height= "58%", out.width="58%", message=FALSE}
-
-path.image <- paste0(path, "/outputs/rf/figures/varimp_LA.png")
-knitr::include_graphics(path.image)
-```
-
-## Variable importance for Output Areas
-
-```{r varimpOA, eval = T, out.height= "58%", out.width="58%", message=FALSE}
-#,  fig.height=5, fig.width=5,message=FALSE
-
-path.image <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
-knitr::include_graphics(path.image)
-```
-
-## Regional similarities
-
-```{r similarity, eval = T, message=FALSE}
-path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
-path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
-
-lad <- read.csv(path.lad) %>% 
-  rename('RSquared LAD' = Rsquared)
-
-oa <- read.csv(path.oa) %>% rename(Region = test.region,
-                                   'RSquared OA' = Rsquared)
-
-lad %>% left_join(oa) %>% 
-  mutate('Rank OA' = rank(desc(`RSquared OA`)),
-         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
-  relocate(5, .after = 2) %>% 
-  arrange(`Rank LAD`) %>% 
-  kable(digits=3)
-```
-
-## Alternative specification: growth, trained on all data
-
-<br> <small>
-$$Website\,Density\,Growth_{t} \sim \color{orange}{Distance\,London} + 
-\color{orange}{Website\,density\,London_{t-1}} + \\
-\color{orange}{Distance\,Nearest\,City} + 
-\color{orange}{Website\,density\,Nearest\,City_{t-1}} + \\
-\color{orange}{Distance\,Nearest\,Retail_{i}} +
-\color{orange}{Website\,density\,Nearest\,Retail_{t-1}} +\\
-\color{olive}{W*\, Website\,density_{t-1}} +\\ 
-\color{violet}{year_{t}}$$ </small>
-
-<br>
-
-|                     | RMSE  | RSquared | MAE   |
-|---------------------|-------|----------|-------|
-| Local Authorities\* | 0.200 | 0.634    | 0.148 |
-
-<small> \*No retail centre predictors for the Local Authority models
-</small>
-
-## Alternative specification: growth, trained on all data
-
-```{r growth LA}
-path.table <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv")
-read.csv(path.table) %>% 
-  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
-  arrange(desc(Rsquared)) %>% 
-  kable()
-```
-
-# Conclusions
-
-------------------------------------------------------------------------
-
--   Established technological diffusion spatial processes still apply
-
-    -   for a digital technology
-    -   at local scales
-
--   Geography matters: spatial dependency, urban gravitation
-
--   Hierarchical diffusion
+-   Some indications of a hierarchical diffusion
 
@@
 
--   Stability and volatility: leapfrogging, early adopters dropping, but
-    also stable positions
+-   Well-established theoretical approaches of diffusion survive even at
+    a granular level
 
--   Spatially consistent mechanisms at local scale
-
--   Regional heterogeneity
+-   Stability and volatility: leapfrogging, early adopters dropping, but also maintaining
 

### 2025-01-21 src/la_RF_corrected.Rmd
early/late

@@
 library(rprojroot)
-library(rgdal)
-library(rgeos)
+# library(rgdal)
+# library(rgeos)
 library(sf)
@@
 indices <- CreateSpacetimeFolds(df,# %>% filter(year > 1999),                   # FILTER BY YEAR FOR GROWTH
-                                spacevar = "ladcd",
+                                spacevar = "RGN21CD",                           # was ladcd in 1st submission
                                 timevar = "year", 
@@
   indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i), #& year > 1999),   # FILTER BY YEAR FOR GROWTH 
-                                spacevar = "ladcd",
+                                spacevar = "RGN21CD",
                                 timevar = "year", 
@@
 indices <- CreateSpacetimeFolds(df %>% filter(year > 1999),                     # FILTER BY YEAR FOR GROWTH
-                                spacevar = "ladcd",
+                                spacevar = "RGN21CD",
                                 timevar = "year", 
@@
   indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i & year > 1999),      # FILTER BY YEAR FOR GROWTH 
-                                spacevar = "ladcd",
+                                spacevar = "RGN21CD",
                                 timevar = "year", 

### 2025-01-21 src/oa_RF_corrected.Rmd
early/late

@@
 indices <- CreateSpacetimeFolds(df, 
-                                spacevar = "oa11cd",
+                                spacevar = "RGN11CD",
                                 timevar = "year", 
@@
   indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD!=i), 
-                                spacevar = "oa11cd",
+                                spacevar = "RGN21CD",
                                 timevar = "year", 

### 2025-01-21 src/oa_RF_corrected_10.Rmd
early/late

@@
 indices <- CreateSpacetimeFolds(df, 
-                                spacevar = "oa11cd",
+                                spacevar = "RGN21CD",
                                 timevar = "year", 

### 2025-01-21 src/ranks_oa_corrected.Rmd
early/late

---
title: "Ranks, OA corrected"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
# library(sf)
# library(spdep)
# library(knitr)
# library(REAT)
# library(tidygeocoder)
# library(geosphere)
# library(broom)
# library(doParallel)
# library(raster)
library(plm)
#library(lmtest)
# library(caret)
# library(randomForest)
# library(CAST)
library(ggbump)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

## Load data

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

## Problem postcodes

```{r}
n = 1 #number of unique postcodes. 
m = 11

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610) %>% 
  filter(
      #V1.domain < m,             # for multiple pc
      V1.domain == n,             # for n == 1
      year > 1995)

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012) %>% 
  filter(
    #V1.domain < m,             # for multiple pc
    V1.domain == n)             # for n == 1

df.long <- bind_rows(df9610, df1112) %>% 
  filter(year > 1995 & year < 2013) %>% 
  group_by(pc, year) %>% 
  summarise(n = n()) %>%
  ungroup() %>% 
  #arrange(desc(n)) %>% 
  pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
  dplyr::select(pc, n_1996, n_1999, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010, n_2011, n_2012) %>%
  arrange(desc(n_2004), desc(n_2005)) %>% 
  slice_head(n=1000) %>% 
  pivot_longer(!pc, names_to = "year", values_to = "n") %>% 
  mutate(year = as.integer(sub("n_", "", year, fixed = TRUE))) %>% 
  replace(is.na(.), 0) %>% 
  mutate(outlier = ifelse(n > 1000 & (year==2004 | year == 2005), pc, ""))

problem.pcs <- df.long %>% filter(year==2004 | year == 2005, 
                   n > 1000) %>% 
  distinct(pc)

# n > 1500: "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"   
# n > 1000: SE24 9HP, CV8 2ED, GL16 7YA, CW1 6GL, M28 2SL, DE21 7BF
```

## Corrected

The above exploratory analysis indicated that there is a huge increase in 2004-2006
in the above postcodes. I turn them to NAs and then use a regression to impute these
gaps.

```{r}
df.corrected <- bind_rows(df9610, df1112) %>% 
  filter(year > 1995 & year < 2013) %>% 
  group_by(pc, year) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2001 & year <2007, NA, n))

df.corrected <- pdata.frame(df.corrected)
predictions <- round(predict(plm(n ~ as.factor(year), effect = "individual", 
                                 model = "within", data = df.corrected), newdata = df.corrected), 2)

df.corrected$n[is.na(df.corrected$n)] <- predictions[is.na(df.corrected$n)]
```

## OA df

```{r}
df <- df.corrected %>% 
  mutate(year = as.integer(as.character(year))) %>% 
  filter(year > 1995 & year < 2013) %>% 
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, oa11cd) %>%
  summarise(n = sum(n)) %>% # **********ATTENTION*********  
  ungroup() %>% 
  complete(oa11cd, year) %>% 
  filter(!is.na(oa11cd)) %>% # drop c. 10 - 200 per year, which are not assigned to a LAD
  replace(is.na(.), 0) 
```

## Average ranks websites per firm t and 2012 only: 05 and 95

Very awkward figure as the ranks are compresses because of the low ranges. For
instance, there are so many 0s, so all these OAs are together in the bottom.

```{r}

t <- 2000

out.path <- paste0(path, "/outputs/ranks/oa", t,"_2012_only0595_av.png")

df %>%

  # left_join(oa.uk %>% as_tibble() %>% dplyr::select(id),
  #           by = c("ladcd" = "id")) %>%

  mutate(period = ifelse(year < 1999, 1, 
                         ifelse(year > 2009, 3, 2))) %>%
  group_by(oa11cd, period) %>%
  mutate(n = ifelse(period != 2, mean(n), n)) %>% 
  filter(year > 1997 & year < 2011) %>% 
  group_by(year) %>% 
  #mutate(rank = rank(desc(n))) %>% # this should be the same as min_rank()
  mutate(rank = min_rank(desc(n))) %>% 
  ungroup() %>% 
  filter(year==1998 | year == 2010) %>% 
  arrange(oa11cd, year) %>% 
  group_by(oa11cd) %>% 
  #mutate(diff_1st = -diff(rank))
  mutate(diff_1st = -(rank-dplyr::lag(rank))) %>% #filter(abs(diff_1st) > 300) #arrange(diff_1st)

  #ggplot(aes(abs(diff_1st))) + geom_histogram()

  #ungroup() %>% summarise(q = quantile(abs(diff_1st), .95, na.rm = T))
  # q.95 = 136884, q.9 = 136884, q.1 = 10469, q.05 = 3915. **abs()** # for dense_rank()
  # q.95 = 302, q.9 = 302, q.1 = 287, q.05 = 277. **abs()** # for min_rank()
  # ungroup() %>% summarise(max = max(abs(diff_1st), na.rm = T),
  #                         min = min(abs(diff_1st), na.rm = T))

  mutate(alpha = ifelse(abs(diff_1st) <= 277 | abs(diff_1st) > 302, 1, 0.5)) %>% 
  mutate(alpha = ifelse(is.na(diff_1st), 0.5, alpha)) %>%

  mutate(q = ifelse(abs(diff_1st) <= 277, "05",
                    ifelse(abs(diff_1st) > 302, "95", "m"))) %>% 
  mutate(q = ifelse(is.na(diff_1st), "m", q)) %>%

  ggplot(aes(year, rank, group = oa11cd, alpha=alpha)) +
  geom_bump(aes(colour = q)) + #colour = "red"
  scale_color_manual(values = c("darkorchid3", "orange", "grey70"),
                     breaks = c("05", "95"),
                     labels = c("Stability","Volatility")) + 
  labs(color = "oa11cd") +
  # scale_y_reverse(labels = c(30000, 60000, 90000, 120000, 161000)) +
  scale_y_reverse() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  #theme(legend.position="none") +
  scale_x_continuous(NULL, 
                     #limits = c(1998, 2010), 
                     breaks=c(1998, 2010),
                     labels = c("1996-98","2010-12")) +
  guides(alpha = "none")

# ggsave(out.path)
```

Plot n

```{r}

t <- 2000

out.path <- paste0(path, "/outputs/ranks/oa_n_", t,"_2012_only0595_av.png")

df %>%

  mutate(period = ifelse(year < 1999, 1, 
                         ifelse(year > 2009, 3, 2))) %>%
  group_by(oa11cd, period) %>%
  mutate(n = ifelse(period != 2, mean(n), n)) %>%

  filter(year==1998 | year == 2010) %>% 
  group_by(year) %>% 
  mutate(n.st = n / max(n)) %>%

  group_by(oa11cd) %>% 
  mutate(diff_1st = -(n.st-dplyr::lag(n.st))) %>%

  # ggplot(aes(abs(diff_1st))) + geom_histogram()

  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .05, na.rm = T))
  # q.95 = 0.01221996, q.9 = 0.006924644, q.1 = 0, q.05 = 0.
  # ungroup() %>% summarise(max = max(abs(diff_1st), na.rm = T),
  #                         min = min(abs(diff_1st), na.rm = T))

  mutate(alpha = ifelse(abs(diff_1st) <= 0 | abs(diff_1st) > 0.01221996, 1, 0.5)) %>% 
  mutate(alpha = ifelse(is.na(diff_1st), 0.5, alpha)) %>%

  mutate(q = ifelse(abs(diff_1st) <= 0, "05",
                    ifelse(abs(diff_1st) > 0.01221996, "95", "m"))) %>% 
  mutate(q = ifelse(is.na(diff_1st), "m", q)) %>%

  ggplot(aes(year, n.st, group = oa11cd, alpha=alpha)) +
  geom_bump(aes(colour = q)) + #colour = "red"
  scale_color_manual(values = c("darkorchid3", "orange", "grey70"),
                     breaks = c("05", "95"),
                     labels = c("Stability","Volatility")) + 
  labs(color = "oa11cd") +
  #scale_y_reverse() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  #theme(legend.position="none") +
  scale_x_continuous(NULL, 
                     #limits = c(1998, 2010), 
                     breaks=c(1998, 2010),
                     labels = c("1996-98","2010-12")) +
  guides(alpha = "none")

ggsave(out.path)
```

## For specific LA

```{r}

t <- 2000
LAD <- "E09000002"

#out.path <- paste0(path, "/outputs/ranks/oa", t,"_2012_only0595_av.png")

# Lookup: OA to regions
path.lookup <- paste0(path, "/data/raw/Output_Area_to_Lower_layer_Super_Output_Area_to_Middle_layer_Super_Output_Area_to_Local_Authority_District_(December_2011)_Lookup_in_England_and_Wales.csv")  
lookup.lad <- read_csv(path.lookup) %>% dplyr::select(OA11CD, LAD11CD, LAD11NM)

df %>% left_join(lookup.lad, by = c("oa11cd" = "OA11CD")) %>% 
  relocate(c(LAD11NM, LAD11CD), .after=year) %>% 
  filter(LAD11CD == LAD) %>% 
  mutate(period = ifelse(year < 1999, 1, 
                         ifelse(year > 2009, 3, 2))) %>%
  group_by(oa11cd, period) %>%
  mutate(n = ifelse(period != 2, mean(n), n)) %>% 
  filter(year > 1997 & year < 2011) %>% 
  group_by(year) %>% 
  #mutate(rank = rank(desc(n))) %>% # this should be the same as min_rank()
  mutate(rank = min_rank(desc(n))) %>% 
  ungroup() %>% 
  filter(year==1998 | year == 2010) %>% 
  arrange(oa11cd, year) %>% 
  group_by(oa11cd) %>% 
  #mutate(diff_1st = -diff(rank))
  mutate(diff_1st = -(rank-dplyr::lag(rank))) %>% #filter(abs(diff_1st) > 300) #arrange(diff_1st)

  #ggplot(aes(abs(diff_1st))) + geom_histogram()

  #ungroup() %>% summarise(q = quantile(abs(diff_1st), .05, na.rm = T))
  # q.95 = 9.2, q.05 = 0. **abs()** # for min_rank()
  # ungroup() %>% summarise(max = max(abs(diff_1st), na.rm = T),
  #                         min = min(abs(diff_1st), na.rm = T))

  mutate(alpha = ifelse(abs(diff_1st) <= 0 | abs(diff_1st) > 9.2, 1, 0.5)) %>% 
  mutate(alpha = ifelse(is.na(diff_1st), 0.5, alpha)) %>%

  mutate(q = ifelse(abs(diff_1st) <= 0, "05",
                    ifelse(abs(diff_1st) > 9.2, "95", "m"))) %>% 
  mutate(q = ifelse(is.na(diff_1st), "m", q)) %>%

  ggplot(aes(year, rank, group = oa11cd, alpha=alpha)) +
  geom_bump(aes(colour = q)) + #colour = "red"
  scale_color_manual(values = c("darkorchid3", "orange", "grey70"),
                     breaks = c("05", "95"),
                     labels = c("Stability","Volatility")) + 
  labs(color = "oa11cd") +
  # scale_y_reverse(labels = c(30000, 60000, 90000, 120000, 161000)) +
  scale_y_reverse() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  #theme(legend.position="none") +
  scale_x_continuous(NULL, 
                     #limits = c(1998, 2010), 
                     breaks=c(1998, 2010),
                     labels = c("1996-98","2010-12")) +
  guides(alpha = "none")

# ggsave(out.path)
```

Plot n

```{r}

t <- 2000
LAD <- "E09000023"

out.path <- paste0(path, "/outputs/ranks/oa_n_", t,"_2012_only0595_av.png")

df %>% left_join(lookup.lad, by = c("oa11cd" = "OA11CD")) %>% 
  relocate(c(LAD11NM, LAD11CD), .after=year) %>% 
  filter(LAD11CD == LAD) %>% 
  mutate(period = ifelse(year < 1999, 1, 
                         ifelse(year > 2009, 3, 2))) %>%
  group_by(oa11cd, period) %>%
  mutate(n = ifelse(period != 2, mean(n), n)) %>%

  filter(year==1998 | year == 2010) %>% 
  group_by(year) %>% 
  mutate(n.st = n / max(n)) %>%

  group_by(oa11cd) %>% 
  mutate(diff_1st = -(n.st-dplyr::lag(n.st))) %>%

  # ggplot(aes(abs(diff_1st))) + geom_histogram()

  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .05, na.rm = T))
  # q.95 = 0.01221996, q.9 = 0.006924644, q.1 = 0, q.05 = 0.
  # ungroup() %>% summarise(max = max(abs(diff_1st), na.rm = T),
  #                         min = min(abs(diff_1st), na.rm = T))

  mutate(alpha = ifelse(abs(diff_1st) <= 0 | abs(diff_1st) > 0.01221996, 1, 0.5)) %>% 
  mutate(alpha = ifelse(is.na(diff_1st), 0.5, alpha)) %>%

  mutate(q = ifelse(abs(diff_1st) <= 0, "05",
                    ifelse(abs(diff_1st) > 0.01221996, "95", "m"))) %>% 
  mutate(q = ifelse(is.na(diff_1st), "m", q)) %>%

  ggplot(aes(year, n.st, group = oa11cd, alpha=alpha)) +
  geom_bump(aes(colour = q)) + #colour = "red"
  scale_color_manual(values = c("darkorchid3", "orange", "grey70"),
                     breaks = c("05", "95"),
                     labels = c("Stability","Volatility")) + 
  labs(color = "oa11cd") +
  #scale_y_reverse() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  #theme(legend.position="none") +
  scale_x_continuous(NULL, 
                     #limits = c(1998, 2010), 
                     breaks=c(1998, 2010),
                     labels = c("1996-98","2010-12")) +
  guides(alpha = "none")

#ggsave(out.path)
```

### 2025-01-21 paper/submission2/response.md

--- 2025-01-03 paper/submission2/response.qmd
@@
 title: "Response to editor and reviewer comments"
-format: pdf
 bibliography: bibliography.bib
+output: html_document
 ---
@@
 ## Editor; theory and justification 
-\color{teal}
 First, you will need to develop a more convincing theoretical argument and contribution.
@@
 The reviewers challenge you to do this using your data and suggest both the argument and structure of the paper will need to change.
-\color{black}
 
 ## Editor; data
-\color{teal}
 Second, you need to justify your data and analysis better.
@@
 As it stands, the paper feels rather overloaded with data, rather than acting as a focused analysis.
-\color{black}
 
 ## Editor; conclusions
-\color{teal}
 The problems identified above result in the paper feeling imprecise, confusing and lacking a contribution.
@@
 There is, then, a route to publication, but it will require significant changes to convince the reviewers.
-\color{black}
 
@@
 
-\color{teal}
-OVERVIEW\
+Comments to the Author
+OVERVIEW
 - focused on innovation diffusion, using the case of the web in the UK
@@
 
-Figure 2 shows the diffusion of domains per company at the scale of the UK matches the S growth model.
+> Figure 2 shows the diffusion of domains per company at the scale of the UK matches the S growth model.
 Figure 3 shows some spatial clustering.
 I think this is for all years in the dataset but this could be better labeled
-\color{black}
+*explain, change title*
 
-**explain, change title**
+> Figure 4, not sure what this demonstrates
 
-\color{teal}
-Figure 4, not sure what this demonstrates
-\color{black}
+> Figure 5&6, shows early clustering and then diffusion (I think)
+*yes*
 
-\color{teal}
-Figure 5&6, shows early clustering and then diffusion (I think)
-\color{black}
-
-**yes**
-
-\color{teal}
-DATA\
-Focusing just on .co.uk website is OK but needs to be better justified given that it only represented about 50 percent of domains in the UK at that time.
+DATA
+> Focusing just on .co.uk website is OK but needs to be better justified given that it only represented about 50 percent of domains in the UK at that time.
 See my back of the envelop calculations: As per Zook (2001, Table 1) .uk represented 66% of all domains in the UK, .co.uk is a subset of that, so this is perhaps 50% of all domains registered in the UK at that time.
@@
 (perhaps this is too low, aggregating from webpages to websites brought the number higher, not entirely clear to me)
-\color{black}
 
@@
 place reference [@hill2009georeferencing] and this reference has no been added 
-in the paper.
+in the paper. **ADD REF**
 
 So, based on the above 35% of all domains with a registered address in the UK
-were represented in the data. Most importantly though the paper does not model 
-the adoption of the Web from the general population but instead from firms. So, 
+were represented in the data. Most importantly though the paper does not model the 
+adoption of the Web from the general population but instead from firms. So, 
 although the data only represents 35% of the population of websites with a 
@@
 
-The third paragraph in the Data and Methods section includes the following text:
+Active/inactive, don;t suffer from multiple registrations
 
-> Instead of using the whole .uk ccTLD, this paper focuses on its
+Third paragraph in the Data and Methods section. 
+"Instead of using the whole .uk ccTLD, this paper focuses on its
 commercial subset, the .co.uk second level domain (SLD). This choice
@@
 support online transactions and share opinions [@THELWALL2000441; 
-@blazquez2018big].
+@blazquez2018big]".
 
-<!-- Active/inactive, don;t suffer from multiple registrations -->
+**FIX TEXT TO EMPHASISE THE COMMERCIAL - LINK TO R3**
 
-\color{teal}
-overall the time period 1996-2012 is a good time to study the diffusion of the commercial internet. But why stop at 2012? Is it when the Library stopped its data series?
-\color{black}
+*rebutal: this is not about general population web adoption, but firms, which more often than not use .co.uk*
 
-Yes, the British Library had obtained a copy of all of the .uk contents of the 
-Internet Archive for that period. From a data perspective, it wouldn't be correct
-to extend this dataset with data from other open archives such as the Common 
-Crawl as they use different crawlers and there will be inconsistencies. Most
-importantly though, this period covers the diffusion of the web from the very
-early stages until its maturity. This is also illustrated by the figure 
-depicting the cumulative adoption and the flattening right end of the line.
+> - overall the time period 1996-2012 is a good time to study the diffusion of the commercial internet. But why stop at 2012? Is it when the Library stopped its data series?
+*yes, explain the internet archive data. But also say that this is okay given that after 2012 the curve was flattened*
 
-There is a specific paragraph justifying how the time period of the data fits
-the aim of the paper and another line is now included in the second paragraph 
-of the introductions ("from the very early days until its maturity (1996-2012)").
+Framing
+>“With the exception of Sinai and Waldfogel (2004), who focused, among other things, on web content and demonstrated that more such content is available in larger markets, this strand of literature has been mostly concerned with the spatiality of the Internet’s hard and tangible infrastructure and shied away from exploring the spatial dimensions of one of the most distinct software layers of the Internet infrastructure, the Web.” This statement is simply not true, the literature reviews misses the entirety of web content diffusion studies: Zook, 2000, Malecki, 2002.
+Townsend, 2000, Gorman, 2002, Feldman, 2002, lots more
 
-\color{teal}
-Framing
-“With the exception of Sinai and Waldfogel (2004), who focused, among other things, on web content and demonstrated that more such content is available in larger markets, this strand of literature has been mostly concerned with the spatiality of the Internet’s hard and tangible infrastructure and shied away from exploring the spatial dimensions of one of the most distinct software layers of the Internet infrastructure, the Web.” This statement is simply not true, the literature reviews misses the entirety of web content diffusion studies: Zook, 2000, Malecki, 2002.
-Townsend, 2000, Gorman, 2002, Feldman, 2002, lots more
-\color{black}
+Para has been changed and added a long time series motivation. @zook2000web, @zook2000web and 
+@moss1997tracking used domain name registration data, but rest (e.g. @malecki2002economic and @moss2000internet) used focused on hard infrastructure.
 
-Point taken. Despite being aware of these studies, I omitted citing them as I 
-was primarily searching for studies testing the diffusion mechanisms over longer
-time periods. This error has now been corrected and the above paragraph has been 
-replaced with the following one, which emphasises the long time series motivation. 
-Regarding the actual references, @zook2000web, @zook2000web and @moss1997tracking 
-used relevant, domain name registration data, but the rest (e.g. @malecki2002economic 
-and @moss2000internet) focused on hard infrastructure. @gorman2002web examined 
-the head and branch locations of the top 40 e-business integration firms in the 
-USA, which although interesting is not directly relevant to this paper. I also
-went through of Feldman's work around 2002, but I did not find papers
-adopting a similar empirical perspective.
+>“for a digital technology as weightless as the Web” – this runs a bit counter to lots of work by digital geographers who have shown again and again how the digital is not weightless or placeless
 
-> Another line of motivation has to do with the very nature of the Web as a 
-digital technology and how it evolved over time. The digital geographies field 
-has debunked premature claims about the Internet's placeless-ness by mapping 
-various of its facets and technological layers [@tranos2013death], but the lack 
-of relevant long time-series data prevented researchers from observing digital
-phenomena overtime. Most of these studies focused on its hard infrastructure 
-[e.g. @malecki2002economic; @moss2000internet], on Internet subscribers 
-[e.g. @blank2018local], or on social media users [e.g. @crampton2013beyond].
-Exceptions include early work from @zook2000web, @zook2000web and 
-@moss1997tracking, which utilised domain name registration data and use it to
-map and illustrate the spatial footprint of the web content. Pivotal as these 
-contributions have been in understanding the economic geographies of the 
-Internet, they were based on cross-sectional, often spatially detailed, mapping 
-and on shallow time comparisons (e.g. 1993-1998) and did not observe the 
-evolution and diffusion of the Web until its maturity. As we know from the 
-relevant literature, different technologies diffuse at different rates and 
-there is no theory that can explain these differences 
-[@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
-Web among firms can offer useful insights.
-
-\color{teal}
-“for a digital technology as weightless as the Web” – this runs a bit counter to lots of work by digital geographers who have shown again and again how the digital is not weightless or placeless
-\color{black}
-
-\color{teal}
 Conclusion
-p.
+>p.
 18 “This is the first time that the importance of granular spatial mechanisms for the diffusion of an intangible, digital technology is exposed.” Honestly, I would not expect the diffusion of the web to be any different than any other kind of innovation diffusion.
@@
 So I’m not convinced it makes much sense to emphasize this, if a granular spatial analysis of innovation diffusion is already novel.
-\color{black}
 
-**for the last two comments, I need to soften the weightless argument throughout.**
+*for the last two comments, I need to soften the weightless argument throughout.* 
 I have added a long time series argument and also said that there is heterogeneity
@@
 
-As per the 'framing' point above, the digital motivation para in the 
-introduction has been amended acordilngly.
+The digital motivation para in the introduction has been ammended accordilngly.
 
@@
 
-\color{teal}
+Comments to the Author
+Review JOEG-2024-369: A multi-scale story of the diffusion of a new technology: the web
+
 This paper addresses a gap in Economic Geography by examining the spatial diffusion of digital technologies, specifically the adoption of commercial websites, from 1996 to 2012.
@@
 
-Most importantly, I suggest developing some clear hypotheses to structure and motivate the paper better. In light of this, I recommend major revisions.
-\color{black}
+> Most importantly, I suggest developing some clear hypotheses to structure and motivate the paper better. In light of this, I recommend major revisions.
 
-This is very helpful. The first two sections have been restructured around two 
-hypotheses. I believe the structure and the flow of the paper is much clearer
-now.
+Hypotheses have been added.
 
-\color{teal}
-\noindent 1. Abstract:\
-\noindent 1.
+>1. Abstract:
+1.
 The abstract needs significant improvement.
@@
 Additionally, the assertion that geographers have "shied away" from analyzing the diffusion of digital technologies is misleading; the issue was primarily due to the lack of available data.
-\color{black}
 
-**fix the abstract**
-
-"shied away" has been replaced in the opening paragraph.
-
-\color{teal}
-\noindent 2. Introduction\
-\noindent 2.
+2. Introduction
+>2.
 General Diffusion Studies: The statement that there are few contemporary general diffusion studies in Economic Geography is not accurate, although it may apply specifically to digital technologies.
@@
 o Feldman, M. P., Kogler, D. F., & Rigby, D. L. (2015). rKnowledge: The Spatial Diffusion and Adoption of rDNA Methods. Regional Studies, 49(5), 798–817. https://doi.org/10.1080/00343404.2014.980799
-\color{black}
 
-@PERIS2021101565 is indeed an interesting paper, but it is not directly relevant 
-as it focuses on information diffusion rather than diffusion of new 
-technologies. Information diffuses through social and, consequently, urban 
-networks, but it is not adopted like new technologies. @bednarz2020pulled was 
-already cited and @feldman2015rknowledge has now been added. These references 
-are included in the second section.
+Peris et al. (2021) is not directly relevant as it focuses on information 
+diffusion rather than diffusion of innovations or new technologies.  Bednarz et 
+al (2020) was already cited in the first submission. Feldman et al (2015) has 
+now been cited. These references are included in the literature section.
 
@@
 new technologies and innovation, I do believe that this is a fair statement. 
-The references listed below are the ones reviewed in the second section and none 
-of them are published in one of the flagship economic geography journals 
-(*Journal of Economic Geography* and *Economic Geography*). There are two papers
+The references listed below are the ones reviewed in the second section and none of 
+them are published in one of the flagship economic geography journals 
+(*Journal of Economic Geography* and *Economic Geography*). There two papers
 published in *The Professional Geographer*, which is a broad-scope geography 
@@
 [@feldman2015rknowledge] paper is focused on inventions as observed by patent 
-data contrary to the broader 'adoption of a new technology' approach of this 
-paper. All other papers listed below are published in business and economics 
-journals.
+data contrary to the broader 'adoption of a new technology' approach of this paper.
+All other papers listed below are published in business and economics journals.
 
@@
 Also, the lack of spatial granularity argument has been highlighted in various 
-instances throughout the paper as the main line of justification.
+instances throughout the paper:
 
-**ADD SPATIAL GRANULARITY IN CONCLUSIONS**
+*ADD SPATIAL GRANULARITY IN CONCLUSIONS*
+Added in intro and lit rev. Made it key justification as highlighted by Reviewer 1.
 
-\color{teal}
-\noindent 3. Novelty: The study needs to be more explicit about its novel contributions to understanding spatial diffusion mechanisms. Currently, the discussion is too general.
-\color{black}
+> 3. Novelty: The study needs to be more explicit about its novel contributions to understanding spatial diffusion mechanisms. Currently, the discussion is too general.
 
-**granularity**
+granularity
 
-**HAVE A THINK, intro has been restructured**
+*HAVE A THINK*
 
-\color{teal}
-\noindent 4.
+>4.
 Motivation: The motivation for the study is unclear.
 The text raises several issues, but it is unclear which are specific to digital technologies and which pertain to technological diffusion more broadly.
-\color{black}
 
-**HAVE A THINK, intro has been restructured**
+*HAVE A THINK*
 
-\color{teal}
-\noindent 5.
+>5.
 Structure and Focus: The discussion on motivations should be streamlined and clarified.
@@
 Consider restructuring this section to clearly differentiate between literature gaps (motivations) and expected outcomes (mechanisms).
-\color{black}
 
-The revised paper consistently illustrates that the focus is on commercial 
-websites.
-
-In the relevant paragrapg in the Data and Methods section, the following line
-has been added: "Henceforward any reference to websites will refer to commercial 
--- .co.uk -- websites".
-
-**fixed for intro, lit rev  and data -- check the rest of the paper.**
+**LINK WITH REVIEWER 2 ABOUT WEB DATA**
 
@@
 
-\color{teal}
-\noindent 6.
+> 6.
 Data Consistency: The terminology for the data should be consistent.
 At one point, the text refers to the “totality of the web,” while elsewhere, it specifies “commercial websites.” Clarify and maintain consistency throughout.
-\color{black}
 
-The terminology is now consistent across the paper. The "totality" point 
-referred to different web technologies and not to the totality of the web. It 
-has now been replaced with "an overarching technology".
+The terminology is consistent across the paper now. The "totality" point referred 
+to different web technologies. Now it has been replaced as "an overarching technology".
 
-\color{teal}
-\noindent 7.
+7.
 Recommended Structure: To align with the common structure in Journal of Economic Geography (JoEG), consider splitting the introduction into a shorter introduction and a dedicated theory section merged with the literature review.
-\color{black}
 
-The introduction is now shorter and the literature review section has been 
-re-structured and re-named as "The mechanisms that shape the diffusion of a new 
-technology".
-
-\color{teal}
-\noindent 3. Literature Review / Theory\
-\noindent 1.
+3. Literature Review / Theory
+> 1.
 Synthesis and Motivation: The literature review should be transformed into a theory section.
 Rather than presenting a historical review, synthesize the arguments to clearly motivate and inform your hypotheses.
-\color{black}
 
@@
 
-\color{teal}
-\noindent 2.
+> 2.
 Knowledge Gaps: The current review does not clearly identify the knowledge gaps that this study aims to address.
 Clarify whether the goal is to confirm known processes with new data and methods or to explore unique diffusion patterns related to websites.
-\color{black}
 
@@
 
-\color{teal}
-\noindent 3. Lack of Hypotheses: To clarify what the paper delivers and what knowledge gaps it fills, I strongly recommend developing a number of hypotheses that will be tested.
-\color{black}
+> 3. Lack of Hypotheses: To clarify what the paper delivers and what knowledge gaps it fills, I strongly recommend developing a number of hypotheses that will be tested.
 
@@
 
-\color{teal}
-\noindent 4. Fit to title: The term “multi-scalar” appears in the title but is never really discussed in the paper.
-\color{black}
+> 4. Fit to title: The term “multi-scalar” appears in the title but is never really discussed in the paper.
 
-This was discussed from the data section onwards, but indeed not before that. 
-Section 2 includes a new paragraph describing the second hypothesis (H~2~), 
-which explicitely discusses the value of the multiscale approach: 
-"Understanding the diffusion of a new technology in multiple, detailed spatial
-scales is of interest to economic geography as ..."
+Added before H~2~.
 
-\color{teal}
-\noindent 4. Data & Methods\
-\noindent 1. Order of Presentation: Introduce the data first, followed by the geographical units used in the study.
-\color{black}
+4. Data & Methods
+1. Order of Presentation: Introduce the data first, followed by the geographical units used in the study.
+2. Postcode Identification: The approach of identifying postcodes on websites is innovative, but potential errors need to be discussed. Specifically:
+>o How do you ensure that a postcode on a website accurately represents the company’s location or its subsidiary?
+*trading address, one to many postcodes*
 
-The order of presentation has been amended accordingly and the two scales of 
-analyses are presented after the data.
+>o How do you handle numbers that resemble postcodes but are not actual postcodes?
+*matched with postcode lookup file*
 
-\color{teal}
-\noindent 2. Postcode Identification: The approach of identifying postcodes on websites is innovative, but potential errors need to be discussed. Specifically:
-- How do you ensure that a postcode on a website accurately represents the 
-company’s location or its subsidiary?
-- How do you handle numbers that resemble postcodes but are not actual postcodes?
-- What about references to events or other organizations unrelated to the website’s owner?
+>o What about references to events or other organizations unrelated to the website’s owner?
+*we cannot distinguish, but 1-10 postcodes*
 Include a brief discussion of these potential issues.
-\color{black}
 
-A number of different strategies have been developed to address the potential 
-issues described above.
-
-\noindent 1. To prevent false positives, the resulted postcodes
-were matched against official postcode lookup files [@ons]. This was not 
-explicitly mentioned in the first submission on the paper and the previous 
-sentence has now been added in the relevant paragraph:
-
-> The text from these webpages was scanned using a regular expression
-(regex) to identify strings of text which resemble UK postcodes and one
-fifth of them included a mention to a postcode [@BL2013geo]. This
-information allows the geolocation of the data and the creation of the
-LAD and OA counts. To prevent false positives, the resulted postcodes
-were matched against official postcode lookup files [@ons].
-
-\noindent 2. Then, to remove the bias from link farms associated with a handful of postcodes
-having irregular high numbers of websites pointing to them, a targeted data 
-imputation procedure was developed. This is detailed in the data section -- see
-also Figure 1.
-
-\noindent 3. Regarding the points about subsidiary company location and event postings, 
-these cannot be directly observed and distinguished. However, a process that has 
-been successfully used in previous studies has been adopted here. Firstly, given 
-that the focus of the paper is commercial websites, the relevant literature 
-illustrates that such websites serve specific purposes:
-
-> they are used to diffuse information, support online transactions and share 
-opinions [@THELWALL2000441; @blazquez2018big].
-
-Secondly, the webpage-to-website aggregation process leads to websites with 
-different number of postcodes as presented in Table 1. To decrease heterogeneity, 
-the analysis firstly focuses on websites with one unique postcode. Given the the
-commercial nature of these websites (.co.uk), it is fair to expect that such 
-websites with one unique postcode in their web text represent companies with a 
-single trading location. The analysis included in the main body of the paper is
-based on this subset of the website population. For year 2000, this includes 
-70% of all the archived, geolocated commercial websites. Then, as a robustness
-check all the different steps of the analysis are repeated for websites with up 
-to 10 postcodes and the results are presented in the appendix. This includes 94%
-of all the archived, geolocated commercial websites. For this step, websites 
-with multiple postcodes were counted multiple times, one for each different 
-postcode. All the results are qualitatively the same and only marginally 
-different, which is indication of the robustness of the website geolocation 
-strategy.
-
-As a further argument for the robustness of the above process, the results of a 
-previous study which used the same approach for a small subset of these data 
-from a specific neighbourhood in London matched in terms of accuracy previous 
-research that was based on extensive qualitative work, interviews and 
-observational studies [@shoreditch]. This last point has now been added in the 
-relevant section:
-
-> Previous research which used the same methodology for a small subset of these 
-data illustrates the robustness of the above approach: its results matched in 
-terms of accuracy past rounds of research based on extensive qualitative work, 
-interviews and observational studies for the same study area [@shoreditch].
-
-\color{teal}
-\noindent 3.
+>3.
 Equation 2: The rationale behind Equation 2 is unclear.
@@
 Additionally, clarify how spatial dependencies are accounted for in the model.
-\color{black}
+*growth as an appendix, spatial dependencies have been included*
 
-**growth as an appendix, spatial dependencies have been included**
-
-\color{teal}
-\noindent 5. Results\
-\noindent 1.
+5. Results
+>1.
 Diffusion Speed: The decision to binarize diffusion speed is questionable.
 Continuous measures, such as growth rates or the time required to reach 50% saturation, would provide more informative insights.
-\color{black}
+*This is only for the map, this is a critical point in diffusion studies*
 
-**This is only for the map, this is a critical point in diffusion studies**
+> 2.
+Geographical Scale: While the effort to model diffusion curves at a fine geographical scale is commendable, the paper does not sufficiently explain the benefits or additional knowledge gained from this approach.
 
-\color{teal}
-\noindent 2.
-Geographical Scale: While the effort to model diffusion curves at a fine geographical scale is commendable, the paper does not sufficiently explain the benefits or additional knowledge gained from this approach.
-\color{black}
-
-**reveal that theories are valid for lcoal scales and inform local tech policies**
-
-\color{teal}
-\noindent 3.
+> 3.
 Early Adopters and Latecomers: The discussion of early-adopter and latecomer advantages is underdeveloped.
@@
 If this aspect is retained, it should be elaborated in the theory section, including how it relates to local capabilities or diffusion hierarchies.
-\color{black}
 
-**relates to local capabilities or diffusion hierarchies**
-
-\color{teal}
-\noindent 4.
+4.
 Descriptive Statistics: The value of the descriptive Moran’s I and Gini coefficient analyses in Section 4.2 is unclear.
 Instead, calculating growth rates of adoption and examining their distribution over time would offer a better sense of adoption speed and magnitude.
-\color{black}
 
-**think, growth rates will not work for OA**
+5. Retail Centers: The sudden shift to retail centers on page 15 is confusing, given that the data pertains to all commercial sites. Clarify the rationale for this focus.
 
-\color{teal}
-\noindent 5. Retail Centers: The sudden shift to retail centers on page 15 is confusing, given that the data pertains to all commercial sites. Clarify the rationale for this focus.
-\color{black}
+6. Modeling Details: The methodological details in Section 4.3 should be moved to the methods section to maintain clarity and coherence. Including these details in the results section is confusing.
 
-**hierarchy**
-
-\color{teal}
-\noindent 6.
-Modeling Details: The methodological details in Section 4.3 should be moved to the methods section to maintain clarity and coherence.
-Including these details in the results section is confusing.
-\color{black}
-
-\color{teal}
-\noindent 7.
+7.
 Clarity and Structure: Section 4.3 is difficult to follow.
@@
 Without this, the section feels like a descriptive summary without clear takeaways for the reader.
-\color{black}
 
-\color{teal}
-\noindent 
 Overall Recommendations
-\noindent 1. Consistency: Ensure consistent terminology and focus throughout the paper, particularly when referring to the web and commercial websites.
-\noindent 2. Structure: Adopt a clearer structure by separating the introduction, theory/literature review, methods, and results sections.
-\noindent 3. Clarity: Develop hypotheses to frame the results and provide clearer explanations for methodological choices and findings.
-\color{black}
-
-## References
+1. Consistency: Ensure consistent terminology and focus throughout the paper, particularly when referring to the web and commercial websites.
+2. Structure: Adopt a clearer structure by separating the introduction, theory/literature review, methods, and results sections.
+3. Clarity: Develop hypotheses to frame the results and provide clearer explanations for methodological choices and findings.

### 2025-01-21 Untitled/Untitled.Rmd

--- 2022-05-26 paper/paper.Rmd
@@
 ---
-title: Title of submission to PLOS journal
+title: |
+  Taylor & Francis Rmarkdown template for authors (\LaTeX\-based 
+  \textsf{Interact} layout + Chicago author-date reference style)
+type: ARTICLE TEMPLATE
 author:
-  - name: Emmanouil Tranos
-    email: e.tranos@bristol.ac.uk
-    affiliation: University of Bristol and The Alan Turing Institute
-    corresponding: e.tranos@bristol.ac.uk
-address:
-  - code: University of Bristol and The Alan Turing Institute
-    address: University of Bristol, School of Geographical Sciences, Bristol, UK; The Alan Turing Institute, London, UK
+  - name: A. N. Author
+    affil: a
+    email: latex.helpdesk@tandf.co.uk
+  - name: John Smith
+    affil: b
+    email: john.smith@uni-freiburg.de
+  - name: Dominik Leutnant
+    affil: c, \dagger, \ddagger
+    email: leutnant@fh-muenster.de
+affiliation:
+  - num: a
+    address: |
+      Taylor & Francis, 4 Park Square, Milton Park, Abingdon, UK
+  - num: b
+    address: |
+      Institut für Informatik, Albert-Ludwigs-Universität, Freiburg, Germany
+  - num: c
+    address: |
+      Muenster University of Applied Sciences - 
+      Institute for Infrastructure, Water, Resources, Environment,
+      Correnstr. 25, 48149 Münster, Germany
+thanks:
+  author:
+  - name: A. N. Author
+    email: latex.helpdesk@tandf.co.uk
+bibliography: interactcadsample.bib
+appendix: appendix.tex
 abstract: |
-  blah blah
-
-author_summary: |
-  blah blah
-
-bibliography: mybibfile.bib
-output: rticles::plos_article
-csl: plos.csl
-header-includes:
-- \usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}
+  This template is for authors who are preparing a manuscript for a Taylor & 
+  Francis journal using the \LaTeX\ document preparation system and the 
+  `interact} class file, which is available via selected journals' home 
+  pages on the Taylor & Francis website.
+keywords: |
+  Sections; lists; figures; tables; mathematics; fonts; references; appendices
+header-includes: |
+  \usepackage{hyperref}
+  \usepackage[utf8]{inputenc}
+  \def\tightlist{}
+output: rticles::tf_article
 ---
 
-```{r echo=FALSE, message = FALSE, warning=FALSE}
-library(stargazer)
-library(tidyverse)
-#library(rmarkdown)
+# Introduction
+
+In order to assist authors in the process of preparing a manuscript for a journal, the Taylor & Francis '\textsf{Interact}' layout style has been implemented as a \LaTeXe\ class file based on the `article` document class.
+A \textsc{Bib}\TeX\ bibliography style file and a sample bibliography are also provided in order to assist with the formatting of your references.
+
+Commands that differ from or are provided in addition to standard \LaTeXe\ are described in this document, which is _not_ a substitute for a \LaTeXe\ tutorial.
+
+The `interactcadsample.tex` file can be used as a template for a manuscript by cutting, pasting, inserting and deleting text as appropriate, using the preamble and the \LaTeX\ environments provided (e.g.\ `\begin{abstract}`, `\begin{keywords}`).
+
+## The \textsf{Interact} class file
+
+The `interact` class file preserves the standard \LaTeXe\ interface such that any document that can be produced using `article.cls` can also be produced with minimal alteration using the `interact` class file as described in this document.
+
+If your article is accepted for publication it will be typeset as the journal requires in Minion Pro and/or Myriad Pro.
+Since most authors will not have these fonts installed, the page make-up is liable to alter slightly with the change of font.
+Also, the `interact` class file produces only single-column format, which is preferred for peer review and will be converted to two-column format by the typesetter if necessary during preparation of the proofs.
+Please therefore do not try to match the typeset format exactly, but use the standard \LaTeX\ fonts instead and ignore details such as slightly long lines of text or figures/tables not appearing in exact synchronization with their citations in the text: these details will be dealt with by the typesetter.
+Similarly, it is unnecessary to spend time addressing warnings in the log file -- if your .tex file compiles to produce a PDF document that correctly shows how you wish your paper to appear, such warnings will not prevent your source files being imported into the typesetter's program.
+
+## Submission of manuscripts prepared using _\LaTeX_
+
+Manuscripts for possible publication should be submitted to the Editors for review as directed in the journal's Instructions for Authors, and in accordance with any technical instructions provided in the journal's ScholarOne Manuscripts or Editorial Manager site.
+Your \LaTeX\ source file(s), the class file and any graphics files will be required in addition to the final PDF version when final, revised versions of accepted manuscripts are submitted.
+
+Please ensure that any author-defined macros used in your article are gathered together in the preamble of your .tex file, i.e.\ before the `\begin{document}` command.
+Note that if serious problems are encountered in the coding of a document (missing author-defined macros, for example), the typesetter may resort to rekeying it.
+
+# Using the `interact` class file
+
+For convenience, simply copy the `interact.cls` file into the same directory as your manuscript files (you do not need to install it in your \TeX\ distribution).
+In order to use the `interact` document class, replace the command `\documentclass{article}` at the beginning of your document with the command `\documentclass{interact}`.
+
+The following document-class options should _not_ be used with the `interact` class file:
+
+* `10pt`, `11pt`, `12pt` -- unavailable;
+* `oneside`, `twoside` -- not necessary, `oneside` is the default;
+* `leqno`, `titlepage` -- should not be used;
+* `twocolumn` -- should not be used;
+* `onecolumn` -- not necessary as it is the default style.
+
+To prepare a manuscript for a journal that is printed in A4 (two column) format, use the `largeformat` document-class option provided by `interact.cls`; otherwise the class file produces pages sized for B5 (single column) format by default.
+The `geometry` package should not be used to make any further adjustments to the page dimensions.
+
+<!-- If your manuscript has supplementary content you can prepare this using the `suppldata` document-class option, which will suppress the `article history' date.
+This option must _not_ be used on any primary content.
+-->
+
+# Additional features of the `interact` class file
+
+## Title, authors' names and affiliations, abstracts and article types
+
+## Abbreviations
+
+A list of abbreviations may be included if required, enclosed within an `abbreviations` environment, i.e.\ `\begin{abbreviations}`\ldots`\end{abbreviations}`, immediately following the `abstract` environment.
+
+## Keywords
+
+A list of keywords may be included if required, enclosed within a `keywords` environment, i.e.\ `\begin{keywords}`\ldots`\end{keywords}`.
+Additional keywords in other languages (preceded by a translation of the word `keywords`) may also be included within the `keywords` environment if required.
+
+## Subject classification codes
+
+AMS, JEL or PACS classification codes may be included if required.
+The `interact` class file provides an `amscode` environment, i.e.\ `\begin{amscode}`\ldots`\end{amscode}`, a `jelcode` environment, i.e.\ `\begin{jelcode}`\ldots`\end{jelcode}`, and a `pacscode` environment, i.e.\ `\begin{pacscode}`\ldots`\end{pacscode}` to assist with this.
+
+## Additional footnotes to the title or authors' names
+
+<!-- The `\thanks` command may be used to create additional footnotes to the title or authors' names if required. Footnote symbols for this purpose should be used in the order -->
+<!-- $^\ast$~(coded as `$^\ast$`), $\dagger$~(`$\dagger$`), $\ddagger$~(`$\ddagger$`), $\S$~(`$\S$`), $\P$~(`$\P$`), $\|$~(`$\|$`), -->
+<!-- $\dagger\dagger$~(`$\dagger\dagger$`), $\ddagger\ddagger$~(`$\ddagger\ddagger$`), $\S\S$~(`$\S\S$`), $\P\P$~(`$\P\P$`). -->
+
+Note that any `footnote`s to the main text will automatically be assigned the superscript symbols 1, 2, 3, etc.
+by the class file.^[If preferred, the `endnotes` package may be used to set the notes at the end of your text, before the bibliography.
+The symbols will be changed to match the style of the journal if necessary by the typesetter.]
+
+# Embedding R code
+
+## Code chunks
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-# Introduction {#sec:1}
+```{r cars}
+summary(cars)
+```
 
-@beardsell1999spatial
-To characterize evolution of the computer industry, we examine the distribution
-of relative employment across cities in 1977 and how that distribution
-changes over time to 1992
+## Including Plots
 
-There is no
-tendency of relative size distributions of urban computer employment to collapse,
-go bimodal, or fully spread. Overall computers exhibit some turbulence,
-with dramatic big winners and losers among cities. In attracting or repelling an
-industry, urban heterogeneity is important. Large, well educated cities near San
-Jose have much greater chances of attracting high-tech employment and less of
-losing it
+You can also embed plots, for example:
 
-@bednarz2020pulled
-spatial innovation diffusion
-By using Bayesian survival models with time-dependent data of wind turbine deployment
-and firm foundation for 402 German regions between the years 1970 and 2015, we show that the spatial
-evolution of the German wind energy industry was more strongly influenced by local demand–
-pull than local supply–push processes.
+```{r pressure, echo=FALSE, out.width='80%'}
+plot(pressure)
+```
 
-The industry’s initial locations are distributed relatively arbitrarily
-and unpredictably, as their needs in terms of resources and skills are diverse and distinct from the older existing
-industries (Boschma and Lambooy, 1999). Consequently, emerging industries are characterized by relatively high
-degrees of freedom in terms of location. In later extensions of the concept, the assumption of the randomness of locations
-was revised with greater importance assigned to regional conditions (Boschma and Lambooy, 1999; Fornahl
-et al., 2012).
+Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
 
-larger local technological niches (at least from a demand perspective) are
-more likely to be formed in urban areas than in rural ones, which is in line with the work of Ha¨gerstrand (1965a,b).
-However, size is not the only regional characteristic that may matter in this context. Regions also differ in accumulated
-experiences and the presence of tacit knowledge with respect to products, as well as in actors’ propensity share
-this knowledge among producers (Martin et al., 2019)
+# Some guidelines for using the standard features of \LaTeX
 
-@bento2015spatial
-The concept of absorptive capacity has also been applied in the TIS context in terms of the con-struction of capacity needed to transfer and implement a new technology in the receiving country (cf.Van Alphen, 2011), with repercussions for the nature of system functions
+## Sections
 
-@das2022diffusion [not good paper]
-Invention, innovation
-and diffusion are considered as the three pillars of
-the theory of technological change and Schumpeter is considered
-as the father of the study of diffusion
+The \textsf{Interact} layout style allows for five levels of section heading, all of which are provided in the `interact` class file using the standard \LaTeX\ commands `\section`, `\subsection`, `\subsubsection`, `\paragraph` and `\subparagraph`.
+Numbering will be automatically generated for all these headings by default.
 
-Epidemic approach: The approach emphasizes that the adoption of a technology
-spreads like an infection among the potential adopters.
-By taking a homogeneous group of potential adopters as
-an assumption, it explains the adoption of any innovation.
+## Lists
 
-To summarize
-the discussion, first, here the diffusion process is selfperpetuating
-as initial use stimulates further use. Second, it
-follows a disequilibrium path as the level of users is
-always lower than the number of potential users (Stoneman
-2002).
+Numbered lists are produced using the `enumerate` environment, which will number each list item with arabic numerals by default. For example,
 
-The
-criticisms are mainly based on the fact that although the
-approach gives an idea of aggregate (industry or household)
-behaviour, it does not focus on the individual’s
-(firm or household) adoption process
+1. first item
+2. second item
+3. third item
 
-Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.) Early adopters: next
-13.5%, (III.) Early Majority: following 34%, (IV.) Late Majority: next 34%, and the rest as (V.) Laggards
+Alternative numbering styles can be achieved by inserting an optional argument in square brackets to each `item`, e.g.\ `\item[(i)] first item`\, to create a list numbered with roman numerals at level one.
 
-@morrill2020spatial
-Spatial diffusion is the process by which behavior or characteristics of the landscape change as a result of
-what happens elsewhere earlier. Spatial diffusion is the spread of the phenomenon, over space and timed,
-from limited origins
+Bulleted lists are produced using the `itemize` environment. For example,
 
-Contagious diffusion=> spatial autocorelation
+* First bulleted item
+* Second bulleted item
+* Third bulleted item
 
-expansion-type diffusion
+## Figures
 
-the three geographic diffusion processes are (1) purely contagious where distance or adjacency
-is the absolute controlling factor, (2) purely hierarchical where size or urban position in the central place
-hierarchy is the absolute controlling factor, or (3) where the location of change is purely random. These are
-the pure or ideal forms of the diffusion process; all real diffusion processes are the result of a combination of
-these extremes.
+```{r pressure-plot, fig.show='hide'}
+plot(pressure)
+```
 
-The work by the communication sociologist Everett Rogers (1962, 1971, 1983) has emphasized the role of
-information, communication, formal and informal media, opinion leaders and social networks, and economic
-and psychological constraints on acceptance. Rogers’s work stresses the decision mechanism of the potential
-adopter; the work of Rogers partitions the adopter’s process of choosing to accept a new phenomenon or
-trait into five stages: stage 1–the potential adopter gains knowledge or awareness of an innovation; stage
-2–persuasion is exercised to adopt; stage 3–a decision is made to adopt; stage 4–the decision to adopt is
-implemented; stage 5–the adopter confirms the decision to adopt. In Rogers’s stage theory there may be a
-substantial time lag between when a potential adopter becomes aware of the new characteristic and when a
-decision is actually made to adopt.
+The `interact` class file will deal with positioning your figures in the same way as standard \LaTeX.
+It should not normally be necessary to use the optional `[htb]` location specifiers of the `figure` environment in your manuscript; you may, however, find the `[p]` placement option or the `endfloat` package useful if a journal insists on the need to separate figures from the text.
 
-Hägerstrand considered diffusion to be a fundamental geographic process: Whatever the phenomenon being
-diffused might be, one may consider it in the context of a larger universal process of spatial diffusion.
+Figure captions appear below the figures themselves, therefore the `\caption` command should appear after the body of the figure.
+For example, Figure~\ref{sample-figure} with caption and sub-captions is produced using the following commands:
+```
+\begin{figure}
+\centering
+\subfigure[An example of an individual figure sub-caption.]{%
+\resizebox*{5cm}{!}{\includegraphics{path/to/fig}}}\hspace{5pt}
+\subfigure[A slightly shorter sub-caption.]{%
+\resizebox*{5cm}{!}{\includegraphics{path/to/fig}}}
+\caption{Example of a two-part figure with individual sub-captions
+ showing that captions are flush left and justified if greater
+ than one line of text.} \label{sample-figure}
+\end{figure}
+```
+\begin{figure}
+\centering
+\subfigure[An example of an individual figure sub-caption.]{%
+\resizebox*{5cm}{!}{\includegraphics{`r knitr::fig_chunk('pressure-plot', 'pdf')`}}}\hspace{5pt}
+\subfigure[A slightly shorter sub-caption.]{%
+\resizebox*{5cm}{!}{\includegraphics{`r knitr::fig_chunk('pressure-plot', 'pdf')`}}}
+\caption{Example of a two-part figure with individual sub-captions
+ showing that captions are flush left and justified if greater
+ than one line of text.} \label{sample-figure}
+\end{figure}
 
-@zook2022mapping
-Because new digital activities are rarely—if
-ever—captured in official state data, researchers
-must rely on information gathered from alternative
-sources. With this in mind, the overarching aim of
-this article is outlining an approach for analyzing the
-“new spaces and geographies” of digital phenomena
-and practices.
+To ensure that figures are correctly numbered automatically, the `\label` command should be included just after the `\caption` command, or in its argument.
 
-@leibowicz2016representing for energy
-There is no generally accepted theory that explains diffusion
-rate heterogeneity across technologies, but several factors are
-considered important. Greater unit scale and larger market size
-contribute to slower diffusion. Requirements for interrelated
-technologies or complex infrastructures also hinder the diffusion
-process (Grubler, 2012).
+The `\subfigure` command requires `subfigure.sty`, which is called in the preamble of the `interacttfssample.tex` file (to allow your choice of an alternative package if preferred) and included in the \textsf{Interact} \LaTeX\ bundle for convenience.
+Please supply any additional figure macros used with your article in the preamble of your .tex file.
 
-Mobile phones benefited from early
-deployment in recreational boats and automobiles, where the
-traditional competitor was not a viable option. In the early
-stages of diffusion, performance is a more important driver of
-adoption than cost competitiveness. Typically, significant cost
-reductions only occur once the technology reaches a deployment
-level capable of supporting standardization and mass
-production (Wilson, 2012).
+The source files of any figures will be required when the final, revised version of a manuscript is submitted.
+Authors should ensure that these are suitable (in terms of lettering size, etc.) for the reductions they envisage.
 
-Diffusion is a spatial (Hagerstrand, 1967) as well as a
-temporal process, and historical evidence confirms that
-technologies diffuse at different times, at different rates, and
-to different extents in different places, and can be significantly
-influenced by policies (Victor, 1993). An illustrative empirically
-based theory for technology diffusion acrossmultiple regions is
-suggested by “Schmidt's Law” (Grubler, 1990). With respect to a
-specific technology, it divides regions into three groups: core,
-rim, and periphery. The core iswhere the technology is invented,
-subject to experimentation, and first deployed commercially.
-To develop the technology, the core must have both sufficient
-innovative ability and economic motivation to make the
-necessary resource commitment. As the technology spreads
-from the core to the rim and eventually to the periphery, later
-adopters have the advantage of benefiting from the experiences
-of earlier adopters via knowledge spillover effects (Verdolini and
-Galeotti, 2011). At this point, the technology has likely been
-refined and costs have likely declined significantly. Schmidt's
-Law claims that, as a result, diffusion proceeds at a faster rate in
-regions which adopt the technology later. However, diffusion
-tends to be less pervasive in regions where adoption is delayed,
-ultimately saturating at a lower extent than in regions where
-adoption occurred earlier. This is often attributed to a lack of
-proper institutions and infrastructures in later adopters, markets
-whichmay not be as favorable to a technology, or the availability
-of yet newer options that invite technological leapfrogging.
+The `epstopdf` package can be used to incorporate encapsulated PostScript (.eps) illustrations when using PDF\LaTeX, etc.
+Please provide the original .eps source files rather than the generated PDF images of those illustrations for production purposes.
 
-Empirical evidence supports the validity of Schmidt's Law
-over a wide range of technologies, time periods, and geographical
-contexts. A recent meta-analysis of technology up-scaling
-found that diffusion accelerated moving from the core to the
-rim and periphery for technologies as diverse as natural gas
-power, oil refineries, and automobiles (Wilson, 2009). One
-historical example that conforms particularly well to Schmidt's
-Law is the diffusion of coal power in Europe (Grubler, 2012).
-England emerged as the core region for coal power because it
-had legal and economic institutions that incentivized scientific
-pursuits, domestic coal reserves, and a clear industrial motivation
-to replace water power with coal. 
-...
+## Tables
 
-Lastly, it can be globally optimal for innovative economies
-to deploy advanced technologies more than what is locally
-optimal if this enables faster diffusion in less advanced regions
-that could benefit from the technology.
+The `interact` class file will deal with positioning your tables in the same way as standard \LaTeX.
+It should not normally be necessary to use the optional `[htb]` location specifiers of the `table` environment in your manuscript; you may, however, find the `[p]` placement option or the `endfloat` package useful if a journal insists on the need to separate tables from the text.
 
-@wilson201281
-Logistic growth describes an initial period of gradual diffusion as a technology is introduced as a new commercial application, moving then through a rapid, exponential growth phase, before slowing and eventually saturating [@grubler1999dynamics].
-The substitution of incumbent technologies by new competitors leads to subsequent decline and eventual obsolescence.
+The `tabular` environment can be used as shown to create tables with single horizontal rules at the head, foot and elsewhere as appropriate.
+The captions appear above the tables in the \textsf{Interact} style, therefore the `\tbl` command should be used before the body of the table.
+For example, Table~\ref{sample-table} is produced using the following commands:
+\begin{table}
+\tbl{Example of a table showing that its caption is as wide as
+ the table itself and justified.}
+{\begin{tabular}{lcccccc} \toprule
+ & \multicolumn{2}{l}{Type} \\ \cmidrule{2-7}
+ Class & One & Two & Three & Four & Five & Six \\ \midrule
+ Alpha\textsuperscript{a} & A1 & A2 & A3 & A4 & A5 & A6 \\
+ Beta & B2 & B2 & B3 & B4 & B5 & B6 \\
+ Gamma & C2 & C2 & C3 & C4 & C5 & C6 \\ \bottomrule
+\end{tabular}}
+\tabnote{\textsuperscript{a}This footnote shows how to include
+ footnotes to a table if required.}
+\label{sample-table}
+\end{table}
+```
+\begin{table}
+\tbl{Example of a table showing that its caption is as wide as
+ the table itself and justified.}
+{\begin{tabular}{lcccccc} \toprule
+ & \multicolumn{2}{l}{Type} \\ \cmidrule{2-7}
+ Class & One & Two & Three & Four & Five & Six \\ \midrule
+ Alpha\textsuperscript{a} & A1 & A2 & A3 & A4 & A5 & A6 \\
+ Beta & B2 & B2 & B3 & B4 & B5 & B6 \\
+ Gamma & C2 & C2 & C3 & C4 & C5 & C6 \\ \bottomrule
+\end{tabular}}
+\tabnote{\textsuperscript{a}This footnote shows how to include
+ footnotes to a table if required.}
+\label{sample-table}
+\end{table}
+```
 
-Early on in their lifecycle, new technologies are crude, imperfect,
-and expensive (Rosenberg, 1994). New energy technologies
-are attractive for their ability to perform a particular task or
-deliver a new or improved energy service (Fouquet, 2010). This is
-often circumscribed by a particular set of needs in a particular
-context: a market ‘niche’. End-users in niche markets are generally
-less sensitive to the effective price of the energy service
-provided or have a higher willingness to pay for its performance
-advantages (Fouquet, 2010). Thus initially, performance dominates
-cost competitiveness (Wilson and Grubler, 2011). Market
-niches afford some protection from competitive pressures, allowing
-technologies to be tested and improved in applied settings,
-reducing uncertainties with performance or market demand
-(Kemp et al., 1998). Costs may only fall substantively after an
-extended period of commercial experimentation, concurrent with
-the establishment of an industrial base and characteristic moves
-towards standardisation and mass production (Grubler, 1998).
-The influence of accumulating production experience on costs is
-captured by the concept of learning.
+To ensure that tables are correctly numbered automatically, the `\label` command should be included just before `\end{table}`.
 
-@grubler1990rise
-Following the definition given by Rogers, 1983, diffusion is the
-process by which an innovation is communicated through certain communication
-channels over time among the members of a social system.
-An innovation may be an idea, object or practice which is perceived
-as new by an individual or another unit of adoption (e.g., an
-organization) .
-Communication via communication channels refers to how
-knowledge about an innovation is transferred between the members of
-a social system. Communication takes time, and is therefore an
-important aspect of innovation diffusion. In fact, all theoretical and
-empirical diffusion studies agree that an innovation does not instantly
-diffuse into a social system. Instead a typical time pattern of diffusion
-along a S-shaped trajectory seems to be the rule. The S-shaped pattern
-of diffusion appears to be a basic anthropologic phenomenon, as
-it is also confirmed by a large number of studies· of primitive preindustrial
-societies.
+The `\toprule`, `\midrule`, `\bottomrule` and `\cmidrule` commands are those used by `booktabs.sty`, which is called by the `interact` class file and included in the \textsf{Interact} \LaTeX\ bundle for convenience.
+Tables produced using the standard commands of the `tabular` environment are also compatible with the `interact` class file.
 
-The scale of spatial innovation diffusion is similar* to that of
-temporal diffusion models: hierarchically decomposed.
+## Landscape pages
 
-Spatial innovation diffusion in Hagerstrand's model consists of
-three stages.
-Stage 1: Local concentrations of initial acceptances (initial
-agglomeration) .
-Stage 2: Radial dissemination outward from the initial agglomeration,
-accompanied by the rise of secondary agglomerations.
-Stage 3: The condensing and saturation stage, in which filling-in
-occurs and diffusion eventually ceases.
+If a figure or table is too wide to fit the page it will need to be rotated, along with its caption, through 90$^{\circ}$ anticlockwise.
+Landscape figures and tables can be produced using the `rotating` package, which is called by the `interact` class file.
+The following commands (for example) can be used to produce such pages.
+```
+\setcounter{figure}{1}
+\begin{sidewaysfigure}
+\centerline{\epsfbox{figname.eps}}
+\caption{Example landscape figure caption.}
+\label{landfig}
+\end{sidewaysfigure}
+```
+```
+\setcounter{table}{1}
+\begin{sidewaystable}
+ \tbl{Example landscape table caption.}
+  {\begin{tabular}{@{}llllcll}
+    .
+    .
+    .
+  \end{tabular}}\label{landtab}
+\end{sidewaystable}
+```
+Before any such float environment, use the `\setcounter` command as above to fix the numbering of the caption (the value of the counter being the number given to the preceding figure or table).
+Subsequent captions will then be automatically renumbered accordingly.
+The `\epsfbox` command requires `epsfig.sty`, which is called by the `interact` class file and is also included in the \textsf{Interact} \LaTeX\ bundle for convenience.
 
-Thus, three main characteristics define the spatial diffusion process:
-(1) As in temporal diffusion models, an S-shaped pattern in the
-cumulative level of adoption.
-(2) A "hierarchy effect" by which diffusion proceeds from initial
-centers to secondary ones. Within the context of urban systems
-(central places) this implies a hierarchical spread of an innovation
-from larger to smaller centers, and from the centers to the surrounding
-hinterland.
-(3) A "neighborhood effect" Le., diffusion proceeds outwards from
-innovation centers, first "hitting" (Brown, 1981) nearby rather
-than far-away locations.
+Please note that if the `endfloat` package is used, one or both of the commands
+```
+\DeclareDelayedFloatFlavor{sidewaysfigure}{figure}
+\DeclareDelayedFloatFlavor{sidewaystable}{table}
+```
+will need to be included in the preamble of your .tex file, after the `endfloat` package is loaded, in order to process any landscape figures and/or tables correctly.
 
-Later Hagerstrand conceptualized
-physical "barrier" effects like lakes or uninhabited areas, which, in
-addition to distance, act as further retarding effects on diffusion.
-These are formalized in the form of "zero" or "half" contact multiplicators
-on the (distance decaying) message flows.
+## Theorem-like structures
 
-With respect to the formalization of the communication flows
-Hagerstrand defines a "mean information field" (MIF), in which the
-probability of communication is a negative function of distance
-between individuals
+A predefined `proof` environment is provided by the `amsthm` package (which is called by the `interact` class file), as follows:
+\begin{proof}
+More recent algorithms for solving the semidefinite programming relaxation are particularly efficient, because they explore the structure of the MAX-CUT problem.
+\end{proof}
+\noindent This was produced by simply typing:
+```
+\begin{proof}
+More recent algorithms for solving the semidefinite programming
+relaxation are particularly efficient, because they explore the
+structure of the MAX-CUT problem.
+\end{proof}
+```
+Other theorem-like environments (theorem, definition, remark, etc.) need to be defined as required, e.g.\ using `\newtheorem{theorem}{Theorem}` in the preamble of your .tex file (see the preamble of `interactcadsample.tex` for more examples).
+You can define the numbering scheme for these structures however suits your article best.
+Please note that the format of the text in these environments may be changed if necessary to match the style of individual journals by the typesetter during preparation of the proofs.
 
-@ding2010modeling model the spatial diffusion of mobile telecommunications in China
-as well as its determinants. regional socioeconomic characteristics play an
-important role in determining the timing,
-speed, and level of mobile telecommunications
-diffusion in China
+## Mathematics
 
-Rogers
-(1995) characterized early adopters as knowledgeable
-risk takers. Griliches (1957) looked
-at the rate of return for early adopters and
-characterized them as profit maximizers. The
-geographer Hagerstrand (1967) was focusing
-on the character of early adopters in the transmittal
-process that produces follow-on participants
-in the process. Space was treated as a
-contiguity system with embedded characteristics
-producing barrier elements to the diffusion
-process.
+### Displayed mathematics
 
-REVIEW ON Technological Diffusion
+The `interact` class file will set displayed mathematical formulas centred on the page without equation numbers if you use the `displaymath` environment or the equivalent `\[...\]` construction.
+For example, the equation
+\[
+ \hat{\theta}_{w_i} = \hat{\theta}(s(t,\mathcal{U}_{w_i}))
+\]
+was typeset using the commands
+```
+\[
+ \hat{\theta}_{w_i} = \hat{\theta}(s(t,\mathcal{U}_{w_i}))
+\]
+```
 
-Perkins and Neumayer (2005)
-examined whether latecomers have a higher diffusion
-rate of three technologies including digital
-telephone mainline
+For those of your equations that you wish to be automatically numbered sequentially throughout the text for future reference, use the `equation` environment, e.g.
+\begin{equation}
+ \hat{\theta}_{w_i} = \hat{\theta}(s(t,\mathcal{U}_{w_i}))
+\end{equation}
+was typeset using the commands
+```
+\begin{equation}
+ \hat{\theta}_{w_i} = \hat{\theta}(s(t,\mathcal{U}_{w_i}))
+\end{equation}
+```
 
-@bento2018time explore 
-What determines the duration of
-formative phases for energy innovations in different markets? We are interested
-both in initial markets (also: core, lead, first mover, early
-adopter) where formative phases prepare technologies for mass commercialization,
-and in follower markets (also: periphery, lag, late
-adopter) where accelerated formative phases may benefit from diffusion
-and spillovers.
+Part numbers for sets of equations may be generated using the `subequations` environment, e.g.
+\begin{subequations} \label{subeqnexample}
+\begin{equation}
+     \varepsilon \rho w_{tt}(s,t) = N[w_{s}(s,t),w_{st}(s,t)]_{s},
+     \label{subeqnparta}
+\end{equation}
+\begin{equation}
+     w_{tt}(1,t)+N[w_{s}(1,t),w_{st}(1,t)] = 0,
+     \label{subeqnpartb}
+\end{equation}
+\end{subequations}
+which was typeset using the commands
+```
+\begin{subequations} \label{subeqnexample}
+\begin{equation}
+     \varepsilon \rho w_{tt}(s,t) = N[w_{s}(s,t),w_{st}(s,t)]_{s},
+     \label{subeqnparta}
+\end{equation}
+\begin{equation}
+     w_{tt}(1,t)+N[w_{s}(1,t),w_{st}(1,t)] = 0,   \label{subeqnpartb}
+\end{equation}
+\end{subequations}
+```
+<!-- This is made possible by the `amsmath` package, which is called by the class file.
+If you put a `\label` just after the `\begin{subequations}` command, references can be made to the collection of equations, i.e.\ `(\ref{subeqnexample})' in the example above.
+Or, as the example also shows, you can label and refer to each equation individually -- i.e.\ `(\ref{subeqnparta})' and `(\ref{subeqnpartb})'.
+-->
 
-Hagerstrand (1968) was the first
-to demonstrate that diffusion typically starts from innovative centers
-(‘core´) and disseminates, through a hierarchy of subcenters, to the
-periphery. This process is sequential rather than simultaneous, and
-tends to accelerate from core to periphery, where diffusion reaches a
-lower intensity than in the core (Hagerstrand, 1968; Grubler, 1990;
-Morrill, 2005). A similar effect of spatial acceleration has been identified
-in marketing research and is known as the “lead-lag effect” (Peres
-et al., 2010).
+Displayed mathematics should be given end-of-line punctuation appropriate to the running text sentence of which it forms a part, if required.
 
-@perkins2005international analyze whether the rate at which new producer
-technologies diffuse is significantly influenced by
-(1) latecomer advantage and (2) engagement with the
-global economy via trade and foreign investment
+### Math fonts
 
-Indeed, precisely
-because of these latecomer advantages, developing
-countries1 are believed to be well placed to catch up with
-developed ones (Gerschenkron 1962; Abramovitz
-1986).
+#### Superscripts and subscripts
+Superscripts and subscripts will automatically come out in the correct size in a math environment (i.e.\ enclosed within `\(...\)` or `$...$` commands in running text, or within `\[...\]` or the `equation` environment for displayed equations).
+Sub/superscripts that are physical variables should be italic, whereas those that are labels should be roman (e.g.\ $C_p$, $T_\mathrm{eff}$).
+If the subscripts or superscripts need to be other than italic, they must be coded individually.
 
-Geographers, of course, were at the forefront of
-quantitative diffusion research. Ha¨gerstrand’s (1967)
-early work on understanding innovation diffusion in
-Sweden using simulation modeling is particularly noteworthy
-in this respect. However, already long before the
-‘‘cultural turn’’ in economic geography (Barnes 2001),
-geographers’ interest in formally modeling the diffusion
-of new innovations waned (Gregory 1985). Indeed,
-much of the recent quantitative research into the diffusion
-of new technology has been undertaken by
-economists, sociologists, and business scholars.
+#### Upright Greek characters and the upright partial derivative sign
+Upright lowercase Greek characters can be obtained by inserting the letter `u` in the control code for the character, e.g.\ `\umu` and `\upi` produce $\umu$ (used, for example, in the symbol for the unit microns -- $\umu\mathrm{m}$) and $\upi$ (the ratio of the circumference of a circle to its diameter).
+Similarly, the control code for the upright partial derivative $\upartial$ is `\upartial`.
+Bold lowercase as well as uppercase Greek characters can be obtained by `{\bm \gamma}`, for example, which gives ${\bm \gamma}$, and `{\bm \Gamma}`, which gives ${\bm \Gamma}$.
 
-Theoretical models disagree as to why firms adopt
-innovations at different times. Epidemic models emphasize
-information (Griliches 1957). Certain firms are hypothesized
-to adopt earlier because they come into
-contact with, and learn from, adopters of the new technology
-before others. Economic models, on the other
-hand, predominantly emphasize firm heterogeneity (Ireland
-and Stoneman 1986). Firms adopt technologies at
-different times because they differ with respect to various
-organizational and environmental variables influencing
-the economic returns from adoption (Blackman 1999).
+# Acknowledgement(s) {-}
 
-Either
-way, the strong assumption is that developing countries
-can acquire modern technology innovated in developed
-economies, often at a fraction of the original research
-and development (R&D) costs, thereby leapfrogging
-many decades of technological progress (Teece 2000).
+An unnumbered section, e.g.\ `\section*{Acknowledgements}`, may be used for thanks, etc.\ if required and included _in the non-anonymous version_ before any Notes or References.
 
-The first involves a country’s geographical location.
-Recent empirical work suggests that diffusion is ‘‘geographically
-localized’’ (Globerman, Kokko, and Sjo¨holm
-2000; Keller 2002; Milner 2003) in that a technology
-diffuses faster in a country where it is already more
-widely diffused in neighboring countries. Underlying
-these regional effects are contagion and contact with
-prior users or producers of technology.
+# Disclosure statement {-}
 
-@perkins2011internet
-Two main mechanisms are
-identified in the literature to explain diffusion:
-(1) epidemic-type dynamics whereby contact
-with previous adopters stimulates uptake as po
-tential
-adopters learn about a new innovation;
-and (2) economic-type mechanisms whereby
-potential users adopt a new innovation as it be
-comesmore
-profitable, useful, or valuable, with
-uptake characteristically spreading as costs be
-come
-lower, performance improves, or the po
-tential
-uses of the innovation grow over time.
+An unnumbered section, e.g.\ `\section*{Disclosure statement}`, may be used to declare any potential conflict of interest and included _in the non-anonymous version_ before any Notes or References, after any Acknowledgements and before any Funding information.
 
-To this extent,
-our results contribute to a growing body of
-work that has sought to caution against claims
-about the supposed novelty of the Internet and
-the suggestion that it is somehow different.
+# Funding {-}
 
-@meade2021modelling
-There are numerous agents who would benefit from accurate predictions of introduction times.
-These include: suppliers of network equipment for production planning; suppliers of mobile phones who need to have inventories of
-appropriately configured handsets; international network planners who need to provide sufficient bandwidth for communications
-between different countries; network operators and regulatory authorities.
+An unnumbered section, e.g.\ `\section*{Funding}`, may be used for grant details, etc.\ if required and included _in the non-anonymous version_ before any Notes or References.
 
-Classifying adopters of a technology along a time axis was pioneered by
-Rogers (1962), who categorized adopters into five groups as innovators, early adopters, early majority, late majority and laggards.
+# Notes on contributor(s) {-}
 
-An a priori approach groups similar countries into segments, where similarity is defined using economic, social, or
-political covariates. Examples include Hofstede (2001) and Lee (1990). An a posteriori analysis considers segments based on realized
-market behaviour, for example see Sood et al. (2009). Some studies investigate changing segment membership over time, see Cannon
-and Yaprak (2011).
+An unnumbered section, e.g.\ `\section*{Notes on contributors}`, may be included _in the non-anonymous version_ if required. A photograph may be added if requested.
 
-@fritsch2015new
+# Nomenclature/Notation {-}
 
-@fritsch2015new analyze the spatial diffusion of laser technology research in West Germany from
-1960, when this technology began, until 2005.
+An unnumbered section, e.g.\ `\section*{Nomenclature}` (or `\section*{Notation}`), may be included if required, before any Notes or References.
 
-Hägerstrand (1952, 1967) was the first scholar to intensively analyze spatial diffusion pro
-cesses.
-Based on the empirical observations of the regional spread of process innovations,
-he hypothesized that new technology is first implemented in the ‘center’ – the large agglom
-erations
-– and then moves ‘down’ the spatial hierarchy, the last stop being the ‘periphery’,
-i.e. remote and sparsely populated regions. There are at least two factors that may explain
-this pattern: first, the transfer of tacit knowledge via face-to-face contact and second, the
-absorptive capacity of a region.
+# Notes {-}
 
-# Literature review {#sec2} MAYBE SKIP THIS??
+An unnumbered `Notes` section may be included before the References (if using the `endnotes` package, use the command `\theendnotes` where the notes are to appear, instead of creating a `\section*`).
 
-# Materials and Methods {#sec3}
+# References
 
-@wilson201281 growth function description in p. 86
+## References cited in the text
 
-from R: Asym/(1+exp((xmid-input)/scal))
-using the terms from Wilson k/(1 + e^(t_0 - t)/scal)
+## The list of references
 
-Wilson: k(1 + e^-b(t - t_0))
-So, scal = - 1/b
+References should be listed at the end of the main text in alphabetical order by authors' surnames, then chronologically (earliest first).
+If references have the same author(s), editor(s), etc., arrange by year of publication, with undated works at the end.
+A single-author entry precedes a multi-author entry that begins with the same name.
+If the reference list contains two or more items by the same author(s) in the same year, add a, b, etc. and list them alphabetically by title.
+Successive entries by two or more authors when only the first author is the same are alphabetized by co-authors' surnames.
+If a reference has more than ten named authors, list only the first seven, followed by 'et al.'.
+If a reference has no author or editor, order by title; if a date of publication is impossible to find, use `n.d.' in its place.
 
-The curve is symmetric* around t_0
+The following list shows some sample references prepared in the Taylor & Francis Chicago author-date style.
 
-**minimum R2 = 95%**, see also @grubler1990rise
-
-The literature usually uses the saturation level as the asymptote. I am using the total
-number of websites as we cannot compute a rate.
-
-Moran's I as @ding2010modeling
-**TODO** for t_0,  diffusion speed
-
-## Multi-country analysis {#sec3.1}
-
-## 3.2	Case study approach {#sec3.2}
-
-# Digital technologies and spatial structure: a global view {#sec4}
-
-# The impact of ICT on the US and the UK spatial structures {#sec5}
-
-# Discussion and conclusions {#sec6}
-
-contrary to results from future studies regarding social media [@lengyel2020role], web technologies 
-did not exclusively spread from a central location.
-
-@lengyel2020role
-We see that the OSN
-spread almost exclusively from the original location (the capital Budapest, with an order of magnitude more inhabitants than
-the next size town) to various parts of the country in the early phase of the life-cycle. Later, diffusion became less mono-centric
-and other towns also emerged as spreaders. Our findings support the idea that spreading initially happens to large distances and
-becomes more local over time. This is illustrated and discussed later in Figure 2I.
-
-# References {#references .unnumbered}
+[@Blair1977; @Alb05]

### 2025-01-21 paper/submission2/tranos2025.qmd

--- 2025-01-03 paper/submission1/tranos2023.qmd
@@
 
+<!-- ADD SOMETHING ABOUT MULTIPLE SCALES -->
+
 <!-- geographers lack of interest for diffusion + motivation -->
@@
 innovations diffuse across space and time with the work of
-@hagerstrand1968innovation, which demonstrated the importance of communications 
-as a mechanism for the diffusion of innovations, being the most impactful. 
-It was followed by various empirical studies -- see for instance @ormrod1990 
-and @iso2005. However, the torch of exploring and modelling such diffusion 
-processes had been passed to other disciplines such as economics, business 
-studies and sociology well before the 'cultural turn' of economic geography
-[@perkins2005international]. A potential explanation of the lack of
-geographical studies exploring the diffusion of new and, more specifically for 
-this paper, digital technologies across both space and time can be attributed to 
-the scarcity of relevant and granular enough data, a problem also highlighted 
-by @iso2005 and @kemeny2011international. As @zook2022mapping pointed, digital 
+@hagerstrand1968innovation, which demonstrated the importance of
+communications as a mechanism for the diffusion of innovations, being
+the most impactful. It was followed by various empirical studies -- see
+for instance @ormrod1990 and @iso2005. However, the torch of exploring
+and modelling such diffusion processes had been passed to other
+disciplines such as economics, business studies and sociology well
+before the 'cultural turn' of economic geography
+[@perkins2005international; @ding2010modeling].
+<!-- NEW spatial granularity --> As a consequence, relevant studies
+did not investigate diffusion mechanisms at granular geographical scales to 
+provide local policy insights, for instance, regarding infrastructural d
+evelopment and local technology policies. 
+A potential explanation of the lack of geographical studies exploring the 
+diffusion of new and, more specifically for this paper, digital technologies 
+across both space and time can be attributed to the scarcity of relevant and 
+granular enough data, a problem also highlighted by @iso2005 and
+@kemeny2011international. As @zook2022mapping pointed, digital
 activities are hardly ever captured in official data.
@@
 <!-- contribution -->
-This paper offers such a contribution: an economic geography study revealing
-the importance of spatial mechanisms for the diffusion of an intangible, 
-digital technology that is the Web in the UK at a high level of spatial 
-granularity during the 1996-2012 period.
-<!-- how a new technology that is the Web diffused over space and time  -->
-It does so by employing a novel source of big data and machine 
-learning algorithms to model the *active* engagement with Web during that 
-period and tests how different spatial and temporal mechanisms shaped its 
-diffusion. This paper also exemplifies how the combination of data sources 
-and state-of-the-art research methods which escape the traditional social 
-science domain can offer new lenses to geographical research regarding the 
-understanding of technological diffusion.
+This paper offers such a contribution: an economic geography study
+revealing the importance of spatial mechanisms for the diffusion of an
+intangible, digital technology that is the Web in the UK at a high level
+of spatial granularity 
+<!-- NEW TO JUSTIFY DATA TIME PERIOD-->
+from the very early days until its maturity (1996-2012).
+<!-- NEW DATA CLARIFICATION -->
+It specifically focuses on the diffusion of commercial websites.
+It does so by employing a novel source of big data and machine learning
+algorithms to model the *active* engagement with Web -- that is creating and 
+maintaining instead of just browsing a website -- during that period
+and tests how different spatial and temporal mechanisms shaped its
+diffusion. This paper also exemplifies how the combination of data
+sources and state-of-the-art research methods which escape the
+traditional social science domain can offer new lenses to geographical
+research regarding the understanding of technological diffusion.
 
@@
 about the diffusion of related future technologies. As per
-@leibowicz2016representing, historical studies agree that technologies diffuse 
-differently in terms of times, rates, and geographies and can be driven
-by related policies [@victor1993]. @meade2021modelling highlight
+@leibowicz2016representing, historical studies agree that technologies
+diffuse differently in terms of times, rates, and geographies and can be
+driven by related policies [@victor1993]. @meade2021modelling highlight
 that a variety of actors have a direct interest in gaining such
@@
 places [@leibowicz2016representing]. Despite the spatial heterogeneity
-of such diffusion mechanisms and the policy relevance, there is a scarcity 
-of studies analysing the diffusion of new digital technologies at a detailed
-geographical level because of the lack of detailed enough data and adequate 
-research methods.
+of such diffusion mechanisms and the policy relevance, there is a
+scarcity of studies analysing the diffusion of new digital technologies
+at a detailed geographical level because of the lack of detailed enough
+data and adequate research methods.
 
@@
 This gap becomes even more prominent if we consider the importance of
-technology and its adoption in economic geography and macroeconomic growth 
-theories. Simply put, differences in technological adoption and sophistication 
-may lead to uneven development as new technologies can increase productivity 
-and allow firms to introduce new products [@solow1957technical; 
-@aghion1990model; @kemeny2011international]. Linked to the subject of this 
-paper, the Web -- and the active engagement with this technology that is 
-creating and maintaining instead of just browsing a website -- can be 
-considered as the underpinning technology for the transformative process 
-identified by @capello2024nexus as the *digital service economy*. Importantly, 
-the focus here is not the Web in general, but instead commercial websites.
-After all, diffusion together with invention and innovation are considered the 
-pillars of technological change [@das2022diffusion]. Findings from previous 
-studies illustrated that early engagement with digital technologies such as 
-the Web led to longstanding positive productivity gains for these regions that 
-hosted early adopters [@tranosuk]. Hence, obtaining a robust understanding of 
-the spatial diffusion mechanisms is crucial in mapping future development 
-trajectories.
-
-<!-- motivation: intangible/digital -->
-Another line of motivation has to do with the intangible nature of the Web 
-and the debate about the footloose nature of digital technologies, which
-contrary to several earlier general purpose technologies are 'light in weight' 
-[@coyle1999weightless]. Following the seminal work of @gaspar1998information, 
-which demonstrated that technological improvements in telecommunications lead 
-to increased demand for face-to-face communications and, consequently, 
-enhance the importance of cities, empirical studies from economic geography 
-and urban economics explored whether digital technologies substitute or 
-complement agglomeration externalities and spatial structure 
-[@kolko_death_2000; @ioannides2008effect; @tranos2021ubiquitous].
-With the exception of @sinai2004geography, who focused, among other things, 
-on web content and demonstrated that more such content is available in larger
-markets, this strand of literature has been mostly concerned with the
-spatiality of the Internet's hard and tangible infrastructure
-and shied away from exploring the spatial dimensions of one of the 
-most distinct software layers of the Internet infrastructure, the Web.
-
-<!-- diffusion and adoption -->
-Technological diffusion, which is by definition an aggregated process,
-can be discussed in parallel with individual adoption mechanisms. On the
-one hand, @rogers2010diffusion identifies early adopter of new
-technologies as 'knowledgeable risk takers' and @griliches1957 as
-'profit maximisers' [@ding2010modeling]. Such individual agents are
-rewarded because of their attitude towards new technologies and
-innovations. On the other hand, @perkins2011internet attribute diffusion
-to two processes: (i) epidemic-like mechanisms, which are governed by
-distance, proximity and social interactions, and (ii) economic
-mechanisms as new innovations are adopted by users as they become more
-profitable, valueable and useful.
-<!-- The criticisms are mainly based on the fact that although the approach -->
-<!-- gives an idea of aggregate (industry or household) behaviour, it does -->
-<!-- not focus on the individual's (firm or household) adoption process -->
+technology and its adoption in economic geography and macroeconomic
+growth theories. Simply put, differences in technological adoption and
+sophistication may lead to uneven development as new technologies can
+increase productivity and allow firms to introduce new products
+[@solow1957technical; @aghion1990model; @kemeny2011international].
+Linked to the subject of this paper, the Web and, more specifically, 
+commercial websites can be considered as the underpinning
+technology for the transformative process identified by
+@capello2024nexus as the *digital service economy*. Importantly, the
+focus here is not the Web in general, but instead commercial websites.
+After all, diffusion together with invention and innovation are
+considered the pillars of technological change [@das2022diffusion].
+Findings from previous studies illustrated that early engagement with
+digital technologies such as the Web led to longstanding positive
+productivity gains for these regions that hosted early adopters
+[@tranosuk]. Hence, obtaining a robust understanding of the spatial
+diffusion mechanisms is crucial in mapping future development trajectories.
+
+<!-- NEW motivation: intangible/digital -->
+Another line of motivation has to do with the very nature of the Web as a 
+digital technology and how it evolved over time. The digital geographies field 
+has debunked premature claims about the Internet's placeless-ness by mapping 
+various of its facets and technological layers [@tranos2013death], but the lack 
+of relevant long time-series data prevented researchers from observing digital
+phenomena overtime. Most of these studies focused on its hard infrastructure 
+[e.g. @malecki2002economic; @moss2000internet], on Internet subscribers 
+[e.g. @blank2018local], or on social media users [e.g. @crampton2013beyond].
+Exceptions include early work from @zook2000web, @zook2000web and 
+@moss1997tracking, which utilised domain name registration data and use it to
+map and illustrate the spatial footprint of the web content. Pivotal as these 
+contributions have been in understanding the economic geographies of the 
+Internet, they were based on cross-sectional, often spatially detailed, mapping 
+and on shallow time comparisons (e.g. 1993-1998) and did not observe the 
+evolution and diffusion of the Web until its maturity. As we know from the 
+relevant literature, different technologies diffuse at different rates and 
+there is no theory that can explain these differences 
+[@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
+Web among firms can offer useful insights.
 
@@
 [@ons2018], the web included mostly static webpages, there were no
-social media and web browsing involved exclusively a desktop PCs
-as there were no smartphones [@tranosuk]. Hence, it is fair to say that
-the study period captures the very early stages of the diffusion of a
-new technology that is the Web until its maturity. The former is a key 
-point in the lifecycle of a new technology. Firstly, during this period new 
-technologies are expensive, crude and imperfect [@rosenberg1994exploring; 
-@wilson201281]. A simple comparison between Web 1.0 and Web 3.0 applications 
-clearly illustrates this [@tranos2020social]: while a static website 
-facilitates one-way information dissemination, a platform like `github` 
-enables cooperation between users and the creation of new information, 
-meaning, and even knowledge [@faraj2016special; @barassi2012does]. 
-During this period the performance of a new technology is the main 
-attraction and not the cost to access and use it [@wilson2011lessons]. 
-There is a broader theoretical discussion in the literature about 
-the early adoption motives. As summarised by @perkins2005international, 
-on one hand, epidemic models highlight the role of interpersonal 
-contacts as a way for new technologies to diffuse. On the other hand, 
-economic models underline the importance of heterogeneity. Different firms 
-have different structures and business plans, which define the potential
+social media and web browsing involved exclusively a desktop PCs as
+there were no smartphones [@tranosuk]. Hence, it is fair to say that the
+study period captures the very early stages of the diffusion of a new
+technology that is the Web until its maturity. The former is a key point
+in the lifecycle of a new technology. Firstly, during this period new
+technologies are expensive, crude and imperfect
+[@rosenberg1994exploring; @wilson201281]. A simple comparison between
+Web 1.0 and Web 3.0 applications clearly illustrates this
+[@tranos2020social]: while a static website facilitates one-way
+information dissemination, a platform like `github` enables cooperation
+between users and the creation of new information, meaning, and even
+knowledge [@faraj2016special; @barassi2012does]. During this period the
+performance of a new technology is the main attraction and not the cost
+to access and use it [@wilson2011lessons]. There is a broader
+theoretical discussion in the literature about the early adoption
+motives. As summarised by @perkins2005international, on one hand,
+epidemic models highlight the role of interpersonal contacts as a way
+for new technologies to diffuse. On the other hand, economic models
+underline the importance of heterogeneity. Different firms have
+different structures and business plans, which define the potential
 economic returns of the adoption of a new technology and, therefore, the
@@
 argument is even more relevant when the focus is on digital technologies
-because of the commonly found lag between investment and economic returns as 
-reflected in the Solow paradox [@acemoglu2014return; 
+because of the commonly found lag between investment and economic
+returns as reflected in the Solow paradox [@acemoglu2014return;
 @brynjolfsson2018artificial].
@@
 <!-- data -->
-The data used here depict the active engagement with the Web as they contain 
-geolocated and time-stamped counts of commercial websites from the UK. Instead 
-of adopting metrics reflecting the passive engagement with the Internet such as 
-Internet adoption and speeds [e.g. @blank2018local; @destefano2022fuzzy], this 
-paper observes yearly counts of geolocated websites, which reflect the creation 
-and maintenance of websites. It needs to be highlighted here that although 
+The data used here depict the active engagement with the Web as they
+contain geolocated and time-stamped counts of commercial websites from
+the UK. Instead of adopting metrics reflecting the passive engagement
+with the Internet such as Internet adoption and speeds [e.g.
+@blank2018local; @destefano2022fuzzy], this paper observes yearly counts
+of geolocated, commercial websites, which reflect the creation and 
+maintenance of such websites. It needs to be highlighted here that although 
 there are various types of underpinning web technologies -- namely markup,
-stylesheet, client- and server-side coding languages as well as different
-web browsers and other related programming language -- the focus here
-is on the diffusion of the Web in its totality. For a more detailed and 
+stylesheet, client- and server-side coding languages as well as
+different web browsers and other related programming language -- the
+focus here is on the diffusion of the Web as an overarching technology
+instead of more specific web technologies. For a such a detailed and 
 cross-country technological diffusion study, see @PAPAGIANNIDIS2015308.
@@
 <!-- description of the methods and results -->
+<!-- AMMEND I AFTER CHANGE THE RESULTS -->
 To analyse these data, a diverse set of methods are employed. Firstly,
-cumulative adoption curves are estimated for the UK as a whole and for the 
-Local Authority Districts (LAD), illustrating, for the first time, such 
-S-shaped diffusion patterns for local areas. This system level analysis also 
-reveals insights about the volatility of the adoption of the Web. Then, 
-exploratory spatial data analysis offers some first indications of the 
-importance of the different spatial diffusion mechanisms in shaping the 
-diffusion of the Web in the UK. Importantly, the analysis adopts two distinct 
-geographical scales: the LAD and the most detailed scale of UK census areas, 
-the Output Areas (OA). Lastly, a novel methodological framework based on 
-Machine Learning (ML) algorithms is developed to test the 
-role of such mechanisms at these two scales. The results reveal, for the first 
-time, the importance of different spatial diffusion mechanisms in the diffusion 
-of an intangible, digital technology at very granular geographical scales.
+cumulative adoption curves are estimated for the UK as a whole and for
+the Local Authority Districts (LAD), illustrating, for the first time,
+such S-shaped diffusion patterns for local areas. This system level
+analysis also reveals insights about the volatility of the adoption of
+the Web. Then, exploratory spatial data analysis offers some first
+indications of the importance of the different spatial diffusion
+mechanisms in shaping the diffusion of the Web in the UK. Importantly,
+the analysis adopts two distinct geographical scales: the LAD and the
+most detailed scale of UK census areas, the Output Areas (OA). Lastly, a
+novel methodological framework based on Machine Learning (ML) algorithms
+is developed to test the role of such mechanisms at these two scales.
+The results reveal, for the first time, the importance of different
+spatial diffusion mechanisms in the diffusion of an intangible, digital
+technology at very granular geographical scales.
 
 <!-- contents -->
-The structure of the paper goes as follows. @sec-litreview reviews the key
-concepts in the diffusion literature as well as relevant empirical studies. 
-Then, @sec-datamethods describes the data and methods used in this paper. 
-The results of the different stages of the analysis are presented in 
-@sec-results. The discussion of the results and the conclusions of the paper
-are offered in @sec-conclusions.
-
-<!-- @grubler1990rise Later Hagerstrand conceptualized physical "barrier" -->
-<!-- effects like lakes or uninhabited areas, which, in addition to distance, -->
-<!-- act as further retarding effects on diffusion. These are formalized in -->
-<!-- the form of "zero" or "half" contact multiplicators on the (distance -->
-<!-- decaying) message flows. -->
-
-<!-- @grubler1990rise With respect to the formalization of the communication -->
-<!-- flows Hagerstrand defines a "mean information field" (MIF), in which the -->
-<!-- probability of communication is a negative function of distance between -->
-<!-- individuals -->
-
-# Literature review {#sec-litreview}
-
-<!-- key concepts -- copied from help.Rmd -->
-Geographical diffusion is a synthesis of different processes. On the one
-hand, a purely spatial or, in other words, contagious processes can be 
-identified. Adjacency and, more broadly, distance are the key drivers of
-diffusion. This perspective draws similarities with epidemics:
-innovations just like pathogens spread because of contagion and,
-consequently, proximity and exposure [@hivner2003facilitating]. On the
-other hand, there is a hierarchical processes. According to this, instead of 
-a horizontal distance-based diffusion mechanism, the top-down hierarchy
-of urban systems shapes technological diffusion. In reality, the synthesis 
-of these two processes better reflects how new technologies diffuse over 
-space [@morrill2020spatial].
-
-These ideas were firstly introduced by Torsten Hägerstrand and his
-thesis entitled 'Innovation Diffusion as a Spatial Process'
-[@hagerstrand1968innovation]. Hägerstrand was the first to identify
-diffusion as a geographical process. The starting point was the idea
-that diffusion is based on passing information through social networks,
-which themselves tend to be defined by geography. Hence, he identified
-a 'neighbourhood' mechanism of how information, and consequently,
-innovation diffuse. He used agricultural innovations to test and model
-his ideas using Monte Carlo simulations. Hägerstrand also incorporated
-the role of hierarchy and how some phenomena maybe firstly adopted in
-larger cities and then diffuse to second tier ones. This is a sequential
-instead of a simultaneous process, which resembles the 'lead-lag'
-spatial acceleration effect in market research [@bento2018time;
-@PERES201091]. Hägerstrand is more widely known though for highlighting
-the role time plays in the diffusion of innovations: an early-pioneering
-period, a middle fast accelerating period and a final saturation one
-[@morrill2020spatial].
-
-The temporal dimension was further explored by Everett Rogers and his
-seminal work on 'Diffusion of Innovations' [@rogers2010diffusion].
-Rogers being a sociologist focused not on the diffusion of
-innovations over space and time, but instead on the adoption of new
-technologies and innovations by individuals and the individual
-mechanisms that drive the decisions behind adoption. He identified five
-groups of individuals regarding their adoption speed: innovators, early
-adopters, early majority, late majority and laggards. The key mechanism of
-diffusion and adoption is communication and how knowledge is transferred 
-within a social system. All above approaches agree that diffusion and the
-cumulative adoption of a new technology are represented by an S-shaped 
-line plot [@grubler1990rise].
-
-Schmidt's Law empirically illustrates a similar pattern. *Core* and
-usually highly agglomerated regions is where new technologies are
-invented and commercially deployed [@grubler1990rise].
-This is where the first adopters tend to be based. Then, technologies
-spread to the *rim* and eventually to the *periphery*. Although adoption
-pace might be higher when new technologies finally arrive to the
-periphery, the saturation levels there may never reach the ones in the
-core because of the lack of infrastructure or other necessary
-institutions [@leibowicz2016representing].
-
-@grubler1990rise effectively summarises the three key spatial diffusion 
-mechanisms: (i) the cumulative level of adoption follows an S-shaped
-pattern just like purely temporal models; (ii) diffusion is shaped by a
-hierarchical mechanism in a form of a centrifugal force: from core to
-periphery; and (iii) diffusion is also shaped by distance and proximity. 
-These are the three mechanisms that the empirical analysis in @sec-results 
-investigates.
-
-<!-- previous studies -->
-The remaining of this section reviews empirical studies which analysed
-the diffusion on new technologies over space and time. Although the
-spatial dimension is present in most of the following studies, the level
-of spatial detail is always more coarse than the one adopted in this
-paper. @beardsell1999spatial studied the evolution of the computer
-industry in 317 US metro areas during the 1977-1992 period using
+<!-- AMMEND I AFTER CHANGE THE RESULTS -->
+The structure of the paper goes as follows. @sec-litreview reviews the
+key concepts in the diffusion literature as well as relevant empirical
+studies. Then, @sec-datamethods describes the data and methods used in
+this paper. The results of the different stages of the analysis are
+presented in @sec-results. The discussion of the results and the
+conclusions of the paper are offered in @sec-conclusions.
+
+# The mechanisms that shape the diffusion of a new technology {#sec-litreview}
+
+<!-- The fundamentals of technology diffusion -->
+The diffusion of new technologies and innovations is driven by three
+different mechanisms [@grubler1990rise; @morrill2020spatial]: (i)
+distance and proximity, (ii) urban hierarchies in the form of a
+centrifugal forces as new technologies diffuse from core to periphery,
+and (iii) over time the cumulative level of adoption follows an S-shaped
+pattern.
+<!-- neighbourhood -->
+@hagerstrand1968innovation was the first to identify diffusion as a
+geographical process. The starting point was the idea that diffusion is
+based on passing information through social networks, which themselves
+tend to be geographically bounded. Hence, he identified a
+'neighbourhood' mechanism of how information, and consequently,
+innovations and new technologies diffuse. This perspective draws
+similarities with epidemics: new technologies and innovations, just like
+pathogens, spread because of contagion and, consequently, proximity and
+exposure [@hivner2003facilitating].
+
+<!-- hierarchy -->
+@hagerstrand1968innovation also emphasised the role of the top-down
+hierarchy of urban systems and highlighted that new technologies and
+innovations are firstly adopted in larger cities and then diffuse to
+second tier ones. This is a sequential instead of a simultaneous
+process, which resembles the 'lead-lag' spatial acceleration effect in
+market research [@bento2018time; @PERES201091]. Schmidt's Law
+empirically illustrates a similar pattern [@grubler1990rise]. Core and
+highly agglomerated regions is where new technologies are invented and
+commercially deployed. This is where the first adopters tend to be
+based. Then, technologies spread to the rim and eventually to the
+periphery. Although adoption pace might be higher when new technologies
+finally arrive to the periphery, the saturation levels there may never
+reach the ones in the core because of the lack of infrastructure or
+other necessary institutions [@leibowicz2016representing].
+
+<!-- time -->
+Diffusion is naturally a dynamic process and Hägerstrand recognised the
+role time plays in the diffusion of innovations and new technologies: an
+early-pioneering period, a middle fast accelerating period and a final
+saturation one [@morrill2020spatial]. This was further explored by
+@rogers2010diffusion, who identified five types of adopters of new
+technologies and innovations: innovators, early adopters, early
+majority, late majority and laggards. Hence, plotting the cumulative
+adoption of a new technology over time often leads to an S-shaped line
+plot [@grubler1990rise].
+
+Although the above processes have been studied in innovation studies,
+the distancing of economic geographers from research questions about the
+diffusion of new technologies and innovations resulted to a knowledge gap
+regarding the explanatory power of these mechanisms at granular, local
+scales [@ding2010modeling]. As discussed in @sec-introduction, such 
+spatially explicit knowledge can directly inform local technology
+and infrastructure policies. Indeed, while the spatial dimension is
+present in the following studies, the geographical level of analysis is
+often too coarse to ascertain the role of the above mechanisms at the
+local scale. @beardsell1999spatial studied the evolution of the computer
+industry in 317 US metropolitan areas during the 1977-1992 period using
 employment data. Their analysis indicated that the relative size
@@
 study, @bednarz2020pulled focused on wind turbines and modelled their
-spatial diffusion across 402 German regions during 1970-2015. Their
-key finding is that local demand than local supply was the main driving
+spatial diffusion across 402 German regions during 1970-2015. Their key
+finding is that local demand than local supply was the main driving
 factor. @haller2011determinants employed firm panel data for Ireland
-during 2003-2005 and illustrated that when firms are located in a
-region or industry with high share of firms having a website, they are
-more likely to have a website too.
-
-At a global scale @perkins2005international explored whether the
-diffusion rate of new technologies is driven by a latecomer advantage
-and the engagement with the global economy via foreign direct
-investments and trade. Their results illustrate that indeed latecomers
-and developing countries experience diffusion of new technologies more
-rapidly than early adopters and developed countries. At the same scale,
-@perkins2011internet explored whether the adoption of previous
-communication technologies that is mail, telegrams and telephones were
-shaped by similar socioeconomic factors as the Internet. Their results
-indicated common patterns regarding the drivers behind the adoption of
-different communication technologies.
-
-Turning to studies that share more technological and scalar similarities
-with this paper, @ding2010modeling modelled the spatial diffusion of
-mobile telecommunications across regions in China. Their analysis
-indicated that socioeconomic characteristics are important determinants
-of the timing, speed and the level of mobile diffusion within China.
-Using data from a Hungarian online social network, @lengyel2020role
-analysed its adoption and the churn at a very granular spatial level.
-Their results were in agreement with the early theoretical and empirical
-contributions reviewed here: assortativity, urban scaling and distance
-are the key drivers of spatial diffusion. @bakher2013diffusion compared 
-different the growth curves as a means to model diffusion of online 
-shopping in Australia during the 1998-2009 period. At a global scale
-@PAPAGIANNIDIS2015308 modelled the diffusion of different web
-technologies and practices across countries. Using similar but less extensive 
-data as the one used here, their analysis illustrated how the diffusion 
-of different web technologies and practices follow an S-shaped pattern 
-as well as the different diffusion rates of the different technologies 
-and practices.
-
-All in all, this paper tests whether the three distinct spatial diffusion 
-mechanisms shaped the diffusion of the Web in the UK at very granular
-spatial scales from the Web's early stages onwards. The next section discusses
-the empirical framework.
-
-<!-- @fritsch2015new analyze the spatial diffusion of laser technology -->
-<!-- research in West Germany from 1960, when this technology began, until -->
-<!-- 2005. -->
-<!-- in one region  -->
-<!-- @leibowicz2016representing for energy There is no generally accepted -->
-<!-- theory that explains diffusion rate heterogeneity across technologies, -->
-<!-- but several factors are considered important. Greater unit scale and -->
-<!-- larger market size contribute to slower diffusion. Requirements for -->
-<!-- interrelated technologies or complex infrastructures also hinder the -->
-<!-- diffusion process (Grubler, 2012). -->
-
-<!-- Mobile phones benefited from early deployment in recreational boats and -->
-<!-- automobiles, where the traditional competitor was not a viable option. -->
-<!-- In the early stages of diffusion, performance is a more important driver -->
-<!-- of adoption than cost competitiveness. Typically, significant cost -->
-<!-- reductions only occur once the technology reaches a deployment level -->
-<!-- capable of supporting standardization and mass production (Wilson, -->
-<!-- 2012). -->
-
-<!-- @leibowicz2016representing Empirical evidence supports the validity of -->
-<!-- Schmidt's Law over a wide range of technologies, time periods, and -->
-<!-- geographical contexts. A recent meta-analysis of technology up-scaling -->
-<!-- found that diffusion accelerated moving from the core to the rim and -->
-<!-- periphery for technologies as diverse as natural gas power, oil -->
-<!-- refineries, and automobiles (Wilson, 2009). One historical example that -->
-<!-- conforms particularly well to Schmidt's Law is the diffusion of coal -->
-<!-- power in Europe (Grubler, 2012). England emerged as the core region for -->
-<!-- coal power because it had legal and economic institutions that -->
-<!-- incentivized scientific pursuits, domestic coal reserves, and a clear -->
-
-<!-- industrial motivation to replace water power with coal. ... -->
-
-<!-- country level, not sure if needed. -->
-
-<!-- @bento2018time explore What determines the duration of formative phases -->
-<!-- for energy innovations in different markets? We are interested both in -->
-<!-- initial markets (also: core, lead, first mover, early adopter) where -->
-<!-- formative phases prepare technologies for mass commercialization, and in -->
-<!-- follower markets (also: periphery, lag, late adopter) where accelerated -->
-<!-- formative phases may benefit from diffusion and spillovers. -->
+during 2003-2005 and illustrated that when firms are located in a region
+or industry with high share of firms having a website, they are more
+likely to have a website too. At a global scale @perkins2011internet
+explored whether the adoption of previous communication technologies
+that is mail, telegrams and telephones were shaped by similar
+socioeconomic factors as the Internet. Their results indicated common
+patterns regarding the drivers behind the adoption of different
+communication technologies. @feldman2015rknowledge focused on inventions
+and used patent data to map how the the rDNA technology diffused among
+inventors across 366 US metropolitan areas. @ding2010modeling modelled
+the spatial diffusion of mobile telecommunications across the highest
+level of administrative division in China (29 regions in total). Their
+analysis indicated that socioeconomic characteristics are important
+determinants of the timing, speed and the level of mobile diffusion
+within China. @bakher2013diffusion compared different the growth curves
+as a means to model the diffusion of online shopping in Australia during
+the 1998-2009 period, but they did not consider any spatial mechanisms.
+At a global scale @PAPAGIANNIDIS2015308 modelled the diffusion of
+different web technologies and practices across countries. Using similar
+but less extensive data as the one used here, their analysis illustrated
+how the diffusion of different web technologies and practices follows an
+S-shaped pattern as well as the different diffusion rates of the
+different technologies and practices. The most detailed consideration of
+the spatial footprint of the diffusion of a new technology comes from
+@lengyel2020role, who analysed the adoption and churn of a Hungarian
+online social network. Their results were in agreement with the early
+theoretical and empirical contributions reviewed here: assortativity,
+urban scaling and distance are the key drivers of spatial diffusion.
+Based on the above, the main empirical hypothesis states:
+
+> H~1~: The three diffusion mechanisms namely distance, urban hierarchy
+and the the S-shaped pattern of the cummulative level of adoption shaped
+the diffusion of the commercial Web in the UK even at granular geographical 
+scales.
+
+<!-- EARLY/LATE adoption-->
+Understanding the diffusion of a new technology in multiple, detailed spatial
+scales is of interest to economic geography as potential local economic 
+benefits may occur because of concentration of early adopters or through 
+technological leapfrogging. On the one hand, early adopters are often rewarded 
+because of their attitudes towards new technologies and innovations. 
+@rogers2010diffusion identifies them as 'knowledgeable risk takers' and 
+@griliches1957 as 'profit maximisers' [@ding2010modeling]. On the other hand,
+latecomers can adopt and benefit from new technologies that have been developed
+elsewhere without incurring the hefty initial R&D costs [@teece2008firm]. At a 
+global scale @perkins2005international explored whether the diffusion rate of 
+new technologies is driven by a latecomer advantage and the engagement with the 
+global economy via foreign direct investments and trade. Their results 
+illustrate that indeed latecomers and developing countries experience diffusion 
+of new technologies more rapidly than early adopters and developed countries. 
+At a regional scale, previous research highlighted the long term and 
+sustained productivity benefit of the early adopters of web technologies in 
+the UK [@tranosuk]. Although the early adoption and leapfrogging literature has 
+not paid much attention on cities and regions [@yu2018sustainability], the 
+potential local economic benefits justify the need to investigate the
+dynamics of the web diffusion in different scales in the UK. Hence, the second 
+empirical hypothesis states:
+
+> H~2~: The diffusion of the commecrial Web is shaped by both early adoption 
+and late comer effects.
+
+These are the two hypotheses are empirically tested in @sec-results using the 
+data and methods described in the next section.
+
+<!-- Technological diffusion, which is by definition an aggregated process, -->
+<!-- can be discussed in parallel with individual adoption mechanisms. On the -->
+<!-- one hand,  -->
+<!-- On the other hand, @perkins2011internet attribute diffusion -->
+<!-- to two processes: (i) epidemic-like mechanisms, which are governed by -->
+<!-- distance, proximity and social interactions, and (ii) economic -->
+<!-- mechanisms as new innovations are adopted by users as they become more -->
+<!-- profitable, valueable and useful. -->
 
@@
 
-<!-- main metric -->
-To capture the diffusion of the Web, a website density metric is developed 
-for two different geographical scales: the Local Authority Districts (LAD) 
-and the Output Areas (OA). The former is an administrative unit and there 
-are 374 LAD in the UK. The latter is the smallest census-based geographical 
-unit and there are c. 230,000 of them in the UK. This methodological choice 
-will allow the mapping of the diffusion of web technologies and the assessment 
-of the diffusion mechanisms at these two very different spatial scales.
-
-<!-- web data -->
-The counts of websites at these scales are calculated using data from the 
-Internet Archive[^2] and, specifically, the JISC UK Web Domain Dataset 
-[@ukwebarchive]. The Internet Archive is one of the most complete and oldest 
-archive of webpages in the world operating since 1996 
-[@ainsworth2011much; @holzmann2016dawn]. It is a web crawler, which discovers 
-webpages by following the hyperlinks of every webpage it archives.
-This dataset, which is curated by the British Library, contains all
-the archived webpages from the UK country code top level domain (ccTLD - .uk) 
-from the 1996–2012 period. In essence, this is a long list of 2.5 billion URLs 
-of archived webpages including also the archival timestamp.
+<!-- main data -->
+To capture the diffusion of the commercial Web, a website density metric is
+developed for two different geographical scales that are discussed later in 
+this section. The main data are counts of commercial websites that are 
+calculated using data from the Internet Archive[^2] and, specifically, the 
+JISC UK Web Domain Dataset [@ukwebarchive]. The Internet Archive is one of the 
+most complete and oldest archive of webpages in the world operating since
+1996 [@ainsworth2011much; @holzmann2016dawn]. It is a web crawler, which
+discovers webpages by following the hyperlinks of every webpage it
+archives. This dataset, which is curated by the British Library,
+contains all the archived webpages from the UK country code top level
+domain (ccTLD - .uk) from the 1996–2012 period. In essence, this is a
+long list of 2.5 billion URLs of archived webpages including also the
+archival timestamp.
 
@@
 
-Instead of using the whole .uk ccTLD, this paper focuses on its commercial 
-subset, the .co.uk second level domain (SLD). This choices decrease the 
-heterogeneity of the web data as such commercial websites have specific 
-aims: they are used to diffuse information, support online transactions 
-and share opinions [@THELWALL2000441; @blazquez2018big]. 
-Although a UK company can adopt a generic TLD such as .com and these 
-cases escape the data used here, such omissions should not affect the 
-validity of the results given the popularity of the .uk ccTLD [@tranosuk]: 
-UK consumers prefer to visit a .uk website when they are searching for 
-products or services [@hope]; and anecdotal evidence indicates that during 
-the first half of 2000, three .co.uk domains were registered every minute 
-[@oecd_coms]. Importantly, previous studies illustrated that .co.uk is the 
-most popular UK SLD [@tranosuk].
-
-The text from these webpages was scanned using a regular expression (regex) to
-identify strings of text which resemble UK postcodes and one fifth of them 
-included a mention to a postcode [@BL2013geo]. This information allows the 
-geolocation of the data and the creation of the LAD and OA counts.
+Instead of using the whole .uk ccTLD, this paper focuses on its
+commercial subset, the .co.uk second level domain (SLD). This choice
+is aligned with the topic of the paper -- the diffusion of the Web for 
+commercial functions -- and decreases the heterogeneity of the web data as such 
+commercial websites have specific aims: they are used to diffuse information, 
+support online transactions and share opinions [@THELWALL2000441; 
+@blazquez2018big]. Although a UK company can adopt a generic TLD such as .com 
+and these cases escape the data used here, such omissions should not affect the
+validity of the results given the popularity of the .uk ccTLD
+[@tranosuk]: UK consumers prefer to visit a .uk website when they are
+searching for products or services [@hope]; and anecdotal evidence
+indicates that during the first half of 2000, three .co.uk domains were
+registered every minute [@oecd_coms]. Importantly, previous studies
+illustrated that .co.uk is the most popular UK SLD [@tranosuk]. Henceforward 
+any reference to websites will refer to commercial -- .co.uk -- websites.
+
+The text from these webpages was scanned using a regular expression
+(regex) to identify strings of text which resemble UK postcodes and one
+fifth of them included a mention to a postcode [@BL2013geo]. This
+information allows the geolocation of the data and the creation of the
+LAD and OA counts. To prevent false positives, the resulted postcodes
+were matched against official postcode lookup files [@ons].
 
 The data cleaning process includes an aggregation step, through which
-the archived webpages are aggregated to the parent websites. This website 
-reconstruction process lend itself to the construction of *website* instead 
-of *webpage* density metrics. Websites tend to represent specific organisations 
-or entities, and, arguably, are more meaningful observational units than 
-webpages, which can ignore the upstream dependency of the website they belong 
-to. Based on the following example, all three webpages are part of the 
-same website ([http://www.website.co.uk](http://www.website.co.uk)). Only 
-the overall website and not the nested webpages are considered for the counts 
-as otherwise the density metrics would have been biased towards large, place 
-specific, websites.
-
-  - [http://www.website.co.uk/webpage_a](http://www.website.co.uk/webpage_a) B15 2TT
-
-  - [http://www.website.co.uk/webpage_b](http://www.website.co.uk/webpage_b) BS8 1TH
-
-  - [http://www.website.co.uk/webpage_c](http://www.website.co.uk/webpage_c) B15 2TT
-
-\noindent What is challenging is that this aggregation approach,
-which has been used elsewhere [@tranosuk; @shoreditch] may lead to websites with 
-multiple postcodes. As per the above example, 
-[www.website.co.uk](http://www.website.co.uk) includes two unique postcodes: 
-B15 2TT and BS8 1SS. The distribution of the number of postcodes per website for 
-2000 is presented in Table \ref{f2000}, which illustrates a wide range. 
-At the left end of the distribution, there are websites anchored to a unique 
-location (72% of all the reconstructed websites in 2000), which may represent a 
-small company with a single trading location. At the right end, there are 
-websites with thousands of different postcodes. Considering the time period of 
-the analysis, such cases can represent directories which used to be popular 
-in the pre-search engines early times of the commercial Internet as well as real 
-estate websites [@tranosuk].
-
-The analysis is firstly conducted on websites, which only contain one unique 
-postcode. As a robustness check, the analysis is replicated for an extended 
-subset of websites, which include up to 10 unique postcodes to capture 
-commercial websites which point to multiple locations. They are geolocated 
-by equally attaching them to these locations. This extended sample includes 94 
-percent of all the archived websites in 2000.
+the archived webpages are aggregated to the parent websites. This
+website reconstruction process lend itself to the construction of
+*website* instead of *webpage* density metrics. Websites tend to
+represent specific organisations or entities, and, arguably, are more
+meaningful observational units than webpages, which can ignore the
+upstream dependency of the website they belong to. Based on the
+following example, all three webpages are part of the same website
+(<http://www.website.co.uk>). Only the overall website and not the
+nested webpages are considered for the counts as otherwise the density
+metrics would have been biased towards large, place specific, websites.
+It is important to note here that websites commonly contain place references, 
+with previous research estimating this share to be around 70% 
+[@hill2009georeferencing].
+
+-   <http://www.website.co.uk/webpage_a> B15 2TT
+
+-   <http://www.website.co.uk/webpage_b> BS8 1TH
+
+-   <http://www.website.co.uk/webpage_c> B15 2TT
+
+\noindent What is challenging is that this aggregation approach, which
+has been used elsewhere [@tranosuk; @shoreditch] may lead to websites
+with multiple postcodes. As per the above example,
+[www.website.co.uk](http://www.website.co.uk) includes two unique
+postcodes: B15 2TT and BS8 1SS. The distribution of the number of
+postcodes per website for 2000 is presented in Table \ref{f2000}, which
+illustrates a wide range. At the left end of the distribution, there are
+websites anchored to a unique location (72% of all the reconstructed
+websites in 2000), which may represent a small company with a single
+trading location. At the right end, there are websites with thousands of
+different postcodes. Considering the time period of the analysis, such
+cases can represent directories which used to be popular in the
+pre-search engines early times of the commercial Internet as well as
+real estate websites [@tranosuk]. Previous research which used the same 
+methodology for a small subset of these data illustrates the robustness of the 
+above approach: its results matched in terms of accuracy past rounds of 
+research based on extensive qualitative work, interviews and observational 
+studies for the same study area [@shoreditch].
+
+The analysis is firstly conducted on websites, which only contain one
+unique postcode. As a robustness check, the analysis is replicated for
+an extended subset of websites, which include up to 10 unique postcodes
+to capture commercial websites which point to multiple locations. They
+are geolocated by equally attaching them to these locations. This
+extended sample includes 94 percent of all the archived websites in
+2000.
 
@@
 Another data cleaning step deals with some extreme outliers. Figure
-\ref{correct} plots the website counts for the top 1000 postcodes. Some obvious 
-outliers can be observed for the 2002-2006 period for a handful of postcodes, 
-which can be attributed to link farms [@BL2013links]. The website counts for 
-these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1 6GL, M28 2SL, DE21 7BF), which 
-in 2004 or 2005 had more than 1000 websites pointing to them, were replaced 
-with predicted values based on a simple panel regression model with postcode 
-fixed effects and yearly dummy variables. These postcodes refer to a residential 
-area with a small park in South London, residential areas outside Warwick and 
-Manchester, a rural area in the south border of England and Wales and small 
-business parks in Crew and Derby. Hence, it is difficult to believe that these
-postcodes genuinely hosted such a large amounts of websites, which were later 
-extensively declined. To put the magnitude of the data imputation into 
-perspective, this process affected 6 out of the 557,808 postcodes included in 
-the data. As the @sec-results illustrates, this led to a small increase of the 
-model predictive power.
+\ref{correct} plots the website counts for the top 1000 postcodes. Some
+obvious outliers can be observed for the 2002-2006 period for a handful
+of postcodes, which can be attributed to link farms [@BL2013links]. The
+website counts for these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1
+6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1000
+websites pointing to them, were replaced with predicted values based on
+a simple panel regression model with postcode fixed effects and yearly
+dummy variables. These postcodes refer to a residential area with a
+small park in South London, residential areas outside Warwick and
+Manchester, a rural area in the south border of England and Wales and
+small business parks in Crew and Derby. Hence, it is difficult to
+believe that these postcodes genuinely hosted such a large amounts of
+websites, which were later extensively declined. To put the magnitude of
+the data imputation into perspective, this process affected 6 out of the
+557,808 postcodes included in the data. As the @sec-results illustrates,
+this led to a small increase of the model predictive power.
 
@@
 
-To create the website density metric, the yearly website counts at the LAD level 
-are standardised by the number of firms in LAD to avoid biases associated with 
-LAD hosting a large number of firms. Given that there is no such readily 
-available statistic for the OA, the actual OA level counts are used. However, 
-because of the the consistent spatial definition of OA (they host 40-250 
-households),[^4] website counts in OA are interpreted as a density metric too.
-
-[^4]: [https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies](https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies)
+<!-- Multi-scale -->
+To empirically test the two hypotheses stated in @sec-litreview, τhe website 
+density metric was calculated for two different scales the Local Authority
+Districts (LAD) and the Output Areas (OA). The former is an administrative unit 
+and there are 374 LAD in the UK. The latter is the smallest census-based 
+geographical unit and there are c. 230,000 of them in the UK. To create the 
+website density metric, the yearly website counts at the LAD level are 
+standardised by the number of firms in LAD to avoid biases associated with LAD 
+hosting a large number of firms. Given that there is no such readily available 
+statistic for the OA, the actual OA level counts are used. However, because of 
+the the consistent spatial definition of OA (they host 40-250 households),[^3] 
+website counts in OA are interpreted as a density metric too.
+
+[^3]: <https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies>
 
 <!-- system-level analysis -->
-These website density metrics are used for the three different stages of the 
-analysis. Firstly, a system-level analysis explores how the diffusion of web 
-technologies in the UK fits with the well-established S-curve of cumulative 
-adoption. To do so, the following logistic function (Equation \ref{eq:s}) is 
-estimated for the whole of the UK and for each LAD separately:
+These website density metrics are used for the three different stages of
+the analysis. Firstly, a system-level analysis explores how the
+diffusion of web technologies in the UK fits with the well-established
+S-curve of cumulative adoption. To do so, the following logistic
+function (Equation \ref{eq:s}) is estimated for the whole of the UK and
+for each LAD separately:
 
@@
 
-\noindent $k$ is the asymptote or, in other words, the saturation level, $b$ 
-the overall growth rate, and $t_{0}$ the *inflection point* of maximum growth 
-at $k/2$, where the logistic function is symmetrical [@wilson201281]. 
-The $t_0$ of each LAD is compared against the $t_0$ of the UK to delineate 
-whether a LAD reached that point faster or slower than the country average. 
-Importantly, an accuracy criterion was imposed and only LAD with $R^2 > 0.9$ 
-were included in this analysis. To estimate Equation \ref{eq:s} a self-starting 
-logistic growth model was employed using the `nls` and `SSlogis` functions in 
-`R`. This logistic growth model depicts the initial period of slow diffusion of 
-a new technology, which is followed by the fast, exponential growth stage to
-then eventually slow down and saturate [@wilson201281; @grubler1999dynamics].
+\noindent $k$ is the asymptote or, in other words, the saturation level,
+$b$ the overall growth rate, and $t_{0}$ the *inflection point* of
+maximum growth at $k/2$, where the logistic function is symmetrical
+[@wilson201281]. The $t_0$ of each LAD is compared against the $t_0$ of
+the UK to delineate whether a LAD reached that point faster or slower
+than the country average. Importantly, an accuracy criterion was imposed
+and only LAD with $R^2 > 0.9$ were included in this analysis. To
+estimate Equation \ref{eq:s} a self-starting logistic growth model was
+employed using the `nls` and `SSlogis` functions in `R`. This logistic
+growth model depicts the initial period of slow diffusion of a new
+technology, which is followed by the fast, exponential growth stage to
+then eventually slow down and saturate [@wilson201281;
+@grubler1999dynamics].
 
 <!-- **minimum R2 = 95%** in @wilson201281, see also @grubler1990rise -->
+
 <!-- The literature usually uses the saturation level as the asymptote. I am -->
+
 <!-- using the total number of websites as we cannot compute a rate. -->
 
-The system-level analysis also focuses on the volatility of the web adoption to 
-depict places with high concentration of early adopters and, equally, 
-latecomers. To do so, the change of the ranking of the UK LAD over time is 
-plotted and discussed. Both the S-curve and the volatility analysis focus only 
-on LAD as the very large number of OA would have made such analysis difficult 
-to visualise and interpret.
+The system-level analysis also focuses on the volatility of the web
+adoption to depict places with high concentration of early adopters and,
+equally, latecomers. To do so, the change of the ranking of the UK LAD
+over time is plotted and discussed. Both the S-curve and the volatility
+analysis focus only on LAD as the very large number of OA would have
+made such analysis difficult to visualise and interpret.
 
@@
 Next, exploratory spatial data analysis depicts whether the two main
-drivers of spatial diffusion -- namely neighbourhood and hierarchy -- underpin 
-the diffusion of the Web in the UK. To capture the former and following 
-@ding2010modeling the Moran's I and the Local Index of Spatial Association 
-(LISA) are estimated for the website density. To address the hierarchy mechanism, 
-the Gini coefficient -- an established metric of inequality -- is 
-calculated. All the above are computed and plotted longitudinally both for the 
-LAD and the OA.
+drivers of spatial diffusion -- namely neighbourhood and hierarchy --
+underpin the diffusion of the Web in the UK. To capture the former and
+following @ding2010modeling the Moran's I and the Local Index of Spatial
+Association (LISA) are estimated for the website density. To address the
+hierarchy mechanism, the Gini coefficient -- an established metric of
+inequality -- is calculated. All the above are computed and plotted
+longitudinally both for the LAD and the OA.
 
@@
 Lastly, a modelling framework is developed to test the above diffusion
-mechanisms. The overarching aim is to build a model that can test how well
-these mechanisms can predict website density:
+mechanisms. The overarching aim is to build a model that can test how
+well these mechanisms can predict website density:
 
@@
 
-\noindent Equation \ref{eq:rf.generic} is estimated using Random Forest (RF).
-This is a popular ML algorithm for both regression and classification
-problems [@biau2012analysis]. It was introduced by @breiman2001random and has 
-become a go-to data science tool. RF can effectively handle skewed distributions 
-and outliers, model non-linear relationships, require minimal hyperparameter 
-tuning, exhibit low sensitivity to these parameters, and have relatively short 
-training times [@Caruana2008; @liaw2002classification; @yan2020using]. These 
-attributes match well with the website density data characteristics including 
-skewness especially for the OA. Also, the large data size (c. 230k data points 
-for each of the 17 years) calls for fast training times. Importantly, RF 
-predictions tend to be more accurate than those from single regression trees and 
-outperform Ordinary Least Squares in out-of-sample predictions, even with 
-moderate-sized training data and a small number of predictors 
-[@mullainathan2017machine; @athey2019machine; @sulaiman2011intelligent; 
-@pourebrahim2019trip; @biau2012analysis].
-
-RF is a tree-based ensemble learning algorithm [@breiman2001random]. It begins 
-by generating random samples of the training data, which are then used to grow 
-regression trees to predict the dependent variable. Data points and predictors 
-are randomly sampled for the different trees. The trees are trained in parallel 
-using their own bootstrapped samples of the training data. A crucial feature of 
-RF is their ability not to overfit, meaning they can generalize well to unseen 
-test data. While each tree may overfit individually, the ensemble of trees does 
-not because the errors of individual trees are averaged, reducing the overall 
-variance and preventing overfitting [@last2002improving]. For regression 
-problems, RF predictions are made by averaging the predictions of all decision 
-trees.
+\noindent Equation \ref{eq:rf.generic} is estimated using Random Forest
+(RF). This is a popular ML algorithm for both regression and
+classification problems [@biau2012analysis]. It was introduced by
+@breiman2001random and has become a go-to data science tool. RF can
+effectively handle skewed distributions and outliers, model non-linear
+relationships, require minimal hyperparameter tuning, exhibit low
+sensitivity to these parameters, and have relatively short training
+times [@Caruana2008; @liaw2002classification; @yan2020using]. These
+attributes match well with the website density data characteristics
+including skewness especially for the OA. Also, the large data size (c.
+230k data points for each of the 17 years) calls for fast training
+times. Importantly, RF predictions tend to be more accurate than those
+from single regression trees and outperform Ordinary Least Squares in
+out-of-sample predictions, even with moderate-sized training data and a
+small number of predictors [@mullainathan2017machine; @athey2019machine;
+@sulaiman2011intelligent; @pourebrahim2019trip; @biau2012analysis].
+
+RF is a tree-based ensemble learning algorithm [@breiman2001random]. It
+begins by generating random samples of the training data, which are then
+used to grow regression trees to predict the dependent variable. Data
+points and predictors are randomly sampled for the different trees. The
+trees are trained in parallel using their own bootstrapped samples of
+the training data. A crucial feature of RF is their ability not to
+overfit, meaning they can generalize well to unseen test data. While
+each tree may overfit individually, the ensemble of trees does not
+because the errors of individual trees are averaged, reducing the
+overall variance and preventing overfitting [@last2002improving]. For
+regression problems, RF predictions are made by averaging the
+predictions of all decision trees.
 
 RF have been widely employed to address regression research problems.
-@pourebrahim2019trip combined a spatial interaction modelling framework with ML 
-algorithms including RF to predict commuting flows in New York City. 
-@sinha2019assessing advocated for adopting spatial ensemble learning approaches, 
-such as RF, to model spatial data with high autocorrelation and heterogeneity. 
-@creditspatial predicted employment density in Los Angeles using spatially 
-explicit RF. @guns2014recommending used RF to build a recommendation system for 
+@pourebrahim2019trip combined a spatial interaction modelling framework
+with ML algorithms including RF to predict commuting flows in New York
+City. @sinha2019assessing advocated for adopting spatial ensemble
+learning approaches, such as RF, to model spatial data with high
+autocorrelation and heterogeneity. @creditspatial predicted employment
+density in Los Angeles using spatially explicit RF.
+@guns2014recommending used RF to build a recommendation system for
 research collaborations. @ren2019predicting trained RF to predict the
-socio-economic status of cities using various online and mobility predictors. 
-@tranos2023using utilised hyperlinks data and RF to make out-of-sample 
-predictions of interregional trade. @zhou2023geography employed such a 
-framework to assess whether key predictors of obesity differ across English 
-cities.
+socio-economic status of cities using various online and mobility
+predictors. @tranos2023using utilised hyperlinks data and RF to make
+out-of-sample predictions of interregional trade. @zhou2023geography
+employed such a framework to assess whether key predictors of obesity
+differ across English cities.
 
@@
 <!-- S-shaped -->
+
 ```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk}Grwoth curve, UK"}
@@
 
-Figure \ref{s_uk} plots the cumulative adoption of the Web in total in the UK 
-during the 1996-2012 period. Not surprisingly, the cumulative adoption is well 
-represented by a curve that resembles the figure S. The vertical 
-line around year 2003 illustrates the point where the modelled cumulative 
-adoption is equal to 50% of the maximum. This $t_0$ *inflection* point signals 
-the maximum adoption speed and is used here to determine whether a UK LAD 
-reached that inflection point earlier or later than the UK average. 
-Specifically, the S curve and the inflection point are estimated for every LAD 
-individually and then compared against the UK average. LAD that reached their 
-inflection point earlier than the country average are labelled as *fast* and 
-the rest as *slow*. Figure \ref{s_map} maps this pattern. There is a clear high 
-concentration of fast adopting LAD in the South and East of England as expected. 
-However, there is also such a concentration in Wales, Scotland and the North 
-West of England. At a more detailed level, there are some expected 
-examples of LAD with relevant industrial backgrounds delineated as fast: the 
-City of London, a world-renowned cluster of finance industries [@cook2007role], 
-and Reading, a town with high-tech service industries in proximity to London 
-and its main airport, Heathrow [@england2005polynet]. However, there are also 
-LAD which were expected to appear as fast -- e.g. Hackney in central London and 
-Bristol, a well-established creative cluster [@oatley1999cultural; 
-@bassett2002cultural] -- but were delineated as slow. Nevertheless, there is 
-obvious spatial clustering of fast and slow LAD. Importantly, out of the ten 
-fastest LAD, nine were located in the South East of England and London and one 
-in Cambridge (see Table \ref{table.s.lads} in the Appendix). The above 
-observations remain almost unchanged when the analysis includes websites 
-with up to 10 unique postcodes -- see Figures \ref{s_uk10} and \ref{s_map10} in 
-the Appendix -- advocating towards the robustness of the analysis.
+Figure \ref{s_uk} plots the cumulative adoption of the Web in total in
+the UK during the 1996-2012 period. Not surprisingly, the cumulative
+adoption is well represented by a curve that resembles the figure S. The
+vertical line around year 2003 illustrates the point where the modelled
+cumulative adoption is equal to 50% of the maximum. This $t_0$
+*inflection* point signals the maximum adoption speed and is used here
+to determine whether a UK LAD reached that inflection point earlier or
+later than the UK average. Specifically, the S curve and the inflection
+point are estimated for every LAD individually and then compared against
+the UK average. LAD that reached their inflection point earlier than the
+country average are labelled as *fast* and the rest as *slow*. Figure
+\ref{s_map} maps this pattern. There is a clear high concentration of
+fast adopting LAD in the South and East of England as expected. However,
+there is also such a concentration in Wales, Scotland and the North West
+of England. At a more detailed level, there are some expected examples
+of LAD with relevant industrial backgrounds delineated as fast: the City
+of London, a world-renowned cluster of finance industries
+[@cook2007role], and Reading, a town with high-tech service industries
+in proximity to London and its main airport, Heathrow
+[@england2005polynet]. However, there are also LAD which were expected
+to appear as fast -- e.g. Hackney in central London and Bristol, a
+well-established creative cluster [@oatley1999cultural;
+@bassett2002cultural] -- but were delineated as slow. Nevertheless,
+there is obvious spatial clustering of fast and slow LAD. Importantly,
+out of the ten fastest LAD, nine were located in the South East of
+England and London and one in Cambridge (see Table \ref{table.s.lads} in
+the Appendix). The above observations remain almost unchanged when the
+analysis includes websites with up to 10 unique postcodes -- see Figures
+\ref{s_uk10} and \ref{s_map10} in the Appendix -- advocating towards the
+robustness of the analysis.
 
@@
 <!-- discussion -->
-The S-curves of the cumulative adoption of Web in the UK in total and in the 
-UK LAD demonstrate a pattern well aligned with all previous studies discussed 
-in @sec-litreview. Importantly, the above findings are in line with previous 
-studies exploring the diffusion of web technologies [@PAPAGIANNIDIS2015308], 
-social media [@lengyel2020role] and online shopping [@bakher2013diffusion]. 
-Similar to @beardsell1999spatial, who analysed the spatial evolution of 
-computer industry employment across 317 US metro areas, spatial heterogeneity
-is evident in the diffusion speed when the scale of analysis is as detailed
-as the LAD. Importantly, no previous studies estimated such cumulative 
-adoption curves for areal units as small as the UK LADS.
+The S-curves of the cumulative adoption of Web in the UK in total and in
+the UK LAD demonstrate a pattern well aligned with all previous studies
+discussed in @sec-litreview. Importantly, the above findings are in line
+with previous studies exploring the diffusion of web technologies
+[@PAPAGIANNIDIS2015308], social media [@lengyel2020role] and online
+shopping [@bakher2013diffusion]. Similar to @beardsell1999spatial, who
+analysed the spatial evolution of computer industry employment across
+317 US metro areas, spatial heterogeneity is evident in the diffusion
+speed when the scale of analysis is as detailed as the LAD. Importantly,
+no previous studies estimated such cumulative adoption curves for areal
+units as small as the UK LADS.
 
 <!-- rankings -->
-The next step assesses the stability and volatility of the LAD in terms of 
-their adoption of web technologies. As highlighted in the literature 
-[@risk_perceptions], different agents have different perceptions about and 
-levels of acceptance of the risks and the potential economic returns associated 
-with the adoption of new technologies -- see for instance the seminal work of
-@venkatesh2000theoretical. To reveal such aggregated patterns, Figure \ref{rank} 
-plots the ranking of UK LAD based on website density. To decrease noise, the 
-average ranking of 1996-1998 and 2010-2012 is plotted instead of the individual 
-years. The colours depict the 5th and 95th percentile of the absolute 
-difference of the LAD ranking between 1996-1998 and 2010-2012. While there are 
-quite a few obvious cases of LAD that maintained their position between the 
-beginning and the end of the study period both at the top or at the bottom of 
-the hierarchy, there are also quite a few LAD that changed drastically 
-their position. Some of these LAD enjoyed a process that at the first instance 
-looks like *leapfrogging* since they managed to jump at the top of the hierarchy 
-despite their slow start. There is extensive literature regarding the potential 
-benefits of technological leapfrogging. The underpinning argument is that 
-latecomers can adopt and benefit from new technologies that have been developed 
-elsewhere without incurring the hefty initial R&D costs [@teece2008firm].
-Although the leapfrogging literature does not pay much attention to cities and 
-regions [@yu2018sustainability], previous research highlighted the long term 
-and sustained productivity benefit of the early adopters of web technologies 
-in the UK [@tranosuk]. So, although the LAD system is volatile, it is not clear 
-whether the LAD with high concentration of late adopters will gain any 
-latecomer-related benefits in a way similar to countries experiencing 
-technological leapfrogging. Still, the opportunity to catch up that 
-@perkins2005international illustrated at the country level, is also visible at 
-the much more detailed LAD scale. As previously, this is consistent with 
-including websites with up to 10 postcodes (see Figure \ref{rank10}).
+The next step assesses the stability and volatility of the LAD in terms
+of their adoption of web technologies. As highlighted in the literature
+[@risk_perceptions], different agents have different perceptions about
+and levels of acceptance of the risks and the potential economic returns
+associated with the adoption of new technologies -- see for instance the
+seminal work of @venkatesh2000theoretical. To reveal such aggregated
+patterns, Figure \ref{rank} plots the ranking of UK LAD based on website
+density. To decrease noise, the average ranking of 1996-1998 and
+2010-2012 is plotted instead of the individual years. The colours depict
+the 5th and 95th percentile of the absolute difference of the LAD
+ranking between 1996-1998 and 2010-2012. While there are quite a few
+obvious cases of LAD that maintained their position between the
+beginning and the end of the study period both at the top or at the
+bottom of the hierarchy, there are also quite a few LAD that changed
+drastically their position. Some of these LAD enjoyed a process that at
+the first instance looks like *leapfrogging* since they managed to jump
+at the top of the hierarchy despite their slow start. There is extensive
+literature regarding the potential benefits of technological
+leapfrogging. The underpinning argument is that latecomers can adopt and
+benefit from new technologies that have been developed elsewhere without
+incurring the hefty initial R&D costs [@teece2008firm]. Although the
+leapfrogging literature does not pay much attention to cities and
+regions [@yu2018sustainability], previous research highlighted the long
+term and sustained productivity benefit of the early adopters of web
+technologies in the UK [@tranosuk]. So, although the LAD system is
+volatile, it is not clear whether the LAD with high concentration of
+late adopters will gain any latecomer-related benefits in a way similar
+to countries experiencing technological leapfrogging. Still, the
+opportunity to catch up that @perkins2005international illustrated at
+the country level, is also visible at the much more detailed LAD scale.
+As previously, this is consistent with including websites with up to 10
+postcodes (see Figure \ref{rank10}).
 
@@
 Figures \ref{morani} and \ref{lisa} offer a first insight into whether a
-neighbourhood mechanism underpins the diffusion of the Web in the UK as they 
-present the Moran's I and the LISA maps of website density respectively for LAD 
-and OA. Starting from the former, spatial autocorrelation was higher in the 
-beginning of the study period and then after 2000 dropped slightly and 
-stabilised around 0.2. This reflects the early concentration of high website 
-density around London, which over time diffused as high website density clusters 
-can be seen in other parts of the country away from London (Figure \ref{lisa}). 
-An almost reverse pattern can be observed for OA. At the beginning of the study 
+neighbourhood mechanism underpins the diffusion of the Web in the UK as
+they present the Moran's I and the LISA maps of website density
+respectively for LAD and OA. Starting from the former, spatial
+autocorrelation was higher in the beginning of the study period and then
+after 2000 dropped slightly and stabilised around 0.2. This reflects the
+early concentration of high website density around London, which over
+time diffused as high website density clusters can be seen in other
+parts of the country away from London (Figure \ref{lisa}). An almost
+reverse pattern can be observed for OA. At the beginning of the study
 period Moran's I was around 0.5 and it plateaued after 2000 around 0.2.
-Because of the very small size of OA, at the early stages of the diffusion 
-of web technologies their adoption was spatially scattered. This is reflected in 
-the lack of any significant clusters in 1996 (Figure \ref{lisa}). Eventually, as 
-the adoption rate increased, more such clusters of high website density were 
-formed and this is reflected both in the Moran's I and the LISA maps in Figures 
-\ref{morani} and \ref{lisa}. A similar pattern is observed for the extended 
-dataset (see Figures\ref{morani10} and \ref{lisa10}). In this case, the 
-magnitude of spatial autocorrelation is slightly higher illustrating a stronger 
-spatial clustering mechanism. All in all, the exploratory spatial data analysis 
-advocates towards an underpinning neighbourhood mechanism and the different 
-scales of analysis illustrate how it evolved differently over time. 
-<!-- discussion -->
-Interestingly, the magnitude of Moran's I is similar to the ones revealed
-by @ding2010modeling about mobile phone adoption in China albeit the much more 
-detailed spatial scale of this paper.
+Because of the very small size of OA, at the early stages of the
+diffusion of web technologies their adoption was spatially scattered.
+This is reflected in the lack of any significant clusters in 1996
+(Figure \ref{lisa}). Eventually, as the adoption rate increased, more
+such clusters of high website density were formed and this is reflected
+both in the Moran's I and the LISA maps in Figures \ref{morani} and
+\ref{lisa}. A similar pattern is observed for the extended dataset (see
+Figures\ref{morani10} and \ref{lisa10}). In this case, the magnitude of
+spatial autocorrelation is slightly higher illustrating a stronger
+spatial clustering mechanism. All in all, the exploratory spatial data
+analysis advocates towards an underpinning neighbourhood mechanism and
+the different scales of analysis illustrate how it evolved differently
+over time. <!-- discussion --> Interestingly, the magnitude of Moran's I
+is similar to the ones revealed by @ding2010modeling about mobile phone
+adoption in China albeit the much more detailed spatial scale of this
+paper.
 
@@
 
-To illustrate whether a hierarchical process also underpins the diffusion of web 
-technologies, the Gini coefficient is calculated yearly both for LAD and OA. As 
-a metric of inequality, the Gini coefficient demonstrates whether website 
-density is concentrated in a small number of LAD or OA, or whether it is more 
-equally spread across the country. Both scales of analysis in Figure \ref{gini} 
-illustrate the same picture. At the beginning of the commercial Internet, 
-website density was extremely unequal, or, in other words, only a few places had 
-websites associated with them. Inequality dropped and plateaued after 2000 for 
-both scales. This is illustrative of a hierarchical diffusion mechanism that 
-led over time to a more equal spread. Interestingly, the year 2000 is again a 
-period of change for this diffusion mechanism as it was for the neighbourhood 
-process. There is a substantial difference between the Gini coefficient 
-magnitude for LAD and OA, but this is expected as the very small size of OA 
-equates to a lot of polygons without any websites pointing to them -- for 
-example residential OA.
+To illustrate whether a hierarchical process also underpins the
+diffusion of web technologies, the Gini coefficient is calculated yearly
+both for LAD and OA. As a metric of inequality, the Gini coefficient
+demonstrates whether website density is concentrated in a small number
+of LAD or OA, or whether it is more equally spread across the country.
+Both scales of analysis in Figure \ref{gini} illustrate the same
+picture. At the beginning of the commercial Internet, website density
+was extremely unequal, or, in other words, only a few places had
+websites associated with them. Inequality dropped and plateaued after
+2000 for both scales. This is illustrative of a hierarchical diffusion
+mechanism that led over time to a more equal spread. Interestingly, the
+year 2000 is again a period of change for this diffusion mechanism as it
+was for the neighbourhood process. There is a substantial difference
+between the Gini coefficient magnitude for LAD and OA, but this is
+expected as the very small size of OA equates to a lot of polygons
+without any websites pointing to them -- for example residential OA.
 
@@
 The next section incorporates the above mechanisms of the spatial
-diffusion of the Web into a modelling framework. The aim is to use variables 
-depicting these spatial processes in order to predict the diffusion of the Web 
-as a new technology in the UK over space and time and across different scales. 
-Specifically, four different models are estimated. Firstly, all the data points 
-for the OA and LAD are utilised in order to build two RF models and assess their 
-capacity to predict the adoption of the Web at the two different scales. 
-These two models will reveal the predictive capacity of the diffusion mechanisms 
-and also demonstrate how the importance of such variables changes across different 
-scales. The next two sets of models are again trained at these two scales: 
-OA and LAD. However, instead of using all the data points, the OA and
-the LAD from one of the twelve UK regions are held out for the model training.
-Then, the trained model is used to predict website density for the OA or the LAD 
-of the held-out region. This process takes place recursively for all twelve UK
-regions. The regional differences in the predictive capacity of the different
-samples will reveal how dissimilar these spatial process are across regions 
-and, importantly, at different scales.
-
-It needs to be highlighted here that the cross validation (CV) for all models
-is spatially and temporally sensitive. Instead of using
-10 random and space- and time-agnostic folds,the `CAST` package is employed 
-which allows holding back data points from specific years and spatial units 
+diffusion of the Web into a modelling framework. The aim is to use
+variables depicting these spatial processes in order to predict the
+diffusion of the Web as a new technology in the UK over space and time
+and across different scales. Specifically, four different models are
+estimated. Firstly, all the data points for the OA and LAD are utilised
+in order to build two RF models and assess their capacity to predict the
+adoption of the Web at the two different scales. These two models will
+reveal the predictive capacity of the diffusion mechanisms and also
+demonstrate how the importance of such variables changes across
+different scales. The next two sets of models are again trained at these
+two scales: OA and LAD. However, instead of using all the data points,
+the OA and the LAD from one of the twelve UK regions are held out for
+the model training. Then, the trained model is used to predict website
+density for the OA or the LAD of the held-out region. This process takes
+place recursively for all twelve UK regions. The regional differences in
+the predictive capacity of the different samples will reveal how
+dissimilar these spatial process are across regions and, importantly, at
+different scales.
+
+It needs to be highlighted here that the cross validation (CV) for all
+models is spatially and temporally sensitive. Instead of using 10 random
+and space- and time-agnostic folds,the `CAST` package is employed which
+allows holding back data points from specific years and spatial units
 and use them for testing in order to estimate the model performance
@@
 
-The models include variables that capture the three processes that the relevant 
-literature and the descriptive analysis highlighted. Namely, the models capture: 
-(i) a hierarchical mechanism with diffusion running from main centres to 
-secondary ones, (ii) a neighborhood mechanism according to which diffusion 
-first hits nearby locations, and (iii) the rather canonical pattern of 
-diffusion over time as reflected in the S-shaped pattern in the cumulative level 
-of adoption.
-
-To capture the hierarchical mechanism the models include as predictors a one
-year lag of website density in London, the largest city in the UK, a
+The models include variables that capture the three processes that the
+relevant literature and the descriptive analysis highlighted. Namely,
+the models capture: (i) a hierarchical mechanism with diffusion running
+from main centres to secondary ones, (ii) a neighborhood mechanism
+according to which diffusion first hits nearby locations, and (iii) the
+rather canonical pattern of diffusion over time as reflected in the
+S-shaped pattern in the cumulative level of adoption.
+
+To capture the hierarchical mechanism the models include as predictors a
+one year lag of website density in London, the largest city in the UK, a
 one year lag of the website density in the nearest city and the same for
@@
 the latter is only relevant for and included in the OA-level models. In
-addition, the models include the distance to London, the nearest city and the 
-nearest retail centre. The underlying logic is that the level of website
-adoption in a spatial unit depends on the level of the adoption in
-places further up in the urban hierarchy the previous year. The inclusion of
-the distance variables incorporates spatial structure into the hierarchy 
-argument. To depict the neighbourhood mechanism, the website density of the 
-neighbouring spatial units in the previous year is employed. This is calculated 
-using rook continuity. Again, the underpinning rationale is that the level of 
-web adoption within a spatial unit depends on the level of web adoption in the 
-neighbouring spatial units the year before. This represents the 'hitting nearby 
-locations first' argument. Therefore, the spatial and temporal lag of the 
-website density in LAD and OA is calculated. Lastly, the temporal pattern of 
-the cumulative adoption is captured by a time trend variable. Hence, all four 
-models will follow the following generic form (Eq. \ref{model}):
-
-```{=tex}
+addition, the models include the distance to London, the nearest city
+and the nearest retail centre. The underlying logic is that the level of
+website adoption in a spatial unit depends on the level of the adoption
+in places further up in the urban hierarchy the previous year. The
+inclusion of the distance variables incorporates spatial structure into
+the hierarchy argument. To depict the neighbourhood mechanism, the
+website density of the neighbouring spatial units in the previous year
+is employed. This is calculated using rook continuity. Again, the
+underpinning rationale is that the level of web adoption within a
+spatial unit depends on the level of web adoption in the neighbouring
+spatial units the year before. This represents the 'hitting nearby
+locations first' argument. Therefore, the spatial and temporal lag of
+the website density in LAD and OA is calculated. Lastly, the temporal
+pattern of the cumulative adoption is captured by a time trend variable.
+Hence, all four models will follow the following generic form (Eq.
+\ref{model}):
+
 \begin{align} \label{model}
@@
 \end{align}
-```
 
@@
 
-```{=tex}
 \begin{align}
@@
 \end{align}
-```
-```{=tex}
 \begin{align}
@@
 \end{align}
-```
-
-\noindent $y_{k}$ is the $k^{th}$ observation of the dataset, which consists of
-$N$ observations in total. $\hat{y_{k}}$ is the $k_{th}$ predicted value
-for the dependent variable and $\overline{y_{k}}$ is the average value
-of $y$. The last two metrics are expressed in the same units as the
-dependent variable -- websites per firm for the LAD modes and the number
-of websites for the OA models -- while the first one is the coefficient
-of determination between the observed and the predicted values of
-website adoption. Regarding $MAE$, it is the absolute difference between
-the observed and the predicted website adoption. While $MAE$ does not
-penalise for large errors, $RSME$ does, as it is proportional to the
-squared difference between the observed and the predicted trade flows.
-Hence larger errors weigh more for $RMSE$ [@pontius2008components].
-
-@tbl-model-metrics presents the model performance for the first set of models, 
-for which all data points are employed for training and testing via CV. The 
-first one is trained and tested on 374 LAD and the second on 
-232,296 OA, both for the 16 year period (1997-2012). The results are remarkably 
-good considering that the are the outcome of space and time sensitive 
-CV, so the the model does not suffer from overfitting. At the LAD level the 
-model predicts 85% of the variation of website density. Both error metrics 
-indicate that the model error is $(\frac{1}{20}, \frac{1}{30})$ of a website per 
-firm. At the OA the $R^2$ drops down to 32%. Considering its granularity, this 
-is still a remarkable performance. To contextualise it, the model results in a 
+
+\noindent $y_{k}$ is the $k^{th}$ observation of the dataset, which
+consists of $N$ observations in total. $\hat{y_{k}}$ is the $k_{th}$
+predicted value for the dependent variable and $\overline{y_{k}}$ is the
+average value of $y$. The last two metrics are expressed in the same
+units as the dependent variable -- websites per firm for the LAD modes
+and the number of websites for the OA models -- while the first one is
+the coefficient of determination between the observed and the predicted
+values of website adoption. Regarding $MAE$, it is the absolute
+difference between the observed and the predicted website adoption.
+While $MAE$ does not penalise for large errors, $RSME$ does, as it is
+proportional to the squared difference between the observed and the
+predicted trade flows. Hence larger errors weigh more for $RMSE$
+[@pontius2008components].
+
+@tbl-model-metrics presents the model performance for the first set of
+models, for which all data points are employed for training and testing
+via CV. The first one is trained and tested on 374 LAD and the second on
+232,296 OA, both for the 16 year period (1997-2012). The results are
+remarkably good considering that the are the outcome of space and time
+sensitive CV, so the the model does not suffer from overfitting. At the
+LAD level the model predicts 85% of the variation of website density.
+Both error metrics indicate that the model error is
+$(\frac{1}{20}, \frac{1}{30})$ of a website per firm. At the OA the
+$R^2$ drops down to 32%. Considering its granularity, this is still a
+remarkable performance. To contextualise it, the model results in a
 $MAE$ equal to one website for areas small enough to host less than 140
-households.
-^[According to the Office for National Statistics, 80% of OA in England and 
-Wales host 110-139 households,
-[www.ons.gov.uk](https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas).]
-Because of the small size of the spatial units, the distribution is highly skewed
-and a significant part of them is not linked to any websites. In 1997 only 1% of 
-the UK OA were associated with at least one website. This should not come as a 
-surprise as this was the very beginning of the commercial Internet and any 
-activities with a digital footprint were concentrated in a handful of areas. 
-This was clearly illustrated in Figure \ref{lisa}. At the end of the study 
-period almost half of the UK OA were not associated with a website. Again, 
-given the granularity of the data this should not come as a surprise.
+households. [^4] Because of the small size of the spatial units, the
+distribution is highly skewed and a significant part of them is not
+linked to any websites. In 1997 only 1% of the UK OA were associated
+with at least one website. This should not come as a surprise as this
+was the very beginning of the commercial Internet and any activities
+with a digital footprint were concentrated in a handful of areas. This
+was clearly illustrated in Figure \ref{lisa}. At the end of the study
+period almost half of the UK OA were not associated with a website.
+Again, given the granularity of the data this should not come as a
+surprise.
+
+[^4]: According to the Office for National Statistics, 80% of OA in
+    England and Wales host 110-139 households,
+    [www.ons.gov.uk](https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas).
 
@@
 
-The results in @tbl-model-metrics are robust against different specification. 
-As noted earlier, the data imputation process resulted in a modest improvement 
-of the predictive capacity of the model. As per
-@tbl-model-metrics-no-correction all accuracy metrics are only slightly worst 
-when the data imputation is not applied. As expected, the improvement is 
-higher for the more detailed scale of analysis. Then, @tbl-model-metrics-10 
-presents the model results for the extended dataset, which includes websites 
-with up to 10 unique postcodes. While the results for the LAD are slightly 
-less accurate, the $R^{2}$ of the OA model is more than double in comparison 
-to the base results in @tbl-model-metrics. All in all, the modelling exercise
-revealed how well can the three diffusion mechanisms predict 
-the diffusion of the Web as a new technology. Importantly, this predictive
-power is highly robust against against different scales and delineations
-of the diffusion of the web.
+The results in @tbl-model-metrics are robust against different
+specification. As noted earlier, the data imputation process resulted in
+a modest improvement of the predictive capacity of the model. As per
+@tbl-model-metrics-no-correction all accuracy metrics are only slightly
+worst when the data imputation is not applied. As expected, the
+improvement is higher for the more detailed scale of analysis. Then,
+@tbl-model-metrics-10 presents the model results for the extended
+dataset, which includes websites with up to 10 unique postcodes. While
+the results for the LAD are slightly less accurate, the $R^{2}$ of the
+OA model is more than double in comparison to the base results in
+@tbl-model-metrics. All in all, the modelling exercise revealed how well
+can the three diffusion mechanisms predict the diffusion of the Web as a
+new technology. Importantly, this predictive power is highly robust
+against against different scales and delineations of the diffusion of
+the web.
 
@@
 
-: Model metrics without data imputation {#tbl-model-metrics-no-correction}
+: Model metrics without data imputation
+{#tbl-model-metrics-no-correction}
 
@@
 
-Figure \ref{var.imp} plots the importance of the different predictors. When the 
-focus is on LAD, the website density in the nearest city, in London and in the 
-neighbouring LAD the year before are the most important predictors. They are 
-followed by the yearly trend, while the spatial configuration as reflected
-in distances to London or the nearest city only plays a minor role. This can be 
-attributed to the relatively coarse spatial scale of LAD. Nevertheless, all 
-previously discussed spatial processes are at play in the diffusion of web 
-technologies at the LAD level: the first two predictors depict the hierarchical 
-mechanism, the spatial and temporal lag of website density the neighbourhood 
-mechanism, and the yearly trend the time-sensitive cumulative adoption pattern.
-
-When the much more granular scale of OA is adopted, the picture is almost 
-reversed. The two most important predictors are the distance to London and to 
-the nearest city followed by the spatial and temporal lag of website density 
-and the distance to the nearest retail centre. They still depict a hierarchical 
-mechanism, but proximity to the different population centres is more important 
-than their lagged web densities in predicting website diffusion. The 
-neighbourhood mechanism is still strong but less important at this scale. What 
-is interesting is the almost negligible role of the yearly trend and London's 
-website density. While the former probably illustrates the large heterogeneity 
-in how the Web has been adopted at this very fine scale, the latter highlights 
-that the importance of past web adoption rates in large population centres is 
+Figure \ref{var.imp} plots the importance of the different predictors.
+When the focus is on LAD, the website density in the nearest city, in
+London and in the neighbouring LAD the year before are the most
+important predictors. They are followed by the yearly trend, while the
+spatial configuration as reflected in distances to London or the nearest
+city only plays a minor role. This can be attributed to the relatively
+coarse spatial scale of LAD. Nevertheless, all previously discussed
+spatial processes are at play in the diffusion of web technologies at
+the LAD level: the first two predictors depict the hierarchical
+mechanism, the spatial and temporal lag of website density the
+neighbourhood mechanism, and the yearly trend the time-sensitive
+cumulative adoption pattern.
+
+When the much more granular scale of OA is adopted, the picture is
+almost reversed. The two most important predictors are the distance to
+London and to the nearest city followed by the spatial and temporal lag
+of website density and the distance to the nearest retail centre. They
+still depict a hierarchical mechanism, but proximity to the different
+population centres is more important than their lagged web densities in
+predicting website diffusion. The neighbourhood mechanism is still
+strong but less important at this scale. What is interesting is the
+almost negligible role of the yearly trend and London's website density.
+While the former probably illustrates the large heterogeneity in how the
+Web has been adopted at this very fine scale, the latter highlights that
+the importance of past web adoption rates in large population centres is
 surpassed by proximity to them and spatial configuration at this scale.
 
-The above pattern of variable importance is largely in accordance with the 
-models estimated for the extended dataset. A comparison with Figure 
-\ref{var.imp10} illustrates that the main difference is the importance of the 
-neighbourhood mechanism for the OA as the space and time lag of the website 
-density is by far the most important predictor when the extended dataset is 
-employed. But still, the variable importance patterns for the two datasets are 
-not dissimilar.
+The above pattern of variable importance is largely in accordance with
+the models estimated for the extended dataset. A comparison with Figure
+\ref{var.imp10} illustrates that the main difference is the importance
+of the neighbourhood mechanism for the OA as the space and time lag of
+the website density is by far the most important predictor when the
+extended dataset is employed. But still, the variable importance
+patterns for the two datasets are not dissimilar.
 
@@
 
-Table \ref{table.regions} presents the results of the recursive hold out models
-and allows to (i) highlight the regional heterogeneity of the spatial 
-processes behind the diffusion of the Web, and (ii) compare this heterogeneity
-across the two different scales of analysis. It illustrates some striking 
-similarities, but also a few significant differences. At the top of the 
-similarity with the rest of the country we see West Midlands and Yorkshire
-and the Humber for both scales. The web density for the LAD and OA of these 
-regions can be very well predicted by models trained on data for the LAD 
-and OA of the rest of the country. In other words, the spatial diffusion 
-mechanisms for these regions represent quite well the UK's average. This
-should not come as as surprise given their urban structure: they both contain
-a large urban centre (Birmingham and Leeds), smaller commuter towns as well 
-as more rural settlements. The bottom of the similarity ranking is populated 
-by Scotland, Northern Ireland, East of England and London for both scales. 
-The diffusion mechanisms for these regions diverge significantly from the 
-country's average and this could be attributed to rurality and remoteness 
-or to high levels of urbanisation and sophistication of economic activities 
-for the case of London and for East of England, which contains the Cambridge 
-area.
+Table \ref{table.regions} presents the results of the recursive hold out
+models and allows to (i) highlight the regional heterogeneity of the
+spatial processes behind the diffusion of the Web, and (ii) compare this
+heterogeneity across the two different scales of analysis. It
+illustrates some striking similarities, but also a few significant
+differences. At the top of the similarity with the rest of the country
+we see West Midlands and Yorkshire and the Humber for both scales. The
+web density for the LAD and OA of these regions can be very well
+predicted by models trained on data for the LAD and OA of the rest of
+the country. In other words, the spatial diffusion mechanisms for these
+regions represent quite well the UK's average. This should not come as
+as surprise given their urban structure: they both contain a large urban
+centre (Birmingham and Leeds), smaller commuter towns as well as more
+rural settlements. The bottom of the similarity ranking is populated by
+Scotland, Northern Ireland, East of England and London for both scales.
+The diffusion mechanisms for these regions diverge significantly from
+the country's average and this could be attributed to rurality and
+remoteness or to high levels of urbanisation and sophistication of
+economic activities for the case of London and for East of England,
+which contains the Cambridge area.
 
@@
 differ a lot between the two scales. North West and South West are very
-close to the UK's average at the more granular OA scale, while the opposite
-applies for South East. Interestingly, the first two regions are well-known 
-tourist destinations (Lake District and Cornwall-Devon) and the exploratory
-analysis revealed clusters of high web density only visible at the OA scale 
-(Figure \ref{lisa}. Then, the South East is very well integrated with the 
-Greater London, hosts the broader Oxford area -- a known hub of research 
-excellence -- and Figure \ref{lisa} revealed high web density clusters at 
-the LAD level.
+close to the UK's average at the more granular OA scale, while the
+opposite applies for South East. Interestingly, the first two regions
+are well-known tourist destinations (Lake District and Cornwall-Devon)
+and the exploratory analysis revealed clusters of high web density only
+visible at the OA scale (Figure \ref{lisa}. Then, the South East is very
+well integrated with the Greater London, hosts the broader Oxford area
+-- a known hub of research excellence -- and Figure \ref{lisa} revealed
+high web density clusters at the LAD level.
 
@@
 
-This paper assessed the importance of spatial mechanisms in the diffusion 
-of a new technology -- the Web -- at a high level of spatial granularity.
-Specifically, it assessed the capacity of the following three spatial 
-diffusion mechanisms in predicting website density in the UK over space and 
-time: (i) the hierarchical mechanism according to which diffusion runs 
-from main urban centres to secondary ones, (ii) the neighborhood mechanism 
-which highlights that diffusion first hits nearby locations, and (iii) the 
-rather canonical pattern of diffusion over time as reflected in the S-shaped 
-pattern of the cumulative level of adoption. This is the first time that 
-the importance of granular spatial mechanisms for the diffusion of an
-intangible, digital technology is exposed. The time period of the study is 
-also important as it captures the very early stages of this technology until 
-its maturity. Importantly, the empirical framework and the data employed here 
-allows to observe the active engagement with the Web that is creating and 
+<!-- ADD MORE ABOUT LOCAL POLICIES: E.G. USE MY RESULTS TO MODEL WHERE INFRASTRUCTURE OR GSM ANTENNAS SHOULD BE LOCALTED -->
+
+This paper assessed the importance of spatial mechanisms in the
+diffusion of a new technology -- the Web -- at a high level of spatial
+granularity. Specifically, it assessed the capacity of the following
+three spatial diffusion mechanisms in predicting website density in the
+UK over space and time: (i) the hierarchical mechanism according to
+which diffusion runs from main urban centres to secondary ones, (ii) the
+neighborhood mechanism which highlights that diffusion first hits nearby
+locations, and (iii) the rather canonical pattern of diffusion over time
+as reflected in the S-shaped pattern of the cumulative level of
+adoption. This is the first time that the importance of granular spatial
+mechanisms for the diffusion of an intangible, digital technology is
+exposed. The time period of the study is also important as it captures
+the very early stages of this technology until its maturity.
+Importantly, the empirical framework and the data employed here allows
+to observe the active engagement with the Web that is creating and
 maintaining a commercial website instead of the more passive action of
@@
 
-The analysis firstly illustrated how well the S-pattern of cumulative adoption 
-fits at the local scale of LAD. It highlighted the heterogeneity of adoption 
-rates between LAD -- with both more and less expected examples -- as well as 
-the spatial clustering of LAD with similar adoption rates. The exploratory 
-analysis also illustrated instances of remarkable stability as well as 
-volatility of the web diffusion system that  is indicative of leapfrogging. 
-Previous research had already indicated the long-term productivity gains of 
-the early adoption of such technologies [@tranosuk]. Then, the paper developed 
-a robust and novel ML modelling framework to predict website diffusion over 
-space and time using the three spatial diffusion mechanisms based on space- 
-and time-sensitive CV without suffering from overfitting. The model revealed 
-the remarkably good predictive capacity of these mechanisms at both scales. 
-For the first time, the power of these mechanisms in predicting the diffusion 
-of a new technology was empirically tested, importantly, at very fine scales.
-
-The modelling framework also revealed that the importance of these spatial 
-mechanisms differ across different scales. While spatial structure as 
-reflected in distances to different urban centres is highly important at the 
-OA scale, it hardly plays any role at the more coarse LAD scale. And while 
-the almost canonical pattern of cumulative adoption is a key predictor at the 
-LAD level, it only plays a small role at the detailed OA level. Then, the 
-analysis revealed another level of spatial heterogeneity. While for some UK 
-regions these spatial diffusion mechanisms are very similar to the rest of 
-the country, other regions diverge significantly from the country average.
-Importantly, the above results are robust against different specifications 
-and subsets of the data.
-
-The paper offers a number of different contributions. Firstly, it assesses, 
-for the first time, the role of these spatial diffusion mechanisms at 
-such granular spatial scales. While previous studies had 
-highlighted these mechanisms [@hagerstrand1968innovation; @rogers2010diffusion; 
-@grubler1990rise], their role had never been tested before at such granular 
-scale and, importantly, for a digital technology as weightless as the Web. 
-The findings talk directly to this literature  [e.g. @beardsell1999spatial; 
-@haller2011determinants; @lengyel2020role]. More broadly, the paper offers 
-a robust methodological framework, which utilises novel data and modern ML 
-algorithms and techniques to model the diffusion of a new technology across 
-space and time. It directly addresses the data problem which prevented economic
-geographers in engaging  with diffusion studies [@iso2005; 
-@kemeny2011international; @zook2022mapping].
-Thus, the paper seeks to renew economic geography's interest in the diffusion 
-of new technologies. Current data science advances including the availability 
+The analysis firstly illustrated how well the S-pattern of cumulative
+adoption fits at the local scale of LAD. It highlighted the
+heterogeneity of adoption rates between LAD -- with both more and less
+expected examples -- as well as the spatial clustering of LAD with
+similar adoption rates. The exploratory analysis also illustrated
+instances of remarkable stability as well as volatility of the web
+diffusion system that is indicative of leapfrogging. Previous research
+had already indicated the long-term productivity gains of the early
+adoption of such technologies [@tranosuk]. Then, the paper developed a
+robust and novel ML modelling framework to predict website diffusion
+over space and time using the three spatial diffusion mechanisms based
+on space- and time-sensitive CV without suffering from overfitting. The
+model revealed the remarkably good predictive capacity of these
+mechanisms at both scales. For the first time, the power of these
+mechanisms in predicting the diffusion of a new technology was
+empirically tested, importantly, at very fine scales.
+
+The modelling framework also revealed that the importance of these
+spatial mechanisms differ across different scales. While spatial
+structure as reflected in distances to different urban centres is highly
+important at the OA scale, it hardly plays any role at the more coarse
+LAD scale. And while the almost canonical pattern of cumulative adoption
+is a key predictor at the LAD level, it only plays a small role at the
+detailed OA level. Then, the analysis revealed another level of spatial
+heterogeneity. While for some UK regions these spatial diffusion
+mechanisms are very similar to the rest of the country, other regions
+diverge significantly from the country average. Importantly, the above
+results are robust against different specifications and subsets of the
+data.
+
+The paper offers a number of different contributions. Firstly, it
+assesses, for the first time, the role of these spatial diffusion
+mechanisms at such granular spatial scales. While previous studies had
+highlighted these mechanisms [@hagerstrand1968innovation;
+@rogers2010diffusion; @grubler1990rise], their role had never been
+tested before at such granular scale and, importantly, for a digital
+technology as weightless as the Web. The findings talk directly to this
+literature [e.g. @beardsell1999spatial; @haller2011determinants;
+@lengyel2020role]. More broadly, the paper offers a robust
+methodological framework, which utilises novel data and modern ML
+algorithms and techniques to model the diffusion of a new technology
+across space and time. It directly addresses the data problem which
+prevented economic geographers in engaging with diffusion studies
+[@iso2005; @kemeny2011international; @zook2022mapping]. Thus, the paper
+seeks to renew economic geography's interest in the diffusion of new
+technologies. Current data science advances including the availability
 of granular enough data and adequate methods can enable economic
@@
 technologies as this paper has demonstrated. This is important as (i)
-various stakeholders are interested in predicting localised demand for new 
-technologies for planning purposes [@leibowicz2016representing; 
+various stakeholders are interested in predicting localised demand for
+new technologies for planning purposes [@leibowicz2016representing;
 @meade2021modelling], and (ii) the adoption of such digital technologies
-is associated with economic gains [@solow1957technical; @aghion1990model; 
-@kemeny2011international; @tranosuk; @capello2024nexus]. In addition, while 
-most related studies focused on tangible technologies, this paper sheds 
-light on the diffusion of a truly weightless digital technology that is the 
-Web. Thus, it adds much more granular findings to the limited related studies
-[@sinai2004geography; @haller2011determinants]. Importantly, although the 
-geographical focus of this paper is the UK, the empirical framework developed 
-here is easily transferable to other geographical contexts.
+is associated with economic gains [@solow1957technical;
+@aghion1990model; @kemeny2011international; @tranosuk;
+@capello2024nexus]. In addition, while most related studies focused on
+tangible technologies, this paper sheds light on the diffusion of a
+truly weightless digital technology that is the Web. Thus, it adds much
+more granular findings to the limited related studies
+[@sinai2004geography; @haller2011determinants]. Importantly, although
+the geographical focus of this paper is the UK, the empirical framework
+developed here is easily transferable to other geographical contexts.
 
 <!-- Future research can further investigate other -->
+
 <!-- determinants of the diffusion of the Web. For instance, the archived web data -->
+
 <!-- can reveal important information about the industrial base of different places -->
+
 <!-- as well as the industries which were faster in adopting the Web. -->
@@
 <!-- reset the figure, table and section numbering -->
+
 \setcounter{section}{0}
 \renewcommand{\thesection}{\Alph{section}}
-
 \setcounter{table}{0}
 \renewcommand{\thetable}{A\arabic{table}}
-
 \setcounter{figure}{0}
@@
 <!-- #sec-appendix -->
+
 <!-- .appendix -->

### 2025-01-21 paper/submission2/response.qmd
paper

@@
 Regarding the statement about economic geography studies on diffusion of 
-new technologies and innovation, I do believe that this is a fair statement. 
-The references listed below are the ones reviewed in the second section and none 
-of them are published in one of the flagship economic geography journals 
-(*Journal of Economic Geography* and *Economic Geography*). There are two papers
-published in *The Professional Geographer*, which is a broad-scope geography 
-journal and then only one paper in *Regional Studies*, which is the closest to
-economic geography. Importantly though, the *Regional Studies* paper 
+new technologies and innovation, respectfully, I do believe that this is a fair
+statement. The references listed below are the ones reviewed in the second 
+section and none of them are published in one of the flagship economic geography
+journals (*Journal of Economic Geography* and *Economic Geography*). There are 
+two papers published in *The Professional Geographer*, which is a broad-scope
+geography journal and then only one paper in *Regional Studies*, which is the 
+closest to economic geography. Importantly though, the *Regional Studies* paper 
 [@feldman2015rknowledge] paper is focused on inventions as observed by patent 
@@
 
-1. @beardsell1999spatial: European Economic Review
-2. @bednarz2020pulled: Industrial and Corporate Change
-3. @haller2011determinants: Applied Economics
-4. @perkins2011internet: The Professional Geographer
-5. @feldman2015rknowledge: Regional Studies
-6. @ding2010modeling: The Professional Geographer
-7. @bakher2013diffusion: Journal of Marketing Analytics
-8. @PAPAGIANNIDIS2015308: Technological Forecasting and Social Change
-9. @lengyel2020role: Scientific reports
+1. @beardsell1999spatial: *European Economic Review*
+2. @bednarz2020pulled: *Industrial and Corporate Change*
+3. @haller2011determinants: *Applied Economics*
+4. @perkins2011internet: *The Professional Geographer*
+5. @feldman2015rknowledge: *Regional Studies*
+6. @ding2010modeling: *The Professional Geographer*
+7. @bakher2013diffusion: *Journal of Marketing Analytics*
+8. @PAPAGIANNIDIS2015308: *Technological Forecasting and Social Change*
+9. @lengyel2020role: *Scientific reports*
 
@@
 
-In the relevant paragrapg in the Data and Methods section, the following line
+In the relevant paragraph in the Data and Methods section, the following line
 has been added: "Henceforward any reference to websites will refer to commercial 
@@
 Section 2 includes a new paragraph describing the second hypothesis (H~2~), 
-which explicitely discusses the value of the multiscale approach: 
+which explicitly discusses the value of the multiscale approach: 
 "Understanding the diffusion of a new technology in multiple, detailed spatial
@@
 
+H1: first descriptives, then modelling
+- S curves
+Figure 1: UK curve, Figure 2: some LAD curves. link to a gif. 
+keep fast slow map, briefly touch upon and say it will be further discussed below
+- neighbourhood
+- hierarchy
+- modelling
+
+H2: link map with ranks. Explore OA stability within LADs
+
 \color{teal}
@@
 
-**reveal that theories are valid for lcoal scales and inform local tech policies**
+**reveal that theories are valid for local scales and inform local tech policies**
 

### 2025-01-21 paper/submission2/tranos2025.qmd
paper

@@
 
-> H~2~: The diffusion of the commecrial Web is shaped by both early adoption 
+> H~2~: The diffusion of the commercial Web is shaped by both early adoption 
 and late comer effects.
@@
 
-<!-- rankings -->
-The next step assesses the stability and volatility of the LAD in terms
-of their adoption of web technologies. As highlighted in the literature
-[@risk_perceptions], different agents have different perceptions about
-and levels of acceptance of the risks and the potential economic returns
-associated with the adoption of new technologies -- see for instance the
-seminal work of @venkatesh2000theoretical. To reveal such aggregated
-patterns, Figure \ref{rank} plots the ranking of UK LAD based on website
-density. To decrease noise, the average ranking of 1996-1998 and
-2010-2012 is plotted instead of the individual years. The colours depict
-the 5th and 95th percentile of the absolute difference of the LAD
-ranking between 1996-1998 and 2010-2012. While there are quite a few
-obvious cases of LAD that maintained their position between the
-beginning and the end of the study period both at the top or at the
-bottom of the hierarchy, there are also quite a few LAD that changed
-drastically their position. Some of these LAD enjoyed a process that at
-the first instance looks like *leapfrogging* since they managed to jump
-at the top of the hierarchy despite their slow start. There is extensive
-literature regarding the potential benefits of technological
-leapfrogging. The underpinning argument is that latecomers can adopt and
-benefit from new technologies that have been developed elsewhere without
-incurring the hefty initial R&D costs [@teece2008firm]. Although the
-leapfrogging literature does not pay much attention to cities and
-regions [@yu2018sustainability], previous research highlighted the long
-term and sustained productivity benefit of the early adopters of web
-technologies in the UK [@tranosuk]. So, although the LAD system is
-volatile, it is not clear whether the LAD with high concentration of
-late adopters will gain any latecomer-related benefits in a way similar
-to countries experiencing technological leapfrogging. Still, the
-opportunity to catch up that @perkins2005international illustrated at
-the country level, is also visible at the much more detailed LAD scale.
-As previously, this is consistent with including websites with up to 10
-postcodes (see Figure \ref{rank10}).
-
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank}Dynamics of wed diffusion (LAD)"}
-path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av.png")
-knitr::include_graphics(path.image)
-```
-
 ## Exploratory analysis
@@
 high web density clusters at the LAD level.
+
+## H2
+
+<!-- H~2~: The diffusion of the commercial Web is shaped by both early adoption  -->
+<!-- and late comer effects. -->
+
+To test the second hypothesis (H~2~) the stability and volatility of the LAD 
+in terms of their adoption of the Web is assessed. As highlighted in 
+the literature [@risk_perceptions], different agents have different perceptions 
+about and levels of acceptance of the risks and the potential economic returns
+associated with the adoption of new technologies -- see for instance the
+seminal work of @venkatesh2000theoretical. To reveal such aggregated
+patterns, Figure \ref{rank} plots the ranking of UK LAD based on website
+density. To decrease noise, the average ranking of 1996-1998 and
+2010-2012 is plotted instead of the individual years. The colours depict
+the 5th and 95th percentile of the absolute difference of the LAD
+ranking between 1996-1998 and 2010-2012. While there are quite a few
+obvious cases of LAD that maintained their position between the
+beginning and the end of the study period both at the top or at the
+bottom of the hierarchy, there are also quite a few LAD that changed
+drastically their position. Some of these LAD enjoyed a process that at
+the first instance looks like *leapfrogging* since they managed to jump
+at the top of the hierarchy despite their slow start.
+
+There is extensive
+literature regarding the potential benefits of technological
+leapfrogging. The underpinning argument is that latecomers can adopt and
+benefit from new technologies that have been developed elsewhere without
+incurring the hefty initial R&D costs [@teece2008firm]. Although the
+leapfrogging literature does not pay much attention to cities and
+regions [@yu2018sustainability], previous research highlighted the long
+term and sustained productivity benefit of the early adopters of web
+technologies in the UK [@tranosuk]. So, although the LAD system is
+volatile, it is not clear whether the LAD with high concentration of
+late adopters will gain any latecomer-related benefits in a way similar
+to countries experiencing technological leapfrogging. Still, the
+opportunity to catch up that @perkins2005international illustrated at
+the country level, is also visible at the much more detailed LAD scale.
+As previously, this is consistent with including websites with up to 10
+postcodes (see Figure \ref{rank10}).
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank}Dynamics of wed diffusion (LAD)"}
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av.png")
+knitr::include_graphics(path.image)
+```
 

### 2025-01-21 paper/submission1/tranos2023.tex

@@
   pdfauthor={Emmanouil Tranos},
-  pdfkeywords={keyword1, keyword2},
+  pdfkeywords={diffusion, technologies, Web, digital},
   colorlinks=true,
@@
 
-Table \ref{table.regions} presents the results of the recursive hold out
-models, which aim to highlight the potential regional heterogeneity of
-the spatial processes behind the diffusion of web technologies. To begin
-with, as highlighted before, there is a difference of magnitude of one
-order between the LAD and the OA prediction errors, which is aligned
-with previous results that employed all data points. What is of interest
-here is the regional comparison. Table \ref{table.regions} illustrates
-some striking similarities, but also a few significant differences. The
-regions the web diffusion of which is better predicted using models
-trained in the rest of the country are the same despite the scale of
-analysis: South East, Wales, Yorkshire and The Humber and the North East
-of England. In other words, these are the regions whose spatial
-diffusion mechanisms of web technologies is closer to the country's
-average. Despite the consistency across scales, this is a diverse set of
-regions: \textbf{ADD CHARACTERISTICS}.
-
-At the other end of the spectrum, Scotland's and the North West's web
-diffusion mechanisms are consistently diverging from the country's
-average. This should not come as a surprise as these regions are
-characterised of high levels of rurality and remoteness. Similarly,
-London diffusion mechanisms diverge from the country's average and this
-is consistent across scales. London's uniqueness in UK's urban system
-and economy is also reflected in the spatial diffusion mechanisms of web
-technologies within its LAD and OA. It needs to be highlighted though
-that the difference between the \(R^2\) of LAD and OA is more that an
-order of magnitude signaling how difficult is to predict diffusion at
-such a small spatial scale. Northern Ireland is an interesting case.
-While it ranks at the bottom of the scale when the models are trained
-and tested on LAD data, when the modelling adopts the more granular OA
-scale, the spatial mechanisms that shape the web diffusion within this
-region appear to be closer to the country's average. At this scale,
-proximity, or lack of, relative to the rest of the country become less
-important and the internal to the region spatial structure predictors
-start playing a more important role \textbf{CHECK NI OA model}.
-
 \begin{longtable}[]{@{}lrrrr@{}}

### 2025-01-21 presentations/web_diffusion.Rmd

--- 2024-06-11 presentations/web_diffusion.Rmd
@@
 
-<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture1.png){width="70%"} -->
-<!-- ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture1.png){width="70%"} -->
 ![](images/Picture1.png){width="70%"}
@@
 
-<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture2.png){width="70%"} -->
-<!-- ![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/figure/Picture2.png){width="70%"} -->
 ![](images/Picture2.png){width="70%"}
@@
 -   Curated by the British Library
-
-<!-- <small> -->
-
-<!-- -   Tranos, E., and C. Stich. 2020. Individual internet usage and the -->
-
-<!--     availability of online content of local interest: A multilevel -->
-
-<!--     approach. *Computers, Environment and Urban Systems*, 79:101371. -->
-
-<!-- -   Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy -->
-
-<!--     in the UK: Regional productivity effects of early adoption. -->
-
-<!--     *Regional Studies*, 55:12, 1924-1938. -->
-
-<!-- -   Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from -->
-
-<!--     the ground up: a web data approach. *Environment and Planning B*, in -->
-
-<!--     press. -->
-
-<!-- -   Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to -->
-
-<!--     predict regional trade flows: data extraction, modelling, and -->
-
-<!--     validation, *Annals of the AAG*, in press. -->
-
-<!-- </small> -->
 
@@
 
-<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png) -->
-
-![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)
+![](images/screenshot_directory_website_2010.png)
 
@@
 
-<!-- ![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png) -->
-
-![](/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)
+![](images/screenshot_2010.png)
 
@@
 ```{r UK S, eval=T, out.width = "60%", out.height="60%"}
-#, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"
 
-path.image <- paste0(path, "/outputs/s_uk_per_firm.png")
+path.image <- paste0(path, "/outputs/s/s_uk_per_firm.png")
 knitr::include_graphics(path.image)
@@
 ```{r LA_S_gif, eval=F, results='asis', out.width = "60%"}
-path.images <- paste0(path, "/outputs/s_lad_per_firm/")
+path.images <- paste0(path, "/outputs/s/s_lad_per_firm/")
 
@@
 ```{r Xmid_frequency_graph, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/xmid_f.png")
+path.image <- paste0(path, "/outputs/s/xmid_f.png")
 knitr::include_graphics(path.image)
@@
 
-```{r diffusion_speed_map, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/speed_pam.png")
+```{r diffusion_speed_map, fig.height=5, fig.width=5, message=FALSE, out.width = "100%"}
+path.image <- paste0(path, "/outputs/s/speed_map.png")
 knitr::include_graphics(path.image)
@@
 
-<!-- fast  -->
-
-```{r Darlington, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000005.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r Nottingham, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000018.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
 ```{r Reading, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000038.png")
+path.image <- paste0(path, "/outputs/s/s_lad_per_firm/lad_E06000038.png")
 knitr::include_graphics(path.image)
@@
 ```{r City_of_London, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000001.png")
+path.image <- paste0(path, "/outputs/s/s_lad_per_firm/lad_E09000001.png")
 knitr::include_graphics(path.image)
@@
 ```{r Middlesbrough, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000002.png")
+path.image <- paste0(path, "/outputs/s/s_lad_per_firm/lad_E06000002.png")
 knitr::include_graphics(path.image)
@@
 
-```{r Leicester, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000016.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
 ```{r Bristol, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000023.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r Plymouth, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000026.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r Liverpool, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E08000012.png")
+path.image <- paste0(path, "/outputs/s/s_lad_per_firm/lad_E06000023.png")
 knitr::include_graphics(path.image)
@@
 ```{r Hackney, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000012.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- ------------------------------------------------------------------------ -->
-
-```{r Edinburgh, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
-path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_S12000036.png")
+path.image <- paste0(path, "/outputs/s/s_lad_per_firm/lad_E09000012.png")
 knitr::include_graphics(path.image)
@@
 ```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
-path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only05.png")
-knitr::include_graphics(path.image)
-```
-
-------------------------------------------------------------------------
-
-```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
-path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only95.png")
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av.png")
 knitr::include_graphics(path.image)
@@
 
-<!-- ------------------------------------------------------------------------ -->
+------------------------------------------------------------------------
 
-```{r, eval=FALSE, heatmaps, results='asis', out.width = "60%"}
-path.images <- paste0(path, "/outputs/")
+> **Neighbourhood effect**: diffusion proceeds outwards from innovation
+> centers, first "hitting" nearby rather than far-away locations
+> [@grubler1990rise]
 
-list.files(path = path.images, pattern = "^hex1", full.names = T) %>% 
-  map(image_read) %>% # reads each path file
-  image_join() %>% # joins image
-  image_animate(delay=50) #%>% # animates, can opt for number of loops
-  #image_write("test.gif") # write to current dir
+-   Spatial dependency (Moran's I & LISA maps)
+
+-   Websites per firm in Local authorities (c. 400)
+
+-   Websites in Output Areas (c. 230,000)
+
+------------------------------------------------------------------------
+
+```{r moransi, results='asis'}
+path.in.la <- paste0(path, "/outputs/lisa/corrected/LA/morani_la.csv")
+path.in.oa <- paste0(path, "/outputs/lisa/corrected/oa/morani_oa.csv")
+
+la.m <- read_csv(path.in.la) %>% 
+  filter(p < 0.001) %>% # all values were significant
+  mutate(spatial.unit = 'LAD')
+oa.m <- read_csv(path.in.oa) %>% 
+  filter(p < 0.001) %>% # all values were significant
+  mutate(spatial.unit = 'OA')
+
+bind_rows(la.m, oa.m) %>% 
+  ggplot(aes(x=year, y=morani, fill= spatial.unit)) +
+  geom_bar(stat = 'identity', position = 'dodge') +
+  ylab('Moran\'s I') +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_fill_manual('Spatial unit', 
+                    values = c("LAD" = "grey80",
+                                "OA"="grey48"),
+                     limits = force) +
+  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
+  theme(legend.position="bottom",
+        text = element_text(size = 12),
+        legend.text=element_text(size=12),
+        legend.title=element_text(size=12),
+        axis.title=element_text(size=12),
+        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
 ```
 
-<!-- ------------------------------------------------------------------------ -->
+------------------------------------------------------------------------
 
-```{r eval = F, heatmaps_log, results='asis', out.width = "60%"}
-path.images <- paste0(path, "/outputs/")
+```{r lisa_la, results='asis', out.width = "60%"}
+path.images <- paste0(path, "/outputs/lisa/corrected/la/")
 
-list.files(path = path.images, pattern = "^hex_log1", full.names = T) %>% 
+list.files(path = path.images, pattern = "^lisa_level_pc1_la", full.names = T)[-c(1,2)] %>%
   map(image_read) %>% # reads each path file
@@
 
-> **Neighbourhood effect**: diffusion proceeds outwards from innovation
-> centers, first "hitting" nearby rather than far-away locations
-> [@grubler1990rise]
-
--   Spatial dependency (Moran's I & LISA maps)
-
-<!-- -   Website density regressions -- distance effect -->
-
--   Websites per firm in Local authorities (c. 400)
-
--   Websites in Output Areas (c. 230,000)
-
-------------------------------------------------------------------------
-
-```{r moransi_la, results='asis'}
-path.in <- paste0(path, "/outputs/lisa/la_ggplot/morani_la.csv")
-
-read.csv(path.in) %>% dplyr::select(year, morani) %>% 
-  ggplot(aes(x=year, y=morani)) +
-  geom_bar(stat = "identity") + 
-  ylab('Moran\'s I') +
-  theme_minimal() +
-  ggtitle("Moran\'s I, websites per firm a Local Authority") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-```
-
-------------------------------------------------------------------------
-
-```{r, eval=T, moransi_oa, results='asis'}
-path.in <- paste0(path, "/outputs/lisa/oa_ggplot/morani_oa.csv")
-# read.csv(path.in) %>% dplyr::select(year, morani) %>% 
-#   rename('Moran\'s I' = morani,
-#          Year = year) %>% 
-#   arrange(Year) %>% 
-#   kable(digits = 3, 
-#         label = "Moran\'s I per year, Output Areas") %>% 
-#     kable_classic(full_width = F, html_font = "Cambria")
-
-read.csv(path.in) %>% dplyr::select(year, morani) %>% 
-  ggplot(aes(x=year, y=morani)) +
-  geom_bar(stat = "identity") + 
-  ylab('Moran\'s I') +
-  theme_minimal() +
-  ggtitle("Moran\'s I, websites per Output Area") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-```
-
-------------------------------------------------------------------------
-
-```{r lisa_la, results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/la_ggplot/lisa/")
-
-list.files(path = path.images, pattern = "^lisa_level_pc1_la", full.names = T) %>% 
-  map(image_read) %>% # reads each path file
-  image_join() %>% # joins image
-  image_animate(delay=50) #%>% # animates, can opt for number of loops
-  #image_write("test.gif") # write to current dir
-```
-
-------------------------------------------------------------------------
-
-```{r lisa_oa, eval=T,results='asis', out.width = "30%"}
-path.images <- paste0(path, "/outputs/oa_ggplot/lisa/")
+```{r lisa_oa, eval=T,results='asis', out.width = "60%"}
+path.images <- paste0(path, "/outputs/lisa/corrected/oa/")
 
@@
 ```
-
-<!-- ## Website density regressions -->
-
-<!-- <br> -->
-
-<!-- $$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$ -->
-
-<!-- <br> -->
-
-<!-- ::: {style="margin-left: 50px !important; text-align:left !important"} -->
-
-<!-- $Website\,Density_{i}$: -->
-
-<!-- -   Websites per firm in a Local Authority $i$, *or* -->
-
-<!-- -   Websites in an Output Area $i$ -->
-
-<!-- ::: -->
-
-<!-- ## Website density regressions -->
-
-<!-- <br> -->
-
-<!-- $$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$ -->
-
-<!-- <br> -->
-
-<!-- ::: {style="margin-left: 50px !important; text-align:left !important"} -->
-
-<!-- $Place$: -->
-
-<!-- -   London, *or* -->
-
-<!-- -   Nearest city, *or* -->
-
-<!-- -   Nearest retail centre -->
-
-<!-- ::: -->
-
-<!-- ## Website density regressions -->
-
-<!-- <br> -->
-
-<!-- $\beta$ interpretation: -->
-
-<!-- -   The lower the $\beta$ is (or the larger the $|\beta|$ is)... -->
-
-<!-- -   ... the larger urban gravitation is for web adoption. -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_London_la, fig.height=5, fig.width=5, message=FALSE} -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_london_la.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_city_la, fig.height=5, fig.width=5, message=FALSE} -->
-
-<!-- #, out.width = "60%" -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_city_la.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_centre_la,  fig.height=5, fig.width=5,message=FALSE} -->
-
-<!-- #, out.height= "60%", out.width="60%" -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_centre_la.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_London_oa, eval=T, fig.height=5, fig.width=5, message=FALSE} -->
-
-<!-- #, out.width = "60%" -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_london_oa.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_centre_oa, eval = T,  fig.height=5, fig.width=5,message=FALSE} -->
-
-<!-- #fig.height=4, fig.width=4,  -->
-
-<!-- #, out.height= "60%", out.width="60%" -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_city_oa.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r b_city_oa, eval = T, fig.height=5, fig.width=5, message=FALSE} -->
-
-<!-- #, out.width = "60%" -->
-
-<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_centre_oa.png") -->
-
-<!-- knitr::include_graphics(path.image) -->
-
-<!-- ``` -->
-
-<!-- ------------------------------------------------------------------------ -->
-
-<!-- ```{r r_squared, eval=T, results='asis'} -->
-
-<!-- # This is for the RSquared of the different website density regressions. -->
-
-<!-- path.in1 <- paste0(path, "/outputs/lisa/r2_all_oa.csv") -->
-
-<!-- path.in2 <- paste0(path, "/outputs/lisa/r2_all_la.csv") -->
-
-<!-- # read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>%  -->
-
-<!-- #   ggplot(aes(x=year)) + -->
-
-<!-- #   geom_line(aes(y=London, color = "OA to London"), size=1) + -->
-
-<!-- #   geom_line(aes(y=cities, color = "OA to nearest city"), size=1) + -->
-
-<!-- #   geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) + -->
-
-<!-- #  -->
-
-<!-- #   geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) + -->
-
-<!-- #   geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) + -->
-
-<!-- #   geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) + -->
-
-<!-- #  -->
-
-<!-- #   scale_color_discrete(name = "R-squared for regression:") + -->
-
-<!-- #   #scale_linetype_manual(name = "Website density regressions111")  -->
-
-<!-- #   ylab("R-squared") +  -->
-
-<!-- #   ggtitle("Website density regressions") + -->
-
-<!-- #   theme_minimal() + -->
-
-<!-- #   theme(axis.title.x=element_blank(), -->
-
-<!-- #         plot.title = element_text(hjust = 0.5)) -->
-
-<!-- read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>%  -->
-
-<!--   pivot_longer(!year, names_to = "regression", values_to = "Rsquared") %>%  -->
-
-<!--   mutate(LA = if_else(str_ends(regression, ".LA"), " to LA", "to OA"), -->
-
-<!--          regression = str_remove_all(regression, ".LA"), -->
-
-<!--          line = paste0(regression, " ", LA)) %>%  -->
-
-<!--   ggplot() + -->
-
-<!--   geom_line(aes(x=year, y = Rsquared, color = line), -->
-
-<!--             size = 1) + -->
-
-<!--   scale_color_discrete(name = "R-squared for regression:") + -->
-
-<!--   #scale_linetype_manual(name = "Website density regressions111")  -->
-
-<!--   ylab("R-squared") +  -->
-
-<!--   ggtitle("Website density regressions") + -->
-
-<!--   theme_minimal() + -->
-
-<!--   theme(axis.title.x=element_blank(), -->
-
-<!--         plot.title = element_text(hjust = 0.5)) -->
-
-<!-- #   scale_color_manual(values = c("red", "blue", "green")) -->
-
-<!-- ``` -->
 
@@
 
-```{r gini_la, results='asis'}
-path.in <- paste0(path, "/outputs/lisa/gini_la.csv")
+```{r gini, results='asis'}
+path.in.la <- paste0(path, "/outputs/gini/gini_la.csv")
+path.in.oa <- paste0(path, "/outputs/gini/gini_oa.csv")
 
-read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
-  ggplot(aes(x=year, y=gini.a.n.)) +
-  geom_bar(stat = "identity") + 
-  ylab("Gini") +
+la.gini <- read_csv(path.in.la) %>% 
+  mutate(spatial.unit = 'LAD')
+oa.gini <- read_csv(path.in.oa) %>% 
+  mutate(spatial.unit = 'OA')
+
+bind_rows(la.gini, oa.gini) %>%
+  rename(Gini = 'gini(a$n)') %>% 
+  ggplot(aes(x=year, y=Gini, fill= spatial.unit)) +
+  geom_bar(stat = 'identity', position = 'dodge') +
+  ylab('Gini') +
   theme_minimal() +
-  ggtitle("Gini coefficient, websites per firm in Local Auhtority") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-```
-
-------------------------------------------------------------------------
-
-```{r gini_oa, eval=T, results='asis'}
-path.in <- paste0(path, "/outputs/lisa/gini_oa.csv")
-
-read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
-  ggplot(aes(x=year, y=gini.a.n.)) +
-  geom_bar(stat = "identity") + 
-  ylab("Gini") +
-  theme_minimal() +
-  ggtitle("Gini coefficient, websites per Output Area") +
-  theme(plot.title = element_text(hjust = 0.5)) 
+  theme(plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_fill_manual('Spatial unit', 
+                    values = c("LAD" = "grey80",
+                                "OA"="grey48"),
+                     limits = force) +
+  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
+  theme(legend.position="bottom",
+        text = element_text(size = 12),
+        legend.text=element_text(size=12),
+        legend.title=element_text(size=12),
+        axis.title=element_text(size=12),
+        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
 ```
@@
 
-|                     | RMSE  | RSquared | MAE   |
-|---------------------|-------|----------|-------|
-| Local Authorities\* | 0.032 | 0.810    | 0.019 |
-| Output Areas        | 5.000 | 0.205    | 1.047 |
+|                    | RMSE  | $R^{2}$   | MAE   |
+|--------------------|-------|-----------|-------|
+| Local Authorities\*| 0.028 | 0.850     | 0.019 |
+| Output Areas       | 3.284 | 0.320     | 1.034 |
 
@@
 
-## Variable importance for Local Authorities
+## Variable importance
 
-```{r varimpLA, eval = T, out.height= "58%", out.width="58%", message=FALSE}
+```{r varimp, eval = T, out.height= "58%", out.width="58%", message=FALSE}
+path.importance.lad <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.csv")
+path.importance.oa <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.csv")
 
-path.image <- paste0(path, "/outputs/rf/figures/varimp_LA.png")
-knitr::include_graphics(path.image)
-```
+importance.oa <- read_csv(path.importance.oa) %>% 
+  mutate(id = c("LD", "NC", "NR", "STL", "Y", "DL", "DNC", "DNR")) %>% 
+  mutate(variable = str_to_sentence(variable))
+importance.lad <- read_csv(path.importance.lad) %>% 
+  mutate(id = c("LD", "NC", "STL", "Y", "DL", "DNC")) %>% 
+  mutate(variable = str_to_sentence(variable))
 
-## Variable importance for Output Areas
-
-```{r varimpOA, eval = T, out.height= "58%", out.width="58%", message=FALSE}
-#,  fig.height=5, fig.width=5,message=FALSE
-
-path.image <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
-knitr::include_graphics(path.image)
+importance.oa %>% 
+  rename(OA = importance) %>% 
+  left_join(importance.lad %>% rename(LAD = importance), by = "id") %>%
+  dplyr::select(-variable.y, -id) %>% 
+  rename(variable = variable.x) %>% 
+  pivot_longer(!variable, names_to = "spatial.unit", values_to = "importance") %>% 
+  mutate(order = ifelse(variable == "Distance to london", 1,
+                        ifelse(variable == "London's wensite density, t-1", 2,
+                               ifelse(variable == "Distance to the nearest city", 3,
+                                      ifelse(variable == "Nearest city's wensite density, t-1", 4,
+                                             ifelse(variable == "Distance to the nearest retail centre", 5,
+                                                    ifelse(variable == "Nearest retail centre's wensite density, t-1", 6,
+                                                           ifelse(variable == "Spatial and temporal lag of wensite density", 7, 8)))))))) %>% 
+  ggplot(aes(x=importance, y=reorder(variable, desc(order)), #forcats::fct_reorder(variable, importance), 
+         fill = spatial.unit)) +
+  geom_col(position = 'dodge') +
+  ylab('') +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5), 
+        axis.text.x = element_text(vjust = 0.5, hjust=1)) +
+  scale_fill_manual('Spatial unit: ', 
+                    values = c("LAD" = "grey80",
+                                "OA"="grey48"),
+                     limits = force) +
+  scale_x_continuous('Variable importance') +
+  theme(legend.position="bottom",
+        text = element_text(size = 14),
+        legend.text=element_text(size=14),
+        legend.title=element_text(size=14),
+        axis.title=element_text(size=14))
 ```
@@
 ```{r similarity, eval = T, message=FALSE}
-path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
-path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
+path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected.csv")
+path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA_corrected.csv")
 
@@
 
-## Alternative specification: growth, trained on all data
+<!-- ## Alternative specification: growth, trained on all data -->
 
-<br> <small>
-$$Website\,Density\,Growth_{t} \sim \color{orange}{Distance\,London} + 
-\color{orange}{Website\,density\,London_{t-1}} + \\
-\color{orange}{Distance\,Nearest\,City} + 
-\color{orange}{Website\,density\,Nearest\,City_{t-1}} + \\
-\color{orange}{Distance\,Nearest\,Retail_{i}} +
-\color{orange}{Website\,density\,Nearest\,Retail_{t-1}} +\\
-\color{olive}{W*\, Website\,density_{t-1}} +\\ 
-\color{violet}{year_{t}}$$ </small>
+<!-- <br> <small> -->
+<!-- $$Website\,Density\,Growth_{t} \sim \color{orange}{Distance\,London} +  -->
+<!-- \color{orange}{Website\,density\,London_{t-1}} + \\ -->
+<!-- \color{orange}{Distance\,Nearest\,City} +  -->
+<!-- \color{orange}{Website\,density\,Nearest\,City_{t-1}} + \\ -->
+<!-- \color{orange}{Distance\,Nearest\,Retail_{i}} + -->
+<!-- \color{orange}{Website\,density\,Nearest\,Retail_{t-1}} +\\ -->
+<!-- \color{olive}{W*\, Website\,density_{t-1}} +\\  -->
+<!-- \color{violet}{year_{t}}$$ </small> -->
 
-<br>
+<!-- <br> -->
 
-|                     | RMSE  | RSquared | MAE   |
-|---------------------|-------|----------|-------|
-| Local Authorities\* | 0.200 | 0.634    | 0.148 |
+<!-- |                     | RMSE  | RSquared | MAE   | -->
+<!-- |---------------------|-------|----------|-------| -->
+<!-- | Local Authorities\* | 0.200 | 0.634    | 0.148 | -->
 
-<small> \*No retail centre predictors for the Local Authority models
-</small>
+<!-- <small> \*No retail centre predictors for the Local Authority models -->
+<!-- </small> -->
 
-## Alternative specification: growth, trained on all data
+<!-- ## Alternative specification: growth, trained on all data -->
 
-```{r growth LA}
-path.table <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv")
-read.csv(path.table) %>% 
-  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
-  arrange(desc(Rsquared)) %>% 
-  kable()
-```
+<!-- ```{r growth LA} -->
+<!-- path.table <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv") -->
+<!-- read.csv(path.table) %>%  -->
+<!--   mutate(Rsquared = round(Rsquared, digits= 3)) %>%  -->
+<!--   arrange(desc(Rsquared)) %>%  -->
+<!--   kable() -->
+<!-- ``` -->
 

### 2025-01-21 src/help_memory_log.Rmd

```{r}
#mem <- 
read_delim("memory.log", delim = "\\t") %>%
  rename(t = 1) %>% 
  filter(!str_detect(t, 'total')) %>% 
  separate(t, sep = ":", into = c("type", "b")) %>% 
  filter(type == "Mem") %>% 
  separate(b, sep = "\\S*\\K\\s+", into = c("total", "used", "free", "shared", "buff/cache", "available")) %>% 
  arrange(shared)
  #extract(b, c("total", "used", "free", "shared", "buff/cache", "available"), "(\\d+)\\s(\\d+)\\s(\\d+)\\s(\\d+)\\s(\\d+)\\s(\\d+)\\s")

read_delim("memory.log", delim = "\\t") %>%
  rename(t = 1) %>% 
  filter(!str_detect(t, 'total')) %>% 
  separate(t, sep = ":", into = c("type", "b")) %>% 
  filter(type != "Mem") %>% 
  separate(b, sep = "\\S*\\K\\s+", into = c("total", "used", "free")) %>% 
  arrange(desc(free))
```

### 2025-01-21 src/la_RF_corrected_10.Rmd

@@
 indices <- CreateSpacetimeFolds(df,# %>% filter(year > 1999),                   # FILTER BY YEAR FOR GROWTH
-                                spacevar = "ladcd",
+                                spacevar = "RGN21CD",
                                 timevar = "year", 

### 2025-01-21 src/old/oa_RF_OLD_rgeos.Rmd

--- 2024-01-27 src/oa_RF.Rmd
@@
 library(rprojroot)
-#library(rgdal)
-#library(rgeos)
+library(rgdal)
+library(rgeos)
 library(sf)
@@
 library(randomForest)
-library(CAST)
 
@@
 
-# Partially complete panel as not all OAs have a website in at least one year
+# Partially complete panel as not all OAs have a website in at least ine year
 df <- df %>% filter(!is.na(oa11cd)) %>% 
@@
 
-## spatial data OLD
-
-```{r eval=F}
+## spatial data
+
+```{r}
 # get OA for England and Wales
@@
 
-## spatial data NEW
-
-```{r}
-library(sf)
-
-# Get OA for England and Wales
-path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
-oa.ew <- st_read(path.geo)
-
-# Spatial transformations
-oa.ew <- st_transform(oa.ew, 4326)  # EPSG code for WGS84
-
-# Rename columns and keep only necessary columns
-oa.ew <- oa.ew %>% dplyr::select(OA11CD, Shape__Area) %>% 
-  rename(id = OA11CD,
-         area = Shape__Area)
-# oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
-# colnames(oa.ew) <- c('id', 'area')
-
-# Check the data structure
-print(summary(oa.ew))
-
-# Get OA for Scotland
-path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
-oa.sc <- st_read(dsn = path.geo.sc, layer = "OutputArea2011_MHW")
-
-# Spatial transformations
-oa.sc <- st_transform(oa.sc, 4326)  # EPSG code for WGS84
-
-# Rename columns and keep only necessary columns
-oa.sc <- oa.sc %>% dplyr::select(code, SHAPE_1_Ar) %>% 
-  rename(id = code,
-         area = SHAPE_1_Ar)
-# oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
-# colnames(oa.sc) <- c('id', 'area')
-
-# Check the data structure
-print(summary(oa.sc))
-
-# NI
-path.geo.ni <- paste0(path, "/data/raw/ni_small_area/")
-oa.ni <- st_read(dsn = path.geo.ni, layer = "SA2011")
-oa.ni <- st_transform(oa.ni, 4326)  # EPSG code for WGS84
-oa.ni <- oa.ni %>% dplyr::select(SA2011, Hectares) %>% 
-  rename(id= SA2011,
-         area = Hectares)
-
-# Build OA for UK
-oa.uk <- rbind(oa.ew, oa.sc, oa.ni)
-rm(oa.ew, oa.sc, oa.ni)
-
-# Get GB
-path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
-gb <- st_read(path.gb)
-
-# Spatial transformations
-gb <- st_transform(gb, 4326)  # EPSG code for WGS84
-```
-
 ## Full panel
@@
 ```{r}
-#oa.sf <- st_as_sf(oa.uk)
-
-oa.full <- replicate(17, oa.uk %>% as_tibble(), simplify = FALSE) %>%
+oa.sf <- st_as_sf(oa.gb)
+
+oa.full <- replicate(17, oa.sf %>% as_tibble(), simplify = FALSE) %>%
   bind_rows(.id = 'id_') %>% 
@@
 
-### Cities as points
-
 First set up the destinations. Origins are all the OA.
@@
 
-I use the centroids from arcgis API as they better reflect the city centres than 
-than the geometric centres of the boundaries, which I use for `sum(n)`
-
 ```{r}
 # oa centroids
-# oa.uk.c <- (rgeos::gCentroid(oa.uk, byid=TRUE))@coords %>% as.data.frame()
-# oa.uk.c$id <- row.names(oa.uk.c)  #as_tibble(rownames = NA) %>% 
+# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
+# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
 #   mutate(id = row.names())
-
-# oa.uk.c <- gCentroid(oa.uk, byid=TRUE)@coords %>% 
-#   as.data.frame() %>% 
-#   add_rownames(var = "id")
-sf_use_s2(FALSE)
-oa.uk.c <- st_centroid(oa.uk) %>%
-  st_coordinates() %>%
+oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
   as.data.frame() %>% 
-  #add_rownames(var = "id") %>% 
-  bind_cols(oa.uk %>% dplyr::select(id))
+  add_rownames(var = "id")
 
@@
 "Edinburgh, UK",
-"Leicester, UK",
-"Belfast, UK")
-
-cities <- geo(city.names, no_query = F, method = "arcgis") %>% 
-  mutate(address = stringr::str_remove(address, ", UK"))
-
-# # Replace London with Inner London centroid
-# # Source: https://data.london.gov.uk/dataset/sub-regions-london-plan-consultation-2009
-# london.path <- paste0(path, "/data/raw/London_shape")
-# london.sf <- st_read(dsn = london.path, layer = "lp-consultation-oct-2009-subregions")
-# # Spatial transformations
-# london.sf <- st_transform(london.sf, 4326)  # EPSG code for WGS84
-# 
-# london.c <- london.sf %>% 
-#   filter(Name == "Central") %>% 
-#   st_centroid() %>% 
-#   st_coordinates() %>%
-#   as.data.frame() %>% 
-#   rename(lat = Y, long = X) %>% 
-#   mutate(address = "London, UK") %>% 
-#   dplyr::select(address, lat, long)
-# 
-# cities <- cities %>% bind_rows(london.c) %>% 
-#   slice(2:n()) %>% # to remove the old London centroid based on the cities geometry
-#   mutate(address = stringr::str_remove(address, ", UK"))
-```
-
-### City boundaries
-
-```{r}
-
-# # cities
-# city.names <- c(
-#   "London", 
-#   "Birmingham",
-#   "Glasgow",
-#   "Liverpool",
-#   "Bristol",
-#   "Manchester",
-#   "Sheffield",
-#   "Leeds",
-#   "Edinburgh",
-#   "Leicester",
-#   )
-
-# City boundaries
-# Source: https://geoportal.statistics.gov.uk/search?collection=Dataset&sort=name&tags=all(BDY_TCITY%2CDEC_2015)
-path.in <- paste0(path, "/data/raw/Major_Towns_and_Cities_Dec_2015_Boundaries_V2_2022_121402779136482658.geojson")
+"Leicester, UK")
+
+cities <- geo(city.names, no_query = F, method = "arcgis")
+
+# retail centres
+geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
+retail <-readOGR(geo.path) 
+#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
+retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
+  as.data.frame() %>% 
+  add_rownames(var = "id") #%>%
+
+retail.major.cetres <- retail.major.cetres.help %>% 
+  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
+  st_transform(4326) %>%
+  st_coordinates() %>%
+  as_tibble() %>% 
+  bind_cols(retail.major.cetres.help$id) %>% 
+  rename(id = '...3') %>% 
+  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
+```
+
+Distance to cities
+
+```{r}
+dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
+dist <- round((dist/1000),2) %>% 
+  as_tibble()
 
 city.names <- city.names %>% stringr::str_remove(", UK")
-
-city.boundaries.ew <- st_read(path.in) %>% 
-  st_transform(4326) %>% 
-  filter(TCITY15NM %in% city.names) %>% 
-  dplyr::select(OBJECTID, TCITY15CD, TCITY15NM) %>% 
-  rename(code = TCITY15CD,
-         name = TCITY15NM)
-
-# Source: https://geoportal.statistics.gov.uk/
-path.in <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
-city.boundaries.sc.ni <- st_read(path.in) %>% 
-  st_transform(4326) %>% 
-  filter(grepl('Edinburgh|Glasgow|Belfast', LAD21NM)) %>% 
-  dplyr::select(OBJECTID, LAD21CD, LAD21NM) %>% 
-  rename(code = LAD21CD,
-         name = LAD21NM)
-
-# merge
-city.boundaries <- rbind(city.boundaries.ew, city.boundaries.sc.ni)
-```
-
-### Retail centres
-
-```{r}
-# retail centres <- 
-geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
-#retail <-readOGR(geo.path) 
-retail <- st_read(geo.path) %>% st_transform(4326)  # EPSG code for WGS84
-
-#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
-# retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
-#   as.data.frame() %>% 
-#   add_rownames(var = "id") #%>%
-retail.major.cetres.help <- st_centroid(retail) %>% 
-  #st_transform(4326) %>%
-  st_coordinates() %>% 
-  as.data.frame() %>% 
-  add_rownames(var = "id")
-
-# retail.major.cetres <- retail.major.cetres.help %>% 
-#   st_as_sf(coords = c("x", "y"), crs = 27700) %>%
-#   st_transform(4326) %>%
-#   st_coordinates() %>%
-#   as_tibble() %>% 
-#   bind_cols(retail.major.cetres.help$id) %>% 
-#   rename(id = '...3') %>% 
-#   left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
-retail.major.cetres <- retail.major.cetres.help %>% 
-  # as_tibble() %>% 
-  # bind_cols(retail.major.cetres.help$id) %>% 
-  # #rename(id = '...3') %>% 
-  left_join(retail %>% rownames_to_column(var = "id"), by = 'id')
-```
-
-### Distance to cities
-
-```{r}
-# NI
-oa.uk.c.ni <- oa.uk.c %>% filter(str_detect(id, "^N"))
-
-dist.ni <- distm(cbind(oa.uk.c.ni$X, oa.uk.c.ni$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist.ni <- round((dist.ni/1000),2) %>% 
-  as_tibble()
-
-names(dist.ni) <- cities$address #city.names
-
-dist.ni <- dist.ni %>% bind_cols(oa.uk.c.ni$id) %>% 
+names(dist) <- city.names
+
+dist <- dist %>% bind_cols(oa.gb.c$id) %>% 
   rename(oa11cd = last_col()) %>% 
@@
 
-# min distance
-dist.ni <- transform(dist.ni, dist = do.call(pmin, dist.ni[-1]))
-
-# City name for minimum distance 
-dist.ni <- dist.ni %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
-
-# GB
-oa.uk.c.gb <- oa.uk.c %>% filter(!str_detect(id, "^N"))
-
-dist.gb <- distm(cbind(oa.uk.c.gb$X, oa.uk.c.gb$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
-dist.gb <- round((dist.gb/1000),2) %>% 
+dist <- transform(dist, dist = do.call(pmin, dist[-1]))
+
+# NOT USED AS IT SCREWS UP THE REGRESSIONS
+# Complete panel, N = 3871903, 17 year * 227759 OA
+# df <- df %>% bind_rows(dist %>% dplyr::select(oa11cd)) %>%
+#   mutate(year = ifelse(is.na(year),1000, year),
+#          country = substring(oa11cd, 1, 1)) %>%
+#   filter(country != "L" & country != "M" & country != "N") %>%
+#   filter(!is.na(country)) %>%
+#   complete(year, oa11cd, fill = list(n = 0)) %>%
+#   filter(year!=1000,
+#          !is.na(oa11cd)) %>%
+#   dplyr::select(-country)
+
+# tests
+# test$country <- substring(test$oa11cd, 1, 1)
+# unique(test$country)
+# length(unique(test$oa11cd))
+
+# calculate area
+oa.gb$area <- area(oa.gb) #/ 1000000
+
+# Join with distance and area
+df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
+  rename(oa11cd=id) %>% 
+  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
+  relocate(area, .after = n) %>% 
+  mutate(area = area / 1000000,
+         density = n / area) %>% 
+  relocate(density, .after = area)
+
+sapply(df, function(x) sum(is.na(x)))
+```
+
+Distance to retail centres
+
+```{r}
+dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
+dist.retail <- round((dist.retail/1000),2) %>% 
   as_tibble()
 
-names(dist.gb) <- cities$address #city.names
-
-dist.gb <- dist.gb %>% bind_cols(oa.uk.c.gb$id) %>% 
+# retail.names <- retail.major.cetres$RC_Name
+# names(dist.retail) <- retail.names
+
+dist.retail <- dist.retail %>% bind_cols(oa.gb.c$id) %>% 
   rename(oa11cd = last_col()) %>% 
@@
 
-# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
-# dist$distMet <- apply(dist[,2:11], 1, min)
-
-# min distance
-dist.gb <- transform(dist.gb, dist = do.call(pmin, dist.gb[-1]))
-
-# City name for minimum distance 
-dist.gb <- dist.gb %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
-
-# UK
-dist <- rbind(dist.ni, dist.gb)
-
-# calculate area
-#oa.uk$area <- st_area(oa.uk) #/ 1000000
-
-# Join with distance and area
-df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
-  rename(oa11cd=id)
-
-sapply(df, function(x) sum(is.na(x)))
-```
-
-### Distance to retail centres
-
-```{r}
-# dist.retail <- distm(cbind(oa.uk.c$X, oa.uk.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-# dist.retail <- round((dist.retail/1000),2) %>% 
-#   as_tibble()
-
-# NI
-oa.uk.c.ni <- oa.uk.c %>% filter(str_detect(id, "^N"))
-dist.retail.ni <- distm(cbind(oa.uk.c.ni$X, oa.uk.c.ni$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail.ni <- round((dist.retail.ni/1000),2) %>% 
-  as_tibble()
-
-names(dist.retail.ni) <- retail.major.cetres$RC_ID
-
-dist.retail.ni <- dist.retail.ni %>% bind_cols(oa.uk.c.ni$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# This efficiently returns the name of the column with the shortest distance
-dist.retail.ni <- dist.retail.ni %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])
-
-# dist.retail_ <- dist.retail %>% slice_sample(n=100)
-# dist.retail_[,6425:6426]
-# dist.retail_ <- transform(dist.retail_, dist.retail = do.call(pmin, dist.retail_[2:6424])) 
-# sum(is.na(dist.retail_$dist.retail.name))
-
 # Minimum distance 
-dist.retail.ni <- transform(dist.retail.ni, dist.retail = do.call(pmin, dist.retail.ni[2:6424])) 
-dist.retail.ni <- dist.retail.ni  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)
-
-# GB
-oa.uk.c.gb <- oa.uk.c %>% filter(!str_detect(id, "^N"))
-dist.retail.gb <- distm(cbind(oa.uk.c.gb$X, oa.uk.c.gb$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
-dist.retail.gb <- round((dist.retail.gb/1000),2) %>% 
-  as_tibble()
-
-names(dist.retail.gb) <- retail.major.cetres$RC_ID
-
-dist.retail.gb <- dist.retail.gb %>% bind_cols(oa.uk.c.gb$id) %>% 
-  rename(oa11cd = last_col()) %>% 
-  relocate(oa11cd)
-
-# This efficiently returns the name of the column with the shortest distance
-dist.retail.gb <- dist.retail.gb %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])
-
-# dist.retail_ <- dist.retail %>% slice_sample(n=100)
-# dist.retail_[,6425:6426]
-# dist.retail_ <- transform(dist.retail_, dist.retail = do.call(pmin, dist.retail_[2:6424])) 
-# sum(is.na(dist.retail_$dist.retail.name))
-
-# Minimum distance 
-dist.retail.gb <- transform(dist.retail.gb, dist.retail = do.call(pmin, dist.retail.gb[2:6424])) 
-dist.retail.gb <- dist.retail.gb  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)
-
-# UK
-dist.retail <- rbind(dist.retail.ni, dist.retail.gb)
+dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
+dist.retail <- dist.retail %>% dplyr::select(oa11cd, dist.retail)
 
@@
 df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd"))
-#sapply(df, function(x) sum(is.na(x)))
-```
-
-## n for London, nearest city and retail
-
-### TODELETE
-
-```{r, eval=FALSE}
-# ## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
-# #df <- df %>% filter(!is.na(area)) 
-# 
-# cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
-# cities.sf <- st_set_crs(cities.sf, 4326)
-# 
-# # help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
-# #   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
-# 
-# sf_use_s2(FALSE)
-# help1 <- st_filter( oa.gb, cities.sf) %>% 
-#   #st_coordinates() %>% 
-#   as_tibble() %>% 
-#   dplyr::select(id) %>% 
-#   mutate(cities = c("London",     # The city names have been manually checked and assigned
-#                     "Leicester",
-#                     "Leeds",
-#                     "Sheffield",
-#                     "Bristol",
-#                     "Birmingham",
-#                     "Manchester",
-#                     "Liverpool",
-#                     "Edinburgh",
-#                     "Glasgow"))
-# 
-# # too slow
-# #dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
-# 
-# # This efficiently returns the name of the column with the shortest distance
-# dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
-# 
-# help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
-#   #rename(nearest.city.oa11cd = id) %>% 
-#   dplyr::select(oa11cd, dist.city.name)
-# 
-# df <- df %>%  left_join(help2, by = c("oa11cd" = "oa11cd")) 
-```
-
-### n for cities
-
-```{R}
-#help.retail <- dist.retail
-
-# n nearest city
-# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-#                          rename(n.nearest.city = n), 
-#                        by = c("nearest.city.oa11cd" = "oa11cd",
-#                               "year" = "year"))
+
+sapply(df, function(x) sum(is.na(x)))
+```
+
+## df for RF
+
+```{r}
+## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
+#df <- df %>% filter(!is.na(area))
+
+# London = Islington 
+cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
+cities.sf <- st_set_crs(cities.sf, 4326)
+
+# help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
+#   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)
 
 sf_use_s2(FALSE)
-city.boundaries.oa <- st_join(oa.uk, city.boundaries, join = st_intersects) %>% 
-  as_tibble() %>% 
-  filter(!is.na(name)) %>% 
-  dplyr::select(id, name) %>% 
-  mutate(name = ifelse(name == "City of Edinburgh", "Edinburgh", name)) %>% 
-  mutate(name = ifelse(name == "Glasgow City", "Glasgow", name))
-
-city.boundaries.help <- df %>% inner_join(city.boundaries.oa, by = c("oa11cd" = "id")) %>% 
-  group_by(name, year) %>% 
-  summarise(n.nearest.city = sum(n)) %>% 
-  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
-  ungroup()
-
-london.boundaries.help <- city.boundaries.help %>% 
-  filter(name == "London") %>% 
-  rename(n.London = n.nearest.city,
-         n.London.lag = n.nearest.city.lag)
-
-df <- df %>% left_join(city.boundaries.help, by = c("dist.city.name" = "name", "year" = "year"))
-
-df <- df %>% left_join(london.boundaries.help, by = c("year" = "year"))
-
-#sapply(df, function(x) sum(is.na(x)))
-```
-
-### DELETE n for London
-
-```{r, eval=FALSE}
-
-# # OLD London definitions
-# # df <- df %>% left_join(df %>% 
-# #                          ungroup() %>% 
-# #                          filter(oa11cd == "E00176659") %>% 
-# #                          dplyr::select(year, n) %>% 
-# #                          rename(n.London = n)) %>% 
-# #   group_by(oa11cd) %>% 
-# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-# #   mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
-# #   ungroup() 
-# # 
-# # london.help <- df %>% group_by(RGN11NM, year) %>% 
-# #   summarise(n.London = sum(n)) %>% 
-# #   filter(RGN11NM == "London") %>%
-# #   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-# #   ungroup() %>% 
-# #   dplyr::select(-RGN11NM) 
-# 
-# london.oa <- st_filter(oa.gb, london.sf %>% 
-#                          filter(Name == "Central")) %>% 
-#   as_tibble() %>% 
-#   dplyr::select(id)
-# 
-# london.help <- df %>% inner_join(london.oa, by = c("oa11cd" = "id")) %>% 
-#   group_by(year) %>% 
-#   summarise(n.London = sum(n)) %>% 
-#   mutate(n.London.lag = dplyr::lag(n.London)) %>% 
-#   ungroup()
-# 
-# df <- df %>% left_join(london.help, by = c("year" = "year"))
-```
-
-### n for nearest retail
-
-```{r}
-# Old definition for retail centre
-# df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-#                          rename(n.nearest.retail = n), 
-#                        by = c("oa11cd.retail" = "oa11cd",
-#                               "year" = "year")) %>% 
-#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
-# 
-# 
-# #n.nearest.retail IS WRONG. FIX UPSTREAM
-# df <- df %>% 
-#   rename(oa11cd.retail = n.nearest.retail) %>% 
-#   dplyr::select(-n.nearest.retail.lag) %>% 
-#   left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
-#                          rename(n.nearest.retail = n), 
-#                        by = c("oa11cd.retail" = "oa11cd",
-#                               "year" = "year")) %>% 
-#   mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))
-# 
-# 
-# df <- df %>% dplyr::select(-n.London, -n.London.lag) %>% 
-#   left_join(london.help, by = c('year' = 'year'))
-
-retail.oa <- st_join(oa.uk, retail, 
-                     join = st_intersects) %>% #, 
-                    #largest = TRUE) %>% # returns the x features augmented with 
-                                         # the fields of y that have the largest 
-                                         # overlap with each of the features of x
-  as_tibble() %>% 
-  filter(!is.na(RC_ID)) %>% 
-  dplyr::select(id, RC_ID) %>% 
-  add_row(id = "E00019779", RC_ID = "RC_EW_5951")
-
-# If `larget = FALSE`, we have multiple RC_IDs for OAs.
-# Also, RC_ID="RC_EW_5951" would fall in Thames as it is the London Bridge Tube Station.
-# To avoid NAs for for n.nearest.retail, I would have assigned dist.retail.name = RC_EW_1935 for E00019779
-
-# st_join(oa.uk, retail, join = st_is_within_distance, dist = 500) %>% 
-#   as_tibble() %>% 
-#   filter(!is.na(RC_ID)) %>% 
-#   dplyr::select(id, RC_ID)
-# 
-# st_filter(oa.uk, retail,  .pred = st_is_within_distance, dist = 100) %>% 
-#   as_tibble() %>% 
-#   filter(!is.na(RC_ID)) %>% 
-#   dplyr::select(id, RC_ID)
-
-# h <- oa.uk %>% st_intersects(retail)
-
-# retail.oa_ %>% filter(id == "E00004828"|id == "E00004829"|id == "E00004827")
-# # RC_EW_5285
-# retail.oa <- retail.oa %>% add_row(id = "E00004828", RC_ID = "RC_EW_5285")
-# retail.oa <- retail.oa %>% arrange(id)
-
-retail.help <- df %>% full_join(retail.oa, by = c("oa11cd" = "id")) %>% 
-  group_by(RC_ID, year) %>% 
-  summarise(n.nearest.retail = sum(n)) %>% 
-  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail)) %>% 
-  ungroup()
-
-# retail.help2 <- retail.oa %>% left_join(retail.help, by = c("RC_ID" = "RC_ID")) %>% 
-#   dplyr::select(-RC_ID)
-
-df <- df %>% left_join(retail.help, by = c("dist.retail.name" = "RC_ID", "year" = "year"))
-#sapply(df, function(x) sum(is.na(x)))
-```
-
-### Urban Rural TODO
-
-## TODO
-
-```{r}
-# Lookup: OA to regions
-# **FIXED< DELETE IN TEST** NEED TO ADD NI. NOT A PROBLEM FOR NOW AS IT IS FIXED IN THE TEST SECTION
-
-# # regional identifier
-# df_ <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
-#   mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
-#   ungroup() %>% 
-#   dplyr::select(-help)
-
-#sapply(df, function(x) sum(is.na(x)))
-```
-
-## Spatial and spatio-temporal lags
-
-```{r}
-# sf_use_s2(FALSE)
-# oa.c <- st_centroid(oa.uk)
-# kn <- knearneigh(oa.c, k = 5)
-# knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
-# knn.l <- nb2listw(knn)
-
-path.out <- paste0(path, "/data/temp/knnl.RData")
-# save(knn.l, file = path.out)
-load(path.out)
+help1 <- st_filter( oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
+  mutate(cities = c("Leicester",
+                    "Bristol",
+                    "Manchester",
+                    "Liverpool",
+                    "Sheffield",
+                    "Birmingham",
+                    "Leeds",
+                    "London",
+                    "Edinburgh",
+                    "Glasgow"))
+
+# too slow
+#dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
+
+# This efficiently returns the name of the column with the shortest distance
+dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
+
+help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
+  rename(nearest.city.oa11cd = id) %>% 
+  dplyr::select(oa11cd, nearest.city.oa11cd)
+
+#####save this#####
+sf_use_s2(FALSE)
+oa.c <- st_centroid(oa.sf)
+kn <- knearneigh(oa.c, k = 5)
+knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
+knn.l <- nb2listw(knn)
+##################
 
@@
   mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
-         n.lag = dplyr::lag(n, n=1, order_by = year))        # duplicate line
-```
-
-## RF with CAST
-
-```{r}
+         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
+  left_join(help2, by = c("oa11cd" = "oa11cd"))
+
+df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
+                         rename(n.nearest.city = n), 
+                       by = c("nearest.city.oa11cd" = "oa11cd",
+                              "year" = "year"))
 
@@
                          ifelse(help=="W", "Wales",
-                                ifelse(help == "S", "Scotland", "NI"))),
+                                ifelse(help == "S", "Scotland", NA))),
          RGN11NM = ifelse(help=="E", RGN11NM,
                          ifelse(help=="W", "Wales",
-                                ifelse(help == "S", "Scotland", "NI")))) %>% 
+                                ifelse(help == "S", "Scotland", NA)))) %>% 
   dplyr::select(-help)
 
-# # North/South
-# #lookup.region %>% distinct(RGN21NM, .keep_all = T)
-# south <- c("London", "South West", "South West", "East of England")
-# df <- df %>% mutate(south = ifelse(RGN11NM %in% south, 1, 0))
+# df %>% ungroup() %>% distinct(RGN11NM)
+# sapply(df, function(x) sum(is.na(x)))
+
+# North/South
+#lookup.region %>% distinct(RGN21NM, .keep_all = T)
+south <- c("London", "South West", "South West", "East of England")
+df <- df %>% mutate(south = ifelse(RGN11NM %in% south, 1, 0))
 
 # For growth:
-df <- df %>% group_by(oa11cd) %>% 
-  mutate(#n.lag = dplyr::lag(n),
-         growth = log(n/n.lag),
-                    abs.growth = n - n.lag) %>% 
-  relocate(n.lag, .after = n)
+#df <- df %>% mutate(growth = log(n/n.lag))
+
+sapply(df, function(x) sum(is.na(x)))
+
+# out.path <- paste0(path, "/data/temp/oa_rf.RData")
+# save.image(file = out.path)
+# load(out.path)
 
 # out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
-# df <- read_csv(out.path)
-# df %>% write_csv(file = out.path)
-
-# rm(list=setdiff(ls(), "df"))
-
-# # This is the project path
-# path <- find_rstudio_root_file()
-
-# remove not used columns
-df <- df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol",
-                        "Manchester", "Sheffield", "Leeds", "Edinburgh",
-                        "Leicester", "dist.city.name", "dist.retail.name", "name"))
-
-# There are no lags for 1996, so it is removed
-df <- df %>% filter(year > 1996) %>% 
-  dplyr::select(oa11cd, year, RGN11CD, n, abs.growth, n.l.slag, n.nearest.city.lag, 
-                n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) #%>% #growth, abs.growth, RGN11NM, 
-  # mutate(London.v = n.London.lag / London,
-  #        city.v = n.nearest.city.lag / dist,
-  #        retail.v = n.nearest.retail.lag / (dist.retail + .1)) %>% 
-  # dplyr::select(-n.London.lag, -London, -n.nearest.city.lag, -dist, -n.nearest.retail.lag, -dist.retail)
-
-df %>% group_by(year) %>% 
-  filter(n == 0) %>% 
-  count(n)
-# in 2012, 120848 0s
-
-# df %>% dplyr::select(area, dist, dist.retail, London, south, n.slag, n.lag, n.l.slag, n.nearest.city, year) %>% 
-#   #sapply(function(x) sum(is.na(x)))
-#   summarise(across(where(is.numeric), .fns = 
-#                      list(#n = is.na,
-#                           min = min,
-#                           median = median,
-#                           mean = mean,
-#                           stdev = sd,
-#                           # q25 = ~quantile(., 0.25),
-#                           # q75 = ~quantile(., 0.75),
-#                           max = max))) %>%
-#   pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))
-# 
-# df %>% dplyr::select(n, n.London.lag, n.nearest.city.lag, n.nearest.retail.lag, n.l.slag, year, London, dist, dist.retail) %>%
-#   #sapply(function(x) sum(is.na(x)))
-#   summarise(across(where(is.numeric), .fns =
-#                      list(#n = is.na,
-#                           min = min,
-#                           median = median,
-#                           mean = mean,
-#                           stdev = sd,
-#                           # q25 = ~quantile(., 0.25),
-#                           # q75 = ~quantile(., 0.75),
-#                           max = max))) %>%
-#   pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))
-# 
-# df %>% slice_max(order_by = n, n = 5, by = year)
-```
-
-### unregister_dopar function
-
-This is helpful when the parallel loop collapses.
-
-```{r}
-unregister_dopar <- function() {
-  env <- foreach:::.foreachGlobals
-  rm(list=ls(name=env), pos=env)
-}
-unregister_dopar()
-```
-
-### train the model in all data
-
-```{r}
-set.seed(123)
-indices <- CreateSpacetimeFolds(df, 
-                                spacevar = "oa11cd",
-                                timevar = "year", 
-                                k = 10)
-
-#length is = (n_repeats*nresampling)+1
-seeds <- vector(mode = "list", length = 11)
-
-#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # It should be RHS variables + 2
-
-#for the last model
-seeds[[11]]<-sample.int(1000, 1)
-
-# CV
-tc <- trainControl(method = "cv",
-                   number = 10,
-                   seeds = seeds,
-                   allowParallel = T,
-                   index = indices$index,
-                   savePredictions = 'final')
-
-# detectCores()
-# cl <- makePSOCKcluster(2)
-# registerDoParallel(cl)
-
-start.time <- Sys.time()
-model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
-                     n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
-                     #London.v + city.v + retail.v + n.l.slag + year,
-                   data = df, 
-                   trControl = tc,
-                   method = "ranger",
-                   na.action = na.omit,
-                   #preProc = c("center", "scale"),
-                   importance = "impurity")
-
-# stopCluster(cl)
-end.time <- Sys.time()
-time.taken <- round(end.time - start.time,2)
-time.taken #20.6h
-
-print(model.all)
-
-# Random Forest 
-# 
-# 3716736 samples
-#       8 predictor
-# 
-# No pre-processing
-# Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 2926910, 2926924, 3136005, 2926924, 3136005, 2926924, ... 
-# Resampling results across tuning parameters:
-# 
-#   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    4.999920  0.2053697  1.047293
-#   2     extratrees  5.167655  0.1540213  1.103048
-#   5     variance    5.187039  0.1710128  1.034382
-#   5     extratrees  5.169570  0.1557868  1.122845
-#   8     variance    5.633427  0.1247189  1.035171
-#   8     extratrees  5.182860  0.1563555  1.137332
-# 
-# Tuning parameter 'min.node.size' was held constant at a value of 5
-# RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule =
-#  variance and min.node.size = 5.
-
-varimp_mars <- varImp(model.all) 
-varimp_mars$importance <- varimp_mars$importance %>% 
-  rownames_to_column() %>% 
-  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
-                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
-                                 ifelse(rowname == "London", "distance to London",
-                                        ifelse(rowname == "n.nearest.retail.lag", "Nearest retail centre's wensite density, t-1",
-                                               ifelse(rowname == "n.London.lag", "London's wensite density, t-1",
-                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag of wensite density",
-                                                             ifelse(rowname == "n.nearest.city", "Nearest city's wensite density, t-1", rowname)))))))) %>% 
-  column_to_rownames()
-
-path.out <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
-ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
-ggsave(path.out)
-
-path.out <- paste0(path, "/outputs/rf/models/oa_rf/all_OAs/all_OAs.RData")
-save(model.all, file = path.out)
-
-# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
-#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
-#                                                         "year")], 
-#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
-#                     trControl = tc,
-#                     method = "ranger",
-#                     #na.action = na.omit,
-#                     #importance = "impurity",
-#                     metric="Rsquared",
-#                     verbose=FALSE)
-# ffsmodel_LLO
-# ffsmodel_LLO$selectedvars
-# plot_ffs(ffsmodel_LLO)
-```
-
-### train the model in a loop for all but one regions
-
-```{r}
-
-#folds
-k <- 10 #
-
-#length is = (n_repeats*nresampling)+1
-seeds <- vector(mode = "list", length = 11)
-
-#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # It should be RHS variables + 2
-
-#for the last model
-seeds[[11]]<-sample.int(1000, 1)
-
-# detectCores()
-# cl <- makePSOCKcluster(2)
-# registerDoParallel(cl)
-
-# train in every region
-start.time <- Sys.time()
-
-regions <- df %>% ungroup() %>% distinct(RGN11CD) 
-regions <- as.vector(regions$RGN11CD)
-
-for (i in regions){
-
-  print(paste0("start loop for ", i))
-
-  set.seed(71)
-  indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD!=i), 
-                                spacevar = "oa11cd",
-                                timevar = "year", 
-                                k = k)
-
-  # CV
-  tc <- trainControl(method = "cv",
-                   number = k,
-                   seeds = seeds,
-                   allowParallel = T,
-                   index = indices$index,
-                   savePredictions = 'final')
-
-  model.all <- train(n ~ n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
-                   data = df %>% filter(RGN11CD!=i), 
-                   trControl = tc,
-                   method = "ranger",
-                   na.action = na.omit,
-                   #preProc = c("center", "scale"),
-                   importance = "impurity")
-  #assign(paste0("model.all",as.character(i)), model.all)
-
-  #file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
-  file <- paste0("/hdd/tmp/regions/region_", as.character(i), ".RData")  
-  save(model.all, file = file)  # file name = region not in the training data
-  rm(model.all)
-
-  print(paste0("end loop for ", i))
-
-}
-
-# stopCluster(cl)
-end.time <- Sys.time()
-time.taken <- round(end.time - start.time,2)
-time.taken
-```
-
-### Resamples for plots
-
-```{r}
-# create a list of the above model outputs
-filenames <- list.files(paste0(path, "/outputs/rf/models/oa_rf/regions"), pattern = "*.RData", full.names = T)
-
-for (i in filenames){
-  load(file = i)
-  i <- gsub("/home/nw19521/projects/web.diffusion/outputs/rf/models/oa_rf/regions/|.RData", "", i)
-  assign(i, model.all)  
-}
-
- fit.model.all <- mget(ls(pattern = "^region_"))
-
-# clear up
-rm(model.all) # this is the last item from the loop
-rm(list = ls(pattern = "^region_"))
-
-# resamples
-fit.model.all.res <- resamples(fit.model.all)
-summary(fit.model.all.res)
-
-# plot metrics
-
-# TODO: fix the year labels
-
-scales <- list(x=list(relation="free"), y=list(relation="free"))
-bwplot(fit.model.all.res, scales=scales)
-
-# TODO: fix the year labels
-```
-
-### Variable importance
-
-```{r , include=TRUE, results= 'markup', message=FALSE}
-lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
-```
-
-```{r}
-help <- lapply(fit.model.all, function(x) varImp(x)) 
-help <- lapply(help, function(x) x[[1]])
-help <- do.call("rbind", help)
-
-importance <- help %>% rownames_to_column() %>% 
-  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
-         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
-  rename(importance = Overall) %>% 
-  relocate(importance, .after = year) %>% 
-  dplyr::select(-rowname)
-
-importance %>% ggplot() +
-  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
-  scale_color_discrete(name = "Features") +
-  #scale_linetype_manual(name = "Website density regressions111") 
-  ylab("Importance") + 
-  ggtitle("Feature importance") +
-  theme_minimal() +
-  theme(axis.title.x=element_blank(),
-        plot.title = element_text(hjust = 0.5))
-
-```
-
-### Test on one region
-
-```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}
-
-# for a reference point
-df %>% group_by(year) %>%
-  summarise(min = min(n), max=max(n),
-            mean = mean(n), median = median(n)) %>%
-  round(2) %>% kable()
-```
-
-```{r}
-# train on all but one region, test on that region
-
-df.predictions <- data.frame()
-for (i in regions){
-  #for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], 
-                    df[df$RGN11CD==i,])
-    pred <- as.data.frame(pred)
-    rownames(pred) <- c()
-    assign(paste0("region.predict.model.", ".on.", i), pred)
-
-    d <- paste0("region.predict.model.", ".on.", i)
-    d <- cbind(get(d), df %>% filter(RGN11CD==i)) #%>% arrange(year, ladcd)
-
-    #***dplyr::select(RGN21CD, ladcd, year, growth, 1)***
-
-    d <- d[,c(5, 2, 4, 6, 1)]     
-    colnames(d)[5] <- "predictions"
-    d <- d %>% mutate(trained.on = i)
-    df.predictions <- rbind(d, df.predictions)
-  #}
-}
-
-df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN11CD, ".", trained.on))
-
-# split to list by region pair
-pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
-
-# calculate metrics for every region pair 
-rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))  # CHANGE n WITH growth
-
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% 
-  dplyr::select(RGN21CD, RGN21NM) %>% 
-  distinct() %>% 
-  rename(RGN11CD = RGN21CD) %>% 
-  add_row(RGN11CD = c("Scotland", "N0", "Wales"), 
-          RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
-
-path.out <- paste0(path, "/outputs/rf/heatmap_OA.png")
-rf.year.all.metrics %>% 
-  as.data.frame() %>%
-  rownames_to_column(var = "metrics") %>% 
-  pivot_longer(!metrics, names_to = "train.test") %>% 
-  pivot_wider(names_from = metrics, values_from = value) %>% 
-  separate(train.test, c("train", "test"), remove = F) %>% 
-  left_join(lookup.region, by = c("train" = "RGN11CD")) %>% 
-  rename(train.region=RGN21NM) %>% 
-  left_join(lookup.region, by = c("test" = "RGN11CD")) %>% 
-  rename(test.region=RGN21NM) %>% 
-  arrange(Rsquared) %>% 
-  dplyr::select(train.region, test.region, Rsquared) %>% 
-  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
-  geom_tile() +
-  coord_fixed() +
-  labs(y = "", x = "") + #y: test.region, x: train.region
-  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
-  theme(#legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust=1),
-        panel.border = element_blank(), 
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.background = element_rect(fill = "white"))
-ggsave(path.out)
-
-  #summarise(r = range(Rsquared))
-  #round(2) %>%
-  #kable()
-
-```
-
-### OLD
-
-```{r, eval=FALSE}
-# train on all but one region, test on that region
-
-df.predictions <- data.frame()
-for (i in regions){
-  for (j in regions[!regions == i]){
-    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], df[df$RGN11CD==j,])
-    pred <- as.data.frame(pred)
-    rownames(pred) <- c()
-    assign(paste0("region.predict.model.", i, ".on.", j), pred)
-
-    d <- paste0("region.predict.model.", i, ".on.", j)
-    d <- cbind(get(d), df %>% filter(RGN11CD==j)) #%>% arrange(year, ladcd)
-    d <- d[,c(5, 2, 4, 6, 1)]
-    colnames(d)[5] <- "predictions"
-    d <- d %>% mutate(trained.on = i)
-    df.predictions <- rbind(d, df.predictions)
-  }
-}
-
-df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN11CD, ".", trained.on))
-
-# split to list by region pair
-pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11
-
-# calculate metrics for every region pair 
-rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
-                                                                     obs = x$n))  # CHANGE n WITH growth
-
-path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
-lookup.region <- read_csv(path.lookup) %>% 
-  dplyr::select(RGN21CD, RGN21NM) %>% 
-  distinct() %>% 
-  rename(RGN11CD = RGN21CD) %>% 
-  add_row(RGN11CD = c("Scotland", "N0", "Wales"), 
-          RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))
-
-path.out <- paste0(path, "/outputs/rf/heatmap_OA.png")
-rf.year.all.metrics %>% 
-  as.data.frame() %>%
-  rownames_to_column(var = "metrics") %>% 
-  pivot_longer(!metrics, names_to = "train.test") %>% 
-  pivot_wider(names_from = metrics, values_from = value) %>% 
-  separate(train.test, c("train", "test"), remove = F) %>% 
-  left_join(lookup.region, by = c("train" = "RGN11CD")) %>% 
-  rename(train.region=RGN21NM) %>% 
-  left_join(lookup.region, by = c("test" = "RGN11CD")) %>% 
-  rename(test.region=RGN21NM) %>% 
-  arrange(Rsquared) %>% 
-  dplyr::select(train.region, test.region, Rsquared) %>% 
-  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
-  geom_tile() +
-  coord_fixed() +
-  labs(y = "", x = "") + #y: test.region, x: train.region
-  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
-  theme(#legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust=1),
-        panel.border = element_blank(), 
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.background = element_rect(fill = "white"))
-ggsave(path.out)
-
-  #summarise(r = range(Rsquared))
-  #round(2) %>%
-  #kable()
-
+# df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol", 
+#                         "Manchester", "Sheffield", "Leeds", "Edinburgh", 
+#                         "Leicester", "nearest.city.oa11cd")) %>% 
+#   write_csv(file = out.path)
 ```
@@
   #assign(paste0("model.all",as.character(t + 1)), model.all)
-  file <- paste0(path, "/data/temp/oa_rf/timeseries/timeseries", as.character(t + 1), ".RData")
+  file <- paste0(path, "/data/temp/oa_rf/timeseries", as.character(t + 1), ".RData")
   save(model.all, file = file)
@@
 
-#load(file = (paste0(path, "/data/temp/oa_rf/timeseries/timeseries2001.RData")))
-
-filenames <- list.files(paste0(path, "/data/temp/oa_rf/timeseries/"), pattern = "*.RData", full.names = T)
+load(file = (paste0(path, "/data/temp/oa_rf/timeseries2001.RData")))
+
+filenames <- list.files(paste0(path, "/data/temp/oa_rf"), pattern = "*.RData", full.names = T)
 

### 2025-01-21 src/s_la_per_firm_corrected.Rmd
0595_10

--- 2024-08-15 src/s_la_per_firm_corrected.Rmd
@@
 ```
+
+## Average ranks websites per firm t and 2012 only: 05 and 95
+
+```{r}
+
+t <- 2000
+
+out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only0595_av.png")
+
+df %>%
+
+  left_join(la %>% as_tibble() %>% dplyr::select(LAD21CD, LAD21NM),
+            by = c("ladcd" = "LAD21CD")) %>%
+
+  mutate(period = ifelse(year < 1999, 1, 
+                         ifelse(year > 2009, 3, 2))) %>%
+  group_by(ladcd, period) %>%
+  mutate(n = ifelse(period != 2, mean(n), n)) %>% 
+  filter(year > 1997 & year < 2011) %>% 
+  group_by(year) %>% 
+  mutate(rank = rank(desc(n))) %>% 
+  ungroup() %>% 
+  filter(year==1998 | year == 2010) %>% 
+  arrange(ladcd, year) %>% 
+  group_by(ladcd) %>% 
+  #mutate(diff_1st = -diff(rank))
+  mutate(diff_1st = -(rank-dplyr::lag(rank))) %>%
+
+  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .05, na.rm = T))
+  # q.95 = 253, q.9 = 214, q.1 = 11, q.05 = 3.75. **abs()**
+
+  mutate(alpha = ifelse(abs(diff_1st) <= 3.75 | abs(diff_1st) > 253, 1, 0.5)) %>% 
+  mutate(alpha = ifelse(is.na(diff_1st), 0.5, alpha)) %>%
+  mutate(q = ifelse(abs(diff_1st) <= 3.75, "05",
+                    ifelse(abs(diff_1st) > 253, "95", "m"))) %>% 
+  mutate(q = ifelse(is.na(diff_1st), "m", q)) %>% #filter(rank < 100 & q == "95")
+  mutate(label = ifelse(abs(diff_1st) > 253  & q == "95", #& rank < 100
+                        LAD21NM, "")) %>% 
+  mutate(label.low = ifelse(abs(diff_1st) <= 3.75  & q == "05", #& rank < 100
+                        LAD21NM, "")) %>%
+
+  filter(abs(diff_1st) > 253 & rank > 100 & q == "95") %>% distinct(LAD21NM)
+
+  ggplot(aes(year, rank, group = ladcd, alpha=alpha)) +
+  geom_bump(aes(colour = q)) + #colour = "red"
+  scale_color_manual(values = c("darkorchid3", "orange", "grey70"),
+                     breaks = c("05", "95"),
+                     labels = c("Stability","Volatility")) + 
+  labs(color = "LAD") +
+  scale_y_reverse(labels = c(1, 100, 200, 300, 400)) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  #theme(legend.position="none") +
+
+  geom_text_repel(aes(x = Inf, label = label), 
+                  size = 3, hjust = 0, color = "darkorange3", 
+                  max.overlaps = Inf,
+                  min.segment.length = Inf,
+                  hjust = "left") + # x = 2012 + .1
+
+  geom_text_repel(aes(x = -Inf, label = label.low), 
+                  size = 3, hjust = 0, color = "darkorchid4", 
+                  max.overlaps = Inf,
+                  min.segment.length = Inf,
+                  hjust = "left") + # x = 2012 + .1
+
+  scale_x_continuous(NULL, 
+                     #limits = c(1998, 2010), 
+                     breaks=c(1998, 2010),
+                     labels = c("1996-98","2010-12")) +
+  guides(alpha = "none")
+
+ggsave(out.path)
+```
+
+```{r}
+path.in <- paste0(path, "/outputs/s/s_la_per_firm.csv")
+
+latecomers <- c("Halton", "Blackpool", "Rutland", "Torbay", "Isles of Scilly", "North Devon", "Gosport", "Thanet", "Na h-Eileanan Siar", "Conwy")
+
+earlyadopters <- c("Slough", "Three Rivers", "Epsom and Ewell", "City of London", "Hillingdon", "Hounslow", "Kensington and Chelsea", "Kingston upon Thames", "Richmond upon Thames", "Wandsworth")
+
+la.df <- read_csv(path.in)
+
+early <- la.df %>% filter(LAD21NM %in% earlyadopters) %>% 
+  dplyr::select(LAD21NM, estimate, fast) %>% 
+  mutate(estimate = round(estimate, 1)) %>% 
+  mutate('early/late' = "early adopting") %>% 
+  rename(LAD = LAD21NM)
+
+late <- la.df %>% filter(LAD21NM %in% latecomers) %>% 
+  dplyr::select(LAD21NM, estimate, fast) %>% 
+  mutate(estimate = round(estimate, 1)) %>% 
+  mutate('early/late' = "late coming") %>% 
+  rename(LAD = LAD21NM)
+
+df.late.early <- bind_rows(late, early)
+
+path.out <- paste0(path, "/outputs/ranks/late_early.csv")
+write_csv(df.late.early, path.out)
+
+# stargazer(df.late.early[1:20,],
+#           type = 'text', 
+#           summary = FALSE)
+```

### 2025-01-21 src/s_la_per_firm_corrected_10.Rmd

@@
 ```
+
+## Average ranks websites per firm t and 2012 only: 05 and 95
+
+```{r}
+
+t <- 2000
+
+out.path <- paste0(path, "/outputs/ranks/web_per_firm", t,"_2012_only0595_av_10.png")
+
+df %>%
+
+  left_join(la %>% as_tibble() %>% dplyr::select(LAD21CD, LAD21NM),
+            by = c("ladcd" = "LAD21CD")) %>%
+
+  mutate(period = ifelse(year < 1999, 1, 
+                         ifelse(year > 2009, 3, 2))) %>%
+  group_by(ladcd, period) %>%
+  mutate(n = ifelse(period != 2, mean(n), n)) %>% 
+  filter(year > 1997 & year < 2011) %>% 
+  group_by(year) %>% 
+  mutate(rank = rank(desc(n))) %>% 
+  ungroup() %>% 
+  filter(year==1998 | year == 2010) %>% 
+  arrange(ladcd, year) %>% 
+  group_by(ladcd) %>% 
+  #mutate(diff_1st = -diff(rank))
+  mutate(diff_1st = -(rank-dplyr::lag(rank))) %>%
+
+  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .05, na.rm = T))
+  # q.95 = 194.5, q.9 = 162.5, q.1 = 10.5, q.05 = 5. **abs()**
+
+  mutate(alpha = ifelse(abs(diff_1st) <= 5 | abs(diff_1st) > 194.5, 1, 0.5)) %>% 
+  mutate(alpha = ifelse(is.na(diff_1st), 0.5, alpha)) %>%
+
+  mutate(q = ifelse(abs(diff_1st) <= 5, "05",
+                    ifelse(abs(diff_1st) > 194.5, "95", "m"))) %>% 
+  mutate(q = ifelse(is.na(diff_1st), "m", q)) %>%
+
+  ggplot(aes(year, rank, group = ladcd, alpha=alpha)) +
+  geom_bump(aes(colour = q)) + #colour = "red"
+  scale_color_manual(values = c("darkorchid3", "orange", "grey70"),
+                     breaks = c("05", "95"),
+                     labels = c("Stability","Volatility")) + 
+  labs(color = "LAD") +
+  scale_y_reverse(labels = c(1, 100, 200, 300, 400)) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  #theme(legend.position="none") +
+  scale_x_continuous(NULL, 
+                     #limits = c(1998, 2010), 
+                     breaks=c(1998, 2010),
+                     labels = c("1996-98","2010-12")) +
+  guides(alpha = "none")
+
+ggsave(out.path)
+```

### 2025-01-23 src/s_la_per_firm_corrected.Rmd
late

@@
 
-  filter(abs(diff_1st) > 253 & rank > 100 & q == "95") %>% distinct(LAD21NM)
+  #filter(abs(diff_1st) > 253 & rank > 100 & q == "95") %>% distinct(LAD21NM)
 
@@
 
-  geom_text_repel(aes(x = Inf, label = label), 
-                  size = 3, hjust = 0, color = "darkorange3", 
-                  max.overlaps = Inf,
-                  min.segment.length = Inf,
-                  hjust = "left") + # x = 2012 + .1
-
-  geom_text_repel(aes(x = -Inf, label = label.low), 
-                  size = 3, hjust = 0, color = "darkorchid4", 
-                  max.overlaps = Inf,
-                  min.segment.length = Inf,
-                  hjust = "left") + # x = 2012 + .1
+  # geom_text_repel(aes(x = Inf, label = label), 
+  #                 size = 3, hjust = 0, color = "darkorange3", 
+  #                 max.overlaps = Inf,
+  #                 min.segment.length = Inf,
+  #                 hjust = "left") + # x = 2012 + .1
+  # 
+  # geom_text_repel(aes(x = -Inf, label = label.low), 
+  #                 size = 3, hjust = 0, color = "darkorchid4", 
+  #                 max.overlaps = Inf,
+  #                 min.segment.length = Inf,
+  #                 hjust = "left") + # x = 2012 + .1
 
@@
 ```
+
+## Table for leapfroging
 
@@
 
+path.out <- paste0(path, "/outputs/ranks/late.csv")
+write_csv(late, path.out)
+
 # stargazer(df.late.early[1:20,],

### 2025-01-23 src/s_la_per_firm_corrected.Rmd
lad curves

@@
 
+## One line plot for all LAD
+
+```{r}
+a.la <- tibble()
+all.la <- df %>% 
+  filter(ladcd != "NA") %>% 
+  distinct(ladcd) %>% 
+  #slice_head(n=100) %>% 
+  simplify2array()
+
+for (i in all.la){
+  data <- df %>% filter(ladcd==i) %>%
+    dplyr::select(year, ladcd, n)
+
+  tryCatch(model <- (nls(n ~ SSlogis(year, Asym, xmid, scal), 
+                         data = data, 
+                         control = nls.control(maxiter = 1000))), error=function(e) NULL)
+
+  model.name <- paste0("model.", i)
+
+  assign(model.name, model)
+
+  tryCatch(b.la <- tidy(model), error=function(e) NULL)
+
+  tryCatch(b.la$r2 <- rsquare(model, data), error=function(e) NULL)
+
+  tryCatch(a.la <-rbind(a.la, b.la), error=function(e) NULL)
+}
+
+a.la <- a.la %>% filter(term=="xmid") %>% 
+  bind_cols(as_tibble(all.la)) %>% 
+  #mutate(estimate = round(estimate, 0)) %>% 
+  #summarise(r = range(r2))
+  filter(p.value < 0.01 & 
+         r2 > 0.9) %>%    # 376 / 376, all sig, 330 r2 > 0.9
+  dplyr::select(ladcd, estimate, r2, std.error) %>%
+  mutate(fast = ifelse(estimate < xmid.uk, "fast", "slow")) # it was 2003
+
+# predictions 
+pred <- tibble()
+help <- tibble(year = 1996:2012)
+for (i in all.la){
+  model.name <- (paste0("model.", i))
+  #assign(col.name, paste0("pred.", i))
+  pred_ <- df %>% filter(ladcd == i) %>% 
+    mutate(prednls = eval(call("predict", as.name(model.name))))  # "pred.{i}" :=
+  help_ <- help %>% left_join(pred_)
+
+  pred <- pred %>% bind_rows(help_) 
+}
+
+# plot
+path.out <- paste0(path, "/outputs/s/lad_per_firm_all.png")
+
+ggplot(pred %>% 
+              left_join(a.la), 
+            aes(x = year, y = n, color = fast, group = ladcd)) +
+  geom_line(aes(y = prednls), size = .1, alpha = .7) +
+  scale_x_continuous(NULL, labels = as.character(df$year), breaks = df$year) +
+  #geom_vline(xintercept = xmid.uk) +
+  ylab("N. of websites / per firm") + 
+  labs(color = "Website \ndiffusion speed") +
+  theme(legend.position=c(.9,.15)) +
+  scale_color_manual(values = c("fast" = "purple",
+                                "slow"="orange"),
+                     limits = force) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5),
+        panel.grid.minor = element_blank()) +
+  geom_vline(xintercept = xmid.uk, colour="grey30", linetype="dashed")
+
+ggsave(filename = path.out, bg="white")
+```
+
 ## Map the xmid

### 2025-01-23 src/s_la_per_firm_corrected.Rmd
lad curves

@@
 library(ggbump)
+library(latex2exp)
 
@@
   ylab("N. of websites / per firm") + 
-  labs(color = "Website \ndiffusion speed") +
+  labs(color = "Website diffusion speed") +
   theme(legend.position=c(.9,.15)) +
@@
   theme(plot.title = element_text(hjust = 0.5),
-        panel.grid.minor = element_blank()) +
-  geom_vline(xintercept = xmid.uk, colour="grey30", linetype="dashed")
+        panel.grid.minor = element_blank(),
+        legend.position = "bottom") +
+  geom_vline(aes(xintercept = xmid.uk, linetype="dashed"), colour="grey30") +
+  scale_linetype_manual(values=c("dashed"), 
+                        labels = c(""), 
+                        name = unname(TeX(c("$t_{0}$ for the UK"))))
 

### 2025-02-02 src/la_RF_corrected.Rmd
all lad + oa without the hold-out

@@
                    #preProc = c("center", "scale"),
-                   importance = "impurity")
-
-stopCluster(cl)
+                   importance = "impurity",
+                   num.threads = 15)
+
+# stopCluster(cl)
 end.time <- Sys.time()
@@
 # Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 4718, 4690, 5040, 4718, 5055, 4704, ... 
+# Summary of sample sizes: 5082, 4354, 4650, 4746, 5130, 4928, ... 
 # Resampling results across tuning parameters:
@@
 #   mtry  splitrule   RMSE        Rsquared   MAE       
-#   2     variance    0.02811777  0.8509273  0.01912432
-#   2     extratrees  0.02782836  0.8508316  0.01901921
-#   4     variance    0.02867344  0.8466191  0.01970102
-#   4     extratrees  0.02795250  0.8509000  0.01911956
-#   6     variance    0.02974784  0.8364575  0.02066007
-#   6     extratrees  0.02830277  0.8482189  0.01944133
+#   2     variance    0.02899374  0.8242687  0.02021469
+#   2     extratrees  0.02919421  0.8158094  0.02039216
+#   4     variance    0.02929606  0.8252793  0.02067827
+#   4     extratrees  0.02922286  0.8194044  0.02056084
+#   6     variance    0.03068719  0.8155331  0.02189821
+#   6     extratrees  0.02938456  0.8190111  0.02074402
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule = extratrees and min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule = variance and min.node.size = 5.
 
@@
 set.seed(831)
-for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # 6 feature + 2
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # 6 feature + 2
 
@@
                    #preProc = c("center", "scale"),
-                   importance = "impurity")
+                   importance = "impurity",
+                   num.threads = 15)
   assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data
@@
 # Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 4044, 3685, 4032, 4044, 4044, 3696, ... 
+# Summary of sample sizes: 4356, 3421, 3720, 4068, 4104, 3872, ... 
 # Resampling results across tuning parameters:
@@
 #   mtry  splitrule   RMSE       Rsquared   MAE      
-#   2     variance    0.2658416  0.5284505  0.1934243
-#   2     extratrees  0.2612133  0.5292689  0.1900297
-#   4     variance    0.2621564  0.5194092  0.1912807
-#   4     extratrees  0.2611575  0.5183202  0.1910029
-#   6     variance    0.2609983  0.5140347  0.1891971
-#   6     extratrees  0.2607463  0.5124544  0.1907650
+#   2     variance    0.2894270  0.5199902  0.2268798
+#   2     extratrees  0.2872946  0.5173582  0.2238973
+#   4     variance    0.2883168  0.5049246  0.2257499
+#   4     extratrees  0.2860372  0.5056609  0.2225991
+#   6     variance    0.2921377  0.4901490  0.2284739
+#   6     extratrees  0.2861368  0.4945032  0.2219559
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 6, splitrule = extratrees and min.node.size = 5.
+# The final values used for the model were mtry = 4, splitrule = extratrees and min.node.size = 5.
 

### 2025-02-02 src/la_RF_corrected_10.Rmd
all lad + oa without the hold-out

@@
                    #preProc = c("center", "scale"),
-                   importance = "impurity")
+                   importance = "impurity",
+                   num.threads = 15)
 
@@
 #   mtry  splitrule   RMSE       Rsquared   MAE      
-#   2     variance    0.1877289  0.7762104  0.1222690
-#   2     extratrees  0.1945882  0.7537303  0.1274455
-#   4     variance    0.1899769  0.7754308  0.1253349
-#   4     extratrees  0.1925299  0.7604062  0.1270665
-#   6     variance    0.1926697  0.7707287  0.1277624
-#   6     extratrees  0.1930151  0.7601929  0.1280934
+#   2     variance    0.1908100  0.7375479  0.1279762
+#   2     extratrees  0.1994918  0.7045691  0.1340791
+#   4     variance    0.1899064  0.7491920  0.1294553
+#   4     extratrees  0.1985176  0.7110348  0.1347026
+#   6     variance    0.1911215  0.7530296  0.1312026
+#   6     extratrees  0.1992530  0.7110950  0.1359168
 # 
-# Tuning parameter 'min.node.size' was held constant at a value of 5
-# RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule = variance and min.node.size = 5.
+# Tuning parameter 'min.node.size' was held constant at
+#  a value of 5
+# RMSE was used to select the optimal model using
+#  the smallest value.
+# The final values used for the model were mtry =
+#  4, splitrule = variance and min.node.size = 5.
 
@@
 time.taken <- round(end.time - start.time,2)
-time.taken # ca. 4 minutes 
+time.taken # ca. 18 minutes 
 ```
@@
 #   mtry  splitrule   RMSE       Rsquared   MAE      
-#   2     variance    0.2773784  0.6171406  0.2025901
-#   2     extratrees  0.2696423  0.6424804  0.1985126
-#   4     variance    0.2834811  0.5991036  0.2058774
-#   4     extratrees  0.2720828  0.6324695  0.2007287
-#   6     variance    0.2902072  0.5801010  0.2101899
-#   6     extratrees  0.2755322  0.6215775  0.2037045
+#   2     variance    0.3044927  0.5868676  0.2369061
+#   2     extratrees  0.2954167  0.6153847  0.2312128
+#   4     variance    0.3170846  0.5455342  0.2439098
+#   4     extratrees  0.3008763  0.5922954  0.2347268
+#   6     variance    0.3222021  0.5224533  0.2464477
+#   6     extratrees  0.3055270  0.5782538  0.2388544
 # 

### 2025-02-02 src/s_la_per_firm_corrected.Rmd
all lad + oa without the hold-out

@@
 
-# xmid = 2003.118
+# xmid.uk = 2003.118
 xmid.uk <- tidy(fit) %>% 
@@
 ```{r}
+# xmid.uk = 2003.118
+
 a.la <- tibble()
@@
   #geom_vline(xintercept = xmid.uk) +
-  ylab("N. of websites / per firm") + 
+  ylab("N. of websites per firm") + 
   labs(color = "Website diffusion speed") +

### 2025-02-02 src/oa_RF_corrected_10.R

@@ -1,17 +1,17 @@
----
-title: "Random Forests", OAs
-date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
-output: 
-  html_document:
-    df_print: paged
-    toc: true
-    toc_float: true
-knit: (function(inputFile, encoding) {
-    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
-  })
----
-
-```{r setup, include=FALSE}
+#' ---
+#' title: "Random Forests", OAs
+#' date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
+#' output: 
+#'   html_document:
+#'     df_print: paged
+#'     toc: true
+#'     toc_float: true
+#' knit: (function(inputFile, encoding) {
+#'     rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
+#'   })
+#' ---
+#' 
+## ----setup, include=FALSE-----------------------------------------------------
 knitr::opts_chunk$set(echo = TRUE)
 
 library(tidyverse)
@@ -41,15 +41,15 @@ options(scipen=10000)
 
 # This is the project path
 path <- find_rstudio_root_file()
-```
 
-## Load data
-
-Postcode lookup.
-
-[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)
-
-```{r}
+#' 
+#' ## Load data
+#' 
+#' Postcode lookup.
+#' 
+#' [source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)
+#' 
+## -----------------------------------------------------------------------------
 path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
 lookup <- read_csv(path.lookup) %>% 
   dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
@@ -58,17 +58,17 @@ lookup <- read_csv(path.lookup) %>%
 # The problems refer to Welsh LAD names. Not a problem for the analysis.
 #sapply(lookup, function(x) sum(is.na(x)))
[…]

### 2025-02-04 src/oa_RF_corrected.Rmd
oa

@@
                     abs.growth = n - n.lag) %>% 
-  relocate(n.lag, .after = n)
+  relocate(n.lag, .after = n) %>% 
+  ungroup()
 
@@
 # Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 2926910, 2926924, 3136005, 2926924, 3136005, 2926924, ... 
+# Summary of sample sizes: 3215700, 2574600, 3128916, 3225315, 2789175, 2618196, ... 
 # Resampling results across tuning parameters:
@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    3.283708  0.3202280  1.034153
-#   2     extratrees  3.516978  0.2323646  1.087689
-#   5     variance    3.240309  0.3354508  1.019184
-#   5     extratrees  3.443571  0.2565672  1.107110
-#   8     variance    3.278693  0.3251487  1.017672
-#   8     extratrees  3.449965  0.2562403  1.124713
+#   2     variance    3.235219  0.3131576  1.211443
+#   2     extratrees  3.452069  0.2176445  1.182582
+#   5     variance    3.217974  0.3337214  1.233777
+#   5     extratrees  3.386142  0.2500448  1.251388
+#   8     variance    3.294155  0.3208588  1.240884
+#   8     extratrees  3.418390  0.2496608  1.281916
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 5, splitrule = variance and min.node.size = 5.
+# The final values used for the model were mtry = 5, splitrule
+#  = variance and min.node.size = 5.
 
@@
                    #preProc = c("center", "scale"),
-                   importance = "impurity")
+                   importance = "impurity",,
+                   num.threads = 15)
 

### 2025-02-04 src/oa_RF_corrected_10.Rmd
oa

--- 2024-09-11 src/oa_RF_corrected_10.Rmd
@@
 #length is = (n_repeats*nresampling)+1
-seeds <- vector(mode = "list", length = 11)
+seeds <- vector(mode = "list", length = 8)
 
@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    22.46734  0.4784290  5.681843
-#   2     extratrees  24.69943  0.3875775  6.049473
-#   5     variance    22.50225  0.4803563  5.664478
-#   5     extratrees  23.95316  0.4143508  6.102471
-#   8     variance    23.14682  0.4644807  5.678126
-#   8     extratrees  24.14426  0.4095256  6.213216
+#   2     variance    21.61442  0.4482161  6.736518
+#   2     extratrees  23.82034  0.3557068  6.584642
+#   5     variance    21.58205  0.4549608  6.845549
+#   5     extratrees  23.00958  0.3848122  6.884375
+#   8     variance    22.07397  0.4426922  6.949996
+#   8     extratrees  23.17989  0.3796106  7.060913
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule = variance and min.node.size = 5.
+# The final values used for the model were mtry = 5, splitrule
+#  = variance and min.node.size = 5.
 
@@
   indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD!=i), 
-                                spacevar = "oa11cd",
+                                spacevar = "RGN11CD",
                                 timevar = "year", 

### 2025-02-04 src/oa_RF_corrected.R

@@ -11,7 +11,7 @@
 #'   })
 #' ---
 #' 
-## ----setup, include=FALSE-----------------------------
+## ----setup, include=FALSE-------------------------------------------
 knitr::opts_chunk$set(echo = TRUE)
 
 library(tidyverse)
@@ -48,7 +48,7 @@ path <- find_rstudio_root_file()
 #' 
 #' [source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)
 #' 
-## -----------------------------------------------------
+## -------------------------------------------------------------------
 path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
 lookup <- read_csv(path.lookup) %>% 
   dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
@@ -64,7 +64,7 @@ lookup <- read_csv(path.lookup) %>%
 #' 
 #' ## Problem postcodes
 #' 
-## -----------------------------------------------------
+## -------------------------------------------------------------------
 n = 1 #number of unique postcodes. 
 m = 11
 
@@ -117,7 +117,7 @@ problem.pcs <- df.long %>% filter(year==2004 | year == 2005,
 #' in the above postcodes. I turn them to NAs and then use a regression to impute these
 #' gaps.
 #' 
-## -----------------------------------------------------
+## -------------------------------------------------------------------
 df.corrected <- bind_rows(df9610, df1112) %>% 
   filter(year > 1995 & year < 2013) %>% 
   group_by(pc, year) %>% 
@@ -134,7 +134,7 @@ df.corrected$n[is.na(df.corrected$n)] <- predictions[is.na(df.corrected$n)]
 #' 
 #' ## OA df
 #' 
-## -----------------------------------------------------
+## -------------------------------------------------------------------
 df <- df.corrected %>% 
   mutate(year = as.integer(as.character(year))) %>% 
   filter(year > 1995 & year < 2013) %>% 
@@ -149,7 +149,7 @@ df <- df.corrected %>%
 #' 
 #' ## spatial data
 #' 
-## -----------------------------------------------------
+## -------------------------------------------------------------------
 # get OA for England and Wales
[…]

### 2025-02-04 src/oa_RF_corrected_10.R

@@ -11,7 +11,7 @@
 #'   })
 #' ---
 #' 
-## ----setup, include=FALSE-----------------------------------------------------
+## ----setup, include=FALSE-------------------------------------------
 knitr::opts_chunk$set(echo = TRUE)
 
 library(tidyverse)
@@ -49,7 +49,7 @@ path <- find_rstudio_root_file()
 #' 
 #' [source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)
 #' 
-## -----------------------------------------------------------------------------
+## -------------------------------------------------------------------
 path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
 lookup <- read_csv(path.lookup) %>% 
   dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
@@ -68,7 +68,7 @@ lookup <- read_csv(path.lookup) %>%
 #' I use the problem postcodes from n = 1 as these are the true outliers as per the
 #' line plots.
 #' 
-## -----------------------------------------------------------------------------
+## -------------------------------------------------------------------
 n = 1 #number of unique postcodes. 
 m = 11
 
@@ -117,7 +117,7 @@ problem.pcs <- c("SE24 9HP", "CV8 2ED", "GL16 7YA", "CW1 6GL", "M28 2SL", "DE21 
 #' in the above postcodes. I turn them to NAs and then use a regression to impute these
 #' gaps.
 #' 
-## -----------------------------------------------------------------------------
+## -------------------------------------------------------------------
 df.corrected <- bind_rows(df9610, df1112) %>% 
   filter(year > 1995 & year < 2013) %>% 
   group_by(pc, year) %>% 
@@ -135,7 +135,7 @@ df.corrected$n[is.na(df.corrected$n)] <- predictions[is.na(df.corrected$n)]
 #' 
 #' ## OA df
 #' 
-## -----------------------------------------------------------------------------
+## -------------------------------------------------------------------
 df <- df.corrected %>% 
   mutate(year = as.integer(as.character(year))) %>% 
   filter(year > 1995 & year < 2013) %>% 
@@ -150,7 +150,7 @@ df <- df.corrected %>%
 #' 
 #' ## spatial data
 #' 
[…]

### 2025-02-05 src/oa_RF_corrected_10.Rmd
oa_10

@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    21.61442  0.4482161  6.736518
-#   2     extratrees  23.82034  0.3557068  6.584642
-#   5     variance    21.58205  0.4549608  6.845549
-#   5     extratrees  23.00958  0.3848122  6.884375
-#   8     variance    22.07397  0.4426922  6.949996
-#   8     extratrees  23.17989  0.3796106  7.060913
+#   2     variance    21.61233  0.4487783  6.733987
+#   2     extratrees  23.80635  0.3564440  6.583231
+#   5     variance    21.59062  0.4546951  6.845095
+#   5     extratrees  23.00905  0.3846503  6.885083
+#   8     variance    22.10042  0.4421229  6.953833
+#   8     extratrees  23.19776  0.3791247  7.061227
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 5, splitrule
-#  = variance and min.node.size = 5.
+# The final values used for the model were mtry = 5, splitrule = variance and
+#  min.node.size = 5.
 

### 2025-02-08 src/la_RF.Rmd
oa

@@
 library(rprojroot)
-library(rgdal)
-library(rgeos)
-library(sf)
+# library(rgdal)
+# library(rgeos)
+# library(sf)
 library(spdep)
@@
                    #preProc = c("center", "scale"),
-                   importance = "impurity")
+                   importance = "impurity",
+                   num.threads = 15)
 
-stopCluster(cl)
+# stopCluster(cl)
 end.time <- Sys.time()
@@
 # Resampling: Cross-Validated (10 fold) 
-# Summary of sample sizes: 5055, 4690, 4732, 5070, 5040, 4718, ... 
+# Summary of sample sizes: 5082, 4354, 4650, 4746, 5130, 4928, ... 
 # Resampling results across tuning parameters:
@@
 #   mtry  splitrule   RMSE        Rsquared   MAE       
-#   2     variance    0.03227897  0.8023185  0.01991724
-#   2     extratrees  0.03150004  0.8057960  0.01912348
-#   4     variance    0.03271887  0.7985855  0.02025685
-#   4     extratrees  0.03174791  0.8047090  0.01932473
-#   6     variance    0.03358260  0.7879953  0.02080616
-#   6     extratrees  0.03207532  0.8018495  0.01964516
+#   2     variance    0.03216513  0.7810704  0.02077849
+#   2     extratrees  0.03220664  0.7752043  0.02087374
+#   4     variance    0.03271488  0.7775723  0.02134707
+#   4     extratrees  0.03229937  0.7760603  0.02102094
+#   6     variance    0.03397678  0.7680207  0.02235307
+#   6     extratrees  0.03248101  0.7767931  0.02132943
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule = extratrees and min.node.size = 5.
+# The final values used for the model were mtry = 2, splitrule = variance
+#  and min.node.size = 5.
 
@@
 ```
+
+### FROM NOW ON, IT HAS NOT BEEN USED FOR SUBMISSION 2
 

### 2025-02-08 src/oa_RF_corrected.Rmd
oa

@@
 
+# [not reproducible despite having the code as the LAD models]
 # Random Forest 
@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    3.235219  0.3131576  1.211443
-#   2     extratrees  3.452069  0.2176445  1.182582
-#   5     variance    3.217974  0.3337214  1.233777
-#   5     extratrees  3.386142  0.2500448  1.251388
-#   8     variance    3.294155  0.3208588  1.240884
-#   8     extratrees  3.418390  0.2496608  1.281916
+#   2     variance    3.236301  0.3128048  1.211779
+#   2     extratrees  3.453903  0.2169766  1.184256
+#   5     variance    3.217652  0.3331498  1.233429
+#   5     extratrees  3.390341  0.2491582  1.254907
+#   8     variance    3.296126  0.3199896  1.241124
+#   8     extratrees  3.418875  0.2485125  1.283134
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 5, splitrule
-#  = variance and min.node.size = 5.
+# The final values used for the model were mtry = 5, splitrule = variance
+#  and min.node.size = 5.
 

### 2025-02-08 src/oa_RF_corrected.R

@@ -11,7 +11,7 @@
 #'   })
 #' ---
 #' 
-## ----setup, include=FALSE-------------------------------------------
+## ----setup, include=FALSE--------------------------------------------------
 knitr::opts_chunk$set(echo = TRUE)
 
 library(tidyverse)
@@ -48,7 +48,7 @@ path <- find_rstudio_root_file()
 #' 
 #' [source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)
 #' 
-## -------------------------------------------------------------------
+## --------------------------------------------------------------------------
 path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
 lookup <- read_csv(path.lookup) %>% 
   dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
@@ -64,7 +64,7 @@ lookup <- read_csv(path.lookup) %>%
 #' 
 #' ## Problem postcodes
 #' 
-## -------------------------------------------------------------------
+## --------------------------------------------------------------------------
 n = 1 #number of unique postcodes. 
 m = 11
 
@@ -117,7 +117,7 @@ problem.pcs <- df.long %>% filter(year==2004 | year == 2005,
 #' in the above postcodes. I turn them to NAs and then use a regression to impute these
 #' gaps.
 #' 
-## -------------------------------------------------------------------
+## --------------------------------------------------------------------------
 df.corrected <- bind_rows(df9610, df1112) %>% 
   filter(year > 1995 & year < 2013) %>% 
   group_by(pc, year) %>% 
@@ -134,7 +134,7 @@ df.corrected$n[is.na(df.corrected$n)] <- predictions[is.na(df.corrected$n)]
 #' 
 #' ## OA df
 #' 
-## -------------------------------------------------------------------
+## --------------------------------------------------------------------------
 df <- df.corrected %>% 
   mutate(year = as.integer(as.character(year))) %>% 
   filter(year > 1995 & year < 2013) %>% 
@@ -149,7 +149,7 @@ df <- df.corrected %>%
 #' 
 #' ## spatial data
 #' 
-## -------------------------------------------------------------------
+## --------------------------------------------------------------------------
[…]

### 2025-02-08 src/oa_RF_corrected_10.R

@@ -1,827 +1,0 @@
-#' ---
-#' title: "Random Forests", OAs
-#' date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
-#' output: 
-#'   html_document:
-#'     df_print: paged
-#'     toc: true
-#'     toc_float: true
-#' knit: (function(inputFile, encoding) {
-#'     rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
-#'   })
-#' ---
-#' 
-## ----setup, include=FALSE-------------------------------------------
-knitr::opts_chunk$set(echo = TRUE)
-
-library(tidyverse)
-library(rprojroot)
-#library(rgdal)
-#library(rgeos)
-library(sf)
-library(spdep)
-#library(tmap)
-#library(maptools)
-library(knitr)
-library(REAT)
-library(tidygeocoder)
-library(geosphere)
-library(broom)
-#library(foreach)
-library(doParallel)
-library(raster)
-library(plm)
-#library(lmtest)
-library(caret)
-library(randomForest)
-library(CAST)
-library(Hmisc)
-
-options(scipen=10000)
-
-# This is the project path
-path <- find_rstudio_root_file()
-
-#' 
-#' ## Load data
-#' 
-#' Postcode lookup.
-#' 
-#' [source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)
-#' 
-## -------------------------------------------------------------------
-path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
-lookup <- read_csv(path.lookup) %>% 
-  dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
-  dplyr::rename(pc = pcds)
-#glimpse(lookup)
-# The problems refer to Welsh LAD names. Not a problem for the analysis.
-#sapply(lookup, function(x) sum(is.na(x)))
-# 10332 missing msoa11cd
-
-#' 
-#' The internet archive $1996$-$2010$ data is saved on /hdd.
-#' The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.
-#' 
-#' ## Problem postcodes
-#' 
[…]

### 2025-02-09 src/oa_RF.Rmd
oa

@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    4.999920  0.2053697  1.047293
-#   2     extratrees  5.167655  0.1540213  1.103048
-#   5     variance    5.187039  0.1710128  1.034382
-#   5     extratrees  5.169570  0.1557868  1.122845
-#   8     variance    5.633427  0.1247189  1.035171
-#   8     extratrees  5.182860  0.1563555  1.137332
+#   2     variance    4.644230  0.2444898  1.236831
+#   2     extratrees  4.874616  0.1642528  1.207521
+#   5     variance    4.716793  0.2415236  1.251176
+#   5     extratrees  4.779489  0.2006192  1.294454
+#   8     variance    5.257495  0.1919632  1.262540
+#   8     extratrees  4.814190  0.1951799  1.311195
 # 
@@
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were mtry = 2, splitrule =
-#  variance and min.node.size = 5.
-
+# The final values used for the model were mtry = 2, splitrule = variance
+#  and min.node.size = 5.
+#  
 varimp_mars <- varImp(model.all) 

### 2025-02-09 src/oa_RF.R

@@ -51,7 +51,7 @@ path <- find_rstudio_root_file()
 ## --------------------------------------------------------------------------
 path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
 lookup <- read_csv(path.lookup) %>% 
-  dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
+  dplyr::select(pcds, oa11cd) %>% 
   dplyr::rename(pc = pcds)
 #glimpse(lookup)
 # The problems refer to Welsh LAD names. Not a problem for the analysis.
@@ -62,120 +62,161 @@ lookup <- read_csv(path.lookup) %>%
 #' The internet archive $1996$-$2010$ data is saved on /hdd.
 #' The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.
 #' 
-#' ## Problem postcodes
-#' 
 ## --------------------------------------------------------------------------
 n = 1 #number of unique postcodes. 
-m = 11
+m = 12
 
 data.folder <- "/hdd/internet_archive/archive/data/"
 data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
 #Created by domain.R, which uses domain instead of host.
 #This is what we use for the hyperlinks paper as per George's script
 
-df9610 <- read_csv(data.path9610) %>% 
-  filter(
-      #V1.domain < m,             # for multiple pc
-      V1.domain == n,             # for n == 1
-      year > 1995)   
+df9610 <- read_csv(data.path9610) 
 
 data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
 #Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
 #This is what we use for the hyperlinks paper as per George's script
 
-df1112 <- read_csv(data.path2011_2012) %>% 
+df1112 <- read_csv(data.path2011_2012)
+
+df <- bind_rows(df9610, df1112) %>% 
   filter(
     #V1.domain < m,             # for multiple pc
-    V1.domain == n)             # for n == 1
-
-df.long <- bind_rows(df9610, df1112) %>% 
-  filter(year > 1995 & year < 2013) %>% 
[…]

### 2025-02-09 src/oa_RF_corrected_10.R

@@ -35,6 +35,7 @@ library(plm)
 library(caret)
 library(randomForest)
 library(CAST)
+library(Hmisc)
 
 options(scipen=10000)
 
@@ -64,6 +65,9 @@ lookup <- read_csv(path.lookup) %>%
 #' 
 #' ## Problem postcodes
 #' 
+#' I use the problem postcodes from n = 1 as these are the true outliers as per the
+#' line plots.
+#' 
 ## --------------------------------------------------------------------------
 n = 1 #number of unique postcodes. 
 m = 11
@@ -75,8 +79,8 @@ data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
 
 df9610 <- read_csv(data.path9610) %>% 
   filter(
-      #V1.domain < m,             # for multiple pc
-      V1.domain == n,             # for n == 1
+      V1.domain < m,               # for multiple pc
+      #V1.domain == n,             # for n == 1
       year > 1995)   
 
 data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
@@ -85,13 +89,13 @@ data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
 
 df1112 <- read_csv(data.path2011_2012) %>% 
   filter(
-    #V1.domain < m,             # for multiple pc
-    V1.domain == n)             # for n == 1
+    V1.domain < m)               # for multiple pc
+    #V1.domain == n)             # for n == 1
 
 df.long <- bind_rows(df9610, df1112) %>% 
   filter(year > 1995 & year < 2013) %>% 
   group_by(pc, year) %>% 
-  summarise(n = n()) %>%
+  summarise(n = sum(V1.domain)) %>%
   ungroup() %>% 
   #arrange(desc(n)) %>% 
   pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
@@ -103,12 +107,8 @@ df.long <- bind_rows(df9610, df1112) %>%
   replace(is.na(.), 0) %>% 
   mutate(outlier = ifelse(n > 1000 & (year==2004 | year == 2005), pc, ""))
 
-problem.pcs <- df.long %>% filter(year==2004 | year == 2005, 
-                   n > 1000) %>% 
-  distinct(pc) 
-
-# n > 1500: "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"   
[…]

### 2025-02-21 src/oa_RF_corrected.R
oa regions

@@ -726,6 +726,105 @@ unregister_dopar <- function() {
 }
 unregister_dopar()
 
+#' 
+#' ### train the model in all data 
+#' 
+## --------------------------------------------------------------------------
+set.seed(123)
+indices <- CreateSpacetimeFolds(df, 
+                                spacevar = "RGN11CD",
+                                timevar = "year", 
+                                k = 10)
+
+#length is = (n_repeats*nresampling)+1
+seeds <- vector(mode = "list", length = 11)
+
+#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
+for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # It should be RHS variables + 2
+
+#for the last model
+seeds[[11]]<-sample.int(1000, 1)
+
+# CV
+tc <- trainControl(method = "cv",
+                   number = 10,
+                   seeds = seeds,
+                   allowParallel = T,
+                   index = indices$index,
+                   savePredictions = 'final')
+
+# detectCores()
+# cl <- makePSOCKcluster(2)
+# registerDoParallel(cl)
+
+start.time <- Sys.time()
+model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
+                     n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
+                     #London.v + city.v + retail.v + n.l.slag + year,
+                   data = df, 
+                   trControl = tc,
+                   method = "ranger",
+                   na.action = na.omit,
+                   #preProc = c("center", "scale"),
+                   importance = "impurity",
+                   num.threads = 15)
+
+# stopCluster(cl)
+end.time <- Sys.time()
+time.taken <- round(end.time - start.time,2)
+time.taken #20.16h
+
+print(model.all)
+
+# [not reproducible despite having the code as the LAD models]
+# Random Forest 
+# 
+# 3716736 samples
+#       8 predictor
+# 
+# No pre-processing
+# Resampling: Cross-Validated (10 fold) 
+# Summary of sample sizes: 3215700, 2574600, 3128916, 3225315, 2789175, 2618196, ... 
+# Resampling results across tuning parameters:
+# 
+#   mtry  splitrule   RMSE      Rsquared   MAE     
+#   2     variance    3.236301  0.3128048  1.211779
[…]

### 2025-03-10 src/oa_RF_corrected.Rmd
oa regions
oa regions 10

@@
 #   mtry  splitrule   RMSE      Rsquared   MAE     
-#   2     variance    3.236301  0.3128048  1.211779
-#   2     extratrees  3.453903  0.2169766  1.184256
-#   5     variance    3.217652  0.3331498  1.233429
-#   5     extratrees  3.390341  0.2491582  1.254907
-#   8     variance    3.296126  0.3199896  1.241124
-#   8     extratrees  3.418875  0.2485125  1.283134
+#   2     variance    3.236887  0.3130255  1.212322
+#   2     extratrees  3.452335  0.2176573  1.182235
+#   5     variance    3.218065  0.3330621  1.233640
+#   5     extratrees  3.390579  0.2490956  1.252252
+#   8     variance    3.295241  0.3206509  1.240731
+#   8     extratrees  3.420014  0.2487011  1.283227
 # 
@@
   indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD!=i), 
-                                spacevar = "RGN21CD",
+                                spacevar = "RGN11CD",
                                 timevar = "year", 

### 2025-03-10 src/oa_RF_corrected_10.R

@@ -11,7 +11,7 @@
 #'   })
 #' ---
 #' 
-## ----setup, include=FALSE--------------------------------------------------
+## ----setup, include=FALSE-----------------------------------------------------
 knitr::opts_chunk$set(echo = TRUE)
 
 library(tidyverse)
@@ -49,7 +49,7 @@ path <- find_rstudio_root_file()
 #' 
 #' [source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)
 #' 
-## --------------------------------------------------------------------------
+## -----------------------------------------------------------------------------
 path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
 lookup <- read_csv(path.lookup) %>% 
   dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
@@ -68,7 +68,7 @@ lookup <- read_csv(path.lookup) %>%
 #' I use the problem postcodes from n = 1 as these are the true outliers as per the
 #' line plots.
 #' 
-## --------------------------------------------------------------------------
+## -----------------------------------------------------------------------------
 n = 1 #number of unique postcodes. 
 m = 11
 
@@ -117,7 +117,7 @@ problem.pcs <- c("SE24 9HP", "CV8 2ED", "GL16 7YA", "CW1 6GL", "M28 2SL", "DE21 
 #' in the above postcodes. I turn them to NAs and then use a regression to impute these
 #' gaps.
 #' 
-## --------------------------------------------------------------------------
+## -----------------------------------------------------------------------------
 df.corrected <- bind_rows(df9610, df1112) %>% 
   filter(year > 1995 & year < 2013) %>% 
   group_by(pc, year) %>% 
@@ -135,7 +135,7 @@ df.corrected$n[is.na(df.corrected$n)] <- predictions[is.na(df.corrected$n)]
 #' 
 #' ## OA df
 #' 
-## --------------------------------------------------------------------------
+## -----------------------------------------------------------------------------
 df <- df.corrected %>% 
   mutate(year = as.integer(as.character(year))) %>% 
   filter(year > 1995 & year < 2013) %>% 
@@ -150,7 +150,7 @@ df <- df.corrected %>%
 #' 
 #' ## spatial data
 #' 
-## --------------------------------------------------------------------------
[…]

### 2025-03-12 paper/submission2/response.qmd
paper v2

@@
 
-18-Dec-2024
+Editor and reviewer comments as well as author responses are presented below. 
+Comments are in \color{teal}teal fonts \color{black} and responses in black.
 
-Dear Emmanouil
-
+\color{teal}
 Thank you for submitting your manuscript to Journal of Economic Geography.
@@
 I am, therefore, putting more weight on reviewers 2 and 3 when identifying priorities for the revision.
+\color{black}
 
-## Editor; theory and justification 
+Thank you for providing the opportunity to revise and resubmit my paper. The 
+reviewer reports were really helpful and I think it is fair to say that the
+reviewers did go the extra mile to offer useful recommendations. I think I have 
+adequately addressed their concerns and my feeling is that the paper has now 
+been drastically improved. I hope the editor and the reviewers will agree.
+
 \color{teal}
@@
 
-## Editor; data
+The paper has been restructured to better illustrate the theoretical
+argumentation, motivation and contribution. It is also better situated within 
+the relevant literatures. Below I offer detailed responses to the reviewer 
+comments.
+
 \color{teal}
@@
 
-## Editor; conclusions
+The data issues have been clarified.
+
 \color{teal}
@@
 
+The restructuring of the paper around two hypotheses as well as the emphasis on
+the granularity of the analysis and the the long time series approach allowed to
+more clearly highlight the contribution in the conclusion section.
+
+As a sidenote, there was a minor error in the previous version of the paper 
+with the implementation of the spatially sensitive CV. This has now been 
+corrected hence the minor changes in the modelling results (second decimal).
+
 ## Referee: 1
 
+\color{teal}
 I have read your paper with interest and I appreciate its quality and clarity. The work is both well-executed and well-written, making it a valuable contribution to the field.
@@
 However, I would like to offer a few minor suggestions that could further enhance the readability and overall coherence of the paper:
+\color{black}
 
-> Simplify the Introduction: I recommend streamlining the introduction by reducing the inclusion of overly technical or highly specific details.
+Thank you for your nice comments and suggestions.
+
+\color{teal}
+Simplify the Introduction: I recommend streamlining the introduction by reducing the inclusion of overly technical or highly specific details.
 Instead, consider expanding on the broader conceptual framework that situates the paper within its scholarly context.
 This approach would help orient readers and provide a more accessible foundation for understanding the study.
-*Simplify the introduction*
+\color{black}
 
+Both the introduction and the literature review sections have been restructured
+and simplified. Some of the concepts initial unpacked in the introduction
+have now moved to the second section.
+
+\color{teal}
 Strengthen the Connection Between Hypotheses and Results: It may be beneficial to establish a clearer and more explicit link between the study's hypotheses and its findings.
 One way to achieve this could be by using section titles that are more conceptually informative, which would guide the reader through the logical progression of the argument and findings more effectively.
-*change section titles?*
+\color{black}
+
+The second section has been restructured and is now based on two hypotheses, 
+which are also used to structure the methodology and the results sections. All 
+section titles have been changed:
+
+- The mechanisms that shape the diffusion of a new technology
+- Results
+  - The three diffusion mechanisms at granular geographical scales
+  - Early adoption and latecomers
+- Conclusions
 
@@
 
-**explain, change title**
+Both figures has been re-titled:
+
+- Cumulative adoption of the Web, UK
+- LAD Web diffusion speed, 1996-2012
+
+Please note that the old Figure 3 has now moved to the end of the Results 
+section (now Figure 8).
 
@@
 
+This figure has also moved towards the end of the Results section as it is now 
+used to test the second hypothesis. The new title reads "Figure 9: Dynamics of 
+Wed diffusion based on LAD ranking" and the following paragraph explains how it 
+was created and what it reveals.
+
+> Next, the stability and volatility of these patterns is analysed to reveal 
+patterns of early adoption and leapfrogging. Figure 
+9 plots the relative ranking of LAD based on website density at the
+beginning and the end of the study period. To decrease noise, the average 
+ranking of 1996-1998 and 2010-2012 is plotted instead of the individual years. 
+Each line represents a single LAD and the colours depict the 
+5\textsuperscript{th} and 95\textsuperscript{th} percentile of the absolute
+difference of the LAD ranking between 1996-1998 and 2010-2012. In other words, 
+the 5\textsuperscript{th} percentile includes the LAD, the relative
+position of which remained unchanged between the beginning and the end of 
+the study period, both at the top and at the bottom of the hierarchy (small
+absolute difference in ranking). The 95th percentile includes LAD that either
+increased or decreased a lot their relative positions over time (high absolute 
+difference in ranking). Thus, Figure 9 reveals cases of 
+*leapfrogging* since the LAD at the top-right part of the figure managed to 
+jump at the top of the hierarchy at the end of the study period.
+
 \color{teal}
@@
 
-**yes**
+Yes. This is also reflected in the text. For instance:
+
+> This reflects the early concentration of high website density around London, which over time diffused as high website density clusters can be seen in other parts of the country away from London (Figure 5)
+
+> Eventually, as the adoption rate increased, more such clusters of high website density were formed and this is reflected both in the Moran’s I and the LISA maps in Figures 4 and 5.
+A similar pattern is observed for the extended dataset (see FiguresA.4 and A.5).
 
@@
 
-Thanks for going the extra mile to check the data used here! I agree with the 
+Thanks for going the extra mile to check the data used here. I agree with the 
 above calculation up until the findable postcode point as 1/5 of the archived 
@@
 Think of the website of a small company with one trading address. I think it is 
-fair to expect that website to have multiple webpages to present the different
+fair to expect a website to have multiple webpages to present the different
 products and services and then one webpage to showcase the address -- the 
@@
 a postcode. Previous estimates indicated that around 70% of all websites contain 
-place reference [@hill2009georeferencing] and this reference has no been added 
+place reference [@hill2009georeferencing] and this reference has now been added 
 in the paper.
@@
 were represented in the data. Most importantly though the paper does not model 
-the adoption of the Web from the general population but instead from firms. So, 
-although the data only represents 35% of the population of websites with a 
-registered address in the UK, it represents almost all the *relevant* archived
-websites -- that is websites using the commercial .co.uk SLD. As stated in the 
-paper, this has been the primary choice for commercial websites in the UK.
+the adoption of the Web from the general population, but instead from firms as 
+only commercial websites are considered. So, although the data only represents 
+35% of the population of websites with a registered address in the UK, it
+represents almost all the *relevant* archived websites -- that is websites using
+the commercial .co.uk SLD. As stated in the paper, this has been the primary 
+choice for commercial websites in the UK.
 
@@
 Yes, the British Library had obtained a copy of all of the .uk contents of the 
-Internet Archive for that period. From a data perspective, it wouldn't be correct
-to extend this dataset with data from other open archives such as the Common 
-Crawl as they use different crawlers and there will be inconsistencies. Most
-importantly though, this period covers the diffusion of the web from the very
-early stages until its maturity. This is also illustrated by the figure 
-depicting the cumulative adoption and the flattening right end of the line.
+Internet Archive for that period. From a data perspective, it wouldn't be 
+correct to attempt to extend this dataset with data from other open web archives,
+such as the Common Crawl, as they use different crawlers and there will be
+inconsistencies. Most importantly though, this period covers the diffusion of 
+the web from the very early stages until its maturity. This is also illustrated 
+by the figures depicting the cumulative adoption and the flattening right end of 
+the curve for both subsets of the data (Figure 2 and A1).
 
@@
 the aim of the paper and another line is now included in the second paragraph 
-of the introductions ("from the very early days until its maturity (1996-2012)").
+of the introduction ("from the very early days until its maturity (1996-2012)").
 
@@
 
-Point taken. Despite being aware of these studies, I omitted citing them as I 
-was primarily searching for studies testing the diffusion mechanisms over longer
-time periods. This error has now been corrected and the above paragraph has been 
-replaced with the following one, which emphasises the long time series motivation. 
-Regarding the actual references, @zook2000web, @zook2000web and @moss1997tracking 
-used relevant, domain name registration data, but the rest (e.g. @malecki2002economic 
-and @moss2000internet) focused on hard infrastructure. @gorman2002web examined 
-the head and branch locations of the top 40 e-business integration firms in the 
-USA, which although interesting is not directly relevant to this paper. I also
-went through of Feldman's work around 2002, but I did not find papers
-adopting a similar empirical perspective.
+Point taken. Despite being well aware of these studies, I omitted citing them as 
+I was primarily searching for studies testing the diffusion mechanisms over 
+longer time periods. This error has now been corrected and the above paragraph 
+has been replaced with the following one, which emphasises the long time series 
+motivation. Regarding the actual references, @zook2000web, @zook2001old and 
+@moss1997tracking used relevant, domain name registration data, but the rest 
+(e.g. @malecki2002economic and @moss2000internet) focused on hard 
+infrastructure. @gorman2002web examined the head and branch locations of the top 
+40 e-business integration firms in the USA, which although interesting is not 
+directly related to this paper. I also went through Feldman's work from around 
+2002, but I did not find papers adopting a similar empirical perspective.
 
@@
 [e.g. @blank2018local], or on social media users [e.g. @crampton2013beyond].
-Exceptions include early work from @zook2000web, @zook2000web and 
+Exceptions include early work from @zook2000web, @zook2001old and 
 @moss1997tracking, which utilised domain name registration data and use it to
@@
 Internet, they were based on cross-sectional, often spatially detailed, mapping 
-and on shallow time comparisons (e.g. 1993-1998) and did not observe the 
+and on short temporal comparisons (e.g. 1993-1998) and did not observe the 
 evolution and diffusion of the Web until its maturity. As we know from the 
@@
 [@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
-Web among firms can offer useful insights.
+Web can offer useful insights.
 
@@
 \color{black}
+
+See response to the next comment.
 
@@
 
-**for the last two comments, I need to soften the weightless argument throughout.**
-I have added a long time series argument and also said that there is heterogeneity
-[@leibowicz2016representing] among diffusion rates of different technologies.
-
-As per the 'framing' point above, the digital motivation para in the 
-introduction has been amended acordilngly.
+Point taken. The emphasis now is on the granularity of the analysis and also on 
+the long time series that allow to observe the technology almost from its 
+beginning until its maturity. The above text as well as other similar text 
+elsewhere in the paper have been amended accordingly.
 
@@
 
-This is very helpful. The first two sections have been restructured around two 
-hypotheses. I believe the structure and the flow of the paper is much clearer
-now.
+This is very helpful, thank you. The whole paper has now been restructured
+around two hypotheses. I believe the structure and the flow of the paper are 
+much clearer now.
 
@@
 
-**fix the abstract**
+The abstract has been amended accordingly:
 
-"shied away" has been replaced in the opening paragraph.
+> This paper investigates the spatial diffusion of a new technology that is the Web in the UK.
+It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms.
+Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012).
+Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales.
+They also highlight spatial heterogeneity and instances of leapfrogging.
 
@@
 
-@PERIS2021101565 is indeed an interesting paper, but it is not directly relevant 
-as it focuses on information diffusion rather than diffusion of new 
-technologies. Information diffuses through social and, consequently, urban 
-networks, but it is not adopted like new technologies. @bednarz2020pulled was 
-already cited and @feldman2015rknowledge has now been added. These references 
-are included in the second section.
+@PERIS2021101565 is indeed an interesting and, to humble opinion, novel paper, 
+but it is not directly relevant as it focuses on information diffusion rather 
+than diffusion of new technologies. Information diffuses through social and, 
+consequently, urban networks, but it is not adopted like new technologies. 
+Then, @bednarz2020pulled was already cited and @feldman2015rknowledge has now 
+been added. These references are included in the second section.
 
@@
 closest to economic geography. Importantly though, the *Regional Studies* paper 
-[@feldman2015rknowledge] paper is focused on inventions as observed by patent 
+[@feldman2015rknowledge] is focused on inventions as observed by patent 
 data contrary to the broader 'adoption of a new technology' approach of this 
@@
 
-**ADD SPATIAL GRANULARITY IN CONCLUSIONS**
-
 \color{teal}
@@
 
-**granularity**
+The novelty of the paper lies in the fact that there are hardly any studies
+that analysed the role of these three spatial mechanisms in the diffusion 
+of a new technology at such granular spatial scales. Importantly, the paper does 
+this by adopting a long time series perspective which captures the commercial 
+beginning of the Web until its maturity. The following paragraph in the 
+introduction highlights this:
 
-**HAVE A THINK, intro has been restructured**
+> This paper offers such a contribution: an economic geography study
+revealing the importance of spatial mechanisms for the diffusion of a
+digital technology that is the Web in the UK at a high level
+of spatial granularity 
+from the very early days until its maturity (1996-2012).
+It specifically focuses on the diffusion of commercial websites.
+It does so by employing a novel source of big data and machine learning
+algorithms to model the *active* engagement with Web -- that is creating and 
+maintaining instead of just browsing a website -- during that period
+and tests how different spatial and temporal mechanisms shaped its
+diffusion. This paper also exemplifies how the combination of data
+sources and state-of-the-art research methods which escape the
+traditional social science domain can offer new lenses to geographical
+research regarding the understanding of technological diffusion.
 
@@
 
-**HAVE A THINK, intro has been restructured**
+There are three paragraphs in the revised introduction that highlight the 
+paper's motivation. The first paragraph highlights the importance of this work 
+from a stakeholder perspective. The second emphasises the importance of 
+technology and its adoption in economic geography and macroeconomic growth 
+theories. The third one, motivates the paper from the standpoint of digital 
+geographies and connects it to that literature.
+
+> <!-- motivation: stakeholders -->
+The motivation for this paper lies in the fact that there are various
+stakeholders who are interested in knowing how new digital technologies
+diffuse over space and time and use this knowledge to make predictions
+about the diffusion of related future technologies. As per
+@leibowicz2016representing, historical studies agree that technologies
+diffuse differently in terms of times, rates, and geographies and can be
+driven by related policies [@victor1993]. @meade2021modelling highlight
+that a variety of actors have a direct interest in gaining such
+knowledge including network equipment suppliers, network operators,
+regulatory and local authorities. These processes and their effects vary
+a lot across scales: although the diffusion of a new technology might
+not be optimal at a local level, it might be beneficial from a global
+perspective as it could lead to faster diffusion to less advantaged
+places [@leibowicz2016representing]. Despite the spatial heterogeneity
+of such diffusion mechanisms and the policy relevance, there is a
+scarcity of studies analysing the diffusion of new technologies
+at a detailed geographical level because of the lack of detailed enough
+data and adequate research methods.
+
+<!-- motivation: regional development -->
+>This gap becomes even more prominent if we consider the importance of
+technology and its adoption in economic geography and macroeconomic
+growth theories. Simply put, differences in technological adoption and
+sophistication may lead to uneven development as new technologies can
+increase productivity and allow firms to introduce new products
+[@solow1957technical; @aghion1990model; @kemeny2011international].
+Linked to the subject of this paper, the Web and, more specifically, 
+commercial websites can be considered as the underpinning
+technology for the transformative process identified by
+@capello2024nexus as the *digital service economy*. Importantly, the
+focus here is not the Web in general, but instead commercial websites.
+After all, diffusion together with invention and innovation are
+considered the pillars of technological change [@das2022diffusion].
+Findings from previous studies illustrated that early engagement with
+digital technologies such as the Web led to longstanding positive
+productivity gains for these regions that hosted early adopters
+[@tranosuk]. Hence, obtaining a robust understanding of the spatial
+diffusion mechanisms is crucial in mapping future development trajectories.
+
+<!-- NEW motivation: intangible/digital -->
+> Another line of motivation has to do with the very nature of the Web as a 
+digital technology and how it evolved over time. The digital geographies field 
+has debunked premature claims about the Internet's placeless-ness by mapping 
+various of its facets and technological layers [@tranos2013death], but the lack 
+of relevant long time-series data prevented researchers from observing digital
+phenomena overtime. Most of these studies focused on its hard infrastructure 
+[e.g. @malecki2002economic; @moss2000internet], on Internet subscribers 
+[e.g. @blank2018local], or on social media users [e.g. @crampton2013beyond].
+Exceptions include early work from @zook2000web, @zook2000web and 
+@moss1997tracking, which utilised domain name registration data and use it to
+map and illustrate the spatial footprint of the web content. Pivotal as these 
+contributions have been in understanding the economic geographies of the 
+Internet, they were based on cross-sectional, often spatially detailed, mapping 
+and on short temporal comparisons (e.g. 1993-1998) and did not observe the 
+evolution and diffusion of the Web until its maturity. As we know from the 
+relevant literature, different technologies diffuse at different rates and 
+there is no theory that can explain these differences 
+[@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
+Web can offer useful insights.
 
@@
 -- .co.uk -- websites".
-
-**fixed for intro, lit rev  and data -- check the rest of the paper.**
 
@@
 
-Done. Restructured: theory, gaps, hypotheses.
+This section has been re-written and structured around the two hypotheses:
+
+> H~1~: The three diffusion mechanisms namely distance, urban hierarchy
+and the S-shaped pattern of the cumulative level of adoption shaped
+the diffusion of the commercial Web in the UK even at granular geographical 
+scales.
+
+> H~2~: The diffusion of the commercial Web is shaped by both early adoption 
+and latecomer effects.
 
@@
 
-Known processes at very detailed scale: explanatory power of these mechanisms at granular, local
-scales
+The knowledge gaps have been explicitly highlighted in the restructured 
+literature review section. For instance:
+
+> Although the above processes have been studied in innovation studies, the distancing of economic geographers from research questions about the diffusion of new technologies and innovations resulted to a knowledge gap regarding the explanatory power of these mechanisms at granular, local scales [@ding2010modeling].
+
+> Although the early adoption and leapfrogging literature has not paid much attention on cities and regions [@yu2018sustainability], the potential local economic benefits justify the need to investigate the dynamics of the web diffusion in different scales
+in the UK. Hence, the second empirical hypothesis states:
 
@@
 
-Added.
+The following two hypotheses have been added and the whole paper is now 
+structured around them:
+
+> H~1~: The three diffusion mechanisms namely distance, urban hierarchy
+and the S-shaped pattern of the cumulative level of adoption shaped
+the diffusion of the commercial Web in the UK even at granular geographical 
+scales.
+
+> H~2~: The diffusion of the commercial Web is shaped by both early adoption 
+and latecomer effects.
 
@@
 scales is of interest to economic geography as ..."
+
+The two scales are firstly introduced in the introduction: "Importantly, the analysis adopts two distinct geographical scales: the LAD and the most detailed scale of UK census areas, the Output Areas (OA)".
 
@@
 were matched against official postcode lookup files [@ons]. This was not 
-explicitly mentioned in the first submission on the paper and the previous 
+explicitly mentioned in the first submission of the paper and the previous 
 sentence has now been added in the relevant paragraph:
@@
 
-\noindent 2. Then, to remove the bias from link farms associated with a handful of postcodes
-having irregular high numbers of websites pointing to them, a targeted data 
-imputation procedure was developed. This is detailed in the data section -- see
-also Figure 1.
+\noindent 2. Then, to remove the bias from link farms associated with a handful 
+of postcodes having irregular high numbers of websites pointing to them, a 
+targeted data imputation procedure was developed. This is detailed in the data 
+section -- see also Figure 1.
 
-\noindent 3. Regarding the points about subsidiary company location and event postings, 
-these cannot be directly observed and distinguished. However, a process that has 
-been successfully used in previous studies has been adopted here. Firstly, given 
-that the focus of the paper is commercial websites, the relevant literature 
-illustrates that such websites serve specific purposes:
+\noindent 3. Regarding the points about subsidiary company location and event
+postings, these cannot be directly observed and distinguished. However, a 
+process that has been successfully used in previous studies has been adopted 
+here. Firstly, given that the focus of the paper is commercial websites, the 
+relevant literature illustrates that such websites serve specific purposes:
 
-> they are used to diffuse information, support online transactions and share 
+> ... they are used to diffuse information, support online transactions and share 
 opinions [@THELWALL2000441; @blazquez2018big].
@@
 to 10 postcodes and the results are presented in the appendix. This includes 94%
-of all the archived, geolocated commercial websites. For this step, websites 
-with multiple postcodes were counted multiple times, one for each different 
-postcode. All the results are qualitatively the same and only marginally 
-different, which is indication of the robustness of the website geolocation 
-strategy.
+of all the archived, geolocated commercial websites for 2000. For this step, 
+websites with multiple postcodes were counted multiple times, one for each 
+different postcode. All the results are qualitatively the same and only 
+marginally different, which is an indication of the robustness of the website 
+geolocation strategy.
 
@@
 
-**growth as an appendix, spatial dependencies have been included**
+Because of the restructuring of the methods section, which now includes all the 
+details of the modelling exercise, Equation 2 is followed by Equation 3 that 
+clarifies the spatial and temporal dependencies. All these details are described
+in the following paragraph, which is placed just before Equation 3:
+
+> To capture the hierarchical mechanism the model include as predictors a
+one year lag of website density in London, the largest city in the UK, a
+one year lag of the website density in the nearest city and the same for
+the nearest retail centre. Due to the small sizes of the retail 
+centres, the latter is only relevant for and included in the OA models and not 
+in the LAD models. In addition, the model includes the distance to London, the 
+nearest city and the nearest retail centre. The underlying logic is that the 
+level of website adoption in a spatial unit depends on the level of the 
+adoption in places further up in the urban hierarchy the previous year. The
+inclusion of the distance variables incorporates spatial structure into
+the hierarchy argument. To depict the neighbourhood mechanism, the
+website density of the neighbouring spatial units in the previous year
+is employed. This is calculated using rook continuity. Again, the
+underpinning rationale is that the level of web adoption within a
+spatial unit depends on the level of web adoption in the neighbouring
+spatial units the year before. This represents the 'hitting nearby
+locations first' argument. Therefore, the spatial and temporal lag of
+the website density in LAD and OA is calculated. Lastly, the temporal
+pattern of the cumulative adoption is captured by a time trend variable.
+Hence, the model will follow the following generic form (Eq. 3):
+
+Regarding the growth model point, growth rates can be and have been calculated
+and modelled for LAD. However, such a strategy would have been problematic for 
+OA, because of the very large frequency of OA without any websites. As 
+mentioned in the text:
+
+> In 1997 only 1% of the UK OA were associated
+with at least one website. This should not come as a surprise given that this
+was the very beginning of the commercial Internet and any activities
+with a digital footprint were concentrated in a handful of areas, as 
+illustrated in Figure 5. At the end of the study
+period almost half of the UK OA were not associated with a website.
+Again, given the granularity of the data this was expected.
+
+Therefore, if it was to to calculate growth rates for OA,  a small constant
+should have been added to all OA. Given how many 0s exist, this would have been
+problematic and add unnecessary noise. Given that a multi-scale and granular
+approach was important for this paper and part of its novelty, the levels 
+modelling framework was adopted. However, the current revised version of the 
+paper now reports such a LAD growth model in an appendix. The following text has 
+been added in the methods section:
+
+> An alternative specification could model the growth of website density. However, 
+such a strategy is problematic because of the large frequency of 
+OA without websites, which is to be expected given their small size and high 
+granularity. Still, as a robustness text, the results of a growth model for
+LAD are presented in the Appendix.
+
+Then, the following text has been added in the Results section:
+
+> In addition, Table A3 in the Appendix
+presents the accuracy results of a LAD model the dependent variable of which 
+is the yearly growth rate of website density. Even for such a specification
+(see Equation 7 in the Appendix) the predictive capacity of
+model is exceptionally high ($R^2 = 0.5$).
 
@@
 
-**This is only for the map, this is a critical point in diffusion studies**
+Please see my response to the previous point regarding modelling growth. 
+Specifically for this point, the map with the 'binarised' speed (now Figure 8) 
+has been moved further down (Section 4.2) to address H~2~. Please note that 
+the modelling part does not use this binarised variable, but instead continuous 
+ones. Regarding the map, its aim is to depict heterogeneity in Web adoption. I 
+could have, of course, plotted the year that each LAD reached their inflection 
+point, which is very similar to "the time required to reach 50% saturation", 
+using 4-5 bins. But I think the comparison with the country average makes it 
+more accesible for the reader. The following text has been added in the first 
+paragraph of section 4.2:
 
-H1: first descriptives, then modelling
-- S curves
-Figure 1: UK curve, Figure 2: some LAD curves. link to a gif. 
-keep fast slow map, briefly touch upon and say it will be further discussed below
-- neighbourhood
-- hierarchy
-- modelling
+> Inflection points have been used in the relevant literature to compare diffusion processes across different settings and technologies [@cherp2021national; @woo2017exploring] as they can enable more insightful comparisons than, for example, the average growth over the 1996-2012 study period.
 
-H2: link map with ranks. Explore OA stability within LADs
+In addition, the inflection ($t_0$) estimates, which used for the above map 
+are presented in detail in the appendix and are also discussed in the main body 
+of the paper:
+
+> Importantly, there is obvious spatial clustering of fast and slow LAD.
+Out of the ten fastest LAD based on when they reached their inflection point (see detailed Table A.2 in the Appendix), nine were located in the South East of England and London and one in Cambridge.
+
+They are also presented in Table 6 for the late adopters discussion.
 
@@
 
-**reveal that theories are valid for local scales and inform local tech policies**
+A new figure (Figure 3) and the following text have been added to address this
+point.
+
+> What escaped from previous studies was εexploring whether such a pattern also 
+describes cumulative adoption of new technologies at sub-national scales. To 
+test this, the same  logistic function is estimated for each of the 374 LAD. The
+detailed results are presented in Table \ref{table.s.lads} in the Appendix. 
+The key finding is that for 342 LAD (91% of all LAD) the $R^2 > 0.9$ for the 
+logistic function, which indicates that even at local scale the cumulative 
+adoption of the Web is characterised by a pattern similar to the one observed 
+for countries. Graphically, all these LAD curves are plotted in Figure 
+\ref{s_lad_uk}. The different colour lines illustrate whether a LAD reached $t_0$ 
+faster or slower than the UK in total. This heterogeneity is further discussed 
+later in this section.
+
+> This is the first time that such cumulative adoption curves have been estimated
+for areal units as small as the UK LAD. Understanding and modelling the dynamics
+of technological diffusion and, ultimately, being able to predict diffusion 
+rates at such small local scales can directly support the deployment of 
+infrastructural networks (e.g. 5G) and local technology policies. The above are 
+in line with previous studies exploring the diffusion of web technologies 
+[@PAPAGIANNIDIS2015308], social media [@lengyel2020role] and online
+shopping [@bakher2013diffusion].
 
@@
 
-**relates to local capabilities or diffusion hierarchies**
+The results section has been restructured in order to match the two hypotheses
+structure. The early adopters part has been further developed with more 
+empirical results and now answers the second hypothesis (H~2~). The key message
+is that the diffusion of the Web at the detailed scale of LAD is 
+characterised by spatial heterogeneity, well-established spatial patterns and 
+both stability and volatility. Although the relevant literature neglected local
+scales, the above findings illustrated that such local dynamics exists and 
+worth further consideration and investigation especially regarding potential 
+local economic effects.
 
@@
 
-**think, growth rates will not work for OA**
+Please see previous responses about growth rates, zeros and OA. Specifically for 
+this point, Moran's I and the Gini coefficient offer some first insights about the
+two of the three diffusion mechanisms that H~1~ tests. Section 3 has now been 
+restructured based on the two hypotheses, and the following text describes their
+value:
+
+> To assess the second diffusion mechanisms, Figures 4 and 5  offer a first insight into whether a neighbourhood or, in other words, a distance mechanism underpins the diffusion of the Web in the UK as they present the Moran's I and the LISA maps of website density respectively for LAD and OA.
+
+More broadly, on top of the fact that in order to calculate growth rates for OA 
+a small constant should have been added to overcome the 0s problem, it is not 
+clear how their distribution could offer insights about the two diffusion 
+mechanisms. It is also unclear to me how such distribution could be used -- the 
+distribution of yearly growth rates for every LAD and OA or for all LAD and OA? 
+Importantly there are c. 230,000 OA.
 
@@
 
-**hierarchy**
+The retail centres discussion has now been moved in the methods sections, so it
+appears much earlier in the paper. In addition, the following text has been 
+modified in the relevant paragraph:
+
+> To capture the hierarchical mechanism the model include as predictors a
+one year lag of website density in London, the largest city in the UK, and a
+one year lag of the website density in the nearest city. Both variables clearly
+illustrate the urban hierarchy in the UK. Especially for the OA models, a 
+third predictor is added to capture hierarchy at a much more granular scale:
+a one year lag of the website density in the the nearest retail centre [^4]. As
+the literature illustrates, retail centres are key components of urban systems
+[@dennis2002central, @jones2021reframing]. Due to the small sizes of the retail 
+centres, the latter is only relevant for and included in the OA models and not 
+in the LAD models.
+
+> [^4]: The retail centres data have been obtained by the Consumer Data Research Centre, an ESRC data investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1: <https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators>.
 
@@
 \color{black}
+
+The modelling details have now been moved to the methods section.
 
@@
 
+This section has been restructured based on the two hypotheses. I believe the 
+key messages are now more clearly communicated.
+
 \color{teal}
@@
 
+Thank you, this was very helpful. The whole paper has been restructured based 
+on the above suggestions.
+
 ## References

### 2025-03-12 paper/submission2/anonymised.qmd
paper v2

--- 2025-01-21 paper/submission2/tranos2025.qmd
@@
 title: "A multi-scale story of the diffusion of a new technology: the web"
-author:
-  - name: Emmanouil Tranos
-    email: e.tranos@bristol.ac.uk
-    affiliations: 
-      - id: some-tech
-        name: University of Bristol
-        department: 
-        address: 
-        city: 
-        state: State
-        postal-code: UK
-    corresponding: e.tranos@bristol.ac.uk
-address:
-  - attributes:
-        corresponding: true
-    note: This is the first author footnote.
 abstract: |
-    Despite the broad consensus that differences in technological adoption and sophistication can lead to uneven trajectories of local development as new technologies can increase productivity and allow firms to introduce new products and services, economic geographers have shied away from exploring the diffusion of digital technologies across space and time.
-This paper fills this gap by exposing the roles of different spatial mechanisms in the diffusion of an intangible, digital technology at very granular geographical scales.
-It does so by employing a novel source of data and machine learning algorithms to model the active engagement with Web during 1996-2012.
+    This paper investigates the spatial diffusion of a new technology that is the Web in the UK.
+It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms.
+Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012).
+Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales.
+They also highlight spatial heterogeneity and instances of leapfrogging.
 keywords: 
@@
 
-<!-- ADD SOMETHING ABOUT MULTIPLE SCALES -->
-
-<!-- geographers lack of interest for diffusion + motivation -->
+<!-- geographers lack of interest for diffusion + gap -->
 Economic geographers were always interested in how new technologies and
@@
 did not investigate diffusion mechanisms at granular geographical scales to 
-provide local policy insights, for instance, regarding infrastructural d
-evelopment and local technology policies. 
+provide local policy insights, for instance, regarding infrastructural 
+development and local technology policies. 
 A potential explanation of the lack of geographical studies exploring the 
@@
 This paper offers such a contribution: an economic geography study
-revealing the importance of spatial mechanisms for the diffusion of an
-intangible, digital technology that is the Web in the UK at a high level
+revealing the importance of spatial mechanisms for the diffusion of a
+digital technology that is the Web in the UK at a high level
 of spatial granularity 
@@
 of such diffusion mechanisms and the policy relevance, there is a
-scarcity of studies analysing the diffusion of new digital technologies
+scarcity of studies analysing the diffusion of new technologies
 at a detailed geographical level because of the lack of detailed enough
@@
 <!-- description of the methods and results -->
-<!-- AMMEND I AFTER CHANGE THE RESULTS -->
 To analyse these data, a diverse set of methods are employed. Firstly,
@@
 the Local Authority Districts (LAD), illustrating, for the first time,
-such S-shaped diffusion patterns for local areas. This system level
-analysis also reveals insights about the volatility of the adoption of
-the Web. Then, exploratory spatial data analysis offers some first
-indications of the importance of the different spatial diffusion
-mechanisms in shaping the diffusion of the Web in the UK. Importantly,
-the analysis adopts two distinct geographical scales: the LAD and the
-most detailed scale of UK census areas, the Output Areas (OA). Lastly, a
+such S-shaped diffusion patterns for local areas. The exploratory spatial data
+analysis offers some first indications of the importance of the different 
+spatial diffusion mechanisms in shaping the diffusion of the Web in the UK.
+Importantly, the analysis adopts two distinct geographical scales: the LAD and 
+the most detailed scale of UK census areas, the Output Areas (OA). Then, a
 novel methodological framework based on Machine Learning (ML) algorithms
-is developed to test the role of such mechanisms at these two scales.
-The results reveal, for the first time, the importance of different
-spatial diffusion mechanisms in the diffusion of an intangible, digital
-technology at very granular geographical scales.
+is developed to test the role of such mechanisms at these two scales. The 
+results reveal, for the first time, the importance of different spatial 
+diffusion mechanisms in the diffusion of a digital technology at very granular
+geographical scales. The analysis also reveals that the diffusion of the Web 
+is characterised by spatial heterogeneity, well-established spatial patterns 
+and both stability and volatility.
 
 <!-- contents -->
-<!-- AMMEND I AFTER CHANGE THE RESULTS -->
 The structure of the paper goes as follows. @sec-litreview reviews the
-key concepts in the diffusion literature as well as relevant empirical
-studies. Then, @sec-datamethods describes the data and methods used in
-this paper. The results of the different stages of the analysis are
-presented in @sec-results. The discussion of the results and the
-conclusions of the paper are offered in @sec-conclusions.
+mechanisms that shape the diffusion of a new technology and sets up the two 
+hypotheses that are empirically tested in @sec-results. @sec-datamethods 
+describes the data and methods used in this paper. The discussion of the 
+results and the conclusions of the paper are offered in @sec-conclusions.
 
@@
 periphery. Although adoption pace might be higher when new technologies
-finally arrive to the periphery, the saturation levels there may never
+finally arrive to the periphery, the saturation levels may never
 reach the ones in the core because of the lack of infrastructure or
@@
 determinants of the timing, speed and the level of mobile diffusion
-within China. @bakher2013diffusion compared different the growth curves
+within China. @bakher2013diffusion compared different growth curves
 as a means to model the diffusion of online shopping in Australia during
@@
 urban scaling and distance are the key drivers of spatial diffusion.
-Based on the above, the main empirical hypothesis states:
+Based on the above, the main empirical hypothesis that is tested in this paper
+states:
 
 > H~1~: The three diffusion mechanisms namely distance, urban hierarchy
-and the the S-shaped pattern of the cummulative level of adoption shaped
+and the S-shaped pattern of the cumulative level of adoption shaped
 the diffusion of the commercial Web in the UK even at granular geographical 
@@
 
-> H~2~: The diffusion of the commecrial Web is shaped by both early adoption 
-and late comer effects.
-
-These are the two hypotheses are empirically tested in @sec-results using the 
-data and methods described in the next section.
+> H~2~: The diffusion of the commercial Web is shaped by both early adoption 
+and latecomer effects.
+
+These two hypotheses are empirically tested in @sec-results employing 
+the data and methods described in the next section.
 
@@
 
-[^2]: [https://archive.org/](See%20https://archive.org/).
+[^2]: [https://archive.org/](https://archive.org/).
 
@@
 extended sample includes 94 percent of all the archived websites in
-2000.
+2000 (see Table \ref{f2000}).
 
@@
          'F (%)' = perc,
-         'Cummulative F' = cumperc)
+         'cumulative F' = cumperc)
 
@@
 website counts for these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1
-6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1000
+6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1,000
 websites pointing to them, were replaced with predicted values based on
@@
 the data imputation into perspective, this process affected 6 out of the
-557,808 postcodes included in the data. As the @sec-results illustrates,
+557,808 postcodes included in the data. As @sec-results illustrates,
 this led to a small increase of the model predictive power.
@@
 <!-- Multi-scale -->
-To empirically test the two hypotheses stated in @sec-litreview, τhe website 
+To empirically test the two hypotheses stated in @sec-litreview, the website 
 density metric was calculated for two different scales the Local Authority
-Districts (LAD) and the Output Areas (OA). The former is an administrative unit 
-and there are 374 LAD in the UK. The latter is the smallest census-based 
-geographical unit and there are c. 230,000 of them in the UK. To create the 
-website density metric, the yearly website counts at the LAD level are 
-standardised by the number of firms in LAD to avoid biases associated with LAD 
-hosting a large number of firms. Given that there is no such readily available 
-statistic for the OA, the actual OA level counts are used. However, because of 
-the the consistent spatial definition of OA (they host 40-250 households),[^3] 
-website counts in OA are interpreted as a density metric too.
+Districts (LAD) and the Output Areas (OA). The former is the lowest
+administrative unit in the UK and there are 374 LAD. The latter is the smallest
+census-based geographical unit and there are c. 230,000 of them in the UK. To 
+create the website density metric, the yearly website counts at the LAD level 
+are standardised by the number of firms in LAD to avoid biases associated with 
+LAD hosting a large number of firms. Given that there is no such readily 
+available statistic for the OA, the actual OA level counts are used. However, 
+because of the the consistent spatial definition of OA (they host 40-250
+households),[^3] website counts in OA are interpreted as a density metric too.
 
@@
 
-<!-- system-level analysis -->
-These website density metrics are used for the three different stages of
-the analysis. Firstly, a system-level analysis explores how the
-diffusion of web technologies in the UK fits with the well-established
-S-curve of cumulative adoption. To do so, the following logistic
-function (Equation \ref{eq:s}) is estimated for the whole of the UK and
-for each LAD separately:
+<!-- S -->
+These website density metrics are used for the different stages of the analysis. 
+Starting with the first hypothesis (H~1~), to test whether the S-curve 
+accurately describes the cumulative adoption of the Web at a local scale, the
+following logistic function (Equation \ref{eq:s}) is estimated for the whole of 
+the UK and then for each LAD separately:
 
@@
 the UK to delineate whether a LAD reached that point faster or slower
-than the country average. Importantly, an accuracy criterion was imposed
+than the country average. This heterogeneity is discussed further when testing 
+the second hypothesis (H~2~). Importantly, an accuracy criterion was imposed
 and only LAD with $R^2 > 0.9$ were included in this analysis. To
@@
 <!-- **minimum R2 = 95%** in @wilson201281, see also @grubler1990rise -->
-
 <!-- The literature usually uses the saturation level as the asymptote. I am -->
-
 <!-- using the total number of websites as we cannot compute a rate. -->
 
-The system-level analysis also focuses on the volatility of the web
-adoption to depict places with high concentration of early adopters and,
-equally, latecomers. To do so, the change of the ranking of the UK LAD
-over time is plotted and discussed. Both the S-curve and the volatility
-analysis focus only on LAD as the very large number of OA would have
-made such analysis difficult to visualise and interpret.
-
 <!-- ESDA -->
-Next, exploratory spatial data analysis depicts whether the two main
-drivers of spatial diffusion -- namely neighbourhood and hierarchy --
-underpin the diffusion of the Web in the UK. To capture the former and
-following @ding2010modeling the Moran's I and the Local Index of Spatial
-Association (LISA) are estimated for the website density. To address the
+Next, exploratory spatial data analysis provides some first insights of whether 
+the other two drivers of spatial diffusion -- namely neighbourhood and 
+hierarchy -- underpin the diffusion of the Web in the UK. To capture the former 
+and following @ding2010modeling the Moran's I and the Local Index of Spatial
+Association (LISA) are estimated for the website density. To capture the
 hierarchy mechanism, the Gini coefficient -- an established metric of
@@
 
+<!-- conceptual model -->
+To formally test H~1~ a modelling framework is developed to assess the 
+above diffusion mechanisms. The overarching aim is to build a model that can 
+test how well these three mechanisms can predict website density:
+
+\begin{align}
+Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S\text-curve_{t}\label{eq:rf.generic}
+\end{align}
+
+<!-- <!-- conceptual model --> 
+<!-- The models include variables that capture the three processes that the -->
+<!-- relevant literature and the descriptive analysis highlighted. Namely, -->
+<!-- the models capture: (i) a hierarchical mechanism with diffusion running -->
+<!-- from main centres to secondary ones, (ii) a neighborhood mechanism -->
+<!-- according to which diffusion first hits nearby locations, and (iii) the -->
+<!-- rather canonical pattern of diffusion over time as reflected in the -->
+<!-- S-shaped pattern in the cumulative level of adoption. -->
+
+<!-- detailed features -->
+To capture the hierarchical mechanism the model include as predictors a
+one year lag of website density in London, the largest city in the UK, and a
+one year lag of the website density in the nearest city. Both variables clearly
+illustrate the urban hierarchy in the UK. Especially for the OA models, a 
+third predictor is added to capture hierarchy at a much more granular scale:
+a one year lag of the website density in the the nearest retail centre [^4]. As
+the literature illustrates, retail centres are key components of urban systems
+[@dennis2002central; @jones2021reframing]. Due to the small sizes of the retail 
+centres, the latter is only relevant for and included in the OA models and not 
+in the LAD models. In addition, the model includes the distance to London, the 
+nearest city and the nearest retail centre. The underlying logic is that the 
+level of website adoption in a spatial unit depends on the level of the 
+adoption in places further up in the urban hierarchy the previous year. The
+inclusion of the distance variables incorporates spatial structure into
+the hierarchy argument. To depict the neighbourhood mechanism, the
+website density of the neighbouring spatial units in the previous year
+is employed. This is calculated using rook continuity. Again, the
+underpinning rationale is that the level of web adoption within a
+spatial unit depends on the level of web adoption in the neighbouring
+spatial units the year before. This represents the 'hitting nearby
+locations first' argument. Therefore, the spatial and temporal lag of
+the website density in LAD and OA is calculated. Lastly, the temporal
+pattern of the cumulative adoption is captured by a time trend variable.
+Hence, the model will follow the following generic form (Eq.
+\ref{eq:model}):
+
+[^4]: The retail centres data have been obtained by the Consumer Data Research Centre, an ESRC data investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1: <https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators>.
+
+\begin{align} \label{eq:model}
+Website\,Density_{t} \sim Distance\,London +
+Website\,density\,London_{t-1} +\notag\\
+Distance\,Nearest\,City +
+Website\,density\,Nearest\,City_{t-1} +\notag\\
+Distance\,Nearest\,Retail_{i} +
+Website\,density\,Nearest\,Retail_{t-1} +\notag\\
+W*\, Website\,density_{t-1} +\notag\\ 
+year_{t}
+\end{align}
+
+<!-- 4 models -->
+Specifically, four different models are estimated. Firstly, all the data points 
+for the OA and LAD are utilised in order to build two models and assess their
+capacity to predict the adoption of the Web at the two different scales. These 
+two models will reveal the predictive capacity of the diffusion mechanisms and 
+also demonstrate how the importance of such variables changes across
+different scales. The next two sets of models are again trained at these
+two scales: OA and LAD. However, instead of using all the data points,
+the OA and the LAD from one of the twelve UK regions are held out for
+the model training. Then, the trained model is used to predict website
+density for the OA or the LAD of the held-out region. This process takes
+place recursively for all twelve UK regions. The regional differences in
+the predictive capacity of the different samples will reveal how
+dissimilar these spatial process are across regions and, importantly, at
+different scales.
+
 <!-- RF -->
-Lastly, a modelling framework is developed to test the above diffusion
-mechanisms. The overarching aim is to build a model that can test how
-well these mechanisms can predict website density:
-
-\begin{align}
-Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S-curve_{t}\label{eq:rf.generic}
-\end{align}
-
-\noindent Equation \ref{eq:rf.generic} is estimated using Random Forest
+Equation \ref{eq:model} is estimated using Random Forest
 (RF). This is a popular ML algorithm for both regression and
@@
 
+<!-- RF -->
 RF is a tree-based ensemble learning algorithm [@breiman2001random]. It
@@
 
+<!-- CV -->
+It needs to be highlighted here that the cross validation (CV) for all
+models is spatially and temporally sensitive. Instead of using 10 random
+and space- and time-agnostic folds, the `CAST` package is employed which
+allows holding back data points from specific years and spatial units (regions)
+and use them for testing in order to estimate the model performance
+[@meyer2018improving].
+
+<!-- assessment -->
+To assess the predictive power of the model, three broadly utilised
+metrics are employed: the coefficient of determination ($R^2$), mean
+absolute error (MAE) and root mean square error (RMSE):
+
+\begin{align}
+R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
+\end{align}
+\begin{align}
+MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
+\end{align}
+\begin{align}
+RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
+\end{align}
+
+\noindent $y_{k}$ is the $k^{th}$ observation of the dataset, which
+consists of $N$ observations in total. $\hat{y_{k}}$ is the $k_{th}$
+predicted value for the dependent variable and $\overline{y_{k}}$ is the
+average value of $y$. The last two metrics are expressed in the same
+units as the dependent variable -- websites per firm for the LAD modes
+and the number of websites for the OA models -- while the first one is
+the coefficient of determination between the observed and the predicted
+values of website adoption. Regarding $MAE$, it is the absolute
+difference between the observed and the predicted website adoption.
+While $MAE$ does not penalise for large errors, $RSME$ does, as it is
+proportional to the squared difference between the observed and the
+predicted trade flows. Hence larger errors weigh more for $RMSE$
+[@pontius2008components].
+
+An alternative specification could model the growth of website density. However, 
+such a strategy is problematic because of the large frequency of 
+OA without websites, which is to be expected given their small size and high 
+granularity. Still, as a robustness text, the results of a growth model for
+LAD are presented in the Appendix.
+
+<!-- H2 -->
+Lastly, to test the second hypothesis (H~2~) about the early adoption and the
+latecomer effects, a different research framework is developed. Firstly, the 
+S-curve and the estimated $t_0$ for every LAD are employed and mapped in order 
+to depict the spatial heterogeneity and specific patterns of the speed of the 
+diffusion of the Web. Then, the volatility and the stability of the diffusion 
+across the different LAD is analysed using tools from the complex systems domain 
+[@Batty2009] namely changes in rankings of LAD based on website density. Both 
+the S-curve estimation and the volatility analysis focus only on LAD as the 
+very large number of OA would have made such analysis difficult to visualise 
+and interpret.
+
 # Results {#sec-results}
 
-## System-level analysis
+## The three diffusion mechanisms at granular geographical scales
 
 <!-- S-shaped -->
-
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk}Grwoth curve, UK"}
+Figure \ref{s_uk} plots the cumulative adoption of the Web in the UK as a whole
+during the 1996-2012 period. In line with the theoretical expectations discussed
+in the previous section, the cumulative adoption of the Web is well represented 
+by a curve that resembles the letter S. The vertical line in the beginning of 
+2003 (2003.158) illustrates the point when the modelled cumulative adoption was 
+equal to 50% of the maximum. This $t_0$ point is known as the *inflection* point 
+and signals the maximum adoption speed. The results remain unchanged when the 
+extended subset of websites, which include up to 10 unique postcodes is used
+(see Figure \ref{s_uk10} in the Appendix). Importantly though, the extended
+subset of websites indicates a saturation level close to 100% at the end of 
+the study period.
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk}Cumulative adoption of the Web, UK"}
 path.image <- paste0(path, "/outputs/s/s_uk_per_firm.png")
@@
 
-Figure \ref{s_uk} plots the cumulative adoption of the Web in total in
-the UK during the 1996-2012 period. Not surprisingly, the cumulative
-adoption is well represented by a curve that resembles the figure S. The
-vertical line around year 2003 illustrates the point where the modelled
-cumulative adoption is equal to 50% of the maximum. This $t_0$
-*inflection* point signals the maximum adoption speed and is used here
-to determine whether a UK LAD reached that inflection point earlier or
-later than the UK average. Specifically, the S curve and the inflection
-point are estimated for every LAD individually and then compared against
-the UK average. LAD that reached their inflection point earlier than the
-country average are labelled as *fast* and the rest as *slow*. Figure
-\ref{s_map} maps this pattern. There is a clear high concentration of
-fast adopting LAD in the South and East of England as expected. However,
-there is also such a concentration in Wales, Scotland and the North West
-of England. At a more detailed level, there are some expected examples
-of LAD with relevant industrial backgrounds delineated as fast: the City
-of London, a world-renowned cluster of finance industries
-[@cook2007role], and Reading, a town with high-tech service industries
-in proximity to London and its main airport, Heathrow
-[@england2005polynet]. However, there are also LAD which were expected
-to appear as fast -- e.g. Hackney in central London and Bristol, a
-well-established creative cluster [@oatley1999cultural;
-@bassett2002cultural] -- but were delineated as slow. Nevertheless,
-there is obvious spatial clustering of fast and slow LAD. Importantly,
-out of the ten fastest LAD, nine were located in the South East of
-England and London and one in Cambridge (see Table \ref{table.s.lads} in
-the Appendix). The above observations remain almost unchanged when the
-analysis includes websites with up to 10 unique postcodes -- see Figures
-\ref{s_uk10} and \ref{s_map10} in the Appendix -- advocating towards the
-robustness of the analysis.
-
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map}LAD adoption rates"}
-path.image <- paste0(path, "/outputs/s/speed_map.png")
+What escaped from previous studies was εexploring whether such a pattern also 
+describes cumulative adoption of new technologies at sub-national scales. To 
+test this, the same  logistic function is estimated for each of the 374 LAD. The
+detailed results are presented in Table \ref{table.s.lads} in the Appendix. 
+The key finding is that for 342 LAD (91% of all LAD) the $R^2 > 0.9$ for the 
+logistic function, which indicates that even at local scale the cumulative 
+adoption of the Web is characterised by a pattern similar to the one observed 
+for countries. Graphically, all these LAD curves are plotted in Figure 
+\ref{s_lad_uk}. The different colour lines illustrate whether a LAD reached $t_0$ 
+faster or slower than the UK in total. This heterogeneity is further discussed 
+later in this section.
+
+This is the first time that such cumulative adoption curves have been estimated
+for areal units as small as the UK LAD. Understanding and modelling the dynamics
+of technological diffusion and, ultimately, being able to predict diffusion 
+rates at such small local scales can directly support the deployment of 
+infrastructural networks (e.g. 5G) and local technology policies. The above are 
+in line with previous studies exploring the diffusion of web technologies 
+[@PAPAGIANNIDIS2015308], social media [@lengyel2020role] and online
+shopping [@bakher2013diffusion].
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_lad_uk}Cumulative adoption of the Web, LAD"}
+path.image <- paste0(path, "/outputs/s/lad_per_firm_all.png")
 knitr::include_graphics(path.image)
@@
 
-<!-- discussion -->
-The S-curves of the cumulative adoption of Web in the UK in total and in
-the UK LAD demonstrate a pattern well aligned with all previous studies
-discussed in @sec-litreview. Importantly, the above findings are in line
-with previous studies exploring the diffusion of web technologies
-[@PAPAGIANNIDIS2015308], social media [@lengyel2020role] and online
-shopping [@bakher2013diffusion]. Similar to @beardsell1999spatial, who
-analysed the spatial evolution of computer industry employment across
-317 US metro areas, spatial heterogeneity is evident in the diffusion
-speed when the scale of analysis is as detailed as the LAD. Importantly,
-no previous studies estimated such cumulative adoption curves for areal
-units as small as the UK LADS.
-
-<!-- rankings -->
-The next step assesses the stability and volatility of the LAD in terms
-of their adoption of web technologies. As highlighted in the literature
-[@risk_perceptions], different agents have different perceptions about
-and levels of acceptance of the risks and the potential economic returns
-associated with the adoption of new technologies -- see for instance the
-seminal work of @venkatesh2000theoretical. To reveal such aggregated
-patterns, Figure \ref{rank} plots the ranking of UK LAD based on website
-density. To decrease noise, the average ranking of 1996-1998 and
-2010-2012 is plotted instead of the individual years. The colours depict
-the 5th and 95th percentile of the absolute difference of the LAD
-ranking between 1996-1998 and 2010-2012. While there are quite a few
-obvious cases of LAD that maintained their position between the
-beginning and the end of the study period both at the top or at the
-bottom of the hierarchy, there are also quite a few LAD that changed
-drastically their position. Some of these LAD enjoyed a process that at
-the first instance looks like *leapfrogging* since they managed to jump
-at the top of the hierarchy despite their slow start. There is extensive
-literature regarding the potential benefits of technological
-leapfrogging. The underpinning argument is that latecomers can adopt and
-benefit from new technologies that have been developed elsewhere without
-incurring the hefty initial R&D costs [@teece2008firm]. Although the
-leapfrogging literature does not pay much attention to cities and
-regions [@yu2018sustainability], previous research highlighted the long
-term and sustained productivity benefit of the early adopters of web
-technologies in the UK [@tranosuk]. So, although the LAD system is
-volatile, it is not clear whether the LAD with high concentration of
-late adopters will gain any latecomer-related benefits in a way similar
-to countries experiencing technological leapfrogging. Still, the
-opportunity to catch up that @perkins2005international illustrated at
-the country level, is also visible at the much more detailed LAD scale.
-As previously, this is consistent with including websites with up to 10
-postcodes (see Figure \ref{rank10}).
-
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank}Dynamics of wed diffusion (LAD)"}
-path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av.png")
-knitr::include_graphics(path.image)
-```
-
-## Exploratory analysis
-
 <!-- Neighbourhood effect -->
-Figures \ref{morani} and \ref{lisa} offer a first insight into whether a
-neighbourhood mechanism underpins the diffusion of the Web in the UK as
-they present the Moran's I and the LISA maps of website density
-respectively for LAD and OA. Starting from the former, spatial
-autocorrelation was higher in the beginning of the study period and then
-after 2000 dropped slightly and stabilised around 0.2. This reflects the
+To assess the second diffusion mechanisms, Figures \ref{morani} and \ref{lisa} 
+offer a first insight into whether a neighbourhood or, in other words, a 
+distance mechanism underpins the diffusion of the Web in the UK as they present 
+the Moran's I and the LISA maps of website density respectively for LAD and OA.
+Starting from the former, spatial autocorrelation was higher in the beginning of 
+the study period and then after 2000 dropped slightly and stabilised around 0.2. 
+This reflects the
 early concentration of high website density around London, which over
@@
 spatial clustering mechanism. All in all, the exploratory spatial data
-analysis advocates towards an underpinning neighbourhood mechanism and
-the different scales of analysis illustrate how it evolved differently
-over time. <!-- discussion --> Interestingly, the magnitude of Moran's I
+analysis advocates towards an underpinning neighbourhood mechanism, which 
+is present at different scales. Interestingly, the magnitude of Moran's I
 is similar to the ones revealed by @ding2010modeling about mobile phone
@@
 To illustrate whether a hierarchical process also underpins the
-diffusion of web technologies, the Gini coefficient is calculated yearly
+diffusion of web technologies -- the third diffusion mechanism that is tested 
+here -- the Gini coefficient is calculated yearly
 both for LAD and OA. As a metric of inequality, the Gini coefficient
@@
 
-## Modelling outputs
-
-The next section incorporates the above mechanisms of the spatial
-diffusion of the Web into a modelling framework. The aim is to use
-variables depicting these spatial processes in order to predict the
-diffusion of the Web as a new technology in the UK over space and time
-and across different scales. Specifically, four different models are
-estimated. Firstly, all the data points for the OA and LAD are utilised
-in order to build two RF models and assess their capacity to predict the
-adoption of the Web at the two different scales. These two models will
-reveal the predictive capacity of the diffusion mechanisms and also
-demonstrate how the importance of such variables changes across
-different scales. The next two sets of models are again trained at these
-two scales: OA and LAD. However, instead of using all the data points,
-the OA and the LAD from one of the twelve UK regions are held out for
-the model training. Then, the trained model is used to predict website
-density for the OA or the LAD of the held-out region. This process takes
-place recursively for all twelve UK regions. The regional differences in
-the predictive capacity of the different samples will reveal how
-dissimilar these spatial process are across regions and, importantly, at
-different scales.
-
-It needs to be highlighted here that the cross validation (CV) for all
-models is spatially and temporally sensitive. Instead of using 10 random
-and space- and time-agnostic folds,the `CAST` package is employed which
-allows holding back data points from specific years and spatial units
-and use them for testing in order to estimate the model performance
-[@meyer2018improving].
-
-The models include variables that capture the three processes that the
-relevant literature and the descriptive analysis highlighted. Namely,
-the models capture: (i) a hierarchical mechanism with diffusion running
-from main centres to secondary ones, (ii) a neighborhood mechanism
-according to which diffusion first hits nearby locations, and (iii) the
-rather canonical pattern of diffusion over time as reflected in the
-S-shaped pattern in the cumulative level of adoption.
-
-To capture the hierarchical mechanism the models include as predictors a
-one year lag of website density in London, the largest city in the UK, a
-one year lag of the website density in the nearest city and the same for
-the nearest retail centre. Due to the small sizes of the retail centres,
-the latter is only relevant for and included in the OA-level models. In
-addition, the models include the distance to London, the nearest city
-and the nearest retail centre. The underlying logic is that the level of
-website adoption in a spatial unit depends on the level of the adoption
-in places further up in the urban hierarchy the previous year. The
-inclusion of the distance variables incorporates spatial structure into
-the hierarchy argument. To depict the neighbourhood mechanism, the
-website density of the neighbouring spatial units in the previous year
-is employed. This is calculated using rook continuity. Again, the
-underpinning rationale is that the level of web adoption within a
-spatial unit depends on the level of web adoption in the neighbouring
-spatial units the year before. This represents the 'hitting nearby
-locations first' argument. Therefore, the spatial and temporal lag of
-the website density in LAD and OA is calculated. Lastly, the temporal
-pattern of the cumulative adoption is captured by a time trend variable.
-Hence, all four models will follow the following generic form (Eq.
-\ref{model}):
-
-\begin{align} \label{model}
-Website\,Density_{t} \sim Distance\,London +
-Website\,density\,London_{t-1} +\notag\\
-Distance\,Nearest\,City +
-Website\,density\,Nearest\,City_{t-1} +\notag\\
-Distance\,Nearest\,Retail_{i} +
-Website\,density\,Nearest\,Retail_{t-1} +\notag\\
-W*\, Website\,density_{t-1} +\notag\\ 
-year_{t}
-\end{align}
-
-To assess the predictive power of the model, three broadly utilised
-metrics are employed: the coefficient of determination ($R^2$), mean
-absolute error (MAE) and root mean square error (RMSE):
-
-\begin{align}
-R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
-\end{align}
-\begin{align}
-MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
-\end{align}
-\begin{align}
-RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
-\end{align}
-
-\noindent $y_{k}$ is the $k^{th}$ observation of the dataset, which
-consists of $N$ observations in total. $\hat{y_{k}}$ is the $k_{th}$
-predicted value for the dependent variable and $\overline{y_{k}}$ is the
-average value of $y$. The last two metrics are expressed in the same
-units as the dependent variable -- websites per firm for the LAD modes
-and the number of websites for the OA models -- while the first one is
-the coefficient of determination between the observed and the predicted
-values of website adoption. Regarding $MAE$, it is the absolute
-difference between the observed and the predicted website adoption.
-While $MAE$ does not penalise for large errors, $RSME$ does, as it is
-proportional to the squared difference between the observed and the
-predicted trade flows. Hence larger errors weigh more for $RMSE$
-[@pontius2008components].
+After discussing the descriptive findings, the validity of the above diffusion
+mechanisms at the local scale is assessed with a predictive modelling framework
+described in @sec-datamethods. The reader is reminded that four models are 
+estimated: for the OA and LAD, firstly all data points are employed to 
+assess the capacity of the three mechanisms to to predict the adoption of the 
+Web at these two scales; then, the LAD/OA from one of the twelve regions are
+recursively held out to assess the spatial heterogeneity of these mechanisms
+across the different UK regions.
 
@@
 sensitive CV, so the the model does not suffer from overfitting. At the
-LAD level the model predicts 85% of the variation of website density.
-Both error metrics indicate that the model error is
-$(\frac{1}{20}, \frac{1}{30})$ of a website per firm. At the OA the
-$R^2$ drops down to 32%. Considering its granularity, this is still a
-remarkable performance. To contextualise it, the model results in a
-$MAE$ equal to one website for areas small enough to host less than 140
-households. [^4] Because of the small size of the spatial units, the
+LAD level the model predicts 83% of the variation of website density.
+Both error metrics indicate that the model error is fraction of a website 
+per firm (0.020 and 0.029). At the OA the $R^2$ drops down to 33%. Considering 
+its granularity, this is still a remarkable performance. To contextualise it, 
+the model results in a $MAE$ equal to 1.2 websites for areas small enough to 
+host 40-250 households. Because of the small size of the spatial units, the
 distribution is highly skewed and a significant part of them is not
 linked to any websites. In 1997 only 1% of the UK OA were associated
-with at least one website. This should not come as a surprise as this
+with at least one website. This should not come as a surprise given that this
 was the very beginning of the commercial Internet and any activities
-with a digital footprint were concentrated in a handful of areas. This
-was clearly illustrated in Figure \ref{lisa}. At the end of the study
+with a digital footprint were concentrated in a handful of areas, as 
+illustrated in Figure \ref{lisa}. At the end of the study
 period almost half of the UK OA were not associated with a website.
-Again, given the granularity of the data this should not come as a
-surprise.
-
-[^4]: According to the Office for National Statistics, 80% of OA in
-    England and Wales host 110-139 households,
-    [www.ons.gov.uk](https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas).
+Again, given the granularity of the data this was expected.
 
@@
 |-------------------|--------|---------|-------|
-| Local Authorities | 0.028  | 0.850   | 0.019 |
-| Output Areas      | 3.284  | 0.320   | 1.034 |
+| Local Authorities | 0.029  | 0.824   | 0.020 |
+| Output Areas      | 3.218  | 0.333   | 1.233 |
 
@@
 The results in @tbl-model-metrics are robust against different
-specification. As noted earlier, the data imputation process resulted in
+specifications. As noted earlier, the data imputation process resulted in
 a modest improvement of the predictive capacity of the model. As per
-@tbl-model-metrics-no-correction all accuracy metrics are only slightly
+@tbl-model-metrics-no-correction, all accuracy metrics are only slightly
 worst when the data imputation is not applied. As expected, the
@@
 dataset, which includes websites with up to 10 unique postcodes. While
-the results for the LAD are slightly less accurate, the $R^{2}$ of the
-OA model is more than double in comparison to the base results in
-@tbl-model-metrics. All in all, the modelling exercise revealed how well
-can the three diffusion mechanisms predict the diffusion of the Web as a
-new technology. Importantly, this predictive power is highly robust
+the results for LAD are only slightly less accurate, the $R^{2}$ of the
+OA model is much higher in comparison to the base results in
+@tbl-model-metrics. In addition, @tbl-model-growth-metrics in the Appendix
+presents the accuracy results of a LAD model, the dependent variable of which 
+is the yearly growth rate of website density. Even for such a specification
+(see Equation \ref{model.growth} in the Appendix) the predictive capacity of
+the model is exceptionally high ($R^2 = 0.5$). All in all, the modelling exercise
+revealed how well the three diffusion mechanisms predict the diffusion of the 
+Web as a new technology. Importantly, this predictive power is highly robust
 against against different scales and delineations of the diffusion of
@@
 |-------------------|--------|---------|-------|
-| Local Authorities | 0.032  | 0.810   | 0.019 |
-| Output Areas      | 5.000  | 0.205   | 1.047 |
-
-: Model metrics without data imputation
-{#tbl-model-metrics-no-correction}
+| Local Authorities | 0.032  | 0.781   | 0.021 |
+| Output Areas      | 4.644  | 0.244   | 1.237 |
+
+: Model metrics without data imputation {#tbl-model-metrics-no-correction}
 
@@
 |-------------------|--------|---------|-------|
-| Local Authorities | 0.188  | 0.77    | 0.122 |
-| Output Areas      | 22.467 | 0.478   | 5.681 |
+| Local Authorities |  0.190 | 0.749   | 0.129 |
+| Output Areas      | 21.591 | 0.455   | 6.845 |
 
@@
 Figure \ref{var.imp} plots the importance of the different predictors.
-When the focus is on LAD, the website density in the nearest city, in
-London and in the neighbouring LAD the year before are the most
-important predictors. They are followed by the yearly trend, while the
-spatial configuration as reflected in distances to London or the nearest
-city only plays a minor role. This can be attributed to the relatively
-coarse spatial scale of LAD. Nevertheless, all previously discussed
-spatial processes are at play in the diffusion of web technologies at
-the LAD level: the first two predictors depict the hierarchical
-mechanism, the spatial and temporal lag of website density the
+When the focus is on LAD, the website density in the neighbouring LAD, in the
+nearest city and in London the year before are the most important predictors. 
+They are followed by the yearly trend, while the spatial configuration as 
+reflected in distances to London or to the nearest city only plays a minor 
+role. This can be attributed to the relatively coarse spatial scale of LAD.
+Nevertheless, all previously discussed spatial processes are at play in the
+diffusion of the Web at the LAD level: the first two predictors depict 
+the hierarchical mechanism, the spatial and temporal lag of website density the
 neighbourhood mechanism, and the yearly trend the time-sensitive
@@
 
-When the much more granular scale of OA is adopted, the picture is
+At the much more granular scale of OA, the picture is
 almost reversed. The two most important predictors are the distance to
@@
 the website density is by far the most important predictor when the
-extended dataset is employed. But still, the variable importance
-patterns for the two datasets are not dissimilar.
-
-```{r varimpLA, eval = F, echo=FALSE, message = F, fig.cap="\\label{var.imp.LAD}Variable importance, LAD"}
-path.image <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.png")
-knitr::include_graphics(path.image)
-```
-
-```{r varimpOA, eval = F, echo=FALSE, message = F, fig.cap="\\label{var.imp.OA}Variable importance, OA"}
-path.image <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.png")
-knitr::include_graphics(path.image)
-```
+extended dataset is employed. For the LAD, the importance of the distance to
+the nearest city increases in favour to its website density the previous
+year. But still, the variable importance patterns for the two datasets 
+are not dissimilar.
 
@@
   mutate(order = ifelse(variable == "Distance to london", 1,
-                        ifelse(variable == "London's wensite density, t-1", 2,
+                        ifelse(variable == "London's website density, t-1", 2,
                                ifelse(variable == "Distance to the nearest city", 3,
-                                      ifelse(variable == "Nearest city's wensite density, t-1", 4,
+                                      ifelse(variable == "Nearest city's website density, t-1", 4,
                                              ifelse(variable == "Distance to the nearest retail centre", 5,
-                                                    ifelse(variable == "Nearest retail centre's wensite density, t-1", 6,
-                                                           ifelse(variable == "Spatial and temporal lag of wensite density", 7, 8)))))))) %>% 
+                                                    ifelse(variable == "Nearest retail centre's website density, t-1", 6,
+                                                           ifelse(variable == "Spatial and temporal lag of website density", 7, 8)))))))) %>% 
   ggplot(aes(x=importance, y=reorder(variable, desc(order)), #forcats::fct_reorder(variable, importance), 
@@
 
+Table \ref{table.regions} presents the results of the recursive hold out
+models and allows to (i) highlight the regional heterogeneity of the
+spatial processes behind the diffusion of the Web, and (ii) compare this
+heterogeneity across the two different scales of analysis. It illustrates both 
+regional similarities and differences. At the top of Table \ref{table.regions}
+the regions of South East and the West Midlands have the highest similarity with 
+the rest of the country. In other words, the web density for the LAD of these 
+regions can be very well predicted by models trained on data from the LAD from 
+the rest of the country and, therefore, the spatial diffusion mechanisms for 
+these regions represent quite well the UK's average. This should not come as a 
+surprise as the spatial structure of both regions resembles the broader UK: 
+both regions contain some highly urbanised LAD (South East is well integrated 
+with the Greater London and West Midlands is the home of the city of Birmingham), 
+but also smaller commuter towns as well as more rural settlements. At the other 
+end of the spectrum, the spatial diffusion mechanisms for Northern Ireland, 
+Scotland and North West diverge the most from the country average. Rurality and 
+remoteness seem to be common attributes for these regions. This hierarchy though 
+is not consistent across scales. For the more granular OA models, the spatial 
+diffusion mechanisms for South West, Yorkshire and The Humber and North West 
+regions are closer to the country average. Interestingly, these regions are 
+well-known tourist destinations (Lake District, Peak District and 
+Cornwall-Devon) and the exploratory analysis revealed clusters of high web 
+density only visible at the OA scale (Figure \ref{lisa}). Then, Scotland, 
+Northern Ireland and London have the lowest similarity with the country 
+average. Again, rurality, peripherality, but also the highest level or 
+urbanisation make the spatial diffusion mechanisms of these regions more 
+distinct.
+
 ```{r regions}
@@
 
-Table \ref{table.regions} presents the results of the recursive hold out
-models and allows to (i) highlight the regional heterogeneity of the
-spatial processes behind the diffusion of the Web, and (ii) compare this
-heterogeneity across the two different scales of analysis. It
-illustrates some striking similarities, but also a few significant
-differences. At the top of the similarity with the rest of the country
-we see West Midlands and Yorkshire and the Humber for both scales. The
-web density for the LAD and OA of these regions can be very well
-predicted by models trained on data for the LAD and OA of the rest of
-the country. In other words, the spatial diffusion mechanisms for these
-regions represent quite well the UK's average. This should not come as
-as surprise given their urban structure: they both contain a large urban
-centre (Birmingham and Leeds), smaller commuter towns as well as more
-rural settlements. The bottom of the similarity ranking is populated by
-Scotland, Northern Ireland, East of England and London for both scales.
-The diffusion mechanisms for these regions diverge significantly from
-the country's average and this could be attributed to rurality and
-remoteness or to high levels of urbanisation and sophistication of
-economic activities for the case of London and for East of England,
-which contains the Cambridge area.
-
-Still, there are three regions the spatial diffusion mechanisms of which
-differ a lot between the two scales. North West and South West are very
-close to the UK's average at the more granular OA scale, while the
-opposite applies for South East. Interestingly, the first two regions
-are well-known tourist destinations (Lake District and Cornwall-Devon)
-and the exploratory analysis revealed clusters of high web density only
-visible at the OA scale (Figure \ref{lisa}. Then, the South East is very
-well integrated with the Greater London, hosts the broader Oxford area
--- a known hub of research excellence -- and Figure \ref{lisa} revealed
-high web density clusters at the LAD level.
+Comparing the above with Table \ref{table.regions.10} in the Appendix, which 
+presents the results for the the extended dataset with websites with up to 10 
+unique postcodes, the results do not differentiate much, signalling the 
+robustness of the analysis. The most salient difference has to do with Northern 
+Ireland. According to the extended sample at the granular OA level, Northern 
+Ireland's spatial diffusion mechanisms are closer to the UK's average than the 
+results based on the websites with a unique postcode. While the latter are 
+websites and, consequently, economic activities solely based in Northern 
+Ireland, websites with multiple postcodes also include websites with postcodes 
+in Britain and, therefore, can be characterised by slightly different spatial 
+diffusion processes.
+
+## Early adoption and latecomers
+
+<!-- H~2~: The diffusion of the commercial Web is shaped by both early adoption  -->
+<!-- and latecomer effects. -->
+
+To test the second hypothesis (H~2~), the spatial heterogeneity of Web adoption 
+speeds is firstly analysed. As highlighted in the literature [@risk_perceptions], 
+different agents have different perceptions about and levels of acceptance of 
+the risks and the potential economic returns associated with the adoption of 
+new technologies -- see for instance the seminal work of @venkatesh2000theoretical.
+To reveal such patterns, Figure \ref{s_map} maps whether LAD reached their 
+inflection point earlier or later than the UK in total and based on this are 
+labelled as *fast* or *slow* accordingly. Inflection points have been used in
+the relevant literature to compare diffusion processes across different settings 
+and technologies [@cherp2021national; @woo2017exploring] as they can enable more 
+insightful comparisons than, for example, the average growth over the 1996-2012
+study period.
+
+The map illustrates a clear high concentration of fast adopting LAD in the 
+South and East of England, as expected. There are also some concentrations in 
+Wales, Scotland and the North West of England. At a more detailed level, there 
+are some expected cases of LAD with relevant industrial backgrounds delineated 
+as fast: the City of London, a world-renowned cluster of finance industries
+[@cook2007role], and Reading, a town with high-tech service industries
+in proximity to London and its main airport, Heathrow
+[@england2005polynet]. However, there are also LAD which were expected
+to appear as fast -- e.g. Hackney in central London and Bristol, a
+well-established creative cluster [@oatley1999cultural;
+@bassett2002cultural] -- but are delineated as slow. Similar to 
+@beardsell1999spatial, who analysed the spatial evolution of computer industry 
+employment across 317 US metro areas, spatial heterogeneity is evident in the 
+diffusion speed when the scale of analysis is as detailed as the LAD.
+
+Importantly, there is obvious spatial clustering of fast and slow LAD. Out of 
+the ten fastest LAD based on when they reached their inflection point (see 
+detailed Table \ref{table.s.lads} in the Appendix), nine were located in the 
+South East of England and London and one in Cambridge. The above observations 
+remain almost unchanged when the analysis includes websites with up to 10 unique
+postcodes -- see Figure \ref{s_map10} in the Appendix -- demonstrating the
+robustness of the analysis. Given (i) the expectation that 
+early adopters of new technologies are often rewarded [@rogers2010diffusion; 
+@griliches1957; @ding2010modeling], and (ii) the previous findings 
+illustrating the long term and sustained productivity benefit for the early 
+adopters of Web in the UK [@tranosuk], the observed spatial clustering of LAD 
+with fast diffusion rates can further reinforce existing and rooted patterns of 
+unequal economic productivity in the UK [@mccann2022productivity].
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map}LAD Web diffusion speed, 1996-2012"}
+path.image <- paste0(path, "/outputs/s/speed_map.png")
+knitr::include_graphics(path.image)
+```
+
+To reveal patterns of early adoption and leapfrogging, the stability and 
+volatility of the above patterns is analysed. Figure \ref{rank} plots the 
+relative ranking of LAD based on website density at the beginning and the end 
+of the study period. To decrease noise, the average ranking of 1996-1998 and 
+2010-2012 is plotted instead of the individual years. Each line represents a 
+single LAD and the colours depict the 
+5\textsuperscript{th} and 95\textsuperscript{th} percentile of the absolute
+difference of the LAD ranking between 1996-1998 and 2010-2012. In other words, 
+the 5\textsuperscript{th} percentile includes the LAD, the relative
+position of which remained unchanged between the beginning and the end of 
+the study period, both at the top and at the bottom of the hierarchy (small
+absolute difference in ranking). The 95th percentile includes LAD that either
+increased or decreased a lot their relative positions over time (high absolute 
+difference in ranking). Thus, Figure \ref{rank} reveals cases of 
+*leapfrogging* since the LAD at the top-right part of the figure managed to 
+jump at the top of the hierarchy at the end of the study period. 
+<!-- which are often associated with latecomers benefits in the  -->
+<!-- literature. -->
+
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank}Dynamics of Web diffusion based on LAD rankings"}
+path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av.png")
+knitr::include_graphics(path.image)
+```
+
+To further analyse the above, Table \ref{table.late} juxtaposes the 
+volatility of the LAD Web diffusion against the early conditions as it reports 
+the $t_0$ estimate (inflection point) for the leapfrogging LAD identified in
+Figure \ref{rank}. Indeed, more than half of the LAD that drastically increased 
+their relative position at the end of the study period reached the inflection 
+point $t_0$ later than the country in total and, therefore, the Web diffusion 
+speed has been relatively slow for them. The fact that these LAD experienced
+relatively faster diffusion rates at the end of the study period, which is 
+aligned with the maturity period of the web, can lead to potential positive 
+economic gains for these LAD. There is extensive literature describing the 
+benefits of such technological leapfrogging and the underpinning argument is 
+that latecomers can adopt and benefit from new technologies that have been 
+developed elsewhere without incurring the hefty initial R&D costs 
+[@teece2008firm].
+
+Based on Figure \ref{rank} and Table \ref{table.late}, the opportunity 
+to catch up that @perkins2005international illustrated at the country level, is 
+also visible at the much more detailed LAD scale. Although the relevant 
+literature did not pay attention to cities and regions [@yu2018sustainability], 
+the above finding advocates towards further exploring leapfrogging effects at 
+such a granular scale. As previously, this is consistent with including websites 
+with up to 10 postcodes (see Figure \ref{rank10}).
+
+All in all, the diffusion of the Web at the detailed scale of LAD is 
+characterised by spatial heterogeneity, well-established spatial patterns and 
+both stability and volatility. Although the relevant literature neglected local
+scales, the above findings illustrated that such local dynamics exists and 
+worth further consideration and investigation especially regarding potential 
+local economic effects.
+
+```{r late/early}
+late.path <- paste0(path, "/outputs/ranks/late.csv")
+
+read_csv(late.path) %>%
+  dplyr::select(-'early/late') %>% 
+  rename('$t_{0}$ estimate' = estimate,
+         "Diffusion speed" = fast) %>% 
+  kable(caption= "Late adopters\\label{table.late}", 
+        digits=3,
+        #col.names = c("LAD", "$t_{0}$ estimate", "Diffusion speed", "Early/late"),
+        align = c("l", "c", "c"),
+        escape = F)  
+  #kableExtra::row_spec(10, hline_after = TRUE)
+```
 
@@
 
-This paper assessed the importance of spatial mechanisms in the
-diffusion of a new technology -- the Web -- at a high level of spatial
-granularity. Specifically, it assessed the capacity of the following
-three spatial diffusion mechanisms in predicting website density in the
-UK over space and time: (i) the hierarchical mechanism according to
-which diffusion runs from main urban centres to secondary ones, (ii) the
-neighborhood mechanism which highlights that diffusion first hits nearby
-locations, and (iii) the rather canonical pattern of diffusion over time
-as reflected in the S-shaped pattern of the cumulative level of
-adoption. This is the first time that the importance of granular spatial
-mechanisms for the diffusion of an intangible, digital technology is
-exposed. The time period of the study is also important as it captures
-the very early stages of this technology until its maturity.
+The paper offers an economic geography study of the diffusion of a new 
+technology -- the Web. Contrary to previous studies, it does so by adopting 
+(i) a high level of spatial granularity, and (ii) a long study period that
+captures the very early stages of this technology until its maturity.
+This is the first time that the importance of granular spatial
+mechanisms for the diffusion of a new technology is exposed. 
 Importantly, the empirical framework and the data employed here allows
@@
 
-The analysis firstly illustrated how well the S-pattern of cumulative
-adoption fits at the local scale of LAD. It highlighted the
-heterogeneity of adoption rates between LAD -- with both more and less
-expected examples -- as well as the spatial clustering of LAD with
-similar adoption rates. The exploratory analysis also illustrated
-instances of remarkable stability as well as volatility of the web
-diffusion system that is indicative of leapfrogging. Previous research
-had already indicated the long-term productivity gains of the early
-adoption of such technologies [@tranosuk]. Then, the paper developed a
-robust and novel ML modelling framework to predict website diffusion
-over space and time using the three spatial diffusion mechanisms based
-on space- and time-sensitive CV without suffering from overfitting. The
-model revealed the remarkably good predictive capacity of these
-mechanisms at both scales. For the first time, the power of these
-mechanisms in predicting the diffusion of a new technology was
+<!-- H1 -->
+Specifically, two hypotheses are tested. Firstly, whether the three diffusion 
+mechanisms namely distance, urban hierarchy and the S-shaped pattern of the 
+cumulative level of adoption shaped the diffusion of the commercial Web in 
+the UK at granular geographical scales. The analysis firstly illustrates how 
+well the S-pattern of cumulative adoption fits at the local scale and offers 
+some descriptive insights about the importance of the distance and 
+hierarchical diffusion mechanisms. Then, the modelling exercise reveals the 
+remarkably good predictive capacity of these mechanisms at granular geographical 
+scales. The modelling framework exposes that the relative importance of 
+these spatial mechanisms differ across different scales. It also reveals 
+spatial heterogeneity. While for some UK regions the importance of the spatial 
+diffusion mechanisms are very similar to the rest of the country, other regions
+diverge significantly from the country average.
+
+To do the above, a robust and novel ML modelling framework was designed to 
+predict website diffusion over space and time using as predictors the spatial 
+diffusion mechanisms. Space- and time-sensitive CV was designed for the model
+to avoid overfitting. Importantly, the results are robust against different 
+specifications and subsets of the data. This is the first time that the power 
+of these mechanisms in predicting the diffusion of a new technology was
 empirically tested, importantly, at very fine scales.
 
-The modelling framework also revealed that the importance of these
-spatial mechanisms differ across different scales. While spatial
-structure as reflected in distances to different urban centres is highly
-important at the OA scale, it hardly plays any role at the more coarse
-LAD scale. And while the almost canonical pattern of cumulative adoption
-is a key predictor at the LAD level, it only plays a small role at the
-detailed OA level. Then, the analysis revealed another level of spatial
-heterogeneity. While for some UK regions these spatial diffusion
-mechanisms are very similar to the rest of the country, other regions
-diverge significantly from the country average. Importantly, the above
-results are robust against different specifications and subsets of the
-data.
+<!-- H2 -->
+<!-- H~2~: The diffusion of the commercial Web is shaped by both early adoption  -->
+<!-- and latecomer effects. -->
+The second hypothesis assessed whether the diffusion of the web is shaped by 
+both early adoption and latecomer effects. On the one hand, the analysis 
+highlighted the heterogeneity of adoption rates between LAD as well as the 
+spatial clustering of LAD with similar adoption rates. Importantly, the 
+observed spatial clustering of LAD with fast diffusion rates can
+reinforce existing and rooted patterns of unequal economic productivity 
+in the UK [@mccann2022productivity]. Previous research had already indicated 
+the long-term productivity gains of the early adoption of such technologies 
+[@tranosuk]. On the other hand, the analysis also exposed instances of 
+remarkable stability as well as volatility of the Web adoption, with the 
+latter being indicative of leapfrogging. While the literature had 
+discussed the benefits of technological leapfrogging at the national scale 
+[@teece2008firm; @perkins2005international], local scales had escaped such 
+studies [@yu2018sustainability]. Thus the leapfrogging evidence presented here 
+advocate towards further exploring such effects at granular scales.
 
@@
 @rogers2010diffusion; @grubler1990rise], their role had never been
-tested before at such granular scale and, importantly, for a digital
-technology as weightless as the Web. The findings talk directly to this
+tested before at such granular scale. The findings talk directly to this
 literature [e.g. @beardsell1999spatial; @haller2011determinants;
@@
 various stakeholders are interested in predicting localised demand for
-new technologies for planning purposes [@leibowicz2016representing;
-@meade2021modelling], and (ii) the adoption of such digital technologies
-is associated with economic gains [@solow1957technical;
-@aghion1990model; @kemeny2011international; @tranosuk;
-@capello2024nexus]. In addition, while most related studies focused on
-tangible technologies, this paper sheds light on the diffusion of a
-truly weightless digital technology that is the Web. Thus, it adds much
-more granular findings to the limited related studies
-[@sinai2004geography; @haller2011determinants]. Importantly, although
-the geographical focus of this paper is the UK, the empirical framework
-developed here is easily transferable to other geographical contexts.
-
-<!-- Future research can further investigate other -->
-
-<!-- determinants of the diffusion of the Web. For instance, the archived web data -->
-
-<!-- can reveal important information about the industrial base of different places -->
-
-<!-- as well as the industries which were faster in adopting the Web. -->
+new technologies for planning purposes, for example 5G networks
+[@leibowicz2016representing;@meade2021modelling], and (ii) the adoption 
+of such digital technologies is associated with local economic gains 
+[@solow1957technical; @aghion1990model; @kemeny2011international; @tranosuk;
+@capello2024nexus]. 
+<!-- In addition, while most related studies focused on -->
+<!-- tangible technologies, this paper sheds light on the diffusion of a -->
+<!-- truly weightless digital technology that is the Web. Thus, it adds much -->
+<!-- more granular findings to the limited related studies -->
+<!-- [@sinai2004geography; @haller2011determinants].  -->
+Importantly, although the geographical focus of this paper is the UK, the 
+empirical framework developed here is easily transferable to other geographical 
+contexts.
 
@@
 
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk10}Grwoth curve, UK; up to 10 postcodes per website"}
-path.image <- paste0(path, "/outputs/s/s_uk_per_firm.png")
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk10}Cumulative adoption of the Web, UK; up to 10 postcodes per website"}
+path.image <- paste0(path, "/outputs/s/s_uk_per_firm_10.png")
 knitr::include_graphics(path.image)
@@
 
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map10}LAD adoption rates; up to 10 postcodes per website"}
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map10}LAD Web diffusion speed, 1996-2012; up to 10 postcodes per website"}
 path.image <- paste0(path, "/outputs/s/speed_map_10.png")
@@
 
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank10}Dynamics of wed diffusion (LAD); up to 10 postcodes per website"}
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank10}Dynamics of Web diffusion based on LAD rankings; up to 10 postcodes per website"}
 path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av_10.png")
@@
 
-```{r regions_no_imp}
-path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
-path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
-
-lad <- read.csv(path.lad) %>% 
-  rename('RSquared LAD' = Rsquared)
-
-oa <- read.csv(path.oa) %>% rename(Region = test.region,
-                                   'RSquared OA' = Rsquared)
-
-lad %>% left_join(oa) %>% 
-  mutate('Rank OA' = rank(desc(`RSquared OA`)),
-         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
-  relocate(5, .after = 2) %>% 
-  arrange(`Rank LAD`) %>% 
-  kable(caption= "Regional differences; up to 10 postcodes per website\\label{table.regions.no.imp}", 
-        digits=3,
-        col.names = c("Region", "$R^2$ LAD", "Rank LAD", "$R^2$ OA", "Rank OA")) 
-```
-
 ```{r, echo=FALSE, message = F, error=FALSE}
@@
          "Diffusion speed" = "fast") %>% 
-  kable(caption= "S-curve estiamtes for LAD; up to 10 postcodes per website\\label{table.s.lads}", 
+  kable(caption= "S-curve estiamtes for LAD\\label{table.s.lads}", 
         digits=3,
@@
 
+```{r regions_no_imp}
+path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected_10.csv")
+path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA_corrected_10.csv")
+
+lad <- read.csv(path.lad) %>% 
+  rename('RSquared LAD' = Rsquared)
+
+oa <- read.csv(path.oa) %>% rename(Region = test.region,
+                                   'RSquared OA' = Rsquared)
+
+lad %>% left_join(oa) %>% 
+  mutate('Rank OA' = rank(desc(`RSquared OA`)),
+         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
+  relocate(5, .after = 2) %>% 
+  arrange(`Rank LAD`) %>% 
+  kable(caption= "Regional differences; up to 10 postcodes per website\\label{table.regions.10}", 
+        digits=3,
+        col.names = c("Region", "$R^2$ LAD", "Rank LAD", "$R^2$ OA", "Rank OA")) 
+```
+
+|                   | $RMSE$ | $R^{2}$ | $MAE$ |
+|-------------------|--------|---------|-------|
+| Local Authorities | 0.286  | 0.506   | 0.223 |
+
+: Model metrics for growth model {#tbl-model-growth-metrics}
+
+@tbl-model-growth-metrics presents the results of a Random Forest model that 
+was trained based on Equation \ref{model.growth} using all LAD data points. In
+essence, this is a model similar to the one estimated based on Equation
+\ref{eq:model} and presented in @tbl-model-metrics, with the main difference 
+being the dependent variable which is the year growth rate of website 
+density for each LAD. To avoid cases with LAD have no websites, 
+$t \in [2000, 2012]$. The high predictive capability of the model reflects 
+the robustness of the modelling approach.
+
+\begin{align} \label{model.growth}
+Website\,Density\,Growth_{t} \sim Distance\,London +
+Website\,density\,London_{t-1} +\notag\\
+Distance\,Nearest\,City +
+Website\,density\,Nearest\,City_{t-1} +\notag\\
+Distance\,Nearest\,Retail_{i} +
+Website\,density\,Nearest\,Retail_{t-1} +\notag\\
+W*\, Website\,density_{t-1} +\notag\\ 
+year_{t}
+\end{align}
+
 # References {#references .unnumbered}

### 2025-03-12 paper/submission2/anonymised.tex
paper v2

--- 2025-01-03 paper/submission2/tranos2025.tex
@@
   pdftitle={A multi-scale story of the diffusion of a new technology: the web},
-  pdfauthor={Emmanouil Tranos},
   pdfkeywords={diffusion, technologies, Web, digital},
@@
 \begin{abstract}
-Despite the broad consensus that differences in technological adoption
-and sophistication can lead to uneven trajectories of local development
-as new technologies can increase productivity and allow firms to
-introduce new products and services, economic geographers have shied
-away from exploring the diffusion of digital technologies across space
-and time. This paper fills this gap by exposing the roles of different
-spatial mechanisms in the diffusion of an intangible, digital technology
-at very granular geographical scales. It does so by employing a novel
-source of data and machine learning algorithms to model the active
-engagement with Web during 1996-2012.
+This paper investigates the spatial diffusion of a new technology that
+is the Web in the UK. It employs novel data and machine learning methods
+to model the influence of well-established diffusion mechanisms.
+Contrary to previous studies, it adopts multiple scales, high spatial
+granularity and a long study period that captures the early stages of
+the Web until its maturity (1996-2012). Findings reveal the importance
+of such spatial mechanisms (namely distance, urban hierarchy and the
+S-shaped pattern of the cumulative level of adoption) even at granular
+scales. They also highlight spatial heterogeneity and instances of
+leapfrogging.
 \end{abstract}
@@
 geographical scales to provide local policy insights, for instance,
-regarding infrastructural d evelopment and local technology policies. A
+regarding infrastructural development and local technology policies. A
 potential explanation of the lack of geographical studies exploring the
@@
 This paper offers such a contribution: an economic geography study
-revealing the importance of spatial mechanisms for the diffusion of an
-intangible, digital technology that is the Web in the UK at a high level
-of spatial granularity from the very early days until its maturity
-(1996-2012). It specifically focuses on the diffusion of commercial
-websites. It does so by employing a novel source of big data and machine
-learning algorithms to model the \emph{active} engagement with Web --
-that is creating and maintaining instead of just browsing a website --
-during that period and tests how different spatial and temporal
-mechanisms shaped its diffusion. This paper also exemplifies how the
-combination of data sources and state-of-the-art research methods which
-escape the traditional social science domain can offer new lenses to
-geographical research regarding the understanding of technological
-diffusion.
+revealing the importance of spatial mechanisms for the diffusion of a
+digital technology that is the Web in the UK at a high level of spatial
+granularity from the very early days until its maturity (1996-2012). It
+specifically focuses on the diffusion of commercial websites. It does so
+by employing a novel source of big data and machine learning algorithms
+to model the \emph{active} engagement with Web -- that is creating and
+maintaining instead of just browsing a website -- during that period and
+tests how different spatial and temporal mechanisms shaped its
+diffusion. This paper also exemplifies how the combination of data
+sources and state-of-the-art research methods which escape the
+traditional social science domain can offer new lenses to geographical
+research regarding the understanding of technological diffusion.
 
@@
 policy relevance, there is a scarcity of studies analysing the diffusion
-of new digital technologies at a detailed geographical level because of
-the lack of detailed enough data and adequate research methods.
+of new technologies at a detailed geographical level because of the lack
+of detailed enough data and adequate research methods.
 
@@
 \citep[e.g.][]{crampton2013beyond}. Exceptions include early work from
-\citet{zook2000web}, \citet{zook2000web} and \citet{moss1997tracking},
+\citet{zook2000web}, \citet{zook2001old} and \citet{moss1997tracking},
 which utilised domain name registration data and use it to map and
@@
 Internet, they were based on cross-sectional, often spatially detailed,
-mapping and on shallow time comparisons (e.g.~1993-1998) and did not
+mapping and on short temporal comparisons (e.g.~1993-1998) and did not
 observe the evolution and diffusion of the Web until its maturity. As we
@@
 differences \citep{leibowicz2016representing}. Hence, mapping and
-modelling the diffusion of the Web among firms can offer useful
-insights.
+modelling the diffusion of the Web can offer useful insights.
 
@@
 the Local Authority Districts (LAD), illustrating, for the first time,
-such S-shaped diffusion patterns for local areas. This system level
-analysis also reveals insights about the volatility of the adoption of
-the Web. Then, exploratory spatial data analysis offers some first
-indications of the importance of the different spatial diffusion
-mechanisms in shaping the diffusion of the Web in the UK. Importantly,
-the analysis adopts two distinct geographical scales: the LAD and the
-most detailed scale of UK census areas, the Output Areas (OA). Lastly, a
-novel methodological framework based on Machine Learning (ML) algorithms
-is developed to test the role of such mechanisms at these two scales.
-The results reveal, for the first time, the importance of different
-spatial diffusion mechanisms in the diffusion of an intangible, digital
-technology at very granular geographical scales.
+such S-shaped diffusion patterns for local areas. The exploratory
+spatial data analysis offers some first indications of the importance of
+the different spatial diffusion mechanisms in shaping the diffusion of
+the Web in the UK. Importantly, the analysis adopts two distinct
+geographical scales: the LAD and the most detailed scale of UK census
+areas, the Output Areas (OA). Then, a novel methodological framework
+based on Machine Learning (ML) algorithms is developed to test the role
+of such mechanisms at these two scales. The results reveal, for the
+first time, the importance of different spatial diffusion mechanisms in
+the diffusion of a digital technology at very granular geographical
+scales. The analysis also reveals that the diffusion of the Web is
+characterised by spatial heterogeneity, well-established spatial
+patterns and both stability and volatility.
 
 The structure of the paper goes as follows. Section~\ref{sec-litreview}
-reviews the key concepts in the diffusion literature as well as relevant
-empirical studies. Then, Section~\ref{sec-datamethods} describes the
-data and methods used in this paper. The results of the different stages
-of the analysis are presented in Section~\ref{sec-results}. The
-discussion of the results and the conclusions of the paper are offered
-in Section~\ref{sec-conclusions}.
+reviews the mechanisms that shape the diffusion of a new technology and
+sets up the two hypotheses that are empirically tested in
+Section~\ref{sec-results}. Section~\ref{sec-datamethods} describes the
+data and methods used in this paper. The discussion of the results and
+the conclusions of the paper are offered in
+Section~\ref{sec-conclusions}.
 
@@
 periphery. Although adoption pace might be higher when new technologies
-finally arrive to the periphery, the saturation levels there may never
-reach the ones in the core because of the lack of infrastructure or
-other necessary institutions \citep{leibowicz2016representing}.
+finally arrive to the periphery, the saturation levels may never reach
+the ones in the core because of the lack of infrastructure or other
+necessary institutions \citep{leibowicz2016representing}.
 
@@
 of the timing, speed and the level of mobile diffusion within China.
-\citet{bakher2013diffusion} compared different the growth curves as a
-means to model the diffusion of online shopping in Australia during the
+\citet{bakher2013diffusion} compared different growth curves as a means
+to model the diffusion of online shopping in Australia during the
 1998-2009 period, but they did not consider any spatial mechanisms. At a
@@
 assortativity, urban scaling and distance are the key drivers of spatial
-diffusion. Based on the above, the main empirical hypothesis states:
+diffusion. Based on the above, the main empirical hypothesis that is
+tested in this paper states:
 
@@
 H\textsubscript{1}: The three diffusion mechanisms namely distance,
-urban hierarchy and the the S-shaped pattern of the cummulative level of
+urban hierarchy and the S-shaped pattern of the cumulative level of
 adoption shaped the diffusion of the commercial Web in the UK even at
@@
 \begin{quote}
-H\textsubscript{2}: The diffusion of the commecrial Web is shaped by
-both early adoption and late comer effects.
+H\textsubscript{2}: The diffusion of the commercial Web is shaped by
+both early adoption and latecomer effects.
 \end{quote}
 
-These are the two hypotheses are empirically tested in
-Section~\ref{sec-results} using the data and methods described in the
-next section.
+These two hypotheses are empirically tested in Section~\ref{sec-results}
+employing the data and methods described in the next section.
 
@@
 later in this section. The main data are counts of commercial websites
-that are calculated using data from the Internet Archive\footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
+that are calculated using data from the Internet Archive\footnote{\url{https://archive.org/}.}
 and, specifically, the JISC UK Web Domain Dataset \citep{ukwebarchive}.
@@
 pre-search engines early times of the commercial Internet as well as
-real estate websites \citep{tranosuk}.
+real estate websites \citep{tranosuk}. Previous research which used the
+same methodology for a small subset of these data illustrates the
+robustness of the above approach: its results matched in terms of
+accuracy past rounds of research based on extensive qualitative work,
+interviews and observational studies for the same study area
+\citep{shoreditch}.
 
@@
 are geolocated by equally attaching them to these locations. This
-extended sample includes 94 percent of all the archived websites in
-2000.
+extended sample includes 94 percent of all the archived websites in 2000
+(see Table \ref{f2000}).
 
@@
 \toprule
-Postcodes & F & F (\%) & Cummulative F\\
+Postcodes & F & F (\%) & cumulative F\\
 \midrule
@@
 The website counts for these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1
-6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1000
+6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1,000
 websites pointing to them, were replaced with predicted values based on
@@
 the data imputation into perspective, this process affected 6 out of the
-557,808 postcodes included in the data. As the Section~\ref{sec-results}
+557,808 postcodes included in the data. As Section~\ref{sec-results}
 illustrates, this led to a small increase of the model predictive power.
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-2-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{anonymised_files/figure-pdf/unnamed-chunk-2-1.pdf}
 
@@
 
-the Local Authority Districts (LAD) and the Output Areas (OA). The
-former is an administrative unit and there are 374 LAD in the UK. The
-latter is the smallest census-based geographical unit and there are
-c.~230,000 of them in the UK. This methodological choice will allow the
-mapping of the diffusion of web technologies and the assessment of the
-diffusion mechanisms at these two very different spatial scales.
-
-To create the website density metric, the yearly website counts at the
-LAD level are standardised by the number of firms in LAD to avoid biases
+To empirically test the two hypotheses stated in
+Section~\ref{sec-litreview}, the website density metric was calculated
+for two different scales the Local Authority Districts (LAD) and the
+Output Areas (OA). The former is the lowest administrative unit in the
+UK and there are 374 LAD. The latter is the smallest census-based
+geographical unit and there are c.~230,000 of them in the UK. To create
+the website density metric, the yearly website counts at the LAD level
+are standardised by the number of firms in LAD to avoid biases
 associated with LAD hosting a large number of firms. Given that there is
@@
 
-These website density metrics are used for the three different stages of
-the analysis. Firstly, a system-level analysis explores how the
-diffusion of web technologies in the UK fits with the well-established
-S-curve of cumulative adoption. To do so, the following logistic
-function (Equation \ref{eq:s}) is estimated for the whole of the UK and
-for each LAD separately:
+These website density metrics are used for the different stages of the
+analysis. Starting with the first hypothesis (H\textsubscript{1}), to
+test whether the S-curve accurately describes the cumulative adoption of
+the Web at a local scale, the following logistic function (Equation
+\ref{eq:s}) is estimated for the whole of the UK and then for each LAD
+separately:
 
@@
 against the \(t_0\) of the UK to delineate whether a LAD reached that
-point faster or slower than the country average. Importantly, an
-accuracy criterion was imposed and only LAD with \(R^2 > 0.9\) were
-included in this analysis. To estimate Equation \ref{eq:s} a
-self-starting logistic growth model was employed using the \texttt{nls}
-and \texttt{SSlogis} functions in \texttt{R}. This logistic growth model
-depicts the initial period of slow diffusion of a new technology, which
-is followed by the fast, exponential growth stage to then eventually
-slow down and saturate \citep{wilson201281, grubler1999dynamics}.
-
-The system-level analysis also focuses on the volatility of the web
-adoption to depict places with high concentration of early adopters and,
-equally, latecomers. To do so, the change of the ranking of the UK LAD
-over time is plotted and discussed. Both the S-curve and the volatility
-analysis focus only on LAD as the very large number of OA would have
-made such analysis difficult to visualise and interpret.
-
-Next, exploratory spatial data analysis depicts whether the two main
-drivers of spatial diffusion -- namely neighbourhood and hierarchy --
-underpin the diffusion of the Web in the UK. To capture the former and
-following \citet{ding2010modeling} the Moran's I and the Local Index of
-Spatial Association (LISA) are estimated for the website density. To
-address the hierarchy mechanism, the Gini coefficient -- an established
-metric of inequality -- is calculated. All the above are computed and
-plotted longitudinally both for the LAD and the OA.
-
-Lastly, a modelling framework is developed to test the above diffusion
-mechanisms. The overarching aim is to build a model that can test how
-well these mechanisms can predict website density:
+point faster or slower than the country average. This heterogeneity is
+discussed further when testing the second hypothesis
+(H\textsubscript{2}). Importantly, an accuracy criterion was imposed and
+only LAD with \(R^2 > 0.9\) were included in this analysis. To estimate
+Equation \ref{eq:s} a self-starting logistic growth model was employed
+using the \texttt{nls} and \texttt{SSlogis} functions in \texttt{R}.
+This logistic growth model depicts the initial period of slow diffusion
+of a new technology, which is followed by the fast, exponential growth
+stage to then eventually slow down and saturate
+\citep{wilson201281, grubler1999dynamics}.
+
+Next, exploratory spatial data analysis provides some first insights of
+whether the other two drivers of spatial diffusion -- namely
+neighbourhood and hierarchy -- underpin the diffusion of the Web in the
+UK. To capture the former and following \citet{ding2010modeling} the
+Moran's I and the Local Index of Spatial Association (LISA) are
+estimated for the website density. To capture the hierarchy mechanism,
+the Gini coefficient -- an established metric of inequality -- is
+calculated. All the above are computed and plotted longitudinally both
+for the LAD and the OA.
+
+To formally test H\textsubscript{1} a modelling framework is developed
+to assess the above diffusion mechanisms. The overarching aim is to
+build a model that can test how well these three mechanisms can predict
+website density:
 
 \begin{align}
-Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S-curve_{t}\label{eq:rf.generic}
+Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S\text-curve_{t}\label{eq:rf.generic}
 \end{align}
 
-\noindent Equation \ref{eq:rf.generic} is estimated using Random Forest
-(RF). This is a popular ML algorithm for both regression and
-classification problems \citep{biau2012analysis}. It was introduced by
-\citet{breiman2001random} and has become a go-to data science tool. RF
-can effectively handle skewed distributions and outliers, model
-non-linear relationships, require minimal hyperparameter tuning, exhibit
-low sensitivity to these parameters, and have relatively short training
-times \citep{Caruana2008, liaw2002classification, yan2020using}. These
+To capture the hierarchical mechanism the model include as predictors a
+one year lag of website density in London, the largest city in the UK,
+and a one year lag of the website density in the nearest city. Both
+variables clearly illustrate the urban hierarchy in the UK. Especially
+for the OA models, a third predictor is added to capture hierarchy at a
+much more granular scale: a one year lag of the website density in the
+the nearest retail centre \footnote{The retail centres data have been
+  obtained by the Consumer Data Research Centre, an ESRC data
+  investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1:
+  \url{https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators}.}.
+As the literature illustrates, retail centres are key components of
+urban systems \citep{dennis2002central, jones2021reframing}. Due to the
+small sizes of the retail centres, the latter is only relevant for and
+included in the OA models and not in the LAD models. In addition, the
+model includes the distance to London, the nearest city and the nearest
+retail centre. The underlying logic is that the level of website
+adoption in a spatial unit depends on the level of the adoption in
+places further up in the urban hierarchy the previous year. The
+inclusion of the distance variables incorporates spatial structure into
+the hierarchy argument. To depict the neighbourhood mechanism, the
+website density of the neighbouring spatial units in the previous year
+is employed. This is calculated using rook continuity. Again, the
+underpinning rationale is that the level of web adoption within a
+spatial unit depends on the level of web adoption in the neighbouring
+spatial units the year before. This represents the `hitting nearby
+locations first' argument. Therefore, the spatial and temporal lag of
+the website density in LAD and OA is calculated. Lastly, the temporal
+pattern of the cumulative adoption is captured by a time trend variable.
+Hence, the model will follow the following generic form (Eq.
+\ref{eq:model}):
+
+\begin{align} \label{eq:model}
+Website\,Density_{t} \sim Distance\,London +
+Website\,density\,London_{t-1} +\notag\\
+Distance\,Nearest\,City +
+Website\,density\,Nearest\,City_{t-1} +\notag\\
+Distance\,Nearest\,Retail_{i} +
+Website\,density\,Nearest\,Retail_{t-1} +\notag\\
+W*\, Website\,density_{t-1} +\notag\\ 
+year_{t}
+\end{align}
+
+Specifically, four different models are estimated. Firstly, all the data
+points for the OA and LAD are utilised in order to build two models and
+assess their capacity to predict the adoption of the Web at the two
+different scales. These two models will reveal the predictive capacity
+of the diffusion mechanisms and also demonstrate how the importance of
+such variables changes across different scales. The next two sets of
+models are again trained at these two scales: OA and LAD. However,
+instead of using all the data points, the OA and the LAD from one of the
+twelve UK regions are held out for the model training. Then, the trained
+model is used to predict website density for the OA or the LAD of the
+held-out region. This process takes place recursively for all twelve UK
+regions. The regional differences in the predictive capacity of the
+different samples will reveal how dissimilar these spatial process are
+across regions and, importantly, at different scales.
+
+Equation \ref{eq:model} is estimated using Random Forest (RF). This is a
+popular ML algorithm for both regression and classification problems
+\citep{biau2012analysis}. It was introduced by \citet{breiman2001random}
+and has become a go-to data science tool. RF can effectively handle
+skewed distributions and outliers, model non-linear relationships,
+require minimal hyperparameter tuning, exhibit low sensitivity to these
+parameters, and have relatively short training times
+\citep{Caruana2008, liaw2002classification, yan2020using}. These
 attributes match well with the website density data characteristics
@@
 
+It needs to be highlighted here that the cross validation (CV) for all
+models is spatially and temporally sensitive. Instead of using 10 random
+and space- and time-agnostic folds, the \texttt{CAST} package is
+employed which allows holding back data points from specific years and
+spatial units (regions) and use them for testing in order to estimate
+the model performance \citep{meyer2018improving}.
+
+To assess the predictive power of the model, three broadly utilised
+metrics are employed: the coefficient of determination (\(R^2\)), mean
+absolute error (MAE) and root mean square error (RMSE):
+
+\begin{align}
+R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
+\end{align} \begin{align}
+MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
+\end{align} \begin{align}
+RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
+\end{align}
+
+\noindent \(y_{k}\) is the \(k^{th}\) observation of the dataset, which
+consists of \(N\) observations in total. \(\hat{y_{k}}\) is the
+\(k_{th}\) predicted value for the dependent variable and
+\(\overline{y_{k}}\) is the average value of \(y\). The last two metrics
+are expressed in the same units as the dependent variable -- websites
+per firm for the LAD modes and the number of websites for the OA models
+-- while the first one is the coefficient of determination between the
+observed and the predicted values of website adoption. Regarding
+\(MAE\), it is the absolute difference between the observed and the
+predicted website adoption. While \(MAE\) does not penalise for large
+errors, \(RSME\) does, as it is proportional to the squared difference
+between the observed and the predicted trade flows. Hence larger errors
+weigh more for \(RMSE\) \citep{pontius2008components}.
+
+An alternative specification could model the growth of website density.
+However, such a strategy is problematic because of the large frequency
+of OA without websites, which is to be expected given their small size
+and high granularity. Still, as a robustness text, the results of a
+growth model for LAD are presented in the Appendix.
+
+Lastly, to test the second hypothesis (H\textsubscript{2}) about the
+early adoption and the latecomer effects, a different research framework
+is developed. Firstly, the S-curve and the estimated \(t_0\) for every
+LAD are employed and mapped in order to depict the spatial heterogeneity
+and specific patterns of the speed of the diffusion of the Web. Then,
+the volatility and the stability of the diffusion across the different
+LAD is analysed using tools from the complex systems domain
+\citep{Batty2009} namely changes in rankings of LAD based on website
+density. Both the S-curve estimation and the volatility analysis focus
+only on LAD as the very large number of OA would have made such analysis
+difficult to visualise and interpret.
+
 \section{Results}\label{sec-results}
 
-\subsection{System-level analysis}\label{system-level-analysis}
+\subsection{The three diffusion mechanisms at granular geographical
+scales}\label{the-three-diffusion-mechanisms-at-granular-geographical-scales}
+
+Figure \ref{s_uk} plots the cumulative adoption of the Web in the UK as
+a whole during the 1996-2012 period. In line with the theoretical
+expectations discussed in the previous section, the cumulative adoption
+of the Web is well represented by a curve that resembles the letter S.
+The vertical line in the beginning of 2003 (2003.158) illustrates the
+point when the modelled cumulative adoption was equal to 50\% of the
+maximum. This \(t_0\) point is known as the \emph{inflection} point and
+signals the maximum adoption speed. The results remain unchanged when
+the extended subset of websites, which include up to 10 unique postcodes
+is used (see Figure \ref{s_uk10} in the Appendix). Importantly though,
+the extended subset of websites indicates a saturation level close to
+100\% at the end of the study period.
 
@@
 
-\caption{\label{s_uk}Grwoth curve, UK}
+\caption{\label{s_uk}Cumulative adoption of the Web, UK}
 
@@
 
-Figure \ref{s_uk} plots the cumulative adoption of the Web in total in
-the UK during the 1996-2012 period. Not surprisingly, the cumulative
-adoption is well represented by a curve that resembles the figure S. The
-vertical line around year 2003 illustrates the point where the modelled
-cumulative adoption is equal to 50\% of the maximum. This \(t_0\)
-\emph{inflection} point signals the maximum adoption speed and is used
-here to determine whether a UK LAD reached that inflection point earlier
-or later than the UK average. Specifically, the S curve and the
-inflection point are estimated for every LAD individually and then
-compared against the UK average. LAD that reached their inflection point
-earlier than the country average are labelled as \emph{fast} and the
-rest as \emph{slow}. Figure \ref{s_map} maps this pattern. There is a
-clear high concentration of fast adopting LAD in the South and East of
-England as expected. However, there is also such a concentration in
-Wales, Scotland and the North West of England. At a more detailed level,
-there are some expected examples of LAD with relevant industrial
-backgrounds delineated as fast: the City of London, a world-renowned
-cluster of finance industries \citep{cook2007role}, and Reading, a town
-with high-tech service industries in proximity to London and its main
-airport, Heathrow \citep{england2005polynet}. However, there are also
-LAD which were expected to appear as fast -- e.g.~Hackney in central
-London and Bristol, a well-established creative cluster
-\citep{oatley1999cultural, bassett2002cultural} -- but were delineated
-as slow. Nevertheless, there is obvious spatial clustering of fast and
-slow LAD. Importantly, out of the ten fastest LAD, nine were located in
-the South East of England and London and one in Cambridge (see Table
-\ref{table.s.lads} in the Appendix). The above observations remain
-almost unchanged when the analysis includes websites with up to 10
-unique postcodes -- see Figures \ref{s_uk10} and \ref{s_map10} in the
-Appendix -- advocating towards the robustness of the analysis.
+What escaped from previous studies was εexploring whether such a pattern
+also describes cumulative adoption of new technologies at sub-national
+scales. To test this, the same logistic function is estimated for each
+of the 374 LAD. The detailed results are presented in Table
+\ref{table.s.lads} in the Appendix. The key finding is that for 342 LAD
+(91\% of all LAD) the \(R^2 > 0.9\) for the logistic function, which
+indicates that even at local scale the cumulative adoption of the Web is
+characterised by a pattern similar to the one observed for countries.
+Graphically, all these LAD curves are plotted in Figure \ref{s_lad_uk}.
+The different colour lines illustrate whether a LAD reached \(t_0\)
+faster or slower than the UK in total. This heterogeneity is further
+discussed later in this section.
+
+This is the first time that such cumulative adoption curves have been
+estimated for areal units as small as the UK LAD. Understanding and
+modelling the dynamics of technological diffusion and, ultimately, being
+able to predict diffusion rates at such small local scales can directly
+support the deployment of infrastructural networks (e.g.~5G) and local
+technology policies. The above are in line with previous studies
+exploring the diffusion of web technologies
+\citep{PAPAGIANNIDIS2015308}, social media \citep{lengyel2020role} and
+online shopping \citep{bakher2013diffusion}.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/speed_map.png}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/lad_per_firm_all.png}
 
@@
 
-\caption{\label{s_map}LAD adoption rates}
+\caption{\label{s_lad_uk}Cumulative adoption of the Web, LAD}
 
@@
 
-The S-curves of the cumulative adoption of Web in the UK in total and in
-the UK LAD demonstrate a pattern well aligned with all previous studies
-discussed in Section~\ref{sec-litreview}. Importantly, the above
-findings are in line with previous studies exploring the diffusion of
-web technologies \citep{PAPAGIANNIDIS2015308}, social media
-\citep{lengyel2020role} and online shopping \citep{bakher2013diffusion}.
-Similar to \citet{beardsell1999spatial}, who analysed the spatial
-evolution of computer industry employment across 317 US metro areas,
-spatial heterogeneity is evident in the diffusion speed when the scale
-of analysis is as detailed as the LAD. Importantly, no previous studies
-estimated such cumulative adoption curves for areal units as small as
-the UK LADS.
-
-The next step assesses the stability and volatility of the LAD in terms
-of their adoption of web technologies. As highlighted in the literature
-\citep{risk_perceptions}, different agents have different perceptions
-about and levels of acceptance of the risks and the potential economic
-returns associated with the adoption of new technologies -- see for
-instance the seminal work of \citet{venkatesh2000theoretical}. To reveal
-such aggregated patterns, Figure \ref{rank} plots the ranking of UK LAD
-based on website density. To decrease noise, the average ranking of
-1996-1998 and 2010-2012 is plotted instead of the individual years. The
-colours depict the 5th and 95th percentile of the absolute difference of
-the LAD ranking between 1996-1998 and 2010-2012. While there are quite a
-few obvious cases of LAD that maintained their position between the
-beginning and the end of the study period both at the top or at the
-bottom of the hierarchy, there are also quite a few LAD that changed
-drastically their position. Some of these LAD enjoyed a process that at
-the first instance looks like \emph{leapfrogging} since they managed to
-jump at the top of the hierarchy despite their slow start. There is
-extensive literature regarding the potential benefits of technological
-leapfrogging. The underpinning argument is that latecomers can adopt and
-benefit from new technologies that have been developed elsewhere without
-incurring the hefty initial R\&D costs \citep{teece2008firm}. Although
-the leapfrogging literature does not pay much attention to cities and
-regions \citep{yu2018sustainability}, previous research highlighted the
-long term and sustained productivity benefit of the early adopters of
-web technologies in the UK \citep{tranosuk}. So, although the LAD system
-is volatile, it is not clear whether the LAD with high concentration of
-late adopters will gain any latecomer-related benefits in a way similar
-to countries experiencing technological leapfrogging. Still, the
-opportunity to catch up that \citet{perkins2005international}
-illustrated at the country level, is also visible at the much more
-detailed LAD scale. As previously, this is consistent with including
-websites with up to 10 postcodes (see Figure \ref{rank10}).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/ranks/web_per_firm2000_2012_only0595_av.png}
-
-}
-
-\caption{\label{rank}Dynamics of wed diffusion (LAD)}
-
-\end{figure}%
-
-\subsection{Exploratory analysis}\label{exploratory-analysis}
-
-Figures \ref{morani} and \ref{lisa} offer a first insight into whether a
-neighbourhood mechanism underpins the diffusion of the Web in the UK as
-they present the Moran's I and the LISA maps of website density
-respectively for LAD and OA. Starting from the former, spatial
+To assess the second diffusion mechanisms, Figures \ref{morani} and
+\ref{lisa} offer a first insight into whether a neighbourhood or, in
+other words, a distance mechanism underpins the diffusion of the Web in
+the UK as they present the Moran's I and the LISA maps of website
+density respectively for LAD and OA. Starting from the former, spatial
 autocorrelation was higher in the beginning of the study period and then
@@
 spatial clustering mechanism. All in all, the exploratory spatial data
-analysis advocates towards an underpinning neighbourhood mechanism and
-the different scales of analysis illustrate how it evolved differently
-over time. Interestingly, the magnitude of Moran's I is similar to the
-ones revealed by \citet{ding2010modeling} about mobile phone adoption in
-China albeit the much more detailed spatial scale of this paper.
+analysis advocates towards an underpinning neighbourhood mechanism,
+which is present at different scales. Interestingly, the magnitude of
+Moran's I is similar to the ones revealed by \citet{ding2010modeling}
+about mobile phone adoption in China albeit the much more detailed
+spatial scale of this paper.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2025_files/figure-pdf/morani-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/morani-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-6-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-5-1.pdf}
 
@@
 To illustrate whether a hierarchical process also underpins the
-diffusion of web technologies, the Gini coefficient is calculated yearly
-both for LAD and OA. As a metric of inequality, the Gini coefficient
-demonstrates whether website density is concentrated in a small number
-of LAD or OA, or whether it is more equally spread across the country.
-Both scales of analysis in Figure \ref{gini} illustrate the same
-picture. At the beginning of the commercial Internet, website density
-was extremely unequal, or, in other words, only a few places had
-websites associated with them. Inequality dropped and plateaued after
-2000 for both scales. This is illustrative of a hierarchical diffusion
-mechanism that led over time to a more equal spread. Interestingly, the
-year 2000 is again a period of change for this diffusion mechanism as it
-was for the neighbourhood process. There is a substantial difference
-between the Gini coefficient magnitude for LAD and OA, but this is
-expected as the very small size of OA equates to a lot of polygons
-without any websites pointing to them -- for example residential OA.
+diffusion of web technologies -- the third diffusion mechanism that is
+tested here -- the Gini coefficient is calculated yearly both for LAD
+and OA. As a metric of inequality, the Gini coefficient demonstrates
+whether website density is concentrated in a small number of LAD or OA,
+or whether it is more equally spread across the country. Both scales of
+analysis in Figure \ref{gini} illustrate the same picture. At the
+beginning of the commercial Internet, website density was extremely
+unequal, or, in other words, only a few places had websites associated
+with them. Inequality dropped and plateaued after 2000 for both scales.
+This is illustrative of a hierarchical diffusion mechanism that led over
+time to a more equal spread. Interestingly, the year 2000 is again a
+period of change for this diffusion mechanism as it was for the
+neighbourhood process. There is a substantial difference between the
+Gini coefficient magnitude for LAD and OA, but this is expected as the
+very small size of OA equates to a lot of polygons without any websites
+pointing to them -- for example residential OA.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2025_files/figure-pdf/gini-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/gini-1.pdf}
 
@@
 
-\subsection{Modelling outputs}\label{modelling-outputs}
-
-The next section incorporates the above mechanisms of the spatial
-diffusion of the Web into a modelling framework. The aim is to use
-variables depicting these spatial processes in order to predict the
-diffusion of the Web as a new technology in the UK over space and time
-and across different scales. Specifically, four different models are
-estimated. Firstly, all the data points for the OA and LAD are utilised
-in order to build two RF models and assess their capacity to predict the
-adoption of the Web at the two different scales. These two models will
-reveal the predictive capacity of the diffusion mechanisms and also
-demonstrate how the importance of such variables changes across
-different scales. The next two sets of models are again trained at these
-two scales: OA and LAD. However, instead of using all the data points,
-the OA and the LAD from one of the twelve UK regions are held out for
-the model training. Then, the trained model is used to predict website
-density for the OA or the LAD of the held-out region. This process takes
-place recursively for all twelve UK regions. The regional differences in
-the predictive capacity of the different samples will reveal how
-dissimilar these spatial process are across regions and, importantly, at
-different scales.
-
-It needs to be highlighted here that the cross validation (CV) for all
-models is spatially and temporally sensitive. Instead of using 10 random
-and space- and time-agnostic folds,the \texttt{CAST} package is employed
-which allows holding back data points from specific years and spatial
-units and use them for testing in order to estimate the model
-performance \citep{meyer2018improving}.
-
-The models include variables that capture the three processes that the
-relevant literature and the descriptive analysis highlighted. Namely,
-the models capture: (i) a hierarchical mechanism with diffusion running
-from main centres to secondary ones, (ii) a neighborhood mechanism
-according to which diffusion first hits nearby locations, and (iii) the
-rather canonical pattern of diffusion over time as reflected in the
-S-shaped pattern in the cumulative level of adoption.
-
-To capture the hierarchical mechanism the models include as predictors a
-one year lag of website density in London, the largest city in the UK, a
-one year lag of the website density in the nearest city and the same for
-the nearest retail centre. Due to the small sizes of the retail centres,
-the latter is only relevant for and included in the OA-level models. In
-addition, the models include the distance to London, the nearest city
-and the nearest retail centre. The underlying logic is that the level of
-website adoption in a spatial unit depends on the level of the adoption
-in places further up in the urban hierarchy the previous year. The
-inclusion of the distance variables incorporates spatial structure into
-the hierarchy argument. To depict the neighbourhood mechanism, the
-website density of the neighbouring spatial units in the previous year
-is employed. This is calculated using rook continuity. Again, the
-underpinning rationale is that the level of web adoption within a
-spatial unit depends on the level of web adoption in the neighbouring
-spatial units the year before. This represents the `hitting nearby
-locations first' argument. Therefore, the spatial and temporal lag of
-the website density in LAD and OA is calculated. Lastly, the temporal
-pattern of the cumulative adoption is captured by a time trend variable.
-Hence, all four models will follow the following generic form (Eq.
-\ref{model}):
-
-\begin{align} \label{model}
-Website\,Density_{t} \sim Distance\,London +
-Website\,density\,London_{t-1} +\notag\\
-Distance\,Nearest\,City +
-Website\,density\,Nearest\,City_{t-1} +\notag\\
-Distance\,Nearest\,Retail_{i} +
-Website\,density\,Nearest\,Retail_{t-1} +\notag\\
-W*\, Website\,density_{t-1} +\notag\\ 
-year_{t}
-\end{align}
-
-To assess the predictive power of the model, three broadly utilised
-metrics are employed: the coefficient of determination (\(R^2\)), mean
-absolute error (MAE) and root mean square error (RMSE):
-
-\begin{align}
-R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
-\end{align} \begin{align}
-MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
-\end{align} \begin{align}
-RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
-\end{align}
-
-\noindent \(y_{k}\) is the \(k^{th}\) observation of the dataset, which
-consists of \(N\) observations in total. \(\hat{y_{k}}\) is the
-\(k_{th}\) predicted value for the dependent variable and
-\(\overline{y_{k}}\) is the average value of \(y\). The last two metrics
-are expressed in the same units as the dependent variable -- websites
-per firm for the LAD modes and the number of websites for the OA models
--- while the first one is the coefficient of determination between the
-observed and the predicted values of website adoption. Regarding
-\(MAE\), it is the absolute difference between the observed and the
-predicted website adoption. While \(MAE\) does not penalise for large
-errors, \(RSME\) does, as it is proportional to the squared difference
-between the observed and the predicted trade flows. Hence larger errors
-weigh more for \(RMSE\) \citep{pontius2008components}.
+After discussing the descriptive findings, the validity of the above
+diffusion mechanisms at the local scale is assessed with a predictive
+modelling framework described in Section~\ref{sec-datamethods}. The
+reader is reminded that four models are estimated: for the OA and LAD,
+firstly all data points are employed to assess the capacity of the three
+mechanisms to to predict the adoption of the Web at these two scales;
+then, the LAD/OA from one of the twelve regions are recursively held out
+to assess the spatial heterogeneity of these mechanisms across the
+different UK regions.
 
@@
 space and time sensitive CV, so the the model does not suffer from
-overfitting. At the LAD level the model predicts 85\% of the variation
+overfitting. At the LAD level the model predicts 83\% of the variation
 of website density. Both error metrics indicate that the model error is
-\((\frac{1}{20}, \frac{1}{30})\) of a website per firm. At the OA the
-\(R^2\) drops down to 32\%. Considering its granularity, this is still a
+fraction of a website per firm (0.020 and 0.029). At the OA the \(R^2\)
+drops down to 33\%. Considering its granularity, this is still a
 remarkable performance. To contextualise it, the model results in a
-\(MAE\) equal to one website for areas small enough to host less than
-140 households. \footnote{According to the Office for National
-  Statistics, 80\% of OA in England and Wales host 110-139 households,
-  \href{https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas}{www.ons.gov.uk}.}
-Because of the small size of the spatial units, the distribution is
-highly skewed and a significant part of them is not linked to any
-websites. In 1997 only 1\% of the UK OA were associated with at least
-one website. This should not come as a surprise as this was the very
-beginning of the commercial Internet and any activities with a digital
-footprint were concentrated in a handful of areas. This was clearly
-illustrated in Figure \ref{lisa}. At the end of the study period almost
-half of the UK OA were not associated with a website. Again, given the
-granularity of the data this should not come as a surprise.
+\(MAE\) equal to 1.2 websites for areas small enough to host 40-250
+households. Because of the small size of the spatial units, the
+distribution is highly skewed and a significant part of them is not
+linked to any websites. In 1997 only 1\% of the UK OA were associated
+with at least one website. This should not come as a surprise given that
+this was the very beginning of the commercial Internet and any
+activities with a digital footprint were concentrated in a handful of
+areas, as illustrated in Figure \ref{lisa}. At the end of the study
+period almost half of the UK OA were not associated with a website.
+Again, given the granularity of the data this was expected.
 
@@
 \endlastfoot
-Local Authorities & 0.028 & 0.850 & 0.019 \\
-Output Areas & 3.284 & 0.320 & 1.034 \\
+Local Authorities & 0.029 & 0.824 & 0.020 \\
+Output Areas & 3.218 & 0.333 & 1.233 \\
 \end{longtable}
@@
 The results in Table~\ref{tbl-model-metrics} are robust against
-different specification. As noted earlier, the data imputation process
+different specifications. As noted earlier, the data imputation process
 resulted in a modest improvement of the predictive capacity of the
-model. As per Table~\ref{tbl-model-metrics-no-correction} all accuracy
+model. As per Table~\ref{tbl-model-metrics-no-correction}, all accuracy
 metrics are only slightly worst when the data imputation is not applied.
@@
 results for the extended dataset, which includes websites with up to 10
-unique postcodes. While the results for the LAD are slightly less
-accurate, the \(R^{2}\) of the OA model is more than double in
-comparison to the base results in Table~\ref{tbl-model-metrics}. All in
-all, the modelling exercise revealed how well can the three diffusion
-mechanisms predict the diffusion of the Web as a new technology.
-Importantly, this predictive power is highly robust against against
-different scales and delineations of the diffusion of the web.
+unique postcodes. While the results for LAD are only slightly less
+accurate, the \(R^{2}\) of the OA model is much higher in comparison to
+the base results in Table~\ref{tbl-model-metrics}. In addition,
+Table~\ref{tbl-model-growth-metrics} in the Appendix presents the
+accuracy results of a LAD model, the dependent variable of which is the
+yearly growth rate of website density. Even for such a specification
+(see Equation \ref{model.growth} in the Appendix) the predictive
+capacity of the model is exceptionally high (\(R^2 = 0.5\)). All in all,
+the modelling exercise revealed how well the three diffusion mechanisms
+predict the diffusion of the Web as a new technology. Importantly, this
+predictive power is highly robust against against different scales and
+delineations of the diffusion of the web.
 
 \begin{longtable}[]{@{}llll@{}}
-\caption{Model metrics without data imputation
-}\label{tbl-model-metrics-no-correction}\tabularnewline
+\caption{Model metrics without data
+imputation}\label{tbl-model-metrics-no-correction}\tabularnewline
 \toprule\noalign{}
@@
 \endlastfoot
-Local Authorities & 0.032 & 0.810 & 0.019 \\
-Output Areas & 5.000 & 0.205 & 1.047 \\
+Local Authorities & 0.032 & 0.781 & 0.021 \\
+Output Areas & 4.644 & 0.244 & 1.237 \\
 \end{longtable}
@@
 \endlastfoot
-Local Authorities & 0.188 & 0.77 & 0.122 \\
-Output Areas & 22.467 & 0.478 & 5.681 \\
+Local Authorities & 0.190 & 0.749 & 0.129 \\
+Output Areas & 21.591 & 0.455 & 6.845 \\
 \end{longtable}
@@
 Figure \ref{var.imp} plots the importance of the different predictors.
-When the focus is on LAD, the website density in the nearest city, in
-London and in the neighbouring LAD the year before are the most
-important predictors. They are followed by the yearly trend, while the
-spatial configuration as reflected in distances to London or the nearest
-city only plays a minor role. This can be attributed to the relatively
-coarse spatial scale of LAD. Nevertheless, all previously discussed
-spatial processes are at play in the diffusion of web technologies at
-the LAD level: the first two predictors depict the hierarchical
-mechanism, the spatial and temporal lag of website density the
-neighbourhood mechanism, and the yearly trend the time-sensitive
-cumulative adoption pattern.
-
-When the much more granular scale of OA is adopted, the picture is
-almost reversed. The two most important predictors are the distance to
-London and to the nearest city followed by the spatial and temporal lag
-of website density and the distance to the nearest retail centre. They
-still depict a hierarchical mechanism, but proximity to the different
-population centres is more important than their lagged web densities in
-predicting website diffusion. The neighbourhood mechanism is still
-strong but less important at this scale. What is interesting is the
-almost negligible role of the yearly trend and London's website density.
-While the former probably illustrates the large heterogeneity in how the
-Web has been adopted at this very fine scale, the latter highlights that
-the importance of past web adoption rates in large population centres is
+When the focus is on LAD, the website density in the neighbouring LAD,
+in the nearest city and in London the year before are the most important
+predictors. They are followed by the yearly trend, while the spatial
+configuration as reflected in distances to London or to the nearest city
+only plays a minor role. This can be attributed to the relatively coarse
+spatial scale of LAD. Nevertheless, all previously discussed spatial
+processes are at play in the diffusion of the Web at the LAD level: the
+first two predictors depict the hierarchical mechanism, the spatial and
+temporal lag of website density the neighbourhood mechanism, and the
+yearly trend the time-sensitive cumulative adoption pattern.
+
+At the much more granular scale of OA, the picture is almost reversed.
+The two most important predictors are the distance to London and to the
+nearest city followed by the spatial and temporal lag of website density
+and the distance to the nearest retail centre. They still depict a
+hierarchical mechanism, but proximity to the different population
+centres is more important than their lagged web densities in predicting
+website diffusion. The neighbourhood mechanism is still strong but less
+important at this scale. What is interesting is the almost negligible
+role of the yearly trend and London's website density. While the former
+probably illustrates the large heterogeneity in how the Web has been
+adopted at this very fine scale, the latter highlights that the
+importance of past web adoption rates in large population centres is
 surpassed by proximity to them and spatial configuration at this scale.
@@
 the website density is by far the most important predictor when the
-extended dataset is employed. But still, the variable importance
-patterns for the two datasets are not dissimilar.
+extended dataset is employed. For the LAD, the importance of the
+distance to the nearest city increases in favour to its website density
+the previous year. But still, the variable importance patterns for the
+two datasets are not dissimilar.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{tranos2025_files/figure-pdf/varimp-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{anonymised_files/figure-pdf/varimp-1.pdf}
 
@@
 \end{figure}%
+
+Table \ref{table.regions} presents the results of the recursive hold out
+models and allows to (i) highlight the regional heterogeneity of the
+spatial processes behind the diffusion of the Web, and (ii) compare this
+heterogeneity across the two different scales of analysis. It
+illustrates both regional similarities and differences. At the top of
+Table \ref{table.regions} the regions of South East and the West
+Midlands have the highest similarity with the rest of the country. In
+other words, the web density for the LAD of these regions can be very
+well predicted by models trained on data from the LAD from the rest of
+the country and, therefore, the spatial diffusion mechanisms for these
+regions represent quite well the UK's average. This should not come as a
+surprise as the spatial structure of both regions resembles the broader
+UK: both regions contain some highly urbanised LAD (South East is well
+integrated with the Greater London and West Midlands is the home of the
+city of Birmingham), but also smaller commuter towns as well as more
+rural settlements. At the other end of the spectrum, the spatial
+diffusion mechanisms for Northern Ireland, Scotland and North West
+diverge the most from the country average. Rurality and remoteness seem
+to be common attributes for these regions. This hierarchy though is not
+consistent across scales. For the more granular OA models, the spatial
+diffusion mechanisms for South West, Yorkshire and The Humber and North
+West regions are closer to the country average. Interestingly, these
+regions are well-known tourist destinations (Lake District, Peak
+District and Cornwall-Devon) and the exploratory analysis revealed
+clusters of high web density only visible at the OA scale (Figure
+\ref{lisa}). Then, Scotland, Northern Ireland and London have the lowest
+similarity with the country average. Again, rurality, peripherality, but
+also the highest level or urbanisation make the spatial diffusion
+mechanisms of these regions more distinct.
 
@@
 \endlastfoot
-South East & 0.945 & 1 & 0.116 & 8 \\
-West Midlands & 0.922 & 2 & 0.125 & 4 \\
-Wales & 0.915 & 3 & 0.125 & 5 \\
-Yorkshire and The Humber & 0.909 & 4 & 0.160 & 2 \\
-East Midlands & 0.903 & 5 & 0.120 & 7 \\
-North East & 0.896 & 6 & 0.124 & 6 \\
-South West & 0.894 & 7 & 0.184 & 1 \\
-London & 0.881 & 8 & 0.093 & 9 \\
-East of England & 0.874 & 9 & 0.092 & 10 \\
-North West & 0.837 & 10 & 0.143 & 3 \\
-Scotland & 0.780 & 11 & 0.055 & 12 \\
-Nortern Ireland & 0.579 & 12 & 0.060 & 11 \\
+South East & 0.952 & 1 & 0.122 & 6 \\
+West Midlands & 0.926 & 2 & 0.126 & 4 \\
+Wales & 0.909 & 3 & 0.122 & 8 \\
+Yorkshire and The Humber & 0.908 & 4 & 0.160 & 2 \\
+East Midlands & 0.904 & 5 & 0.122 & 7 \\
+South West & 0.894 & 6 & 0.185 & 1 \\
+North East & 0.887 & 7 & 0.124 & 5 \\
+London & 0.880 & 8 & 0.090 & 11 \\
+East of England & 0.876 & 9 & 0.106 & 9 \\
+North West & 0.837 & 10 & 0.142 & 3 \\
+Scotland & 0.774 & 11 & 0.059 & 12 \\
+Nortern Ireland & 0.590 & 12 & 0.101 & 10 \\
 \end{longtable}
 
-Table \ref{table.regions} presents the results of the recursive hold out
-models and allows to (i) highlight the regional heterogeneity of the
-spatial processes behind the diffusion of the Web, and (ii) compare this
-heterogeneity across the two different scales of analysis. It
-illustrates some striking similarities, but also a few significant
-differences. At the top of the similarity with the rest of the country
-we see West Midlands and Yorkshire and the Humber for both scales. The
-web density for the LAD and OA of these regions can be very well
-predicted by models trained on data for the LAD and OA of the rest of
-the country. In other words, the spatial diffusion mechanisms for these
-regions represent quite well the UK's average. This should not come as
-as surprise given their urban structure: they both contain a large urban
-centre (Birmingham and Leeds), smaller commuter towns as well as more
-rural settlements. The bottom of the similarity ranking is populated by
-Scotland, Northern Ireland, East of England and London for both scales.
-The diffusion mechanisms for these regions diverge significantly from
-the country's average and this could be attributed to rurality and
-remoteness or to high levels of urbanisation and sophistication of
-economic activities for the case of London and for East of England,
-which contains the Cambridge area.
-
-Still, there are three regions the spatial diffusion mechanisms of which
-differ a lot between the two scales. North West and South West are very
-close to the UK's average at the more granular OA scale, while the
-opposite applies for South East. Interestingly, the first two regions
-are well-known tourist destinations (Lake District and Cornwall-Devon)
-and the exploratory analysis revealed clusters of high web density only
-visible at the OA scale (Figure \ref{lisa}. Then, the South East is very
-well integrated with the Greater London, hosts the broader Oxford area
--- a known hub of research excellence -- and Figure \ref{lisa} revealed
-high web density clusters at the LAD level.
+Comparing the above with Table \ref{table.regions.10} in the Appendix,
+which presents the results for the the extended dataset with websites
+with up to 10 unique postcodes, the results do not differentiate much,
+signalling the robustness of the analysis. The most salient difference
+has to do with Northern Ireland. According to the extended sample at the
+granular OA level, Northern Ireland's spatial diffusion mechanisms are
+closer to the UK's average than the results based on the websites with a
+unique postcode. While the latter are websites and, consequently,
+economic activities solely based in Northern Ireland, websites with
+multiple postcodes also include websites with postcodes in Britain and,
+therefore, can be characterised by slightly different spatial diffusion
+processes.
+
+\subsection{Early adoption and
+latecomers}\label{early-adoption-and-latecomers}
+
+To test the second hypothesis (H\textsubscript{2}), the spatial
+heterogeneity of Web adoption speeds is firstly analysed. As highlighted
+in the literature \citep{risk_perceptions}, different agents have
+different perceptions about and levels of acceptance of the risks and
+the potential economic returns associated with the adoption of new
+technologies -- see for instance the seminal work of
+\citet{venkatesh2000theoretical}. To reveal such patterns, Figure
+\ref{s_map} maps whether LAD reached their inflection point earlier or
+later than the UK in total and based on this are labelled as \emph{fast}
+or \emph{slow} accordingly. Inflection points have been used in the
+relevant literature to compare diffusion processes across different
+settings and technologies \citep{cherp2021national, woo2017exploring} as
+they can enable more insightful comparisons than, for example, the
+average growth over the 1996-2012 study period.
+
+The map illustrates a clear high concentration of fast adopting LAD in
+the South and East of England, as expected. There are also some
+concentrations in Wales, Scotland and the North West of England. At a
+more detailed level, there are some expected cases of LAD with relevant
+industrial backgrounds delineated as fast: the City of London, a
+world-renowned cluster of finance industries \citep{cook2007role}, and
+Reading, a town with high-tech service industries in proximity to London
+and its main airport, Heathrow \citep{england2005polynet}. However,
+there are also LAD which were expected to appear as fast -- e.g.~Hackney
+in central London and Bristol, a well-established creative cluster
+\citep{oatley1999cultural, bassett2002cultural} -- but are delineated as
+slow. Similar to \citet{beardsell1999spatial}, who analysed the spatial
+evolution of computer industry employment across 317 US metro areas,
+spatial heterogeneity is evident in the diffusion speed when the scale
+of analysis is as detailed as the LAD.
+
+Importantly, there is obvious spatial clustering of fast and slow LAD.
+Out of the ten fastest LAD based on when they reached their inflection
+point (see detailed Table \ref{table.s.lads} in the Appendix), nine were
+located in the South East of England and London and one in Cambridge.
+The above observations remain almost unchanged when the analysis
+includes websites with up to 10 unique postcodes -- see Figure
+\ref{s_map10} in the Appendix -- demonstrating the robustness of the
+analysis. Given (i) the expectation that early adopters of new
+technologies are often rewarded
+\citep{rogers2010diffusion, griliches1957, ding2010modeling}, and (ii)
+the previous findings illustrating the long term and sustained
+productivity benefit for the early adopters of Web in the UK
+\citep{tranosuk}, the observed spatial clustering of LAD with fast
+diffusion rates can further reinforce existing and rooted patterns of
+unequal economic productivity in the UK \citep{mccann2022productivity}.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/speed_map.png}
+
+}
+
+\caption{\label{s_map}LAD Web diffusion speed, 1996-2012}
+
+\end{figure}%
+
+To reveal patterns of early adoption and leapfrogging, the stability and
+volatility of the above patterns is analysed. Figure \ref{rank} plots
+the relative ranking of LAD based on website density at the beginning
+and the end of the study period. To decrease noise, the average ranking
+of 1996-1998 and 2010-2012 is plotted instead of the individual years.
+Each line represents a single LAD and the colours depict the
+5\textsuperscript{th} and 95\textsuperscript{th} percentile of the
+absolute difference of the LAD ranking between 1996-1998 and 2010-2012.
+In other words, the 5\textsuperscript{th} percentile includes the LAD,
+the relative position of which remained unchanged between the beginning
+and the end of the study period, both at the top and at the bottom of
+the hierarchy (small absolute difference in ranking). The 95th
+percentile includes LAD that either increased or decreased a lot their
+relative positions over time (high absolute difference in ranking).
+Thus, Figure \ref{rank} reveals cases of \emph{leapfrogging} since the
+LAD at the top-right part of the figure managed to jump at the top of
+the hierarchy at the end of the study period.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/ranks/web_per_firm2000_2012_only0595_av.png}
+
+}
+
+\caption{\label{rank}Dynamics of Web diffusion based on LAD rankings}
+
+\end{figure}%
+
+To further analyse the above, Table \ref{table.late} juxtaposes the
+volatility of the LAD Web diffusion against the early conditions as it
+reports the \(t_0\) estimate (inflection point) for the leapfrogging LAD
+identified in Figure \ref{rank}. Indeed, more than half of the LAD that
+drastically increased their relative position at the end of the study
+period reached the inflection point \(t_0\) later than the country in
+total and, therefore, the Web diffusion speed has been relatively slow
+for them. The fact that these LAD experienced relatively faster
+diffusion rates at the end of the study period, which is aligned with
+the maturity period of the web, can lead to potential positive economic
+gains for these LAD. There is extensive literature describing the
+benefits of such technological leapfrogging and the underpinning
+argument is that latecomers can adopt and benefit from new technologies
+that have been developed elsewhere without incurring the hefty initial
+R\&D costs \citep{teece2008firm}.
+
+Based on Figure \ref{rank} and Table \ref{table.late}, the opportunity
+to catch up that \citet{perkins2005international} illustrated at the
+country level, is also visible at the much more detailed LAD scale.
+Although the relevant literature did not pay attention to cities and
+regions \citep{yu2018sustainability}, the above finding advocates
+towards further exploring leapfrogging effects at such a granular scale.
+As previously, this is consistent with including websites with up to 10
+postcodes (see Figure \ref{rank10}).
+
+All in all, the diffusion of the Web at the detailed scale of LAD is
+characterised by spatial heterogeneity, well-established spatial
+patterns and both stability and volatility. Although the relevant
+literature neglected local scales, the above findings illustrated that
+such local dynamics exists and worth further consideration and
+investigation especially regarding potential local economic effects.
+
+\begin{longtable}[]{@{}lcc@{}}
+\caption{Late adopters\label{table.late}}\tabularnewline
+\toprule\noalign{}
+LAD & \(t_{0}\) estimate & Diffusion speed \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+LAD & \(t_{0}\) estimate & Diffusion speed \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Halton & 2007.1 & slow \\
+Blackpool & 2002.4 & fast \\
+Rutland & 2006.5 & slow \\
+Torbay & 2002.2 & fast \\
+Isles of Scilly & 2012.2 & slow \\
+North Devon & 2002.9 & fast \\
+Gosport & 2003.8 & slow \\
+Thanet & 2004.0 & slow \\
+Na h-Eileanan Siar & NA & NA \\
+Conwy & 2002.6 & fast \\
+\end{longtable}
 
@@
 
-This paper assessed the importance of spatial mechanisms in the
-diffusion of a new technology -- the Web -- at a high level of spatial
-granularity. Specifically, it assessed the capacity of the following
-three spatial diffusion mechanisms in predicting website density in the
-UK over space and time: (i) the hierarchical mechanism according to
-which diffusion runs from main urban centres to secondary ones, (ii) the
-neighborhood mechanism which highlights that diffusion first hits nearby
-locations, and (iii) the rather canonical pattern of diffusion over time
-as reflected in the S-shaped pattern of the cumulative level of
-adoption. This is the first time that the importance of granular spatial
-mechanisms for the diffusion of an intangible, digital technology is
-exposed. The time period of the study is also important as it captures
-the very early stages of this technology until its maturity.
+The paper offers an economic geography study of the diffusion of a new
+technology -- the Web. Contrary to previous studies, it does so by
+adopting (i) a high level of spatial granularity, and (ii) a long study
+period that captures the very early stages of this technology until its
+maturity. This is the first time that the importance of granular spatial
+mechanisms for the diffusion of a new technology is exposed.
 Importantly, the empirical framework and the data employed here allows
@@
 
-The analysis firstly illustrated how well the S-pattern of cumulative
-adoption fits at the local scale of LAD. It highlighted the
-heterogeneity of adoption rates between LAD -- with both more and less
-expected examples -- as well as the spatial clustering of LAD with
-similar adoption rates. The exploratory analysis also illustrated
-instances of remarkable stability as well as volatility of the web
-diffusion system that is indicative of leapfrogging. Previous research
+Specifically, two hypotheses are tested. Firstly, whether the three
+diffusion mechanisms namely distance, urban hierarchy and the S-shaped
+pattern of the cumulative level of adoption shaped the diffusion of the
+commercial Web in the UK at granular geographical scales. The analysis
+firstly illustrates how well the S-pattern of cumulative adoption fits
+at the local scale and offers some descriptive insights about the
+importance of the distance and hierarchical diffusion mechanisms. Then,
+the modelling exercise reveals the remarkably good predictive capacity
+of these mechanisms at granular geographical scales. The modelling
+framework exposes that the relative importance of these spatial
+mechanisms differ across different scales. It also reveals spatial
+heterogeneity. While for some UK regions the importance of the spatial
+diffusion mechanisms are very similar to the rest of the country, other
+regions diverge significantly from the country average.
+
+To do the above, a robust and novel ML modelling framework was designed
+to predict website diffusion over space and time using as predictors the
+spatial diffusion mechanisms. Space- and time-sensitive CV was designed
+for the model to avoid overfitting. Importantly, the results are robust
+against different specifications and subsets of the data. This is the
+first time that the power of these mechanisms in predicting the
+diffusion of a new technology was empirically tested, importantly, at
+very fine scales.
+
+The second hypothesis assessed whether the diffusion of the web is
+shaped by both early adoption and latecomer effects. On the one hand,
+the analysis highlighted the heterogeneity of adoption rates between LAD
+as well as the spatial clustering of LAD with similar adoption rates.
+Importantly, the observed spatial clustering of LAD with fast diffusion
+rates can reinforce existing and rooted patterns of unequal economic
+productivity in the UK \citep{mccann2022productivity}. Previous research
 had already indicated the long-term productivity gains of the early
-adoption of such technologies \citep{tranosuk}. Then, the paper
-developed a robust and novel ML modelling framework to predict website
-diffusion over space and time using the three spatial diffusion
-mechanisms based on space- and time-sensitive CV without suffering from
-overfitting. The model revealed the remarkably good predictive capacity
-of these mechanisms at both scales. For the first time, the power of
-these mechanisms in predicting the diffusion of a new technology was
-empirically tested, importantly, at very fine scales.
-
-The modelling framework also revealed that the importance of these
-spatial mechanisms differ across different scales. While spatial
-structure as reflected in distances to different urban centres is highly
-important at the OA scale, it hardly plays any role at the more coarse
-LAD scale. And while the almost canonical pattern of cumulative adoption
-is a key predictor at the LAD level, it only plays a small role at the
-detailed OA level. Then, the analysis revealed another level of spatial
-heterogeneity. While for some UK regions these spatial diffusion
-mechanisms are very similar to the rest of the country, other regions
-diverge significantly from the country average. Importantly, the above
-results are robust against different specifications and subsets of the
-data.
+adoption of such technologies \citep{tranosuk}. On the other hand, the
+analysis also exposed instances of remarkable stability as well as
+volatility of the Web adoption, with the latter being indicative of
+leapfrogging. While the literature had discussed the benefits of
+technological leapfrogging at the national scale
+\citep{teece2008firm, perkins2005international}, local scales had
+escaped such studies \citep{yu2018sustainability}. Thus the leapfrogging
+evidence presented here advocate towards further exploring such effects
+at granular scales.
 
@@
 \citep{hagerstrand1968innovation, rogers2010diffusion, grubler1990rise},
-their role had never been tested before at such granular scale and,
-importantly, for a digital technology as weightless as the Web. The
+their role had never been tested before at such granular scale. The
 findings talk directly to this literature
@@
 (i) various stakeholders are interested in predicting localised demand
-for new technologies for planning purposes
+for new technologies for planning purposes, for example 5G networks
 \citep{leibowicz2016representing, meade2021modelling}, and (ii) the
-adoption of such digital technologies is associated with economic gains
+adoption of such digital technologies is associated with local economic
+gains
 \citep{solow1957technical, aghion1990model, kemeny2011international, tranosuk, capello2024nexus}.
-In addition, while most related studies focused on tangible
-technologies, this paper sheds light on the diffusion of a truly
-weightless digital technology that is the Web. Thus, it adds much more
-granular findings to the limited related studies
-\citep{sinai2004geography, haller2011determinants}. Importantly,
-although the geographical focus of this paper is the UK, the empirical
-framework developed here is easily transferable to other geographical
-contexts.
+Importantly, although the geographical focus of this paper is the UK,
+the empirical framework developed here is easily transferable to other
+geographical contexts.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/s_uk_per_firm.png}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/s_uk_per_firm_10.png}
 
@@
 
-\caption{\label{s_uk10}Grwoth curve, UK; up to 10 postcodes per website}
+\caption{\label{s_uk10}Cumulative adoption of the Web, UK; up to 10
+postcodes per website}
 
@@
 
-\caption{\label{s_map10}LAD adoption rates; up to 10 postcodes per
-website}
+\caption{\label{s_map10}LAD Web diffusion speed, 1996-2012; up to 10
+postcodes per website}
 
@@
 
-\caption{\label{rank10}Dynamics of wed diffusion (LAD); up to 10
-postcodes per website}
+\caption{\label{rank10}Dynamics of Web diffusion based on LAD rankings;
+up to 10 postcodes per website}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2025_files/figure-pdf/morani10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/morani10-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-11-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{tranos2025_files/figure-pdf/gini10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/gini10-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{tranos2025_files/figure-pdf/varimp10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{anonymised_files/figure-pdf/varimp10-1.pdf}
 
@@
 \end{figure}%
-
-\begin{longtable}[]{@{}lrrrr@{}}
-\caption{Regional differences; up to 10 postcodes per
-website\label{table.regions.no.imp}}\tabularnewline
-\toprule\noalign{}
-Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-South East & 0.947 & 1 & 0.134 & 2 \\
-Wales & 0.916 & 2 & 0.131 & 3 \\
-Yorkshire and The Humber & 0.906 & 3 & 0.144 & 1 \\
-North East & 0.895 & 4 & 0.128 & 4 \\
-West Midlands & 0.883 & 5 & 0.070 & 9 \\
-East Midlands & 0.882 & 6 & 0.088 & 8 \\
-East of England & 0.876 & 7 & 0.106 & 6 \\
-South West & 0.864 & 8 & 0.117 & 5 \\
-London & 0.805 & 9 & 0.055 & 10 \\
-Scotland & 0.770 & 10 & 0.035 & 11 \\
-North West & 0.664 & 11 & 0.017 & 12 \\
-Nortern Ireland & 0.576 & 12 & 0.101 & 7 \\
-\end{longtable}
 
@@
   >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1504}}@{}}
-\caption{S-curve estiamtes for LAD; up to 10 postcodes per
-website\label{table.s.lads}}\tabularnewline
+\caption{S-curve estiamtes for LAD\label{table.s.lads}}\tabularnewline
 \toprule\noalign{}
@@
 
+\begin{longtable}[]{@{}lrrrr@{}}
+\caption{Regional differences; up to 10 postcodes per
+website\label{table.regions.10}}\tabularnewline
+\toprule\noalign{}
+Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Yorkshire and The Humber & 0.864 & 1 & 0.191 & 5 \\
+West Midlands & 0.860 & 2 & 0.148 & 8 \\
+South East & 0.848 & 3 & 0.171 & 7 \\
+North West & 0.827 & 4 & 0.221 & 1 \\
+London & 0.826 & 5 & 0.118 & 9 \\
+South West & 0.801 & 6 & 0.194 & 4 \\
+Wales & 0.746 & 7 & 0.114 & 11 \\
+East Midlands & 0.732 & 8 & 0.117 & 10 \\
+East of England & 0.731 & 9 & 0.178 & 6 \\
+North East & 0.695 & 10 & 0.215 & 2 \\
+Scotland & 0.588 & 11 & 0.052 & 12 \\
+Nortern Ireland & 0.488 & 12 & 0.197 & 3 \\
+\end{longtable}
+
+\begin{longtable}[]{@{}llll@{}}
+\caption{Model metrics for growth
+model}\label{tbl-model-growth-metrics}\tabularnewline
+\toprule\noalign{}
+& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Local Authorities & 0.286 & 0.506 & 0.223 \\
+\end{longtable}
+
+Table~\ref{tbl-model-growth-metrics} presents the results of a Random
+Forest model that was trained based on Equation \ref{model.growth} using
+all LAD data points. In essence, this is a model similar to the one
+estimated based on Equation \ref{eq:model} and presented in
+Table~\ref{tbl-model-metrics}, with the main difference being the
+dependent variable which is the year growth rate of website density for
+each LAD. To avoid cases with LAD have no websites,
+\(t \in [2000, 2012]\). The high predictive capability of the model
+reflects the robustness of the modelling approach.
+
+\begin{align} \label{model.growth}
+Website\,Density\,Growth_{t} \sim Distance\,London +
+Website\,density\,London_{t-1} +\notag\\
+Distance\,Nearest\,City +
+Website\,density\,Nearest\,City_{t-1} +\notag\\
+Distance\,Nearest\,Retail_{i} +
+Website\,density\,Nearest\,Retail_{t-1} +\notag\\
+W*\, Website\,density_{t-1} +\notag\\ 
+year_{t}
+\end{align}
+
 \renewcommand\refname{References}

### 2025-03-12 paper/submission2/tranos2025.qmd
paper v2

--- 2025-01-21 paper/submission2/tranos2025.qmd
@@
 abstract: |
-    Despite the broad consensus that differences in technological adoption and sophistication can lead to uneven trajectories of local development as new technologies can increase productivity and allow firms to introduce new products and services, economic geographers have shied away from exploring the diffusion of digital technologies across space and time.
-This paper fills this gap by exposing the roles of different spatial mechanisms in the diffusion of an intangible, digital technology at very granular geographical scales.
-It does so by employing a novel source of data and machine learning algorithms to model the active engagement with Web during 1996-2012.
+    This paper investigates the spatial diffusion of a new technology that is the Web in the UK.
+It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms.
+Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012).
+Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales.
+They also highlight spatial heterogeneity and instances of leapfrogging.
 keywords: 

### 2025-03-12 paper/submission2/tranos2025.tex
paper v2

@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-6-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-5-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-10-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-11-1.pdf}
 

### 2025-06-08 paper/submission3/response.qmd
R2

--- 2025-01-03 paper/submission2/response.qmd
@@
 
-18-Dec-2024
+\color{teal}
+Thanks a lot for revising the paper, which has improved considerably (from its already high level).
+Apart from the paper's still-not-perfect motivation, there are just a couple of minor things that remain
+to be adapted. As for the motivation, I would (still) like to see a better argumentation for what the paper
+adds to our understanding of diffusion processes, which, in my eyes, can be best achieved by discussing
+some expectations of the relationship between diffusion mechanisms and the considered spatial scale.
+In light of the magnitude of the required changes, I recommend minor revisions. Please see my detailed
+responses to the author’s responses below.\color{blue}
 
-Dear Emmanouil
+Thanks again for the constructive comments and suggestions. Really appreciated!
+Indeed, there is such a theoretical motivation and I agree that highlighting it
+increases the value of the work. Briefly, this is the Modifiable Areal Unit 
+Problem.
 
-Thank you for submitting your manuscript to Journal of Economic Geography.
+The below only includes the comments raised at the last round. I grouped 
+together the points linked to the theoretical motivation. For clarity, the first 
+round comments are in \color{teal}teal \color{blue} and my responses in 
+\color{black} black \color{blue} fonts. Then, the second round comments are in 
+\color{red} red \color{blue} and my responses in blue fonts. I hope the below 
+address all the reviewer comments.
 
-I have now received comments from reviewers.
-Their comments are at the end of this email.
-The reviewers raise a range of concerns and have questions about the suitability of the paper for publication.
-The overall feeling is that the paper has probably been submitted too early and needs major work on most sections if it is to become publishable.
-I concur with the reviewers and see a number of significant issues that need to be addressed.
-This left me with a dilemma in terms of the decision.
-I have decided to invite you to revise and resubmit the paper so that it can be considered further for publication.
-However, this is a very high-risk revision, and I was bordering on rejection given the scale of the work needed and the limitations that the reviewers highlight.
-I am saying this to be open about the risk that after the next round the paper could be rejected unless reviewers believe you have addressed the range of issues identified.
-You may want to factor this into your decision about whether to revise the paper, given how much work is needed.
-
-Having said this, let me offer some suggestions about the main areas to focus on if you decide to revise the paper.
-Reviewer 1 is more positive but very brief in their report.
-I am, therefore, putting more weight on reviewers 2 and 3 when identifying priorities for the revision.
-
-## Editor; theory and justification 
 \color{teal}
-First, you will need to develop a more convincing theoretical argument and contribution.
-You will see that both reviewers 2 and 3 raise questions about your claims in relation to the gap in the literature and suggest you fail to build on existing work appropriately and do not convincingly identify the motivation for this analysis.
-In general, claims that an issue has never been addressed are hard to sustain and it is better to build on existing work and extend knowledge relating to a particular theoretical issue.
-The reviewers challenge you to do this using your data and suggest both the argument and structure of the paper will need to change.
+\noindent 3. Novelty: The study needs to be more explicit about its novel contributions to understanding spatial diffusion mechanisms. Currently, the discussion is too general.
 \color{black}
 
-## Editor; data
+The novelty of the paper lies in the fact that there are hardly any studies
+that analysed the role of these three spatial mechanisms in the diffusion 
+of a new technology at such granular spatial scales. Importantly, the paper does 
+this by adopting a long time series perspective which captures the commercial 
+beginning of the Web until its maturity. The following paragraph in the 
+introduction highlights this:
+
+> This paper offers such a contribution: an economic geography study
+revealing the importance of spatial mechanisms for the diffusion of a
+digital technology that is the Web in the UK at a high level
+of spatial granularity 
+from the very early days until its maturity (1996-2012).
+It specifically focuses on the diffusion of commercial websites.
+It does so by employing a novel source of big data and machine learning
+algorithms to model the *active* engagement with Web -- that is creating and 
+maintaining instead of just browsing a website -- during that period
+and tests how different spatial and temporal mechanisms shaped its
+diffusion. This paper also exemplifies how the combination of data
+sources and state-of-the-art research methods which escape the
+traditional social science domain can offer new lenses to geographical
+research regarding the understanding of technological diffusion.
+
+\color{red}
+Reply: I understand the empirical novelty of the paper, which is unquestioned. However, I miss
+a deeper discussion of why it is important to understand diffusion at such a granular scale and
+why existing studies' results lack such data, are “insufficient” or require this investigation at a
+finer scale.
+\color{blue}
+
+See response below.
+
 \color{teal}
-Second, you need to justify your data and analysis better.
-Reviewer 2 fairly questions the end point of the analysis – which is not explained and given the 10+ years gap to your submission raises questions about the timeliness of the analysis.
-The questions about coverage of the UK website population also need to be addressed convincingly.
-More generally, the reviewers ask for more careful presentation of the data to make clear the significance of each figure and finding.
-As it stands, the paper feels rather overloaded with data, rather than acting as a focused analysis.
+\noindent 4.
+Motivation: The motivation for the study is unclear.
+The text raises several issues, but it is unclear which are specific to digital technologies and which pertain to technological diffusion more broadly.
 \color{black}
 
-## Editor; conclusions
-\color{teal}
-The problems identified above result in the paper feeling imprecise, confusing and lacking a contribution.
-The end point is a conclusion that the reviewers do not find justified or original.
-The reviewers do see potential.
-As I noted, the feeling is a paper that is underdeveloped and submitted prematurely.
-There is, then, a route to publication, but it will require significant changes to convince the reviewers.
-\color{black}
+There are three paragraphs in the revised introduction that highlight the 
+paper's motivation. The first paragraph highlights the importance of this work 
+from a stakeholder perspective. The second emphasises the importance of 
+technology and its adoption in economic geography and macroeconomic growth 
+theories. The third one, motivates the paper from the standpoint of digital 
+geographies and connects it to that literature.
 
-## Referee: 1
+> <!-- motivation: stakeholders -->
+The motivation for this paper lies in the fact that there are various
+stakeholders who are interested in knowing how new digital technologies
+diffuse over space and time and use this knowledge to make predictions
+about the diffusion of related future technologies. As per
+@leibowicz2016representing, historical studies agree that technologies
+diffuse differently in terms of times, rates, and geographies and can be
+driven by related policies [@victor1993]. @meade2021modelling highlight
+that a variety of actors have a direct interest in gaining such
+knowledge including network equipment suppliers, network operators,
+regulatory and local authorities. These processes and their effects vary
+a lot across scales: although the diffusion of a new technology might
+not be optimal at a local level, it might be beneficial from a global
+perspective as it could lead to faster diffusion to less advantaged
+places [@leibowicz2016representing]. Despite the spatial heterogeneity
+of such diffusion mechanisms and the policy relevance, there is a
+scarcity of studies analysing the diffusion of new technologies
+at a detailed geographical level because of the lack of detailed enough
+data and adequate research methods.
 
-I have read your paper with interest and I appreciate its quality and clarity. The work is both well-executed and well-written, making it a valuable contribution to the field.
+<!-- motivation: regional development -->
+>This gap becomes even more prominent if we consider the importance of
+technology and its adoption in economic geography and macroeconomic
+growth theories. Simply put, differences in technological adoption and
+sophistication may lead to uneven development as new technologies can
+increase productivity and allow firms to introduce new products
+[@solow1957technical; @aghion1990model; @kemeny2011international].
+Linked to the subject of this paper, the Web and, more specifically, 
+commercial websites can be considered as the underpinning
+technology for the transformative process identified by
+@capello2024nexus as the *digital service economy*. Importantly, the
+focus here is not the Web in general, but instead commercial websites.
+After all, diffusion together with invention and innovation are
+considered the pillars of technological change [@das2022diffusion].
+Findings from previous studies illustrated that early engagement with
+digital technologies such as the Web led to longstanding positive
+productivity gains for these regions that hosted early adopters
+[@tranosuk]. Hence, obtaining a robust understanding of the spatial
+diffusion mechanisms is crucial in mapping future development trajectories.
 
-However, I would like to offer a few minor suggestions that could further enhance the readability and overall coherence of the paper:
-
-> Simplify the Introduction: I recommend streamlining the introduction by reducing the inclusion of overly technical or highly specific details.
-Instead, consider expanding on the broader conceptual framework that situates the paper within its scholarly context.
-This approach would help orient readers and provide a more accessible foundation for understanding the study.
-*Simplify the introduction*
-
-Strengthen the Connection Between Hypotheses and Results: It may be beneficial to establish a clearer and more explicit link between the study's hypotheses and its findings.
-One way to achieve this could be by using section titles that are more conceptually informative, which would guide the reader through the logical progression of the argument and findings more effectively.
-*change section titles?*
-
-## Referee: 2
-
-\color{teal}
-OVERVIEW\
-- focused on innovation diffusion, using the case of the web in the UK
-- “this paper tests whether the three distinct spatial diffusion mechanisms shaped the diffusion of the Web in the UK at very granular spatial scales from the Web’s early stages onwards”; presumably “S-shaped pattern”/temporal models, core to periphery; distance/proximity.
-
-Figure 2 shows the diffusion of domains per company at the scale of the UK matches the S growth model.
-Figure 3 shows some spatial clustering.
-I think this is for all years in the dataset but this could be better labeled
-\color{black}
-
-**explain, change title**
-
-\color{teal}
-Figure 4, not sure what this demonstrates
-\color{black}
-
-\color{teal}
-Figure 5&6, shows early clustering and then diffusion (I think)
-\color{black}
-
-**yes**
-
-\color{teal}
-DATA\
-Focusing just on .co.uk website is OK but needs to be better justified given that it only represented about 50 percent of domains in the UK at that time.
-See my back of the envelop calculations: As per Zook (2001, Table 1) .uk represented 66% of all domains in the UK, .co.uk is a subset of that, so this is perhaps 50% of all domains registered in the UK at that time.
-If 1/5 of .co.uk domains had findable postal codes (p.
-5), this study is based on 10% of the domains in the UK.
-(perhaps this is too low, aggregating from webpages to websites brought the number higher, not entirely clear to me)
-\color{black}
-
-Thanks for going the extra mile to check the data used here! I agree with the 
-above calculation up until the findable postcode point as 1/5 of the archived 
-*webpages* -- not *websites*, i.e. domains -- included a postcode in the web text. 
-Think of the website of a small company with one trading address. I think it is 
-fair to expect that website to have multiple webpages to present the different
-products and services and then one webpage to showcase the address -- the 
-'Contact us' page. So, the 1/5 share of *webpages* containing a postcode should
-translate to a much higher share of *websites* -- i.e. domains -- containing 
-a postcode. Previous estimates indicated that around 70% of all websites contain 
-place reference [@hill2009georeferencing] and this reference has no been added 
-in the paper.
-
-So, based on the above 35% of all domains with a registered address in the UK
-were represented in the data. Most importantly though the paper does not model 
-the adoption of the Web from the general population but instead from firms. So, 
-although the data only represents 35% of the population of websites with a 
-registered address in the UK, it represents almost all the *relevant* archived
-websites -- that is websites using the commercial .co.uk SLD. As stated in the 
-paper, this has been the primary choice for commercial websites in the UK.
-
-The third paragraph in the Data and Methods section includes the following text:
-
-> Instead of using the whole .uk ccTLD, this paper focuses on its
-commercial subset, the .co.uk second level domain (SLD). This choice
-is aligned with the topic of the paper -- the diffusion of the Web for 
-commercial functions and decreases the heterogeneity of the web data as such 
-commercial websites have specific aims: they are used to diffuse information, 
-support online transactions and share opinions [@THELWALL2000441; 
-@blazquez2018big].
-
-<!-- Active/inactive, don;t suffer from multiple registrations -->
-
-\color{teal}
-overall the time period 1996-2012 is a good time to study the diffusion of the commercial internet. But why stop at 2012? Is it when the Library stopped its data series?
-\color{black}
-
-Yes, the British Library had obtained a copy of all of the .uk contents of the 
-Internet Archive for that period. From a data perspective, it wouldn't be correct
-to extend this dataset with data from other open archives such as the Common 
-Crawl as they use different crawlers and there will be inconsistencies. Most
-importantly though, this period covers the diffusion of the web from the very
-early stages until its maturity. This is also illustrated by the figure 
-depicting the cumulative adoption and the flattening right end of the line.
-
-There is a specific paragraph justifying how the time period of the data fits
-the aim of the paper and another line is now included in the second paragraph 
-of the introductions ("from the very early days until its maturity (1996-2012)").
-
-\color{teal}
-Framing
-“With the exception of Sinai and Waldfogel (2004), who focused, among other things, on web content and demonstrated that more such content is available in larger markets, this strand of literature has been mostly concerned with the spatiality of the Internet’s hard and tangible infrastructure and shied away from exploring the spatial dimensions of one of the most distinct software layers of the Internet infrastructure, the Web.” This statement is simply not true, the literature reviews misses the entirety of web content diffusion studies: Zook, 2000, Malecki, 2002.
-Townsend, 2000, Gorman, 2002, Feldman, 2002, lots more
-\color{black}
-
-Point taken. Despite being aware of these studies, I omitted citing them as I 
-was primarily searching for studies testing the diffusion mechanisms over longer
-time periods. This error has now been corrected and the above paragraph has been 
-replaced with the following one, which emphasises the long time series motivation. 
-Regarding the actual references, @zook2000web, @zook2000web and @moss1997tracking 
-used relevant, domain name registration data, but the rest (e.g. @malecki2002economic 
-and @moss2000internet) focused on hard infrastructure. @gorman2002web examined 
-the head and branch locations of the top 40 e-business integration firms in the 
-USA, which although interesting is not directly relevant to this paper. I also
-went through of Feldman's work around 2002, but I did not find papers
-adopting a similar empirical perspective.
-
+<!-- NEW motivation: intangible/digital -->
 > Another line of motivation has to do with the very nature of the Web as a 
@@
 [@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
-Web among firms can offer useful insights.
+Web can offer useful insights.
 
-\color{teal}
-“for a digital technology as weightless as the Web” – this runs a bit counter to lots of work by digital geographers who have shown again and again how the digital is not weightless or placeless
-\color{black}
+\color{red}
+Reply: Ok, I understand these points and especially appreciate the praxis-oriented motivation.
+What is still a bit thin is why insights from previous studies (which are limited) are not enough
+and require an updated study at a finer granular level (this is the same as the previous point, so
+the two can be combined.
+\color{blue}
 
-\color{teal}
-Conclusion
-p.
-18 “This is the first time that the importance of granular spatial mechanisms for the diffusion of an intangible, digital technology is exposed.” Honestly, I would not expect the diffusion of the web to be any different than any other kind of innovation diffusion.
-The spatiality of digital life, and its connection to material spaces is a constant refrain within digital geography.
-So I’m not convinced it makes much sense to emphasize this, if a granular spatial analysis of innovation diffusion is already novel.
-\color{black}
-
-**for the last two comments, I need to soften the weightless argument throughout.**
-I have added a long time series argument and also said that there is heterogeneity
-[@leibowicz2016representing] among diffusion rates of different technologies.
-
-As per the 'framing' point above, the digital motivation para in the 
-introduction has been amended acordilngly.
-
-## Referee: 3
-
-\color{teal}
-This paper addresses a gap in Economic Geography by examining the spatial diffusion of digital technologies, specifically the adoption of commercial websites, from 1996 to 2012.
-Recognizing the potential of digital technologies to shape local development trajectories through increased productivity and innovation, the study employs novel data sources and machine learning techniques to analyze diffusion patterns at highly granular geographical scales.
-The authors explore the roles of various spatial mechanisms driving this process, offering insights into how digital technology adoption unfolds across space and time.
-The paper has the potential to make an interesting contribution to the literature as it combines a fascinating topic with novel data and methodologies.
-However, in its current form, it is not ready for publication, as it never becomes clear what it actually delivers in terms of new knowledge and what knowledge gap it seeks to fill.
-Despite some analytical elements, it is rather descriptive in nature, which doesn’t fit well with this journal.
-
-Most importantly, I suggest developing some clear hypotheses to structure and motivate the paper better. In light of this, I recommend major revisions.
-\color{black}
-
-This is very helpful. The first two sections have been restructured around two 
-hypotheses. I believe the structure and the flow of the paper is much clearer
-now.
-
-\color{teal}
-\noindent 1. Abstract:\
-\noindent 1.
-The abstract needs significant improvement.
-It needs to adequately report the main findings, relevance, or implications of the study.
-Additionally, the assertion that geographers have "shied away" from analyzing the diffusion of digital technologies is misleading; the issue was primarily due to the lack of available data.
-\color{black}
-
-**fix the abstract**
-
-"shied away" has been replaced in the opening paragraph.
-
-\color{teal}
-\noindent 2. Introduction\
-\noindent 2.
-General Diffusion Studies: The statement that there are few contemporary general diffusion studies in Economic Geography is not accurate, although it may apply specifically to digital technologies.
-Consider the following recent examples:
-o Peris, A., Meijers, E., & van Ham, M.
-(2021).
-Information diffusion between Dutch cities: Revisiting Zipf and Pred using a computational social science approach.
-Computers, Environment and Urban Systems, 85(November), 101565.
-https://doi.org/10.1016/j.compenvurbsys.2020.101565
-o Bednarz, M., & Broekel, T.
-(2020).
-Pulled or pushed?
-The spatial diffusion of wind energy between local demand and supply.
-Industrial and Corporate Change, 29(4), 893–916.
-https://doi.org/10.1093/icc/dtaa012
-o Feldman, M. P., Kogler, D. F., & Rigby, D. L. (2015). rKnowledge: The Spatial Diffusion and Adoption of rDNA Methods. Regional Studies, 49(5), 798–817. https://doi.org/10.1080/00343404.2014.980799
-\color{black}
-
-@PERIS2021101565 is indeed an interesting paper, but it is not directly relevant 
-as it focuses on information diffusion rather than diffusion of new 
-technologies. Information diffuses through social and, consequently, urban 
-networks, but it is not adopted like new technologies. @bednarz2020pulled was 
-already cited and @feldman2015rknowledge has now been added. These references 
-are included in the second section.
-
-Regarding the statement about economic geography studies on diffusion of 
-new technologies and innovation, I do believe that this is a fair statement. 
-The references listed below are the ones reviewed in the second section and none 
-of them are published in one of the flagship economic geography journals 
-(*Journal of Economic Geography* and *Economic Geography*). There are two papers
-published in *The Professional Geographer*, which is a broad-scope geography 
-journal and then only one paper in *Regional Studies*, which is the closest to
-economic geography. Importantly though, the *Regional Studies* paper 
-[@feldman2015rknowledge] paper is focused on inventions as observed by patent 
-data contrary to the broader 'adoption of a new technology' approach of this 
-paper. All other papers listed below are published in business and economics 
-journals.
-
-1. @beardsell1999spatial: European Economic Review
-2. @bednarz2020pulled: Industrial and Corporate Change
-3. @haller2011determinants: Applied Economics
-4. @perkins2011internet: The Professional Geographer
-5. @feldman2015rknowledge: Regional Studies
-6. @ding2010modeling: The Professional Geographer
-7. @bakher2013diffusion: Journal of Marketing Analytics
-8. @PAPAGIANNIDIS2015308: Technological Forecasting and Social Change
-9. @lengyel2020role: Scientific reports
-
-Also, the lack of spatial granularity argument has been highlighted in various 
-instances throughout the paper as the main line of justification.
-
-**ADD SPATIAL GRANULARITY IN CONCLUSIONS**
-
-\color{teal}
-\noindent 3. Novelty: The study needs to be more explicit about its novel contributions to understanding spatial diffusion mechanisms. Currently, the discussion is too general.
-\color{black}
-
-**granularity**
-
-**HAVE A THINK, intro has been restructured**
-
-\color{teal}
-\noindent 4.
-Motivation: The motivation for the study is unclear.
-The text raises several issues, but it is unclear which are specific to digital technologies and which pertain to technological diffusion more broadly.
-\color{black}
-
-**HAVE A THINK, intro has been restructured**
-
-\color{teal}
-\noindent 5.
-Structure and Focus: The discussion on motivations should be streamlined and clarified.
-For example, after specifying that the study focuses on commercial websites rather than the web, subsequent references revert to the general term “web.” Additionally, the discussion confuses diffusion patterns, firm perspectives, and early adopter advantages.
-Consider restructuring this section to clearly differentiate between literature gaps (motivations) and expected outcomes (mechanisms).
-\color{black}
-
-The revised paper consistently illustrates that the focus is on commercial 
-websites.
-
-In the relevant paragrapg in the Data and Methods section, the following line
-has been added: "Henceforward any reference to websites will refer to commercial 
--- .co.uk -- websites".
-
-**fixed for intro, lit rev  and data -- check the rest of the paper.**
-
-The introduction and literature review sections have been restructured, so 
-the firm perspectives and early adopter advantages discussions have been moved.
-
-\color{teal}
-\noindent 6.
-Data Consistency: The terminology for the data should be consistent.
-At one point, the text refers to the “totality of the web,” while elsewhere, it specifies “commercial websites.” Clarify and maintain consistency throughout.
-\color{black}
-
-The terminology is now consistent across the paper. The "totality" point 
-referred to different web technologies and not to the totality of the web. It 
-has now been replaced with "an overarching technology".
-
-\color{teal}
-\noindent 7.
-Recommended Structure: To align with the common structure in Journal of Economic Geography (JoEG), consider splitting the introduction into a shorter introduction and a dedicated theory section merged with the literature review.
-\color{black}
-
-The introduction is now shorter and the literature review section has been 
-re-structured and re-named as "The mechanisms that shape the diffusion of a new 
-technology".
-
-\color{teal}
-\noindent 3. Literature Review / Theory\
-\noindent 1.
-Synthesis and Motivation: The literature review should be transformed into a theory section.
-Rather than presenting a historical review, synthesize the arguments to clearly motivate and inform your hypotheses.
-\color{black}
-
-Done. Restructured: theory, gaps, hypotheses.
+See response below.
 
@@
 
-Known processes at very detailed scale: explanatory power of these mechanisms at granular, local
-scales
+The knowledge gaps have been explicitly highlighted in the restructured 
+literature review section. For instance:
 
-\color{teal}
-\noindent 3. Lack of Hypotheses: To clarify what the paper delivers and what knowledge gaps it fills, I strongly recommend developing a number of hypotheses that will be tested.
-\color{black}
+> Although the above processes have been studied in innovation studies, the distancing of economic geographers from research questions about the diffusion of new technologies and innovations resulted to a knowledge gap regarding the explanatory power of these mechanisms at granular, local scales [@ding2010modeling].
 
-Added.
+> Although the early adoption and leapfrogging literature has not paid much attention on cities and regions [@yu2018sustainability], the potential local economic benefits justify the need to investigate the dynamics of the web diffusion in different scales
+in the UK. Hence, the second empirical hypothesis states:
 
-\color{teal}
-\noindent 4. Fit to title: The term “multi-scalar” appears in the title but is never really discussed in the paper.
-\color{black}
+\color{red}
+Reply: Please see my previous points concerning the motivation. While this is a valid argument,
+I still miss a strong claim as to why existing studies leave a knowledge gap and why filling this
+gap matters.
+\color{blue}
 
-This was discussed from the data section onwards, but indeed not before that. 
-Section 2 includes a new paragraph describing the second hypothesis (H~2~), 
-which explicitely discusses the value of the multiscale approach: 
-"Understanding the diffusion of a new technology in multiple, detailed spatial
-scales is of interest to economic geography as ..."
-
-\color{teal}
-\noindent 4. Data & Methods\
-\noindent 1. Order of Presentation: Introduce the data first, followed by the geographical units used in the study.
-\color{black}
-
-The order of presentation has been amended accordingly and the two scales of 
-analyses are presented after the data.
-
-\color{teal}
-\noindent 2. Postcode Identification: The approach of identifying postcodes on websites is innovative, but potential errors need to be discussed. Specifically:
-- How do you ensure that a postcode on a website accurately represents the 
-company’s location or its subsidiary?
-- How do you handle numbers that resemble postcodes but are not actual postcodes?
-- What about references to events or other organizations unrelated to the website’s owner?
-Include a brief discussion of these potential issues.
-\color{black}
-
-A number of different strategies have been developed to address the potential 
-issues described above.
-
-\noindent 1. To prevent false positives, the resulted postcodes
-were matched against official postcode lookup files [@ons]. This was not 
-explicitly mentioned in the first submission on the paper and the previous 
-sentence has now been added in the relevant paragraph:
-
-> The text from these webpages was scanned using a regular expression
-(regex) to identify strings of text which resemble UK postcodes and one
-fifth of them included a mention to a postcode [@BL2013geo]. This
-information allows the geolocation of the data and the creation of the
-LAD and OA counts. To prevent false positives, the resulted postcodes
-were matched against official postcode lookup files [@ons].
-
-\noindent 2. Then, to remove the bias from link farms associated with a handful of postcodes
-having irregular high numbers of websites pointing to them, a targeted data 
-imputation procedure was developed. This is detailed in the data section -- see
-also Figure 1.
-
-\noindent 3. Regarding the points about subsidiary company location and event postings, 
-these cannot be directly observed and distinguished. However, a process that has 
-been successfully used in previous studies has been adopted here. Firstly, given 
-that the focus of the paper is commercial websites, the relevant literature 
-illustrates that such websites serve specific purposes:
-
-> they are used to diffuse information, support online transactions and share 
-opinions [@THELWALL2000441; @blazquez2018big].
-
-Secondly, the webpage-to-website aggregation process leads to websites with 
-different number of postcodes as presented in Table 1. To decrease heterogeneity, 
-the analysis firstly focuses on websites with one unique postcode. Given the the
-commercial nature of these websites (.co.uk), it is fair to expect that such 
-websites with one unique postcode in their web text represent companies with a 
-single trading location. The analysis included in the main body of the paper is
-based on this subset of the website population. For year 2000, this includes 
-70% of all the archived, geolocated commercial websites. Then, as a robustness
-check all the different steps of the analysis are repeated for websites with up 
-to 10 postcodes and the results are presented in the appendix. This includes 94%
-of all the archived, geolocated commercial websites. For this step, websites 
-with multiple postcodes were counted multiple times, one for each different 
-postcode. All the results are qualitatively the same and only marginally 
-different, which is indication of the robustness of the website geolocation 
-strategy.
-
-As a further argument for the robustness of the above process, the results of a 
-previous study which used the same approach for a small subset of these data 
-from a specific neighbourhood in London matched in terms of accuracy previous 
-research that was based on extensive qualitative work, interviews and 
-observational studies [@shoreditch]. This last point has now been added in the 
-relevant section:
-
-> Previous research which used the same methodology for a small subset of these 
-data illustrates the robustness of the above approach: its results matched in 
-terms of accuracy past rounds of research based on extensive qualitative work, 
-interviews and observational studies for the same study area [@shoreditch].
-
-\color{teal}
-\noindent 3.
-Equation 2: The rationale behind Equation 2 is unclear.
-Shouldn’t the primary predictor of website density at time tbe the location density at t-1?
-Why was this approach chosen over a growth model?
-Additionally, clarify how spatial dependencies are accounted for in the model.
-\color{black}
-
-**growth as an appendix, spatial dependencies have been included**
-
-\color{teal}
-\noindent 5. Results\
-\noindent 1.
-Diffusion Speed: The decision to binarize diffusion speed is questionable.
-Continuous measures, such as growth rates or the time required to reach 50% saturation, would provide more informative insights.
-\color{black}
-
-**This is only for the map, this is a critical point in diffusion studies**
+See response below.
 
@@
 
-**reveal that theories are valid for lcoal scales and inform local tech policies**
+A new figure (Figure 3) and the following text have been added to address this
+point.
+
+> What escaped from previous studies was εexploring whether such a pattern also 
+describes cumulative adoption of new technologies at sub-national scales. To 
+test this, the same  logistic function is estimated for each of the 374 LAD. The
+detailed results are presented in Table \ref{table.s.lads} in the Appendix. 
+The key finding is that for 342 LAD (91% of all LAD) the $R^2 > 0.9$ for the 
+logistic function, which indicates that even at local scale the cumulative 
+adoption of the Web is characterised by a pattern similar to the one observed 
+for countries. Graphically, all these LAD curves are plotted in Figure 
+\ref{s_lad_uk}. The different colour lines illustrate whether a LAD reached $t_0$ 
+faster or slower than the UK in total. This heterogeneity is further discussed 
+later in this section.
+
+> This is the first time that such cumulative adoption curves have been estimated
+for areal units as small as the UK LAD. Understanding and modelling the dynamics
+of technological diffusion and, ultimately, being able to predict diffusion 
+rates at such small local scales can directly support the deployment of 
+infrastructural networks (e.g. 5G) and local technology policies. The above are 
+in line with previous studies exploring the diffusion of web technologies 
+[@PAPAGIANNIDIS2015308], social media [@lengyel2020role] and online
+shopping [@bakher2013diffusion].
+
+\color{red}
+Reply: Ok, that is interesting and relevant and can be partially used to address my comment about the
+motivation of the study. Yet, what is still not perfectly clear to me is whether the similarity of the
+findings at the finer spatial level is to be expected or if there are theoretical reasons to expect otherwise.
+In my eyes, using such a fine granular scale is the central (but not only) contribution of the paper.
+Motivating it better, i.e., why it matters from a theoretical point and understanding of diffusion patterns,
+can elevate the paper’s value. Put differently, are the diffusion mechanisms and drivers (spatially) scale
+invariant? Discussing this briefly would greatly contribute to the paper’s message.
+\color{blue}
+
+I think all the above points can be addressed together, hence I am responding to 
+them here. Indeed, there is a deeper theoretical justification for testing
+how sensitive the diffusion mechanisms are to the geographical scale of observation 
+and this is the Modifiable Unit Areal Problem. We know that various key theoretical concepts
+in economic geography are sensitive to the spatial scale at which they are 
+observed, but given the lack of previous studies, we do not know whether 
+the diffusion mechanisms of a new technology are also sensitive to the spatial 
+scale as well. The following paragraph has been added in the introduction.
+
+> The lack of previous granular studies about the diffusion of new technologies 
+deprives us from a reliable prior regarding whether the diffusion mechanisms 
+and drivers are sensitive or not to the spatial scale that the diffusion of a 
+new technology is observed. As per other key concepts in economic geography
+[@briant2010dots; @meliciani2015determinants; @andini2019social; @rosenthal2001determinants],
+diffusion mechanisms can indeed suffer from Modifiable Areal Unit Problem 
+(MAUP): although the adoption of a new technology is at its core an individual 
+process, it is hardly ever observed at this, or at a close enough, scale.
+
+---
 
 \color{teal}
-\noindent 3.
-Early Adopters and Latecomers: The discussion of early-adopter and latecomer advantages is underdeveloped.
-How do these advantages manifest locally?
-What specific benefits arise if a company registers its website six months earlier than competitors in other locations?
-If this aspect is retained, it should be elaborated in the theory section, including how it relates to local capabilities or diffusion hierarchies.
+\noindent 3. Lack of Hypotheses: To clarify what the paper delivers and what knowledge gaps it fills, I strongly recommend developing a number of hypotheses that will be tested.
 \color{black}
 
-**relates to local capabilities or diffusion hierarchies**
+The following two hypotheses have been added and the whole paper is now 
+structured around them:
 
-\color{teal}
-\noindent 4.
-Descriptive Statistics: The value of the descriptive Moran’s I and Gini coefficient analyses in Section 4.2 is unclear.
-Instead, calculating growth rates of adoption and examining their distribution over time would offer a better sense of adoption speed and magnitude.
-\color{black}
+> H~1~: The three diffusion mechanisms namely distance, urban hierarchy
+and the S-shaped pattern of the cumulative level of adoption shaped
+the diffusion of the commercial Web in the UK even at granular geographical 
+scales.
 
-**think, growth rates will not work for OA**
+> H~2~: The diffusion of the commercial Web is shaped by both early adoption 
+and latecomer effects.
+
+\color{red}
+Reply: I do wonder a bit what the alternatives are. I mean, what factors other than the ones named in
+the hypothesis can be expected to share the diffusion? Please clarify in the text leading to the
+hypotheses, explaining why these factors can be expected to be dominant.
+\color{blue}
+
+The below sentence has been added after the description of the three mechanisms:
+
+> The diffusion of new technologies and innovations is driven by three
+different mechanisms [@grubler1990rise; @morrill2020spatial]: (i)
+distance and proximity, (ii) urban hierarchies in the form of a
+centrifugal forces as new technologies diffuse from core to periphery,
+and (iii) over time the cumulative level of adoption follows an S-shaped
+pattern.
+<!-- R2 addition -->
+Although some studies focused on other, case-specific drivers -- for example 
+@lengyel2020role on urban scaling and @bednarz2020pulled on regional demand --
+the literature is in agreement that these are the three fundamental diffusion 
+mechanisms.
 
@@
 
-**hierarchy**
+The retail centres discussion has now been moved in the methods sections, so it
+appears much earlier in the paper. In addition, the following text has been 
+modified in the relevant paragraph:
 
-\color{teal}
-\noindent 6.
-Modeling Details: The methodological details in Section 4.3 should be moved to the methods section to maintain clarity and coherence.
-Including these details in the results section is confusing.
-\color{black}
+> To capture the hierarchical mechanism the model includes as predictors a
+one year lag of website density in London, the largest city in the UK, and a
+one year lag of the website density in the nearest city. Both variables clearly
+illustrate the urban hierarchy in the UK. Especially for the OA models, a 
+third predictor is added to capture hierarchy at a much more granular scale:
+a one year lag of the website density in the the nearest retail centre [^4]. As
+the literature illustrates, retail centres are key components of urban systems
+[@dennis2002central, @jones2021reframing]. Due to the small sizes of the retail 
+centres, the latter is only relevant for and included in the OA models and not 
+in the LAD models.
 
-\color{teal}
-\noindent 7.
-Clarity and Structure: Section 4.3 is difficult to follow.
-Clearly state the main message and highlight how diffusion mechanisms vary spatially.
-Develop hypotheses to structure the results and provide potential explanations for observed heterogeneity.
-Without this, the section feels like a descriptive summary without clear takeaways for the reader.
-\color{black}
+> [^4]: The retail centres data have been obtained by the Consumer Data Research Centre, an ESRC data investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1: <https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators>.
 
-\color{teal}
-\noindent 
-Overall Recommendations
-\noindent 1. Consistency: Ensure consistent terminology and focus throughout the paper, particularly when referring to the web and commercial websites.
-\noindent 2. Structure: Adopt a clearer structure by separating the introduction, theory/literature review, methods, and results sections.
-\noindent 3. Clarity: Develop hypotheses to frame the results and provide clearer explanations for methodological choices and findings.
+\color{red}
+Reply: Ok. I would suggest refining the text a bit to make clear that this is an illustration of a specific
+empirical case, i.e., the consideration of an additional factor, as an illustration of the robustness of the
+results as well as potential adaptability of the empirical model to consider more factors in other
+circumstances.
+\color{blue}
+
+The below sentence has been added in the above paragraph:
+
+> By including this predictor, the model is better tuned towards the spatial 
+characteristic of the OA scale. Hence, it increases the robustness of the model.
+
 \color{black}

### 2025-06-08 paper/submission3/anonymised.qmd
R2

--- 2025-03-12 paper/submission2/anonymised.qmd
@@
 and tests how different spatial and temporal mechanisms shaped its
-diffusion. This paper also exemplifies how the combination of data
+diffusion [@tranosuk]. This paper also exemplifies how the combination of data
 sources and state-of-the-art research methods which escape the
@@
 Web can offer useful insights.
+
+<!-- R2 addition -->
+The lack of previous granular studies about the diffusion of new technologies 
+deprives us from a reliable prior regarding whether the diffusion mechanisms 
+and drivers are sensitive or not to the spatial scale that the diffusion of a 
+new technology is observed. As per other key concepts in economic geography
+[@briant2010dots; @meliciani2015determinants; @andini2019social; @rosenthal2001determinants],
+diffusion mechanisms can indeed suffer from Modifiable Areal Unit Problem 
+(MAUP): although the adoption of a new technology is at its core an individual 
+process, it is hardly ever observed at this, or at a close enough, scale.
 
@@
 pattern.
+<!-- R2 addition -->
+Although some studies focused on other, case-specific drivers -- for example 
+@lengyel2020role on urban scaling and @bednarz2020pulled on regional demand --
+the literature is in agreement that these are the three fundamental diffusion 
+mechanisms. 
 <!-- neighbourhood -->
@@
 
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{correct}Yearly website counts per postcode (top 1000) before (top) and after (down) data imputation", out.height="70%"}
-path.image1 <- paste0(path, "/outputs/pc_year_1.png")
-path.image2 <- paste0(path, "/outputs/pc_year_1_corrected.png")
-
-p1 <- ggdraw() + draw_image(path.image1)  
-p2 <- ggdraw() + draw_image(path.image2)
-
-par(mar=rep(0,4)) # no margins
-
-p1 + plot_spacer() + p2 + plot_layout(ncol = 2)#, widths = c(1, -0.8, 1))
-```
-
 <!-- Multi-scale -->
@@
 
+```{r echo=FALSE, message=FALSE, fig.cap="\\label{correct}Yearly website counts per postcode (top 1000) before (top) and after (down) data imputation", out.height="70%"}
+path.image1 <- paste0(path, "/outputs/pc_year_1.png")
+path.image2 <- paste0(path, "/outputs/pc_year_1_corrected.png")
+
+p1 <- ggdraw() + draw_image(path.image1)  
+p2 <- ggdraw() + draw_image(path.image2)
+
+par(mar=rep(0,4)) # no margins
+
+p1 + plot_spacer() + p2 + plot_layout(ncol = 2)#, widths = c(1, -0.8, 1))
+```
+
 <!-- ESDA -->
@@
 <!-- detailed features -->
-To capture the hierarchical mechanism the model include as predictors a
+To capture the hierarchical mechanism the model includes as predictors a
 one year lag of website density in London, the largest city in the UK, and a
@@
 centres, the latter is only relevant for and included in the OA models and not 
-in the LAD models. In addition, the model includes the distance to London, the 
+in the LAD models. By including this predictor, the model is better tuned 
+towards the spatial characteristic of the OA scale. Hence, it increases the 
+robustness of the model.
+
+In addition, the model includes the distance to London, the 
 nearest city and the nearest retail centre. The underlying logic is that the 

### 2025-06-08 paper/submission3/anonymised.tex
R2

--- 2025-03-12 paper/submission2/anonymised.tex
@@
 maintaining instead of just browsing a website -- during that period and
-tests how different spatial and temporal mechanisms shaped its
-diffusion. This paper also exemplifies how the combination of data
-sources and state-of-the-art research methods which escape the
+tests how different spatial and temporal mechanisms shaped its diffusion
+\citep{tranosuk}. This paper also exemplifies how the combination of
+data sources and state-of-the-art research methods which escape the
 traditional social science domain can offer new lenses to geographical
@@
 modelling the diffusion of the Web can offer useful insights.
+
+The lack of previous granular studies about the diffusion of new
+technologies deprives us from a reliable prior regarding whether the
+diffusion mechanisms and drivers are sensitive or not to the spatial
+scale that the diffusion of a new technology is observed. As per other
+key concepts in economic geography
+\citep{briant2010dots, meliciani2015determinants, andini2019social, rosenthal2001determinants},
+diffusion mechanisms can indeed suffer from Modifiable Areal Unit
+Problem (MAUP): although the adoption of a new technology is at its core
+an individual process, it is hardly ever observed at this, or at a close
+enough, scale.
 
@@
 and (iii) over time the cumulative level of adoption follows an S-shaped
-pattern. \citet{hagerstrand1968innovation} was the first to identify
-diffusion as a geographical process. The starting point was the idea
-that diffusion is based on passing information through social networks,
-which themselves tend to be geographically bounded. Hence, he identified
-a `neighbourhood' mechanism of how information, and consequently,
+pattern. Although some studies focused on other, case-specific drivers
+-- for example \citet{lengyel2020role} on urban scaling and
+\citet{bednarz2020pulled} on regional demand -- the literature is in
+agreement that these are the three fundamental diffusion mechanisms.
+\citet{hagerstrand1968innovation} was the first to identify diffusion as
+a geographical process. The starting point was the idea that diffusion
+is based on passing information through social networks, which
+themselves tend to be geographically bounded. Hence, he identified a
+`neighbourhood' mechanism of how information, and consequently,
 innovations and new technologies diffuse. This perspective draws
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{anonymised_files/figure-pdf/unnamed-chunk-2-1.pdf}
-
-}
-
-\caption{\label{correct}Yearly website counts per postcode (top 1000)
-before (top) and after (down) data imputation}
-
-\end{figure}%
-
 To empirically test the two hypotheses stated in
@@
 
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{anonymised_files/figure-pdf/unnamed-chunk-2-1.pdf}
+
+}
+
+\caption{\label{correct}Yearly website counts per postcode (top 1000)
+before (top) and after (down) data imputation}
+
+\end{figure}%
+
 Next, exploratory spatial data analysis provides some first insights of
@@
 
-To capture the hierarchical mechanism the model include as predictors a
+To capture the hierarchical mechanism the model includes as predictors a
 one year lag of website density in London, the largest city in the UK,
@@
 small sizes of the retail centres, the latter is only relevant for and
-included in the OA models and not in the LAD models. In addition, the
-model includes the distance to London, the nearest city and the nearest
-retail centre. The underlying logic is that the level of website
-adoption in a spatial unit depends on the level of the adoption in
-places further up in the urban hierarchy the previous year. The
+included in the OA models and not in the LAD models. By including this
+predictor, the model is better tuned towards the spatial characteristic
+of the OA scale. Hence, it increases the robustness of the model.
+
+In addition, the model includes the distance to London, the nearest city
+and the nearest retail centre. The underlying logic is that the level of
+website adoption in a spatial unit depends on the level of the adoption
+in places further up in the urban hierarchy the previous year. The
 inclusion of the distance variables incorporates spatial structure into

### 2025-06-08 paper/submission3/tranos2025.qmd
R2

--- 2025-06-08 paper/submission3/anonymised.qmd
@@
 title: "A multi-scale story of the diffusion of a new technology: the web"
+author:
+  - name: Emmanouil Tranos
+    email: e.tranos@bristol.ac.uk
+    affiliations: 
+      - id: some-tech
+        name: University of Bristol
+        department: 
+        address: 
+        city: 
+        state: State
+        postal-code: UK
+    corresponding: e.tranos@bristol.ac.uk
+address:
+  - attributes:
+        corresponding: true
+    note: This is the first author footnote.
 abstract: |

### 2025-06-08 paper/submission3/tranos2025.tex
R2

--- 2025-06-08 paper/submission3/anonymised.tex
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-5-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-5-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-11-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{tranos2025_files/figure-pdf/unnamed-chunk-11-1.pdf}
 

### 2025-07-29 paper/submission3/title.tex
submission 3

--- 2025-03-12 paper/submission2/anonymised.tex
@@
 
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 \makeatletter
@@
 \makeatother
-\journal{Journal Name}
+\journal{Journal of Economic Geography}
 
@@
 web}
+\author[1]{Emmanouil Tranos%
+\corref{cor1}%
+}
+ \ead{e.tranos@bristol.ac.uk}
+
+\affiliation[1]{organization={University of Bristol, School of
+Geographical Sciences},addressline={University Road, BS8
+1SS},city={Bristol},country={UK},countrysep={,},postcodesep={}}
 
@@
 
-\section{Introduction}\label{sec-introduction}
+\section{Acknowledgements}\label{acknowledgements}
 
-Economic geographers were always interested in how new technologies and
-innovations diffuse across space and time with the work of
-\citet{hagerstrand1968innovation}, which demonstrated the importance of
-communications as a mechanism for the diffusion of innovations, being
-the most impactful. It was followed by various empirical studies -- see
-for instance \citet{ormrod1990} and \citet{iso2005}. However, the torch
-of exploring and modelling such diffusion processes had been passed to
-other disciplines such as economics, business studies and sociology well
-before the `cultural turn' of economic geography
-\citep{perkins2005international, ding2010modeling}. As a consequence,
-relevant studies did not investigate diffusion mechanisms at granular
-geographical scales to provide local policy insights, for instance,
-regarding infrastructural development and local technology policies. A
-potential explanation of the lack of geographical studies exploring the
-diffusion of new and, more specifically for this paper, digital
-technologies across both space and time can be attributed to the
-scarcity of relevant and granular enough data, a problem also
-highlighted by \citet{iso2005} and \citet{kemeny2011international}. As
-\citet{zook2022mapping} pointed, digital activities are hardly ever
-captured in official data.
+Many thanks to Dr Levi Wolf and Professor Dani Arribas-Bel for their
+constructive comments. Also many thanks to the participants of the 6th
+Global Conference on Economic Geography and the Winter Seminar of GfR.
+Then, many thanks to the organisers of the seminar series and the
+members of the Quantitative Spatial Science (QuSS) Research Group at the
+University of Bristol, the Centre for Business Prosperity Research Group
+at Aston University (Dr Tasos Kitsos), and the Centre for Innovation
+Research (Professor Tom Broekel) at the University of Stavanger.
+Finally, many thanks to the editor and the three anonymous reviewers for
+the very helpful comments and suggestions.
 
-This paper offers such a contribution: an economic geography study
-revealing the importance of spatial mechanisms for the diffusion of a
-digital technology that is the Web in the UK at a high level of spatial
-granularity from the very early days until its maturity (1996-2012). It
-specifically focuses on the diffusion of commercial websites. It does so
-by employing a novel source of big data and machine learning algorithms
-to model the \emph{active} engagement with Web -- that is creating and
-maintaining instead of just browsing a website -- during that period and
-tests how different spatial and temporal mechanisms shaped its
-diffusion. This paper also exemplifies how the combination of data
-sources and state-of-the-art research methods which escape the
-traditional social science domain can offer new lenses to geographical
-research regarding the understanding of technological diffusion.
+\section{Funding}\label{funding}
 
-The motivation for this paper lies in the fact that there are various
-stakeholders who are interested in knowing how new digital technologies
-diffuse over space and time and use this knowledge to make predictions
-about the diffusion of related future technologies. As per
-\citet{leibowicz2016representing}, historical studies agree that
-technologies diffuse differently in terms of times, rates, and
-geographies and can be driven by related policies \citep{victor1993}.
-\citet{meade2021modelling} highlight that a variety of actors have a
-direct interest in gaining such knowledge including network equipment
-suppliers, network operators, regulatory and local authorities. These
-processes and their effects vary a lot across scales: although the
-diffusion of a new technology might not be optimal at a local level, it
-might be beneficial from a global perspective as it could lead to faster
-diffusion to less advantaged places \citep{leibowicz2016representing}.
-Despite the spatial heterogeneity of such diffusion mechanisms and the
-policy relevance, there is a scarcity of studies analysing the diffusion
-of new technologies at a detailed geographical level because of the lack
-of detailed enough data and adequate research methods.
-
-This gap becomes even more prominent if we consider the importance of
-technology and its adoption in economic geography and macroeconomic
-growth theories. Simply put, differences in technological adoption and
-sophistication may lead to uneven development as new technologies can
-increase productivity and allow firms to introduce new products
-\citep{solow1957technical, aghion1990model, kemeny2011international}.
-Linked to the subject of this paper, the Web and, more specifically,
-commercial websites can be considered as the underpinning technology for
-the transformative process identified by \citet{capello2024nexus} as the
-\emph{digital service economy}. Importantly, the focus here is not the
-Web in general, but instead commercial websites. After all, diffusion
-together with invention and innovation are considered the pillars of
-technological change \citep{das2022diffusion}. Findings from previous
-studies illustrated that early engagement with digital technologies such
-as the Web led to longstanding positive productivity gains for these
-regions that hosted early adopters \citep{tranosuk}. Hence, obtaining a
-robust understanding of the spatial diffusion mechanisms is crucial in
-mapping future development trajectories.
-
-Another line of motivation has to do with the very nature of the Web as
-a digital technology and how it evolved over time. The digital
-geographies field has debunked premature claims about the Internet's
-placeless-ness by mapping various of its facets and technological layers
-\citep{tranos2013death}, but the lack of relevant long time-series data
-prevented researchers from observing digital phenomena overtime. Most of
-these studies focused on its hard infrastructure
-\citep[e.g.][]{malecki2002economic, moss2000internet}, on Internet
-subscribers \citep[e.g.][]{blank2018local}, or on social media users
-\citep[e.g.][]{crampton2013beyond}. Exceptions include early work from
-\citet{zook2000web}, \citet{zook2001old} and \citet{moss1997tracking},
-which utilised domain name registration data and use it to map and
-illustrate the spatial footprint of the web content. Pivotal as these
-contributions have been in understanding the economic geographies of the
-Internet, they were based on cross-sectional, often spatially detailed,
-mapping and on short temporal comparisons (e.g.~1993-1998) and did not
-observe the evolution and diffusion of the Web until its maturity. As we
-know from the relevant literature, different technologies diffuse at
-different rates and there is no theory that can explain these
-differences \citep{leibowicz2016representing}. Hence, mapping and
-modelling the diffusion of the Web can offer useful insights.
-
-This paper focuses on the diffusion of the Web as new technology during
-the 1996-2012 period. This was an exciting period for digital
-technologies as it corresponds with the commercialisation of the
-Internet and, consequently, its almost universal adoption. The reader is
-reminded that it was only in 1994 when Netscape Navigator was
-introduced, a year before Microsoft's Internet Explorer.\footnote{\url{https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft}}
-Also, only 9 per cent of UK's households had access to the Internet in
-1998 \citep{ons2018}, the web included mostly static webpages, there
-were no social media and web browsing involved exclusively a desktop PCs
-as there were no smartphones \citep{tranosuk}. Hence, it is fair to say
-that the study period captures the very early stages of the diffusion of
-a new technology that is the Web until its maturity. The former is a key
-point in the lifecycle of a new technology. Firstly, during this period
-new technologies are expensive, crude and imperfect
-\citep{rosenberg1994exploring, wilson201281}. A simple comparison
-between Web 1.0 and Web 3.0 applications clearly illustrates this
-\citep{tranos2020social}: while a static website facilitates one-way
-information dissemination, a platform like \texttt{github} enables
-cooperation between users and the creation of new information, meaning,
-and even knowledge \citep{faraj2016special, barassi2012does}. During
-this period the performance of a new technology is the main attraction
-and not the cost to access and use it \citep{wilson2011lessons}. There
-is a broader theoretical discussion in the literature about the early
-adoption motives. As summarised by \citet{perkins2005international}, on
-one hand, epidemic models highlight the role of interpersonal contacts
-as a way for new technologies to diffuse. On the other hand, economic
-models underline the importance of heterogeneity. Different firms have
-different structures and business plans, which define the potential
-economic returns of the adoption of a new technology and, therefore, the
-choice to adopt a new technology becomes an individual option. From a
-broader and evolutionary perspective, initial conditions are essential
-for the creation and evolution of path-dependent technological
-development trajectories \citep{neffke2011regions, simmie2014new}. This
-argument is even more relevant when the focus is on digital technologies
-because of the commonly found lag between investment and economic
-returns as reflected in the Solow paradox
-\citep{acemoglu2014return, brynjolfsson2018artificial}.
-
-The data used here depict the active engagement with the Web as they
-contain geolocated and time-stamped counts of commercial websites from
-the UK. Instead of adopting metrics reflecting the passive engagement
-with the Internet such as Internet adoption and speeds
-\citep[e.g.][]{blank2018local, destefano2022fuzzy}, this paper observes
-yearly counts of geolocated, commercial websites, which reflect the
-creation and maintenance of such websites. It needs to be highlighted
-here that although there are various types of underpinning web
-technologies -- namely markup, stylesheet, client- and server-side
-coding languages as well as different web browsers and other related
-programming language -- the focus here is on the diffusion of the Web as
-an overarching technology instead of more specific web technologies. For
-a such a detailed and cross-country technological diffusion study, see
-\citet{PAPAGIANNIDIS2015308}.
-
-To analyse these data, a diverse set of methods are employed. Firstly,
-cumulative adoption curves are estimated for the UK as a whole and for
-the Local Authority Districts (LAD), illustrating, for the first time,
-such S-shaped diffusion patterns for local areas. The exploratory
-spatial data analysis offers some first indications of the importance of
-the different spatial diffusion mechanisms in shaping the diffusion of
-the Web in the UK. Importantly, the analysis adopts two distinct
-geographical scales: the LAD and the most detailed scale of UK census
-areas, the Output Areas (OA). Then, a novel methodological framework
-based on Machine Learning (ML) algorithms is developed to test the role
-of such mechanisms at these two scales. The results reveal, for the
-first time, the importance of different spatial diffusion mechanisms in
-the diffusion of a digital technology at very granular geographical
-scales. The analysis also reveals that the diffusion of the Web is
-characterised by spatial heterogeneity, well-established spatial
-patterns and both stability and volatility.
-
-The structure of the paper goes as follows. Section~\ref{sec-litreview}
-reviews the mechanisms that shape the diffusion of a new technology and
-sets up the two hypotheses that are empirically tested in
-Section~\ref{sec-results}. Section~\ref{sec-datamethods} describes the
-data and methods used in this paper. The discussion of the results and
-the conclusions of the paper are offered in
-Section~\ref{sec-conclusions}.
-
-\section{The mechanisms that shape the diffusion of a new
-technology}\label{sec-litreview}
-
-The diffusion of new technologies and innovations is driven by three
-different mechanisms \citep{grubler1990rise, morrill2020spatial}: (i)
-distance and proximity, (ii) urban hierarchies in the form of a
-centrifugal forces as new technologies diffuse from core to periphery,
-and (iii) over time the cumulative level of adoption follows an S-shaped
-pattern. \citet{hagerstrand1968innovation} was the first to identify
-diffusion as a geographical process. The starting point was the idea
-that diffusion is based on passing information through social networks,
-which themselves tend to be geographically bounded. Hence, he identified
-a `neighbourhood' mechanism of how information, and consequently,
-innovations and new technologies diffuse. This perspective draws
-similarities with epidemics: new technologies and innovations, just like
-pathogens, spread because of contagion and, consequently, proximity and
-exposure \citep{hivner2003facilitating}.
-
-\citet{hagerstrand1968innovation} also emphasised the role of the
-top-down hierarchy of urban systems and highlighted that new
-technologies and innovations are firstly adopted in larger cities and
-then diffuse to second tier ones. This is a sequential instead of a
-simultaneous process, which resembles the `lead-lag' spatial
-acceleration effect in market research
-\citep{bento2018time, PERES201091}. Schmidt's Law empirically
-illustrates a similar pattern \citep{grubler1990rise}. Core and highly
-agglomerated regions is where new technologies are invented and
-commercially deployed. This is where the first adopters tend to be
-based. Then, technologies spread to the rim and eventually to the
-periphery. Although adoption pace might be higher when new technologies
-finally arrive to the periphery, the saturation levels may never reach
-the ones in the core because of the lack of infrastructure or other
-necessary institutions \citep{leibowicz2016representing}.
-
-Diffusion is naturally a dynamic process and Hägerstrand recognised the
-role time plays in the diffusion of innovations and new technologies: an
-early-pioneering period, a middle fast accelerating period and a final
-saturation one \citep{morrill2020spatial}. This was further explored by
-\citet{rogers2010diffusion}, who identified five types of adopters of
-new technologies and innovations: innovators, early adopters, early
-majority, late majority and laggards. Hence, plotting the cumulative
-adoption of a new technology over time often leads to an S-shaped line
-plot \citep{grubler1990rise}.
-
-Although the above processes have been studied in innovation studies,
-the distancing of economic geographers from research questions about the
-diffusion of new technologies and innovations resulted to a knowledge
-gap regarding the explanatory power of these mechanisms at granular,
-local scales \citep{ding2010modeling}. As discussed in
-Section~\ref{sec-introduction}, such spatially explicit knowledge can
-directly inform local technology and infrastructure policies. Indeed,
-while the spatial dimension is present in the following studies, the
-geographical level of analysis is often too coarse to ascertain the role
-of the above mechanisms at the local scale. \citet{beardsell1999spatial}
-studied the evolution of the computer industry in 317 US metropolitan
-areas during the 1977-1992 period using employment data. Their analysis
-indicated that the relative size distribution holds for urban computer
-employment and also urban heterogeneity is essential in explaining this
-distribution. In a recent study, \citet{bednarz2020pulled} focused on
-wind turbines and modelled their spatial diffusion across 402 German
-regions during 1970-2015. Their key finding is that local demand than
-local supply was the main driving factor. \citet{haller2011determinants}
-employed firm panel data for Ireland during 2003-2005 and illustrated
-that when firms are located in a region or industry with high share of
-firms having a website, they are more likely to have a website too. At a
-global scale \citet{perkins2011internet} explored whether the adoption
-of previous communication technologies that is mail, telegrams and
-telephones were shaped by similar socioeconomic factors as the Internet.
-Their results indicated common patterns regarding the drivers behind the
-adoption of different communication technologies.
-\citet{feldman2015rknowledge} focused on inventions and used patent data
-to map how the the rDNA technology diffused among inventors across 366
-US metropolitan areas. \citet{ding2010modeling} modelled the spatial
-diffusion of mobile telecommunications across the highest level of
-administrative division in China (29 regions in total). Their analysis
-indicated that socioeconomic characteristics are important determinants
-of the timing, speed and the level of mobile diffusion within China.
-\citet{bakher2013diffusion} compared different growth curves as a means
-to model the diffusion of online shopping in Australia during the
-1998-2009 period, but they did not consider any spatial mechanisms. At a
-global scale \citet{PAPAGIANNIDIS2015308} modelled the diffusion of
-different web technologies and practices across countries. Using similar
-but less extensive data as the one used here, their analysis illustrated
-how the diffusion of different web technologies and practices follows an
-S-shaped pattern as well as the different diffusion rates of the
-different technologies and practices. The most detailed consideration of
-the spatial footprint of the diffusion of a new technology comes from
-\citet{lengyel2020role}, who analysed the adoption and churn of a
-Hungarian online social network. Their results were in agreement with
-the early theoretical and empirical contributions reviewed here:
-assortativity, urban scaling and distance are the key drivers of spatial
-diffusion. Based on the above, the main empirical hypothesis that is
-tested in this paper states:
-
-\begin{quote}
-H\textsubscript{1}: The three diffusion mechanisms namely distance,
-urban hierarchy and the S-shaped pattern of the cumulative level of
-adoption shaped the diffusion of the commercial Web in the UK even at
-granular geographical scales.
-\end{quote}
-
-Understanding the diffusion of a new technology in multiple, detailed
-spatial scales is of interest to economic geography as potential local
-economic benefits may occur because of concentration of early adopters
-or through technological leapfrogging. On the one hand, early adopters
-are often rewarded because of their attitudes towards new technologies
-and innovations. \citet{rogers2010diffusion} identifies them as
-`knowledgeable risk takers' and \citet{griliches1957} as `profit
-maximisers' \citep{ding2010modeling}. On the other hand, latecomers can
-adopt and benefit from new technologies that have been developed
-elsewhere without incurring the hefty initial R\&D costs
-\citep{teece2008firm}. At a global scale
-\citet{perkins2005international} explored whether the diffusion rate of
-new technologies is driven by a latecomer advantage and the engagement
-with the global economy via foreign direct investments and trade. Their
-results illustrate that indeed latecomers and developing countries
-experience diffusion of new technologies more rapidly than early
-adopters and developed countries. At a regional scale, previous research
-highlighted the long term and sustained productivity benefit of the
-early adopters of web technologies in the UK \citep{tranosuk}. Although
-the early adoption and leapfrogging literature has not paid much
-attention on cities and regions \citep{yu2018sustainability}, the
-potential local economic benefits justify the need to investigate the
-dynamics of the web diffusion in different scales in the UK. Hence, the
-second empirical hypothesis states:
-
-\begin{quote}
-H\textsubscript{2}: The diffusion of the commercial Web is shaped by
-both early adoption and latecomer effects.
-\end{quote}
-
-These two hypotheses are empirically tested in Section~\ref{sec-results}
-employing the data and methods described in the next section.
-
-\section{Data and Methods}\label{sec-datamethods}
-
-To capture the diffusion of the commercial Web, a website density metric
-is developed for two different geographical scales that are discussed
-later in this section. The main data are counts of commercial websites
-that are calculated using data from the Internet Archive\footnote{\url{https://archive.org/}.}
-and, specifically, the JISC UK Web Domain Dataset \citep{ukwebarchive}.
-The Internet Archive is one of the most complete and oldest archive of
-webpages in the world operating since 1996
-\citep{ainsworth2011much, holzmann2016dawn}. It is a web crawler, which
-discovers webpages by following the hyperlinks of every webpage it
-archives. This dataset, which is curated by the British Library,
-contains all the archived webpages from the UK country code top level
-domain (ccTLD - .uk) from the 1996--2012 period. In essence, this is a
-long list of 2.5 billion URLs of archived webpages including also the
-archival timestamp.
-
-Instead of using the whole .uk ccTLD, this paper focuses on its
-commercial subset, the .co.uk second level domain (SLD). This choice is
-aligned with the topic of the paper -- the diffusion of the Web for
-commercial functions -- and decreases the heterogeneity of the web data
-as such commercial websites have specific aims: they are used to diffuse
-information, support online transactions and share opinions
-\citep{THELWALL2000441, blazquez2018big}. Although a UK company can
-adopt a generic TLD such as .com and these cases escape the data used
-here, such omissions should not affect the validity of the results given
-the popularity of the .uk ccTLD \citep{tranosuk}: UK consumers prefer to
-visit a .uk website when they are searching for products or services
-\citep{hope}; and anecdotal evidence indicates that during the first
-half of 2000, three .co.uk domains were registered every minute
-\citep{oecd_coms}. Importantly, previous studies illustrated that .co.uk
-is the most popular UK SLD \citep{tranosuk}. Henceforward any reference
-to websites will refer to commercial -- .co.uk -- websites.
-
-The text from these webpages was scanned using a regular expression
-(regex) to identify strings of text which resemble UK postcodes and one
-fifth of them included a mention to a postcode \citep{BL2013geo}. This
-information allows the geolocation of the data and the creation of the
-LAD and OA counts. To prevent false positives, the resulted postcodes
-were matched against official postcode lookup files \citep{ons}.
-
-The data cleaning process includes an aggregation step, through which
-the archived webpages are aggregated to the parent websites. This
-website reconstruction process lend itself to the construction of
-\emph{website} instead of \emph{webpage} density metrics. Websites tend
-to represent specific organisations or entities, and, arguably, are more
-meaningful observational units than webpages, which can ignore the
-upstream dependency of the website they belong to. Based on the
-following example, all three webpages are part of the same website
-(\url{http://www.website.co.uk}). Only the overall website and not the
-nested webpages are considered for the counts as otherwise the density
-metrics would have been biased towards large, place specific, websites.
-It is important to note here that websites commonly contain place
-references, with previous research estimating this share to be around
-70\% \citep{hill2009georeferencing}.
-
-\begin{itemize}
-\item
-  \url{http://www.website.co.uk/webpage_a} B15 2TT
-\item
-  \url{http://www.website.co.uk/webpage_b} BS8 1TH
-\item
-  \url{http://www.website.co.uk/webpage_c} B15 2TT
-\end{itemize}
-
-\noindent What is challenging is that this aggregation approach, which
-has been used elsewhere \citep{tranosuk, shoreditch} may lead to
-websites with multiple postcodes. As per the above example,
-\href{http://www.website.co.uk}{www.website.co.uk} includes two unique
-postcodes: B15 2TT and BS8 1SS. The distribution of the number of
-postcodes per website for 2000 is presented in Table \ref{f2000}, which
-illustrates a wide range. At the left end of the distribution, there are
-websites anchored to a unique location (72\% of all the reconstructed
-websites in 2000), which may represent a small company with a single
-trading location. At the right end, there are websites with thousands of
-different postcodes. Considering the time period of the analysis, such
-cases can represent directories which used to be popular in the
-pre-search engines early times of the commercial Internet as well as
-real estate websites \citep{tranosuk}. Previous research which used the
-same methodology for a small subset of these data illustrates the
-robustness of the above approach: its results matched in terms of
-accuracy past rounds of research based on extensive qualitative work,
-interviews and observational studies for the same study area
-\citep{shoreditch}.
-
-The analysis is firstly conducted on websites, which only contain one
-unique postcode. As a robustness check, the analysis is replicated for
-an extended subset of websites, which include up to 10 unique postcodes
-to capture commercial websites which point to multiple locations. They
-are geolocated by equally attaching them to these locations. This
-extended sample includes 94 percent of all the archived websites in 2000
-(see Table \ref{f2000}).
-
-\begin{table}
-
-\caption{Number of unique postcodes per .co.uk website, 2000.\label{f2000}}
-\centering
-\begin{tabular}[t]{lrrr}
-\toprule
-Postcodes & F & F (\%) & cumulative F\\
-\midrule
-(0,1] & 41,596 & 0.718 & 0.718\\
-(1,2] & 6,451 & 0.111 & 0.830\\
-(2,10] & 6,163 & 0.106 & 0.936\\
-(10,100] & 2,975 & 0.051 & 0.988\\
-(100,1000] & 646 & 0.011 & 0.999\\
-\addlinespace
-(1000,10000] & 62 & 0.001 & 1.000\\
-(10000,100000] & 4 & 0.000 & 1.000\\
-\bottomrule
-\multicolumn{4}{l}{\rule{0pt}{1em}\textit{Source: } Tranos et al. 2021}\\
-\end{tabular}
-\end{table}
-
-Another data cleaning step deals with some extreme outliers. Figure
-\ref{correct} plots the website counts for the top 1000 postcodes. Some
-obvious outliers can be observed for the 2002-2006 period for a handful
-of postcodes, which can be attributed to link farms \citep{BL2013links}.
-The website counts for these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1
-6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1,000
-websites pointing to them, were replaced with predicted values based on
-a simple panel regression model with postcode fixed effects and yearly
-dummy variables. These postcodes refer to a residential area with a
-small park in South London, residential areas outside Warwick and
-Manchester, a rural area in the south border of England and Wales and
-small business parks in Crew and Derby. Hence, it is difficult to
-believe that these postcodes genuinely hosted such a large amounts of
-websites, which were later extensively declined. To put the magnitude of
-the data imputation into perspective, this process affected 6 out of the
-557,808 postcodes included in the data. As Section~\ref{sec-results}
-illustrates, this led to a small increase of the model predictive power.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{anonymised_files/figure-pdf/unnamed-chunk-2-1.pdf}
-
-}
-
-\caption{\label{correct}Yearly website counts per postcode (top 1000)
-before (top) and after (down) data imputation}
-
-\end{figure}%
-
-To empirically test the two hypotheses stated in
-Section~\ref{sec-litreview}, the website density metric was calculated
-for two different scales the Local Authority Districts (LAD) and the
-Output Areas (OA). The former is the lowest administrative unit in the
-UK and there are 374 LAD. The latter is the smallest census-based
-geographical unit and there are c.~230,000 of them in the UK. To create
-the website density metric, the yearly website counts at the LAD level
-are standardised by the number of firms in LAD to avoid biases
-associated with LAD hosting a large number of firms. Given that there is
-no such readily available statistic for the OA, the actual OA level
-counts are used. However, because of the the consistent spatial
-definition of OA (they host 40-250 households),\footnote{\url{https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies}}
-website counts in OA are interpreted as a density metric too.
-
-These website density metrics are used for the different stages of the
-analysis. Starting with the first hypothesis (H\textsubscript{1}), to
-test whether the S-curve accurately describes the cumulative adoption of
-the Web at a local scale, the following logistic function (Equation
-\ref{eq:s}) is estimated for the whole of the UK and then for each LAD
-separately:
-
-\begin{align}
-y = k /(1 + e^{-(t-t_{o})})\label{eq:s}
-\end{align}
-
-\noindent \(k\) is the asymptote or, in other words, the saturation
-level, \(b\) the overall growth rate, and \(t_{0}\) the \emph{inflection
-point} of maximum growth at \(k/2\), where the logistic function is
-symmetrical \citep{wilson201281}. The \(t_0\) of each LAD is compared
-against the \(t_0\) of the UK to delineate whether a LAD reached that
-point faster or slower than the country average. This heterogeneity is
-discussed further when testing the second hypothesis
-(H\textsubscript{2}). Importantly, an accuracy criterion was imposed and
-only LAD with \(R^2 > 0.9\) were included in this analysis. To estimate
-Equation \ref{eq:s} a self-starting logistic growth model was employed
-using the \texttt{nls} and \texttt{SSlogis} functions in \texttt{R}.
-This logistic growth model depicts the initial period of slow diffusion
-of a new technology, which is followed by the fast, exponential growth
-stage to then eventually slow down and saturate
-\citep{wilson201281, grubler1999dynamics}.
-
-Next, exploratory spatial data analysis provides some first insights of
-whether the other two drivers of spatial diffusion -- namely
-neighbourhood and hierarchy -- underpin the diffusion of the Web in the
-UK. To capture the former and following \citet{ding2010modeling} the
-Moran's I and the Local Index of Spatial Association (LISA) are
-estimated for the website density. To capture the hierarchy mechanism,
-the Gini coefficient -- an established metric of inequality -- is
-calculated. All the above are computed and plotted longitudinally both
-for the LAD and the OA.
-
-To formally test H\textsubscript{1} a modelling framework is developed
-to assess the above diffusion mechanisms. The overarching aim is to
-build a model that can test how well these three mechanisms can predict
-website density:
-
-\begin{align}
-Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S\text-curve_{t}\label{eq:rf.generic}
-\end{align}
-
-To capture the hierarchical mechanism the model include as predictors a
-one year lag of website density in London, the largest city in the UK,
-and a one year lag of the website density in the nearest city. Both
-variables clearly illustrate the urban hierarchy in the UK. Especially
-for the OA models, a third predictor is added to capture hierarchy at a
-much more granular scale: a one year lag of the website density in the
-the nearest retail centre \footnote{The retail centres data have been
-  obtained by the Consumer Data Research Centre, an ESRC data
-  investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1:
-  \url{https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators}.}.
-As the literature illustrates, retail centres are key components of
-urban systems \citep{dennis2002central, jones2021reframing}. Due to the
-small sizes of the retail centres, the latter is only relevant for and
-included in the OA models and not in the LAD models. In addition, the
-model includes the distance to London, the nearest city and the nearest
-retail centre. The underlying logic is that the level of website
-adoption in a spatial unit depends on the level of the adoption in
-places further up in the urban hierarchy the previous year. The
-inclusion of the distance variables incorporates spatial structure into
-the hierarchy argument. To depict the neighbourhood mechanism, the
-website density of the neighbouring spatial units in the previous year
-is employed. This is calculated using rook continuity. Again, the
-underpinning rationale is that the level of web adoption within a
-spatial unit depends on the level of web adoption in the neighbouring
-spatial units the year before. This represents the `hitting nearby
-locations first' argument. Therefore, the spatial and temporal lag of
-the website density in LAD and OA is calculated. Lastly, the temporal
-pattern of the cumulative adoption is captured by a time trend variable.
-Hence, the model will follow the following generic form (Eq.
-\ref{eq:model}):
-
-\begin{align} \label{eq:model}
-Website\,Density_{t} \sim Distance\,London +
-Website\,density\,London_{t-1} +\notag\\
-Distance\,Nearest\,City +
-Website\,density\,Nearest\,City_{t-1} +\notag\\
-Distance\,Nearest\,Retail_{i} +
-Website\,density\,Nearest\,Retail_{t-1} +\notag\\
-W*\, Website\,density_{t-1} +\notag\\ 
-year_{t}
-\end{align}
-
-Specifically, four different models are estimated. Firstly, all the data
-points for the OA and LAD are utilised in order to build two models and
-assess their capacity to predict the adoption of the Web at the two
-different scales. These two models will reveal the predictive capacity
-of the diffusion mechanisms and also demonstrate how the importance of
-such variables changes across different scales. The next two sets of
-models are again trained at these two scales: OA and LAD. However,
-instead of using all the data points, the OA and the LAD from one of the
-twelve UK regions are held out for the model training. Then, the trained
-model is used to predict website density for the OA or the LAD of the
-held-out region. This process takes place recursively for all twelve UK
-regions. The regional differences in the predictive capacity of the
-different samples will reveal how dissimilar these spatial process are
-across regions and, importantly, at different scales.
-
-Equation \ref{eq:model} is estimated using Random Forest (RF). This is a
-popular ML algorithm for both regression and classification problems
-\citep{biau2012analysis}. It was introduced by \citet{breiman2001random}
-and has become a go-to data science tool. RF can effectively handle
-skewed distributions and outliers, model non-linear relationships,
-require minimal hyperparameter tuning, exhibit low sensitivity to these
-parameters, and have relatively short training times
-\citep{Caruana2008, liaw2002classification, yan2020using}. These
-attributes match well with the website density data characteristics
-including skewness especially for the OA. Also, the large data size (c.
-230k data points for each of the 17 years) calls for fast training
-times. Importantly, RF predictions tend to be more accurate than those
-from single regression trees and outperform Ordinary Least Squares in
-out-of-sample predictions, even with moderate-sized training data and a
-small number of predictors
-\citep{mullainathan2017machine, athey2019machine, sulaiman2011intelligent, pourebrahim2019trip, biau2012analysis}.
-
-RF is a tree-based ensemble learning algorithm
-\citep{breiman2001random}. It begins by generating random samples of the
-training data, which are then used to grow regression trees to predict
-the dependent variable. Data points and predictors are randomly sampled
-for the different trees. The trees are trained in parallel using their
-own bootstrapped samples of the training data. A crucial feature of RF
-is their ability not to overfit, meaning they can generalize well to
-unseen test data. While each tree may overfit individually, the ensemble
-of trees does not because the errors of individual trees are averaged,
-reducing the overall variance and preventing overfitting
-\citep{last2002improving}. For regression problems, RF predictions are
-made by averaging the predictions of all decision trees.
-
-RF have been widely employed to address regression research problems.
-\citet{pourebrahim2019trip} combined a spatial interaction modelling
-framework with ML algorithms including RF to predict commuting flows in
-New York City. \citet{sinha2019assessing} advocated for adopting spatial
-ensemble learning approaches, such as RF, to model spatial data with
-high autocorrelation and heterogeneity. \citet{creditspatial} predicted
-employment density in Los Angeles using spatially explicit RF.
-\citet{guns2014recommending} used RF to build a recommendation system
-for research collaborations. \citet{ren2019predicting} trained RF to
-predict the socio-economic status of cities using various online and
-mobility predictors. \citet{tranos2023using} utilised hyperlinks data
-and RF to make out-of-sample predictions of interregional trade.
-\citet{zhou2023geography} employed such a framework to assess whether
-key predictors of obesity differ across English cities.
-
-It needs to be highlighted here that the cross validation (CV) for all
-models is spatially and temporally sensitive. Instead of using 10 random
-and space- and time-agnostic folds, the \texttt{CAST} package is
-employed which allows holding back data points from specific years and
-spatial units (regions) and use them for testing in order to estimate
-the model performance \citep{meyer2018improving}.
-
-To assess the predictive power of the model, three broadly utilised
-metrics are employed: the coefficient of determination (\(R^2\)), mean
-absolute error (MAE) and root mean square error (RMSE):
-
-\begin{align}
-R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
-\end{align} \begin{align}
-MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
-\end{align} \begin{align}
-RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
-\end{align}
-
-\noindent \(y_{k}\) is the \(k^{th}\) observation of the dataset, which
-consists of \(N\) observations in total. \(\hat{y_{k}}\) is the
-\(k_{th}\) predicted value for the dependent variable and
-\(\overline{y_{k}}\) is the average value of \(y\). The last two metrics
-are expressed in the same units as the dependent variable -- websites
-per firm for the LAD modes and the number of websites for the OA models
--- while the first one is the coefficient of determination between the
-observed and the predicted values of website adoption. Regarding
-\(MAE\), it is the absolute difference between the observed and the
-predicted website adoption. While \(MAE\) does not penalise for large
-errors, \(RSME\) does, as it is proportional to the squared difference
-between the observed and the predicted trade flows. Hence larger errors
-weigh more for \(RMSE\) \citep{pontius2008components}.
-
-An alternative specification could model the growth of website density.
-However, such a strategy is problematic because of the large frequency
-of OA without websites, which is to be expected given their small size
-and high granularity. Still, as a robustness text, the results of a
-growth model for LAD are presented in the Appendix.
-
-Lastly, to test the second hypothesis (H\textsubscript{2}) about the
-early adoption and the latecomer effects, a different research framework
-is developed. Firstly, the S-curve and the estimated \(t_0\) for every
-LAD are employed and mapped in order to depict the spatial heterogeneity
-and specific patterns of the speed of the diffusion of the Web. Then,
-the volatility and the stability of the diffusion across the different
-LAD is analysed using tools from the complex systems domain
-\citep{Batty2009} namely changes in rankings of LAD based on website
-density. Both the S-curve estimation and the volatility analysis focus
-only on LAD as the very large number of OA would have made such analysis
-difficult to visualise and interpret.
-
-\section{Results}\label{sec-results}
-
-\subsection{The three diffusion mechanisms at granular geographical
-scales}\label{the-three-diffusion-mechanisms-at-granular-geographical-scales}
-
-Figure \ref{s_uk} plots the cumulative adoption of the Web in the UK as
-a whole during the 1996-2012 period. In line with the theoretical
-expectations discussed in the previous section, the cumulative adoption
-of the Web is well represented by a curve that resembles the letter S.
-The vertical line in the beginning of 2003 (2003.158) illustrates the
-point when the modelled cumulative adoption was equal to 50\% of the
-maximum. This \(t_0\) point is known as the \emph{inflection} point and
-signals the maximum adoption speed. The results remain unchanged when
-the extended subset of websites, which include up to 10 unique postcodes
-is used (see Figure \ref{s_uk10} in the Appendix). Importantly though,
-the extended subset of websites indicates a saturation level close to
-100\% at the end of the study period.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/s_uk_per_firm.png}
-
-}
-
-\caption{\label{s_uk}Cumulative adoption of the Web, UK}
-
-\end{figure}%
-
-What escaped from previous studies was εexploring whether such a pattern
-also describes cumulative adoption of new technologies at sub-national
-scales. To test this, the same logistic function is estimated for each
-of the 374 LAD. The detailed results are presented in Table
-\ref{table.s.lads} in the Appendix. The key finding is that for 342 LAD
-(91\% of all LAD) the \(R^2 > 0.9\) for the logistic function, which
-indicates that even at local scale the cumulative adoption of the Web is
-characterised by a pattern similar to the one observed for countries.
-Graphically, all these LAD curves are plotted in Figure \ref{s_lad_uk}.
-The different colour lines illustrate whether a LAD reached \(t_0\)
-faster or slower than the UK in total. This heterogeneity is further
-discussed later in this section.
-
-This is the first time that such cumulative adoption curves have been
-estimated for areal units as small as the UK LAD. Understanding and
-modelling the dynamics of technological diffusion and, ultimately, being
-able to predict diffusion rates at such small local scales can directly
-support the deployment of infrastructural networks (e.g.~5G) and local
-technology policies. The above are in line with previous studies
-exploring the diffusion of web technologies
-\citep{PAPAGIANNIDIS2015308}, social media \citep{lengyel2020role} and
-online shopping \citep{bakher2013diffusion}.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/lad_per_firm_all.png}
-
-}
-
-\caption{\label{s_lad_uk}Cumulative adoption of the Web, LAD}
-
-\end{figure}%
-
-To assess the second diffusion mechanisms, Figures \ref{morani} and
-\ref{lisa} offer a first insight into whether a neighbourhood or, in
-other words, a distance mechanism underpins the diffusion of the Web in
-the UK as they present the Moran's I and the LISA maps of website
-density respectively for LAD and OA. Starting from the former, spatial
-autocorrelation was higher in the beginning of the study period and then
-after 2000 dropped slightly and stabilised around 0.2. This reflects the
-early concentration of high website density around London, which over
-time diffused as high website density clusters can be seen in other
-parts of the country away from London (Figure \ref{lisa}). An almost
-reverse pattern can be observed for OA. At the beginning of the study
-period Moran's I was around 0.5 and it plateaued after 2000 around 0.2.
-Because of the very small size of OA, at the early stages of the
-diffusion of web technologies their adoption was spatially scattered.
-This is reflected in the lack of any significant clusters in 1996
-(Figure \ref{lisa}). Eventually, as the adoption rate increased, more
-such clusters of high website density were formed and this is reflected
-both in the Moran's I and the LISA maps in Figures \ref{morani} and
-\ref{lisa}. A similar pattern is observed for the extended dataset (see
-Figures\ref{morani10} and \ref{lisa10}). In this case, the magnitude of
-spatial autocorrelation is slightly higher illustrating a stronger
-spatial clustering mechanism. All in all, the exploratory spatial data
-analysis advocates towards an underpinning neighbourhood mechanism,
-which is present at different scales. Interestingly, the magnitude of
-Moran's I is similar to the ones revealed by \citet{ding2010modeling}
-about mobile phone adoption in China albeit the much more detailed
-spatial scale of this paper.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/morani-1.pdf}
-
-}
-
-\caption{\label{morani}Website density Moran's I}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-5-1.pdf}
-
-}
-
-\caption{\label{lisa}\centering Website density LISA maps}
-
-\end{figure}%
-
-To illustrate whether a hierarchical process also underpins the
-diffusion of web technologies -- the third diffusion mechanism that is
-tested here -- the Gini coefficient is calculated yearly both for LAD
-and OA. As a metric of inequality, the Gini coefficient demonstrates
-whether website density is concentrated in a small number of LAD or OA,
-or whether it is more equally spread across the country. Both scales of
-analysis in Figure \ref{gini} illustrate the same picture. At the
-beginning of the commercial Internet, website density was extremely
-unequal, or, in other words, only a few places had websites associated
-with them. Inequality dropped and plateaued after 2000 for both scales.
-This is illustrative of a hierarchical diffusion mechanism that led over
-time to a more equal spread. Interestingly, the year 2000 is again a
-period of change for this diffusion mechanism as it was for the
-neighbourhood process. There is a substantial difference between the
-Gini coefficient magnitude for LAD and OA, but this is expected as the
-very small size of OA equates to a lot of polygons without any websites
-pointing to them -- for example residential OA.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/gini-1.pdf}
-
-}
-
-\caption{\label{gini}Website density Gini coefficient}
-
-\end{figure}%
-
-After discussing the descriptive findings, the validity of the above
-diffusion mechanisms at the local scale is assessed with a predictive
-modelling framework described in Section~\ref{sec-datamethods}. The
-reader is reminded that four models are estimated: for the OA and LAD,
-firstly all data points are employed to assess the capacity of the three
-mechanisms to to predict the adoption of the Web at these two scales;
-then, the LAD/OA from one of the twelve regions are recursively held out
-to assess the spatial heterogeneity of these mechanisms across the
-different UK regions.
-
-Table~\ref{tbl-model-metrics} presents the model performance for the
-first set of models, for which all data points are employed for training
-and testing via CV. The first one is trained and tested on 374 LAD and
-the second on 232,296 OA, both for the 16 year period (1997-2012). The
-results are remarkably good considering that the are the outcome of
-space and time sensitive CV, so the the model does not suffer from
-overfitting. At the LAD level the model predicts 83\% of the variation
-of website density. Both error metrics indicate that the model error is
-fraction of a website per firm (0.020 and 0.029). At the OA the \(R^2\)
-drops down to 33\%. Considering its granularity, this is still a
-remarkable performance. To contextualise it, the model results in a
-\(MAE\) equal to 1.2 websites for areas small enough to host 40-250
-households. Because of the small size of the spatial units, the
-distribution is highly skewed and a significant part of them is not
-linked to any websites. In 1997 only 1\% of the UK OA were associated
-with at least one website. This should not come as a surprise given that
-this was the very beginning of the commercial Internet and any
-activities with a digital footprint were concentrated in a handful of
-areas, as illustrated in Figure \ref{lisa}. At the end of the study
-period almost half of the UK OA were not associated with a website.
-Again, given the granularity of the data this was expected.
-
-\begin{longtable}[]{@{}llll@{}}
-\caption{Model metrics}\label{tbl-model-metrics}\tabularnewline
-\toprule\noalign{}
-& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Local Authorities & 0.029 & 0.824 & 0.020 \\
-Output Areas & 3.218 & 0.333 & 1.233 \\
-\end{longtable}
-
-The results in Table~\ref{tbl-model-metrics} are robust against
-different specifications. As noted earlier, the data imputation process
-resulted in a modest improvement of the predictive capacity of the
-model. As per Table~\ref{tbl-model-metrics-no-correction}, all accuracy
-metrics are only slightly worst when the data imputation is not applied.
-As expected, the improvement is higher for the more detailed scale of
-analysis. Then, Table~\ref{tbl-model-metrics-10} presents the model
-results for the extended dataset, which includes websites with up to 10
-unique postcodes. While the results for LAD are only slightly less
-accurate, the \(R^{2}\) of the OA model is much higher in comparison to
-the base results in Table~\ref{tbl-model-metrics}. In addition,
-Table~\ref{tbl-model-growth-metrics} in the Appendix presents the
-accuracy results of a LAD model, the dependent variable of which is the
-yearly growth rate of website density. Even for such a specification
-(see Equation \ref{model.growth} in the Appendix) the predictive
-capacity of the model is exceptionally high (\(R^2 = 0.5\)). All in all,
-the modelling exercise revealed how well the three diffusion mechanisms
-predict the diffusion of the Web as a new technology. Importantly, this
-predictive power is highly robust against against different scales and
-delineations of the diffusion of the web.
-
-\begin{longtable}[]{@{}llll@{}}
-\caption{Model metrics without data
-imputation}\label{tbl-model-metrics-no-correction}\tabularnewline
-\toprule\noalign{}
-& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Local Authorities & 0.032 & 0.781 & 0.021 \\
-Output Areas & 4.644 & 0.244 & 1.237 \\
-\end{longtable}
-
-\begin{longtable}[]{@{}llll@{}}
-\caption{Model metrics for the extended
-dataset}\label{tbl-model-metrics-10}\tabularnewline
-\toprule\noalign{}
-& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Local Authorities & 0.190 & 0.749 & 0.129 \\
-Output Areas & 21.591 & 0.455 & 6.845 \\
-\end{longtable}
-
-Figure \ref{var.imp} plots the importance of the different predictors.
-When the focus is on LAD, the website density in the neighbouring LAD,
-in the nearest city and in London the year before are the most important
-predictors. They are followed by the yearly trend, while the spatial
-configuration as reflected in distances to London or to the nearest city
-only plays a minor role. This can be attributed to the relatively coarse
-spatial scale of LAD. Nevertheless, all previously discussed spatial
-processes are at play in the diffusion of the Web at the LAD level: the
-first two predictors depict the hierarchical mechanism, the spatial and
-temporal lag of website density the neighbourhood mechanism, and the
-yearly trend the time-sensitive cumulative adoption pattern.
-
-At the much more granular scale of OA, the picture is almost reversed.
-The two most important predictors are the distance to London and to the
-nearest city followed by the spatial and temporal lag of website density
-and the distance to the nearest retail centre. They still depict a
-hierarchical mechanism, but proximity to the different population
-centres is more important than their lagged web densities in predicting
-website diffusion. The neighbourhood mechanism is still strong but less
-important at this scale. What is interesting is the almost negligible
-role of the yearly trend and London's website density. While the former
-probably illustrates the large heterogeneity in how the Web has been
-adopted at this very fine scale, the latter highlights that the
-importance of past web adoption rates in large population centres is
-surpassed by proximity to them and spatial configuration at this scale.
-
-The above pattern of variable importance is largely in accordance with
-the models estimated for the extended dataset. A comparison with Figure
-\ref{var.imp10} illustrates that the main difference is the importance
-of the neighbourhood mechanism for the OA as the space and time lag of
-the website density is by far the most important predictor when the
-extended dataset is employed. For the LAD, the importance of the
-distance to the nearest city increases in favour to its website density
-the previous year. But still, the variable importance patterns for the
-two datasets are not dissimilar.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{anonymised_files/figure-pdf/varimp-1.pdf}
-
-}
-
-\caption{\label{var.imp}Variable importance}
-
-\end{figure}%
-
-Table \ref{table.regions} presents the results of the recursive hold out
-models and allows to (i) highlight the regional heterogeneity of the
-spatial processes behind the diffusion of the Web, and (ii) compare this
-heterogeneity across the two different scales of analysis. It
-illustrates both regional similarities and differences. At the top of
-Table \ref{table.regions} the regions of South East and the West
-Midlands have the highest similarity with the rest of the country. In
-other words, the web density for the LAD of these regions can be very
-well predicted by models trained on data from the LAD from the rest of
-the country and, therefore, the spatial diffusion mechanisms for these
-regions represent quite well the UK's average. This should not come as a
-surprise as the spatial structure of both regions resembles the broader
-UK: both regions contain some highly urbanised LAD (South East is well
-integrated with the Greater London and West Midlands is the home of the
-city of Birmingham), but also smaller commuter towns as well as more
-rural settlements. At the other end of the spectrum, the spatial
-diffusion mechanisms for Northern Ireland, Scotland and North West
-diverge the most from the country average. Rurality and remoteness seem
-to be common attributes for these regions. This hierarchy though is not
-consistent across scales. For the more granular OA models, the spatial
-diffusion mechanisms for South West, Yorkshire and The Humber and North
-West regions are closer to the country average. Interestingly, these
-regions are well-known tourist destinations (Lake District, Peak
-District and Cornwall-Devon) and the exploratory analysis revealed
-clusters of high web density only visible at the OA scale (Figure
-\ref{lisa}). Then, Scotland, Northern Ireland and London have the lowest
-similarity with the country average. Again, rurality, peripherality, but
-also the highest level or urbanisation make the spatial diffusion
-mechanisms of these regions more distinct.
-
-\begin{longtable}[]{@{}lrrrr@{}}
-\caption{Regional differences\label{table.regions}}\tabularnewline
-\toprule\noalign{}
-Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-South East & 0.952 & 1 & 0.122 & 6 \\
-West Midlands & 0.926 & 2 & 0.126 & 4 \\
-Wales & 0.909 & 3 & 0.122 & 8 \\
-Yorkshire and The Humber & 0.908 & 4 & 0.160 & 2 \\
-East Midlands & 0.904 & 5 & 0.122 & 7 \\
-South West & 0.894 & 6 & 0.185 & 1 \\
-North East & 0.887 & 7 & 0.124 & 5 \\
-London & 0.880 & 8 & 0.090 & 11 \\
-East of England & 0.876 & 9 & 0.106 & 9 \\
-North West & 0.837 & 10 & 0.142 & 3 \\
-Scotland & 0.774 & 11 & 0.059 & 12 \\
-Nortern Ireland & 0.590 & 12 & 0.101 & 10 \\
-\end{longtable}
-
-Comparing the above with Table \ref{table.regions.10} in the Appendix,
-which presents the results for the the extended dataset with websites
-with up to 10 unique postcodes, the results do not differentiate much,
-signalling the robustness of the analysis. The most salient difference
-has to do with Northern Ireland. According to the extended sample at the
-granular OA level, Northern Ireland's spatial diffusion mechanisms are
-closer to the UK's average than the results based on the websites with a
-unique postcode. While the latter are websites and, consequently,
-economic activities solely based in Northern Ireland, websites with
-multiple postcodes also include websites with postcodes in Britain and,
-therefore, can be characterised by slightly different spatial diffusion
-processes.
-
-\subsection{Early adoption and
-latecomers}\label{early-adoption-and-latecomers}
-
-To test the second hypothesis (H\textsubscript{2}), the spatial
-heterogeneity of Web adoption speeds is firstly analysed. As highlighted
-in the literature \citep{risk_perceptions}, different agents have
-different perceptions about and levels of acceptance of the risks and
-the potential economic returns associated with the adoption of new
-technologies -- see for instance the seminal work of
-\citet{venkatesh2000theoretical}. To reveal such patterns, Figure
-\ref{s_map} maps whether LAD reached their inflection point earlier or
-later than the UK in total and based on this are labelled as \emph{fast}
-or \emph{slow} accordingly. Inflection points have been used in the
-relevant literature to compare diffusion processes across different
-settings and technologies \citep{cherp2021national, woo2017exploring} as
-they can enable more insightful comparisons than, for example, the
-average growth over the 1996-2012 study period.
-
-The map illustrates a clear high concentration of fast adopting LAD in
-the South and East of England, as expected. There are also some
-concentrations in Wales, Scotland and the North West of England. At a
-more detailed level, there are some expected cases of LAD with relevant
-industrial backgrounds delineated as fast: the City of London, a
-world-renowned cluster of finance industries \citep{cook2007role}, and
-Reading, a town with high-tech service industries in proximity to London
-and its main airport, Heathrow \citep{england2005polynet}. However,
-there are also LAD which were expected to appear as fast -- e.g.~Hackney
-in central London and Bristol, a well-established creative cluster
-\citep{oatley1999cultural, bassett2002cultural} -- but are delineated as
-slow. Similar to \citet{beardsell1999spatial}, who analysed the spatial
-evolution of computer industry employment across 317 US metro areas,
-spatial heterogeneity is evident in the diffusion speed when the scale
-of analysis is as detailed as the LAD.
-
-Importantly, there is obvious spatial clustering of fast and slow LAD.
-Out of the ten fastest LAD based on when they reached their inflection
-point (see detailed Table \ref{table.s.lads} in the Appendix), nine were
-located in the South East of England and London and one in Cambridge.
-The above observations remain almost unchanged when the analysis
-includes websites with up to 10 unique postcodes -- see Figure
-\ref{s_map10} in the Appendix -- demonstrating the robustness of the
-analysis. Given (i) the expectation that early adopters of new
-technologies are often rewarded
-\citep{rogers2010diffusion, griliches1957, ding2010modeling}, and (ii)
-the previous findings illustrating the long term and sustained
-productivity benefit for the early adopters of Web in the UK
-\citep{tranosuk}, the observed spatial clustering of LAD with fast
-diffusion rates can further reinforce existing and rooted patterns of
-unequal economic productivity in the UK \citep{mccann2022productivity}.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/speed_map.png}
-
-}
-
-\caption{\label{s_map}LAD Web diffusion speed, 1996-2012}
-
-\end{figure}%
-
-To reveal patterns of early adoption and leapfrogging, the stability and
-volatility of the above patterns is analysed. Figure \ref{rank} plots
-the relative ranking of LAD based on website density at the beginning
-and the end of the study period. To decrease noise, the average ranking
-of 1996-1998 and 2010-2012 is plotted instead of the individual years.
-Each line represents a single LAD and the colours depict the
-5\textsuperscript{th} and 95\textsuperscript{th} percentile of the
-absolute difference of the LAD ranking between 1996-1998 and 2010-2012.
-In other words, the 5\textsuperscript{th} percentile includes the LAD,
-the relative position of which remained unchanged between the beginning
-and the end of the study period, both at the top and at the bottom of
-the hierarchy (small absolute difference in ranking). The 95th
-percentile includes LAD that either increased or decreased a lot their
-relative positions over time (high absolute difference in ranking).
-Thus, Figure \ref{rank} reveals cases of \emph{leapfrogging} since the
-LAD at the top-right part of the figure managed to jump at the top of
-the hierarchy at the end of the study period.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/ranks/web_per_firm2000_2012_only0595_av.png}
-
-}
-
-\caption{\label{rank}Dynamics of Web diffusion based on LAD rankings}
-
-\end{figure}%
-
-To further analyse the above, Table \ref{table.late} juxtaposes the
-volatility of the LAD Web diffusion against the early conditions as it
-reports the \(t_0\) estimate (inflection point) for the leapfrogging LAD
-identified in Figure \ref{rank}. Indeed, more than half of the LAD that
-drastically increased their relative position at the end of the study
-period reached the inflection point \(t_0\) later than the country in
-total and, therefore, the Web diffusion speed has been relatively slow
-for them. The fact that these LAD experienced relatively faster
-diffusion rates at the end of the study period, which is aligned with
-the maturity period of the web, can lead to potential positive economic
-gains for these LAD. There is extensive literature describing the
-benefits of such technological leapfrogging and the underpinning
-argument is that latecomers can adopt and benefit from new technologies
-that have been developed elsewhere without incurring the hefty initial
-R\&D costs \citep{teece2008firm}.
-
-Based on Figure \ref{rank} and Table \ref{table.late}, the opportunity
-to catch up that \citet{perkins2005international} illustrated at the
-country level, is also visible at the much more detailed LAD scale.
-Although the relevant literature did not pay attention to cities and
-regions \citep{yu2018sustainability}, the above finding advocates
-towards further exploring leapfrogging effects at such a granular scale.
-As previously, this is consistent with including websites with up to 10
-postcodes (see Figure \ref{rank10}).
-
-All in all, the diffusion of the Web at the detailed scale of LAD is
-characterised by spatial heterogeneity, well-established spatial
-patterns and both stability and volatility. Although the relevant
-literature neglected local scales, the above findings illustrated that
-such local dynamics exists and worth further consideration and
-investigation especially regarding potential local economic effects.
-
-\begin{longtable}[]{@{}lcc@{}}
-\caption{Late adopters\label{table.late}}\tabularnewline
-\toprule\noalign{}
-LAD & \(t_{0}\) estimate & Diffusion speed \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-LAD & \(t_{0}\) estimate & Diffusion speed \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Halton & 2007.1 & slow \\
-Blackpool & 2002.4 & fast \\
-Rutland & 2006.5 & slow \\
-Torbay & 2002.2 & fast \\
-Isles of Scilly & 2012.2 & slow \\
-North Devon & 2002.9 & fast \\
-Gosport & 2003.8 & slow \\
-Thanet & 2004.0 & slow \\
-Na h-Eileanan Siar & NA & NA \\
-Conwy & 2002.6 & fast \\
-\end{longtable}
-
-\section{Conclusions}\label{sec-conclusions}
-
-The paper offers an economic geography study of the diffusion of a new
-technology -- the Web. Contrary to previous studies, it does so by
-adopting (i) a high level of spatial granularity, and (ii) a long study
-period that captures the very early stages of this technology until its
-maturity. This is the first time that the importance of granular spatial
-mechanisms for the diffusion of a new technology is exposed.
-Importantly, the empirical framework and the data employed here allows
-to observe the active engagement with the Web that is creating and
-maintaining a commercial website instead of the more passive action of
-browsing webpages and accessing the Internet.
-
-Specifically, two hypotheses are tested. Firstly, whether the three
-diffusion mechanisms namely distance, urban hierarchy and the S-shaped
-pattern of the cumulative level of adoption shaped the diffusion of the
-commercial Web in the UK at granular geographical scales. The analysis
-firstly illustrates how well the S-pattern of cumulative adoption fits
-at the local scale and offers some descriptive insights about the
-importance of the distance and hierarchical diffusion mechanisms. Then,
-the modelling exercise reveals the remarkably good predictive capacity
-of these mechanisms at granular geographical scales. The modelling
-framework exposes that the relative importance of these spatial
-mechanisms differ across different scales. It also reveals spatial
-heterogeneity. While for some UK regions the importance of the spatial
-diffusion mechanisms are very similar to the rest of the country, other
-regions diverge significantly from the country average.
-
-To do the above, a robust and novel ML modelling framework was designed
-to predict website diffusion over space and time using as predictors the
-spatial diffusion mechanisms. Space- and time-sensitive CV was designed
-for the model to avoid overfitting. Importantly, the results are robust
-against different specifications and subsets of the data. This is the
-first time that the power of these mechanisms in predicting the
-diffusion of a new technology was empirically tested, importantly, at
-very fine scales.
-
-The second hypothesis assessed whether the diffusion of the web is
-shaped by both early adoption and latecomer effects. On the one hand,
-the analysis highlighted the heterogeneity of adoption rates between LAD
-as well as the spatial clustering of LAD with similar adoption rates.
-Importantly, the observed spatial clustering of LAD with fast diffusion
-rates can reinforce existing and rooted patterns of unequal economic
-productivity in the UK \citep{mccann2022productivity}. Previous research
-had already indicated the long-term productivity gains of the early
-adoption of such technologies \citep{tranosuk}. On the other hand, the
-analysis also exposed instances of remarkable stability as well as
-volatility of the Web adoption, with the latter being indicative of
-leapfrogging. While the literature had discussed the benefits of
-technological leapfrogging at the national scale
-\citep{teece2008firm, perkins2005international}, local scales had
-escaped such studies \citep{yu2018sustainability}. Thus the leapfrogging
-evidence presented here advocate towards further exploring such effects
-at granular scales.
-
-The paper offers a number of different contributions. Firstly, it
-assesses, for the first time, the role of these spatial diffusion
-mechanisms at such granular spatial scales. While previous studies had
-highlighted these mechanisms
-\citep{hagerstrand1968innovation, rogers2010diffusion, grubler1990rise},
-their role had never been tested before at such granular scale. The
-findings talk directly to this literature
-\citep[e.g.][]{beardsell1999spatial, haller2011determinants, lengyel2020role}.
-More broadly, the paper offers a robust methodological framework, which
-utilises novel data and modern ML algorithms and techniques to model the
-diffusion of a new technology across space and time. It directly
-addresses the data problem which prevented economic geographers in
-engaging with diffusion studies
-\citep{iso2005, kemeny2011international, zook2022mapping}. Thus, the
-paper seeks to renew economic geography's interest in the diffusion of
-new technologies. Current data science advances including the
-availability of granular enough data and adequate methods can enable
-economic geographers to reveal interesting spatial patterns of diffusion
-of new technologies as this paper has demonstrated. This is important as
-(i) various stakeholders are interested in predicting localised demand
-for new technologies for planning purposes, for example 5G networks
-\citep{leibowicz2016representing, meade2021modelling}, and (ii) the
-adoption of such digital technologies is associated with local economic
-gains
-\citep{solow1957technical, aghion1990model, kemeny2011international, tranosuk, capello2024nexus}.
-Importantly, although the geographical focus of this paper is the UK,
-the empirical framework developed here is easily transferable to other
-geographical contexts.
-
-\setcounter{section}{0}
-\renewcommand{\thesection}{\Alph{section}}
-\setcounter{table}{0}
-\renewcommand{\thetable}{A\arabic{table}}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{A\arabic{figure}}
-
-\section{Appendix}\label{appendix}
-
-\counterwithin{figure}{section}
-\counterwithin{table}{section}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/s_uk_per_firm_10.png}
-
-}
-
-\caption{\label{s_uk10}Cumulative adoption of the Web, UK; up to 10
-postcodes per website}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/speed_map_10.png}
-
-}
-
-\caption{\label{s_map10}LAD Web diffusion speed, 1996-2012; up to 10
-postcodes per website}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/ranks/web_per_firm2000_2012_only0595_av_10.png}
-
-}
-
-\caption{\label{rank10}Dynamics of Web diffusion based on LAD rankings;
-up to 10 postcodes per website}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/morani10-1.pdf}
-
-}
-
-\caption{\label{morani10}Website density Moran's I; up to 10 postcodes
-per website}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-11-1.pdf}
-
-}
-
-\caption{\label{lisa10}\centering Website density LISA maps; up to 10
-postcodes per website}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/gini10-1.pdf}
-
-}
-
-\caption{\label{gini10}Website density Gini coefficient; up to 10
-postcodes per website}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{anonymised_files/figure-pdf/varimp10-1.pdf}
-
-}
-
-\caption{\label{var.imp10}Variable importance; up to 10 postcodes per
-website}
-
-\end{figure}%
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.3186}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2212}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1416}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1062}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0619}}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1504}}@{}}
-\caption{S-curve estiamtes for LAD\label{table.s.lads}}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-LAD
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Region
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-\(t_0\) estimate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Std. error
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-\(R^2\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Diffusion speed
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-LAD
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Region
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-\(t_0\) estimate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Std. error
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-\(R^2\)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Diffusion speed
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Horsham & South East & 2001.423 & 0.254 & 0.950 & fast \\
-Fareham & South East & 2001.443 & 0.346 & 0.925 & fast \\
-Kingston upon Thames & London & 2001.513 & 0.354 & 0.933 & fast \\
-Kensington and Chelsea & London & 2001.581 & 0.341 & 0.943 & fast \\
-Runnymede & South East & 2001.582 & 0.266 & 0.943 & fast \\
-Bracknell Forest & South East & 2001.586 & 0.324 & 0.928 & fast \\
-Elmbridge & South East & 2001.656 & 0.333 & 0.932 & fast \\
-Reigate and Banstead & South East & 2001.665 & 0.253 & 0.964 & fast \\
-South Cambridgeshire & East of England & 2001.680 & 0.433 & 0.905 &
-fast \\
-Walsall & West Midlands & 2001.715 & 0.413 & 0.901 & fast \\
-Surrey Heath & South East & 2001.718 & 0.333 & 0.933 & fast \\
-Woking & South East & 2001.759 & 0.292 & 0.951 & fast \\
-South Norfolk & East of England & 2001.874 & 0.358 & 0.931 & fast \\
-City of London & London & 2001.874 & 0.499 & 0.930 & fast \\
-Wokingham & South East & 2001.891 & 0.287 & 0.958 & fast \\
-Reading & South East & 2001.900 & 0.318 & 0.948 & fast \\
-Sevenoaks & South East & 2001.904 & 0.407 & 0.926 & fast \\
-Huntingdonshire & East of England & 2001.912 & 0.402 & 0.925 & fast \\
-Pendle & North West & 2001.924 & 0.407 & 0.923 & fast \\
-St Albans & East of England & 2001.933 & 0.396 & 0.927 & fast \\
-Perth and Kinross & NA & 2001.943 & 0.228 & 0.970 & fast \\
-Bromley & London & 2001.945 & 0.396 & 0.919 & fast \\
-Rushmoor & South East & 2001.962 & 0.518 & 0.912 & fast \\
-Powys & NA & 2001.969 & 0.278 & 0.963 & fast \\
-Swindon & South West & 2001.996 & 0.390 & 0.916 & fast \\
-Chelmsford & East of England & 2002.008 & 0.545 & 0.911 & fast \\
-Crawley & South East & 2002.008 & 0.379 & 0.949 & fast \\
-Dumfries and Galloway & NA & 2002.012 & 0.251 & 0.967 & fast \\
-Inverclyde & NA & 2002.030 & 0.330 & 0.915 & fast \\
-North Hertfordshire & East of England & 2002.031 & 0.475 & 0.920 &
-fast \\
-Moray & NA & 2002.050 & 0.358 & 0.947 & fast \\
-Orkney Islands & NA & 2002.070 & 0.283 & 0.958 & fast \\
-Fife & NA & 2002.070 & 0.363 & 0.945 & fast \\
-Mole Valley & South East & 2002.071 & 0.460 & 0.925 & fast \\
-Guildford & South East & 2002.113 & 0.413 & 0.941 & fast \\
-Buckinghamshire & South East & 2002.114 & 0.390 & 0.942 & fast \\
-Cotswold & South West & 2002.116 & 0.375 & 0.946 & fast \\
-Shetland Islands & NA & 2002.120 & 0.332 & 0.949 & fast \\
-South Oxfordshire & South East & 2002.146 & 0.367 & 0.950 & fast \\
-Watford & East of England & 2002.157 & 0.451 & 0.929 & fast \\
-Aberdeenshire & NA & 2002.168 & 0.416 & 0.931 & fast \\
-Southampton & South East & 2002.189 & 0.472 & 0.922 & fast \\
-Mid Suffolk & East of England & 2002.196 & 0.442 & 0.922 & fast \\
-Eden & North West & 2002.196 & 0.334 & 0.949 & fast \\
-South Lakeland & North West & 2002.197 & 0.238 & 0.974 & fast \\
-Stoke-on-Trent & West Midlands & 2002.203 & 0.419 & 0.927 & fast \\
-Babergh & East of England & 2002.209 & 0.449 & 0.927 & fast \\
-Bexley & London & 2002.213 & 0.493 & 0.909 & fast \\
-Torbay & South West & 2002.215 & 0.374 & 0.942 & fast \\
-Brentwood & East of England & 2002.219 & 0.336 & 0.936 & fast \\
-Ceredigion & NA & 2002.227 & 0.241 & 0.971 & fast \\
-Spelthorne & South East & 2002.232 & 0.451 & 0.933 & fast \\
-Bath and North East Somerset & South West & 2002.238 & 0.402 & 0.941 &
-fast \\
-Falkirk & NA & 2002.245 & 0.367 & 0.938 & fast \\
-Broxtowe & East Midlands & 2002.258 & 0.435 & 0.932 & fast \\
-East Hertfordshire & East of England & 2002.261 & 0.453 & 0.928 &
-fast \\
-Stirling & NA & 2002.268 & 0.251 & 0.975 & fast \\
-Ealing & London & 2002.294 & 0.523 & 0.922 & fast \\
-Mid Sussex & South East & 2002.295 & 0.453 & 0.936 & fast \\
-Barrow-in-Furness & North West & 2002.303 & 0.284 & 0.959 & fast \\
-West Oxfordshire & South East & 2002.322 & 0.467 & 0.929 & fast \\
-Sutton & London & 2002.326 & 0.544 & 0.911 & fast \\
-Portsmouth & South East & 2002.356 & 0.479 & 0.921 & fast \\
-Great Yarmouth & East of England & 2002.369 & 0.265 & 0.968 & fast \\
-East Hampshire & South East & 2002.370 & 0.474 & 0.931 & fast \\
-Wyre Forest & West Midlands & 2002.376 & 0.561 & 0.916 & fast \\
-Angus & NA & 2002.378 & 0.396 & 0.939 & fast \\
-Argyll and Bute & NA & 2002.427 & 0.321 & 0.962 & fast \\
-Scottish Borders & NA & 2002.428 & 0.298 & 0.962 & fast \\
-East Suffolk & East of England & 2002.431 & 0.533 & 0.919 & fast \\
-Blackpool & North West & 2002.443 & 0.294 & 0.962 & fast \\
-Cherwell & South East & 2002.448 & 0.533 & 0.923 & fast \\
-Wandsworth & London & 2002.449 & 0.570 & 0.921 & fast \\
-Tendring & East of England & 2002.456 & 0.456 & 0.935 & fast \\
-Westminster & London & 2002.459 & 0.437 & 0.952 & fast \\
-Richmond upon Thames & London & 2002.462 & 0.389 & 0.956 & fast \\
-Test Valley & South East & 2002.463 & 0.553 & 0.920 & fast \\
-Isle of Wight & South East & 2002.478 & 0.377 & 0.953 & fast \\
-Vale of White Horse & South East & 2002.496 & 0.408 & 0.949 & fast \\
-Croydon & London & 2002.510 & 0.504 & 0.927 & fast \\
-Hounslow & London & 2002.521 & 0.426 & 0.952 & fast \\
-Copeland & North West & 2002.524 & 0.237 & 0.971 & fast \\
-Wiltshire & South West & 2002.528 & 0.450 & 0.940 & fast \\
-Gwynedd & NA & 2002.536 & 0.372 & 0.954 & fast \\
-Hillingdon & London & 2002.560 & 0.470 & 0.940 & fast \\
-Conwy & NA & 2002.564 & 0.372 & 0.954 & fast \\
-Highland & NA & 2002.566 & 0.270 & 0.972 & fast \\
-Basingstoke and Deane & South East & 2002.572 & 0.461 & 0.947 & fast \\
-Harrow & London & 2002.587 & 0.509 & 0.936 & fast \\
-West Berkshire & South East & 2002.595 & 0.696 & 0.902 & fast \\
-North Lincolnshire & Yorkshire and The Humber & 2002.596 & 0.501 & 0.933
-& fast \\
-East Dunbartonshire & NA & 2002.598 & 0.528 & 0.925 & fast \\
-Bradford & Yorkshire and The Humber & 2002.598 & 0.613 & 0.916 & fast \\
-Waverley & South East & 2002.606 & 0.512 & 0.928 & fast \\
-Stroud & South West & 2002.608 & 0.559 & 0.916 & fast \\
-Herefordshire, County of & West Midlands & 2002.622 & 0.388 & 0.956 &
-fast \\
-North Norfolk & East of England & 2002.624 & 0.422 & 0.945 & fast \\
-Nottingham & East Midlands & 2002.630 & 0.531 & 0.936 & fast \\
-Castle Point & East of England & 2002.645 & 0.470 & 0.936 & fast \\
-North Warwickshire & West Midlands & 2002.657 & 0.380 & 0.955 & fast \\
-Gravesham & South East & 2002.680 & 0.473 & 0.933 & fast \\
-Lancaster & North West & 2002.692 & 0.336 & 0.963 & fast \\
-West Suffolk & East of England & 2002.699 & 0.536 & 0.930 & fast \\
-Stafford & West Midlands & 2002.719 & 0.499 & 0.933 & fast \\
-Tunbridge Wells & South East & 2002.720 & 0.401 & 0.958 & fast \\
-Oxford & South East & 2002.723 & 0.583 & 0.934 & fast \\
-East Lothian & NA & 2002.731 & 0.537 & 0.927 & fast \\
-Redcar and Cleveland & North East & 2002.750 & 0.509 & 0.929 & fast \\
-Tamworth & West Midlands & 2002.755 & 0.551 & 0.929 & fast \\
-South Hams & South West & 2002.773 & 0.422 & 0.952 & fast \\
-Allerdale & North West & 2002.784 & 0.376 & 0.958 & fast \\
-Central Bedfordshire & East of England & 2002.825 & 0.620 & 0.921 &
-fast \\
-East Cambridgeshire & East of England & 2002.843 & 0.636 & 0.927 &
-fast \\
-Havant & South East & 2002.853 & 0.541 & 0.932 & fast \\
-South Lanarkshire & NA & 2002.874 & 0.486 & 0.938 & fast \\
-Stratford-on-Avon & West Midlands & 2002.879 & 0.470 & 0.945 & fast \\
-Arun & South East & 2002.882 & 0.526 & 0.938 & fast \\
-Fenland & East of England & 2002.891 & 0.632 & 0.918 & fast \\
-Monmouthshire & NA & 2002.893 & 0.501 & 0.935 & fast \\
-Denbighshire & NA & 2002.901 & 0.457 & 0.947 & fast \\
-Southend-on-Sea & East of England & 2002.903 & 0.551 & 0.932 & fast \\
-High Peak & East Midlands & 2002.903 & 0.488 & 0.946 & fast \\
-Pembrokeshire & NA & 2002.905 & 0.387 & 0.958 & fast \\
-Malvern Hills & West Midlands & 2002.907 & 0.524 & 0.940 & fast \\
-Camden & London & 2002.915 & 0.599 & 0.941 & fast \\
-North Devon & South West & 2002.926 & 0.426 & 0.950 & fast \\
-Wyre & North West & 2002.927 & 0.553 & 0.927 & fast \\
-Darlington & North East & 2002.933 & 0.615 & 0.929 & fast \\
-Swale & South East & 2002.940 & 0.447 & 0.948 & fast \\
-Welwyn Hatfield & East of England & 2002.951 & 0.628 & 0.922 & fast \\
-Mid Devon & South West & 2002.960 & 0.485 & 0.946 & fast \\
-Lewes & South East & 2002.964 & 0.507 & 0.945 & fast \\
-Dorset & South West & 2002.968 & 0.407 & 0.958 & fast \\
-Brighton and Hove & South East & 2002.976 & 0.578 & 0.934 & fast \\
-Braintree & East of England & 2002.978 & 0.534 & 0.938 & fast \\
-Epsom and Ewell & South East & 2002.979 & 0.746 & 0.908 & fast \\
-Sefton & North West & 2002.980 & 0.616 & 0.928 & fast \\
-Eastleigh & South East & 2002.992 & 0.595 & 0.929 & fast \\
-Craven & Yorkshire and The Humber & 2002.993 & 0.379 & 0.965 & fast \\
-Hammersmith and Fulham & London & 2002.997 & 0.585 & 0.944 & fast \\
-Greenwich & London & 2003.014 & 0.571 & 0.930 & slow \\
-Cheltenham & South West & 2003.025 & 0.602 & 0.937 & slow \\
-Worcester & West Midlands & 2003.041 & 0.659 & 0.920 & slow \\
-Tower Hamlets & London & 2003.043 & 0.660 & 0.926 & slow \\
-New Forest & South East & 2003.046 & 0.554 & 0.941 & slow \\
-East Renfrewshire & NA & 2003.046 & 0.739 & 0.900 & slow \\
-Fylde & North West & 2003.056 & 0.501 & 0.946 & slow \\
-Carmarthenshire & NA & 2003.067 & 0.390 & 0.958 & slow \\
-West Lancashire & North West & 2003.076 & 0.547 & 0.937 & slow \\
-Hastings & South East & 2003.086 & 0.685 & 0.911 & slow \\
-Carlisle & North West & 2003.111 & 0.541 & 0.934 & slow \\
-Wychavon & West Midlands & 2003.115 & 0.570 & 0.935 & slow \\
-Isle of Anglesey & NA & 2003.144 & 0.403 & 0.960 & slow \\
-Dudley & West Midlands & 2003.150 & 0.678 & 0.925 & slow \\
-Ryedale & Yorkshire and The Humber & 2003.152 & 0.380 & 0.961 & slow \\
-Rother & South East & 2003.162 & 0.366 & 0.966 & slow \\
-Leicester & East Midlands & 2003.167 & 0.686 & 0.924 & slow \\
-Breckland & East of England & 2003.182 & 0.567 & 0.938 & slow \\
-Tandridge & South East & 2003.190 & 0.747 & 0.918 & slow \\
-Shropshire & West Midlands & 2003.206 & 0.621 & 0.928 & slow \\
-West Devon & South West & 2003.228 & 0.489 & 0.949 & slow \\
-Milton Keynes & South East & 2003.231 & 0.700 & 0.929 & slow \\
-Sandwell & West Midlands & 2003.234 & 0.760 & 0.916 & slow \\
-Winchester & South East & 2003.234 & 0.443 & 0.956 & slow \\
-Redditch & West Midlands & 2003.238 & 0.706 & 0.923 & slow \\
-Rugby & West Midlands & 2003.265 & 0.709 & 0.922 & slow \\
-North Somerset & South West & 2003.278 & 0.558 & 0.939 & slow \\
-West Lindsey & East Midlands & 2003.285 & 0.674 & 0.924 & slow \\
-Richmondshire & Yorkshire and The Humber & 2003.286 & 0.425 & 0.958 &
-slow \\
-Melton & East Midlands & 2003.292 & 0.534 & 0.946 & slow \\
-Chichester & South East & 2003.295 & 0.582 & 0.942 & slow \\
-Torridge & South West & 2003.301 & 0.532 & 0.942 & slow \\
-Dundee City & NA & 2003.303 & 0.568 & 0.939 & slow \\
-South Ayrshire & NA & 2003.303 & 0.626 & 0.931 & slow \\
-East Devon & South West & 2003.308 & 0.492 & 0.947 & slow \\
-Bolton & North West & 2003.323 & 0.774 & 0.913 & slow \\
-Hertsmere & East of England & 2003.326 & 0.726 & 0.924 & slow \\
-Bournemouth, Christchurch and Poole & South West & 2003.347 & 0.662 &
-0.936 & slow \\
-Maldon & East of England & 2003.349 & 0.495 & 0.951 & slow \\
-Barnet & London & 2003.352 & 0.766 & 0.919 & slow \\
-Brent & London & 2003.389 & 0.762 & 0.925 & slow \\
-Midlothian & NA & 2003.413 & 0.732 & 0.921 & slow \\
-Boston & East Midlands & 2003.416 & 0.593 & 0.933 & slow \\
-Scarborough & Yorkshire and The Humber & 2003.421 & 0.431 & 0.963 &
-slow \\
-Kirklees & Yorkshire and The Humber & 2003.435 & 0.725 & 0.925 & slow \\
-Tameside & North West & 2003.441 & 0.819 & 0.907 & slow \\
-Newport & NA & 2003.449 & 0.635 & 0.930 & slow \\
-Glasgow City & NA & 2003.451 & 0.759 & 0.922 & slow \\
-Clackmannanshire & NA & 2003.454 & 0.585 & 0.936 & slow \\
-Three Rivers & East of England & 2003.458 & 0.628 & 0.947 & slow \\
-Harborough & East Midlands & 2003.488 & 0.549 & 0.950 & slow \\
-Rochdale & North West & 2003.496 & 0.685 & 0.937 & slow \\
-Tonbridge and Malling & South East & 2003.497 & 0.710 & 0.927 & slow \\
-Somerset West and Taunton & South West & 2003.502 & 0.491 & 0.955 &
-slow \\
-Wrexham & NA & 2003.512 & 0.553 & 0.942 & slow \\
-Worthing & South East & 2003.522 & 0.622 & 0.938 & slow \\
-Tewkesbury & South West & 2003.540 & 0.580 & 0.946 & slow \\
-City of Edinburgh & NA & 2003.541 & 0.685 & 0.941 & slow \\
-Coventry & West Midlands & 2003.592 & 0.564 & 0.948 & slow \\
-Northumberland & North East & 2003.607 & 0.508 & 0.954 & slow \\
-Mid and East Antrim & NA & 2003.609 & 0.893 & 0.900 & slow \\
-Wigan & North West & 2003.617 & 0.755 & 0.929 & slow \\
-Lichfield & West Midlands & 2003.629 & 0.625 & 0.939 & slow \\
-Ashford & South East & 2003.633 & 0.535 & 0.944 & slow \\
-Cornwall & South West & 2003.638 & 0.443 & 0.963 & slow \\
-Mendip & South West & 2003.652 & 0.505 & 0.955 & slow \\
-Stevenage & East of England & 2003.665 & 0.525 & 0.958 & slow \\
-Blaenau Gwent & NA & 2003.666 & 0.643 & 0.933 & slow \\
-Lincoln & East Midlands & 2003.666 & 0.634 & 0.940 & slow \\
-South Somerset & South West & 2003.676 & 0.553 & 0.948 & slow \\
-Charnwood & East Midlands & 2003.677 & 0.699 & 0.934 & slow \\
-North Ayrshire & NA & 2003.682 & 0.607 & 0.942 & slow \\
-Folkestone and Hythe & South East & 2003.684 & 0.792 & 0.905 & slow \\
-Eastbourne & South East & 2003.684 & 0.697 & 0.930 & slow \\
-Blackburn with Darwen & North West & 2003.685 & 0.680 & 0.938 & slow \\
-Dover & South East & 2003.699 & 0.574 & 0.948 & slow \\
-Adur & South East & 2003.735 & 0.551 & 0.954 & slow \\
-Haringey & London & 2003.739 & 0.667 & 0.943 & slow \\
-Gosport & South East & 2003.751 & 0.600 & 0.934 & slow \\
-Trafford & North West & 2003.752 & 0.640 & 0.949 & slow \\
-Norwich & East of England & 2003.756 & 0.742 & 0.939 & slow \\
-West Lothian & NA & 2003.758 & 0.652 & 0.943 & slow \\
-North East Lincolnshire & Yorkshire and The Humber & 2003.811 & 0.858 &
-0.920 & slow \\
-Bristol, City of & South West & 2003.828 & 0.831 & 0.928 & slow \\
-Rushcliffe & East Midlands & 2003.839 & 0.591 & 0.954 & slow \\
-Hinckley and Bosworth & East Midlands & 2003.857 & 0.749 & 0.934 &
-slow \\
-Torfaen & NA & 2003.861 & 0.841 & 0.922 & slow \\
-Swansea & NA & 2003.875 & 0.650 & 0.947 & slow \\
-Wolverhampton & West Midlands & 2003.882 & 0.793 & 0.931 & slow \\
-Kingston upon Hull, City of & Yorkshire and The Humber & 2003.885 &
-0.757 & 0.936 & slow \\
-Calderdale & Yorkshire and The Humber & 2003.889 & 0.714 & 0.939 &
-slow \\
-North Northamptonshire & East Midlands & 2003.902 & 0.746 & 0.935 &
-slow \\
-Stockton-on-Tees & North East & 2003.902 & 0.955 & 0.916 & slow \\
-Wealden & South East & 2003.906 & 0.736 & 0.938 & slow \\
-Thanet & South East & 2003.953 & 0.667 & 0.947 & slow \\
-Broadland & East of England & 2003.962 & 0.727 & 0.938 & slow \\
-West Northamptonshire & East Midlands & 2003.972 & 0.736 & 0.940 &
-slow \\
-North Kesteven & East Midlands & 2003.976 & 0.633 & 0.942 & slow \\
-Neath Port Talbot & NA & 2003.981 & 0.767 & 0.933 & slow \\
-Rhondda Cynon Taf & NA & 2003.991 & 0.769 & 0.925 & slow \\
-Enfield & London & 2003.992 & 0.588 & 0.954 & slow \\
-East Lindsey & East Midlands & 2004.010 & 0.577 & 0.952 & slow \\
-Teignbridge & South West & 2004.015 & 0.681 & 0.941 & slow \\
-York & Yorkshire and The Humber & 2004.025 & 0.739 & 0.944 & slow \\
-Harlow & East of England & 2004.030 & 0.803 & 0.936 & slow \\
-Vale of Glamorgan & NA & 2004.033 & 0.592 & 0.951 & slow \\
-Newark and Sherwood & East Midlands & 2004.034 & 0.668 & 0.946 & slow \\
-Merton & London & 2004.041 & 0.932 & 0.922 & slow \\
-Hambleton & Yorkshire and The Humber & 2004.065 & 0.636 & 0.953 &
-slow \\
-Bedford & East of England & 2004.068 & 0.852 & 0.933 & slow \\
-Stockport & North West & 2004.075 & 0.757 & 0.942 & slow \\
-South Ribble & North West & 2004.126 & 0.674 & 0.948 & slow \\
-Uttlesford & East of England & 2004.144 & 0.905 & 0.926 & slow \\
-Rossendale & North West & 2004.160 & 0.976 & 0.923 & slow \\
-Oldham & North West & 2004.162 & 0.801 & 0.938 & slow \\
-Sheffield & Yorkshire and The Humber & 2004.189 & 0.770 & 0.944 &
-slow \\
-Ribble Valley & North West & 2004.204 & 0.718 & 0.945 & slow \\
-Maidstone & South East & 2004.222 & 0.854 & 0.932 & slow \\
-Cheshire West and Chester & North West & 2004.234 & 0.742 & 0.947 &
-slow \\
-Thurrock & East of England & 2004.241 & 0.841 & 0.924 & slow \\
-Telford and Wrekin & West Midlands & 2004.241 & 1.024 & 0.917 & slow \\
-Cardiff & NA & 2004.246 & 0.823 & 0.937 & slow \\
-County Durham & North East & 2004.250 & 0.626 & 0.953 & slow \\
-Plymouth & South West & 2004.252 & 0.785 & 0.934 & slow \\
-Bridgend & NA & 2004.256 & 0.547 & 0.956 & slow \\
-Rotherham & Yorkshire and The Humber & 2004.273 & 0.722 & 0.938 &
-slow \\
-Hyndburn & North West & 2004.298 & 0.881 & 0.932 & slow \\
-Newcastle upon Tyne & North East & 2004.305 & 0.906 & 0.932 & slow \\
-Wakefield & Yorkshire and The Humber & 2004.314 & 0.905 & 0.923 &
-slow \\
-Havering & London & 2004.345 & 0.822 & 0.938 & slow \\
-South Kesteven & East Midlands & 2004.353 & 1.026 & 0.915 & slow \\
-Canterbury & South East & 2004.354 & 0.602 & 0.960 & slow \\
-Broxbourne & East of England & 2004.363 & 0.817 & 0.944 & slow \\
-Peterborough & East of England & 2004.382 & 0.937 & 0.932 & slow \\
-Redbridge & London & 2004.411 & 1.028 & 0.925 & slow \\
-Bury & North West & 2004.431 & 0.735 & 0.953 & slow \\
-Harrogate & Yorkshire and The Humber & 2004.433 & 0.971 & 0.927 &
-slow \\
-Birmingham & West Midlands & 2004.435 & 1.182 & 0.912 & slow \\
-South Tyneside & North East & 2004.446 & 0.636 & 0.948 & slow \\
-Chorley & North West & 2004.462 & 1.093 & 0.913 & slow \\
-North East Derbyshire & East Midlands & 2004.483 & 0.703 & 0.953 &
-slow \\
-Lambeth & London & 2004.485 & 0.958 & 0.937 & slow \\
-Doncaster & Yorkshire and The Humber & 2004.496 & 0.672 & 0.953 &
-slow \\
-South Gloucestershire & South West & 2004.499 & 0.881 & 0.937 & slow \\
-Caerphilly & NA & 2004.499 & 0.819 & 0.938 & slow \\
-East Staffordshire & West Midlands & 2004.501 & 1.132 & 0.910 & slow \\
-North Tyneside & North East & 2004.509 & 0.864 & 0.933 & slow \\
-South Holland & East Midlands & 2004.534 & 0.851 & 0.941 & slow \\
-Gateshead & North East & 2004.535 & 0.869 & 0.934 & slow \\
-St.~Helens & North West & 2004.542 & 1.097 & 0.923 & slow \\
-Flintshire & NA & 2004.549 & 0.764 & 0.951 & slow \\
-Selby & Yorkshire and The Humber & 2004.549 & 0.716 & 0.947 & slow \\
-Belfast & NA & 2004.556 & 1.332 & 0.906 & slow \\
-Sedgemoor & South West & 2004.613 & 0.540 & 0.966 & slow \\
-Exeter & South West & 2004.617 & 0.964 & 0.936 & slow \\
-South Derbyshire & East Midlands & 2004.629 & 0.934 & 0.938 & slow \\
-Derbyshire Dales & East Midlands & 2004.668 & 0.648 & 0.963 & slow \\
-East Riding of Yorkshire & Yorkshire and The Humber & 2004.683 & 0.728 &
-0.955 & slow \\
-Basildon & East of England & 2004.691 & 1.067 & 0.925 & slow \\
-Barking and Dagenham & London & 2004.800 & 1.198 & 0.916 & slow \\
-Causeway Coast and Glens & NA & 2004.834 & 1.024 & 0.928 & slow \\
-North Lanarkshire & NA & 2004.839 & 0.878 & 0.945 & slow \\
-Chesterfield & East Midlands & 2004.887 & 1.173 & 0.927 & slow \\
-Newcastle-under-Lyme & West Midlands & 2004.933 & 1.172 & 0.925 &
-slow \\
-East Ayrshire & NA & 2004.934 & 1.280 & 0.903 & slow \\
-Nuneaton and Bedworth & West Midlands & 2004.956 & 0.950 & 0.941 &
-slow \\
-Blaby & East Midlands & 2004.959 & 1.059 & 0.929 & slow \\
-Leeds & Yorkshire and The Humber & 2004.979 & 0.836 & 0.956 & slow \\
-Erewash & East Midlands & 2004.987 & 0.939 & 0.944 & slow \\
-Bassetlaw & East Midlands & 2004.998 & 0.982 & 0.944 & slow \\
-Hackney & London & 2005.026 & 1.157 & 0.934 & slow \\
-Barnsley & Yorkshire and The Humber & 2005.055 & 0.912 & 0.942 & slow \\
-Amber Valley & East Midlands & 2005.105 & 1.017 & 0.941 & slow \\
-Colchester & East of England & 2005.144 & 1.250 & 0.927 & slow \\
-Ashfield & East Midlands & 2005.176 & 1.073 & 0.942 & slow \\
-Wirral & North West & 2005.184 & 1.084 & 0.939 & slow \\
-Knowsley & North West & 2005.186 & 1.160 & 0.923 & slow \\
-Windsor and Maidenhead & South East & 2005.195 & 1.746 & 0.902 & slow \\
-Bromsgrove & West Midlands & 2005.494 & 1.144 & 0.941 & slow \\
-Liverpool & North West & 2005.526 & 1.276 & 0.936 & slow \\
-Staffordshire Moorlands & West Midlands & 2005.583 & 1.426 & 0.924 &
-slow \\
-Islington & London & 2005.970 & 1.223 & 0.955 & slow \\
-Warrington & North West & 2006.015 & 1.387 & 0.942 & slow \\
-Newham & London & 2006.078 & 1.908 & 0.907 & slow \\
-Bolsover & East Midlands & 2006.188 & 1.143 & 0.957 & slow \\
-Lisburn and Castlereagh & NA & 2006.193 & 1.970 & 0.903 & slow \\
-Manchester & North West & 2006.449 & 1.434 & 0.950 & slow \\
-Sunderland & North East & 2006.479 & 1.679 & 0.928 & slow \\
-South Staffordshire & West Midlands & 2006.490 & 1.837 & 0.922 & slow \\
-Rutland & East Midlands & 2006.529 & 0.986 & 0.967 & slow \\
-Cannock Chase & West Midlands & 2006.700 & 2.231 & 0.917 & slow \\
-Mansfield & East Midlands & 2006.889 & 1.579 & 0.946 & slow \\
-Ards and North Down & NA & 2006.947 & 2.485 & 0.907 & slow \\
-Halton & North West & 2007.069 & 0.932 & 0.967 & slow \\
-Burnley & North West & 2007.372 & 2.286 & 0.919 & slow \\
-Middlesbrough & North East & 2010.074 & 5.297 & 0.900 & slow \\
-Isles of Scilly & South West & 2012.163 & 4.257 & 0.953 & slow \\
-\end{longtable}
-
-\begin{longtable}[]{@{}lrrrr@{}}
-\caption{Regional differences; up to 10 postcodes per
-website\label{table.regions.10}}\tabularnewline
-\toprule\noalign{}
-Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Yorkshire and The Humber & 0.864 & 1 & 0.191 & 5 \\
-West Midlands & 0.860 & 2 & 0.148 & 8 \\
-South East & 0.848 & 3 & 0.171 & 7 \\
-North West & 0.827 & 4 & 0.221 & 1 \\
-London & 0.826 & 5 & 0.118 & 9 \\
-South West & 0.801 & 6 & 0.194 & 4 \\
-Wales & 0.746 & 7 & 0.114 & 11 \\
-East Midlands & 0.732 & 8 & 0.117 & 10 \\
-East of England & 0.731 & 9 & 0.178 & 6 \\
-North East & 0.695 & 10 & 0.215 & 2 \\
-Scotland & 0.588 & 11 & 0.052 & 12 \\
-Nortern Ireland & 0.488 & 12 & 0.197 & 3 \\
-\end{longtable}
-
-\begin{longtable}[]{@{}llll@{}}
-\caption{Model metrics for growth
-model}\label{tbl-model-growth-metrics}\tabularnewline
-\toprule\noalign{}
-& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Local Authorities & 0.286 & 0.506 & 0.223 \\
-\end{longtable}
-
-Table~\ref{tbl-model-growth-metrics} presents the results of a Random
-Forest model that was trained based on Equation \ref{model.growth} using
-all LAD data points. In essence, this is a model similar to the one
-estimated based on Equation \ref{eq:model} and presented in
-Table~\ref{tbl-model-metrics}, with the main difference being the
-dependent variable which is the year growth rate of website density for
-each LAD. To avoid cases with LAD have no websites,
-\(t \in [2000, 2012]\). The high predictive capability of the model
-reflects the robustness of the modelling approach.
-
-\begin{align} \label{model.growth}
-Website\,Density\,Growth_{t} \sim Distance\,London +
-Website\,density\,London_{t-1} +\notag\\
-Distance\,Nearest\,City +
-Website\,density\,Nearest\,City_{t-1} +\notag\\
-Distance\,Nearest\,Retail_{i} +
-Website\,density\,Nearest\,Retail_{t-1} +\notag\\
-W*\, Website\,density_{t-1} +\notag\\ 
-year_{t}
-\end{align}
-
-\renewcommand\refname{References}
-  \bibliography{bibliography}
+No external funding was received for this work.
 

### 2025-07-29 paper/submission4/anonymised.qmd
submission 3

--- 2025-06-08 paper/submission3/anonymised.qmd
@@
 ---
-title: "A multi-scale story of the diffusion of a new technology: the web"
+title: "A multi-scale story of the diffusion of a new technology: the Web"
 abstract: |
@@
 new technology is observed. As per other key concepts in economic geography
-[@briant2010dots; @meliciani2015determinants; @andini2019social; @rosenthal2001determinants],
+[@briant2010dots; @meliciani2015determinants; @andini2019social],
 diffusion mechanisms can indeed suffer from Modifiable Areal Unit Problem 
@@
 process, it is hardly ever observed at this, or at a close enough, scale. 
+<!-- response to editor -->
+Hence, this paper exposes the importance of spatially-explicit diffusion 
+mechanisms at granular geographical scales without the risk of MAUP. The results
+are linked to core debates in economic geography about agglomeration 
+externalities, centripetal forces and the spatially clustered demand for new 
+technologies. Although the paper focuses on the Web, the framework offered here 
+and the findings which illustrate the micro-geographical patterns of the 
+diffusion of a new technology are applicable to other new technologies.
 
@@
 captures the very early stages of this technology until its maturity.
-This is the first time that the importance of granular spatial
-mechanisms for the diffusion of a new technology is exposed. 
-Importantly, the empirical framework and the data employed here allows
-to observe the active engagement with the Web that is creating and
-maintaining a commercial website instead of the more passive action of
-browsing webpages and accessing the Internet.
+This is the first time that the importance of granular spatial mechanisms for 
+the diffusion of a new technology is exposed. Importantly, the empirical 
+framework and the data employed here allows to observe the active engagement 
+with the Web that is creating and maintaining a commercial website instead of 
+the more passive action of browsing webpages and accessing the Internet.
 
@@
 remarkably good predictive capacity of these mechanisms at granular geographical 
-scales. The modelling framework exposes that the relative importance of 
-these spatial mechanisms differ across different scales. It also reveals 
-spatial heterogeneity. While for some UK regions the importance of the spatial 
-diffusion mechanisms are very similar to the rest of the country, other regions
-diverge significantly from the country average.
+scales. It also exposes that the relative importance of these spatial mechanisms 
+differ across different scales and the spatial heterogeneity of these findings. 
+While for some UK regions the importance of the spatial diffusion mechanisms are 
+very similar to the rest of the country, other regions diverge significantly 
+from the country average.
 
@@
 <!-- H2 -->
-<!-- H~2~: The diffusion of the commercial Web is shaped by both early adoption  -->
-<!-- and latecomer effects. -->
 The second hypothesis assessed whether the diffusion of the web is shaped by 
@@
 highlighted the heterogeneity of adoption rates between LAD as well as the 
-spatial clustering of LAD with similar adoption rates. Importantly, the 
-observed spatial clustering of LAD with fast diffusion rates can
-reinforce existing and rooted patterns of unequal economic productivity 
+spatial clustering of LAD with similar adoption rates. 
+<!-- response to editor --> 
+This empirical finding speaks directly to the spatial clustering literature. 
+Just like the clustering of firms [@boschma2011emerging; 
+@audretsch1998innovation], the adoption of a new technology is shaped by 
+geographically bounded spillovers and path dependency visible even at very
+local scales.
+Importantly, the observed spatial clustering of LAD with fast diffusion rates 
+can reinforce existing and rooted patterns of unequal economic productivity 
 in the UK [@mccann2022productivity]. Previous research had already indicated 
@@
 
-The paper offers a number of different contributions. Firstly, it
-assesses, for the first time, the role of these spatial diffusion
-mechanisms at such granular spatial scales. While previous studies had
-highlighted these mechanisms [@hagerstrand1968innovation;
-@rogers2010diffusion; @grubler1990rise], their role had never been
-tested before at such granular scale. The findings talk directly to this
-literature [e.g. @beardsell1999spatial; @haller2011determinants;
-@lengyel2020role]. More broadly, the paper offers a robust
-methodological framework, which utilises novel data and modern ML
-algorithms and techniques to model the diffusion of a new technology
-across space and time. It directly addresses the data problem which
-prevented economic geographers in engaging with diffusion studies
+The paper offers a number of different contributions. Firstly, it assesses, for 
+the first time, the role of these spatial diffusion mechanisms at granular 
+spatial scales. While previous studies had highlighted these mechanisms
+[@hagerstrand1968innovation; @rogers2010diffusion; @grubler1990rise], their role 
+had never been tested before at such a granular scale 
+<!-- response to editor -->
+to overcome the potential risk of a MAUP. The spatial heterogeneity of 
+the diffusion mechanisms the analysis revealed demonstrates the need for 
+observing the diffusion of a new technology at a scale as close as possible to 
+the individual level as otherwise significant adoption patterns and, importantly
+for economic geography, spatially clustered patterns can be missed. 
+The economic geography literature has found evidence of a MAUP in various 
+concepts (see for example @briant2010dots for spatial concentration, 
+agglomeration economies, and trade determinants and @andini2019social for social 
+capital), but this is the first paper that revealed how the scale of observation 
+can affect our understanding of the diffusion of a new technology.
+
+In addition, the paper directly addresses the data problem which prevented 
+economic geographers in engaging with diffusion studies
 [@iso2005; @kemeny2011international; @zook2022mapping]. Thus, the paper
 seeks to renew economic geography's interest in the diffusion of new 
-technologies. Current data science advances including the availability
-of granular enough data and adequate methods can enable economic
-geographers to reveal interesting spatial patterns of diffusion of new
-technologies as this paper has demonstrated. This is important as (i)
-various stakeholders are interested in predicting localised demand for
-new technologies for planning purposes, for example 5G networks
-[@leibowicz2016representing;@meade2021modelling], and (ii) the adoption 
-of such digital technologies is associated with local economic gains 
-[@solow1957technical; @aghion1990model; @kemeny2011international; @tranosuk;
-@capello2024nexus]. 
-<!-- In addition, while most related studies focused on -->
-<!-- tangible technologies, this paper sheds light on the diffusion of a -->
-<!-- truly weightless digital technology that is the Web. Thus, it adds much -->
-<!-- more granular findings to the limited related studies -->
-<!-- [@sinai2004geography; @haller2011determinants].  -->
-Importantly, although the geographical focus of this paper is the UK, the 
-empirical framework developed here is easily transferable to other geographical 
-contexts.
+technologies.
+
+More broadly, the paper offers a robust methodological framework, which utilises 
+novel data and modern ML algorithms and techniques to model the diffusion of a 
+new technology across space and time. Current data science advances including 
+the availability of granular enough data and adequate methods enable economic
+geographers to reveal interesting spatial patterns about the diffusion of new 
+technologies as this paper has demonstrated. This is important as (i) various 
+stakeholders are interested in predicting localised demand for new technologies 
+for planning purposes [@leibowicz2016representing; @meade2021modelling], and 
+(ii) the adoption of such digital technologies is associated with local economic 
+gains [@solow1957technical; @aghion1990model; @kemeny2011international; 
+@tranosuk; @capello2024nexus].
+
+<!-- response to editor -->
+These contributions are transferable to different geographical and technological 
+contexts. Although the focus of this paper is the Web, the framework and the 
+findings are applicable to different technologies. For example, the three 
+diffusion mechanisms -- that is distance, urban hierarchy and the S-shaped 
+pattern of the cumulative level of adoption -- and their spatial heterogeneity 
+can be used to explain and, importantly from a policy perspective, predict local 
+demand for different new technologies. While, for instance, it is well expected 
+to firstly install new infrastructure such as 5G antennas in central urban areas, 
+the results indicate that proximity to these core urban areas can also drive 
+local demand for such a technology, probably at a second stage. In terms of 
+the empirical framework, it can be used with any point data representing 
+technological adoption. Also, archived web data can be obtain for other country 
+domains.
+
+<!-- Such an approach could assist  -->
+<!-- in mapping the footprints of "technological niches" and, consequently, the role  -->
+<!-- of proximity in the creation of such niche markets [@schot2013strategic;  -->
+<!-- @bednarz2020pulled].  -->
 

### 2025-07-29 paper/submission4/anonymised.tex
submission 3

--- 2025-06-08 paper/submission3/anonymised.tex
@@
 \hypersetup{
-  pdftitle={A multi-scale story of the diffusion of a new technology: the web},
+  pdftitle={A multi-scale story of the diffusion of a new technology: the Web},
   pdfkeywords={diffusion, technologies, Web, digital},
@@
 \title{A multi-scale story of the diffusion of a new technology: the
-web}
+Web}
 
@@
 key concepts in economic geography
-\citep{briant2010dots, meliciani2015determinants, andini2019social, rosenthal2001determinants},
+\citep{briant2010dots, meliciani2015determinants, andini2019social},
 diffusion mechanisms can indeed suffer from Modifiable Areal Unit
@@
 an individual process, it is hardly ever observed at this, or at a close
-enough, scale.
+enough, scale. Hence, this paper exposes the importance of
+spatially-explicit diffusion mechanisms at granular geographical scales
+without the risk of MAUP. The results are linked to core debates in
+economic geography about agglomeration externalities, centripetal forces
+and the spatially clustered demand for new technologies. Although the
+paper focuses on the Web, the framework offered here and the findings
+which illustrate the micro-geographical patterns of the diffusion of a
+new technology are applicable to other new technologies.
 
@@
 the modelling exercise reveals the remarkably good predictive capacity
-of these mechanisms at granular geographical scales. The modelling
-framework exposes that the relative importance of these spatial
-mechanisms differ across different scales. It also reveals spatial
-heterogeneity. While for some UK regions the importance of the spatial
-diffusion mechanisms are very similar to the rest of the country, other
-regions diverge significantly from the country average.
+of these mechanisms at granular geographical scales. It also exposes
+that the relative importance of these spatial mechanisms differ across
+different scales and the spatial heterogeneity of these findings. While
+for some UK regions the importance of the spatial diffusion mechanisms
+are very similar to the rest of the country, other regions diverge
+significantly from the country average.
 
@@
 as well as the spatial clustering of LAD with similar adoption rates.
-Importantly, the observed spatial clustering of LAD with fast diffusion
-rates can reinforce existing and rooted patterns of unequal economic
-productivity in the UK \citep{mccann2022productivity}. Previous research
-had already indicated the long-term productivity gains of the early
-adoption of such technologies \citep{tranosuk}. On the other hand, the
-analysis also exposed instances of remarkable stability as well as
-volatility of the Web adoption, with the latter being indicative of
-leapfrogging. While the literature had discussed the benefits of
-technological leapfrogging at the national scale
-\citep{teece2008firm, perkins2005international}, local scales had
-escaped such studies \citep{yu2018sustainability}. Thus the leapfrogging
-evidence presented here advocate towards further exploring such effects
-at granular scales.
+This empirical finding speaks directly to the spatial clustering
+literature. Just like the clustering of firms
+\citep{boschma2011emerging, audretsch1998innovation}, the adoption of a
+new technology is shaped by geographically bounded spillovers and path
+dependency visible even at very local scales. Importantly, the observed
+spatial clustering of LAD with fast diffusion rates can reinforce
+existing and rooted patterns of unequal economic productivity in the UK
+\citep{mccann2022productivity}. Previous research had already indicated
+the long-term productivity gains of the early adoption of such
+technologies \citep{tranosuk}. On the other hand, the analysis also
+exposed instances of remarkable stability as well as volatility of the
+Web adoption, with the latter being indicative of leapfrogging. While
+the literature had discussed the benefits of technological leapfrogging
+at the national scale \citep{teece2008firm, perkins2005international},
+local scales had escaped such studies \citep{yu2018sustainability}. Thus
+the leapfrogging evidence presented here advocate towards further
+exploring such effects at granular scales.
 
@@
 assesses, for the first time, the role of these spatial diffusion
-mechanisms at such granular spatial scales. While previous studies had
+mechanisms at granular spatial scales. While previous studies had
 highlighted these mechanisms
 \citep{hagerstrand1968innovation, rogers2010diffusion, grubler1990rise},
-their role had never been tested before at such granular scale. The
-findings talk directly to this literature
-\citep[e.g.][]{beardsell1999spatial, haller2011determinants, lengyel2020role}.
+their role had never been tested before at such a granular scale to
+overcome the potential risk of a MAUP. The spatial heterogeneity of the
+diffusion mechanisms the analysis revealed demonstrates the need for
+observing the diffusion of a new technology at a scale as close as
+possible to the individual level as otherwise significant adoption
+patterns and, importantly for economic geography, spatially clustered
+patterns can be missed. The economic geography literature has found
+evidence of a MAUP in various concepts (see for example
+\citet{briant2010dots} for spatial concentration, agglomeration
+economies, and trade determinants and \citet{andini2019social} for
+social capital), but this is the first paper that revealed how the scale
+of observation can affect our understanding of the diffusion of a new
+technology.
+
+In addition, the paper directly addresses the data problem which
+prevented economic geographers in engaging with diffusion studies
+\citep{iso2005, kemeny2011international, zook2022mapping}. Thus, the
+paper seeks to renew economic geography's interest in the diffusion of
+new technologies.
+
 More broadly, the paper offers a robust methodological framework, which
 utilises novel data and modern ML algorithms and techniques to model the
-diffusion of a new technology across space and time. It directly
-addresses the data problem which prevented economic geographers in
-engaging with diffusion studies
-\citep{iso2005, kemeny2011international, zook2022mapping}. Thus, the
-paper seeks to renew economic geography's interest in the diffusion of
-new technologies. Current data science advances including the
-availability of granular enough data and adequate methods can enable
-economic geographers to reveal interesting spatial patterns of diffusion
-of new technologies as this paper has demonstrated. This is important as
-(i) various stakeholders are interested in predicting localised demand
-for new technologies for planning purposes, for example 5G networks
-\citep{leibowicz2016representing, meade2021modelling}, and (ii) the
-adoption of such digital technologies is associated with local economic
-gains
+diffusion of a new technology across space and time. Current data
+science advances including the availability of granular enough data and
+adequate methods enable economic geographers to reveal interesting
+spatial patterns about the diffusion of new technologies as this paper
+has demonstrated. This is important as (i) various stakeholders are
+interested in predicting localised demand for new technologies for
+planning purposes \citep{leibowicz2016representing, meade2021modelling},
+and (ii) the adoption of such digital technologies is associated with
+local economic gains
 \citep{solow1957technical, aghion1990model, kemeny2011international, tranosuk, capello2024nexus}.
-Importantly, although the geographical focus of this paper is the UK,
-the empirical framework developed here is easily transferable to other
-geographical contexts.
+
+These contributions are transferable to different geographical and
+technological contexts. Although the focus of this paper is the Web, the
+framework and the findings are applicable to different technologies. For
+example, the three diffusion mechanisms -- that is distance, urban
+hierarchy and the S-shaped pattern of the cumulative level of adoption
+-- and their spatial heterogeneity can be used to explain and,
+importantly from a policy perspective, predict local demand for
+different new technologies. While, for instance, it is well expected to
+firstly install new infrastructure such as 5G antennas in central urban
+areas, the results indicate that proximity to these core urban areas can
+also drive local demand for such a technology, probably at a second
+stage. In terms of the empirical framework, it can be used with any
+point data representing technological adoption. Also, archived web data
+can be obtain for other country domains.
 

### 2025-07-29 paper/submission4/title.qmd
submission 3

--- 2025-07-29 paper/submission3/title.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-  authoryear,
-  preprint,
-  3p]{elsarticle}
+---
+title: "A multi-scale story of the diffusion of a new technology: the web"
+author:
+  - name: Emmanouil Tranos
+    email: e.tranos@bristol.ac.uk
+    affiliations: 
+      - id: some-tech
+        name: University of Bristol
+        department: School of Geographical Sciences
+        address: University Road, BS8 1SS
+        city: Bristol
+        country: UK
+    corresponding: e.tranos@bristol.ac.uk
+address:
+  - attributes:
+        corresponding: true
+    note: This is the first author footnote.
+abstract: |
+    This paper investigates the spatial diffusion of a new technology that is the Web in the UK.
+It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms.
+Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012).
+Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales.
+They also highlight spatial heterogeneity and instances of leapfrogging.
+keywords: 
+  - diffusion
+  - technologies
+  - Web
+  - digital
+date: last-modified
+format:
+  elsevier-pdf:
+    keep-tex: true
+    text: |
+      \usepackage[demo]{graphicx} % "demo" option just for this example
+      \usepackage{subcaption}
+    journal:
+      name: Journal of Economic Geography
+      formatting: preprint
+      model: 3p
+      cite-style: authoryear
+editor: 
+  markdown: 
+    wrap: 72
+---
 
-\usepackage{amsmath,amssymb}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else  
-    % xetex/luatex font selection
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
-% Make \paragraph and \subparagraph free-standing
-\makeatletter
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-\makeatother
+# Acknowledgements
 
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
+Many thanks to Dr Levi Wolf and Professor Dani Arribas-Bel for their constructive 
+comments. Also many thanks to the participants of the 6th Global Conference on 
+Economic Geography and the Winter Seminar of GfR. Then, many thanks to the 
+organisers of the seminar series and the members of the Quantitative Spatial 
+Science (QuSS) Research Group at the University of Bristol, the Centre for 
+Business Prosperity Research Group at Aston University (Dr Tasos Kitsos), and 
+the Centre for Innovation Research (Professor Tom Broekel) at the University of 
+Stavanger. Finally, many thanks to the editor and the three anonymous reviewers 
+for the very helpful comments and suggestions.
 
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\AtBeginDocument{%
-\ifdefined\contentsname
-  \renewcommand*\contentsname{Table of contents}
-\else
-  \newcommand\contentsname{Table of contents}
-\fi
-\ifdefined\listfigurename
-  \renewcommand*\listfigurename{List of Figures}
-\else
-  \newcommand\listfigurename{List of Figures}
-\fi
-\ifdefined\listtablename
-  \renewcommand*\listtablename{List of Tables}
-\else
-  \newcommand\listtablename{List of Tables}
-\fi
-\ifdefined\figurename
-  \renewcommand*\figurename{Figure}
-\else
-  \newcommand\figurename{Figure}
-\fi
-\ifdefined\tablename
-  \renewcommand*\tablename{Table}
-\else
-  \newcommand\tablename{Table}
-\fi
-}
-\@ifpackageloaded{float}{}{\usepackage{float}}
-\floatstyle{ruled}
-\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
-\floatname{codelisting}{Listing}
-\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\journal{Journal of Economic Geography}
-
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\usepackage[]{natbib}
-\bibliographystyle{elsarticle-harv}
-\usepackage{bookmark}
-
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  pdftitle={A multi-scale story of the diffusion of a new technology: the web},
-  pdfauthor={Emmanouil Tranos},
-  pdfkeywords={diffusion, technologies, Web, digital},
-  colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
-
-\setlength{\parindent}{6pt}
-\begin{document}
-
-\begin{frontmatter}
-\title{A multi-scale story of the diffusion of a new technology: the
-web}
-\author[1]{Emmanouil Tranos%
-\corref{cor1}%
-}
- \ead{e.tranos@bristol.ac.uk}
-
-\affiliation[1]{organization={University of Bristol, School of
-Geographical Sciences},addressline={University Road, BS8
-1SS},city={Bristol},country={UK},countrysep={,},postcodesep={}}
-
-\cortext[cor1]{Corresponding author}
-
-\begin{abstract}
-This paper investigates the spatial diffusion of a new technology that
-is the Web in the UK. It employs novel data and machine learning methods
-to model the influence of well-established diffusion mechanisms.
-Contrary to previous studies, it adopts multiple scales, high spatial
-granularity and a long study period that captures the early stages of
-the Web until its maturity (1996-2012). Findings reveal the importance
-of such spatial mechanisms (namely distance, urban hierarchy and the
-S-shaped pattern of the cumulative level of adoption) even at granular
-scales. They also highlight spatial heterogeneity and instances of
-leapfrogging.
-\end{abstract}
-
-\begin{keyword}
-    diffusion \sep technologies \sep Web \sep 
-    digital
-\end{keyword}
-\end{frontmatter}
-
-\section{Acknowledgements}\label{acknowledgements}
-
-Many thanks to Dr Levi Wolf and Professor Dani Arribas-Bel for their
-constructive comments. Also many thanks to the participants of the 6th
-Global Conference on Economic Geography and the Winter Seminar of GfR.
-Then, many thanks to the organisers of the seminar series and the
-members of the Quantitative Spatial Science (QuSS) Research Group at the
-University of Bristol, the Centre for Business Prosperity Research Group
-at Aston University (Dr Tasos Kitsos), and the Centre for Innovation
-Research (Professor Tom Broekel) at the University of Stavanger.
-Finally, many thanks to the editor and the three anonymous reviewers for
-the very helpful comments and suggestions.
-
-\section{Funding}\label{funding}
+# Funding
 
 No external funding was received for this work.
-
-\end{document}

### 2025-07-29 paper/submission4/tranos2025.qmd
submission 3

--- 2025-07-29 paper/submission4/anonymised.qmd
@@
 title: "A multi-scale story of the diffusion of a new technology: the Web"
+author:
+  - name: Emmanouil Tranos
+    email: e.tranos@bristol.ac.uk
+    affiliations: 
+      - id: some-tech
+        name: University of Bristol
+        department: 
+        address: 
+        city: 
+        state: State
+        postal-code: UK
+    corresponding: e.tranos@bristol.ac.uk
+address:
+  - attributes:
+        corresponding: true
+    note: This is the first author footnote.
 abstract: |

### 2025-07-29 paper/submission4/tranos2025.tex
submission 3

--- 2025-07-29 paper/submission4/anonymised.tex
@@
   pdftitle={A multi-scale story of the diffusion of a new technology: the Web},
+  pdfauthor={Emmanouil Tranos},
   pdfkeywords={diffusion, technologies, Web, digital},
@@
 Web}
+\author[1]{Emmanouil Tranos%
+\corref{cor1}%
+}
+ \ead{e.tranos@bristol.ac.uk}
+
+\affiliation[1]{organization={University of
+Bristol},postcode={UK},postcodesep={}}
 

### 2025-07-30 abstracts_notes/help.R
rm directories from cache

@@ -1,1801 +1,0 @@
-# from https://github.com/rsbivand/survival_katrina/blob/master/bivand%2Bgomez-rubio_SMij.R
-# linked to Bivand_Gomez-Rubio_2021.pdf
-
-### R code from vignette source 'bivand+gomez-rubio_SMij.Rnw'
-
-zz <- file("sunk_output.txt")
-sink(zz)
-sink(zz, type="message")
-
-###################################################
-### code chunk number 13: bivand+gomez-rubio_SMij.Rnw:242-254
-###################################################
-library(spatialprobit)
-data(Katrina)
-library(sf)
-sf_katrina <- st_as_sf(Katrina.raw, coords=c("long", "lat"))
-st_crs(sf_katrina) <- "+proj=longlat +datum=WGS84"
-is.na(sf_katrina$days) <- sf_katrina$days == 99999
-sf_katrina$Days <- as.Date("2005-08-29") + sf_katrina$days
-t <- table(sf_katrina$Days)
-diff_days <- diff(as.Date(names(t)))
-names(diff_days) <- names(t)[-length(t)]
-yrs <- format(as.Date(names(diff_days)), "%Y")
-# o
-
-
-###################################################
-### code chunk number 14: bivand+gomez-rubio_SMij.Rnw:265-268
-###################################################
-suppressMessages(library(xtable))
-mat <- table(yrs, diff_days)
-#print(xtable(mat, align=c("l","r","r","r","r","r","r","r"), digits=c(NA, 0, 0, 0, 0, 0, 0, 0), display=c("s", "d", "d", "d", "d", "d", "d", "d")), floating=FALSE, comment=FALSE, sanitize.text.function=function(x){x}, hline.after=c(-1,0,2))
-
-
-###################################################
-### code chunk number 15: bivand+gomez-rubio_SMij.Rnw:283-292
-###################################################
-sf_katrina$days_mod <- sf_katrina$days
-sf_katrina$days[is.na(sf_katrina$days)] <- 99999
-sf_katrina$y0_90 <- as.numeric(!sf_katrina$days < 90)
-sf_katrina$y0_180 <- as.numeric(!sf_katrina$days < 180)
-sf_katrina$y0_360 <- as.numeric(!sf_katrina$days < 360)
-sf_katrina$sliced <- (sf_katrina$y0_90 == 0) + (sf_katrina$y0_180 == 0) + (sf_katrina$y0_360 == 0) + as.numeric(sf_katrina$days < 99999)
[…]

### 2025-07-30 paper/submission1/help.R

@@ -1,35 +1,0 @@
-## Old ideas
-
-Hexagon density maps: reflect the spatial structure of Britain. Websites
-are associated in places where people live and work.
-
-**TODO** maps at the local authority level per firms and discuss
-patterns
-
-Neighbourhood effect: diffusion proceeds outwards from innovation
-centers, first "hitting" nearby rather than far-away locations (Grubler
-                                                                1990)
-
--   Moran's I: for OA and LAD over time
-
--   LISA maps: for OA and LAD over time More and less expected clusters.
-    Different scales show different results
-
--   Website density regressions: for OA and LAD over time **TODO** add
-    0s Similar pattern
-
-Hierarchy effect: from main centers to secondary ones -- central places
-
--   Gini coefficient. Almost perfect polarisation of web adoption in the
-    early stages at a granular level More equally diffused at the Local
-    Authority level Plateau overtime
-
-S-shaped diffusion curves
-
--   S for LADs per firm and OA.
-
--   Fast and slow LAs map. There is clustering
-
--   s_uk_firm_11
-
--   ranks: there is stability and movement

### 2025-09-28 paper/submission4/appendix.qmd
appendix

--- 2025-06-08 paper/submission3/tranos2025.qmd
@@
 ---
-title: "A multi-scale story of the diffusion of a new technology: the web"
-author:
-  - name: Emmanouil Tranos
-    email: e.tranos@bristol.ac.uk
-    affiliations: 
-      - id: some-tech
-        name: University of Bristol
-        department: 
-        address: 
-        city: 
-        state: State
-        postal-code: UK
-    corresponding: e.tranos@bristol.ac.uk
-address:
-  - attributes:
-        corresponding: true
-    note: This is the first author footnote.
-abstract: |
-    This paper investigates the spatial diffusion of a new technology that is the Web in the UK.
-It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms.
-Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012).
-Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales.
-They also highlight spatial heterogeneity and instances of leapfrogging.
-keywords: 
-  - diffusion
-  - technologies
-  - Web
-  - digital
+title: "A multi-scale story of the diffusion of a new technology: the Web"
 date: last-modified
-bibliography: bibliography
 format:
@@
     journal:
-      name: Journal Name
-      formatting: preprint
+      name: Journal of Economic Geography
       model: 3p
@@
 
-# Introduction {#sec-introduction}
+<!-- This file is for creating a stand alone appendix. It was created from tranos2025.qmd after deleting everything but the appendix part and adding an equation set counter.-->
 
-<!-- geographers lack of interest for diffusion + gap -->
-Economic geographers were always interested in how new technologies and
-innovations diffuse across space and time with the work of
-@hagerstrand1968innovation, which demonstrated the importance of
-communications as a mechanism for the diffusion of innovations, being
-the most impactful. It was followed by various empirical studies -- see
-for instance @ormrod1990 and @iso2005. However, the torch of exploring
-and modelling such diffusion processes had been passed to other
-disciplines such as economics, business studies and sociology well
-before the 'cultural turn' of economic geography
-[@perkins2005international; @ding2010modeling].
-<!-- NEW spatial granularity --> As a consequence, relevant studies
-did not investigate diffusion mechanisms at granular geographical scales to 
-provide local policy insights, for instance, regarding infrastructural 
-development and local technology policies. 
-A potential explanation of the lack of geographical studies exploring the 
-diffusion of new and, more specifically for this paper, digital technologies 
-across both space and time can be attributed to the scarcity of relevant and 
-granular enough data, a problem also highlighted by @iso2005 and
-@kemeny2011international. As @zook2022mapping pointed, digital
-activities are hardly ever captured in official data.
-
-<!-- contribution -->
-This paper offers such a contribution: an economic geography study
-revealing the importance of spatial mechanisms for the diffusion of a
-digital technology that is the Web in the UK at a high level
-of spatial granularity 
-<!-- NEW TO JUSTIFY DATA TIME PERIOD-->
-from the very early days until its maturity (1996-2012).
-<!-- NEW DATA CLARIFICATION -->
-It specifically focuses on the diffusion of commercial websites.
-It does so by employing a novel source of big data and machine learning
-algorithms to model the *active* engagement with Web -- that is creating and 
-maintaining instead of just browsing a website -- during that period
-and tests how different spatial and temporal mechanisms shaped its
-diffusion [@tranosuk]. This paper also exemplifies how the combination of data
-sources and state-of-the-art research methods which escape the
-traditional social science domain can offer new lenses to geographical
-research regarding the understanding of technological diffusion.
-
-<!-- motivation: stakeholders -->
-The motivation for this paper lies in the fact that there are various
-stakeholders who are interested in knowing how new digital technologies
-diffuse over space and time and use this knowledge to make predictions
-about the diffusion of related future technologies. As per
-@leibowicz2016representing, historical studies agree that technologies
-diffuse differently in terms of times, rates, and geographies and can be
-driven by related policies [@victor1993]. @meade2021modelling highlight
-that a variety of actors have a direct interest in gaining such
-knowledge including network equipment suppliers, network operators,
-regulatory and local authorities. These processes and their effects vary
-a lot across scales: although the diffusion of a new technology might
-not be optimal at a local level, it might be beneficial from a global
-perspective as it could lead to faster diffusion to less advantaged
-places [@leibowicz2016representing]. Despite the spatial heterogeneity
-of such diffusion mechanisms and the policy relevance, there is a
-scarcity of studies analysing the diffusion of new technologies
-at a detailed geographical level because of the lack of detailed enough
-data and adequate research methods.
-
-<!-- motivation: regional development -->
-This gap becomes even more prominent if we consider the importance of
-technology and its adoption in economic geography and macroeconomic
-growth theories. Simply put, differences in technological adoption and
-sophistication may lead to uneven development as new technologies can
-increase productivity and allow firms to introduce new products
-[@solow1957technical; @aghion1990model; @kemeny2011international].
-Linked to the subject of this paper, the Web and, more specifically, 
-commercial websites can be considered as the underpinning
-technology for the transformative process identified by
-@capello2024nexus as the *digital service economy*. Importantly, the
-focus here is not the Web in general, but instead commercial websites.
-After all, diffusion together with invention and innovation are
-considered the pillars of technological change [@das2022diffusion].
-Findings from previous studies illustrated that early engagement with
-digital technologies such as the Web led to longstanding positive
-productivity gains for these regions that hosted early adopters
-[@tranosuk]. Hence, obtaining a robust understanding of the spatial
-diffusion mechanisms is crucial in mapping future development trajectories.
-
-<!-- NEW motivation: intangible/digital -->
-Another line of motivation has to do with the very nature of the Web as a 
-digital technology and how it evolved over time. The digital geographies field 
-has debunked premature claims about the Internet's placeless-ness by mapping 
-various of its facets and technological layers [@tranos2013death], but the lack 
-of relevant long time-series data prevented researchers from observing digital
-phenomena overtime. Most of these studies focused on its hard infrastructure 
-[e.g. @malecki2002economic; @moss2000internet], on Internet subscribers 
-[e.g. @blank2018local], or on social media users [e.g. @crampton2013beyond].
-Exceptions include early work from @zook2000web, @zook2001old and 
-@moss1997tracking, which utilised domain name registration data and use it to
-map and illustrate the spatial footprint of the web content. Pivotal as these 
-contributions have been in understanding the economic geographies of the 
-Internet, they were based on cross-sectional, often spatially detailed, mapping 
-and on short temporal comparisons (e.g. 1993-1998) and did not observe the 
-evolution and diffusion of the Web until its maturity. As we know from the 
-relevant literature, different technologies diffuse at different rates and 
-there is no theory that can explain these differences 
-[@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
-Web can offer useful insights.
-
-<!-- R2 addition -->
-The lack of previous granular studies about the diffusion of new technologies 
-deprives us from a reliable prior regarding whether the diffusion mechanisms 
-and drivers are sensitive or not to the spatial scale that the diffusion of a 
-new technology is observed. As per other key concepts in economic geography
-[@briant2010dots; @meliciani2015determinants; @andini2019social; @rosenthal2001determinants],
-diffusion mechanisms can indeed suffer from Modifiable Areal Unit Problem 
-(MAUP): although the adoption of a new technology is at its core an individual 
-process, it is hardly ever observed at this, or at a close enough, scale.
-
-<!-- early engagement with the web -->
-This paper focuses on the diffusion of the Web as new technology during
-the 1996-2012 period. This was an exciting period for digital
-technologies as it corresponds with the commercialisation of the
-Internet and, consequently, its almost universal adoption. The reader is
-reminded that it was only in 1994 when Netscape Navigator was
-introduced, a year before Microsoft's Internet Explorer.[^1] Also, only
-9 per cent of UK's households had access to the Internet in 1998
-[@ons2018], the web included mostly static webpages, there were no
-social media and web browsing involved exclusively a desktop PCs as
-there were no smartphones [@tranosuk]. Hence, it is fair to say that the
-study period captures the very early stages of the diffusion of a new
-technology that is the Web until its maturity. The former is a key point
-in the lifecycle of a new technology. Firstly, during this period new
-technologies are expensive, crude and imperfect
-[@rosenberg1994exploring; @wilson201281]. A simple comparison between
-Web 1.0 and Web 3.0 applications clearly illustrates this
-[@tranos2020social]: while a static website facilitates one-way
-information dissemination, a platform like `github` enables cooperation
-between users and the creation of new information, meaning, and even
-knowledge [@faraj2016special; @barassi2012does]. During this period the
-performance of a new technology is the main attraction and not the cost
-to access and use it [@wilson2011lessons]. There is a broader
-theoretical discussion in the literature about the early adoption
-motives. As summarised by @perkins2005international, on one hand,
-epidemic models highlight the role of interpersonal contacts as a way
-for new technologies to diffuse. On the other hand, economic models
-underline the importance of heterogeneity. Different firms have
-different structures and business plans, which define the potential
-economic returns of the adoption of a new technology and, therefore, the
-choice to adopt a new technology becomes an individual option. From a
-broader and evolutionary perspective, initial conditions are essential
-for the creation and evolution of path-dependent technological
-development trajectories [@neffke2011regions; @simmie2014new]. This
-argument is even more relevant when the focus is on digital technologies
-because of the commonly found lag between investment and economic
-returns as reflected in the Solow paradox [@acemoglu2014return;
-@brynjolfsson2018artificial].
-
-[^1]: <https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft>
-
-<!-- data -->
-The data used here depict the active engagement with the Web as they
-contain geolocated and time-stamped counts of commercial websites from
-the UK. Instead of adopting metrics reflecting the passive engagement
-with the Internet such as Internet adoption and speeds [e.g.
-@blank2018local; @destefano2022fuzzy], this paper observes yearly counts
-of geolocated, commercial websites, which reflect the creation and 
-maintenance of such websites. It needs to be highlighted here that although 
-there are various types of underpinning web technologies -- namely markup,
-stylesheet, client- and server-side coding languages as well as
-different web browsers and other related programming language -- the
-focus here is on the diffusion of the Web as an overarching technology
-instead of more specific web technologies. For a such a detailed and 
-cross-country technological diffusion study, see @PAPAGIANNIDIS2015308.
-
-<!-- description of the methods and results -->
-To analyse these data, a diverse set of methods are employed. Firstly,
-cumulative adoption curves are estimated for the UK as a whole and for
-the Local Authority Districts (LAD), illustrating, for the first time,
-such S-shaped diffusion patterns for local areas. The exploratory spatial data
-analysis offers some first indications of the importance of the different 
-spatial diffusion mechanisms in shaping the diffusion of the Web in the UK.
-Importantly, the analysis adopts two distinct geographical scales: the LAD and 
-the most detailed scale of UK census areas, the Output Areas (OA). Then, a
-novel methodological framework based on Machine Learning (ML) algorithms
-is developed to test the role of such mechanisms at these two scales. The 
-results reveal, for the first time, the importance of different spatial 
-diffusion mechanisms in the diffusion of a digital technology at very granular
-geographical scales. The analysis also reveals that the diffusion of the Web 
-is characterised by spatial heterogeneity, well-established spatial patterns 
-and both stability and volatility.
-
-<!-- contents -->
-The structure of the paper goes as follows. @sec-litreview reviews the
-mechanisms that shape the diffusion of a new technology and sets up the two 
-hypotheses that are empirically tested in @sec-results. @sec-datamethods 
-describes the data and methods used in this paper. The discussion of the 
-results and the conclusions of the paper are offered in @sec-conclusions.
-
-# The mechanisms that shape the diffusion of a new technology {#sec-litreview}
-
-<!-- The fundamentals of technology diffusion -->
-The diffusion of new technologies and innovations is driven by three
-different mechanisms [@grubler1990rise; @morrill2020spatial]: (i)
-distance and proximity, (ii) urban hierarchies in the form of a
-centrifugal forces as new technologies diffuse from core to periphery,
-and (iii) over time the cumulative level of adoption follows an S-shaped
-pattern.
-<!-- R2 addition -->
-Although some studies focused on other, case-specific drivers -- for example 
-@lengyel2020role on urban scaling and @bednarz2020pulled on regional demand --
-the literature is in agreement that these are the three fundamental diffusion 
-mechanisms. 
-<!-- neighbourhood -->
-@hagerstrand1968innovation was the first to identify diffusion as a
-geographical process. The starting point was the idea that diffusion is
-based on passing information through social networks, which themselves
-tend to be geographically bounded. Hence, he identified a
-'neighbourhood' mechanism of how information, and consequently,
-innovations and new technologies diffuse. This perspective draws
-similarities with epidemics: new technologies and innovations, just like
-pathogens, spread because of contagion and, consequently, proximity and
-exposure [@hivner2003facilitating].
-
-<!-- hierarchy -->
-@hagerstrand1968innovation also emphasised the role of the top-down
-hierarchy of urban systems and highlighted that new technologies and
-innovations are firstly adopted in larger cities and then diffuse to
-second tier ones. This is a sequential instead of a simultaneous
-process, which resembles the 'lead-lag' spatial acceleration effect in
-market research [@bento2018time; @PERES201091]. Schmidt's Law
-empirically illustrates a similar pattern [@grubler1990rise]. Core and
-highly agglomerated regions is where new technologies are invented and
-commercially deployed. This is where the first adopters tend to be
-based. Then, technologies spread to the rim and eventually to the
-periphery. Although adoption pace might be higher when new technologies
-finally arrive to the periphery, the saturation levels may never
-reach the ones in the core because of the lack of infrastructure or
-other necessary institutions [@leibowicz2016representing].
-
-<!-- time -->
-Diffusion is naturally a dynamic process and Hägerstrand recognised the
-role time plays in the diffusion of innovations and new technologies: an
-early-pioneering period, a middle fast accelerating period and a final
-saturation one [@morrill2020spatial]. This was further explored by
-@rogers2010diffusion, who identified five types of adopters of new
-technologies and innovations: innovators, early adopters, early
-majority, late majority and laggards. Hence, plotting the cumulative
-adoption of a new technology over time often leads to an S-shaped line
-plot [@grubler1990rise].
-
-Although the above processes have been studied in innovation studies,
-the distancing of economic geographers from research questions about the
-diffusion of new technologies and innovations resulted to a knowledge gap
-regarding the explanatory power of these mechanisms at granular, local
-scales [@ding2010modeling]. As discussed in @sec-introduction, such 
-spatially explicit knowledge can directly inform local technology
-and infrastructure policies. Indeed, while the spatial dimension is
-present in the following studies, the geographical level of analysis is
-often too coarse to ascertain the role of the above mechanisms at the
-local scale. @beardsell1999spatial studied the evolution of the computer
-industry in 317 US metropolitan areas during the 1977-1992 period using
-employment data. Their analysis indicated that the relative size
-distribution holds for urban computer employment and also urban
-heterogeneity is essential in explaining this distribution. In a recent
-study, @bednarz2020pulled focused on wind turbines and modelled their
-spatial diffusion across 402 German regions during 1970-2015. Their key
-finding is that local demand than local supply was the main driving
-factor. @haller2011determinants employed firm panel data for Ireland
-during 2003-2005 and illustrated that when firms are located in a region
-or industry with high share of firms having a website, they are more
-likely to have a website too. At a global scale @perkins2011internet
-explored whether the adoption of previous communication technologies
-that is mail, telegrams and telephones were shaped by similar
-socioeconomic factors as the Internet. Their results indicated common
-patterns regarding the drivers behind the adoption of different
-communication technologies. @feldman2015rknowledge focused on inventions
-and used patent data to map how the the rDNA technology diffused among
-inventors across 366 US metropolitan areas. @ding2010modeling modelled
-the spatial diffusion of mobile telecommunications across the highest
-level of administrative division in China (29 regions in total). Their
-analysis indicated that socioeconomic characteristics are important
-determinants of the timing, speed and the level of mobile diffusion
-within China. @bakher2013diffusion compared different growth curves
-as a means to model the diffusion of online shopping in Australia during
-the 1998-2009 period, but they did not consider any spatial mechanisms.
-At a global scale @PAPAGIANNIDIS2015308 modelled the diffusion of
-different web technologies and practices across countries. Using similar
-but less extensive data as the one used here, their analysis illustrated
-how the diffusion of different web technologies and practices follows an
-S-shaped pattern as well as the different diffusion rates of the
-different technologies and practices. The most detailed consideration of
-the spatial footprint of the diffusion of a new technology comes from
-@lengyel2020role, who analysed the adoption and churn of a Hungarian
-online social network. Their results were in agreement with the early
-theoretical and empirical contributions reviewed here: assortativity,
-urban scaling and distance are the key drivers of spatial diffusion.
-Based on the above, the main empirical hypothesis that is tested in this paper
-states:
-
-> H~1~: The three diffusion mechanisms namely distance, urban hierarchy
-and the S-shaped pattern of the cumulative level of adoption shaped
-the diffusion of the commercial Web in the UK even at granular geographical 
-scales.
-
-<!-- EARLY/LATE adoption-->
-Understanding the diffusion of a new technology in multiple, detailed spatial
-scales is of interest to economic geography as potential local economic 
-benefits may occur because of concentration of early adopters or through 
-technological leapfrogging. On the one hand, early adopters are often rewarded 
-because of their attitudes towards new technologies and innovations. 
-@rogers2010diffusion identifies them as 'knowledgeable risk takers' and 
-@griliches1957 as 'profit maximisers' [@ding2010modeling]. On the other hand,
-latecomers can adopt and benefit from new technologies that have been developed
-elsewhere without incurring the hefty initial R&D costs [@teece2008firm]. At a 
-global scale @perkins2005international explored whether the diffusion rate of 
-new technologies is driven by a latecomer advantage and the engagement with the 
-global economy via foreign direct investments and trade. Their results 
-illustrate that indeed latecomers and developing countries experience diffusion 
-of new technologies more rapidly than early adopters and developed countries. 
-At a regional scale, previous research highlighted the long term and 
-sustained productivity benefit of the early adopters of web technologies in 
-the UK [@tranosuk]. Although the early adoption and leapfrogging literature has 
-not paid much attention on cities and regions [@yu2018sustainability], the 
-potential local economic benefits justify the need to investigate the
-dynamics of the web diffusion in different scales in the UK. Hence, the second 
-empirical hypothesis states:
-
-> H~2~: The diffusion of the commercial Web is shaped by both early adoption 
-and latecomer effects.
-
-These two hypotheses are empirically tested in @sec-results employing 
-the data and methods described in the next section.
-
-<!-- Technological diffusion, which is by definition an aggregated process, -->
-<!-- can be discussed in parallel with individual adoption mechanisms. On the -->
-<!-- one hand,  -->
-<!-- On the other hand, @perkins2011internet attribute diffusion -->
-<!-- to two processes: (i) epidemic-like mechanisms, which are governed by -->
-<!-- distance, proximity and social interactions, and (ii) economic -->
-<!-- mechanisms as new innovations are adopted by users as they become more -->
-<!-- profitable, valueable and useful. -->
-
-# Data and Methods {#sec-datamethods}
-
-<!-- main data -->
-To capture the diffusion of the commercial Web, a website density metric is
-developed for two different geographical scales that are discussed later in 
-this section. The main data are counts of commercial websites that are 
-calculated using data from the Internet Archive[^2] and, specifically, the 
-JISC UK Web Domain Dataset [@ukwebarchive]. The Internet Archive is one of the 
-most complete and oldest archive of webpages in the world operating since
-1996 [@ainsworth2011much; @holzmann2016dawn]. It is a web crawler, which
-discovers webpages by following the hyperlinks of every webpage it
-archives. This dataset, which is curated by the British Library,
-contains all the archived webpages from the UK country code top level
-domain (ccTLD - .uk) from the 1996–2012 period. In essence, this is a
-long list of 2.5 billion URLs of archived webpages including also the
-archival timestamp.
-
-[^2]: [https://archive.org/](https://archive.org/).
-
-Instead of using the whole .uk ccTLD, this paper focuses on its
-commercial subset, the .co.uk second level domain (SLD). This choice
-is aligned with the topic of the paper -- the diffusion of the Web for 
-commercial functions -- and decreases the heterogeneity of the web data as such 
-commercial websites have specific aims: they are used to diffuse information, 
-support online transactions and share opinions [@THELWALL2000441; 
-@blazquez2018big]. Although a UK company can adopt a generic TLD such as .com 
-and these cases escape the data used here, such omissions should not affect the
-validity of the results given the popularity of the .uk ccTLD
-[@tranosuk]: UK consumers prefer to visit a .uk website when they are
-searching for products or services [@hope]; and anecdotal evidence
-indicates that during the first half of 2000, three .co.uk domains were
-registered every minute [@oecd_coms]. Importantly, previous studies
-illustrated that .co.uk is the most popular UK SLD [@tranosuk]. Henceforward 
-any reference to websites will refer to commercial -- .co.uk -- websites.
-
-The text from these webpages was scanned using a regular expression
-(regex) to identify strings of text which resemble UK postcodes and one
-fifth of them included a mention to a postcode [@BL2013geo]. This
-information allows the geolocation of the data and the creation of the
-LAD and OA counts. To prevent false positives, the resulted postcodes
-were matched against official postcode lookup files [@ons].
-
-The data cleaning process includes an aggregation step, through which
-the archived webpages are aggregated to the parent websites. This
-website reconstruction process lend itself to the construction of
-*website* instead of *webpage* density metrics. Websites tend to
-represent specific organisations or entities, and, arguably, are more
-meaningful observational units than webpages, which can ignore the
-upstream dependency of the website they belong to. Based on the
-following example, all three webpages are part of the same website
-(<http://www.website.co.uk>). Only the overall website and not the
-nested webpages are considered for the counts as otherwise the density
-metrics would have been biased towards large, place specific, websites.
-It is important to note here that websites commonly contain place references, 
-with previous research estimating this share to be around 70% 
-[@hill2009georeferencing].
-
--   <http://www.website.co.uk/webpage_a> B15 2TT
-
--   <http://www.website.co.uk/webpage_b> BS8 1TH
-
--   <http://www.website.co.uk/webpage_c> B15 2TT
-
-\noindent What is challenging is that this aggregation approach, which
-has been used elsewhere [@tranosuk; @shoreditch] may lead to websites
-with multiple postcodes. As per the above example,
-[www.website.co.uk](http://www.website.co.uk) includes two unique
-postcodes: B15 2TT and BS8 1SS. The distribution of the number of
-postcodes per website for 2000 is presented in Table \ref{f2000}, which
-illustrates a wide range. At the left end of the distribution, there are
-websites anchored to a unique location (72% of all the reconstructed
-websites in 2000), which may represent a small company with a single
-trading location. At the right end, there are websites with thousands of
-different postcodes. Considering the time period of the analysis, such
-cases can represent directories which used to be popular in the
-pre-search engines early times of the commercial Internet as well as
-real estate websites [@tranosuk]. Previous research which used the same 
-methodology for a small subset of these data illustrates the robustness of the 
-above approach: its results matched in terms of accuracy past rounds of 
-research based on extensive qualitative work, interviews and observational 
-studies for the same study area [@shoreditch].
-
-The analysis is firstly conducted on websites, which only contain one
-unique postcode. As a robustness check, the analysis is replicated for
-an extended subset of websites, which include up to 10 unique postcodes
-to capture commercial websites which point to multiple locations. They
-are geolocated by equally attaching them to these locations. This
-extended sample includes 94 percent of all the archived websites in
-2000 (see Table \ref{f2000}).
-
-```{r eval=T, echo=FALSE, results='asis'}
-
-# load 2000 co.uk
-# path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"
-# path.2000 <- "/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/all2000couk.csv"
-path.2000 <- paste0(path, "/data/temp/all2000couk.csv")
-
-all2000.duplicates <- data.table::fread(path.2000) #
-# dim(all2000.duplicates) # 3336162
-# it includes duplicated URLs: if one webpages includes multiple postcodes
-# then it appears multiple times. This is ok for the nuts aggregation, but
-# not for the frequencies
-# This is only .co.uk
-
-# one line for every host
-all2000 <- unique(all2000.duplicates, by = c("host")) # 57897
-
-# unique postcodes per website f table
-f.websites.pc <- DescTools::Freq(all2000$V1, breaks = c(0, 1,2, 10,100,1000,10000,100000), ord = "desc")
-f.websites.pc$level <- factor(f.websites.pc$level, levels = c("[0,1]","(1,2]", "(2,10]", "(10,100]",
-                                                              "(100,1e+03]", "(1e+03,1e+04]",
-                                                              "(1e+04,1e+05]"))
-levels(f.websites.pc$level) <- c("(0,1]","(1,2]", "(2,10]", "(10,100]", "(100,1000]", "(1000,10000]", "(10000,100000]")
-
-f.websites.pc <- f.websites.pc %>% 
-  dplyr::select(-cumfreq) %>% 
-  rename(Postcodes = level,
-         F = freq,
-         'F (%)' = perc,
-         'cumulative F' = cumperc)
-
-kable(f.websites.pc,
-      format = "latex",
-      digits = 3,
-      booktabs = T,
-      format.args = list(big.mark = ","),
-      caption = "Number of unique postcodes per .co.uk website, 2000.\\label{f2000}") %>%
-  kableExtra::footnote(general = "Tranos et al. 2021",
-                       general_title = "Source: ",
-                       footnote_as_chunk = T)  
-   # kable_classic(full_width = F, html_font = "Cambria")
-```
-
-Another data cleaning step deals with some extreme outliers. Figure
-\ref{correct} plots the website counts for the top 1000 postcodes. Some
-obvious outliers can be observed for the 2002-2006 period for a handful
-of postcodes, which can be attributed to link farms [@BL2013links]. The
-website counts for these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1
-6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1,000
-websites pointing to them, were replaced with predicted values based on
-a simple panel regression model with postcode fixed effects and yearly
-dummy variables. These postcodes refer to a residential area with a
-small park in South London, residential areas outside Warwick and
-Manchester, a rural area in the south border of England and Wales and
-small business parks in Crew and Derby. Hence, it is difficult to
-believe that these postcodes genuinely hosted such a large amounts of
-websites, which were later extensively declined. To put the magnitude of
-the data imputation into perspective, this process affected 6 out of the
-557,808 postcodes included in the data. As @sec-results illustrates,
-this led to a small increase of the model predictive power.
-
-<!-- Multi-scale -->
-To empirically test the two hypotheses stated in @sec-litreview, the website 
-density metric was calculated for two different scales the Local Authority
-Districts (LAD) and the Output Areas (OA). The former is the lowest
-administrative unit in the UK and there are 374 LAD. The latter is the smallest
-census-based geographical unit and there are c. 230,000 of them in the UK. To 
-create the website density metric, the yearly website counts at the LAD level 
-are standardised by the number of firms in LAD to avoid biases associated with 
-LAD hosting a large number of firms. Given that there is no such readily 
-available statistic for the OA, the actual OA level counts are used. However, 
-because of the the consistent spatial definition of OA (they host 40-250
-households),[^3] website counts in OA are interpreted as a density metric too.
-
-[^3]: <https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies>
-
-<!-- S -->
-These website density metrics are used for the different stages of the analysis. 
-Starting with the first hypothesis (H~1~), to test whether the S-curve 
-accurately describes the cumulative adoption of the Web at a local scale, the
-following logistic function (Equation \ref{eq:s}) is estimated for the whole of 
-the UK and then for each LAD separately:
-
-\begin{align}
-y = k /(1 + e^{-(t-t_{o})})\label{eq:s}
-\end{align}
-
-\noindent $k$ is the asymptote or, in other words, the saturation level,
-$b$ the overall growth rate, and $t_{0}$ the *inflection point* of
-maximum growth at $k/2$, where the logistic function is symmetrical
-[@wilson201281]. The $t_0$ of each LAD is compared against the $t_0$ of
-the UK to delineate whether a LAD reached that point faster or slower
-than the country average. This heterogeneity is discussed further when testing 
-the second hypothesis (H~2~). Importantly, an accuracy criterion was imposed
-and only LAD with $R^2 > 0.9$ were included in this analysis. To
-estimate Equation \ref{eq:s} a self-starting logistic growth model was
-employed using the `nls` and `SSlogis` functions in `R`. This logistic
-growth model depicts the initial period of slow diffusion of a new
-technology, which is followed by the fast, exponential growth stage to
-then eventually slow down and saturate [@wilson201281;
-@grubler1999dynamics].
-
-<!-- **minimum R2 = 95%** in @wilson201281, see also @grubler1990rise -->
-<!-- The literature usually uses the saturation level as the asymptote. I am -->
-<!-- using the total number of websites as we cannot compute a rate. -->
-
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{correct}Yearly website counts per postcode (top 1000) before (top) and after (down) data imputation", out.height="70%"}
-path.image1 <- paste0(path, "/outputs/pc_year_1.png")
-path.image2 <- paste0(path, "/outputs/pc_year_1_corrected.png")
-
-p1 <- ggdraw() + draw_image(path.image1)  
-p2 <- ggdraw() + draw_image(path.image2)
-
-par(mar=rep(0,4)) # no margins
-
-p1 + plot_spacer() + p2 + plot_layout(ncol = 2)#, widths = c(1, -0.8, 1))
-```
-
-<!-- ESDA -->
-Next, exploratory spatial data analysis provides some first insights of whether 
-the other two drivers of spatial diffusion -- namely neighbourhood and 
-hierarchy -- underpin the diffusion of the Web in the UK. To capture the former 
-and following @ding2010modeling the Moran's I and the Local Index of Spatial
-Association (LISA) are estimated for the website density. To capture the
-hierarchy mechanism, the Gini coefficient -- an established metric of
-inequality -- is calculated. All the above are computed and plotted
-longitudinally both for the LAD and the OA.
-
-<!-- conceptual model -->
-To formally test H~1~ a modelling framework is developed to assess the 
-above diffusion mechanisms. The overarching aim is to build a model that can 
-test how well these three mechanisms can predict website density:
-
-\begin{align}
-Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S\text-curve_{t}\label{eq:rf.generic}
-\end{align}
-
-<!-- <!-- conceptual model --> 
-<!-- The models include variables that capture the three processes that the -->
-<!-- relevant literature and the descriptive analysis highlighted. Namely, -->
-<!-- the models capture: (i) a hierarchical mechanism with diffusion running -->
-<!-- from main centres to secondary ones, (ii) a neighborhood mechanism -->
-<!-- according to which diffusion first hits nearby locations, and (iii) the -->
-<!-- rather canonical pattern of diffusion over time as reflected in the -->
-<!-- S-shaped pattern in the cumulative level of adoption. -->
-
-<!-- detailed features -->
-To capture the hierarchical mechanism the model includes as predictors a
-one year lag of website density in London, the largest city in the UK, and a
-one year lag of the website density in the nearest city. Both variables clearly
-illustrate the urban hierarchy in the UK. Especially for the OA models, a 
-third predictor is added to capture hierarchy at a much more granular scale:
-a one year lag of the website density in the the nearest retail centre [^4]. As
-the literature illustrates, retail centres are key components of urban systems
-[@dennis2002central; @jones2021reframing]. Due to the small sizes of the retail 
-centres, the latter is only relevant for and included in the OA models and not 
-in the LAD models. By including this predictor, the model is better tuned 
-towards the spatial characteristic of the OA scale. Hence, it increases the 
-robustness of the model.
-
-In addition, the model includes the distance to London, the 
-nearest city and the nearest retail centre. The underlying logic is that the 
-level of website adoption in a spatial unit depends on the level of the 
-adoption in places further up in the urban hierarchy the previous year. The
-inclusion of the distance variables incorporates spatial structure into
-the hierarchy argument. To depict the neighbourhood mechanism, the
-website density of the neighbouring spatial units in the previous year
-is employed. This is calculated using rook continuity. Again, the
-underpinning rationale is that the level of web adoption within a
-spatial unit depends on the level of web adoption in the neighbouring
-spatial units the year before. This represents the 'hitting nearby
-locations first' argument. Therefore, the spatial and temporal lag of
-the website density in LAD and OA is calculated. Lastly, the temporal
-pattern of the cumulative adoption is captured by a time trend variable.
-Hence, the model will follow the following generic form (Eq.
-\ref{eq:model}):
-
-[^4]: The retail centres data have been obtained by the Consumer Data Research Centre, an ESRC data investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1: <https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators>.
-
-\begin{align} \label{eq:model}
-Website\,Density_{t} \sim Distance\,London +
-Website\,density\,London_{t-1} +\notag\\
-Distance\,Nearest\,City +
-Website\,density\,Nearest\,City_{t-1} +\notag\\
-Distance\,Nearest\,Retail_{i} +
-Website\,density\,Nearest\,Retail_{t-1} +\notag\\
-W*\, Website\,density_{t-1} +\notag\\ 
-year_{t}
-\end{align}
-
-<!-- 4 models -->
-Specifically, four different models are estimated. Firstly, all the data points 
-for the OA and LAD are utilised in order to build two models and assess their
-capacity to predict the adoption of the Web at the two different scales. These 
-two models will reveal the predictive capacity of the diffusion mechanisms and 
-also demonstrate how the importance of such variables changes across
-different scales. The next two sets of models are again trained at these
-two scales: OA and LAD. However, instead of using all the data points,
-the OA and the LAD from one of the twelve UK regions are held out for
-the model training. Then, the trained model is used to predict website
-density for the OA or the LAD of the held-out region. This process takes
-place recursively for all twelve UK regions. The regional differences in
-the predictive capacity of the different samples will reveal how
-dissimilar these spatial process are across regions and, importantly, at
-different scales.
-
-<!-- RF -->
-Equation \ref{eq:model} is estimated using Random Forest
-(RF). This is a popular ML algorithm for both regression and
-classification problems [@biau2012analysis]. It was introduced by
-@breiman2001random and has become a go-to data science tool. RF can
-effectively handle skewed distributions and outliers, model non-linear
-relationships, require minimal hyperparameter tuning, exhibit low
-sensitivity to these parameters, and have relatively short training
-times [@Caruana2008; @liaw2002classification; @yan2020using]. These
-attributes match well with the website density data characteristics
-including skewness especially for the OA. Also, the large data size (c.
-230k data points for each of the 17 years) calls for fast training
-times. Importantly, RF predictions tend to be more accurate than those
-from single regression trees and outperform Ordinary Least Squares in
-out-of-sample predictions, even with moderate-sized training data and a
-small number of predictors [@mullainathan2017machine; @athey2019machine;
-@sulaiman2011intelligent; @pourebrahim2019trip; @biau2012analysis].
-
-<!-- RF -->
-RF is a tree-based ensemble learning algorithm [@breiman2001random]. It
-begins by generating random samples of the training data, which are then
-used to grow regression trees to predict the dependent variable. Data
-points and predictors are randomly sampled for the different trees. The
-trees are trained in parallel using their own bootstrapped samples of
-the training data. A crucial feature of RF is their ability not to
-overfit, meaning they can generalize well to unseen test data. While
-each tree may overfit individually, the ensemble of trees does not
-because the errors of individual trees are averaged, reducing the
-overall variance and preventing overfitting [@last2002improving]. For
-regression problems, RF predictions are made by averaging the
-predictions of all decision trees.
-
-RF have been widely employed to address regression research problems.
-@pourebrahim2019trip combined a spatial interaction modelling framework
-with ML algorithms including RF to predict commuting flows in New York
-City. @sinha2019assessing advocated for adopting spatial ensemble
-learning approaches, such as RF, to model spatial data with high
-autocorrelation and heterogeneity. @creditspatial predicted employment
-density in Los Angeles using spatially explicit RF.
-@guns2014recommending used RF to build a recommendation system for
-research collaborations. @ren2019predicting trained RF to predict the
-socio-economic status of cities using various online and mobility
-predictors. @tranos2023using utilised hyperlinks data and RF to make
-out-of-sample predictions of interregional trade. @zhou2023geography
-employed such a framework to assess whether key predictors of obesity
-differ across English cities.
-
-<!-- CV -->
-It needs to be highlighted here that the cross validation (CV) for all
-models is spatially and temporally sensitive. Instead of using 10 random
-and space- and time-agnostic folds, the `CAST` package is employed which
-allows holding back data points from specific years and spatial units (regions)
-and use them for testing in order to estimate the model performance
-[@meyer2018improving].
-
-<!-- assessment -->
-To assess the predictive power of the model, three broadly utilised
-metrics are employed: the coefficient of determination ($R^2$), mean
-absolute error (MAE) and root mean square error (RMSE):
-
-\begin{align}
-R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
-\end{align}
-\begin{align}
-MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
-\end{align}
-\begin{align}
-RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
-\end{align}
-
-\noindent $y_{k}$ is the $k^{th}$ observation of the dataset, which
-consists of $N$ observations in total. $\hat{y_{k}}$ is the $k_{th}$
-predicted value for the dependent variable and $\overline{y_{k}}$ is the
-average value of $y$. The last two metrics are expressed in the same
-units as the dependent variable -- websites per firm for the LAD modes
-and the number of websites for the OA models -- while the first one is
-the coefficient of determination between the observed and the predicted
-values of website adoption. Regarding $MAE$, it is the absolute
-difference between the observed and the predicted website adoption.
-While $MAE$ does not penalise for large errors, $RSME$ does, as it is
-proportional to the squared difference between the observed and the
-predicted trade flows. Hence larger errors weigh more for $RMSE$
-[@pontius2008components].
-
-An alternative specification could model the growth of website density. However, 
-such a strategy is problematic because of the large frequency of 
-OA without websites, which is to be expected given their small size and high 
-granularity. Still, as a robustness text, the results of a growth model for
-LAD are presented in the Appendix.
-
-<!-- H2 -->
-Lastly, to test the second hypothesis (H~2~) about the early adoption and the
-latecomer effects, a different research framework is developed. Firstly, the 
-S-curve and the estimated $t_0$ for every LAD are employed and mapped in order 
-to depict the spatial heterogeneity and specific patterns of the speed of the 
-diffusion of the Web. Then, the volatility and the stability of the diffusion 
-across the different LAD is analysed using tools from the complex systems domain 
-[@Batty2009] namely changes in rankings of LAD based on website density. Both 
-the S-curve estimation and the volatility analysis focus only on LAD as the 
-very large number of OA would have made such analysis difficult to visualise 
-and interpret.
-
-# Results {#sec-results}
-
-## The three diffusion mechanisms at granular geographical scales
-
-<!-- S-shaped -->
-Figure \ref{s_uk} plots the cumulative adoption of the Web in the UK as a whole
-during the 1996-2012 period. In line with the theoretical expectations discussed
-in the previous section, the cumulative adoption of the Web is well represented 
-by a curve that resembles the letter S. The vertical line in the beginning of 
-2003 (2003.158) illustrates the point when the modelled cumulative adoption was 
-equal to 50% of the maximum. This $t_0$ point is known as the *inflection* point 
-and signals the maximum adoption speed. The results remain unchanged when the 
-extended subset of websites, which include up to 10 unique postcodes is used
-(see Figure \ref{s_uk10} in the Appendix). Importantly though, the extended
-subset of websites indicates a saturation level close to 100% at the end of 
-the study period.
-
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk}Cumulative adoption of the Web, UK"}
-path.image <- paste0(path, "/outputs/s/s_uk_per_firm.png")
-knitr::include_graphics(path.image)
-```
-
-What escaped from previous studies was εexploring whether such a pattern also 
-describes cumulative adoption of new technologies at sub-national scales. To 
-test this, the same  logistic function is estimated for each of the 374 LAD. The
-detailed results are presented in Table \ref{table.s.lads} in the Appendix. 
-The key finding is that for 342 LAD (91% of all LAD) the $R^2 > 0.9$ for the 
-logistic function, which indicates that even at local scale the cumulative 
-adoption of the Web is characterised by a pattern similar to the one observed 
-for countries. Graphically, all these LAD curves are plotted in Figure 
-\ref{s_lad_uk}. The different colour lines illustrate whether a LAD reached $t_0$ 
-faster or slower than the UK in total. This heterogeneity is further discussed 
-later in this section.
-
-This is the first time that such cumulative adoption curves have been estimated
-for areal units as small as the UK LAD. Understanding and modelling the dynamics
-of technological diffusion and, ultimately, being able to predict diffusion 
-rates at such small local scales can directly support the deployment of 
-infrastructural networks (e.g. 5G) and local technology policies. The above are 
-in line with previous studies exploring the diffusion of web technologies 
-[@PAPAGIANNIDIS2015308], social media [@lengyel2020role] and online
-shopping [@bakher2013diffusion].
-
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_lad_uk}Cumulative adoption of the Web, LAD"}
-path.image <- paste0(path, "/outputs/s/lad_per_firm_all.png")
-knitr::include_graphics(path.image)
-```
-
-<!-- Neighbourhood effect -->
-To assess the second diffusion mechanisms, Figures \ref{morani} and \ref{lisa} 
-offer a first insight into whether a neighbourhood or, in other words, a 
-distance mechanism underpins the diffusion of the Web in the UK as they present 
-the Moran's I and the LISA maps of website density respectively for LAD and OA.
-Starting from the former, spatial autocorrelation was higher in the beginning of 
-the study period and then after 2000 dropped slightly and stabilised around 0.2. 
-This reflects the
-early concentration of high website density around London, which over
-time diffused as high website density clusters can be seen in other
-parts of the country away from London (Figure \ref{lisa}). An almost
-reverse pattern can be observed for OA. At the beginning of the study
-period Moran's I was around 0.5 and it plateaued after 2000 around 0.2.
-Because of the very small size of OA, at the early stages of the
-diffusion of web technologies their adoption was spatially scattered.
-This is reflected in the lack of any significant clusters in 1996
-(Figure \ref{lisa}). Eventually, as the adoption rate increased, more
-such clusters of high website density were formed and this is reflected
-both in the Moran's I and the LISA maps in Figures \ref{morani} and
-\ref{lisa}. A similar pattern is observed for the extended dataset (see
-Figures\ref{morani10} and \ref{lisa10}). In this case, the magnitude of
-spatial autocorrelation is slightly higher illustrating a stronger
-spatial clustering mechanism. All in all, the exploratory spatial data
-analysis advocates towards an underpinning neighbourhood mechanism, which 
-is present at different scales. Interestingly, the magnitude of Moran's I
-is similar to the ones revealed by @ding2010modeling about mobile phone
-adoption in China albeit the much more detailed spatial scale of this
-paper.
-
-```{r, morani, echo=FALSE, message=FALSE, fig.cap="\\label{morani}Website density Moran's I"}
-
-path.in.la <- paste0(path, "/outputs/lisa/corrected/LA/morani_la.csv")
-path.in.oa <- paste0(path, "/outputs/lisa/corrected/oa/morani_oa.csv")
-
-la.m <- read_csv(path.in.la) %>% 
-  filter(p < 0.001) %>% # all values were significant
-  mutate(spatial.unit = 'LAD')
-oa.m <- read_csv(path.in.oa) %>% 
-  filter(p < 0.001) %>% # all values were significant
-  mutate(spatial.unit = 'OA')
-
-bind_rows(la.m, oa.m) %>% 
-  ggplot(aes(x=year, y=morani, fill= spatial.unit)) +
-  geom_bar(stat = 'identity', position = 'dodge') +
-  ylab('Moran\'s I') +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5), 
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  scale_fill_manual('Spatial unit', 
-                    values = c("LAD" = "grey80",
-                                "OA"="grey48"),
-                     limits = force) +
-  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
-  theme(legend.position="bottom",
-        text = element_text(size = 7),
-        legend.text=element_text(size=7),
-        legend.title=element_text(size=7),
-        axis.title=element_text(size=7),
-        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
-```
-
-```{r, echo=FALSE, message=FALSE, fig.cap="\\label{lisa}\\centering Website density LISA maps", out.height="80%"}
-
-path.image1 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la1996.png")
-path.image2 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la2000.png")
-path.image3 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la2012.png")
-path.image4 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa1996.png")
-path.image5 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa2000.png")
-path.image6 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa2012.png")
-path.image7 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la_legend.png")
-
-p1 <- ggdraw() + draw_image(path.image1)  
-p2 <- ggdraw() + draw_image(path.image2)  
-p3 <- ggdraw() + draw_image(path.image3)  
-p4 <- ggdraw() + draw_image(path.image4) 
-p5 <- ggdraw() + draw_image(path.image5)
-p6 <- ggdraw() + draw_image(path.image6)  
-p7 <- ggdraw() + draw_image(path.image7)
-
-par(mar=rep(0,4)) # no margins
-
-library(patchwork)
-# (p1 | p4)/(p2 | p5)/(p3 | p6)/ p7 +
-#   #plot_layout(widths = 1)
-#     plot_layout(widths = c(1, -0.5))
-
-# p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 / p7 +
-#   plot_layout(ncol = 3, widths = c(1, -0.85, 1), heights = c(1,1,1,2))
-
-p <- p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 +
-  plot_layout(ncol = 3, widths = c(1, -0.8, 1), heights = c(1,1,1))
-
-# p + inset_element(p7, left = 0.45, bottom = -0.5, right = 0.6, top = 0.1)
-p + inset_element(p7, left = 0.29, bottom = 0.8, right = 0.49, top = 1.2)
-```
-
-To illustrate whether a hierarchical process also underpins the
-diffusion of web technologies -- the third diffusion mechanism that is tested 
-here -- the Gini coefficient is calculated yearly
-both for LAD and OA. As a metric of inequality, the Gini coefficient
-demonstrates whether website density is concentrated in a small number
-of LAD or OA, or whether it is more equally spread across the country.
-Both scales of analysis in Figure \ref{gini} illustrate the same
-picture. At the beginning of the commercial Internet, website density
-was extremely unequal, or, in other words, only a few places had
-websites associated with them. Inequality dropped and plateaued after
-2000 for both scales. This is illustrative of a hierarchical diffusion
-mechanism that led over time to a more equal spread. Interestingly, the
-year 2000 is again a period of change for this diffusion mechanism as it
-was for the neighbourhood process. There is a substantial difference
-between the Gini coefficient magnitude for LAD and OA, but this is
-expected as the very small size of OA equates to a lot of polygons
-without any websites pointing to them -- for example residential OA.
-
-```{r, gini, echo=FALSE, message=FALSE, fig.cap="\\label{gini}Website density Gini coefficient"}
-
-path.in.la <- paste0(path, "/outputs/gini/gini_la.csv")
-path.in.oa <- paste0(path, "/outputs/gini/gini_oa.csv")
-
-la.gini <- read_csv(path.in.la) %>% 
-  mutate(spatial.unit = 'LAD')
-oa.gini <- read_csv(path.in.oa) %>% 
-  mutate(spatial.unit = 'OA')
-
-bind_rows(la.gini, oa.gini) %>%
-  rename(Gini = 'gini(a$n)') %>% 
-  ggplot(aes(x=year, y=Gini, fill= spatial.unit)) +
-  geom_bar(stat = 'identity', position = 'dodge') +
-  ylab('Gini') +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5), 
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  scale_fill_manual('Spatial unit', 
-                    values = c("LAD" = "grey80",
-                                "OA"="grey48"),
-                     limits = force) +
-  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
-  theme(legend.position="bottom",
-        text = element_text(size = 7),
-        legend.text=element_text(size=7),
-        legend.title=element_text(size=7),
-        axis.title=element_text(size=7),
-        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
-```
-
-After discussing the descriptive findings, the validity of the above diffusion
-mechanisms at the local scale is assessed with a predictive modelling framework
-described in @sec-datamethods. The reader is reminded that four models are 
-estimated: for the OA and LAD, firstly all data points are employed to 
-assess the capacity of the three mechanisms to to predict the adoption of the 
-Web at these two scales; then, the LAD/OA from one of the twelve regions are
-recursively held out to assess the spatial heterogeneity of these mechanisms
-across the different UK regions.
-
-@tbl-model-metrics presents the model performance for the first set of
-models, for which all data points are employed for training and testing
-via CV. The first one is trained and tested on 374 LAD and the second on
-232,296 OA, both for the 16 year period (1997-2012). The results are
-remarkably good considering that the are the outcome of space and time
-sensitive CV, so the the model does not suffer from overfitting. At the
-LAD level the model predicts 83% of the variation of website density.
-Both error metrics indicate that the model error is fraction of a website 
-per firm (0.020 and 0.029). At the OA the $R^2$ drops down to 33%. Considering 
-its granularity, this is still a remarkable performance. To contextualise it, 
-the model results in a $MAE$ equal to 1.2 websites for areas small enough to 
-host 40-250 households. Because of the small size of the spatial units, the
-distribution is highly skewed and a significant part of them is not
-linked to any websites. In 1997 only 1% of the UK OA were associated
-with at least one website. This should not come as a surprise given that this
-was the very beginning of the commercial Internet and any activities
-with a digital footprint were concentrated in a handful of areas, as 
-illustrated in Figure \ref{lisa}. At the end of the study
-period almost half of the UK OA were not associated with a website.
-Again, given the granularity of the data this was expected.
-
-|                   | $RMSE$ | $R^{2}$ | $MAE$ |
-|-------------------|--------|---------|-------|
-| Local Authorities | 0.029  | 0.824   | 0.020 |
-| Output Areas      | 3.218  | 0.333   | 1.233 |
-
-: Model metrics {#tbl-model-metrics}
-
-The results in @tbl-model-metrics are robust against different
-specifications. As noted earlier, the data imputation process resulted in
-a modest improvement of the predictive capacity of the model. As per
-@tbl-model-metrics-no-correction, all accuracy metrics are only slightly
-worst when the data imputation is not applied. As expected, the
-improvement is higher for the more detailed scale of analysis. Then,
-@tbl-model-metrics-10 presents the model results for the extended
-dataset, which includes websites with up to 10 unique postcodes. While
-the results for LAD are only slightly less accurate, the $R^{2}$ of the
-OA model is much higher in comparison to the base results in
-@tbl-model-metrics. In addition, @tbl-model-growth-metrics in the Appendix
-presents the accuracy results of a LAD model, the dependent variable of which 
-is the yearly growth rate of website density. Even for such a specification
-(see Equation \ref{model.growth} in the Appendix) the predictive capacity of
-the model is exceptionally high ($R^2 = 0.5$). All in all, the modelling exercise
-revealed how well the three diffusion mechanisms predict the diffusion of the 
-Web as a new technology. Importantly, this predictive power is highly robust
-against against different scales and delineations of the diffusion of
-the web.
-
-|                   | $RMSE$ | $R^{2}$ | $MAE$ |
-|-------------------|--------|---------|-------|
-| Local Authorities | 0.032  | 0.781   | 0.021 |
-| Output Areas      | 4.644  | 0.244   | 1.237 |
-
-: Model metrics without data imputation {#tbl-model-metrics-no-correction}
-
-|                   | $RMSE$ | $R^{2}$ | $MAE$ |
-|-------------------|--------|---------|-------|
-| Local Authorities |  0.190 | 0.749   | 0.129 |
-| Output Areas      | 21.591 | 0.455   | 6.845 |
-
-: Model metrics for the extended dataset {#tbl-model-metrics-10}
-
-Figure \ref{var.imp} plots the importance of the different predictors.
-When the focus is on LAD, the website density in the neighbouring LAD, in the
-nearest city and in London the year before are the most important predictors. 
-They are followed by the yearly trend, while the spatial configuration as 
-reflected in distances to London or to the nearest city only plays a minor 
-role. This can be attributed to the relatively coarse spatial scale of LAD.
-Nevertheless, all previously discussed spatial processes are at play in the
-diffusion of the Web at the LAD level: the first two predictors depict 
-the hierarchical mechanism, the spatial and temporal lag of website density the
-neighbourhood mechanism, and the yearly trend the time-sensitive
-cumulative adoption pattern.
-
-At the much more granular scale of OA, the picture is
-almost reversed. The two most important predictors are the distance to
-London and to the nearest city followed by the spatial and temporal lag
-of website density and the distance to the nearest retail centre. They
-still depict a hierarchical mechanism, but proximity to the different
-population centres is more important than their lagged web densities in
-predicting website diffusion. The neighbourhood mechanism is still
-strong but less important at this scale. What is interesting is the
-almost negligible role of the yearly trend and London's website density.
-While the former probably illustrates the large heterogeneity in how the
-Web has been adopted at this very fine scale, the latter highlights that
-the importance of past web adoption rates in large population centres is
-surpassed by proximity to them and spatial configuration at this scale.
-
-The above pattern of variable importance is largely in accordance with
-the models estimated for the extended dataset. A comparison with Figure
-\ref{var.imp10} illustrates that the main difference is the importance
-of the neighbourhood mechanism for the OA as the space and time lag of
-the website density is by far the most important predictor when the
-extended dataset is employed. For the LAD, the importance of the distance to
-the nearest city increases in favour to its website density the previous
-year. But still, the variable importance patterns for the two datasets 
-are not dissimilar.
-
-```{r varimp, eval = T, echo=FALSE, message = F, warning=FALSE, fig.cap="\\label{var.imp}Variable importance", out.height="60%"}
-path.importance.lad <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.csv")
-path.importance.oa <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.csv")
-
-importance.oa <- read_csv(path.importance.oa) %>% 
-  mutate(id = c("LD", "NC", "NR", "STL", "Y", "DL", "DNC", "DNR")) %>% 
-  mutate(variable = str_to_sentence(variable))
-importance.lad <- read_csv(path.importance.lad) %>% 
-  mutate(id = c("LD", "NC", "STL", "Y", "DL", "DNC")) %>% 
-  mutate(variable = str_to_sentence(variable))
-
-importance.oa %>% 
-  rename(OA = importance) %>% 
-  left_join(importance.lad %>% rename(LAD = importance), by = "id") %>%
-  dplyr::select(-variable.y, -id) %>% 
-  rename(variable = variable.x) %>% 
-  pivot_longer(!variable, names_to = "spatial.unit", values_to = "importance") %>% 
-  mutate(order = ifelse(variable == "Distance to london", 1,
-                        ifelse(variable == "London's website density, t-1", 2,
-                               ifelse(variable == "Distance to the nearest city", 3,
-                                      ifelse(variable == "Nearest city's website density, t-1", 4,
-                                             ifelse(variable == "Distance to the nearest retail centre", 5,
-                                                    ifelse(variable == "Nearest retail centre's website density, t-1", 6,
-                                                           ifelse(variable == "Spatial and temporal lag of website density", 7, 8)))))))) %>% 
-  ggplot(aes(x=importance, y=reorder(variable, desc(order)), #forcats::fct_reorder(variable, importance), 
-         fill = spatial.unit)) +
-  geom_col(position = 'dodge') +
-  ylab('') +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5), 
-        axis.text.x = element_text(vjust = 0.5, hjust=1)) +
-  scale_fill_manual('Spatial unit: ', 
-                    values = c("LAD" = "grey80",
-                                "OA"="grey48"),
-                     limits = force) +
-  scale_x_continuous('Variable importance') +
-  theme(legend.position="bottom",
-        text = element_text(size = 7),
-        legend.text=element_text(size=7),
-        legend.title=element_text(size=7),
-        axis.title=element_text(size=7))
-```
-
-Table \ref{table.regions} presents the results of the recursive hold out
-models and allows to (i) highlight the regional heterogeneity of the
-spatial processes behind the diffusion of the Web, and (ii) compare this
-heterogeneity across the two different scales of analysis. It illustrates both 
-regional similarities and differences. At the top of Table \ref{table.regions}
-the regions of South East and the West Midlands have the highest similarity with 
-the rest of the country. In other words, the web density for the LAD of these 
-regions can be very well predicted by models trained on data from the LAD from 
-the rest of the country and, therefore, the spatial diffusion mechanisms for 
-these regions represent quite well the UK's average. This should not come as a 
-surprise as the spatial structure of both regions resembles the broader UK: 
-both regions contain some highly urbanised LAD (South East is well integrated 
-with the Greater London and West Midlands is the home of the city of Birmingham), 
-but also smaller commuter towns as well as more rural settlements. At the other 
-end of the spectrum, the spatial diffusion mechanisms for Northern Ireland, 
-Scotland and North West diverge the most from the country average. Rurality and 
-remoteness seem to be common attributes for these regions. This hierarchy though 
-is not consistent across scales. For the more granular OA models, the spatial 
-diffusion mechanisms for South West, Yorkshire and The Humber and North West 
-regions are closer to the country average. Interestingly, these regions are 
-well-known tourist destinations (Lake District, Peak District and 
-Cornwall-Devon) and the exploratory analysis revealed clusters of high web 
-density only visible at the OA scale (Figure \ref{lisa}). Then, Scotland, 
-Northern Ireland and London have the lowest similarity with the country 
-average. Again, rurality, peripherality, but also the highest level or 
-urbanisation make the spatial diffusion mechanisms of these regions more 
-distinct.
-
-```{r regions}
-path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected.csv")
-path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA_corrected.csv")
-
-lad <- read.csv(path.lad) %>% 
-  rename('RSquared LAD' = Rsquared)
-
-oa <- read.csv(path.oa) %>% rename(Region = test.region,
-                                   'RSquared OA' = Rsquared)
-
-lad %>% left_join(oa) %>% 
-  mutate('Rank OA' = rank(desc(`RSquared OA`)),
-         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
-  relocate(5, .after = 2) %>% 
-  arrange(`Rank LAD`) %>% 
-  kable(caption= "Regional differences\\label{table.regions}", 
-        digits=3,
-        col.names = c("Region", "$R^2$ LAD", "Rank LAD", "$R^2$ OA", "Rank OA")) 
-```
-
-Comparing the above with Table \ref{table.regions.10} in the Appendix, which 
-presents the results for the the extended dataset with websites with up to 10 
-unique postcodes, the results do not differentiate much, signalling the 
-robustness of the analysis. The most salient difference has to do with Northern 
-Ireland. According to the extended sample at the granular OA level, Northern 
-Ireland's spatial diffusion mechanisms are closer to the UK's average than the 
-results based on the websites with a unique postcode. While the latter are 
-websites and, consequently, economic activities solely based in Northern 
-Ireland, websites with multiple postcodes also include websites with postcodes 
-in Britain and, therefore, can be characterised by slightly different spatial 
-diffusion processes.
-
-## Early adoption and latecomers
-
-<!-- H~2~: The diffusion of the commercial Web is shaped by both early adoption  -->
-<!-- and latecomer effects. -->
-
-To test the second hypothesis (H~2~), the spatial heterogeneity of Web adoption 
-speeds is firstly analysed. As highlighted in the literature [@risk_perceptions], 
-different agents have different perceptions about and levels of acceptance of 
-the risks and the potential economic returns associated with the adoption of 
-new technologies -- see for instance the seminal work of @venkatesh2000theoretical.
-To reveal such patterns, Figure \ref{s_map} maps whether LAD reached their 
-inflection point earlier or later than the UK in total and based on this are 
-labelled as *fast* or *slow* accordingly. Inflection points have been used in
-the relevant literature to compare diffusion processes across different settings 
-and technologies [@cherp2021national; @woo2017exploring] as they can enable more 
-insightful comparisons than, for example, the average growth over the 1996-2012
-study period.
-
-The map illustrates a clear high concentration of fast adopting LAD in the 
-South and East of England, as expected. There are also some concentrations in 
-Wales, Scotland and the North West of England. At a more detailed level, there 
-are some expected cases of LAD with relevant industrial backgrounds delineated 
-as fast: the City of London, a world-renowned cluster of finance industries
-[@cook2007role], and Reading, a town with high-tech service industries
-in proximity to London and its main airport, Heathrow
-[@england2005polynet]. However, there are also LAD which were expected
-to appear as fast -- e.g. Hackney in central London and Bristol, a
-well-established creative cluster [@oatley1999cultural;
-@bassett2002cultural] -- but are delineated as slow. Similar to 
-@beardsell1999spatial, who analysed the spatial evolution of computer industry 
-employment across 317 US metro areas, spatial heterogeneity is evident in the 
-diffusion speed when the scale of analysis is as detailed as the LAD.
-
-Importantly, there is obvious spatial clustering of fast and slow LAD. Out of 
-the ten fastest LAD based on when they reached their inflection point (see 
-detailed Table \ref{table.s.lads} in the Appendix), nine were located in the 
-South East of England and London and one in Cambridge. The above observations 
-remain almost unchanged when the analysis includes websites with up to 10 unique
-postcodes -- see Figure \ref{s_map10} in the Appendix -- demonstrating the
-robustness of the analysis. Given (i) the expectation that 
-early adopters of new technologies are often rewarded [@rogers2010diffusion; 
-@griliches1957; @ding2010modeling], and (ii) the previous findings 
-illustrating the long term and sustained productivity benefit for the early 
-adopters of Web in the UK [@tranosuk], the observed spatial clustering of LAD 
-with fast diffusion rates can further reinforce existing and rooted patterns of 
-unequal economic productivity in the UK [@mccann2022productivity].
-
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map}LAD Web diffusion speed, 1996-2012"}
-path.image <- paste0(path, "/outputs/s/speed_map.png")
-knitr::include_graphics(path.image)
-```
-
-To reveal patterns of early adoption and leapfrogging, the stability and 
-volatility of the above patterns is analysed. Figure \ref{rank} plots the 
-relative ranking of LAD based on website density at the beginning and the end 
-of the study period. To decrease noise, the average ranking of 1996-1998 and 
-2010-2012 is plotted instead of the individual years. Each line represents a 
-single LAD and the colours depict the 
-5\textsuperscript{th} and 95\textsuperscript{th} percentile of the absolute
-difference of the LAD ranking between 1996-1998 and 2010-2012. In other words, 
-the 5\textsuperscript{th} percentile includes the LAD, the relative
-position of which remained unchanged between the beginning and the end of 
-the study period, both at the top and at the bottom of the hierarchy (small
-absolute difference in ranking). The 95th percentile includes LAD that either
-increased or decreased a lot their relative positions over time (high absolute 
-difference in ranking). Thus, Figure \ref{rank} reveals cases of 
-*leapfrogging* since the LAD at the top-right part of the figure managed to 
-jump at the top of the hierarchy at the end of the study period. 
-<!-- which are often associated with latecomers benefits in the  -->
-<!-- literature. -->
-
-```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank}Dynamics of Web diffusion based on LAD rankings"}
-path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av.png")
-knitr::include_graphics(path.image)
-```
-
-To further analyse the above, Table \ref{table.late} juxtaposes the 
-volatility of the LAD Web diffusion against the early conditions as it reports 
-the $t_0$ estimate (inflection point) for the leapfrogging LAD identified in
-Figure \ref{rank}. Indeed, more than half of the LAD that drastically increased 
-their relative position at the end of the study period reached the inflection 
-point $t_0$ later than the country in total and, therefore, the Web diffusion 
-speed has been relatively slow for them. The fact that these LAD experienced
-relatively faster diffusion rates at the end of the study period, which is 
-aligned with the maturity period of the web, can lead to potential positive 
-economic gains for these LAD. There is extensive literature describing the 
-benefits of such technological leapfrogging and the underpinning argument is 
-that latecomers can adopt and benefit from new technologies that have been 
-developed elsewhere without incurring the hefty initial R&D costs 
-[@teece2008firm].
-
-Based on Figure \ref{rank} and Table \ref{table.late}, the opportunity 
-to catch up that @perkins2005international illustrated at the country level, is 
-also visible at the much more detailed LAD scale. Although the relevant 
-literature did not pay attention to cities and regions [@yu2018sustainability], 
-the above finding advocates towards further exploring leapfrogging effects at 
-such a granular scale. As previously, this is consistent with including websites 
-with up to 10 postcodes (see Figure \ref{rank10}).
-
-All in all, the diffusion of the Web at the detailed scale of LAD is 
-characterised by spatial heterogeneity, well-established spatial patterns and 
-both stability and volatility. Although the relevant literature neglected local
-scales, the above findings illustrated that such local dynamics exists and 
-worth further consideration and investigation especially regarding potential 
-local economic effects.
-
-```{r late/early}
-late.path <- paste0(path, "/outputs/ranks/late.csv")
-
-read_csv(late.path) %>%
-  dplyr::select(-'early/late') %>% 
-  rename('$t_{0}$ estimate' = estimate,
-         "Diffusion speed" = fast) %>% 
-  kable(caption= "Late adopters\\label{table.late}", 
-        digits=3,
-        #col.names = c("LAD", "$t_{0}$ estimate", "Diffusion speed", "Early/late"),
-        align = c("l", "c", "c"),
-        escape = F)  
-  #kableExtra::row_spec(10, hline_after = TRUE)
-```
-
-# Conclusions {#sec-conclusions}
-
-<!-- ADD MORE ABOUT LOCAL POLICIES: E.G. USE MY RESULTS TO MODEL WHERE INFRASTRUCTURE OR GSM ANTENNAS SHOULD BE LOCALTED -->
-
-The paper offers an economic geography study of the diffusion of a new 
-technology -- the Web. Contrary to previous studies, it does so by adopting 
-(i) a high level of spatial granularity, and (ii) a long study period that
-captures the very early stages of this technology until its maturity.
-This is the first time that the importance of granular spatial
-mechanisms for the diffusion of a new technology is exposed. 
-Importantly, the empirical framework and the data employed here allows
-to observe the active engagement with the Web that is creating and
-maintaining a commercial website instead of the more passive action of
-browsing webpages and accessing the Internet.
-
-<!-- H1 -->
-Specifically, two hypotheses are tested. Firstly, whether the three diffusion 
-mechanisms namely distance, urban hierarchy and the S-shaped pattern of the 
-cumulative level of adoption shaped the diffusion of the commercial Web in 
-the UK at granular geographical scales. The analysis firstly illustrates how 
-well the S-pattern of cumulative adoption fits at the local scale and offers 
-some descriptive insights about the importance of the distance and 
-hierarchical diffusion mechanisms. Then, the modelling exercise reveals the 
-remarkably good predictive capacity of these mechanisms at granular geographical 
-scales. The modelling framework exposes that the relative importance of 
-these spatial mechanisms differ across different scales. It also reveals 
-spatial heterogeneity. While for some UK regions the importance of the spatial 
-diffusion mechanisms are very similar to the rest of the country, other regions
-diverge significantly from the country average.
-
-To do the above, a robust and novel ML modelling framework was designed to 
-predict website diffusion over space and time using as predictors the spatial 
-diffusion mechanisms. Space- and time-sensitive CV was designed for the model
-to avoid overfitting. Importantly, the results are robust against different 
-specifications and subsets of the data. This is the first time that the power 
-of these mechanisms in predicting the diffusion of a new technology was
-empirically tested, importantly, at very fine scales.
-
-<!-- H2 -->
-<!-- H~2~: The diffusion of the commercial Web is shaped by both early adoption  -->
-<!-- and latecomer effects. -->
-The second hypothesis assessed whether the diffusion of the web is shaped by 
-both early adoption and latecomer effects. On the one hand, the analysis 
-highlighted the heterogeneity of adoption rates between LAD as well as the 
-spatial clustering of LAD with similar adoption rates. Importantly, the 
-observed spatial clustering of LAD with fast diffusion rates can
-reinforce existing and rooted patterns of unequal economic productivity 
-in the UK [@mccann2022productivity]. Previous research had already indicated 
-the long-term productivity gains of the early adoption of such technologies 
-[@tranosuk]. On the other hand, the analysis also exposed instances of 
-remarkable stability as well as volatility of the Web adoption, with the 
-latter being indicative of leapfrogging. While the literature had 
-discussed the benefits of technological leapfrogging at the national scale 
-[@teece2008firm; @perkins2005international], local scales had escaped such 
-studies [@yu2018sustainability]. Thus the leapfrogging evidence presented here 
-advocate towards further exploring such effects at granular scales.
-
-The paper offers a number of different contributions. Firstly, it
-assesses, for the first time, the role of these spatial diffusion
-mechanisms at such granular spatial scales. While previous studies had
-highlighted these mechanisms [@hagerstrand1968innovation;
-@rogers2010diffusion; @grubler1990rise], their role had never been
-tested before at such granular scale. The findings talk directly to this
-literature [e.g. @beardsell1999spatial; @haller2011determinants;
-@lengyel2020role]. More broadly, the paper offers a robust
-methodological framework, which utilises novel data and modern ML
-algorithms and techniques to model the diffusion of a new technology
-across space and time. It directly addresses the data problem which
-prevented economic geographers in engaging with diffusion studies
-[@iso2005; @kemeny2011international; @zook2022mapping]. Thus, the paper
-seeks to renew economic geography's interest in the diffusion of new
-technologies. Current data science advances including the availability
-of granular enough data and adequate methods can enable economic
-geographers to reveal interesting spatial patterns of diffusion of new
-technologies as this paper has demonstrated. This is important as (i)
-various stakeholders are interested in predicting localised demand for
-new technologies for planning purposes, for example 5G networks
-[@leibowicz2016representing;@meade2021modelling], and (ii) the adoption 
-of such digital technologies is associated with local economic gains 
-[@solow1957technical; @aghion1990model; @kemeny2011international; @tranosuk;
-@capello2024nexus]. 
-<!-- In addition, while most related studies focused on -->
-<!-- tangible technologies, this paper sheds light on the diffusion of a -->
-<!-- truly weightless digital technology that is the Web. Thus, it adds much -->
-<!-- more granular findings to the limited related studies -->
-<!-- [@sinai2004geography; @haller2011determinants].  -->
-Importantly, although the geographical focus of this paper is the UK, the 
-empirical framework developed here is easily transferable to other geographical 
-contexts.
-
-<!-- reset the figure, table and section numbering -->
+<!-- To remove the corresponding author footnote, the appendix.tex should be manually tweaked: delete the \cortext[cor1]{Corresponding author} part. -->
 
@@
 \renewcommand{\thefigure}{A\arabic{figure}}
+\setcounter{equation}{0}
+\renewcommand{\theequation}{A\arabic{equation}}
 
@@
 essence, this is a model similar to the one estimated based on Equation
-\ref{eq:model} and presented in @tbl-model-metrics, with the main difference 
+3 and presented in Table 2, with the main difference 
 being the dependent variable which is the year growth rate of website 
@@
 
-# References {#references .unnumbered}
+<!-- # References {#references .unnumbered} -->

### 2025-09-28 paper/submission4/tranos2025.qmd
appendix

@@
 calculated using data from the Internet Archive[^2] and, specifically, the 
-JISC UK Web Domain Dataset [@ukwebarchive]. The Internet Archive is one of the 
+JISC UK Web Domain Dataset [@ukwebarchive][^3]. The Internet Archive is one of the 
 most complete and oldest archive of webpages in the world operating since
@@
 [^2]: [https://archive.org/](https://archive.org/).
+[^3]: The data used here is directly available from @tranos_2025_16610235.
 
@@
 because of the the consistent spatial definition of OA (they host 40-250
-households),[^3] website counts in OA are interpreted as a density metric too.
-
-[^3]: <https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies>
+households),[^4] website counts in OA are interpreted as a density metric too.
+
+[^4]: <https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies>
 
@@
 third predictor is added to capture hierarchy at a much more granular scale:
-a one year lag of the website density in the the nearest retail centre [^4]. As
+a one year lag of the website density in the the nearest retail centre [^5]. As
 the literature illustrates, retail centres are key components of urban systems
@@
 
-[^4]: The retail centres data have been obtained by the Consumer Data Research Centre, an ESRC data investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1: <https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators>.
+[^5]: The retail centres data have been obtained by the Consumer Data Research Centre, an ESRC data investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1: <https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators>.
 

### 2025-09-28 paper/submission4/tranos2025.tex
appendix

@@
 that are calculated using data from the Internet Archive\footnote{\url{https://archive.org/}.}
-and, specifically, the JISC UK Web Domain Dataset \citep{ukwebarchive}.
-The Internet Archive is one of the most complete and oldest archive of
-webpages in the world operating since 1996
-\citep{ainsworth2011much, holzmann2016dawn}. It is a web crawler, which
-discovers webpages by following the hyperlinks of every webpage it
-archives. This dataset, which is curated by the British Library,
-contains all the archived webpages from the UK country code top level
-domain (ccTLD - .uk) from the 1996--2012 period. In essence, this is a
-long list of 2.5 billion URLs of archived webpages including also the
-archival timestamp.
+and, specifically, the JISC UK Web Domain Dataset
+\citep{ukwebarchive}\footnote{The data used here is directly available
+  from \citet{tranos_2025_16610235}.}. The Internet Archive is one of
+the most complete and oldest archive of webpages in the world operating
+since 1996 \citep{ainsworth2011much, holzmann2016dawn}. It is a web
+crawler, which discovers webpages by following the hyperlinks of every
+webpage it archives. This dataset, which is curated by the British
+Library, contains all the archived webpages from the UK country code top
+level domain (ccTLD - .uk) from the 1996--2012 period. In essence, this
+is a long list of 2.5 billion URLs of archived webpages including also
+the archival timestamp.
 
